### 2016-06-13 redaction/dhp_draft_new.tex
changement nom

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
%\usepackage[nolists, figuresfirst]{endfloat}
%\usepackage{sidefloat}

\setlength{\abovecaptionskip}{-8pt}

%\usepackage[usenames]{color}
%\definecolor{grey}{rgb}{0.35,0.35,0.35}
%\definecolor{webdarkblue}{rgb}{0,0,0.4}
%\definecolor{orange}{rgb}{0.7,0.2,0.05}

%\usepackage[pdfcreator={PDFLaTeX}, pdfproducer={PDFLaTeX}, pdfstartview=FitH, pdfpagemode=UseOutlines, pagebackref=false, colorlinks={true},
%citecolor={webdarkblue}, linkcolor={webdarkblue},
%urlcolor={webdarkblue}]{hyperref}

%\addtolength{\oddsidemargin}{-0.4in}
%\addtolength{\evensidemargin}{-0.4in}
%\addtolength{\textwidth}{0.8in} \addtolength{\topmargin}{-0.85in}
%\addtolength{\textheight}{1.7in}
\renewcommand{\baselinestretch}{1.2}

\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}

% commands

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}
\newcommand{\beqa}{\begin{eqnarray}}
\newcommand{\eeqa}{\end{eqnarray}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bs}{\bigskip}
\newcommand{\A}{$^a$}
\newcommand{\B}{$^b$}
\newcommand{\C}{$^c$}
\begin{document}

\date{\today}
\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX. Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01. Finally, any remaining errors are ours.}}}
%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
\author{Guillaume \textsc{Daud}\thanks{%
Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa), Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa)} } \maketitle

\bigskip

\begin{abstract}

\begin{spacing}{1.2}

{\normalsize XXXX}

{\normalsize \bigskip }

\end{spacing}
\end{abstract}

\thispagestyle{empty} \pagestyle{plain} \setcounter{page}{1}

{\normalsize JEL classification: F14, N70, R40\newline
Keywords: Transport costs estimates, transport costs determinants, non-linear econometrics, period 1974-2013, additive transport costs }

{\normalsize \vspace{0cm} }

{\normalsize \titlepage }

{\normalsize \newpage }

%\begin{spacing}{1.5}

\section{Introduction}

\noindent Trade costs remain central in international economic analysis. Defined as the costs associated with the exchange of goods across national borders, they are a major obstacle to international economic integration. In addition, they might be the key to the explanation of major puzzles in international macroeconomics (\citealp{rogoff}). Historical perspective show that indeed, trade costs variations are a major determinant of the level of trade flows. According to the estimates by \citet{Jacks08}, trade cost declines explain around 55 percent of the pre World War I trade boom and 33 percent of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter war trade collapse. After 1950, average trade costs fell by 16\%. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.

Trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff-costs), time costs (time to ship goods) and transport costs \emph{per se}. Many believe that the latter have dramatically decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \citet{Glaeser04} find that, over the twentieth century, the costs of moving goods have declined by over 90\% in real terms. However, \citet{hummels2007} shows that the bulk of price declines in transportation comes from air shipping, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004; concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, even if the rise in containerization lowered shipping costs from 3 to 13\%. Studies overviewed by \citet{Behar_Venables} are consistent with those results: the fall in measured transport costs has been relatively small. The importance of fuel costs and technical progress (which improved speed and reliability rather than decreased costs) in transport explain that result.

Despite the decrease in distance-related costs\footnote{\citet{Behar_Venables} suggest freight costs account for two thirds of the effect of distance, the remaining third ebmoyding more or less the effect of time}, several papers (mostly based on empirical estimates of the gravity equation) in the last decade have shown that distance remained a major obstacle to trade (e.g. \citealp{Head_Mayer04} and \citealp{Disdier_Head08}). \citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs (1.7 $=$ 1.55 $\times$ 1.74 - 1). Again, the latter comprises 44\% border-related trade barriers and 21\% transportation costs. That means that transportation costs are higher than each component of border costs taken separately (namely, 8\% for tariff and non-tariff costs, 7\% for language differences, 14\% for for currency differences and 9\% for other costs).\footnote{1.44 $=$ 1.08 $\times$ 1.07 $\times$ 1.14 $\times$  1.09)}. Therefore, the share of transport costs in the consumer price of manufactured goods remains high (\citealp{Lafourcade_Thisse}). Besides, \citet{combes_lafourcade} emphasize that the figures provided by \citet{anderson_wincoop_jel} should be considered as a a downward benchmark, and could be higher in developing economies. Despite technical progress and infrastructure development, transport costs remain large and deserve attention. In that respect, \citet{Behar_Venables} support that the elasticity of trade with respect to freight costs is around -3, which remains a large number.

Following \citet{samuelson1954}, standard models of international trade have usually relied on the iceberg cost hypothesis, implicitly reducing trade costs to their transport part, and modelling the latter as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded). This implies that more expensive goods are costlier to trade. However, many common empirical facts support the existence of additive costs: pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure (see \citealp{Irrazabal_2015}). The additive versus iceberg trade costs is not a new debate in the international trade literature. The famous Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: the demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out''). \citet{martin2012} gives a strong empirical support to this conjecture: based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite. \citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: the elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption. Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs. Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level. Based on US imports data, he finds that a 10\% increase in the price of an item increases the shipping cost by 9.5\%. The additive structure of trade costs emerges only when weight is forced to be independent of price. Besides, \citet{Lashka15} supports that higher-priced, still exported to faraway markets, varieties are heavier and involve proportionally higher shipping costs, the latter increasing the relative demand for high-markup varieties. Markups would therefore be the driving force for explaining the ``shipping the good apples out'' effect in the presence of \emph{ad-valorem} trade costs.

A correct measure of transport costs is also important for understanding several key features of international trade. For example, the existence of additive costs may explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows. Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e. more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments. Besides, several recent papers show that the existence of additive trade costs has important non trivial effects when assessing the welfare effects of trade liberalization. \citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive (per-unit) trade costs, in addition to iceberg/multiplicative costs. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs. The difference arises from dissimilar adjustments at the extensive margin - fewer firms enter into the export market following a reduction in additive costs. Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff.

Our purpose in this paper is to provide results on the size of transportation costs over time, explicitly distinguishing between multiplicative and additive parts. Our paper stands at the crossroad of these two strands of the literature, the first one dealing with the dynamics of transport, and more generally trade costs over time, the second one focusing on the measurement and nature of the latter. More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over a large time dimension - several decades. To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}. Therefore, our period of study runs from 1974 to 2013. Closely related to this paper is the work by \citet{Irrazabal_2015}, who develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level data, their results suggest that additive costs are on average 14 percent, relatively to the median price, and that they are strongly correlated with standard proxies for trade costs (like e. g., distance). However, while our data requires that we focus only on transportation costs, our approach departs from theirs in several key aspects.

First, our theoretically agnostic approach aims at letting the data speak, and provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models; to our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the shares of multiplicative and additive costs in total transport costs. Second, we provide, through standard measures of ``goodness-of-fit'', an empirical assessment of what standard international trade models lose by skipping additive transport costs. Third, the time dimension of our data allows us to characterize the evolution of transportation costs, and more precisely the dynamics of the relative share of each specific (multiplicative and additive) cost over a forty years time span; to our knowledge, no firm-level database is able to provide such a time coverage. Fourth, we are also able to discriminate our results by transport mode (air and vessel).

Finally, our analysis is also indirectly related to a large strand of the international trade literature assessing the impact of trade costs on international exchanges. Following \citet{anderson_wincoop} and \citet{anderson_wincoop_jel}, there has been a huge bunch of articles relying on gravity equations to infer indirectly trade costs as tariff equivalents, or losses related to various kinds of trade barriers - many of these works are surveyed in \citet{anderson_wincoop}; among the various subsequent works, see in particular \citet{anderson_yotov}, \citet{novy13} or \citealp{chen_novy}). However, our purpose in the present paper is different. We first infer directly transport costs from price data, discriminating between multiplicative and additive parts. In a second step, following an approach similar to \citet{Behar_Venables}, we relate the previous measures to traditional determinants of transport costs (like e. g., distance), in order to understand how differently the latter impact the multiplicative and additive transport costs.

%time as a trade barrier, on risque d'avoir des problèmes avec ça... \citet{Lafourcade_Thisse},

XXX Here a summary of our results and contribution XXX

The next section presents the data and our fundamental equation. Section 3 reports our baseline estimates for multiplicative and additive measures of transport costs across time, as well as a number of robustness checks. Section 4 discusses the results of estimates relating traditional determinants of transport costs to the previous measures, while Section 5 XXX has yet to be discussed!!! XXX. The last section concludes.

%\noindent More generally,  Among the closest to our own work, Eaton et al. (2002) develop a Ricardian general equilibrium trade model to estimate parameters relating to absolute advantage, to comparative advantage, and to geographic barriers. Anderson and Wincoop (2003) elaborate a theory-founded gravity equation to assess the cost of trade frictions, and use it to provide a direct measure of the trade losses due to national borders (between 20 and 50 percent).

\section{Data Sources and Empirical Methodology}

\subsection{A measure of Transportation Costs}

\noindent

Our strategy for representing transportation costs is relatively standard, and consists in exploiting the difference between commodity-level export and prices. We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, origins and transportation mode. More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs. The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price. Strictly higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.

To have the import price of the good (CIF price), next to the export price (FOB price) is highly valuable. Indeed, this allows us thus to estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs. This differentiates us from \citet{Irrazabal_2015}, that can only estimate the ratio of additive costs as a share of the tola price. However, note that this measure based on \emph{freight} cost ignores another dimension of transportation costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.

\subsection{Database}

%laius sur pourquoi pas de donnes de firmes (triop comutational), ni comtrade (trop mauvais, citer RIE) XXX

\noindent Even if the data of transportation costs we rely could be available at the firm-level, the use of a nonlinear estimator (see below) triggers computational limitations that do not make them a likely option. A product-level bilateral database of CIF-FOB ratios is theoretically possible, and could be built from the well-known UN's COMTRADE data. Unfortunately, \citet{hummels_rie} show that issues inherent to the construction of data (especially the use of ``mirror'' flows) make this database unreliable for our purpose.

We decide therefore to use database for a single country, but with strong statistical reliability arising from a single, trustworthy customs origin. More precisely, the empirical analysis relies on yearly data from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013. Based on customs declarations, this dataset inventories all imports (both values and quantities) by origin to the United states at the HS 10-digit highly disaggregated level, with a concordance code to the SITC 5-digit coding system. In addition, the database reports information regarding freight expenditures and transportation mode (Ocean Vessel and Air). The first will be crucial to compute transport costs (see below), the second will alow us enlightening potential different dynamics of transport costs across transportation mode.

Due to the abovementioned computational limitations, data are reaggregated at the 3- and 4-digit level: comparing different levels of aggregation will be useful to check differences and the presence of biases precisely due to aggregation. Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products.

% Why the gap CIF-FOB is a good measure of trade costs (cf mimeo by Gaulier and Zignago)? JH: to me, This is not the right question to ask. We have the freight costs in our data. The true question is: how well these freight costs proxy correctly transportation costs as a whole.

\subsection{Key empirical specification}

\noindent Our purpose is to provide estimates over time of the relative shares of multiplicative and additive costs among total transport costs. To do so, we start from a very simple equation, quite alike the one proposed by \citet{hummels2010}:

\begin{equation}\label{eq:base}
p = \tau \widetilde{p} + t
\end{equation}

or as a ratio

\begin{equation}\label{eq:base_bis}
\frac{p}{\widetilde{p}} = \tau  + \frac{t}{\widetilde{p}}
\end{equation}

This equation relates the ratio of the consumer price (ie, cif price, denoted $p$) to the producer price (ie, fob price, denoted $\widetilde{p}$) to the two types of costs (multiplicative and additive). Iceberg trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs) while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs).
%\begin{equation*}
%p =\tau \widetilde{p}+t
%\end{equation*}
We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis. Let us denote $i$ the origin country and $k$, the sector. Hence, the equation to be estimated is given by:

\begin{equation}\label{eq:base_estimee}
\frac{p_{ik}}{\widetilde{p}_{ik}} = \tau_{ik}+\frac{t_{ik}}{ \widetilde{p}_{ik}}
\end{equation}

\section{Decomposing the Dynamics of Transport Costs over Time}

\subsection{Estimation Strategy}

% nonlinear least squares: problem with heteroskedasticity???

We transpose the approach by \citet{Irrazabal_2015} by considering that both multiplicative and additive costs are separable between origin country and products, in a multiplicative way for the former and an additive way for the latter. In other words, $\tau$ and $t$ from equation~\ref{eq:base_estimee} become:

%\begin{eqnarray}
%\tau_{ik} &=& dist^{\rho_I}F^I_i \label{eq:iceberg}\\
%t_{ik} &=& dist^{\rho_A}F^A_i-1 \label{eq:add}
%\end{eqnarray}

\begin{eqnarray}
\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:iceberg}\\
t_{ik} &=& t_{i} + t_{k} \label{eq:add}
\end{eqnarray}

As a result, our underlying theoretical equation is specified as:

%\begin{equation}
%\frac{p_{ik}}{\widetilde{p}_{ik}} =dist^{\rho_I}F^I_ik+\frac{dist^{\rho_A}F^A_ik-1}{ \widetilde{p}_{k}}\label{eq:estimatedequation}
%\end{equation}

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} =\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
\end{equation}

The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the \emph{cif} price $p$ cannot be lower than the \emph{fob} price: $p>\widetilde{p}_{ik}$. Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term. More precisely:

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right)\times \varepsilon_{ik}
\end{equation}

Log-linearization gets the following, tractable empirical equation:

\begin{equation}
ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= ln(\varepsilon_{ik})$ .

Still following \citet{Irrazabal_2015}, we report a ``synthetic estimate'' of each type of transport cost, by averaging all $\widehat{\tau}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}_{ik} = t_{i} + t_{k}$ using the values of each trade flow over total yearly trade as a weighting scheme. Therefore, we recover for each year and transportation mode a $\widehat{\tau}$ and $\widehat{t}$ that will be the results we will report for step 1.\\

\subsection{Econometric Issues}

\noindent Our primary interest relates to the importance of additive costs relatively to iceberg costs. Put differently, what traditional models of international trade lose by ignoring additive costs? A natural way to answer this question is to perform estimations of equation~\ref{eq:estimatedequation} constraining $t$ to be equal to zero, and compare the fitting properties/explanatory power of the restricted and complete models. This is done bye computing several standard diagnostic statistics (R$^{2}$ and statistics based on the likelihood function).

The shape of our main equation \ref{eq:estimatedequation} is such that estimations cannot be performed using standard linear estimators. Therefore, all estimates are performed using non linear least squares. The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next. In our case, due to computational limitations implied by the size of our dataset, we implement 100 iterations and set the convergence criterion for successive parameter estimates and for the residual sum of squares at 0.01.

Finally, to control for the potential influence of outliers, we excluded observations in the 5 percent from the upper and lower tails of the distribution in the regression variables, and all our three measures of trade costs are bounded by 0 as minimal value. These cut-offs are aimed at eliminating reporting or coding errors. %(XXX develop a little bit?XXX)

%\noindent with the exponents $I$ and $A$ respectively identifying the iceberg and the additive trade costs, $\rho_I$ and $\rho_A$ being the parameters of interest to be estimated. The adding of $-1 $ in the specification for the additive costs (\ref{eq:add}) is necessary to be consistent with $t$ starting from 0. In the absence of additive trade costs, ie $\rho_A=F^A_i=F^A_j=0$, the specification (\ref{eq:add}) consistently implies $t=0$. As well, if iceberg costs are null ($\rho_I=F^I_i=F^I_j=0$), the specification (\ref{eq:iceberg})) correctly implies $\tau=1$.

\subsection{Results: Distinguishing Additive and Multiplicative Parts of Trade Costs}

\subsubsection{Goodness-of-fit}

Here, we proceed to a pure statistical analysis of our results, i.e., we investigate the properties of our two models (with and without an additive term) in terms of goodness-of-fit. Our idea is to assess what loss of explanatory power is implied by the omission of the additive term. We distinguish between ``Air'' and ``Vessel'', in order to see if and how the quality-of-fit varies across transports modes. We start with a simple visual inspection of Figure \ref{fig:good_fit}, where are reported the yearly median \emph{ad valorem} estimates of the iceberg costs ($\widehat{\tau}_{i}$) estimated in equation \ref{eq:estimatedequation}\footnote{Qualitatively similar results are obtained when the mean values are reported instead. More details of those results available upon request to the authors.}. The dotted line plots the estimated iceberg cost ($\widehat{\tau}_{i}$) when the additive term $t$ is restricted to 0, and the plain dark line represents the iceberg cost in the model including a non-null additive term. In both cases, multiplicative costs are reported minus one, for a better reading comfort and comparability with Figure \ref{fig:mult_vs_add}.

For both transport modes, it appears that the omission of the additive term seriously biases the multiplicative term upward. For Air, the size of the bias seems to decrease over time, while it seems pretty stationary for Vessel. In any case, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2. On average over the whole period, biased estimate for iceberg is 1.04 for Air and 1.05 for Vessel, while unbiased estimates, are, respectively 1.02 and 1.03.

In order to deliver a more systematic diagnosis, we use several standard measures of fit. Of course R$^{2}$ comes naturally to mind, but its use is far from being straightforward when evaluating non-linear estimates.\footnote{R-squared is based on the underlying assumption that the adjusted model is a linear one. In a non-linear context, R-squared is therefore inappropriate, strictly speaking. However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.)}. Therefore, we provide as a complement the Standard Error of Regression (SER) which represents the average distance that the observed values fall from the regression line. Smaller values are better because it indicates that the observations are closer to the fitted line. We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test. A decrease in the log-likelihood function points to a better quality-of-fit. However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic, so that AIC stat $= 2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $. Once again, the preferred model is the one with the minimum AIC value. Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, see equation~\ref{eq:estimatedequation}) and the Restricted one (\emph{R}, i.e. equation~\ref{eq:estimatedequation} with $t=0$). The null tested is that the two models are statistically equivalent. Results are reported in Tables~\ref{tab:good_fit_air} and \ref{tab:good_fit_ves}.

%If we keep the R² as a measure of GoF, we will have to provide some evidence of normality for residuals. In addition, a plot of observations and the adjustment curve might be useful.
Unsurprisingly, it appears that the inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or transport mode. On average over the whole period, $R^{2}$ doubles when restricting to Air Transport, and increases by 50\% for Vessel. Similar qualitative conclusions arise from the comparisons of SERs, but may be more reliable on the quantitative side, since the use of R-squared on non-linear estimates may be qualified. Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes. Both AIC and Log-Likelihood statistics decrease with the inclusion of the additive term, and the LL test unambiguously reject the null of statistical equivalence of the two models. This is true whatever the considered year.

However, going into the examination of the levels of these goodness-of-fit statistics reveals intriguing differences. Multiplicative costs appear more important for the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector; the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air. Besides, the dynamics over time also reveals intriguing differences. Quality of fit decreases strongly after 2000, for both transport modes. For Air, it seems that the additive term explains less and less variance over time: in 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the SER by 14 percentage points ;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER. The picture is quite different for Vessel. For that transport mode, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years. It appears that the deteriorating performance of the model in terms of goodness-of-fit from 2000 comes mainly from the decreasing performance of the iceberg term. In order to enlighten the roots of these differences, the following section elaborates further on the quantitative dynamics of each type of cost.

For comparison purposes, we provide in the appendix similar results for some years based on a more disaggregated dataset at the product level (4-digits).\footnote{Since we use a non-linear estimator, estimations are highly computer-intensive. This explains why we decided to focus on the 3-digit product classification, which requires less time-consuming calculations.}. If anything the quality of fitting appears slightly higher when estimations are based on the 4-digits classification. This is especially true for the model restricting trade cost to their iceberg dimension, whatever the transport mode considered. When the additive part is taken into account however, the difference becomes very small, whatever the considered criterion.

Using a more disaggregated classification unsurprisingly adds some statistical precision, but not to an extent which disqualifies the use of slightly more aggregated data.
% need for SER on all years to be complete.

\subsubsection{Interpretation}

We now more specifically investigate the long-run behavior of the estimated iceberg and additive parameters. Tables\label{tab:result_air} and \label{tab:result_ves} report, respectively for the Air and Vessel sector, our estimates for Iceberg trade costs, first from equation \label{eq:iceberg} (upper panel), then from equation \label{eq:estimatedequation} where they are estimated jointly with an additive term, which estimate is also reported (bottom panel). Remember that estimation is performed on a yearly basis, consequently, for the sake of clarity, we do not report estimates for each year, but for the beginning and the end of our sample, as well as starting year for each decade. To complement these tables, Figure\label{fig:mult_vs_add} reports the estimates of equation \label{eq:estimatedequation}, this time for each year: the dotted line plots estimated for the additive term, and the plain, dark line plots the iceberg cost (-1, for a better reading comfort). As we did for the quality-of-fit, we report in Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} in the Appendix the estimated for coefficients for products disaggregated at the 4-digits level. In any case, these estimates are very close, and do not show any major difference with the baseline ones.

Unsurprisingly, the trend in transportation costs is clearly decreasing whatever the mode (air or vessel) or measure (with or without additive term) we consider. Starting from 1974, the decrease appears more important for sea than air transportation: transportation costs are divided by three for the first mode, and two for the second one. That said, it is widely know the two oil shocks inflated artificially transportation costs that specific year. If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for Air, and 2-3 percentage points for Vessel - depending if we consider mean or median estimates. These figures are consistent with the ones by Hummels (2007). However, our methodology allows distinguishing in this trend, what comes from multiplicative and additive costs. To make a long story short, it appears that the additive costs represent the bulk of the decrease for Air, whereas the multiplicative component is the main responsible of the decreasing trend for Vessel, except for the years 2012-2103, witnessing an abrupt decrease in additive costs. To give a few average numbers, starting from 1980 and ending in 2013, the iceberg term decreases by 0.6 percentage point (pp), and the additive term is reduced by 1 pp for Air; for Ocean shipping, the reduction is roughly 2 pp for additive costs, and 1 pp for multiplicative costs.

%Itrrazbal

%Composition effects ??

In section \ref{sec:determinants}, we provide rationales for those differences.

\section{Reassessing the Traditional Determinants of Transport Costs}\label{sec:determinants}

\subsection{Methodological Concerns}

%In step 2, we ask if the traditional determinants of transportation costs impact differently the multiplicative and additive costs estimated in the first step.
%Therefore, we consistently focus only on origin-country-specific transportation costs, taking the estimated $\tau_{i}$ and $t_{i}$ from step 1 as dependent variables. In line with the trade literature, we proxy transport costs by (log of) distance and other proxies for trade costs (mainly, geographic and cultural factors).\interfootnotelinepenalty=10000\footnote{All those variables come from the well-known GeoDist database built by the CEPII: \url{http://www.cepii.fr/CEPII/fr/bdd_modele/presentation.asp?id=6}.} More formally, we estimate, still on a yearly basis and for each transportation mode:

%\begin{eqnarray}
%ln(\widehat{\tau}_{i}) &=& \beta_{1} \textrm{ln(distance)} + \Gamma_{1} \textrm{Gravity} + \mu_{i} \label{eq:step2_iceberg}\\
%ln(\widehat{t}_{i}) &=& \beta_{2} \textrm{ln(distance)} + \Gamma_{2} \textrm{Gravity} + \nu_{i} \label{eq:step2_additive}
%\end{eqnarray}

%where the ln operator denotes that variables have been log-linearized, $distance$ is the bilateral distance between US and the origin country $i$  and $Gravity$ is a vector of controls including the languages spoken in the country under different definitions, whether the country is landlocked, and their colonial links.

In this second stage of the paper, we thus keep on questioning the way the literature models the trade costs determinants. Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs? Or total? Usually, people use a measure of distance as a way to capture iceberg trade costs. What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs? Using our decomposition of transport costs obtained in the first stage, we can assess the contributions of typical gravity variables to each component of transport costs (additive / iceberg). To put this simply, consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and find that this fits reasonably well, the estimated coefficient stands in the average. We then apply the same regression on each multiplicative / additive component of trade costs. Suppose that we find a high correlation with the additive cost, and almost 0 with the multiplicative cost. This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error. \bigskip

\paragraph{Data treatment} Before presenting the estimated equation, let us mention how we deal with the data. We regress 1°) the ``overall trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{ik}^{nlI}$ hereafter ($i$ for the origin country, $k$ for the product and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{ik}$ and $t_{ik}$ respectively, on some common determinants. Precisely, we re-treat the iceberg component (both $\tau_{ik}^{nlI}$ and $\tau_{ik}$) to have it expressed in percentage of the fob price, ie we build the variable $\tau_{ik} =  100(\widehat{\tau}_{ik}-1)$ (with $\widehat{\tau}_{ik}$ the estimated variable obtained at stage 1). As for the additive cost, from stage 1 of the paper, we have obtained the value of (country-specific) trade cost as a fraction of the fob price (denoted $\widehat{t}_{ik}$); we multiply it to have it expressed as a percentage of the fob price (considering $t_{ik} =  100 \widehat{t}_{ik}$). Further, and in line with our first-step estimation method, we estimate the determinants of trade costs on a yearly basis, and conditional to the transport mode (air or vessel).

In the three cases, we want to explain trade costs expressed as a percentage of the fob price. All our three measures of trade costs are bounded by 0 as minimal value, and we drop outliers by only keeping the 95 first percentile of each distribution.
\bigskip

\paragraph{Functional form} To derive the estimated equation, we start recalling what is behind the gap between cif-fob prices. In the US import data we use, the cif price is based on applying import charges on the custom value. What is defined as import charges? As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S. import duties) incurred in bringing the merchandise from alongside the carrier''. So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''. This means that the seller has paid to bring the goods to the port, but not to put them on the boat. In this respect, all hauling charges are excluded from what we call the ``fob'' price. In comparison with the fas price, the cif price is ``cost, insurance, freight'' included.\footnote{From OECD stats, the c.i.f. price (i.e. cost, insurance and freight price) is the price of a good delivered at the frontier of the importing country, including any insurance and freight charges incurred to that point, or the price of a service delivered to a resident, before the payment of any import duties or other taxes on imports or trade and transport margins within the country.}

Accordingly, if we denote $TC_{ik}$ the total transport cost paid to import the product $k$ from country $i$ (on a given year, conditional to a certain transport mode), expressed in USD, we can decompose it in three elements:

$$TC_{ik} = \text{Freight + Insurance + Handling costs}$$

Our second-stage estimation is then guided by two questions. For each of these three dimensions (cost, insurance, freight) at the root of the cif-fas prices gap, 1°) what are its determinants, and 2°) do we expect this dimension to play rather on the additive or the multiplicative dimension?

\paragraph{Insurance}: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ??? for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood). The database provides us with the quantity (in terms of kilograms) of the shipment ($q_{ik}$), as well as the price, so that we have the value of the shipment. However, we have no dedicated variable to measure the insurance cost in the database. Accordingly, we capture this dimension by including product fixed effects. Insurance grids retain a degree of product categorization much broader than the 3-digit level we retain as benchmark sectorial degree (our $k$ index for the product dimension). In most pieces of information we have collected, insurance grids adopt a ranking in \textbf{20 (??)} types of products. Accordingly, we include product fixed effects of dimension \textbf{20}, that we refer to with the subscript $k^\prime$, i.e. $FE_{k^\prime}$. If we denote $p_{ik}^{fob}$ and $q_{ik}$ the (fas) price and the quantity of the shipment, total insurance costs (in USD) are accordingly measured by $FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}$. As insurance cost depend on the value of the shipment, we expect it to be proportional to the price, ie rather play on the iceberg part of the transport costs.

\paragraph{Costs} (for handling costs). The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs). If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs. We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD. As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$). Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular). We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following. Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$. As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.

 \paragraph{Freight}. Quite intuitively, freight costs should depend on the distance between the origin country and the U.S. (denoted $dist_i$). Further, they likely depend on the quantity ($q_{ik}$) and the bulk of the shipment. Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs. Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$. Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.

As a result, total transport costs by $TC_{ik}$ (expressed in USD) can be decomposed in the three components of insurance, handling costs and freight, according to:

$$TC_{ik} =  \underbrace{EC_i\times V_{k}\times q_{ik}}_{\text{Costs}} +\underbrace{FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}}_{\text{Insurance}}+ \underbrace{V_k\times dist_i\times q_{ik}}_{Freight}$$

By definition, such transport costs can be written as $TC_{ik} = (p^{cif}_{ik} - p^{fas}_{ik})q_{ik}$. Accordingly, we can write:

\begin{eqnarray*}
&&(p^{cif}_{ik} - p^{fas}_{ik})q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_{k}\times dist_i\times q_{ik} \\
\Leftrightarrow && \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)p^{fas}_{ik} q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_k \times dist_i\times q_{ik}
\end{eqnarray*}

From the above equation, it comes that $\frac{TC_{ik}}{p^{fas}_{ik} q_{ik}} = \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)$, which is exactly the measure of trade costs obtained in the first stage.\footnote{Recall that our first stage starts from the equation: $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau^{nlI}_{ik}$ in the ``raw'' equation and $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau_{ik}+ \frac{t_{ik}}{p^{fas}_{ik}}$ when decomposing between iceberg and additive transport costs.}

Accordingly, we can decompose each of our three measures of trade costs ($\tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, expressed in the three cases as a percentage of the fob price, and denoted $tc_{ik}$ to gain in generality, according to:

\begin{equation}
tc_{ik} = \underbrace{\frac{EC_i\times V_{k}}{p^{fas}_{ik}}}_{\text{C}} + \underbrace{FE_{k^\prime}}_{\text{I}}+ \underbrace{\frac{V_k\times dist_i}{p_{ik}^{fas}}}_{\text{F}} \label{eq:theory}
\end{equation}

Based on Equation (\ref{eq:theory}), we specify the estimated equation as follows:

\begin{equation}
tc_{ik} = \left[\frac{EC_i\times V_k}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k\right]e^{\varepsilon_{ik}} \label{eq:estim}
\end{equation}

with $\varepsilon_{ik}$ the error term, centered on 0. This calls for two comments. First, we add the country fixed effects separately in the regression, in order to correctly identify the associated interaction term. Second, we specify the error term multiplicatively. If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points). In this case, it is very likely that the variance of the error term depends on the size of the observed costs. E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high. Adopting this functional form would then be subject to a potential bias in point estimates due to the heteroscedasticity of error terms. This drives us to specify the error term in a multiplicative manner (such that the error term is in percentage of each explicative variable and centered on 0), and to take the log of the equation.

Accordingly, the functional form that we estimate, for each of the three transport costs obtained at the first stage ($tc_{ik} = \tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, conditional on the year and the transport mode), is expressed as:

\begin{equation}
\log tc_{ik} = \log\left[\frac{EC_i\times V_{k}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k \right]+ \varepsilon_{ik} \label{eq:estim1}
\end{equation}

%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively. So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ... \bigskip

%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !

\subsection{Results and Interpretation}
%\section{One step further into interpretation: transportation costs as an indicator of market structure.}

\section{Conclusion}

\newpage
\bibliographystyle{essaien}
\bibliography{biblio}

\newpage

\appendix

\section*{Data Appendix \label{app:data}}

\subsection{Fob-cif prices}
From the US Census website:

The Customs value is the value of imports as appraised by the U.S. Customs and Border Protection in accordance with the legal requirements of the Tariff Act of 1930, as amended. This value is generally defined as the price actually paid or payable for merchandise when sold for exportation to the United States, excluding U.S. import duties, freight, insurance, and other charges incurred in bringing the merchandise to the United States. The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller. In the case of transactions between related parties, the relationship between buyer and seller should not influence the Customs value.

In those instances where assistance was furnished to a foreign manufacturer for use in producing an article which is imported into the United States, the value of the assistance is required to be included in the value reported for the merchandise. Such ``assists'' include both tangible and intangible assistance, such as machinery, tools, dies and molds, blue prints, copyrights, research and development, and engineering and consulting services. If the value of these ``assists'' is identified and separately reported, it is subtracted from the value during statistical processing. However, where it is not possible to isolate the value of ``assists'', they are included. In these cases the unit values may be increased due to the inclusion of such ``assists''.
Import Charges

The import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S. import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States. In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States) in Canada or Mexico to the first port of entry.
C.I.F. Import Value

The C.I.F. (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States. It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S. import duties.

\subsection{Product classification for Insurance fixed effects}
\textbf{to be done}

\subsection{Volumetric weight and the Maritime Transport Costs database}
The Maritime Transport Costs (MTC) database built by the OECD\footnote{See \texttt{http://stats.oecd.org/Index.aspx?datasetcode=MTC}.} contains data from 1991 to the most recent available year of bilateral maritime transport costs. Transport costs are available for 43 importing countries (including EU15 countries as a custom union) from 218 countries of origin at the detailed commodity (6 digit) level of the Harmonized System 1988.

At commodity level (6 digit HS), maritime transport costs are displayed using three measures:
\begin{itemize}
\item Transport cost: the total cost expressed in USD (insurance plus freight) of transporting all of the given product during the given year,
\item  Unit transport cost: transport cost per kilogramme, or in other words, the cost in USD required to transport one kilogramme of merchandise, and
\item Ad valorem equivalent: transport cost divided by the  total import value, i.e. the share transport cost represents in the total import value of the product.
\end{itemize}
At chapter level (2 digit HS), the transport costs are the sum of the corresponding 6 digits commodities. The unit cost and ad valorem cost are then the transport costs at chapter level divided respectively by sum of the weights at 6 digits, and by the sum of the import values at 6 digits.

\subsection{Country-specific handling costs and the ``Cost to export'' variable}

Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport. The most recent round of data collection for the project was completed in June 2014. Among the four variables provided, we retain the variable ``Cost to export'' (US\$ per container), as it represents the ``cost associated with all procedures required to export goods[...,] including the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' We use the variable observed \textbf{in year ??? attention la variable est-elle destination spefici? A priori non, a verifier.}
\newpage

\section*{Tables and Figures}

%%%Figures
% Goodness-of-fit

\begin{figure}[htbp]
\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
\end{tabular}
\end{center}
\end{figure}

%Economic interpretation
%usepackage{float} + H is great to keep images on the right page
\begin{figure}[H]
\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
\end{tabular}
\end{center}
\end{figure}

%%%%Tables
%Goodness-of-fit
\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit}
\begin{center}
    \begin{tabular}{lccccccc}
  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}}\\
\hline
\textbf{R$^{2}$} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.31} & \multicolumn{1}{c}{\textbf{0.31}} \\
Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{\textbf{0.60}} \\
\textbf{SER} &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & \multicolumn{1}{c}{\textbf{0.85}} \\
Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & \multicolumn{1}{c}{\textbf{0.73}} \\
\textbf{Log-likelihood} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & -17530.49 & -20253.49 & -29977.79 & -43341.27 & \multicolumn{1}{c}{-50746.83} & \multicolumn{1}{c}{-52690.29} & \multicolumn{1}{c}{\textbf{-34888.64}} \\
Terms A \& I & -15125.65 & -17263.20 & -25393.46 & -36788.44 & \multicolumn{1}{c}{-47277.53} & \multicolumn{1}{c}{-49419.70} & \multicolumn{1}{c}{\textbf{-30508.29}} \\
\textbf{AIC criteria} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{106130.58} & \multicolumn{1}{c}{\textbf{70498.08}} \\
Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{100155.41} & \multicolumn{1}{c}{\textbf{62284.99}} \\
\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       &  \\
2$\times$(ll(UR) -ll(R)) & 4809.68 & 5980.59 & 9168.67 & 13105.67 & \multicolumn{1}{c}{6938.61} & \multicolumn{1}{c}{6541.17} & \multicolumn{1}{c}{\textbf{8760.69}} \\
\# restrictions  & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426.00} & \multicolumn{1}{c}{422} & \multicolumn{1}{c}{\textbf{401.93}} \\
p-value & 0.000 & 0.000 & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{\textbf{}} \\

\hline\hline
   % \bottomrule

    \end{tabular}%
    \end{center}
  \label{tab:good_fit_air}%
  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Vessel, Measures of Goodness-of-fit}
\begin{center}
    \begin{tabular}{lccccccc}
  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\
          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \textbf{Mean stat} \\
\hline
\textbf{R$^{2}$}\unboldmath{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &  \\
Term I only & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{0.34} & \multicolumn{1}{c}{\textbf{0.39}} \\
Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.56}} \\
\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.76} & \multicolumn{1}{c}{\textbf{0.66}} \\
Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{\textbf{0.57}} \\
\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & \multicolumn{1}{c}{-16287.40} & \multicolumn{1}{c}{-16129.13} & \multicolumn{1}{c}{-25169.31} & \multicolumn{1}{c}{-35263.95} & \multicolumn{1}{c}{-41998.95} & \multicolumn{1}{c}{-43692.93} & \multicolumn{1}{c}{\textbf{-28534.30}} \\
Terms A\& I & \multicolumn{1}{c}{-12985.76} & \multicolumn{1}{c}{-13353.65} & \multicolumn{1}{c}{-21171.37} & \multicolumn{1}{c}{-29490.96} & \multicolumn{1}{c}{-37418.66} & \multicolumn{1}{c}{-39751.86} & \multicolumn{1}{c}{\textbf{-24151.31}} \\
\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & \multicolumn{1}{c}{33328.81} & \multicolumn{1}{c}{33010.27} & \multicolumn{1}{c}{51142.62} & \multicolumn{1}{c}{71365.89} & \multicolumn{1}{c}{84789.89} & \multicolumn{1}{c}{88191.87} & \multicolumn{1}{c}{\textbf{57848.60}} \\
Terms A \& I & \multicolumn{1}{c}{27331.52} & \multicolumn{1}{c}{28067.31} & \multicolumn{1}{c}{43664.74} & \multicolumn{1}{c}{60475.91} & \multicolumn{1}{c}{76161.33} & \multicolumn{1}{c}{80873.72} & \multicolumn{1}{c}{\textbf{49683.32}} \\
\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{\textbf{}} \\
2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & \multicolumn{1}{c}{9160.56} & \multicolumn{1}{c}{7882.15} & \multicolumn{1}{c}{\textbf{8765.97}} \\
\# restrictions  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{427} & \multicolumn{1}{c}{\textbf{417.05}} \\
p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\

\hline\hline
   % \bottomrule

    \end{tabular}%
    \end{center}
  \label{tab:good_fit_ves}%
  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of Regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

%Economic Interpretation

\begin{table}[htbp]
  \centering
  \caption{Air: Dynamics of Trade Costs over Time}
\begin{center}
    \begin{tabular}{lccccccc}

  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
 \hline
\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034 & \multicolumn{1}{c}{\textbf{1.051}} \\
Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029 & \multicolumn{1}{c}{\textbf{1.042}} \\
Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
\hline
\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010 & \multicolumn{1}{c}{\textbf{0.019}} \\
\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005 & \multicolumn{1}{c}{\textbf{0.007}} \\
\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017 & \multicolumn{1}{c}{\textbf{1.025}} \\
\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 & \multicolumn{1}{c}{\textbf{1.018}} \\
Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
\hline
\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\

\hline\hline
    \end{tabular}%
  \end{center}
  \label{tab:result_air}%
\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Vessel: Dynamics of Trade Costs over Time}
\begin{center}
    \begin{tabular}{lccccccc}

  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
 \hline
   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036 & \textbf{1.058} \\
Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033 & \textbf{1.051} \\
Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
\hline
\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015 & \textbf{0.029} \\
Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 & \textbf{0.019} \\
Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 & \textbf{1.032} \\
Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018 & \textbf{1.028} \\
Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
\hline
\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\

\hline\hline
    \end{tabular}%
  \end{center}
  \label{tab:result_ves}%
\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
\end{table}%

\newpage

\appendix

\section{Results with a Higher Level of Products Disaggregation }
%\vspace{-0.5cm}
\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit, 4-digits}
\begin{center}
\label{tab:good_fit_air_rob}%
%\vspace{0.5cm}
\scalebox{0.97}{
\begin{tabular}{lccccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}                      &  \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\
\hline
\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} & \textbf{0.47} \\
Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} & \textbf{0.63} \\
\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93} & \textbf{0.89} \\
Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86} & \textbf{0.79} \\
\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87} & \textbf{-34744.40} \\
Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14} & \textbf{-30243.95} \\
\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} & \textbf{70940.07} \\
Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28} & \textbf{62955.73} \\
\textbf{Test LL} &       &       &       &       &       &       & \textbf{} \\
2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45} & \textbf{9000.89} \\
\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818} & \textbf{755.73} \\
p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \textbf{} \\
\hline\hline
\end{tabular}}%
\end{center}
 \vspace{-10pt} \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

\begin{table}[H]
  \centering
  \caption{Vessel: Measures of Goodness-of-fit, 4-digits}
\begin{center}
\label{tab:good_fit_ves_rob}%

\scalebox{0.97}{\begin{tabular}{lccccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}                      &  \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\ \hline
\boldmath{}\textbf{R$^{2}$}\unboldmath{} &       &       &       &       &       &       &  \\
Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{\textbf{0.44}} \\
Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.59}} \\
\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} & \multicolumn{1}{c}{\textbf{0.77}} \\
Terms A \& I &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} & \multicolumn{1}{c}{\textbf{0.68}} \\
\textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} & \multicolumn{1}{c}{\textbf{-29883.62}} \\
Terms A\& I & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22} & \multicolumn{1}{c}{\textbf{-25303.92}} \\
\textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23} & \multicolumn{1}{c}{\textbf{61425.60}} \\
Terms A \& I & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} & \multicolumn{1}{c}{\textbf{53573.29}} \\
\textbf{Test LL} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.75 & \multicolumn{1}{c}{17354.65} & \multicolumn{1}{c}{20113.52} & \multicolumn{1}{c}{16608.16} & \multicolumn{1}{c}{12589.59} & \multicolumn{1}{c}{15704.63} \\
\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} & \multicolumn{1}{c}{860} \\
p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\

\hline\hline
\end{tabular}}
\end{center}
  \vspace{-10pt}\tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Air: Dynamics of Trade Costs over Time}
\begin{center}
    \begin{tabular}{lccccccc}

  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\

\hline
\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       &  \\
Mean  & 1.066 & 1.058 & 1.052 & \multicolumn{1}{c}{1.033} & \multicolumn{1}{c}{1.037} & \multicolumn{1}{c}{1.032} & \multicolumn{1}{c}{\textbf{1.049}} \\
Median & 1.052 & 1.044 & 1.041 & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.027} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{\textbf{1.037}} \\
Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{\textbf{0.045}} \\
\hline
\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{} \\
\textit{Additive term} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} & \multicolumn{1}{c}{\textbf{0.019}} \\
Median & 0.012 & 0.006 & 0.006 & \multicolumn{1}{c}{0.005} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{\textbf{0.008}} \\
Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{\textbf{0.034}} \\
\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016} & \multicolumn{1}{c}{\textbf{1.024}} \\
Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014} & \multicolumn{1}{c}{\textbf{1.016}} \\
Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{\textbf{0.026}} \\
\hline
\# obs & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460} & \multicolumn{1}{c}{\textbf{}} \\

\hline\hline
    \end{tabular}%
  \end{center}
  \label{tab:result_air_rob}%
\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
\end{table}%

\begin{table}[htbp]
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
  \centering
\caption{Vessel: Dynamics of Trade Costs over Time}
\begin{center}
    \begin{tabular}{lccccccc}
  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\
& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
\hline
\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.098 & 1.061 & \multicolumn{1}{c}{1.058} & \multicolumn{1}{c}{1.051} & \multicolumn{1}{c}{1.042} & \multicolumn{1}{c}{1.036} & \textbf{1.060} \\
Median & 1.094 & 1.051 & \multicolumn{1}{c}{1.048} & \multicolumn{1}{c}{1.045} & \multicolumn{1}{c}{1.038} & \multicolumn{1}{c}{1.031} & \textbf{1.052} \\
Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} & \textbf{0.036} \\
\hline
\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{}\\
\textit{Additive term} &       &       &       &       &       &       & \textbf{} \\
Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015} & \textbf{0.028} \\
Median & 0.029 & 0.013 & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{0.013} & \multicolumn{1}{c}{0.008} & \textbf{0.017} \\
Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} & \textbf{0.039} \\
\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.054 & 1.034 & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.021} & \textbf{1.033} \\
Median & 1.049 & 1.030 & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.025} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{1.018} & \textbf{1.028} \\
Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013} & \textbf{0.025} \\
\# obs & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} & \textbf{}\\
\hline\hline
    \end{tabular}%
  \end{center}
  \label{tab:result_ves_rob}%
\vspace{-10pt}\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
\end{table}%

\end{document}

### 2016-06-13 redaction/dhp_draft_new.tex

@@
 %We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
-\author{Guillaume \textsc{Daud}\thanks{%
+\author{Guillaume \textsc{Daudin}\thanks{%
 Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa), Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa)} } \maketitle

### 2016-07-20 redaction/dhp_draft_july2016.tex
Rédaction dans le github

--- 2016-06-13 redaction/dhp_draft_new.tex
@@
 
-\date{\today}
 \title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX.
@@
 %We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
-\author{Guillaume \textsc{Daud}\thanks{%
-Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa), Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa)} } \maketitle
+\author{Guillaume \textsc{Daudin}\thanks{%
+Universit\'{e} Paris Dauphine, LEDa \& PSL, Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine, LEDa \& PSL} }
+ \maketitle
 
+\date{July 2016}
 \bigskip
@@
 \begin{abstract}
-
-\begin{spacing}{1.2}
-
-{\normalsize XXXX}
-
-{\normalsize \bigskip }
-
-\end{spacing}
+Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad valorem tax equivalent.
+However, many common empirical facts support the existence of additive costs.
+This paper provides results on the size and composition of transport costs over time, explicitly distinguishing between its multiplicative and additive components, using sectorial cif-fob prices over 1974-2013 taken from US imports data.
+First, we estimate the two components of transport costs, by transport mode (air or ocean).
+We thus find that ocean transport costs decompose in an ad-valorem tax equivalent of mean value 3.22\% and a per-unit cost that amounts to 2.85\% of the mean fob price over the period.
+These values are respectively equal to 2.5\% and 1.9\% in air shipping.
+Second, we provide a quantitative assessment what standard international trade models lose by skipping additive transport costs.
+Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
+On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
+Third, the time dimension of our data allows us to characterize the evolution of transport costs.
+After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012.
+While this order of magnitude applies for both the additive and multiplicative components of vessel transport costs, air transport costs exhibit more contrasted trends.
 \end{abstract}
@@
 \noindent Trade costs remain central in international economic analysis.
-Defined as the costs associated with the exchange of goods across national borders, they are a major obstacle to international economic integration.
-In addition, they might be the key to the explanation of major puzzles in international macroeconomics (\citealp{rogoff}).
-Historical perspective show that indeed, trade costs variations are a major determinant of the level of trade flows.
-According to the estimates by \citet{Jacks08}, trade cost declines explain around 55 percent of the pre World War I trade boom and 33 percent of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter war trade collapse.
-After 1950, average trade costs fell by 16\%.
-Based on panel data, \citet{novy13} finds  that U.S.
-trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-
-Trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff-costs), time costs (time to ship goods) and transport costs \emph{per se}.
+Defined as the costs associated with the exchange of goods across national borders, they are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
 Many believe that the latter have dramatically decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
@@
 However, \citet{hummels2007} shows that the bulk of price declines in transportation comes from air shipping, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004; concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, even if the rise in containerization lowered shipping costs from 3 to 13\%.
-Studies overviewed by \citet{Behar_Venables} are consistent with those results: the fall in measured transport costs has been relatively small.
-The importance of fuel costs and technical progress (which improved speed and reliability rather than decreased costs) in transport explain that result.
+Studies overviewed by \citet{Behar_Venables} are consistent with those results: The fall in measured transport costs has been relatively small.
+According to these authors, this is attributable to the growing importance of fuel costs and technical progress, which improved speed and reliability rather than decreased costs.\smallskip
 
-Despite the decrease in distance-related costs\footnote{\citet{Behar_Venables} suggest freight costs account for two thirds of the effect of distance, the remaining third ebmoyding more or less the effect of time}, several papers (mostly based on empirical estimates of the gravity equation) in the last decade have shown that distance remained a major obstacle to trade (e.g.
+Trade costs are considered as a major obstacle to international economic integration and international trade flows.
+According to the estimates by \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
+After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements.
+Based on panel data, \citet{novy13} thus finds  that U.S.
+trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
+Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs (typically captured through distance) still remain a major obstacle to trade (e.g.
 \citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
-\citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs (1.7 $=$ 1.55 $\times$ 1.74 - 1).
-Again, the latter comprises 44\% border-related trade barriers and 21\% transportation costs.
-That means that transportation costs are higher than each component of border costs taken separately (namely, 8\% for tariff and non-tariff costs, 7\% for language differences, 14\% for for currency differences and 9\% for other costs).\footnote{1.44 $=$ 1.08 $\times$ 1.07 $\times$ 1.14 $\times$  1.09)}.
-Therefore, the share of transport costs in the consumer price of manufactured goods remains high (\citealp{Lafourcade_Thisse}).
-Besides, \citet{combes_lafourcade} emphasize that the figures provided by \citet{anderson_wincoop_jel} should be considered as a a downward benchmark, and could be higher in developing economies.
-Despite technical progress and infrastructure development, transport costs remain large and deserve attention.
-In that respect, \citet{Behar_Venables} support that the elasticity of trade with respect to freight costs is around -3, which remains a large number.
+\citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs.
+Within this latter dimension, 44\% are border-related trade barriers and 21\% transportation costs.
+\citet{Lafourcade_Thisse} also find that the share of transport costs in the consumer price of manufactured goods remains high, while \citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
+If much of trade-policy barriers have been removed over the second half of the 20th century, these findings suggest that the transport costs component of the overall trade costs remain large and deserve attention.
+This is accordingly the focus of the paper.\smallskip
 
-Following \citet{samuelson1954}, standard models of international trade have usually relied on the iceberg cost hypothesis, implicitly reducing trade costs to their transport part, and modelling the latter as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded).
-This implies that more expensive goods are costlier to trade.
-However, many common empirical facts support the existence of additive costs: pricing structure in shipping, additive tariffs, distribution costs...
-often exhibit (at least partly) an additive structure (see \citealp{Irrazabal_2015}).
-The additive versus iceberg trade costs is not a new debate in the international trade literature.
-The famous Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: the demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
-\citet{martin2012} gives a strong empirical support to this conjecture: based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite.
-\citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: the elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption.
+Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, the ``iceberg cost'' hypothesis).
+However, many common empirical facts support the existence of additive costs.
+As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
+often exhibit (at least partly) an additive structure.
+The structure (additive vs multiplicative) of transport costs is far from being anecdotal, as the literature has long pointed out its role in shaping the pattern of trade flows.
+The Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
+\citet{martin2012} gives a strong empirical support to this conjecture: Based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite.
+\citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: The elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption.
 Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
-Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level.
+\textbf{je ne sais pas quoi faire de la reference a lashkaripour.
+Complique d'en parler, car on ne peut pas le contredire car on n'a pas les donnees qui le permettent.
+Ne pas en parler?
+C'est un peu limite.} The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
+Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e.
+more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments.
+
+%Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level.
 Based on US imports data, he finds that a 10\% increase in the price of an item increases the shipping cost by 9.5\%.
@@
 
-A correct measure of transport costs is also important for understanding several key features of international trade.
-For example, the existence of additive costs may explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
-Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e.
-more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments.
-Besides, several recent papers show that the existence of additive trade costs has important non trivial effects when assessing the welfare effects of trade liberalization.
-\citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive (per-unit) trade costs, in addition to iceberg/multiplicative costs.
+Beyond the positive aspect of understanding trade patterns, several recent papers also point out the normative implications of additive trade costs.
+\citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the iceberg component.
 A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs.
-The difference arises from dissimilar adjustments at the extensive margin - fewer firms enter into the export market following a reduction in additive costs.
 Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff.
+While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much can be done in quantifying such gains though, by lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap.
+\bigskip
 
-Our purpose in this paper is to provide results on the size of transportation costs over time, explicitly distinguishing between multiplicative and additive parts.
-Our paper stands at the crossroad of these two strands of the literature, the first one dealing with the dynamics of transport, and more generally trade costs over time, the second one focusing on the measurement and nature of the latter.
-More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over a large time dimension - several decades.
-To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}.
-Therefore, our period of study runs from 1974 to 2013.
-Closely related to this paper is the work by \citet{Irrazabal_2015}, who develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level data, their results suggest that additive costs are on average 14 percent, relatively to the median price, and that they are strongly correlated with standard proxies for trade costs (like e.
+Our paper contributes to the literature by providing results on the size of transport costs over time, explicitly distinguishing between multiplicative and additive parts.
+More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over several decades.
+To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013.
+Closely related to this paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
+Based on Norwegian firm-level data, their results suggest that additive costs are on average 30-45\% relative to the median consumer price, and that they are strongly correlated with standard proxies for trade costs (like e.
 g., distance).
-However, while our data requires that we focus only on transportation costs, our approach departs from theirs in several key aspects.
+However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects.
 
-First, our theoretically agnostic approach aims at letting the data speak, and provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models; to our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the shares of multiplicative and additive costs in total transport costs.
+First, our theoretically agnostic approach provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models.
+We thus obtain that the mean values over 1974-2012 of iceberg costs are equal to 2.5\% in air and 3.2\% in ocean shipping, whereas the additive component amounts to 1.9\% and 2.85\% of the fob price in air and ocean shipping respectively.
+To our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the magnitude of both multiplicative and additive costs in total transport costs.
 Second, we provide, through standard measures of ``goodness-of-fit'', an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Third, the time dimension of our data allows us to characterize the evolution of transportation costs, and more precisely the dynamics of the relative share of each specific (multiplicative and additive) cost over a forty years time span; to our knowledge, no firm-level database is able to provide such a time coverage.
-Fourth, we are also able to discriminate our results by transport mode (air and vessel).
+Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
+On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
+Third, the time dimension of our data allows us to characterize the evolution of transport costs, by transport mode (air or vessel) and for each specific (multiplicative and additive) cost over a forty years time span.
+After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012.
+Decomposing by transport mode and by type of transport costs allows to go deeper into this picture.
+First, as in \cite{hummels2007}, we find that the transport costs decrease is more pronounced in air than in ocean shipping, starting in 1984.
+However, while ocean shipping costs display a regular decreasing pattern until 2012, air transport costs rather get stabilized over 2005-2012.
+Second, while both the additive and multiplicative components of ocean transport costs have roughly decreased by 40\% over the period, air transport costs exhibit more contrasted trends.
+If the reduction of air shipping costs can most be attributed to the decrease in its multiplicative component over 1974-2005, the opposite scheme prevails afterwards, the reduction in per-unit air transport costs being compensated by an increase in the ad-valorem component.
+\smallskip
 
-Finally, our analysis is also indirectly related to a large strand of the international trade literature assessing the impact of trade costs on international exchanges.
-Following \citet{anderson_wincoop} and \citet{anderson_wincoop_jel}, there has been a huge bunch of articles relying on gravity equations to infer indirectly trade costs as tariff equivalents, or losses related to various kinds of trade barriers - many of these works are surveyed in \citet{anderson_wincoop}; among the various subsequent works, see in particular \citet{anderson_yotov}, \citet{novy13} or \citealp{chen_novy}).
-However, our purpose in the present paper is different.
-We first infer directly transport costs from price data, discriminating between multiplicative and additive parts.
-In a second step, following an approach similar to \citet{Behar_Venables}, we relate the previous measures to traditional determinants of transport costs (like e.
-g., distance), in order to understand how differently the latter impact the multiplicative and additive transport costs.
+Section \ref{sec:data_method} explains the data sources and the empirical methodology retained in the paper.
+Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results.
+In Section \ref{sec:results_decomposition}, we characterize the role of the additive component of transport costs.
+After reporting the mean values over the period (by transport mode), we show the improved performance in including additive trade costs in the measure of transport costs.
+This being established, Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2012.
+Section \ref{sec:conclu} concludes.
 
-%time as a trade barrier, on risque d'avoir des problèmes avec ça... \citet{Lafourcade_Thisse},
-
-XXX Here a summary of our results and contribution XXX
-
-The next section presents the data and our fundamental equation.
-Section 3 reports our baseline estimates for multiplicative and additive measures of transport costs across time, as well as a number of robustness checks.
-Section 4 discusses the results of estimates relating traditional determinants of transport costs to the previous measures, while Section 5 XXX has yet to be discussed!!!
-XXX.
-The last section concludes.
-
-%\noindent More generally,  Among the closest to our own work, Eaton et al.
-(2002) develop a Ricardian general equilibrium trade model to estimate parameters relating to absolute advantage, to comparative advantage, and to geographic barriers.
-Anderson and Wincoop (2003) elaborate a theory-founded gravity equation to assess the cost of trade frictions, and use it to provide a direct measure of the trade losses due to national borders (between 20 and 50 percent).
-
-\section{Data Sources and Empirical Methodology}
+\section{Data Sources and Empirical Methodology \label{sec:data_method}}
 
@@
 
-\noindent
-
-Our strategy for representing transportation costs is relatively standard, and consists in exploiting the difference between commodity-level export and prices.
-We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, origins and transportation mode.
+As in \cite{hummels2007} (among others), our measure of transportation costs consists in exploiting the difference between commodity-level export and import prices.
+We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode.
 More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs.
@@
 
-To have the import price of the good (CIF price), next to the export price (FOB price) is highly valuable.
-Indeed, this allows us thus to estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs.
-This differentiates us from \citet{Irrazabal_2015}, that can only estimate the ratio of additive costs as a share of the tola price.
-However, note that this measure based on \emph{freight} cost ignores another dimension of transportation costs related to the time value of goods in transit.
-According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
-
-\subsection{Database}
-
-%laius sur pourquoi pas de donnes de firmes (triop comutational), ni comtrade (trop mauvais, citer RIE) XXX
-
-\noindent Even if the data of transportation costs we rely could be available at the firm-level, the use of a nonlinear estimator (see below) triggers computational limitations that do not make them a likely option.
-A product-level bilateral database of CIF-FOB ratios is theoretically possible, and could be built from the well-known UN's COMTRADE data.
-Unfortunately, \citet{hummels_rie} show that issues inherent to the construction of data (especially the use of ``mirror'' flows) make this database unreliable for our purpose.
-
-We decide therefore to use database for a single country, but with strong statistical reliability arising from a single, trustworthy customs origin.
-More precisely, the empirical analysis relies on yearly data from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
+The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
+Using this dataset has (at least) three main advantages.
+First, this dataset has been used by \cite{hummels2007}, which enables us to compare our results to his findings.
+However, we complement Hummels' (2007) findings as we extend the time coverage to 2013 (while \cite{hummels2007} stops in 2004).
+As we show in Section \ref{sec:results_trends}, extending the time period over he recent years delivers interesting insights regarding the trends in air shipping costs.
+Second, and importantly, this dataset delivers a strong statistical reliability arising from a single, trustworthy customs origin.
 Based on customs declarations, this dataset inventories all imports (both values and quantities) by origin to the United states at the HS 10-digit highly disaggregated level, with a concordance code to the SITC 5-digit coding system.
 In addition, the database reports information regarding freight expenditures and transportation mode (Ocean Vessel and Air).
-The first will be crucial to compute transport costs (see below), the second will alow us enlightening potential different dynamics of transport costs across transportation mode.
+The first will be crucial to compute transport costs (see below), the second will allow us enlightening substantial differences in the dynamics of transport costs across transportation mode.
+Third, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FOB price).
+This is highly valuable, as we can estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs.
+This differentiates us from \citet{Irrazabal_2015}, which can only estimate the ratio of additive costs as a share of the total consumer price (the only price they can encover).
 
-Due to the abovementioned computational limitations, data are reaggregated at the 3- and 4-digit level: comparing different levels of aggregation will be useful to check differences and the presence of biases precisely due to aggregation.
-Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products.
+It is also true that using this dataset has drawbacks.
+First, our measure of the cif-fob price gap only covers transportation costs by nature, thereby being mute about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
+Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit.
+According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
+\textbf{en faire un contre-argument positif}
 
-% Why the gap CIF-FOB is a good measure of trade costs (cf mimeo by Gaulier and Zignago)?
-JH: to me, This is not the right question to ask.
-We have the freight costs in our data.
-The true question is: how well these freight costs proxy correctly transportation costs as a whole.
+Even if the data of transportation costs is available at a more disaggregated level, we use sectorial price data at the 3-digit level, primarily for technical reasons.
+As detailed below, the use of a nonlinear estimator triggers computational limitations that do not make them a likely option, especially when covering a long period of time.
+Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level (for some selected years).
+Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
+Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.
 
-\subsection{Key empirical specification}
+\subsection{Empirical specification}
 
-\noindent Our purpose is to provide estimates over time of the relative shares of multiplicative and additive costs among total transport costs.
-To do so, we start from a very simple equation, quite alike the one proposed by \citet{hummels2010}:
+\paragraph{The estimated equation} Our purpose is to provide estimates over time of the sizes of multiplicative and additive costs among total transport costs.
+To do so, we start from a very simple equation (similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others):
 
-\begin{equation}\label{eq:base}
-p = \tau \widetilde{p} + t
+\begin{equation}
+p = \tau \widetilde{p} + t \label{eq:base}
 \end{equation}
 
-or as a ratio
+\noindent This equation expressed the consumer (import, or cif) price $p$ as a function of the producer (export, or fob) price $\widetilde{p}$ given both per-unit ($t$) and ad-valorem ($\tau$) transport costs.
+As usual in the literature, the so-called ``iceberg'' trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs), while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs).
+We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.
+Let us denote $i$ the origin country, and $k$, the sector (or product).
+Transforming the above equation (\ref{eq:base}) as ratio, we thus get the equation to be estimated as given by:\footnote{Also keeping in mind that the estimated equation is transport-mode and year specific (air or ocean shipping).
+We skip the year and transport-mode dimensions here to alleviate notations.}
 
-\begin{equation}\label{eq:base_bis}
-\frac{p}{\widetilde{p}} = \tau  + \frac{t}{\widetilde{p}}
+\begin{equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation}
 
-This equation relates the ratio of the consumer price (ie, cif price, denoted $p$) to the producer price (ie, fob price, denoted $\widetilde{p}$) to the two types of costs (multiplicative and additive).
-Iceberg trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs) while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs).
-%\begin{equation*}
-%p =\tau \widetilde{p}+t
-%\end{equation*}
-We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.
-Let us denote $i$ the origin country and $k$, the sector.
-Hence, the equation to be estimated is given by:
-
-\begin{equation}\label{eq:base_estimee}
-\frac{p_{ik}}{\widetilde{p}_{ik}} = \tau_{ik}+\frac{t_{ik}}{ \widetilde{p}_{ik}}
-\end{equation}
-
-\section{Decomposing the Dynamics of Transport Costs over Time}
-
-\subsection{Estimation Strategy}
-
-% nonlinear least squares: problem with heteroskedasticity???
-
-We transpose the approach by \citet{Irrazabal_2015} by considering that both multiplicative and additive costs are separable between origin country and products, in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau$ and $t$ from equation~\ref{eq:base_estimee} become:
-
-%\begin{eqnarray}
-%\tau_{ik} &=& dist^{\rho_I}F^I_i \label{eq:iceberg}\\
-%t_{ik} &=& dist^{\rho_A}F^A_i-1 \label{eq:add}
-%\end{eqnarray}
+\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1°)} both multiplicative and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2))} in a multiplicative way for the former and an additive way for the latter.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) become:
 
@@
 
-%\begin{equation}
-%\frac{p_{ik}}{\widetilde{p}_{ik}} =dist^{\rho_I}F^I_ik+\frac{dist^{\rho_A}F^A_ik-1}{ \widetilde{p}_{k}}\label{eq:estimatedequation}
-%\end{equation}
-
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} =\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
 \end{equation}
@@
 The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the \emph{cif} price $p$ cannot be lower than the \emph{fob} price: $p>\widetilde{p}_{ik}$.
-Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term.
-More precisely:
+Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term, according to:
+
+\begin{equation*}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \varepsilon_{ik}
+\end{equation*}
+
+Taking in log, this finally drives us to estimate the following equation:
 
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right)\times \varepsilon_{ik}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
 \end{equation}
-
-Log-linearization gets the following, tractable empirical equation:
-
-\begin{equation}
-ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
-\end{equation}
-where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= ln(\varepsilon_{ik})$ .
-
-Still following \citet{Irrazabal_2015}, we report a ``synthetic estimate'' of each type of transport cost, by averaging all $\widehat{\tau}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}_{ik} = t_{i} + t_{k}$ using the values of each trade flow over total yearly trade as a weighting scheme.
-Therefore, we recover for each year and transportation mode a $\widehat{\tau}$ and $\widehat{t}$ that will be the results we will report for step 1.\\
-
-\subsection{Econometric Issues}
-
-\noindent Our primary interest relates to the importance of additive costs relatively to iceberg costs.
-Put differently, what traditional models of international trade lose by ignoring additive costs?
-A natural way to answer this question is to perform estimations of equation~\ref{eq:estimatedequation} constraining $t$ to be equal to zero, and compare the fitting properties/explanatory power of the restricted and complete models.
-This is done bye computing several standard diagnostic statistics (R$^{2}$ and statistics based on the likelihood function).
+where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= \ln(\varepsilon_{ik})$.
 
@@
 In our case, due to computational limitations implied by the size of our dataset, we implement 100 iterations and set the convergence criterion for successive parameter estimates and for the residual sum of squares at 0.01.
+Finally, to eliminate the potential influence of outliers, we excluded observations in the 5 percent from the upper and lower tails of the distribution in the regression variables, and all our three measures of trade costs are bounded by 0 as minimal value.
+These cut-offs are aimed at eliminating reporting or coding errors.\medskip
 
-Finally, to control for the potential influence of outliers, we excluded observations in the 5 percent from the upper and lower tails of the distribution in the regression variables, and all our three measures of trade costs are bounded by 0 as minimal value.
-These cut-offs are aimed at eliminating reporting or coding errors.
-%(XXX develop a little bit?XXX)
+Further, one key objective of the paper is to characterize the importance of additive costs relatively to iceberg costs.
+Put differently, what traditional models of international trade lose by ignoring additive costs?
+A natural way to answer this question is to perform estimations of equation (\ref{eq:estimatedequation}) constraining $t$ to be equal to zero, and compare the fitting properties and the explanatory power of the restricted and complete models.
+This is done by computing several standard diagnostic statistics (R$^{2}$ and statistics based on the likelihood function).
+Accordingly, for each year and transport mode, we estimate two equations, depending on additive transport costs being included (Equation (\ref{eq:model_with_add}) or not (Equation (\ref{eq:model_nlI})):
 
-%\noindent with the exponents $I$ and $A$ respectively identifying the iceberg and the additive trade costs, $\rho_I$ and $\rho_A$ being the parameters of interest to be estimated.
-The adding of $-1 $ in the specification for the additive costs (\ref{eq:add}) is necessary to be consistent with $t$ starting from 0.
-In the absence of additive trade costs, ie $\rho_A=F^A_i=F^A_j=0$, the specification (\ref{eq:add}) consistently implies $t=0$.
-As well, if iceberg costs are null ($\rho_I=F^I_i=F^I_j=0$), the specification (\ref{eq:iceberg})) correctly implies $\tau=1$.
+\begin{eqnarray}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\underbrace{\tau_{i} \times \tau_{k}}_{\tau^{ice}_{ik}}-1 +\underbrace{\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}}_{t^{add}_{ik}} \right) + \epsilon_{ik} \label{eq:model_with_add} \\
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\underbrace{\tau_{i} \times \tau_{k}}_{\tau^{nlI}_{ik}}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI}
+\end{eqnarray}
 
-\subsection{Results: Distinguishing Additive and Multiplicative Parts of Trade Costs}
+After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI}), we proceed similarly to get $\widehat{\tau}^{nlI}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
+We thus recover a ``synthetic estimate'' of each type of transport cost $\widehat{\tau}$ and $\widehat{t}$, for each year and transportation mode.
+These results are reported in Section \ref{sec:results_decomposition}.
 
-\subsubsection{Goodness-of-fit}
+\section{Decomposing Transport Costs: The importance of the additive component \label{sec:results_decomposition}}
 
-Here, we proceed to a pure statistical analysis of our results, i.e., we investigate the properties of our two models (with and without an additive term) in terms of goodness-of-fit.
-Our idea is to assess what loss of explanatory power is implied by the omission of the additive term.
-We distinguish between ``Air'' and ``Vessel'', in order to see if and how the quality-of-fit varies across transports modes.
+The objective of this section is twofold.
+First, we characterize the magnitudes of transport costs over time (by transport mode), distinguishing whether the additive component $t_ik$ is included or not in the estimated equation (\ref{eq:estimatedequation}).
+Second, we quantitatively assess the importance of the per-unit component of in overall transport costs through the means of goodness-of-fit measures.
+
+\subsection{Decomposing transport costs over 1974-2013}
+
+Our first contribution to the literature is to provide estimates for the size of both the multiplicative and the additive components of transport costs.
+Tables \ref{tab:result_air} and \ref{tab:result_ves} report a summary of our results.
+Precisely, they display the mean and median values of each type of trade costs (multiplicative estimated alone, estimated along with additive costs and the additive component), expressed in percentage of the fob price, as well as the associated standard deviation, averaged over the period 1974-2013 and for selected years.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Air (3 digits): Mean values over the period and selected years}
+\begin{center}
+    \begin{tabular}{lccccccc}
+
+  %  \toprule
+   \hline\hline
+          & \multicolumn{6}{c}{Year}                      &  \\
+
+          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
+ \hline
+\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
+Mean  & 6.9 & 5.4 & 5.0 & 3.6 & 4.2 & 3.4 & \multicolumn{1}{c}{\textbf{5.1}} \\
+Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9 & \multicolumn{1}{c}{\textbf{4.2}} \\
+Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
+\hline
+\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
+\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
+\multicolumn{1}{l}{Mean } & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0 & \multicolumn{1}{c}{\textbf{1.9}} \\
+\multicolumn{1}{l}{Median} & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5 & \multicolumn{1}{c}{\textbf{0.7}} \\
+\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
+\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
+\multicolumn{1}{l}{Mean } & 3.6 & 2.3 & 2.4 & 1.7 & 2.6 & 1.7 & \multicolumn{1}{c}{\textbf{2.5}} \\
+\multicolumn{1}{l}{Median} & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 & \multicolumn{1}{c}{\textbf{1.8}} \\
+Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
+\hline
+\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\
+\hline\hline
+    \end{tabular}%
+  \end{center}
+  \label{tab:result_air}%
+\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Mean and median values of all type of transport costs are expressed in percentage of the fob price.
+}
+\end{table}%
+
+\begin{table}[htbp]
+  \centering
+  \caption{Vessel (3 digits): Mean values over the period and selected years}
+\begin{center}
+    \begin{tabular}{lccccccc}
+
+  %  \toprule
+   \hline\hline
+          & \multicolumn{6}{c}{Year}                      &  \\
+
+          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
+ \hline
+   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
+Mean  & 9.8 & 6.5 & 5.7 &5.1 & 4.0 & 3.6 & \textbf{5.8} \\
+Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3 & \textbf{5.1} \\
+Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
+\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
+\hline
+\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
+Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5 & \textbf{2.9} \\
+Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 & \textbf{1.9} \\
+Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
+\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
+Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 & \textbf{3.2} \\
+Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8 & \textbf{2.8} \\
+Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
+\hline
+\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\
+\hline\hline
+    \end{tabular}%
+  \end{center}
+  \label{tab:result_ves}%
+\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Mean and median values of all type of transport costs are expressed in percentage of the fob price.}
+\end{table}%
+
+Tables \ref{tab:result_air} and \ref{tab:result_ves} yield the following results.
+First, the magnitude of overall transport costs is sizeable, either considering the average value over the period or even in the most recent years.
+Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5?1\% for air shipping.
+\textbf{comparer avec la litterature?}
+
+\subsection{Asessing the importance of per-unit trade costs}
+In this section, we investigate the properties of our two models (with and without an additive term), to assess what loss of explanatory power is implied by the omission of the additive term.
 We start with a simple visual inspection of Figure \ref{fig:good_fit}, where are reported the yearly median \emph{ad valorem} estimates of the iceberg costs ($\widehat{\tau}_{i}$) estimated in equation \ref{eq:estimatedequation}\footnote{Qualitatively similar results are obtained when the mean values are reported instead.
 More details of those results available upon request to the authors.}.
-The dotted line plots the estimated iceberg cost ($\widehat{\tau}_{i}$) when the additive term $t$ is restricted to 0, and the plain dark line represents the iceberg cost in the model including a non-null additive term.
+The dotted line plots the estimated iceberg cost ($\widehat{\tau}^{nlI}$) when the additive term $t$ is restricted to 0 (Equation (\ref{eq:model_nlI}) being estimated), and the plain dark line represents the iceberg cost in the model including a non-null additive term (Equation (\ref{eq:model_with_add}) being estimated).
 In both cases, multiplicative costs are reported minus one, for a better reading comfort and comparability with Figure \ref{fig:mult_vs_add}.
 
-For both transport modes, it appears that the omission of the additive term seriously biases the multiplicative term upward.
+\begin{figure}[htbp]
+\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
+& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+In Figure \ref{fig:mult_vs_add}, we report the yearly median value for each component of transport cost (expressed in terms of the fob price, by transport mode) ) after running estimation of Equation (\ref{eq:model_with_add}).
+
+%Economic interpretation
+%usepackage{float} + H is great to keep images on the right page
+\begin{figure}[H]
+\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
+& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+For both transport modes, we can infer from Figure \ref{fig:good_fit} that the omission of the additive term seriously biases the multiplicative term upward.
 For Air, the size of the bias seems to decrease over time, while it seems pretty stationary for Vessel.
 In any case, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
-On average over the whole period, biased estimate for iceberg is 1.04 for Air and 1.05 for Vessel, while unbiased estimates, are, respectively 1.02 and 1.03.
 
@@
 We now more specifically investigate the long-run behavior of the estimated iceberg and additive parameters.
-Tables\label{tab:result_air} and \label{tab:result_ves} report, respectively for the Air and Vessel sector, our estimates for Iceberg trade costs, first from equation \label{eq:iceberg} (upper panel), then from equation \label{eq:estimatedequation} where they are estimated jointly with an additive term, which estimate is also reported (bottom panel).
+Tables \ref{tab:result_air} and \ref{tab:result_ves} report, respectively for the Air and Vessel sector, our estimates for Iceberg trade costs, first from equation \label{eq:iceberg} (upper panel), then from equation \ref{eq:estimatedequation} where they are estimated jointly with an additive term, which estimate is also reported (bottom panel).
 Remember that estimation is performed on a yearly basis, consequently, for the sake of clarity, we do not report estimates for each year, but for the beginning and the end of our sample, as well as starting year for each decade.
-To complement these tables, Figure\label{fig:mult_vs_add} reports the estimates of equation \label{eq:estimatedequation}, this time for each year: the dotted line plots estimated for the additive term, and the plain, dark line plots the iceberg cost (-1, for a better reading comfort).
-As we did for the quality-of-fit, we report in Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} in the Appendix the estimated for coefficients for products disaggregated at the 4-digits level.
+we report in Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} in the Appendix the estimated for coefficients for products disaggregated at the 4-digits level.
 In any case, these estimates are very close, and do not show any major difference with the baseline ones.
@@
 
-%Itrrazbal
+\section{Decomposing Transport Costs: Characterizing the trends \label{sec:results_trends}}
 
-%Composition effects ??
-
-In section \ref{sec:determinants}, we provide rationales for those differences.
-
-\section{Reassessing the Traditional Determinants of Transport Costs}\label{sec:determinants}
-
-\subsection{Methodological Concerns}
-
-%In step 2, we ask if the traditional determinants of transportation costs impact differently the multiplicative and additive costs estimated in the first step.
-%Therefore, we consistently focus only on origin-country-specific transportation costs, taking the estimated $\tau_{i}$ and $t_{i}$ from step 1 as dependent variables.
-In line with the trade literature, we proxy transport costs by (log of) distance and other proxies for trade costs (mainly, geographic and cultural factors).\interfootnotelinepenalty=10000\footnote{All those variables come from the well-known GeoDist database built by the CEPII: \url{http://www.cepii.fr/CEPII/fr/bdd_modele/presentation.asp?id=6}.} More formally, we estimate, still on a yearly basis and for each transportation mode:
-
-%\begin{eqnarray}
-%ln(\widehat{\tau}_{i}) &=& \beta_{1} \textrm{ln(distance)} + \Gamma_{1} \textrm{Gravity} + \mu_{i} \label{eq:step2_iceberg}\\
-%ln(\widehat{t}_{i}) &=& \beta_{2} \textrm{ln(distance)} + \Gamma_{2} \textrm{Gravity} + \nu_{i} \label{eq:step2_additive}
-%\end{eqnarray}
-
-%where the ln operator denotes that variables have been log-linearized, $distance$ is the bilateral distance between US and the origin country $i$  and $Gravity$ is a vector of controls including the languages spoken in the country under different definitions, whether the country is landlocked, and their colonial links.
-
-In this second stage of the paper, we thus keep on questioning the way the literature models the trade costs determinants.
-Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs?
-Or total?
-Usually, people use a measure of distance as a way to capture iceberg trade costs.
-What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs?
-Using our decomposition of transport costs obtained in the first stage, we can assess the contributions of typical gravity variables to each component of transport costs (additive / iceberg).
-To put this simply, consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and find that this fits reasonably well, the estimated coefficient stands in the average.
-We then apply the same regression on each multiplicative / additive component of trade costs.
-Suppose that we find a high correlation with the additive cost, and almost 0 with the multiplicative cost.
-This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error.
-\bigskip
-
-\paragraph{Data treatment} Before presenting the estimated equation, let us mention how we deal with the data.
-We regress 1°) the ``overall trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{ik}^{nlI}$ hereafter ($i$ for the origin country, $k$ for the product and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{ik}$ and $t_{ik}$ respectively, on some common determinants.
-Precisely, we re-treat the iceberg component (both $\tau_{ik}^{nlI}$ and $\tau_{ik}$) to have it expressed in percentage of the fob price, ie we build the variable $\tau_{ik} =  100(\widehat{\tau}_{ik}-1)$ (with $\widehat{\tau}_{ik}$ the estimated variable obtained at stage 1).
-As for the additive cost, from stage 1 of the paper, we have obtained the value of (country-specific) trade cost as a fraction of the fob price (denoted $\widehat{t}_{ik}$); we multiply it to have it expressed as a percentage of the fob price (considering $t_{ik} =  100 \widehat{t}_{ik}$).
-Further, and in line with our first-step estimation method, we estimate the determinants of trade costs on a yearly basis, and conditional to the transport mode (air or vessel).
-
-In the three cases, we want to explain trade costs expressed as a percentage of the fob price.
-All our three measures of trade costs are bounded by 0 as minimal value, and we drop outliers by only keeping the 95 first percentile of each distribution.
-\bigskip
-
-\paragraph{Functional form} To derive the estimated equation, we start recalling what is behind the gap between cif-fob prices.
-In the US import data we use, the cif price is based on applying import charges on the custom value.
-What is defined as import charges?
-As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
-import duties) incurred in bringing the merchandise from alongside the carrier''.
-So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''.
-This means that the seller has paid to bring the goods to the port, but not to put them on the boat.
-In this respect, all hauling charges are excluded from what we call the ``fob'' price.
-In comparison with the fas price, the cif price is ``cost, insurance, freight'' included.\footnote{From OECD stats, the c.i.f.
-price (i.e.
-cost, insurance and freight price) is the price of a good delivered at the frontier of the importing country, including any insurance and freight charges incurred to that point, or the price of a service delivered to a resident, before the payment of any import duties or other taxes on imports or trade and transport margins within the country.}
-
-Accordingly, if we denote $TC_{ik}$ the total transport cost paid to import the product $k$ from country $i$ (on a given year, conditional to a certain transport mode), expressed in USD, we can decompose it in three elements:
-
-$$TC_{ik} = \text{Freight + Insurance + Handling costs}$$
-
-Our second-stage estimation is then guided by two questions.
-For each of these three dimensions (cost, insurance, freight) at the root of the cif-fas prices gap, 1°) what are its determinants, and 2°) do we expect this dimension to play rather on the additive or the multiplicative dimension?
-
-\paragraph{Insurance}: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ???
-for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood).
-The database provides us with the quantity (in terms of kilograms) of the shipment ($q_{ik}$), as well as the price, so that we have the value of the shipment.
-However, we have no dedicated variable to measure the insurance cost in the database.
-Accordingly, we capture this dimension by including product fixed effects.
-Insurance grids retain a degree of product categorization much broader than the 3-digit level we retain as benchmark sectorial degree (our $k$ index for the product dimension).
-In most pieces of information we have collected, insurance grids adopt a ranking in \textbf{20 (??)} types of products.
-Accordingly, we include product fixed effects of dimension \textbf{20}, that we refer to with the subscript $k^\prime$, i.e.
-$FE_{k^\prime}$.
-If we denote $p_{ik}^{fob}$ and $q_{ik}$ the (fas) price and the quantity of the shipment, total insurance costs (in USD) are accordingly measured by $FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}$.
-As insurance cost depend on the value of the shipment, we expect it to be proportional to the price, ie rather play on the iceberg part of the transport costs.
-
-\paragraph{Costs} (for handling costs).
-The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs).
-If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs.
-We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD.
-As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$).
-Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular).
-We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following.
-Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$.
-As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.
-
- \paragraph{Freight}.
-Quite intuitively, freight costs should depend on the distance between the origin country and the U.S.
-(denoted $dist_i$).
-Further, they likely depend on the quantity ($q_{ik}$) and the bulk of the shipment.
-Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs.
-Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$.
-Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.
-
-As a result, total transport costs by $TC_{ik}$ (expressed in USD) can be decomposed in the three components of insurance, handling costs and freight, according to:
-
-$$TC_{ik} =  \underbrace{EC_i\times V_{k}\times q_{ik}}_{\text{Costs}} +\underbrace{FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}}_{\text{Insurance}}+ \underbrace{V_k\times dist_i\times q_{ik}}_{Freight}$$
-
-By definition, such transport costs can be written as $TC_{ik} = (p^{cif}_{ik} - p^{fas}_{ik})q_{ik}$. Accordingly, we can write:
-
-\begin{eqnarray*}
-&&(p^{cif}_{ik} - p^{fas}_{ik})q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_{k}\times dist_i\times q_{ik} \\
-\Leftrightarrow && \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)p^{fas}_{ik} q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_k \times dist_i\times q_{ik}
-\end{eqnarray*}
-
-From the above equation, it comes that $\frac{TC_{ik}}{p^{fas}_{ik} q_{ik}} = \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)$, which is exactly the measure of trade costs obtained in the first stage.\footnote{Recall that our first stage starts from the equation: $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau^{nlI}_{ik}$ in the ``raw'' equation and $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau_{ik}+ \frac{t_{ik}}{p^{fas}_{ik}}$ when decomposing between iceberg and additive transport costs.}
-
-Accordingly, we can decompose each of our three measures of trade costs ($\tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, expressed in the three cases as a percentage of the fob price, and denoted $tc_{ik}$ to gain in generality, according to:
-
-\begin{equation}
-tc_{ik} = \underbrace{\frac{EC_i\times V_{k}}{p^{fas}_{ik}}}_{\text{C}} + \underbrace{FE_{k^\prime}}_{\text{I}}+ \underbrace{\frac{V_k\times dist_i}{p_{ik}^{fas}}}_{\text{F}} \label{eq:theory}
-\end{equation}
-
-Based on Equation (\ref{eq:theory}), we specify the estimated equation as follows:
-
-\begin{equation}
-tc_{ik} = \left[\frac{EC_i\times V_k}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k\right]e^{\varepsilon_{ik}} \label{eq:estim}
-\end{equation}
-
-with $\varepsilon_{ik}$ the error term, centered on 0.
-This calls for two comments.
-First, we add the country fixed effects separately in the regression, in order to correctly identify the associated interaction term.
-Second, we specify the error term multiplicatively.
-If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points).
-In this case, it is very likely that the variance of the error term depends on the size of the observed costs.
-E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high.
-Adopting this functional form would then be subject to a potential bias in point estimates due to the heteroscedasticity of error terms.
-This drives us to specify the error term in a multiplicative manner (such that the error term is in percentage of each explicative variable and centered on 0), and to take the log of the equation.
-
-Accordingly, the functional form that we estimate, for each of the three transport costs obtained at the first stage ($tc_{ik} = \tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, conditional on the year and the transport mode), is expressed as:
-
-\begin{equation}
-\log tc_{ik} = \log\left[\frac{EC_i\times V_{k}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k \right]+ \varepsilon_{ik} \label{eq:estim1}
-\end{equation}
-
-%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
-So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ...
-\bigskip
-
-%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !
-
-\subsection{Results and Interpretation}
-%\section{One step further into interpretation: transportation costs as an indicator of market structure.}
-
-\section{Conclusion}
+\section{Conclusion \label{sec:conclu}}
 
@@
 \subsection{Fob-cif prices}
-From the US Census website:
 
@@
 
-\subsection{Product classification for Insurance fixed effects}
-\textbf{to be done}
-
-\subsection{Volumetric weight and the Maritime Transport Costs database}
-The Maritime Transport Costs (MTC) database built by the OECD\footnote{See \texttt{http://stats.oecd.org/Index.aspx?datasetcode=MTC}.} contains data from 1991 to the most recent available year of bilateral maritime transport costs.
-Transport costs are available for 43 importing countries (including EU15 countries as a custom union) from 218 countries of origin at the detailed commodity (6 digit) level of the Harmonized System 1988.
-
-At commodity level (6 digit HS), maritime transport costs are displayed using three measures:
-\begin{itemize}
-\item Transport cost: the total cost expressed in USD (insurance plus freight) of transporting all of the given product during the given year,
-\item  Unit transport cost: transport cost per kilogramme, or in other words, the cost in USD required to transport one kilogramme of merchandise, and
-\item Ad valorem equivalent: transport cost divided by the  total import value, i.e. the share transport cost represents in the total import value of the product.
-\end{itemize}
-At chapter level (2 digit HS), the transport costs are the sum of the corresponding 6 digits commodities.
-The unit cost and ad valorem cost are then the transport costs at chapter level divided respectively by sum of the weights at 6 digits, and by the sum of the import values at 6 digits.
-
-\subsection{Country-specific handling costs and the ``Cost to export'' variable}
-
-Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport.
-The most recent round of data collection for the project was completed in June 2014.
-Among the four variables provided, we retain the variable ``Cost to export'' (US\$ per container), as it represents the ``cost associated with all procedures required to export goods[...,] including the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' We use the variable observed \textbf{in year ???
-attention la variable est-elle destination spefici?
-A priori non, a verifier.}
-\newpage
-
-\section*{Tables and Figures}
-
 %%%Figures
 % Goodness-of-fit
-
-\begin{figure}[htbp]
-\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-%Economic interpretation
-%usepackage{float} + H is great to keep images on the right page
-\begin{figure}[H]
-\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
 
@@
 
-%Economic Interpretation
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034 & \multicolumn{1}{c}{\textbf{1.051}} \\
-Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029 & \multicolumn{1}{c}{\textbf{1.042}} \\
-Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
-\hline
-\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010 & \multicolumn{1}{c}{\textbf{0.019}} \\
-\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005 & \multicolumn{1}{c}{\textbf{0.007}} \\
-\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
-\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017 & \multicolumn{1}{c}{\textbf{1.025}} \\
-\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 & \multicolumn{1}{c}{\textbf{1.018}} \\
-Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
-\hline
-\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036 & \textbf{1.058} \\
-Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033 & \textbf{1.051} \\
-Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
-\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-\hline
-\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
-Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015 & \textbf{0.029} \\
-Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 & \textbf{0.019} \\
-Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 & \textbf{1.032} \\
-Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018 & \textbf{1.028} \\
-Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
-\hline
-\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\newpage
+%
 
@@
 \end{document}
+
+In section \ref{sec:determinants}, we provide rationales for those differences.
+
+\section{Reassessing the Traditional Determinants of Transport Costs}\label{sec:determinants}
+
+\subsection{Methodological Concerns}
+
+%In step 2, we ask if the traditional determinants of transportation costs impact differently the multiplicative and additive costs estimated in the first step.
+%Therefore, we consistently focus only on origin-country-specific transportation costs, taking the estimated $\tau_{i}$ and $t_{i}$ from step 1 as dependent variables.
+In line with the trade literature, we proxy transport costs by (log of) distance and other proxies for trade costs (mainly, geographic and cultural factors).\interfootnotelinepenalty=10000\footnote{All those variables come from the well-known GeoDist database built by the CEPII: \url{http://www.cepii.fr/CEPII/fr/bdd_modele/presentation.asp?id=6}.} More formally, we estimate, still on a yearly basis and for each transportation mode:
+
+%\begin{eqnarray}
+%ln(\widehat{\tau}_{i}) &=& \beta_{1} \textrm{ln(distance)} + \Gamma_{1} \textrm{Gravity} + \mu_{i} \label{eq:step2_iceberg}\\
+%ln(\widehat{t}_{i}) &=& \beta_{2} \textrm{ln(distance)} + \Gamma_{2} \textrm{Gravity} + \nu_{i} \label{eq:step2_additive}
+%\end{eqnarray}
+
+%where the ln operator denotes that variables have been log-linearized, $distance$ is the bilateral distance between US and the origin country $i$  and $Gravity$ is a vector of controls including the languages spoken in the country under different definitions, whether the country is landlocked, and their colonial links.
+
+In this second stage of the paper, we thus keep on questioning the way the literature models the trade costs determinants.
+Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs?
+Or total?
+Usually, people use a measure of distance as a way to capture iceberg trade costs.
+What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs?
+Using our decomposition of transport costs obtained in the first stage, we can assess the contributions of typical gravity variables to each component of transport costs (additive / iceberg).
+To put this simply, consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and find that this fits reasonably well, the estimated coefficient stands in the average.
+We then apply the same regression on each multiplicative / additive component of trade costs.
+Suppose that we find a high correlation with the additive cost, and almost 0 with the multiplicative cost.
+This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error.
+\bigskip
+
+\paragraph{Data treatment} Before presenting the estimated equation, let us mention how we deal with the data.
+We regress 1°) the ``overall trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{ik}^{nlI}$ hereafter ($i$ for the origin country, $k$ for the product and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{ik}$ and $t_{ik}$ respectively, on some common determinants.
+Precisely, we re-treat the iceberg component (both $\tau_{ik}^{nlI}$ and $\tau_{ik}$) to have it expressed in percentage of the fob price, ie we build the variable $\tau_{ik} =  100(\widehat{\tau}_{ik}-1)$ (with $\widehat{\tau}_{ik}$ the estimated variable obtained at stage 1).
+As for the additive cost, from stage 1 of the paper, we have obtained the value of (country-specific) trade cost as a fraction of the fob price (denoted $\widehat{t}_{ik}$); we multiply it to have it expressed as a percentage of the fob price (considering $t_{ik} =  100 \widehat{t}_{ik}$).
+Further, and in line with our first-step estimation method, we estimate the determinants of trade costs on a yearly basis, and conditional to the transport mode (air or vessel).
+
+In the three cases, we want to explain trade costs expressed as a percentage of the fob price.
+All our three measures of trade costs are bounded by 0 as minimal value, and we drop outliers by only keeping the 95 first percentile of each distribution.
+\bigskip
+
+\paragraph{Functional form} To derive the estimated equation, we start recalling what is behind the gap between cif-fob prices.
+In the US import data we use, the cif price is based on applying import charges on the custom value.
+What is defined as import charges?
+As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
+import duties) incurred in bringing the merchandise from alongside the carrier''.
+So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''.
+This means that the seller has paid to bring the goods to the port, but not to put them on the boat.
+In this respect, all hauling charges are excluded from what we call the ``fob'' price.
+In comparison with the fas price, the cif price is ``cost, insurance, freight'' included.\footnote{From OECD stats, the c.i.f.
+price (i.e.
+cost, insurance and freight price) is the price of a good delivered at the frontier of the importing country, including any insurance and freight charges incurred to that point, or the price of a service delivered to a resident, before the payment of any import duties or other taxes on imports or trade and transport margins within the country.}
+
+Accordingly, if we denote $TC_{ik}$ the total transport cost paid to import the product $k$ from country $i$ (on a given year, conditional to a certain transport mode), expressed in USD, we can decompose it in three elements:
+
+$$TC_{ik} = \text{Freight + Insurance + Handling costs}$$
+
+Our second-stage estimation is then guided by two questions.
+For each of these three dimensions (cost, insurance, freight) at the root of the cif-fas prices gap, 1°) what are its determinants, and 2°) do we expect this dimension to play rather on the additive or the multiplicative dimension?
+
+\paragraph{Insurance}: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ???
+for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood).
+The database provides us with the quantity (in terms of kilograms) of the shipment ($q_{ik}$), as well as the price, so that we have the value of the shipment.
+However, we have no dedicated variable to measure the insurance cost in the database.
+Accordingly, we capture this dimension by including product fixed effects.
+Insurance grids retain a degree of product categorization much broader than the 3-digit level we retain as benchmark sectorial degree (our $k$ index for the product dimension).
+In most pieces of information we have collected, insurance grids adopt a ranking in \textbf{20 (??)} types of products.
+Accordingly, we include product fixed effects of dimension \textbf{20}, that we refer to with the subscript $k^\prime$, i.e.
+$FE_{k^\prime}$.
+If we denote $p_{ik}^{fob}$ and $q_{ik}$ the (fas) price and the quantity of the shipment, total insurance costs (in USD) are accordingly measured by $FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}$.
+As insurance cost depend on the value of the shipment, we expect it to be proportional to the price, ie rather play on the iceberg part of the transport costs.
+
+\paragraph{Costs} (for handling costs).
+The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs).
+If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs.
+We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD.
+As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$).
+Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular).
+We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following.
+Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$.
+As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.
+
+ \paragraph{Freight}.
+Quite intuitively, freight costs should depend on the distance between the origin country and the U.S.
+(denoted $dist_i$).
+Further, they likely depend on the quantity ($q_{ik}$) and the bulk of the shipment.
+Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs.
+Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$.
+Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.
+
+As a result, total transport costs by $TC_{ik}$ (expressed in USD) can be decomposed in the three components of insurance, handling costs and freight, according to:
+
+$$TC_{ik} =  \underbrace{EC_i\times V_{k}\times q_{ik}}_{\text{Costs}} +\underbrace{FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}}_{\text{Insurance}}+ \underbrace{V_k\times dist_i\times q_{ik}}_{Freight}$$
+
+By definition, such transport costs can be written as $TC_{ik} = (p^{cif}_{ik} - p^{fas}_{ik})q_{ik}$. Accordingly, we can write:
+
+\begin{eqnarray*}
+&&(p^{cif}_{ik} - p^{fas}_{ik})q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_{k}\times dist_i\times q_{ik} \\
+\Leftrightarrow && \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)p^{fas}_{ik} q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_k \times dist_i\times q_{ik}
+\end{eqnarray*}
+
+From the above equation, it comes that $\frac{TC_{ik}}{p^{fas}_{ik} q_{ik}} = \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)$, which is exactly the measure of trade costs obtained in the first stage.\footnote{Recall that our first stage starts from the equation: $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau^{nlI}_{ik}$ in the ``raw'' equation and $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau_{ik}+ \frac{t_{ik}}{p^{fas}_{ik}}$ when decomposing between iceberg and additive transport costs.}
+
+Accordingly, we can decompose each of our three measures of trade costs ($\tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, expressed in the three cases as a percentage of the fob price, and denoted $tc_{ik}$ to gain in generality, according to:
+
+\begin{equation}
+tc_{ik} = \underbrace{\frac{EC_i\times V_{k}}{p^{fas}_{ik}}}_{\text{C}} + \underbrace{FE_{k^\prime}}_{\text{I}}+ \underbrace{\frac{V_k\times dist_i}{p_{ik}^{fas}}}_{\text{F}} \label{eq:theory}
+\end{equation}
+
+Based on Equation (\ref{eq:theory}), we specify the estimated equation as follows:
+
+\begin{equation}
+tc_{ik} = \left[\frac{EC_i\times V_k}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k\right]e^{\varepsilon_{ik}} \label{eq:estim}
+\end{equation}
+
+with $\varepsilon_{ik}$ the error term, centered on 0.
+This calls for two comments.
+First, we add the country fixed effects separately in the regression, in order to correctly identify the associated interaction term.
+Second, we specify the error term multiplicatively.
+If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points).
+In this case, it is very likely that the variance of the error term depends on the size of the observed costs.
+E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high.
+Adopting this functional form would then be subject to a potential bias in point estimates due to the heteroscedasticity of error terms.
+This drives us to specify the error term in a multiplicative manner (such that the error term is in percentage of each explicative variable and centered on 0), and to take the log of the equation.
+
+Accordingly, the functional form that we estimate, for each of the three transport costs obtained at the first stage ($tc_{ik} = \tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, conditional on the year and the transport mode), is expressed as:
+
+\begin{equation}
+\log tc_{ik} = \log\left[\frac{EC_i\times V_{k}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k \right]+ \varepsilon_{ik} \label{eq:estim1}
+\end{equation}
+
+%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
+So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ...
+\bigskip
+
+%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !
+
+\subsection{Results and Interpretation}
+%\section{One step further into interpretation: transportation costs as an indicator of market structure.}
+
+\end{document}

### 2016-07-20 redaction/poubelle/note_secondstage.tex
Rédaction dans le github

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
%\usepackage[nolists, figuresfirst]{endfloat}
%\usepackage{sidefloat}

%\usepackage[usenames]{color}
%\definecolor{grey}{rgb}{0.35,0.35,0.35}
%\definecolor{webdarkblue}{rgb}{0,0,0.4}
%\definecolor{orange}{rgb}{0.7,0.2,0.05}

%\usepackage[pdfcreator={PDFLaTeX}, pdfproducer={PDFLaTeX}, pdfstartview=FitH, pdfpagemode=UseOutlines, pagebackref=false, colorlinks={true},
%citecolor={webdarkblue}, linkcolor={webdarkblue},
%urlcolor={webdarkblue}]{hyperref}

%\addtolength{\oddsidemargin}{-0.4in}
%\addtolength{\evensidemargin}{-0.4in}
%\addtolength{\textwidth}{0.8in} \addtolength{\topmargin}{-0.85in}
%\addtolength{\textheight}{1.7in}
\renewcommand{\baselinestretch}{1.2}

\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}

% commands

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}
\newcommand{\beqa}{\begin{eqnarray}}
\newcommand{\eeqa}{\end{eqnarray}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bs}{\bigskip}
\newcommand{\A}{$^a$}
\newcommand{\B}{$^b$}
\newcommand{\C}{$^c$}
\begin{document}

\begin{center}

\textsc{On the relative importance of iceberg  and additive transport costs in international trade} \\ \vspace{1 cm}

\textsc{Note: Second-stage regression}

\end{center}

In this second stage regression, the key question is: Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs? Or total? Usually, people use a measure of distance as a way to capture iceberg trade costs. Using our decomposition of transport costs, we want to assess the contributions of such proxies to each component (additive / iceberg). What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs?

This question drives our second-step analysis. Consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and found that this first reasonably well, the estimated coefficient stands in the average. We then apply the same regression on each multiplicative / additive component of trade costs. Suppose that we find that we find a high correlation with the additive cost, almost 0 with the multiplicative cost. This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error.

To investigate this point, we thus regress 1°) the ``overal trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{iy}^{nlI}$ hereafter ($i$ for the origin country, $y$ for the estimation year and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{iy}$ and $t_{iy}$ respectively, on some common determinants. \medskip

Precisely, what are these dependent variables? We use our first-stage regression results, from which we extract the fixed effect of transport costs associated to the country dimension (recall that we assumed that the iceberg trade cost from country $i$, good $k$, year $y$, $\tau_{iky} = \tau_{iy}\tau_{ky}$, and  similarly for the additive transport cost $t_{ik} = t_{iy}+ t_{ky}$. Both are expressed in percentage (of the fob price).

\textbf{Careful}: This supposes some re-treatment of our first-stage estimation results (recalling our general estimated equation):
$$p_{iky}^{cif} = \left[\widehat{\tau}_{iy}\widehat{\tau}_{ky}\right] p^{fob}_{iky} +  \left[\widehat{t}_{iy}\widehat{t}_{ky}\right]$$

From this, we re-treat the iceberg component (both estimated alone and with the additive part) to consider $\tau_{iy} =  100(\widehat{\tau}_{iy}-1)$ (leaving aside the ``product'' dimension as unrelated to the question here), to have the (country-specific) iceberg trade cost expressed in percentage of the fob price.

As for the additive cost, we have obtained the value of (country-specific) trade cost as a fraction of the fob price, that we re-scale to consider it in percentage, ie considering $t_{iy} =  100 t_{iy}$, so that we have the additive cost in percentage of the fob price to explain. \medskip

We use data from 2004 to 2013, because we have information provided by Doing Business about exports formalities on an annual basis on this period.\bigskip

Extracted from the Doing Business website:

Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport. The time and cost necessary to complete 4 predefined stages (document preparation; customs clearance and inspections; inland transport and handling; and port and terminal handling) for exporting and importing the goods are recorded; however, the time and cost for sea transport are not included. All documents needed by the trader to export or import the goods across the border are also recorded. The most recent round of data collection for the project was completed in June 2014.

In particular, we are interested in the variable ``Cost to export'' (US\$ per container). From DB: It is the ``cost associated with all procedures required to export goods. Includes the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' \textbf{Careful}: inland transport, transport between the warehouse (assumed to be in the main economic city of the country) and the port. So, this part of the export formality cost is not in the cif-fob price, because the fob price already includes the cost incurred to bring the shipment from the initial warehouse to the boat (even includes handling at the port to bring the merchandises on the boat itself). So, there is a measurement error, the export formality cost is higher than it should be; a priori, bias the associated coefficient downwards.

\textbf{What do we call ``fob'' price exactly?} In the US import data we use, the cif price is based on applying import charges on the custom value. What is defined as import charges? As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S. import duties) incurred in bringing the merchandise from alongside the carrier''. So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''. This means that the seller has paid to bring the goods to the port, but not to put them on the boat. In this respect, all hauling charges are excluded from what we call the ``fob'' price. Our ``export formality cost'' measure is thus adequate as potential explanation of the gap between the cif price and the fob (in fact, fas) price.\bigskip

One way of explaining our two-stage reasoning is to start recalling the equation at the root of the analysis:

\begin{equation}
p^{cif}_{iy} q_{iy} = p^{fob}_{iy} q_{iy} \tau_{iy} + t_{iy} q_{iy} \label{eq:cif-fob}
\end{equation}

This equation states that the value of an import by the US, from country $i$, in year $y$ (we neglect the product dimension here) ($p^{cif}_{iy} q_{iy}$), is equal to the value of the good exported ($p^{fob}_{iy} q_{iy}$), raised by a given multiplicative cost $\tau_{iy}$, plus an additive cost component, that depends on the quantity of the good exported $ t_{iy} q_{iy}$. Manipulating the equation, we get:

\begin{equation}
\frac{p^{cif}_{iy}}{p^{fob}_{iy}} = \tau_{iy} +\frac{t_{iy}}{p^{fob}_{iy}}
\label{eq:cif-fob2}
\end{equation}

In the first step, we decompose the gap between the cif and the fob price between its two multiplicative and additive components. In the second stage, we raise the question, what is behind the gap between the fob and the cif prices? As clear from Equation (\ref{eq:cif-fob}), the multiplicative cost is tied to the price of the good exported (the fob price), while the additive cost is rather applied on the quantity exported. This drives us to refine the above question: What is behind each component of the cif-fob gap? In particular, do the gravity variables, primarily distance, that are commonly used as a proxy for iceberg trade costs, really matter on iceberg (i.e., multiplicative) transport costs, or rather on the additive part? \medskip

To answer the question, it is worth recalling what the ``cif'' acronym mean, i.e. what is at the root of the gap between the cif-fob prices. In comparison with the fob price, the cif price is ``cost, insurance, freight'' included. Our second-stage estimation is then guided by two questions. For each dimension of the three dimension, 1°) how can we proxy this dimension, and 2°) do we expect this dimension to play rather on the additive or the multiplicative dimension?

From OECD stats: The c.i.f. price (i.e. cost, insurance and freight price) is the price of a good delivered at the frontier of the importing country, including any insurance and freight charges incurred to that point, or the price of a service delivered to a resident, before the payment of any import duties or other taxes on imports or trade and transport margins within the country.

\begin{itemize}

\item Freight. 1°) how to measure it: the distance from the origin country, as well the energy cost implied by this distance. We accordingly use the ''distance'' variable, and the oil price per kilometer covered. 2°) additive or multiplicative? We expect that this cost depends more on the quantity of goods exported, rather than their value. In terms of the estimated equation, we expect $\beta_2$ and $\alpha$ to be non significantly different from 0 when Equation (\ref{eq:2dstage}) is estimated on the iceberg transport cost $\tau_{ij}$, while significantly positive when considering the additive component $t_{ij}$ (in percentage of the fob price).
\item Costs (for handling costs). 1°) how to measure it: Our first measure is the ``export formality cost''  provided by the Doing Business database. Our second measure is the oil price (per USD exported) involved in the shipment, that we view as measuring the maintenance cost of the shipment (independent of the distance covered). 2°) additive or multiplicative? We expect this dimension to be additive (related to the quantities of goods that are exported, but not to their value (ie, their price)). In terms of the estimated equation, we expect $\gamma$ and $\beta_1$ to be non significantly different from 0 when Equation (\ref{eq:2dstage}) is estimated on the iceberg transport cost $\tau_{ij}$, while significantly positive when considering the additive component $t_{ij}$ (in percentage of the fob price).
\item Insurance: 1°) how to measure it: We have no specific variable to measure the insurance cost. Accordingly, it is likely to be captured by the constant term. You may also argue that insurance cost is related to the distance covered. 2°) additive or multiplicative? By contrast to the two other dimensions of the cif-fob price, we expect insurance costs to be proportional to the unit price; accordingly, it is likely to imply a coefficient $\alpha$ significantly different from 0 when estimation is on the iceberg part.

\end{itemize}

\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively. So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' Even if we carefully explain that indeed, we are aware of that, the (second) counter-argument could be, ok but then what's the point...

\textbf{Questions raised} \medskip
\begin{itemize}
\item Which variable to consider for export formality costs? In the DB database, we also have information about the \textit{1°)} number of documents to fulfill (``The total number of documents required per shipment to export goods. Documents required for clearance by government ministries, customs authorities, port and container terminal authorities, health and technical control agencies and banks are taken into account), \textit{2°)} the time to export (in days) (``The time necessary to comply with all procedures required to export goods. If a procedure can be accelerated for an additional cost, the fastest legal procedure is chosen.'') and \textit{3°)} the cost to export (deflated US\$). Ie, ``the Cost to export for each year is divided by GDP deflator, to take the general price level into account when benchmarking this absolute-cost indicator across economies with different inflation trends. In Doing Business 2015 the deflated costs are identical to the non-deflated (base year for the deflator).''

\item[$\Rightarrow$] For now we have chosen the ``cost of export'' . Deflated or not? The answer is no, since we have on the left, the transport costs in percentage of the fob price, not deflated. Consider also the ``administrative costs'' (ie nb of documents and time delay)?, but pb of compatibility of measures? Answer: No. The reason is that the custom delay and the workload required to fill the documents are not included in the declaration fob-caf, which is what we use as measure of transport costs. Even if such time delay or administrative burden are likely to affect the magnitude of trade flows.

\item Which other variables to consider? The literature commonly uses gravity variables (such as distance, common language, colonial linkages, common border, etc.) as a proxy for trade costs. In the second-step estimation, we only consider distance. This choice relies on the fact that we explain the transport gap measured by the gap between the fob price and the caf price. Our measure of transport cost is only part of the larger size of trade costs, andthere is no a priori reason why the other gravity variables except distance might intervene in the diffence between the price at the export departure point and the import arrival point.

\item Question of dimension comparability. On the leftwards of the equation, we have a percentage (iceberg or additive transport costs, in \% of the for price), in the country-specific dimension. As for the export formalities costs, in the raw DB database, expressed in US\$ per container transported. We want to have it expressed in terms of percentage of the fob price, to have the same dimension on both sides of the equation. We thus calculate a (weighted) mean fob price per country/year, and we divide the formality cost by the mean fob price, so that this variable represents the formality cost per US\$ exported.

\item Oil price and distance Raw data for oil prices are from BP website (crude oil prices, US\$ per barrel).\footnote{See more details in Appendix \ref{app:data}.} The oil price is expressed in US\$ per barrel. To ensure comparability, we transform this variable to express it as percentage of the mean fob price (per country/year). We consider this variable alone (the cost of oil per USD exported) and in interaction with distance (the oil cost burden per kilometer made to export). We come back later on the interpretation of the associated estimated coefficients below.

\item Time dimension. How to treat it? Two options, run the regression year by year, but not many point in the database (160 to 210 countries depending on the year considered). So, in panel, way to increase the number of observations. Besides, also exploit the idea that technological improvements in aircraft and vessel might have changed the relation between oil prices and trade cost. To capture this dimension, we interact a year fixed effect with the ``oil cost per kilometer exported'' variable.

\item Make a difference between the estimated additive trade cost and the multiplicative trade costs. Ie, should we suppose that the explicative variables enter additively when considering the additive trade costs, and multiplicatively regarding the iceberg component? No a priori reason. In both cases, the dependent variable is expressed as a percentage (of the fob price). We decompose this percentage in various dimension (the cost of oil, the cost of distance, etc.) , which all added, compose the overall trade cost component. This justifies an additive specification, for both the additive and the iceberg trade costs components.

\item The specification of error. If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points). If we specify the error term in a multiplicative manner, it means that we estimate an error term in percentage of each explicative variable. See below with the estimated equation.

\item The transport mode. We have run estimation separating air from vessel. Question, how to treat the ``mode'' dimension? If we consider them jointly (on a yearly basis), and distinguish between air and vessel by including a dummy for, say, air, it amounts assuming that the only difference between trade costs associated to air and to vessel is in the constant term. This seems to us a quite restrictive assumption. The transport mode is rather to interact with the distance dimension, the oil price dimension, etc. with a specific manner. So, if all explicative variables are interacted with the mode dummy, there is no gain in having a single equation. This drives us to estimate the equation by transport mode. Then, we will compare the results (of distance, of oil price) between both modes, to explore whether there is (or not), a significant difference (regarding the roles of the various explicative variables, as well as in the constant terms) between both modes.

\end{itemize}

If we adopt the specification that the error term enters additively, on top of the other explicative variables, the estimated equation is the following (conditional on the transport mode, vessel or air, and the type of transport costs, iceberg or additive):
\begin{eqnarray}
TC_{iy} &=& a + \alpha dist_{i}+ \beta_1\frac{oil_{y}}{\bar{pfob}_{iy}} + \beta_2\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i} + \beta_3\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i}\times FE_y \notag \\
 &&+ FE_y + \gamma\frac{formalities_{iy}}{\bar{pfob}_{iy}} + \varepsilon_{iy} \label{eq:2dstage}
\end{eqnarray}

with the subscript $i$ referring to the origin country, $y$ the year of estimation. Under this specification,
\begin{itemize}
\item $TC_{iy}$ is the transport cost, expressed as a fraction of the fob price, that varies over time and with the origin country. That is, for 1 US\$ exported by country $i$ to the US in year $y$, it costs $TC_{iy}$ in trade costs (specific to the transport mode $x$ ($x=$ air, vessel) and the trade cost type $z$ (iceberg, additive or total\footnote{Meaning, retaining an iceberg formulation without modeling an additive component.}, denoted $\tau_{iy}^{nlI}$)).
\item $dist_i$ is the distance (in km) from the US,
\item $oil_{y}$ is the oil price in US\$, per year
\item $formalities_{iy}$ is the export formality costs in US\$, for country $i$ in year $y$ (from Doing Business database).
\end{itemize}
As explained above, we transform the last two variables to have them expressed in percentage of the mean fob price observed for the pair (origin country, year). Reasoning: $oil_{y}$ is the price of the barrel (in US\$, per year). Then $\frac{oil_{y}}{\bar{pfob}_{iy}}$ is the price of a barrel per US\$ exported. The variable $\frac{formalities_{iy}}{\bar{pfob}_{iy}}$ represents the export formalities cost (in US\$) per USD\$ exported (by year-country). $FE_y$ are year fixed effects (the reference year being 2004), $\varepsilon_{iy}$ is the residual, and $a$ is a constant term.\smallskip

Under this specification, the explicative variables and the associated coefficients can be interpreted as follows.

\begin{itemize}
\item $\alpha$ measures the cost for each kilometer made, in percentage of the fob price, i.e. for each US\$ exported, that is involved in the transport cost of mode $x$, type $z$. Such that $\alpha dist$ represents the share of transport cost ($x,z$) (in percentage of the fob price) that is attributable to distance (alone). It is expressed in percentage points.
\item $\beta_1$ represents the number of barrels involved per US\$ exported (by country-year, conditional on the transport mode and the transport cost type), such that $\beta_1\frac{oil_{y}}{\bar{pfob}_{iy}}$ measures the cost in oil per US\$ exported (the number of oil barrels per \$ exported ($\beta_1$), times the oil price of one barrel per \$ exported ($\frac{oil_{y}}{\bar{pfob}_{iy}}$)). In a sense, this measures the energy cost of freight involved in the export procedure.
\item The variable $\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i}$ represents the oil price cost per barrel, of a kilometer covered by the shipment. $\beta_2$ can thus be interpreted as the number of barrels associated with a kilometer ``exported'', such that $\beta_2\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i}$ measures the cost in oil involved in the shipment (the number of barrels, times the cost of an oil barrel par USD exported, times the number of kilometers crossed.)

\item $\beta_3\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i}\times FE_y$. Accordingly, $\beta_3$ in a vector of dimension (1,11) (2005-2013). The idea here is to capture the possibility that the oil cost of international trade might have changed over the years, because aircrafts and vessels have gained in energy needs, so that everything else equal the cost in oil per km covered for exports is lower in 2013 than in 2004 because the machines are less demanding.  If this intuition is correct, we expect (the vector of) $\beta_3$ to be negative (and all the more negative as we are on recebt years).
\item Given that $formalities_{iy}$ represents the cost of export formality per container (in USD), and $\frac{formalities_{iy}}{\bar{pfob}_{iy}}$ the export formalities of a container per USD exported, $\gamma$ can be interpreted as the number of containers per USD exported, such that $\gamma\frac{formalities_{iy}}{\bar{pfob}_{iy}}$ scales the overall cost of export formality, per USD exported.
\item $\varepsilon_{iy}$ is the trade cost per USD exported that is left unexplained (for instance, if on the left-hand side we have a value for a given pair (country-year) of 0.008 (for this counry-year pair, trade costs is 0.8\% of the fob price, or 0.8 USD for 1 USD exported), and that the sum of all explicative terms (from $a$ to $\gamma\frac{formalities_{iy}}{\bar{pfob}_{iy}}$) amounts to 0.007, it means that the error term is 0.001, equal to 0.1 percentage point of the trade cost (in \% of the fob price).

\end{itemize}
\bigskip
\textbf{rediger totpo sur forme de l'erreur}
It is not impossible that the variance of the error term depends on the size of the observed costs. E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high. To correct for the potential heteroskedasticity, we report Huber?White standard errors.

\appendix

\section{Data Appendix \label{app:data}}

\subsection{Fob-cif prices}
From the US Census website:

The Customs value is the value of imports as appraised by the U.S. Customs and Border Protection in accordance with the legal requirements of the Tariff Act of 1930, as amended. This value is generally defined as the price actually paid or payable for merchandise when sold for exportation to the United States, excluding U.S. import duties, freight, insurance, and other charges incurred in bringing the merchandise to the United States. The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller. In the case of transactions between related parties, the relationship between buyer and seller should not influence the Customs value.

In those instances where assistance was furnished to a foreign manufacturer for use in producing an article which is imported into the United States, the value of the assistance is required to be included in the value reported for the merchandise. Such ``assists'' include both tangible and intangible assistance, such as machinery, tools, dies and molds, blue prints, copyrights, research and development, and engineering and consulting services. If the value of these ``assists'' is identified and separately reported, it is subtracted from the value during statistical processing. However, where it is not possible to isolate the value of ``assists'', they are included. In these cases the unit values may be increased due to the inclusion of such ``assists''.
Import Charges

The import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S. import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States. In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States) in Canada or Mexico to the first port of entry.
C.I.F. Import Value

The C.I.F. (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States. It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S. import duties.

\subsection{The second-stage explicative variables}

\paragraph{Oil prices} It may come from two sources, and for two measures. 
\begin{itemize}
\item From the BP Statistical Review of World Energy, we get an Oil Price series from 1974 (annual data, in dollars per barrel). 
\item From the Federal reserve of Saint Louis (\texttt{https://www.stlouisfed.org/}), we get the ``Crude Oil Prices: West Texas Intermediate (WTI) - Cushing, Oklahoma'' series, expressed in dollars per Barrel (Annual, Not Seasonally Adjusted data). It is available from 1986 to 2013. We could also use the ``Crude Oil Prices: Brent - Europe'' series (still in dollars per Barrel, from 1987 to 2013).

\end{itemize}

We check that these 3 series are in fact very close to one another over 1987-2013. Because we don't want to limit ourselves to data post 2004 (the availability of the ``export formality cost'' from Doing Business, in order to exploit all information over the entire period 1974-2013, we retain the BP series as benchmark (very frequently retained in the literature), as it is the sole which is available over the whole period.

\end{document}

### 2016-07-20 redaction/poubelle/note_secondstage_v0812.tex
Rédaction dans le github

--- 2016-06-13 redaction/dhp_draft_new.tex
@@
 \usepackage{hyperref}
-\usepackage{float}
-\usepackage{caption}
-\usepackage[tableposition=top]{caption}
 %\usepackage[nolists, figuresfirst]{endfloat}
 %\usepackage{sidefloat}
-
-\setlength{\abovecaptionskip}{-8pt}
 
@@
 
-\date{\today}
-\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX.
-Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01.
-Finally, any remaining errors are ours.}}}
-%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
-\author{Guillaume \textsc{Daud}\thanks{%
-Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa), Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa)} } \maketitle
+\begin{center}
+\begin{Large}
+\textsc{Opening the transport costs black box} \\ \vspace{1 cm}
 
-\bigskip
+\textsc{Note: Second-stage regression}
+\end{Large}
+\end{center}
 
-\begin{abstract}
+One way of explaining our two-stage reasoning is to start recalling the equation at the root of the analysis:
 
-\begin{spacing}{1.2}
-
-{\normalsize XXXX}
-
-{\normalsize \bigskip }
-
-\end{spacing}
-\end{abstract}
-
-\thispagestyle{empty} \pagestyle{plain} \setcounter{page}{1}
-
-{\normalsize JEL classification: F14, N70, R40\newline
-Keywords: Transport costs estimates, transport costs determinants, non-linear econometrics, period 1974-2013, additive transport costs }
-
-{\normalsize \vspace{0cm} }
-
-{\normalsize \titlepage }
-
-{\normalsize \newpage }
-
-%\begin{spacing}{1.5}
-
-\section{Introduction}
-
-\noindent Trade costs remain central in international economic analysis.
-Defined as the costs associated with the exchange of goods across national borders, they are a major obstacle to international economic integration.
-In addition, they might be the key to the explanation of major puzzles in international macroeconomics (\citealp{rogoff}).
-Historical perspective show that indeed, trade costs variations are a major determinant of the level of trade flows.
-According to the estimates by \citet{Jacks08}, trade cost declines explain around 55 percent of the pre World War I trade boom and 33 percent of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter war trade collapse.
-After 1950, average trade costs fell by 16\%.
-Based on panel data, \citet{novy13} finds  that U.S.
-trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-
-Trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff-costs), time costs (time to ship goods) and transport costs \emph{per se}.
-Many believe that the latter have dramatically decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-\citet{Glaeser04} find that, over the twentieth century, the costs of moving goods have declined by over 90\% in real terms.
-However, \citet{hummels2007} shows that the bulk of price declines in transportation comes from air shipping, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004; concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, even if the rise in containerization lowered shipping costs from 3 to 13\%.
-Studies overviewed by \citet{Behar_Venables} are consistent with those results: the fall in measured transport costs has been relatively small.
-The importance of fuel costs and technical progress (which improved speed and reliability rather than decreased costs) in transport explain that result.
-
-Despite the decrease in distance-related costs\footnote{\citet{Behar_Venables} suggest freight costs account for two thirds of the effect of distance, the remaining third ebmoyding more or less the effect of time}, several papers (mostly based on empirical estimates of the gravity equation) in the last decade have shown that distance remained a major obstacle to trade (e.g.
-\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
-\citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs (1.7 $=$ 1.55 $\times$ 1.74 - 1).
-Again, the latter comprises 44\% border-related trade barriers and 21\% transportation costs.
-That means that transportation costs are higher than each component of border costs taken separately (namely, 8\% for tariff and non-tariff costs, 7\% for language differences, 14\% for for currency differences and 9\% for other costs).\footnote{1.44 $=$ 1.08 $\times$ 1.07 $\times$ 1.14 $\times$  1.09)}.
-Therefore, the share of transport costs in the consumer price of manufactured goods remains high (\citealp{Lafourcade_Thisse}).
-Besides, \citet{combes_lafourcade} emphasize that the figures provided by \citet{anderson_wincoop_jel} should be considered as a a downward benchmark, and could be higher in developing economies.
-Despite technical progress and infrastructure development, transport costs remain large and deserve attention.
-In that respect, \citet{Behar_Venables} support that the elasticity of trade with respect to freight costs is around -3, which remains a large number.
-
-Following \citet{samuelson1954}, standard models of international trade have usually relied on the iceberg cost hypothesis, implicitly reducing trade costs to their transport part, and modelling the latter as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded).
-This implies that more expensive goods are costlier to trade.
-However, many common empirical facts support the existence of additive costs: pricing structure in shipping, additive tariffs, distribution costs...
-often exhibit (at least partly) an additive structure (see \citealp{Irrazabal_2015}).
-The additive versus iceberg trade costs is not a new debate in the international trade literature.
-The famous Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: the demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
-\citet{martin2012} gives a strong empirical support to this conjecture: based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite.
-\citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: the elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption.
-Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
-Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level.
-Based on US imports data, he finds that a 10\% increase in the price of an item increases the shipping cost by 9.5\%.
-The additive structure of trade costs emerges only when weight is forced to be independent of price.
-Besides, \citet{Lashka15} supports that higher-priced, still exported to faraway markets, varieties are heavier and involve proportionally higher shipping costs, the latter increasing the relative demand for high-markup varieties.
-Markups would therefore be the driving force for explaining the ``shipping the good apples out'' effect in the presence of \emph{ad-valorem} trade costs.
-
-A correct measure of transport costs is also important for understanding several key features of international trade.
-For example, the existence of additive costs may explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
-Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e.
-more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments.
-Besides, several recent papers show that the existence of additive trade costs has important non trivial effects when assessing the welfare effects of trade liberalization.
-\citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive (per-unit) trade costs, in addition to iceberg/multiplicative costs.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs.
-The difference arises from dissimilar adjustments at the extensive margin - fewer firms enter into the export market following a reduction in additive costs.
-Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff.
-
-Our purpose in this paper is to provide results on the size of transportation costs over time, explicitly distinguishing between multiplicative and additive parts.
-Our paper stands at the crossroad of these two strands of the literature, the first one dealing with the dynamics of transport, and more generally trade costs over time, the second one focusing on the measurement and nature of the latter.
-More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over a large time dimension - several decades.
-To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}.
-Therefore, our period of study runs from 1974 to 2013.
-Closely related to this paper is the work by \citet{Irrazabal_2015}, who develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level data, their results suggest that additive costs are on average 14 percent, relatively to the median price, and that they are strongly correlated with standard proxies for trade costs (like e.
-g., distance).
-However, while our data requires that we focus only on transportation costs, our approach departs from theirs in several key aspects.
-
-First, our theoretically agnostic approach aims at letting the data speak, and provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models; to our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the shares of multiplicative and additive costs in total transport costs.
-Second, we provide, through standard measures of ``goodness-of-fit'', an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Third, the time dimension of our data allows us to characterize the evolution of transportation costs, and more precisely the dynamics of the relative share of each specific (multiplicative and additive) cost over a forty years time span; to our knowledge, no firm-level database is able to provide such a time coverage.
-Fourth, we are also able to discriminate our results by transport mode (air and vessel).
-
-Finally, our analysis is also indirectly related to a large strand of the international trade literature assessing the impact of trade costs on international exchanges.
-Following \citet{anderson_wincoop} and \citet{anderson_wincoop_jel}, there has been a huge bunch of articles relying on gravity equations to infer indirectly trade costs as tariff equivalents, or losses related to various kinds of trade barriers - many of these works are surveyed in \citet{anderson_wincoop}; among the various subsequent works, see in particular \citet{anderson_yotov}, \citet{novy13} or \citealp{chen_novy}).
-However, our purpose in the present paper is different.
-We first infer directly transport costs from price data, discriminating between multiplicative and additive parts.
-In a second step, following an approach similar to \citet{Behar_Venables}, we relate the previous measures to traditional determinants of transport costs (like e.
-g., distance), in order to understand how differently the latter impact the multiplicative and additive transport costs.
-
-%time as a trade barrier, on risque d'avoir des problèmes avec ça... \citet{Lafourcade_Thisse},
-
-XXX Here a summary of our results and contribution XXX
-
-The next section presents the data and our fundamental equation.
-Section 3 reports our baseline estimates for multiplicative and additive measures of transport costs across time, as well as a number of robustness checks.
-Section 4 discusses the results of estimates relating traditional determinants of transport costs to the previous measures, while Section 5 XXX has yet to be discussed!!!
-XXX.
-The last section concludes.
-
-%\noindent More generally,  Among the closest to our own work, Eaton et al.
-(2002) develop a Ricardian general equilibrium trade model to estimate parameters relating to absolute advantage, to comparative advantage, and to geographic barriers.
-Anderson and Wincoop (2003) elaborate a theory-founded gravity equation to assess the cost of trade frictions, and use it to provide a direct measure of the trade losses due to national borders (between 20 and 50 percent).
-
-\section{Data Sources and Empirical Methodology}
-
-\subsection{A measure of Transportation Costs}
-
-\noindent
-
-Our strategy for representing transportation costs is relatively standard, and consists in exploiting the difference between commodity-level export and prices.
-We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, origins and transportation mode.
-More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs.
-The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price.
-Strictly higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.
-
-To have the import price of the good (CIF price), next to the export price (FOB price) is highly valuable.
-Indeed, this allows us thus to estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs.
-This differentiates us from \citet{Irrazabal_2015}, that can only estimate the ratio of additive costs as a share of the tola price.
-However, note that this measure based on \emph{freight} cost ignores another dimension of transportation costs related to the time value of goods in transit.
-According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
-
-\subsection{Database}
-
-%laius sur pourquoi pas de donnes de firmes (triop comutational), ni comtrade (trop mauvais, citer RIE) XXX
-
-\noindent Even if the data of transportation costs we rely could be available at the firm-level, the use of a nonlinear estimator (see below) triggers computational limitations that do not make them a likely option.
-A product-level bilateral database of CIF-FOB ratios is theoretically possible, and could be built from the well-known UN's COMTRADE data.
-Unfortunately, \citet{hummels_rie} show that issues inherent to the construction of data (especially the use of ``mirror'' flows) make this database unreliable for our purpose.
-
-We decide therefore to use database for a single country, but with strong statistical reliability arising from a single, trustworthy customs origin.
-More precisely, the empirical analysis relies on yearly data from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
-Based on customs declarations, this dataset inventories all imports (both values and quantities) by origin to the United states at the HS 10-digit highly disaggregated level, with a concordance code to the SITC 5-digit coding system.
-In addition, the database reports information regarding freight expenditures and transportation mode (Ocean Vessel and Air).
-The first will be crucial to compute transport costs (see below), the second will alow us enlightening potential different dynamics of transport costs across transportation mode.
-
-Due to the abovementioned computational limitations, data are reaggregated at the 3- and 4-digit level: comparing different levels of aggregation will be useful to check differences and the presence of biases precisely due to aggregation.
-Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products.
-
-% Why the gap CIF-FOB is a good measure of trade costs (cf mimeo by Gaulier and Zignago)?
-JH: to me, This is not the right question to ask.
-We have the freight costs in our data.
-The true question is: how well these freight costs proxy correctly transportation costs as a whole.
-
-\subsection{Key empirical specification}
-
-\noindent Our purpose is to provide estimates over time of the relative shares of multiplicative and additive costs among total transport costs.
-To do so, we start from a very simple equation, quite alike the one proposed by \citet{hummels2010}:
-
-\begin{equation}\label{eq:base}
-p = \tau \widetilde{p} + t
+\begin{equation}
+p^{cif}_{ikt} q_{ikt} = p^{fob}_{ikt} q_{ikt} \tau_{ikt} + t_{ikt} q_{ikt} \label{eq:cif-fob}
 \end{equation}
 
-or as a ratio
+This equation states that the value of an import by the US, from country $i$, of product $k$ in year $t$ ($p^{cif}_{ikt} q_{ikt}$), is equal to the value of the good exported ($p^{fob}_{ikt} q_{ikt}$), raised by a given multiplicative cost $\tau_{ikt}$, plus an additive cost component, that depends on the quantity of the good exported $ t_{ikt} q_{ikt}$.
+Manipulating the equation, we get:
 
-\begin{equation}\label{eq:base_bis}
-\frac{p}{\widetilde{p}} = \tau  + \frac{t}{\widetilde{p}}
+\begin{equation}
+\frac{p^{cif}_{ikt}}{p^{fob}_{ikt}} = \tau_{ikt} +\frac{t_{ikt}}{p^{fob}_{ikt}}
+\label{eq:cif-fob2}
 \end{equation}
 
-This equation relates the ratio of the consumer price (ie, cif price, denoted $p$) to the producer price (ie, fob price, denoted $\widetilde{p}$) to the two types of costs (multiplicative and additive).
-Iceberg trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs) while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs).
-%\begin{equation*}
-%p =\tau \widetilde{p}+t
-%\end{equation*}
-We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.
-Let us denote $i$ the origin country and $k$, the sector.
-Hence, the equation to be estimated is given by:
-
-\begin{equation}\label{eq:base_estimee}
-\frac{p_{ik}}{\widetilde{p}_{ik}} = \tau_{ik}+\frac{t_{ik}}{ \widetilde{p}_{ik}}
-\end{equation}
-
-\section{Decomposing the Dynamics of Transport Costs over Time}
-
-\subsection{Estimation Strategy}
-
-% nonlinear least squares: problem with heteroskedasticity???
-
-We transpose the approach by \citet{Irrazabal_2015} by considering that both multiplicative and additive costs are separable between origin country and products, in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau$ and $t$ from equation~\ref{eq:base_estimee} become:
-
-%\begin{eqnarray}
-%\tau_{ik} &=& dist^{\rho_I}F^I_i \label{eq:iceberg}\\
-%t_{ik} &=& dist^{\rho_A}F^A_i-1 \label{eq:add}
-%\end{eqnarray}
-
-\begin{eqnarray}
-\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:iceberg}\\
-t_{ik} &=& t_{i} + t_{k} \label{eq:add}
-\end{eqnarray}
-
-As a result, our underlying theoretical equation is specified as:
-
-%\begin{equation}
-%\frac{p_{ik}}{\widetilde{p}_{ik}} =dist^{\rho_I}F^I_ik+\frac{dist^{\rho_A}F^A_ik-1}{ \widetilde{p}_{k}}\label{eq:estimatedequation}
-%\end{equation}
-
-\begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} =\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
-\end{equation}
-
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the \emph{cif} price $p$ cannot be lower than the \emph{fob} price: $p>\widetilde{p}_{ik}$.
-Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term.
-More precisely:
-
-\begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right)\times \varepsilon_{ik}
-\end{equation}
-
-Log-linearization gets the following, tractable empirical equation:
-
-\begin{equation}
-ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
-\end{equation}
-where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= ln(\varepsilon_{ik})$ .
-
-Still following \citet{Irrazabal_2015}, we report a ``synthetic estimate'' of each type of transport cost, by averaging all $\widehat{\tau}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}_{ik} = t_{i} + t_{k}$ using the values of each trade flow over total yearly trade as a weighting scheme.
-Therefore, we recover for each year and transportation mode a $\widehat{\tau}$ and $\widehat{t}$ that will be the results we will report for step 1.\\
-
-\subsection{Econometric Issues}
-
-\noindent Our primary interest relates to the importance of additive costs relatively to iceberg costs.
-Put differently, what traditional models of international trade lose by ignoring additive costs?
-A natural way to answer this question is to perform estimations of equation~\ref{eq:estimatedequation} constraining $t$ to be equal to zero, and compare the fitting properties/explanatory power of the restricted and complete models.
-This is done bye computing several standard diagnostic statistics (R$^{2}$ and statistics based on the likelihood function).
-
-The shape of our main equation \ref{eq:estimatedequation} is such that estimations cannot be performed using standard linear estimators.
-Therefore, all estimates are performed using non linear least squares.
-The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
-The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
-In our case, due to computational limitations implied by the size of our dataset, we implement 100 iterations and set the convergence criterion for successive parameter estimates and for the residual sum of squares at 0.01.
-
-Finally, to control for the potential influence of outliers, we excluded observations in the 5 percent from the upper and lower tails of the distribution in the regression variables, and all our three measures of trade costs are bounded by 0 as minimal value.
-These cut-offs are aimed at eliminating reporting or coding errors.
-%(XXX develop a little bit?XXX)
-
-%\noindent with the exponents $I$ and $A$ respectively identifying the iceberg and the additive trade costs, $\rho_I$ and $\rho_A$ being the parameters of interest to be estimated.
-The adding of $-1 $ in the specification for the additive costs (\ref{eq:add}) is necessary to be consistent with $t$ starting from 0.
-In the absence of additive trade costs, ie $\rho_A=F^A_i=F^A_j=0$, the specification (\ref{eq:add}) consistently implies $t=0$.
-As well, if iceberg costs are null ($\rho_I=F^I_i=F^I_j=0$), the specification (\ref{eq:iceberg})) correctly implies $\tau=1$.
-
-\subsection{Results: Distinguishing Additive and Multiplicative Parts of Trade Costs}
-
-\subsubsection{Goodness-of-fit}
-
-Here, we proceed to a pure statistical analysis of our results, i.e., we investigate the properties of our two models (with and without an additive term) in terms of goodness-of-fit.
-Our idea is to assess what loss of explanatory power is implied by the omission of the additive term.
-We distinguish between ``Air'' and ``Vessel'', in order to see if and how the quality-of-fit varies across transports modes.
-We start with a simple visual inspection of Figure \ref{fig:good_fit}, where are reported the yearly median \emph{ad valorem} estimates of the iceberg costs ($\widehat{\tau}_{i}$) estimated in equation \ref{eq:estimatedequation}\footnote{Qualitatively similar results are obtained when the mean values are reported instead.
-More details of those results available upon request to the authors.}.
-The dotted line plots the estimated iceberg cost ($\widehat{\tau}_{i}$) when the additive term $t$ is restricted to 0, and the plain dark line represents the iceberg cost in the model including a non-null additive term.
-In both cases, multiplicative costs are reported minus one, for a better reading comfort and comparability with Figure \ref{fig:mult_vs_add}.
-
-For both transport modes, it appears that the omission of the additive term seriously biases the multiplicative term upward.
-For Air, the size of the bias seems to decrease over time, while it seems pretty stationary for Vessel.
-In any case, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
-On average over the whole period, biased estimate for iceberg is 1.04 for Air and 1.05 for Vessel, while unbiased estimates, are, respectively 1.02 and 1.03.
-
-In order to deliver a more systematic diagnosis, we use several standard measures of fit.
-Of course R$^{2}$ comes naturally to mind, but its use is far from being straightforward when evaluating non-linear estimates.\footnote{R-squared is based on the underlying assumption that the adjusted model is a linear one.
-In a non-linear context, R-squared is therefore inappropriate, strictly speaking.
-However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.)}.
-Therefore, we provide as a complement the Standard Error of Regression (SER) which represents the average distance that the observed values fall from the regression line.
-Smaller values are better because it indicates that the observations are closer to the fitted line.
-We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test.
-A decrease in the log-likelihood function points to a better quality-of-fit.
-However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic, so that AIC stat $= 2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $.
-Once again, the preferred model is the one with the minimum AIC value.
-Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, see equation~\ref{eq:estimatedequation}) and the Restricted one (\emph{R}, i.e.
-equation~\ref{eq:estimatedequation} with $t=0$).
-The null tested is that the two models are statistically equivalent.
-Results are reported in Tables~\ref{tab:good_fit_air} and \ref{tab:good_fit_ves}.
-
-%If we keep the R² as a measure of GoF, we will have to provide some evidence of normality for residuals. In addition, a plot of observations and the adjustment curve might be useful.
-Unsurprisingly, it appears that the inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or transport mode.
-On average over the whole period, $R^{2}$ doubles when restricting to Air Transport, and increases by 50\% for Vessel.
-Similar qualitative conclusions arise from the comparisons of SERs, but may be more reliable on the quantitative side, since the use of R-squared on non-linear estimates may be qualified.
-Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes.
-Both AIC and Log-Likelihood statistics decrease with the inclusion of the additive term, and the LL test unambiguously reject the null of statistical equivalence of the two models.
-This is true whatever the considered year.
-
-However, going into the examination of the levels of these goodness-of-fit statistics reveals intriguing differences.
-Multiplicative costs appear more important for the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector; the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air.
-Besides, the dynamics over time also reveals intriguing differences.
-Quality of fit decreases strongly after 2000, for both transport modes.
-For Air, it seems that the additive term explains less and less variance over time: in 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the SER by 14 percentage points ;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
-The picture is quite different for Vessel.
-For that transport mode, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years.
-It appears that the deteriorating performance of the model in terms of goodness-of-fit from 2000 comes mainly from the decreasing performance of the iceberg term.
-In order to enlighten the roots of these differences, the following section elaborates further on the quantitative dynamics of each type of cost.
-
-For comparison purposes, we provide in the appendix similar results for some years based on a more disaggregated dataset at the product level (4-digits).\footnote{Since we use a non-linear estimator, estimations are highly computer-intensive.
-This explains why we decided to focus on the 3-digit product classification, which requires less time-consuming calculations.}.
-If anything the quality of fitting appears slightly higher when estimations are based on the 4-digits classification.
-This is especially true for the model restricting trade cost to their iceberg dimension, whatever the transport mode considered.
-When the additive part is taken into account however, the difference becomes very small, whatever the considered criterion.
-
-Using a more disaggregated classification unsurprisingly adds some statistical precision, but not to an extent which disqualifies the use of slightly more aggregated data.
-% need for SER on all years to be complete.
-
-\subsubsection{Interpretation}
-
-We now more specifically investigate the long-run behavior of the estimated iceberg and additive parameters.
-Tables\label{tab:result_air} and \label{tab:result_ves} report, respectively for the Air and Vessel sector, our estimates for Iceberg trade costs, first from equation \label{eq:iceberg} (upper panel), then from equation \label{eq:estimatedequation} where they are estimated jointly with an additive term, which estimate is also reported (bottom panel).
-Remember that estimation is performed on a yearly basis, consequently, for the sake of clarity, we do not report estimates for each year, but for the beginning and the end of our sample, as well as starting year for each decade.
-To complement these tables, Figure\label{fig:mult_vs_add} reports the estimates of equation \label{eq:estimatedequation}, this time for each year: the dotted line plots estimated for the additive term, and the plain, dark line plots the iceberg cost (-1, for a better reading comfort).
-As we did for the quality-of-fit, we report in Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} in the Appendix the estimated for coefficients for products disaggregated at the 4-digits level.
-In any case, these estimates are very close, and do not show any major difference with the baseline ones.
-
-Unsurprisingly, the trend in transportation costs is clearly decreasing whatever the mode (air or vessel) or measure (with or without additive term) we consider.
-Starting from 1974, the decrease appears more important for sea than air transportation: transportation costs are divided by three for the first mode, and two for the second one.
-That said, it is widely know the two oil shocks inflated artificially transportation costs that specific year.
-If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for Air, and 2-3 percentage points for Vessel - depending if we consider mean or median estimates.
-These figures are consistent with the ones by Hummels (2007).
-However, our methodology allows distinguishing in this trend, what comes from multiplicative and additive costs.
-To make a long story short, it appears that the additive costs represent the bulk of the decrease for Air, whereas the multiplicative component is the main responsible of the decreasing trend for Vessel, except for the years 2012-2103, witnessing an abrupt decrease in additive costs.
-To give a few average numbers, starting from 1980 and ending in 2013, the iceberg term decreases by 0.6 percentage point (pp), and the additive term is reduced by 1 pp for Air; for Ocean shipping, the reduction is roughly 2 pp for additive costs, and 1 pp for multiplicative costs.
-
-%Itrrazbal
-
-%Composition effects ??
-
-In section \ref{sec:determinants}, we provide rationales for those differences.
-
-\section{Reassessing the Traditional Determinants of Transport Costs}\label{sec:determinants}
-
-\subsection{Methodological Concerns}
-
-%In step 2, we ask if the traditional determinants of transportation costs impact differently the multiplicative and additive costs estimated in the first step.
-%Therefore, we consistently focus only on origin-country-specific transportation costs, taking the estimated $\tau_{i}$ and $t_{i}$ from step 1 as dependent variables.
-In line with the trade literature, we proxy transport costs by (log of) distance and other proxies for trade costs (mainly, geographic and cultural factors).\interfootnotelinepenalty=10000\footnote{All those variables come from the well-known GeoDist database built by the CEPII: \url{http://www.cepii.fr/CEPII/fr/bdd_modele/presentation.asp?id=6}.} More formally, we estimate, still on a yearly basis and for each transportation mode:
-
-%\begin{eqnarray}
-%ln(\widehat{\tau}_{i}) &=& \beta_{1} \textrm{ln(distance)} + \Gamma_{1} \textrm{Gravity} + \mu_{i} \label{eq:step2_iceberg}\\
-%ln(\widehat{t}_{i}) &=& \beta_{2} \textrm{ln(distance)} + \Gamma_{2} \textrm{Gravity} + \nu_{i} \label{eq:step2_additive}
-%\end{eqnarray}
-
-%where the ln operator denotes that variables have been log-linearized, $distance$ is the bilateral distance between US and the origin country $i$  and $Gravity$ is a vector of controls including the languages spoken in the country under different definitions, whether the country is landlocked, and their colonial links.
+In the first step, we decompose the gap between the cif and the fob price between its two multiplicative and additive components.
+In the second stage, we raise the question, what is behind the gap between the fob and the cif prices?
+As clear from Equation (\ref{eq:cif-fob}), the multiplicative cost is tied to the price of the good exported (the fob price), while the additive cost is rather applied on the quantity exported.
+This drives us to refine the above question: What is behind each component of the cif-fob gap?
+In particular, do the gravity variables, primarily distance, that are commonly used as a proxy for iceberg trade costs, really matter on iceberg (i.e., multiplicative) transport costs, or rather on the additive part?
 
@@
 
-\paragraph{Insurance}: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ???
+\begin{itemize}
+\item Insurance: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ???
 for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood).
@@
 
-\paragraph{Costs} (for handling costs).
-The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs).
+\item Costs (for handling costs).
+The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping (A ton of cars does not represent the same volume on board as a ton of apparel); further, handling costs also depend on the country of origin (labor and various other costs).
 If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs.
-We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD.
-As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$).
-Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular).
-We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following.
-Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$.
+We capture this dimension by including fixed effects by product category.
+In doing so, we adopt the same product categorization as for insurance costs, for similar reasons (it is very unlikely that the handling costs varies within the textile category).
+Last, handling costs are likely to be country-dependent, which we capture by including country fixed effects ($FE_i$).
+Denoting the fixed effects associated to handling costs by $\widehat{FE}_{k^\prime}$, we accordingly measure total handling costs (in USD) by $FE_i\times\widehat{FE}_{k^\prime}\times q_{ik}$.
 As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.
 
- \paragraph{Freight}.
+\item Freight.
 Quite intuitively, freight costs should depend on the distance between the origin country and the U.S.
@@
 Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs.
-Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$.
+Accordingly, we capture freight costs through $\widetilde{FE}_{k^\prime}\times dist_i\times q_{ik}$.
 Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.
+
+\end{itemize}
 
@@
 
-$$TC_{ik} =  \underbrace{EC_i\times V_{k}\times q_{ik}}_{\text{Costs}} +\underbrace{FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}}_{\text{Insurance}}+ \underbrace{V_k\times dist_i\times q_{ik}}_{Freight}$$
+$$TC_{ik} =  \underbrace{FE_i\times\widehat{FE}_{k^\prime}\times q_{ik}}_{\text{Costs}} +\underbrace{FE_{k^\prime \prime}\times p_{ik}^{fob}\times q_{ik}}_{\text{Insurance}}+ \underbrace{\widetilde{FE}_{k^\prime}\times dist_i\times q_{ik}}_{Freight}$$
 
@@
 \begin{eqnarray*}
-&&(p^{cif}_{ik} - p^{fas}_{ik})q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_{k}\times dist_i\times q_{ik} \\
-\Leftrightarrow && \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)p^{fas}_{ik} q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_k \times dist_i\times q_{ik}
+&&(p^{cif}_{ik} - p^{fas}_{ik})q_{ik} = FE_i\times\widehat{FE}_{k^\prime}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ \widetilde{FE}_{k^\prime}\times dist_i\times q_{ik} \\
+\Leftrightarrow && \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)p^{fas}_{ik} q_{ik} = FE_i\times\widehat{FE}_{k^\prime}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ \widetilde{FE}_{k^\prime}\times dist_i\times q_{ik}
 \end{eqnarray*}
@@
 \begin{equation}
-tc_{ik} = \underbrace{\frac{EC_i\times V_{k}}{p^{fas}_{ik}}}_{\text{C}} + \underbrace{FE_{k^\prime}}_{\text{I}}+ \underbrace{\frac{V_k\times dist_i}{p_{ik}^{fas}}}_{\text{F}} \label{eq:theory}
+tc_{ik} = \underbrace{\frac{FE_i\times\widehat{FE}_{k^\prime}}{p^{fas}_{ik}}}_{\text{C}} + \underbrace{FE_{k^\prime}}_{\text{I}}+ \underbrace{\frac{\widetilde{FE}_{k^\prime}\times dist_i}{p_{ik}^{fas}}}_{\text{F}} \label{eq:theory}
 \end{equation}
@@
 \begin{equation}
-tc_{ik} = \left[\frac{EC_i\times V_k}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k\right]e^{\varepsilon_{ik}} \label{eq:estim}
+tc_{ik} = \left[\frac{FE_i\times\widehat{FE}_{k^\prime}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{\widetilde{FE}_{k^\prime}\times dist_i}{p_{ik}^{fas}} + FE_i\right]e^{\varepsilon_{ik}} \label{eq:estim}
 \end{equation}
@@
 This calls for two comments.
-First, we add the country fixed effects separately in the regression, in order to correctly identify the associated interaction term.
+First, we add the country fixed effects seperately in the regression, in order to correctly identify the associated interaction term.
 Second, we specify the error term multiplicatively.
@@
 \begin{equation}
-\log tc_{ik} = \log\left[\frac{EC_i\times V_{k}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k \right]+ \varepsilon_{ik} \label{eq:estim1}
+\log tc_{ik} = \log\left[\frac{FE_i\times\widehat{FE}_{k^\prime}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{\widetilde{FE}_{k^\prime}\times dist_i}{p_{ik}^{fas}} + FE_i\right]+ \varepsilon_{ik} \label{eq:estim1}
 \end{equation}
 
-%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
+\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
 So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ...
@@
 
-%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !
+\textbf{Questions raised} \medskip
+\begin{itemize}
+\item The literature commonly uses gravity variables (such as distance, common language, colonial linkages, common border, etc.) as a proxy for trade costs.
+In the second-step estimation, we only consider distance.
+This choice relies on the fact that we explain the transport gap measured by the gap between the fas price and the caf price.
+Our measure of transport cost is only part of the larger size of trade costs, and there is no a priori reason why the other gravity variables except distance might intervene in the difference between the price at the export departure point and the import arrival point.
 
-\subsection{Results and Interpretation}
-%\section{One step further into interpretation: transportation costs as an indicator of market structure.}
+\item Make a difference between the estimated additive trade cost and the multiplicative trade costs.
+Ie, should we suppose that the explicative variables enter additively when considering the additive trade costs, and multiplicatively regarding the iceberg component?
+No a priori reason.
+In both cases, the dependent variable is expressed as a percentage (of the fob price).
+We decompose this percentage in various dimension (the cost of oil, the cost of distance, etc.) , which all added, compose the overall trade cost component.
+This justifies an additive specification, for both the additive and the iceberg trade costs components.
 
-\section{Conclusion}
-
-\newpage
-\bibliographystyle{essaien}
-\bibliography{biblio}
-
-\newpage
+\end{itemize}
 
@@
 
-\section*{Data Appendix \label{app:data}}
+\section{Data Appendix \label{app:data}}
 
@@
 
-\subsection{Product classification for Insurance fixed effects}
+\subsection{Product classification for C,I, F fixed effects}
 \textbf{to be done}
 
-\subsection{Volumetric weight and the Maritime Transport Costs database}
-The Maritime Transport Costs (MTC) database built by the OECD\footnote{See \texttt{http://stats.oecd.org/Index.aspx?datasetcode=MTC}.} contains data from 1991 to the most recent available year of bilateral maritime transport costs.
-Transport costs are available for 43 importing countries (including EU15 countries as a custom union) from 218 countries of origin at the detailed commodity (6 digit) level of the Harmonized System 1988.
-
-At commodity level (6 digit HS), maritime transport costs are displayed using three measures:
-\begin{itemize}
-\item Transport cost: the total cost expressed in USD (insurance plus freight) of transporting all of the given product during the given year,
-\item  Unit transport cost: transport cost per kilogramme, or in other words, the cost in USD required to transport one kilogramme of merchandise, and
-\item Ad valorem equivalent: transport cost divided by the  total import value, i.e. the share transport cost represents in the total import value of the product.
-\end{itemize}
-At chapter level (2 digit HS), the transport costs are the sum of the corresponding 6 digits commodities.
-The unit cost and ad valorem cost are then the transport costs at chapter level divided respectively by sum of the weights at 6 digits, and by the sum of the import values at 6 digits.
-
-\subsection{Country-specific handling costs and the ``Cost to export'' variable}
-
-Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport.
-The most recent round of data collection for the project was completed in June 2014.
-Among the four variables provided, we retain the variable ``Cost to export'' (US\$ per container), as it represents the ``cost associated with all procedures required to export goods[...,] including the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' We use the variable observed \textbf{in year ???
-attention la variable est-elle destination spefici?
-A priori non, a verifier.}
-\newpage
-
-\section*{Tables and Figures}
-
-%%%Figures
-% Goodness-of-fit
-
-\begin{figure}[htbp]
-\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-%Economic interpretation
-%usepackage{float} + H is great to keep images on the right page
-\begin{figure}[H]
-\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-%%%%Tables
-%Goodness-of-fit
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of Goodness-of-fit}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}}\\
-\hline
-\textbf{R$^{2}$} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.31} & \multicolumn{1}{c}{\textbf{0.31}} \\
-Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{\textbf{0.60}} \\
-\textbf{SER} &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & \multicolumn{1}{c}{\textbf{0.85}} \\
-Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & \multicolumn{1}{c}{\textbf{0.73}} \\
-\textbf{Log-likelihood} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & -17530.49 & -20253.49 & -29977.79 & -43341.27 & \multicolumn{1}{c}{-50746.83} & \multicolumn{1}{c}{-52690.29} & \multicolumn{1}{c}{\textbf{-34888.64}} \\
-Terms A \& I & -15125.65 & -17263.20 & -25393.46 & -36788.44 & \multicolumn{1}{c}{-47277.53} & \multicolumn{1}{c}{-49419.70} & \multicolumn{1}{c}{\textbf{-30508.29}} \\
-\textbf{AIC criteria} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{106130.58} & \multicolumn{1}{c}{\textbf{70498.08}} \\
-Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{100155.41} & \multicolumn{1}{c}{\textbf{62284.99}} \\
-\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       &  \\
-2$\times$(ll(UR) -ll(R)) & 4809.68 & 5980.59 & 9168.67 & 13105.67 & \multicolumn{1}{c}{6938.61} & \multicolumn{1}{c}{6541.17} & \multicolumn{1}{c}{\textbf{8760.69}} \\
-\# restrictions  & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426.00} & \multicolumn{1}{c}{422} & \multicolumn{1}{c}{\textbf{401.93}} \\
-p-value & 0.000 & 0.000 & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{\textbf{}} \\
-
-\hline\hline
-   % \bottomrule
-
-    \end{tabular}%
-    \end{center}
-  \label{tab:good_fit_air}%
-  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel, Measures of Goodness-of-fit}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \textbf{Mean stat} \\
-\hline
-\textbf{R$^{2}$}\unboldmath{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &  \\
-Term I only & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{0.34} & \multicolumn{1}{c}{\textbf{0.39}} \\
-Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.56}} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.76} & \multicolumn{1}{c}{\textbf{0.66}} \\
-Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{\textbf{0.57}} \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{-16287.40} & \multicolumn{1}{c}{-16129.13} & \multicolumn{1}{c}{-25169.31} & \multicolumn{1}{c}{-35263.95} & \multicolumn{1}{c}{-41998.95} & \multicolumn{1}{c}{-43692.93} & \multicolumn{1}{c}{\textbf{-28534.30}} \\
-Terms A\& I & \multicolumn{1}{c}{-12985.76} & \multicolumn{1}{c}{-13353.65} & \multicolumn{1}{c}{-21171.37} & \multicolumn{1}{c}{-29490.96} & \multicolumn{1}{c}{-37418.66} & \multicolumn{1}{c}{-39751.86} & \multicolumn{1}{c}{\textbf{-24151.31}} \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{33328.81} & \multicolumn{1}{c}{33010.27} & \multicolumn{1}{c}{51142.62} & \multicolumn{1}{c}{71365.89} & \multicolumn{1}{c}{84789.89} & \multicolumn{1}{c}{88191.87} & \multicolumn{1}{c}{\textbf{57848.60}} \\
-Terms A \& I & \multicolumn{1}{c}{27331.52} & \multicolumn{1}{c}{28067.31} & \multicolumn{1}{c}{43664.74} & \multicolumn{1}{c}{60475.91} & \multicolumn{1}{c}{76161.33} & \multicolumn{1}{c}{80873.72} & \multicolumn{1}{c}{\textbf{49683.32}} \\
-\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{\textbf{}} \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & \multicolumn{1}{c}{9160.56} & \multicolumn{1}{c}{7882.15} & \multicolumn{1}{c}{\textbf{8765.97}} \\
-\# restrictions  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{427} & \multicolumn{1}{c}{\textbf{417.05}} \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-
-\hline\hline
-   % \bottomrule
-
-    \end{tabular}%
-    \end{center}
-  \label{tab:good_fit_ves}%
-  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of Regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-%Economic Interpretation
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034 & \multicolumn{1}{c}{\textbf{1.051}} \\
-Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029 & \multicolumn{1}{c}{\textbf{1.042}} \\
-Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
-\hline
-\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010 & \multicolumn{1}{c}{\textbf{0.019}} \\
-\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005 & \multicolumn{1}{c}{\textbf{0.007}} \\
-\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
-\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017 & \multicolumn{1}{c}{\textbf{1.025}} \\
-\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 & \multicolumn{1}{c}{\textbf{1.018}} \\
-Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
-\hline
-\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036 & \textbf{1.058} \\
-Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033 & \textbf{1.051} \\
-Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
-\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-\hline
-\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
-Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015 & \textbf{0.029} \\
-Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 & \textbf{0.019} \\
-Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 & \textbf{1.032} \\
-Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018 & \textbf{1.028} \\
-Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
-\hline
-\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\newpage
-
-\appendix
-
-\section{Results with a Higher Level of Products Disaggregation }
-%\vspace{-0.5cm}
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of Goodness-of-fit, 4-digits}
-\begin{center}
-\label{tab:good_fit_air_rob}%
-%\vspace{0.5cm}
-\scalebox{0.97}{
-\begin{tabular}{lccccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                      &  \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\
-\hline
-\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} & \textbf{0.47} \\
-Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} & \textbf{0.63} \\
-\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93} & \textbf{0.89} \\
-Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86} & \textbf{0.79} \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87} & \textbf{-34744.40} \\
-Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14} & \textbf{-30243.95} \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} & \textbf{70940.07} \\
-Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28} & \textbf{62955.73} \\
-\textbf{Test LL} &       &       &       &       &       &       & \textbf{} \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45} & \textbf{9000.89} \\
-\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818} & \textbf{755.73} \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \textbf{} \\
-\hline\hline
-\end{tabular}}%
-\end{center}
- \vspace{-10pt} \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-\begin{table}[H]
-  \centering
-  \caption{Vessel: Measures of Goodness-of-fit, 4-digits}
-\begin{center}
-\label{tab:good_fit_ves_rob}%
-
-\scalebox{0.97}{\begin{tabular}{lccccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                      &  \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\ \hline
-\boldmath{}\textbf{R$^{2}$}\unboldmath{} &       &       &       &       &       &       &  \\
-Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{\textbf{0.44}} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.59}} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} & \multicolumn{1}{c}{\textbf{0.77}} \\
-Terms A \& I &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} & \multicolumn{1}{c}{\textbf{0.68}} \\
-\textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} & \multicolumn{1}{c}{\textbf{-29883.62}} \\
-Terms A\& I & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22} & \multicolumn{1}{c}{\textbf{-25303.92}} \\
-\textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23} & \multicolumn{1}{c}{\textbf{61425.60}} \\
-Terms A \& I & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} & \multicolumn{1}{c}{\textbf{53573.29}} \\
-\textbf{Test LL} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.75 & \multicolumn{1}{c}{17354.65} & \multicolumn{1}{c}{20113.52} & \multicolumn{1}{c}{16608.16} & \multicolumn{1}{c}{12589.59} & \multicolumn{1}{c}{15704.63} \\
-\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} & \multicolumn{1}{c}{860} \\
-p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-
-\hline\hline
-\end{tabular}}
-\end{center}
-  \vspace{-10pt}\tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
-
-\hline
-\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       &  \\
-Mean  & 1.066 & 1.058 & 1.052 & \multicolumn{1}{c}{1.033} & \multicolumn{1}{c}{1.037} & \multicolumn{1}{c}{1.032} & \multicolumn{1}{c}{\textbf{1.049}} \\
-Median & 1.052 & 1.044 & 1.041 & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.027} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{\textbf{1.037}} \\
-Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{\textbf{0.045}} \\
-\hline
-\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{} \\
-\textit{Additive term} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} & \multicolumn{1}{c}{\textbf{0.019}} \\
-Median & 0.012 & 0.006 & 0.006 & \multicolumn{1}{c}{0.005} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{\textbf{0.008}} \\
-Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{\textbf{0.034}} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016} & \multicolumn{1}{c}{\textbf{1.024}} \\
-Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014} & \multicolumn{1}{c}{\textbf{1.016}} \\
-Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{\textbf{0.026}} \\
-\hline
-\# obs & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460} & \multicolumn{1}{c}{\textbf{}} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air_rob}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\begin{table}[htbp]
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-  \centering
-\caption{Vessel: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
-\hline
-\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.098 & 1.061 & \multicolumn{1}{c}{1.058} & \multicolumn{1}{c}{1.051} & \multicolumn{1}{c}{1.042} & \multicolumn{1}{c}{1.036} & \textbf{1.060} \\
-Median & 1.094 & 1.051 & \multicolumn{1}{c}{1.048} & \multicolumn{1}{c}{1.045} & \multicolumn{1}{c}{1.038} & \multicolumn{1}{c}{1.031} & \textbf{1.052} \\
-Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} & \textbf{0.036} \\
-\hline
-\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{}\\
-\textit{Additive term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015} & \textbf{0.028} \\
-Median & 0.029 & 0.013 & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{0.013} & \multicolumn{1}{c}{0.008} & \textbf{0.017} \\
-Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} & \textbf{0.039} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.054 & 1.034 & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.021} & \textbf{1.033} \\
-Median & 1.049 & 1.030 & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.025} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{1.018} & \textbf{1.028} \\
-Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013} & \textbf{0.025} \\
-\# obs & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} & \textbf{}\\
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves_rob}%
-\vspace{-10pt}\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
 \end{document}

### 2016-07-20 redaction/poubelle/note_secondstage_v3011.tex
Rédaction dans le github

--- 2016-07-20 redaction/poubelle/note_secondstage.tex
@@
 
-In this second stage regression, the key question is: Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs?
-Or total?
-Usually, people use a measure of distance as a way to capture iceberg trade costs.
-Using our decomposition of transport costs, we want to assess the contributions of such proxies to each component (additive / iceberg).
-What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs?
-
-This question drives our second-step analysis.
-Consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and found that this first reasonably well, the estimated coefficient stands in the average.
-We then apply the same regression on each multiplicative / additive component of trade costs.
-Suppose that we find that we find a high correlation with the additive cost, almost 0 with the multiplicative cost.
-This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error.
-
-To investigate this point, we thus regress 1°) the ``overal trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{iy}^{nlI}$ hereafter ($i$ for the origin country, $y$ for the estimation year and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{iy}$ and $t_{iy}$ respectively, on some common determinants.
-\medskip
-
-Precisely, what are these dependent variables?
-We use our first-stage regression results, from which we extract the fixed effect of transport costs associated to the country dimension (recall that we assumed that the iceberg trade cost from country $i$, good $k$, year $y$, $\tau_{iky} = \tau_{iy}\tau_{ky}$, and  similarly for the additive transport cost $t_{ik} = t_{iy}+ t_{ky}$.
-Both are expressed in percentage (of the fob price).
-
-\textbf{Careful}: This supposes some re-treatment of our first-stage estimation results (recalling our general estimated equation):
-$$p_{iky}^{cif} = \left[\widehat{\tau}_{iy}\widehat{\tau}_{ky}\right] p^{fob}_{iky} +  \left[\widehat{t}_{iy}\widehat{t}_{ky}\right]$$
-
-From this, we re-treat the iceberg component (both estimated alone and with the additive part) to consider $\tau_{iy} =  100(\widehat{\tau}_{iy}-1)$ (leaving aside the ``product'' dimension as unrelated to the question here), to have the (country-specific) iceberg trade cost expressed in percentage of the fob price.
-
-As for the additive cost, we have obtained the value of (country-specific) trade cost as a fraction of the fob price, that we re-scale to consider it in percentage, ie considering $t_{iy} =  100 t_{iy}$, so that we have the additive cost in percentage of the fob price to explain.
-\medskip
-
-We use data from 2004 to 2013, because we have information provided by Doing Business about exports formalities on an annual basis on this period.\bigskip
-
-Extracted from the Doing Business website:
-
-Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport.
-The time and cost necessary to complete 4 predefined stages (document preparation; customs clearance and inspections; inland transport and handling; and port and terminal handling) for exporting and importing the goods are recorded; however, the time and cost for sea transport are not included.
-All documents needed by the trader to export or import the goods across the border are also recorded.
-The most recent round of data collection for the project was completed in June 2014.
-
-In particular, we are interested in the variable ``Cost to export'' (US\$ per container).
-From DB: It is the ``cost associated with all procedures required to export goods.
-Includes the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' \textbf{Careful}: inland transport, transport between the warehouse (assumed to be in the main economic city of the country) and the port.
-So, this part of the export formality cost is not in the cif-fob price, because the fob price already includes the cost incurred to bring the shipment from the initial warehouse to the boat (even includes handling at the port to bring the merchandises on the boat itself).
-So, there is a measurement error, the export formality cost is higher than it should be; a priori, bias the associated coefficient downwards.
-
-\textbf{What do we call ``fob'' price exactly?} In the US import data we use, the cif price is based on applying import charges on the custom value.
-What is defined as import charges?
-As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
-import duties) incurred in bringing the merchandise from alongside the carrier''.
-So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''.
-This means that the seller has paid to bring the goods to the port, but not to put them on the boat.
-In this respect, all hauling charges are excluded from what we call the ``fob'' price.
-Our ``export formality cost'' measure is thus adequate as potential explanation of the gap between the cif price and the fob (in fact, fas) price.\bigskip
-
 One way of explaining our two-stage reasoning is to start recalling the equation at the root of the analysis:
@@
 \begin{equation}
-p^{cif}_{iy} q_{iy} = p^{fob}_{iy} q_{iy} \tau_{iy} + t_{iy} q_{iy} \label{eq:cif-fob}
+p^{cif}_{ijy} q_{ijy} = p^{fob}_{ijy} q_{ijy} \tau_{ijy} + t_{ijy} q_{ijy} \label{eq:cif-fob}
 \end{equation}
 
-This equation states that the value of an import by the US, from country $i$, in year $y$ (we neglect the product dimension here) ($p^{cif}_{iy} q_{iy}$), is equal to the value of the good exported ($p^{fob}_{iy} q_{iy}$), raised by a given multiplicative cost $\tau_{iy}$, plus an additive cost component, that depends on the quantity of the good exported $ t_{iy} q_{iy}$.
+This equation states that the value of an import by the US, from country $i$, of product $j$ in year $y$ ($p^{cif}_{ijy} q_{ijy}$), is equal to the value of the good exported ($p^{fob}_{ijy} q_{ijy}$), raised by a given multiplicative cost $\tau_{ijy}$, plus an additive cost component, that depends on the quantity of the good exported $ t_{ijy} q_{ijy}$.
 Manipulating the equation, we get:
@@
 \begin{equation}
-\frac{p^{cif}_{iy}}{p^{fob}_{iy}} = \tau_{iy} +\frac{t_{iy}}{p^{fob}_{iy}}
+\frac{p^{cif}_{ijy}}{p^{fob}_{ijy}} = \tau_{ijy} +\frac{t_{ijy}}{p^{fob}_{ijy}}
 \label{eq:cif-fob2}
@@
 
-From OECD stats: The c.i.f.
+In this second stage regression, the key question is: Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs?
+Or total?
+Usually, people use a measure of distance as a way to capture iceberg trade costs.
+Using our decomposition of transport costs, we want to assess the contributions of such proxies to each component (additive / iceberg).
+What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs?
+
+This question drives our second-step analysis.
+Consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and found that this first reasonably well, the estimated coefficient stands in the average.
+We then apply the same regression on each multiplicative / additive component of trade costs.
+Suppose that we find that we find a high correlation with the additive cost, almost 0 with the multiplicative cost.
+This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error.\bigskip
+
+To investigate this point, we thus regress 1°) the ``overall trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{ijy}^{nlI}$ hereafter ($i$ for the origin country, $y$ for the estimation year, $j$ for the product and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{ijy}$ and $t_{ijy}$ respectively, on some common determinants.
+\medskip
+
+\paragraph{Data treatment} We re-treat the iceberg component (both estimated alone and with the additive part) to consider $\tau_{ijy} =  100(\widehat{\tau}_{ijy}-1)$, to have the iceberg trade cost expressed in percentage of the fob price.
+As for the additive cost, we have obtained the value of (country-specific) trade cost as a fraction of the fob price, that we re-scale to consider it in percentage, ie considering $t_{ijy} =  100 t_{ijy}$, so that we have the additive cost in percentage of the fob price to explain.
+All our three measures of trade costs are bounded by 0 as minimal value, and we drop outliers by only keeping the 95 first percentile of each distribution.\bigskip
+
+\textbf{What do we call ``fob'' price exactly?} In the US import data we use, the cif price is based on applying import charges on the custom value.
+What is defined as import charges?
+As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
+import duties) incurred in bringing the merchandise from alongside the carrier''.
+So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''.
+This means that the seller has paid to bring the goods to the port, but not to put them on the boat.
+In this respect, all hauling charges are excluded from what we call the ``fob'' price.
+\bigskip
+
+\textbf{What do we call ``cif'' price exactly?} From OECD stats, the c.i.f.
 price (i.e.
@@
 \item Costs (for handling costs).
-1°) how to measure it: Our first measure is the ``export formality cost''  provided by the Doing Business database.
+1°) how to measure it: Our first measure is the weight of the product \textbf{to be continued}.
 Our second measure is the oil price (per USD exported) involved in the shipment, that we view as measuring the maintenance cost of the shipment (independent of the distance covered).
@@
 \textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
-So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' Even if we carefully explain that indeed, we are aware of that, the (second) counter-argument could be, ok but then what's the point...
+So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' Even if we carefully explain that indeed, we are aware of that, the (second) counter-argument could be, ok but then what's the point...\bigskip
 
@@
 \begin{itemize}
-\item Which variable to consider for export formality costs?
-In the DB database, we also have information about the \textit{1°)} number of documents to fulfill (``The total number of documents required per shipment to export goods.
-Documents required for clearance by government ministries, customs authorities, port and container terminal authorities, health and technical control agencies and banks are taken into account), \textit{2°)} the time to export (in days) (``The time necessary to comply with all procedures required to export goods.
-If a procedure can be accelerated for an additional cost, the fastest legal procedure is chosen.'') and \textit{3°)} the cost to export (deflated US\$).
-Ie, ``the Cost to export for each year is divided by GDP deflator, to take the general price level into account when benchmarking this absolute-cost indicator across economies with different inflation trends.
-In Doing Business 2015 the deflated costs are identical to the non-deflated (base year for the deflator).''
-
-\item[$\Rightarrow$] For now we have chosen the ``cost of export'' .
-Deflated or not?
-The answer is no, since we have on the left, the transport costs in percentage of the fob price, not deflated.
-Consider also the ``administrative costs'' (ie nb of documents and time delay)?, but pb of compatibility of measures?
-Answer: No.
-The reason is that the custom delay and the workload required to fill the documents are not included in the declaration fob-caf, which is what we use as measure of transport costs.
-Even if such time delay or administrative burden are likely to affect the magnitude of trade flows.
-
-\item Which other variables to consider?
-The literature commonly uses gravity variables (such as distance, common language, colonial linkages, common border, etc.) as a proxy for trade costs.
+\item The literature commonly uses gravity variables (such as distance, common language, colonial linkages, common border, etc.) as a proxy for trade costs.
 In the second-step estimation, we only consider distance.
@@
 \item Question of dimension comparability.
-On the leftwards of the equation, we have a percentage (iceberg or additive transport costs, in \% of the for price), in the country-specific dimension.
-As for the export formalities costs, in the raw DB database, expressed in US\$ per container transported.
-We want to have it expressed in terms of percentage of the fob price, to have the same dimension on both sides of the equation.
-We thus calculate a (weighted) mean fob price per country/year, and we divide the formality cost by the mean fob price, so that this variable represents the formality cost per US\$ exported.
-
-\item Oil price and distance Raw data for oil prices are from BP website (crude oil prices, US\$ per barrel).\footnote{See more details in Appendix \ref{app:data}.} The oil price is expressed in US\$ per barrel.
-To ensure comparability, we transform this variable to express it as percentage of the mean fob price (per country/year).
+On the leftwards of the equation, we have a percentage (iceberg or additive transport costs, in \% of the fob price), in the country-specific dimension.
+On the RHS: Raw data for oil prices are from BP website (crude oil prices, US\$ per barrel).\footnote{See more details in Appendix \ref{app:data}.} The oil price is expressed in US\$ per barrel.
+To ensure comparability, we transform this variable to express it as percentage of the fob price.
 We consider this variable alone (the cost of oil per USD exported) and in interaction with distance (the oil cost burden per kilometer made to export).
@@
 See below with the estimated equation.
+\textbf{To be continued.
+Pb of heteroscedasticity.
+Go to log form?}
 
@@
 \begin{eqnarray}
-TC_{iy} &=& a + \alpha dist_{i}+ \beta_1\frac{oil_{y}}{\bar{pfob}_{iy}} + \beta_2\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i} + \beta_3\frac{oil_{y}}{\bar{pfob}_{iy}}\times dist_{i}\times FE_y \notag \\
- &&+ FE_y + \gamma\frac{formalities_{iy}}{\bar{pfob}_{iy}} + \varepsilon_{iy} \label{eq:2dstage}
+TC_{ijy} &=& a + \alpha dist_{i}+ \beta_1\frac{oil_{y}}{p^{fob}_{ijy}} + \beta_2\frac{oil_{y}}{p^{fob}_{ijy}}\times dist_{i} + \beta_3\frac{oil_{y}}{p^{fob}_{ijy}}\times dist_{i}\times FE_y \notag \\
+ &&+ + \gamma\frac{weight_{jy}}{p^{fob}_{ijy}} + \varepsilon_{iy} \label{eq:2dstage}
 \end{eqnarray}
 
-with the subscript $i$ referring to the origin country, $y$ the year of estimation. Under this specification,
+with the subscript $i$ referring to the origin country, $y$ the year of estimation and $j$ the product dimension. \textbf{to be continued, how to take weight into account}
+
+Under this specification,
 \begin{itemize}
-\item $TC_{iy}$ is the transport cost, expressed as a fraction of the fob price, that varies over time and with the origin country.
-That is, for 1 US\$ exported by country $i$ to the US in year $y$, it costs $TC_{iy}$ in trade costs (specific to the transport mode $x$ ($x=$ air, vessel) and the trade cost type $z$ (iceberg, additive or total\footnote{Meaning, retaining an iceberg formulation without modeling an additive component.}, denoted $\tau_{iy}^{nlI}$)).
+\item $TC^{x,z}_{ijy}$ is the transport cost, expressed as a fraction of the fob price, that varies over time and with the origin country.
+That is, for 1 US\$ of product $j$ exported by country $i$ to the US in year $y$, it costs $TC_{ijy}$ in trade costs (specific to the transport mode $x$ ($x=$ air, vessel) and the trade cost type $z$ (iceberg, additive or total\footnote{Meaning, retaining an iceberg formulation without modeling an additive component.}, denoted $\tau_{iy}^{nlI}$)).
 \item $dist_i$ is the distance (in km) from the US,
 \item $oil_{y}$ is the oil price in US\$, per year
-\item $formalities_{iy}$ is the export formality costs in US\$, for country $i$ in year $y$ (from Doing Business database).
+\item $weight_{jy}$ is the weight of the shipment of product $j$, year $y$. 
 \end{itemize}
-As explained above, we transform the last two variables to have them expressed in percentage of the mean fob price observed for the pair (origin country, year).
-Reasoning: $oil_{y}$ is the price of the barrel (in US\$, per year).
-Then $\frac{oil_{y}}{\bar{pfob}_{iy}}$ is the price of a barrel per US\$ exported.
-The variable $\frac{formalities_{iy}}{\bar{pfob}_{iy}}$ represents the export formalities cost (in US\$) per USD\$ exported (by year-country).
+
 $FE_y$ are year fixed effects (the reference year being 2004), $\varepsilon_{iy}$ is the residual, and $a$ is a constant term.\smallskip
@@
 for each US\$ exported, that is involved in the transport cost of mode $x$, type $z$.
-Such that $\alpha dist$ represents the share of transport cost ($x,z$) (in percentage of the fob price) that is attributable to distance (alone).
+Such that $\alpha \times dist$ represents the share of transport cost ($x,z$) (in percentage of the fob price) that is attributable to distance (alone).
 It is expressed in percentage points.
@@
 The idea here is to capture the possibility that the oil cost of international trade might have changed over the years, because aircrafts and vessels have gained in energy needs, so that everything else equal the cost in oil per km covered for exports is lower in 2013 than in 2004 because the machines are less demanding.
-If this intuition is correct, we expect (the vector of) $\beta_3$ to be negative (and all the more negative as we are on recebt years).
-\item Given that $formalities_{iy}$ represents the cost of export formality per container (in USD), and $\frac{formalities_{iy}}{\bar{pfob}_{iy}}$ the export formalities of a container per USD exported, $\gamma$ can be interpreted as the number of containers per USD exported, such that $\gamma\frac{formalities_{iy}}{\bar{pfob}_{iy}}$ scales the overall cost of export formality, per USD exported.
+If this intuition is correct, we expect (the vector of) $\beta_3$ to be negative (and all the more negative as we are on recent years).
+\item \textbf{explain the role and the interpretation of the variable associated to the weight.}
+
 \item $\varepsilon_{iy}$ is the trade cost per USD exported that is left unexplained (for instance, if on the left-hand side we have a value for a given pair (country-year) of 0.008 (for this counry-year pair, trade costs is 0.8\% of the fob price, or 0.8 USD for 1 USD exported), and that the sum of all explicative terms (from $a$ to $\gamma\frac{formalities_{iy}}{\bar{pfob}_{iy}}$) amounts to 0.007, it means that the error term is 0.001, equal to 0.1 percentage point of the trade cost (in \% of the fob price).
@@
 
-We check that these 3 series are in fact very close to one another over 1987-2013.
+We check that these three series are in fact very close to one another over 1987-2013.
 Because we don't want to limit ourselves to data post 2004 (the availability of the ``export formality cost'' from Doing Business, in order to exploit all information over the entire period 1974-2013, we retain the BP series as benchmark (very frequently retained in the literature), as it is the sole which is available over the whole period.

### 2016-07-21 redaction/dhp_draft_july2016.tex
Modif intro

@@
 However, many common empirical facts support the existence of additive costs.
-This paper provides results on the size and composition of transport costs over time, explicitly distinguishing between its multiplicative and additive components, using sectorial cif-fob prices over 1974-2013 taken from US imports data.
-First, we estimate the two components of transport costs, by transport mode (air or ocean).
-We thus find that ocean transport costs decompose in an ad-valorem tax equivalent of mean value 3.22\% and a per-unit cost that amounts to 2.85\% of the mean fob price over the period.
-These values are respectively equal to 2.5\% and 1.9\% in air shipping.
-Second, we provide a quantitative assessment what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
-On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
-Third, the time dimension of our data allows us to characterize the evolution of transport costs.
-After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012.
-While this order of magnitude applies for both the additive and multiplicative components of vessel transport costs, air transport costs exhibit more contrasted trends.
+This paper measures the size of additive costs in transport, using SITC 3 and 4 digit cif-fob unit values over 1974-2013 taken from US imports data.
+We estimate the two components of transport costs, by transport mode (air or ocean).
+We find that additive costs are 2.85\% of fob unit values for ocean transport costs (and ad-valorem ones 3.22\%).
+These values are respectively equal to 1.9\% and 2.5\% for air transport costs.
+We show that taking additive costs into account improves the fit of the modelling of transport costs.
+The time dimension of our data allows us to characterize the evolution of transport costs.
+After correcting for composition effects, we obtain that all types of transport costs have been constant from 1974 to 1984 and then  decreased by 40\% over the period 1984-2013.
+Most of the early decline in air transport costs can be explained by the ad-valorem component.
+By contrast, this component nearly doubled in the 2000s.
 \end{abstract}

### 2016-07-21 redaction/dhp_draft_july2016.tex
Abstract et Lashkaripour

@@
 The time dimension of our data allows us to characterize the evolution of transport costs.
-After correcting for composition effects, we obtain that all types of transport costs have been constant from 1974 to 1984 and then  decreased by 40\% over the period 1984-2013.
-Most of the early decline in air transport costs can be explained by the ad-valorem component.
+After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
+Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
+E.g.
+most of the early decline in air transport costs can be explained by the ad-valorem component.
 By contrast, this component nearly doubled in the 2000s.
@@
 Ne pas en parler?
-C'est un peu limite.} The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
+C'est un peu limite.}
+In contrast with most of the litterature, \citet{lashkaripour} finds supporting evidence for the iceberg assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
+His study is restricted to 60\% of US imports that are enumerated by items in the statistics.
+Furethermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
+The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
 Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e.
@@
   \centering
-  \caption{Air: Dynamics of Trade Costs over Time}
+  \caption{Air: Dynamics of Trade Costs over Time, 4-digit}
 \begin{center}
@@
   \centering
-\caption{Vessel: Dynamics of Trade Costs over Time}
+\caption{Vessel: Dynamics of Trade Costs over Time, 4-digit}
 \begin{center}

### 2016-07-21 redaction/dhp_draft_july2016.tex
Explication de la procédure de fin

@@
 \usepackage[tableposition=top]{caption}
+\usepackage{ bbold }
+\usepackage{ upgreek }
 %\usepackage[nolists, figuresfirst]{endfloat}
@@
 
+The evolution of transport costs depends both on the evolution of per product and per partner costs and on the evolution of the composition of trade.
+To isolate the evolution of per product and per partner cost, we estimate the following equations.
+
+\begin{eqnarray}
+\tau_{ikt}&=&\exp\left(\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_k\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(\tau_{ikt})&=&\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i + \sum_k\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt}
+\end{eqnarray}
+
+\begin{eqnarray}
+t_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_k\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_k\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt}
+\end{eqnarray}
+
+Both equations are estimated using the non-linear least square procedure in Stata. 
+In the iceberg and multiplicative case, we report $\Gamma_t$, defined as:
+
+\begin{eqnarray}
+\Gamma_t &=& 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{eqnarray}
+
+Figure ???? reports the time effect as the $\Gamma_t$ in the iceberg and multiplicative case and as $100.\exp(\gamma_t)$ in the additive case (assuming $\gamma_{1974}=0$).
+
 \section{Conclusion \label{sec:conclu}}

### 2016-07-21 redaction/dhp_draft_july2016.tex
Petite modif dans la rédaction

@@
 \begin{eqnarray}
-\tau_{ikt}&=&\exp\left(\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_k\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(\tau_{ikt})&=&\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i + \sum_k\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt}
+\tau_{ikt}&=&cst.\exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(\tau_{ikt})&=&cst+\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt}
 \end{eqnarray}
@@
 
-Figure ???? reports the time effect as the $\Gamma_t$ in the iceberg and multiplicative case and as $100.\exp(\gamma_t)$ in the additive case (assuming $\gamma_{1974}=0$).
+Figure ???? reports the time effect as the $\Gamma_t$ in the iceberg and multiplicative case and as $100.\exp(\gamma_t)$ in the additive case (assuming in both cases that $\gamma_{1974}=0$).
 

### 2016-07-21 redaction/dhp_draft_july2016.tex
changement version lise

@@
 However, many common empirical facts support the existence of additive costs.
-This paper measures the size of additive costs in transport, using SITC 3 and 4 digit cif-fob unit values over 1974-2013 taken from US imports data.
-We estimate the two components of transport costs, by transport mode (air or ocean).
-We find that additive costs are 2.85\% of fob unit values for ocean transport costs (and ad-valorem ones 3.22\%).
-These values are respectively equal to 1.9\% and 2.5\% for air transport costs.
-We show that taking additive costs into account improves the fit of the modelling of transport costs.
-The time dimension of our data allows us to characterize the evolution of transport costs.
+This paper contributes to this debate.
+Precisely, we provide a quantitative assessment of the size and the importance of the additive component in international transport costs.
+Using SITC 3 and 4 digit cif-fob unit values over 1974-2013 taken from US imports data, we estimate the two components of transport costs, by transport mode (air or ocean).
+We find that additive costs are 2.85\% of fob unit values for ocean shipping (and ad-valorem ones 3.22\%).
+These values are respectively equal to 1.9\% and 2.5\% for air transport.
+Further, we show that taking additive costs into account improves the fit of the modelling of transport costs.
+We also use the time dimension of our data to characterize the evolution of transport costs.
 After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
 Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
-E.g.
-most of the early decline in air transport costs can be explained by the ad-valorem component.
-By contrast, this component nearly doubled in the 2000s.
+While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s.
+These results therefore confirm the importance of the additive component in accounting for international transport costs.
 \end{abstract}
@@
 \noindent Trade costs remain central in international economic analysis.
-Defined as the costs associated with the exchange of goods across national borders, they are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-Many believe that the latter have dramatically decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-\citet{Glaeser04} find that, over the twentieth century, the costs of moving goods have declined by over 90\% in real terms.
-However, \citet{hummels2007} shows that the bulk of price declines in transportation comes from air shipping, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004; concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, even if the rise in containerization lowered shipping costs from 3 to 13\%.
-Studies overviewed by \citet{Behar_Venables} are consistent with those results: The fall in measured transport costs has been relatively small.
-According to these authors, this is attributable to the growing importance of fuel costs and technical progress, which improved speed and reliability rather than decreased costs.\smallskip
-
-Trade costs are considered as a major obstacle to international economic integration and international trade flows.
+Notably, they are considered as a major obstacle to international economic integration and international trade flows.
 According to the estimates by \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
@@
 \citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
+Defined as the costs associated with the exchange of goods across national borders, they are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
 \citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs.
@@
 often exhibit (at least partly) an additive structure.
-The structure (additive vs multiplicative) of transport costs is far from being anecdotal, as the literature has long pointed out its role in shaping the pattern of trade flows.
+The structure (additive vs multiplicative) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows.
 The Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
-\citet{martin2012} gives a strong empirical support to this conjecture: Based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite.
+\citet{lashkaripour} challenges this view though.
+He finds supporting evidence for the iceberg assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
+However, his study is restricted to 60\% of US imports that are enumerated by items in the statistics.
+Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
+Besides, a number eo empirical papers provide a strong empirical support to the role of additive trade costs.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite.
 \citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: The elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption.
 Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
-\textbf{je ne sais pas quoi faire de la reference a lashkaripour.
-Complique d'en parler, car on ne peut pas le contredire car on n'a pas les donnees qui le permettent.
-Ne pas en parler?
-C'est un peu limite.}
-In contrast with most of the litterature, \citet{lashkaripour} finds supporting evidence for the iceberg assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
-His study is restricted to 60\% of US imports that are enumerated by items in the statistics.
-Furethermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
 The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
-Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e.
-more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments.
-
-%Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level.
-Based on US imports data, he finds that a 10\% increase in the price of an item increases the shipping cost by 9.5\%.
-The additive structure of trade costs emerges only when weight is forced to be independent of price.
-Besides, \citet{Lashka15} supports that higher-priced, still exported to faraway markets, varieties are heavier and involve proportionally higher shipping costs, the latter increasing the relative demand for high-markup varieties.
-Markups would therefore be the driving force for explaining the ``shipping the good apples out'' effect in the presence of \emph{ad-valorem} trade costs.
+Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments.
+With additive costs, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments, thereby inducing more ``lumpiness'' in international trade.
+\smallskip
 
 Beyond the positive aspect of understanding trade patterns, several recent papers also point out the normative implications of additive trade costs.
-\citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the iceberg component.
+\citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the iceberg component.
 A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs.
 Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff.
-While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much can be done in quantifying such gains though, by lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap.
+While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much can be done in quantifying such gains though, one potential reason being the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap.
 \bigskip
@@
 Our paper contributes to the literature by providing results on the size of transport costs over time, explicitly distinguishing between multiplicative and additive parts.
-More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over several decades.
+Precisely, we quantitatively assess the size and the importance of the additive component in international transport costs.
 To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013.
-Closely related to this paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level data, their results suggest that additive costs are on average 30-45\% relative to the median consumer price, and that they are strongly correlated with standard proxies for trade costs (like e.
-g., distance).
+Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
+Based on Norwegian firm-level data, their results suggest that additive costs are on average 30-45\% relative to the median consumer price.
 However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects.
 
-First, our theoretically agnostic approach provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models.
-We thus obtain that the mean values over 1974-2012 of iceberg costs are equal to 2.5\% in air and 3.2\% in ocean shipping, whereas the additive component amounts to 1.9\% and 2.85\% of the fob price in air and ocean shipping respectively.
-To our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the magnitude of both multiplicative and additive costs in total transport costs.
-Second, we provide, through standard measures of ``goodness-of-fit'', an empirical assessment of what standard international trade models lose by skipping additive transport costs.
+First, our theoretically agnostic approach provides a fairly simple framework for measuring both multiplicative and additive parts of transportation costs, which represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} only allows them to identify the ratio of the additive to the multiplicative cost, ie $\frac{t_{ik}}{\tau_{ik}}$ in our terminology, that they interpret by expressing it in terms of the median fob price by country-product ($\bar{p}^{med}_{ik}$).
+By contrast, our estimation strategy enables to uncover both values of the iceberg and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fob price.} We thus obtain that the mean values over 1974-2012 of iceberg costs are equal to 2.5\% and 3.2\% in air and ocean respectively, whereas the additive component amounts to 1.9\% and 2.85\% of the fob price (in air and ocean).
+To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both multiplicative and additive costs in total transport costs.
+Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
 Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
 On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
-Third, the time dimension of our data allows us to characterize the evolution of transport costs, by transport mode (air or vessel) and for each specific (multiplicative and additive) cost over a forty years time span.
+We also show that taking additive costs into account significantly improves the fit of the modelling of transport costs, through various measures of ``goodness-of-fit'', for both transport modes and years.
+
+Third, our paper contributes to the literature documenting the patterns of transport costs over time.
+Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}).
+\cite{Glaeser04} find that, over the twentieth century.
+\textbf{stop} he time dimension of our data allows us to characterize the evolution of transport costs, by transport mode (air or vessel) and for each specific (multiplicative and additive) cost over a forty years time span.
 After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012.

### 2016-07-21 redaction/dhp_draft_july2016.tex
changement version lise, suite

@@
 
-Third, our paper contributes to the literature documenting the patterns of transport costs over time.
+Third, we exploit the time dimension of our database to document the patterns of transport costs over time.
 Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}).
-\cite{Glaeser04} find that, over the twentieth century.
-\textbf{stop} he time dimension of our data allows us to characterize the evolution of transport costs, by transport mode (air or vessel) and for each specific (multiplicative and additive) cost over a forty years time span.
-After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012.
-Decomposing by transport mode and by type of transport costs allows to go deeper into this picture.
-First, as in \cite{hummels2007}, we find that the transport costs decrease is more pronounced in air than in ocean shipping, starting in 1984.
-However, while ocean shipping costs display a regular decreasing pattern until 2012, air transport costs rather get stabilized over 2005-2012.
-Second, while both the additive and multiplicative components of ocean transport costs have roughly decreased by 40\% over the period, air transport costs exhibit more contrasted trends.
-If the reduction of air shipping costs can most be attributed to the decrease in its multiplicative component over 1974-2005, the opposite scheme prevails afterwards, the reduction in per-unit air transport costs being compensated by an increase in the ad-valorem component.
-\smallskip
+\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
+However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004.
+Concernin ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, a conclusion in accordnace with the studies reviewed by \cite{Behar_Venables}.
+Our paper contributes to this debate.
+As claimed by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time.
+After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
+Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
+Air transport in particular exhibit contrasted trends.
+While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost.
+These results therefore confirm the importance of the additive component in accounting for international transport costs.
 
-Section \ref{sec:data_method} explains the data sources and the empirical methodology retained in the paper.
+The paper is built as follows.
+\ref{sec:data_method} explains the data sources and the empirical methodology retained in the paper.
 Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results.
@@
 \section{Data Sources and Empirical Methodology \label{sec:data_method}}
+
+\textbf{reprendre completement la section 2}
 

### 2016-07-22 redaction/dhp_draft_july2016.tex
Nouvelle version, pour soumission ETSG

@@
 \noindent Trade costs remain central in international economic analysis.
-Notably, they are considered as a major obstacle to international economic integration and international trade flows.
+In particular, they are considered as a major obstacle to international economic integration and international trade flows.
 According to the estimates by \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
 After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements.
-Based on panel data, \citet{novy13} thus finds  that U.S.
+Based on panel data, \citet{novy13} finds  that U.S.
 trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
@@
 \citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
-Defined as the costs associated with the exchange of goods across national borders, they are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-\citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs.
-Within this latter dimension, 44\% are border-related trade barriers and 21\% transportation costs.
-\citet{Lafourcade_Thisse} also find that the share of transport costs in the consumer price of manufactured goods remains high, while \citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
-If much of trade-policy barriers have been removed over the second half of the 20th century, these findings suggest that the transport costs component of the overall trade costs remain large and deserve attention.
+Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries put up to 74\% markups over production costs.
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
+In this vein, \citet{anderson_wincoop_jel} obtain that the 74\% markup made by international trade costs can be divided in 44\% for border-related trade barriers and 21\% for transportation costs.
+This last figure suggests that the share of transport costs in the consumer price of manufactured goods is substantial.
+\citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
+If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
 This is accordingly the focus of the paper.\smallskip
@@
 The Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
-\citet{lashkaripour} challenges this view though.
+\citet{Lashkaripour-mimeo-2016} challenges this view though.
 He finds supporting evidence for the iceberg assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
-However, his study is restricted to 60\% of US imports that are enumerated by items in the statistics.
+One can yet be concerned by the generality of his results though.
+By nature, his study is restricted to goods that are enumerated by items in the statistics (that represent 60\% of US imports for instance).
 Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
-Besides, a number eo empirical papers provide a strong empirical support to the role of additive trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite.
-\citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: The elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption.
-Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
+Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fob unit values on exports to more remote countries, in contradiction with the iceberg hypothesis.
+\citet{hummels_skiba} estimate the elasticity of freight rates with respect to price is estimated to be well below unity.
+Also, their estimates imply that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
+These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
 The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
@@
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level data, their results suggest that additive costs are on average 30-45\% relative to the median consumer price.
-However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects.
+Based on Norwegian exports in 2004, their results suggest that additive costs are about 14\% of the median consumer price.
+However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects, as we develop below.
+One important difference is that we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013.
+In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
+They can be summarized in three main points.
 
-First, our theoretically agnostic approach provides a fairly simple framework for measuring both multiplicative and additive parts of transportation costs, which represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} only allows them to identify the ratio of the additive to the multiplicative cost, ie $\frac{t_{ik}}{\tau_{ik}}$ in our terminology, that they interpret by expressing it in terms of the median fob price by country-product ($\bar{p}^{med}_{ik}$).
-By contrast, our estimation strategy enables to uncover both values of the iceberg and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fob price.} We thus obtain that the mean values over 1974-2012 of iceberg costs are equal to 2.5\% and 3.2\% in air and ocean respectively, whereas the additive component amounts to 1.9\% and 2.85\% of the fob price (in air and ocean).
+First, our theoretically agnostic approach provides a fairly simple framework for measuring both multiplicative and additive parts of transportation costs, which represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the multiplicative cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median fob price by country-product ($\bar{\widetilde{p}}_{ik}$).
+By contrast, our estimation strategy enables us to uncover both values of the iceberg and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fob price $\widetilde{p}_{ik}$.} We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% in air and ocean respectively, whereas the additive component amounts to 1.8\% and 2.9\% of the fob price (in air and ocean).
 To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both multiplicative and additive costs in total transport costs.
@@
 However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004.
-Concernin ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, a conclusion in accordnace with the studies reviewed by \cite{Behar_Venables}.
+Concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}.
 Our paper contributes to this debate.
-As claimed by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time.
+As pointed out by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time.
 After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
@@
 Air transport in particular exhibit contrasted trends.
-While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost.
-These results therefore confirm the importance of the additive component in accounting for international transport costs.
+While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost.
+These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
 
@@
 
-\textbf{reprendre completement la section 2}
-
 \subsection{A measure of Transportation Costs}
@@
 Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price.
-Strictly higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.
-
+Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.
 The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
-Using this dataset has (at least) three main advantages.
-First, this dataset has been used by \cite{hummels2007}, which enables us to compare our results to his findings.
-However, we complement Hummels' (2007) findings as we extend the time coverage to 2013 (while \cite{hummels2007} stops in 2004).
-As we show in Section \ref{sec:results_trends}, extending the time period over he recent years delivers interesting insights regarding the trends in air shipping costs.
-Second, and importantly, this dataset delivers a strong statistical reliability arising from a single, trustworthy customs origin.
-Based on customs declarations, this dataset inventories all imports (both values and quantities) by origin to the United states at the HS 10-digit highly disaggregated level, with a concordance code to the SITC 5-digit coding system.
-In addition, the database reports information regarding freight expenditures and transportation mode (Ocean Vessel and Air).
-The first will be crucial to compute transport costs (see below), the second will allow us enlightening substantial differences in the dynamics of transport costs across transportation mode.
-Third, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FOB price).
-This is highly valuable, as we can estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs.
-This differentiates us from \citet{Irrazabal_2015}, which can only estimate the ratio of additive costs as a share of the total consumer price (the only price they can encover).
-
-It is also true that using this dataset has drawbacks.
-First, our measure of the cif-fob price gap only covers transportation costs by nature, thereby being mute about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
+It is worth acknowledging that using this dataset encounters some limitations.
+First, our measure of the cif-fob price gap only covers transportation costs by nature, thereby being silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
 Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit.
 According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
-\textbf{en faire un contre-argument positif}
+In this respect, it is true that using this dataset embraces a partial view of international transport barriers.
 
-Even if the data of transportation costs is available at a more disaggregated level, we use sectorial price data at the 3-digit level, primarily for technical reasons.
+Yet, we believe that the results we get deliver valuable insights on this topic.
+Indeed, using this dataset has (at least) three main advantages.
+First, this dataset delivers a strong statistical reliability arising from a single, trustworthy customs origin.
+Based on customs declarations, this dataset inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
+In addition, the database reports information regarding freight expenditures and transportation mode (Ocean vessel and Air).
+The first will be crucial to compute transport costs (see below), the second will allow us enlightening substantial differences in the dynamics of transport costs across transportation mode.
+Second, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FOB price).
+This is highly valuable, as we can estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs.
+This differentiates us from \citet{Irrazabal_2015}, which can only estimate the ratio of additive costs as a share of the median total consumer price.
+Third, this dataset having been used by \cite{hummels2007}, this enables us to compare our results to his findings.
+Further, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004.
+As we show in Section \ref{sec:results_trends}, covering the time period over the recent years delivers interesting insights regarding the trends in air shipping costs.
+
+We make use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level.
+This is primarily due to technical reasons.
 As detailed below, the use of a nonlinear estimator triggers computational limitations that do not make them a likely option, especially when covering a long period of time.
-Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level (for some selected years).
-Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
+Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level (for some selected years).\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.} Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
 Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.
@@
 \paragraph{The estimated equation} Our purpose is to provide estimates over time of the sizes of multiplicative and additive costs among total transport costs.
-To do so, we start from a very simple equation (similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others):
+To do so, we start from the equation that expresses the price paid by the consumer (import, or cif) price $p$ as a function of the producer (export, or fob) price $\widetilde{p}$ given both per-unit ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}
 
@@
 
-\noindent This equation expressed the consumer (import, or cif) price $p$ as a function of the producer (export, or fob) price $\widetilde{p}$ given both per-unit ($t$) and ad-valorem ($\tau$) transport costs.
-As usual in the literature, the so-called ``iceberg'' trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs), while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs).
-We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.
-Let us denote $i$ the origin country, and $k$, the sector (or product).
-Transforming the above equation (\ref{eq:base}) as ratio, we thus get the equation to be estimated as given by:\footnote{Also keeping in mind that the estimated equation is transport-mode and year specific (air or ocean shipping).
-We skip the year and transport-mode dimensions here to alleviate notations.}
+\noindent As usual in the literature, the so-called ``iceberg'' trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no iceberg trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive trade costs).
+We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.} Let us denote $i$ the origin country, and $k$, the sector (or product).
+Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:
 
-\begin{equation}
+\begin{equation*}
 \frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
-\end{equation}
+\end{equation*}
 
-\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1°)} both multiplicative and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2))} in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) become:
+\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1°)} both multiplicative and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} in a multiplicative way for the former and an additive way for the latter.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:
 
@@
 
-As a result, our underlying theoretical equation is specified as:
+\noindent As a result, our underlying theoretical equation is specified as:
 
-\begin{equation}
+\begin{equation*}
 \frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
-\end{equation}
+\end{equation*}
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the \emph{cif} price $p$ cannot be lower than the \emph{fob} price: $p>\widetilde{p}_{ik}$.
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the cif price $p$ cannot be lower than the fob price ($p>\widetilde{p}_{ik}$).
 Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term, according to:
@@
 \end{equation*}
-
-Taking in log, this finally drives us to estimate the following equation:
+\noindent Taking in log, this finally drives us to estimate the following equation:
 
@@
 
-The shape of our main equation \ref{eq:estimatedequation} is such that estimations cannot be performed using standard linear estimators.
-Therefore, all estimates are performed using non linear least squares.
+The non-linearity of Equation (\ref{eq:estimatedequation}) implies that estimations cannot be performed using standard linear estimators.
+Therefore, all estimates are performed using non-linear least squares.
 The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
@@
 This is done by computing several standard diagnostic statistics (R$^{2}$ and statistics based on the likelihood function).
-Accordingly, for each year and transport mode, we estimate two equations, depending on additive transport costs being included (Equation (\ref{eq:model_with_add}) or not (Equation (\ref{eq:model_nlI})):
+Accordingly, for each year and transport mode, we estimate two equations, depending on additive transport costs being included (Equation (\ref{eq:model_with_add})) or not (Equation (\ref{eq:model_nlI})):
 
 \begin{eqnarray}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\underbrace{\tau_{i} \times \tau_{k}}_{\tau^{ice}_{ik}}-1 +\underbrace{\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}}_{t^{add}_{ik}} \right) + \epsilon_{ik} \label{eq:model_with_add} \\
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\underbrace{\tau_{i} \times \tau_{k}}_{\tau^{nlI}_{ik}}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add} \\
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI}
 \end{eqnarray}
@@
 After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI}), we proceed similarly to get $\widehat{\tau}^{nlI}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
 To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
@@
 The objective of this section is twofold.
-First, we characterize the magnitudes of transport costs over time (by transport mode), distinguishing whether the additive component $t_ik$ is included or not in the estimated equation (\ref{eq:estimatedequation}).
-Second, we quantitatively assess the importance of the per-unit component of in overall transport costs through the means of goodness-of-fit measures.
+First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (\ref{eq:estimatedequation}).
+Second, we evaluate the importance of the per-unit component of in overall transport costs through the means of goodness-of-fit measures.
 
@@
 Our first contribution to the literature is to provide estimates for the size of both the multiplicative and the additive components of transport costs.
-Tables \ref{tab:result_air} and \ref{tab:result_ves} report a summary of our results.
-Precisely, they display the mean and median values of each type of trade costs (multiplicative estimated alone, estimated along with additive costs and the additive component), expressed in percentage of the fob price, as well as the associated standard deviation, averaged over the period 1974-2013 and for selected years.
+Table \ref{tab:summary_results} reports a summary of our results.
+It displays the mean and median values of each type of trade costs (multiplicative estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix , we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+Results for all years are of course, available upon request to the authors.}$^{,}$\footnote{We present the results for Air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
+Overall results (over the whole period) are not substantially affected if this year is included though.
+These results are available upon request to the authors.}
 
@@
   \centering
-  \caption{Air (3 digits): Mean values over the period and selected years}
+  \footnotesize{
+  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
   \begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 6.9 & 5.4 & 5.0 & 3.6 & 4.2 & 3.4 & \multicolumn{1}{c}{\textbf{5.1}} \\
-Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9 & \multicolumn{1}{c}{\textbf{4.2}} \\
-Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
-\hline
-\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0 & \multicolumn{1}{c}{\textbf{1.9}} \\
-\multicolumn{1}{l}{Median} & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5 & \multicolumn{1}{c}{\textbf{0.7}} \\
-\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
-\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 3.6 & 2.3 & 2.4 & 1.7 & 2.6 & 1.7 & \multicolumn{1}{c}{\textbf{2.5}} \\
-\multicolumn{1}{l}{Median} & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 & \multicolumn{1}{c}{\textbf{1.8}} \\
-Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
-\hline
-\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Mean and median values of all type of transport costs are expressed in percentage of the fob price.
-}
+    \begin{tabular}{l|cc|cc}
+      \hline \hline
+    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
+    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
+    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
+    \multicolumn{5}{l}{With only Iceberg Trade Costs }  \\ \hline
+    Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
+    Median & 1.051 & 1.042 & 1.052 & 1.037 \\
+    Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
+    Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
+    Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
+    \multicolumn{5}{l}{With Additive \& Iceberg Trade Costs } \\ \hline
+   \textit{Iceberg term} & & & & \\ \hline
+    Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
+    Median & 1.028 & 1.018 & 1.028 & 1.016 \\
+    Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
+    Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
+    Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
+    \textit{Additive term }& & & &   \\ \hline
+    Mean  & 0.029 & 0.018 & 0.028 & 0.019 \\
+    Median & 0.019 & 0.007 & 0.017 & 0.008 \\
+    Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
+    Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
+    Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+        \# obs. & 29279 & 28207 & 29317 & 27680 \\
+    \# origin country & 188 & 191 & 188 & 189 \\
+    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
+  \end{tabular}
+    \end{center}}
+\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Term A expressed in fraction of fob price.
+($^\ast$): Four 4-digit estimation: 0n selected years.
+($^{\ast \ast}$): 1989 omitted in 3 digit estimation for Air.}}
 \end{table}%
 
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel (3 digits): Mean values over the period and selected years}
-\begin{center}
-    \begin{tabular}{lccccccc}
+Table \ref{tab:summary_results} yields the following comments.
+First, the magnitude of overall transport costs is sizeable, either considering the average value over the period or even in the most recent years.
+Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
+\textbf{comparer avec la litterature?}.
+Second, when decomposing transport costs into an additive and a multiplicative component, we find that the per-unit cost dimension is sizeable.
+Over the whole period, additive costs are 2.85\% of fob unit values for ocean shipping  - and ad-valorem ones being equal to 3.22\%.
+These values are respectively equal to 1.9\% and 2.5\% for air transport.
+For both transport modes, the omission of the additive term seriously biases the multiplicative term upward.
+In quantitative terms, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
 
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 9.8 & 6.5 & 5.7 &5.1 & 4.0 & 3.6 & \textbf{5.8} \\
-Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3 & \textbf{5.1} \\
-Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
-\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-\hline
-\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
-Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5 & \textbf{2.9} \\
-Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 & \textbf{1.9} \\
-Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 & \textbf{3.2} \\
-Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8 & \textbf{2.8} \\
-Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
-\hline
-\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Mean and median values of all type of transport costs are expressed in percentage of the fob price.}
-\end{table}%
-
-Tables \ref{tab:result_air} and \ref{tab:result_ves} yield the following results.
-First, the magnitude of overall transport costs is sizeable, either considering the average value over the period or even in the most recent years.
-Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5?1\% for air shipping.
-\textbf{comparer avec la litterature?}
-
-\subsection{Asessing the importance of per-unit trade costs}
-In this section, we investigate the properties of our two models (with and without an additive term), to assess what loss of explanatory power is implied by the omission of the additive term.
-We start with a simple visual inspection of Figure \ref{fig:good_fit}, where are reported the yearly median \emph{ad valorem} estimates of the iceberg costs ($\widehat{\tau}_{i}$) estimated in equation \ref{eq:estimatedequation}\footnote{Qualitatively similar results are obtained when the mean values are reported instead.
-More details of those results available upon request to the authors.}.
-The dotted line plots the estimated iceberg cost ($\widehat{\tau}^{nlI}$) when the additive term $t$ is restricted to 0 (Equation (\ref{eq:model_nlI}) being estimated), and the plain dark line represents the iceberg cost in the model including a non-null additive term (Equation (\ref{eq:model_with_add}) being estimated).
-In both cases, multiplicative costs are reported minus one, for a better reading comfort and comparability with Figure \ref{fig:mult_vs_add}.
+We investigate the robustness of this conclusion on a yearly basis.
+Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the multiplicative component, estimated alone (Equation (\ref{eq:model_nlI}), dashed line) and along with the additive component (Equation (\ref{eq:model_with_add}, plain line), by transport mode.
 
 \begin{figure}[htbp]
-\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
+\caption{Multiplicative Costs (3 digits}\label{fig:mult_alone_withadd}
 \begin{center}
@@
 {\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
+\includegraphics[width=3in, height=2.5in]{Fig1a_mult_air_3d.pdf}
+& \includegraphics[width=3in,height=2.5in]{Fig1b_mult_vessel_3d.pdf} \\
 \end{tabular}
@@
 
-In Figure \ref{fig:mult_vs_add}, we report the yearly median value for each component of transport cost (expressed in terms of the fob price, by transport mode) ) after running estimation of Equation (\ref{eq:model_with_add}).
+From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds on a yearly basis.
+For Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
+Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air transport.
+To get a better picture on this, we study the shares of both multiplicative and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.
 
-%Economic interpretation
-%usepackage{float} + H is great to keep images on the right page
-\begin{figure}[H]
-\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
+\begin{figure}[htbp]
+\caption{Decomposing Transport costs (3 digits}\label{fig:decomp_TC_3d}
 \begin{center}
@@
 {\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
+\includegraphics[width=3.5in, height=3in]{Fig2a_decompTC_air_3d.pdf}
+& \includegraphics[width=3.5in,height=3in]{Fig2b_decompTC_vessel_3d.pdf} \\
 \end{tabular}
@@
 
-For both transport modes, we can infer from Figure \ref{fig:good_fit} that the omission of the additive term seriously biases the multiplicative term upward.
-For Air, the size of the bias seems to decrease over time, while it seems pretty stationary for Vessel.
-In any case, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
+At this stage, two interesting comments can be made from Figure \ref{fig:decomp_TC_3d}.
+First, the magnitude of transport costs is higher for ocean shipping than for air transport.
+This result also holds considering iceberg costs alone (Figure \ref{fig:mult_alone_withadd}).
+Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
+We come back to this aspect in Section \ref{sec:results_trends}.}
 
-In order to deliver a more systematic diagnosis, we use several standard measures of fit.
-Of course R$^{2}$ comes naturally to mind, but its use is far from being straightforward when evaluating non-linear estimates.\footnote{R-squared is based on the underlying assumption that the adjusted model is a linear one.
+\subsection{Assessing the importance of per-unit trade costs}
+
+In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fob prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs.
+To do so, we rely on several standard measures of fit.
+The first indicator is through comparing $R^{2}$.
+However, its use is far from being straightforward when evaluating non-linear estimates.
+\textbf{mettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
 In a non-linear context, R-squared is therefore inappropriate, strictly speaking.
-However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.)}.
-Therefore, we provide as a complement the Standard Error of Regression (SER) which represents the average distance that the observed values fall from the regression line.
+However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.}.
+As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
 Smaller values are better because it indicates that the observations are closer to the fitted line.
@@
 Once again, the preferred model is the one with the minimum AIC value.
-Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, see equation~\ref{eq:estimatedequation}) and the Restricted one (\emph{R}, i.e.
-equation~\ref{eq:estimatedequation} with $t=0$).
-The null tested is that the two models are statistically equivalent.
-Results are reported in Tables~\ref{tab:good_fit_air} and \ref{tab:good_fit_ves}.
+Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e.
+Equation (\ref{eq:model_with_add}) and the Restricted one (\emph{R}, i.e.
+Equation (\ref{eq:model_nlI})).
+The null tested is that the two models are statistically equivalent.\medskip
 
-%If we keep the R² as a measure of GoF, we will have to provide some evidence of normality for residuals. In addition, a plot of observations and the adjustment curve might be useful.
-Unsurprisingly, it appears that the inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or transport mode.
-On average over the whole period, $R^{2}$ doubles when restricting to Air Transport, and increases by 50\% for Vessel.
-Similar qualitative conclusions arise from the comparisons of SERs, but may be more reliable on the quantitative side, since the use of R-squared on non-linear estimates may be qualified.
+Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_ves}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level).
+The results for all years are available upon request to the authors.}\medskip
+
+\begin{table}[htbp]
+  \centering
+  \caption{Air: Measures of Goodness-of-fit (3 digits)}
+  \footnotesize{
+\begin{center}
+    \begin{tabular}{l|cccccc|c}
+    \hline \hline
+    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat($^{\ast}$) \\ \hline
+    \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
+    Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & \textbf{0.31} \\
+    Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
+    \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
+    Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
+    Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
+   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
+    Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
+    Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
+    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
+    Term I only & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
+    Terms A \& I & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
+    \multicolumn{8}{l}{\textit{Test LL }}  \\
+    2*(ll(UR) -ll(R)) & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
+    nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
+    p-value(d) & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
+    \hline \hline
+ \end{tabular}%
+    \end{center}}
+  \label{tab:good_fit_air}%
+ \parbox[l]{12cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
+($^\ast$): The mean statistics calculated as the average value over all years.
+}}
+\end{table}%
+
+\begin{table}[htbp]
+  \centering
+  \caption{Vessel: Measures of Goodness-of-fit (3 digits)}
+    \footnotesize{
+\begin{center}
+\begin{tabular}{l|cccccc|c}
+\hline \hline
+Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat($^{\ast}$) \\ \hline
+\multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
+Term I only & \multicolumn{1}{c}{0.450} & \multicolumn{1}{c}{0,415} & \multicolumn{1}{c}{0.456} & \multicolumn{1}{c}{0.401} & 0,350 & \multicolumn{1}{c}{0.339} & 0.39 \\
+Terms A \& I & \multicolumn{1}{c}{0.612} & \multicolumn{1}{c}{0.575} & \multicolumn{1}{c}{0.590} & \multicolumn{1}{c}{0.571} & 0.491 & \multicolumn{1}{c}{0.462} & 0.56 \\ \hline
+\multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
+    Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
+    Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
+   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
+    Term I only & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
+    Terms A \& I & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
+    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
+    Term I only & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
+    Terms A \& I & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
+    \multicolumn{8}{l}{\textit{Test LL }}  \\
+    2*(ll(UR) -ll(R )) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
+    nb of restrictions =q  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
+    p-value(d) & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
+    \hline \hline
+    \end{tabular}%
+    \end{center}}
+  \label{tab:good_fit_vessel}%
+  \parbox[l]{8cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
+($^\ast$): The mean statistics calculated as the average value over all years.
+}}
+\end{table}%
+
+Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_ves} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
+On average over the whole period, the $R^{2}$ doubles when per-unit costs are included for Air, and increases by 50\% for Vessel.
+Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
 Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes.
-Both AIC and Log-Likelihood statistics decrease with the inclusion of the additive term, and the LL test unambiguously reject the null of statistical equivalence of the two models.
-This is true whatever the considered year.
+Both AIC and Log-Likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
+These results holds whatever the considered year.\smallskip
 
-However, going into the examination of the levels of these goodness-of-fit statistics reveals intriguing differences.
-Multiplicative costs appear more important for the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector; the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air.
-Besides, the dynamics over time also reveals intriguing differences.
-Quality of fit decreases strongly after 2000, for both transport modes.
-For Air, it seems that the additive term explains less and less variance over time: in 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the SER by 14 percentage points ;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
+Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences.
+First, multiplicative costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes).
+As well, the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air.
+\textbf{et pourtant, la part des additifs dans vessel est plus elevée, 48.2\% en moyenne vs 42.3\% pour AIr, cf le .xls.
+A dire?
+Faire reference a la figure 2.}
+
+Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000, for both transport modes.
+For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
 The picture is quite different for Vessel.
-For that transport mode, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years.
-It appears that the deteriorating performance of the model in terms of goodness-of-fit from 2000 comes mainly from the decreasing performance of the iceberg term.
-In order to enlighten the roots of these differences, the following section elaborates further on the quantitative dynamics of each type of cost.
+In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included.
+Rather, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the iceberg term.
+\textbf{so what?
+pas tres punchy}\smallskip
 
-For comparison purposes, we provide in the appendix similar results for some years based on a more disaggregated dataset at the product level (4-digits).\footnote{Since we use a non-linear estimator, estimations are highly computer-intensive.
-This explains why we decided to focus on the 3-digit product classification, which requires less time-consuming calculations.}.
-If anything the quality of fitting appears slightly higher when estimations are based on the 4-digits classification.
+For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in in Appendix \ref{app:robust_4d}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
+If anything, the quality of fitting appears slightly higher when estimations are based on the 4-digit classification.
 This is especially true for the model restricting trade cost to their iceberg dimension, whatever the transport mode considered.
-When the additive part is taken into account however, the difference becomes very small, whatever the considered criterion.
-
-Using a more disaggregated classification unsurprisingly adds some statistical precision, but not to an extent which disqualifies the use of slightly more aggregated data.
-% need for SER on all years to be complete.
-
-\subsubsection{Interpretation}
-
-We now more specifically investigate the long-run behavior of the estimated iceberg and additive parameters.
-Tables \ref{tab:result_air} and \ref{tab:result_ves} report, respectively for the Air and Vessel sector, our estimates for Iceberg trade costs, first from equation \label{eq:iceberg} (upper panel), then from equation \ref{eq:estimatedequation} where they are estimated jointly with an additive term, which estimate is also reported (bottom panel).
-Remember that estimation is performed on a yearly basis, consequently, for the sake of clarity, we do not report estimates for each year, but for the beginning and the end of our sample, as well as starting year for each decade.
-we report in Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} in the Appendix the estimated for coefficients for products disaggregated at the 4-digits level.
-In any case, these estimates are very close, and do not show any major difference with the baseline ones.
-
-Unsurprisingly, the trend in transportation costs is clearly decreasing whatever the mode (air or vessel) or measure (with or without additive term) we consider.
-Starting from 1974, the decrease appears more important for sea than air transportation: transportation costs are divided by three for the first mode, and two for the second one.
-That said, it is widely know the two oil shocks inflated artificially transportation costs that specific year.
-If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for Air, and 2-3 percentage points for Vessel - depending if we consider mean or median estimates.
-These figures are consistent with the ones by Hummels (2007).
-However, our methodology allows distinguishing in this trend, what comes from multiplicative and additive costs.
-To make a long story short, it appears that the additive costs represent the bulk of the decrease for Air, whereas the multiplicative component is the main responsible of the decreasing trend for Vessel, except for the years 2012-2103, witnessing an abrupt decrease in additive costs.
-To give a few average numbers, starting from 1980 and ending in 2013, the iceberg term decreases by 0.6 percentage point (pp), and the additive term is reduced by 1 pp for Air; for Ocean shipping, the reduction is roughly 2 pp for additive costs, and 1 pp for multiplicative costs.
+When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
+In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not with an extent that would disqualify the use of slightly more aggregated data.
+Further, the same conclusion emerges regarding the significant role of the additive components in fitting international transport costs, that established at the 3-digit level.
 
@@
 
-The evolution of transport costs depends both on the evolution of per product and per partner costs and on the evolution of the composition of trade.
-To isolate the evolution of per product and per partner cost, we estimate the following equations.
+In this section, we investigate the role of the additive component in total transport costs by exploring the time dimension of our data.
+As first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
+Let us consider first the trend in overall transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
+Two main comments can be made.
+First, both air and ocean shipping exhibit a downward trend in overall transport costs.
+Second, starting from 1974, the decrease appears more important for sea than air transportation: Transportation costs are divided by 3 for the ocean shipping, versus by 2 for air transport.
+That said, it may be the case the oil shock inflated transportation costs that specific year.
+If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for Air, and 2-3 percentage points for Vessel (depending if we consider mean or median estimates).
+These results are consistent with \citet{hummels2007}.
+If we now consider the respective shares of additive and multiplicative components, we observe that the share of additive costs in air transport costs has declined between 1974 and 2013, in particular starting in 1988 (panel (a) of Figure \ref{fig:decomp_TC_3d}).
+By contrast, this share has been roughly constant throughout the period in ocean shipping (panel (b)).
+
+Before making any definite statement about this though, it is worth emphasizing that the evolution of 
+transport costs depends both on the evolution of per product and per partner costs and on the evolution of the composition of trade.
+Put it simply, total transport costs may have decreased over time either the country either trades with countries with which trade is cheaper, or trades products which are cheaper to transport, independently of any change in tranport costs \textit{per se}.
+As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
+In this aim, we estimate the following equation for the multiplicative cost, based on the iceberg cost series obtained when estimated alone ($\widehat{\tau}_{ikt}^{nlI}$ and the iceberg cost series estimated along with additive costs ($\widehat{\tau}_{ikt}^{ice}$:\footnote{We thus estimate Equation (\ref{eq:compeffects_mult}) twice, depending on the type of iceberg cost series considered ($\widehat{\tau}_{ikt}^{nlI}$ or $\widehat{\tau}_{ikt}^{ice}$), even if it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
+For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
 
 \begin{eqnarray}
-\tau_{ikt}&=&cst.\exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(\tau_{ikt})&=&cst+\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt}
+\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(\tau_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
 
+As for the additive cost, we estimate the following equation, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
 \begin{eqnarray}
-t_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_k\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_k\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt}
+\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
 \end{eqnarray}
 
-Both equations are estimated using the non-linear least square procedure in Stata.
-In the iceberg and multiplicative case, we report $\Gamma_t$, defined as:
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_mult}) preserve our specification of the multiplicative and additive costs of Equations (\ref{eq:iceberg}) and (\ref{eq:add}).
+That is, we consider that the multiplicative cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+In the three estimations (iceberg costs alone, iceberg costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
+As before, both equations are estimated using the non-linear least square procedure (by transport mode).
 
+In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
+From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{nlI}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:
 \begin{eqnarray}
@@
 
-Figure ???? reports the time effect as the $\Gamma_t$ in the iceberg and multiplicative case and as $100.\exp(\gamma_t)$ in the additive case (assuming in both cases that $\gamma_{1974}=0$).
+As for the additive cost, we built the variable $\Gamma^{add}_t = 100.\exp(\gamma_t)$, with the reference year being 1974 in all cases (ie, with $\gamma_{1974}=0$).
+As a result, the series for $\Gamma_t$ (with an initial value of 100 for $t=1974$) have a straightforward interpretation in percentage changes.
+\medskip
+
+We first report the changes over time of the transport costs, estimated iceberg alone, in Figure \ref{fig:totalTC_compeffects_excl}, by transport mode (Panel (a) for Air and Panel (b) for Vessel).
+
+\begin{figure}[htbp]
+\caption{Total transport costs (estimated iceberg alone), \\
+composition effects excluded
+}
+\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[width=7cm, height=5cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
+First, whatever the transport mode considered,\footnote{Note that this is also true for all type of transport costs, additive and iceberg, as reported in Figure \ref{fig:TC_add_mult_compeffects_excl}, as we come back later on.} they have been roughly constant from 1974 to 1984.
+This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that transport costs fall immediately started in 1974.
+Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 is rather attributable to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
+In line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
+Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased by 40\% over the period 1984-2013.
+Further, if we only assume iceberg costs (Figure \ref{fig:totalTC_compeffects_excl}), we infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005, in accordance with the findings of \cite{hummels2007} and \cite{Behar_Venables}.
+Yet, going deeper in the decomposition between the additive and the multiplicative components substantially modifies the picture.
+
+In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the multiplicative and the additive costs (re-scaled to start from the reference value 100), for each transport mode.
+
+\begin{figure}[htbp]
+\caption{Total transport costs (additive \& multiplicative combined), 
+composition effects excluded}
+\label{fig:TC_addplusmult_compeffects_excl}
+\begin{center}
+\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+In line with Figure \ref{fig:totalTC_compeffects_excl}, Figure \ref{fig:TC_addplusmult_compeffects_excl} reports that transport costs have been roughly constant over 1974-1984 in both modes, decreasing afterwards.
+Yet, another result emerges, that conversely stands in sharp contrast.
+There is now not much difference in the reduction of transport costs between air transport and ocean shipping, and both display very similar trends.
+Taking into account the additive component therefore substantially modifies the analysis of the dynamics of transport costs over time.
+In Figure \ref{fig:TC_add_mult_compeffects_excl}, we investigate this result further, by reporting the changes in both the additive and multiplicative components of transport costs, by transport model.
+
+\begin{figure}[htbp]
+\caption{Decomposing transport costs over time,
+composition effects excluded}
+\label{fig:TC_add_mult_compeffects_excl}
+\begin{center}
+\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+Again, the decomposition between both components of transport costs brings valuable insights.
+Comparing Figure \ref{fig:TC_add_mult_compeffects_excl} with Figure \ref{fig:TC_addplusmult_compeffects_excl} (or Figure \ref{fig:totalTC_compeffects_excl}) indicates that the steady decline of overall transport costs hides shifts in the relative importance of additive and ad-valorem.
+Air transport in particular exhibit contrasted trends.
+While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost.
+These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
 
 \section{Conclusion \label{sec:conclu}}
+
+This paper empirically studies the magnitude of additive (or per-unit) costs in international transport costs, by exploiting the differences between the import and the export prices.
+Using SITC 3 and 4- digit cif-fob unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
+Our results may be summarized in three main findings.
+First, we obtain that additive costs are 2.8\% of fob unit values for ocean shipping (and ad-valorem ones 3.2\%).
+These values are respectively equal to 1.8\% and 2.5\% for air transport.
+Second, we show that taking additive costs into account improves the fit of the modelling of transport costs.
+All goodness-of-fit measures point out to this conclusion, which holds for both transport modes and all years considered.
+Third, we also use the time dimension of our data to characterize the evolution of transport costs.
+After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
+Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
+While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s.
+Further, the inclusion of additive costs yields to the conclusion that overall trade costs are decreased in air transport and ocean shipping roughly the same path and the same magnitude of order.
+This last result stands in contrast with related studies (\citealp{hummels2007} (that omit the role of additive costs), which rather obtain a larger decrease in air transport costs, relative to ocean shipping costs.
+\citealp{Behar_Venables}).
+In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+
+Our results could be extended in two main ways.
+On the empirical side, one may want to ge deeper in the ``structural'' determinants of trade costs, i.e.
+identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
+On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) as well as regarding the international transmission of business cycles.
+This is left for further research.
 
@@
 
-%%%Figures
-% Goodness-of-fit
+\section{Some detailed results}
 
-%%%%Tables
-%Goodness-of-fit
+In this section, we report the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_with_add})) or not (Equation (\ref{eq:model_nlI})).
+For sake of space saving, we only report the results for a sample of years.
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation.
+
 \begin{table}[htbp]
   \centering
-  \caption{Air: Measures of Goodness-of-fit}
+  \caption{Air: Dynamics of Trade Costs over Time, 3 digits (selected years)}
 \begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
+    \begin{tabular}{l|cccccc}
 \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
 
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}}\\
+Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\ \hline
+\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs}}     \\
+Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034  \\
+Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029  \\
+Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
 \hline
-\textbf{R$^{2}$} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.31} & \multicolumn{1}{c}{\textbf{0.31}} \\
-Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{\textbf{0.60}} \\
-\textbf{SER} &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & \multicolumn{1}{c}{\textbf{0.85}} \\
-Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & \multicolumn{1}{c}{\textbf{0.73}} \\
-\textbf{Log-likelihood} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & -17530.49 & -20253.49 & -29977.79 & -43341.27 & \multicolumn{1}{c}{-50746.83} & \multicolumn{1}{c}{-52690.29} & \multicolumn{1}{c}{\textbf{-34888.64}} \\
-Terms A \& I & -15125.65 & -17263.20 & -25393.46 & -36788.44 & \multicolumn{1}{c}{-47277.53} & \multicolumn{1}{c}{-49419.70} & \multicolumn{1}{c}{\textbf{-30508.29}} \\
-\textbf{AIC criteria} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{106130.58} & \multicolumn{1}{c}{\textbf{70498.08}} \\
-Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{100155.41} & \multicolumn{1}{c}{\textbf{62284.99}} \\
-\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       &  \\
-2$\times$(ll(UR) -ll(R)) & 4809.68 & 5980.59 & 9168.67 & 13105.67 & \multicolumn{1}{c}{6938.61} & \multicolumn{1}{c}{6541.17} & \multicolumn{1}{c}{\textbf{8760.69}} \\
-\# restrictions  & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426.00} & \multicolumn{1}{c}{422} & \multicolumn{1}{c}{\textbf{401.93}} \\
-p-value & 0.000 & 0.000 & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{\textbf{}} \\
-
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}    \\
+\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &      \\
+\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010  \\
+\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005  \\
+\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
+\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &        \\
+\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017  \\
+\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 \\
+Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\ \hline
+\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351  \\ 
 \hline\hline
-   % \bottomrule
-
     \end{tabular}%
   \end{center}
-  \label{tab:good_fit_air}%
-  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
+  \label{tab:result_air_3d_detail}%
+\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
 \end{table}%
@@
   \centering
-  \caption{Vessel, Measures of Goodness-of-fit}
+  \caption{Vessel: Dynamics of Trade Costs over Time (3 digits), selected years}
 \begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
+    \begin{tabular}{l|cccccc}
    \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \textbf{Mean stat} \\
+Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
  \hline
-\textbf{R$^{2}$}\unboldmath{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &  \\
-Term I only & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{0.34} & \multicolumn{1}{c}{\textbf{0.39}} \\
-Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.56}} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.76} & \multicolumn{1}{c}{\textbf{0.66}} \\
-Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{\textbf{0.57}} \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{-16287.40} & \multicolumn{1}{c}{-16129.13} & \multicolumn{1}{c}{-25169.31} & \multicolumn{1}{c}{-35263.95} & \multicolumn{1}{c}{-41998.95} & \multicolumn{1}{c}{-43692.93} & \multicolumn{1}{c}{\textbf{-28534.30}} \\
-Terms A\& I & \multicolumn{1}{c}{-12985.76} & \multicolumn{1}{c}{-13353.65} & \multicolumn{1}{c}{-21171.37} & \multicolumn{1}{c}{-29490.96} & \multicolumn{1}{c}{-37418.66} & \multicolumn{1}{c}{-39751.86} & \multicolumn{1}{c}{\textbf{-24151.31}} \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{33328.81} & \multicolumn{1}{c}{33010.27} & \multicolumn{1}{c}{51142.62} & \multicolumn{1}{c}{71365.89} & \multicolumn{1}{c}{84789.89} & \multicolumn{1}{c}{88191.87} & \multicolumn{1}{c}{\textbf{57848.60}} \\
-Terms A \& I & \multicolumn{1}{c}{27331.52} & \multicolumn{1}{c}{28067.31} & \multicolumn{1}{c}{43664.74} & \multicolumn{1}{c}{60475.91} & \multicolumn{1}{c}{76161.33} & \multicolumn{1}{c}{80873.72} & \multicolumn{1}{c}{\textbf{49683.32}} \\
-\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{\textbf{}} \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & \multicolumn{1}{c}{9160.56} & \multicolumn{1}{c}{7882.15} & \multicolumn{1}{c}{\textbf{8765.97}} \\
-\# restrictions  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{427} & \multicolumn{1}{c}{\textbf{417.05}} \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-
+   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       \\
+Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036  \\
+Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033  \\
+Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018  \\
+\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &     \\
+\hline
+\textit{Additive term } &       &       &       &       &       &        \\
+Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015  \\
+Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 \\
+Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
+\textit{Iceberg term} &       &       &       &       &       &        \\
+Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 \\
+Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018  \\
+Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
+\hline
+ \# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\ \hline
 \hline\hline
-   % \bottomrule
-
     \end{tabular}%
   \end{center}
-  \label{tab:good_fit_ves}%
-  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of Regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
+  \label{tab:result_ves_3d_detail}%
+\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
 \end{table}%
 
-%
+\begin{table}[htbp]
+  \centering
+  \caption{Air: Dynamics of Trade Costs over Time, 4-digit}
+\begin{center}
+    \begin{tabular}{l|cccccc}
+   \hline\hline
+Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\
+\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
+Mean  & 1.066 & 1.058 & 1.052 & \multicolumn{1}{c}{1.033} & \multicolumn{1}{c}{1.037} & \multicolumn{1}{c}{1.032} \\
+Median & 1.052 & 1.044 & 1.041 & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.027} & \multicolumn{1}{c}{1.026}  \\
+Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
+\hline
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}  \\
+\textit{Iceberg term} &       &       &       &       &       &     \\
+Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016}  \\
+Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014}  \\
+Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\ 
+\textit{Additive term} &       &       &       &       &       &     \\
+Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} \\
+Median & 0.012 & 0.006 & 0.006 & \multicolumn{1}{c}{0.005} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{0.004}  \\
+Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\ \hline
+\# obs & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
+\hline\hline
+    \end{tabular}%
+  \end{center}
+  \label{tab:result_air_rob}%
+\tiny{Notes:Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
+\end{table}%
 
-\appendix
+\begin{table}[htbp]
+  \centering
+\caption{Vessel: Dynamics of Trade Costs over Time, 4-digit}
+\begin{center}
+    \begin{tabular}{l|cccccc}
+   \hline\hline
+Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ 
+\hline
+\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs}} \\
+Mean  & 1.098 & 1.061 & \multicolumn{1}{c}{1.058} & \multicolumn{1}{c}{1.051} & \multicolumn{1}{c}{1.042} & \multicolumn{1}{c}{1.036}  \\
+Median & 1.094 & 1.051 & \multicolumn{1}{c}{1.048} & \multicolumn{1}{c}{1.045} & \multicolumn{1}{c}{1.038} & \multicolumn{1}{c}{1.031} \\
+Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
+\hline
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs } }\\
+\textit{Iceberg term} &       &       &       &       &       &     \\
+Mean  & 1.054 & 1.034 & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.021}  \\
+Median & 1.049 & 1.030 & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.025} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{1.018} \\
+Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
+\textit{Iceberg term} &       &       &       &       &       &     \\
+Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015}  \\
+Median & 0.029 & 0.013 & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{0.013} & \multicolumn{1}{c}{0.008} \\
+Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\ \hline
+\# obs & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
+\hline\hline
+    \end{tabular}%
+  \end{center}
+  \label{tab:result_ves_rob}%
+\vspace{-10pt}\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
+\end{table}%
 
-\section{Results with a Higher Level of Products Disaggregation }
-%\vspace{-0.5cm}
+\section{Robustness checks \label{app:robust_4d}}
+
+In this section, we report the goodness-of-fit exercise conducted by transport mode, for selected years at the 4-digit product classification level.
+The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel), for a sample of some years.
 \begin{table}[htbp]
@@
 \end{center}
- \vspace{-10pt} \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
+ \vspace{-10pt} \tiny{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
 The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
@@
 
-\begin{table}[H]
+\begin{table}[htbp]
   \centering
@@
 \end{center}
-  \vspace{-10pt}\tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
+  \vspace{-10pt}\tiny{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
 The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
@@
 
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Dynamics of Trade Costs over Time, 4-digit}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
-
-\hline
-\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       &  \\
-Mean  & 1.066 & 1.058 & 1.052 & \multicolumn{1}{c}{1.033} & \multicolumn{1}{c}{1.037} & \multicolumn{1}{c}{1.032} & \multicolumn{1}{c}{\textbf{1.049}} \\
-Median & 1.052 & 1.044 & 1.041 & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.027} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{\textbf{1.037}} \\
-Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{\textbf{0.045}} \\
-\hline
-\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{} \\
-\textit{Additive term} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} & \multicolumn{1}{c}{\textbf{0.019}} \\
-Median & 0.012 & 0.006 & 0.006 & \multicolumn{1}{c}{0.005} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{\textbf{0.008}} \\
-Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{\textbf{0.034}} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016} & \multicolumn{1}{c}{\textbf{1.024}} \\
-Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014} & \multicolumn{1}{c}{\textbf{1.016}} \\
-Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{\textbf{0.026}} \\
-\hline
-\# obs & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460} & \multicolumn{1}{c}{\textbf{}} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air_rob}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\begin{table}[htbp]
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-  \centering
-\caption{Vessel: Dynamics of Trade Costs over Time, 4-digit}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
-\hline
-\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.098 & 1.061 & \multicolumn{1}{c}{1.058} & \multicolumn{1}{c}{1.051} & \multicolumn{1}{c}{1.042} & \multicolumn{1}{c}{1.036} & \textbf{1.060} \\
-Median & 1.094 & 1.051 & \multicolumn{1}{c}{1.048} & \multicolumn{1}{c}{1.045} & \multicolumn{1}{c}{1.038} & \multicolumn{1}{c}{1.031} & \textbf{1.052} \\
-Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} & \textbf{0.036} \\
-\hline
-\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{}\\
-\textit{Additive term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015} & \textbf{0.028} \\
-Median & 0.029 & 0.013 & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{0.013} & \multicolumn{1}{c}{0.008} & \textbf{0.017} \\
-Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} & \textbf{0.039} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.054 & 1.034 & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.021} & \textbf{1.033} \\
-Median & 1.049 & 1.030 & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.025} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{1.018} & \textbf{1.028} \\
-Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013} & \textbf{0.025} \\
-\# obs & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} & \textbf{}\\
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves_rob}%
-\vspace{-10pt}\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
 \end{document}
-
-In section \ref{sec:determinants}, we provide rationales for those differences.
-
-\section{Reassessing the Traditional Determinants of Transport Costs}\label{sec:determinants}
-
-\subsection{Methodological Concerns}
-
-%In step 2, we ask if the traditional determinants of transportation costs impact differently the multiplicative and additive costs estimated in the first step.
-%Therefore, we consistently focus only on origin-country-specific transportation costs, taking the estimated $\tau_{i}$ and $t_{i}$ from step 1 as dependent variables.
-In line with the trade literature, we proxy transport costs by (log of) distance and other proxies for trade costs (mainly, geographic and cultural factors).\interfootnotelinepenalty=10000\footnote{All those variables come from the well-known GeoDist database built by the CEPII: \url{http://www.cepii.fr/CEPII/fr/bdd_modele/presentation.asp?id=6}.} More formally, we estimate, still on a yearly basis and for each transportation mode:
-
-%\begin{eqnarray}
-%ln(\widehat{\tau}_{i}) &=& \beta_{1} \textrm{ln(distance)} + \Gamma_{1} \textrm{Gravity} + \mu_{i} \label{eq:step2_iceberg}\\
-%ln(\widehat{t}_{i}) &=& \beta_{2} \textrm{ln(distance)} + \Gamma_{2} \textrm{Gravity} + \nu_{i} \label{eq:step2_additive}
-%\end{eqnarray}
-
-%where the ln operator denotes that variables have been log-linearized, $distance$ is the bilateral distance between US and the origin country $i$  and $Gravity$ is a vector of controls including the languages spoken in the country under different definitions, whether the country is landlocked, and their colonial links.
-
-In this second stage of the paper, we thus keep on questioning the way the literature models the trade costs determinants.
-Do the proxies frequently used as measures for trade costs, typically distance, correspond more to additive cots or to multiplicative trade costs?
-Or total?
-Usually, people use a measure of distance as a way to capture iceberg trade costs.
-What is the size of the error we make (what is the size of the approximation) when we suppose that distance is a good proxy of (total) transport costs?
-Using our decomposition of transport costs obtained in the first stage, we can assess the contributions of typical gravity variables to each component of transport costs (additive / iceberg).
-To put this simply, consider first our ``raw'' estimate of trade costs (meaning, only modeling a multiplicative form); we estimate the role of distance (as quite frequent) and find that this fits reasonably well, the estimated coefficient stands in the average.
-We then apply the same regression on each multiplicative / additive component of trade costs.
-Suppose that we find a high correlation with the additive cost, and almost 0 with the multiplicative cost.
-This means that, using this proxy to capture the size of iceberg trade costs amounts making a sizeable approximation error.
-\bigskip
-
-\paragraph{Data treatment} Before presenting the estimated equation, let us mention how we deal with the data.
-We regress 1°) the ``overall trade costs'' (i.e.), modeled as an iceberg transport cost (denoted $\tau_{ik}^{nlI}$ hereafter ($i$ for the origin country, $k$ for the product and $nlI$ for the ``non-linear iceberg'' estimation method) and 2°) the two additive and multiplicative transport costs, $\tau_{ik}$ and $t_{ik}$ respectively, on some common determinants.
-Precisely, we re-treat the iceberg component (both $\tau_{ik}^{nlI}$ and $\tau_{ik}$) to have it expressed in percentage of the fob price, ie we build the variable $\tau_{ik} =  100(\widehat{\tau}_{ik}-1)$ (with $\widehat{\tau}_{ik}$ the estimated variable obtained at stage 1).
-As for the additive cost, from stage 1 of the paper, we have obtained the value of (country-specific) trade cost as a fraction of the fob price (denoted $\widehat{t}_{ik}$); we multiply it to have it expressed as a percentage of the fob price (considering $t_{ik} =  100 \widehat{t}_{ik}$).
-Further, and in line with our first-step estimation method, we estimate the determinants of trade costs on a yearly basis, and conditional to the transport mode (air or vessel).
-
-In the three cases, we want to explain trade costs expressed as a percentage of the fob price.
-All our three measures of trade costs are bounded by 0 as minimal value, and we drop outliers by only keeping the 95 first percentile of each distribution.
-\bigskip
-
-\paragraph{Functional form} To derive the estimated equation, we start recalling what is behind the gap between cif-fob prices.
-In the US import data we use, the cif price is based on applying import charges on the custom value.
-What is defined as import charges?
-As detailed in Appendix \ref{app:data}, ``import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
-import duties) incurred in bringing the merchandise from alongside the carrier''.
-So in fact what we call ``fob price'' is ``fas price'', for ``free alongside''.
-This means that the seller has paid to bring the goods to the port, but not to put them on the boat.
-In this respect, all hauling charges are excluded from what we call the ``fob'' price.
-In comparison with the fas price, the cif price is ``cost, insurance, freight'' included.\footnote{From OECD stats, the c.i.f.
-price (i.e.
-cost, insurance and freight price) is the price of a good delivered at the frontier of the importing country, including any insurance and freight charges incurred to that point, or the price of a service delivered to a resident, before the payment of any import duties or other taxes on imports or trade and transport margins within the country.}
-
-Accordingly, if we denote $TC_{ik}$ the total transport cost paid to import the product $k$ from country $i$ (on a given year, conditional to a certain transport mode), expressed in USD, we can decompose it in three elements:
-
-$$TC_{ik} = \text{Freight + Insurance + Handling costs}$$
-
-Our second-stage estimation is then guided by two questions.
-For each of these three dimensions (cost, insurance, freight) at the root of the cif-fas prices gap, 1°) what are its determinants, and 2°) do we expect this dimension to play rather on the additive or the multiplicative dimension?
-
-\paragraph{Insurance}: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ???
-for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood).
-The database provides us with the quantity (in terms of kilograms) of the shipment ($q_{ik}$), as well as the price, so that we have the value of the shipment.
-However, we have no dedicated variable to measure the insurance cost in the database.
-Accordingly, we capture this dimension by including product fixed effects.
-Insurance grids retain a degree of product categorization much broader than the 3-digit level we retain as benchmark sectorial degree (our $k$ index for the product dimension).
-In most pieces of information we have collected, insurance grids adopt a ranking in \textbf{20 (??)} types of products.
-Accordingly, we include product fixed effects of dimension \textbf{20}, that we refer to with the subscript $k^\prime$, i.e.
-$FE_{k^\prime}$.
-If we denote $p_{ik}^{fob}$ and $q_{ik}$ the (fas) price and the quantity of the shipment, total insurance costs (in USD) are accordingly measured by $FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}$.
-As insurance cost depend on the value of the shipment, we expect it to be proportional to the price, ie rather play on the iceberg part of the transport costs.
-
-\paragraph{Costs} (for handling costs).
-The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs).
-If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs.
-We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD.
-As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$).
-Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular).
-We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following.
-Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$.
-As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.
-
- \paragraph{Freight}.
-Quite intuitively, freight costs should depend on the distance between the origin country and the U.S.
-(denoted $dist_i$).
-Further, they likely depend on the quantity ($q_{ik}$) and the bulk of the shipment.
-Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs.
-Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$.
-Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.
-
-As a result, total transport costs by $TC_{ik}$ (expressed in USD) can be decomposed in the three components of insurance, handling costs and freight, according to:
-
-$$TC_{ik} =  \underbrace{EC_i\times V_{k}\times q_{ik}}_{\text{Costs}} +\underbrace{FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}}_{\text{Insurance}}+ \underbrace{V_k\times dist_i\times q_{ik}}_{Freight}$$
-
-By definition, such transport costs can be written as $TC_{ik} = (p^{cif}_{ik} - p^{fas}_{ik})q_{ik}$. Accordingly, we can write:
-
-\begin{eqnarray*}
-&&(p^{cif}_{ik} - p^{fas}_{ik})q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_{k}\times dist_i\times q_{ik} \\
-\Leftrightarrow && \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)p^{fas}_{ik} q_{ik} = EC_i\times V_{k}\times q_{ik}+ FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}+ V_k \times dist_i\times q_{ik}
-\end{eqnarray*}
-
-From the above equation, it comes that $\frac{TC_{ik}}{p^{fas}_{ik} q_{ik}} = \left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right)$, which is exactly the measure of trade costs obtained in the first stage.\footnote{Recall that our first stage starts from the equation: $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau^{nlI}_{ik}$ in the ``raw'' equation and $\left(\frac{p^{cif}_{ik}}{p^{fas}_{ik}} -1 \right) = \tau_{ik}+ \frac{t_{ik}}{p^{fas}_{ik}}$ when decomposing between iceberg and additive transport costs.}
-
-Accordingly, we can decompose each of our three measures of trade costs ($\tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, expressed in the three cases as a percentage of the fob price, and denoted $tc_{ik}$ to gain in generality, according to:
-
-\begin{equation}
-tc_{ik} = \underbrace{\frac{EC_i\times V_{k}}{p^{fas}_{ik}}}_{\text{C}} + \underbrace{FE_{k^\prime}}_{\text{I}}+ \underbrace{\frac{V_k\times dist_i}{p_{ik}^{fas}}}_{\text{F}} \label{eq:theory}
-\end{equation}
-
-Based on Equation (\ref{eq:theory}), we specify the estimated equation as follows:
-
-\begin{equation}
-tc_{ik} = \left[\frac{EC_i\times V_k}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k\right]e^{\varepsilon_{ik}} \label{eq:estim}
-\end{equation}
-
-with $\varepsilon_{ik}$ the error term, centered on 0.
-This calls for two comments.
-First, we add the country fixed effects separately in the regression, in order to correctly identify the associated interaction term.
-Second, we specify the error term multiplicatively.
-If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points).
-In this case, it is very likely that the variance of the error term depends on the size of the observed costs.
-E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high.
-Adopting this functional form would then be subject to a potential bias in point estimates due to the heteroscedasticity of error terms.
-This drives us to specify the error term in a multiplicative manner (such that the error term is in percentage of each explicative variable and centered on 0), and to take the log of the equation.
-
-Accordingly, the functional form that we estimate, for each of the three transport costs obtained at the first stage ($tc_{ik} = \tau^{nlI}_{ik}$, $\tau_{ik}$ and $\frac{t_{ik}}{p^{fas}_{ik}}$, conditional on the year and the transport mode), is expressed as:
-
-\begin{equation}
-\log tc_{ik} = \log\left[\frac{EC_i\times V_{k}}{p^{fas}_{ik}} + FE_{k^\prime}+ \frac{V_k\times dist_i}{p_{ik}^{fas}} + EC_i + V_k \right]+ \varepsilon_{ik} \label{eq:estim1}
-\end{equation}
-
-%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
-So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ...
-\bigskip
-
-%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !
-
-\subsection{Results and Interpretation}
-%\section{One step further into interpretation: transportation costs as an indicator of market structure.}
-
-\end{document}

### 2016-07-22 redaction/dhp_subETSG.tex
Version sans gras soumise ETSG

--- 2016-07-22 redaction/dhp_draft_july2016.tex
@@
 Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
-\textbf{comparer avec la litterature?}.
 Second, when decomposing transport costs into an additive and a multiplicative component, we find that the per-unit cost dimension is sizeable.
@@
 {\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{Fig2a_decompTC_air_3d.pdf}
-& \includegraphics[width=3.5in,height=3in]{Fig2b_decompTC_vessel_3d.pdf} \\
+\includegraphics[width=3in, height=2.5in]{Fig2a_decompTC_air_3d.pdf}
+& \includegraphics[width=3in,height=2.5in]{Fig2b_decompTC_vessel_3d.pdf} \\
 \end{tabular}
@@
 The first indicator is through comparing $R^{2}$.
-However, its use is far from being straightforward when evaluating non-linear estimates.
-\textbf{mettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
+However, its use is far from being straightforward when evaluating non-linear estimates.\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
 In a non-linear context, R-squared is therefore inappropriate, strictly speaking.
@@
 As well, the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air.
-\textbf{et pourtant, la part des additifs dans vessel est plus elevée, 48.2\% en moyenne vs 42.3\% pour AIr, cf le .xls.
-A dire?
-Faire reference a la figure 2.}
 
@@
 Rather, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the iceberg term.
-\textbf{so what?
-pas tres punchy}\smallskip
+\smallskip
 

### 2016-07-25 redaction/dhp_draft_july2016.tex
Modifications Guillaume

@@
 %\usepackage{hyperref}
+%\usepackage[T1]{fontenc}
+%\usepackage[utf8]{inputenc}
+%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
+%J'ai essayé de débuger : je n'y suis pas arrivé
 \usepackage{rotating}
@@
 \usepackage{ upgreek }
+\usepackage{comment}
+\includecomment{commentGD}
+\usepackage[draft]{todonotes}
+%%GDPour cacher les notes, \usepackage[disable]{todonotes}
+
 %\usepackage[nolists, figuresfirst]{endfloat}
@@
 
-\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}
+\setmarginsrb{3cm}{2cm}{3cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}
 
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris Dauphine, LEDa \& PSL, Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine, LEDa \& PSL} }
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE \newline
+SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), 75007, PARIS, FRANCE}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE \newline
+ } }
  \maketitle
@@
 \begin{abstract}
-Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad valorem tax equivalent.
+Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad-valorem tax equivalent.
 However, many common empirical facts support the existence of additive costs.
-This paper contributes to this debate.
-Precisely, we provide a quantitative assessment of the size and the importance of the additive component in international transport costs.
-Using SITC 3 and 4 digit cif-fob unit values over 1974-2013 taken from US imports data, we estimate the two components of transport costs, by transport mode (air or ocean).
-We find that additive costs are 2.85\% of fob unit values for ocean shipping (and ad-valorem ones 3.22\%).
+This paper provides a quantitative assessment of the size and the importance of the additive component in international transport costs.
+Using SITC 3 and 4 digit cif-fab unit values over 1974-2013 taken from US imports data, we estimate the two components of transport costs by transport mode (air or ocean).
+We find that additive costs are 2.85\% of fab unit values for ocean shipping (and ad-valorem ones 3.22\%).
 These values are respectively equal to 1.9\% and 2.5\% for air transport.
@@
 
-\noindent Trade costs remain central in international economic analysis.
-In particular, they are considered as a major obstacle to international economic integration and international trade flows.
-According to the estimates by \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
+\noindent Trade costs are central in international economic analysis.
+In particular, they are considered a major obstacle to international economic integration and international trade flows.
+According to \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
 After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements.
-Based on panel data, \citet{novy13} finds  that U.S.
-trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
+Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
 Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs (typically captured through distance) still remain a major obstacle to trade (e.g.
 \citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
-Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries put up to 74\% markups over production costs.
+Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
 Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-In this vein, \citet{anderson_wincoop_jel} obtain that the 74\% markup made by international trade costs can be divided in 44\% for border-related trade barriers and 21\% for transportation costs.
+In this vein, \citet{anderson_wincoop_jel} obtain that the 74\% markup made by international trade costs can be divided between 44\% for border-related trade barriers and 21\% for transportation costs.
 This last figure suggests that the share of transport costs in the consumer price of manufactured goods is substantial.
 \citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
+\todo{GD 230716 Euh... je crois que c'est un non-sequitur, non ?}
 If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
@@
 
-Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, the ``iceberg cost'' hypothesis).
+Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).
+\todo{GD 230716 iceberg, c'est deux choses : ad-valorem + le fait que les coûts sont payés avec des bouts du bien transporté}
 However, many common empirical facts support the existence of additive costs.
@@
 often exhibit (at least partly) an additive structure.
-The structure (additive vs multiplicative) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows.
-The Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
-\citet{Lashkaripour-mimeo-2016} challenges this view though.
-He finds supporting evidence for the iceberg assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
+The structure (additive vs ad-valorem) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows.
+The Alchian and Allen conjecture (\citealp{alchian}), points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
+\citet{Lashkaripour-mimeo-2016} challenges this view.
+He finds supporting evidence for the \emph{ad-valorem} assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
 One can yet be concerned by the generality of his results though.
-By nature, his study is restricted to goods that are enumerated by items in the statistics (that represent 60\% of US imports for instance).
+By nature, his study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports).
 Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
 Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fob unit values on exports to more remote countries, in contradiction with the iceberg hypothesis.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fab unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
 \citet{hummels_skiba} estimate the elasticity of freight rates with respect to price is estimated to be well below unity.
-Also, their estimates imply that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
+Also, their estimates imply that doubling freight costs increases average fab export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
 These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
@@
 With additive costs, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments, thereby inducing more ``lumpiness'' in international trade.
-\smallskip
+\todo{GD 230716 pas sur que cela soit le sujet. Plutôt «per shipment» / «fixed costs» que «additive»} \smallskip
 
-Beyond the positive aspect of understanding trade patterns, several recent papers also point out the normative implications of additive trade costs.
-\citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the iceberg component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs.
-Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff.
-While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much can be done in quantifying such gains though, one potential reason being the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap.
+Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+\citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
+A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.
+Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
+While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
+One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap.
 \bigskip
 
-Our paper contributes to the literature by providing results on the size of transport costs over time, explicitly distinguishing between multiplicative and additive parts.
+Our paper contributes to the literature by providing results on the size of transport costs in long run, explicitly distinguishing between ad-valorem and additive parts.
 Precisely, we quantitatively assess the size and the importance of the additive component in international transport costs.
 To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013.
-Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian exports in 2004, their results suggest that additive costs are about 14\% of the median consumer price.
-However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects, as we develop below.
-One important difference is that we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013.
+Closely related to our paper is the work by \citet{Irrazabal_2015}, that develops a structural framework for inferring additive trade costs from firm-level trade data.
+Based on Norwegian exports in 2004, this work finds that additive costs are about 14\% of the median consumer price.
+However our approach departs from theirs in several key aspects, as we develop below.
+Our data are restricted to transport, insurance and handling costs.
+Still, we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
@@
 
-First, our theoretically agnostic approach provides a fairly simple framework for measuring both multiplicative and additive parts of transportation costs, which represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the multiplicative cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median fob price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the iceberg and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fob price $\widetilde{p}_{ik}$.} We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% in air and ocean respectively, whereas the additive component amounts to 1.8\% and 2.9\% of the fob price (in air and ocean).
-To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both multiplicative and additive costs in total transport costs.
+First, our theoretically agnostic approach provides a fairly simple framework for measuring both ad-valorem and additive parts of transportation costs.
+This represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median fab price by country-product ($\bar{\widetilde{p}}_{ik}$).
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fab price $\widetilde{p}_{ik}$.} We thus obtain that the mean values over 1974-2013 of ad-valorem costs are equal to 2.5\% and 3.2\% of the fab price in air and ocean transport respectively, whereas the additive component amounts to 1.8\% and 2.9\% .
+To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
+
 Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
-On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
+Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
+On average over the whole period, biased estimate for ad-valorem is 5\% for air and 6\% for vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
 We also show that taking additive costs into account significantly improves the fit of the modelling of transport costs, through various measures of ``goodness-of-fit'', for both transport modes and years.
@@
 However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004.
-Concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}.
+Concerning ocean shipping, which represents the major part of world trade, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}.
 Our paper contributes to this debate.
@@
 After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
-Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
+Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem costs.
 Air transport in particular exhibit contrasted trends.
-While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost.
+While most of the early decline in air transport costs can be explained by ad-valorem costs, this component nearly doubled in the 2000s, all along with a substantial decline in the per-kg cost.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
@@
 The paper is built as follows.
-\ref{sec:data_method} explains the data sources and the empirical methodology retained in the paper.
+\ref{sec:data_method} presents our data and our  methodology.
 Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results.
 In Section \ref{sec:results_decomposition}, we characterize the role of the additive component of transport costs.
-After reporting the mean values over the period (by transport mode), we show the improved performance in including additive trade costs in the measure of transport costs.
-This being established, Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2012.
+After reporting the mean values over the period (by transport mode), we show how introducing additive transport costs improves the modelling of transport costs.
+Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2013.
 Section \ref{sec:conclu} concludes.
@@
 As in \cite{hummels2007} (among others), our measure of transportation costs consists in exploiting the difference between commodity-level export and import prices.
-We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode.
-More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs.
+We first use customs values, quantities and freight costs to recover free-alongside-board (FAB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode.
+More precisely, the (unit) FAB price is computed as the total "customs value" in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
 The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price.
-Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.
+Our dependant variable is finally computed as the ratio of the CIF price divided by the FAB price.
+Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
 The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
 It is worth acknowledging that using this dataset encounters some limitations.
-First, our measure of the cif-fob price gap only covers transportation costs by nature, thereby being silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
+First, our measure of the cif-fab price gap only covers freight, insurance and handling costs.
+It is silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
 Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit.
@@
 
-Yet, we believe that the results we get deliver valuable insights on this topic.
+Yet, we believe that our deliver valuable insights on this topic.
 Indeed, using this dataset has (at least) three main advantages.
-First, this dataset delivers a strong statistical reliability arising from a single, trustworthy customs origin.
+First, this dataset delivers a strong statistical reliability arising from a single, trustworthy, customs origin.
 Based on customs declarations, this dataset inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
-In addition, the database reports information regarding freight expenditures and transportation mode (Ocean vessel and Air).
-The first will be crucial to compute transport costs (see below), the second will allow us enlightening substantial differences in the dynamics of transport costs across transportation mode.
-Second, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FOB price).
-This is highly valuable, as we can estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs.
-This differentiates us from \citet{Irrazabal_2015}, which can only estimate the ratio of additive costs as a share of the median total consumer price.
-Third, this dataset having been used by \cite{hummels2007}, this enables us to compare our results to his findings.
+In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode (ocean (or "vessel") and air).
+The first will be crucial to compute transport costs (see below), the second will allow to enlight substantial differences in the dynamics of transport costs across transportation mode.
+Second, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FAB price).
+This is highly valuable, as we can estimate both the \textit{levels} of the ad-valorem trade costs and of the additive trade costs.
+This differentiates us from \citet{Irrazabal_2015}, who can only estimate the ratio of additive costs as a share of the median total consumer price.
+Third, this dataset has been used by \cite{hummels2007}, this enables us to compare our results to his findings.
 Further, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004.
@@
 
-We make use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level.
-This is primarily due to technical reasons.
-As detailed below, the use of a nonlinear estimator triggers computational limitations that do not make them a likely option, especially when covering a long period of time.
-Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level (for some selected years).\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.} Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
+We use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level.
+As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.} Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
 Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.
@@
 
-\paragraph{The estimated equation} Our purpose is to provide estimates over time of the sizes of multiplicative and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price paid by the consumer (import, or cif) price $p$ as a function of the producer (export, or fob) price $\widetilde{p}$ given both per-unit ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}
+\paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
+To do so, we start from the equation that expresses the price $p$ paid by the importer
+\todo{GD 240716 importer plutôt que consumer : il reste des coûts entre les deux...}
+ (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fab), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}
 
@@
 
-\noindent As usual in the literature, the so-called ``iceberg'' trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no iceberg trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive trade costs).
-We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.} Let us denote $i$ the origin country, and $k$, the sector (or product).
-Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:
+\noindent As usual in the literature, the so-called ``ad-valorem'' trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive trade costs).
+Let us denote $i$ the origin country, and $k$, the sector (or product).
+Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}:
 
@@
 
-\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1°)} both multiplicative and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} in a multiplicative way for the former and an additive way for the latter.
+We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.
+
+\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.
 In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:
@@
 \begin{eqnarray}
-\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:iceberg}\\
+\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:ad-valorem}\\
 t_{ik} &=& t_{i} + t_{k} \label{eq:add}
@@
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the cif price $p$ cannot be lower than the fob price ($p>\widetilde{p}_{ik}$).
-Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term, according to:
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fab price ($p>\widetilde{p}_{ik}$).
+Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:
 
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \varepsilon_{ik}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
 \end{equation*}
-\noindent Taking in log, this finally drives us to estimate the following equation:
+\noindent Where $\epsilon_{ik}$ follows an normal law centered on 0. Taking the logs, this drives us to estimate the following equation:
 
@@
 
-The non-linearity of Equation (\ref{eq:estimatedequation}) implies that estimations cannot be performed using standard linear estimators.
+The non-linearity of Equation (\ref{eq:estimatedequation}) implies that it cannot be estimated using standard linear estimators.
 Therefore, all estimates are performed using non-linear least squares.
@@
 The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
-In our case, due to computational limitations implied by the size of our dataset, we implement 100 iterations and set the convergence criterion for successive parameter estimates and for the residual sum of squares at 0.01.
-Finally, to eliminate the potential influence of outliers, we excluded observations in the 5 percent from the upper and lower tails of the distribution in the regression variables, and all our three measures of trade costs are bounded by 0 as minimal value.
+% In our case, due to computational limitations implied by the size of our dataset, we implement 100 iterations and set the convergence criterion for successive parameter estimates and for the residual sum of squares at 0.01.
+%GD : cela me semble être de la cuisine sans grand intérêt.
+Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables
+%%and all our three measures of trade costs are bounded by 0 as minimal value. 
+%%Pas compris. Parfois un, non ?
 These cut-offs are aimed at eliminating reporting or coding errors.\medskip
 
-Further, one key objective of the paper is to characterize the importance of additive costs relatively to iceberg costs.
+Further, one key objective of the paper is to characterize the importance of additive costs relatively to ad-valorem costs.
 Put differently, what traditional models of international trade lose by ignoring additive costs?
@@
 
-After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
+After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
+When assuming ad-valorem costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$
+\todo{GD 240716 je n'aime pas beaucoup l'abbréviation nlI, mais je n'ai pas de meilleure idée...}.
+\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}
+\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}
+Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost $\widehat{\tau}$ and $\widehat{t}$, for each year and transportation mode.
@@
 First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (\ref{eq:estimatedequation}).
-Second, we evaluate the importance of the per-unit component of in overall transport costs through the means of goodness-of-fit measures.
+Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.
 
@@
 
-Our first contribution to the literature is to provide estimates for the size of both the multiplicative and the additive components of transport costs.
+Our first contribution to the literature is to provide estimates for the size of both the ad-valorem and the additive components of transport costs.
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs (multiplicative estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix , we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
-Results for all years are of course, available upon request to the authors.}$^{,}$\footnote{We present the results for Air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
+It displays the mean and median values of each type of trade costs (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix , we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+Results for all years are of course, available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
 Overall results (over the whole period) are not substantially affected if this year is included though.
@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{With only Iceberg Trade Costs }  \\ \hline
+    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs }  \\ \hline
     Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
@@
     Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{With Additive \& Iceberg Trade Costs } \\ \hline
-   \textit{Iceberg term} & & & & \\ \hline
+    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
+   \textit{Ad-valorem term} & & & & \\ \hline
     Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
@@
     \end{center}}
-\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
+\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.
+The additive term is expressed in fraction of fab price.
 ($^\ast$): Four 4-digit estimation: 0n selected years.
-($^{\ast \ast}$): 1989 omitted in 3 digit estimation for Air.}}
+($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
 \end{table}%
@@
 Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
-\textbf{comparer avec la litterature?}.
-Second, when decomposing transport costs into an additive and a multiplicative component, we find that the per-unit cost dimension is sizeable.
-Over the whole period, additive costs are 2.85\% of fob unit values for ocean shipping  - and ad-valorem ones being equal to 3.22\%.
+\todo{LP comparer avec la litterature?}.
+Air transport costs are smaller than sea transport costs.
+This is counter-intuitive, but certainly linked to selection effects.
+Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
+Over the whole period, additive costs are 2.85\% of fab unit values for ocean shipping  - and ad-valorem ones being equal to 3.22\%.
 These values are respectively equal to 1.9\% and 2.5\% for air transport.
-For both transport modes, the omission of the additive term seriously biases the multiplicative term upward.
-In quantitative terms, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.
+For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
+In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
 
 We investigate the robustness of this conclusion on a yearly basis.
-Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the multiplicative component, estimated alone (Equation (\ref{eq:model_nlI}), dashed line) and along with the additive component (Equation (\ref{eq:model_with_add}, plain line), by transport mode.
+Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Equation (\ref{eq:model_nlI}), dashed line) and along with the additive component (Equation (\ref{eq:model_with_add}, plain line), by transport mode.
+
+\todo{GD 250716 préciser mean or median?}
 
 \begin{figure}[htbp]
-\caption{Multiplicative Costs (3 digits}\label{fig:mult_alone_withadd}
+\caption{Ad-valorem Costs (3 digits)}
+\label{fig:mult_alone_withadd}
 \begin{center}
@@
 Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air transport.
-To get a better picture on this, we study the shares of both multiplicative and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.
+To get a better picture on this, we study the shares of both ad-valorem and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.
+
+\todo{GD 250716 préciser mean or median?}
 
 \begin{figure}[htbp]
-\caption{Decomposing Transport costs (3 digits}\label{fig:decomp_TC_3d}
+\caption{Decomposing Transport costs (3 digits)}
+\label{fig:decomp_TC_3d}
 \begin{center}
@@
 First, the magnitude of transport costs is higher for ocean shipping than for air transport.
-This result also holds considering iceberg costs alone (Figure \ref{fig:mult_alone_withadd}).
+This result also holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}).
 Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
@@
 
-\subsection{Assessing the importance of per-unit trade costs}
+\subsection{Assessing the importance of per-kg trade costs}
 
-In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fob prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs.
+In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fab prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs.
 To do so, we rely on several standard measures of fit.
@@
 However, its use is far from being straightforward when evaluating non-linear estimates.
-\textbf{mettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
-In a non-linear context, R-squared is therefore inappropriate, strictly speaking.
+\todo{LPmettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
+In a non-linear context, R-squared is  strictly speaking inappropriate.
 However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.}.
@@
 
-Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_ves}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level).
+Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level).
 The results for all years are available upon request to the authors.}\medskip
@@
 
-Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_ves} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
-On average over the whole period, the $R^{2}$ doubles when per-unit costs are included for Air, and increases by 50\% for Vessel.
+Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
+On average over the whole period, the $R^{2}$ doubles when per-kg costs are included for Air, and increases by 50\% for Vessel.
 Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
@@
 Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences.
-First, multiplicative costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes).
-As well, the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air.
-\textbf{et pourtant, la part des additifs dans vessel est plus elevée, 48.2\% en moyenne vs 42.3\% pour AIr, cf le .xls.
+First, ad-valorem costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FAB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes).
+The standard error of the regression based only on the ad-valorem term is consistently lower for Vessel than for Air.
+\todo{LP et pourtant, la part des additifs dans vessel est plus elevée, 48.2\% en moyenne vs 42.3\% pour AIr, cf le .xls.
 A dire?
@@
 
-Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000, for both transport modes.
-For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
+Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000 for both transport modes.
+For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term increases the explanatory power by 30\%, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
 The picture is quite different for Vessel.
 In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included.
-Rather, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the iceberg term.
-\textbf{so what?
+the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term.
+This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs \todo{LPso what?
 pas tres punchy}\smallskip
@@
 If anything, the quality of fitting appears slightly higher when estimations are based on the 4-digit classification.
-This is especially true for the model restricting trade cost to their iceberg dimension, whatever the transport mode considered.
+This is especially true for the model restricting trade cost to their ad-valorem dimension, whatever the transport mode considered.
 When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
 In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not with an extent that would disqualify the use of slightly more aggregated data.
-Further, the same conclusion emerges regarding the significant role of the additive components in fitting international transport costs, that established at the 3-digit level.
+Further, the same conclusion established at the 3-digit level regarding the significant role of the additive components in fitting international transport costs emerges at the 4-digit level.
 
@@
 In this section, we investigate the role of the additive component in total transport costs by exploring the time dimension of our data.
-As first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
+As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and ad-valorem components in transport costs over time by transport mode.
 Let us consider first the trend in overall transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
@@
 First, both air and ocean shipping exhibit a downward trend in overall transport costs.
-Second, starting from 1974, the decrease appears more important for sea than air transportation: Transportation costs are divided by 3 for the ocean shipping, versus by 2 for air transport.
+Second, starting from 1974, the decrease appears more important for sea than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport.
 That said, it may be the case the oil shock inflated transportation costs that specific year.
@@
 These results are consistent with \citet{hummels2007}.
-If we now consider the respective shares of additive and multiplicative components, we observe that the share of additive costs in air transport costs has declined between 1974 and 2013, in particular starting in 1988 (panel (a) of Figure \ref{fig:decomp_TC_3d}).
+If we now consider the respective shares of additive and ad-valorem components, we observe that the share of additive costs in air transport costs has declined between 1974 and 2013, in particular starting in 1988 (panel (a) of Figure \ref{fig:decomp_TC_3d}).
 By contrast, this share has been roughly constant throughout the period in ocean shipping (panel (b)).
 
-Before making any definite statement about this though, it is worth emphasizing that the evolution of 
-transport costs depends both on the evolution of per product and per partner costs and on the evolution of the composition of trade.
-Put it simply, total transport costs may have decreased over time either the country either trades with countries with which trade is cheaper, or trades products which are cheaper to transport, independently of any change in tranport costs \textit{per se}.
+Before making any definite statement about this though, it is worth emphasizing that the evolution of transport costs depends both of the evolution of per product and per partner costs and of the evolution of the composition of trade.
+Total transport costs may have decreased over time either the country either trades with countries with which trade is cheaper, or trades products which are cheaper to transport, independently of any change in tranport costs \textit{per se}.
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
-In this aim, we estimate the following equation for the multiplicative cost, based on the iceberg cost series obtained when estimated alone ($\widehat{\tau}_{ikt}^{nlI}$ and the iceberg cost series estimated along with additive costs ($\widehat{\tau}_{ikt}^{ice}$:\footnote{We thus estimate Equation (\ref{eq:compeffects_mult}) twice, depending on the type of iceberg cost series considered ($\widehat{\tau}_{ikt}^{nlI}$ or $\widehat{\tau}_{ikt}^{ice}$), even if it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
+To do this, we estimate the following equation for the ad-valorem cost, based on the ad-valorem cost series obtained when estimated alone ($\widehat{\tau}_{ikt}^{nlI}$ and the ad-valorem cost series estimated along with additive costs ($\widehat{\tau}_{ikt}^{adv}$):
+\footnote{We thus estimate Equation (\ref{eq:compeffects_mult}) twice, depending on the type of ad-valorem cost series considered ($\widehat{\tau}_{ikt}^{nlI}$ or $\widehat{\tau}_{ikt}^{adv}$), even if it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
 For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
@@
 \begin{eqnarray}
-\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
+\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
 \end{eqnarray}
-
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_mult}) preserve our specification of the multiplicative and additive costs of Equations (\ref{eq:iceberg}) and (\ref{eq:add}).
-That is, we consider that the multiplicative cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
-In the three estimations (iceberg costs alone, iceberg costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
-As before, both equations are estimated using the non-linear least square procedure (by transport mode).
+\todo{GD 260716 il n'y a pas d'indicatrice exclue pour les biens dans  l'équation pour les additifs, car je n'ai pas mis de constante}
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}).
+That is, we consider that the ad-valorem cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+%%In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
+Equation (\ref{eq:compeffects_mult}) is estimated using OLS and (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).
 
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{nlI}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:
+From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{nlI}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:
 \begin{eqnarray}
@@
 
-We first report the changes over time of the transport costs, estimated iceberg alone, in Figure \ref{fig:totalTC_compeffects_excl}, by transport mode (Panel (a) for Air and Panel (b) for Vessel).
+We first report the changes over time of the transport costs, estimated ad-valorem alone, in Figure \ref{fig:totalTC_compeffects_excl}, by transport mode (Panel (a) for Air and Panel (b) for Vessel).
 
 \begin{figure}[htbp]
-\caption{Total transport costs (estimated iceberg alone), \\
+\caption{Total transport costs (estimated ad-valorem alone), \\
 composition effects excluded
@@
 Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
-First, whatever the transport mode considered,\footnote{Note that this is also true for all type of transport costs, additive and iceberg, as reported in Figure \ref{fig:TC_add_mult_compeffects_excl}, as we come back later on.} they have been roughly constant from 1974 to 1984.
+First, whatever the transport mode considered,\footnote{Note that this is also true for all type of transport costs, additive and ad-valorem, as reported in Figure \ref{fig:TC_add_mult_compeffects_excl}} they have been roughly constant from 1974 to 1984.
 This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that transport costs fall immediately started in 1974.
-Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 is rather attributable to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
+Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 was due to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
 In line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
 Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased by 40\% over the period 1984-2013.
-Further, if we only assume iceberg costs (Figure \ref{fig:totalTC_compeffects_excl}), we infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005, in accordance with the findings of \cite{hummels2007} and \cite{Behar_Venables}.
-Yet, going deeper in the decomposition between the additive and the multiplicative components substantially modifies the picture.
+Further, if we only assume ad-valorem costs (Figure \ref{fig:totalTC_compeffects_excl}), we infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005, in accordance with the findings of \cite{hummels2007} and \cite{Behar_Venables}.
+Yet, going deeper in the decomposition between the additive and the ad-valorem components substantially modifies the picture.
 
-In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the multiplicative and the additive costs (re-scaled to start from the reference value 100), for each transport mode.
+In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode. 
+\todo{Je ne crois pas qu'on puisse faire cela : tu ne peux pas additionner des indices sans les pondérer. Par les poids de 1974, par exemple ? Même cela je ne suis pas sûr que cela soit casher.}
 
 \begin{figure}[htbp]
-\caption{Total transport costs (additive \& multiplicative combined), 
+\caption{Total transport costs (additive \& ad-valorem combined), 
 composition effects excluded}
@@
 \begin{center}
-\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
+\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_sum of both.pdf}
 %{\footnotesize {OECD data} }
@@
 Taking into account the additive component therefore substantially modifies the analysis of the dynamics of transport costs over time.
-In Figure \ref{fig:TC_add_mult_compeffects_excl}, we investigate this result further, by reporting the changes in both the additive and multiplicative components of transport costs, by transport model.
+In Figure \ref{fig:TC_add_mult_compeffects_excl}, we investigate this result further, by reporting the changes in both the additive and ad-valorem components of transport costs, by transport model.
 
@@
 Air transport in particular exhibit contrasted trends.
-While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost.
+While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-kg cost.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
@@
 
-This paper empirically studies the magnitude of additive (or per-unit) costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fob unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
+This paper empirically studies the magnitude of additive (or per-kg) costs in international transport costs, by exploiting the differences between the import and the export prices.
+Using SITC 3 and 4- digit cif-fab unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
 Our results may be summarized in three main findings.
-First, we obtain that additive costs are 2.8\% of fob unit values for ocean shipping (and ad-valorem ones 3.2\%).
+First, we obtain that additive costs are 2.8\% of fab unit values for ocean shipping (and ad-valorem ones 3.2\%).
 These values are respectively equal to 1.8\% and 2.5\% for air transport.
@@
 While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s.
-Further, the inclusion of additive costs yields to the conclusion that overall trade costs are decreased in air transport and ocean shipping roughly the same path and the same magnitude of order.
+Further, the inclusion of additive costs yields to the conclusion the decrease of overall trade costs are decreased in air transport and ocean shipping roughly the same path and the same magnitude of order.
 This last result stands in contrast with related studies (\citealp{hummels2007} (that omit the role of additive costs), which rather obtain a larger decrease in air transport costs, relative to ocean shipping costs.
 \citealp{Behar_Venables}).
+\todo{GD 260716 ce dernier point ne me semble pas justifié}
 In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) as well as regarding the international transmission of business cycles.
+On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in coordinating the international transmission of business cycles.
 This is left for further research.
@@
 
-\subsection{Fob-cif prices}
+\subsection{Fab-cif prices}
 
@@
   \label{tab:result_air_3d_detail}%
-\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
+\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
 \end{table}%
@@
   \label{tab:result_ves_3d_detail}%
-\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
+\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
 \end{table}%
@@
   \label{tab:result_air_rob}%
-\tiny{Notes:Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
+\tiny{Notes:Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
 \end{table}%
@@
   \label{tab:result_ves_rob}%
-\vspace{-10pt}\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
+\vspace{-10pt}\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
 \end{table}%

### 2016-07-28 redaction/dhp_draft_july2016.tex
Modification version
Suite aux remarques de Guillaume 27 juillet 2016

@@
 Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).
-\todo{GD 230716 iceberg, c'est deux choses : ad-valorem + le fait que les coûts sont payés avec des bouts du bien transporté}
 However, many common empirical facts support the existence of additive costs.
@@
 These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
-The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows.
-Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments.
-With additive costs, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments, thereby inducing more ``lumpiness'' in international trade.
-\todo{GD 230716 pas sur que cela soit le sujet. Plutôt «per shipment» / «fixed costs» que «additive»} \smallskip
+\smallskip
 
@@
 Our paper contributes to the literature by providing results on the size of transport costs in long run, explicitly distinguishing between ad-valorem and additive parts.
-Precisely, we quantitatively assess the size and the importance of the additive component in international transport costs.
+In this respect, the paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} We quantitatively assess the size and the importance of the additive component in international transport costs.
 To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013.
@@
 \paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p$ paid by the importer
-\todo{GD 240716 importer plutôt que consumer : il reste des coûts entre les deux...}
- (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fab), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}
+To do so, we start from the equation that expresses the price $p$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fab), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}
 
@@
 \ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add} \\
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
 \end{eqnarray}
@@
 After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
-When assuming ad-valorem costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$
-\todo{GD 240716 je n'aime pas beaucoup l'abbréviation nlI, mais je n'ai pas de meilleure idée...}.
-\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}
-\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}
-Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
+When assuming ad-valorem costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.} Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost $\widehat{\tau}$ and $\widehat{t}$, for each year and transportation mode.
@@
 We investigate the robustness of this conclusion on a yearly basis.
-Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Equation (\ref{eq:model_nlI}), dashed line) and along with the additive component (Equation (\ref{eq:model_with_add}, plain line), by transport mode.
-
-\todo{GD 250716 préciser mean or median?}
+Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Equation (\ref{eq:model_nlI}), dashed line) and along with the mean value of the additive component (Equation (\ref{eq:model_with_add}, plain line), by transport mode.
 
 \begin{figure}[htbp]
-\caption{Ad-valorem Costs (3 digits)}
+\caption{Ad-valorem Costs (Yearly mean value, 3 digits)}
 \label{fig:mult_alone_withadd}
@@
 
-\todo{GD 250716 préciser mean or median?}
-
 \begin{figure}[htbp]
-\caption{Decomposing Transport costs (3 digits)}
+\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
 \label{fig:decomp_TC_3d}
@@
 However, its use is far from being straightforward when evaluating non-linear estimates.
-\todo{LPmettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
+\todo{LP mettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
 In a non-linear context, R-squared is  strictly speaking inappropriate.
@@
   \label{tab:good_fit_vessel}%
-  \parbox[l]{8cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+  \parbox[l]{12cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
@@
 the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term.
-This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs \todo{LPso what?
+This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs \todo{LP so what?
 pas tres punchy}\smallskip
@@
 \end{eqnarray}
-\todo{GD 260716 il n'y a pas d'indicatrice exclue pour les biens dans  l'équation pour les additifs, car je n'ai pas mis de constante}
 Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}).
@@
 %%In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
-Equation (\ref{eq:compeffects_mult}) is estimated using OLS and (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).
+Equation (\ref{eq:compeffects_mult}) is estimated using OLS and Equation (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).
 
@@
 \begin{figure}[htbp]
-\caption{Total transport costs (estimated ad-valorem alone), \\
-composition effects excluded
-}
+\caption{Total transport costs (estimated ad-valorem alone), composition effects excluded}
 \label{fig:totalTC_compeffects_excl}
 \begin{center}
-\includegraphics[width=7cm, height=5cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
+\includegraphics[width=14cm, height=7cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
 %{\footnotesize {OECD data} }
@@
 
-In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode. 
-\todo{Je ne crois pas qu'on puisse faire cela : tu ne peux pas additionner des indices sans les pondérer. Par les poids de 1974, par exemple ? Même cela je ne suis pas sûr que cela soit casher.}
+In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode.\footnote{For each mode, and by year, we build the total value of transport costs as the weighted sum of both the ad valorem index and the additive index, taking as weighting scheme the relative share of each type of costs in the overall total cost in 1974 (using the figures at the root of Figure \ref{fig:mult_alone_withadd}), i.e.
+respectively 48.2 \% for additive costs in Vessel and 42?3\% for Air in 1974.}
 
@@
 \begin{center}
-\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_sum of both.pdf}
+\includegraphics[width=14cm, height=7cm]{TC_addplusmult_compeffects_excl.pdf}
 %{\footnotesize {OECD data} }
@@
 
-In line with Figure \ref{fig:totalTC_compeffects_excl}, Figure \ref{fig:TC_addplusmult_compeffects_excl} reports that transport costs have been roughly constant over 1974-1984 in both modes, decreasing afterwards.
-Yet, another result emerges, that conversely stands in sharp contrast.
+In line with Figure \ref{fig:totalTC_compeffects_excl}, Figure \ref{fig:TC_addplusmult_compeffects_excl} reports that transport costs have been approximatively constant over 1974-1984 in both modes, decreasing afterwards.
+Yet, another result emerges, that conversely stands in sharp contrast with Figure \ref{fig:totalTC_compeffects_excl}.
 There is now not much difference in the reduction of transport costs between air transport and ocean shipping, and both display very similar trends.
 Taking into account the additive component therefore substantially modifies the analysis of the dynamics of transport costs over time.
-In Figure \ref{fig:TC_add_mult_compeffects_excl}, we investigate this result further, by reporting the changes in both the additive and ad-valorem components of transport costs, by transport model.
+In Figure \ref{fig:TC_add_mult_compeffects_excl}, we investigate this result further, by reporting the changes in both the additive and ad-valorem components of transport costs, by transport mode.
 
@@
 \citealp{Behar_Venables}).
-\todo{GD 260716 ce dernier point ne me semble pas justifié}
-In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+\todo{LP 28-07-2016: En reponse aux doutes de GD sur ce point, est-ce plus convaincant ainsi?} In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
 
@@
 For sake of space saving, we only report the results for a sample of years.
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation.
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation.\todo{LP 28-07-2016: Completer les chiffres qui manquent dans les tableaux}
 
@@
 \hline
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}  \\
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}  \\ \hline
 \textit{Iceberg term} &       &       &       &       &       &     \\
@@
 \hline
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs } }\\
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs } }\\ \hline
 \textit{Iceberg term} &       &       &       &       &       &     \\
@@
 \scalebox{0.97}{
-\begin{tabular}{lccccccc}
+\begin{tabular}{lcccccc}
 \hline
 \hline
-      & \multicolumn{6}{c}{Year}                      &  \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\
+      & \multicolumn{5}{c}{Year}                      &  \\
+      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
 \hline
-\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} & \textbf{0.47} \\
-Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} & \textbf{0.63} \\
-\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93} & \textbf{0.89} \\
-Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86} & \textbf{0.79} \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87} & \textbf{-34744.40} \\
-Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14} & \textbf{-30243.95} \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} & \textbf{70940.07} \\
-Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28} & \textbf{62955.73} \\
-\textbf{Test LL} &       &       &       &       &       &       & \textbf{} \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45} & \textbf{9000.89} \\
-\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818} & \textbf{755.73} \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \textbf{} \\
+\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &      \\
+Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
+Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
+\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
+Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
+Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
+\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87}  \\
+Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14}  \\
+\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
+Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} \\
+Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28}  \\
+\textbf{Test LL} &       &       &       &       &       &       \\
+2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45}  \\
+\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
+p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
 \hline\hline
@@
 
-\scalebox{0.97}{\begin{tabular}{lccccccc}
+\scalebox{0.97}{\begin{tabular}{l|cccccc}
 \hline
 \hline
-      & \multicolumn{6}{c}{Year}                      &  \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\ \hline
-\boldmath{}\textbf{R$^{2}$}\unboldmath{} &       &       &       &       &       &       &  \\
-Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{\textbf{0.44}} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.59}} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} & \multicolumn{1}{c}{\textbf{0.77}} \\
-Terms A \& I &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} & \multicolumn{1}{c}{\textbf{0.68}} \\
-\textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} & \multicolumn{1}{c}{\textbf{-29883.62}} \\
-Terms A\& I & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22} & \multicolumn{1}{c}{\textbf{-25303.92}} \\
-\textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23} & \multicolumn{1}{c}{\textbf{61425.60}} \\
-Terms A \& I & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} & \multicolumn{1}{c}{\textbf{53573.29}} \\
-\textbf{Test LL} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.75 & \multicolumn{1}{c}{17354.65} & \multicolumn{1}{c}{20113.52} & \multicolumn{1}{c}{16608.16} & \multicolumn{1}{c}{12589.59} & \multicolumn{1}{c}{15704.63} \\
-\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} & \multicolumn{1}{c}{860} \\
-p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-
+      & \multicolumn{6}{c}{Year}                   \\
+      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
+\boldmath{}\textbf{R$^{2}$}\unboldmath{} &       &       &       &       &       &         \\
+Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
+Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\
+\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Term I only &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
+Terms A \& I &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
+\textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Term I only & -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} \\
+Terms A\& I & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22}  \\
+\textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Term I only & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23}\\
+Terms A \& I & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} \\
+\textbf{Test LL} &       &       & & &  & \\
+2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.75 & \multicolumn{1}{c}{17354.65} & \multicolumn{1}{c}{20113.52} & \multicolumn{1}{c}{16608.16} & \multicolumn{1}{c}{12589.59} \\
+\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
+p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
 \hline\hline

### 2016-07-29 redaction/dhp_draft_july2016.tex
Petites modifications

@@
 Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
-Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international.
+Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international costs.
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fab unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
@@
 Our paper contributes to the literature by providing results on the size of transport costs in long run, explicitly distinguishing between ad-valorem and additive parts.
-In this respect, the paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} We quantitatively assess the size and the importance of the additive component in international transport costs.
+In this respect, the paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.}
+We quantitatively assess the size and the importance of the additive component in international transport costs.
 To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013.
@@
 The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
-% In our case, due to computational limitations implied by the size of our dataset, we implement 100 iterations and set the convergence criterion for successive parameter estimates and for the residual sum of squares at 0.01.
-%GD : cela me semble être de la cuisine sans grand intérêt.
-Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables
-%%and all our three measures of trade costs are bounded by 0 as minimal value.
-%%Pas compris. Parfois un, non ?
+Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
 These cut-offs are aimed at eliminating reporting or coding errors.\medskip
@@
 This result also holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}).
-Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+Second, in both transport modes, the additive component appears of sizeable importance: as mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
@@
 In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode.\footnote{For each mode, and by year, we build the total value of transport costs as the weighted sum of both the ad valorem index and the additive index, taking as weighting scheme the relative share of each type of costs in the overall total cost in 1974 (using the figures at the root of Figure \ref{fig:mult_alone_withadd}), i.e.
-respectively 48.2 \% for additive costs in Vessel and 42?3\% for Air in 1974.}
+respectively 48.2 \% for additive costs in Vessel and 42.3\% for Air in 1974.}
 
@@
 Further, the inclusion of additive costs yields to the conclusion the decrease of overall trade costs are decreased in air transport and ocean shipping roughly the same path and the same magnitude of order.
-This last result stands in contrast with related studies (\citealp{hummels2007} (that omit the role of additive costs), which rather obtain a larger decrease in air transport costs, relative to ocean shipping costs.
-\citealp{Behar_Venables}).
-\todo{LP 28-07-2016: En reponse aux doutes de GD sur ce point, est-ce plus convaincant ainsi?} In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+This last result stands in contrast with related studies (\citealp{hummels2007}) (that omit the role of additive costs), which rather obtain a larger decrease in air transport costs, relative to ocean shipping costs (\citealp{Behar_Venables}).
+In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
 

### 2016-07-29 redaction/dhp_subETSG_2.tex
changement en vue seconde soumission ETSG

--- 2016-07-29 redaction/dhp_draft_july2016.tex
@@
 \citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
-\todo{GD 230716 Euh... je crois que c'est un non-sequitur, non ?}
 If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
@@
 Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
-\todo{LP comparer avec la litterature?}.
 Air transport costs are smaller than sea transport costs.
@@
 However, its use is far from being straightforward when evaluating non-linear estimates.
-\todo{LP mettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
+\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
 In a non-linear context, R-squared is  strictly speaking inappropriate.
@@
 The standard error of the regression based only on the ad-valorem term is consistently lower for Vessel than for Air.
-\todo{LP et pourtant, la part des additifs dans vessel est plus elevée, 48.2\% en moyenne vs 42.3\% pour AIr, cf le .xls.
-A dire?
-Faire reference a la figure 2.}
 
@@
 the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term.
-This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs \todo{LP so what?
-pas tres punchy}\smallskip
+This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs.\smallskip
 
@@
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
-To do this, we estimate the following equation for the ad-valorem cost, based on the ad-valorem cost series obtained when estimated alone ($\widehat{\tau}_{ikt}^{nlI}$ and the ad-valorem cost series estimated along with additive costs ($\widehat{\tau}_{ikt}^{adv}$):
+To do this, we estimate the following Equation for the ad-valorem cost, based on the ad-valorem cost series obtained when estimated alone ($\widehat{\tau}_{ikt}^{nlI}$ and the ad-valorem cost series estimated along with additive costs ($\widehat{\tau}_{ikt}^{adv}$):
 \footnote{We thus estimate Equation (\ref{eq:compeffects_mult}) twice, depending on the type of ad-valorem cost series considered ($\widehat{\tau}_{ikt}^{nlI}$ or $\widehat{\tau}_{ikt}^{adv}$), even if it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
@@
 For sake of space saving, we only report the results for a sample of years.
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation.\todo{LP 28-07-2016: Completer les chiffres qui manquent dans les tableaux}
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation.
 

### 2016-08-26 redaction/dhp_sept2016.tex
Première version des slides pour ETSG et modification du papier

--- 2016-07-29 redaction/dhp_subETSG_2.tex
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE \newline
-SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), 75007, PARIS, FRANCE}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE \newline
- } }
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
+SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France } }
+
+\date{September 2016}
  \maketitle
-
-\date{July 2016}
 \bigskip
@@
 Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs (typically captured through distance) still remain a major obstacle to trade (e.g.
+Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (e.g.
 \citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
+
 Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
@@
 
-Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).
+Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis\footnote{Rigourously speaking, the ``iceberg cost'' includes the ad-valorem costs and the fact that the costs are paid in terms of the good that is traded.
+As this last element is irrelevant in our case, we use the terme ``iceberg'' and ``ad-valorem'' interchangeably, as commonly made in the literature.}).
 However, many common empirical facts support the existence of additive costs.
@@
 He finds supporting evidence for the \emph{ad-valorem} assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
-One can yet be concerned by the generality of his results though.
+One can yet be concerned by the generality of this result.
 By nature, his study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports).
@@
 Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fab unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fas unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
 \citet{hummels_skiba} estimate the elasticity of freight rates with respect to price is estimated to be well below unity.
-Also, their estimates imply that doubling freight costs increases average fab export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
+Also, their estimates imply that doubling freight costs increases average fas export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
 These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
@@
 
-Our paper contributes to the literature by providing results on the size of transport costs in long run, explicitly distinguishing between ad-valorem and additive parts.
+Our paper provides an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
 In this respect, the paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.}
-We quantitatively assess the size and the importance of the additive component in international transport costs.
-To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013.
-Closely related to our paper is the work by \citet{Irrazabal_2015}, that develops a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian exports in 2004, this work finds that additive costs are about 14\% of the median consumer price.
-However our approach departs from theirs in several key aspects, as we develop below.
-Our data are restricted to transport, insurance and handling costs.
-Still, we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013.
+We quantitatively assess the size and the importance of the additive component in international transport costs, that we measure as the difference between the import and the export prices, based on the US Imports flows over 1974-2013.
+Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring additive trade costs from firm-level trade data.
+Based on Norwegian exports in 2004, they find that additive costs are about 14\% of the median consumer price.
+Our paper complements their findings in many respects.
+While they study trade costs in general, our database implies that we focus on international transport costs.
+Similarly to them, our various results emphasize the important role of the additive component of international transport costs.
+Further, our empirical analysis allows us to provide a quantitative measure of the levels of both the iceberg and the additive trade costs.
+Last, we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-They can be summarized in three main points.
+By exploiting the time coverage of our database, our paper is also related to the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+We also share in common with these papers to investigate the time trends of transport costs by transport mode (i.e., air or sea).
+Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}).
+\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
+However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004.
+Concerning ocean shipping, which represents the major part of world trade, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}.
+Our paper contributes to this debate.
+In particular, we show the importance of taking into account the additive component in characterizing the time trends of international transport costs.
+Our findings can be summarized in three main points.
 
 First, our theoretically agnostic approach provides a fairly simple framework for measuring both ad-valorem and additive parts of transportation costs.
-This represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median fab price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fab price $\widetilde{p}_{ik}$.} We thus obtain that the mean values over 1974-2013 of ad-valorem costs are equal to 2.5\% and 3.2\% of the fab price in air and ocean transport respectively, whereas the additive component amounts to 1.8\% and 2.9\% .
-To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
+We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% of the export price in air and ocean transport respectively, whereas the additive component amounts to 1.8\% and 2.9\%.
+To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}  This represents a valuable insight for calibrating related models.
 
@@
 Third, we exploit the time dimension of our database to document the patterns of transport costs over time.
-Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}).
-\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004.
-Concerning ocean shipping, which represents the major part of world trade, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}.
-Our paper contributes to this debate.
 As pointed out by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time.
 After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
-Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem costs.
-Air transport in particular exhibit contrasted trends.
-While most of the early decline in air transport costs can be explained by ad-valorem costs, this component nearly doubled in the 2000s, all along with a substantial decline in the per-kg cost.
+Further, our results point out an important bias in the time trends of transport costs when additive costs are omitted.
+While transport costs seem to exhibit a larger decrease in air than in sea transport between 1985 and 2005 when only iceberg costs are modeled, this difference vanishes when additive costs are taken into account.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
@@
 
-As in \cite{hummels2007} (among others), our measure of transportation costs consists in exploiting the difference between commodity-level export and import prices.
-We first use customs values, quantities and freight costs to recover free-alongside-board (FAB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode.
-More precisely, the (unit) FAB price is computed as the total "customs value" in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
-The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-Our dependant variable is finally computed as the ratio of the CIF price divided by the FAB price.
+Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others).
+The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
+We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
+While both are closely related, we refer to the fas price as the price considered by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
+The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
 Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
-The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
+
 It is worth acknowledging that using this dataset encounters some limitations.
-First, our measure of the cif-fab price gap only covers freight, insurance and handling costs.
-It is silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
-Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit.
-According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
+First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
+It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
+Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap omits the other dimension of transport costs related to the time value of goods in transit.
+According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
 In this respect, it is true that using this dataset embraces a partial view of international transport barriers.
 
-Yet, we believe that our deliver valuable insights on this topic.
+Yet, we believe that our analysis delivers valuable insights on this topic.
 Indeed, using this dataset has (at least) three main advantages.
 First, this dataset delivers a strong statistical reliability arising from a single, trustworthy, customs origin.
-Based on customs declarations, this dataset inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
-In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode (ocean (or "vessel") and air).
-The first will be crucial to compute transport costs (see below), the second will allow to enlight substantial differences in the dynamics of transport costs across transportation mode.
-Second, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FAB price).
-This is highly valuable, as we can estimate both the \textit{levels} of the ad-valorem trade costs and of the additive trade costs.
-This differentiates us from \citet{Irrazabal_2015}, who can only estimate the ratio of additive costs as a share of the median total consumer price.
-Third, this dataset has been used by \cite{hummels2007}, this enables us to compare our results to his findings.
-Further, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004.
+Based on customs declarations, the US Imports database inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
+In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode (ocean (or ``vessel'') and air).
+The first will be crucial to compute transport costs (see below), the second will allow to enlighten substantial differences in the dynamics of transport costs across transportation mode.
+Second, using this dataset allows us to have the import price of the good, next to the export price.
+This is highly valuable, as we can estimate both the \textit{levels} of the iceberg costs and of the additive costs.
+This strongly differentiates us from \citet{Irrazabal_2015}, who can only estimate the ratio of additive costs as a share of the median total consumer price.
+Third, this dataset is available over a long time span.
+We are thus able to exploit information on a yearly basis from 1974 to 2013.
+\cite{hummels2007} preceded us in investigating the time dimension of this database.
+However, our paper differentiates from his work in two main respects.
+First, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004.
 As we show in Section \ref{sec:results_trends}, covering the time period over the recent years delivers interesting insights regarding the trends in air shipping costs.
+Second, we study the time trends of transport costs by disentangling the additive from the iceberg components.
+This proves to be important in the characterization of transport costs changes over time, in particular in air transport.\smallskip
 
@@
 \paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fab), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}
+To do so, we start from the equation that expresses the price $p$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
@@
 
-\noindent As usual in the literature, the so-called ``ad-valorem'' trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive trade costs).
+\noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
 Let us denote $i$ the origin country, and $k$, the sector (or product).
-Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}:
+Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}
 
@@
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fab price ($p>\widetilde{p}_{ik}$).
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
 Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:
@@
 \end{equation*}
-\noindent Where $\epsilon_{ik}$ follows an normal law centered on 0. Taking the logs, this drives us to estimate the following equation:
+\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Taking the logs, this drives us to estimate the following equation:
 
@@
 The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
-The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
-Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
+The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.\todo{Dire plutot: sum of squares of residuals; to increase rather than to decrease?
+Not clear to me.} Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
 These cut-offs are aimed at eliminating reporting or coding errors.\medskip
@@
 Put differently, what traditional models of international trade lose by ignoring additive costs?
-A natural way to answer this question is to perform estimations of equation (\ref{eq:estimatedequation}) constraining $t$ to be equal to zero, and compare the fitting properties and the explanatory power of the restricted and complete models.
-This is done by computing several standard diagnostic statistics (R$^{2}$ and statistics based on the likelihood function).
+A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining $t$ to be equal to zero, and compare the fitting properties and the explanatory power of the restricted and complete models.
 Accordingly, for each year and transport mode, we estimate two equations, depending on additive transport costs being included (Equation (\ref{eq:model_with_add})) or not (Equation (\ref{eq:model_nlI})):
@@
 After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
-When assuming ad-valorem costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a non-linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.} Still following \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost $\widehat{\tau}$ and $\widehat{t}$, for each year and transportation mode.
+
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
+We thus recover a ``synthetic estimate'' of each type of transport cost $\widehat{\tau}^{ice}$, $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 The objective of this section is twofold.
-First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (\ref{eq:estimatedequation}).
+First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (Equation (\ref{eq:model_with_add}) or (\ref{eq:model_nlI})).
 Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix , we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
-Results for all years are of course, available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
+It displays the mean and median values of each type of trade costs (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
 Overall results (over the whole period) are not substantially affected if this year is included though.
@@
 \parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.
-The additive term is expressed in fraction of fab price.
+The additive term is expressed in fraction of fas price.
 ($^\ast$): Four 4-digit estimation: 0n selected years.
@@
 Table \ref{tab:summary_results} yields the following comments.
-First, the magnitude of overall transport costs is sizeable, either considering the average value over the period or even in the most recent years.
+First, the magnitude of overall transport costs is sizeable.
 Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
@@
 Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
-Over the whole period, additive costs are 2.85\% of fab unit values for ocean shipping  - and ad-valorem ones being equal to 3.22\%.
-These values are respectively equal to 1.9\% and 2.5\% for air transport.
+Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%.
+These values are respectively equal to 1.8\% and 2.5\% for air transport.
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
@@
 We investigate the robustness of this conclusion on a yearly basis.
-Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Equation (\ref{eq:model_nlI}), dashed line) and along with the mean value of the additive component (Equation (\ref{eq:model_with_add}, plain line), by transport mode.
+Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_with_add}), by transport mode.
 
@@
 
-At this stage, two interesting comments can be made from Figure \ref{fig:decomp_TC_3d}.
+At this stage, two interesting comments can be made.
 First, the magnitude of transport costs is higher for ocean shipping than for air transport.
-This result also holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}).
-Second, in both transport modes, the additive component appears of sizeable importance: as mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period.
+Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
@@
 
-\subsection{Assessing the importance of per-kg trade costs}
+\subsection{Assessing the importance of additive transport costs}
 
-In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fab prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs.
+In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fas prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs.
 To do so, we rely on several standard measures of fit.
 The first indicator is through comparing $R^{2}$.
-However, its use is far from being straightforward when evaluating non-linear estimates.
-\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
-In a non-linear context, R-squared is  strictly speaking inappropriate.
-However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.}.
-As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
+However, its use is far from being straightforward when evaluating non-linear estimates.\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
+In a non-linear context, $R^2$ is  strictly speaking inappropriate.
+However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.} As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
 Smaller values are better because it indicates that the observations are closer to the fitted line.
@@
 Equation (\ref{eq:model_nlI})).
-The null tested is that the two models are statistically equivalent.\medskip
-
+The null tested is that the two models are statistically equivalent.
 Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level).
@@
     \hline \hline
-    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat($^{\ast}$) \\ \hline
+    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
     \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
-    Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & \textbf{0.31} \\
+    Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
     Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
@@
     Terms A \& I & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
-    \multicolumn{8}{l}{\textit{Test LL }}  \\
-    2*(ll(UR) -ll(R)) & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
+    LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
     nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
-    p-value(d) & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
+    p-value & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
     \hline \hline
@@
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
-($^\ast$): The mean statistics calculated as the average value over all years.
+For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
+The mean statistics calculated as the average value over all years.
 }}
@@
 \hline \hline
-Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat($^{\ast}$) \\ \hline
+Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
 \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
-Term I only & \multicolumn{1}{c}{0.450} & \multicolumn{1}{c}{0,415} & \multicolumn{1}{c}{0.456} & \multicolumn{1}{c}{0.401} & 0,350 & \multicolumn{1}{c}{0.339} & 0.39 \\
+Term I only & \multicolumn{1}{c}{0.450} & \multicolumn{1}{c}{0.415} & \multicolumn{1}{c}{0.456} & \multicolumn{1}{c}{0.401} & 0.350 & \multicolumn{1}{c}{0.339} & 0.39 \\
 Terms A \& I & \multicolumn{1}{c}{0.612} & \multicolumn{1}{c}{0.575} & \multicolumn{1}{c}{0.590} & \multicolumn{1}{c}{0.571} & 0.491 & \multicolumn{1}{c}{0.462} & 0.56 \\ \hline
@@
     Terms A \& I & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
-    \multicolumn{8}{l}{\textit{Test LL }}  \\
-    2*(ll(UR) -ll(R )) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
-    nb of restrictions =q  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
-    p-value(d) & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
+    LL ratio & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
+    nb of restrictions & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
+    p-value& \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
     \hline \hline
@@
 
-Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences.
-First, ad-valorem costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FAB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes).
-The standard error of the regression based only on the ad-valorem term is consistently lower for Vessel than for Air.
-
-Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000 for both transport modes.
-For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term increases the explanatory power by 30\%, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
-The picture is quite different for Vessel.
-In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included.
-the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term.
-This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs.\smallskip
-
 For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in in Appendix \ref{app:robust_4d}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
@@
 
+Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences.
+First, iceberg costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FAB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes).
+The standard error of the regression based only on the ad-valorem term is consistently lower for Vessel than for Air.
+
+Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000 for both transport modes.
+Interestingly, this seems to be attributable to different reasons depending of the transport mode considered.
+For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term increases the explanatory power by 30\%, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
+Accordingly, the lower quality of fit for air transport seems to be attributable to a loss of explanatory power of the additive component.
+The picture is quite different for Vessel.
+In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included.
+In contrast to air transport, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term.
+This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs.
+We henceforth go deeper into this time-trend analysis in the next section.\smallskip
+
 \section{Decomposing Transport Costs: Characterizing the trends \label{sec:results_trends}}
@@
 In this section, we investigate the role of the additive component in total transport costs by exploring the time dimension of our data.
-As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and ad-valorem components in transport costs over time by transport mode.
-Let us consider first the trend in overall transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
+As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
+Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
 Two main comments can be made.
-First, both air and ocean shipping exhibit a downward trend in overall transport costs.
-Second, starting from 1974, the decrease appears more important for sea than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport.
+First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974.
+Second, the decrease appears more important for sea than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport.
 That said, it may be the case the oil shock inflated transportation costs that specific year.
@@
 These results are consistent with \citet{hummels2007}.
-If we now consider the respective shares of additive and ad-valorem components, we observe that the share of additive costs in air transport costs has declined between 1974 and 2013, in particular starting in 1988 (panel (a) of Figure \ref{fig:decomp_TC_3d}).
-By contrast, this share has been roughly constant throughout the period in ocean shipping (panel (b)).
 
@@
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
-To do this, we estimate the following Equation for the ad-valorem cost, based on the ad-valorem cost series obtained when estimated alone ($\widehat{\tau}_{ikt}^{nlI}$ and the ad-valorem cost series estimated along with additive costs ($\widehat{\tau}_{ikt}^{adv}$):
-\footnote{We thus estimate Equation (\ref{eq:compeffects_mult}) twice, depending on the type of ad-valorem cost series considered ($\widehat{\tau}_{ikt}^{nlI}$ or $\widehat{\tau}_{ikt}^{adv}$), even if it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
+To do this, we estimate the following equation for the ad-valorem cost, based on the series obtained $\widehat{\tau}_{ikt}^{ice}$ and $\widehat{\tau}_{ikt}^{adv}$ that have been previously estimated:
+\footnote{Even if Equation (\ref{eq:compeffects_mult}) is estimated on two different series, it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
 For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
@@
 \end{eqnarray}
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}).
-That is, we consider that the ad-valorem cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
 %%In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
@@
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{nlI}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:
-\begin{eqnarray}
-\Gamma_t &=& 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{eqnarray}
+From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{ice}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:
+\begin{equation*}
+\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
 
@@
 As a result, the series for $\Gamma_t$ (with an initial value of 100 for $t=1974$) have a straightforward interpretation in percentage changes.
-\medskip
-
-We first report the changes over time of the transport costs, estimated ad-valorem alone, in Figure \ref{fig:totalTC_compeffects_excl}, by transport mode (Panel (a) for Air and Panel (b) for Vessel).
+Figure \ref{fig:totalTC_compeffects_excl} reports the results.
+Panel (a) displays the time changes in transport costs when only iceberg costs are modeled.
+In Panel (b), we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode.\footnote{For each mode, and by year, we build the total value of transport costs as the weighted sum of both the ad valorem index and the additive index, taking as weighting scheme the relative share of each type of costs in the overall total cost in 1974 (using the figures at the root of Figure \ref{fig:mult_alone_withadd}), i.e.
+respectively 48.2 \% for additive costs in Vessel and 42.3\% for Air in 1974.}
 
 \begin{figure}[htbp]
-\caption{Total transport costs (estimated ad-valorem alone), composition effects excluded}
+\caption{Transport costs (composition effects excluded)}
 \label{fig:totalTC_compeffects_excl}
 \begin{center}
-\includegraphics[width=14cm, height=7cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
-%{\footnotesize {OECD data} }
+\begin{tabular}{cc}
+{\small (a) Iceberg alone } & {\small (b) With additive costs}\\
+\includegraphics[width=3in, height=2.5in]{Fig3a_TC_overtime_comp_effects_excl.pdf}
+& \includegraphics[width=3in,height=2.5in]{TC_addplusmult_compeffects_excl.pdf} \\
+\end{tabular}
 \end{center}
@@
 Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
-First, whatever the transport mode considered,\footnote{Note that this is also true for all type of transport costs, additive and ad-valorem, as reported in Figure \ref{fig:TC_add_mult_compeffects_excl}} they have been roughly constant from 1974 to 1984.
-This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that transport costs fall immediately started in 1974.
+First, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
+This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that the fall in transport costs immediately started in 1974.
 Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 was due to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
 In line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
-Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased by 40\% over the period 1984-2013.
-Further, if we only assume ad-valorem costs (Figure \ref{fig:totalTC_compeffects_excl}), we infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005, in accordance with the findings of \cite{hummels2007} and \cite{Behar_Venables}.
-Yet, going deeper in the decomposition between the additive and the ad-valorem components substantially modifies the picture.
 
-In Figure \ref{fig:TC_addplusmult_compeffects_excl}, we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode.\footnote{For each mode, and by year, we build the total value of transport costs as the weighted sum of both the ad valorem index and the additive index, taking as weighting scheme the relative share of each type of costs in the overall total cost in 1974 (using the figures at the root of Figure \ref{fig:mult_alone_withadd}), i.e.
-respectively 48.2 \% for additive costs in Vessel and 42.3\% for Air in 1974.}
-
-\begin{figure}[htbp]
-\caption{Total transport costs (additive \& ad-valorem combined),
-composition effects excluded}
-\label{fig:TC_addplusmult_compeffects_excl}
-\begin{center}
-\includegraphics[width=14cm, height=7cm]{TC_addplusmult_compeffects_excl.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
-In line with Figure \ref{fig:totalTC_compeffects_excl}, Figure \ref{fig:TC_addplusmult_compeffects_excl} reports that transport costs have been approximatively constant over 1974-1984 in both modes, decreasing afterwards.
-Yet, another result emerges, that conversely stands in sharp contrast with Figure \ref{fig:totalTC_compeffects_excl}.
-There is now not much difference in the reduction of transport costs between air transport and ocean shipping, and both display very similar trends.
+Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased, to reach a reduction of 40\% over the period 1984-2011, in accordance with the literature.
+Further, if we only assume ad-valorem costs (Panel (a)), we can infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005.
+This in accordance with the findings of \cite{hummels2007} and \cite{Behar_Venables}.
+Yet, going deeper in the decomposition between the additive and the ad-valorem components substantially modifies the picture, as it can be inferred from Panel (b).
+When the additive costs are taken into account, there is now not much difference in the reduction of transport costs between air transport and ocean shipping, and both display very similar time changes.
+The downward trend in transport costs indeed is roughly of the same order of magnitude throughout the years, for both transport modes.
 Taking into account the additive component therefore substantially modifies the analysis of the dynamics of transport costs over time.
@@
 Again, the decomposition between both components of transport costs brings valuable insights.
-Comparing Figure \ref{fig:TC_add_mult_compeffects_excl} with Figure \ref{fig:TC_addplusmult_compeffects_excl} (or Figure \ref{fig:totalTC_compeffects_excl}) indicates that the steady decline of overall transport costs hides shifts in the relative importance of additive and ad-valorem.
-Air transport in particular exhibit contrasted trends.
-While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-kg cost.
-These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
+For ocean shipping, both additive and iceberg components exhibit a downward pattern since 1985.
+Air transport exhibit much more contrasted trends.
+While the ad-valorem component has been steadily declining over 1985-2000, the additive cost has been rather stable on the same period.
+Such a difference therefore most likely explains the difference of results between Panels (a) et (b) of Figure \ref{fig:totalTC_compeffects_excl}.
+Skipping the additive component leads to an important omission bias in the characterization of the time changes of transport costs.
+When correctly measured (ie, when the additive component is included), there is no noticeable difference in the downward trend of transport costs between ocean and air transport, in contradiction with the findings of \citet{hummels2007} and \cite{Behar_Venables}.
+Again, this finding points out to the importance of taking into account the additive component of international transport costs.
+\smallskip
+
+Our analysis also brings valuable results to the findings of \citet{hummels2007}, as we extend his time coverage (end year 2004) by one decade (to stop in 2013).
+On this decade, our analysis reveals substantial changes in the time trend of transport costs in the air sector.
+As reported in Figure \ref{fig:totalTC_compeffects_excl}, Panel (b), transport costs have raised between 2005 and 2010, to substantially decline on the very recent years (2011-2013), by more than 20\%.
+As reported in Figure \ref{fig:TC_add_mult_compeffects_excl}, this can be attributed to the iceberg component (while the additive cost has conversely been drastically reduced since 2005).
+These results suggest that important changes in the structure of costs ate at work in the air transport in the very recent period, that should be investigated in further research.
 
@@
 This paper empirically studies the magnitude of additive (or per-kg) costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fab unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
+Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
 Our results may be summarized in three main findings.
-First, we obtain that additive costs are 2.8\% of fab unit values for ocean shipping (and ad-valorem ones 3.2\%).
-These values are respectively equal to 1.8\% and 2.5\% for air transport.
+First, we provide a quantitative measure of both the additive and the iceberg transport cost.
+We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%.
+These values are respectively equal to 1.8 and 2.5\% for air transport.
 Second, we show that taking additive costs into account improves the fit of the modelling of transport costs.
@@
 Further, the inclusion of additive costs yields to the conclusion the decrease of overall trade costs are decreased in air transport and ocean shipping roughly the same path and the same magnitude of order.
-This last result stands in contrast with related studies (\citealp{hummels2007}) (that omit the role of additive costs), which rather obtain a larger decrease in air transport costs, relative to ocean shipping costs (\citealp{Behar_Venables}).
+This last result stands in contrast with related studies (\citealp{hummels2007}, \citealp{Behar_Venables}).
 In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in coordinating the international transmission of business cycles.
+On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles.
 This is left for further research.
@@
 The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller.
-In the case of transactions between related parties, the relationship between buyer and seller should not influence the Customs value.
-
-In those instances where assistance was furnished to a foreign manufacturer for use in producing an article which is imported into the United States, the value of the assistance is required to be included in the value reported for the merchandise.
-Such ``assists'' include both tangible and intangible assistance, such as machinery, tools, dies and molds, blue prints, copyrights, research and development, and engineering and consulting services.
-If the value of these ``assists'' is identified and separately reported, it is subtracted from the value during statistical processing.
-However, where it is not possible to isolate the value of ``assists'', they are included.
-In these cases the unit values may be increased due to the inclusion of such ``assists''.
-Import Charges
+In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
 
@@
 import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States.
-In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States) in Canada or Mexico to the first port of entry.
-C.I.F. Import Value
+In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States in Canada or Mexico to the first port of entry.
 
-The C.I.F.
-(cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
+The cif (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
 It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S.
@@
 
-\section{Some detailed results}
+\section{Some detailed results \label{app:more_results}}
 
@@
   \label{tab:result_air_3d_detail}%
-\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
+\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
 \end{table}%
@@
   \label{tab:result_ves_3d_detail}%
-\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
+\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
 \end{table}%
@@
   \label{tab:result_air_rob}%
-\tiny{Notes:Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
+\tiny{Notes:Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
 \end{table}%
@@
   \label{tab:result_ves_rob}%
-\vspace{-10pt}\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fab price. }
+\vspace{-10pt}\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
 \end{table}%

### 2016-08-26 redaction/slides_ETSG.tex
Première version des slides pour ETSG et modification du papier

\documentclass[10 pt,Helvetica, french]{beamer}
% option compress pour réduire la taille des bars
\mode<presentation>
%[compress,hideallsubsections,width=<0.5cm>]
%\usetheme{Darmstadt}
%\usetheme{Malmoe}
%\usetheme{Warsaw}
%\useoutertheme{split}
%\usecolortheme{whale}

%\usepackage[dark]{beamerthemesidebar}
%\usepackage{beamerthemeshadow}
%\usepackage{beamerthemesplit}
%\usepackage{beamerthemebars}
%\usepackage{beamerthemeclassic}
\newcommand{\spac}{\rule[0mm]{0mm}{2mm}}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{amsmath,amssymb}
\usepackage[english]{babel}
\usepackage{epic,eepic,bez123}
%\usepackage[T1]{fontenc}
%\usepackage[ansinew]{inputenc}
%\usepackage[latin1]{inputenc}
%\usepackage{babel,varioref}

%\usetheme{Luebeck}%(encadrements rectangulaires)
%\usetheme{Madrid}%(pas de plan au-dessus)
%\usetheme{Berkeley}%(colonne à gauche)
%\usetheme{Warsaw}%(plan détaillé en haut, utilise la navigation)
%\usetheme{Dresden}%(encadrements rectangulaires)
%\usetheme{Frankfurt}%(pas de paln détaillé en haut)
%\usetheme{Marburg}%(colonne à droite)
%\usetheme{Berlin}%(encadrements rectangulaires)
%\usetheme{Montpellier}

%\usecolortheme{crane}
%\usecolortheme{seahorse}
%\usecolortheme{green}
%\usecolortheme{beetle}
%\usecolortheme{default}
%\usecolortheme{rose}
%\usecolortheme{Berkeley}(inconnu)

%\useoutertheme{smoothtree}
%\useoutertheme{split}
%\usecolortheme{whale}

\setbeamercovered{transparent}
\newtheorem{defi}{Definition}
\newtheorem{theo}{Theorem}
\newtheorem{prop}{Proposition}
\newtheorem{property}{Property}
\newtheorem{coro}{Corollary}
\newtheorem{propannex}{Proposition}
\newtheorem{objective}{Objective}
\newtheorem{res}{Results}

\setbeamertemplate{footline}[text line]{  }
%\setbeamertemplate{footline}[page number]{ }

\newcounter{countAnnexA}
\setcounter{countAnnexA}{0}
\renewcommand{\propannex}{\addtocounter{countAnnexA}{1}
\noindent \large \textbf{Proposition \thesection.\thecountAnnexA}
\normalsize}

\newcommand{\beginbackup}{
   \newcounter{framenumbervorappendix}
   \setcounter{framenumbervorappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
   \addtocounter{framenumbervorappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumbervorappendix}}
}

%\colorlet{orange}{red!70!yellow}% mes petites couleurs que j'aime bien
%\colorlet{mauve}{blue!70!red} \colorlet{brouge}{red!70!blue}
\addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\date[September 2016]{ETSG, Helsinki, September 2016}
\title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
\author[Daudin et al.]{Guillaume Daudin\\
{\footnotesize Universit\'{e} Paris-Dauphine, PSL \& OFCE }\\ \smallskip
J\'{e}rome H\'{e}ricourt \\
{\footnotesize Universit\'{e} de Lille I \& CEPII }\\  \smallskip
Lise Patureau \\
{\footnotesize  Universit\'{e} Paris-Dauphine, PSL}}

\setcounter{framenumber}{-1}
\begin{document}
\begin{frame}[plain]
\titlepage
\end{frame}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
\item Trade costs: A central role in international economic analysis \vspace{0.1cm}
\begin{itemize}
\item[-] Declining over the second half of the 20$^{th}$ century (Jacks et al., 2008, Novy, 2013) \vspace{0.1cm}
\item[-] But still significant: Average international trade costs = a 74\% markup over production costs (Anderson \& Van Wincoop, 2004)
\end{itemize}
\item What exactly are ``trade costs''?  \vspace{0.1cm}
\begin{itemize}
\item[-] Transaction costs, policy costs, time costs, and transport costs \textit{per se}
\end{itemize}
\item Transport costs: A sizeable share of international trade costs \vspace{0.1cm}
\begin{itemize}
\item[-] Account for 21\% of international trade costs (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] If much trade policy barriers have been removed, the transport cost component of trade costs remains sizeable \vspace{0.1cm}
\end{itemize}
\textbf{The paper: On international transport costs}

\end{frame}

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Wouldn't trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
\item The structure (additive vs iceberg) of transport costs, far from being anecdotal \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Additive trade costs, an important role in shaping the pattern of trade flows (the Alchian \& Allen (1964) conjecture) \vspace{0.1cm}
\item[-] A bunch of empirical papers in support of the additive hypothesis: Martin (2012), Hummels \& Skiba (2004), Irrarazabal et al. (2015) \vspace{0.1cm}
\item[-] With strong normative implications: Sorensen (2014), Irrarazabal et al. (2015) \vspace{0.1cm}}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.1cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}
\end{frame}

\begin{frame}
\frametitle{Our paper}
\begin{itemize}
\item An empirical decomposition of the structure of transport costs \vspace{0.1cm}
\item Quantitatively assess the size and the importance of the additive component in international transport costs
\begin{itemize}
%\footnotesize{
\item[-] Exploit the differences between export and imports prices (by transport mode, ocean and air)\vspace{0.1cm}
\item[-] From the US imports flows, on a yearly basis from 1974 to 2013  \vspace{0.1cm}
%}
\end{itemize}
\item Our contribution to the literature
\begin{itemize}
\item[-] Confirm the literature about the importance of the additive component of trade costs (Martin, 2012, Hummels \& Skiba, 2004)\vspace{0.1cm}
\item[-] But, we quantify it \vspace{0.1cm}
\item[-] Over a large spectrum of time (1974 to 2013) and distinguishing between air and sea transport \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] A broad view of the magnitude of additive costs in international trade over time
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Three questions, three answers}
\begin{itemize}
\item What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
\end{itemize}
\item What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] We lose much \vspace{0.1cm}
\begin{itemize}
\item[-] With the additive term included, the ad-valorem component reduces by a factor of 2 \vspace{0.1cm}
\item[-] With the additive term taken into account, a significantly better ``goodness-of-fit'' \vspace{0.1cm}
\end{itemize}
\item How have international transport costs evolved over time? \vspace{0.1cm}
\begin{itemize}
\item[-] Transport costs \textit{per se} have fallen since 1985, by 40\% \vspace{0.1cm}
\item[-] When additive costs are included, not much difference between air and sea transport
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Plan of the talk}
\begin{itemize}
\item Data Sources  \vspace{0.1cm}
\item Empirical Methodology \vspace{0.1cm}
\item Assessing the importance of the additive component in transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Measure both the ad-valorem and the additive components  \vspace{0.1cm}
\item[-] Goodness-of-fit evaluation\vspace{0.1cm}
\end{itemize}
\item Characterizing the trends of transport costs \vspace{0.1cm}
\item Conclusion
\end{itemize}
\end{frame}

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
%\item Measure international transport costs as the difference between the export price and the import price \vspace{0.1cm}
\item Database: US annual imports of merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level \vspace{0.1cm}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the US import gate \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The estimated equation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$ (and by transport mode, air or vessel) \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
 \end{equation}
\end{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The estimated equation becomes:
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
\end{itemize}

\end{itemize}
\end{frame}

\begin{frame} [label=slide_method]
\begin{itemize}
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\item A non-linear equation (due to the additive costs) $\rightarrow$ Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.1cm}
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\begin{itemize}
\item[-] For each year and by transport mode, estimate two equations, depending on additive costs included or not \vspace{0.1cm}
\end{itemize}
\item When additive costs are included:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}
\normalsize
\begin{itemize}
\item[-] From which we re-built:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item When assuming iceberg costs only
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\begin{itemize}
\item[-] From which we re-built:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{\tau}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Decomposing transport costs: Additive vs Iceberg}
\textbf{Result 1}: Provide estimates for the size of both the ad-valorem and the additive components of transport costs
\begin{itemize}
\item Transport costs expressed in percent of the export price, average values over 1974-2013 \hyperlink{app_results_summary}{\beamergotobutton{More}}  \vspace{0.1cm}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{table}%

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Comments}
\begin{itemize}
\item An important magnitude of overall transport costs  \vspace{0.1cm}
\begin{itemize}
\item[-] Over 1974-2013, correspond to a mean increase of the export price by 5.8\% for ocean shipping, by 5.1\% for air transport  \vspace{0.1cm}
\end{itemize}
\item When decomposing the structure of transport costs: A sizeable additive component   \vspace{0.1cm}
\begin{itemize}
\item[-] Additive costs amount to 2.9\% of the export price in ocean shipping, 1.8\% in air transport  \vspace{0.1cm}
\end{itemize}
\item Omitting the additive term substantially biases the iceberg component upwards  \vspace{0.1cm}
\begin{itemize}
\item[-] The ad-valorem cost reduces by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[-] 3.2\% and 2.5\% as mean values in vessel and air respectively (in presence of additive costs) \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_fig1]
\frametitle{More on the importance of additive transport costs}
\begin{itemize}
\item Using the time dimension of our database, study the shares of both ad-valorem and additive components in total transport costs
\item[] \hyperlink{slide_moreonfig1}{\beamergotobutton{More on this figure later}}
\end{itemize}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
\includegraphics[width=5cm, height=5cm]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=5cm,height=5cm]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[label=slide_comments_fig1]
\frametitle{Comments}
\begin{itemize}
\item The size of transport costs is larger for ocean shipping than from air transport \vspace{0.1cm}
\begin{itemize}
\item[-] On average over the period (Table 1) and on a yearly basis (Figure 1) \hyperlink{app_comments_fig1}{\beamergotobutton{More}}  \vspace{0.1cm}
\end{itemize}
\item A sizeable share of the additive component in total transport costs  \vspace{0.1cm}
\begin{itemize}
\item[-] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
\item[-] A decreasing share of additive costs over time for air transport \vspace{0.1cm}
\item[-] Roughly constant for ocean shipping
\end{itemize}
\end{itemize}
$\Rightarrow$ \textbf{Result 2}: Additive costs, an important dimension of international transport costs \vspace{0.1cm}
\begin{itemize}
\item Further investigate the robustness of this result
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The role of additive costs: Goodness-of-fit evaluations}
\begin{itemize}
\item Provide a more systematic diagnosis about the importance of additive costs  \vspace{0.1cm}
\item By comparing the goodness-of-fit of the regressions
\begin{itemize}
\item[-] Obtained under Specification (\ref{eq:model_nlI}) (no additive costs)
\item[-] vs Specification (\ref{eq:model_with_add}) (with additive costs) \vspace{0.1cm}
\end{itemize}
\item Various measures of goodness of fit
\begin{itemize}
\item[-] The $R^2$ (the larger the value, the better the fit) \vspace{0.1cm}
\item[-] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
\item[-] The log-likelihood function and two derived measures \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] The Akaike Information Criterion (the lower AIC, the better the fit) \vspace{0.1cm}
\item[$\ast$] The log-likelihood ratio test ($H_0$: both models are equivalent)
\end{itemize}
\end{itemize}

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Goodness of fit comparison}
\begin{itemize}
\item Air, 3 digit - level, selected years
\end{itemize}
\begin{table}[htbp]
  \centering
 % \caption{Air: Measures of Goodness-of-fit (3 digits)}
  \scriptsize{
\begin{center}
    \begin{tabular}{l|ccccc|c}
    \hline \hline
    Year  &1980  & 1990  & 2000  & 2010 & 2013 & Mean stat \\ \hline
    \multicolumn{7}{l}{\bf{$R^2$} }\\ \hline
    Term I only &  0.27  & 0.25  & 0.32  & 0.42 & 0.34 & 0.31 \\
    Terms A \& I &  0.65  & 0.63  & 0.64  & 0.51 & 0.46 & 0.60 \\ \hline
    \multicolumn{7}{l}{\textbf{SER}  }  \\ \hline
    Term I only &  0.86  & 0.81  & 0.84  & 0.86 & 0.92 & 0.85 \\
    Terms A \& I &  0.71  & 0.67  & 0.70  & 0.79 & 0.85 & 0.73 \\ \hline
   \multicolumn{7}{l}{\textbf{AIC criteria}}  \\ \hline
    Term I only &  41171.0 & 60715.6 & 87492.6 & 102297.6 & 88191.9 & 70498.1 \\
    Terms A \& I & 35738.4 & 52098.9 & 74954.9 & 95887.1 & 80873.7 & 62285.0 \\ \hline
    \multicolumn{7}{l}{\textbf{Log-likelihood}} \\ \hline
    Term I only &  -20253.5 & -29977.8 & -43341.3 & -50746.8 & -43692.9 & -34888.6 \\
    Terms A \& I &  -17263.2 & -25393.5 & -36788.4 & -47277.5 & -39751.9 & -30508.3 \\
    LL ratio &  5980.6 & 9168.7 & 13105.7 & 6938.6 & 7882.1 & 8760.69 \\
    nb of restrictions & 369   & 393   & 426   & 426 & 427 & 402 \\
    p-value & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 \\
    \hline \hline
 \end{tabular}%
    \end{center}}
  \label{tab:good_fit_air}%
 \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion. $R^{2}$ between the log of predicted ratio and the log of the observed ratio. For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products. The mean statistics is calculated as the average value over all years. }}
\end{table}%
\end{frame}

\begin{frame}
\frametitle{Goodness of fit comparison (cont')}
\begin{itemize}
\item Vessel, 3 digit - level, selected years
\end{itemize}
\begin{table}[htbp]
  \centering
  %\caption{Vessel: Measures of Goodness-of-fit (3 digits)}
    \scriptsize{
\begin{center}
\begin{tabular}{l|ccccc|c}
\hline \hline
Year  &  1980 & 1990 & 2000 & 2010  & 2013 & Mean stat \\ \hline
\multicolumn{7}{l}{\bf{$R^2$} }\\ \hline
Term I only &  0.415 & 0.456 & 0.401 & 0.350 & 0.339 & 0.39 \\
Terms A \& I &  0.575 & 0.590 & 0.571 & 0.491 & 0.462 & 0.56 \\ \hline
\multicolumn{7}{l}{\textbf{SER}  }  \\ \hline
    Term I only &  0.62 & 0.59 & 0.65 & 0.74  & 0.76 & 0.66 \\
    Terms A \& I &  0.53 & 0.51 & 0.55 & 0.66  & 0.68 & 0.57 \\ \hline
   \multicolumn{7}{l}{\textbf{AIC criteria}}  \\ \hline
    Term I only & 33010.3 & 51142.6 & 71365.9 & 84789.9 & 88191.9 & 57848.6 \\
    Terms A \& I &  28067.3 & 43664.7 & 60475.9 & 76161.3 & 80873.7 & 49682.3 \\ \hline
    \multicolumn{7}{l}{\textbf{Log-likelihood}} \\ \hline
    Term I only &  -16129.1 & -25169.3 & -35263.9 & -41998.9 & -43692.9 & -28534.3 \\
    Terms A \& I &  -13353.7 & -21171.4 & -29491.0 & -37418.7 & -39751.9 & -24151.3 \\
    LL ratio & 5550.96 & 7995.88 & 11545.98 & 9160.56 & 7882.15 & 8766.0 \\
    nb of restrictions  &  395 & 411 & 436 & 424   & 427 & 417 \\
    p-value & 0.00 & 0.00 & 0.00 & 0.00  & 0.00 & 0.00 \\
    \hline \hline
    \end{tabular}%
    \end{center}}
  \label{tab:good_fit_vessel}%
  \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion. $R^{2}$ between the log of predicted ratio and the log of the observed ratio. For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products. The mean statistics calculated as the average value over all years. }}
\end{table}%
\end{frame}

\begin{frame}
\frametitle{Comments}
\begin{itemize}
\item \textbf{Result 2 (confirmed)}: Including the additive component improves the fit of the model, whatever the considered criterion or the transport mode \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] On average over the period, the $R^2$ doubles for air transport, increases by 50\% for vessel (also on a yearly basis) \vspace{0.1cm}
}
\end{itemize}
\item A decrease in the quality of fit strongly after 2000, for both transport modes \vspace{0.1cm}
\item Contrasted results between air and ocean transports \vspace{0.1cm}
\begin{itemize}
\item[-] Ad-valorem costs have more explanative power in ocean shipping than in air transport \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] On average, account for 39\% of the variance of the cif-fab ratio, vs 31\% in Air \vspace{0.1cm}
\end{itemize}
\item[-] But their explanatory power seems to have decreased over time \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] A roughly constant contribution of the additive component to the goodness of fit over time \vspace{0.1cm}
\end{itemize}
\item[$\neq$] For Air: The decreasing role of the explanatory power of the additive component over the recent years (consistent with Figure 1)\vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Go deeper in the analysis of the time-trends
\end{itemize}
\end{frame}

\begin{frame}[label=slide_moreonfig1]
\frametitle{Decomposing transport costs: Characterizing the trends}
\begin{itemize}
\item Coming back to Figure 1 \hyperlink{slide_fig1}{\beamergotobutton{Go}}: Downward trend in overall transport costs, for both air and ocean shipping \vspace{0.1cm}
\begin{itemize}
\item[-] Considering 1980 as starting date, a 1.5- 2 percentage points decrease in Air, a 2-3 percentage point decrease in Vessel \vspace{0.1cm}
\end{itemize}
\item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.1cm}
\item The change in overall transport costs over time: Depend on the evolution of per product- per partner costs, but also on the composition of trade flows \vspace{0.1cm}
\begin{itemize}
\item[-] Over time, import more goods that are cheaper to transport, and/or from countries with which it is cheaper to trade \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}  \vspace{0.1cm}
    \begin{itemize}
\item[-] What we do, in accordance with Hummels (2007) \hyperlink{app_compositioneffects}{\beamergotobutton{More on the method}}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Time trends in transport costs \textit{per se}}
\begin{itemize}
\item Total transport costs (composition effects excluded) over time
\end{itemize}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Iceberg alone } & {\scriptsize  (b) With additive }\\
\includegraphics[width=5cm, height=5cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
& \includegraphics[width=5cm,height=5cm]{TC_addplusmult_compeffects_excl.pdf} \\
\end{tabular}
\end{center}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Comments}
\textbf{Result 3: About the time trends in transport costs}
\begin{itemize}
\item The reduction of the \textit{pure} transport costs : Starts in 1985 \vspace{0.1cm}
\begin{itemize}
\item[-] For both transport modes \vspace{0.1cm}
\item[-] The reduction between 1974 and 1984 (Figure 1), attributable to change in the composition of trade patterns \vspace{0.1cm}
\end{itemize}
\item Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
\item Comparing air and ocean shipping \vspace{0.1cm}
\begin{itemize}
\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in air over the 1985-2005 period \vspace{0.1cm}
\item[-] A result in accordance with Hummels (2007), Behar and Venables (2011) \vspace{0.1cm}
\item[-] No more substantial difference when additive costs are included (Panel (b))
\end{itemize}
\item[$\Rightarrow$] The importance of taking into account the additive component of trade costs  \vspace{0.1cm}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Decomposing (\textit{pure}) transport costs over time }
\begin{figure}[htbp]
%\caption{Decomposing transport costs over time,
%composition effects excluded}
%\label{fig:TC_add_mult_compeffects_excl}
\begin{center}
\includegraphics[width=8cm, height=4cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
%{\footnotesize {OECD data} }
\end{center}
\end{figure}
\begin{itemize}
\item For Vessel: Similar downward trend for both $\tau$ and $t$
\item For Air: Much more contrasted trends
\begin{itemize}
\item[-] Substantial decrease in the ad-valorem costs over 1985-2005, but roughly constant additive costs
\item[$\Rightarrow$] An explanation to the difference between Panels (a) and (b) above
\end{itemize}
\item[$\Rightarrow$] The importance of modeling additive costs
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Conclusion}
\textbf{Our paper: Empirical evidence about the role of the additive component in international transport costs} \vspace{0.1cm}
\begin{itemize}
\item Provide a quantitative evaluation of both the additive and the ad-valorem components \vspace{0.1cm}
\begin{itemize}
\item[-] Additive cost: amount to 2.8\% of the export price in ocean shipping, 1.8\% in air transport \vspace{0.1cm}
\item[-] Iceberg cost: 3.2\% and 2.5\% for air and ocean respectively \vspace{0.1cm}
\end{itemize}
\item The importance of taking into account additive transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Additive costs are far from negligible quantitatively \vspace{0.1cm}
\item[-] A better fit of the model when taken into account    \vspace{0.1cm}
\end{itemize}
\item Characterize the evolution of transport costs over time \vspace{0.1cm}
\begin{itemize}
\item[-] Importance of the composition effects \vspace{0.1cm}
\item[-] Biased picture of the time trends of transport costs in air when omitting the additive dimension
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Conclusion}

\textbf{Two main possible extensions}
\begin{itemize}
\item On the empirical side: Go deeper in the structural determinants of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
\end{itemize}
\item On the theoretical side: Use our results to explore the role of additive costs \vspace{0.1cm}
\begin{itemize}
\item[-] In shaping international trade flows (trade theory) \vspace{0.1cm}
\item[-] In affecting the international transmission of business cycles (business cycle theory)
\end{itemize}
\end{itemize}
\end{frame}

\beginbackup

\appendix
\begin{frame}[label=app_data]
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, large time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a larger period of time \vspace{0.1cm}
\item[-] Have both the import and the export prices: Estimate both the levels of the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
\end{itemize}
\end{itemize}
\hyperlink{slide_data}{\beamergotobutton{Back to slide}}
\end{frame}

\begin{frame}[label=app_method]
\frametitle{More on the estimation method}
\begin{itemize}
\item The non-linear least squares method \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not decrease \textbf{increase?} from one iteration to the next. \vspace{0.1cm}
\end{itemize}
\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}

\end{itemize}
\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
\end{frame}

\begin{frame}[label=app_results_summary]
\frametitle{Decomposing transport costs: Summary \hyperlink{slide_results_summary}{\beamergotobutton{Back to slide}}}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs }  \\ \hline
    Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
    Median & 1.051 & 1.042 & 1.052 & 1.037 \\
    Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \textit{Ad-valorem term} & & & & \\ \hline
    Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
    Median & 1.028 & 1.018 & 1.028 & 1.016 \\
    Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term }& & & &   \\ \hline
    Mean  & 0.029 & 0.018 & 0.028 & 0.019 \\
    Median & 0.019 & 0.007 & 0.017 & 0.008 \\
    Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{table}%
\end{frame}

\begin{frame}[label=app_comments_fig1]
\frametitle{Ad-valorem costs over time \hyperlink{slide_comments_fig1}{\beamergotobutton{Back to slide}}}
\begin{figure}[htbp]
\caption{Ad-valorem Costs (Yearly mean value, 3 digits)}
\label{fig:mult_alone_withadd}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=5cm, height=2.5in]{Fig1a_mult_air_3d.pdf}
& \includegraphics[width=5cm,height=2.5in]{Fig1b_mult_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}

\end{frame}

\begin{frame}[label=app_compositioneffects]
\frametitle{Excluding the composition effects of transport costs changes}
\begin{itemize}
\item For the ad-valorem component, we estimate the following equation:
\scriptsize
\begin{eqnarray}
\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\Leftrightarrow \ln(\tau_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}
\normalsize
\begin{itemize}
\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
\end{itemize}
\item For the additive component:
\scriptsize
\begin{eqnarray}
\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
\normalsize
\end{itemize}

\end{frame}

\begin{frame}
\begin{itemize}
\item More on the estimation method
\begin{itemize}
\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, \vspace{0.1cm}
\item[-] Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode) \vspace{0.1cm}
\end{itemize}
\item Exclude the composition effects of transport costs changes $\Leftrightarrow$ Isolate the change in the time dimension  \vspace{0.1cm}
\begin{itemize}
\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
\begin{equation*}
\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
\end{equation*}
\item[-] For the additive cost, we build the variable ($\forall~t \geq 1974$)
$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
\end{itemize}
\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
\begin{itemize}
\item[] with an initial value of 100 for $t=1974$
\end{itemize}

\end{itemize}
 \hyperlink{slide_moreonfig1}{\beamergotobutton{Back to slide}}
\end{frame}

\end{document}

### 2016-09-05 redaction/slides_ETSG1.tex
Nouvell version des slides

--- 2016-08-26 redaction/slides_ETSG.tex
@@
 \item[-] Account for 21\% of international trade costs (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
-\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
+%\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
 \end{itemize}
@@
 \footnotesize{
-\item[-] Additive trade costs, an important role in shaping the pattern of trade flows (the Alchian \& Allen (1964) conjecture) \vspace{0.1cm}
-\item[-] A bunch of empirical papers in support of the additive hypothesis: Martin (2012), Hummels \& Skiba (2004), Irrarazabal et al. (2015) \vspace{0.1cm}
+\item[-] Additive trade costs, an important role in shaping the pattern of trade flows: Martin (2012), Hummels \& Skiba (2004), Irrarazabal et al. (2015) \vspace{0.1cm}
 \item[-] With strong normative implications: Sorensen (2014), Irrarazabal et al. (2015) \vspace{0.1cm}}
@@
 \begin{frame}
-\frametitle{Our paper}
+\frametitle{Our paper in 3 questions (and 3 answers)}
+An empirical decomposition of the structure of transport costs \vspace{0.1cm}
 \begin{itemize}
-\item An empirical decomposition of the structure of transport costs \vspace{0.1cm}
-\item Quantitatively assess the size and the importance of the additive component in international transport costs
-\begin{itemize}
-%\footnotesize{
-\item[-] Exploit the differences between export and imports prices (by transport mode, ocean and air)\vspace{0.1cm}
-\item[-] From the US imports flows, on a yearly basis from 1974 to 2013  \vspace{0.1cm}
-%}
-\end{itemize}
-\item Our contribution to the literature
-\begin{itemize}
-\item[-] Confirm the literature about the importance of the additive component of trade costs (Martin, 2012, Hummels \& Skiba, 2004)\vspace{0.1cm}
-\item[-] But, we quantify it \vspace{0.1cm}
-\item[-] Over a large spectrum of time (1974 to 2013) and distinguishing between air and sea transport \vspace{0.1cm}
-\end{itemize}
-\item[$\Rightarrow$] A broad view of the magnitude of additive costs in international trade over time
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Three questions, three answers}
-\begin{itemize}
-\item What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
+\item[(1)] What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
 \item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
 \begin{itemize}
+\footnotesize{
 \item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
 \item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
+}
 \end{itemize}
-\item What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
-\item[$\Rightarrow$] We lose much \vspace{0.1cm}
+\item[(2)] What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
+\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
 \begin{itemize}
-\item[-] With the additive term included, the ad-valorem component reduces by a factor of 2 \vspace{0.1cm}
-\item[-] With the additive term taken into account, a significantly better ``goodness-of-fit'' \vspace{0.1cm}
+\item[-] The estimated iceberg component reduces by a factor of 2 \vspace{0.1cm}
+\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
 \end{itemize}
-\item How have international transport costs evolved over time? \vspace{0.1cm}
+\item[(3)] How have international transport costs evolved over time? \vspace{0.1cm}
 \begin{itemize}
-\item[-] Transport costs \textit{per se} have fallen since 1985, by 40\% \vspace{0.1cm}
-\item[-] When additive costs are included, not much difference between air and sea transport
+\item[-] Transport costs \textit{per se} have fallen since 1985, by $\simeq$ 40\% \vspace{0.1cm}
+\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011)
 \end{itemize}
@@
 \item Empirical Methodology \vspace{0.1cm}
-\item Assessing the importance of the additive component in transport costs \vspace{0.1cm}
-\begin{itemize}
-\item[-] Measure both the ad-valorem and the additive components  \vspace{0.1cm}
-\item[-] Goodness-of-fit evaluation\vspace{0.1cm}
-\end{itemize}
-\item Characterizing the trends of transport costs \vspace{0.1cm}
+\item Results \vspace{0.1cm}
 \item Conclusion
@@
 \begin{itemize}
-%\item Measure international transport costs as the difference between the export price and the import price \vspace{0.1cm}
-\item Database: US annual imports of merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
+\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
+\item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
 \begin{itemize}
-\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level \vspace{0.1cm}
 \item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
-\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the US import gate \vspace{0.1cm}
+\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
+\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Empirical specification (1)}
-\textbf{The estimated equation}
+\frametitle{Empirical specification: The estimated equation}
 \begin{itemize}
@@
 $$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
-\item For product $k$, from country $i$ (and by transport mode, air or vessel) \vspace{0.1cm}
+\item For product $k$, from country $i$  \vspace{0.1cm}
 \item Rewrite to get:
 $$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
-\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level)
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Empirical specification (2)}
-\textbf{The estimation strategy}
+\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 \vspace{0.1cm}
 \begin{itemize}
-\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
-\begin{itemize}
-\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
-\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
-\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
-\begin{equation}
- \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
- \end{equation}
-\end{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
-\item[$\Rightarrow$] The estimated equation becomes:
-$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
-\begin{itemize}
-\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
+\item[-] The equation being also year- and mode (air or vessel)- specific
 \end{itemize}
@@
 \begin{frame}  [label=slide_method]
+\frametitle{Empirical specification: The estimation strategy }
 \begin{itemize}
-\item Taking logs, we finally estimate the following equation
+\item With some assumptions on the specification of transport costs \& the error term, and taking logs  \hyperlink{app_method_1}{\beamergotobutton{More}}\vspace{0.1cm}
+\item[$\Rightarrow$] Estimate the following equation \vspace{0.1cm}
 \footnotesize
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_with_add}
 \end{equation}
+\begin{itemize}
+\item[-] Estimates performed using non-linear least squares \hyperlink{app_method_2}{\beamergotobutton{More}} \vspace{0.1cm}
+\end{itemize}
 \normalsize
-\item A non-linear equation (due to the additive costs) $\rightarrow$ Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.1cm}
 \item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
-\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
-\begin{itemize}
-\item[-] For each year and by transport mode, estimate two equations, depending on additive costs included or not \vspace{0.1cm}
-\end{itemize}
-\item When additive costs are included:
+\item[$\Rightarrow$] Estimate Equation (\ref{eq:model_with_add}) constraining $t=0$:
 \footnotesize
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add}
-\end{equation}
-\normalsize
-\begin{itemize}
-\item[-] From which we re-built:
-$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item When assuming iceberg costs only
 \begin{equation}
@@
 \end{equation}
+\normalsize
+\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
 \begin{itemize}
-\item[-] From which we re-built:
-$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
-\end{itemize}
-\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
-\begin{itemize}
-\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{\tau}^{add}$  \vspace{0.1cm}
+\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
 \item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
@@
 \begin{frame}[label=slide_results_summary]
-\frametitle{Decomposing transport costs: Additive vs Iceberg}
-\textbf{Result 1}: Provide estimates for the size of both the ad-valorem and the additive components of transport costs
+\frametitle{Result 1: Estimating transport costs over time}
+For both the ad-valorem and the additive components of transport costs (by year \& transport mode)\vspace{0.1cm}
 \begin{itemize}
-\item Transport costs expressed in percent of the export price, average values over 1974-2013 \hyperlink{app_results_summary}{\beamergotobutton{More}}  \vspace{0.1cm}
+\item[-] \footnotesize{ Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
+\end{itemize}
 \begin{table}[htbp]
@@
     %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}$)} \\ \hline
+    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
     Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
@@
         \# obs. & 29279 & 28207 & 29317 & 27680 \\
-    \# origin country & 188 & 191 & 188 & 189 \\
-    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
+   % \# origin country & 188 & 191 & 188 & 189 \\
+   % \# products & 230 & 211 & 666 & 567 \\  \hline \hline
   \end{tabular}
@@
 \end{table}%
-
+\begin{itemize}
+\item Transport costs are sizeable \vspace{0.1cm}
+\begin{itemize}
+\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
+\end{itemize}
+\item Opening the black box of transport costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air \& ocean resp. \vspace{0.1cm}
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
+\end{itemize}
 \end{itemize}
@@
 
-\begin{frame}
-\frametitle{Comments}
+\begin{frame}[label=slide_result2]
+\frametitle{Result 2: Additive transport costs do matter}
 \begin{itemize}
-\item An important magnitude of overall transport costs  \vspace{0.1cm}
+\item International transport costs: A sizeable additive component \vspace{0.1cm}
 \begin{itemize}
-\item[-] Over 1974-2013, correspond to a mean increase of the export price by 5.8\% for ocean shipping, by 5.1\% for air transport  \vspace{0.1cm}
+\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.1cm}
+\begin{itemize}
+\item[$\star$] The ad-valorem cost reduces by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
+\item[$\star$] Amounts to 3.2\% and 2.5\% as mean values in vessel and air respectively in presence of additive costs \vspace{0.1cm}
+\item[$\star$] vs 5.8\% and 5.1\% when estimated alone \vspace{0.1cm}
 \end{itemize}
-\item When decomposing the structure of transport costs: A sizeable additive component   \vspace{0.1cm}
+\item[-] A sizeable share of the additive component in total transport costs \vspace{0.1cm}
 \begin{itemize}
-\item[-] Additive costs amount to 2.9\% of the export price in ocean shipping, 1.8\% in air transport  \vspace{0.1cm}
+\item[$\star$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
+\item[$\star$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1 later}} \vspace{0.1cm}
 \end{itemize}
-\item Omitting the additive term substantially biases the iceberg component upwards  \vspace{0.1cm}
+\end{itemize}
+\item A result confirmed by goodness of fit comparisons \vspace{0.1cm}
 \begin{itemize}
-\item[-] The ad-valorem cost reduces by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
-\item[-] 3.2\% and 2.5\% as mean values in vessel and air respectively (in presence of additive costs) \vspace{0.1cm}
+\item[-] 4 measures of goodness of fit for Models (\ref{eq:model_with_add}) (with additive TC) and (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
+\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}}
 \end{itemize}
@@
 \begin{frame}[label=slide_fig1]
-\frametitle{More on the importance of additive transport costs}
+\frametitle{Result 3: Characterizing the trends of transport costs }
 \begin{itemize}
-\item Using the time dimension of our database, study the shares of both ad-valorem and additive components in total transport costs
-\item[] \hyperlink{slide_moreonfig1}{\beamergotobutton{More on this figure later}}
-\end{itemize}
+\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Result 3}}
 \begin{figure}[htbp]
@@
 \end{figure}
+\item[$\Rightarrow$] Downward trend in overall transport costs, for both air and ocean shipping since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
+\begin{itemize}
+\footnotesize{
+\item[-] From 1980, a 1.5- 2 percentage points decrease in Air, a 2-3 percentage point decrease in Vessel }\vspace{0.1cm}
+\end{itemize}
+\
+\end{itemize}
 \end{frame}
 
-\begin{frame}[label=slide_comments_fig1]
-\frametitle{Comments}
+\begin{frame}[label=slide_compositioneffects]
+\frametitle{Time trends in transport costs: Excluding the composition effects }
 \begin{itemize}
-\item The size of transport costs is larger for ocean shipping than from air transport \vspace{0.1cm}
+\item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.2cm}
+\item The change in overall transport costs over time: \vspace{0.1cm}
 \begin{itemize}
-\item[-] On average over the period (Table 1) and on a yearly basis (Figure 1) \hyperlink{app_comments_fig1}{\beamergotobutton{More}}  \vspace{0.1cm}
-\end{itemize}
-\item A sizeable share of the additive component in total transport costs  \vspace{0.1cm}
+\item[-] Depend on the evolution of per product- per partner costs, \vspace{0.1cm}
+\item[-] But also on the composition of trade flows \vspace{0.2cm}
 \begin{itemize}
-\item[-] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
-\item[-] A decreasing share of additive costs over time for air transport \vspace{0.1cm}
-\item[-] Roughly constant for ocean shipping
+\item[$\star$] Over time, import more goods that are cheaper to transport, and/or from countries with which it is cheaper to trade \vspace{0.1cm}
 \end{itemize}
 \end{itemize}
-$\Rightarrow$ \textbf{Result 2}: Additive costs, an important dimension of international transport costs \vspace{0.1cm}
-\begin{itemize}
-\item Further investigate the robustness of this result
+\item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
+\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) \hyperlink{app_compositioneffects}{\beamergotobutton{More on the method}} 
 \end{itemize}
@@
 \begin{frame}
-\frametitle{The role of additive costs: Goodness-of-fit evaluations}
-\begin{itemize}
-\item Provide a more systematic diagnosis about the importance of additive costs  \vspace{0.1cm}
-\item By comparing the goodness-of-fit of the regressions
-\begin{itemize}
-\item[-] Obtained under Specification (\ref{eq:model_nlI}) (no additive costs)
-\item[-] vs Specification (\ref{eq:model_with_add}) (with additive costs) \vspace{0.1cm}
-\end{itemize}
-\item Various measures of goodness of fit
-\begin{itemize}
-\item[-] The $R^2$ (the larger the value, the better the fit) \vspace{0.1cm}
-\item[-] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
-\item[-] The log-likelihood function and two derived measures \vspace{0.1cm}
-\begin{itemize}
-\item[$\ast$] The Akaike Information Criterion (the lower AIC, the better the fit) \vspace{0.1cm}
-\item[$\ast$] The log-likelihood ratio test ($H_0$: both models are equivalent)
-\end{itemize}
-\end{itemize}
-
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Goodness of fit comparison}
-\begin{itemize}
-\item Air, 3 digit - level, selected years
-\end{itemize}
-\begin{table}[htbp]
-  \centering
- % \caption{Air: Measures of Goodness-of-fit (3 digits)}
-  \scriptsize{
-\begin{center}
-    \begin{tabular}{l|ccccc|c}
-    \hline \hline
-    Year  &1980  & 1990  & 2000  & 2010 & 2013 & Mean stat \\ \hline
-    \multicolumn{7}{l}{\bf{$R^2$} }\\ \hline
-    Term I only &  0.27  & 0.25  & 0.32  & 0.42 & 0.34 & 0.31 \\
-    Terms A \& I &  0.65  & 0.63  & 0.64  & 0.51 & 0.46 & 0.60 \\ \hline
-    \multicolumn{7}{l}{\textbf{SER}  }  \\ \hline
-    Term I only &  0.86  & 0.81  & 0.84  & 0.86 & 0.92 & 0.85 \\
-    Terms A \& I &  0.71  & 0.67  & 0.70  & 0.79 & 0.85 & 0.73 \\ \hline
-   \multicolumn{7}{l}{\textbf{AIC criteria}}  \\ \hline
-    Term I only &  41171.0 & 60715.6 & 87492.6 & 102297.6 & 88191.9 & 70498.1 \\
-    Terms A \& I & 35738.4 & 52098.9 & 74954.9 & 95887.1 & 80873.7 & 62285.0 \\ \hline
-    \multicolumn{7}{l}{\textbf{Log-likelihood}} \\ \hline
-    Term I only &  -20253.5 & -29977.8 & -43341.3 & -50746.8 & -43692.9 & -34888.6 \\
-    Terms A \& I &  -17263.2 & -25393.5 & -36788.4 & -47277.5 & -39751.9 & -30508.3 \\
-    LL ratio &  5980.6 & 9168.7 & 13105.7 & 6938.6 & 7882.1 & 8760.69 \\
-    nb of restrictions & 369   & 393   & 426   & 426 & 427 & 402 \\
-    p-value & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 \\
-    \hline \hline
- \end{tabular}%
-    \end{center}}
-  \label{tab:good_fit_air}%
- \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
-The mean statistics is calculated as the average value over all years.
-}}
-\end{table}%
-\end{frame}
-
-\begin{frame}
-\frametitle{Goodness of fit comparison (cont')}
-\begin{itemize}
-\item Vessel, 3 digit - level, selected years
-\end{itemize}
-\begin{table}[htbp]
-  \centering
-  %\caption{Vessel: Measures of Goodness-of-fit (3 digits)}
-    \scriptsize{
-\begin{center}
-\begin{tabular}{l|ccccc|c}
-\hline \hline
-Year  &  1980 & 1990 & 2000 & 2010  & 2013 & Mean stat \\ \hline
-\multicolumn{7}{l}{\bf{$R^2$} }\\ \hline
-Term I only &  0.415 & 0.456 & 0.401 & 0.350 & 0.339 & 0.39 \\
-Terms A \& I &  0.575 & 0.590 & 0.571 & 0.491 & 0.462 & 0.56 \\ \hline
-\multicolumn{7}{l}{\textbf{SER}  }  \\ \hline
-    Term I only &  0.62 & 0.59 & 0.65 & 0.74  & 0.76 & 0.66 \\
-    Terms A \& I &  0.53 & 0.51 & 0.55 & 0.66  & 0.68 & 0.57 \\ \hline
-   \multicolumn{7}{l}{\textbf{AIC criteria}}  \\ \hline
-    Term I only & 33010.3 & 51142.6 & 71365.9 & 84789.9 & 88191.9 & 57848.6 \\
-    Terms A \& I &  28067.3 & 43664.7 & 60475.9 & 76161.3 & 80873.7 & 49682.3 \\ \hline
-    \multicolumn{7}{l}{\textbf{Log-likelihood}} \\ \hline
-    Term I only &  -16129.1 & -25169.3 & -35263.9 & -41998.9 & -43692.9 & -28534.3 \\
-    Terms A \& I &  -13353.7 & -21171.4 & -29491.0 & -37418.7 & -39751.9 & -24151.3 \\
-    LL ratio & 5550.96 & 7995.88 & 11545.98 & 9160.56 & 7882.15 & 8766.0 \\
-    nb of restrictions  &  395 & 411 & 436 & 424   & 427 & 417 \\
-    p-value & 0.00 & 0.00 & 0.00 & 0.00  & 0.00 & 0.00 \\
-    \hline \hline
-    \end{tabular}%
-    \end{center}}
-  \label{tab:good_fit_vessel}%
-  \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
-The mean statistics calculated as the average value over all years.
-}}
-\end{table}%
-\end{frame}
-
-\begin{frame}
-\frametitle{Comments}
-\begin{itemize}
-\item \textbf{Result 2 (confirmed)}: Including the additive component improves the fit of the model, whatever the considered criterion or the transport mode \vspace{0.1cm}
-\begin{itemize}
-\footnotesize{
-\item[-] On average over the period, the $R^2$ doubles for air transport, increases by 50\% for vessel (also on a yearly basis) \vspace{0.1cm}
-}
-\end{itemize}
-\item A decrease in the quality of fit strongly after 2000, for both transport modes \vspace{0.1cm}
-\item Contrasted results between air and ocean transports \vspace{0.1cm}
-\begin{itemize}
-\item[-] Ad-valorem costs have more explanative power in ocean shipping than in air transport \vspace{0.1cm}
-\begin{itemize}
-\item[$\ast$] On average, account for 39\% of the variance of the cif-fab ratio, vs 31\% in Air \vspace{0.1cm}
-\end{itemize}
-\item[-] But their explanatory power seems to have decreased over time \vspace{0.1cm}
-\begin{itemize}
-\item[$\ast$] A roughly constant contribution of the additive component to the goodness of fit over time \vspace{0.1cm}
-\end{itemize}
-\item[$\neq$] For Air: The decreasing role of the explanatory power of the additive component over the recent years (consistent with Figure 1)\vspace{0.1cm}
-\end{itemize}
-\item[$\Rightarrow$] Go deeper in the analysis of the time-trends
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label=slide_moreonfig1]
-\frametitle{Decomposing transport costs: Characterizing the trends}
-\begin{itemize}
-\item Coming back to Figure 1 \hyperlink{slide_fig1}{\beamergotobutton{Go}}: Downward trend in overall transport costs, for both air and ocean shipping \vspace{0.1cm}
-\begin{itemize}
-\item[-] Considering 1980 as starting date, a 1.5- 2 percentage points decrease in Air, a 2-3 percentage point decrease in Vessel \vspace{0.1cm}
-\end{itemize}
-\item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.1cm}
-\item The change in overall transport costs over time: Depend on the evolution of per product- per partner costs, but also on the composition of trade flows \vspace{0.1cm}
-\begin{itemize}
-\item[-] Over time, import more goods that are cheaper to transport, and/or from countries with which it is cheaper to trade \vspace{0.1cm}
-\end{itemize}
-\item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}  \vspace{0.1cm}
-    \begin{itemize}
-\item[-] What we do, in accordance with Hummels (2007) \hyperlink{app_compositioneffects}{\beamergotobutton{More on the method}}
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Time trends in transport costs \textit{per se}}
 \begin{itemize}
@@
 
-\begin{frame}
+\begin{frame}[label=slide_comment_compositioneffects]
 \frametitle{Comments}
-\textbf{Result 3: About the time trends in transport costs}
 \begin{itemize}
-\item The reduction of the \textit{pure} transport costs : Starts in 1985 \vspace{0.1cm}
+\item The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
 \begin{itemize}
 \item[-] For both transport modes \vspace{0.1cm}
-\item[-] The reduction between 1974 and 1984 (Figure 1), attributable to change in the composition of trade patterns \vspace{0.1cm}
+\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}), attributable to change in the composition of trade patterns \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in air over the 1985-2005 period \vspace{0.1cm}
-\item[-] A result in accordance with Hummels (2007), Behar and Venables (2011) \vspace{0.1cm}
+\item[-] A result in accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
 \item[-] No more substantial difference when additive costs are included (Panel (b))
@@
 
-\begin{frame}[label=app_method]
-\frametitle{More on the estimation method}
+\begin{frame}[label=app_method_1]
+\frametitle{More on the estimation strategy (1)}
+\begin{itemize}
+\item Assumptions on the specification of transport costs (as in Irrarazabal et al., 2015) \vspace{0.1cm}
+\begin{itemize}
+\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
+\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
+\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
+\begin{equation}
+ \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
+ \end{equation}
+\end{itemize}
+\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item[$\Rightarrow$] The estimated equation becomes:
+$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
+\begin{itemize}
+\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
+\end{itemize}
+\end{itemize}
+\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}[label=app_method_2]
+\frametitle{More on the estimation strategy (2)}
 \begin{itemize}
@@
 \item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not decrease \textbf{increase?} from one iteration to the next. \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not increase? from one iteration to the next \vspace{0.1cm}
 \end{itemize}
 \item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
-
+\item Obtaining the final estimates: More details  \vspace{0.1cm}
+\begin{itemize}
+\item[-] After estimating Equations (\ref{eq:model_with_add}) and (\ref{eq:model_nlI}), re-built \vspace{0.1cm}
+\begin{itemize}
+\item[$\star$] With additive costs:
+$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
+\item[$\star$] With only iceberg costs:
+$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
+\end{itemize}
+\item[-] Then, take the (weighted) average by country-product
+\end{itemize}
 \end{itemize}
@@
 
-\begin{frame}[label=app_comments_fig1]
-\frametitle{Ad-valorem costs over time \hyperlink{slide_comments_fig1}{\beamergotobutton{Back to slide}}}
+\begin{frame}[label=app_goodnessfit]
+\frametitle{The role of additive costs: Goodness-of-fit evaluations}
+\begin{itemize}
+\item Provide a systematic diagnosis about the importance of additive costs  \vspace{0.1cm}
+\item By comparing the goodness-of-fit of the regressions
+\begin{itemize}
+\item[-] Obtained under Specification (\ref{eq:model_nlI}) (no additive costs)
+\item[-] vs Specification (\ref{eq:model_with_add}) (with additive costs) \vspace{0.1cm}
+\end{itemize}
+\item Various measures of goodness of fit
+\begin{itemize}
+\item[-] The $R^2$ (the larger the value, the better the fit) \vspace{0.1cm}
+\item[-] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
+\item[-] The log-likelihood function and two derived measures \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] The Akaike Information Criterion (the lower AIC, the better the fit) \vspace{0.1cm}
+\item[$\ast$] The log-likelihood ratio test ($H_0$: both models are equivalent)
+\end{itemize}
+\end{itemize}
+
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Goodness of fit comparison}
+\begin{itemize}
+\item Air, 3 digit - level, selected years
+\end{itemize}
+\begin{table}[htbp]
+  \centering
+ % \caption{Air: Measures of Goodness-of-fit (3 digits)}
+  \scriptsize{
+\begin{center}
+    \begin{tabular}{l|ccccc|c}
+    \hline \hline
+    Year  &1980  & 1990  & 2000  & 2010 & 2013 & Mean stat \\ \hline
+    \multicolumn{7}{l}{\bf{$R^2$} }\\ \hline
+    Term I only &  0.27  & 0.25  & 0.32  & 0.42 & 0.34 & 0.31 \\
+    Terms A \& I &  0.65  & 0.63  & 0.64  & 0.51 & 0.46 & 0.60 \\ \hline
+    \multicolumn{7}{l}{\textbf{SER}  }  \\ \hline
+    Term I only &  0.86  & 0.81  & 0.84  & 0.86 & 0.92 & 0.85 \\
+    Terms A \& I &  0.71  & 0.67  & 0.70  & 0.79 & 0.85 & 0.73 \\ \hline
+   \multicolumn{7}{l}{\textbf{AIC criteria}}  \\ \hline
+    Term I only &  41171.0 & 60715.6 & 87492.6 & 102297.6 & 88191.9 & 70498.1 \\
+    Terms A \& I & 35738.4 & 52098.9 & 74954.9 & 95887.1 & 80873.7 & 62285.0 \\ \hline
+    \multicolumn{7}{l}{\textbf{Log-likelihood}} \\ \hline
+    Term I only &  -20253.5 & -29977.8 & -43341.3 & -50746.8 & -43692.9 & -34888.6 \\
+    Terms A \& I &  -17263.2 & -25393.5 & -36788.4 & -47277.5 & -39751.9 & -30508.3 \\
+    LL ratio &  5980.6 & 9168.7 & 13105.7 & 6938.6 & 7882.1 & 8760.69 \\
+    nb of restrictions & 369   & 393   & 426   & 426 & 427 & 402 \\
+    p-value & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 \\
+    \hline \hline
+ \end{tabular}%
+    \end{center}}
+  \label{tab:good_fit_air}%
+ \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
+The mean statistics is calculated as the average value over all years.
+}}
+\end{table}%
+\hyperlink{slide_result2}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}
+\frametitle{Goodness of fit comparison (cont')}
+\begin{itemize}
+\item Vessel, 3 digit - level, selected years
+\end{itemize}
+\begin{table}[htbp]
+  \centering
+  %\caption{Vessel: Measures of Goodness-of-fit (3 digits)}
+    \scriptsize{
+\begin{center}
+\begin{tabular}{l|ccccc|c}
+\hline \hline
+Year  &  1980 & 1990 & 2000 & 2010  & 2013 & Mean stat \\ \hline
+\multicolumn{7}{l}{\bf{$R^2$} }\\ \hline
+Term I only &  0.415 & 0.456 & 0.401 & 0.350 & 0.339 & 0.39 \\
+Terms A \& I &  0.575 & 0.590 & 0.571 & 0.491 & 0.462 & 0.56 \\ \hline
+\multicolumn{7}{l}{\textbf{SER}  }  \\ \hline
+    Term I only &  0.62 & 0.59 & 0.65 & 0.74  & 0.76 & 0.66 \\
+    Terms A \& I &  0.53 & 0.51 & 0.55 & 0.66  & 0.68 & 0.57 \\ \hline
+   \multicolumn{7}{l}{\textbf{AIC criteria}}  \\ \hline
+    Term I only & 33010.3 & 51142.6 & 71365.9 & 84789.9 & 88191.9 & 57848.6 \\
+    Terms A \& I &  28067.3 & 43664.7 & 60475.9 & 76161.3 & 80873.7 & 49682.3 \\ \hline
+    \multicolumn{7}{l}{\textbf{Log-likelihood}} \\ \hline
+    Term I only &  -16129.1 & -25169.3 & -35263.9 & -41998.9 & -43692.9 & -28534.3 \\
+    Terms A \& I &  -13353.7 & -21171.4 & -29491.0 & -37418.7 & -39751.9 & -24151.3 \\
+    LL ratio & 5550.96 & 7995.88 & 11545.98 & 9160.56 & 7882.15 & 8766.0 \\
+    nb of restrictions  &  395 & 411 & 436 & 424   & 427 & 417 \\
+    p-value & 0.00 & 0.00 & 0.00 & 0.00  & 0.00 & 0.00 \\
+    \hline \hline
+    \end{tabular}%
+    \end{center}}
+  \label{tab:good_fit_vessel}%
+  \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
+The mean statistics calculated as the average value over all years.
+}}
+\end{table}%
+\hyperlink{slide_result2}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}[label=app_fig1]
+\frametitle{Ad-valorem costs over time \hyperlink{slide_fig1}{\beamergotobutton{Back to slide}}}
 \begin{figure}[htbp]
@@
 \end{itemize}
- \hyperlink{slide_moreonfig1}{\beamergotobutton{Back to slide}}
+ \hyperlink{slide_compositioneffects}{\beamergotobutton{Back to slide}}
 \end{frame}

### 2016-09-06 redaction/slides_LEDa.tex
Nouvelle version encore

--- 2016-08-26 redaction/slides_ETSG.tex
@@
 
-\date[September 2016]{ETSG, Helsinki, September 2016}
+\date[September 2016]{S\'{e}minaire LEDa, Université Paris Dauphine, Septembre 2016}
 \title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
@@
 \item[-] Transport costs \textit{per se} have fallen since 1985, by 40\% \vspace{0.1cm}
-\item[-] When additive costs are included, not much difference between air and sea transport
+\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011)
 \end{itemize}
@@
 $$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
-\item For product $k$, from country $i$ (and by transport mode, air or vessel) \vspace{0.1cm}
+\item For product $k$, from country $i$  \vspace{0.1cm}
 \item Rewrite to get:
 $$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
-\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level)
+\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
+\begin{itemize}
+\item[-] Notice the equation is also year- and mode (air or vessel)- specific
+\end{itemize}
 \end{itemize}
@@
 \end{equation}
-\normalsize
-\begin{itemize}
-\item[-] From which we re-built:
-$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
-\end{itemize}
+
 \end{itemize}
@@
 \end{equation}
+
+\item From which we re-built: \vspace{0.1cm}
 \begin{itemize}
-\item[-] From which we re-built:
+\item[-] With additive costs:
+$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
+\item[-] With only iceberg costs:
 $$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$

### 2016-09-06 redaction/slides_ETSG1.tex
encore...

@@
 \usepackage{epic,eepic,bez123}
+\usepackage{ bbold }
+\usepackage{ upgreek }
+\usepackage{amsmath}
 %\usepackage[T1]{fontenc}
@@
 \begin{itemize}
-\item[-] Account for 21\% of international trade costs (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
+\item[-] Account for 21\% of international trade costs $\Leftrightarrow$ A 15\% markup (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
 %\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
 \end{itemize}
-\item[$\Rightarrow$] If much trade policy barriers have been removed, the transport cost component of trade costs remains sizeable \vspace{0.1cm}
+\item[$\Rightarrow$] If many trade policy barriers have been removed, the transport cost component of trade costs does not seem to be declining\vspace{0.1cm}
 \end{itemize}
@@
 \item Yet... A debated question \vspace{0.1cm}
-\item Wouldn't trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
-\item The structure (additive vs iceberg) of transport costs, far from being anecdotal \vspace{0.1cm}
+\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
+\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.1cm}
 \begin{itemize}
 \footnotesize{
-\item[-] Additive trade costs, an important role in shaping the pattern of trade flows: Martin (2012), Hummels \& Skiba (2004), Irrarazabal et al. (2015) \vspace{0.1cm}
-\item[-] With strong normative implications: Sorensen (2014), Irrarazabal et al. (2015) \vspace{0.1cm}}
+\item[-] Additive trade costs \& the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
+\item[-] Strong normative implications (Sorensen, 2014)\vspace{0.1cm}
+\item[-] The additive structure of trade costs is supported by recent empirical evidence (Hummels \& Skiba, 2004)  \vspace{0.1cm}}
 \end{itemize}
@@
 \item[(1)] What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
-\item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
+\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
 \begin{itemize}
 \footnotesize{
-\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
+\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013, US imports) \vspace{0.1cm}
 \item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
@@
 \begin{itemize}
-\item[-] The estimated iceberg component reduces by a factor of 2 \vspace{0.1cm}
+\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
 \item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
@@
 \item[-] Transport costs \textit{per se} have fallen since 1985, by $\simeq$ 40\% \vspace{0.1cm}
-\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011)
+\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011) \vspace{0.1cm}
+%\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Plan of the talk}
+\frametitle{Outline of the talk}
 \begin{itemize}
@@
 \begin{itemize}
-\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
+\item Our measure of international transport costs: The difference between the export (fas) price and the import (cif) price \vspace{0.1cm}
 \item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
 \begin{itemize}
-\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
+\item[-] The export ($\simeq$ fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
 \item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
 \item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
+\item[-] A dataset also used by Hummels (2007) (until 2004) \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] At the 3-digit classification level  \vspace{0.1cm}
-\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
+\item[-] At the 4-digit level on some selected years as robustness check \vspace{0.1cm}
 \item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
@@
 \item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
-\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
+\item[-] With only ad-valorem costs: $\widehat{\tau}^{ice}$
 \end{itemize}
@@
 \frametitle{Result 1: Estimating transport costs over time}
-For both the ad-valorem and the additive components of transport costs (by year \& transport mode)\vspace{0.1cm}
+For both the ad-valorem and the additive components (by year \& transport mode)\vspace{0.1cm}
 \begin{itemize}
-\item[-] \footnotesize{ Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
+\item[-] \footnotesize{Average weighted values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
 \end{itemize}
@@
     %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
+    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
     Mode  & Vessel & Air & Vessel & Air \\ \hline
@@
     %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-        \# obs. & 29279 & 28207 & 29317 & 27680 \\
+        \# obs. & 29279 & 28207 & 29317 & 27680 \\ \hline \hline 
    % \# origin country & 188 & 191 & 188 & 189 \\
-   % \# products & 230 & 211 & 666 & 567 \\  \hline \hline
+   % \# products & 230 & 211 & 666 & 567 \\  
   \end{tabular}
+%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
+%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
 \end{center}}
-%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.
-The additive term is expressed in fraction of fab price.
-($^\ast$): Four 4-digit estimation: 0n selected years.
-($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
   \end{table}%
@@
 \begin{itemize}
-\item[$\star$] The ad-valorem cost reduces by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
-\item[$\star$] Amounts to 3.2\% and 2.5\% as mean values in vessel and air respectively in presence of additive costs \vspace{0.1cm}
+\item[$\star$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
+\item[$\star$] An iceberg cost equal 3.2\% and 2.5\% as mean values in vessel and air respectively in presence of additive costs \vspace{0.1cm}
 \item[$\star$] vs 5.8\% and 5.1\% when estimated alone \vspace{0.1cm}
@@
 \item[-] 4 measures of goodness of fit for Models (\ref{eq:model_with_add}) (with additive TC) and (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
-\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}}
+\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}} \vspace{0.1cm}
+\begin{itemize}
+\item[$\star$] Taking into account the additional degrees of freedom
+\end{itemize}
 \end{itemize}
@@
 {\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
-\includegraphics[width=5cm, height=5cm]{Fig2a_decompTC_air_3d.pdf}
-& \includegraphics[width=5cm,height=5cm]{Fig2b_decompTC_vessel_3d.pdf} \\
+\includegraphics[width=5cm, height=4cm]{Fig2a_decompTC_air_3d.pdf}
+& \includegraphics[width=5cm,height=4cm]{Fig2b_decompTC_vessel_3d.pdf} \\
 \end{tabular}
@@
 \end{figure}
-\item[$\Rightarrow$] Downward trend in overall transport costs, for both air and ocean shipping since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
 \begin{itemize}
-\footnotesize{
-\item[-] From 1980, a 1.5- 2 percentage points decrease in Air, a 2-3 percentage point decrease in Vessel }\vspace{0.1cm}
+\item[-] Lower overall transport costs in Air than in Vessel \vspace{0.1cm}
+\item[-] Downward trend for both modes since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
+%\item[-] Slighlty ore pronounced in ocean shipping \vspace{0.1cm}
+\begin{itemize}
+\item[$\star$] A 50\% decrease in Air, a 60\% decrease in Vessel \vspace{0.1cm}
 \end{itemize}
-\
+\end{itemize}
 \end{itemize}
@@
 \begin{frame}[label=slide_compositioneffects]
-\frametitle{Time trends in transport costs: Excluding the composition effects }
+\frametitle{Time trends in transport costs \& composition effects }
 \begin{itemize}
@@
 \begin{frame}[label=slide_comment_compositioneffects]
-\frametitle{Comments}
+\frametitle{Three main findings}
 \begin{itemize}
-\item The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
+\item The importance of excluding composition effects  \vspace{0.1cm}
 \begin{itemize}
-\item[-] For both transport modes \vspace{0.1cm}
-\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}), attributable to change in the composition of trade patterns \vspace{0.1cm}
+\item[-] The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
+\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}) is attributable to change in the composition of trade patterns \vspace{0.1cm}
 \end{itemize}
 \item Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
-\item Comparing air and ocean shipping \vspace{0.1cm}
+\item The importance of the additive component of transport costs (again) \vspace{0.1cm}
 \begin{itemize}
-\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in air over the 1985-2005 period \vspace{0.1cm}
-\item[-] A result in accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
-\item[-] No more substantial difference when additive costs are included (Panel (b))
+\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in Air transport over the 1985-2005 period \vspace{0.1cm}
+\item[-] In accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
+\item[-] But... No substantial difference when additive costs are included (Panel (b))
 \end{itemize}
-\item[$\Rightarrow$] The importance of taking into account the additive component of trade costs  \vspace{0.1cm}
 \end{itemize}
@@
 \begin{center}
-\includegraphics[width=8cm, height=4cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
+\includegraphics[width=8cm, height=3.5cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
 %{\footnotesize {OECD data} }
@@
 \begin{itemize}
-\item For Vessel: Similar downward trend for both $\tau$ and $t$
-\item For Air: Much more contrasted trends
+\item For Vessel: Similar downward trend for both $\tau$ and $t$  \vspace{0.1cm}
+\item For Air: Much more contrasted trends  \vspace{0.1cm}
 \begin{itemize}
-\item[-] Substantial decrease in the ad-valorem costs over 1985-2005, but roughly constant additive costs
-\item[$\Rightarrow$] An explanation to the difference between Panels (a) and (b) above
+\item[-] Substantial decrease in the ad-valorem costs over 1985-2005, but roughly constant additive costs  \vspace{0.1cm}
+\item[$\Rightarrow$] An explanation to the difference between Panels (a) and (b) above  \vspace{0.1cm}
+
+\item[$\Rightarrow$] The importance of modeling additive costs  \vspace{0.1cm}
+\item[-] Trend reversal around 2005: Any suggestion?
 \end{itemize}
-\item[$\Rightarrow$] The importance of modeling additive costs
 \end{itemize}
@@
 \begin{itemize}
+\item[-] Based on the US imports flows from 1974 to 2013 \vspace{0.1cm}
 \item[-] Additive cost: amount to 2.8\% of the export price in ocean shipping, 1.8\% in air transport \vspace{0.1cm}
@@
 \item[-] Additive costs are far from negligible quantitatively \vspace{0.1cm}
-\item[-] A better fit of the model when taken into account    \vspace{0.1cm}
+\item[-] A better fit of the model when they are taken into account    \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
+\item[-] Compare the transports costs between air and vessel (excluding composition effects) \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] Only cover international transport costs \vspace{0.1cm}
-\item[-] Among transport costs, quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
+\item[-] Among transport costs, only quantitative costs (freight, insurance and handling - nothing about those related to the time value of goods) \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] US imports, large time period: Broad view of international trade flows \vspace{0.1cm}
-\item[-] A reliable database, already used by Hummels (2007), but on a larger period of time \vspace{0.1cm}
+\item[-] A reliable database, already used by Hummels (2007). But we have a longer period of time \vspace{0.1cm}
 \item[-] Have both the import and the export prices: Estimate both the levels of the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
@@
 \end{itemize}
-\item[-] Then, take the (weighted) average by country-product
+\item[-] Then, take the weighted average by country-product, \vspace{0.1cm}
+\item[-] Using the values of each trade flow (ik-specific) over total yearly trade as a weighting scheme
 \end{itemize}
@@
 \item[-] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
-\item[-] The log-likelihood function and two derived measures \vspace{0.1cm}
+\item[-] The log-likelihood function and two derived measures, which take into account the degrees of freedom \vspace{0.1cm}
 \begin{itemize}

### 2016-09-06 redaction/slides_ETSG1.tex
Last ?

@@
 \begin{itemize}
-\item Our measure of international transport costs: The difference between the export (fas) price and the import (cif) price \vspace{0.1cm}
+\item Our measure of international transport costs: The difference between the export ($\simeq$ fas) price and the import (cif) price \vspace{0.1cm}
 \item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
 \begin{itemize}
-\item[-] The export ($\simeq$ fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
+\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
 \item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
@@
 \normalsize
-\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
+\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
 \begin{itemize}
@@
 \frametitle{Result 1: Estimating transport costs over time}
-For both the ad-valorem and the additive components (by year \& transport mode)\vspace{0.1cm}
+For both the ad-valorem and the additive components \vspace{0.1cm}
 \begin{itemize}
-\item[-] \footnotesize{Average weighted values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
+\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
 \end{itemize}
@@
 \begin{itemize}
-\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air \& ocean resp. \vspace{0.1cm}
-\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
+\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp. 
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price 
 \end{itemize}
@@
 \item[$\star$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
-\item[$\star$] An iceberg cost equal 3.2\% and 2.5\% as mean values in vessel and air respectively in presence of additive costs \vspace{0.1cm}
-\item[$\star$] vs 5.8\% and 5.1\% when estimated alone \vspace{0.1cm}
+\item[$\star$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.1cm}
+\item[$\star$] From 5.1\% to 2.5\% in Air transport \vspace{0.1cm}
 \end{itemize}
@@
 & \includegraphics[width=5cm,height=4cm]{Fig2b_decompTC_vessel_3d.pdf} \\
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
 \end{figure}
@@
 \begin{frame}[label=slide_comment_compositioneffects]
-\frametitle{Three main findings}
+\frametitle{Two main findings}
 \begin{itemize}
@@
 \item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}) is attributable to change in the composition of trade patterns \vspace{0.1cm}
+\item[-] Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
 \end{itemize}
-\item Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
 \item The importance of the additive component of transport costs (again) \vspace{0.1cm}
@@
 
-\textbf{Two main possible extensions}
+\textbf{Three main possible extensions}
 \begin{itemize}
-\item On the empirical side: Go deeper in the structural determinants of transport costs \vspace{0.1cm}
+\item On the empirical side: \vspace{0.1cm}
 \begin{itemize}
-\item[-] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
-\item[-] Compare the transports costs between air and vessel (excluding composition effects) \vspace{0.1cm}
+\item[(1)] Compare the trends in transports costs \textit{between} air and vessel (excluding composition effects) \vspace{0.1cm}
+\item[(2)] Go deeper in the structural determinants of transport costs \vspace{0.1cm}
+\begin{itemize} 
+\item[$\star$] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
 \end{itemize}
-\item On the theoretical side: Use our results to explore the role of additive costs \vspace{0.1cm}
+\end{itemize}
+\item On the theoretical side: 
 \begin{itemize}
-\item[-] In shaping international trade flows (trade theory) \vspace{0.1cm}
-\item[-] In affecting the international transmission of business cycles (business cycle theory)
+\item[(3)]Use our results to explore the role of additive costs \vspace{0.1cm}
+\begin{itemize}
+\item[$\star$] In shaping international trade flows (trade theory) \vspace{0.1cm}
+\item[$\star$] In affecting the international transmission of business cycles (business cycle theory)
+\end{itemize}
 \end{itemize}
@@
 \item[-] Then, take the weighted average by country-product, \vspace{0.1cm}
-\item[-] Using the values of each trade flow (ik-specific) over total yearly trade as a weighting scheme
+\item[-] Using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme
 \end{itemize}

### 2016-09-06 redaction/slides_ETSG1.tex
cette fois c'est mineur

@@
 \item[-] Transport costs \textit{per se} have fallen since 1985, by $\simeq$ 40\% \vspace{0.1cm}
-\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011) \vspace{0.1cm}
+\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.1cm}
 %\item[-] Marked difference in the evolution of additive and multiplicative costs in Air

### 2016-09-07 redaction/slides_ETSG1.tex
encore et encore

@@
 \item[-] Declining over the second half of the 20$^{th}$ century (Jacks et al., 2008, Novy, 2013) \vspace{0.1cm}
-\item[-] But still significant: Average international trade costs = a 74\% markup over production costs (Anderson \& Van Wincoop, 2004)
+\item[-] But still significant: International trade costs $\simeq$ a 74\% markup over production costs (Early 90s, Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
 \end{itemize}
@@
 \begin{itemize}
-\item[-] Account for 21\% of international trade costs $\Leftrightarrow$ A 15\% markup (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
+\item[-] Amount to $\simeq 30\%$ of international trade costs  $\Leftrightarrow$ A 21\% markup over production costs  \vspace{0.1cm}
 %\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}

### 2016-09-11 redaction/slides_ETSG1.tex
version définitive

@@
 \item[-] Strong normative implications (Sorensen, 2014)\vspace{0.1cm}
-\item[-] The additive structure of trade costs is supported by recent empirical evidence (Hummels \& Skiba, 2004)  \vspace{0.1cm}}
+\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}}
 \end{itemize}
@@
 \begin{itemize}
-\item[(1)] What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
+\item[(1)] What is the size of the iceberg and the additive transport costs? \vspace{0.1cm}
 \item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
@@
 \item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
-$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
+$$p = \tau \widetilde{p} + t, \quad \text{with }~~\tau \geq 1, \quad t \geq 0$$
 \item For product $k$, from country $i$  \vspace{0.1cm}
@@
     Mode  & Vessel & Air & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
+    \multicolumn{5}{l}{With only Ad-Valorem Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
     Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
@@
     %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
+    \multicolumn{5}{l}{With Additive \& Ad-Valorem Costs } \\ \hline
    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
@@
 \end{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} \geq 1$, the error term should be always positive and multiplicative \vspace{0.1cm}
 \item[$\Rightarrow$] The estimated equation becomes:
@@
 \item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not increase? from one iteration to the next \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
 \end{itemize}
-\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
+\item Eliminate potential influence of outliers: Exclude the 5 \% of the upper and lower tails of the distribution \vspace{0.1cm}
 \item Obtaining the final estimates: More details  \vspace{0.1cm}

### 2016-09-22 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex

--- 2016-09-06 redaction/slides_ETSG1.tex
@@
 \usepackage{ upgreek }
-\usepackage{amsmath}
 %\usepackage[T1]{fontenc}
@@
 
-\date[September 2016]{ETSG, Helsinki, September 2016}
+\date[September 2016]{S\'{e}minaire LEDa, Universit\'{e} Paris Dauphine, Septembre 2016}
 \title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
@@
 \begin{itemize}
-\item[-] Account for 21\% of international trade costs $\Leftrightarrow$ A 15\% markup (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
+\item[-]  Amount to $\simeq 30\%$ of international trade costs  $\Leftrightarrow$ A 21\% markup over production costs  \vspace{0.1cm}
 %\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
 \end{itemize}
-\item[$\Rightarrow$] If many trade policy barriers have been removed, the transport cost component of trade costs does not seem to be declining\vspace{0.1cm}
+\item[$\Rightarrow$] If much trade policy barriers have been removed, the transport cost component of trade costs remains sizeable \vspace{0.1cm}
 \end{itemize}
@@
 \footnotesize{
-\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013, US imports) \vspace{0.1cm}
-\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
+\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.1cm}
 }
@@
 \begin{itemize}
+\footnotesize{
 \item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
 \item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
+}
 \end{itemize}
 \item[(3)] How have international transport costs evolved over time? \vspace{0.1cm}
+\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.1cm}
 \begin{itemize}
-\item[-] Transport costs \textit{per se} have fallen since 1985, by $\simeq$ 40\% \vspace{0.1cm}
-\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.1cm}
+\footnotesize{
+\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.1cm}
+}
+\end{itemize}
+\item[$\Rightarrow$] The importance of including the additive component \vspace{0.1cm}
+\begin{itemize}
+\footnotesize{
+\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.1cm}
 %\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
+%\end{itemize}
+}
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Outline of the talk}
+\frametitle{Plan of the talk}
 \begin{itemize}
-\item Data Sources  \vspace{0.1cm}
-\item Empirical Methodology \vspace{0.1cm}
-\item Results \vspace{0.1cm}
+\item Data Sources  \vspace{0.2cm}
+\item Empirical Methodology \vspace{0.2cm}
+\item Results \vspace{0.2cm}
 \item Conclusion
@@
 \begin{itemize}
-\item Our measure of international transport costs: The difference between the export ($\simeq$ fas) price and the import (cif) price \vspace{0.1cm}
-\item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
+\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
+\item Database: US Imports of Merchandise database \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
-\item[-] A dataset also used by Hummels (2007) (until 2004) \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] At the 3-digit classification level  \vspace{0.1cm}
-\item[-] At the 4-digit level on some selected years as robustness check \vspace{0.1cm}
+\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
 \item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
@@
 \begin{frame}
-\frametitle{Empirical specification: The estimated equation}
+\frametitle{More on our database}
+\begin{itemize}
+\item Implications (and limitations) \vspace{0.1cm}
+\begin{itemize}
+\item[-] Only cover international transport costs \vspace{0.1cm}
+\item[-] Among transport costs, quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
+%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
+\end{itemize}
+\item A rich database to exploit \vspace{0.1cm}
+\begin{itemize}
+\item[-] US imports, large time period: Broad view of international trade flows \vspace{0.1cm}
+\item[-] A reliable database, already used by Hummels (2007), but on a larger period of time \vspace{0.1cm}
+\item[-] Have both the import and the export prices: Estimate both the levels of the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Empirical specification (1)}
+\textbf{The estimated equation}
 \begin{itemize}
@@
 $$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
-\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 \vspace{0.1cm}
+\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
 \begin{itemize}
-\item[-] The equation being also year- and mode (air or vessel)- specific
+\item[-] The equation is also year- and mode (air or vessel)- specific
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Empirical specification (2)}
+\textbf{The estimation strategy}
+\begin{itemize}
+\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
+\begin{itemize}
+\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
+\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
+\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
+\begin{equation}
+ \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
+ \end{equation}
+\end{itemize}
+\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item[$\Rightarrow$] The estimated equation becomes:
+\footnotesize
+$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
+\normalsize
+\begin{itemize}
+\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
 \end{itemize}
@@
 \begin{frame} [label=slide_method]
-\frametitle{Empirical specification: The estimation strategy }
 \begin{itemize}
-\item With some assumptions on the specification of transport costs \& the error term, and taking logs  \hyperlink{app_method_1}{\beamergotobutton{More}}\vspace{0.1cm}
-\item[$\Rightarrow$] Estimate the following equation \vspace{0.1cm}
+\item Taking logs, we finally estimate the following equation
 \footnotesize
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_with_add}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
 \end{equation}
-\begin{itemize}
-\item[-] Estimates performed using non-linear least squares \hyperlink{app_method_2}{\beamergotobutton{More}} \vspace{0.1cm}
-\end{itemize}
 \normalsize
+\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
+\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.2cm}
 \item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
-\item[$\Rightarrow$] Estimate Equation (\ref{eq:model_with_add}) constraining $t=0$:
+\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
 \footnotesize
@@
 \normalsize
-\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
+\item Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
+\begin{itemize}
+\item[-] With additive costs included: Equation (\ref{eq:est_equation})  \vspace{0.1cm}
+\begin{itemize}
+\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
+\end{itemize}
+\item[-] With only ad-valorem costs: Equation (\ref{eq:model_nlI})
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame} 
+\begin{itemize}
+\item After running the estimates, we re-built: \vspace{0.1cm}
+\begin{itemize}
+\item[-] With additive costs:
+$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
+\item[-] With only iceberg costs:
+$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
+\end{itemize}
+\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
 \begin{itemize}
 \item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
-\item[-] With only ad-valorem costs: $\widehat{\tau}^{ice}$
+\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
 \end{itemize}
@@
     %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-        \# obs. & 29279 & 28207 & 29317 & 27680 \\ \hline \hline
-   % \# origin country & 188 & 191 & 188 & 189 \\
-   % \# products & 230 & 211 & 666 & 567 \\
+       \multicolumn{5}{l}{\textit{Data} }\\ \hline
+    Mean &5.6 &4.0 &5.7 & 4.0\\
+    Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline 
+    \# obs. & 29279 & 28207 & 29317 & 27680 \\ 
+    \# origin country & 188 & 191 & 188 & 189 \\
+    \# products & 230 & 211 & 666 & 567 \\    \hline \hline
   \end{tabular}
@@
   \end{table}%
+ \end{frame}
+
+ \begin{frame}
+\frametitle{Result 1: Comments}
 \begin{itemize}
@@
 \item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
+\item[-] Lower than the 21\% markup taken from AVW, but  \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
+\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
+\end{itemize}
+\item[$\Rightarrow$] Our work: A more precise and exhaustive estimation of international transport costs \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp. 
-\item[-] Additive cost: 1.8\% and 2.9\% of the export price 
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
 \end{itemize}
+\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration 
 \end{itemize}
@@
 \begin{itemize}
-\item International transport costs: A sizeable additive component \vspace{0.1cm}
+\item International transport costs: A sizeable additive component  \vspace{0.2cm}
 \begin{itemize}
-\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.1cm}
+\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
 \begin{itemize}
-\item[$\star$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
-\item[$\star$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.1cm}
-\item[$\star$] From 5.1\% to 2.5\% in Air transport \vspace{0.1cm}
+\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.2cm}
+\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.2cm}
+\item[$\ast$] From 5.1\% to 2.5\% in Air transport \vspace{0.2cm}
 \end{itemize}
-\item[-] A sizeable share of the additive component in total transport costs \vspace{0.1cm}
+\item[-] A sizeable share of the additive component in total transport costs \vspace{0.2cm}
 \begin{itemize}
-\item[$\star$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
-\item[$\star$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1 later}} \vspace{0.1cm}
-\end{itemize}
-\end{itemize}
-\item A result confirmed by goodness of fit comparisons \vspace{0.1cm}
-\begin{itemize}
-\item[-] 4 measures of goodness of fit for Models (\ref{eq:model_with_add}) (with additive TC) and (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
-\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}} \vspace{0.1cm}
-\begin{itemize}
-\item[$\star$] Taking into account the additional degrees of freedom
+\item[$\ast$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.2cm}
+\item[$\ast$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
 \end{itemize}
@@
 
-\begin{frame}[label=slide_fig1]
+\begin{frame}[label=slide_goodnessfit]
+\begin{itemize}
+\item A better quality of fit with the additive component included \vspace{0.2cm}
+\begin{itemize}
+\item[-] Compare the goodness-of-fit of Model (\ref{eq:est_equation}) (with additive TC)  \vspace{0.1cm}
+\item[ ] vs Model (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
+\item[-] Various measures of goodness of fit \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] The $R^2$ (the larger the value, the better the fit) \vspace{0.1cm}
+\item[$\ast$] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
+\item[$\ast$] The Akaike Information Criterion (the lower AIC, the better the fit) \vspace{0.1cm}
+\item[$\ast$] The log-likelihood ratio test ($H_0$: both models are equivalent) \vspace{0.1cm}
+\end{itemize}
+\end{itemize}
+\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component  \vspace{0.1cm}
+
+\begin{itemize}
+\item[-] \hyperlink{app_goodnessfit}{\beamergotobutton{See the results}} \vspace{0.1cm}
+\item[-] Even when taking into account the additional degrees of freedom
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label = slide_fig1]
 \frametitle{Result 3: Characterizing the trends of transport costs }
@@
 
-\begin{frame}[label=slide_compositioneffects]
+\begin{frame}
 \frametitle{Time trends in transport costs \& composition effects }
@@
 \item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
-\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) \hyperlink{app_compositioneffects}{\beamergotobutton{More on the method}}
+\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) 
 \end{itemize}
@@
 \begin{frame}
+\frametitle{Excluding the composition effects of transport costs changes}
+\begin{itemize}
+\item For the ad-valorem component, we estimate the following equation:
+\scriptsize
+\begin{eqnarray}
+%\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\simeq \ln(\tau_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{+ \simeq \ln(\tau_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\end{eqnarray}
+\normalsize
+\begin{itemize}
+\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
+\end{itemize}
+\item For the additive component:
+\scriptsize
+\begin{eqnarray}
+%\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\simeq \ln(t_i)}+\underbrace{\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k}_{+ \simeq t_k}\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
+\end{eqnarray}
+\normalsize
+\item Underlying rationale
+\begin{itemize}
+\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
+\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode) \vspace{0.1cm}
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item Exclude the composition effects of transport costs changes  \vspace{0.1cm}
+\item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
+\begin{itemize}
+\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
+\begin{equation*}
+\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
+\begin{itemize}
+\item[$\ast$] with $\bar{\tau}_{1974} = ???$
+\end{itemize}
+\item[-] For the additive cost, we build the variable ($\forall~t \geq 1974$)
+$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
+\end{itemize}
+\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
+\begin{itemize}
+\item[] with an initial value of 100 for $t=1974$
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Time trends in transport costs \textit{per se}}
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{Decomposing (\textit{pure}) transport costs over time }
+\frametitle{Decomposing \textit{pure} transport costs over time }
 \begin{figure}[htbp]
@@
 \begin{frame}
-\frametitle{Conclusion}
+\frametitle{Conclusion (1)}
 \textbf{Our paper: Empirical evidence about the role of the additive component in international transport costs} \vspace{0.1cm}
@@
 \begin{frame}
-\frametitle{Conclusion}
+\frametitle{Conclusion (2)}
 
@@
 \appendix
-\begin{frame}[label=app_data]
-\frametitle{More on our database}
-\begin{itemize}
-\item Implications (and limitations) \vspace{0.1cm}
-\begin{itemize}
-\item[-] Only cover international transport costs \vspace{0.1cm}
-\item[-] Among transport costs, only quantitative costs (freight, insurance and handling - nothing about those related to the time value of goods) \vspace{0.1cm}
-\end{itemize}
-\item A rich database to exploit \vspace{0.1cm}
-\begin{itemize}
-\item[-] US imports, large time period: Broad view of international trade flows \vspace{0.1cm}
-\item[-] A reliable database, already used by Hummels (2007). But we have a longer period of time \vspace{0.1cm}
-\item[-] Have both the import and the export prices: Estimate both the levels of the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
-\end{itemize}
-\end{itemize}
-\hyperlink{slide_data}{\beamergotobutton{Back to slide}}
-\end{frame}
 
-\begin{frame}[label=app_method_1]
-\frametitle{More on the estimation strategy (1)}
-\begin{itemize}
-\item Assumptions on the specification of transport costs (as in Irrarazabal et al., 2015) \vspace{0.1cm}
-\begin{itemize}
-\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
-\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
-\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
-\begin{equation}
- \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
- \end{equation}
-\end{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
-\item[$\Rightarrow$] The estimated equation becomes:
-$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
-\begin{itemize}
-\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
-\end{itemize}
-\end{itemize}
-\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
-\end{frame}
-
-\begin{frame}[label=app_method_2]
-\frametitle{More on the estimation strategy (2)}
+\begin{frame}[label=app_method]
+\frametitle{More on the estimation method}
 \begin{itemize}
@@
 \item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not increase? from one iteration to the next \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
 \end{itemize}
 \item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
-\item Obtaining the final estimates: More details  \vspace{0.1cm}
-\begin{itemize}
-\item[-] After estimating Equations (\ref{eq:model_with_add}) and (\ref{eq:model_nlI}), re-built \vspace{0.1cm}
-\begin{itemize}
-\item[$\star$] With additive costs:
-$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
-\item[$\star$] With only iceberg costs:
-$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
-\end{itemize}
-\item[-] Then, take the weighted average by country-product, \vspace{0.1cm}
-\item[-] Using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme
-\end{itemize}
+
 \end{itemize}
@@
 
-\begin{frame}[label=app_goodnessfit]
-\frametitle{The role of additive costs: Goodness-of-fit evaluations}
-\begin{itemize}
-\item Provide a systematic diagnosis about the importance of additive costs  \vspace{0.1cm}
-\item By comparing the goodness-of-fit of the regressions
-\begin{itemize}
-\item[-] Obtained under Specification (\ref{eq:model_nlI}) (no additive costs)
-\item[-] vs Specification (\ref{eq:model_with_add}) (with additive costs) \vspace{0.1cm}
-\end{itemize}
-\item Various measures of goodness of fit
-\begin{itemize}
-\item[-] The $R^2$ (the larger the value, the better the fit) \vspace{0.1cm}
-\item[-] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
-\item[-] The log-likelihood function and two derived measures, which take into account the degrees of freedom \vspace{0.1cm}
-\begin{itemize}
-\item[$\ast$] The Akaike Information Criterion (the lower AIC, the better the fit) \vspace{0.1cm}
-\item[$\ast$] The log-likelihood ratio test ($H_0$: both models are equivalent)
-\end{itemize}
-\end{itemize}
-
-\end{itemize}
+\begin{frame}[label=app_fig1]
+\frametitle{Ad-valorem costs over time \hyperlink{slide_fig1}{\beamergotobutton{Back to slide}}}
+\begin{figure}[htbp]
+\caption{Ad-valorem Costs (Yearly mean value, 3 digits)}
+\label{fig:mult_alone_withadd}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=5cm, height=2.5in]{Fig1a_mult_air_3d.pdf}
+& \includegraphics[width=5cm,height=2.5in]{Fig1b_mult_vessel_3d.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
 
@@
 
-\begin{frame}
-\frametitle{Goodness of fit comparison}
+\begin{frame}[label=app_goodnessfit]
+\frametitle{More on Result 2: Goodness of fit comparison}
 \begin{itemize}
@@
 \end{table}%
-\hyperlink{slide_result2}{\beamergotobutton{Back to slide}}
+\hyperlink{slide_goodnessfit}{\beamergotobutton{Back to slide}}
 \end{frame}
@@
 \end{table}%
-\hyperlink{slide_result2}{\beamergotobutton{Back to slide}}
-\end{frame}
-
-\begin{frame}[label=app_fig1]
-\frametitle{Ad-valorem costs over time \hyperlink{slide_fig1}{\beamergotobutton{Back to slide}}}
-\begin{figure}[htbp]
-\caption{Ad-valorem Costs (Yearly mean value, 3 digits)}
-\label{fig:mult_alone_withadd}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=5cm, height=2.5in]{Fig1a_mult_air_3d.pdf}
-& \includegraphics[width=5cm,height=2.5in]{Fig1b_mult_vessel_3d.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}[label=app_compositioneffects]
-\frametitle{Excluding the composition effects of transport costs changes}
-\begin{itemize}
-\item For the ad-valorem component, we estimate the following equation:
-\scriptsize
-\begin{eqnarray}
-\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(\tau_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
-\end{eqnarray}
-\normalsize
-\begin{itemize}
-\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
-\end{itemize}
-\item For the additive component:
-\scriptsize
-\begin{eqnarray}
-\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
-\end{eqnarray}
-\normalsize
-\end{itemize}
-
+\hyperlink{slide_goodnessfit}{\beamergotobutton{Back to slide}}
 \end{frame}
@@
 \begin{frame}
+\frametitle{Goodness of fit: Comments}
 \begin{itemize}
-\item More on the estimation method
+\item \textbf{Result 2}: Including the additive component improves the fit of the model, whatever the considered criterion or the transport mode \vspace{0.1cm}
 \begin{itemize}
-\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
-\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, \vspace{0.1cm}
-\item[-] Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode) \vspace{0.1cm}
+\footnotesize{
+\item[-] On average over the period, the $R^2$ doubles for air transport, increases by 50\% for vessel (also on a yearly basis) \vspace{0.1cm}
+}
 \end{itemize}
-\item Exclude the composition effects of transport costs changes $\Leftrightarrow$ Isolate the change in the time dimension  \vspace{0.1cm}
+\item A decrease in the quality of fit strongly after 2000, for both transport modes \vspace{0.1cm}
+\item Contrasted results between air and ocean transports \vspace{0.1cm}
 \begin{itemize}
-\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
-\begin{equation*}
-\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
-\item[-] For the additive cost, we build the variable ($\forall~t \geq 1974$)
-$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
+\item[-] Ad-valorem costs have more explanative power in ocean shipping than in air transport \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] On average, account for 39\% of the variance of the cif-fas ratio, vs 31\% in Air \vspace{0.1cm}
 \end{itemize}
-\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
+\item[-] But their explanatory power seems to have decreased over time \vspace{0.1cm}
 \begin{itemize}
-\item[] with an initial value of 100 for $t=1974$
+\item[$\ast$] For vessel: A roughly constant contribution of the additive component to the goodness of fit over time\vspace{0.1cm}
 \end{itemize}
-
+\item[$\neq$] For Air: The decreasing role of the explanatory power of the additive component over the recent years (consistent with Figure 1)\vspace{0.1cm}
 \end{itemize}
- \hyperlink{slide_compositioneffects}{\beamergotobutton{Back to slide}}
+\item[$\Rightarrow$] Go deeper in the analysis of the time-trends
+\end{itemize}
+\hyperlink{slide_goodnessfit}{\beamergotobutton{Back to slide}}
 \end{frame}

### 2016-09-23 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
changements

@@
 \usepackage{ upgreek }
+\usepackage{eurosym}
 %\usepackage[T1]{fontenc}
@@
 \item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
+\begin{itemize}
+\item[-] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
+\footnotesize
+$$pq = \tau \widetilde{p}q$$
+\normalsize
+\end{itemize}
 \end{itemize}
 \item Yet... A debated question \vspace{0.1cm}
-\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
-\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.1cm}
+\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
 \begin{itemize}
-\footnotesize{
-\item[-] Additive trade costs \& the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
-\item[-] Strong normative implications (Sorensen, 2014)\vspace{0.1cm}
-\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}}
+\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
+\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
+\begin{itemize}
+\item[$\star$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2014) \vspace{0.1cm}
 \end{itemize}
-\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.1cm}
+\item[-] Some trade policy instruments as well (quota license price)
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.2cm}
+\begin{itemize}
+\item[-] Additive trade costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
+\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
+\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
+\end{itemize}
+\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
 \end{itemize}
 \textbf{One objective of the paper: Provide an answer to this question}
+
 \end{frame}
@@
 \frametitle{Our paper in 3 questions (and 3 answers)}
-An empirical decomposition of the structure of transport costs \vspace{0.1cm}
+An empirical decomposition of the structure of transport costs \vspace{0.2cm}
 \begin{itemize}
-\item[(1)] What is the size of the iceberg and the additive transport costs? \vspace{0.1cm}
+\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
 \item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
 \begin{itemize}
-\footnotesize{
+%\footnotesize{
 \item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
-\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.1cm}
-}
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
+%}
 \end{itemize}
-\item[(2)] What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
+\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
 \item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
 \begin{itemize}
-\footnotesize{
+%\footnotesize{
 \item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
 \item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
-}
+%}
 \end{itemize}
-\item[(3)] How have international transport costs evolved over time? \vspace{0.1cm}
-\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.1cm}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \begin{itemize}
-\footnotesize{
-\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.1cm}
-}
+\item[(3)] \textbf{ How have international transport costs evolved over time?} \vspace{0.2cm}
+\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.2cm}
+\begin{itemize}
+%\footnotesize{
+\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
+%}
 \end{itemize}
-\item[$\Rightarrow$] The importance of including the additive component \vspace{0.1cm}
+\item[$\Rightarrow$] The importance of including the additive component \vspace{0.2cm}
 \begin{itemize}
-\footnotesize{
-\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.1cm}
-%\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
+%\footnotesize{
+\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.2cm}
+\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
 %\end{itemize}
-}
+%}
 \end{itemize}
@@
 \frametitle{Empirical specification (1)}
-\textbf{The estimated equation}
+\textbf{The equation at the root of the estimation}
 \begin{itemize}
@@
 \item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
-\item[$\Rightarrow$] The estimated equation becomes:
+\item[$\Rightarrow$] The equation of interest becomes:
 \footnotesize
@@
 
-\begin{frame} [label=slide_method]
+\begin{frame} 
 \begin{itemize}
 \item Taking logs, we finally estimate the following equation
-\footnotesize
+
 \begin{equation}
@@
 \end{equation}
-\normalsize
+\vspace{0.1cm}
 \item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
-\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.2cm}
-\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
-\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
-\footnotesize
+\item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
+\begin{itemize}
+\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
+\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Empirical specification (3)}
+\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
+\begin{itemize}
+\item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
+\item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
+\begin{itemize}
+\item[-] With additive costs included: 
+%\footnotesize
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
+\end{equation}  
+%\normalsize
+\begin{itemize}
+\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
+\end{itemize}
+\item[-] With only ad-valorem costs:
+%\footnotesize
 \begin{equation}
@@
 \end{equation}
-\normalsize
-\item Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
-\begin{itemize}
-\item[-] With additive costs included: Equation (\ref{eq:est_equation})  \vspace{0.1cm}
-\begin{itemize}
-\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
-\end{itemize}
-\item[-] With only ad-valorem costs: Equation (\ref{eq:model_nlI})
+%\normalsize
 \end{itemize}
@@
 \end{itemize}
-\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
+\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
 \begin{itemize}
@@
     %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-       \multicolumn{5}{l}{\textit{Data} }\\ \hline
-    Mean &5.6 &4.0 &5.7 & 4.0\\
-    Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline 
+    %   \multicolumn{5}{l}{\textit{Data} }\\ \hline
+    %Mean &5.6 &4.0 &5.7 & 4.0\\
+    %Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline
     \# obs. & 29279 & 28207 & 29317 & 27680 \\
@@
 \begin{frame}
-\frametitle{Excluding the composition effects of transport costs changes}
+\frametitle{Excluding the composition effects}
 \begin{itemize}
 \item For the ad-valorem component, we estimate the following equation:
-\scriptsize
+\footnotesize
 \begin{eqnarray}
 %\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\simeq \ln(\tau_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{+ \simeq \ln(\tau_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
 \item For the additive component:
-\scriptsize
+\footnotesize
 \begin{eqnarray}
 %\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\simeq \ln(t_i)}+\underbrace{\prod_{k\neq \text{011}}\beta_k.\mathbb{1}_k}_{+ \simeq t_k}\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\prod_{k}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ijt} \label{eq:compeffects_add}
 \end{eqnarray}
@@
 \item Underlying rationale
+\footnotesize
 \begin{itemize}
 \item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
-\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode) \vspace{0.1cm}
+\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
+    \normalsize
 \end{itemize}
@@
 \begin{itemize}
-\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
-\begin{equation*}
-\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
+\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t > 1974$):
+\begin{eqnarray*}
+&&\Gamma_t = 100\frac {\bar{\tau}_{t}-1} {\bar{\tau}_{1974}-1} \\
+\Leftrightarrow && \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{eqnarray*}
 \begin{itemize}
-\item[$\ast$] with $\bar{\tau}_{1974} = ???$
+\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
 \end{itemize}
-\item[-] For the additive cost, we build the variable ($\forall~t \geq 1974$)
+\item[-] For the additive cost, we build the variable ($\forall~t > 1974$)
 $$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
@@
 \begin{frame}
-\frametitle{Conclusion (1)}
+\frametitle{Conclusion: Main findings}
 \textbf{Our paper: Empirical evidence about the role of the additive component in international transport costs} \vspace{0.1cm}
@@
 \begin{frame}
-\frametitle{Conclusion (2)}
+\frametitle{Conclusion : What to do next?}
 
@@
 \appendix
-
-\begin{frame}[label=app_method]
-\frametitle{More on the estimation method}
-\begin{itemize}
-\item The non-linear least squares method \vspace{0.1cm}
-\begin{itemize}
-\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
-\end{itemize}
-\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
-
-\end{itemize}
-\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
-\end{frame}
 

### 2016-09-23 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
Changements slides LEDa + pgm stats des

@@
     %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-    %   \multicolumn{5}{l}{\textit{Data} }\\ \hline
-    %Mean &5.6 &4.0 &5.7 & 4.0\\
-    %Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline
+       \multicolumn{5}{l}{\textit{Data} }\\ \hline
+    Mean &5.3 &5.0 & 5.6&3.9 \\
+    Median  &4.3 & 2.0 &4.4 & 1.9 \\  \hline
     \# obs. & 29279 & 28207 & 29317 & 27680 \\
@@
 %\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
 %\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\prod_{k}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ijt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ijt})&=&\ln\left(\delta +  \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\sum_{k \neq \text{011}}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ijt} \label{eq:compeffects_add}
 \end{eqnarray}
@@
 
-\begin{frame}
+\begin{frame}[label = slide_compeffects]
 \begin{itemize}
-\item Exclude the composition effects of transport costs changes  \vspace{0.1cm}
+\item Exclude the composition effects of transport costs changes  \hyperlink{app_compeffects}{\beamergotobutton{More details}}\vspace{0.1cm}
 \item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
@@
 \item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t > 1974$):
-\begin{eqnarray*}
-&&\Gamma_t = 100\frac {\bar{\tau}_{t}-1} {\bar{\tau}_{1974}-1} \\
-\Leftrightarrow && \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{eqnarray*}
+\begin{equation*}
+ \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
 \begin{itemize}
@@
 
+\begin{frame}[label=app_compeffects]
+\frametitle{Excluding the composition effects: More details}
+\begin{itemize}
+\item \textbf{For the ad-valorem cost}, rewriting Equation (\ref{eq:compeffects_mult}) :
+\footnotesize
+\begin{equation*}
+\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
+\end{equation*}
+\normalsize
+\begin{itemize}
+\item[-] From which we deduce after estimation: $\widehat{\tau}_{ik74} = \exp(\delta +\alpha_i +\beta_k)$ \vspace{0.1cm}
+\item[-] And, for any year $t > 1974$: $\widehat{\tau}_{ikt} = \exp(\delta +\alpha_i +\beta_k)\times \exp(\gamma_t)$ \vspace{0.1cm}
+\item[-] Which implies:
+\begin{equation*}
+\widehat{\tau}_{ikt} = \widehat{\tau}_{ik74}\times \exp(\gamma_t)
+\end{equation*}
+\end{itemize}
+\item With $\tau>1$, rewrite things to get the percentage change between years 1974 and $t$:
+\footnotesize
+$$\Gamma_{ikt} = 100.\frac{\widehat{\tau}_{ikt}-1}{\widehat{\tau}_{ik74}-1} = 100.\frac{\widehat{\tau}_{ik74}\exp(\gamma_t)-1}{\widehat{\tau}_{ik74}-1}$$
+\normalsize
+\begin{itemize}
+\item[-] With $\Gamma_{ikt}$ an index of transport costs in $t$ (relative to 1974) \vspace{0.1cm}
+\item[-] Only depend on the cost observed in 1974 and the time trend \vspace{0.1cm}
+\item[-] But, specific to a product-origin country pair  \vspace{0.1cm}
+
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item Choose to pick as reference the mean value of the cost in 1974
+\begin{itemize}
+\item[-] Build the index $\Gamma_t$ such that:
+\footnotesize
+\begin{equation*}
+ \Gamma_t= 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
+\item[-] With $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974 \vspace{0.2cm}
+\end{itemize}
+\item \textbf{For the additive component}, after estimating Equation (\ref{eq:compeffects_add}):
+\footnotesize
+\begin{eqnarray*}
+\widehat{t}_{ik74}&= & \delta + \alpha_i+ \beta_k \\
+\widehat{t}_{ikt}&=&\left(\delta + \alpha_i+ \beta_k\right).\exp(\gamma_t),\qquad \forall t > 1974
+\end{eqnarray*}
+\normalsize
+
+\begin{itemize}
+\item[-] From which we deduce: 
+$$\widehat{t}_{ikt} = \widehat{t}_{ik74} \times \exp(\gamma_t)$$
+\item[-] With $t >0$, obtain the percentage change from 1974 through:
+\footnotesize
+\begin{equation*}
+\Gamma^{add}_{ikt} = 100\frac{\widehat{t}_{ikt}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)
+\end{equation*}
+\normalsize
+\item Independent of the $i,k$ dimension, such that we can rewrite:
+$$\Gamma^{add}  = 100\exp(\gamma_t) $$
+\end{itemize}
+\end{itemize}
+\hyperlink{slide_compeffects}{\beamergotobutton{Back to slide}}
+\end{frame}
+
 \end{document}

### 2016-10-05 redaction/dhp_sept2016.tex
modif_test_JH

@@
 Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
-SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France } }
+SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - Sciences et Technologies, LEM-CNRS UMR 9221, France \& CEPII, France}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France } }
 

### 2016-10-11 redaction/dhp_sept2016.tex
Graphique composition et correction commentaire

@@
 \includecomment{commentGD}
-\usepackage[draft]{todonotes}
-%%GDPour cacher les notes, \usepackage[disable]{todonotes}
+%\usepackage[draft]{todonotes}
+\usepackage[disable]{todonotes}
 
@@
 Figure \ref{fig:totalTC_compeffects_excl} reports the results.
-Panel (a) displays the time changes in transport costs when only iceberg costs are modeled.
-In Panel (b), we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode.\footnote{For each mode, and by year, we build the total value of transport costs as the weighted sum of both the ad valorem index and the additive index, taking as weighting scheme the relative share of each type of costs in the overall total cost in 1974 (using the figures at the root of Figure \ref{fig:mult_alone_withadd}), i.e.
-respectively 48.2 \% for additive costs in Vessel and 42.3\% for Air in 1974.}
 
@@
 \begin{tabular}{cc}
-{\small (a) Iceberg alone } & {\small (b) With additive costs}\\
-\includegraphics[width=3in, height=2.5in]{Fig3a_TC_overtime_comp_effects_excl.pdf}
-& \includegraphics[width=3in,height=2.5in]{TC_addplusmult_compeffects_excl.pdf} \\
+ \includegraphics[width=6in]{graph_composition_all.eps} \\
 \end{tabular}
@@
 
-Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
+[To be commented]
+
+%Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
 First, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
@@
 
-Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased, to reach a reduction of 40\% over the period 1984-2011, in accordance with the literature.
+%Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased, to reach a reduction of 40\% over the period 1984-2011, in accordance with the literature.
 Further, if we only assume ad-valorem costs (Panel (a)), we can infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005.
@@
 
-\begin{figure}[htbp]
-\caption{Decomposing transport costs over time,
-composition effects excluded}
-\label{fig:TC_add_mult_compeffects_excl}
-\begin{center}
-\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
+%\begin{figure}[htbp]
+%\caption{Decomposing transport costs over time,composition effects excluded}
+%\label{fig:TC_add_mult_compeffects_excl}
+%\begin{center}
+%\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
 %{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
+%\end{center}
+%\end{figure}
 
-Again, the decomposition between both components of transport costs brings valuable insights.
+%Again, the decomposition between both components of transport costs brings valuable insights.
 For ocean shipping, both additive and iceberg components exhibit a downward pattern since 1985.
@@
 
-Our analysis also brings valuable results to the findings of \citet{hummels2007}, as we extend his time coverage (end year 2004) by one decade (to stop in 2013).
+%Our analysis also brings valuable results to the findings of \citet{hummels2007}, as we extend his time coverage (end year 2004) by one decade (to stop in 2013).
 On this decade, our analysis reveals substantial changes in the time trend of transport costs in the air sector.

### 2016-10-19 redaction/Presentations/SeminaireIrvine/slides_Irvine.tex
Nouveau dossier pour Irvine

--- 2016-09-23 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
@@
 \end{itemize}
-\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
+\item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] A reliable database, already used by Hummels (2007), but on a larger period of time \vspace{0.1cm}
-\item[-] Have both the import and the export prices: Estimate both the levels of the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
+\item[-] Have both the import and the export prices: Estimate the levels of both the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
 \end{itemize}
@@
 $$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
-\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
+\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
 \begin{itemize}
-\item[-] The equation is also year- and mode (air or vessel)- specific
+\item[-] The equation is also mode (air or vessel)- specific
 \end{itemize}
@@
 \end{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
 \item[$\Rightarrow$] The equation of interest becomes:
@@
 \begin{itemize}
-\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
+\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
 \end{itemize}
@@
 \item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
-\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
+\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution 
 \end{itemize}
@@
 \begin{itemize}
-\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Result 3}}
+\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Main findings}}
 \begin{figure}[htbp]
@@
 \item Independent of the $i,k$ dimension, such that we can rewrite:
-$$\Gamma^{add}  = 100\exp(\gamma_t) $$
+$$\Gamma^{add}_t  = 100\exp(\gamma_t) $$
 \end{itemize}

### 2016-10-19 redaction/Presentations/SeminaireIrvine/slides_Irvine.tex
Début Irvine again

@@
 \begin{itemize}
-\item Trade costs: A central role in international economic analysis \vspace{0.1cm}
+\item Trade costs have a central role in international economic analysis \vspace{0.1cm}
 \begin{itemize}
@@
 \end{itemize}
-\item Transport costs: A sizeable share of international trade costs \vspace{0.1cm}
+\item Transport costs are a sizeable share of international trade costs \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Some trade policy instruments as well (quota license price)
+\footnotesize
+$$pq = (t+\widetilde{p})q  $$
 \end{itemize}
@@
 \begin{itemize}
-\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.2cm}
+\item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.2cm}
 \begin{itemize}
+\item[-] Many models rely on the tractability of iceberg costs \vspace{0.1cm}
 \item[-] Additive trade costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
-\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
+\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization because they distort relative prices (Sorensen, 2014)\vspace{0.1cm}
 \item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
@@
 \end{itemize}
-\textbf{One objective of the paper: Provide an answer to this question}
+\textbf{One objective of the paper is to provide an answer to this question}
 
@@
 %\footnotesize{
-\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
+\item[-] The estimated iceberg component is reduced by a factor linked to the importance of additive costs (which is kind of obvious) \vspace{0.1cm}
 \item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
@@
 \item[-] Only cover international transport costs \vspace{0.1cm}
-\item[-] Among transport costs, quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
+\item[-] Among transport costs, insurance + handling + quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
 %\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
@@
 \begin{tabular}{cc}
-{\scriptsize (a) Iceberg alone } & {\scriptsize  (b) With additive }\\
-\includegraphics[width=5cm, height=5cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
-& \includegraphics[width=5cm,height=5cm]{TC_addplusmult_compeffects_excl.pdf} \\
+\\
+ \includegraphics[width=10cm,height=7.5cm]
+{graph_composition_all-eps-converted-to.pdf} \\
 \end{tabular}

### 2016-11-17 redaction/dhp_sept2016.tex
changements dans le texte

--- 2016-08-26 redaction/dhp_sept2016.tex
@@
 \paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
+To do so, we start from the equation that expresses the price $p(z)$ of a good $z$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}(z)$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
 \begin{equation}
-p = \tau \widetilde{p} + t \label{eq:base}
+p(z) = \tau \widetilde{p}(z) + t \label{eq:base}
 \end{equation}
@@
 \noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
-Let us denote $i$ the origin country, and $k$, the sector (or product).
+Let us denote $i$ the origin country, and $k$, the sector.
 Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ikz}}{\widetilde{p}_{ikz}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ikz}} \label{eq:base_estimee}
 \end{equation*}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
+\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ikz}} \label{eq:theory_equation}
 \end{equation*}
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
+The ratio $\frac{p_{ikz}}{\widetilde{p}_{ikz}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ikz}>\widetilde{p}_{ikz}$).
 Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
+\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}} \right)\times \exp(\epsilon_{ikz})
 \end{equation*}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
+\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}}-1 \right) + \epsilon_{ikz} \label{eq:estimatedequation}
 \end{equation}
@@
 The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
-The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.\todo{Dire plutot: sum of squares of residuals; to increase rather than to decrease?
-Not clear to me.} Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
+The intuitive criterion for convergence is that the sum of squares does not increase from one iteration to the next.
+Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
 These cut-offs are aimed at eliminating reporting or coding errors.\medskip
@@
 \begin{eqnarray}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add} \\
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
+\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}}\right) + \epsilon_{ikz} \label{eq:model_with_add} \\
+\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ikz} \label{eq:model_nlI}
 \end{eqnarray}
@@
     Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+        \multicolumn{5}{l}{Data } \\ \hline
+        Mean & 1.053 & 1.05& 1.056&1.039 \\
+        Median & 1.043 & 1.02 & 1.044& 1.019 \\ \hline
         \# obs. & 29279 & 28207 & 29317 & 27680 \\
@@
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
-In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-
+In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
 We investigate the robustness of this conclusion on a yearly basis.
@@
 From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds on a yearly basis.
-For Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
+Further, Figure \ref{fig:mult_alone_withadd} reports that for Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
 Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air transport.
@@
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
-We come back to this aspect in Section \ref{sec:results_trends}.}
+We come back to this aspect in Section \ref{sec:results_trends}.}\todo{Comparer avec la litterature.
+Cf AvW 2004 pure transport costs 11\% markup over production costs.}
 
@@
 In a non-linear context, $R^2$ is  strictly speaking inappropriate.
-However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.} As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
+However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.} This drives us to complement the goodness of fit diagnosis with three alternative measures.
+We provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
 Smaller values are better because it indicates that the observations are closer to the fitted line.
@@
 A decrease in the log-likelihood function points to a better quality-of-fit.
-However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic, so that AIC stat $= 2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $.
-Once again, the preferred model is the one with the minimum AIC value.
+However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $.} The preferred model is the one with the minimum AIC value.
 Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e.
@@
 
-Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences.
-First, iceberg costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FAB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes).
-The standard error of the regression based only on the ad-valorem term is consistently lower for Vessel than for Air.
-
-Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000 for both transport modes.
-Interestingly, this seems to be attributable to different reasons depending of the transport mode considered.
-For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term increases the explanatory power by 30\%, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER.
-Accordingly, the lower quality of fit for air transport seems to be attributable to a loss of explanatory power of the additive component.
-The picture is quite different for Vessel.
-In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included.
-In contrast to air transport, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term.
-This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs.
-We henceforth go deeper into this time-trend analysis in the next section.\smallskip
+\smallskip
 
@@
 \widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(\tau_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
+\Leftrightarrow \ln(\widehat{\tau}_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
+\widehat{t}_{ikt}&=&\left( \sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\sum_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\sum_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
 \end{eqnarray}
@@
 \end{figure}
+
+\textbf{Faire l'analyse mais sur l'ecart cif-fob observe et non plus reconstruit. Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
 
@@
 These results suggest that important changes in the structure of costs ate at work in the air transport in the very recent period, that should be investigated in further research.
+
+\section{Robustness checks}
+
+\subsection{Robustness to the specification of the estimated equation}
+
+\paragraph{Estimating the difference between fas and cif prices in level}
+
+Ie, estimate
+
+$$\log(p_{ikz})= \log(\tau_i\times \tau_k \times \widetilde{p}_{ikz}+ t_i+ t_k)+ \exp(\epsilon_{ikz})$$
+
+\paragraph{Robustness to the separability assumption}
+
+Ie, estimate
+
+$$\log(\frac{p_{ikz}}{\widetilde{p}_{ikz}} -1)= \log(\tau_{ik} -1+ \frac{t_{ik}}{\widetilde{p}_{ikz}})+ \exp(\epsilon_{ikz})$$
 
@@
 
-\section{Robustness checks \label{app:robust_4d}}
+\subsection{Goodness-of-fit: Detailed results \label{app:robust_4d}}
 

### 2016-11-17 redaction/Presentations/SeminaireIrvine/slides_Irvine.tex

@@
 \item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
-\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
+\item[$\Leftrightarrow$] Part of the ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization because they distort relative prices (Sorensen, 2014)\vspace{0.1cm}
-\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
+\end{itemize}
+\item Empiric studies
+\begin{itemize}
+\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2015, Hummels \& Skiba, 2004)  \vspace{0.1cm}
+\item[-] Contrarian view  Lashkaripour (2016) taking into account the fact that more expensive goods are heavier and hence more expensive to transport. Yet  \vspace{0.1cm}
+\begin{itemize}
+\item[-] This can only apply to goods that are enumerated by items (60\% of US trade)
+\item[-] Reasonable hypothesis for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour (2016))
+\end{itemize}
 \end{itemize}
@@
 \item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
-\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
+\item[$\Rightarrow$] We lose much. With the additive term included:\vspace{0.1cm}
 \begin{itemize}
@@
 \item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.2cm}
-\begin{itemize}
+%\begin{itemize}
 %\footnotesize{
-\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
+%\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
 %}
-\end{itemize}
+%\end{itemize}
 \item[$\Rightarrow$] The importance of including the additive component \vspace{0.2cm}
@@
 %\footnotesize{
-\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.2cm}
-\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
+\item[-] Composition effect is especially strong for the additive costs \vspace{0.2cm}
+\item[-] In the air case, the composition effec was not the same on additive and multiplicative costs
 %\end{itemize}
@@
 \item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
-\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
+\item[-] Yearly basis, from 1974 to 2013, SITC 5-digit classification level, by transport mode (air or vessel) (Could go up to HS-10 with recent data)\vspace{0.1cm}
 \end{itemize}
@@
 \begin{itemize}
-\item[-] At the 3-digit classification level  \vspace{0.1cm}
+\item[-] The RHS are at the 3-digit classification level  \vspace{0.1cm}
 \item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
-\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
+\item[-] Approximatively 200 sectors (3 digits), from around 200 countries  \vspace{0.1cm}
 \begin{itemize}
-\item[$\ast$] Around 600-700 products at the 4-digit level
+\item[$\ast$] Around 600-700 sectors at the 4-digit level
 \end{itemize}
@@
 \item[-] Only cover international transport costs \vspace{0.1cm}
-\item[-] Among transport costs, insurance + handling + quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
+\item[-] Among transport costs, insurance + handling + quantitative freight costs (e.g. not those related to the time value of goods) \vspace{0.1cm}
 %\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
@@
 \begin{itemize}
-\item[-] US imports, large time period: Broad view of international trade flows \vspace{0.1cm}
-\item[-] A reliable database, already used by Hummels (2007), but on a larger period of time \vspace{0.1cm}
-\item[-] Have both the import and the export prices: Estimate the levels of both the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
+\item[-] US imports, long time period: Broad view of international trade flows \vspace{0.1cm}
+\item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
+\item[-] We use both the import and the export prices.
+That allows the estimate the levels of both the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015, that uses one year of Norwegian data)
 \end{itemize}
@@
 $$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
-\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
+\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
-\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
+\item Additional assumption : all products in a 3-digit sector (s) share the same structure of costs \vspace{0.1cm}
+\item[$\Leftrightarrow$] Write $t_{ik^s}$ and $\tau_{ik^s}$ as:
 \begin{equation}
- \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
+ \tau_{ik^s} = \tau_{i} \times \tau_{s}, \qquad t_{ik} = t_{i} + t_{s} \label{eq:specifTC}
  \end{equation}
@@
 \footnotesize
-$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
+$$\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 =\left(\tau_{i} \times \tau_{s} -1+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}} \right)\times \exp(\epsilon_{ik^s})$$
 \normalsize
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
+\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}}-1 \right) + \epsilon_{ik^s} \label{eq:est_equation}
 \end{equation}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
+\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}}-1 \right) + \epsilon_{ik^s}
 \end{equation}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
+\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}-1 \right) + \epsilon^{ice}_{ik^s} \label{eq:model_nlI}
 \end{equation}
@@
 \item[-] With additive costs:
-$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
+$$\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}, \qquad \widehat{t}^{add}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$$
 \item[-] With only iceberg costs:
-$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
+$$\widehat{\tau}^{ice}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$$
 \end{itemize}
-\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
+\item  Taking the weighted average over the sector-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
 \begin{itemize}
@@
 \item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
+\item[$\ast$] Our data are only for the US, which presumably has lower trade costs \vspace{0.1cm}
+
 \end{itemize}
@@
 \begin{itemize}
-\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
+\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) (which is kind of obvios) \vspace{0.2cm}
 \begin{itemize}
@@
 %\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ist})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{s\neq \text{011}}\beta_s.\mathbb{1}_s }_{\ln(\widetilde{\tau}_s)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ist} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
 \begin{itemize}
-\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
+\item[-] With $\widehat{\tau}_{ist} = \widehat{\tau}^{ice}_{ist}, \widehat{\tau}^{adv}_{ist}$ previously obtained  \vspace{0.1cm}
 \end{itemize}
@@
 %\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{t}_{ijt})&=&\ln\left(\delta +  \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\sum_{k \neq \text{011}}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ijt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ist})&=&\ln\left(\delta +  \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\sum_{s \neq \text{011}}\beta_s.\mathbb{1}_s}_{\widetilde{t}_s}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ist} \label{eq:compeffects_add}
 \end{eqnarray}
@@
 \begin{frame}[label=slide_comment_compositioneffects]
-\frametitle{Two main findings}
+\frametitle{What do we find}
 \begin{itemize}
-\item The importance of excluding composition effects  \vspace{0.1cm}
+\item At the aggregate level  \vspace{0.1cm}
 \begin{itemize}
-\item[-] The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
-\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}) is attributable to change in the composition of trade patterns \vspace{0.1cm}
-\item[-] Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
+\item[-] Air transport costs were reduced by 50\% and vessel transport costs by 60\%. (as already seen) \vspace{0.1cm}
+\item[-] In the case of vessel, a large part of that is a composition effect (reduction in the share of bulky goods because of the relative decline of primary goods) \vspace{0.1cm}
+\item[-] The reverse happened for air transport (increase in the shipping of bulky goods) \vspace{0.1cm}
 \end{itemize}
-\item The importance of the additive component of transport costs (again) \vspace{0.1cm}
+\item Additive and multiplicative costs \vspace{0.1cm}
 \begin{itemize}
-\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in Air transport over the 1985-2005 period \vspace{0.1cm}
-\item[-] In accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
-\item[-] But... No substantial difference when additive costs are included (Panel (b))
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Decomposing \textit{pure} transport costs over time }
-\begin{figure}[htbp]
-%\caption{Decomposing transport costs over time,
-%composition effects excluded}
-%\label{fig:TC_add_mult_compeffects_excl}
-\begin{center}
-\includegraphics[width=8cm, height=3.5cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-\begin{itemize}
-\item For Vessel: Similar downward trend for both $\tau$ and $t$  \vspace{0.1cm}
-\item For Air: Much more contrasted trends  \vspace{0.1cm}
-\begin{itemize}
-\item[-] Substantial decrease in the ad-valorem costs over 1985-2005, but roughly constant additive costs  \vspace{0.1cm}
-\item[$\Rightarrow$] An explanation to the difference between Panels (a) and (b) above  \vspace{0.1cm}
-
-\item[$\Rightarrow$] The importance of modeling additive costs  \vspace{0.1cm}
-\item[-] Trend reversal around 2005: Any suggestion?
+\item[-] The composition effect is especially strong for the additive costs. Maybe bulkiness is more important for additive costs than for multiplicative costs ? \vspace{0.1cm}
+\item[-] One might want to think of multiplicative as insurance and additive as shipping of handling. Will that help thinking about the composition effect? \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] Importance of the composition effects \vspace{0.1cm}
-\item[-] Biased picture of the time trends of transport costs in air when omitting the additive dimension
 \end{itemize}

### 2016-11-25 redaction/dhp_sept2016.tex
new ref
limao venables

@@
 The structure (additive vs ad-valorem) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows.
-The Alchian and Allen conjecture (\citealp{alchian}), points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
+The Alchian and Allen conjecture (\citealp{alchian}), that points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
 \citet{Lashkaripour-mimeo-2016} challenges this view.
@@
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fas unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
-\citet{hummels_skiba} estimate the elasticity of freight rates with respect to price is estimated to be well below unity.
+\citet{hummels_skiba} estimate the elasticity of freight rates with respect to price to be well below unity.
 Also, their estimates imply that doubling freight costs increases average fas export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
@@
 While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
-One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap.
+One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to fill this gap.
 \bigskip
@@
 We quantitatively assess the size and the importance of the additive component in international transport costs, that we measure as the difference between the import and the export prices, based on the US Imports flows over 1974-2013.
+Therefore, we focus only on a subset of trade costs.
+Besides, in terms of transport costs \textit{per se}, our measure based on the gap between export and import prices omits the other dimension of transport costs related to the time value of goods in transit.
+According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
+However, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: they remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
+In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
+Furthermore, our data is based on a single and reliable customs origin, and allows for a wide time and geographical coverage, while limiting the measurement errors coming from strategies inferring trade costs.
+Therefore, we strongly believe they may deliver useful and reliable insights on the structure of international trade costs.
+
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring additive trade costs from firm-level trade data.
@@
 Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
+This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
 
@@
 This is counter-intuitive, but certainly linked to selection effects.
+
+%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
 Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
@@
 \begin{eqnarray}
-\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(\widehat{\tau}_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
+\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
+\Leftrightarrow \ln(\widehat{\tau}_{ikt})&=&\delta +\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}

### 2016-12-12 redaction/dhp_sept2016.tex
readction121216

@@
 However, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: they remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
+\citet{Hummels_1999} mentions a bunch of papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
 In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
@@
 We use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level.
-As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.} Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.
+As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.\footnote{Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.
+The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
+Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
+We thus obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
+Estimation results at the 4-digit classification level are reported in Appendix ??.} Depending on the considered year, this leaves us with around 200 sectors sectors, from approximately 200 countries of origin.
 
@@
 \paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p(z)$ of a good $z$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}(z)$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
+To do so, we start from the equation that expresses the price $p$ of a good $k$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
 \begin{equation}
-p(z) = \tau \widetilde{p}(z) + t \label{eq:base}
+p = \tau \widetilde{p}+ t \label{eq:base}
 \end{equation}
@@
 \noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
-Let us denote $i$ the origin country, and $k$, the sector.
+Let us denote $i$ the origin country, and $k$, the product.
 Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}
@@
 \begin{equation*}
-\frac{p_{ikz}}{\widetilde{p}_{ikz}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ikz}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ikz}} \label{eq:base_estimee}
 \end{equation*}
-
-We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.
 
@@
 \begin{equation*}
-\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ikz}} \label{eq:theory_equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ikz}} \label{eq:theory_equation}
 \end{equation*}
 
-The ratio $\frac{p_{ikz}}{\widetilde{p}_{ikz}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ikz}>\widetilde{p}_{ikz}$).
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
 Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:
@@
 \begin{equation*}
-\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}} \right)\times \exp(\epsilon_{ikz})
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
 \end{equation*}
-\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Taking the logs, this drives us to estimate the following equation:
+\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Considered in logs, the above equation becomes:
 
 \begin{equation}
-\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}}-1 \right) + \epsilon_{ikz} \label{eq:estimatedequation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
 \end{equation}
-where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= \ln(\varepsilon_{ik})$.
 
-The non-linearity of Equation (\ref{eq:estimatedequation}) implies that it cannot be estimated using standard linear estimators.
-Therefore, all estimates are performed using non-linear least squares.
-The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
-The intuitive criterion for convergence is that the sum of squares does not increase from one iteration to the next.
-Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
-These cut-offs are aimed at eliminating reporting or coding errors.\medskip
+The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators.
+All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
+The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
+See ???
+for more details on the non-linear estimation method.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+Confronted to this arbitrage, we retain the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
+This drives us to estimate a modified version of Equation (\ref{eq:equation0}), according to:
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
+\end{equation}
+where $\tau_{i}$, $\tau_{s}$, $t_{i}$ and $t_{s}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.
+To eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
+These cut-offs are aimed at eliminating reporting or coding errors.
+We estimate this equation for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$).\smallskip
 
-Further, one key objective of the paper is to characterize the importance of additive costs relatively to ad-valorem costs.
-Put differently, what traditional models of international trade lose by ignoring additive costs?
-A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining $t$ to be equal to zero, and compare the fitting properties and the explanatory power of the restricted and complete models.
-Accordingly, for each year and transport mode, we estimate two equations, depending on additive transport costs being included (Equation (\ref{eq:model_with_add})) or not (Equation (\ref{eq:model_nlI})):
+One key objective of the paper is to characterize the importance of additive costs relatively to ad-valorem costs.
+Put it differently, what traditional models of international trade lose by ignoring additive costs?
+A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining either additive costs $t$ or ad-valorem costs $\tau$ to be equal to zero, and compare the fitting properties and the explanatory power of the models.
+Accordingly, for each year and transport mode, we estimate the three models: \textit{(a)} when transport costs are modeled as iceberg costs (Equation (\ref{eq:model_nlI})), \textit{(b)} when transport costs are modeled only as additive costs (Equation (\ref{eq:model_nlA})), and \textit{(c)} when transport costs are modeled as decomposed in the two dimensions of additive and iceberg (Equation (\ref{eq:model_IetA}))
 
 \begin{eqnarray}
-\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}}\right) + \epsilon_{ikz} \label{eq:model_with_add} \\
-\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ikz} \label{eq:model_nlI}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI} \\
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA} 
 \end{eqnarray}
 
-After estimating Equation (\ref{eq:model_nlI}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$ and $\widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$, that is country-product specific, by year and transport mode.
-
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost $\widehat{\tau}^{ice}$, $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
+After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$ and $\widehat{t}^{add}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$, that is country-product specific, by year and transport mode.
+Under Model (a), when assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}, while under Model (b), we re-build $\widehat{t}^{nlA}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$.
+Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{nlI}$, \textit{b)} $\widehat{t}^{nlA}$, \textit{c)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 
+% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
+More precisely, for this specific year, we find...
+something slightly higher...
+
 %How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
@@
 In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
-We investigate the robustness of this conclusion on a yearly basis.
+
+We extend further our analysis by using explicitly the time dimension of our analysis in order to assess the dynamics of our estimates over time.
 Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_with_add}), by transport mode.

### 2016-12-12 redaction/dhp_sept2016.tex
Obtention résultats 3 modèles + nouvelle version tex

@@
 However, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: they remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
-\citet{Hummels_1999} mentions a bunch of papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
 In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
@@
 
-% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
-More precisely, for this specific year, we find...
-something slightly higher...
-
 %How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
@@
 In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
-
-We extend further our analysis by using explicitly the time dimension of our analysis in order to assess the dynamics of our estimates over time.
+We investigate the robustness of this conclusion on a yearly basis.
 Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_with_add}), by transport mode.

### 2016-12-13 redaction/dhp_sept2016.tex
modification tex
footnote sur 3e modèle avec que des additifs; footnote sur robustesse à
décomposition en somme ou produit des termes i et k

--- 2016-12-12 redaction/dhp_sept2016.tex
@@
 
-\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:
+\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
 
@@
 The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
-See ???
-for more details on the non-linear estimation method.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
 Confronted to this arbitrage, we retain the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
@@
 Put it differently, what traditional models of international trade lose by ignoring additive costs?
-A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining either additive costs $t$ or ad-valorem costs $\tau$ to be equal to zero, and compare the fitting properties and the explanatory power of the models.
-Accordingly, for each year and transport mode, we estimate the three models: \textit{(a)} when transport costs are modeled as iceberg costs (Equation (\ref{eq:model_nlI})), \textit{(b)} when transport costs are modeled only as additive costs (Equation (\ref{eq:model_nlA})), and \textit{(c)} when transport costs are modeled as decomposed in the two dimensions of additive and iceberg (Equation (\ref{eq:model_IetA}))
+A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining additive costs $t$ to be equal to zero, and compare the fitting properties and the explanatory power of the models.
+For each year and transport mode, we thus estimate the two models: \textit{(a)} when transport costs are modeled as iceberg costs (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are modeled as decomposed in the two dimensions of additive and iceberg (Equation (\ref{eq:model_IetA})):
 
@@
 \ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI} \\
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
+%\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
 \ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA}
@@
 Under Model (a), when assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}, while under Model (b), we re-build $\widehat{t}^{nlA}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$.
-Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{nlI}$, \textit{b)} $\widehat{t}^{nlA}$, \textit{c)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
+This has driven us to estimate this case as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik}$.
+The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (which is itself dominated by the complete model (b), anticipating on further results.
+For sake of space saving, the results of this model are not reported in the paper.
+Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{nlI}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 The objective of this section is twofold.
-First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (Equation (\ref{eq:model_with_add}) or (\ref{eq:model_nlI})).
+First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
 Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.
@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs }  \\ \hline
-    Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
-    Median & 1.051 & 1.042 & 1.052 & 1.037 \\
-    Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
-    Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
-    Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
-   \textit{Ad-valorem term} & & & & \\ \hline
-    Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
-    Median & 1.028 & 1.018 & 1.028 & 1.016 \\
-    Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
-    Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
-    Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term }& & & &   \\ \hline
-    Mean  & 0.029 & 0.018 & 0.028 & 0.019 \\
-    Median & 0.019 & 0.007 & 0.017 & 0.008 \\
-    Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
-    Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
-    Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-        \multicolumn{5}{l}{Data } \\ \hline
-        Mean & 1.053 & 1.05& 1.056&1.039 \\
-        Median & 1.043 & 1.02 & 1.044& 1.019 \\ \hline
+    \multicolumn{5}{l}{\textbf{With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
+    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
+    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
+    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
+    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
+    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
+    \multicolumn{5}{l}{\textbf{With Additive \& Ad-Valorem Trade Costs} } \\ \hline
+   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
+    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
+    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
+    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
+    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
+    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
+    \textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}& & & &   \\ \hline
+    Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
+    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
+    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
+    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
+    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+    \multicolumn{5}{l}{\textbf{Data} ($p/\widetilde{p}$, in \%) } \\ \hline
+        Mean & 5.3 & 5.0& 5.6&3.9 \\
+        Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
         \# obs. & 29279 & 28207 & 29317 & 27680 \\
@@
     \end{center}}
-\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.
+\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
 The additive term is expressed in fraction of fas price.
 ($^\ast$): Four 4-digit estimation: 0n selected years.
-($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
+($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
 \end{table}%
@@
 %How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
+
 Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
@@
 We extend further our analysis by using explicitly the time dimension of our analysis in order to assess the dynamics of our estimates over time.
-Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_with_add}), by transport mode.
+Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_IetA}), by transport mode.
 
@@
 Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e.
-Equation (\ref{eq:model_with_add}) and the Restricted one (\emph{R}, i.e.
+Equation (\ref{eq:model_IetA}) and the Restricted one (\emph{R}, i.e.
 Equation (\ref{eq:model_nlI})).
@@
 
-\paragraph{Estimating the difference between fas and cif prices in level}
-
-Ie, estimate
-
-$$\log(p_{ikz})= \log(\tau_i\times \tau_k \times \widetilde{p}_{ikz}+ t_i+ t_k)+ \exp(\epsilon_{ikz})$$
-
 \paragraph{Robustness to the separability assumption}
@@
 
-In this section, we report the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_with_add})) or not (Equation (\ref{eq:model_nlI})).
+In this section, we report the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})).
 For sake of space saving, we only report the results for a sample of years.
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation.
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation (again, for selected years).
 

### 2016-12-19 redaction/dhp_sept2016.tex

@@
 
+\todo{reprendre slides presentation LEDa pour la motivation de la question.}
+
 Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis\footnote{Rigourously speaking, the ``iceberg cost'' includes the ad-valorem costs and the fact that the costs are paid in terms of the good that is traded.
@@
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.
-Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
+A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{\textbf{donner l'intuition.
+Deformation des prix relatifs, TBC}} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
 While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
@@
 These cut-offs are aimed at eliminating reporting or coding errors.
-We estimate this equation for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$).\smallskip
+We estimate this equation for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$).
+Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode (at the 3-digit level).
+\smallskip
 
@@
 These results are reported in Section \ref{sec:results_decomposition}.
+
+\todo{A un moment, parler du risque d'endogeneite entre tauik tik et ptildeik, dire que pas de pb. Dire ou?}
 
@@
 Before making any definite statement about this though, it is worth emphasizing that the evolution of transport costs depends both of the evolution of per product and per partner costs and of the evolution of the composition of trade.
-Total transport costs may have decreased over time either the country either trades with countries with which trade is cheaper, or trades products which are cheaper to transport, independently of any change in tranport costs \textit{per se}.
+Total transport costs may have decreased over time either the country either trades with countries with which trade is cheaper, or trades products which are cheaper to transport, independently of any change in transport costs \textit{per se}.
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
@@
 \begin{eqnarray}
-\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(\widehat{\tau}_{ikt})&=&\delta +\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s\neq \text{011}}\beta_s.\mathbb{1}_s}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
 
-As for the additive cost, we estimate the following equation, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
+Applying a similar reasoning for the additive cost, we estimate the following equation, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
 \begin{eqnarray}
-\widehat{t}_{ikt}&=&\left( \sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\sum_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\sum_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
+\ln(t_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s}\beta_s.\mathbb{1}_s}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
 \end{eqnarray}
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
-%%In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``pure transport costs time trend'', that we identify through Term (c).
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}
 Equation (\ref{eq:compeffects_mult}) is estimated using OLS and Equation (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).
@@
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{ice}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:
+From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{ice}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
 \begin{equation*}
@@
 
-As for the additive cost, we built the variable $\Gamma^{add}_t = 100.\exp(\gamma_t)$, with the reference year being 1974 in all cases (ie, with $\gamma_{1974}=0$).
-As a result, the series for $\Gamma_t$ (with an initial value of 100 for $t=1974$) have a straightforward interpretation in percentage changes.
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s$ the mean (iceberg) transport cost in 1974.
+As for the additive cost, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
+$$\Gamma^{add}_t = 100.\exp(\gamma_t)$$
+
+\noindent As a result, the three series (for $\Gamma_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+
 Figure \ref{fig:totalTC_compeffects_excl} reports the results.
@@
 
-\subsection{Robustness to the specification of the estimated equation}
+Robustness to the specification of the estimated equation.
 
@@
 
-$$\log(\frac{p_{ikz}}{\widetilde{p}_{ikz}} -1)= \log(\tau_{ik} -1+ \frac{t_{ik}}{\widetilde{p}_{ikz}})+ \exp(\epsilon_{ikz})$$
+$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
+
+To do on 100 products ($s$), 50 countries (the most important?)
+
+\paragraph{what else?}
 
@@
 
+\section{More on eliminating the composition effects \label{app:comp-effects}}
+
+In this section, we explain with more details the method employed to eliminate the country- and product- dimensions of the estimated transport cost.
+
+\paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component. Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
+
+\begin{equation*}
+\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
+\end{equation*}
+
+Based on this equation, we deduce after estimation that: 
+\begin{eqnarray*}
+\text{For the year 1974}&& \widehat{\tau}_{is74} = \exp(\delta +\alpha_i+\beta_s), \\
+\text{For any year}~t> 1974&& \widehat{\tau}_{ist} = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t)
+\end{eqnarray*}
+
+From this, we obtain the foolowing recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
+Given that $\tau >1$, we can rewrite things to get the percentage change between year 1974 and any year $t>1974$:
+$$\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}$$
+
+As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trade.
+At this stage though, it remains specific to a product-origin country pair.
+Next step is to build the index $\Gamma_t$ such that:
+\begin{equation*}
+ \Gamma_t= 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
+\noindent with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
+
+\paragraph{For the additive component} After estimating Equation (\ref{eq:compeffects_add}), we can re-build the additive component according to:
+\begin{eqnarray*}
+\text{For the year 1974}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
+\text{For any year}~t> 1974&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
+\end{eqnarray*}
+
+From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$. Given the contraint $t>0$, we then obtain the percentage change from 1974 from:
+
+\begin{equation*}
+\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)
+\end{equation*}
+
+Noticing that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
+$$\Gamma^{add}_t  = 100\exp(\gamma_t) $$
+
 \end{document}

### 2017-01-05 redaction/dhp_sept2016.tex
Jerome050117

--- 2016-12-19 redaction/dhp_sept2016.tex
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+It displays the mean and median values of each type of trade costs, as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
+The first panel reports the results for a specification based on ad-valorem costs estimated alone, while the second panel presents estimates for a specification involving both ad-valorem and multiplicative component.
+Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
 Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
@@
 
-Table \ref{tab:summary_results} yields the following comments.
-First, the magnitude of overall transport costs is sizeable.
+Table \ref{tab:summary_results} calls for two types of comments.
+Starting with pure statistical concerns, it appears that mean values are systematically higher than medians - by around one percentage point.
+This is no surprise at all: once again, our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median.
+Besides, one can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point.
+This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects.
+
+Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable.
 Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
-Air transport costs are smaller than sea transport costs.
-This is counter-intuitive, but certainly linked to selection effects.
-
-% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
-More precisely, for this specific year, we find...
-something slightly higher...
-
-%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
-
-Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
+Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
 Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%.
@@
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
-In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
+In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.%\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
 
@@
 
-At this stage, two interesting comments can be made.
+The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996.
+Our own results point to a total decrease from 6.9 to 4.2\% on average for the same years.\footnote{For the sole 1994 year, \citet{Hummels_2001} finds an all-commodities trade-weighted average transport cost of 3.8 percent of the fob.
+price for the United States }\todo{This has to be clarified: What data uses Hummels?
+Does he distinghuish at some point between air and vessel.
+}
+
+XXX but our contribution goes beyond: we distinguish in this trend what comes from additive, and what comes form iceberg XXX
+
 First, the magnitude of transport costs is higher for ocean shipping than for air transport.
@@
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
-We come back to this aspect in Section \ref{sec:results_trends}.}\todo{Comparer avec la litterature.
-Cf AvW 2004 pure transport costs 11\% markup over production costs.}
+We come back to this aspect in Section \ref{sec:results_trends}.}
+
+\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
+
+% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
+More precisely, for this specific year, we find...
+something slightly higher...
+
+%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
 

### 2017-01-05 redaction/dhp_sept2016.tex
Jerome050117add
Sur l’effet de composition

--- 2016-12-19 redaction/dhp_sept2016.tex
@@
 First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974.
-Second, the decrease appears more important for sea than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport.
+Second, the decrease appears more important for ocean than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport.
 That said, it may be the case the oil shock inflated transportation costs that specific year.
-If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for Air, and 2-3 percentage points for Vessel (depending if we consider mean or median estimates).
+If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates).
 These results are consistent with \citet{hummels2007}.
 
-Before making any definite statement about this though, it is worth emphasizing that the evolution of transport costs depends both of the evolution of per product and per partner costs and of the evolution of the composition of trade.
-Total transport costs may have decreased over time either the country either trades with countries with which trade is cheaper, or trades products which are cheaper to transport, independently of any change in transport costs \textit{per se}.
+Before making any definite statement about this, it is worth emphasizing that the evolution of transport costs depends on the one hand of the evolution of per product and per partner costs and on the other hand of the evolution of the composition of trade.
+Total transport costs may have decreased over time because the share of neighbour countries in total US trade or the share of cheaper to transport goods have increased, independently of any change in transport costs \textit{per se}.
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
-To do this, we estimate the following equation for the ad-valorem cost, based on the series obtained $\widehat{\tau}_{ikt}^{ice}$ and $\widehat{\tau}_{ikt}^{adv}$ that have been previously estimated:
-\footnote{Even if Equation (\ref{eq:compeffects_mult}) is estimated on two different series, it is denoted $\widehat{\tau}_{ikt}$ for sake of notational simplicity.
-For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
+To do this, we estimate the following equations for observed costs and the estimated ad-valorem component using OLS:
+%the ad-valorem cost, based on the series obtained $\widehat{\tau}_{ikt}$ and $$$%\widehat{t}_{ikt}$ that have been previously estimated:
+\footnote{Even if Equation (\ref{eq:compeffects_mult}) is estimated both for observed costs and the estimated ad-valorem componen we write it only for  $\widehat{\tau}_{iskt}$ for sake of notational simplicity.
+For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}
+%$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
 
 \begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s\neq \text{011}}\beta_s.\mathbb{1}_s}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
 
-Applying a similar reasoning for the additive cost, we estimate the following equation, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
+We cannot use the same equation for the estimated additive component, as by construction, the sector fixed effect and the country fixed effect are additive rather than multiplicative.
+We estimate the following equation using non-linear least squares:
+%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
 \begin{eqnarray}
-\ln(t_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s}\beta_s.\mathbb{1}_s}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
 \end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``pure transport costs time trend'', that we identify through Term (c).
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: the country dimension (Term (a)), the product dimension (Term (b)) and the ``pure transport costs time trend'' (Term (c)).
 Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}
-Equation (\ref{eq:compeffects_mult}) is estimated using OLS and Equation (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).
+All equations are estimated using by transport mode.
 
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-From the estimation of Equation (\ref{eq:compeffects_mult}) (on both $\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{ice}$), we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
+From the estimation of Equation (\ref{eq:compeffects_mult}) (on both observed trade costs and estimated multiplicative component
+%$\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{ice}$)
+, we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
 \begin{equation*}
@@
 
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s$ the mean (iceberg) transport cost in 1974.
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean transport cost or the mean multiplicative transport cost in 1974.
+
 As for the additive cost, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
@@
 \begin{figure}[htbp]
-\caption{Transport costs (composition effects excluded)}
+\caption{Transport costs (with and without composition effects)}
 \label{fig:totalTC_compeffects_excl}
 \begin{center}
-\begin{tabular}{cc}
-{\small (a) Iceberg alone } & {\small (b) With additive costs}\\
-\includegraphics[width=3in, height=2.5in]{Fig3a_TC_overtime_comp_effects_excl.pdf}
-& \includegraphics[width=3in,height=2.5in]{TC_addplusmult_compeffects_excl.pdf} \\
-\end{tabular}
+\includegraphics[height=4in]
+{graph_composition_all.pdf}
 \end{center}
@@
 
-\textbf{Faire l'analyse mais sur l'ecart cif-fob observe et non plus reconstruit. Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
+\textbf{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
 
-Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
-First, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
+Figure \ref{fig:totalTC_compeffects_excl} drives the following comments. Air transport costs were reduced by 50\% and vessel transport costs
+by 60\%.
+(as already seen).
+However, in the case of vessel, a large part of that is a composition effect, probably because of the reduction in the share of bulky goods linked to the declining share of primary goods in total trade.
+The reverse happened for air transport : the goods shiped by air were increasingly more costly to transport.
+n line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
+
+(((At the agregate levelFirst, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
 This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that the fall in transport costs immediately started in 1974.
 Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 was due to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
-In line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
+I
 
@@
 As reported in Figure \ref{fig:TC_add_mult_compeffects_excl}, this can be attributed to the iceberg component (while the additive cost has conversely been drastically reduced since 2005).
-These results suggest that important changes in the structure of costs ate at work in the air transport in the very recent period, that should be investigated in further research.
+These results suggest that important changes in the structure of costs ate at work in the air transport in the very recent period, that should be investigated in further research.)))
 

### 2017-01-05 redaction/dhp_sept2016.tex
Changements texte

--- 2017-01-05 redaction/dhp_sept2016.tex
@@
 Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others).
-The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013.
+The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau,\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}} spanning from 1974 to 2013.
 We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
@@
 
-We use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level.
-As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.\footnote{Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.
-The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
+We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level.
+As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
 We thus obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix ??.} Depending on the considered year, this leaves us with around 200 sectors sectors, from approximately 200 countries of origin.
+Estimation results at the 4-digit classification level are reported in Appendix ??.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.
 
@@
 \paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p$ of a good $k$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
+To do so, we start from the equation that expresses the price $p$ of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}_{isk}$ (export, or fas price), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ikz}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation*}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ikz}} \label{eq:theory_equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
 \end{equation*}
@@
 All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
-The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.
+The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
 See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Confronted to this arbitrage, we retain the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
-This drives us to estimate a modified version of Equation (\ref{eq:equation0}), according to:
+Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
+Put it differently, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
+same structure of costs.\footnote{We ensure the robustness of our estimation results to this assumption in two ways.
+First, we conduct the estimations at the 4-digit level for some selected years.
+As reported in Appendix \ref{app:4digit}, the results are not substantially affected.
+Second, we provide a decomposition variance exercise on the observed cif-fas price.
+As reported in Appendix ???, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), according to:
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
 \end{equation}
-where $\tau_{i}$, $\tau_{s}$, $t_{i}$ and $t_{s}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.
+where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.
 To eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
@@
 \begin{eqnarray}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI} \\
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI} \\
 %\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA}
 \end{eqnarray}
 
-After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$ and $\widehat{t}^{add}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$, that is country-product specific, by year and transport mode.
-Under Model (a), when assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
+After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau_{s(k)}}$ and $\widehat{t}^{add}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode.
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau_{s(k)}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
 To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this case as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik}$.
-The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (which is itself dominated by the complete model (b), anticipating on further results.
+This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik}$.
+The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
 For sake of space saving, the results of this model are not reported in the paper.
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs, as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
-The first panel reports the results for a specification based on ad-valorem costs estimated alone, while the second panel presents estimates for a specification involving both ad-valorem and multiplicative component.
-Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
-Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
+It displays the mean and median values of each type of trade costs (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+Results for all years (available at the 3-digit level) are available upon request to the authors.
+We investigate the evolution of transport costs over the period in more details in Section \ref{sec:results_decomposition}.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
 Overall results (over the whole period) are not substantially affected if this year is included though.
@@
 
-Table \ref{tab:summary_results} calls for two types of comments.
-Starting with pure statistical concerns, it appears that mean values are systematically higher than medians - by around one percentage point.
-This is no surprise at all: once again, our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median.
-Besides, one can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point.
-This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects.
+Table \ref{tab:summary_results} yields the following comments.
+First, the magnitude of overall transport costs is sizeable.
+Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
+Air transport costs are smaller than sea transport costs.
+This is counter-intuitive, but certainly linked to selection effects.
 
-Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable.
-Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
-Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
+% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
+More precisely, for this specific year, we find...
+something slightly higher...
+
+%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
+
+Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
 Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%.
@@
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
-In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.%\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
+In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
 
@@
 
-The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996.
-Our own results point to a total decrease from 6.9 to 4.2\% on average for the same years.\footnote{For the sole 1994 year, \citet{Hummels_2001} finds an all-commodities trade-weighted average transport cost of 3.8 percent of the fob.
-price for the United States }\todo{This has to be clarified: What data uses Hummels?
-Does he distinghuish at some point between air and vessel.
-}
-
-XXX but our contribution goes beyond: we distinguish in this trend what comes from additive, and what comes form iceberg XXX
-
+At this stage, two interesting comments can be made.
 First, the magnitude of transport costs is higher for ocean shipping than for air transport.
@@
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
-We come back to this aspect in Section \ref{sec:results_trends}.}
-
-\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
-
-% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
-More precisely, for this specific year, we find...
-something slightly higher...
-
-%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
+We come back to this aspect in Section \ref{sec:results_trends}.}\todo{Comparer avec la litterature.
+Cf AvW 2004 pure transport costs 11\% markup over production costs.}
 
@@
 
-\section{Some detailed results \label{app:more_results}}
+\section{More on the estimation results \label{app:more_results}}
 
-In this section, we report the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})).
-For sake of space saving, we only report the results for a sample of years.
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} display the results for the 4-digit product-level estimation (again, for selected years).
+In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})).
+
+\subsection{Detailed results at the 3-digit classification level}
+
+In this section, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
+Table \ref{tab:result_air_3d_detail} reports the results for Air transport.
+The results for Ocean transport are displayed in Table \ref{tab:result_ves_3d_detail}.
 
@@
   \centering
-  \caption{Air: Dynamics of Trade Costs over Time, 3 digits (selected years)}
+  \caption{Air: Transport costs estimates, 3-digits (selected years)}
 \begin{center}
@@
 Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\ \hline
-\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs}}     \\
+\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs}}     \\ \hline
 Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034  \\
@@
 \hline
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}    \\
-\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &      \\
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}    \\ \hline
+\multicolumn{7}{l}{\textit{Additive term }}     \\ \hline
 \multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010  \\
 \multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005  \\
-\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
-\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &        \\
+\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\ \hline
+\multicolumn{7}{l}{\textit{Iceberg term}}  \\ \hline
 \multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017  \\
@@
 \hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
 \end{tabular}%
 \end{center} 
-  \label{tab:result_air_3d_detail}%
-\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
+\label{tab:result_air_3d_detail}
 \end{table}%
@@
  \hline
-   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       \\
+   \multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
 Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036  \\
@@
 Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018  \\
-\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &     \\
+\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs} }    \\
 \hline
-\textit{Additive term } &       &       &       &       &       &        \\
+\multicolumn{7}{l}{\textit{Additive term } }  \\ \hline
 Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015  \\
 Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 \\
-Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
-\textit{Iceberg term} &       &       &       &       &       &        \\
+Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\ \hline
+\multicolumn{7}{l}{\textit{Iceberg term} } \\ \hline
 Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 \\
@@
 \hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
 \end{tabular}%
-  \end{center}
-  \label{tab:result_ves_3d_detail}%
-\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
+\end{center} \label{tab:result_ves_3d_detail}
 \end{table}%
+
+\subsection{Estimation at the 4-digit level \label{app:4digit}}
+
+In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
+
+\paragraph{Transport cost estimates} Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in Air and Ocean transport respectively.
 
@@
   \centering
-  \caption{Air: Dynamics of Trade Costs over Time, 4-digit}
+  \caption{Air: Transport costs estimates, Selected years, 4-digit}
 \begin{center}
@@
    \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\
+Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
 \multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
@@
 \multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}  \\ \hline
-\textit{Iceberg term} &       &       &       &       &       &     \\
+\multicolumn{7}{l}{\textit{Iceberg term} }   \\ \hline
 Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016}  \\
 Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014}  \\
-Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\
-\textit{Additive term} &       &       &       &       &       &     \\
+Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\ \hline
+\multicolumn{7}{l}{\textit{Additive term} }   \\ \hline
 Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} \\
@@
 \hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
 \end{tabular}%
 \end{center}
-  \label{tab:result_air_rob}%
-\tiny{Notes:Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
+  \label{tab:result_air_rob}
 \end{table}%
@@
   \centering
-\caption{Vessel: Dynamics of Trade Costs over Time, 4-digit}
+\caption{Vessel: Transport costs estimates, Selected years, 4-digit}
 \begin{center}
@@
 Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
-\textit{Iceberg term} &       &       &       &       &       &     \\
+\textit{Additive term} &       &       &       &       &       &     \\
 Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015}  \\
@@
 \hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
 \end{tabular}%
@@
 \label{tab:result_ves_rob}%
-\vspace{-10pt}\tiny{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price. }
 \end{table}%
 
-\subsection{Goodness-of-fit: Detailed results \label{app:robust_4d}}
-
-In this section, we report the goodness-of-fit exercise conducted by transport mode, for selected years at the 4-digit product classification level.
-The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel), for a sample of some years.
+\paragraph{Goodness-of-fit tests at the 4-digit level} We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
+The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel).
 \begin{table}[htbp]
@@
 %\vspace{0.5cm}
-\scalebox{0.97}{
-\begin{tabular}{lcccccc}
+%\scalebox{0.97}{
+\begin{tabular}{l|cccccc}
 \hline
 \hline
-      & \multicolumn{5}{c}{Year}                      &  \\
+      & \multicolumn{6}{c}{Year}              \\
       & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
@@
 \hline\hline
-\end{tabular}}%
+\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
+\end{tabular}%
 \end{center}
- \vspace{-10pt} \tiny{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
 \end{table}%
@@
 \label{tab:good_fit_ves_rob}%
-
-\scalebox{0.97}{\begin{tabular}{l|cccccc}
+\begin{tabular}{l|cccccc}
 \hline
@@
 \hline\hline
-\end{tabular}}
+\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
+\end{tabular}%
 \end{center}
-  \vspace{-10pt}\tiny{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
 \end{table}%
+
+\subsection{Variance decomposition exercise \label{app:decomp_variance}}
+
+In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
+Precisely, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$), \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
+This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
+We also determine the share of the observed variance that can be attributed to the between-country variance.
+Results are reported in Figure \ref{fig:decomp_variance}.
+
+\begin{figure}[htbp]
+\caption{Variance decomposition (observed cif-fas price gap)}
+\label{fig:decomp_variance}
+\begin{center}
+\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+Two interesting results emerge from Figure \ref{fig:decomp_variance}.
+First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of same magnitude of order at the variance between sectors at the 3-digit level.
+Both account for between 30 and 40\% of the total variance in Air transport, depending on the years considered.
+This is also the case for Ocean transport, even if the difference between the between-country variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
+This delivers an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
+Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger that the between-country variance.
+This holds throughout the period and for both transport modes.
+This suggests that what primarily matters in international transport costs is mostly attributable to the product per se, rather than to the country where it comes from.
+By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
 

### 2017-01-05 redaction/dhp_sept2016.tex

@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
-Results for all years (available at the 3-digit level) are available upon request to the authors.
-We investigate the evolution of transport costs over the period in more details in Section \ref{sec:results_decomposition}.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
+It displays the mean and median values of each type of trade costs  (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
+The first panel reports the results for a specification based on ad-valorem costs estimated alone, while the second panel presents estimates for a specification involving both ad-valorem and multiplicative component.
+Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
 Overall results (over the whole period) are not substantially affected if this year is included though.
@@
 
-\textbf{Faire l'analyse mais sur l'ecart cif-fob observe et non plus reconstruit. Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
-
-Figure \ref{fig:totalTC_compeffects_excl} drives the following comments.
-First, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
+\textbf{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
+
+Figure \ref{fig:totalTC_compeffects_excl} drives the following comments. Air transport costs were reduced by 50\% and vessel transport costs
+by 60\%.
+(as already seen).
+However, in the case of vessel, a large part of that is a composition effect, probably because of the reduction in the share of bulky goods linked to the declining share of primary goods in total trade.
+The reverse happened for air transport : the goods shiped by air were increasingly more costly to transport.
+n line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
+
+(((At the agregate levelFirst, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
 This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that the fall in transport costs immediately started in 1974.
 Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 was due to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
-In line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
+I
 

### 2017-01-20 redaction/dhp_sept2016.tex
changes_JH_2001

@@
 The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996.
-Our own results point to a total decrease from 6.9 to 4.2\% on average for the same years.\footnote{For the sole 1994 year, \citet{Hummels_2001} finds an all-commodities trade-weighted average transport cost of 3.8 percent of the fob.
-price for the United States }\todo{This has to be clarified: What data uses Hummels?
-Does he distinghuish at some point between air and vessel.
-}
-
-XXX but our contribution goes beyond: we distinguish in this trend what comes from additive, and what comes form iceberg XXX
-
-First, the magnitude of transport costs is higher for ocean shipping than for air transport.
-This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period.
-Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air transport, and from 9.8\% to 4.8\% for ocean transport.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport.
+\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where quality XXX} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
+But the fact that trends are very close are comforting.
+%\todo{This has to be clarified: What data uses Hummels? Does he distinguish at some point between air and vessel? }
+
+It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air transport.
+This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period.
+Second, in both transport modes, the additive component appears of sizeable importance As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.

### 2017-01-20 redaction/dhp_sept2016.tex
chnagement section 2

@@
 We thus obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix ??.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.
+Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.
 
@@
 
-\begin{equation*}
+\begin{equation}
 \frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
-\end{equation*}
-
-\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
+\end{equation}
+
+\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
 
@@
 Second, we provide a decomposition variance exercise on the observed cif-fas price.
-As reported in Appendix ???, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), according to:
+As reported in Appendix ???, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
 \begin{equation}
@@
 \begin{eqnarray}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI} \\
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI} \\
 %\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
@@
 
-After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau_{s(k)}}$ and $\widehat{t}^{add}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau_{s(k)}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
+After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode.
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
 To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik}$.
+This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
 The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
 For sake of space saving, the results of this model are not reported in the paper.
-Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{nlI}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
+Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 The objective of this section is twofold.
-First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
 Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.
@@
 
-\subsection{Fab-cif prices}
-
 The Customs value is the value of imports as appraised by the U.S.
@@
 
-\section{More on the estimation results \label{app:more_results}}
-
-In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})).
-
-\subsection{Detailed results at the 3-digit classification level}
-
-In this section, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
+\section{Estimation at the 3-digit classification level \label{app:more_results}}
+
+\subsection{Transport costs estimates: More detailed results}
+
+In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
+Precisely, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
 Table \ref{tab:result_air_3d_detail} reports the results for Air transport.
@@
   \centering
-  \caption{Vessel: Dynamics of Trade Costs over Time (3 digits), selected years}
+  \caption{Vessel: Transport costs estimates, 3 digit (selected years)}
 \begin{center}
@@
 
-\subsection{Estimation at the 4-digit level \label{app:4digit}}
+\subsection{Variance decomposition exercise \label{app:decomp_variance}}
+
+In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
+Precisely, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$), \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
+This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
+We also determine the share of the observed variance that can be attributed to the between-country variance.
+Results are reported in Figure \ref{fig:decomp_variance}.
+
+\begin{figure}[htbp]
+\caption{Variance decomposition (observed cif-fas price gap)}
+\label{fig:decomp_variance}
+\begin{center}
+\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+Two interesting results emerge from Figure \ref{fig:decomp_variance}.
+First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of same magnitude of order at the variance between sectors at the 3-digit level.
+Both account for between 30 and 40\% of the total variance in Air transport, depending on the years considered.
+This is also the case for Ocean transport, even if the difference between the between-country variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
+This delivers an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
+Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger that the between-country variance.
+This holds throughout the period and for both transport modes.
+This suggests that what primarily matters in international transport costs is mostly attributable to the product per se, rather than to the country where it comes from.
+By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
+
+\section{Estimation at the 4-digit level \label{app:4digit}}
 
@@
 
-\paragraph{Transport cost estimates} Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in Air and Ocean transport respectively.
+\subsection{Transport cost estimates}
+
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in Air and Ocean transport respectively.
 
@@
 
-\paragraph{Goodness-of-fit tests at the 4-digit level} We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
+\subsection{Goodness-of-fit tests at the 4-digit level}
+
+We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
 The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel).
@@
 
-\subsection{Variance decomposition exercise \label{app:decomp_variance}}
-
-In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
-Precisely, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$), \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
-This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
-We also determine the share of the observed variance that can be attributed to the between-country variance.
-Results are reported in Figure \ref{fig:decomp_variance}.
-
-\begin{figure}[htbp]
-\caption{Variance decomposition (observed cif-fas price gap)}
-\label{fig:decomp_variance}
-\begin{center}
-\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
-Two interesting results emerge from Figure \ref{fig:decomp_variance}.
-First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of same magnitude of order at the variance between sectors at the 3-digit level.
-Both account for between 30 and 40\% of the total variance in Air transport, depending on the years considered.
-This is also the case for Ocean transport, even if the difference between the between-country variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
-This delivers an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
-Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger that the between-country variance.
-This holds throughout the period and for both transport modes.
-This suggests that what primarily matters in international transport costs is mostly attributable to the product per se, rather than to the country where it comes from.
-By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
-
 \section{More on eliminating the composition effects \label{app:comp-effects}}

### 2017-01-20 redaction/dhp_sept2016.tex
Changes_redac_2001_bis

@@
 For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air transport, and from 9.8\% to 4.8\% for ocean transport.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where quality XXX} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
+\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on freight rates.} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
 But the fact that trends are very close are comforting.
-%\todo{This has to be clarified: What data uses Hummels? Does he distinguish at some point between air and vessel? }
-
 It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air transport.
-This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period.
-Second, in both transport modes, the additive component appears of sizeable importance As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+
+Besides, we go beyond by distinguishing within these evolutions between multiplicative and additive transport costs.
+First, the decreasing trend holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period.
+Second, in both transport modes, the additive component appears of sizeable importance.
+As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
 We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
@@
 
-\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
-
-% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994.
-More precisely, for this specific year, we find...
-something slightly higher...
-
-%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.
+Before, we check how our estimates of additive costs compare to those of \citet{Irrazabal_2015}, which provide the most recent evidence on that matter.
+Based on Norwegian firm-level trade data for 2004, \citet{Irrazabal_2015} find that additive costs are on average 14\% of the median price.
+Expressed in our own terminology, this means they find that $\frac{t/\tau}{\widetilde{p}}=14\%$.
+Let us define the ratio of additive costs over total costs as $\frac{t}{\tau\widetilde{p}+t}$, and divide both numerators and denominators by $\tau\widetilde{p}$.
+We get: $\frac{t}{\tau\widetilde{p}+t}=\frac{t/\tau\widetilde{p}}{1+t/\tau\widetilde{p}}$.
+Plugging \citet{Irrazabal_2015} result gives 0,14/1+0,14= 12,5\%.
+Additive trade costs represents 12.5\% of total trade costs.
+Based on \citet{anderson_wincoop_jel}, we know that transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.
+Given that in 2004, our estimates for additive (multiplicative) transport costs is 2.9\% (2.7\%) for Vessel, the same reasoning as before leads to 2.9/(2.7+2.9)*11/74 = 7.7\% of total trade costs, that is 62\% (7.7/12.5=62\%) of \citet{Irrazabal_2015}'s estimates.
+For Air transportation, our estimates for additive (multiplicative) transport costs is 1.5\% (1.9\%), and the same computation brings that additive costs represent 6.6\% of total trade costs, that is 53\% of \citet{Irrazabal_2015}'s estimates.
+In other words, we find that additive transport costs represent between half and two thirds of the total additive trade costs, which seems totally consistent with the magnitudes reported by e.g.
+\citet{anderson_wincoop_jel}.
+
+%\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
 

### 2017-01-20 redaction/dhp_sept2016.tex
Rédaction point endogénéité

@@
 Second, we provide a decomposition variance exercise on the observed cif-fas price.
-As reported in Appendix ???, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
+As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
 \begin{equation}
@@
 
+One may be concerned that the specification of Equations (\ref{eq:model_nlI}}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport costs burden.
+Notably based on the pricing-to-market behavior of firms (see \citealp{Krugman-87}), we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
+However, we do not view this as raising an endogenity problem as this does not change our dependent variable of interest, the gap between the cif and the fas prices.
+Except in the choice between air and ocean transport (that is beyond our scope of analysis), the exporting firm has no influence on the cif-fas price gap.
+Otherwise stated, the firm chooses an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond her control.
+In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
+\smallskip
+
 After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode.
@@
 The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
-For sake of space saving, the results of this model are not reported in the paper.
+For sake of space saving, the results of this third model are not reported in the paper.
 Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
@@
 These results are reported in Section \ref{sec:results_decomposition}.
-
-\todo{A un moment, parler du risque d'endogeneite entre tauik tik et ptildeik, dire que pas de pb. Dire ou?}
 

### 2017-01-20 redaction/dhp_sept2016.tex
changement mineur

@@
 
-One may be concerned that the specification of Equations (\ref{eq:model_nlI}}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport costs burden.
+One may be concerned that the specification of Equations (\ref{eq:model_nlI}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport costs burden.
 Notably based on the pricing-to-market behavior of firms (see \citealp{Krugman-87}), we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).

### 2017-01-20 redaction/dhp_sept2016.tex
ai changé encore

@@
 
-One may be concerned that the specification of Equations (\ref{eq:model_nlI}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport costs burden.
+One may be concerned that the specification of Equations (\ref{eq:model_nlI}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
 Notably based on the pricing-to-market behavior of firms (see \citealp{Krugman-87}), we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-However, we do not view this as raising an endogenity problem as this does not change our dependent variable of interest, the gap between the cif and the fas prices.
-Except in the choice between air and ocean transport (that is beyond our scope of analysis), the exporting firm has no influence on the cif-fas price gap.
-Otherwise stated, the firm chooses an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond her control.
+However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply some a direct influence between the firm's prices and the size of transport costs, except in the choice between air and ocean transport (that is beyond our scope of analysis).
+Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
 In this respect, we are confident that our estimation strategy is immune from endogeneity problems.

### 2017-02-24 redaction/dhp_sept2016.tex
ajout intro

@@
 
-\todo{reprendre slides presentation LEDa pour la motivation de la question.}
+%\todo{reprendre slides presentation LEDa pour la motivation de la question.}
+%JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????
 
@@
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{\textbf{donner l'intuition.
-Deformation des prix relatifs, TBC}} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
+A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{This is due to the alteration of relative prices in a heterogeneous-firms trade framework.
+Lower additive trade costs reduce the export market prices of the highly performing exporters relative to lowly performing exporters and thus shift relative market shares among exporters towards the highly performing exporters - this of course does not happen with iceberg costs.
+Total export sales are then reallocated between a smaller number of high productivity firms.
+Therefore, the expected profits from exporting prior to entry increase all other things being equal by more in the case of lower per-unit costs, driving up incentives to enter the industry and bringing eventually the welfare enhancing intra-industry reallocations highlighted in \citet{melitz} framework.} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
 While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.

### 2017-03-24 redaction/dhp_sept2016.tex
Se positionner par rapport à Hummels sur les effets de composition

@@
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
+
+\textbf{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}
+
+\textbf{argument general}
+In comparison with \citet{hummels2007}, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
+Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.\footnote{In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}
+
+\textbf{argument sur les effets de composition}
+If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
+First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
+In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} eliminates the changes in the price per kilo dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
+146.} By contrast, we provide an explicit decomposition of this additive cost component over time: By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+As well, we are able to do so for the multiplicative component of trade costs.
+We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects.
+This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}).
+Last, for sake of comparison though, we do the same exercise as \citet{hummels2007}.
+
+\textbf{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps.
+Ligne 2, ce qu'on a sur la dimension iceberg cost estimated; ligne 3, sur les couts additifs estimes.
+Ligne 4, a faire, la somme des deux, estimee?
+Ou alors, partir simplement du TC observe, ie ce qu'on a deja?}
+
+Second, we characterize the evolution of the ``pure'' transport costs over time by measuring how these costs have changed over time, blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample), while \citet{hummels2007} captures the trend evolution of by the yearly dummy.
+Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components.
+\textbf{Pertinence de cet argument?
+A reformuler.}
+
 To do this, we estimate the following equations for observed costs and the estimated ad-valorem component using OLS:

### 2017-05-02 redaction/dhp_sept2016.tex
Pour soumission DEGIT

@@
 \usepackage[draft]{todonotes}
+%\usepackage[disable]{todonotes}
 %%GDPour cacher les notes, \usepackage[disable]{todonotes}
@@
 
-%\todo{reprendre slides presentation LEDa pour la motivation de la question.}
-%JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????
+\todo{reprendre slides presentation LEDa pour la motivation de la question.}
+\todo{JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????}
 
@@
 
-%\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
+\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
 
@@
 
-\textbf{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}
-
-\textbf{argument general}
+\todo{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}
+
+\todo{argument general}
 In comparison with \citet{hummels2007}, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
@@
 
-\textbf{argument sur les effets de composition}
+\todo{argument sur les effets de composition}
 If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
@@
 
-\textbf{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps.
+\todo{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps.
 Ligne 2, ce qu'on a sur la dimension iceberg cost estimated; ligne 3, sur les couts additifs estimes.
@@
 Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components.
-\textbf{Pertinence de cet argument?
+\todo{Pertinence de cet argument?
 A reformuler.}
@@
 
-\textbf{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
+\todo{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
 
@@
 
-\paragraph{what else?}
+\todo{what else?}
 

### 2017-05-17 redaction/dhp_sept2016.tex
Changements redaction

@@
 
-\date{September 2016}
+\date{May 2017}
  \maketitle
@@
 Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-In this vein, \citet{anderson_wincoop_jel} obtain that the 74\% markup made by international trade costs can be divided between 44\% for border-related trade barriers and 21\% for transportation costs.
-This last figure suggests that the share of transport costs in the consumer price of manufactured goods is substantial.
-\citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
+In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.
+Equivalently, international transportation costs would represent a 21\% markup over production costs.
+\citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3, suggesting a non-negligible impact of transport costs on trade flows.
 If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
@@
 
-\todo{reprendre slides presentation LEDa pour la motivation de la question.}
-\todo{JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????}
-
 Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis\footnote{Rigourously speaking, the ``iceberg cost'' includes the ad-valorem costs and the fact that the costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terme ``iceberg'' and ``ad-valorem'' interchangeably, as commonly made in the literature.}).
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly made in the literature.}).
 However, many common empirical facts support the existence of additive costs.
@@
 
+\todo{mettre une phrase pour dire qu'on confirme la litterature sur le role des additifs, mais de maniere differente.
+Hummels et Skiba sur des donnes de 1994, on va beaucoup plus loin qu'eux dans cette voie.}
+
 Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
@@
 As pointed out by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time.
+In contrast to his work though, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
+Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.
 After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
@@
 \end{equation}
+
+\todo{expliciter ici notre strategie d'estimation par rapport a Hummels et Skiba, 2004.
+In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}
 
@@
 
-\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
-
 \subsection{Assessing the importance of additive transport costs}
@@
 
-Before making any definite statement about this, it is worth emphasizing that the evolution of transport costs depends on the one hand of the evolution of per product and per partner costs and on the other hand of the evolution of the composition of trade.
-Total transport costs may have decreased over time because the share of neighbour countries in total US trade or the share of cheaper to transport goods have increased, independently of any change in transport costs \textit{per se}.
-As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.
-
-\todo{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}
-
-\todo{argument general}
-In comparison with \citet{hummels2007}, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
-Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.\footnote{In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}
-
-\todo{argument sur les effets de composition}
+Before making any definite statement about this, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i°)} the evolution of per product and per partner transport costs and \textit{ii°)} the evolution of the composition of trade.
+Total transport costs may thus have decreased over time because the share of neighbour countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii°)}), independently of any change in transport costs \textit{per se}.
+As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
+per product- and per partner- transport costs.
+
 If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
@@
 In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} eliminates the changes in the price per kilo dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
-146.} By contrast, we provide an explicit decomposition of this additive cost component over time: By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+146.} By contrast, our estimation strategy allows for an explicit decomposition of the additive cost component over time/product/country partner.
+\textbf{stop here} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
 As well, we are able to do so for the multiplicative component of trade costs.
-We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects.
+We also agglomerate the two components in an unified measure of transport costs from which we exclude the composition effects.
 This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}).

### 2017-05-17 redaction/dhp_sept2016.tex
changement

@@
 
-In this section, we investigate the role of the additive component in total transport costs by exploring the time dimension of our data.
+In this section, we investigate the role of the additive component in total transport costs in a historical perspective.
 As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
@@
 First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974.
-Second, the decrease appears more important for ocean than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport.
-That said, it may be the case the oil shock inflated transportation costs that specific year.
-If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates).
+\todo{Guillaume, ameliorer les choses} If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates).
 These results are consistent with \citet{hummels2007}.
 
-Before making any definite statement about this, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i°)} the evolution of per product and per partner transport costs and \textit{ii°)} the evolution of the composition of trade.
+Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i°)} the evolution of per product and per partner transport costs and \textit{ii°)} the evolution of the composition of trade.
 Total transport costs may thus have decreased over time because the share of neighbour countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii°)}), independently of any change in transport costs \textit{per se}.

### 2017-05-17 redaction/dhp_sept2016.tex
comments_Table1

--- 2017-05-17 redaction/dhp_sept2016.tex
@@
 This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects.
+Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearizations.
+Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
+Descriptive statistics showing that it is indeed the case available upon request to the authors.
 
@@
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
-In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
+In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
 

### 2017-05-17 redaction/dhp_sept2016.tex
chnagement encore

--- 2017-05-17 redaction/dhp_sept2016.tex
@@
 \end{equation}
-
-\todo{expliciter ici notre strategie d'estimation par rapport a Hummels et Skiba, 2004.
-In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}
 
@@
 First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
-In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} eliminates the changes in the price per kilo dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
-146.} By contrast, our estimation strategy allows for an explicit decomposition of the additive cost component over time/product/country partner.
-\textbf{stop here} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
+146.} By contrast, our estimation strategy allows for an explicit decomposition of the additive cost component over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner.
+In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
+By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
 As well, we are able to do so for the multiplicative component of trade costs.

### 2017-05-17 redaction/dhp_sept2016.tex
changement encore

--- 2017-05-17 redaction/dhp_sept2016.tex
@@
 If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
-First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
-In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
-146.} By contrast, our estimation strategy allows for an explicit decomposition of the additive cost component over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner.
-In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
-By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+First, we provide a broader characterization of the time trends in transport costs, by identifying the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for the two additive and ad-valorem dimensions.
+Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner.
+In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
+By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account.
+By contrast, in its decomposition exercise to characterize the time trend in transport costs, \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
+146.
+} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
 As well, we are able to do so for the multiplicative component of trade costs.
@@
 This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}).
-Last, for sake of comparison though, we do the same exercise as \citet{hummels2007}.
+Last, for sake of comparison, we do the same exercise as \citet{hummels2007}.
+
+Our second difference with \citet{hummels2007} relates to the way we characterize the evolution of the ``pure'' transport costs over time.
+\citet{hummels2007} captures the trend evolution of by the yearly dummy.
+Rather, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
+This notably enables us to obtain measures of transport costs that are easily comparable between transport modes and transport costs components.
 
@@
 Ou alors, partir simplement du TC observe, ie ce qu'on a deja?}
-
-Second, we characterize the evolution of the ``pure'' transport costs over time by measuring how these costs have changed over time, blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample), while \citet{hummels2007} captures the trend evolution of by the yearly dummy.
-Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components.
-\todo{Pertinence de cet argument?
-A reformuler.}
 

### 2017-05-17 redaction/dhp_sept2016.tex
Discussion sur le choix du début
1974 ou 1980 ? Cela ne change pas grand chose.

@@
 Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
-Two main comments can be made.
-First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974.
-\todo{Guillaume, ameliorer les choses} If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates).
+Both air and ocean shipping exhibit a downward trend in overall transport costs since 1974.
+(-2.1\% a year for mean air transport costs and -2.0\% a year for mean ocean transport costs).
+\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
+However, computing the time trends from 1980 does not change dramatically the picture.
+The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs.
+In order to go as far as possible into the past, this paper will focus on the results from 1974}
+ \todo{GD : j'ai amélioré. Qu'en penses-tu ?}
 These results are consistent with \citet{hummels2007}.

### 2017-05-17 redaction/dhp_sept2016.tex
Chnagement effet de composition

@@
 
-\smallskip
-
 \section{Decomposing Transport Costs: Characterizing the trends \label{sec:results_trends}}
@@
 146.
-} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
-As well, we are able to do so for the multiplicative component of trade costs.
-We also agglomerate the two components in an unified measure of transport costs from which we exclude the composition effects.
-This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}).
-Last, for sake of comparison, we do the same exercise as \citet{hummels2007}.
+} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+We proceed similarly for the multiplicative component of trade costs and, for sake of completeness, on the observed cif-fas price gap.
 
@@
 This notably enables us to obtain measures of transport costs that are easily comparable between transport modes and transport costs components.
-
-\todo{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps.
-Ligne 2, ce qu'on a sur la dimension iceberg cost estimated; ligne 3, sur les couts additifs estimes.
-Ligne 4, a faire, la somme des deux, estimee?
-Ou alors, partir simplement du TC observe, ie ce qu'on a deja?}
 
@@
 Figure \ref{fig:totalTC_compeffects_excl} reports the results.
-Panel (a) displays the time changes in transport costs when only iceberg costs are modeled.
-In Panel (b), we report the sum of the ad-valorem and the additive costs (re-scaled to start from the reference value 100), for each transport mode.\footnote{For each mode, and by year, we build the total value of transport costs as the weighted sum of both the ad valorem index and the additive index, taking as weighting scheme the relative share of each type of costs in the overall total cost in 1974 (using the figures at the root of Figure \ref{fig:mult_alone_withadd}), i.e.
-respectively 48.2 \% for additive costs in Vessel and 42.3\% for Air in 1974.}
+Panel (a) displays the time changes in the observed transport costs, by transport mode.
+In Panels (b) and (c), we report the time changes of the ad-valorem costs and the additive costs respectively (re-scaled to start from the reference value 100), for each transport mode.
 
@@
 \todo{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
+
+\todo{Laisser tomber l'observe et faire la somme I+A en pondérant par le poids respectif de chacun dans le total des TC en 1974.}
 

### 2017-05-17 redaction/dhp_sept2016.tex
changement encore

@@
 Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
-Both air and ocean shipping exhibit a downward trend in overall transport costs since 1974.
-(-2.1\% a year for mean air transport costs and -2.0\% a year for mean ocean transport costs).
-\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
-However, computing the time trends from 1980 does not change dramatically the picture.
+Two main comments can be made.
+First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
+However, computing the time trends from 1980 does not dramatically change the picture.
 The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs.
-In order to go as far as possible into the past, this paper will focus on the results from 1974}
- \todo{GD : j'ai amélioré. Qu'en penses-tu ?}
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} Second, the decrease appears more pronounced for sea than air transportation.
 These results are consistent with \citet{hummels2007}.
@@
 } By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
-We proceed similarly for the multiplicative component of trade costs and, for sake of completeness, on the observed cif-fas price gap.
+We proceed similarly for the multiplicative component of trade costs.
+We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects.
+This
+allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
+
+\todo{Laisser tomber l'observe et faire la somme I+A en pondérant par le poids respectif de chacun dans le total des TC en 1974.}
 
@@
 \todo{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
-
-\todo{Laisser tomber l'observe et faire la somme I+A en pondérant par le poids respectif de chacun dans le total des TC en 1974.}
 

### 2017-05-19 redaction/dhp_sept2016.tex
Rédaction effets de composition + annexe

@@
 We proceed similarly for the multiplicative component of trade costs.
-We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects.
+We also agglomerate the two components (additive and iceberg) in a unified measure of transport costs.
 This
 allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
-
-\todo{Laisser tomber l'observe et faire la somme I+A en pondérant par le poids respectif de chacun dans le total des TC en 1974.}
 
@@
 
-To do this, we estimate the following equations for observed costs and the estimated ad-valorem component using OLS:
-%the ad-valorem cost, based on the series obtained $\widehat{\tau}_{ikt}$ and $$$%\widehat{t}_{ikt}$ that have been previously estimated:
-\footnote{Even if Equation (\ref{eq:compeffects_mult}) is estimated both for observed costs and the estimated ad-valorem componen we write it only for  $\widehat{\tau}_{iskt}$ for sake of notational simplicity.
-For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}
-%$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
-
+To do this, we estimate the following equations for the estimated ad-valorem component using OLS:
 \begin{eqnarray}
@@
 We cannot use the same equation for the estimated additive component, as by construction, the sector fixed effect and the country fixed effect are additive rather than multiplicative.
-We estimate the following equation using non-linear least squares:
+We estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}
 %, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
@@
 As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: the country dimension (Term (a)), the product dimension (Term (b)) and the ``pure transport costs time trend'' (Term (c)).
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}
-All equations are estimated using by transport mode.
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.} Both equations are estimated using by transport mode.
 
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-From the estimation of Equation (\ref{eq:compeffects_mult}) (on both observed trade costs and estimated multiplicative component
-%$\widehat{\tau}_{ikt}$ and $\widehat{\tau}_{ikt}^{ice}$)
-, we built the variable $\Gamma_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
+From the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
 \begin{equation*}
-\Gamma_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
 \end{equation*}
@@
 
-\noindent As a result, the three series (for $\Gamma_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+\noindent As a result, the three series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+
+Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the ``raw'' (without composition effects excluded) and the ``pure'' estimated transport costs (composition effects excluded, see Appendix \ref{app:comp-effects}).
 
 Figure \ref{fig:totalTC_compeffects_excl} reports the results.
-Panel (a) displays the time changes in the observed transport costs, by transport mode.
+Panel (a) displays the time changes in the overall transport costs, by transport mode.
 In Panels (b) and (c), we report the time changes of the ad-valorem costs and the additive costs respectively (re-scaled to start from the reference value 100), for each transport mode.
@@
 
-\todo{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}
-
-Figure \ref{fig:totalTC_compeffects_excl} drives the following comments. Air transport costs were reduced by 50\% and vessel transport costs
-by 60\%.
-(as already seen).
-However, in the case of vessel, a large part of that is a composition effect, probably because of the reduction in the share of bulky goods linked to the declining share of primary goods in total trade.
-The reverse happened for air transport : the goods shiped by air were increasingly more costly to transport.
-n line with \citet{hummels2007}, this confirms the importance of excluding these composition effects before assessing the time trends of transport costs.
-
-(((At the agregate levelFirst, whatever the transport mode considered, transport costs have been roughly constant from 1974 to 1984.
-This stands in sharp contrast with Figures \ref{fig:mult_alone_withadd} and \ref{fig:decomp_TC_3d}, which rather conveyed the view that the fall in transport costs immediately started in 1974.
-Combining these results therefore indicates that the reduction of transport costs between 1974 and 1984 was due to changes in the composition of trade patterns (either in the country or the product dimension, or both) of the US, rather than a decrease in transport costs \textit{per se}.
-I
-
-Second, after excluding the composition effects, we find that all types of transport costs have steadily decreased, to reach a reduction of 40\% over the period 1984-2011, in accordance with the literature.
-Further, if we only assume ad-valorem costs (Panel (a)), we can infer that the decrease is stronger for air transport than for air shipping over the period 1984-2005.
-This in accordance with the findings of \cite{hummels2007} and \cite{Behar_Venables}.
-Yet, going deeper in the decomposition between the additive and the ad-valorem components substantially modifies the picture, as it can be inferred from Panel (b).
-When the additive costs are taken into account, there is now not much difference in the reduction of transport costs between air transport and ocean shipping, and both display very similar time changes.
-The downward trend in transport costs indeed is roughly of the same order of magnitude throughout the years, for both transport modes.
-Taking into account the additive component therefore substantially modifies the analysis of the dynamics of transport costs over time.
-In Figure \ref{fig:TC_add_mult_compeffects_excl}, we investigate this result further, by reporting the changes in both the additive and ad-valorem components of transport costs, by transport mode.
+We also proceed to the same analysis at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.\footnote{See more details in Appendix \ref{app:comp-effects}.
+We also run the same exercise at the 1-digit level.
+For sake of conciseness, these results are not reported.
+They are available upon request to the authors.} The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
 
 \begin{figure}[htbp]
-\caption{Decomposing transport costs over time,
-composition effects excluded}
-\label{fig:TC_add_mult_compeffects_excl}
-\begin{center}
-\includegraphics[width=14cm, height=7cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
-%{\footnotesize {OECD data} }
+\caption{Transport costs (with and without composition effects), Manufacturing}
+\label{fig:totalTC_compeffects_excl_manuf}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_manuf.pdf}
 \end{center}
@@
 
-Again, the decomposition between both components of transport costs brings valuable insights.
-For ocean shipping, both additive and iceberg components exhibit a downward pattern since 1985.
-Air transport exhibit much more contrasted trends.
-While the ad-valorem component has been steadily declining over 1985-2000, the additive cost has been rather stable on the same period.
-Such a difference therefore most likely explains the difference of results between Panels (a) et (b) of Figure \ref{fig:totalTC_compeffects_excl}.
-Skipping the additive component leads to an important omission bias in the characterization of the time changes of transport costs.
-When correctly measured (ie, when the additive component is included), there is no noticeable difference in the downward trend of transport costs between ocean and air transport, in contradiction with the findings of \citet{hummels2007} and \cite{Behar_Venables}.
-Again, this finding points out to the importance of taking into account the additive component of international transport costs.
-\smallskip
-
-Our analysis also brings valuable results to the findings of \citet{hummels2007}, as we extend his time coverage (end year 2004) by one decade (to stop in 2013).
-On this decade, our analysis reveals substantial changes in the time trend of transport costs in the air sector.
-As reported in Figure \ref{fig:totalTC_compeffects_excl}, Panel (b), transport costs have raised between 2005 and 2010, to substantially decline on the very recent years (2011-2013), by more than 20\%.
-As reported in Figure \ref{fig:TC_add_mult_compeffects_excl}, this can be attributed to the iceberg component (while the additive cost has conversely been drastically reduced since 2005).
-These results suggest that important changes in the structure of costs ate at work in the air transport in the very recent period, that should be investigated in further research.)))
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects), Primary goods}
+\label{fig:totalTC_compeffects_excl_primary}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_primary.pdf}
+\end{center}
+\end{figure}
+
+Consider first the time trends in transport costs at the aggregate level (Figure \ref{fig:totalTC_compeffects_excl}).
+Two main results emerge.
+First, we do not find evidence of a significant difference between the ``raw'' transport costs and the `pure'' transport costs in Air transport.
+This result for each of the two components, additive and ad-valorem (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
+Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``pure'' transport costs.
+This stands in contrast with \citet{hummels2007}, which obtained that the ``pure'' transport costs had decreased more (over 1973- 2004) than the unfitted ones, suggesting an important role to composition effects.
+As noted above, this difference of results can be explained by the difference in methodology.
+Whereas \citet{hummels2007} attributes all changes that are not changes in the ad-valorem components to composition effects, we show that part (and as a matter of fact, most) of these changes come from a reduction in the ``pure'' transport costs, in their additive dimension.
 
@@
 
+\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
 \begin{table}[htbp]
@@
 
-\section{More on eliminating the composition effects \label{app:comp-effects}}
+\section{Eliminating the composition effects: More details \label{app:comp-effects}}
 
@@
 
+\subsection{Methodolgy}
 \paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component. Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
@@
 
-From this, we obtain the foolowing recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
+From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
 Given that $\tau >1$, we can rewrite things to get the percentage change between year 1974 and any year $t>1974$:
@@
 At this stage though, it remains specific to a product-origin country pair.
-Next step is to build the index $\Gamma_t$ such that:
-\begin{equation*}
- \Gamma_t= 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
+Next step is to build the index $\Gamma^{adv}_t$ such that:
+\begin{equation}
+ \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}  \label{eq:tcadv_compoeffect}
+\end{equation}
 \noindent with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
@@
 
-From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$. Given the contraint $t>0$, we then obtain the percentage change from 1974 from:
+From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$. Given the constraint $t>0$, we then obtain the percentage change from 1974 from:
 
@@
 Noticing that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
-$$\Gamma^{add}_t  = 100\exp(\gamma_t) $$
+\begin{equation}
+\Gamma^{add}_t  = 100\exp(\gamma_t)  \label{eq:tcadd_compoeffect}
+\end{equation}
+
+\paragraph{For the total cost measure} We also build a measure of the ``overall'' transport costs, that agglomerates our estimates of the two additive and iceberg components.
+We construct this measure (by transport mode) on both the unfitted series and the ``pure'' transport cost series (composition effects excluded).
+Even if obeying to the same logic, we proceed slightly differently for the unfitted and the fitted measures though, as we now explain.\smallskip
+
+For the unfitted measure, based on Equation (\ref{eq:base_estimee}), we build for each transport mode, the ``overall'' transport cost as:
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ have been estimated (by year) as explained in Section \ref{sec:data_method}, $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\widehat{t}_t$ the additive component, both expressed in percentage of the fas price.
+For sake of comparison, we transform this ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
+
+We apply a slightly different procedure for constructing the fitted measure of total transport cost (i.e., composition effect excluded).
+In this case, we start directly from the two indices previously obtained (Equations (\ref{eq:tcadv_compoeffect}) and (\ref{eq:tcadd_compoeffect}) (by transport mode).
+We build the ``overall'' transport cost measure as a weighted sum of the two indices, using the relative weight of the additive cost component in total cost in 1974, $\omega_{air}= 0.423$ and $\omega_{ocean}= 0.482$ (obtained from our estimates in Section \ref{sec:results_decomposition}, see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), according to:
+$$\Gamma^{tc, pure}_t = \omega_{x}\Gamma^{add}_t +(1-\omega_{x})\Gamma^{adv}_t $$
+
+\noindent with $x=\{air, ocean\}$.
+
+\subsection{At the sector level}
+
+\todo{expliquer comment la repartition entre all, primary, manuf a ete faite.}
 

### 2017-05-19 redaction/dhp_sept2016.tex
changement nouvelle methode

@@
 Figure \ref{fig:totalTC_compeffects_excl} reports the results.
-Panel (a) displays the time changes in the overall transport costs, by transport mode.
-In Panels (b) and (c), we report the time changes of the ad-valorem costs and the additive costs respectively (re-scaled to start from the reference value 100), for each transport mode.
+In Panels (a) and (b), we report the time changes of the ad-valorem costs and the additive costs respectively.
+Panel (c) displays the time changes in the total transport costs (built as the sum of the two components as detailed above).
+In the three cases, the series are reported as indices, starting from the reference value 100 in 1974 and by transport mode.
 
@@
 Consider first the time trends in transport costs at the aggregate level (Figure \ref{fig:totalTC_compeffects_excl}).
-Two main results emerge.
-First, we do not find evidence of a significant difference between the ``raw'' transport costs and the `pure'' transport costs in Air transport.
-This result for each of the two components, additive and ad-valorem (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
+Three main results emerge.
+First, we do not find evidence of a significant difference between the raw transport costs and the ``pure'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``pure'' transport costs.
@@
 As noted above, this difference of results can be explained by the difference in methodology.
-Whereas \citet{hummels2007} attributes all changes that are not changes in the ad-valorem components to composition effects, we show that part (and as a matter of fact, most) of these changes come from a reduction in the ``pure'' transport costs, in their additive dimension.
+Whereas \citet{hummels2007} attributes all changes that are not in the ad-valorem component to composition effects, we show that part (and as a matter of fact, most) of these changes do come from a reduction in the ``pure'' transport costs, in their additive dimension.
+
+Second, composition effects are more pronounced in ocean transport.
+Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects.
+Precisely, the role of composition effects is significant for the multiplicative component, whereas the reduction in the additive costs is mostly attributable to changes in ``pure'' additive costs.
+In line with \citet{hummels2007}, these results confirm the important role of composition effects in maritime transport costs.
+The direction of the effects differs though.
+Whereas  \citet{hummels2007} obtains a more pronounced reduction in the ``pure'' transport costs than the unfitted series, we obtain the opposite result.
+On top of the reduction of the ``pure'' transport costs, overall maritime transport costs have reduced over time because of the composition of flows, \textbf{one potential explanation being the reduction in the share of bulky goods linked to the declining share of primary goods in total trade.
+Aller plus loin?
+mettre la part du primary dans le temps?
+Pertinence de cette hypothese?}
 

### 2017-06-16 redaction/dhp_sept2016.tex
Petites modif Guillaume

@@
 %\usepackage[T1]{fontenc}
-%\usepackage[utf8]{inputenc}
+\usepackage[utf8]{inputenc}
 %%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
@@
 \newcommand{\C}{$^c$}
+
 \begin{document}
@@
 In this section, we investigate the role of the additive component in total transport costs in a historical perspective.
-As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
-Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}).
+As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time by transport mode.
+Let us consider first the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}).
 Two main comments can be made.
@@
 
-Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i°)} the evolution of per product and per partner transport costs and \textit{ii°)} the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighbour countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii°)}), independently of any change in transport costs \textit{per se}.
+Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner transport costs and \textit{ii)} the evolution of the composition of trade.
+Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
 As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
@@
 In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
+\todo{Est-ce qu'on peut faire ça sans présenter l'équation d'Hummels pour de vrai ?}
 By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account.

### 2017-07-06 redaction/dhp_july2017.tex

--- 2017-06-16 redaction/dhp_sept2016.tex
@@
 %\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
 %%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
@@
 
-\date{May 2017}
+\date{July 2017}
  \maketitle
@@
 However, many common empirical facts support the existence of additive costs.
-This paper provides a quantitative assessment of the size and the importance of the additive component in international transport costs.
-Using SITC 3 and 4 digit cif-fab unit values over 1974-2013 taken from US imports data, we estimate the two components of transport costs by transport mode (air or ocean).
-We find that additive costs are 2.85\% of fab unit values for ocean shipping (and ad-valorem ones 3.22\%).
-These values are respectively equal to 1.9\% and 2.5\% for air transport.
-Further, we show that taking additive costs into account improves the fit of the modelling of transport costs.
-We also use the time dimension of our data to characterize the evolution of transport costs.
-After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
-Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
-While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s.
+This paper provides a quantitative assessment of the size and the importance of the additive component in international transport costs, measured as the cif-fab price gap, using US imports flows over 1974-2013.
+We thus find that additive transport costs are not negligible, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
+Further, taking additive costs into account significantly improves the fit of the modelling of transport costs.
+We also use the time dimension of our data to characterize the patterns of transport costs over time.
+After correcting for composition effects, we find that international transport costs have decreased by 50\% in air transport and 60\% in maritime transport over the period, and of same order of magnitude for both additive and iceberg components.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.
+
+\textbf{pas tres sexy la fin. Dire tout de meme que ca reste non negligeable? Avec la valeur moyenne sur 2010-2013 par ex, pour additif et pour multiplicative.}
 \end{abstract}
@@
 
+\begin{figure}[htbp]
+\caption{Share of primary goods in the value of total US imports}
+\label{fig:Share_prim_goods}
+\begin{center}
+\includegraphics[height=4in]
+{Share_of_primary.pdf}
+\end{center}
+\end{figure}
+
 \section{Robustness checks}

### 2017-07-06 redaction/dhp_july2017.tex

@@
 We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} Second, the decrease appears more pronounced for sea than air transportation.
-These results are consistent with \citet{hummels2007}.
+These results are consistent with \cite{hummels2007}.
 
@@
 Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
+As argued by \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
 per product- and per partner- transport costs.
 
-If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
+If we share this view with \cite{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
 First, we provide a broader characterization of the time trends in transport costs, by identifying the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for the two additive and ad-valorem dimensions.
-Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner.
-In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
-\todo{Est-ce qu'on peut faire ça sans présenter l'équation d'Hummels pour de vrai ?}
+Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba},
+ assumes a constant share of additive transport costs over time, product and country partner. In \cite{hummels_skiba}, 
+ this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
 By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account.
-By contrast, in its decomposition exercise to characterize the time trend in transport costs, \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p.
+By contrast, in its decomposition exercise to characterize the time trend in transport costs, \cite{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \cite{hummels2007}, p.
 146.
@@
 allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
+
+\todo{Est-ce qu'on peut faire ça sans présenter l'équation d'Hummels pour de vrai ?}
 

### 2017-07-06 redaction/dhp_july2017.tex
typo_table1

@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
+    \multicolumn{5}{l}{\textbf{With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
     Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
@@
     %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{\textbf{With Additive \& Ad-Valorem Trade Costs} } \\ \hline
+    \multicolumn{5}{l}{\textbf{With Additive \& Ad-Valorem Transport Costs} } \\ \hline
    \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline

### 2017-07-06 redaction/online_appendix_dhp2017.tex
creation_OA

\documentclass[11pt,twoside, authoryear]{elsarticle}
\usepackage[frenchb, english]{babel}
\usepackage[utf8]{inputenc}
%\usepackage[utf8]{fontenc}
%\usepackage[ansinew]{inputenc}
\usepackage{setspace}
\usepackage{babel,varioref}
\usepackage{supertabular}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsExtensions{.pdf,.mps,.png,.jpg,.eps}
\usepackage{lscape}
\usepackage[authoryear]{natbib}
%\usepackage[frenchb,english]{babel}
\usepackage{multirow}
\usepackage{ifthen}
%\usepackage{harvard}
\usepackage{vmargin}
\usepackage{verbatim}
\usepackage{array}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=Black, citecolor=blue, urlcolor=blue}
%\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{0cm}
\usepackage{booktabs,caption,fixltx2e}

\usepackage[flushleft]{threeparttable}
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
\usepackage{adjustbox}
\usepackage{chngcntr}
\counterwithin{table}{section}
%\usepackage[round]{natbib}

\renewcommand{\thesection}{\Alph{section}}

% \newcommand\section[1]{%
%   \refstepcounter{section}%
%   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
%   \sectionmark{#1}}
%   }

% % \refstepcounter{section}%
% %   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
% %   \sectionmark{#1}}

%   \newcommand\subsection[1]{%
%   \refstepcounter{subsection}%
%   \addcontentsline{toc}{subsection}{\protect\numberline{\thesubsection}#1}%
%   %\subsectionmark{#1}
%   }

%\EnableSectionsInLOFT

\begin{document}
\title{\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix (Not for publication)}

\urlstyle{rm}
\author{Guillaume Daudin\footnote{\noindent Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Corresponding author. Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille1.fr}} \qquad \quad Lise Patureau\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France ; email: \url{lise.patureau@dauphine.fr}}}
%\date{\vspace{-5ex}}
%
%%\author[lille,brest,cepii]{J\'{e}r\^{o}me H\'{e}ricourt\fnref{fn1}}
%\fntext[fn1]{Corresponding author. Address: D\'{e}partement de Droit, \'{E}conomie, Gestion et AES, 12, Rue de Kergoat - CS 93837, 29238 Brest Cedex 3, France. Tel: (0033) 1 53 68 55 14. E-mail: jerome.hericourt@univ-brest.fr}
%\address[lille]{Univ. Lille, CNRS, IESEG School of Management, Institut Catholique de Lille, UMR 9221 - LEM - Lille Économie Management}
%\address[brest]{University of Brest-ICI (EA 2652), France}
%\address[cepii]{CEPII, France}
%
%\author[lille]{Clément Nedoncelle\fnref{fn2}}
%%\address[lille]{University of Lille - LEM-CNRS (UMR 9221), France}
%\fntext[fn2]{Address: Université de Lille, Facult\'{e} de Sciences \'{E}conomiques et Sociales, Cit\'{e} Scientifique - B\^{a}t SH2, 59655 Villeneuve d'Ascq Cedex, France. E-mail: clement.nedoncelle@ed.univ-lille1.fr}
%

\maketitle

%\newpage

\tableofcontents
\vspace{1cm}
\newpage
\listoftables

\newpage
%\section{\label{sec:addi}Additional Tables}

	\renewcommand\thesubsubsection{\Alph{subsection}.\arabic{subsubsection}}

	\renewcommand\thesubsection{\Alph{subsection}}

\subsection{Additive-costs only Model \label{oa_sec:additive_only}}

%%%% RAW TABLE TO BE FOUND AT: "...\Dropbox\Papier_Lise_Guillaume\trade_cost\results\3_models"
% modele iceberg tout seul : terme nlI
% colonnes L à U: modèle avec les deux
% termes nla, de V à Z, modèles avec que de l'additif
%nlI quer iceberg
%nla que additif
%nl : les deux
\setcounter{table}{0}
\renewcommand{\thetable}{A.\arabic{table}}
%\subsubsection{Nominal Exchange Rate Volatility}

\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Transport costs estimates: Summary \label{sec_oa:add_only}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{\textbf{With only Additive Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{\textbf{With Additive \& Ad-Valorem Transport Costs} } \\ \hline
   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}& & & &   \\ \hline
    Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
    \multicolumn{5}{l}{\textbf{Data} ($p/\widetilde{p}$, in \%) } \\ \hline
        Mean & 5.3 & 5.0& 5.6&3.9 \\
        Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows. The additive term is expressed in fraction of fas price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
\end{table}%

\begin{landscape}
\begin{table}[htbp]\centering
 \begin{threeparttable}
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{\label{NER_ext}Alternative Measure of Exchange-rate Volatility - Nominal Exchange Rate (NER) Volatility - Extensive Margin}
\begin{tabular}{l*{8}{c}}
\hline\hline
               Dep. Variable &\multicolumn{4}{c}{Entry}&\multicolumn{4}{c}{Participation}\\
                                &\multicolumn{4}{c}{$Pr(X_{ijt}>0\mid X_{ijt-1}=0)$}&\multicolumn{4}{c}{$Pr(X_{ijt}>0$)}\\
                 \cmidrule(lr){2-5}
  \cmidrule(lr){6-9} \\
                    &\multicolumn{1}{c}{(1)}&\multicolumn{1}{c}{(2)}&\multicolumn{1}{c}{(3)}&\multicolumn{1}{c}{(4)}&\multicolumn{1}{c}{(5)}&\multicolumn{1}{c}{(6)}&\multicolumn{1}{c}{(7)}&\multicolumn{1}{c}{(8)}\\

\hline

Ln Bil NER Volatility                 &      -0.002      &       0.006      &                  &                  &       0.004      &       0.013      &                  &                  \\
                    &     (0.002)      &     (0.005)      &                  &                  &     (0.003)      &     (0.011)      &                  &                  \\
[1em]
Ln GDP              &       0.133{$^a$}&       0.133{$^a$}&                  &                  &       0.149{$^a$}&       0.148{$^a$}&                  &                  \\
                    &     (0.011)      &     (0.011)      &                  &                  &     (0.019)      &     (0.019)      &                  &                  \\
[1em]
Ln country price index&       0.012{$^a$}&       0.012{$^a$}&                  &                  &       0.021{$^a$}&       0.021{$^a$}&                  &                  \\
                    &     (0.002)      &     (0.002)      &                  &                  &     (0.004)      &     (0.004)      &                  &                  \\
[1em]
Ln Bil NER Volatility $\times$ Ln Nb Dest$_{t-1}$          &                  &      -0.002      &      -0.002      &      -0.001      &                  &      -0.003      &      -0.001      &      -0.002{$^a$}\\
                    &                  &     (0.002)      &     (0.002)      &     (0.001)      &                  &     (0.003)      &     (0.002)      &     (0.001)      \\
\hline
Observations        &     2672139      &     2672139      &     2672139      &     2672139      &     4819302      &     4819302      &     4819302      &     4819302      \\
\(R^{2}\)           &       0.181      &       0.181      &       0.183      &       0.402      &       0.198      &       0.198      &       0.207      &       0.440      \\

\hline
%Firm-year dyads &     370522      &     370522      &     370522      & 554350& 554350&554350     \\
Firm-year FE  & X & X & X & X & X & X & X & X \\
Country FE  & X & X &  &  & X & X &  &  \\
Country-year FE &  &  & X & X &  &  & X & X \\
Firm-country FE &  &  &  & X &  &  &  & X \\
\hline\hline
\end{tabular}
\begin{tablenotes}
\small
\item Note: In this table, exchange-rate volatility is computed as follows. We compute the standard deviation of monthly log deviation of \emph{nominal} exchange rates, instead of the real exchange-rate levels as we did within the paper. We exclude Euro Area observations because they exhibit zero NER volatility after 1999, which may generate a bias in the estimation. Robust standard errors are clustered by destination-year in parentheses with {$^a$}, {$^b$} and {$^c$} respectively denoting significance at the 1\%, 5\% and 10\% levels.
\end{tablenotes}
\end{threeparttable}
\end{table}

\end{landscape}

%
%
%XXXXXX
%
%
%			SECTION 2 / ALTERNATIVE MEASURES OF FIRM PERFORMANCE
%			
%			
%					
%XXXXXX
%
%
%
%

\subsection{Alternative Measures of Firm Performance \label{secoa:firmperformance}}
\setcounter{table}{0}
\renewcommand{\thetable}{B.\arabic{table}}

\subsubsection{Additional Measures of Firm Performance}

\end{document}

### 2017-07-06 redaction/online_appendix_dhp2017.tex
OA_060717

@@
 
-\subsection{Additive-costs only Model \label{oa_sec:additive_only}}
+\subsection{Additive-costs only Model \label{secoa:additive_only}}
 
@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{With only Additive Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
-    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
-    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
-    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
-    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
-    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{\textbf{With Additive \& Ad-Valorem Transport Costs} } \\ \hline
-   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
-    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
-    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
-    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
-    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
-    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}& & & &   \\ \hline
-    Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
-    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
-    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
-    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
-    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+    \multicolumn{5}{l}{\textbf{With only Additive Transport Costs} ($\widehat{t}^{add}$, in \%)}  \\ \hline
+    Mean  & 9.2 & 5.1 & 8.0 & 5.8 \\
+    Median & 5.7 & 2.4 & 5.5 & 2.5 \\ \hline
+
        \multicolumn{5}{l}{\textbf{Data} ($p/\widetilde{p}$, in \%) } \\ \hline
@@
 
-\begin{landscape}
-\begin{table}[htbp]\centering
- \begin{threeparttable}
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-\caption{\label{NER_ext}Alternative Measure of Exchange-rate Volatility - Nominal Exchange Rate (NER) Volatility - Extensive Margin}
-\begin{tabular}{l*{8}{c}}
-\hline\hline
-               Dep. Variable &\multicolumn{4}{c}{Entry}&\multicolumn{4}{c}{Participation}\\
-                                &\multicolumn{4}{c}{$Pr(X_{ijt}>0\mid X_{ijt-1}=0)$}&\multicolumn{4}{c}{$Pr(X_{ijt}>0$)}\\
-                 \cmidrule(lr){2-5}
-  \cmidrule(lr){6-9} \\
-                    &\multicolumn{1}{c}{(1)}&\multicolumn{1}{c}{(2)}&\multicolumn{1}{c}{(3)}&\multicolumn{1}{c}{(4)}&\multicolumn{1}{c}{(5)}&\multicolumn{1}{c}{(6)}&\multicolumn{1}{c}{(7)}&\multicolumn{1}{c}{(8)}\\
-
-\hline
-
-Ln Bil NER Volatility                 &      -0.002      &       0.006      &                  &                  &       0.004      &       0.013      &                  &                  \\
-                    &     (0.002)      &     (0.005)      &                  &                  &     (0.003)      &     (0.011)      &                  &                  \\
-[1em]
-Ln GDP              &       0.133{$^a$}&       0.133{$^a$}&                  &                  &       0.149{$^a$}&       0.148{$^a$}&                  &                  \\
-                    &     (0.011)      &     (0.011)      &                  &                  &     (0.019)      &     (0.019)      &                  &                  \\
-[1em]
-Ln country price index&       0.012{$^a$}&       0.012{$^a$}&                  &                  &       0.021{$^a$}&       0.021{$^a$}&                  &                  \\
-                    &     (0.002)      &     (0.002)      &                  &                  &     (0.004)      &     (0.004)      &                  &                  \\
-[1em]
-Ln Bil NER Volatility $\times$ Ln Nb Dest$_{t-1}$          &                  &      -0.002      &      -0.002      &      -0.001      &                  &      -0.003      &      -0.001      &      -0.002{$^a$}\\
-                    &                  &     (0.002)      &     (0.002)      &     (0.001)      &                  &     (0.003)      &     (0.002)      &     (0.001)      \\
-\hline
-Observations        &     2672139      &     2672139      &     2672139      &     2672139      &     4819302      &     4819302      &     4819302      &     4819302      \\
-\(R^{2}\)           &       0.181      &       0.181      &       0.183      &       0.402      &       0.198      &       0.198      &       0.207      &       0.440      \\
-
-\hline
-%Firm-year dyads &     370522      &     370522      &     370522      & 554350& 554350&554350     \\
-Firm-year FE  & X & X & X & X & X & X & X & X \\
-Country FE  & X & X &  &  & X & X &  &  \\
-Country-year FE &  &  & X & X &  &  & X & X \\
-Firm-country FE &  &  &  & X &  &  &  & X \\
-\hline\hline
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Note: In this table, exchange-rate volatility is computed as follows.
-We compute the standard deviation of monthly log deviation of \emph{nominal} exchange rates, instead of the real exchange-rate levels as we did within the paper.
-We exclude Euro Area observations because they exhibit zero NER volatility after 1999, which may generate a bias in the estimation.
-Robust standard errors are clustered by destination-year in parentheses with {$^a$}, {$^b$} and {$^c$} respectively denoting significance at the 1\%, 5\% and 10\% levels.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\end{landscape}
-
-%
 %
@@
 
-\subsection{Alternative Measures of Firm Performance \label{secoa:firmperformance}}
+\subsection{Detailed Results \label{secoa:detailed}}
 \setcounter{table}{0}
@@
 
-\subsubsection{Additional Measures of Firm Performance}
+\subsubsection{3-Digits Level}
+
+\subsubsection{4-Digits Level}
 

### 2017-07-12 redaction/dhp_july2017.tex
changement rédaction

@@
 Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad-valorem tax equivalent.
-However, many common empirical facts support the existence of additive costs.
-This paper provides a quantitative assessment of the size and the importance of the additive component in international transport costs, measured as the cif-fab price gap, using US imports flows over 1974-2013.
-We thus find that additive transport costs are not negligible, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
+However, many common empirical facts suggest the existence of additive costs.
+This paper provides new empirical results that support this view.
+To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and ocean).
+Two main results emerge.
+First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
 Further, taking additive costs into account significantly improves the fit of the modelling of transport costs.
-We also use the time dimension of our data to characterize the patterns of transport costs over time.
-After correcting for composition effects, we find that international transport costs have decreased by 50\% in air transport and 60\% in maritime transport over the period, and of same order of magnitude for both additive and iceberg components.
+Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
+We thus show that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.
 
-\textbf{pas tres sexy la fin. Dire tout de meme que ca reste non negligeable? Avec la valeur moyenne sur 2010-2013 par ex, pour additif et pour multiplicative.}
 \end{abstract}
@@
 In particular, they are considered a major obstacle to international economic integration and international trade flows.
-According to \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
-After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements.
+According to \cite{Jacks08}, the trade cost decline explains around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
+After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO as off 1995) multilateral agreements.
 Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (e.g.
-\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
+Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
@@
 Equivalently, international transportation costs would represent a 21\% markup over production costs.
-\citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3, suggesting a non-negligible impact of transport costs on trade flows.
+Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables}, by around -3, suggests a non-negligible impact of transport costs on trade flows.
 If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
@@
 
-Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis\footnote{Rigourously speaking, the ``iceberg cost'' includes the ad-valorem costs and the fact that the costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly made in the literature.}).
-However, many common empirical facts support the existence of additive costs.
+Following \citet{samuelson1954}, standard models of international trade have usually modelled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly made in the literature.} However, many common empirical facts support the existence of additive costs (i.e., cost per unit exported).
 As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
@@
 The structure (additive vs ad-valorem) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows.
-The Alchian and Allen conjecture (\citealp{alchian}), that points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
+As pointed out by the seminal paper of \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
+In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
 \citet{Lashkaripour-mimeo-2016} challenges this view.
@@
 Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
-Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international costs.
+Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international trade costs.
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fas unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
@@
 We quantitatively assess the size and the importance of the additive component in international transport costs, that we measure as the difference between the import and the export prices, based on the US Imports flows over 1974-2013.
-Therefore, we focus only on a subset of trade costs.
-Besides, in terms of transport costs \textit{per se}, our measure based on the gap between export and import prices omits the other dimension of transport costs related to the time value of goods in transit.
+Obviously, exploiting this database is not immune of limitations, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
+Our measure being based on the gap between export and import prices, it necessarily omits the other dimension of transport costs related to the time value of goods in transit.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
-However, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: they remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
+However, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
 \citet{Hummels_1999} mentions a bunch of papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
 In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
-Furthermore, our data is based on a single and reliable customs origin, and allows for a wide time and geographical coverage, while limiting the measurement errors coming from strategies inferring trade costs.
-Therefore, we strongly believe they may deliver useful and reliable insights on the structure of international trade costs.
-
+Furthermore, our data is based on a single and reliable customs origin with a wide time and geographical coverage, and it allows us extracting a direct transport cost measure thereby limiting the measurement error bias.
+Therefore, we strongly believe exploiting this database may deliver useful and reliable insights on the structure of international trade costs.
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian exports in 2004, they find that additive costs are about 14\% of the median consumer price.
-Our paper complements their findings in many respects.
-While they study trade costs in general, our database implies that we focus on international transport costs.
-Similarly to them, our various results emphasize the important role of the additive component of international transport costs.
-Further, our empirical analysis allows us to provide a quantitative measure of the levels of both the iceberg and the additive trade costs.
-Last, we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013.
+Based on Norwegian exports in 2004, they find that additive trade costs are about 14\% of the median consumer price.
+If our study covers a narrower set of trade costs focused on transport costs, we reach a similar conclusion to them, that points out the important role of the additive component of international transport costs.
+Our work differentiates from them though, in two main aspects.
+First, our empirical analysis allows us to provide a quantitative measure of the levels of both the iceberg and the additive trade costs.
+Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-By exploiting the time coverage of our database, our paper is also related to the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
-We also share in common with these papers to investigate the time trends of transport costs by transport mode (i.e., air or sea).
-Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}).
+
+By exploiting the time coverage of our database, our paper also relates to a second strand of the international trade literature, that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
 \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004.
-Concerning ocean shipping, which represents the major part of world trade, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}.
+However, \cite{hummels2007} points out the importance of the composition effects in the characterization of time trends of international transport costs.
+He thus obtains a stronger decline in transportation costs between 1974 and 2004 in both air and ocean transport when the composition effects by origin country and by product have been removed.
 Our paper contributes to this debate.
-In particular, we show the importance of taking into account the additive component in characterizing the time trends of international transport costs.
+In particular, we show the importance of taking into account the additive component in characterizing the time trends of international transport costs, as it substantially modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
+\textbf{le dire autrement/ mieux}
+
 Our findings can be summarized in three main points.
@@
 To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}  This represents a valuable insight for calibrating related models.
-
-\todo{mettre une phrase pour dire qu'on confirme la litterature sur le role des additifs, mais de maniere differente.
-Hummels et Skiba sur des donnes de 1994, on va beaucoup plus loin qu'eux dans cette voie.}
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}  This represents a valuable insight for calibrating related international trade models.
 
@@
 Third, we exploit the time dimension of our database to document the patterns of transport costs over time.
-As pointed out by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time.
-In contrast to his work though, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs.
-Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.
-After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
-Further, our results point out an important bias in the time trends of transport costs when additive costs are omitted.
-While transport costs seem to exhibit a larger decrease in air than in sea transport between 1985 and 2005 when only iceberg costs are modeled, this difference vanishes when additive costs are taken into account.
+Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air transport and 60\% in maritime transport, and of same order of magnitude for both additive and iceberg components.
+Importantly, and in contrast to \cite{hummels2007}, we do not find that composition effects do make a substantial difference in the time trends of international transport costs.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
+
+\textbf{paragraphe precedent a reprendre}
 
@@
 Section \ref{sec:conclu} concludes.
+
+\textbf{stop here}
 

### 2017-07-12 redaction/dhp_july2017.tex
Changement comparaison Hummels

@@
 Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
-In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
+In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.\smallskip
 
@@
 However, \cite{hummels2007} points out the importance of the composition effects in the characterization of time trends of international transport costs.
-He thus obtains a stronger decline in transportation costs between 1974 and 2004 in both air and ocean transport when the composition effects by origin country and by product have been removed.
-Our paper contributes to this debate.
-In particular, we show the importance of taking into account the additive component in characterizing the time trends of international transport costs, as it substantially modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-\textbf{le dire autrement/ mieux}
+If we share this view with \cite{hummels2007}, we reach a different result though.
+In contrast to \cite{hummels2007},\footnote{Over 1974-2004, \cite{hummels2007} obtains a stronger decline in transportation costs when the composition effects by origin country and by product have been removed, in both air and ocean transports.} our results tend to downplay the role of trade composition effects in accounting for the decline of international transport costs observed since 1974.
+Importantly, this can be attributed to the role of the additive component, which modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
+In both air and ocean transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time.
 
@@
 Third, we exploit the time dimension of our database to document the patterns of transport costs over time.
-Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air transport and 60\% in maritime transport, and of same order of magnitude for both additive and iceberg components.
-Importantly, and in contrast to \cite{hummels2007}, we do not find that composition effects do make a substantial difference in the time trends of international transport costs.
+Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in maritime shipping, and of same order of magnitude for both additive and iceberg components.
+Importantly, we find that most of this decline can be attributed to a reduction in the ``pure'' transport costs.
+Investigating this result further, we show that taking into account the additive component plays a key role in assessing the role of composition effects in the time trends of international transport costs.
 These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
-
-\textbf{paragraphe precedent a reprendre}
 
@@
 \ref{sec:data_method} presents our data and our  methodology.
-Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results.
-In Section \ref{sec:results_decomposition}, we characterize the role of the additive component of transport costs.
+In Section \ref{sec:results_decomposition}, we provide results about the size and the role of the additive component of transport costs.
 After reporting the mean values over the period (by transport mode), we show how introducing additive transport costs improves the modelling of transport costs.
-Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2013.
+Section \ref{sec:results_trends} characterizes the trends of transport costs (by transport mode) over the period 1974-2013.
+Section \ref{sec:robustness} provides some robustness checks of our results.
 Section \ref{sec:conclu} concludes.
 
-\textbf{stop here}
-
 \section{Data Sources and Empirical Methodology \label{sec:data_method}}
+
+\textbf{integrer relecture Jerome sur cette section}
 
@@
 Second, we study the time trends of transport costs by disentangling the additive from the iceberg components.
-This proves to be important in the characterization of transport costs changes over time, in particular in air transport.\smallskip
+This proves to be important in the characterization of transport costs changes over time, in particular in air shipping.\smallskip
 
@@
 To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
+This has driven us to estimate this model as well.
 The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
-For sake of space saving, the results of this third model are not reported in the paper.
-Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+These results are reported in the Online Appendix, available on the authors' webpages.} \textbf{mettre les 3 modeles dans la online appendix.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
@@
 \section{Decomposing Transport Costs: The importance of the additive component \label{sec:results_decomposition}}
+
+\textbf{integrer relecture Jerome sur cette section}
 
@@
 This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects.
-Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearizations.
+Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearization.
 Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
@@
 Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%.
-These values are respectively equal to 1.8\% and 2.5\% for air transport.
+These values are respectively equal to 1.8\% and 2.5\% for air shipping.
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
@@
 \begin{figure}[htbp]
-\caption{Ad-valorem Costs (Yearly mean value, 3 digits)}
+\caption{Ad-valorem Costs (Yearly mean value, 3 digits, in \% of the fas price)}
 \label{fig:mult_alone_withadd}
@@
 Further, Figure \ref{fig:mult_alone_withadd} reports that for Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
-Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air transport.
+Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air shipping.
 To get a better picture on this, we study the shares of both ad-valorem and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.
@@
 \begin{figure}[htbp]
-\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
+\caption{Decomposing Transport costs (Yearly mean value, 3 digits, in \% of the fas price)}
 \label{fig:decomp_TC_3d}
@@
 The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996.
-For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air transport, and from 9.8\% to 4.8\% for ocean transport.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport.
+For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport.
 \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on freight rates.} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
 But the fact that trends are very close are comforting.
-It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air transport.
+It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air shipping.
 
@@
 
-\section{Decomposing Transport Costs: Characterizing the trends \label{sec:results_trends}}
-
+\section{Decomposing Transport Costs: Characterizing the trends }\label{sec:results_trends}
+\subsection{Estimation strategy}
 In this section, we investigate the role of the additive component in total transport costs in a historical perspective.
 As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time by transport mode.
-Let us consider first the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}).
-Two main comments can be made.
-First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
+Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and ocean shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
 However, computing the time trends from 1980 does not dramatically change the picture.
 The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} Second, the decrease appears more pronounced for sea than air transportation.
-These results are consistent with \cite{hummels2007}.
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.}
 
@@
 
-If we share this view with \cite{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects.
-First, we provide a broader characterization of the time trends in transport costs, by identifying the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for the two additive and ad-valorem dimensions.
-Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba},
- assumes a constant share of additive transport costs over time, product and country partner. In \cite{hummels_skiba},
- this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension).
-By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account.
-By contrast, in its decomposition exercise to characterize the time trend in transport costs, \cite{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \cite{hummels2007}, p.
-146.
-} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the two additive and ad-valorem measures of international transport costs (Equation (\ref{eq:estimatedequation})), that vary over time, product and origin country.
+Starting from these values, we extract the ``pure'' transport cost component (additive and multiplicative, by transport mode) by the mean of a time fixed effect (see details below).
+By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
 We proceed similarly for the multiplicative component of trade costs.
 We also agglomerate the two components (additive and iceberg) in a unified measure of transport costs.
-This
-allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
-
-\todo{Est-ce qu'on peut faire ça sans présenter l'équation d'Hummels pour de vrai ?}
-
-Our second difference with \citet{hummels2007} relates to the way we characterize the evolution of the ``pure'' transport costs over time.
-\citet{hummels2007} captures the trend evolution of by the yearly dummy.
-Rather, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
-This notably enables us to obtain measures of transport costs that are easily comparable between transport modes and transport costs components.
-
-To do this, we estimate the following equations for the estimated ad-valorem component using OLS:
+This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
+
+Doing so, our empirical analysis differentiates from from \cite{hummels2007} in two main dimensions.
+First, our characterization of the time trends in transport costs identifies the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for both the additive and ad-valorem costs, as well as for the overall transport cost.
+Second, our methodology assumes a share between the additive and the ad-valorem components of transport costs, that varies over the three time/product/country partner dimensions.
+By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{Rigourously speaking, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``pure'' transport costs over time.
+We discuss this point with more details in Appendix \ref{app:comp-effects}.}
+
+\subsection{Empirical specification}
+For the estimated ad-valorem component, we estimate the following equation using OLS:
 \begin{eqnarray}
@@
 
+\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.
 We cannot use the same equation for the estimated additive component, as by construction, the sector fixed effect and the country fixed effect are additive rather than multiplicative.
@@
 As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: the country dimension (Term (a)), the product dimension (Term (b)) and the ``pure transport costs time trend'' (Term (c)).
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.} Both equations are estimated using by transport mode.
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.} Both equations are estimated by transport mode.
 
@@
 From the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
-\begin{equation*}
-\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
-
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean transport cost or the mean multiplicative transport cost in 1974.
+\begin{equation}
+\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
+\end{equation}
+
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean multiplicative transport cost in 1974.
+In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
+This notably enables us to obtain measures of transport costs that are easily comparable between transport modes and transport costs components.
 
 As for the additive cost, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
-$$\Gamma^{add}_t = 100.\exp(\gamma_t)$$
+\begin{equation}
+\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
+\end{equation}
 
@@
 
-Figure \ref{fig:totalTC_compeffects_excl} reports the results.
-In Panels (a) and (b), we report the time changes of the ad-valorem costs and the additive costs respectively.
+\subsection{Characterizing the time trends in transport costs}
+Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In Panels (a) and (b), we report the time changes of the ad-valorem costs and the additive costs respectively.
 Panel (c) displays the time changes in the total transport costs (built as the sum of the two components as detailed above).
@@
 
-We also proceed to the same analysis at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.\footnote{See more details in Appendix \ref{app:comp-effects}.
-We also run the same exercise at the 1-digit level.
-For sake of conciseness, these results are not reported.
-They are available upon request to the authors.} The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects), Manufacturing}
-\label{fig:totalTC_compeffects_excl_manuf}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_manuf.pdf}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects), Primary goods}
-\label{fig:totalTC_compeffects_excl_primary}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_primary.pdf}
-\end{center}
-\end{figure}
-
-Consider first the time trends in transport costs at the aggregate level (Figure \ref{fig:totalTC_compeffects_excl}).
-Three main results emerge.
-First, we do not find evidence of a significant difference between the raw transport costs and the ``pure'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
+Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
+First, echoing the related literature (see \citealp{Lafourcade_Thisse}), we find that international transport costs hae substantially decreased over the period, and in both transport modes.
+International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
+Second, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
+Third, we do not find evidence of a significant difference between the raw transport costs and the ``pure'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``pure'' transport costs.
-This stands in contrast with \citet{hummels2007}, which obtained that the ``pure'' transport costs had decreased more (over 1973- 2004) than the unfitted ones, suggesting an important role to composition effects.
-As noted above, this difference of results can be explained by the difference in methodology.
-Whereas \citet{hummels2007} attributes all changes that are not in the ad-valorem component to composition effects, we show that part (and as a matter of fact, most) of these changes do come from a reduction in the ``pure'' transport costs, in their additive dimension.
-
-Second, composition effects are more pronounced in ocean transport.
+Composition effects are more pronounced in ocean transport.
 Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects.
 Precisely, the role of composition effects is significant for the multiplicative component, whereas the reduction in the additive costs is mostly attributable to changes in ``pure'' additive costs.
-In line with \citet{hummels2007}, these results confirm the important role of composition effects in maritime transport costs.
-The direction of the effects differs though.
-Whereas  \citet{hummels2007} obtains a more pronounced reduction in the ``pure'' transport costs than the unfitted series, we obtain the opposite result.
-On top of the reduction of the ``pure'' transport costs, overall maritime transport costs have reduced over time because of the composition of flows, \textbf{one potential explanation being the reduction in the share of bulky goods linked to the declining share of primary goods in total trade.
-Aller plus loin?
-mettre la part du primary dans le temps?
-Pertinence de cette hypothese?}
-
-\begin{figure}[htbp]
-\caption{Share of primary goods in the value of total US imports}
-\label{fig:Share_prim_goods}
-\begin{center}
-\includegraphics[height=4in]
-{Share_of_primary.pdf}
-\end{center}
-\end{figure}
-
-\section{Robustness checks}
+
+This last set of results stands in sharp contrast with \citet{hummels2007}, which obtained that the ``pure'' transport costs had decreased more than the unfitted ones (over 1973-2004), suggesting an important role to composition effects.
+If anything, we find the opposite result here.
+Investigating this difference of result further (see Appendix \ref{app:comp-effects}), we find that this can be explained by the difference in methodology, and primarily the treatment of the additive component transport costs.
+Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
+In both air and ocean transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in ocean transport.
+
+\section{Robustness checks \label{sec:robustness}}
 
@@
 Third, we also use the time dimension of our data to characterize the evolution of transport costs.
-After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013.
-Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem.
-While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s.
-Further, the inclusion of additive costs yields to the conclusion the decrease of overall trade costs are decreased in air transport and ocean shipping roughly the same path and the same magnitude of order.
-This last result stands in contrast with related studies (\citealp{hummels2007}, \citealp{Behar_Venables}).
-In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+[...] \textbf{completer} In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
 
@@
 
-\subsection{Methodolgy}
+\subsection{More on our methodology}
 \paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component. Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
@@
 Next step is to build the index $\Gamma^{adv}_t$ such that:
-\begin{equation}
+\begin{equation*}
  \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}  \label{eq:tcadv_compoeffect}
-\end{equation}
-\noindent with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
+\end{equation*}
+\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
 
@@
 
-Noticing that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
-\begin{equation}
-\Gamma^{add}_t  = 100\exp(\gamma_t)  \label{eq:tcadd_compoeffect}
-\end{equation}
+\noindent Noticing that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
+\begin{equation*}
+\Gamma^{add}_t  = 100\exp(\gamma_t)  
+\end{equation*}
+
+that is, Equation (\ref{eq:comp_effects_add}).
 
@@
 We apply a slightly different procedure for constructing the fitted measure of total transport cost (i.e., composition effect excluded).
-In this case, we start directly from the two indices previously obtained (Equations (\ref{eq:tcadv_compoeffect}) and (\ref{eq:tcadd_compoeffect}) (by transport mode).
+In this case, we start directly from the two indices previously obtained (Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) (by transport mode).
 We build the ``overall'' transport cost measure as a weighted sum of the two indices, using the relative weight of the additive cost component in total cost in 1974, $\omega_{air}= 0.423$ and $\omega_{ocean}= 0.482$ (obtained from our estimates in Section \ref{sec:results_decomposition}, see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), according to:
@@
 
-\subsection{At the sector level}
-
-\todo{expliquer comment la repartition entre all, primary, manuf a ete faite.}
+\subsection{Comparing with \cite{hummels2007}}
+
+In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of excluding composition effects in characterizing the time trends of international transport costs.
+In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' stata codes provided on his webpage), so as to better identify the precise points of difference between our estimation strategies.
+
+Let us start from Hummels's quotation (p.
+146) according to which (for air), the ``unadjusted measure of ad valorem air shipping costs [is] the aggregate expenditures on air shipping divided by the value of airborne imports''.
+From this, we get the ``raw'' ad-valorem measure of transport costs, as the ratio between the imported value and the exported value ($(p^{cif}_{ikt} - p^{cif}_{ikt})q_{ikt}/p^{fas}_{ikt}q_{ikt})$, the mean yearly value being obtained by weighting each flow by its value of total trade flows (ie, yielding the observed aggregate value of $\tau_t$ (for air).
+
+Consider now the ``fitted ad-valorem rate''.
+\cite{hummels2007} uses ``a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels' (2007) terminology to $i$ (our terminology).} at time $t$.
+The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables''.
+With the value of the shipment equal to $p^{fas}_{ikt}q_{ikt}$, and denoting the air freight cost as $TC_{ikt}$, we can formulate the above sentence according to the following equation (suppressing the transport mode index to alleviate notation):
+
+\begin{eqnarray}
+&&\ln TC_{ikt} = \delta+ \beta \ln \frac{q_{ikt}}{p^{fas}_{ikt}q_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \notag \\
+\Leftrightarrow && \ln TC_{ikt} = \delta+ \beta \ln \frac{1}{p^{fas}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \label{eq:Hummels1}
+\end{eqnarray}
+
+\noindent where $TC_{ikt}$ is measured by the ratio between (air) charges, i.e.
+$(p^{cif}_{ikt} - p^{fas}_{ikt})q^{fas}_{ikt}$ and the value of the shipment, i.e.
+$p^{fas}_{ikt}q^{fas}_{ikt}$, or equivalently : $TC_{ikt} = \frac{p^{cif}_{ikt} - p^{fas}_{ikt}}{p^{fas}_{ikt}}$.
+
+After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} takes the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, from which he obtains the unweighed average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$ to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).
+
+What are the main differences with our estimation strategy?
+Five differences may be underlined.
+First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
+However, this is a slight difference, as in both cases it ultimately amounts taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
+A second minor difference is that we separate the country-product fixed effects in two components (origin country and product dimensions).
+As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
+However, we do not view this as the major cause of difference between our tow methods, as it can be inferred from the robustness analysis driven on this point in Section \ref{sec:robustness}.
+Third, in contrast to us, \cite{hummels2007} does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effects, in contrast to us.
+This is not likely to make a substantial difference though, as these fixed effects are by nature constant over time.
+Accordingly, this only plays as a scale effect.
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of Hummels' (2007) method, the switch from $\widehat{\tau}_{ikt}$ to $\widehat{\tau}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add} in our case).
+\cite{hummels2007} taking the unweighed average value over the $i,k$ dimension implicitly means that he attributes a weight equal to 1 to each flow.
+We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974, so as to not overweight the small flows in value.
+Provided that the weighting scheme in value is constant over time, this difference of treatment should only play as a scale effect.
+However, we ensure that our time trend results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}.
+
+\textbf{si on devait le mettre en equation? Results are reported? Not reported? Online appendix probablement la meilleure solution non?}
+
+One last difference remains to be commented, which in fact turns our to be key in accounting for the difference of results about the importance (or not) of composition effects between \cite{hummels2007} and us.
+This relies on the way the additive component of international transport costs.
+Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
+To see it clearly, rewrite Equation (\ref{eq:Hummels1}) in a much simpler form (getting rid of fixed effects for clarity) as
+
+$$\ln TC_{ikt} \equiv \ln \left(\frac{p^{cif}_{ikt}- p^{fas}_{ikt}}{p^{fas}_{ikt}} \right) = \Delta_{ikt}- \beta \ln p^{fas}_{ikt} $$
+
+Under $\beta = 1$, it becomes:
+\begin{eqnarray*}
+&&\ln \left(\frac{p^{cif}_{ikt}- p^{fas}_{ikt}}{p^{fas}_{ikt}} \right) = \Delta_{ikt}- \ln p^{fas}_{ikt} \\
+\leftrightarrow && \ln (p^{cif}_{ikt}- p^{fas}_{ikt}) = \Delta_{ikt} \\
+\end{eqnarray*}
+\noindent Equivalently, we can write:
+
+\begin{eqnarray*}
+&&p^{cif}_{ikt} = p^{fas}_{ikt} + exp(\Delta_{ikt})\\
+\Leftrightarrow && p^{cif}_{ikt} = p^{fas}_{ikt} + t_{ikt}
+\end{eqnarray*}
+\noindent with $t_{ikt}$ appropriately defined. This exactly corresponds to the case of additive costs only. Conversely, in the case $\beta = 0$, Equation (\ref{eq:Hummels1}) becomes:
+\begin{eqnarray*}
+&&\ln \left(\frac{p^{cif}_{ikt}- p^{fas}_{ikt}}{p^{fas}_{ikt}} \right) = \Delta_{ikt} \\
+\leftrightarrow && \ln (\frac{p^{cif}_{ikt}}{p^{fas}_{ikt}}-1) = \Delta_{ikt}\\
+\leftrightarrow && \frac{p^{cif}_{ikt}}{p^{fas}_{ikt}}  = exp \Delta_{ikt}+1
+\end{eqnarray*}
+\noindent that is,
+
+$$p^{cif}_{ikt} = \tau_{ikt} p^{fas}_{ikt}$$
+
+\noindent with $\tau_{ikt}$ appropriately defined.
+This exactly corresponds to the case of ad-valorem transport costs only.
+As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, which therefore measures the relative eight of the additive component in intrenational transport costs.
+By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
+This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (ie, somehow allowing for a varying $\beta$ across time, country partner and product).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007} and developed above, on our database (which is the same as his until 2004).
+The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$), as in \cite{hummels2007}.
+
+\begin{figure}[htbp]
+\caption{Characterizing the time trends; Applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+\textbf{enlever figure 5 et figure 6}.
+
+The results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
+Applying Hummels' (2007) method, we obtain that the composition effects tend to mitigate the decrease in transport costs in both air and ocean shipping, in line with his results.
+This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
+Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
+
+\subsection{At a more disaggregated level}
+
+\textbf{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
+
+In this Section, we characterize the time trend in international transport costs at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.
+The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
+
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects), Manufacturing}
+\label{fig:totalTC_compeffects_excl_manuf}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_manuf.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects), Primary goods}
+\label{fig:totalTC_compeffects_excl_primary}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_primary.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Share of primary goods in the value of total US imports}
+\label{fig:Share_prim_goods}
+\begin{center}
+\includegraphics[height=4in]
+{Share_of_primary.pdf}
+\end{center}
+\end{figure}
+
+\textbf{A commenter. A mettre en Online Appendix?}
 

### 2017-07-13 redaction/dhp_july2017.tex
Chnagement rédaction (bis)

@@
 \begin{abstract}
-Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad-valorem tax equivalent.
+Following \cite{samuelson1954}, standard models of international trade have usually modelled trade costs as an ad-valorem tax equivalent.
 However, many common empirical facts suggest the existence of additive costs.
@@
 
-\textbf{integrer relecture Jerome sur cette section}
-
 \subsection{A measure of Transportation Costs}
@@
 
-It is worth acknowledging that using this dataset encounters some limitations.
+Obviously, this dataset is not immune of some limitations.
 First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
@@
 
-Yet, we believe that our analysis delivers valuable insights on this topic.
-Indeed, using this dataset has (at least) three main advantages.
-First, this dataset delivers a strong statistical reliability arising from a single, trustworthy, customs origin.
+Yet, this dataset also has (at least) three main advantages.
+First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs origin.
 Based on customs declarations, the US Imports database inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
-In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode (ocean (or ``vessel'') and air).
-The first will be crucial to compute transport costs (see below), the second will allow to enlighten substantial differences in the dynamics of transport costs across transportation mode.
-Second, using this dataset allows us to have the import price of the good, next to the export price.
-This is highly valuable, as we can estimate both the \textit{levels} of the iceberg costs and of the additive costs.
-This strongly differentiates us from \citet{Irrazabal_2015}, who can only estimate the ratio of additive costs as a share of the median total consumer price.
+This will be crucial to compute transport costs, as we detail below.
+In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air.
+We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode.
+Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country).
+This is highly valuable, as this gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs.
 Third, this dataset is available over a long time span.
@@
 First, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004.
-As we show in Section \ref{sec:results_trends}, covering the time period over the recent years delivers interesting insights regarding the trends in air shipping costs.
-Second, we study the time trends of transport costs by disentangling the additive from the iceberg components.
-This proves to be important in the characterization of transport costs changes over time, in particular in air shipping.\smallskip
+Second, and most importantly, we study the time trends of transport costs by disentangling the additive from the iceberg components.
+This proves to be key in the characterization of transport costs changes over time, in particular regarding the role of the composition effects.\smallskip
 
@@
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-We thus obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
+However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
 Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.
@@
 Let us denote $i$ the origin country, and $k$, the product.
-Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}
+Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
 
@@
 
-\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.
+\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
 In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
@@
 Put it differently, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
-same structure of costs.\footnote{We ensure the robustness of our estimation results to this assumption in two ways.
-First, we conduct the estimations at the 4-digit level for some selected years.
-As reported in Appendix \ref{app:4digit}, the results are not substantially affected.
-Second, we provide a decomposition variance exercise on the observed cif-fas price.
+same structure of costs.\footnote{We provide a decomposition variance exercise on the observed cif-fas price.
 As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
@@
 One key objective of the paper is to characterize the importance of additive costs relatively to ad-valorem costs.
-Put it differently, what traditional models of international trade lose by ignoring additive costs?
-A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining additive costs $t$ to be equal to zero, and compare the fitting properties and the explanatory power of the models.
-For each year and transport mode, we thus estimate the two models: \textit{(a)} when transport costs are modeled as iceberg costs (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are modeled as decomposed in the two dimensions of additive and iceberg (Equation (\ref{eq:model_IetA})):
+A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining additive costs to be equal to zero, and compare the fitting properties and the explanatory power of the models.
+For each year and transport mode, we thus estimate the two models: \textit{(a)} when transport costs are modeled as iceberg costs (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are modeled as decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})):\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
+This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
+The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
+These results are reported in the Online Appendix, available on the authors' webpages.}
 
@@
 
-After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode.
+After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
 When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this model as well.
-The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
-These results are reported in the Online Appendix, available on the authors' webpages.} \textbf{mettre les 3 modeles dans la online appendix.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} \textbf{mettre les 3 modeles dans la online appendix.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
@@
 \section{Decomposing Transport Costs: The importance of the additive component \label{sec:results_decomposition}}
-
-\textbf{integrer relecture Jerome sur cette section}
 
@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
+    \multicolumn{5}{l}{\textbf{Model (A) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
     Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
@@
     %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{\textbf{With Additive \& Ad-Valorem Transport Costs} } \\ \hline
+    \multicolumn{5}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
    \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
@@
 Starting with pure statistical concerns, it appears that mean values are systematically higher than medians - by around one percentage point.
-This is no surprise at all: once again, our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median.
-Besides, one can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point.
-This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects.
+This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median.
+One can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point.
+This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - Section \ref{sec:results_trends} will go into more details of these compositions effects.
 Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearization.
 Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
-Descriptive statistics showing that it is indeed the case available upon request to the authors.
+Descriptive statistics showing that it is indeed the case are available upon request to the authors.
 
 Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable.
-Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping.
+When modeled as ad-valorem costs only (Model (A)), they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping over 1974-2013.
 Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
-Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%.
+Over the whole period, additive costs are 2.9\% of the export price for ocean shipping - with a lower estimated value for the ad-valorem freight rate of 3.2\%.
 These values are respectively equal to 1.8\% and 2.5\% for air shipping.
@@
 
-We extend further our analysis by using explicitly the time dimension of our analysis in order to assess the dynamics of our estimates over time.
+We extend further our analysis by using explicitly the time dimension of our dataset in order to assess the dynamics of our estimates over time.
 Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_IetA}), by transport mode.
@@
 
-From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds on a yearly basis.
-Further, Figure \ref{fig:mult_alone_withadd} reports that for Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
+From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds across years.
+Further, Figure \ref{fig:mult_alone_withadd} reports that for air shipping, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
 Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air shipping.
@@
 The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996.
-For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on freight rates.} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
-But the fact that trends are very close are comforting.
-It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air shipping.
+For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for ocean transport.
+\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on freight rates.} The fact that our estimates are on average higher than those by \citet{Hummels_1999} can be attributed to ours being simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
+But the fact that trends are very close is comforting.
 
@@
 As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
-We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs.
-We come back to this aspect in Section \ref{sec:results_trends}.}
+We go on investigating the role of the additive dimension of international transport costs further in the next subsection.
 
@@
 
-For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in in Appendix \ref{app:robust_4d}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
+For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in in Appendix \ref{app:4digit}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
 If anything, the quality of fitting appears slightly higher when estimations are based on the 4-digit classification.

### 2017-07-17 redaction/dhp_july2017.tex
Changement texte

@@
 Following \cite{samuelson1954}, standard models of international trade have usually modelled trade costs as an ad-valorem tax equivalent.
-However, many common empirical facts suggest the existence of additive costs.
+However, many common empirical facts point out the importance of additive costs.
 This paper provides new empirical results that support this view.
 To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and ocean).
-Two main results emerge.
+Three main results emerge.
 First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
-Further, taking additive costs into account significantly improves the fit of the modelling of transport costs.
-Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-We thus show that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
-These results therefore confirm the importance of the additive component in accounting for international transport costs.
+Second, taking additive costs into account significantly improves the fit of the modelling of transport costs.
+Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
+Allowing for a share of additive costs varying in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
+In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
 
@@
 
-{\normalsize JEL classification: F14, N70, R40\newline
+{\normalsize JEL classification: F14, N70, R40 \\
 Keywords: Transport costs estimates, transport costs determinants, non-linear econometrics, period 1974-2013, additive transport costs }
@@
 According to \cite{Jacks08}, the trade cost decline explains around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
-After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO as off 1995) multilateral agreements.
+After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO after 1995) multilateral agreements.
 Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
@@
 
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
 In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs would represent a 21\% markup over production costs.
-Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables}, by around -3, suggests a non-negligible impact of transport costs on trade flows.
+Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows.
 If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
@@
 often exhibit (at least partly) an additive structure.
-The structure (additive vs ad-valorem) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows.
+The structure (additive vs ad-valorem) of transport costs is far from being anecdotal.
+The literature has long emphasized its role in shaping the pattern of trade flows.
 As pointed out by the seminal paper of \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
@@
 He finds supporting evidence for the \emph{ad-valorem} assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
-One can yet be concerned by the generality of this result.
+Yet, one can be concerned by the generality of this result.
 By nature, his study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports).
@@
 Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fas unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.
 \citet{hummels_skiba} estimate the elasticity of freight rates with respect to price to be well below unity.
@@
 Our paper provides an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
-In this respect, the paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.}
-We quantitatively assess the size and the importance of the additive component in international transport costs, that we measure as the difference between the import and the export prices, based on the US Imports flows over 1974-2013.
-Obviously, exploiting this database is not immune of limitations, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
-Our measure being based on the gap between export and import prices, it necessarily omits the other dimension of transport costs related to the time value of goods in transit.
+To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a yearly/product/country partner basis.
+Obviously, exploiting this database is not immune of drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
+Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs related to the time value of goods in transit.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
-However, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
+Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
 \citet{Hummels_1999} mentions a bunch of papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
 In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
-Furthermore, our data is based on a single and reliable customs origin with a wide time and geographical coverage, and it allows us extracting a direct transport cost measure thereby limiting the measurement error bias.
-Therefore, we strongly believe exploiting this database may deliver useful and reliable insights on the structure of international trade costs.
-Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian exports in 2004, they find that additive trade costs are about 14\% of the median consumer price.
-If our study covers a narrower set of trade costs focused on transport costs, we reach a similar conclusion to them, that points out the important role of the additive component of international transport costs.
-Our work differentiates from them though, in two main aspects.
-First, our empirical analysis allows us to provide a quantitative measure of the levels of both the iceberg and the additive trade costs.
+Furthermore, our data is based on a single and reliable customs origin with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure.
+This limits the measurement error bias.
+Therefore, we strongly believe that exploiting this database may deliver useful and reliable insights on the structure of international trade costs.
+
+By investigating the additive dimension of transport costs, our paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
+Based on Norwegian exports in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price.
+Our paper complements their findings in two main respects.
+First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs.
 Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
-In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.\smallskip
-
-By exploiting the time coverage of our database, our paper also relates to a second strand of the international trade literature, that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
+If driven differently, all our results share in common with \citet{Irrazabal_2015} to point out the important role of the additive component of international transport costs.\smallskip
+
+By exploiting the time coverage of our database, our paper also relates to a second strand of the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
 Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
@@
 However, \cite{hummels2007} points out the importance of the composition effects in the characterization of time trends of international transport costs.
-If we share this view with \cite{hummels2007}, we reach a different result though.
-In contrast to \cite{hummels2007},\footnote{Over 1974-2004, \cite{hummels2007} obtains a stronger decline in transportation costs when the composition effects by origin country and by product have been removed, in both air and ocean transports.} our results tend to downplay the role of trade composition effects in accounting for the decline of international transport costs observed since 1974.
-Importantly, this can be attributed to the role of the additive component, which modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and ocean transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time.
+Over 1974-2004, \cite{hummels2007} thus obtains a stronger decline in transportation costs when the composition effects by origin country and by product have been removed, in both air and ocean transports.
+Our results tend to mitigate this view though.
+In contrast to \cite{hummels2007}, we do not find that composition effects have played against the decrease of the ``pure'' transport costs over time.
+They do not matter much in accounting for the decline of transport costs in air shipping; they even seem to have amplified the decrease in the pure transport costs in ocean shipping.
+Importantly, we show that the difference of results can be attributed to the role of the additive component, which modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
+In both air and ocean transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time.\medskip
 
@@
 Investigating this result further, we show that taking into account the additive component plays a key role in assessing the role of composition effects in the time trends of international transport costs.
-These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip
+Allowing for additive costs that vary over time, product and country partner, we find that composition effects do not play against the decrease in the pure transport costs observed from 1974 in the time trend of overall transport costs.
+Put it differently, the reduction of the pure transport costs is not that strong once the role of additive costs is appropriately modeled.
+In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
 
@@
 
-The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996.
-For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for ocean transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on freight rates.} The fact that our estimates are on average higher than those by \citet{Hummels_1999} can be attributed to ours being simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
-But the fact that trends are very close is comforting.
-
-Besides, we go beyond by distinguishing within these evolutions between multiplicative and additive transport costs.
-First, the decreasing trend holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period.
+\textbf{je ne comprends pas les chiffres mis dans le commentaire en 1974, par rapport aux graphiques.}
+
+As reported on Figure \ref{fig:decomp_TC_3d}, total transport costs have regularly declined over time, by an order of magnitude of 3 percentage points in air shipping and 5 percentage points in ocean shipping (basically, divided by 2 in both cases).
+The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who points out that overall transport cost for the United States declined from 6 \% to 4 \% between 1974 and 1996.
+For the same years, we obtain a total decrease \textbf{from 6.9 to 4.2\% on average} for air shipping, and from \textbf{9.8\% to 4.8\%} for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.} The fact that our estimates are on average higher than those by \citet{Hummels_1999} (in ocean shipping mostly) is not necessarily surprising, given that ours being simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
+Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.
+
+Our empirical strategy allows us to go beyond these results, by distinguishing within these evolutions between multiplicative and additive transport costs.
+First, the decreasing trend holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}).
 Second, in both transport modes, the additive component appears of sizeable importance.
 As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
-We go on investigating the role of the additive dimension of international transport costs further in the next subsection.
+Further, the downward trend overall transport costs over time can be attributed to both the additive and the ad-valorem components, with a share of additive costs roughly constant over time.
+These results thus suggest that additive costs are neither negligible in magnitude nor an erratic phenomenon.
+By contrast, they represent an sizeable and structural dimension of international transport costs.
+
+\textbf{discuter du paragraphe qui suit. Je mets ma version et ensuite l'ancienne.}
+
+At this point, let us compare our results to the literature.
+First, consider our estimation of the overall size of transport costs.
+From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
+Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
+With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons.
+\citet{anderson_wincoop_jel} obtain this 11\% markup as a rough value by aggregating various sources over a bunch of industrialized countries, with data mostly based on the second half of the 1990s.
+By contrast, our estimation of a 6\% markup relies on a single source of data based on the US imports flows over 1974-2013.
+The country coverage, or the time sample can accordingly be invoked as candidate explanations.
+Over the 1990 decade, which is covered by \citet{anderson_wincoop_jel}, we obtain an estimated size of overall transport costs of approximatively 5\% in both air and ocean shipping.
+This leads us to favor the role of country coverage as main explanation, making us consider our estimate for the US as the lower bound within the range of transport costs for industrialized countries.
+This also suggests further research to investigate the question of country-heterogeneity in the magnitude of transport costs deeper.}
+
+Second, consider our results specific to the additive costs dimension, which are a key contribution of the paper.
+These can be compared to \citet{Irrazabal_2015}, which provide the most recent evidence on the size of the additive costs.
+More precisely, based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology.
+In contrast to them, we can provide separate estimates for each cost component (additive and ad-valorem), from which we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
+Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport.\footnote{On average over the whole period, from the figures in Table \ref{tab:summary_results}, we have $\frac{t}{\tau\widetilde{p}}= \frac{t\widetilde{p}}{\tau} = \frac{0.029}{1.032}= 2.8\%$ in vessel; and 1.8\% in air shipping.
+} This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}.
+However, while the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
+From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into Irrazabal et al.'s (2015) results, allows to deduce a ratio $t/\tau$ of international transport costs equal to 2.1\% ($0.15\times 0.14$), which is in line with our own results.
+This makes us confident in our quantitative assessment of the size of the additive component in international transport costs.
+In the next section, we analyze their importance through another angle based on quality-of-fit diagnostic tests.
+\bigskip
+
+\textbf{ancienne version}
 
@@
 We provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
-Smaller values are better because it indicates that the observations are closer to the fitted line.
+The smaller the SER value, the better the quality of fit, as it indicates that the observations are closer to the fitted line.
 We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test.
 A decrease in the log-likelihood function points to a better quality-of-fit.
-However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $.} The preferred model is the one with the minimum AIC value.
-Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e.
+However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.} The preferred model is the one with the minimum AIC value.
+Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including the additive term, i.e.
 Equation (\ref{eq:model_IetA}) and the Restricted one (\emph{R}, i.e.
@@
 The null tested is that the two models are statistically equivalent.
-Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level).
-The results for all years are available upon request to the authors.}\medskip
+Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.
 
@@
 If anything, the quality of fitting appears slightly higher when estimations are based on the 4-digit classification.
-This is especially true for the model restricting trade cost to their ad-valorem dimension, whatever the transport mode considered.
+This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
 When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
@@
 Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As argued by \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
+As emphasized by \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
 per product- and per partner- transport costs.
@@
 
-Doing so, our empirical analysis differentiates from from \cite{hummels2007} in two main dimensions.
+This empirical strategy differentiates from \cite{hummels2007} in two (interrelated) main dimensions.
 First, our characterization of the time trends in transport costs identifies the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for both the additive and ad-valorem costs, as well as for the overall transport cost.
-Second, our methodology assumes a share between the additive and the ad-valorem components of transport costs, that varies over the three time/product/country partner dimensions.
-By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{Rigourously speaking, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``pure'' transport costs over time.
+Second, and as direct consequence, our methodology assumes a share between the additive and the ad-valorem components of transport costs, that varies over the three time/product/country partner dimensions.
+By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+This turns out to be of primary importance in the disentangling in the time trend of transport costs, what comes from the composition effects and what comes the ``pure'' transport cost dimension, as we show below.
+\footnote{Rigourously speaking, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``pure'' transport costs over time.
 We discuss this point with more details in Appendix \ref{app:comp-effects}.}
@@
 \subsection{Empirical specification}
-For the estimated ad-valorem component, we estimate the following equation using OLS:
+Our strategy to eliminate the composition effects obeys the following rationale.
+As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in our first stage.
+We measure how these costs have changed over time by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
+
+This can be exposed as a two-stage process.
+First, we decompose the measure of transport costs component in the three product/country/time dimensions, using fixed effects.
+For the estimated ad-valorem component, we thus estimate the following equation:
 \begin{eqnarray}
@@
 \noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.
+Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
 We cannot use the same equation for the estimated additive component, as by construction, the sector fixed effect and the country fixed effect are additive rather than multiplicative.
-We estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}
+We estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
+Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
 %, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
@@
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-From the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
+This constitutes our second stage.
+As for the ad-valorem component, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
 \begin{equation}
@@
 In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
-This notably enables us to obtain measures of transport costs that are easily comparable between transport modes and transport costs components.
 
@@
 
-\noindent As a result, the three series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+\noindent As a result, the two series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
 
@@
 Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, echoing the related literature (see \citealp{Lafourcade_Thisse}), we find that international transport costs hae substantially decreased over the period, and in both transport modes.
+First, echoing the related literature (see \citealp{Lafourcade_Thisse}), we find that international transport costs have substantially decreased over the period, and in both transport modes.
 International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
@@
 First, we provide a quantitative measure of both the additive and the iceberg transport cost.
-We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%.
+We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%, as mean values over 1974-2013.
 These values are respectively equal to 1.8 and 2.5\% for air transport.
@@
 All goodness-of-fit measures point out to this conclusion, which holds for both transport modes and all years considered.
-Third, we also use the time dimension of our data to characterize the evolution of transport costs.
-[...] \textbf{completer} In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
+Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
+In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
 
 Our results could be extended in two main ways.
-On the empirical side, one may want to ge deeper in the ``structural'' determinants of trade costs, i.e.
+On the empirical side, one may want to go deeper in the ``structural'' determinants of trade costs, i.e.
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
@@
 We apply a slightly different procedure for constructing the fitted measure of total transport cost (i.e., composition effect excluded).
-In this case, we start directly from the two indices previously obtained (Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) (by transport mode).
+In this case, we start directly from the two indices previously obtained (Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}), by transport mode).
 We build the ``overall'' transport cost measure as a weighted sum of the two indices, using the relative weight of the additive cost component in total cost in 1974, $\omega_{air}= 0.423$ and $\omega_{ocean}= 0.482$ (obtained from our estimates in Section \ref{sec:results_decomposition}, see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), according to:
@@
 
-In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of excluding composition effects in characterizing the time trends of international transport costs.
-In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' stata codes provided on his webpage), so as to better identify the precise points of difference between our estimation strategies.
-
-Let us start from Hummels's quotation (p.
-146) according to which (for air), the ``unadjusted measure of ad valorem air shipping costs [is] the aggregate expenditures on air shipping divided by the value of airborne imports''.
-From this, we get the ``raw'' ad-valorem measure of transport costs, as the ratio between the imported value and the exported value ($(p^{cif}_{ikt} - p^{cif}_{ikt})q_{ikt}/p^{fas}_{ikt}q_{ikt})$, the mean yearly value being obtained by weighting each flow by its value of total trade flows (ie, yielding the observed aggregate value of $\tau_t$ (for air).
+In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
+In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' Stata codes provided on his webpage), to better identify the precise points of difference between our estimation strategies.
+
+Let us start from Hummels' (2007) quotation (p.
+146) according to which (for air), the ``\textit{unadjusted measure of ad valorem air shipping costs [is] the aggregate expenditures on air shipping divided by the value of airborne imports}''.
+From this, we get the ``raw'' ad-valorem measure of transport costs as the ratio between the imported total value and the exported total value ($(p_{ikt} - \widetilde{p}_{ikt})q_{ikt}/\widetilde{p}_{ikt}q_{ikt})$, the mean yearly value being obtained by weighting each flow by its value in total trade flows (ie, yielding the observed aggregate value of $\tau_t$ for air).
 
 Consider now the ``fitted ad-valorem rate''.
-\cite{hummels2007} uses ``a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels' (2007) terminology to $i$ (our terminology).} at time $t$.
-The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables''.
-With the value of the shipment equal to $p^{fas}_{ikt}q_{ikt}$, and denoting the air freight cost as $TC_{ikt}$, we can formulate the above sentence according to the following equation (suppressing the transport mode index to alleviate notation):
+\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels' (2007) terminology) to $i$ (our terminology).} \textit{at time $t$.
+The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables}''.
+With the value of the shipment equal to $\widetilde{p}_{ikt}q_{ikt}$, and denoting the air freight cost as $TC_{ikt}$, we can formulate the above sentence according to the following equation (suppressing the transport mode index to alleviate notation):
 
 \begin{eqnarray}
-&&\ln TC_{ikt} = \delta+ \beta \ln \frac{q_{ikt}}{p^{fas}_{ikt}q_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \notag \\
-\Leftrightarrow && \ln TC_{ikt} = \delta+ \beta \ln \frac{1}{p^{fas}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \label{eq:Hummels1}
+&&\ln TC_{ikt} = \delta+ \beta \ln \frac{q_{ikt}}{\widetilde{p}_{ikt}q_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \notag \\
+\Leftrightarrow && \ln TC_{ikt} = \delta+ \beta \ln \frac{1}{\widetilde{p}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \label{eq:Hummels1}
 \end{eqnarray}
@@
 \noindent where $TC_{ikt}$ is measured by the ratio between (air) charges, i.e.
-$(p^{cif}_{ikt} - p^{fas}_{ikt})q^{fas}_{ikt}$ and the value of the shipment, i.e.
-$p^{fas}_{ikt}q^{fas}_{ikt}$, or equivalently : $TC_{ikt} = \frac{p^{cif}_{ikt} - p^{fas}_{ikt}}{p^{fas}_{ikt}}$.
-
-After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} takes the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, from which he obtains the unweighed average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$ to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).
+$(p_{ikt} - \widetilde{p}_{ikt})q^{fas}_{ikt}$ and the value of the shipment, i.e.
+$\widetilde{p}_{ikt}q^{fas}_{ikt}$, or equivalently : $TC_{ikt} = \frac{p_{ikt} - \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}$.\medskip
+
+After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighed average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
 
 What are the main differences with our estimation strategy?
-Five differences may be underlined.
+Five points may be underlined.
 First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
 However, this is a slight difference, as in both cases it ultimately amounts taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-A second minor difference is that we separate the country-product fixed effects in two components (origin country and product dimensions).
+
+A second minor difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}$ in Equation (\ref{eq:Hummels1})) in two separate components (origin country and product dimensions).
 As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
-However, we do not view this as the major cause of difference between our tow methods, as it can be inferred from the robustness analysis driven on this point in Section \ref{sec:robustness}.
-Third, in contrast to us, \cite{hummels2007} does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effects, in contrast to us.
+However, we do not view this as the major cause of difference between our tow methods, as it can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
+
+Third, in contrast to us, \cite{hummels2007} does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effects.
 This is not likely to make a substantial difference though, as these fixed effects are by nature constant over time.
 Accordingly, this only plays as a scale effect.
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of Hummels' (2007) method, the switch from $\widehat{\tau}_{ikt}$ to $\widehat{\tau}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add} in our case).
-\cite{hummels2007} taking the unweighed average value over the $i,k$ dimension implicitly means that he attributes a weight equal to 1 to each flow.
-We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974, so as to not overweight the small flows in value.
-Provided that the weighting scheme in value is constant over time, this difference of treatment should only play as a scale effect.
-However, we ensure that our time trend results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}.
-
-\textbf{si on devait le mettre en equation? Results are reported? Not reported? Online appendix probablement la meilleure solution non?}
-
-One last difference remains to be commented, which in fact turns our to be key in accounting for the difference of results about the importance (or not) of composition effects between \cite{hummels2007} and us.
-This relies on the way the additive component of international transport costs.
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of Hummels' (2007) method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
+\cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
+This choice is made for to reasons.
+First, this weighting scheme does not overweight the small flows in value.
+Second, considering the 1974 weighting scheme makes sense as starting point of our time trend analysis.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}.
+\textbf{These results are reported in ???}.
+
+One last difference remains to be commented, which relies on the way the additive component of international transport costs is treated.
+As we show below, it turns out to be key in accounting for the difference of results with \cite{hummels2007}.
 Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
-To see it clearly, rewrite Equation (\ref{eq:Hummels1}) in a much simpler form (getting rid of fixed effects for clarity) as
-
-$$\ln TC_{ikt} \equiv \ln \left(\frac{p^{cif}_{ikt}- p^{fas}_{ikt}}{p^{fas}_{ikt}} \right) = \Delta_{ikt}- \beta \ln p^{fas}_{ikt} $$
-
-Under $\beta = 1$, it becomes:
+To see it clearly, rewrite Equation (\ref{eq:Hummels1}) in a simpler form as:
+
+$$\ln TC_{ikt} \equiv \ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt}- \beta \ln \widetilde{p}_{ikt} $$
+
+\noindent with $\Delta_{ikt} \equiv \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t}$ agglomerates the various fixed effects for reading clarity.
+
+Under the first polar case with $\beta = 1$, it becomes:
 \begin{eqnarray*}
-&&\ln \left(\frac{p^{cif}_{ikt}- p^{fas}_{ikt}}{p^{fas}_{ikt}} \right) = \Delta_{ikt}- \ln p^{fas}_{ikt} \\
-\leftrightarrow && \ln (p^{cif}_{ikt}- p^{fas}_{ikt}) = \Delta_{ikt} \\
+&&\ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt}- \ln \widetilde{p}_{ikt} \\
+\Leftrightarrow && \ln (p_{ikt}- \widetilde{p}_{ikt}) = \Delta_{ikt} \\
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-&&p^{cif}_{ikt} = p^{fas}_{ikt} + exp(\Delta_{ikt})\\
-\Leftrightarrow && p^{cif}_{ikt} = p^{fas}_{ikt} + t_{ikt}
+&&p_{ikt} = \widetilde{p}_{ikt} + \exp(\Delta_{ikt})\\
+\Leftrightarrow && p_{ikt} = \widetilde{p}_{ikt} + t_{ikt}
 \end{eqnarray*}
-\noindent with $t_{ikt}$ appropriately defined. This exactly corresponds to the case of additive costs only. Conversely, in the case $\beta = 0$, Equation (\ref{eq:Hummels1}) becomes:
+\noindent with $t_{ikt}$ appropriately defined. This exactly corresponds to the case of additive costs only.
+
+In the second polar case with $\beta = 0$, Equation (\ref{eq:Hummels1}) becomes:
 \begin{eqnarray*}
-&&\ln \left(\frac{p^{cif}_{ikt}- p^{fas}_{ikt}}{p^{fas}_{ikt}} \right) = \Delta_{ikt} \\
-\leftrightarrow && \ln (\frac{p^{cif}_{ikt}}{p^{fas}_{ikt}}-1) = \Delta_{ikt}\\
-\leftrightarrow && \frac{p^{cif}_{ikt}}{p^{fas}_{ikt}}  = exp \Delta_{ikt}+1
+&&\ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt} \\
+\Leftrightarrow && \ln \left(\frac{p_{ikt}}{\widetilde{p}_{ikt}}-1\right) = \Delta_{ikt}\\
+\Leftrightarrow && \frac{p_{ikt}}{\widetilde{p}_{ikt}}  = \exp (\Delta_{ikt})+1
 \end{eqnarray*}
@@
 
-$$p^{cif}_{ikt} = \tau_{ikt} p^{fas}_{ikt}$$
-
-\noindent with $\tau_{ikt}$ appropriately defined.
-This exactly corresponds to the case of ad-valorem transport costs only.
-As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, which therefore measures the relative eight of the additive component in intrenational transport costs.
+$$p_{ikt} = \tau_{ikt} \widetilde{p}_{ikt}$$
+
+\noindent with $\tau_{ikt}$ appropriately defined. This exactly corresponds to the case of ad-valorem transport costs only.
+
+As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
 By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
 This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (ie, somehow allowing for a varying $\beta$ across time, country partner and product).
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007} and developed above, on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$), as in \cite{hummels2007}.
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).
 
@@
 
-\textbf{enlever figure 5 et figure 6}.
+\textbf{enlever les termes figure 5 et figure 6 des figures}.
 
 The results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying Hummels' (2007) method, we obtain that the composition effects tend to mitigate the decrease in transport costs in both air and ocean shipping, in line with his results.
+Applying Hummels' (2007) method, we obtain that the composition effects tend to compensate the decrease in transport costs in both air and ocean shipping (unsurprisingly in accordance with Hummels' (2007) results, see his Figures 5 (for Air) and 6 (for Ocean)).
 This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.

### 2017-07-25 redaction/dhp_july2017.tex
modifsJH250717
Début révision texte et OA.

@@
 \usepackage[tableposition=top]{caption}
-\usepackage{ bbold }
-\usepackage{ upgreek }
+\usepackage{amsfonts}
+%\usepackage{bbold}
+% JH : impossible de compiler avec le package bbold, remplacé par amsfonts
+\usepackage{upgreek}
 \usepackage{comment}
@@
 This paper provides new empirical results that support this view.
-To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and ocean).
+To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel).
 Three main results emerge.
-First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
+First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime vessel transport respectively.
 Second, taking additive costs into account significantly improves the fit of the modelling of transport costs.
 Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a share of additive costs varying in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
+Allowing for a share of additive costs varying in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{ceteris paribus} transport costs rather than to trade pattern composition effects.
 In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
     \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
-    Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
-    Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
+    Model (A)& 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
+    Model (B)\& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
     \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
-    Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
-    Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
+     Model (A) & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
+    Model (B)\& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
    \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-    Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
-    Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
+    Model (A) & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
+    Model (B) \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
     \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
-    Term I only & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
-    Terms A \& I & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
+    Model (A) & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
+    Model (B) & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
     LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
@@
   \label{tab:good_fit_air}%
- \parbox[l]{12cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+ \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs.
+Model (B) = with additive \& Ad-Valorem transport costs.
+SER = Standard Error of regression; AIC = Akaike Information Criterion.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
@@
 \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
-Term I only & \multicolumn{1}{c}{0.450} & \multicolumn{1}{c}{0.415} & \multicolumn{1}{c}{0.456} & \multicolumn{1}{c}{0.401} & 0.350 & \multicolumn{1}{c}{0.339} & 0.39 \\
-Terms A \& I & \multicolumn{1}{c}{0.612} & \multicolumn{1}{c}{0.575} & \multicolumn{1}{c}{0.590} & \multicolumn{1}{c}{0.571} & 0.491 & \multicolumn{1}{c}{0.462} & 0.56 \\ \hline
+Model (A)& \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & 0.35 & \multicolumn{1}{c}{0.34} & 0.39 \\
+Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & 0.49 & \multicolumn{1}{c}{0.46} & 0.56 \\ \hline
 \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
-    Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
-    Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
+    Model (A) & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
+    Model (B)\& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
    \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-    Term I only & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
-    Terms A \& I & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
+   Model (A) & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
+   Model (B) & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
     \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
-    Term I only & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
-    Terms A \& I & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
+    Model (A) & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
+    Model (B) & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
     LL ratio & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
@@
   \label{tab:good_fit_vessel}%
-  \parbox[l]{12cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+  \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs.
+Model (B) = with additive \& Ad-Valorem transport costs.
+SER = Standard Error of regression; AIC = Akaike Information Criterion.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.

### 2017-07-26 redaction/dhp_july2017.tex
modifsJH2607
révision texte et OA

@@
 
-\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX.
-Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01.
-Finally, any remaining errors are ours.}}}
-%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
+\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, and participants at University of California - Irvine and Paris-Dauphine for very useful discussions.
+Finally, any remaining errors are ours.
+This paper features an online appendix containing additional results and available on the authors' websites.}}
+%We also thank the EEA 27\textsuperscript{th} Congress (Oslo), , INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and .
 \author{Guillaume \textsc{Daudin}\thanks{%
@@
 
-\noindent Trade costs are central in international economic analysis.
+\noindent Trade costs are central to international economic analysis.
 In particular, they are considered a major obstacle to international economic integration and international trade flows.
@@
 After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO after 1995) multilateral agreements.
-Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
+Based on panel data, \citet{novy13} finds  that U.S.
+trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
 Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
@@
 Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows.
-If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
+If much of trade policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
 This is accordingly the focus of the paper.\smallskip
@@
 Following \citet{samuelson1954}, standard models of international trade have usually modelled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly made in the literature.} However, many common empirical facts support the existence of additive costs (i.e., cost per unit exported).
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, many common empirical facts support the existence of additive costs (i.e., cost per unit exported).
 As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
@@
 The literature has long emphasized its role in shaping the pattern of trade flows.
-As pointed out by the seminal paper of \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
+As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
 In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
-\citet{Lashkaripour-mimeo-2016} challenges this view.
+\citet{Lashkaripour-mimeo-2017} challenges this view.
 He finds supporting evidence for the \emph{ad-valorem} assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
 Yet, one can be concerned by the generality of this result.
-By nature, his study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports).
+\citet{Lashkaripour-mimeo-2017}'s study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports).
 Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
@@
 \citet{hummels_skiba} estimate the elasticity of freight rates with respect to price to be well below unity.
-Also, their estimates imply that doubling freight costs increases average fas export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs.
+Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs.
 These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
@@
 Our paper provides an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
-To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a yearly/product/country partner basis.
-Obviously, exploiting this database is not immune of drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
+To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
+Obviously, exploiting this database has drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
 Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs related to the time value of goods in transit.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
-Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers: They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
-\citet{Hummels_1999} mentions a bunch of papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
+Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
+They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
+\citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
 In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
-Furthermore, our data is based on a single and reliable customs origin with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure.
+Furthermore, our data is based on a single and reliable customs source with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure.
 This limits the measurement error bias.
@@
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-If driven differently, all our results share in common with \citet{Irrazabal_2015} to point out the important role of the additive component of international transport costs.\smallskip
+Despite these differences, our results share with \citet{Irrazabal_2015} the important role of the additive component of international transport costs.\smallskip
 
@@
 However, \cite{hummels2007} points out the importance of the composition effects in the characterization of time trends of international transport costs.
-Over 1974-2004, \cite{hummels2007} thus obtains a stronger decline in transportation costs when the composition effects by origin country and by product have been removed, in both air and ocean transports.
-Our results tend to mitigate this view though.
-In contrast to \cite{hummels2007}, we do not find that composition effects have played against the decrease of the ``pure'' transport costs over time.
-They do not matter much in accounting for the decline of transport costs in air shipping; they even seem to have amplified the decrease in the pure transport costs in ocean shipping.
-Importantly, we show that the difference of results can be attributed to the role of the additive component, which modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and ocean transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time.\medskip
+Over 1974-2004, \cite{hummels2007} thus obtains a stronger decline in transportation costs when the composition effects by origin country and by product is removed, in both air and ocean transports.
+Our results tend to mitigate this view.
+In contrast to \cite{hummels2007}, we do not find that composition effects have offset the decrease of the ``ceteris paribus'' component of transport costs over time.
+They do not matter much in accounting for the decline of transport costs in air shipping; they even seem to have amplified the decrease in the pure transport costs in vessel shipping.
+Importantly, we show that the difference of results can be attributed to the role of the additive component, which modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time.\medskip
 
@@
 
-First, our theoretically agnostic approach provides a fairly simple framework for measuring both ad-valorem and additive parts of transportation costs.
+First, our theoretically agnostic approach provides a fairly simple (if computationally heavy) framework for measuring both ad-valorem and additive parts of transportation costs.
 We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% of the export price in air and ocean transport respectively, whereas the additive component amounts to 1.8\% and 2.9\%.
-To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
+To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
 By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}  This represents a valuable insight for calibrating related international trade models.
@@
 Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-On average over the whole period, biased estimate for ad-valorem is 5\% for air and 6\% for vessel, while unbiased estimate is respectively 2.5\% and 3.2\%.
+On average over the whole period, biased estimates for ad-valorem are 5\% for air and 6\% for vessel, while unbiased estimates are respectively 2.5\% and 3.2\%.
 We also show that taking additive costs into account significantly improves the fit of the modelling of transport costs, through various measures of ``goodness-of-fit'', for both transport modes and years.
@@
 Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in maritime shipping, and of same order of magnitude for both additive and iceberg components.
-Importantly, we find that most of this decline can be attributed to a reduction in the ``pure'' transport costs.
+Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs.
 Investigating this result further, we show that taking into account the additive component plays a key role in assessing the role of composition effects in the time trends of international transport costs.
-Allowing for additive costs that vary over time, product and country partner, we find that composition effects do not play against the decrease in the pure transport costs observed from 1974 in the time trend of overall transport costs.
+Allowing for additive costs that vary over time, product and partner country, we find that composition effects do not offset the decrease in the pure transport costs observed from 1974 in the time trend of overall transport costs.
 Put it differently, the reduction of the pure transport costs is not that strong once the role of additive costs is appropriately modeled.
@@
 
-\subsection{A measure of Transportation Costs}
+\subsection{A measure of transportation costs}
 
@@
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
-Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
+Higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
 This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
 
-Obviously, this dataset is not immune of some limitations.
+Obviously, this dataset is not immune of limitations.
 First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
@@
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
-In this respect, it is true that using this dataset embraces a partial view of international transport barriers.
-
-Yet, this dataset also has (at least) three main advantages.
-First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs origin.
-Based on customs declarations, the US Imports database inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
+In this respect, this dataset embrace only a partial view of international transport barriers.
+
+Yet, this dataset also has (at least) three advantages.
+First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source.
+Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
 This will be crucial to compute transport costs, as we detail below.
@@
 Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as this gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs.
+This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs.
 Third, this dataset is available over a long time span.
@@
 \cite{hummels2007} preceded us in investigating the time dimension of this database.
-However, our paper differentiates from his work in two main respects.
-First, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004.
+However, our paper differs from his work in two main respects.
+First, we extend the time coverage up to 2013, while \cite{hummels2007} stops in 2004.
 Second, and most importantly, we study the time trends of transport costs by disentangling the additive from the iceberg components.
@@
 
-\paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs.
+\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem and additive costs among total transport costs.
 To do so, we start from the equation that expresses the price $p$ of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}_{isk}$ (export, or fas price), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
@@
 
-\noindent As a result, our underlying theoretical equation is specified as:
+\noindent As a result, our underlying structural equation is specified as:
 
@@
 The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
-Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:
+Taking into account this constraint in the estimate suggests that the error term should be always positive and multiplicative, as in:
 
@@
 Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
-Put it differently, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
+In other words, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
 same structure of costs.\footnote{We provide a decomposition variance exercise on the observed cif-fas price.
@@
 One may be concerned that the specification of Equations (\ref{eq:model_nlI}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
-Notably based on the pricing-to-market behavior of firms (see \citealp{Krugman-87}), we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply some a direct influence between the firm's prices and the size of transport costs, except in the choice between air and ocean transport (that is beyond our scope of analysis).
+Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
+However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, as that would imply some a direct influence between the firm's prices and the size of transport costs, except in the choice between air and vessel transport (that is beyond our scope of analysis).
 Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
@@
 When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} \textbf{mettre les 3 modeles dans la online appendix.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
@@
 
+%\textbf{mettre les 3 modeles dans la online appendix.}
+
 \section{Decomposing Transport Costs: The importance of the additive component \label{sec:results_decomposition}}
@@
 First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
-Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.
+Second, we assess the importance of the additive component of in overall transport costs through goodness-of-fit measures.
 
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs  (ad-valorem estimated alone, estimated along with additive costs and the additive component), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
+It displays the mean and median values of each type of trade costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
 The first panel reports the results for a specification based on ad-valorem costs estimated alone, while the second panel presents estimates for a specification involving both ad-valorem and multiplicative component.
@@
 One can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point.
-This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - Section \ref{sec:results_trends} will go into more details of these compositions effects.
+This is first due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - Section \ref{sec:results_trends} will go into more details of these compositions effects.
 Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearization.
@@
 
-Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable.
-When modeled as ad-valorem costs only (Model (A)), they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping over 1974-2013.
+Coming now to interpretation, Table \ref{tab:summary_results} shows that the magnitude of overall transport costs is sizeable.
+When modeled as ad-valorem costs only (Model (A)), they correspond to a mean increase of the export price by 5.8\% for vessel shipping, and by 5.1\% for air shipping over 1974-2013.
 Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
@@
 From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds across years.
-Further, Figure \ref{fig:mult_alone_withadd} reports that for air shipping, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping.
+Further, Figure \ref{fig:mult_alone_withadd} reports that for air shipping, the size of the bias seems to decrease over time, while it appears rather stable for vessel shipping.
 Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air shipping.
@@
 
-\textbf{je ne comprends pas les chiffres mis dans le commentaire en 1974, par rapport aux graphiques.}
+%\textbf{je ne comprends pas les chiffres mis dans le commentaire en 1974, par rapport aux graphiques.}
+%JH: Voir Tableaux 4 et 5, années et modes de transport correspondants.
+Ils correspondent à la mesure "Iceberg alone", parce que cela semblait permettre une comparaison plus directe avec Hummels.
+Bref, à discuter.
 
@@
 The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who points out that overall transport cost for the United States declined from 6 \% to 4 \% between 1974 and 1996.
-For the same years, we obtain a total decrease \textbf{from 6.9 to 4.2\% on average} for air shipping, and from \textbf{9.8\% to 4.8\%} for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+For the same years, we obtain a total decrease from 6.9 to 4.2\% on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
 \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.} The fact that our estimates are on average higher than those by \citet{Hummels_1999} (in ocean shipping mostly) is not necessarily surprising, given that ours being simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
@@
 As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
-Further, the downward trend overall transport costs over time can be attributed to both the additive and the ad-valorem components, with a share of additive costs roughly constant over time.
+Further, the downward trend in overall transport costs over time can be attributed to both the additive and the ad-valorem components, with a share of additive costs roughly constant over time.
 These results thus suggest that additive costs are neither negligible in magnitude nor an erratic phenomenon.
@@
 
-\textbf{discuter du paragraphe qui suit. Je mets ma version et ensuite l'ancienne.}
+%\textbf{discuter du paragraphe qui suit. Je mets ma version et ensuite l'ancienne.}
+% JH: I checked Irrara et al, footnote 4 & 1er paragraphe de la section 4.4 Results page 21), il semble que ce soit Lise qui ait raison.
 
@@
 By contrast, our estimation of a 6\% markup relies on a single source of data based on the US imports flows over 1974-2013.
-The country coverage, or the time sample can accordingly be invoked as candidate explanations.
-Over the 1990 decade, which is covered by \citet{anderson_wincoop_jel}, we obtain an estimated size of overall transport costs of approximatively 5\% in both air and ocean shipping.
-This leads us to favor the role of country coverage as main explanation, making us consider our estimate for the US as the lower bound within the range of transport costs for industrialized countries.
-This also suggests further research to investigate the question of country-heterogeneity in the magnitude of transport costs deeper.}
-
-Second, consider our results specific to the additive costs dimension, which are a key contribution of the paper.
+Beyond country coverage, or the time sample, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+
+Second, consider our results specific to the additive costs dimension.
 These can be compared to \citet{Irrazabal_2015}, which provide the most recent evidence on the size of the additive costs.
@@
 However, while the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
-From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into Irrazabal et al.'s (2015) results, allows to deduce a ratio $t/\tau$ of international transport costs equal to 2.1\% ($0.15\times 0.14$), which is in line with our own results.
+From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs equal to 2.1\% ($0.15\times 0.14$), which is in line with our own results.
 This makes us confident in our quantitative assessment of the size of the additive component in international transport costs.
@@
 
-\textbf{ancienne version}
-
-Before, we check how our estimates of additive costs compare to those of \citet{Irrazabal_2015}, which provide the most recent evidence on that matter.
+%\textbf{ancienne version}
+
+%Before, we check how our estimates of additive costs compare to those of \citet{Irrazabal_2015}, which provide the most recent evidence on that matter.
 Based on Norwegian firm-level trade data for 2004, \citet{Irrazabal_2015} find that additive costs are on average 14\% of the median price.
@@
     Model (A)& 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
-    Model (B)\& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
+    Model (B) & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
     \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
      Model (A) & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
-    Model (B)\& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
+    Model (B) & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
    \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
     Model (A) & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
-    Model (B) \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
+    Model (B)  & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
     \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
@@
  \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& Ad-Valorem transport costs.
+Model (B) = with additive \& ad-valorem transport costs.
 SER = Standard Error of regression; AIC = Akaike Information Criterion.
@@
 Model (A)& \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & 0.35 & \multicolumn{1}{c}{0.34} & 0.39 \\
-Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & 0.49 & \multicolumn{1}{c}{0.46} & 0.56 \\ \hline
+  Model (B) & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & 0.49 & \multicolumn{1}{c}{0.46} & 0.56 \\ \hline
 \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
     Model (A) & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
-    Model (B)\& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
+    Model (B) & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
    \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
@@
   \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& Ad-Valorem transport costs.
+Model (B) = with additive \& ad-valorem transport costs.
 SER = Standard Error of regression; AIC = Akaike Information Criterion.
@@
 For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in in Appendix \ref{app:4digit}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
-If anything, the quality of fitting appears slightly higher when estimations are based on the 4-digit classification.
+If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.
 This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
 When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
-In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not with an extent that would disqualify the use of slightly more aggregated data.
+In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
 Further, the same conclusion established at the 3-digit level regarding the significant role of the additive components in fitting international transport costs emerges at the 4-digit level.
@@
 In this section, we investigate the role of the additive component in total transport costs in a historical perspective.
-As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time by transport mode.
-Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and ocean shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
+As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, that displays the respective shares of additive and iceberg components in transport costs over time by transport mode.
+Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
 However, computing the time trends from 1980 does not dramatically change the picture.
-The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs.
+The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
 We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.}
 
-Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner transport costs and \textit{ii)} the evolution of the composition of trade.
+Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner ``ceteris paribus'' transport costs and \textit{ii)} the evolution of the composition of trade.
 Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As emphasized by \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e.
+As emphasized by \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
 per product- and per partner- transport costs.
 
-If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the two additive and ad-valorem measures of international transport costs (Equation (\ref{eq:estimatedequation})), that vary over time, product and origin country.
-Starting from these values, we extract the ``pure'' transport cost component (additive and multiplicative, by transport mode) by the mean of a time fixed effect (see details below).
-By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time.
+If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:estimatedequation})), that vary over time, product and origin country.
+Starting from these values, we extract the ``ceteris paribus'' transport cost component (additive and multiplicative, by transport mode) by the mean of a time fixed effect (see details below).
+By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``ceteris paribus'' additive costs (for instance, handling costs) that have reduced over time.
 We proceed similarly for the multiplicative component of trade costs.
 We also agglomerate the two components (additive and iceberg) in a unified measure of transport costs.
-This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Notice that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported in another perspective (basis 100 in year 1974).}
-
-This empirical strategy differentiates from \cite{hummels2007} in two (interrelated) main dimensions.
-First, our characterization of the time trends in transport costs identifies the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for both the additive and ad-valorem costs, as well as for the overall transport cost.
-Second, and as direct consequence, our methodology assumes a share between the additive and the ad-valorem components of transport costs, that varies over the three time/product/country partner dimensions.
-By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling in the time trend of transport costs, what comes from the composition effects and what comes the ``pure'' transport cost dimension, as we show below.
-\footnote{Rigourously speaking, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``pure'' transport costs over time.
+This allows us to compare how ``ceteris paribus'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Note that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}
+
+This empirical strategy differs from \cite{hummels2007} in two (interrelated) main dimensions.
+First, our characterization of the time trends in transport costs identifies the trend patterns of the transport costs (composition effects excluded) for both the additive and ad-valorem costs, as well as for the overall transport cost.
+Second, and as direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
+%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+This turns out to be of primary importance in the disentangling in the time trend of transport costs, what comes from the composition effects and what comes the ``ceteris paribus'' transport cost dimension, as we show below.
+\footnote{Rigourously speaking, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
 We discuss this point with more details in Appendix \ref{app:comp-effects}.}
@@
 \subsection{Empirical specification}
-Our strategy to eliminate the composition effects obeys the following rationale.
+Our strategy to eliminate the composition effects follows the following rationale.
 As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in our first stage.
@@
 
-This can be exposed as a two-stage process.
+This can be described as a two-stage process.
 First, we decompose the measure of transport costs component in the three product/country/time dimensions, using fixed effects.
@@
 \end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: the country dimension (Term (a)), the product dimension (Term (b)) and the ``pure transport costs time trend'' (Term (c)).
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: the country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus transport costs time trend'' (Term (c)).
 Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.} Both equations are estimated by transport mode.
@@
 This constitutes our second stage.
-As for the ad-valorem component, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
+As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
 \begin{equation}
@@
 
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean multiplicative transport cost in 1974.
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974.
 In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
 
-As for the additive cost, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
+As for the additive cost defined in $[0;+\infty]$, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
 \begin{equation}
@@
 
-Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the ``raw'' (without composition effects excluded) and the ``pure'' estimated transport costs (composition effects excluded, see Appendix \ref{app:comp-effects}).
+Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the ``raw'' (without composition effects excluded) and the ``ceteris paribus'' estimated transport costs (composition effects excluded, see Appendix \ref{app:comp-effects}).
 
@@
 Second, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
-Third, we do not find evidence of a significant difference between the raw transport costs and the ``pure'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
-Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``pure'' transport costs.
-Composition effects are more pronounced in ocean transport.
+Third, we do not find evidence of a significant difference between the raw transport costs and the ``ceteris paribus'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
+Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
+Composition effects are more pronounced in vessel transport.
 Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects.
-Precisely, the role of composition effects is significant for the multiplicative component, whereas the reduction in the additive costs is mostly attributable to changes in ``pure'' additive costs.
-
-This last set of results stands in sharp contrast with \citet{hummels2007}, which obtained that the ``pure'' transport costs had decreased more than the unfitted ones (over 1973-2004), suggesting an important role to composition effects.
+Precisely, the role of composition effects is significant for the multiplicative component, whereas the reduction in the additive costs is mostly attributable to changes in ``ceteris paribus'' additive costs.
+
+This last set of results stands in sharp contrast with \citet{hummels2007}, which obtained that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1973-2004), suggesting an important role to composition effects.
 If anything, we find the opposite result here.
 Investigating this difference of result further (see Appendix \ref{app:comp-effects}), we find that this can be explained by the difference in methodology, and primarily the treatment of the additive component transport costs.
-Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and ocean transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in ocean transport.
-
-\section{Robustness checks \label{sec:robustness}}
-
-Robustness to the specification of the estimated equation.
-
-\paragraph{Robustness to the separability assumption}
-
-Ie, estimate
-
-$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
-
-To do on 100 products ($s$), 50 countries (the most important?)
-
-\todo{what else?}
+Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in vessel transport.
+
+%\section{Robustness checks \label{sec:robustness}}
+
+%Robustness to the specification of the estimated equation.
+
+%\paragraph{Robustness to the separability assumption}
+
+%Ie, estimate
+
+%$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
+
+%To do on 100 products ($s$), 50 countries (the most important?)
+
+%\todo{what else?}
 
@@
 
-\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
+%\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
 \begin{table}[htbp]
@@
 
-Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\ \hline
-\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs}}     \\ \hline
-Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034  \\
-Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029  \\
-Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
-\hline
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}    \\ \hline
-\multicolumn{7}{l}{\textit{Additive term }}     \\ \hline
-\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010  \\
-\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005  \\
-\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\ \hline
-\multicolumn{7}{l}{\textit{Iceberg term}}  \\ \hline
-\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017  \\
-\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 \\
-Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\ \hline
+Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\ 
+\hline
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}     \\ 
+\hline
+Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.2 & 3.4  \\
+Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9  \\
+%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Transport Costs}}    \\ 
+\hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)}}  \\ \hline
+Mean & 3.6 & 2.3 & 2.4 &1.7 & 2.6 & 1.7  \\
+Median & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 \\
+%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
+
+\hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}     \\ \hline
+Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0  \\
+Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5  \\
+%\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\ 
+\hline
 \# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351  \\
@@
 \end{table}%
+
+%%%JH 26/07 : Reprendre XXX ICI XXX
 

### 2017-07-27 redaction/dhp_july2017.tex
modifsJH270717
dernières modifs paper.

@@
 \renewcommand{\baselinestretch}{1.2}
+
+\newcommand\cites[1]{\citeauthor{#1}'s\ (\citeyear{#1})}
+
+\newcommand\citeh[1]{\citeauthor{#1}'\ (\citeyear{#1})}
 
@@
 \hline\hline
-
 Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
 \hline
-\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}     \\ 
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
 \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Transport Costs}}    \\ 
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
 \hline
@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
 Additive term expressed in fraction of fas price.}}}
@@
 
-%%%JH 26/07 : Reprendre XXX ICI XXX
-
 \begin{table}[htbp]
@@
  \hline
-   \multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
-Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036  \\
-Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033  \\
-Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018  \\
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs} }    \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term } }  \\ \hline
-Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015  \\
-Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 \\
-Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\ \hline
-\multicolumn{7}{l}{\textit{Iceberg term} } \\ \hline
-Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 \\
-Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018  \\
-Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
-\hline
- \# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\ \hline
+   \multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
+   \hline
+Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 4.0 & 3.6  \\
+Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3  \\
+%Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
+\hline
+Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 \\
+Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8  \\
+%Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
+\hline
+Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5  \\
+Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 \\
+%Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
+\hline
+ \# observations & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
 Additive term expressed in fraction of fas price.}}}
@@
 Both account for between 30 and 40\% of the total variance in Air transport, depending on the years considered.
-This is also the case for Ocean transport, even if the difference between the between-country variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
+This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
 This delivers an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
@@
 Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
-Mean  & 1.066 & 1.058 & 1.052 & \multicolumn{1}{c}{1.033} & \multicolumn{1}{c}{1.037} & \multicolumn{1}{c}{1.032} \\
-Median & 1.052 & 1.044 & 1.041 & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.027} & \multicolumn{1}{c}{1.026}  \\
-Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
-\hline
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs }}  \\ \hline
-\multicolumn{7}{l}{\textit{Iceberg term} }   \\ \hline
-Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016}  \\
-Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014}  \\
-Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\ \hline
-\multicolumn{7}{l}{\textit{Additive term} }   \\ \hline
-Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} \\
-Median & 0.012 & 0.006 & 0.006 & \multicolumn{1}{c}{0.005} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{0.004}  \\
-Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\ \hline
-\# obs & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%) }  \\
+\hline
+Mean  & 6.6 & 5.8 & 5.2 & 3.3 & 3.7 & 3.2 \\
+Median & 1.052 & 1.044 & 1.041 & 2.1 & 2.7 & 2.6 \\
+%Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}  \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%} }   \\ \hline
+Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
+Median & 2.5 & 1.7 & 1.9 & 1.0 & 1.7 & 1.4  \\
+%Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\ 
+\hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%} }   \\ \hline
+Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
+Median & 1.2 & 0.6 & 0.6 & 0.5 & 0.4 & 0.4  \\
+%Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\ 
+\hline
+\# observations & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
 Additive term expressed in fraction of fas price.}}}
@@
 \hline
-\multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs}} \\
-Mean  & 1.098 & 1.061 & \multicolumn{1}{c}{1.058} & \multicolumn{1}{c}{1.051} & \multicolumn{1}{c}{1.042} & \multicolumn{1}{c}{1.036}  \\
-Median & 1.094 & 1.051 & \multicolumn{1}{c}{1.048} & \multicolumn{1}{c}{1.045} & \multicolumn{1}{c}{1.038} & \multicolumn{1}{c}{1.031} \\
-Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
-\hline
-\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs } }\\ \hline
-\textit{Iceberg term} &       &       &       &       &       &     \\
-Mean  & 1.054 & 1.034 & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.021}  \\
-Median & 1.049 & 1.030 & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.025} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{1.018} \\
-Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
-\textit{Additive term} &       &       &       &       &       &     \\
-Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015}  \\
-Median & 0.029 & 0.013 & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{0.013} & \multicolumn{1}{c}{0.008} \\
-Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\ \hline
-\# obs & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)} \\
+Mean  & 9.8 & 6.1 & 5. & 5.1 & 4.2 & 3.6  \\
+Median & 9.4 & 5.1 & 4.8 & 4.5 & 3.8 & 3.1 \\
+%Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC} }\\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%} }   \\ \hline
+Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
+Median & 4.9 & 3.0 & 2.4 & 2.5 & 2.6 & 1.8 \\
+%Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%} }   \\ \hline
+Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\
+Median & 2.9 & 1.3 & 1.9 & 1.5 & 1.3 & 0.8 \\
+%Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\ 
+\hline
+\# observations & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
 Additive term expressed in fraction of fas price.}}}
@@
 \textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &      \\
-Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
+Model (A)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
+Model (B) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
 \textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
-Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
+Model (A)& \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
+Model (B) & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
 \textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87}  \\
-Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14}  \\
+Model (A) & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87}  \\
+Model (B) & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14}  \\
 \textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
-Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} \\
-Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28}  \\
+Model (A)& \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} \\
+Model (B) & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28}  \\
 \textbf{Test LL} &       &       &       &       &       &       \\
@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
+\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& ad-valorem
+transport costs.
+R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
 The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
@@
 \textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Term I only &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Terms A \& I &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
+Model (A) &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
+Model (B) &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
 \textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Term I only & -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} \\
-Terms A\& I & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22}  \\
+Model (A)& -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} \\
+Model (B) & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22}  \\
 \textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Term I only & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23}\\
-Terms A \& I & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} \\
+Model (A) & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23}\\
+Model (B) & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} \\
 \textbf{Test LL} &       &       & & &  & \\
@@
 From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
-Given that $\tau >1$, we can rewrite things to get the percentage change between year 1974 and any year $t>1974$:
+Given that $\tau >1$, we can rewrite to get the percentage change between year 1974 and any year $t>1974$:
 $$\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}$$
@@
 
-\noindent Noticing that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
+\noindent Note that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
 \begin{equation*}
@@
 
-\noindent with $x=\{air, ocean\}$.
+\noindent with $x=\{air, vessel\}$.
 
@@
 In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
-In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' Stata codes provided on his webpage), to better identify the precise points of difference between our estimation strategies.
+In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' Stata codes provided on his webpage\footnote{\url{http://www.krannert.purdue.edu/faculty/hummelsd/research/jep-transport-cost-data.php}}), to better identify the precise points of difference between our estimation strategies.
 
@@
 This is not likely to make a substantial difference though, as these fixed effects are by nature constant over time.
-Accordingly, this only plays as a scale effect.
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of Hummels' (2007) method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
+Accordingly, they only play as a scale effect as the estimate is in log.
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeh{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
 \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
@@
 Second, considering the 1974 weighting scheme makes sense as starting point of our time trend analysis.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}.
-\textbf{These results are reported in ???}.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007} (Results available upon request to the authors).
+
+%\textbf{These results are reported in ???}.
 
@@
 
-The results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying Hummels' (2007) method, we obtain that the composition effects tend to compensate the decrease in transport costs in both air and ocean shipping (unsurprisingly in accordance with Hummels' (2007) results, see his Figures 5 (for Air) and 6 (for Ocean)).
+As expected, these figures are strictly identical for pre-2004 years to Figures 5 and 6 in \citet{hummels2007}.
+More importantly, these results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
+Applying \citeh{hummels2007} method, we obtain that the composition effects tend to compensate the decrease in transport costs in both air and ocean shipping (unsurprisingly in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Ocean)).
 This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
@@
 
-\subsection{At a more disaggregated level}
-
-\textbf{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
-
-In this Section, we characterize the time trend in international transport costs at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.
-The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects), Manufacturing}
-\label{fig:totalTC_compeffects_excl_manuf}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_manuf.pdf}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects), Primary goods}
-\label{fig:totalTC_compeffects_excl_primary}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_primary.pdf}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Share of primary goods in the value of total US imports}
-\label{fig:Share_prim_goods}
-\begin{center}
-\includegraphics[height=4in]
-{Share_of_primary.pdf}
-\end{center}
-\end{figure}
-
-\textbf{A commenter. A mettre en Online Appendix?}
-
 \end{document}

### 2017-07-27 redaction/online_appendix_dhp2017.tex
modifsJH250717
Début révision texte et OA.
modifsJH270717
dernières modifs paper.

@@
 %\date{\vspace{-5ex}}
-%
-%%\author[lille,brest,cepii]{J\'{e}r\^{o}me H\'{e}ricourt\fnref{fn1}}
-%\fntext[fn1]{Corresponding author.
-Address: D\'{e}partement de Droit, \'{E}conomie, Gestion et AES, 12, Rue de Kergoat - CS 93837, 29238 Brest Cedex 3, France.
-Tel: (0033) 1 53 68 55 14.
-E-mail: jerome.hericourt@univ-brest.fr}
-%\address[lille]{Univ. Lille, CNRS, IESEG School of Management, Institut Catholique de Lille, UMR 9221 - LEM - Lille Économie Management}
-%\address[brest]{University of Brest-ICI (EA 2652), France}
-%\address[cepii]{CEPII, France}
-%
-%\author[lille]{Clément Nedoncelle\fnref{fn2}}
-%%\address[lille]{University of Lille - LEM-CNRS (UMR 9221), France}
-%\fntext[fn2]{Address: Université de Lille, Facult\'{e} de Sciences \'{E}conomiques et Sociales, Cit\'{e} Scientifique - B\^{a}t SH2, 59655 Villeneuve d'Ascq Cedex, France.
-E-mail: clement.nedoncelle@ed.univ-lille1.fr}
-%
+
+%Modèle additif seul
+%+ableaux avec toutes les années en 3 digit et en 4 digit. Harmoniser la présentation des résultats (cf table 1 du papier)
+%es résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix?
 
@@
 %
-%
+%\subsection{At a more disaggregated level}
+
+\textbf{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
+
+In this Section, we characterize the time trend in international transport costs at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.
+The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
+
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects), Manufacturing}
+\label{fig:totalTC_compeffects_excl_manuf}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_manuf.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects), Primary goods}
+\label{fig:totalTC_compeffects_excl_primary}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_primary.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Share of primary goods in the value of total US imports}
+\label{fig:Share_prim_goods}
+\begin{center}
+\includegraphics[height=4in]
+{Share_of_primary.pdf}
+\end{center}
+\end{figure}
+
+\textbf{A commenter. A mettre en Online Appendix?}
 

### 2017-07-27 redaction/dhp_july2017.tex
addendum_modifsJH2707

@@
 
-\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, and participants at University of California - Irvine and Paris-Dauphine for very useful discussions.
+\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, and participants at seminars at University of California-Irvine and Paris-Dauphine for very useful discussions.
 Finally, any remaining errors are ours.

### 2017-07-28 redaction/dhp_july2017.tex

@@
 Mean  & 6.6 & 5.8 & 5.2 & 3.3 & 3.7 & 3.2 \\
-Median & 1.052 & 1.044 & 1.041 & 2.1 & 2.7 & 2.6 \\
+Median & 5.2 & 4.4 & 4.1 & 2.1 & 2.7 & 2.6 \\
 %Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
@@
 \begin{figure}[htbp]
-\caption{Characterizing the time trends; Applying Hummel's (2007) method }
+\caption{Characterizing the time trends: Applying Hummel's (2007) method }
 \label{fig:comp_effects_as_in_Hummels}
@@
 \begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels.pdf} \\
+{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_notitle.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_notitle.pdf} \\
 \end{tabular}
@@
 \end{figure}
-
-\textbf{enlever les termes figure 5 et figure 6 des figures}.
 

### 2017-07-28 redaction/dhp_july2017.tex
modifsJH280717
Coquilles main text + redaction OA

@@
 Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
-SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France } }
+SciencesPo, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France } }
 
@@
 
-{\normalsize JEL classification: F14, N70, R40 \\
-Keywords: Transport costs estimates, transport costs determinants, non-linear econometrics, period 1974-2013, additive transport costs }
+\bigskip
+
+\noindent \emph{JEL classification}: F14, N70, R40 \\
+\noindent \emph{Keywords}: Transport costs estimates, transport costs determinants, non-linear econometrics, period 1974-2013, additive transport costs
 
@@
 However, \cite{hummels2007} points out the importance of the composition effects in the characterization of time trends of international transport costs.
-Over 1974-2004, \cite{hummels2007} thus obtains a stronger decline in transportation costs when the composition effects by origin country and by product is removed, in both air and ocean transports.
+Over 1974-2004, \cite{hummels2007} thus obtains a stronger decline in transportation costs when the composition effects by origin country and by product is removed, in both air and vessel transports.
 Our results tend to mitigate this view.
@@
 First, our theoretically agnostic approach provides a fairly simple (if computationally heavy) framework for measuring both ad-valorem and additive parts of transportation costs.
-We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% of the export price in air and ocean transport respectively, whereas the additive component amounts to 1.8\% and 2.9\%.
+We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% of the export price in air and vessel transport respectively, whereas the additive component amounts to 1.8\% and 2.9\%.
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
@@
 \multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}  \\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%} }   \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%) }}   \\ \hline
 Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
@@
 \hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%} }   \\ \hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$), in \%} }   \\ \hline
 Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
@@
 \multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)} \\
+\hline
 Mean  & 9.8 & 6.1 & 5. & 5.1 & 4.2 & 3.6  \\
@@
 \multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC} }\\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%} }   \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\ \hline
 Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
@@
 \hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%} }   \\ \hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\ \hline
 Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\

### 2017-07-28 redaction/online_appendix_dhp2017.tex
addendum_modifsJH2707
modifsJH280717
Coquilles main text + redaction OA

@@
 \usepackage[flushleft]{threeparttable}
+
+%\usepackage{titling}
+
+%\setlength{\droptitle}{-10em}   % This is your set screw
+
 \makeatletter
@@
 
-%Modèle additif seul
-%+ableaux avec toutes les années en 3 digit et en 4 digit. Harmoniser la présentation des résultats (cf table 1 du papier)
-%es résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix?
+%Modèle additif seul DONE
+%+Tableaux avec toutes les années en 3 digit DONE
+% Tableaux avec toutes les années en 4 digit ???
+%Harmoniser la présentation des résultats (cf table 1 du papier) DONE
+%Les résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix? DONE, mais à compléter
 
@@
 \newpage
-%\section{\label{sec:addi}Additional Tables}
+
+%
+%
+%
+%
+%			SECTION A / ADDITIVE ONLY
+%			
+%			
+%				
+%
+%
 
@@
   \footnotesize{
-  \caption{Transport costs estimates: Summary \label{sec_oa:add_only}}
+  \caption{Transport costs estimates : Summary\label{sec_oa:add_only}}
   \begin{center}
@@
 The additive term is expressed in fraction of fas price.
-($^\ast$): Four 4-digit estimation: 0n selected years.
+($^\ast$): Four 4-digit estimation: on selected years.
 ($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
@@
 %
-%XXXXXX
-%
-%
-%			SECTION 2 / ALTERNATIVE MEASURES OF FIRM PERFORMANCE
 %
@@
 %
-%XXXXXX
+%			SECTION B / DETAILED RESULTS
 %			
@@
 %				
-%\subsection{At a more disaggregated level}
+%
+%
 
-\textbf{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
+\newpage
+%\subsection{Detailed Results \label{secoa:detailed}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{B.\arabic{table}}
+
+\subsection{Transport Cost Estimates: Yearly Detailed Results}
+
+\subsubsection{3-Digits Level Product Classification}
+\begin{landscape}
+\begin{table}[htbp]
+\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
+  %\centering
+\caption{Air: Transport costs estimates, all years, 3-digit}
+\begin{center}
+\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
+\hline
+\hline
+Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
+\hline
+Mean  & 6.9   & 7.5   & 7.2   & 7.7   & 6.9   & 6.1   & 5.4   & 6.0   & 6.4   & 6.9   & 7.2   & 6.1   & 6.4   & 6.6   & 5.7   & 5.3   & 5.0   & 5.1   & 4.9   & 5.1 \\
+Median & 5.4   & 6.4   & 6.9   & 7.1   & 6.3   & 5.3   & 3.8   & 4.9   & 5.3   & 6.1   & 6.7   & 5.5   & 5.9   & 6.3   & 5.3   & 4.6   & 4.4   & 4.5   & 4.5   & 4.4 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
+\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\ 
+\hline
+Mean  & 3.6   & 3.7   & 3.9   & 3.8   & 3.2   & 3.0   & 2.3   & 2.8   & 2.8   & 2.6   & 3.3   & 2.5   & 3.2   & 2.6   & 3.1   & 3.1   & 2.4   & 2.7   & 2.2   & 2.4 \\
+Median & 2.7   & 2.7   & 2.9   & 2.7   & 2.1   & 2.4   & 1.6   & 1.8   & 1.9   & 1.9   & 2.7   & 1.8   & 2.1   & 2.0   & 2.0   & 1.9   & 1.6   & 1.5   & 1.5   & 1.6 \\
+
+\hline
+\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\ 
+\hline
+Mean  & 2.6   & 3.0   & 2.3   & 3.1   & 2.6   & 2.1   & 2.0   & 2.0   & 2.3   & 2.8   & 2.5   & 2.8   & 2.6   & 2.9   & 1.7   & 4.6   & 1.8   & 1.8   & 1.9   & 1.9 \\
+Median & 1.1   & 1.2   & 0.9   & 1.3   & 1.1   & 0.7   & 0.5   & 0.6   & 0.8   & 1.0   & 1.0   & 1.3   & 1.3   & 1.5   & 1.0   & 0.7   & 0.8   & 0.6   & 0.9   & 0.8 \\
+\hline
+\# observations & 14955 & 15299 & 11397 & 10707 & 15222 & 15684 & 16118 & 16864 & 17322 & 18180 & 20644 & 19908 & 20695 & 20793 & 24663 & 25197 & 24958 & 25156 & 26191 & 28296 \\
+
+%%%%%%%%%%%%%%%%2nd part of the table
+
+\hline\hline
+\multicolumn{20}{c}{ }  \\
+\multicolumn{20}{c}{Continued}  \\
+\hline\hline
+
+Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
+\hline
+Mean  & 4.6   & 4.6   & 4.2   & 4.1   & 3.8   & 3.8   & 3.6   & 3.5   & 3.8   & 3.9   & 4.0   & 4.1   & 3.9   & 4.1   & 4.1   & 4.0   & 4.2   & 3.9   & 3.7   & 3.4 \\
+Median & 3.7   & 3.8   & 3.1   & 3.0   & 2.7   & 2.8   & 2.5   & 2.4   & 2.7   & 2.6   & 2.9   & 3.1   & 2.7   & 3.0   & 3.2   & 3.0   & 3.4   & 3.1   & 3.0   & 2.9 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
+\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\ 
+\hline
+Mean  & 2.3   & 2.1   & 1.9   & 1.8   & 1.8   & 1.8   & 1.7   & 1.6   & 1.6   & 1.9   & 1.9   & 2.0   & 1.8   & 2.3   & 2.3   & 2.3   & 2.6   & 2.2   & 2.2   & 1.7 \\
+Median & 1.3   & 1.4   & 1.4   & 1.3   & 1.3   & 1.5   & 1.2   & 1.1   & 1.2   & 1.4   & 1.4   & 1.6   & 1.4   & 1.9   & 1.9   & 1.8   & 2.2   & 1.7   & 1.9   & 1.7 \\
+\hline
+\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
+\hline
+Mean  & 1.7   & 1.6   & 1.5   & 1.5   & 1.4   & 1.4   & 1.3   & 1.3   & 1.6   & 1.4   & 1.5   & 1.4   & 1.3   & 1.2   & 1.2   & 1.2   & 1.1   & 1.1   & 0.9   & 1.0 \\
+Median & 0.8   & 0.7   & 0.6   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.4   & 0.4   & 0.4   & 0.5 \\
+\hline
+\# observations & 29948 & 31037 & 32187 & 33502 & 33492 & 33523 & 35027 & 34885 & 35159 & 35891 & 36990 & 41806 & 42554 & 40858 & 40159 & 38275 & 40279 & 41190 & 40909 & 39351 \\
+
+\hline
+\hline
+\end{tabular}}%
+
+\end{center}
+\label{tab_oa:result_air_ally3}%
+\end{table}%
+\end{landscape}
+
+\begin{landscape}
+\begin{table}[htbp]
+\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
+\caption{Vessel: Transport costs estimates, all years, 3-digit}
+\begin{center}
+\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
+\hline
+\hline
+Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
+\hline
+Mean  & 9.8   & 9.9   & 8.9   & 8.3   & 8.1   & 7.5   & 6.5   & 6.0   & 6.3   & 7.0   & 7.0   & 6.9   & 6.7   & 6.2   & 6.1   & 5.7   & 5.7   & 5.5   & 5.0   & 5.2 \\
+Median & 9.6   & 8.5   & 8.0   & 7.3   & 7.1   & 6.5   & 5.5   & 5.0   & 5.9   & 5.7   & 6.1   & 6.3   & 7.0   & 6.3   & 5.7   & 4.8   & 4.6   & 4.4   & 4.2   & 4.7 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
+\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\  
+\hline
+Mean  & 5.4   & 4.8   & 5.4   & 5.2   & 5.9   & 4.6   & 3.1   & 3.3   & 3.4   & 4.2   & 4.1   & 4.0   & 3.9   & 3.6   & 4.0   & 3.0   & 3.3   & 3.0   & 2.6   & 2.9 \\
+Median & 4.9   & 4.1   & 4.8   & 4.4   & 5.4   & 4.0   & 2.4   & 2.9   & 2.9   & 3.9   & 3.5   & 3.6   & 3.6   & 3.0   & 3.5   & 2.6   & 2.8   & 2.7   & 2.3   & 2.6 \\
+\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\         
+\hline
+Mean  & 5.1   & 5.5   & 3.5   & 3.5   & 2.5   & 3.1   & 3.4   & 2.9   & 3.5   & 2.9   & 3.2   & 3.2   & 2.9   & 2.8   & 2.4   & 2.9   & 2.7   & 2.8   & 2.7   & 2.7 \\
+Median & 2.9   & 3.6   & 1.9   & 1.7   & 1.2   & 1.7   & 2.3   & 1.5   & 2.3   & 2.0   & 2.3   & 2.1   & 1.8   & 1.8   & 1.3   & 2.0   & 1.7   & 1.7   & 1.8   & 1.6 \\
+\hline
+\# observations & 19007 & 18710 & 13615 & 12826 & 16601 & 17274 & 17356 & 17788 & 18075 & 18883 & 21650 & 23348 & 23729 & 23626 & 27661 & 29106 & 28383 & 28095 & 29050 & 30839 \\ %%%%%%%%%%%%%%%%2nd part of the table
+
+\hline\hline
+\multicolumn{20}{c}{ }  \\
+\multicolumn{20}{c}{Continued}  \\
+\hline\hline
+
+Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
+\hline
+Mean  & 5.2   & 5.1   & 4.8   & 4.7   & 4.8   & 5.0   & 5.1   & 5.0   & 4.8   & 5.2   & 5.4   & 5.5   & 4.8   & 4.7   & 4.4   & 4.3   & 4.0   & 3.5   & 3.6   & 3.6 \\
+Median & 4.1   & 4.3   & 3.9   & 3.9   & 3.9   & 4.5   & 4.9   & 4.6   & 4.1   & 4.8   & 5.1   & 4.9   & 4.2   & 4.2   & 3.8   & 4.1   & 3.6   & 3.0   & 3.1   & 3.3 \\
+\hline
+\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
+\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\        
+\hline
+Mean  & 2.6   & 2.8   & 2.6   & 2.5   & 2.2   & 2.5   & 2.5   & 2.7   & 2.4   & 2.4   & 2.7   & 2.6   & 2.3   & 2.5   & 2.1   & 2.2   & 1.9   & 1.8   & 1.8   & 2.2 \\
+Median & 2.2   & 2.5   & 2.2   & 2.2   & 1.9   & 2.1   & 2.1   & 2.6   & 2.3   & 1.9   & 2.8   & 2.2   & 1.9   & 2.3   & 1.8   & 2.0   & 1.8   & 1.6   & 1.4   & 1.8 \\
+\hline
+\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
+\hline
+Mean  & 2.9   & 2.7   & 2.5   & 2.5   & 3.2   & 2.8   & 2.8   & 2.4   & 2.6   & 3.2   & 2.9   & 3.0   & 2.8   & 2.4   & 2.4   & 2.1   & 2.5   & 1.9   & 1.9   & 1.5 \\
+Median & 2.0   & 1.8   & 1.6   & 1.3   & 2.0   & 2.0   & 2.2   & 1.6   & 2.0   & 2.5   & 1.9   & 2.2   & 1.9   & 1.8   & 2.1   & 1.7   & 1.9   & 1.6   & 1.6   & 0.8 \\
+\hline
+\# observations & 31865 & 32146 & 32344 & 33181 & 33986 & 34585 & 36090 & 36407 & 37255 & 37672 & 37757 & 41431 & 41763 & 39604 & 38950 & 37332 & 37748 & 38562 & 38387 & 38473 \\
+\hline
+\hline
+\end{tabular}}
+
+\end{center}
+\label{tab_oa:result_vessel_ally3}%
+\end{table}
+\end{landscape}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{C.\arabic{table}}
+
+\subsection{Eliminating the composition effects: Primary vs. Maufacturing sector} \label{sec_oa:comp-effects}
+
+%\todo{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
 
@@
 
-\textbf{A commenter. A mettre en Online Appendix?}
+%\textbf{A commenter. A mettre en Online Appendix?}
 
-\subsection{Detailed Results \label{secoa:detailed}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{B.\arabic{table}}
-
-\subsubsection{3-Digits Level}
-
-\subsubsection{4-Digits Level}
+[TO BE COMPLETED]
 
 \end{document}
+
+\subsubsection{4-Digits Level Product Classification}
+
+\begin{landscape}
+\begin{table}[htbp]
+\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
+\caption{Air: Transport costs estimates, all years, 4-digit}
+\begin{center}
+\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
+\hline
+\hline
+
+%%%%%%%%%%%%%%%%2nd part of the table
+
+\hline\hline
+\multicolumn{20}{c}{ }  \\
+\multicolumn{20}{c}{Continued}  \\
+\hline\hline
+
+\hline
+\hline
+\end{tabular}}
+
+\end{center}
+\label{tab_oa:result_vessel_ally3}%
+\end{table}
+\end{landscape}

### 2017-09-05 redaction/Presentations/DEGIT_Conference/slides_DEGIT.tex
Version des slides DEGIT conference

--- 2016-11-17 redaction/Presentations/SeminaireIrvine/slides_Irvine.tex
@@
 \usepackage{eurosym}
+\usepackage{color}
 %\usepackage[T1]{fontenc}
@@
 
-\date[September 2016]{S\'{e}minaire LEDa, Universit\'{e} Paris Dauphine, Septembre 2016}
+\date[September 2017]{DEGIT Conference, September 2017}
 \title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
@@
 \begin{itemize}
-\item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.2cm}
+\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.2cm}
 \begin{itemize}
-\item[-] Many models rely on the tractability of iceberg costs \vspace{0.1cm}
-\item[-] Additive trade costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
-\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization because they distort relative prices (Sorensen, 2014)\vspace{0.1cm}
+\item[-] Additive costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
+\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
+\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
 \end{itemize}
-\item Empiric studies
-\begin{itemize}
-\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2015, Hummels \& Skiba, 2004)  \vspace{0.1cm}
-\item[-] Contrarian view  Lashkaripour (2016) taking into account the fact that more expensive goods are heavier and hence more expensive to transport. Yet  \vspace{0.1cm}
-\begin{itemize}
-\item[-] This can only apply to goods that are enumerated by items (60\% of US trade)
-\item[-] Reasonable hypothesis for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour (2016))
+\item[$\Rightarrow$] Transport costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
 \end{itemize}
-\end{itemize}
-\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
-\end{itemize}
-\textbf{One objective of the paper is to provide an answer to this question}
+\textbf{One objective of the paper: Provide an answer to this question}
 
@@
 \begin{frame}
-\frametitle{Our paper in 3 questions (and 3 answers)}
-An empirical decomposition of the structure of transport costs \vspace{0.2cm}
+\frametitle{Our paper in one question (and 3 answers)}
 \begin{itemize}
-\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
-\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
+\item \textbf{Do additive transport costs matter?} \\ \vspace{0.2cm}
+\item[$\Rightarrow$] \textbf{Yes, they do} \vspace{0.1cm}
+\begin{itemize}
+\item[-] An empirical decomposition of the structure of transport costs \vspace{0.1cm}
+\item[-] Using the US imports database over 1974-2013 (air/vessel)\vspace{0.1cm}
+\end{itemize}
+\item[(1)] \textbf{In terms of size}: Additive costs are sizeable \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
-\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel \vspace{0.1cm}
 %}
 \end{itemize}
-\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
-\item[$\Rightarrow$] We lose much. With the additive term included:\vspace{0.1cm}
+\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
 \begin{itemize}
 %\footnotesize{
-\item[-] The estimated iceberg component is reduced by a factor linked to the importance of additive costs (which is kind of obvious) \vspace{0.1cm}
-\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
+\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
+\item[-] A significantly better ``goodness-of-fit'' %}
 %}
@@
 \begin{itemize}
-\item[(3)] \textbf{ How have international transport costs evolved over time?} \vspace{0.2cm}
-\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.2cm}
-%\begin{itemize}
-%\footnotesize{
-%\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
-%}
-%\end{itemize}
-\item[$\Rightarrow$] The importance of including the additive component \vspace{0.2cm}
+\item[(3)] \textbf{In accounting for the time trend of transport costs} \vspace{0.1cm}
 \begin{itemize}
-%\footnotesize{
-\item[-] Composition effect is especially strong for the additive costs \vspace{0.2cm}
-\item[-] In the air case, the composition effec was not the same on additive and multiplicative costs
-%\end{itemize}
-%}
+\item[-] The importance of excluding composition effects when characterizing the time trend (Hummels, 2007) \vspace{0.1cm}
+\item[-] Our contribution: The importance of modelling the additive component of transport costs \vspace{0.1cm}
+\item[$\Rightarrow$] The decrease in TC over time is mostly attributable to the decrease in the ``pure'' transport costs,  \vspace{0.1cm}
+\item[-] Rather than to trade composition effects  \vspace{0.1cm}
+\item[-] A result in sharp contrast with Hummels (2007) \vspace{0.2cm}
 \end{itemize}
 \end{itemize}
+\textbf{All our results: Provide new quantitative evidence about the importance of the additive component in international transport costs}
+
 \end{frame}
@@
 \item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
-\item[-] Yearly basis, from 1974 to 2013, SITC 5-digit classification level, by transport mode (air or vessel) (Could go up to HS-10 with recent data)\vspace{0.1cm}
+\item[-] Yearly basis, from 1974 to 2013, SITC 5-digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
+%(Could go up to HS-10 with recent data)\vspace{0.1cm}
 \end{itemize}
@@
 \item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
-\item[-] We use both the import and the export prices.
-That allows the estimate the levels of both the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015, that uses one year of Norwegian data)
+\item[-] Have both the import and the export prices  \vspace{0.1cm}
+\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive trade costs  \vspace{0.1cm}
+\begin{itemize}
+\item[$\neq$] Irrarazabal et al. (2015)
+\end{itemize}
 \end{itemize}
@@
 \item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
-\item Additional assumption : all products in a 3-digit sector (s) share the same structure of costs \vspace{0.1cm}
-\item[$\Leftrightarrow$] Write $t_{ik^s}$ and $\tau_{ik^s}$ as:
+\end{itemize}
+\footnotesize
+\begin{eqnarray*}
+\tau_{ik} = \tau_i \times \tau_k &\text{and}& t_{ik} = t_i + t_k
+\end{eqnarray*}
+\normalsize
+\item Additional assumption: All products in a 3-digit sector ($s$) share the same structure of costs \vspace{0.1cm}
+\item[$\Leftrightarrow$] Write $t_{is(k)}$ and $\tau_{is(k)}$ as:
+\footnotesize
 \begin{equation}
- \tau_{ik^s} = \tau_{i} \times \tau_{s}, \qquad t_{ik} = t_{i} + t_{s} \label{eq:specifTC}
+ \tau_{is(k)} = \tau_{i} \times \tau_{s(k)}, \qquad t_{is(k)} = t_{i} + t_{s(k)} \label{eq:specifTC}
  \end{equation}
+ \normalsize
 \end{itemize}
+\end{frame}
+
+\begin{frame}[label=slide_estimation_method]
+\begin{itemize}
 \item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
@@
 \footnotesize
-$$\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 =\left(\tau_{i} \times \tau_{s} -1+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}} \right)\times \exp(\epsilon_{ik^s})$$
+$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
+\normalsize
+\item Taking logs, we finally estimate the following equation
+\footnotesize
+\begin{equation*}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
+\end{equation*}
 \normalsize
 \begin{itemize}
-\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
+\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
+\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
 \end{itemize}
+\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
+\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_estimation_method}{\beamergotobutton{More}}\vspace{0.1cm}
 
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item Taking logs, we finally estimate the following equation
-
-\begin{equation}
-\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}}-1 \right) + \epsilon_{ik^s} \label{eq:est_equation}
-\end{equation}
-\vspace{0.1cm}
-\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
-\item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
-\begin{itemize}
-\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
-\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution 
-\end{itemize}
 \end{itemize}
@@
 \item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
+%\begin{itemize}
+\item[(A)] With only ad-valorem costs:
+\footnotesize
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
+\end{equation}
+\normalsize
+\item[(B)] With additive costs included:
+\footnotesize
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
+\end{equation}
+\normalsize
 \begin{itemize}
-\item[-] With additive costs included:
-%\footnotesize
-\begin{equation}
-\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}}-1 \right) + \epsilon_{ik^s}
-\end{equation}
-%\normalsize
-\begin{itemize}
-\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
-\end{itemize}
-\item[-] With only ad-valorem costs:
-%\footnotesize
-\begin{equation}
-\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}-1 \right) + \epsilon^{ice}_{ik^s} \label{eq:model_nlI}
-\end{equation}
-%\normalsize
+\item[$\ast$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
+%\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-\item[-] With additive costs:
-$$\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}, \qquad \widehat{t}^{add}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$$
-\item[-] With only iceberg costs:
+\item[(A)] With only iceberg costs (from Equation (\ref{eq:model_nlI})):
 $$\widehat{\tau}^{ice}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$$
+\item[(B)] With additive costs included (from Equation (\ref{eq:est_equation})):
+$$\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}, \qquad \widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$$
 \end{itemize}
-\item  Taking the weighted average over the sector-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
+\item  Taking the weighted average over the sector-country dimension, we finally get, by year and transport mode: \vspace{0.1cm}
 \begin{itemize}
-\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
+\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}$  \vspace{0.1cm}
 \item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
@@
 \begin{frame}[label=slide_results_summary]
-\frametitle{Result 1: Estimating transport costs over time}
-For both the ad-valorem and the additive components \vspace{0.1cm}
+\frametitle{Do additive costs matter? Our answers}
+\begin{itemize}
+\item Estimate the values of both the ad-valorem and the additive components \vspace{0.1cm}
 \begin{itemize}
 \item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
+\end{itemize}
 \end{itemize}
@@
       \hline \hline
-    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
-    Mode  & Vessel & Air & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{With only Ad-Valorem Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
-    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
+    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
+    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
+    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
+    \multicolumn{5}{l}{\textbf{Model (A) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
+    Mean  & \textcolor{red}{5.8} & \textcolor{red}{5.1} & 6.0 & 4.9 \\
     Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
@@
     %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{With Additive \& Ad-Valorem Costs } \\ \hline
-   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
-    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
+    \multicolumn{5}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
+   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
+    Mean  & \textcolor{red}{3.2} & \textcolor{red}{2.5} & 3.3 & 2.4 \\
     Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
@@
     %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
-    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
+    \textit{Additive term ($\widehat{t}/\widetilde{p}$, in \%)}& & & &   \\ \hline
+    Mean  & \textcolor{blue}{2.9} & \textcolor{blue}{1.8} & 2.8 & 1.9 \\
     Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
@@
     %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-       \multicolumn{5}{l}{\textit{Data} }\\ \hline
-    Mean &5.3 &5.0 & 5.6&3.9 \\
-    Median  &4.3 & 2.0 &4.4 & 1.9 \\  \hline
+    \multicolumn{5}{l}{\textbf{Data} ($p/\widetilde{p}$, \textit{in \%})} \\ \hline
+        Mean & 5.3 & 5.0& 5.6&3.9 \\
+        Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
         \# obs. & 29279 & 28207 & 29317 & 27680 \\
@@
   \end{tabular}
-%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
-%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
     \end{center}}
@@
 
- \begin{frame}
-\frametitle{Result 1: Comments}
+\begin{frame}[label=slide_result1]
+\frametitle{Result 1: Size of additive transport costs \hyperlink{app_result1}{\beamergotobutton{More on this}} }
 \begin{itemize}
-\item Transport costs are sizeable \vspace{0.1cm}
+\item \textbf{International transport costs: A sizeable additive component}  \vspace{0.2cm}
 \begin{itemize}
-\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
-\item[-] Lower than the 21\% markup taken from AVW, but  \vspace{0.1cm}
+\item[-] 48.2\% of total costs in average for ocean, 42.3\% for air  \vspace{0.1cm}
+\item[-] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
+\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
 \begin{itemize}
-\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
-\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
-\item[$\ast$] Our data are only for the US, which presumably has lower trade costs \vspace{0.1cm}
-
+\footnotesize{
+\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
+\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.1cm}
+\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
 \end{itemize}
-\item[$\Rightarrow$] Our work: A more precise and exhaustive estimation of international transport costs \vspace{0.1cm}
 \end{itemize}
@@
 \begin{itemize}
-\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.
+\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.  \vspace{0.1cm}
 \item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
@@
 
-\begin{frame}[label=slide_result2]
-\frametitle{Result 2: Additive transport costs do matter}
+\begin{frame}[label=slide_goodnessfit]
+\frametitle{Result 2: Quality of fit}
 \begin{itemize}
-\item International transport costs: A sizeable additive component  \vspace{0.2cm}
+\item \textbf{ A better quality of fit with the additive component included} \vspace{0.2cm}
 \begin{itemize}
-\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) (which is kind of obvios) \vspace{0.2cm}
-\begin{itemize}
-\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.2cm}
-\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.2cm}
-\item[$\ast$] From 5.1\% to 2.5\% in Air transport \vspace{0.2cm}
-\end{itemize}
-\item[-] A sizeable share of the additive component in total transport costs \vspace{0.2cm}
-\begin{itemize}
-\item[$\ast$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.2cm}
-\item[$\ast$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
-\end{itemize}
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label=slide_goodnessfit]
-\begin{itemize}
-\item A better quality of fit with the additive component included \vspace{0.2cm}
-\begin{itemize}
-\item[-] Compare the goodness-of-fit of Model (\ref{eq:est_equation}) (with additive TC)  \vspace{0.1cm}
-\item[ ] vs Model (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
+\item[-] Compare the goodness-of-fit of Model (B) (with additive TC)  \vspace{0.1cm}
+\item[ ] vs Model (A) (without additive TC) \vspace{0.1cm}
 \item[-] Various measures of goodness of fit \vspace{0.1cm}
@@
 \begin{frame}[label = slide_fig1]
-\frametitle{Result 3: Characterizing the trends of transport costs }
+\frametitle{Result 3: Characterizing the time trend of TC }
 \begin{itemize}
-\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Main findings}}
+\item Exploit the time-dimension of our database  \hyperlink{slide_result1}{\beamergotobutton{Go back to Result 1}} %\hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Main findings}}
 \begin{figure}[htbp]
@@
 \item[-] Lower overall transport costs in Air than in Vessel \vspace{0.1cm}
-\item[-] Downward trend for both modes since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
+\item[-] Downward trend for both modes since 1974 \vspace{0.1cm}
 %\item[-] Slighlty ore pronounced in ocean shipping \vspace{0.1cm}
 \begin{itemize}
-\item[$\star$] A 50\% decrease in Air, a 60\% decrease in Vessel \vspace{0.1cm}
+\item[$\ast$] A 50\% decrease in Air, a 60\% decrease in Vessel over the period
 \end{itemize}
@@
 
-\begin{frame}
-\frametitle{Excluding the composition effects}
+\begin{frame}[label = slide_compeffects_strategy]
+\frametitle{Eliminating composition effects: Our strategy}
 \begin{itemize}
-\item For the ad-valorem component, we estimate the following equation:
-\footnotesize
-\begin{eqnarray}
-%\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{\tau}_{ist})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{s\neq \text{011}}\beta_s.\mathbb{1}_s }_{\ln(\widetilde{\tau}_s)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ist} \label{eq:compeffects_mult}
-\end{eqnarray}
-\normalsize
+\item Based on our estimates of additive/multiplicative transport costs (TC) \hyperlink{app_compeffects_strategy}{\beamergotobutton{Comparison with Hummels (2007)}}\vspace{0.1cm}
 \begin{itemize}
-\item[-] With $\widehat{\tau}_{ist} = \widehat{\tau}^{ice}_{ist}, \widehat{\tau}^{adv}_{ist}$ previously obtained  \vspace{0.1cm}
+\item[-] That vary over time ($t$)/product ($k$) /origin country ($i$) (by transport mode): $\widehat{\tau}_{ikt}, \widehat{t}_{ikt}$ \vspace{0.1cm}
 \end{itemize}
-\item For the additive component:
-\footnotesize
-\begin{eqnarray}
-%\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
-\ln(\widehat{t}_{ist})&=&\ln\left(\delta +  \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\sum_{s \neq \text{011}}\beta_s.\mathbb{1}_s}_{\widetilde{t}_s}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ist} \label{eq:compeffects_add}
-\end{eqnarray}
-\normalsize
-\item Underlying rationale
-\footnotesize
+\item Starting from these values, extract the ``pure'' TC component \hyperlink{app_compeffects_details1}{\beamergotobutton{More details}}\vspace{0.1cm}
 \begin{itemize}
-\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
-\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
-    \normalsize
+\item[-] By the mean of a time fixed-effect \vspace{0.1cm}
+\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
+\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.1cm}
+%\item[-] From which we extract the fitted ``total'' transport cost measure \vspace{0.1cm}
 \end{itemize}
+\item What explains the decrease of transport costs over time?
+\begin{itemize}
+\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
+\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
+\end{itemize}
+\item[$\Rightarrow$] Compare the unfitted TC component to the fitted one (composition effects excluded) \vspace{0.1cm}
 \end{itemize}
@@
 
-\begin{frame}[label = slide_compeffects]
+\begin{frame}[label = slide_compeffects_figure]
+\frametitle{Time trends in transport costs: Results}
 \begin{itemize}
-\item Exclude the composition effects of transport costs changes  \hyperlink{app_compeffects}{\beamergotobutton{More details}}\vspace{0.1cm}
-\item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
-\begin{itemize}
-\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t > 1974$):
-\begin{equation*}
- \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
-\begin{itemize}
-\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
-\end{itemize}
-\item[-] For the additive cost, we build the variable ($\forall~t > 1974$)
-$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
-\end{itemize}
-\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
-\begin{itemize}
-\item[] with an initial value of 100 for $t=1974$
-\end{itemize}
+\item Transport costs over time (with and without composition effects)
 \end{itemize}
 
+\begin{figure}[htbp]
+%\caption{Transport costs (with and without composition effects)}
+%\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[height=7cm]
+{graph_composition_all.pdf}
+\end{center}
+\end{figure}
+\hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Time trends in transport costs \textit{per se}}
+\frametitle{Time trends in transport costs: Comments}
 \begin{itemize}
-\item Total transport costs (composition effects excluded) over time
+\item International transport costs have substantially reduced over the period (1974-2013) \vspace{0.1cm}
+\begin{itemize}
+\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
+\item[-] Consistent with the literature, see Lafourcade \& Thisse (2011) \vspace{0.1cm}
+\item[-] Roughly same order of magnitude for both the ad-valorem and the additive component  \vspace{0.1cm}
 \end{itemize}
-\begin{figure}[htbp]
-%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
-%\label{fig:decomp_TC_3d}
-\begin{center}
-\begin{tabular}{cc}
-\\
- \includegraphics[width=10cm,height=7.5cm]
-{graph_composition_all-eps-converted-to.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
+\item \textbf{Not much evidence of a difference between the ``raw'' and the ``pure'' TC components }\vspace{0.1cm}
+\item In Air transport \vspace{0.1cm}
+\begin{itemize}
+\footnotesize
+\item[-] The 50\% decrease, mainly attributable to the reduction of ``pure'' transport costs \vspace{0.1cm}
+\item[-] For both the additive and the ad-valorem components  \vspace{0.1cm}
+\end{itemize}
+\normalsize
+\item Composition effects (slightly) more pronounced in maritime transport \vspace{0.1cm}
+\begin{itemize}
+\footnotesize
+\item[-] The 60\% decrease: Roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
+\item[-] But for the multiplicative component only \vspace{0.1cm}
+\end{itemize}
+\normalsize
+\end{itemize}
 \end{frame}
 
-\begin{frame}[label=slide_comment_compositioneffects]
-\frametitle{What do we find}
+\begin{frame}[label = slide_compeffects_results]
+\frametitle{Time trends in transport costs: The role of additive costs}
 \begin{itemize}
-\item At the aggregate level  \vspace{0.1cm}
+\item A result in sharp contrast with Hummels (2007) \vspace{0.1cm}
 \begin{itemize}
-\item[-] Air transport costs were reduced by 50\% and vessel transport costs by 60\%. (as already seen) \vspace{0.1cm}
-\item[-] In the case of vessel, a large part of that is a composition effect (reduction in the share of bulky goods because of the relative decline of primary goods) \vspace{0.1cm}
-\item[-] The reverse happened for air transport (increase in the shipping of bulky goods) \vspace{0.1cm}
+\item[-] According to whom pure transport costs have reduced much more than the unfitted ones (over 1974-2004) \vspace{0.1cm}
+\item[-] Suggesting an important role to trade composition effects  \vspace{0.1cm}
 \end{itemize}
-\item Additive and multiplicative costs \vspace{0.1cm}
+\item Not the case here. \vspace{0.1cm}
+\item[$\Rightarrow$] \textbf{Where does the difference come from?} \vspace{0.1cm}
+\item \textbf{From the treatment of the additive component of transport costs} \vspace{0.1cm}
 \begin{itemize}
-\item[-] The composition effect is especially strong for the additive costs. Maybe bulkiness is more important for additive costs than for multiplicative costs ? \vspace{0.1cm}
-\item[-] One might want to think of multiplicative as insurance and additive as shipping of handling. Will that help thinking about the composition effect? \vspace{0.1cm}
+\item[-] Assuming (as we do, $\neq$ Hummels (2007)) a varying share of the additive component over time, product and country partner is key \vspace{0.1cm}
+\item[-] Strong impact in the decomposition of the time trend of transport costs, between composition effects and ``pure'' transport costs changes \vspace{0.1cm}
+\item[-] See  \hyperlink{app_compeffects_results}{\beamergotobutton{Figure 3}}
+\end{itemize}\vspace{0.1cm}
+\item[$\Rightarrow$] The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs
 \end{itemize}
-\end{itemize}
+
 \end{frame}
@@
 \begin{itemize}
-\item[-] Additive costs are far from negligible quantitatively \vspace{0.1cm}
+\item[-] Additive costs are sizeable quantitatively \vspace{0.1cm}
 \item[-] A better fit of the model when they are taken into account    \vspace{0.1cm}
+\item[-] A key role in the understanding of the time trend of transport costs
 \end{itemize}
-\item Characterize the evolution of transport costs over time \vspace{0.1cm}
-\begin{itemize}
-\item[-] Importance of the composition effects \vspace{0.1cm}
-\end{itemize}
+
 \end{itemize}
@@
 
-\textbf{Three main possible extensions}
+\textbf{Two main possible extensions}
 \begin{itemize}
-\item On the empirical side: \vspace{0.1cm}
+\item On the empirical side: Go deeper in the structural determinants of transport costs\vspace{0.1cm}
 \begin{itemize}
-\item[(1)] Compare the trends in transports costs \textit{between} air and vessel (excluding composition effects) \vspace{0.1cm}
-\item[(2)] Go deeper in the structural determinants of transport costs \vspace{0.1cm}
+\item[-] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.2cm}
+\end{itemize}
+\item On the theoretical side: Use our results to explore the role of additive costs  \vspace{0.1cm}
 \begin{itemize}
-\item[$\star$] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
-\end{itemize}
-\end{itemize}
-\item On the theoretical side:
-\begin{itemize}
-\item[(3)]Use our results to explore the role of additive costs \vspace{0.1cm}
-\begin{itemize}
-\item[$\star$] In shaping international trade flows (trade theory) \vspace{0.1cm}
-\item[$\star$] In affecting the international transmission of business cycles (business cycle theory)
-\end{itemize}
+\item[-] In shaping international trade flows (trade theory) \vspace{0.1cm}
+\item[-] In affecting the international transmission of business cycles (business cycle theory)
 \end{itemize}
@@
 \appendix
+
+\begin{frame}
+\large
+\center
+\textsc{Appendix}
+\normalsize
+\end{frame}
+
+\begin{frame}[label=app_estimation_method]
+\frametitle{More on the estimation method}
+\begin{itemize}
+\item More on the non-linear estimation method
+\begin{itemize}
+\item At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
+\item The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
+\end{itemize}
+\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution
+\end{itemize}
+
+\hyperlink{slide_estimation_method}{\beamergotobutton{Back to slide}}
+\end{frame}
 
@@
 
-\begin{frame}[label=app_fig1]
-\frametitle{Ad-valorem costs over time \hyperlink{slide_fig1}{\beamergotobutton{Back to slide}}}
-\begin{figure}[htbp]
-\caption{Ad-valorem Costs (Yearly mean value, 3 digits)}
-\label{fig:mult_alone_withadd}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=5cm, height=2.5in]{Fig1a_mult_air_3d.pdf}
-& \includegraphics[width=5cm,height=2.5in]{Fig1b_mult_vessel_3d.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
+\begin{frame}[label=app_result1]
+\frametitle{Size of transport costs: Comparison with the literature (1)}
 
+\begin{itemize}
+\item Size of overall transport costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
+\item[-] Lower than the 21\% markup taken from Anderson \& Van Wincoop (2004) (AWM), but  \vspace{0.1cm}
+\begin{itemize}
+\footnotesize{
+\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
+\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
+\item[$\ast$] Our data are only for the US, which presumably has lower trade costs } \vspace{0.1cm}
+\end{itemize}
+\end{itemize}
+\item[$\Rightarrow$] Our work: A more precise and exhaustive estimation of international transport costs \vspace{0.1cm}
+\end{itemize}
+\hyperlink{slide_result1}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}
+\frametitle{Size of transport costs: Comparison with the literature (2)}
+\begin{itemize}
+\item Size of additive transport costs: Comparison with Irarrazabal et al. (2015) \vspace{0.1cm}
+\begin{itemize}
+\item[-] Irarrazabal et al. (2015): Estimate the trade costs ratio $t/\tau$ = 14\% of the export price $\widetilde{p}$ \vspace{0.1cm}
+\item[-] In 2004, our estimates lead to $t/\tau = 7.7\% \times \widetilde{p}$ (Air), $t/\tau = 1.5\%\times \widetilde{p}$ (Ocean) \vspace{0.1cm}
+\item[$\Rightarrow$] Much lower !
+\end{itemize}
+\item But, our work: Estimate transport costs, not trade costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] According to AVW, transport costs $\simeq$ 15\% of trade costs \vspace{0.1cm}
+\item[-] Simple calculus thus implies that our estimates amount to 62\% (Air) and 53\% (Vessel) of Irarrazabal \& al.'s estimates \vspace{0.1cm}
+\end{itemize}
+\item[$\Rightarrow$] Additive transport costs represent between 1/2 and 2/3 of total additive trade costs
+\end{itemize}
+\hyperlink{slide_result1}{\beamergotobutton{Back to slide}}
 \end{frame}
@@
     \end{center}}
-  \label{tab:good_fit_air}%
- \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
+ % \label{tab:good_fit_air}%
+% \parbox[l]{10cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
@@
 
-\begin{frame}[label=app_compeffects]
-\frametitle{Excluding the composition effects: More details}
+\begin{frame}[label = app_compeffects_strategy]
+\frametitle{Composition effects: Comparing with Hummels (2007)}
+Reformulating Hummels's (2007) method  \vspace{0.1cm}
 \begin{itemize}
-\item \textbf{For the ad-valorem cost}, rewriting Equation (\ref{eq:compeffects_mult}) :
+\item Start from the observed cif-fas price gap $TC_{ikt} = \frac{p_{ikt} - \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}$ (by transport mode) \vspace{0.1cm}
+\item \textbf{The unfitted TC measure}, $TC_t$: Take the mean yearly value by weighting each flow by its (relative) total value \vspace{0.1cm}
+\item To get the fitted TC measure, start from the regression \vspace{0.1cm}
 \footnotesize
-\begin{equation*}
-\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
-\end{equation*}
+\begin{equation}
+\ln TC_{ikt} = \delta+ \beta \ln \frac{1}{\widetilde{p}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt}
+\label{eq:hummels}\end{equation}
+\normalsize
+\item Take the predict of the regression $\rightarrow$ $\widehat{TC}_{ikt}$   \vspace{0.1cm}
+\item \textbf{The fitted TC measure}, $\widehat{TC}_t$: The unweighed average in the product/origin country dimension   \vspace{0.1cm}
+\item To be compared to the unfitted ad-valorem rate $TC_t$ (by transport mode)
+
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+%\frametitle{Composition effects: Comparing with Hummels (2007) (cont')}
+Five differences with our method  \vspace{0.1cm}
+\footnotesize
+\begin{enumerate}
+
+\item Hummels (2007) obtains the fitted ad-valorem rate in one step (the LHS is the observed price gap) while we use a 2-step approach (extract the fitted TC from the \textit{estimated} unfitted ad-valorem rate) - but a slight difference \vspace{0.1cm}
+\item Hummels (2007) does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effect  - but only a scale effect \vspace{0.1cm}
+\item We separate the country-product fixed effects in two components (country / product), due to the number of fixed effects to estimate in the non-linear setting - robustness analysis to this point (in progress) \vspace{0.1cm}
+\item We differ in the weighting scheme to obtain the yearly value of the fitted TC measure.  \vspace{0.1cm}
+\begin{itemize}
+\footnotesize
+\item[-] Hummels (2007): The unweighed average value over the $i,k$ dimension \vspace{0.1cm}
+\item[$\neq$] We weight each flow by its relative value in total trade flows observed in 1974 \vspace{0.1cm}
+\item[-] Not much of a difference provided the weighting scheme in value is constant over time \vspace{0.1cm}
+\item[-] Robustness to the alternative weighting scheme (apply Hummels's (2007) method) in progress
+\end{itemize}
+
+\end{enumerate}
+\normalsize
+
+\end{frame}
+
+\begin{frame}
+\textcolor{blue}{5.} Last difference - the major one !
+\begin{itemize}
+\item Coming back to Hummels's (2007) equation (\ref{eq:hummels}), rewritten as:
+\footnotesize
+\begin{equation}
+\ln TC_{ikt} \equiv \ln\left(\frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}  \right) = \Delta_{ikt} - \beta \ln \widetilde{p}_{ikt} \label{eq:hummels2}
+\end{equation}
+\begin{itemize}
+\item[-] with $\Delta_{ikt} \equiv \delta+ \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt}$ 
+\end{itemize}
+\normalsize
+\item Encompasses the two extreme cases of only ad-valorem costs / only additive costs
+\item Only multiplicative costs under $\beta=0$
+\begin{itemize}
+\item[-] In which case Equation (\ref{eq:hummels2}) rewrites as:
+$$\ln\left(\frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}  \right) = \Delta_{ikt} ~~\leftrightarrow~~ p_{ikt} = \underbrace{[\exp(\Delta_{ikt})+1]}_{\tau_{ikt}}\times \widetilde{p}_{ikt} $$
+%\item[$\Rightarrow$] Only multiplicative costs
+\end{itemize}
+\item Only additive costs under $\beta=1$
+\begin{itemize}
+\item[-] In which case Equation (\ref{eq:hummels2}) rewrites as:
+$$\ln\left(p_{ikt}-\widetilde{p}_{ikt}\right) = \Delta_{ikt} ~~\leftrightarrow~~ p_{ikt} = \widetilde{p}_{ikt} + \underbrace{[\exp(\Delta_{ikt})+1]}_{t_{ikt}}$$
+%\item[$\Rightarrow$] Only additive costs
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item As noted by Hummels \& Skiba (2004), $0<\beta<1$ is the elasticity of freight costs to the export price \vspace{0.1cm}
+\item Measures the relative weight of the additive component in international transport costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] The higher $\beta$ close to 1, the larger the share of additive costs \vspace{0.1cm}
+\end{itemize}
+\item By estimating Equation (\ref{eq:hummels}) \textbf{with a constant $\beta$}, Hummels (2007) assumes a constant share of additive costs over time/country partner/product \vspace{0.1cm}
+\item A key difference with our method \vspace{0.1cm}
+\begin{itemize}
+\item[-] Our strategy to obtain the transport costs measures explicitly allows for an additive component that varies in the 3 dimensions \vspace{0.1cm}
+\end{itemize}
+\item A key importance in the decomposition of time trends of transport costs (between the role of trade composition effects and the ``pure'' transport costs dimension)
+\end{itemize}
+\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}[label = app_compeffects_results]
+\frametitle{Time trend of transport costs \& the modeling of the additive component}
+
+\begin{itemize}
+\item Assuming (as Hummels, 2007) a constant share of the additive component over time/country partner/product
+\item[$\Rightarrow$] Transport cost changes decompose as:
+\begin{figure}[htbp]
+%\caption{Transport costs (with and without composition effects)}
+%\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\begin{tabular}{cc}
+{\footnotesize (a) Air } & {\footnotesize (b) Vessel}\\
+\includegraphics[width=5cm, height=4cm]{figure5_comme_hummels_air.pdf}
+& \includegraphics[width=5cm,height=4cm]{figure6_comme_hummels_ocean.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+\item[$\Rightarrow$] Suggest a much more important role to trade composition effects...  \vspace{0.1cm}
+\item Than they actually do, see \hyperlink{slide_compeffects_figure}{\beamergotobutton{Figure 2}}
+\end{itemize}
+
+\hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}[label = app_compeffects_details1]
+\frametitle{Excluding the composition effects: Details}
+\begin{itemize}
+\item For the ad-valorem component, we estimate the following equation:
+\footnotesize
+\begin{eqnarray}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\end{eqnarray}
+\normalsize
+\item For the additive component:
+\footnotesize
+\begin{eqnarray}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\end{eqnarray}
 \normalsize
 \begin{itemize}
-\item[-] From which we deduce after estimation: $\widehat{\tau}_{ik74} = \exp(\delta +\alpha_i +\beta_k)$ \vspace{0.1cm}
-\item[-] And, for any year $t > 1974$: $\widehat{\tau}_{ikt} = \exp(\delta +\alpha_i +\beta_k)\times \exp(\gamma_t)$ \vspace{0.1cm}
-\item[-] Which implies:
-\begin{equation*}
-\widehat{\tau}_{ikt} = \widehat{\tau}_{ik74}\times \exp(\gamma_t)
-\end{equation*}
+\item[-] With $\mathbb{1}_i$, $\mathbb{1}_{s(k)}$ and $\mathbb{1}_{t}$ country-, sector- and time- fixed effects \vspace{0.1cm}
 \end{itemize}
-\item With $\tau>1$, rewrite things to get the percentage change between years 1974 and $t$:
-\footnotesize
-$$\Gamma_{ikt} = 100.\frac{\widehat{\tau}_{ikt}-1}{\widehat{\tau}_{ik74}-1} = 100.\frac{\widehat{\tau}_{ik74}\exp(\gamma_t)-1}{\widehat{\tau}_{ik74}-1}$$
-\normalsize
+\item Decompose the estimated TC component in three parts:  \vspace{0.1cm}
 \begin{itemize}
-\item[-] With $\Gamma_{ikt}$ an index of transport costs in $t$ (relative to 1974) \vspace{0.1cm}
-\item[-] Only depend on the cost observed in 1974 and the time trend \vspace{0.1cm}
-\item[-] But, specific to a product-origin country pair  \vspace{0.1cm}
-
+\item[-] The country dimension (Term (a)), the product dimension (Term (b)) and the ``pure TC time trend'' (Term (c)).  \vspace{0.1cm}
+\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC}))
+%\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
 \end{itemize}
@@
 
+\begin{frame}[label = app_compeffects_details2]
+\begin{itemize}
+\item Isolate the change in the time dimension of each TC component \vspace{0.1cm}
+\begin{itemize}
+\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
+\footnotesize
+\begin{equation*}
+ \Gamma^{adv}_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
+\normalsize
+\begin{itemize}
+\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
+\end{itemize}
+\item[-] For the additive cost, build the variable ($\forall~t > 1974$)
+\footnotesize
+$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
+\normalsize
+\end{itemize}
+\item The $\Gamma^{add}_t$ and $\Gamma^{adv}_t$ series:
+\begin{itemize}
+\item[-] How TC components have changed over time if we stick the composition of trade flows by product/ country partner to the one observed in 1974  \vspace{0.1cm}
+\item[-] Interpretation in percentage changes, with an initial value of 100 for $t=1974$    \vspace{0.1cm}
+\item[-] As indices, measures easily comparable
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
 \begin{frame}
 \begin{itemize}
-\item Choose to pick as reference the mean value of the cost in 1974
+\item Rebuild a measure of ``total'' transport costs  \vspace{0.1cm}
 \begin{itemize}
-\item[-] Build the index $\Gamma_t$ such that:
+\item[-] As the sum of the two components (additive and ad-valorem) \vspace{0.1cm}
+\item[-] On both the ``raw'' (unfitted) measures and the ``pure'' (fitted) measures \vspace{0.1cm}
+\end{itemize}
+\item For the unfitted measure, build (by transport mode)
 \footnotesize
-\begin{equation*}
- \Gamma_t= 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
-\item[-] With $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974 \vspace{0.2cm}
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
+\normalsize
+\begin{itemize}
+\item[-] with $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ estimated by year (see before) \vspace{0.1cm}
+\item[-] From this, we build:
+\footnotesize
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
+\normalsize
 \end{itemize}
-\item \textbf{For the additive component}, after estimating Equation (\ref{eq:compeffects_add}):
+\item For the fitted measure, start directly from the two indices just obtained, $\Gamma^{add}_t$ and $\Gamma_{t}^{adv}$:
 \footnotesize
-\begin{eqnarray*}
-\widehat{t}_{ik74}&= & \delta + \alpha_i+ \beta_k \\
-\widehat{t}_{ikt}&=&\left(\delta + \alpha_i+ \beta_k\right).\exp(\gamma_t),\qquad \forall t > 1974
-\end{eqnarray*}
+$$\Gamma^{tc, pure}_t = \omega_{x}\Gamma^{add}_t +(1-\omega_{x})\Gamma^{adv}_t $$
 \normalsize
-
 \begin{itemize}
-\item[-] From which we deduce:
-$$\widehat{t}_{ikt} = \widehat{t}_{ik74} \times \exp(\gamma_t)$$
-\item[-] With $t >0$, obtain the percentage change from 1974 through:
-\footnotesize
-\begin{equation*}
-\Gamma^{add}_{ikt} = 100\frac{\widehat{t}_{ikt}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)
-\end{equation*}
-\normalsize
-\item Independent of the $i,k$ dimension, such that we can rewrite:
-$$\Gamma^{add}_t  = 100\exp(\gamma_t) $$
+\item[-] With $x={air, ~vessel}$ and $\omega_x$ the relative weight of the additive TC component in total cost in 1974 \vspace{0.1cm}
+\item[-] Precisely, $\omega_{air}= 0.423$ and $\omega_{vessel}= 0.482$
 \end{itemize}
 \end{itemize}
-\hyperlink{slide_compeffects}{\beamergotobutton{Back to slide}}
+\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to slide}}
 \end{frame}

### 2017-09-05 redaction/Presentations/DEGIT_Conference/slides_DEGIT.tex
Nouvelle version des slides DEGIT

@@
 \begin{itemize}
-\item[-] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
-\footnotesize
-$$pq = \tau \widetilde{p}q$$
+\item[$\ast$] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
+\item[$\Rightarrow$] Multiplicative costs:     \footnotesize $pq = \tau \widetilde{p}q$
 \normalsize
@@
 \begin{itemize}
-\item[$\star$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2014) \vspace{0.1cm}
-\end{itemize}
-\item[-] Some trade policy instruments as well (quota license price)
-\footnotesize
-$$pq = (t+\widetilde{p})q  $$
+\item[$\ast$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2015) \vspace{0.1cm}
+\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
+\item[$\Rightarrow$] Additive costs: \footnotesize $pq = (t+\widetilde{p})q  $
+\end{itemize}
 \end{itemize}
@@
 \frametitle{Our paper in one question (and 3 answers)}
-\begin{itemize}
-\item \textbf{Do additive transport costs matter?} \\ \vspace{0.2cm}
+
+\begin{itemize}
+\item  \textbf{Do additive transport costs matter?} \vspace{0.1cm}
+\begin{itemize}
+\item[-] Provide an empirical decomposition of the structure of transport costs \vspace{0.1cm}
+\item[-] Using the US imports database over 1974-2013 (air/vessel)\vspace{0.2cm}
+\end{itemize}
 \item[$\Rightarrow$] \textbf{Yes, they do} \vspace{0.1cm}
-\begin{itemize}
-\item[-] An empirical decomposition of the structure of transport costs \vspace{0.1cm}
-\item[-] Using the US imports database over 1974-2013 (air/vessel)\vspace{0.1cm}
-\end{itemize}
+
 \item[(1)] \textbf{In terms of size}: Additive costs are sizeable \vspace{0.1cm}
@@
 %\footnotesize{
-\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
 \item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel \vspace{0.1cm}
+%\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
+\item[-] Roughly 50\% of the overall transport costs \vspace{0.1cm}
+
 %}
@@
 \item[-] Have both the import and the export prices  \vspace{0.1cm}
-\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive trade costs  \vspace{0.1cm}
+\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive transport costs  \vspace{0.1cm}
 \begin{itemize}
@@
 \item Rewrite to get:
-$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
+\begin{equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:basis_equation}
+\end{equation}
 \item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
@@
 \begin{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
-\item[$\Rightarrow$] The equation of interest becomes:
+\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} >1$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item[$\Rightarrow$] The equation of interest (\ref{eq:basis_equation}) becomes:
 \footnotesize
@@
 
-\begin{frame}[label = slide_compeffects_strategy]
+\begin{frame} [label = slide_compeffects_strategy]
 \frametitle{Eliminating composition effects: Our strategy}
@@
 \item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
-\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.1cm}
+\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.2cm}
 %\item[-] From which we extract the fitted ``total'' transport cost measure \vspace{0.1cm}
 \end{itemize}
-\item What explains the decrease of transport costs over time?
-\begin{itemize}
-\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
-\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
+\item What explains the decrease of transport costs over time?  \vspace{0.1cm}
+\begin{itemize}
+\item[-] Composition effects vs ``Pure'' TC changes? \vspace{0.1cm}
+%\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
+%\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
 \end{itemize}
@@
 \begin{itemize}
-\item International transport costs have substantially reduced over the period (1974-2013) \vspace{0.1cm}
-\begin{itemize}
+\item A strong reduction of transport costs over the period (1974-2013) \vspace{0.1cm}
+\begin{itemize}
+\footnotesize
 \item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
@@
 \item[-] Roughly same order of magnitude for both the ad-valorem and the additive component  \vspace{0.1cm}
-\end{itemize}
-\item \textbf{Not much evidence of a difference between the ``raw'' and the ``pure'' TC components }\vspace{0.1cm}
+\normalsize
+\end{itemize}
+\item \textbf{Trade composition effects: A minor role in accounting for the time trend of TC}\vspace{0.1cm}
+\begin{itemize}
+\footnotesize
+\item[-] Not much evidence of a difference between the ``raw'' and the ``pure'' TC components \vspace{0.1cm}
+\normalsize
+\end{itemize}
 \item In Air transport \vspace{0.1cm}

### 2017-09-05 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
Nouvelle version des slides DEGIT

@@
 %\footnotesize{
-\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
-\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
+\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.1cm}
+\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.2cm}
+
 %}
@@
 \item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next. \vspace{0.1cm}
-\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
+\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution
 \end{itemize}

### 2017-09-06 redaction/Presentations/DEGIT_Conference/slides_DEGIT_v1.tex
Nouvelle version des slides

--- 2017-09-05 redaction/Presentations/DEGIT_Conference/slides_DEGIT.tex
@@
 \item[$\ast$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2015) \vspace{0.1cm}
-\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
+%\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
 \item[$\Rightarrow$] Additive costs: \footnotesize $pq = (t+\widetilde{p})q  $
@@
 \begin{itemize}
-\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.2cm}
+\item Recent empirical evidence in support of the additive structure of trade costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] Irrarazabal et al. (2015), Hummels \& Skiba (2004), Martin (2012)  \vspace{0.2cm}
+\end{itemize}
+\item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
-\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
+
 \end{itemize}
@@
 \item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
-\item[-] A significantly better ``goodness-of-fit'' %}
+\item[-] A substantially better ``goodness-of-fit'' %}
 %}
@@
 \begin{itemize}
-\item[(3)] \textbf{In accounting for the time trend of transport costs} \vspace{0.1cm}
-\begin{itemize}
-\item[-] The importance of excluding composition effects when characterizing the time trend (Hummels, 2007) \vspace{0.1cm}
-\item[-] Our contribution: The importance of modelling the additive component of transport costs \vspace{0.1cm}
-\item[$\Rightarrow$] The decrease in TC over time is mostly attributable to the decrease in the ``pure'' transport costs,  \vspace{0.1cm}
-\item[-] Rather than to trade composition effects  \vspace{0.1cm}
+\item[(3)] \textbf{In accounting for the time trend of transport costs} \vspace{0.2cm}
+\begin{itemize}
+\item[-] The importance of modelling the additive component of transport costs \vspace{0.1cm}
+\item[-] Because it changes the result of excluding the trade composition effects when characterizing the time trend \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] The decrease in TC over time is mostly attributable to the decrease in the ``pure'' transport costs,  \vspace{0.1cm}
+\item[$\ast$] Rather than to trade composition effects  \vspace{0.1cm}
+\end{itemize}
 \item[-] A result in sharp contrast with Hummels (2007) \vspace{0.2cm}
@@
 \begin{itemize}
-\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
+\item Our measure of international transport costs: The difference between the ``export'' price and the ``import'' price of US imports\vspace{0.1cm}
 \item Database: US Imports of Merchandise database \vspace{0.1cm}
@@
 \item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
-\begin{itemize}
-\item[-] The RHS are at the 3-digit classification level  \vspace{0.1cm}
+\item The RHS is at the 3-digit classification level  \vspace{0.1cm}
+\begin{itemize}
 \item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
@@
 \begin{itemize}
-\item[$\neq$] Irrarazabal et al. (2015)
+\item[$\neq$] Irarrazabal et al. (2015)
 \end{itemize}
@@
 $$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
-\item For product $k$, from country $i$  \vspace{0.1cm}
+\item For a product $k$, from country $i$ (for a given transport mode)   \vspace{0.1cm}
 \item Rewrite to get:
@@
 \begin{itemize}
-\item[-] The equation is also mode (air or vessel)- specific
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
+\item[-] The equation is also mode (air or vessel) - specific
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label = slide_empirical_strategy]
 \frametitle{Empirical specification (2)}
-\textbf{The estimation strategy}
-\begin{itemize}
-\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
-\begin{itemize}
-\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
-\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
-\end{itemize}
-\footnotesize
-\begin{eqnarray*}
-\tau_{ik} = \tau_i \times \tau_k &\text{and}& t_{ik} = t_i + t_k
-\end{eqnarray*}
-\normalsize
-\item Additional assumption: All products in a 3-digit sector ($s$) share the same structure of costs \vspace{0.1cm}
-\item[$\Leftrightarrow$] Write $t_{is(k)}$ and $\tau_{is(k)}$ as:
-\footnotesize
-\begin{equation}
- \tau_{is(k)} = \tau_{i} \times \tau_{s(k)}, \qquad t_{is(k)} = t_{i} + t_{s(k)} \label{eq:specifTC}
- \end{equation}
- \normalsize
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label=slide_estimation_method]
-\begin{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} >1$, the error term should be always positive and multiplicative \vspace{0.1cm}
-\item[$\Rightarrow$] The equation of interest (\ref{eq:basis_equation}) becomes:
-\footnotesize
-$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
-\normalsize
-\item Taking logs, we finally estimate the following equation
+\begin{itemize}
+\item Make some simplifying assumptions \hyperlink{app_empirical_strategy}{\beamergotobutton{More details}}  \vspace{0.1cm}
+\begin{itemize}
+\item[-] About the specification of both additive and multiplicative costs  \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] Separability between the product and the country dimensions (as Irarrazabal et al., 2015)   \vspace{0.1cm}
+\item[$\ast$] All products $k$ in a 3-digit sector ($s$) share the same structure of costs  \vspace{0.1cm}
+\end{itemize}
+\item[-] About the specification of the error term \vspace{0.1cm}
+\end{itemize}
+\item Taking logs, we finally estimate the following equation:
 \footnotesize
@@
 \end{equation*}
-\normalsize
-\begin{itemize}
+
+\begin{itemize}
+\footnotesize
 \item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
@@
 \end{itemize}
-\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
-\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_estimation_method}{\beamergotobutton{More}}\vspace{0.1cm}
-
+\normalsize
+\item A non-linear equation (due to the additive costs)  
+\item[$\Rightarrow$] Estimation using non-linear squares 
 \end{itemize}
@@
 \item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
-\item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
+\item[$\Rightarrow$] Estimate \textit{two} models\vspace{0.1cm}
 %\begin{itemize}
@@
 \begin{itemize}
-\item[$\ast$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
+\item[$\ast$] Under Model (B), $\simeq$ 800 fixed effects to estimate  \vspace{0.1cm}
+\item[$\ast$] For each year over 1974-2013, by transport mode (air/vessel)
 %\end{itemize}
@@
 \begin{itemize}
-\item[-] 48.2\% of total costs in average for ocean, 42.3\% for air  \vspace{0.1cm}
+\item[-] 48.2\% of total costs in average for vessel, 42.3\% for air  \vspace{0.1cm}
 \item[-] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
@@
 \item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
-\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.1cm}
+\item[$\ast$] From 5.8\% to 3.2\% in maritime transport (mean value over the period) \vspace{0.1cm}
 \item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
@@
 \begin{itemize}
-\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.  \vspace{0.1cm}
+\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.  
 \item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
@@
 
-\begin{frame}
+\begin{frame} [label = slide_compeffects_strategy]
 \frametitle{Time trends in transport costs \& composition effects }
@@
 \item[-] Depend on the evolution of per product- per partner costs, \vspace{0.1cm}
-\item[-] But also on the composition of trade flows \vspace{0.2cm}
+\item[-] But also on the composition of trade flows \vspace{0.1cm}
 \begin{itemize}
@@
 \item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
-\item[$\Rightarrow$] What we do, in accordance with Hummels (2007)
-\end{itemize}
-\end{frame}
-
-\begin{frame} [label = slide_compeffects_strategy]
-\frametitle{Eliminating composition effects: Our strategy}
-\begin{itemize}
-\item Based on our estimates of additive/multiplicative transport costs (TC) \hyperlink{app_compeffects_strategy}{\beamergotobutton{Comparison with Hummels (2007)}}\vspace{0.1cm}
-\begin{itemize}
-\item[-] That vary over time ($t$)/product ($k$) /origin country ($i$) (by transport mode): $\widehat{\tau}_{ikt}, \widehat{t}_{ikt}$ \vspace{0.1cm}
-\end{itemize}
-\item Starting from these values, extract the ``pure'' TC component \hyperlink{app_compeffects_details1}{\beamergotobutton{More details}}\vspace{0.1cm}
-\begin{itemize}
-\item[-] By the mean of a time fixed-effect \vspace{0.1cm}
-\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
-\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.2cm}
-%\item[-] From which we extract the fitted ``total'' transport cost measure \vspace{0.1cm}
-\end{itemize}
-\item What explains the decrease of transport costs over time?  \vspace{0.1cm}
-\begin{itemize}
-\item[-] Composition effects vs ``Pure'' TC changes? \vspace{0.1cm}
-%\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
-%\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
-\end{itemize}
-\item[$\Rightarrow$] Compare the unfitted TC component to the fitted one (composition effects excluded) \vspace{0.1cm}
+\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) \vspace{0.1cm}
+\item Some insights on our methodology \hyperlink{app_compeffects_strategy}{\beamergotobutton{More details}} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{Comparison with Hummels (2007)}} \vspace{0.1cm}
+\begin{itemize}
+\item[-] Start with estimates of additive/multiplicative transport costs  \vspace{0.1cm}
+\item[-] Extract the ``pure'' TC component by the mean of a time fixed-effect \vspace{0.1cm}
+%\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
+\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted
+\end{itemize}
 \end{itemize}
@@
 \footnotesize
-\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
+%\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
 \item[-] Consistent with the literature, see Lafourcade \& Thisse (2011) \vspace{0.1cm}
@@
 \footnotesize
-\item[-] The 50\% decrease, mainly attributable to the reduction of ``pure'' transport costs \vspace{0.1cm}
-\item[-] For both the additive and the ad-valorem components  \vspace{0.1cm}
+\item[-] The 50\% decrease is mainly attributable to the reduction of ``pure'' transport costs, for both components  \vspace{0.1cm}
 \end{itemize}
@@
 \footnotesize
-\item[-] The 60\% decrease: Roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
+\item[-] The 60\% decrease is roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
 \item[-] But for the multiplicative component only \vspace{0.1cm}
@@
 \begin{itemize}
-\item[-] According to whom pure transport costs have reduced much more than the unfitted ones (over 1974-2004) \vspace{0.1cm}
+\item[-] According to whom pure transport costs have decreased much more than the unfitted ones (over 1974-2004) \vspace{0.1cm}
 \item[-] Suggesting an important role to trade composition effects  \vspace{0.1cm}
@@
 \item[-] Assuming (as we do, $\neq$ Hummels (2007)) a varying share of the additive component over time, product and country partner is key \vspace{0.1cm}
-\item[-] Strong impact in the decomposition of the time trend of transport costs, between composition effects and ``pure'' transport costs changes \vspace{0.1cm}
-\item[-] See  \hyperlink{app_compeffects_results}{\beamergotobutton{Figure 3}}
+%\item[-] Strong impact in the decomposition of the time trend of transport costs, between composition effects and ``pure'' transport costs changes \vspace{0.1cm}
+\item[-] What if we relax this assumption?
 \end{itemize}\vspace{0.1cm}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+%\frametitle{Time trend of transport costs \& the modeling of the additive component}
+\begin{itemize}
+\item Assuming (as Hummels, 2007) a constant share of the additive component over time/country partner/product \vspace{0.1cm}
+\item[$\Rightarrow$] Transport cost changes decompose as:
+\begin{figure}[htbp]
+%\caption{Transport costs (with and without composition effects)}
+%\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\begin{tabular}{cc}
+{\scriptsize (a) Air } & {\scriptsize (b) Vessel}\\
+\includegraphics[width=5cm, height=4cm]{figure5_comme_hummels_air.pdf}
+& \includegraphics[width=5cm,height=4cm]{figure6_comme_hummels_ocean.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+\item[$\Rightarrow$] Suggest a much more important role to trade composition effects...  \vspace{0.1cm}
+\item Than they actually do, see \hyperlink{slide_compeffects_figure}{\beamergotobutton{Figure 2}} \vspace{0.1cm}
+
 \item[$\Rightarrow$] The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs
 \end{itemize}
-
 \end{frame}
@@
 
-\begin{frame}[label=app_estimation_method]
+\begin{frame}[label = app_empirical_strategy]
+\frametitle{More on the empirical specification }
+\textbf{The estimation strategy}
+\begin{itemize}
+\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
+\begin{itemize}
+\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
+\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
+\end{itemize}
+\footnotesize
+\begin{eqnarray*}
+\tau_{ik} = \tau_i \times \tau_k &\text{and}& t_{ik} = t_i + t_k
+\end{eqnarray*}
+\normalsize
+\item Additional assumption: All products in a 3-digit sector ($s$) share the same structure of costs \vspace{0.1cm}
+\item[$\Leftrightarrow$] Write $t_{is(k)}$ and $\tau_{is(k)}$ as:
+\footnotesize
+\begin{equation}
+ \tau_{is(k)} = \tau_{i} \times \tau_{s(k)}, \qquad t_{is(k)} = t_{i} + t_{s(k)} \label{eq:specifTC}
+ \end{equation}
+ \normalsize
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} >1$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item[$\Rightarrow$] The equation of interest (\ref{eq:basis_equation}) becomes:
+\footnotesize
+$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
+\normalsize
+\item Taking logs, we finally estimate the following equation
+\footnotesize
+\begin{equation*}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
+\end{equation*}
+\normalsize
+\begin{itemize}
+\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
+\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
+\end{itemize}
+\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
+\item[$\Rightarrow$] Estimation using non-linear squares
+
+\end{itemize}
+\hyperlink{slide_empirical_strategy}{\beamergotobutton{Back to Slide}}
+\end{frame}
+
+\begin{frame}
 \frametitle{More on the estimation method}
 \begin{itemize}
-\item More on the non-linear estimation method
+\item Non-linear least square estimation method: Some insights \vspace{0.1cm}
 \begin{itemize}
 \item At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
+\item The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.2cm}
 \end{itemize}
@@
 
-\hyperlink{slide_estimation_method}{\beamergotobutton{Back to slide}}
+\hyperlink{slide_empirical_strategy}{\beamergotobutton{Back to slide}}
 \end{frame}
@@
 \begin{frame}[label = app_compeffects_strategy]
-\frametitle{Composition effects: Comparing with Hummels (2007)}
+\frametitle{Composition effects: Our strategy}
+\begin{itemize}
+\item Start with our estimates of add. \& adv. transport costs (TC) 
+\begin{itemize}
+\footnotesize
+\item[-] That vary over time ($t$)/product ($k$) /origin country ($i$) (by transport mode): $\widehat{\tau}_{ikt}, \widehat{t}_{ikt}$ \vspace{0.1cm}
+\end{itemize}
+\item[$\Rightarrow$] Decompose the estimated TC component in three parts:  \vspace{0.1cm}
+\begin{itemize}
+\footnotesize
+\item[-] Country dimension (Term (a)), product dimension (Term (b)) and the ``pure TC time trend'' (Term (c)).  \vspace{0.1cm}
+%\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC}))
+%\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
+\end{itemize}
+\item For the ad-valorem component, we estimate the following equation:
+\footnotesize
+\begin{eqnarray}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\end{eqnarray}
+\normalsize
+\item For the additive component:
+\footnotesize
+\begin{eqnarray}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\end{eqnarray}
+\normalsize
+\begin{itemize}
+\item[-] With $\mathbb{1}_i$, $\mathbb{1}_{s(k)}$ and $\mathbb{1}_{t}$ country-, sector- and time- fixed effects \vspace{0.1cm}
+\end{itemize}
+
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item Isolate the change in the time dimension of each TC component \vspace{0.1cm}
+\begin{itemize}
+\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
+\footnotesize
+\begin{equation*}
+ \Gamma^{adv}_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
+\end{equation*}
+\normalsize
+\begin{itemize}
+\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
+\end{itemize}
+\item[-] For the additive cost, build the variable ($\forall~t > 1974$)
+\footnotesize
+$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
+\normalsize
+\end{itemize}
+\item The $\Gamma^{add}_t$ and $\Gamma^{adv}_t$ series:
+\begin{itemize}
+\item[-] How TC components have changed over time if we stick the composition of trade flows by product/ country partner to the one observed in 1974  \vspace{0.1cm}
+\item[-] Interpretation in percentage changes, with an initial value of 100 for $t=1974$    \vspace{0.1cm}
+\item[-] As indices, measures easily comparable
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item Rebuild a measure of ``total'' transport costs  \vspace{0.1cm}
+\begin{itemize}
+\item[-] As the sum of the two components (additive and ad-valorem) \vspace{0.1cm}
+\item[-] On both the ``raw'' (unfitted) measures and the ``pure'' (fitted) measures \vspace{0.1cm}
+\end{itemize}
+\item For the unfitted measure, build (by transport mode)
+\footnotesize
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
+\normalsize
+\begin{itemize}
+\item[-] with $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ estimated by year (see before) \vspace{0.1cm}
+\item[-] From this, we build:
+\footnotesize
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
+\normalsize
+\end{itemize}
+\item For the fitted measure, start directly from the two indices just obtained, $\Gamma^{add}_t$ and $\Gamma_{t}^{adv}$:
+\footnotesize
+$$\Gamma^{tc, pure}_t = \omega_{x}\Gamma^{add}_t +(1-\omega_{x})\Gamma^{adv}_t $$
+\normalsize
+\begin{itemize}
+\item[-] With $x={air, ~vessel}$ and $\omega_x$ the relative weight of the additive TC component in total cost in 1974 \vspace{0.1cm}
+\item[-] Precisely, $\omega_{air}= 0.423$ and $\omega_{vessel}= 0.482$
+\end{itemize}
+\end{itemize}
+\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to slide}}
+\end{frame}
+
+\begin{frame}[label = app_compeffects_Hummels]
+\frametitle{Composition effects: Comparing with Hummels (2007) (1)}
 Reformulating Hummels's (2007) method  \vspace{0.1cm}
@@
 \end{itemize}
-\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to slide}}
-\end{frame}
-
-\begin{frame}[label = app_compeffects_results]
-\frametitle{Time trend of transport costs \& the modeling of the additive component}
-
-\begin{itemize}
-\item Assuming (as Hummels, 2007) a constant share of the additive component over time/country partner/product
-\item[$\Rightarrow$] Transport cost changes decompose as:
-\begin{figure}[htbp]
-%\caption{Transport costs (with and without composition effects)}
-%\label{fig:totalTC_compeffects_excl}
-\begin{center}
-\begin{tabular}{cc}
-{\footnotesize (a) Air } & {\footnotesize (b) Vessel}\\
-\includegraphics[width=5cm, height=4cm]{figure5_comme_hummels_air.pdf}
-& \includegraphics[width=5cm,height=4cm]{figure6_comme_hummels_ocean.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-\item[$\Rightarrow$] Suggest a much more important role to trade composition effects...  \vspace{0.1cm}
-\item Than they actually do, see \hyperlink{slide_compeffects_figure}{\beamergotobutton{Figure 2}}
-\end{itemize}
-
-\hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}}
-\end{frame}
-
-\begin{frame}[label = app_compeffects_details1]
-\frametitle{Excluding the composition effects: Details}
-\begin{itemize}
-\item For the ad-valorem component, we estimate the following equation:
-\footnotesize
-\begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
-\end{eqnarray}
-\normalsize
-\item For the additive component:
-\footnotesize
-\begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
-\end{eqnarray}
-\normalsize
-\begin{itemize}
-\item[-] With $\mathbb{1}_i$, $\mathbb{1}_{s(k)}$ and $\mathbb{1}_{t}$ country-, sector- and time- fixed effects \vspace{0.1cm}
-\end{itemize}
-\item Decompose the estimated TC component in three parts:  \vspace{0.1cm}
-\begin{itemize}
-\item[-] The country dimension (Term (a)), the product dimension (Term (b)) and the ``pure TC time trend'' (Term (c)).  \vspace{0.1cm}
-\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC}))
-%\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label = app_compeffects_details2]
-\begin{itemize}
-\item Isolate the change in the time dimension of each TC component \vspace{0.1cm}
-\begin{itemize}
-\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t \geq 1974$):
-\footnotesize
-\begin{equation*}
- \Gamma^{adv}_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
-\end{equation*}
-\normalsize
-\begin{itemize}
-\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
-\end{itemize}
-\item[-] For the additive cost, build the variable ($\forall~t > 1974$)
-\footnotesize
-$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
-\normalsize
-\end{itemize}
-\item The $\Gamma^{add}_t$ and $\Gamma^{adv}_t$ series:
-\begin{itemize}
-\item[-] How TC components have changed over time if we stick the composition of trade flows by product/ country partner to the one observed in 1974  \vspace{0.1cm}
-\item[-] Interpretation in percentage changes, with an initial value of 100 for $t=1974$    \vspace{0.1cm}
-\item[-] As indices, measures easily comparable
-\end{itemize}
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item Rebuild a measure of ``total'' transport costs  \vspace{0.1cm}
-\begin{itemize}
-\item[-] As the sum of the two components (additive and ad-valorem) \vspace{0.1cm}
-\item[-] On both the ``raw'' (unfitted) measures and the ``pure'' (fitted) measures \vspace{0.1cm}
-\end{itemize}
-\item For the unfitted measure, build (by transport mode)
-\footnotesize
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
-\normalsize
-\begin{itemize}
-\item[-] with $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ estimated by year (see before) \vspace{0.1cm}
-\item[-] From this, we build:
-\footnotesize
-$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
-\normalsize
-\end{itemize}
-\item For the fitted measure, start directly from the two indices just obtained, $\Gamma^{add}_t$ and $\Gamma_{t}^{adv}$:
-\footnotesize
-$$\Gamma^{tc, pure}_t = \omega_{x}\Gamma^{add}_t +(1-\omega_{x})\Gamma^{adv}_t $$
-\normalsize
-\begin{itemize}
-\item[-] With $x={air, ~vessel}$ and $\omega_x$ the relative weight of the additive TC component in total cost in 1974 \vspace{0.1cm}
-\item[-] Precisely, $\omega_{air}= 0.423$ and $\omega_{vessel}= 0.482$
-\end{itemize}
-\end{itemize}
-\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to slide}}
+\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to main slide}}
 \end{frame}

### 2017-09-06 redaction/Presentations/DEGIT_Conference/slides_DEGIT_v1.tex
Dernière version (?)

@@
 \begin{itemize}
-\item[-] The importance of modelling the additive component of transport costs \vspace{0.1cm}
+\item[-] The modelling of the additive component of transport costs is key \vspace{0.1cm}
 \item[-] Because it changes the result of excluding the trade composition effects when characterizing the time trend \vspace{0.1cm}
@@
 $$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
-\item For a product $k$, from country $i$ (for a given transport mode)   \vspace{0.1cm}
+\item From country $i$, for a product $k$ (at the $k=$ 5-digit classification level, and for a given transport mode)   \vspace{0.1cm}
 \item Rewrite to get:
@@
 \end{equation}
-\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
+\item[$\Rightarrow$] For each year over 1974-2013\vspace{0.1cm}
 \begin{itemize}
@@
 \frametitle{Empirical specification (2)}
+\textbf{Obtain the estimated equation}
 \begin{itemize}
@@
 \frametitle{Empirical specification (3)}
-\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
+\textbf{How to characterize the importance of additive costs relatively to ad-valorem?} \vspace{0.1cm}
 \begin{itemize}
@@
 %\begin{itemize}
-\item[(A)] With only ad-valorem costs:
+\item[(A)] With additive costs excluded (only ad-valorem costs):
 \footnotesize
@@
 \begin{itemize}
-\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}$  \vspace{0.1cm}
-\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label=slide_results_summary]
-\frametitle{Do additive costs matter? Our answers}
+\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$ \vspace{0.1cm}
+\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}$
+
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Question, and answers}
+\begin{itemize}
+\item Do additive costs matter? \vspace{0.5cm}
+\item How we answer (yes) to this question
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label=slide_result1]
+\frametitle{Result 1: Size of additive transport costs \hyperlink{app_result1}{\beamergotobutton{More on this}} }
 \begin{itemize}
@@
 
-\begin{frame}[label=slide_result1]
-\frametitle{Result 1: Size of additive transport costs \hyperlink{app_result1}{\beamergotobutton{More on this}} }
+\begin{frame}
 \begin{itemize}

### 2017-09-07 redaction/Presentations/DEGIT_Conference/slides_DEGIT_v1.tex

@@
 \end{figure}
-\hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}}
+\hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}} \hyperlink{app_compeffects_Hummels_rob}{\beamergotobutton{Robustness check}}
 \end{frame}
@@
 
-\begin{frame}
+\begin{frame}[label = app_compeffects_Hummels2]
 %\frametitle{Composition effects: Comparing with Hummels (2007) (cont')}
@@
 \item[-] Not much of a difference provided the weighting scheme in value is constant over time \vspace{0.1cm}
-\item[-] Robustness to the alternative weighting scheme (apply Hummels's (2007) method) in progress
+\item[-] Results robust to the alternative weighting scheme (apply Hummels's (2007) method) \hyperlink{app_compeffects_Hummels_rob}{\beamergotobutton{See}}
 \end{itemize}
@@
 
+\begin{frame}[label = app_compeffects_Hummels_rob]
+\textbf{Excluding the composition effects: Robustness check}
+\begin{itemize}
+\item Applying Hummel's (2007) weighting scheme rather than ours \hyperlink{slide_compeffects_figure}{\beamergotobutton{Compare}}
+\begin{itemize}
+\item[-] Take the unweighed average value over the $i,k$ dimension to get the fitted $TC$ measure
+\end{itemize}
+\end{itemize}
+\begin{figure}[htbp]
+%\caption{Transport costs (with and without composition effects)}
+%\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[height=6cm]
+{graph_composition_all_np.pdf}
+\end{center}
+\end{figure}
+
+\hyperlink{app_compeffects_Hummels2}{\beamergotobutton{Back to slide}}
+\end{frame}
+
 \end{document}

### 2017-09-08 redaction/dhp_july2017.tex
addendumJH2807_maintext
Rajout de la mention "Preliminary version, do not quote..."
Changement texte

@@
 
-\date{July 2017}
+\date{July 2017\\PRELIMINARY AND INCOMPLETE \\ DO NOT QUOTE WITHOUT PERMISSION}
  \maketitle
@@
 However, many common empirical facts point out the importance of additive costs.
-This paper provides new empirical results that support this view.
+This paper provides new empirical results on international transport costs that support this view.
 To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel).
 Three main results emerge.
-First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime vessel transport respectively.
+First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
 Second, taking additive costs into account significantly improves the fit of the modelling of transport costs.
-Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a share of additive costs varying in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{ceteris paribus} transport costs rather than to trade pattern composition effects.
+Third, the modeling of the additive component is key in accounting for the time trend of international transport costs, as it affects the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs ``per se''.
 In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
 \noindent \emph{JEL classification}: F14, N70, R40 \\
-\noindent \emph{Keywords}: Transport costs estimates, transport costs determinants, non-linear econometrics, period 1974-2013, additive transport costs
+\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs
 
@@
 
-\section{Introduction}
+\section{Introduction \label{sec:Intro}}
 
@@
 A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{This is due to the alteration of relative prices in a heterogeneous-firms trade framework.
-Lower additive trade costs reduce the export market prices of the highly performing exporters relative to lowly performing exporters and thus shift relative market shares among exporters towards the highly performing exporters - this of course does not happen with iceberg costs.
-Total export sales are then reallocated between a smaller number of high productivity firms.
+Lower additive trade costs reduce the export market prices of the highly performing exporters relative to lowly performing exporters.
+This, in turn, shifts relative market shares among exporters towards the highly performing ones - a mechanism that can not occur under ad-valorem costs.
+Total export sales are then reallocated between a smaller number of high-productivity firms.
 Therefore, the expected profits from exporting prior to entry increase all other things being equal by more in the case of lower per-unit costs, driving up incentives to enter the industry and bringing eventually the welfare enhancing intra-industry reallocations highlighted in \citet{melitz} framework.} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
@@
 
-Our paper provides an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
+Do additive costs matter?
+To answer this question, we provide an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
 To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
@@
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-Despite these differences, our results share with \citet{Irrazabal_2015} the important role of the additive component of international transport costs.\smallskip
+Despite these differences, our results share with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.\smallskip
 
@@
 \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} points out the importance of the composition effects in the characterization of time trends of international transport costs.
-Over 1974-2004, \cite{hummels2007} thus obtains a stronger decline in transportation costs when the composition effects by origin country and by product is removed, in both air and vessel transports.
-Our results tend to mitigate this view.
-In contrast to \cite{hummels2007}, we do not find that composition effects have offset the decrease of the ``ceteris paribus'' component of transport costs over time.
-They do not matter much in accounting for the decline of transport costs in air shipping; they even seem to have amplified the decrease in the pure transport costs in vessel shipping.
-Importantly, we show that the difference of results can be attributed to the role of the additive component, which modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time.\medskip
+However, \cite{hummels2007} points out the role of trade composition effects in the characterization of time trends of international transport costs.
+If we share this view with \cite{hummels2007}, our results about the underlying sources of the transport costs time trend strikingly differ though.
+In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs.
+Importantly, we show that the difference of results can be attributed to the modelling of the additive component.
+\medskip
 
@@
 
-First, our theoretically agnostic approach provides a fairly simple (if computationally heavy) framework for measuring both ad-valorem and additive parts of transportation costs.
-We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% of the export price in air and vessel transport respectively, whereas the additive component amounts to 1.8\% and 2.9\%.
+First, we show that additive costs are quantitatively sizeable.
+On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period).
+Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model.
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}  This represents a valuable insight for calibrating related international trade models.
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}.
+This represents a valuable insight for calibrating related international trade or business cycle models.
 
@@
 On average over the whole period, biased estimates for ad-valorem are 5\% for air and 6\% for vessel, while unbiased estimates are respectively 2.5\% and 3.2\%.
-We also show that taking additive costs into account significantly improves the fit of the modelling of transport costs, through various measures of ``goodness-of-fit'', for both transport modes and years.
-
-Third, we exploit the time dimension of our database to document the patterns of transport costs over time.
+Further, we obtain a systematically better goodness of fit when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
+
+Third, additive costs matter in accounting for the time trend of transport costs, as we show by exploiting the time dimension of our database to document the patterns of transport costs over time.
 Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in maritime shipping, and of same order of magnitude for both additive and iceberg components.
-Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs.
-Investigating this result further, we show that taking into account the additive component plays a key role in assessing the role of composition effects in the time trends of international transport costs.
-Allowing for additive costs that vary over time, product and partner country, we find that composition effects do not offset the decrease in the pure transport costs observed from 1974 in the time trend of overall transport costs.
-Put it differently, the reduction of the pure transport costs is not that strong once the role of additive costs is appropriately modeled.
+Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Importantly, we show that the difference of results can be attributed to the modelling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
 In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
@@
 It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
-Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap omits the other dimension of transport costs related to the time value of goods in transit.
+However, it is also clear that transport costs do represent a sizeable share of trade costs as pointed out in Section \ref{sec:Intro}.
+Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
 In this respect, this dataset embrace only a partial view of international transport barriers.
+\textbf{Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on the ``actual'' trade flows, i.e.
+ignoring those flows that did not happen because of presumably prohibitive transport costs.
+In light of this, the estimated values that we obtain can be viewed as the lower bound of the ``true'' transport costs.}
 
@@
 \paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p$ of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}_{isk}$ (export, or fas price), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
+To do so, we start from the equation that expresses the price $p$ of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}$ (export, or fas price), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
@@
 \noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
-Let us denote $i$ the origin country, and $k$, the product.
-Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
+Let us denote $i$ the origin country, and $k$, the product (at the 5-digit level).
+Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):\footnote{\textbf{Why do we consider the ratio not the difference, with cif price on the LHS and p fas RHS?
+To be answered?}}
 
@@
 \paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
 
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
 \end{equation}
@@
 One key objective of the paper is to characterize the importance of additive costs relatively to ad-valorem costs.
-A natural way to answer this question is to perform estimations of Equation (\ref{eq:estimatedequation}) constraining additive costs to be equal to zero, and compare the fitting properties and the explanatory power of the models.
-For each year and transport mode, we thus estimate the two models: \textit{(a)} when transport costs are modeled as iceberg costs (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are modeled as decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})):\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
+A natural way to answer this question is to perform estimations of Equation (\ref{eq:model_IetA}) constraining additive costs to be equal to zero, and compare the fitting properties and the explanatory power of the models.
+For each year and transport mode, we thus estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:estimatedequation})).
+Under Model (\textit{(a)}), the estimated equation simplifies as:\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
 This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
@@
 
-\begin{eqnarray}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI} \\
-%\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA}
-\end{eqnarray}
-
-One may be concerned that the specification of Equations (\ref{eq:model_nlI}) or (\ref{eq:model_IetA}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
+\end{equation}
+
+One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (A) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
+    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
     Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
@@
     %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
+    \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
    \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
@@
     %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}& & & &   \\ \hline
+    \textit{Additive term ($\widehat{t}/\widetilde{p}$, in \%)}& & & &   \\ \hline
     Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
@@
 
+\textbf{Mettre en avant mieux le resultat 1 sur size additifs, on ne se vend pas bien. Stop here.}
+
 Coming now to interpretation, Table \ref{tab:summary_results} shows that the magnitude of overall transport costs is sizeable.
-When modeled as ad-valorem costs only (Model (A)), they correspond to a mean increase of the export price by 5.8\% for vessel shipping, and by 5.1\% for air shipping over 1974-2013.
+When modeled as ad-valorem costs only (Model (b)), they correspond to a mean increase of the export price by 5.8\% for vessel shipping, and by 5.1\% for air shipping over 1974-2013.
 Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
@@
 
-%\section{Robustness checks \label{sec:robustness}}
-
-%Robustness to the specification of the estimated equation.
-
-%\paragraph{Robustness to the separability assumption}
-
-%Ie, estimate
-
-%$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
-
-%To do on 100 products ($s$), 50 countries (the most important?)
-
-%\todo{what else?}
+\section{Robustness checks \label{sec:robustness}}
+
+\subsection{Measuring transport costs components: Robustness to the separability assumption}
+
+Ie, estimate
+
+$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
+
+To do on 100 products ($s$), 50 countries (the most important?)
+
+\subsection{Time trends in transport costs: Robustness to the weighting scheme}
 
@@
 {\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_notitle.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_notitle.pdf} \\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_air.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_ocean.pdf} \\
 \end{tabular}

### 2017-09-11 redaction/dhp_july2017.tex
Nouvelle version du papier

@@
 \usepackage{amsfonts}
-%\usepackage{bbold}
+\usepackage{bbold}
 % JH : impossible de compiler avec le package bbold, remplacé par amsfonts
@@
 In this respect, this dataset embrace only a partial view of international transport barriers.
-\textbf{Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on the ``actual'' trade flows, i.e.
+\textbf{Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bound of the ``true'' transport costs.}
+In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.}
 
@@
 A natural way to answer this question is to perform estimations of Equation (\ref{eq:model_IetA}) constraining additive costs to be equal to zero, and compare the fitting properties and the explanatory power of the models.
-For each year and transport mode, we thus estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:estimatedequation})).
+For each year and transport mode, we thus estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
 Under Model (\textit{(a)}), the estimated equation simplifies as:\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
@@
 
-The objective of this section is twofold.
+Do additive costs matter?
+This section provides two elements of answer to this question.
 First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+This allows us to estimate whether additive costs represent a sizeable component of transport costs.
 Second, we assess the importance of the additive component of in overall transport costs through goodness-of-fit measures.
@@
 It displays the mean and median values of each type of trade costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
-The first panel reports the results for a specification based on ad-valorem costs estimated alone, while the second panel presents estimates for a specification involving both ad-valorem and multiplicative component.
+The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})).
 Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
-Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the results for air at the 3-digit level removing the year 1989, as the estimation results reveal the presence of strong outliers that bias the estimates of transport costs upwards.
+Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
 Overall results (over the whole period) are not substantially affected if this year is included though.
@@
 Table \ref{tab:summary_results} calls for two types of comments.
-Starting with pure statistical concerns, it appears that mean values are systematically higher than medians - by around one percentage point.
+The first type is of statistical order.
+As displayed in Table \ref{tab:summary_results}, mean values are systematically higher than medians, by around one percentage point.
 This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median.
-One can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point.
+One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
 This is first due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - Section \ref{sec:results_trends} will go into more details of these compositions effects.
@@
 Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
-Descriptive statistics showing that it is indeed the case are available upon request to the authors.
-
-\textbf{Mettre en avant mieux le resultat 1 sur size additifs, on ne se vend pas bien. Stop here.}
-
-Coming now to interpretation, Table \ref{tab:summary_results} shows that the magnitude of overall transport costs is sizeable.
-When modeled as ad-valorem costs only (Model (b)), they correspond to a mean increase of the export price by 5.8\% for vessel shipping, and by 5.1\% for air shipping over 1974-2013.
-Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable.
-Over the whole period, additive costs are 2.9\% of the export price for ocean shipping - with a lower estimated value for the ad-valorem freight rate of 3.2\%.
-These values are respectively equal to 1.8\% and 2.5\% for air shipping.
+Descriptive statistics showing that it is indeed the case are available upon request to the authors.\medskip
+
+Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}.
+First, it provides estimation results about the size of the overall transport costs.
+From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping in average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
+Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
+With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.
+\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons.
+\citet{anderson_wincoop_jel} obtain this 11\% markup as a rough value by aggregating various sources over a bunch of industrialized countries, with data mostly based on the second half of the 1990s.
+By contrast, our estimation of a 6\% markup relies on a single source of data based on the US imports flows over 1974-2013.
+Beyond country coverage and the time sample, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+
+Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
+Our methodology indeed allows us to go deeper into the structure of transport costs, that can be decomposed into an additive and an ad-valorem components.
+We thus find that the per-kg cost dimension is sizeable.
+On average over the whole period, additive costs are estimated to 2.9\% of the export price for ocean shipping, and 1.8\%.
+This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\/ and 1.8\% for vessel and air shipping respectively (on average over the period).
+Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
 In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-
-We extend further our analysis by using explicitly the time dimension of our dataset in order to assess the dynamics of our estimates over time.
-Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_IetA}), by transport mode.
-
-\begin{figure}[htbp]
-\caption{Ad-valorem Costs (Yearly mean value, 3 digits, in \% of the fas price)}
-\label{fig:mult_alone_withadd}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3in, height=2.5in]{Fig1a_mult_air_3d.pdf}
-& \includegraphics[width=3in,height=2.5in]{Fig1b_mult_vessel_3d.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds across years.
-Further, Figure \ref{fig:mult_alone_withadd} reports that for air shipping, the size of the bias seems to decrease over time, while it appears rather stable for vessel shipping.
-Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air shipping.
-To get a better picture on this, we study the shares of both ad-valorem and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.
+These elements point out that additive costs are quantitatively sizeable.
+
+We can use the time dimension of our dataset in order to assess the validity of this result on a year-to-year basis over the period.
+To this aim, Figure \ref{fig:decomp_TC_3d} reports the shares of both the ad-valorem costs (grey area) and the additive costs (black area) in total costs (the total of the two, expressed in percentage of the fas price), by transport mode for each year between 1974 and 2013.
 
@@
 As reported on Figure \ref{fig:decomp_TC_3d}, total transport costs have regularly declined over time, by an order of magnitude of 3 percentage points in air shipping and 5 percentage points in ocean shipping (basically, divided by 2 in both cases).
-The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who points out that overall transport cost for the United States declined from 6 \% to 4 \% between 1974 and 1996.
-For the same years, we obtain a total decrease from 6.9 to 4.2\% on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.} The fact that our estimates are on average higher than those by \citet{Hummels_1999} (in ocean shipping mostly) is not necessarily surprising, given that ours being simple averages, whereas \citet{Hummels_1999}'s are trade-weighted.
-Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.
-
-Our empirical strategy allows us to go beyond these results, by distinguishing within these evolutions between multiplicative and additive transport costs.
-First, the decreasing trend holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}).
-Second, in both transport modes, the additive component appears of sizeable importance.
-As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
-Further, the downward trend in overall transport costs over time can be attributed to both the additive and the ad-valorem components, with a share of additive costs roughly constant over time.
+At this point though, we postpone this discussion about the downward trend for transport costs for Section \ref{sec:results_trends}, to focus on the importance of the additive component.
+In both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time.
 These results thus suggest that additive costs are neither negligible in magnitude nor an erratic phenomenon.
 By contrast, they represent an sizeable and structural dimension of international transport costs.
-
-%\textbf{discuter du paragraphe qui suit. Je mets ma version et ensuite l'ancienne.}
-% JH: I checked Irrara et al, footnote 4 & 1er paragraphe de la section 4.4 Results page 21), il semble que ce soit Lise qui ait raison.
-
-At this point, let us compare our results to the literature.
-First, consider our estimation of the overall size of transport costs.
-From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
-Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
-With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons.
-\citet{anderson_wincoop_jel} obtain this 11\% markup as a rough value by aggregating various sources over a bunch of industrialized countries, with data mostly based on the second half of the 1990s.
-By contrast, our estimation of a 6\% markup relies on a single source of data based on the US imports flows over 1974-2013.
-Beyond country coverage, or the time sample, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
-
-Second, consider our results specific to the additive costs dimension.
-These can be compared to \citet{Irrazabal_2015}, which provide the most recent evidence on the size of the additive costs.
-More precisely, based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology.
-In contrast to them, we can provide separate estimates for each cost component (additive and ad-valorem), from which we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
-Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport.\footnote{On average over the whole period, from the figures in Table \ref{tab:summary_results}, we have $\frac{t}{\tau\widetilde{p}}= \frac{t\widetilde{p}}{\tau} = \frac{0.029}{1.032}= 2.8\%$ in vessel; and 1.8\% in air shipping.
-} This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}.
-However, while the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
+Further, one side result of our work is to provide quantitative estimates for the values of both additive and ad-valorem transport costs,which can be very useful for the related more theoretical papers, which need to calibrate their models.\medskip
+
+Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
+Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology.
+One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
+From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
+Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport.
+This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}.
+This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
+While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
 From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs equal to 2.1\% ($0.15\times 0.14$), which is in line with our own results.
@@
 In the next section, we analyze their importance through another angle based on quality-of-fit diagnostic tests.
-\bigskip
-
-%\textbf{ancienne version}
-
-%Before, we check how our estimates of additive costs compare to those of \citet{Irrazabal_2015}, which provide the most recent evidence on that matter.
-Based on Norwegian firm-level trade data for 2004, \citet{Irrazabal_2015} find that additive costs are on average 14\% of the median price.
-Expressed in our own terminology, this means they find that $\frac{t/\tau}{\widetilde{p}}=14\%$.
-Let us define the ratio of additive costs over total costs as $\frac{t}{\tau\widetilde{p}+t}$, and divide both numerators and denominators by $\tau\widetilde{p}$.
-We get: $\frac{t}{\tau\widetilde{p}+t}=\frac{t/\tau\widetilde{p}}{1+t/\tau\widetilde{p}}$.
-Plugging \citet{Irrazabal_2015} result gives 0,14/1+0,14= 12,5\%.
-Additive trade costs represents 12.5\% of total trade costs.
-Based on \citet{anderson_wincoop_jel}, we know that transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.
-Given that in 2004, our estimates for additive (multiplicative) transport costs is 2.9\% (2.7\%) for Vessel, the same reasoning as before leads to 2.9/(2.7+2.9)*11/74 = 7.7\% of total trade costs, that is 62\% (7.7/12.5=62\%) of \citet{Irrazabal_2015}'s estimates.
-For Air transportation, our estimates for additive (multiplicative) transport costs is 1.5\% (1.9\%), and the same computation brings that additive costs represent 6.6\% of total trade costs, that is 53\% of \citet{Irrazabal_2015}'s estimates.
-In other words, we find that additive transport costs represent between half and two thirds of the total additive trade costs, which seems totally consistent with the magnitudes reported by e.g.
-\citet{anderson_wincoop_jel}.
 
@@
 Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including the additive term, i.e.
-Equation (\ref{eq:model_IetA}) and the Restricted one (\emph{R}, i.e.
+Equation (\ref{eq:model_IetA})) and the Restricted one (\emph{R}, i.e.
 Equation (\ref{eq:model_nlI})).
@@
    \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-    Model (A) & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
-    Model (B)  & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
+    Model (A) & 35675.0 & 41171.0 & 60715.6 & 87492.6 & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{88191.9} & 70498.1 \\
+    Model (B)  & 31387.3 & 35738.4 & 52098.9 & 74954.9 & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{80873.7} & 62285.0 \\ \hline
     \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
@@
 In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
-Further, the same conclusion established at the 3-digit level regarding the significant role of the additive components in fitting international transport costs emerges at the 4-digit level.
-
-\section{Decomposing Transport Costs: Characterizing the trends }\label{sec:results_trends}
+Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.
+
+\section{Transport Costs time trends: The role of the additive component}\label{sec:results_trends}
 \subsection{Estimation strategy}
@@
 As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, that displays the respective shares of additive and iceberg components in transport costs over time by transport mode.
-Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974 (-2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs).\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974.
+Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
+In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974.
 However, computing the time trends from 1980 does not dramatically change the picture.
 The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.}
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996.
+For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
+Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.
 
@@
 Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As emphasized by \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
+As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
 per product- and per partner- transport costs.
 
-If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:estimatedequation})), that vary over time, product and origin country.
-Starting from these values, we extract the ``ceteris paribus'' transport cost component (additive and multiplicative, by transport mode) by the mean of a time fixed effect (see details below).
-By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``ceteris paribus'' additive costs (for instance, handling costs) that have reduced over time.
-We proceed similarly for the multiplicative component of trade costs.
-We also agglomerate the two components (additive and iceberg) in a unified measure of transport costs.
-This allows us to compare how ``ceteris paribus'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones.\footnote{Note that these ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}
-
-This empirical strategy differs from \cite{hummels2007} in two (interrelated) main dimensions.
-First, our characterization of the time trends in transport costs identifies the trend patterns of the transport costs (composition effects excluded) for both the additive and ad-valorem costs, as well as for the overall transport cost.
-Second, and as direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling in the time trend of transport costs, what comes from the composition effects and what comes the ``ceteris paribus'' transport cost dimension, as we show below.
-\footnote{Rigourously speaking, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
-We discuss this point with more details in Appendix \ref{app:comp-effects}.}
+If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
+Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below).
+For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time.
+We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}
 
 \subsection{Empirical specification}
-Our strategy to eliminate the composition effects follows the following rationale.
-As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in our first stage.
-We measure how these costs have changed over time by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+Our strategy to eliminate the composition effects obeys the following rationale.
+As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in oSection \ref{sec:results_decomposition}.
+We extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
 One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
@@
 This can be described as a two-stage process.
-First, we decompose the measure of transport costs component in the three product/country/time dimensions, using fixed effects.
+First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
 For the estimated ad-valorem component, we thus estimate the following equation:
@@
 
-\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.
-Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
-We cannot use the same equation for the estimated additive component, as by construction, the sector fixed effect and the country fixed effect are additive rather than multiplicative.
-We estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
+\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercice, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, ee estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
 Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
@@
 \end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: the country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus transport costs time trend'' (Term (c)).
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.\footnote{In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.} Both equations are estimated by transport mode.
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)).
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+Both equations are estimated by transport mode.
 
@@
 
-Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the ``raw'' (without composition effects excluded) and the ``ceteris paribus'' estimated transport costs (composition effects excluded, see Appendix \ref{app:comp-effects}).
+Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded).
+Details are reported in Appendix \ref{app:comp-effects}).
 
 \subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In Panels (a) and (b), we report the time changes of the ad-valorem costs and the additive costs respectively.
-Panel (c) displays the time changes in the total transport costs (built as the sum of the two components as detailed above).
-In the three cases, the series are reported as indices, starting from the reference value 100 in 1974 and by transport mode.
+Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{Figure \ref{fig:totalTC_compeffects_excl} reports the results, all types of goods included.
+In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} Panels (a), (b) and (c), we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
+Panels (d), (e) and (f) report the results for Vessel.
+In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
 
@@
 Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, echoing the related literature (see \citealp{Lafourcade_Thisse}), we find that international transport costs have substantially decreased over the period, and in both transport modes.
+First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes.
 International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
-Second, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
-Third, we do not find evidence of a significant difference between the raw transport costs and the ``ceteris paribus'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels).
+This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
+Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
+Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs.
+This is particularly true in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
 Composition effects are more pronounced in vessel transport.
-Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects.
-Precisely, the role of composition effects is significant for the multiplicative component, whereas the reduction in the additive costs is mostly attributable to changes in ``ceteris paribus'' additive costs.
-
-This last set of results stands in sharp contrast with \citet{hummels2007}, which obtained that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1973-2004), suggesting an important role to composition effects.
+Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects, in particular for the multiplicative component (panel (d)).\smallskip
+
+\subsection{Time trends in transport costs and the modeling of the additive component}
+
+This last result stands in sharp contrast with \citet{hummels2007}, which obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects.
 If anything, we find the opposite result here.
-Investigating this difference of result further (see Appendix \ref{app:comp-effects}), we find that this can be explained by the difference in methodology, and primarily the treatment of the additive component transport costs.
+This drives us to investigate this difference of result further.
+As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
+Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
+As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
+%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below.
+\footnote{As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
+Robustness analysis to this point is driven in Section \ref{sec:robustness}.} To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).
+
+\begin{figure}[htbp]
+\caption{Characterizing the time trends: Applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_air.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_ocean.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
+Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partially offset the decrease in transport costs in both air and vessel shipping (as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel).\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel)} This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
 Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
 In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in vessel transport.
+Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
 
@@
 \subsection{Time trends in transport costs: Robustness to the weighting scheme}
+
+In this section, we provide a robustness test to the weighting scheme implemented to extract the trade composition effects implemented in Section \ref{sec:results_trends}.
+When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
+In this section, our objective is to assess that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
+
+To this aim, we rebuild the fitted values for each of of our transport cost measures (ad-valorem, additive and overall) applying Hummels' s (2007) weighting scheme (ie, taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
+The results are reported in Figure \ref{fig:compeffects_robustness}.
+
+\begin{figure}[htbp]
+\caption{Transport cost time trends: Robustness to the weighting scheme}
+\label{fig:compeffects_robustness}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_composition_all_np.pdf}
+\end{center}
+\end{figure}
+
+The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives tow comments.
+First, differences do show up, in particular for the multiplicative component of Air transport.
+This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}.
+Rather than the weighing scheme, this confirms that the functional form and precisely the modeling of the additive component, turns out to be key in understanding the underlying determinants of the overall transport costs time trends, as we pointed out in Section \ref{sec:results_trends}.
 
@@
 \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)}}  \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
 Mean & 3.6 & 2.3 & 2.4 &1.7 & 2.6 & 1.7  \\
@@
 \hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}     \\ \hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
 Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0  \\
@@
 Precisely, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$), \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
+The variance decomposition expression at the product level $k$ (5-digit level) is obtained by applying the following formula (by year and transport mode):
+
+$$\underbrace{\sum_{k=1}^K \sum_{i=1}^I \left(x_{ij} - \bar{x}_g  \right)^2}_{\text{Total variability}} = \underbrace{\sum_k \sum_i \left(x_{ik} - \bar{x}_k  \right)^2}_{\text{Within-product variability}} + \underbrace{\sum_k \left(\bar{x}_{k} - \bar{x}_g  \right)^2}_{\text{Between-product variability}}$$
+
+with $x$ the observed cif-fas price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
+
+\begin{eqnarray*}
+\bar{x}_g \equiv \frac{1}{n} \sum_{k=1}^K \sum_{i=1}^I x_{ik},&& \bar{x}_k \equiv \frac{1}{K}\sum_{k=1}^K x_{ik}
+\end{eqnarray*}
+
+\noindent with $n$ the total number of observations, $K$ the total number of products and $I$ the total number of country partners.
+We apply the same variance decomposition exercise at the sector level, in which case the sector $s$ index (at the 3-digit level) replaces the $k$ index (at the 5-digit level).
 This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
-We also determine the share of the observed variance that can be attributed to the between-country variance.
+We also determine the share of the observed variance that can be attributed to the between-country variance, adapting the variance decomposition formula written above accordingly.
 Results are reported in Figure \ref{fig:decomp_variance}.
@@
 This delivers an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
-Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger that the between-country variance.
+Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
 This holds throughout the period and for both transport modes.
-This suggests that what primarily matters in international transport costs is mostly attributable to the product per se, rather than to the country where it comes from.
+This suggests that what primarily matters in international transport costs is mostly attributable to the product \textit{per se}, rather than to the country where it comes from.
 By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
@@
 \multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}  \\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%) }}   \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}$, in \%) }   \\ \hline
 Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
@@
 \hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$), in \%} }   \\ \hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$), in \%}    \\ \hline
 Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
@@
 \hline
-Mean  & 9.8 & 6.1 & 5. & 5.1 & 4.2 & 3.6  \\
+Mean  & 9.8 & 6.1 & 5.8 & 5.1 & 4.2 & 3.6  \\
 Median & 9.4 & 5.1 & 4.8 & 4.5 & 3.8 & 3.1 \\
@@
 \multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC} }\\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%) }   \\ \hline
 Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
@@
 \hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\ \hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}^{add}/\widetilde{p}$, in \%) }   \\ \hline
 Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\
@@
 %\scalebox{0.97}{
+  \footnotesize{
 \begin{tabular}{l|cccccc}
@@
 \textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
-Model (A)& \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Model (B) & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
+Model (A)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
+Model (B) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
 \textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A) & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87}  \\
-Model (B) & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14}  \\
+Model (A) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
+Model (B) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\
 \textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
-Model (A)& \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} \\
-Model (B) & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28}  \\
+Model (A)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
+Model (B) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\
 \textbf{Test LL} &       &       &       &       &       &       \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45}  \\
+2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
 \# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
+p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& ad-valorem
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& ad-valorem
 transport costs.
@@
 \end{tabular}%
+}
 \end{center}
@@
 \label{tab:good_fit_ves_rob}%
+  \footnotesize{
 \begin{tabular}{l|cccccc}
@@
 \textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A) &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (B) &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
+Model (A) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
+Model (B) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
 \textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A)& -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} \\
-Model (B) & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22}  \\
+Model (A)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
+Model (B) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\
 \textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A) & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23}\\
-Model (B) & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} \\
+Model (A) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
+Model (B) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\
 \textbf{Test LL} &       &       & & &  & \\
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.75 & \multicolumn{1}{c}{17354.65} & \multicolumn{1}{c}{20113.52} & \multicolumn{1}{c}{16608.16} & \multicolumn{1}{c}{12589.59} \\
+2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
 \# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& ad-valorem
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& ad-valorem
 transport costs.
@@
 
-In this section, we explain with more details the method employed to eliminate the country- and product- dimensions of the estimated transport cost.
+In this section, we explain in more details the method employed to eliminate the country- and product- dimensions of the estimated transport cost.
 
 \subsection{More on our methodology}
+
+\subsubsection{Additive and ad-valorem transport costs: Excluding the composition effects}
+
+In this section, we detail our methodology to extract the time trend in the ``ceteris paribus'' (or fitted) transport cost component, for each the ad-valorem and the additive component.
+
 \paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component. Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
@@
 Given that $\tau >1$, we can rewrite to get the percentage change between year 1974 and any year $t>1974$:
-$$\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}$$
-
-As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trade.
+\begin{equation*}
+\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}
+\end{equation*}
+
+As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trend.
 At this stage though, it remains specific to a product-origin country pair.
 Next step is to build the index $\Gamma^{adv}_t$ such that:
-\begin{equation*}
+\begin{equation}
  \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}  \label{eq:tcadv_compoeffect}
-\end{equation*}
+\end{equation}
 \noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
@@
 
-\begin{equation*}
-\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)
-\end{equation*}
+\begin{equation}
+\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t) \label{eq:indice_add}
+\end{equation}
 
 \noindent Note that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
-\begin{equation*}
-\Gamma^{add}_t  = 100\exp(\gamma_t)
-\end{equation*}
-
-that is, Equation (\ref{eq:comp_effects_add}).
-
-\paragraph{For the total cost measure} We also build a measure of the ``overall'' transport costs, that agglomerates our estimates of the two additive and iceberg components.
+\begin{equation}
+\Gamma^{add}_t  = 100\exp(\gamma_t) \label{eq:tcadd_compoeffect}
+\end{equation}
+
+\paragraph{Obtaining the unfitted measures} It consists in expressing the unfitted transport cost component as as index, with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
+That is, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
+
+$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
+
+\subsubsection{For the total cost measure}
+
+We also build a measure of the ``overall'' transport costs, that agglomerates our estimates of the two additive and iceberg components.
 We construct this measure (by transport mode) on both the unfitted series and the ``pure'' transport cost series (composition effects excluded).
@@
 
-We apply a slightly different procedure for constructing the fitted measure of total transport cost (i.e., composition effect excluded).
-In this case, we start directly from the two indices previously obtained (Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}), by transport mode).
-We build the ``overall'' transport cost measure as a weighted sum of the two indices, using the relative weight of the additive cost component in total cost in 1974, $\omega_{air}= 0.423$ and $\omega_{ocean}= 0.482$ (obtained from our estimates in Section \ref{sec:results_decomposition}, see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), according to:
-$$\Gamma^{tc, pure}_t = \omega_{x}\Gamma^{add}_t +(1-\omega_{x})\Gamma^{adv}_t $$
-
-\noindent with $x=\{air, vessel\}$.
-
-\subsection{Comparing with \cite{hummels2007}}
+We apply the same logic to construct the fitted measure of total transport cost (i.e., composition effect excluded), with a first step that consists in getting the fitted measures of each transport cost component in value (rather than as an index).
+For the ad-valorem component, this can be obtained starting from Equation (\ref{eq:tcadv_compoeffect}), rewritten to get:
+
+$$\widehat{\tau}^{cp, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
+
+with $\widehat{\tau}^{cp, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) ad-valorem cost.
+A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
+
+$$\widehat{t}^{cp}_t = \frac{\Gamma^{add}_t}{100}$$
+
+with $\widehat{t}^{cp}_t \equiv \exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) additive cost.
+
+We then deduce the fitted value of the overall cost according to:
+$$\widehat{tc}^{cp}_t= \widehat{\tau}^{cp, adv}_t  + \widehat{t}^{cp}_t$$
+
+Last, we transform this (fitted) ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{cp}_t -1 }{\widehat{tc}^{cp}_{1974}-1}$$
+
+This finally leaves us with six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and overall resp.), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and overall respectively).
+
+\subsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
 
@@
 We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
-This choice is made for to reasons.
+This choice is made for two reasons.
 First, this weighting scheme does not overweight the small flows in value.
@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).
-
-\begin{figure}[htbp]
-\caption{Characterizing the time trends: Applying Hummel's (2007) method }
-\label{fig:comp_effects_as_in_Hummels}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_air.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_ocean.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-As expected, these figures are strictly identical for pre-2004 years to Figures 5 and 6 in \citet{hummels2007}.
-More importantly, these results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we obtain that the composition effects tend to compensate the decrease in transport costs in both air and ocean shipping (unsurprisingly in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Ocean)).
-This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
-Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
 

### 2017-09-12 redaction/online_appendix_dhp2017.tex
modifs_cosmetiquesJH020817
3 fois rien + pb de compilation lié à l'installation de Miktex sur mon nouveau laptop
Chnagement annexe technique

@@
 \begin{document}
-\title{\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix (Not for publication)}
+\title{\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix}
 
@@
 
-\subsection{Additive-costs only Model \label{secoa:additive_only}}
+\subsection{Three models: Comparison \label{secoa:additive_only}}
 
@@
   \centering
-  \footnotesize{
-  \caption{Transport costs estimates : Summary\label{sec_oa:add_only}}
-  \begin{center}
-    \begin{tabular}{l|cc|cc}
+    \caption{Estimation results of the three models (Air, 3-digit level)}
+  \label{tab:3models_estimation_results_air}%
+    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
     \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
-    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{With only Additive Transport Costs} ($\widehat{t}^{add}$, in \%)}  \\ \hline
-    Mean  & 9.2 & 5.1 & 8.0 & 5.8 \\
-    Median & 5.7 & 2.4 & 5.5 & 2.5 \\ \hline
-
-       \multicolumn{5}{l}{\textbf{Data} ($p/\widetilde{p}$, in \%) } \\ \hline
-        Mean & 5.3 & 5.0& 5.6&3.9 \\
-        Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
-        \# obs. & 29279 & 28207 & 29317 & 27680 \\
-    \# origin country & 188 & 191 & 188 & 189 \\
-    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
+    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
+   $\#$ obs.
+& \multicolumn{1}{c}{14955} & \multicolumn{1}{c}{16118} & \multicolumn{1}{c}{19908} & \multicolumn{1}{c}{24958} & \multicolumn{1}{c}{31037} & \multicolumn{1}{c}{35027} & \multicolumn{1}{c}{41806} & \multicolumn{1}{c}{40279} & \multicolumn{1}{c}{39351} \\
+   $\#$  origin countries & \multicolumn{1}{c}{152} & \multicolumn{1}{c}{165} & \multicolumn{1}{c}{169} & \multicolumn{1}{c}{181} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{211} & \multicolumn{1}{c}{210} & \multicolumn{1}{c}{210} \\
+   $\#$  sectors & \multicolumn{1}{c}{203} & \multicolumn{1}{c}{204} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{212} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{218} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{216} & \multicolumn{1}{c}{212} \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
+    Mean (in \%) & \multicolumn{1}{c}{6.93} & \multicolumn{1}{c}{5.41} & \multicolumn{1}{c}{6.08} & \multicolumn{1}{c}{5.03} & \multicolumn{1}{c}{4.61} & \multicolumn{1}{c}{3.60} & \multicolumn{1}{c}{4.10} & \multicolumn{1}{c}{4.19} & \multicolumn{1}{c}{3.36} \\
+    Median (in \%)& \multicolumn{1}{c}{5.43} & \multicolumn{1}{c}{3.79} & \multicolumn{1}{c}{5.47} & \multicolumn{1}{c}{4.37} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{2.47} & \multicolumn{1}{c}{3.12} & \multicolumn{1}{c}{3.41} & \multicolumn{1}{c}{2.92} \\
+    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (B) - With additive and ad-valorem transport costs}} \\ \hline
+     \multicolumn{10}{l}{ \textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
+    Mean (in \%) & \multicolumn{1}{c}{3.64} & \multicolumn{1}{c}{2.32} & \multicolumn{1}{c}{2.46} & \multicolumn{1}{c}{2.38} & \multicolumn{1}{c}{2.05} & \multicolumn{1}{c}{1.66} & \multicolumn{1}{c}{2.00} & \multicolumn{1}{c}{2.57} & \multicolumn{1}{c}{1.70} \\
+    Median (in \%)& \multicolumn{1}{c}{2.71} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{1.79} & \multicolumn{1}{c}{1.60} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.20} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{2.24} & \multicolumn{1}{c}{1.72} \\
+    Std & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline
+      \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)} \\ \hline
+    Mean (in \%) & \multicolumn{1}{c}{2.56} & \multicolumn{1}{c}{2.04} & \multicolumn{1}{c}{2.83} & \multicolumn{1}{c}{1.83} & \multicolumn{1}{c}{1.64} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{1.43} & \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{1.01} \\
+    Median (in \%)& \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.43} & \multicolumn{1}{c}{0.47} \\
+    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
+    Mean (in \%) & \multicolumn{1}{c}{6.88} & \multicolumn{1}{c}{4.82} & \multicolumn{1}{c}{5.99} & \multicolumn{1}{c}{4.44} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{3.09} & \multicolumn{1}{c}{4.00} & \multicolumn{1}{c}{4.49} & \multicolumn{1}{c}{3.29} \\
+    Median (in \%)& \multicolumn{1}{c}{4.45} & \multicolumn{1}{c}{1.82} & \multicolumn{1}{c}{3.36} & \multicolumn{1}{c}{2.27} & \multicolumn{1}{c}{1.72} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.93} & \multicolumn{1}{c}{2.72} & \multicolumn{1}{c}{2.06} \\
+    Std & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline
     \end{tabular}
-    \end{center}}
-\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
-The additive term is expressed in fraction of fas price.
-($^\ast$): Four 4-digit estimation: on selected years.
-($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
 \end{table}%
 
-%
-%
-%
-%
-%			SECTION B / DETAILED RESULTS
-%			
-%			
-%				
-%
-%
+\begin{landscape}
+\begin{table}[htbp]
+  \centering
+  \footnotesize{
+  \caption{Quality-of-fit diagnostic tests of the three models (Air, 3-digit level)}
+    \label{tab:3models_diagnosis_air}%
+   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \hline \hline
+    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
+    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
+    Model (A) & \multicolumn{1}{c}{0.297} & \multicolumn{1}{c}{0.267} & \multicolumn{1}{c}{0.302} & \multicolumn{1}{c}{0.251} & \multicolumn{1}{c}{0.142} & \multicolumn{1}{c}{0.318} & \multicolumn{1}{c}{0.460} & \multicolumn{1}{c}{0.421} & \multicolumn{1}{c}{0.313} \\
+    Model (B ) & \multicolumn{1}{c}{0.594} & \multicolumn{1}{c}{0.646} & \multicolumn{1}{c}{0.635} & \multicolumn{1}{c}{0.627} & \multicolumn{1}{c}{0.658} & \multicolumn{1}{c}{0.640} & \multicolumn{1}{c}{0.593} & \multicolumn{1}{c}{0.513} & \multicolumn{1}{c}{0.419} \\
+    Model (C) & \multicolumn{1}{c}{0.489} & \multicolumn{1}{c}{0.543} & \multicolumn{1}{c}{0.531} & \multicolumn{1}{c}{0.517} & \multicolumn{1}{c}{0.546} & \multicolumn{1}{c}{0.518} & \multicolumn{1}{c}{0.464} & \multicolumn{1}{c}{0.339} & \multicolumn{1}{c}{0.295} \\ \hline
+    \multicolumn{10}{l}{\textbf{SER}}\\ \hline
+    Model (A) & \multicolumn{1}{c}{0.791} & \multicolumn{1}{c}{0.860} & \multicolumn{1}{c}{0.831} & \multicolumn{1}{c}{0.811} & \multicolumn{1}{c}{0.798} & \multicolumn{1}{c}{0.844} & \multicolumn{1}{c}{0.837} & \multicolumn{1}{c}{0.857} & \multicolumn{1}{c}{0.920} \\
+    Model (B) & \multicolumn{1}{c}{0.674} & \multicolumn{1}{c}{0.715} & \multicolumn{1}{c}{0.692} & \multicolumn{1}{c}{0.675} & \multicolumn{1}{c}{0.641} & \multicolumn{1}{c}{0.697} & \multicolumn{1}{c}{0.727} & \multicolumn{1}{c}{0.787} & \multicolumn{1}{c}{0.847} \\
+    Model (C) & \multicolumn{1}{c}{1.610} & \multicolumn{1}{c}{1.778} & \multicolumn{1}{c}{1.736} & \multicolumn{1}{c}{1.699} & \multicolumn{1}{c}{1.700} & \multicolumn{1}{c}{1.786} & \multicolumn{1}{c}{1.783} & \multicolumn{1}{c}{1.776} & \multicolumn{1}{c}{1.723} \\ \hline
+   \multicolumn{10}{l}{ \textbf{AIC criteria}} \\ \hline
+    Model (A) & \multicolumn{1}{c}{35675.0} & \multicolumn{1}{c}{41171.0} & \multicolumn{1}{c}{49315.0} & \multicolumn{1}{c}{60715.6} & \multicolumn{1}{c}{74386.4} & \multicolumn{1}{c}{87492.5} & \multicolumn{1}{c}{103983.0} & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{106130.6} \\
+    Model (B ) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\
+    Model (C) & \multicolumn{1}{c}{40808.1} & \multicolumn{1}{c}{45138.5} & \multicolumn{1}{c}{55214.8} & \multicolumn{1}{c}{69458.5} & \multicolumn{1}{c}{83958.6} & \multicolumn{1}{c}{100040.8} & \multicolumn{1}{c}{123592.1} & \multicolumn{1}{c}{129359.0} & \multicolumn{1}{c}{127399.2} \\ \hline
+    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
+    Model (A) & \multicolumn{1}{c}{-17530.5} & \multicolumn{1}{c}{-20253.5} & \multicolumn{1}{c}{-24315.5} & \multicolumn{1}{c}{-29977.8} & \multicolumn{1}{c}{-36811.2} & \multicolumn{1}{c}{-43341.3} & \multicolumn{1}{c}{-51648.5} & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-52690.3} \\
+    Model (B ) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\
+    Model (C) & \multicolumn{1}{c}{-20074.1} & \multicolumn{1}{c}{-22217.2} & \multicolumn{1}{c}{-27251.4} & \multicolumn{1}{c}{-34355.3} & \multicolumn{1}{c}{-41634.3} & \multicolumn{1}{c}{-49625.4} & \multicolumn{1}{c}{-61533.0} & \multicolumn{1}{c}{-64339.5} & \multicolumn{1}{c}{-63316.6} \\ \hline
+    \multicolumn{10}{l}{\textbf{Test LL} }\\ \hline
+    Stat LL ratio (B vs A) & \multicolumn{1}{c}{4809.7} & \multicolumn{1}{c}{5980.6} & \multicolumn{1}{c}{7257.3} & \multicolumn{1}{c}{9168.7} & \multicolumn{1}{c}{13548.7} & \multicolumn{1}{c}{13105.7} & \multicolumn{1}{c}{11760.4} & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{6541.2} \\
+    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
+    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline
+    Stat LL ratio (B vs C) & \multicolumn{1}{c}{4948.4} & \multicolumn{1}{c}{4954.0} & \multicolumn{1}{c}{6564.5} & \multicolumn{1}{c}{8961.8} & \multicolumn{1}{c}{11597.4} & \multicolumn{1}{c}{12837.0} & \multicolumn{1}{c}{15764.7} & \multicolumn{1}{c}{17062.0} & \multicolumn{1}{c}{13896.9} \\
+    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
+    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline \hline
+\end{tabular}%
+}
+\end{table}%
+\end{landscape}
 
-\newpage
+\begin{table}[htbp]
+  \centering
+    \caption{Estimation results of the three models (Vessel, 3-digit level)}
+  \label{tab:3models_estimation_results_vessel}%
+    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \hline \hline
+    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
+    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
+    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
+    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
+    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
+    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
+    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
+     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
+    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
+    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
+    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
+       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
+    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
+    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
+    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
+    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
+    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
+    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
+   \end{tabular}%
+\end{table}%
+
+\begin{landscape}
+\begin{table}[htbp]
+  \centering
+  \footnotesize{
+  \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
+    \label{tab:3models_diagnosis_air}%
+   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \hline \hline
+    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
+    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
+    Model (A) & 0,450 & 0,415 & 0,427 & 0,456 & 0,438 & 0,401 & 0,378 & 0,350 & 0,339 \\
+    Model (B) & 0,612 & 0,575 & 0,571 & 0,590 & 0,611 & 0,571 & 0,541 & 0,491 & 0,462 \\
+    Model (C ) & 0,424 & 0,401 & 0,374 & 0,429 & 0,456 & 0,431 & 0,417 & 0,358 & 0,349 \\ \hline
+    \multicolumn{10}{l}{\textbf{SER} } \\ \hline
+    Model (A) & 0,576 & 0,620 & 0,569 & 0,592 & 0,615 & 0,652 & 0,673 & 0,740 & 0,758 \\
+    Model (B) & 0,484 & 0,528 & 0,493 & 0,514 & 0,512 & 0,551 & 0,578 & 0,656 & 0,684 \\
+    Model (C ) & 1,271 & 1,339 & 1,283 & 1,326 & 1,302 & 1,319 & 1,336 & 1,392 & 1,410 \\ \hline
+    \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
+    Model (A) & 33328,81 & 33010,27 & 40275,70 & 51142,62 & 60414,92 & 71365,89 & 85051,02 & 84789,89 & 88191,87 \\
+    Model (B) & 27331,52 & 28067,31 & 34170,52 & 43664,74 & 49275,33 & 60475,91 & 73020,09 & 76161,33 & 80873,72 \\
+    Model (C ) & 46082,40 & 44370,26 & 58829,71 & 71461,52 & 77052,41 & 88746,51 & 103310,93 & 101166,91 & 104290,27 \\ \hline
+    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
+    Model (A) & -16287,40 & -16129,13 & -19767,85 & -25169,31 & -29790,46 & -35263,95 & -42122,51 & -41998,95 & -43692,93 \\
+    Model (B) & -12985,76 & -13353,65 & -16398,26 & -21171,37 & -23905,66 & -29490,96 & -35844,04 & -37418,66 & -39751,86 \\
+    Model (C ) & -22674,20 & -21814,13 & -29045,86 & -35403,76 & -38125,20 & -43963,25 & -51245,46 & -50348,45 & -51783,14 \\ \hline
+   \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
+    Stat LL ratio (B vs A) & 6603,28 & 5550,96 & 6739,18 & 7995,88 & 11769,59 & 11545,98 & 12556,94 & 9160,56 & 7882,15 \\
+    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
+    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\ \hline
+    Stat LL ratio (B vs C) & 19376,88 & 16920,95 & 25295,20 & 28464,79 & 28439,08 & 28944,59 & 30802,84 & 25859,58 & 24062,55 \\
+    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
+    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\
+\hline \hline
+    \end{tabular}%
+}
+\end{table}%
+\end{landscape}
+
 %\subsection{Detailed Results \label{secoa:detailed}}

### 2017-09-13 redaction/dhp_july2017.tex
Nouvelle version du papier et de l'annexe online

@@
 Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
-Results for all years (available at the 3-digit level) are available upon request to the authors.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
+Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
 Overall results (over the whole period) are not substantially affected if this year is included though.
-These results are available upon request to the authors.}
+This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}
 
@@
 
-\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercice, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
+\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
 As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, ee estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
@@
 Second, considering the 1974 weighting scheme makes sense as starting point of our time trend analysis.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007} (Results available upon request to the authors).
-
-%\textbf{These results are reported in ???}.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in Section \ref{sec:robustness}.
 

### 2017-09-13 redaction/online_appendix_dhp2017.tex
Nouvelle version du papier et de l'annexe online

@@
 
+In the paper, we compare the empirical performances of two models, where only ad-valorem costs are modeled (Model (A)) and where both ad-valorem and additive costs are modeled (Model (B)).
+One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs (Model (C)).
+This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
+The results are reported in this section.
+We start reporting the estimates of the transport costs components under the three models.
+In a second step, we report the various tests of quality of fit implemented on these models.
+
+\subsubsection{Estimation results}
+
+In this section, we report the estimation results of the three models.
+Results for Models (A) (ad-valorem costs only) and (B) (both ad-valorem and additive costs) are identical to those reported in the paper.
+We also report the estimation results for Model (C), where only additive costs are modeled.
+Table \ref{tab:3models_estimation_results_air} reports the results for Air transport, Table \ref{tab:3models_estimation_results_vessel} for maritime transport.
+
 %%%% RAW TABLE TO BE FOUND AT: "...\Dropbox\Papier_Lise_Guillaume\trade_cost\results\3_models"
@@
   \centering
+   \footnotesize{
     \caption{Estimation results of the three models (Air, 3-digit level)}
@@
     \end{tabular}
+    }
 \end{table}%
+
+\begin{table}[htbp]
+  \centering
+   \footnotesize{
+    \caption{Estimation results of the three models (Vessel, 3-digit level)}
+  \label{tab:3models_estimation_results_vessel}%
+    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \hline \hline
+    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
+    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
+    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
+    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
+    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
+    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
+    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
+     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
+    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
+    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
+    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
+       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
+    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
+    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
+    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
+    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
+    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
+    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
+   \end{tabular}%
+   }
+\end{table}%
+
+Somehow unsurprisingly, the estimated size of transport costs under Model (C) is of same order of magnitude than when transport costs are modeled as ad-valorem alone (Model (A) or of both types (Model (B)).
+As reported in Tables \ref{tab:3models_estimation_results_air} and \ref{tab:3models_estimation_results_vessel}, we also get the downward trend of transport costs over time, in particular since 1980.
+In this respect, it is necessary to go further to characterize the empirical relevance of the model with addiitve costs only (Model (C)), by the mean of quality of fit diagnostic tests.
+This is reported in the next section.
+
+\subsubsection{Quality of fit diagnostic tests}
+
+In this section, we report the results various quality of fit diagnostic tests implemented on the three models.
+Table \ref{tab:3models_diagnosis_air} reports the results for Air transport, those for maritime transport being displayed in Table \ref{tab:3models_diagnosis_vessel}.
+In both tables, we report the values of the $R^2$, the Standard Error of Regression (SER), the AIC criterion  and the log-likelihood (LL) value for each of the three models.
+We also report the value of the log-likelihood ratio, that tests the quality of fit of the global model (Model (B)) versus either the truncated model with only iceberg cost (Model (A)) or with only additive costs (Model (C)).
 
@@
 
-\begin{table}[htbp]
-  \centering
-    \caption{Estimation results of the three models (Vessel, 3-digit level)}
-  \label{tab:3models_estimation_results_vessel}%
-    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
-    \hline \hline
-    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
-    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
-    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
-    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
-    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
-    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
-    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
-     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
-    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
-    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
-    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
-       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
-    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
-    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
-    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
-    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
-    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
-    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
-   \end{tabular}%
-\end{table}%
-
 \begin{landscape}
@@
   \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
-    \label{tab:3models_diagnosis_air}%
+    \label{tab:3models_diagnosis_vessel}%
    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
@@
     \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
-    Model (A) & 0,450 & 0,415 & 0,427 & 0,456 & 0,438 & 0,401 & 0,378 & 0,350 & 0,339 \\
-    Model (B) & 0,612 & 0,575 & 0,571 & 0,590 & 0,611 & 0,571 & 0,541 & 0,491 & 0,462 \\
-    Model (C ) & 0,424 & 0,401 & 0,374 & 0,429 & 0,456 & 0,431 & 0,417 & 0,358 & 0,349 \\ \hline
+    Model (A) & 0.450 & 0.415 & 0.427 & 0.456 & 0.438 & 0.401 & 0.378 & 0.350 & 0.339 \\
+    Model (B) & 0.612 & 0.575 & 0.571 & 0.590 & 0.611 & 0.571 & 0.541 & 0.491 & 0.462 \\
+    Model (C ) & 0.424 & 0.401 & 0.374 & 0.429 & 0.456 & 0.431 & 0.417 & 0.358 & 0.349 \\ \hline
     \multicolumn{10}{l}{\textbf{SER} } \\ \hline
-    Model (A) & 0,576 & 0,620 & 0,569 & 0,592 & 0,615 & 0,652 & 0,673 & 0,740 & 0,758 \\
-    Model (B) & 0,484 & 0,528 & 0,493 & 0,514 & 0,512 & 0,551 & 0,578 & 0,656 & 0,684 \\
-    Model (C ) & 1,271 & 1,339 & 1,283 & 1,326 & 1,302 & 1,319 & 1,336 & 1,392 & 1,410 \\ \hline
+    Model (A) & 0.576 & 0.620 & 0.569 & 0.592 & 0.615 & 0.652 & 0.673 & 0.740 & 0.758 \\
+    Model (B) & 0.484 & 0.528 & 0.493 & 0.514 & 0.512 & 0.551 & 0.578 & 0.656 & 0.684 \\
+    Model (C ) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
     \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
-    Model (A) & 33328,81 & 33010,27 & 40275,70 & 51142,62 & 60414,92 & 71365,89 & 85051,02 & 84789,89 & 88191,87 \\
-    Model (B) & 27331,52 & 28067,31 & 34170,52 & 43664,74 & 49275,33 & 60475,91 & 73020,09 & 76161,33 & 80873,72 \\
-    Model (C ) & 46082,40 & 44370,26 & 58829,71 & 71461,52 & 77052,41 & 88746,51 & 103310,93 & 101166,91 & 104290,27 \\ \hline
+    Model (A) & 33328.81 & 33010.27 & 40275.70 & 51142.62 & 60414.92 & 71365.89 & 85051.02 & 84789.89 & 88191.87 \\
+    Model (B) & 27331.52 & 28067.31 & 34170.52 & 43664.74 & 49275.33 & 60475.91 & 73020.09 & 76161.33 & 80873.72 \\
+    Model (C ) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
     \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
-    Model (A) & -16287,40 & -16129,13 & -19767,85 & -25169,31 & -29790,46 & -35263,95 & -42122,51 & -41998,95 & -43692,93 \\
-    Model (B) & -12985,76 & -13353,65 & -16398,26 & -21171,37 & -23905,66 & -29490,96 & -35844,04 & -37418,66 & -39751,86 \\
-    Model (C ) & -22674,20 & -21814,13 & -29045,86 & -35403,76 & -38125,20 & -43963,25 & -51245,46 & -50348,45 & -51783,14 \\ \hline
+    Model (A) & -16287.40 & -16129.13 & -19767.85 & -25169.31 & -29790.46 & -35263.95 & -42122.51 & -41998.95 & -43692.93 \\
+    Model (B) & -12985.76 & -13353.65 & -16398.26 & -21171.37 & -23905.66 & -29490.96 & -35844.04 & -37418.66 & -39751.86 \\
+    Model (C ) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
    \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
-    Stat LL ratio (B vs A) & 6603,28 & 5550,96 & 6739,18 & 7995,88 & 11769,59 & 11545,98 & 12556,94 & 9160,56 & 7882,15 \\
+    Stat LL ratio (B vs A) & 6603.28 & 5550.96 & 6739.18 & 7995.88 & 11769.59 & 11545.98 & 12556.94 & 9160.56 & 7882.15 \\
     $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
@@
 
+The main result that emerges is that the model with additive costs only (Model (C)) is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself dominated by the complete model (Model (B)), whatever the type of diagnostic test considered.
+
 %\subsection{Detailed Results \label{secoa:detailed}}
@@
 \subsection{Transport Cost Estimates: Yearly Detailed Results}
+In this section, we report the detailed results of the estimation driven at the 3-digit classification level.
+Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for Air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for Vessel.
+In both cases, we report the estimated values of the transport costs (weighed mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).
 
-\subsubsection{3-Digits Level Product Classification}
 \begin{landscape}
@@
 
+As mentioned in the paper, the estimates for Air transport costs in 1989 lead to a surprising high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000.
+This can be attributed to the presence of outliers in the distribution of the additive costs estimates, with a maximum value for $\widehat{t}/\widetilde{p}$ = 10,000\%, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000.
+Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in Air transport.
+One can yet mention that this does not make much difference.
+When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component ($\widehat{\tau}^{adv} = 1.9\%$, vs 1.8\% when 1989 is excluded.
+The weighed median value is left unchanged, as the additive component (with weighed mean and median values of $\widehat{t}/\widetilde{p} = 2.5\%$ and 1.8\%, whether 1989 is included or not.
+
 \begin{landscape}
@@
 
-\subsection{Eliminating the composition effects: Primary vs. Maufacturing sector} \label{sec_oa:comp-effects}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{C.\arabic{figure}}
 
-%\todo{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
+\subsection{Eliminating the composition effects: Primary vs. Manufacturing sector} \label{sec_oa:comp-effects}
 
-In this Section, we characterize the time trend in international transport costs at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.
-The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
+In this section, we refine the characterization of the trend patterns of international transport costs, by distinguishing the trade flows for primary goods and manufactured goods.
+The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
+For sake of reading clarity, we also report the results obtained on the whole range of trade flows, in Figure \ref{fig:totalTC_compeffects_excl} (i.e., Figure 2 of the paper).
+
+The classification retained to categorize trade flows follows the UNCTAD classification (on STIC Revision 3).
+Are considered as ``primary goods'' all flows recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).
 
@@
 \begin{center}
-\includegraphics[height=4in]
+\includegraphics[height=8cm]
 {graph_composition_manuf.pdf}
@@
 \begin{center}
-\includegraphics[height=4in]
+\includegraphics[height=8cm]
 {graph_composition_primary.pdf}
@@
 \begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects)}
+\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_composition_all.pdf}
+\end{center}
+\end{figure}
+
+As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, the ``raw'' transport costs are (or the unfitted transport costs, in plain blue line) have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in Air, 60\% in Vessel).
+However, contrasting these two broad sectors conveys a contrasted message about the role of trade composition effects in accounting for this trend pattern.
+Comparing Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
+In Air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the dotted red line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
+Trade composition effects matter more in maritime transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)), primarily in the ad-valorem component.
+Similarly as the conclusion drawn on the whole range of flows, the 60\% decrease in the unfitted transport costs in Vessel can de decomposed in a 50\% decrease in the ``ceteris paribus'' transport costs (fitted), the 10\% remaining to trade composition effects.
+
+The message is strikingly different when the same decomposition is driven for primary goods only.
+In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in maritime transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)).
+In air transport further, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).
+
+One way to reconcile the contrasted results displayed in Figure \ref{fig:totalTC_compeffects_excl_primary} (primary goods), with respect to Figure \ref{fig:totalTC_compeffects_excl_manuf} (manufacturing) and Figure \ref{fig:totalTC_compeffects_excl} (all flows) is to investigate the share of primary goods in total flows.
+The results are reported in Figure \ref{fig:Share_prim_goods}.
+In Air transport, the share of primary goods in the total value of US imports is negligible, by around 10\% all over the period from 1974 to 2004.
+Primary goods make a higher proportion of trade flows in maritime transport, especially over 1974-1982 (between 40\% and 60\%).
+On the following sub-period though, their share has fallen to 20-30\%.
+Given the modest proportion of primary goods in total imports flows of the US economy, in favor of the manufactured sector, it is hence not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered, is driven by the trend patterns that occur within the manufacturing sector.
+
+\begin{figure}[htbp]
 \caption{Share of primary goods in the value of total US imports}
@@
 \begin{center}
-\includegraphics[height=4in]
+\includegraphics[height=6cm]
 {Share_of_primary.pdf}
@@
 
-%\textbf{A commenter. A mettre en Online Appendix?}
-
-[TO BE COMPLETED]
-
 \end{document}
-
-\subsubsection{4-Digits Level Product Classification}
-
-\begin{landscape}
-\begin{table}[htbp]
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-\caption{Air: Transport costs estimates, all years, 4-digit}
-\begin{center}
-\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
-\hline
-\hline
-
-%%%%%%%%%%%%%%%%2nd part of the table
-
-\hline\hline
-\multicolumn{20}{c}{ }  \\
-\multicolumn{20}{c}{Continued}  \\
-\hline\hline
-
-\hline
-\hline
-\end{tabular}}
-
-\end{center}
-\label{tab_oa:result_vessel_ally3}%
-\end{table}
-\end{landscape}

### 2017-09-15 redaction/dhp_july2017.tex
relecture après DEGIT et modifs Lise

@@
 
-\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, and participants at seminars at University of California-Irvine and Paris-Dauphine for very useful discussions.
+\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars at University of California-Irvine and Paris-Dauphine for very useful discussions.
 Finally, any remaining errors are ours.
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
-SciencesPo, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France } }
-
-\date{July 2017\\PRELIMINARY AND INCOMPLETE \\ DO NOT QUOTE WITHOUT PERMISSION}
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \& Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }
+
+%\date{July 2017\\PRELIMINARY AND INCOMPLETE \\ DO NOT QUOTE WITHOUT PERMISSION}
+
+\date{September 2017}
  \maketitle
@@
 
-Our findings can be summarized in three main points.
+Our three main findings can be summarized as follows.
 
@@
 The paper is built as follows.
-\ref{sec:data_method} presents our data and our  methodology.
+Section \ref{sec:data_method} presents our data and our methodology.
 In Section \ref{sec:results_decomposition}, we provide results about the size and the role of the additive component of transport costs.
@@
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
-In this respect, this dataset embrace only a partial view of international transport barriers.
+In this respect, this dataset embraces only a partial view of international transport barriers.
 \textbf{Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.}
+In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.} %JH: OK, parfaitement clair pour moi !
 
@@
 Let us denote $i$ the origin country, and $k$, the product (at the 5-digit level).
-Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):\footnote{\textbf{Why do we consider the ratio not the difference, with cif price on the LHS and p fas RHS?
-To be answered?}}
+Transforming the above equation (\ref{eq:base}) as ratio in order to get the percentage change in prices induced by transportation, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
+%\footnote{\textbf{Why do we consider the ratio not the difference, with cif price on the LHS and p fas RHS? To be answered?}}
 
@@
 Our strategy to eliminate the composition effects obeys the following rationale.
-As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in oSection \ref{sec:results_decomposition}.
+As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}.
 We extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.

### 2017-10-06 redaction/dhp_july2017.tex
Lashkaripour and others
Modification de l'intro pour prendre en compte Lashkaripour (2017) + neutralisation de la section séparabilité.

@@
 This paper features an online appendix containing additional results and available on the authors' websites.}}
-%We also thank the EEA 27\textsuperscript{th} Congress (Oslo), , INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and .
 \author{Guillaume \textsc{Daudin}\thanks{%
@@
 
-\date{September 2017}
+\date{October 2017}
  \maketitle
@@
 Following \citet{samuelson1954}, standard models of international trade have usually modelled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, many common empirical facts support the existence of additive costs (i.e., cost per unit exported).
-As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
-often exhibit (at least partly) an additive structure.
-The structure (additive vs ad-valorem) of transport costs is far from being anecdotal.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, the structure (additive vs ad-valorem) of transport costs is far from being anecdotal.
 The literature has long emphasized its role in shaping the pattern of trade flows.
@@
 In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
-\citet{Lashkaripour-mimeo-2017} challenges this view.
-He finds supporting evidence for the \emph{ad-valorem} assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport.
-Yet, one can be concerned by the generality of this result.
-\citet{Lashkaripour-mimeo-2017}'s study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports).
-Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
-Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international trade costs.
+This is all the more important since many common empirical facts support the existence of additive costs (i.e., cost per unit exported).
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.
@@
 These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
-\smallskip
-
+Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs are significant, representing 14\% of the median price.
+In contrast, highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
+The positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}, \citet{Lashkaripour-mimeo-2017} finds, based on firm-transaction-level data on Colombian imports, an average transport cost elasticity of 0.9, supporting the \emph{ad-valorem} assumption.
+The result if further confirmed on more aggregate US, sector-level trade data for the period 1989-1994.
+Based on very similar data, but for a longer a period of time, our own estimates suggest however a more important part for additive costs, consistently with \citet{Irrazabal_2015}'s findings at the firm-level.
+
+%\citet{Lashkaripour-mimeo-2017} takes into account the fact  
+%Yet, one can be concerned by the generality of this result.
+The positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
+Measurement error in Columbian data??
 Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
@@
 Section \ref{sec:results_trends} characterizes the trends of transport costs (by transport mode) over the period 1974-2013.
-Section \ref{sec:robustness} provides some robustness checks of our results.
+Section \ref{sec:robustness} provides some additional robustness checks of our results.
 Section \ref{sec:conclu} concludes.
@@
 In this respect, this dataset embraces only a partial view of international transport barriers.
-\textbf{Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
+Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.} %JH: OK, parfaitement clair pour moi !
+In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.
+%JH: OK, parfaitement clair pour moi !
 
@@
 
-\section{Robustness checks \label{sec:robustness}}
-
-\subsection{Measuring transport costs components: Robustness to the separability assumption}
-
-Ie, estimate
-
-$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
-
-To do on 100 products ($s$), 50 countries (the most important?)
-
-\subsection{Time trends in transport costs: Robustness to the weighting scheme}
+%\section{Robustness checks \label{sec:robustness}}
+
+%\subsection{Measuring transport costs components: Robustness to the separability assumption}
+
+%Ie, estimate
+
+%$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
+
+%To do on 100 products ($s$), 50 countries (the most important?)
+
+\section{Time trends in transport costs: Robustness to the weighting scheme}
 

### 2017-10-11 redaction/dhp_july2017.tex
Paragraphe_Lashkaripour
Modif supplémentaire du paragraphe autour de Lashkaripour (2017), suite eux échanges avec Lise.

@@
 In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
-This is all the more important since many common empirical facts support the existence of additive costs (i.e., cost per unit exported).
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.
-\citet{hummels_skiba} estimate the elasticity of freight rates with respect to price to be well below unity.
+This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
 Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs.
 These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
-Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs are significant, representing 14\% of the median price.
 In contrast, highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
@@
 The result if further confirmed on more aggregate US, sector-level trade data for the period 1989-1994.
-Based on very similar data, but for a longer a period of time, our own estimates suggest however a more important part for additive costs, consistently with \citet{Irrazabal_2015}'s findings at the firm-level.
+However, this contradicts several recent empirical studies based on micro-level data, supporting the existence of additive costs (i.e., cost per unit exported).
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.
+Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs are significant, representing 14\% of the median price.
+Based on US sectoral similar to the one used by \citet{Lashkaripour-mimeo-2017}, but for a longer a period of time, our own estimates suggest also a more important part for additive costs.
 
@@
 Measurement error in Columbian data??
+
 Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.

### 2017-10-12 redaction/dhp_july2017.tex
changement paragraphe lashkaripour

@@
 Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs.
-These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
-In contrast, highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
-The positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}, \citet{Lashkaripour-mimeo-2017} finds, based on firm-transaction-level data on Colombian imports, an average transport cost elasticity of 0.9, supporting the \emph{ad-valorem} assumption.
-The result if further confirmed on more aggregate US, sector-level trade data for the period 1989-1994.
-However, this contradicts several recent empirical studies based on micro-level data, supporting the existence of additive costs (i.e., cost per unit exported).
+\citet{Lashkaripour-mimeo-2017} challenges this view.
+Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
+While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports.
+This result is also confirmed on more aggregate US, sector-level trade data.
+The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.
-Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs are significant, representing 14\% of the median price.
-Based on US sectoral similar to the one used by \citet{Lashkaripour-mimeo-2017}, but for a longer a period of time, our own estimates suggest also a more important part for additive costs.
+Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs represent a sizeable share of international trade costs.
+Based on US sectoral data, our own estimates also suggest an important part for additive costs in international transport costs.
 

### 2017-10-16 redaction/Presentations/SeminaireBETA_Nancy_Oct2017/slides_BETA.tex
présentation Nancy

--- 2017-09-07 redaction/Presentations/DEGIT_Conference/slides_DEGIT_v1.tex
@@
 
-\date[September 2017]{DEGIT Conference, September 2017}
+\date[October 2017]{BETA Seminar, University of Nancy, October 2017}
 \title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
@@
 \textbf{All our results: Provide new quantitative evidence about the importance of the additive component in international transport costs}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Related literature: Two related strands}
+\begin{itemize}
+\item Challenging the dominant role of iceberg costs in international trade \vspace{0.1cm}
+\begin{itemize}
+\footnotesize{
+\item[-] Hummels and Skiba (2004) (sectoral data), Martin (2012) (firm-level data) 
+\item[-] The closest to us: Irarrazabal, Moxnes and Opromolla (2015) \vspace{0.1cm}
+\item[$\Rightarrow$] In support of the additive cost component assumption \vspace{0.1cm}
+\item[$\neq$] Our results: Quantitatively estimate the \textit{value} of the additive costs \vspace{0.1cm}
+\item[-] A broader view of the magnitude of additive costs \vspace{0.1cm}
+%\begin{itemize}
+%\item[$\star$] US import flows over 1974-2013 \vspace{0.1cm}
+%\end{itemize}
+\item[$\Rightarrow$] Additive costs do matter quantitatively, throughout the period } \vspace{0.1cm}
+\end{itemize}
+\item Studying the patterns of trade costs over time \vspace{0.1cm}
+\begin{itemize}
+\footnotesize{
+\item[-] Hummels (2007), Behar and Venables (2011) \vspace{0.1cm}
+\item[-] What lies behind the decrease in transportation costs observed in the recent decades? \vspace{0.1cm}
+\item[-] Hummels (2007): Mostly due to trade composition effects \vspace{0.1cm}
+\item[$\neq$] Our results: Trade composition effects only a minor role \vspace{0.1cm}
+\item[-] Why such a difference?  \vspace{0.1cm}
+\item[$\Rightarrow$]Key role of the modelling of the additive component}
+\end{itemize}
+
+\end{itemize}
 
@@
 
-\begin{frame}[label = slide_empirical_strategy]
+\begin{frame}
 \frametitle{Empirical specification (2)}
-\textbf{Obtain the estimated equation}
-\begin{itemize}
-\item Make some simplifying assumptions \hyperlink{app_empirical_strategy}{\beamergotobutton{More details}}  \vspace{0.1cm}
-\begin{itemize}
-\item[-] About the specification of both additive and multiplicative costs  \vspace{0.1cm}
-\begin{itemize}
-\item[$\ast$] Separability between the product and the country dimensions (as Irarrazabal et al., 2015)   \vspace{0.1cm}
-\item[$\ast$] All products $k$ in a 3-digit sector ($s$) share the same structure of costs  \vspace{0.1cm}
-\end{itemize}
-\item[-] About the specification of the error term \vspace{0.1cm}
-\end{itemize}
-\item Taking logs, we finally estimate the following equation:
-\footnotesize
-\begin{equation*}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
-\end{equation*}
-
-\begin{itemize}
-\footnotesize
-\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
-\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
-\end{itemize}
-\normalsize
-\item A non-linear equation (due to the additive costs)
-\item[$\Rightarrow$] Estimation using non-linear squares
+\textbf{The estimation strategy} \vspace{0.1cm}
+\begin{itemize}
+\item Make some simplifying assumptions to get the estimated equation \vspace{0.1cm}
+\item About the specification of both additive and multiplicative costs  \vspace{0.1cm}
+\begin{itemize}
+\item[(1)] Separability between the origin country $i$ and the product $k$ dimensions (ss in Irarrazabal et al., 2015)\vspace{0.1cm}
+\item[(2)] All products $k$ in a 3-digit sector $s$ share the same structure of costs \vspace{0.1cm}
+\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
+\footnotesize
+\begin{equation}
+ \tau_{ik} = \tau_{i} \times \tau_{s(k)}, \qquad t_{ik} = t_{i} + t_{s(k)} \label{eq:specifTC}
+ \end{equation}
+ \normalsize
+\end{itemize}
+\item About the specification of the error term \vspace{0.1cm}
+\begin{itemize}
+\item[-] Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
+\item[$\Rightarrow$] The equation of interest becomes:
+\footnotesize
+$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
+\normalsize
+\begin{itemize}
+\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
+\end{itemize}
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item Taking logs, we finally estimate the following equation 
+\footnotesize
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
+\end{equation}
+\normalsize
+\begin{itemize}
+\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e., fixed effects) to be estimated \vspace{0.1cm}
+\end{itemize}
+\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
+\item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
+\begin{itemize}
+\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
+\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
+\end{itemize}
+\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution
 \end{itemize}
@@
 \item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
-\item[$\Rightarrow$] Estimate \textit{two} models\vspace{0.1cm}
-%\begin{itemize}
+\end{itemize}
+$\Rightarrow$ Estimate \textit{two} models\vspace{0.1cm}
+\begin{itemize}
 \item[(A)] With additive costs excluded (only ad-valorem costs):
@@
 
-\begin{frame}[label=slide_result1]
-\frametitle{Result 1: Size of additive transport costs \hyperlink{app_result1}{\beamergotobutton{More on this}} }
+\begin{frame}[label = slide_result1_more]
+\frametitle{Result 1: Size of additive transport costs }
 \begin{itemize}
@@
 \begin{itemize}
-\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
+\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More details}}  } \vspace{0.1cm}
 \end{itemize}
@@
 \begin{frame}
+\frametitle{Interpretation}
+\begin{itemize}
+\item \textbf{Size of overall transport costs} \vspace{0.1cm}
+\begin{itemize}
+\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
+\item[-] Lower than the 21\% markup taken from Anderson \& Van Wincoop (2004) (AWM), but  \vspace{0.1cm}
+\begin{itemize}
+\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
+\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
+\item[$\ast$] Our data are only for the US, which presumably has lower trade costs  \vspace{0.1cm}
+\end{itemize}
+\end{itemize}
+\item[$\Rightarrow$] Our work: A precise estimation of international transport costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] Going deeper into the structure of international transport costs 
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label = slide_result1]
 \begin{itemize}
@@
 \end{itemize}
-\item Opening the black box of transport costs \vspace{0.1cm}
+\item Opening the black box of transport costs \hyperlink{app_comparison}{\beamergotobutton{More on this}} \vspace{0.1cm}
 \begin{itemize}
@@
 \item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration
+
 \end{itemize}
@@
 
-\begin{frame} [label = slide_compeffects_strategy]
+\begin{frame} 
 \frametitle{Time trends in transport costs \& composition effects }
@@
 \item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
-\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) \vspace{0.1cm}
-\item Some insights on our methodology \hyperlink{app_compeffects_strategy}{\beamergotobutton{More details}} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{Comparison with Hummels (2007)}} \vspace{0.1cm}
-\begin{itemize}
-\item[-] Start with estimates of additive/multiplicative transport costs  \vspace{0.1cm}
+\item[$\Rightarrow$] What we do, in accordance with Hummels (2007)
+
+\end{itemize}
+\end{frame}
+
+\end{document}
+
+\begin{frame} [label = slide_compeffects_strategy]
+Some insights on our methodology \hyperlink{app_compeffects_strategy}{\beamergotobutton{More details}} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{Comparison with Hummels (2007)}} \vspace{0.1cm}
+\begin{itemize}
+\item Start with our estimates of additive/multiplicative transport costs (TC) ($\widehat{\tau_{ikt}$, $\widehat{t_{ikt}$}\vspace{0.1cm}
+\item Decompose the estimated TC component in three parts
+\begin{itemize}
+\item[-] Country dimension / product dimension / ``pure TC time trend''
+\end{itemize}
 \item[-] Extract the ``pure'' TC component by the mean of a time fixed-effect \vspace{0.1cm}
-%\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
+\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
 \item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted
@@
 
-\begin{frame}[label = app_empirical_strategy]
-\frametitle{More on the empirical specification }
-\textbf{The estimation strategy}
-\begin{itemize}
-\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
-\begin{itemize}
-\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
-\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
-\end{itemize}
-\footnotesize
-\begin{eqnarray*}
-\tau_{ik} = \tau_i \times \tau_k &\text{and}& t_{ik} = t_i + t_k
-\end{eqnarray*}
-\normalsize
-\item Additional assumption: All products in a 3-digit sector ($s$) share the same structure of costs \vspace{0.1cm}
-\item[$\Leftrightarrow$] Write $t_{is(k)}$ and $\tau_{is(k)}$ as:
-\footnotesize
-\begin{equation}
- \tau_{is(k)} = \tau_{i} \times \tau_{s(k)}, \qquad t_{is(k)} = t_{i} + t_{s(k)} \label{eq:specifTC}
- \end{equation}
- \normalsize
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} >1$, the error term should be always positive and multiplicative \vspace{0.1cm}
-\item[$\Rightarrow$] The equation of interest (\ref{eq:basis_equation}) becomes:
-\footnotesize
-$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
-\normalsize
-\item Taking logs, we finally estimate the following equation
-\footnotesize
-\begin{equation*}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
-\end{equation*}
-\normalsize
-\begin{itemize}
-\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
-\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
-\end{itemize}
-\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
-\item[$\Rightarrow$] Estimation using non-linear squares
-
-\end{itemize}
-\hyperlink{slide_empirical_strategy}{\beamergotobutton{Back to Slide}}
-\end{frame}
-
-\begin{frame}
-\frametitle{More on the estimation method}
-\begin{itemize}
-\item Non-linear least square estimation method: Some insights \vspace{0.1cm}
-\begin{itemize}
-\item At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.2cm}
-\end{itemize}
-\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution
-\end{itemize}
-
-\hyperlink{slide_empirical_strategy}{\beamergotobutton{Back to slide}}
-\end{frame}
-
 \begin{frame}[label=app_results_summary]
-\frametitle{Decomposing transport costs: Summary \hyperlink{slide_results_summary}{\beamergotobutton{Back to slide}}}
+\frametitle{Decomposing transport costs: Summary \hyperlink{slide_result1_more}{\beamergotobutton{Back to slide}}}
 \begin{table}[htbp]
@@
 
-\begin{frame}[label=app_result1]
-\frametitle{Size of transport costs: Comparison with the literature (1)}
-
-\begin{itemize}
-\item Size of overall transport costs \vspace{0.1cm}
-\begin{itemize}
-\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
-\item[-] Lower than the 21\% markup taken from Anderson \& Van Wincoop (2004) (AWM), but  \vspace{0.1cm}
-\begin{itemize}
-\footnotesize{
-\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
-\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
-\item[$\ast$] Our data are only for the US, which presumably has lower trade costs } \vspace{0.1cm}
-\end{itemize}
-\end{itemize}
-\item[$\Rightarrow$] Our work: A more precise and exhaustive estimation of international transport costs \vspace{0.1cm}
-\end{itemize}
-\hyperlink{slide_result1}{\beamergotobutton{Back to slide}}
-\end{frame}
-
-\begin{frame}
+\begin{frame}[label = app_comparison]
 \frametitle{Size of transport costs: Comparison with the literature (2)}

### 2017-10-16 redaction/Presentations/SeminaireBETA_Nancy_Oct2017/slides_BETA.tex
présentation Nancy

@@
 \begin{itemize}
-\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More details}}  } \vspace{0.1cm}
+\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{On some years}}  } 
 \end{itemize}
@@
 
-\end{document}
-
 \begin{frame} [label = slide_compeffects_strategy]
-Some insights on our methodology \hyperlink{app_compeffects_strategy}{\beamergotobutton{More details}} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{Comparison with Hummels (2007)}} \vspace{0.1cm}
-\begin{itemize}
-\item Start with our estimates of additive/multiplicative transport costs (TC) ($\widehat{\tau_{ikt}$, $\widehat{t_{ikt}$}\vspace{0.1cm}
-\item Decompose the estimated TC component in three parts
-\begin{itemize}
-\item[-] Country dimension / product dimension / ``pure TC time trend''
-\end{itemize}
-\item[-] Extract the ``pure'' TC component by the mean of a time fixed-effect \vspace{0.1cm}
+Some insights on our methodology \hyperlink{app_compeffects_strategy}{\beamergotobutton{More details}} \vspace{0.1cm}
+\begin{itemize}
+\item Start with our estimates of additive/multiplicative transport costs (TC) ($\widehat{\tau}_{ikt}$, $\widehat{t}_{ikt}$)\vspace{0.1cm}
+\item Decompose the estimated TC component in three parts \vspace{0.1cm}
+\begin{itemize}
+\item[-] Country dimension / product dimension / ``pure TC time trend'' \vspace{0.1cm}
+\end{itemize}
+\item Extract the ``pure'' TC component by the mean of a time fixed-effect \vspace{0.1cm}
+\begin{itemize}
 \item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
-\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted
+%\item[$\Rightarrow$] Obtain the fitted measures $\Gamma_{t}^{add}$ and $\Gamma_{t}^{adv}$ \vspace{0.1cm}
+\end{itemize}
+\item Build a unified measure of ``total'' transport costs (agglomerate the two components), both fitted and unfitted \vspace{0.1cm}
+%\begin{itemize}
+%\item[-] Also both fitted ($\Gamma_t^{TC, pure}$) and unfitted ($\Gamma_t^{TC, raw}$) 
+%\end{itemize}
+\item Series built as indices with initial value 100 in 1974 \vspace{0.1cm}
+\begin{itemize}
+\item[-] For the ``pure'' TC series: How transport costs components have changed over time if we stick the composition of trade flows by product/ country partner to the one observed in 1974
 \end{itemize}
@@
 \end{figure}
-\hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}} \hyperlink{app_compeffects_Hummels_rob}{\beamergotobutton{Robustness check}}
+\hyperlink{app_compeffects_Hummels_rob}{\beamergotobutton{Robustness check}} \hyperlink{slide_compeffects_results}{\beamergotobutton{Back to slide}}
 \end{frame}
@@
 \end{itemize}
-\item \textbf{Trade composition effects: A minor role in accounting for the time trend of TC}\vspace{0.1cm}
-\begin{itemize}
-\footnotesize
-\item[-] Not much evidence of a difference between the ``raw'' and the ``pure'' TC components \vspace{0.1cm}
+\item \textbf{Trade composition effects: A minor role in accounting for the time trend of transport costs}\vspace{0.1cm}
+\begin{itemize}
+\footnotesize
+\item[-] Not much difference between the ``raw'' and the ``pure'' TC components \vspace{0.1cm}
 \normalsize
@@
 
-\begin{frame}[label = slide_compeffects_results]
+\begin{frame}[label = slide_compeffects_comparison]
 \frametitle{Time trends in transport costs: The role of additive costs}
@@
 \end{itemize}
-\item Not the case here. \vspace{0.1cm}
+\item Not the case here \vspace{0.1cm}
 \item[$\Rightarrow$] \textbf{Where does the difference come from?} \vspace{0.1cm}
-\item \textbf{From the treatment of the additive component of transport costs} \vspace{0.1cm}
+\item \textbf{From the treatment of the additive component of transport costs} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{More on this}}\vspace{0.1cm}
 \begin{itemize}
@@
 {\scriptsize (a) Air } & {\scriptsize (b) Vessel}\\
-\includegraphics[width=5cm, height=4cm]{figure5_comme_hummels_air.pdf}
-& \includegraphics[width=5cm,height=4cm]{figure6_comme_hummels_ocean.pdf} \\
+\includegraphics[width=4cm, height=3cm]{figure5_comme_hummels_air.pdf}
+& \includegraphics[width=4cm,height=3cm]{figure6_comme_hummels_ocean.pdf} \\
 \end{tabular}
@@
 \end{figure}
-\item[$\Rightarrow$] Suggest a much more important role to trade composition effects...  \vspace{0.1cm}
-\item Than they actually do, see \hyperlink{slide_compeffects_figure}{\beamergotobutton{Figure 2}} \vspace{0.1cm}
-
+\item Composition effects tend to partially offset the decrease in the fitted transport costs, in both Air and Vessel \vspace{0.1cm}
+\begin{itemize}
+\item[-] The downward trend being of higher magnitude for the fitted rate, especially for Vessel
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label = slide_compeffects_results]
+\begin{itemize}
+\item A result overturned when we allow for more flexibility in the role of the additive component  \vspace{0.1cm}
+\begin{itemize}
+\item[-] See \hyperlink{slide_compeffects_figure}{\beamergotobutton{Figure 2}} \vspace{0.1cm}
+\end{itemize}
+\item Assuming a varying share of the additive component over time/country partner/product  \vspace{0.1cm}
+
+\begin{itemize}
+\item[-] Modifies the decomposition of the trend reduction of transport costs \vspace{0.1cm}
+\item[-] Between what comes from trade composition effects vs ``Pure'' TC changes \vspace{0.1cm}
+\item[-] ``Pure'' TC reduction, the main driver of the overall downward trend over time \vspace{0.1cm}
+\end{itemize}
 \item[$\Rightarrow$] The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs
@@
 \begin{frame}[label=app_results_summary]
-\frametitle{Decomposing transport costs: Summary \hyperlink{slide_result1_more}{\beamergotobutton{Back to slide}}}
+\frametitle{Decomposing transport costs: On some (selected) years}
+\footnotesize
 \begin{table}[htbp]
   \centering
-  \scriptsize{
-  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
+  \caption{Air: Transport costs estimates, 3-digit level }
 \begin{center}
-    \begin{tabular}{l|cc|cc}
-      \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
-    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs }  \\ \hline
-    Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
-    Median & 1.051 & 1.042 & 1.052 & 1.037 \\
-    Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
-    Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
-    Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
-   \textit{Ad-valorem term} & & & & \\ \hline
-    Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
-    Median & 1.028 & 1.018 & 1.028 & 1.016 \\
-    Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
-    Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
-    Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term }& & & &   \\ \hline
-    Mean  & 0.029 & 0.018 & 0.028 & 0.019 \\
-    Median & 0.019 & 0.007 & 0.017 & 0.008 \\
-    Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
-    Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
-    Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-        \# obs. & 29279 & 28207 & 29317 & 27680 \\
-    \# origin country & 188 & 191 & 188 & 189 \\
-    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
-  \end{tabular}
-    \end{center}}
-\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.
-The additive term is expressed in fraction of fab price.
-($^\ast$): Four 4-digit estimation: 0n selected years.
-($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
+    \begin{tabular}{l|cccccc}
+\hline\hline
+Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
+\hline
+Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.2 & 3.4  \\
+Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9  \\
+%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
+Mean & 3.6 & 2.3 & 2.4 &1.7 & 2.6 & 1.7  \\
+Median & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 \\
+%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
+Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0  \\
+Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5  \\
+%\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
+\hline
+\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351  \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{8cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
+\end{tabular}%
+\end{center}
+\label{tab:result_air_3d_detail}
 \end{table}%
+\hyperlink{slide_result1_more}{\beamergotobutton{Back to slide}}
+\normalsize
+\end{frame}
+
+\begin{frame}
+\frametitle{Decomposing transport costs: On some (selected) years }
+\footnotesize
+\begin{table}[htbp]
+  \centering
+  \caption{Vessel: Transport costs estimates, 3-digit level}
+\begin{center}
+    \begin{tabular}{l|cccccc}
+   \hline\hline
+Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
+ \hline
+   \multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
+   \hline
+Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 4.0 & 3.6  \\
+Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3  \\
+%Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
+\hline
+Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 \\
+Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8  \\
+%Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
+\hline
+Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5  \\
+Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 \\
+%Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
+\hline
+ \# observations & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{8cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
+\end{tabular}%
+\end{center} \label{tab:result_ves_3d_detail}
+\end{table}%
+\hyperlink{slide_result1_more}{\beamergotobutton{Back to slide}}
+\normalsize
+
 \end{frame}
@@
 \begin{frame}[label = app_comparison]
-\frametitle{Size of transport costs: Comparison with the literature (2)}
-\begin{itemize}
-\item Size of additive transport costs: Comparison with Irarrazabal et al. (2015) \vspace{0.1cm}
+\frametitle{More on the size of additive transport costs}
+\begin{itemize}
+\item Comparison with Irarrazabal et al. (2015) \vspace{0.1cm}
 \begin{itemize}
 \item[-] Irarrazabal et al. (2015): Estimate the trade costs ratio $t/\tau$ = 14\% of the export price $\widetilde{p}$ \vspace{0.1cm}
-\item[-] In 2004, our estimates lead to $t/\tau = 7.7\% \times \widetilde{p}$ (Air), $t/\tau = 1.5\%\times \widetilde{p}$ (Ocean) \vspace{0.1cm}
+\item[-] In 2004, our estimates lead to $t/\tau = 1.8\% \times \widetilde{p}$ (Air), $t/\tau = 2.8\%\times \widetilde{p}$ (Ocean) \vspace{0.1cm}
 \item[$\Rightarrow$] Much lower !
@@
 \item[-] According to AVW, transport costs $\simeq$ 15\% of trade costs \vspace{0.1cm}
-\item[-] Simple calculus thus implies that our estimates amount to 62\% (Air) and 53\% (Vessel) of Irarrazabal \& al.'s estimates \vspace{0.1cm}
+\item[-] Plugging this into Irarrazabal et al.'s (2015) estimates yields a ratio $t/\tau$ of transport costs equal to $2.1\% \times \widetilde{p}$
+\item[$\Rightarrow$] Our estimates amount to 62\% (Air) and 53\% (Vessel) of Irarrazabal \& al.'s estimates (of transport costs)\vspace{0.1cm}
 \end{itemize}
@@
 \begin{itemize}
-\item \textbf{Result 2}: Including the additive component improves the fit of the model, whatever the considered criterion or the transport mode \vspace{0.1cm}
+\item Including the additive component improves the fit of the model, whatever the considered criterion or the transport mode \vspace{0.1cm}
 \begin{itemize}
@@
 \begin{frame}[label = app_compeffects_Hummels]
-\frametitle{Composition effects: Comparing with Hummels (2007) (1)}
+\frametitle{Composition effects: Comparing with Hummels (2007)}
 Reformulating Hummels's (2007) method  \vspace{0.1cm}
@@
 \item[-] Not much of a difference provided the weighting scheme in value is constant over time \vspace{0.1cm}
-\item[-] Results robust to the alternative weighting scheme (apply Hummels's (2007) method) \hyperlink{app_compeffects_Hummels_rob}{\beamergotobutton{See}}
+\item[-] Results robust to the alternative weighting scheme (apply Hummels's (2007) method) 
 \end{itemize}
@@
 
-\end{frame}
-
-\begin{frame}
-\textcolor{blue}{5.} Last difference - the major one !
-\begin{itemize}
-\item Coming back to Hummels's (2007) equation (\ref{eq:hummels}), rewritten as:
-\footnotesize
-\begin{equation}
-\ln TC_{ikt} \equiv \ln\left(\frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}  \right) = \Delta_{ikt} - \beta \ln \widetilde{p}_{ikt} \label{eq:hummels2}
-\end{equation}
-\begin{itemize}
-\item[-] with $\Delta_{ikt} \equiv \delta+ \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt}$
-\end{itemize}
-\normalsize
-\item Encompasses the two extreme cases of only ad-valorem costs / only additive costs
-\item Only multiplicative costs under $\beta=0$
-\begin{itemize}
-\item[-] In which case Equation (\ref{eq:hummels2}) rewrites as:
-$$\ln\left(\frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}  \right) = \Delta_{ikt} ~~\leftrightarrow~~ p_{ikt} = \underbrace{[\exp(\Delta_{ikt})+1]}_{\tau_{ikt}}\times \widetilde{p}_{ikt} $$
-%\item[$\Rightarrow$] Only multiplicative costs
-\end{itemize}
-\item Only additive costs under $\beta=1$
-\begin{itemize}
-\item[-] In which case Equation (\ref{eq:hummels2}) rewrites as:
-$$\ln\left(p_{ikt}-\widetilde{p}_{ikt}\right) = \Delta_{ikt} ~~\leftrightarrow~~ p_{ikt} = \widetilde{p}_{ikt} + \underbrace{[\exp(\Delta_{ikt})+1]}_{t_{ikt}}$$
-%\item[$\Rightarrow$] Only additive costs
-\end{itemize}
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item As noted by Hummels \& Skiba (2004), $0<\beta<1$ is the elasticity of freight costs to the export price \vspace{0.1cm}
-\item Measures the relative weight of the additive component in international transport costs \vspace{0.1cm}
-\begin{itemize}
-\item[-] The higher $\beta$ close to 1, the larger the share of additive costs \vspace{0.1cm}
-\end{itemize}
-\item By estimating Equation (\ref{eq:hummels}) \textbf{with a constant $\beta$}, Hummels (2007) assumes a constant share of additive costs over time/country partner/product \vspace{0.1cm}
-\item A key difference with our method \vspace{0.1cm}
-\begin{itemize}
-\item[-] Our strategy to obtain the transport costs measures explicitly allows for an additive component that varies in the 3 dimensions \vspace{0.1cm}
-\end{itemize}
-\item A key importance in the decomposition of time trends of transport costs (between the role of trade composition effects and the ``pure'' transport costs dimension)
-\end{itemize}
-\hyperlink{slide_compeffects_strategy}{\beamergotobutton{Back to main slide}}
 \end{frame}
@@
 \begin{itemize}
-\item Applying Hummel's (2007) weighting scheme rather than ours \hyperlink{slide_compeffects_figure}{\beamergotobutton{Compare}}
+\item Applying Hummel's (2007) weighting scheme rather than ours 
 \begin{itemize}
@@
 
-\hyperlink{app_compeffects_Hummels2}{\beamergotobutton{Back to slide}}
+\hyperlink{slide_compeffects_figure}{\beamergotobutton{Back to main slide}}
+\end{frame}
+
+\begin{frame}
+\textcolor{blue}{5.} Last difference - the major one !
+\begin{itemize}
+\item Coming back to Hummels's (2007) equation (\ref{eq:hummels}), rewritten as:
+\footnotesize
+\begin{equation}
+\ln TC_{ikt} \equiv \ln\left(\frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}  \right) = \Delta_{ikt} - \beta \ln \widetilde{p}_{ikt} \label{eq:hummels2}
+\end{equation}
+\begin{itemize}
+\item[-] with $\Delta_{ikt} \equiv \delta+ \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt}$
+\end{itemize}
+\normalsize
+\item Encompasses the two extreme cases of only ad-valorem costs / only additive costs
+\item Only multiplicative costs under $\beta=0$
+\begin{itemize}
+\item[-] In which case Equation (\ref{eq:hummels2}) rewrites as:
+$$\ln\left(\frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}  \right) = \Delta_{ikt} ~~\leftrightarrow~~ p_{ikt} = \underbrace{[\exp(\Delta_{ikt})+1]}_{\tau_{ikt}}\times \widetilde{p}_{ikt} $$
+%\item[$\Rightarrow$] Only multiplicative costs
+\end{itemize}
+\item Only additive costs under $\beta=1$
+\begin{itemize}
+\item[-] In which case Equation (\ref{eq:hummels2}) rewrites as:
+$$\ln\left(p_{ikt}-\widetilde{p}_{ikt}\right) = \Delta_{ikt} ~~\leftrightarrow~~ p_{ikt} = \widetilde{p}_{ikt} + \underbrace{[\exp(\Delta_{ikt})+1]}_{t_{ikt}}$$
+%\item[$\Rightarrow$] Only additive costs
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+\item As noted by Hummels \& Skiba (2004), $0<\beta<1$ is the elasticity of freight costs to the export price \vspace{0.1cm}
+\item Measures the relative weight of the additive component in international transport costs \vspace{0.1cm}
+\begin{itemize}
+\item[-] The higher $\beta$ close to 1, the larger the share of additive costs \vspace{0.1cm}
+\end{itemize}
+\item By estimating Equation (\ref{eq:hummels}) \textbf{with a constant $\beta$}, Hummels (2007) assumes a constant share of additive costs over time/country partner/product \vspace{0.1cm}
+\item A key difference with our method \vspace{0.1cm}
+\begin{itemize}
+\item[-] Our strategy to obtain the transport costs measures explicitly allows for an additive component that varies in the 3 dimensions \vspace{0.1cm}
+\end{itemize}
+\item A key importance in the decomposition of time trends of transport costs (between the role of trade composition effects and the ``pure'' transport costs dimension)
+\end{itemize}
+\hyperlink{slide_compeffects_comparison}{\beamergotobutton{Back to main slide}}
 \end{frame}

### 2017-11-13 redaction/dhp_july2017.tex
Changement rédaction

@@
 
-\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars at University of California-Irvine and Paris-Dauphine for very useful discussions.
+\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \& Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \& Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }
 
@@
 
-\date{October 2017}
+\date{November 2017}
  \maketitle
@@
 Our three main findings can be summarized as follows.
-
 First, we show that additive costs are quantitatively sizeable.
@@
 It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
-However, it is also clear that transport costs do represent a sizeable share of trade costs as pointed out in Section \ref{sec:Intro}.
 Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit.
-According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.
 In this respect, this dataset embraces only a partial view of international transport barriers.
+However, it is also clear that direct transport costs do represent a sizeable share of trade costs as pointed out in Section \ref{sec:Intro}.
 Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
@@
 This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs.
-Third, this dataset is available over a long time span.
-We are thus able to exploit information on a yearly basis from 1974 to 2013.
-\cite{hummels2007} preceded us in investigating the time dimension of this database.
-However, our paper differs from his work in two main respects.
-First, we extend the time coverage up to 2013, while \cite{hummels2007} stops in 2004.
-Second, and most importantly, we study the time trends of transport costs by disentangling the additive from the iceberg components.
-This proves to be key in the characterization of transport costs changes over time, in particular regarding the role of the composition effects.\smallskip
+Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013.
+\smallskip
 
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of trade costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
+It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
 The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})).
@@
 One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
-This is first due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - Section \ref{sec:results_trends} will go into more details of these compositions effects.
-Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearization.
+The key explanation for this is of statistical order.
+Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization.
 Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
-Descriptive statistics showing that it is indeed the case are available upon request to the authors.\medskip
+Descriptive statistics showing that that both observed and predicted values do match in logarithms are available upon request to the authors.\medskip
 
@@
 \noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, ee estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
 Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
@@
 Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded).
-Details are reported in Appendix \ref{app:comp-effects}).
+Details are reported in Appendix \ref{app:comp-effects}.
 
@@
 
-\section{Time trends in transport costs: Robustness to the weighting scheme}
+\section{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}
 
@@
 
-The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives tow comments.
+The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives two comments.
 First, differences do show up, in particular for the multiplicative component of Air transport.

### 2017-11-29 redaction/dhp_july2017.tex
Comments Vicard
Retours Vincent Vicard

@@
 One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
-The key explanation for this is of statistical order.
-Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization.
+This is first due to the fact that observed data are \textit{de facto} trade-weighed (i.e., by the weight of each good in total trade) while our estimates are based on simple arithmetic means within sector and country.
+Furthermore, Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization.
 Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.

### 2017-12-11 redaction/Presentations/SeminaireBETA_Nancy_Oct2017/slides_BETA_v1.tex
présentation Nancy

--- 2017-10-16 redaction/Presentations/SeminaireBETA_Nancy_Oct2017/slides_BETA.tex
@@
 
-\begin{frame}
+\begin{frame}[label = slide_variancedecomp]
 \frametitle{Empirical specification (2)}
@@
 \begin{itemize}
-\item[(1)] Separability between the origin country $i$ and the product $k$ dimensions (ss in Irarrazabal et al., 2015)\vspace{0.1cm}
-\item[(2)] All products $k$ in a 3-digit sector $s$ share the same structure of costs \vspace{0.1cm}
+\item[(1)] Separability between the origin country $i$ and the product $k$ dimensions (as in Irarrazabal et al., 2015)\vspace{0.1cm}
+\item[(2)] All products $k$ in a 3-digit sector $s$ share the same structure of costs \hyperlink{app_variancedecomp}{\beamergotobutton{More on this}}\vspace{0.1cm}
 \item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
@@
 \item[-] 48.2\% of total costs in average for vessel, 42.3\% for air  \vspace{0.1cm}
-\item[-] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
 \item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
@@
 \end{itemize}
+\item[$\Rightarrow$] Additive costs, a sizeable and structural dimension of international transportation costs \vspace{0.1cm}
 \item Opening the black box of transport costs \hyperlink{app_comparison}{\beamergotobutton{More on this}} \vspace{0.1cm}
@@
 
+\begin{frame}[label=app_variancedecomp]
+\frametitle{Variance decomposition exercise}
+\begin{figure}[htbp]
+\caption{Variance decomposition (observed cif-fas price gap)}
+\label{fig:decomp_variance}
+\begin{center}
+\includegraphics[width=8cm, height=6cm]{variance_decomposition.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+\hyperlink{slide_variancedecomp}{\beamergotobutton{Back to slide}}
+
+\end{frame}
+
 \begin{frame}[label=app_results_summary]
@@
 \item[-] Plugging this into Irarrazabal et al.'s (2015) estimates yields a ratio $t/\tau$ of transport costs equal to $2.1\% \times \widetilde{p}$
-\item[$\Rightarrow$] Our estimates amount to 62\% (Air) and 53\% (Vessel) of Irarrazabal \& al.'s estimates (of transport costs)\vspace{0.1cm}
-\end{itemize}
-\item[$\Rightarrow$] Additive transport costs represent between 1/2 and 2/3 of total additive trade costs
+\item[$\Rightarrow$] In line with our results\vspace{0.1cm}
+\end{itemize}
 \end{itemize}
@@
 \item[-] Hummels (2007): The unweighed average value over the $i,k$ dimension \vspace{0.1cm}
-\item[$\neq$] We weight each flow by its relative value in total trade flows observed in 1974 \vspace{0.1cm}
+\item[$\neq$] In estimating Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we weight each flow by its relative value in total trade flows  \vspace{0.1cm}
 \item[-] Not much of a difference provided the weighting scheme in value is constant over time \vspace{0.1cm}

### 2017-12-19 redaction/dhp_july2017.tex
J’ai juste changé ma signature
Je ne sais pas pourquoi cela donne tout comme changé.

@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \& Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
+Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
+Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
+; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
 Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

### 2018-01-03 redaction/dhp_july2017.tex
Premier jet

@@
 
-\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
+\title{\textsc{Transport costs did decline since 1974:\\Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
@@
 \begin{abstract}
-Following \cite{samuelson1954}, standard models of international trade have usually modelled trade costs as an ad-valorem tax equivalent.
-However, many common empirical facts point out the importance of additive costs.
-This paper provides new empirical results on international transport costs that support this view.
-To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel).
-Three main results emerge.
-First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively.
-Second, taking additive costs into account significantly improves the fit of the modelling of transport costs.
-Third, the modeling of the additive component is key in accounting for the time trend of international transport costs, as it affects the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs ``per se''.
-In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+ In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports.
+Transport cost ``per se'' did decline.
+This difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
+Exploiting information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel), three main results emerge.
+First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+Second, taking additive costs into account significantly improves the fit of the modeling of transport costs.
+Third, the modeling of the additive component is key to computing  the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs for given goods and origins.
 
@@
 
-Following \citet{samuelson1954}, standard models of international trade have usually modelled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+This focus is shared with part of the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
+\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
+However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that, paradoxically, transport costs for given goods and origins have not declined much.
+We show this result is not robust to a more flexible treatment of additive transport costs.
+In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs.
+\medskip
+
+Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
 As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, the structure (additive vs ad-valorem) of transport costs is far from being anecdotal.
@@
 
-By exploiting the time coverage of our database, our paper also relates to a second strand of the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
-Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} points out the role of trade composition effects in the characterization of time trends of international transport costs.
-If we share this view with \cite{hummels2007}, our results about the underlying sources of the transport costs time trend strikingly differ though.
-In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs.
-Importantly, we show that the difference of results can be attributed to the modelling of the additive component.
-\medskip
-
 Our three main findings can be summarized as follows.
@@
 Third, additive costs matter in accounting for the time trend of transport costs, as we show by exploiting the time dimension of our database to document the patterns of transport costs over time.
-Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in maritime shipping, and of same order of magnitude for both additive and iceberg components.
+Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components.
 Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
-Importantly, we show that the difference of results can be attributed to the modelling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
 In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
@@
 In Section \ref{sec:results_decomposition}, we provide results about the size and the role of the additive component of transport costs.
-After reporting the mean values over the period (by transport mode), we show how introducing additive transport costs improves the modelling of transport costs.
+After reporting the mean values over the period (by transport mode), we show how introducing additive transport costs improves the modeling of transport costs.
 Section \ref{sec:results_trends} characterizes the trends of transport costs (by transport mode) over the period 1974-2013.
@@
 These values are respectively equal to 1.8 and 2.5\% for air transport.
-Second, we show that taking additive costs into account improves the fit of the modelling of transport costs.
+Second, we show that taking additive costs into account improves the fit of the modeling of transport costs.
 All goodness-of-fit measures point out to this conclusion, which holds for both transport modes and all years considered.

### 2018-01-05 redaction/dhp_july2017.tex
relecture_JH
relecture modifs Guillaume

@@
 
-\title{\textsc{Transport costs did decline since 1974:\\Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
+\title{\textsc{Transport costs did decline since 1974:\\Findings from an explicit consideration of additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
@@
 Transport cost ``per se'' did decline.
-This difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
+This difference of results can be attributed to the new way of accounting for the additive component of transport costs we offer.
 Exploiting information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel), three main results emerge.
@@
 Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
-Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+We also show that the difference of results can be attributed to the explicit consideration of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
 In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip

### 2018-02-07 redaction/dhp_july2017.tex

--- 2018-01-03 redaction/dhp_july2017.tex
@@
 
-\title{\textsc{Transport costs did decline since 1974:\\Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
@@
 \begin{abstract}
- In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports.
+ This paper provides a quantitative assessment of the size and the importance of the additive component
+in international transport costs, from exploiting information contained in the US imports flows database over 1974-2013 (distinguishing by transport mode, air and vessel).
+Two main new findings emerge.
+First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+Second, and in contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports.
 Transport cost ``per se'' did decline.
-This difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
-Exploiting information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel), three main results emerge.
-First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
-Second, taking additive costs into account significantly improves the fit of the modeling of transport costs.
-Third, the modeling of the additive component is key to computing  the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs for given goods and origins.
-
+Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
 \end{abstract}
@@
 \noindent \emph{JEL classification}: F14, N70, R40 \\
-\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs
+\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs, trade composition effects
 
@@
 If much of trade policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
-This is accordingly the focus of the paper.\smallskip
+International transport costs accordingly stand at the heart of the paper.\smallskip
 
@@
 \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that, paradoxically, transport costs for given goods and origins have not declined much.
+However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that, somehow paradoxically, transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.
 We show this result is not robust to a more flexible treatment of additive transport costs.
-In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs.
+In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs, once we allow for a modelling of the additive cost component that varies over time, country partner and product.
 \medskip
 
-Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis)\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, the structure (additive vs ad-valorem) of transport costs is far from being anecdotal.
-The literature has long emphasized its role in shaping the pattern of trade flows.
+The modelling of transport costs accordingly plays an important role in the paper.
+Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, in line with the Alchian and Allen's (1964)\nocite{alchian} conjecture, a recent strand of the literature points out to the existence of an additive component (i.e., a cost per unit traded).\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
+often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
+In this respect, our paper contributes to this literature that challenges the dominant role of iceberg costs in international trade.
 As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
@@
 Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
-While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports.
+While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports.
 This result is also confirmed on more aggregate US, sector-level trade data.
 The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.
-Recently, \citet{Irrazabal_2015} document on Norwegian firm-level trade data that additive costs represent a sizeable share of international trade costs.
-Based on US sectoral data, our own estimates also suggest an important part for additive costs in international transport costs.
-
-%\citet{Lashkaripour-mimeo-2017} takes into account the fact
-%Yet, one can be concerned by the generality of this result.
-The positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
-Measurement error in Columbian data??
-
-Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{This is due to the alteration of relative prices in a heterogeneous-firms trade framework.
-Lower additive trade costs reduce the export market prices of the highly performing exporters relative to lowly performing exporters.
-This, in turn, shifts relative market shares among exporters towards the highly performing ones - a mechanism that can not occur under ad-valorem costs.
-Total export sales are then reallocated between a smaller number of high-productivity firms.
-Therefore, the expected profits from exporting prior to entry increase all other things being equal by more in the case of lower per-unit costs, driving up incentives to enter the industry and bringing eventually the welfare enhancing intra-industry reallocations highlighted in \citet{melitz} framework.} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
+A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
+Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
 While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
-One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to fill this gap.
-\bigskip
+One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}).
+One contribution of the paper is to fill this gap.}
+Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
+Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price.
+Our paper complements their findings in two main respects.
+First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs.
+We thus obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively.
+Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
+In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
 
 Do additive costs matter?
-To answer this question, we provide an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
-To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
+To answer to this question, we provide an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
+This relies on exploiting information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
 Obviously, exploiting this database has drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
-Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs related to the time value of goods in transit.
+Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs notably related to the time value of goods in transit.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
@@
 
-By investigating the additive dimension of transport costs, our paper is related to the literature that challenges the dominant role of iceberg costs in international trade.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian exports in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price.
-Our paper complements their findings in two main respects.
-First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs.
-Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
-In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-Despite these differences, our results share with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.\smallskip
-
-Our three main findings can be summarized as follows.
+Our results can be summarized in two main new findings.
 First, we show that additive costs are quantitatively sizeable.
@@
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}.
-This represents a valuable insight for calibrating related international trade or business cycle models.
-
-Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.} This represents a valuable insight for calibrating related international trade or business cycle models.
+We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
 Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-On average over the whole period, biased estimates for ad-valorem are 5\% for air and 6\% for vessel, while unbiased estimates are respectively 2.5\% and 3.2\%.
-Further, we obtain a systematically better goodness of fit when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
-
-Third, additive costs matter in accounting for the time trend of transport costs, as we show by exploiting the time dimension of our database to document the patterns of transport costs over time.
-Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components.
-Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
+
+Second, we establish that the modelling of the additive costs crucially matters in accounting for the trend patterns of international transport costs.
+Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components.
+Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects.
+This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
 Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
+In both aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
 

### 2018-02-08 redaction/dhp_july2017.tex
changement texte

@@
 
-\date{November 2017}
+\date{February 2018}
  \maketitle
@@
 This paper provides a quantitative assessment of the size and the importance of the additive component
-in international transport costs, from exploiting information contained in the US imports flows database over 1974-2013 (distinguishing by transport mode, air and vessel).
+in international transport costs.
+To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode, air and vessel.
 Two main new findings emerge.
@@
 
-\noindent Trade costs are central to international economic analysis.
+Trade costs are central to international economic analysis.
 In particular, they are considered a major obstacle to international economic integration and international trade flows.
@@
 
-This focus is shared with part of the international trade literature that studies the patterns of trade costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
 Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
 \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that, somehow paradoxically, transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.
-We show this result is not robust to a more flexible treatment of additive transport costs.
-In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs, once we allow for a modelling of the additive cost component that varies over time, country partner and product.
-\medskip
-
-The modelling of transport costs accordingly plays an important role in the paper.
+However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.
+In this paper, we investigate the robustness of this result.
+The major contribution of our paper is then to show that it is not robust to a more flexible treatment of the additive component of transport costs (i.e., a cost per unit traded).
+In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs, once we allow for a modelling of the additive cost component that varies over country partner and product.
+More than trade composition effects, it is the reduction of the transport costs ``per se'' that overwhelmingly explains the downward trend of overall transport costs observed over the period.\medskip
+
+To prove this point requires a careful modelling of transport costs.
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, in line with the Alchian and Allen's (1964)\nocite{alchian} conjecture, a recent strand of the literature points out to the existence of an additive component (i.e., a cost per unit traded).\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
 often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
-In this respect, our paper contributes to this literature that challenges the dominant role of iceberg costs in international trade.
+This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts.
+To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
+This database is the same as \cite{hummels2007}, extended over the more recent years through (2005-2013).
+\smallskip
+
+Two main new findings emerge.
+First, additive transport costs are quantitatively sizeable.
+On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period).
+Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model.
+To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{As we discuss below, the closest paper to us in this respect is \citet{Irrazabal_2015}.
+However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.} This represents a valuable insight for calibrating related international trade or business cycle models.
+We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
+Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
+Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
+
+Second, and most importantly, we find that the main source of the downward trend in overall transport costs over 1974-2013 does come from the reduction of transport cost ``per se'', rather than from trade composition effects.\footnote{Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping.} This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary  across products and country partners (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
+
+In both aspects, our results thus point the importance of the additive component in accounting for international transport costs.
+In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade.
 As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
@@
 One contribution of the paper is to fill this gap.}
+
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
 Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price.
-Our paper complements their findings in two main respects.
+Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
+Yet, our paper complements their findings in two main respects.
 First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs.
@@
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
-
-Do additive costs matter?
-To answer to this question, we provide an empirical decomposition of the structure of transport costs over time, by explicitly distinguishing between ad-valorem and additive parts.
-This relies on exploiting information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-Obviously, exploiting this database has drawbacks, in particular as it restricts our scope of study to the monetary dimension of international transport costs.
-Since our measure is based on the gap between export and import prices, it necessarily omits the other dimension of transport costs notably related to the time value of goods in transit.
-According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
-Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
-They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).
-\citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
-In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.
-Furthermore, our data is based on a single and reliable customs source with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure.
-This limits the measurement error bias.
-Therefore, we strongly believe that exploiting this database may deliver useful and reliable insights on the structure of international trade costs.
-
-Our results can be summarized in two main new findings.
-First, we show that additive costs are quantitatively sizeable.
-On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
-This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period).
-Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model.
-To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.} This represents a valuable insight for calibrating related international trade or business cycle models.
-We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
-
-Second, we establish that the modelling of the additive costs crucially matters in accounting for the trend patterns of international transport costs.
-Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components.
-Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects.
-This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
-Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-In both aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
 
@@
 
-\textbf{stop here}
+\textbf{VV: Trop de ``thus'' dans l'intro.}
+
+\textbf{VV: Affirmer nettement la difference avec Hummels !!
+Pour Vincent, notre vrai resultat est la (il y a une baisse des couts de transport, les effets de compo n’expliquent pas tout), et c’est cela que nous devrions vendre, davantage que l’histoire multiplicatif/additif, une question de geek en trade qui limitera forcément le type de revue auquel on pourra soumettre.
+EST-CE BON AINSI?}
 
@@
 We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
-While both are closely related, we refer to the fas price as the price considered by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
+The fas price (for ``Free Alongside Ship'') means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
+It is also the seller's responsibility to clear the goods for export.
+The fob price (for ``Free on Board'') means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
+Once the goods clear the railing of the vessel the buyer assumes the risk.
+Note that this term is used exclusively for maritime and inland waterway transport.
+While both terms are closely related, we refer to the fas price as the price considered by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
 The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
@@
 In this respect, this dataset embraces only a partial view of international transport barriers.
-However, it is also clear that direct transport costs do represent a sizeable share of trade costs as pointed out in Section \ref{sec:Intro}.
+However, it is also clear that direct transport costs do represent a sizeable share of trade costs.
+According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
+Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
+They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
+In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.} Furthermore, our data is based on a single and reliable customs source with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure.
+This limits the measurement error bias.
+Therefore, we strongly believe that exploiting this database delivers useful and reliable insights on the structure of international trade costs.
 Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
@@
 \smallskip
+
+\textbf{VV: Les explications accompagnant l’equation estimee 6 ne sont pas claires : si $\tau_i$, $tau_{s(k)}$, $t_i$ et $t_{s(k)}$ sont des effets fixes, alors $tau_i = t_i$ et $tau_{s(k)} = t_{s(k)}$ !!!
+Clarifier.}
+
+\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu’il y a quand mime de la variation between sectors au niveau 5-digits.
+Or, nous riagrigeons au niveau 3 ou 4d a cause des NLLS.
+Pourquoi ne pas estimer l’aquation initiale (juste avant la (5)) en niveau avec des OLS ?
+A cet agard, l’explication pour la structure multiplicative des rasidus n’est ni tras claire, ni tras convaincante.}
 
@@
 
+\textbf{VV: Nos resultats sont en ad-valorem (multiplicatif du prix), ce qui explique qu’on a des pourcentages superieurs pour Vessel par rapport a Air.
+Spontanement, on s’attend au contraire (les couts de fret sont plus importants dans l’aerien), mais en fait ces chiffres refletent simplement que les prix sont plus eleves dans l’aerien.
+En un mot, ces chiffres ne signifient pas que les couts monétaires dans l’aerien sont inférieurs au vessel.
+Ceci devrait etre précise.}
+
 Table \ref{tab:summary_results} calls for two types of comments.
@@
 In the next section, we analyze their importance through another angle based on quality-of-fit diagnostic tests.
+
+\textbf{VV: Les comparaisons chiffrees sont parfois délicates.
+Le paragraphe de comparaison avec Irarrazabal and co page 12 est tres clair, mais cela vient un peu tard.
+Il faut que des l’intro on comprenne bien la difference entre notre mesure de coûts additifs (representant entre 40 et 60\% des coûts de transport totaux) et les 14\% d’Irrara and co, qui sont un pourcentage du prix fab.}
 
@@
 
+\textbf{VV: Figure 2 : 1) peu lisible en l’état.
+Soit on conserve ces graphs, et dans ce cas-la, on commente la dynamique, pas seulement le debut et la fin; si on veut rester sur seulement le debut et la fin, mettre un tableau.
+2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas...
+et les couts totaux sont en-dessous !
+Déconcertant.
+Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.
+3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas.}
+
 \subsection{Time trends in transport costs and the modeling of the additive component}

### 2018-02-09 redaction/dhp_july2017.tex
Changement rédaction

@@
 In this paper, we investigate the robustness of this result.
-The major contribution of our paper is then to show that it is not robust to a more flexible treatment of the additive component of transport costs (i.e., a cost per unit traded).
-In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs, once we allow for a modelling of the additive cost component that varies over country partner and product.
-More than trade composition effects, it is the reduction of the transport costs ``per se'' that overwhelmingly explains the downward trend of overall transport costs observed over the period.\medskip
+As major contribution of the paper, we find that composition effects play a minor role in accounting for the trend patterns of transport costs, in contrast to \cite{hummels2007}.
+Importantly, we show that this difference of result can be accounted for by the modelling of the additive component of international transport costs (i.e., a transport cost per unit traded).
+Once we allow for a more flexible (and empirically relevant) modelling of the additive cost component that varies over country partner and product, it is the reduction of the transport costs ``per se'' that overwhelmingly explains the downward trend of overall transport costs observed over the period, rather than trade composition effects.
+\medskip
 
@@
 We thus obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively.
+From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
+Straightforward calculation yields a ratio of additive to multiplicative international transport costs around 2\% (expressed in terms of the export price) in both \citet{Irrazabal_2015} and our paper.\footnote{See Section \ref{sec:results_decomposition} for more details.} This makes us confident in our quantitative assessment of the sizes of international transport costs.
 Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
@@
 
-\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu’il y a quand mime de la variation between sectors au niveau 5-digits.
-Or, nous riagrigeons au niveau 3 ou 4d a cause des NLLS.
-Pourquoi ne pas estimer l’aquation initiale (juste avant la (5)) en niveau avec des OLS ?
-A cet agard, l’explication pour la structure multiplicative des rasidus n’est ni tras claire, ni tras convaincante.}
+\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu il y a quand meme de la variation between sectors au niveau 5-digits.
+Or, nous reagrigeons au niveau 3 ou 4d a cause des NLLS.
+Pourquoi ne pas estimer l'equation initiale (juste avant la (5)) en niveau avec des OLS ?
+A cet egard, l'explication pour la structure multiplicative des residus n'est ni tras claire, ni tres convaincante.}
 
@@
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, as that would imply some a direct influence between the firm's prices and the size of transport costs, except in the choice between air and vessel transport (that is beyond our scope of analysis).
+However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
 Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
@@
 
-Do additive costs matter?
-This section provides two elements of answer to this question.
-First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
-This allows us to estimate whether additive costs represent a sizeable component of transport costs.
-Second, we assess the importance of the additive component of in overall transport costs through goodness-of-fit measures.
+Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
+The first step of analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of transport costs.
+In this respect, these results \textit{per se} also constitute our first original contribution to the literature.
+In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.
 
@@
 
-Our first contribution to the literature is to provide estimates for the size of both the ad-valorem and the additive components of transport costs.
 Table \ref{tab:summary_results} reports a summary of our results.
@@
 En un mot, ces chiffres ne signifient pas que les couts monétaires dans l’aerien sont inférieurs au vessel.
-Ceci devrait etre précise.}
+Ceci devrait etre precise.}
+
+\textbf{Reponse de Lise: Une facon de repondre a ca est de montrer le prix fob moyen dans l'aerien / vessel.
+Pour montrer que c'est plus eleve dans l'aerien.
+Fournir une estimation en \$ du cout par unit (kg) additif pourrait etre interessant aussi, apres tout c'est fondamentalement ca qu'ils sont.}
 
@@
 Our methodology indeed allows us to go deeper into the structure of transport costs, that can be decomposed into an additive and an ad-valorem components.
-We thus find that the per-kg cost dimension is sizeable.
+We thus find that the per-kg cost dimension is quantitatively sizeable.
 On average over the whole period, additive costs are estimated to 2.9\% of the export price for ocean shipping, and 1.8\%.
-This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\/ and 1.8\% for vessel and air shipping respectively (on average over the period).
+This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period).
 Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
@@
 In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-These elements point out that additive costs are quantitatively sizeable.
-
-We can use the time dimension of our dataset in order to assess the validity of this result on a year-to-year basis over the period.
-To this aim, Figure \ref{fig:decomp_TC_3d} reports the shares of both the ad-valorem costs (grey area) and the additive costs (black area) in total costs (the total of the two, expressed in percentage of the fas price), by transport mode for each year between 1974 and 2013.
+Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for the related more theoretical papers, which need to calibrate their models.\smallskip
+
+Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
+Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology.
+One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
+From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
+Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport.
+This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}.
+This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
+While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
+From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs to the export price $\widetilde{p}$ equal to 2.1\% ($0.15\times 0.14$).
+This stands in line with our own results, and makes us confident in our quantitative assessment of the size of the additive component in international transport costs.\footnote{See also Appendix \ref{app:diagnostic_test}, which confirms the importance of the additive cost component through another perspective angle based on quality-of-fit diagnostic tests.}\smallskip
+
+\medskip
+
+\textbf{VV: Les comparaisons chiffrees sont parfois délicates.
+Le paragraphe de comparaison avec Irarrazabal and co page 12 est tres clair, mais cela vient un peu tard.
+Il faut que des l’intro on comprenne bien la difference entre notre mesure de coûts additifs (representant entre 40 et 60\% des coûts de transport totaux) et les 14\% d’Irrara and co, qui sont un pourcentage du prix fab.
+MIS UN TOPO LA DESSUS EN INTRO.
+Ok AINSI?}
+
+\section{Transport Costs time trends: The role of the additive component}\label{sec:results_trends}
+
+In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
+This focus is shared with part of the international trade literature, such as \cite{hummels2007} and \cite{Behar_Venables}.
+Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
+\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
+However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.
+In this paper, we investigate the robustness of this result.
+
+\subsection{A first look at the trend patterns}
+
+As a first step, Figure \ref{fig:decomp_TC_3d} displays the respective shares of additive (black area) and iceberg (grey area) components in total transport costs in total costs (the total of the two, expressed in percentage of the fas price), by transport mode for each year between 1974 and 2013.
 
@@
 
-%\textbf{je ne comprends pas les chiffres mis dans le commentaire en 1974, par rapport aux graphiques.}
-%JH: Voir Tableaux 4 et 5, années et modes de transport correspondants.
-Ils correspondent à la mesure "Iceberg alone", parce que cela semblait permettre une comparaison plus directe avec Hummels.
-Bref, à discuter.
-
-As reported on Figure \ref{fig:decomp_TC_3d}, total transport costs have regularly declined over time, by an order of magnitude of 3 percentage points in air shipping and 5 percentage points in ocean shipping (basically, divided by 2 in both cases).
-At this point though, we postpone this discussion about the downward trend for transport costs for Section \ref{sec:results_trends}, to focus on the importance of the additive component.
-In both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time.
-These results thus suggest that additive costs are neither negligible in magnitude nor an erratic phenomenon.
+Two main comments can be made.
+First, in both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time.
+Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
 By contrast, they represent an sizeable and structural dimension of international transport costs.
-Further, one side result of our work is to provide quantitative estimates for the values of both additive and ad-valorem transport costs,which can be very useful for the related more theoretical papers, which need to calibrate their models.\medskip
-
-Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology.
-One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
-From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
-Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport.
-This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}.
-This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
-While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
-From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs equal to 2.1\% ($0.15\times 0.14$), which is in line with our own results.
-This makes us confident in our quantitative assessment of the size of the additive component in international transport costs.
-In the next section, we analyze their importance through another angle based on quality-of-fit diagnostic tests.
-
-\textbf{VV: Les comparaisons chiffrees sont parfois délicates.
-Le paragraphe de comparaison avec Irarrazabal and co page 12 est tres clair, mais cela vient un peu tard.
-Il faut que des l’intro on comprenne bien la difference entre notre mesure de coûts additifs (representant entre 40 et 60\% des coûts de transport totaux) et les 14\% d’Irrara and co, qui sont un pourcentage du prix fab.}
-
-\subsection{Assessing the importance of additive transport costs}
+
+Second, considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
+In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974.
+However, computing the time trends from 1980 does not dramatically change the picture.
+The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996.
+For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
+Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.\smallskip
+
+Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner ``ceteris paribus'' transport costs and \textit{ii)} the evolution of the composition of trade.
+Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
+As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
+per product- and per partner- transport costs.
+
+\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}, that starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped.
+Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p_{ikt}}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:
+
+\begin{equation}
+TC_{ikt}\equiv \frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} = \widetilde{p}_{ikt}^{-\beta}X_{ikt} \label{eq:Hummel_benchmarkeq}
+\end{equation}
+
+Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value).
+As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem.
+At the other extreme, $\beta = 1$ means that all transport costs are additive.
+Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
+Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which Hummels (2007) decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}.
+At this point though, let us mention one key assumption retained by Hummels (2007), which assumes $\beta$ constant across the triplet origin country origin/product/year.
+In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.
+
+However, in light of our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and Figure \ref{fig:decomp_TC_3d}, one may wonder about the empirical relevance of this assumption.
+As reported in Figure \ref{fig:decomp_TC_3d}, it seems that if overall transport costs have decreased over time, its is also the case for each additive and multiplicative components.
+Further, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
+
+We investigate this point deeper by...
+
+\textbf{stop here, A FINIR}
+
+\subsection{Estimation strategy}
+
+The above descriptive statistics analysis drives us to adopt a different strategy from \cite{hummels2007} for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
+Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below).
+For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time.
+We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}
+
+\subsection{Empirical specification}
+Our strategy to eliminate the composition effects obeys the following rationale.
+As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}.
+We extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
+
+This can be described as a two-stage process.
+First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
+For the estimated ad-valorem component, we thus estimate the following equation:
+\begin{eqnarray}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\end{eqnarray}
+
+\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
+Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
+%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
+\begin{eqnarray}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\end{eqnarray}
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)).
+Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+Both equations are estimated by transport mode.
+
+In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
+This constitutes our second stage.
+As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
+\begin{equation}
+\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
+\end{equation}
+
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974.
+In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
+
+As for the additive cost defined in $[0;+\infty]$, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
+\begin{equation}
+\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
+\end{equation}
+
+\noindent As a result, the two series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+
+Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded).
+Details are reported in Appendix \ref{app:comp-effects}.
+
+\subsection{Characterizing the time trends in transport costs}
+Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{Figure \ref{fig:totalTC_compeffects_excl} reports the results, all types of goods included.
+In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} Panels (a), (b) and (c), we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
+Panels (d), (e) and (f) report the results for Vessel.
+In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
+
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects)}
+\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_all.pdf}
+\end{center}
+\end{figure}
+
+Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
+First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes.
+International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
+This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
+Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
+Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs.
+This is particularly true in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
+Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
+Composition effects are more pronounced in vessel transport.
+Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects, in particular for the multiplicative component (panel (d)).\smallskip
+
+\textbf{VV: Figure 2 : 1) peu lisible en l’état.
+Soit on conserve ces graphs, et dans ce cas-la, on commente la dynamique, pas seulement le debut et la fin; si on veut rester sur seulement le debut et la fin, mettre un tableau.
+2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas...
+et les couts totaux sont en-dessous !
+Déconcertant.
+Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.
+3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas.}
+
+\subsection{Time trends in transport costs and the modeling of the additive component}
+
+This last result stands in sharp contrast with \citet{hummels2007}, which obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects.
+If anything, we find the opposite result here.
+This drives us to investigate this difference of result further.
+As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
+Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
+As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
+%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below.
+\footnote{As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
+Robustness analysis to this point is driven in Section \ref{sec:robustness}.} To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).
+
+\begin{figure}[htbp]
+\caption{Characterizing the time trends: Applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_air.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_ocean.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
+Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partially offset the decrease in transport costs in both air and vessel shipping (as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel).\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel)} This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
+Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in vessel transport.
+Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
+
+%\section{Robustness checks \label{sec:robustness}}
+
+%\subsection{Measuring transport costs components: Robustness to the separability assumption}
+
+%Ie, estimate
+
+%$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
+
+%To do on 100 products ($s$), 50 countries (the most important?)
+
+\section{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}
+
+In this section, we provide a robustness test to the weighting scheme implemented to extract the trade composition effects implemented in Section \ref{sec:results_trends}.
+When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
+In this section, our objective is to assess that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
+
+To this aim, we rebuild the fitted values for each of of our transport cost measures (ad-valorem, additive and overall) applying Hummels' s (2007) weighting scheme (ie, taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
+The results are reported in Figure \ref{fig:compeffects_robustness}.
+
+\begin{figure}[htbp]
+\caption{Transport cost time trends: Robustness to the weighting scheme}
+\label{fig:compeffects_robustness}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_composition_all_np.pdf}
+\end{center}
+\end{figure}
+
+The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives two comments.
+First, differences do show up, in particular for the multiplicative component of Air transport.
+This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}.
+Rather than the weighing scheme, this confirms that the functional form and precisely the modeling of the additive component, turns out to be key in understanding the underlying determinants of the overall transport costs time trends, as we pointed out in Section \ref{sec:results_trends}.
+
+\section{Conclusion \label{sec:conclu}}
+
+This paper empirically studies the magnitude of additive (or per-kg) costs in international transport costs, by exploiting the differences between the import and the export prices.
+Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
+Our results may be summarized in three main findings.
+First, we provide a quantitative measure of both the additive and the iceberg transport cost.
+We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%, as mean values over 1974-2013.
+These values are respectively equal to 1.8 and 2.5\% for air transport.
+Second, we show that taking additive costs into account improves the fit of the modeling of transport costs.
+All goodness-of-fit measures point out to this conclusion, which holds for both transport modes and all years considered.
+Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
+Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
+In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+
+Our results could be extended in two main ways.
+On the empirical side, one may want to go deeper in the ``structural'' determinants of trade costs, i.e.
+identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
+On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles.
+This is left for further research.
+
+\newpage
+\bibliographystyle{essaien}
+\bibliography{biblio}
+
+\newpage
+
+\appendix
+
+\section{Data Appendix \label{app:data}}
+
+The Customs value is the value of imports as appraised by the U.S.
+Customs and Border Protection in accordance with the legal requirements of the Tariff Act of 1930, as amended.
+This value is generally defined as the price actually paid or payable for merchandise when sold for exportation to the United States, excluding U.S.
+import duties, freight, insurance, and other charges incurred in bringing the merchandise to the United States.
+The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller.
+In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
+
+The import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
+import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States.
+In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States in Canada or Mexico to the first port of entry.
+
+The cif (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
+It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S.
+import duties.
+
+\section{Estimation at the 3-digit classification level \label{app:more_results}}
+
+\subsection{Transport costs estimates: More detailed results}
+
+In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
+Precisely, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
+Table \ref{tab:result_air_3d_detail} reports the results for Air transport.
+The results for Ocean transport are displayed in Table \ref{tab:result_ves_3d_detail}.
+
+%\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
+\begin{table}[htbp]
+  \centering
+  \caption{Air: Transport costs estimates, 3-digits (selected years)}
+\begin{center}
+    \begin{tabular}{l|cccccc}
+\hline\hline
+Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
+\hline
+Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.2 & 3.4  \\
+Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9  \\
+%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
+Mean & 3.6 & 2.3 & 2.4 &1.7 & 2.6 & 1.7  \\
+Median & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 \\
+%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
+
+\hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
+Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0  \\
+Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5  \\
+%\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
+\hline
+\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351  \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
+\end{tabular}%
+\end{center}
+\label{tab:result_air_3d_detail}
+\end{table}%
+
+\begin{table}[htbp]
+  \centering
+  \caption{Vessel: Transport costs estimates, 3 digit (selected years)}
+\begin{center}
+    \begin{tabular}{l|cccccc}
+   \hline\hline
+Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
+ \hline
+   \multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
+   \hline
+Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 4.0 & 3.6  \\
+Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3  \\
+%Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
+\hline
+Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 \\
+Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8  \\
+%Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
+\hline
+Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5  \\
+Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 \\
+%Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
+\hline
+ \# observations & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
+\end{tabular}%
+\end{center} \label{tab:result_ves_3d_detail}
+\end{table}%
+
+\subsection{Assessing the importance of additive transport costs \label{app:diagnostic_test}}
 
@@
 
-\section{Transport Costs time trends: The role of the additive component}\label{sec:results_trends}
-\subsection{Estimation strategy}
-In this section, we investigate the role of the additive component in total transport costs in a historical perspective.
-As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, that displays the respective shares of additive and iceberg components in transport costs over time by transport mode.
-Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
-In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974.
-However, computing the time trends from 1980 does not dramatically change the picture.
-The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996.
-For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
-Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.
-
-Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner ``ceteris paribus'' transport costs and \textit{ii)} the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
-per product- and per partner- transport costs.
-
-If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
-Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below).
-For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time.
-We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}
-
-\subsection{Empirical specification}
-Our strategy to eliminate the composition effects obeys the following rationale.
-As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}.
-We extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-
-This can be described as a two-stage process.
-First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-For the estimated ad-valorem component, we thus estimate the following equation:
-\begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
-\end{eqnarray}
-
-\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
-Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
-%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
-\begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
-\end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)).
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
-Both equations are estimated by transport mode.
-
-In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-This constitutes our second stage.
-As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
-\begin{equation}
-\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
-\end{equation}
-
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974.
-In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
-
-As for the additive cost defined in $[0;+\infty]$, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
-\begin{equation}
-\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
-\end{equation}
-
-\noindent As a result, the two series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
-
-Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded).
-Details are reported in Appendix \ref{app:comp-effects}.
-
-\subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{Figure \ref{fig:totalTC_compeffects_excl} reports the results, all types of goods included.
-In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} Panels (a), (b) and (c), we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
-Panels (d), (e) and (f) report the results for Vessel.
-In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects)}
-\label{fig:totalTC_compeffects_excl}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_all.pdf}
-\end{center}
-\end{figure}
-
-Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes.
-International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
-This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
-Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
-Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs.
-This is particularly true in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
-Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
-Composition effects are more pronounced in vessel transport.
-Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects, in particular for the multiplicative component (panel (d)).\smallskip
-
-\textbf{VV: Figure 2 : 1) peu lisible en l’état.
-Soit on conserve ces graphs, et dans ce cas-la, on commente la dynamique, pas seulement le debut et la fin; si on veut rester sur seulement le debut et la fin, mettre un tableau.
-2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas...
-et les couts totaux sont en-dessous !
-Déconcertant.
-Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.
-3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas.}
-
-\subsection{Time trends in transport costs and the modeling of the additive component}
-
-This last result stands in sharp contrast with \citet{hummels2007}, which obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects.
-If anything, we find the opposite result here.
-This drives us to investigate this difference of result further.
-As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
-As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below.
-\footnote{As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
-Robustness analysis to this point is driven in Section \ref{sec:robustness}.} To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
-The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).
-
-\begin{figure}[htbp]
-\caption{Characterizing the time trends: Applying Hummel's (2007) method }
-\label{fig:comp_effects_as_in_Hummels}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_air.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_ocean.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partially offset the decrease in transport costs in both air and vessel shipping (as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel).\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel)} This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
-Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in vessel transport.
-Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
-
-%\section{Robustness checks \label{sec:robustness}}
-
-%\subsection{Measuring transport costs components: Robustness to the separability assumption}
-
-%Ie, estimate
-
-%$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
-
-%To do on 100 products ($s$), 50 countries (the most important?)
-
-\section{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}
-
-In this section, we provide a robustness test to the weighting scheme implemented to extract the trade composition effects implemented in Section \ref{sec:results_trends}.
-When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
-In this section, our objective is to assess that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
-
-To this aim, we rebuild the fitted values for each of of our transport cost measures (ad-valorem, additive and overall) applying Hummels' s (2007) weighting scheme (ie, taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
-The results are reported in Figure \ref{fig:compeffects_robustness}.
-
-\begin{figure}[htbp]
-\caption{Transport cost time trends: Robustness to the weighting scheme}
-\label{fig:compeffects_robustness}
-\begin{center}
-\includegraphics[height=8cm]
-{graph_composition_all_np.pdf}
-\end{center}
-\end{figure}
-
-The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives two comments.
-First, differences do show up, in particular for the multiplicative component of Air transport.
-This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}.
-Rather than the weighing scheme, this confirms that the functional form and precisely the modeling of the additive component, turns out to be key in understanding the underlying determinants of the overall transport costs time trends, as we pointed out in Section \ref{sec:results_trends}.
-
-\section{Conclusion \label{sec:conclu}}
-
-This paper empirically studies the magnitude of additive (or per-kg) costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
-Our results may be summarized in three main findings.
-First, we provide a quantitative measure of both the additive and the iceberg transport cost.
-We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%, as mean values over 1974-2013.
-These values are respectively equal to 1.8 and 2.5\% for air transport.
-Second, we show that taking additive costs into account improves the fit of the modeling of transport costs.
-All goodness-of-fit measures point out to this conclusion, which holds for both transport modes and all years considered.
-Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
-In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
-
-Our results could be extended in two main ways.
-On the empirical side, one may want to go deeper in the ``structural'' determinants of trade costs, i.e.
-identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles.
-This is left for further research.
-
-\newpage
-\bibliographystyle{essaien}
-\bibliography{biblio}
-
-\newpage
-
-\appendix
-
-\section{Data Appendix \label{app:data}}
-
-The Customs value is the value of imports as appraised by the U.S.
-Customs and Border Protection in accordance with the legal requirements of the Tariff Act of 1930, as amended.
-This value is generally defined as the price actually paid or payable for merchandise when sold for exportation to the United States, excluding U.S.
-import duties, freight, insurance, and other charges incurred in bringing the merchandise to the United States.
-The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller.
-In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
-
-The import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
-import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States.
-In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States in Canada or Mexico to the first port of entry.
-
-The cif (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
-It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S.
-import duties.
-
-\section{Estimation at the 3-digit classification level \label{app:more_results}}
-
-\subsection{Transport costs estimates: More detailed results}
-
-In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
-Precisely, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
-Table \ref{tab:result_air_3d_detail} reports the results for Air transport.
-The results for Ocean transport are displayed in Table \ref{tab:result_ves_3d_detail}.
-
-%\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Transport costs estimates, 3-digits (selected years)}
-\begin{center}
-    \begin{tabular}{l|cccccc}
-\hline\hline
-Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
-\hline
-Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.2 & 3.4  \\
-Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9  \\
-%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
-Mean & 3.6 & 2.3 & 2.4 &1.7 & 2.6 & 1.7  \\
-Median & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 \\
-%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
-
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
-Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0  \\
-Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5  \\
-%\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
-\hline
-\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351  \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-\end{center}
-\label{tab:result_air_3d_detail}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Transport costs estimates, 3 digit (selected years)}
-\begin{center}
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
- \hline
-   \multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
-   \hline
-Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 4.0 & 3.6  \\
-Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3  \\
-%Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
-\hline
-Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 \\
-Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8  \\
-%Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
-\hline
-Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5  \\
-Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 \\
-%Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
-\hline
- \# observations & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-\end{center} \label{tab:result_ves_3d_detail}
-\end{table}%
-
 \subsection{Variance decomposition exercise \label{app:decomp_variance}}

### 2018-02-15 redaction/dhp_july2017.tex

@@
 
-\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}, that starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped.
+\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
+He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped.
 Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p_{ikt}}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:
@@
 Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which Hummels (2007) decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}.
-At this point though, let us mention one key assumption retained by Hummels (2007), which assumes $\beta$ constant across the triplet origin country origin/product/year.
+
+In doing so, Hummels (2007) implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
 In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.
-
-However, in light of our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and Figure \ref{fig:decomp_TC_3d}, one may wonder about the empirical relevance of this assumption.
-As reported in Figure \ref{fig:decomp_TC_3d}, it seems that if overall transport costs have decreased over time, its is also the case for each additive and multiplicative components.
-Further, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
-
-We investigate this point deeper by...
-
-\textbf{stop here, A FINIR}
-
-\subsection{Estimation strategy}
-
-The above descriptive statistics analysis drives us to adopt a different strategy from \cite{hummels2007} for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
+However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, yield to cast some doubt on the empirical relevance of this assumption.
+As reported in Figure \ref{fig:decomp_TC_3d}, if overall transport costs have decreased over time, it is also the case for each additive and multiplicative components.
+In particular, the share of additive costs appears to be varying over time, in both transport modes.
+On top of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
+We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs
+
+\textbf{stop here}
+
+\subsection{Empirical specification}
+
+The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption.
+One may speculate that it is is not innocuous regarding the analysis of the underlying determinants of transport costs trend patterns.
+This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.}
+
+\paragraph{Estimation strategy} Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
 Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below).
@@
 
-\subsection{Empirical specification}
-Our strategy to eliminate the composition effects obeys the following rationale.
-As explained above, we start from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}.
-We extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+The remaining question is then, how to eliminate the composition effects?
+Our reasoning obeys the following rationale.
+Starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
 One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
 
-This can be described as a two-stage process.
+\paragraph{Estimation strategy} This empirical strategy can be described as a two-stage process.
 First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
+
 For the estimated ad-valorem component, we thus estimate the following equation:
@@
 
-\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
+\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
 As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
@@
 Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
-Both equations are estimated by transport mode.
+Both equations are estimated by transport mode.\smallskip
 
 In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
-This constitutes our second stage.
+This constitutes the second stage of our procedure.
 As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}

### 2018-02-16 redaction/dhp_july2017.tex
changement de texte, encore

@@
 \documentclass[a4paper,11pt]{article}
+
 %\usepackage[pdftex]{graphicx}
@@
 \usepackage{bbold}
+\usepackage{bbm}
+%\usepackage{bbold}
+%\usepackage[T1]{fontenc}
 % JH : impossible de compiler avec le package bbold, remplacé par amsfonts
+% LP : impossible de compiler avec le package bbold, remplacé par bbm
 \usepackage{upgreek}
@@
 
+\begin{abstract}
+This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel, in line with \cite{hummels2007}.
+This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components.
+In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports.
+Transport cost ``per se'' did decline.
+Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
+In quantitative terms further, we find that the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+In both aspects, our results point the importance of the additive component in accounting for international transport costs.
+\end{abstract}
+
 \thispagestyle{empty} \pagestyle{plain} \setcounter{page}{1}
@@
 Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows.
-If much of trade policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention.
+If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention.
 International transport costs accordingly stand at the heart of the paper.\smallskip
 
-This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others).
 Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
@@
 In this paper, we investigate the robustness of this result.
-As major contribution of the paper, we find that composition effects play a minor role in accounting for the trend patterns of transport costs, in contrast to \cite{hummels2007}.
-Importantly, we show that this difference of result can be accounted for by the modelling of the additive component of international transport costs (i.e., a transport cost per unit traded).
-Once we allow for a more flexible (and empirically relevant) modelling of the additive cost component that varies over country partner and product, it is the reduction of the transport costs ``per se'' that overwhelmingly explains the downward trend of overall transport costs observed over the period, rather than trade composition effects.
-\medskip
-
-To prove this point requires a careful modelling of transport costs.
+As major contribution of the paper, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
+Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip
+
+To prove this point drives us to provide a careful modelling of transport costs.
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
@@
 
-Two main new findings emerge.
+Our results may be summarized in two main new findings.
 First, additive transport costs are quantitatively sizeable.
@@
 
-Second, and most importantly, we find that the main source of the downward trend in overall transport costs over 1974-2013 does come from the reduction of transport cost ``per se'', rather than from trade composition effects.\footnote{Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping.} This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
-Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary  across products and country partners (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
-
-In both aspects, our results thus point the importance of the additive component in accounting for international transport costs.
+Second, we find that the main source of the downward trend in overall transport costs over 1974-2013 does come from the reduction of transport cost ``per se'', rather than from trade composition effects.\footnote{Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping.} This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
+Allowing the additive costs to vary  across products and country partners (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
+
+In both aspects, our results point the importance of the additive component in accounting for international transport costs.
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade.
@@
 First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs.
-We thus obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively.
+We obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively.
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-Straightforward calculation yields a ratio of additive to multiplicative international transport costs around 2\% (expressed in terms of the export price) in both \citet{Irrazabal_2015} and our paper.\footnote{See Section \ref{sec:results_decomposition} for more details.} This makes us confident in our quantitative assessment of the sizes of international transport costs.
+Straightforward calculation yields a ratio of additive to multiplicative international transport costs around 2\% (expressed in terms of the export price) in both \citet{Irrazabal_2015} and our paper.\footnote{See Section \ref{sec:results_decomposition} for more details.} This makes us confident in our quantitative assessment of the size of international transport costs.
 Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
 
 The paper is built as follows.
-Section \ref{sec:data_method} presents our data and our methodology.
-In Section \ref{sec:results_decomposition}, we provide results about the size and the role of the additive component of transport costs.
-After reporting the mean values over the period (by transport mode), we show how introducing additive transport costs improves the modeling of transport costs.
-Section \ref{sec:results_trends} characterizes the trends of transport costs (by transport mode) over the period 1974-2013.
+First, we estimate the values of international transport costs annually throughout the period 1974-2013 (and by transport mode), explicitly distinguishing between the additive and the ad-valorem components.
+This is made in Section \ref{sec:data_method}, which reports the mean values of the transport costs estimated over the period (by transport mode).
+Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''.
+Section \ref{sec:results_trends} is devoted to this analysis.
 Section \ref{sec:robustness} provides some additional robustness checks of our results.
 Section \ref{sec:conclu} concludes.
-
-\textbf{VV: Trop de ``thus'' dans l'intro.}
 
@@
 
-\section{Data Sources and Empirical Methodology \label{sec:data_method}}
-
-\subsection{A measure of transportation costs}
+\textbf{AI CHANGE LES SECTIONS}
+
+\section{Estimating international transports\label{sec:data_method}}
+
+\subsection{Data sources}
 
@@
 It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
-Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit.
+Second, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit.
 In this respect, this dataset embraces only a partial view of international transport barriers.
 However, it is also clear that direct transport costs do represent a sizeable share of trade costs.
-According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (10.7\%) and 9\% tax equivalent of ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
-Still, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
+According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...), that amount to a  9\% tax equivalent.
+Further, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
 They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
-In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.} Furthermore, our data is based on a single and reliable customs source with a wide time and geographical coverage, and it allows the computation of a direct transport cost measure.
+In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}
+
+Exploiting this dataset also has (at least) three advantages.
+First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source.
 This limits the measurement error bias.
-Therefore, we strongly believe that exploiting this database delivers useful and reliable insights on the structure of international trade costs.
-Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
-ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.
-%JH: OK, parfaitement clair pour moi !
-
-Yet, this dataset also has (at least) three advantages.
-First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source.
 Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
@@
 Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs.
+This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to \cite{Irrazabal_2015}.
 Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013.
-\smallskip
+Notice that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
+ignoring those flows that did not happen because of presumably prohibitive transport costs.
+In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.\smallskip
 
@@
 
-\subsection{Empirical specification}
-
-\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem and additive costs among total transport costs.
+\subsection{Estimating transport costs: Empirical specification}
+
+\paragraph{The estimated equation} In this section, our purpose is to provide time-varying estimates of the size of ad-valorem and additive costs among total transport costs.
 To do so, we start from the equation that expresses the price $p$ of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}$ (export, or fas price), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
@@
 \noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
-Let us denote $i$ the origin country, and $k$, the product (at the 5-digit level).
+Let us denote $i$ the origin country, and $k$, the product at the 5-digit level.
 Transforming the above equation (\ref{eq:base}) as ratio in order to get the percentage change in prices induced by transportation, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
@@
 \paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
 
@@
 The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
-Taking into account this constraint in the estimate suggests that the error term should be always positive and multiplicative, as in:
+Taking into account this constraint in the estimate implies that the error term should be always positive.
+We ensure that this constraint is fulfilled by specifying the error term as follows:
 
@@
 In other words, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
-same structure of costs.\footnote{We provide a decomposition variance exercise on the observed cif-fas price.
-As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
+same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
+As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
+on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
+This would not solve the problem though, for three main complementary reasons.
+First, at the 5-digit level the number of fixed effects to include in the estimation would be \textbf{very high - UN PEU MOU, dire combien?}, making the estimation computationally extremely burdensome, even in OLS.
+Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
+Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
+Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}
+
+This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
 \begin{equation}
@@
 \end{equation}
-where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.
-To eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
+where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, rigourously speaking, the precise equation that we estimate is the following:
+$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
+After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode).
+For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5 percent of the upper and lower tails of the distribution in the regression variables.
 These cut-offs are aimed at eliminating reporting or coding errors.
-We estimate this equation for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$).
-Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode (at the 3-digit level).
-\smallskip
+We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$).
+Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
+\medskip
 
 \textbf{VV: Les explications accompagnant l’equation estimee 6 ne sont pas claires : si $\tau_i$, $tau_{s(k)}$, $t_i$ et $t_{s(k)}$ sont des effets fixes, alors $tau_i = t_i$ et $tau_{s(k)} = t_{s(k)}$ !!!
-Clarifier.}
+Clarifier.
+Cf FOOTNOTE AJOUTEE?
+EST-CE OK AINSI?}
 
@@
 Pourquoi ne pas estimer l'equation initiale (juste avant la (5)) en niveau avec des OLS ?
-A cet egard, l'explication pour la structure multiplicative des residus n'est ni tras claire, ni tres convaincante.}
-
-One key objective of the paper is to characterize the importance of additive costs relatively to ad-valorem costs.
-A natural way to answer this question is to perform estimations of Equation (\ref{eq:model_IetA}) constraining additive costs to be equal to zero, and compare the fitting properties and the explanatory power of the models.
-For each year and transport mode, we thus estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
+A cet egard, l'explication pour la structure multiplicative des residus n'est ni tras claire, ni tres convaincante.
+AJOUTE UNE FOOTNOTE ET MODIFIE FACON DE JUSTIFIER FORME ERREUR MAIN TEXT.
+EST-CE OK AINSI?}
+
+As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international trade costs, that decomposes into both an additive and an ad-valorem component.
+In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\cite{Irrazabal_2015}, among others).
+We contribute to this issue by adopting the following strategy.
+For each year and transport mode, we estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
+From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.
+
 Under Model (\textit{(a)}), the estimated equation simplifies as:\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
@@
 In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
-\smallskip
+
+\textbf{pas encore completement clair.
+Veut-on dire que ce qui est pbtique, c'est que TC reagisse a ptilde?
+Je crois que ce qu'on veut dire, c'est que le fait que ptilde change en fonction des TC ne change pas les TC, le cif price s'ajuste proportionnellement?
+Si c'est ca, c'est vrai seulement si les couts sont advalorem...
+Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus?
+Par produit 3d/pays, par secteur 3d tous pays, etc...?}\smallskip
 
@@
 
-%\textbf{mettre les 3 modeles dans la online appendix.}
-
-\section{Decomposing Transport Costs: The importance of the additive component \label{sec:results_decomposition}}
+\textbf{stop here relecture}
+
+\subsection{Estimating results: The importance of the additive component \label{sec:results_decomposition}}
 
@@
 In this respect, these results \textit{per se} also constitute our first original contribution to the literature.
-In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.
-
-\subsection{Decomposing transport costs over 1974-2013}
+In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip
 
@@
     %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term ($\widehat{t}/\widetilde{p}$, in \%)}& & & &   \\ \hline
+    \textit{Additive term }& & & &   \\ \hline
+    \textit{In \% of the export price} &&&& \\ \hline
         Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
     Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
+   \textit{ In USD per kg traded }&&&& \\ \hline
+    Mean & 0.09	&1.12	&0.08&	2.16 \\
+    Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
     %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
@@
     %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-    \multicolumn{5}{l}{\textbf{Data} ($p/\widetilde{p}$, in \%) } \\ \hline
+    \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
+    Transport costs ($p/\widetilde{p}$, in \%) & & & & \\ \hline
     Mean & 5.3 & 5.0& 5.6&3.9 \\
     Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
+    Export price ($\widetilde{p}$), in USD & & & & \\
+    Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
+    Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
     \# obs. & 29279 & 28207 & 29317 & 27680 \\
@@
 ($^\ast$): Four 4-digit estimation: 0n selected years.
-($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
+($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}\\
+\parbox[l]{10cm}{\tiny{Notes: For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
+1974, 1981, 1989, 2001, 2009, 2013.}}
 \end{table}%
@@
 En un mot, ces chiffres ne signifient pas que les couts monétaires dans l’aerien sont inférieurs au vessel.
-Ceci devrait etre precise.}
-
-\textbf{Reponse de Lise: Une facon de repondre a ca est de montrer le prix fob moyen dans l'aerien / vessel.
-Pour montrer que c'est plus eleve dans l'aerien.
-Fournir une estimation en \$ du cout par unit (kg) additif pourrait etre interessant aussi, apres tout c'est fondamentalement ca qu'ils sont.}
+Ceci devrait etre precise.
+AMENDE LES RESULTATS DU TABLEAU POUR ESTIMER UNE VALEUR MONETAIRE + COMMENTAIRE (LIGHT) DESSUS.
+Ok ainsi?}
 
@@
 Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-Our methodology indeed allows us to go deeper into the structure of transport costs, that can be decomposed into an additive and an ad-valorem components.
 We thus find that the per-kg cost dimension is quantitatively sizeable.
@@
 Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+Expressed in monetary terms, the additive cost of exporting amounts to around 10 cents and 1.12 USD per kilogram exported, in maritime and air transport cost respectively (at the 3-digit level), indicating that freight rates are much higher in air transport.
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
@@
 
-\medskip
-
 \textbf{VV: Les comparaisons chiffrees sont parfois délicates.
@@
 On top of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
-We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs
-
-\textbf{stop here}
+
+We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
+To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, ie $\beta$, according to:
+\begin{eqnarray*}
+\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
+&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}}
+\end{eqnarray*}
+
+\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
+We can henceforth rebuild the value of the elasticity on a per product-year-country basis.
+The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
+
+\begin{figure}[htbp]
+\caption{Share of additive costs in total costs: Histogram}
+\label{fig:histogram_beta}
+\begin{center}
+\begin{tabular}{cc}
+%{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=3.0in, height=2.5in]{Etude_beta_pondere_air.pdf}
+& \includegraphics[width=3.0in,height=2.5in]{Etude_beta_pondere_ves.pdf} \\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution stands around 0.3-0.4 depending of the transport mode, consistently with our previous findings.
+In sharp contradiction with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, that would rather imply a distribution concentrated on a single point, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
+These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the  underlying sources of the overall transport costs trend patterns (between trade composition effects and transport costs ``per se'').
+This is made in next section.
 
@@
 The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption.
-One may speculate that it is is not innocuous regarding the analysis of the underlying determinants of transport costs trend patterns.
-This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.}
+This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} One may speculate that making the (empirically relevant) assumption of a varying share of additive costs over time, country and product is not anecdotal.
+Importantly, we confirm the key role of this assumption in analyzing the underlying determinants of transport costs trend patterns.
+We start presenting our estimation strategy.\smallskip
 
@@
 \begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbb{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
 
-\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
+\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
 As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
 Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
+
 %, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
 \begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
 \end{eqnarray}
@@
 Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes.
+First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes (Panels (c) and (f)).
 International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
 This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
-Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components.
+Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components (Panels (a) to (d)).
+Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs.
-This is particularly true in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport.
+For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line are very close to each other, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
-Composition effects are more pronounced in vessel transport.
-Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects, in particular for the multiplicative component (panel (d)).\smallskip
+Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)).
+Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+This is particularly the case for the multiplicative component (panel (d)).\smallskip
 
@@
 Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.
-3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas.}
+3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas.
+Ok AI REFORMULE LES CHOSES (SAUF POUR REMARQUE 2).
+OK ainsi?
+}
 
@@
 \begin{equation*}
-\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
+\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbbm{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbbm{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
 \end{equation*}
@@
 \begin{eqnarray}
-&&\ln TC_{ikt} = \delta+ \beta \ln \frac{q_{ikt}}{\widetilde{p}_{ikt}q_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \notag \\
-\Leftrightarrow && \ln TC_{ikt} = \delta+ \beta \ln \frac{1}{\widetilde{p}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t} + \epsilon_{ikt} \label{eq:Hummels1}
+&&\ln TC_{ikt} = \delta+ \beta \ln \frac{q_{ikt}}{\widetilde{p}_{ikt}q_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t} + \epsilon_{ikt} \notag \\
+\Leftrightarrow && \ln TC_{ikt} = \delta+ \beta \ln \frac{1}{\widetilde{p}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t} + \epsilon_{ikt} \label{eq:Hummels1}
 \end{eqnarray}
@@
 
-A second minor difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}$ in Equation (\ref{eq:Hummels1})) in two separate components (origin country and product dimensions).
+A second minor difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}$ in Equation (\ref{eq:Hummels1})) in two separate components (origin country and product dimensions).
 As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
@@
 
-\noindent with $\Delta_{ikt} \equiv \sum_{i,k}\alpha_{ik}.\mathbb{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbb{1}_{t}$ agglomerates the various fixed effects for reading clarity.
+\noindent with $\Delta_{ikt} \equiv \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t}$ agglomerates the various fixed effects for reading clarity.
 

### 2018-02-19 redaction/dhp_july2017.tex
changement

@@
 This paper features an online appendix containing additional results and available on the authors' websites.}}
+
 \author{Guillaume \textsc{Daudin}\thanks{%
@@
 Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
-; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }
+; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
 
@@
 \bigskip
-
-\begin{abstract}
-This paper provides a quantitative assessment of the size and the importance of the additive component
-in international transport costs.
-To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode, air and vessel.
-Two main new findings emerge.
-First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
-Second, and in contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports.
-Transport cost ``per se'' did decline.
-Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
-\end{abstract}
 
@@
 Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
-In quantitative terms further, we find that the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs.
 In both aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
 To prove this point drives us to provide a careful modelling of transport costs.
-Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
 As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
@@
 Yet, our paper complements their findings in two main respects.
-First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs.
-We obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively.
+First, while our study covers a narrower set of trade costs focused on transport costs, their data only allows recovering the ratio of additive to ad-valorem costs as a fraction of export price; in contrast, our empirical framework provides a distinct measure of each additive and ad-valorem cost component.
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-Straightforward calculation yields a ratio of additive to multiplicative international transport costs around 2\% (expressed in terms of the export price) in both \citet{Irrazabal_2015} and our paper.\footnote{See Section \ref{sec:results_decomposition} for more details.} This makes us confident in our quantitative assessment of the size of international transport costs.
-Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
+Based on the estimated share of transport costs in total trade costs provided by \citet{anderson_wincoop_jel} (around 15\%), straightforward calculation yields a ratio of additive to multiplicative transport (and not trade) costs of 2.1\% (14\% $\times$ 15\%) in \citet{Irrazabal_2015}.
+This falls in the range of our own estimates, and makes us confident in our quantitative assessment of the size of international transport costs.
+\footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
@@
 Section \ref{sec:conclu} concludes.
-
-\textbf{VV: Affirmer nettement la difference avec Hummels !!
-Pour Vincent, notre vrai resultat est la (il y a une baisse des couts de transport, les effets de compo n’expliquent pas tout), et c’est cela que nous devrions vendre, davantage que l’histoire multiplicatif/additif, une question de geek en trade qui limitera forcément le type de revue auquel on pourra soumettre.
-EST-CE BON AINSI?}
-
-\textbf{AI CHANGE LES SECTIONS}
 
@@
 \end{equation*}
-\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Considered in logs, the above equation becomes:
+\noindent where $\epsilon_{ik}$ follows a normal law centered on 0. Considered in logs, the above equation becomes:
 
@@
 This would not solve the problem though, for three main complementary reasons.
-First, at the 5-digit level the number of fixed effects to include in the estimation would be \textbf{very high - UN PEU MOU, dire combien?}, making the estimation computationally extremely burdensome, even in OLS.
+First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
 Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
@@
 \end{equation}
-where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, rigourously speaking, the precise equation that we estimate is the following:
+where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, strictly speaking, the exact equation that we estimate is the following:
 $$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
@@
 
-\textbf{VV: Les explications accompagnant l’equation estimee 6 ne sont pas claires : si $\tau_i$, $tau_{s(k)}$, $t_i$ et $t_{s(k)}$ sont des effets fixes, alors $tau_i = t_i$ et $tau_{s(k)} = t_{s(k)}$ !!!
-Clarifier.
-Cf FOOTNOTE AJOUTEE?
-EST-CE OK AINSI?}
-
-\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu il y a quand meme de la variation between sectors au niveau 5-digits.
-Or, nous reagrigeons au niveau 3 ou 4d a cause des NLLS.
-Pourquoi ne pas estimer l'equation initiale (juste avant la (5)) en niveau avec des OLS ?
-A cet egard, l'explication pour la structure multiplicative des residus n'est ni tras claire, ni tres convaincante.
-AJOUTE UNE FOOTNOTE ET MODIFIE FACON DE JUSTIFIER FORME ERREUR MAIN TEXT.
-EST-CE OK AINSI?}
-
 As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international trade costs, that decomposes into both an additive and an ad-valorem component.
@@
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
+This is not an issue here, as we are not interested in causal inference.
+Rather, our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
+
+%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
 Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
@@
 
-\textbf{pas encore completement clair.
+%\textbf{pas encore completement clair.
 Veut-on dire que ce qui est pbtique, c'est que TC reagisse a ptilde?

### 2018-02-19 redaction/dhp_july2017.tex
Changement biblio

@@
 In particular, they are considered a major obstacle to international economic integration and international trade flows.
-According to \cite{Jacks08}, the trade cost decline explains around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse.
+According to \cite{Jacks08}, the trade cost decline explains around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse (but see \cite{corlay_2017}).
 After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO after 1995) multilateral agreements.
@@
 trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
+Several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
 
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se} (\citet{daudin_logistique_2003, daudin_transactions_2005}).
 In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.

### 2018-02-19 redaction/dhp_july2017.tex
changement texte, encore

--- 2018-02-19 redaction/dhp_july2017.tex
@@
 These results are reported in Section \ref{sec:results_decomposition}.
-
-\textbf{stop here relecture}
 
@@
 
-\textbf{VV: Nos resultats sont en ad-valorem (multiplicatif du prix), ce qui explique qu’on a des pourcentages superieurs pour Vessel par rapport a Air.
-Spontanement, on s’attend au contraire (les couts de fret sont plus importants dans l’aerien), mais en fait ces chiffres refletent simplement que les prix sont plus eleves dans l’aerien.
-En un mot, ces chiffres ne signifient pas que les couts monétaires dans l’aerien sont inférieurs au vessel.
-Ceci devrait etre precise.
-AMENDE LES RESULTATS DU TABLEAU POUR ESTIMER UNE VALEUR MONETAIRE + COMMENTAIRE (LIGHT) DESSUS.
-Ok ainsi?}
-
 Table \ref{tab:summary_results} calls for two types of comments.
@@
 With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.
-\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons.
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons.
 \citet{anderson_wincoop_jel} obtain this 11\% markup as a rough value by aggregating various sources over a bunch of industrialized countries, with data mostly based on the second half of the 1990s.
@@
 Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-We thus find that the per-kg cost dimension is quantitatively sizeable.
-On average over the whole period, additive costs are estimated to 2.9\% of the export price for ocean shipping, and 1.8\%.
+On average over the whole period, additive costs are estimated to around 10 cents and 1.12 USD per kilogram exported, in maritime and air transport cost respectively (at the 3-digit level), indicating that freight rates are much higher in air transport.
+Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively.
 This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period).
 Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
-Expressed in monetary terms, the additive cost of exporting amounts to around 10 cents and 1.12 USD per kilogram exported, in maritime and air transport cost respectively (at the 3-digit level), indicating that freight rates are much higher in air transport.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward.
-In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
+This shows that the per-kg cost dimension is quantitatively sizeable.
+For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.
 Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for the related more theoretical papers, which need to calibrate their models.\smallskip
@@
 This stands in line with our own results, and makes us confident in our quantitative assessment of the size of the additive component in international transport costs.\footnote{See also Appendix \ref{app:diagnostic_test}, which confirms the importance of the additive cost component through another perspective angle based on quality-of-fit diagnostic tests.}\smallskip
-
-\textbf{VV: Les comparaisons chiffrees sont parfois délicates.
-Le paragraphe de comparaison avec Irarrazabal and co page 12 est tres clair, mais cela vient un peu tard.
-Il faut que des l’intro on comprenne bien la difference entre notre mesure de coûts additifs (representant entre 40 et 60\% des coûts de transport totaux) et les 14\% d’Irrara and co, qui sont un pourcentage du prix fab.
-MIS UN TOPO LA DESSUS EN INTRO.
-Ok AINSI?}
 

### 2018-02-19 redaction/dhp_july2017.tex
Levinshon The Box

--- 2018-02-19 redaction/dhp_july2017.tex
@@
 Equivalently, international transportation costs would represent a 21\% markup over production costs.
-Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) suggests a non-negligible impact of transport costs on trade flows.
+Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible.
 If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention.
@@
 Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
+%According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
+The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}).
 However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.

### 2018-02-19 redaction/dhp_july2017.tex
Petites modifications

@@
 The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}).
-However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.
+However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have not declined much.
 In this paper, we investigate the robustness of this result.
-As major contribution of the paper, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
+As a major contribution of the paper, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
 Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip
 
-To prove this point drives us to provide a careful modelling of transport costs.
+To prove this point we provide a careful modelling of transport costs.
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
 often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
 This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts.
-To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-This database is the same as \cite{hummels2007}, extended over the more recent years through (2005-2013).
+To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
+This database is the same as \cite{hummels2007}, extended over the more recent years (2005-2013).
 \smallskip
@@
 First, additive transport costs are quantitatively sizeable.
-On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
 This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period).
-Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model.
+Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode.
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{As we discuss below, the closest paper to us in this respect is \citet{Irrazabal_2015}.
@@
 We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
-Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
-
-Second, we find that the main source of the downward trend in overall transport costs over 1974-2013 does come from the reduction of transport cost ``per se'', rather than from trade composition effects.\footnote{Between 1974 and 2013, transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping.} This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.
+Additionnaly, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
+
+Second, we find that the main source of the downward trend in overall transport costs over 1974-2013 does come from the reduction of transport cost ``per se'', rather than from trade composition effects.\footnote{Between 1974 and 2013, transport costs as a share of export prices have decreased by roughly 50\% in air shipping and 60\% in vessel shipping.} This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
 Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
-Allowing the additive costs to vary  across products and country partners (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
+Allowing the additive costs to vary  across products and country partners (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+Our result is more in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip
 
 In both aspects, our results point the importance of the additive component in accounting for international transport costs.
-In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade.
+In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport: see \citealp{bosker2018ice}).
 As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
@@
 Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
-While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the 20\textsuperscript{th} century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports.
+While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the twentieth century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports.
 This result is also confirmed on more aggregate US, sector-level trade data.
@@
 A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
-Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff.
+This is confirmed by \cite{Irrazabal_2015}.
 While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.

### 2018-02-20 redaction/dhp_july2017.tex
changement texte début

@@
 \begin{abstract}
-This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel, in line with \cite{hummels2007}.
+This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
 This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components.
@@
 In this paper, we investigate the robustness of this result.
-As a major contribution of the paper, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
+As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
 Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip
@@
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs...
-often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see the references below).
+Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
 This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts.
@@
 Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode.
-To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{As we discuss below, the closest paper to us in this respect is \citet{Irrazabal_2015}.
-However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.} This represents a valuable insight for calibrating related international trade or business cycle models.
+To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
+This represents a valuable insight for calibrating related international trade or business cycle models.
 We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
 Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.
-Additionnaly, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
-
-Second, we find that the main source of the downward trend in overall transport costs over 1974-2013 does come from the reduction of transport cost ``per se'', rather than from trade composition effects.\footnote{Between 1974 and 2013, transport costs as a share of export prices have decreased by roughly 50\% in air shipping and 60\% in vessel shipping.} This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Additionally, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
+\textbf{attention maintenant en annexe.
+Le dire en intro, est ce pertinent?}
+
+Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se'', rather than from trade composition effects.
+This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
 Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
-Allowing the additive costs to vary  across products and country partners (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
 Our result is more in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip
@@
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport: see \citealp{bosker2018ice}).
-As pointed out by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
+As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
 In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
@@
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-Based on the estimated share of transport costs in total trade costs provided by \citet{anderson_wincoop_jel} (around 15\%), straightforward calculation yields a ratio of additive to multiplicative transport (and not trade) costs of 2.1\% (14\% $\times$ 15\%) in \citet{Irrazabal_2015}.
-This falls in the range of our own estimates, and makes us confident in our quantitative assessment of the size of international transport costs.
+Based on the estimated share of transport costs in total trade costs provided by \citet{anderson_wincoop_jel} (around 15\%), straightforward calculation yields a ratio of additive to multiplicative transport (not trade) costs of 2.1\% (14\% $\times$ 15\%) in \citet{Irrazabal_2015}.
+This falls in the range of our own estimates, making us confident in our quantitative assessment of the size of international transport costs.
 \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
@@
 In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+
+\textbf{enleve ancienne footnote plus haut qui reprenait cette difference, a mettre en footnote ici? en commentaire en dessous}
+%\footnote{As we discuss below, the closest paper to us in this respect is \citet{Irrazabal_2015}.
+However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}
 
@@
 Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others).
-The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau,\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}} spanning from 1974 to 2013.
+The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau, spanning from 1974 to 2013.
+Details on the database are provided in Appendix \ref{app:data}.
 We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
@@
 However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.
+Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors at the 3-digit level, from approximately 200 countries of origin.
+
+\textbf{stop here}
 
@@
 In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
+More information on this database is available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}.
 

### 2018-02-22 redaction/dhp_july2017.tex
Changement texte, encore et toujours

@@
 In particular, they are considered a major obstacle to international economic integration and international trade flows.
-According to \cite{Jacks08}, the trade cost decline explains around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse (but see \cite{corlay_2017}).
-After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO after 1995) multilateral agreements.
+After the Second World War though, the reduction of policy trade costs have been the workhorse of the GATT (WTO after 1995) rounds of multilateral negotiations.
 Based on panel data, \citet{novy13} finds  that U.S.
 trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} and \citealp{Disdier_Head08}).
+Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\cite{Head_Mayer04} or \cite{Disdier_Head08}, to name a few).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
 
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se} (\citet{daudin_logistique_2003, daudin_transactions_2005}).
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
 In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.
@@
 International transport costs accordingly stand at the heart of the paper.\smallskip
+
+\textbf{j'ai enleve les references a Guillaume parce que ca me semble pas bien place ici, je ne sais pas bien ce que ca veut dire a cet endroit. On peut rediscuter de ou les mettre.}
 
@@
 Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.
-Additionally, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom.
-\textbf{attention maintenant en annexe.
-Le dire en intro, est ce pertinent?}
 
@@
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price.
+Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns).
 Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
 Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data only allows recovering the ratio of additive to ad-valorem costs as a fraction of export price; in contrast, our empirical framework provides a distinct measure of each additive and ad-valorem cost component.
+First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-Based on the estimated share of transport costs in total trade costs provided by \citet{anderson_wincoop_jel} (around 15\%), straightforward calculation yields a ratio of additive to multiplicative transport (not trade) costs of 2.1\% (14\% $\times$ 15\%) in \citet{Irrazabal_2015}.
-This falls in the range of our own estimates, making us confident in our quantitative assessment of the size of international transport costs.
+While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid for additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping.
+This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
 \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
@@
 
-\textbf{enleve ancienne footnote plus haut qui reprenait cette difference, a mettre en footnote ici? en commentaire en dessous}
-%\footnote{As we discuss below, the closest paper to us in this respect is \citet{Irrazabal_2015}.
-However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}
+%\textbf{enleve ancienne footnote plus haut qui reprenait cette difference, a mettre en footnote ici? en commentaire en dessous}
+%However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
+By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.
 
@@
 
-\textbf{stop here}
-
 \subsection{Estimating transport costs: Empirical specification}
@@
 
-Under Model (\textit{(a)}), the estimated equation simplifies as:\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
+Under Model  \textit{(a)}, the estimated equation simplifies as:\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
 This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
@@
 After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.
+In this case, notice that the equation could be estimated relying on a linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.
+Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
@@
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
+    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
     Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
@@
     \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
-    Transport costs ($p/\widetilde{p}$, in \%) & & & & \\ \hline
+    Transport costs ($p/\widetilde{p} -1$, in \%) & & & & \\ \hline
     Mean & 5.3 & 5.0& 5.6&3.9 \\
@@
 Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-On average over the whole period, additive costs are estimated to around 10 cents and 1.12 USD per kilogram exported, in maritime and air transport cost respectively (at the 3-digit level), indicating that freight rates are much higher in air transport.
+On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
 Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively.
@@
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.
-Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for the related more theoretical papers, which need to calibrate their models.\smallskip
+Additionally, we run quality-of-fit diagnostic tests that confirms the importance of the additive cost component.
+The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom, as reported in Appendix \ref{app:diagnostic_test}.
+Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip
 
 Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology.
+Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), ie $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and picking up their estimated weighted mean value similarly to us).
 One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
 From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
-Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport.
-This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}.
+Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport.
+This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
 This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
-While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices).
-From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs to the export price $\widetilde{p}$ equal to 2.1\% ($0.15\times 0.14$).
-This stands in line with our own results, and makes us confident in our quantitative assessment of the size of the additive component in international transport costs.\footnote{See also Appendix \ref{app:diagnostic_test}, which confirms the importance of the additive cost component through another perspective angle based on quality-of-fit diagnostic tests.}\smallskip
+While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage).
+In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
+This obeys the following reasoning.
+\citet{Irrazabal_2015}'s results imply that, for a import price of 100 USD, 6 USD are paid in additive trade costs.
+For the same import price of 100 USD, our results yield the value of 2.7 (1.4) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
+With $x = 0.028$ and $0.014$ in vessel and air transport  in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport.
+This, on top of our results summarized in Table \ref{tab:summary_results}, point out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.
+
+\textbf{stop here.
+Grosse interrogation sur notre positionnement par rapport a Hummels dans la section suivante.
+Ai l'impresssion qu'on explique mal nos differences de resultats.
+On a l'air de dire que, contrairement a Hummels, les transport costs per se ont bien baisse, et que c'est ca qui compte.
+En regardant la figure "Characterizing the time trends: Applying Hummel's (2007) method", il me semble que c'est mal dit.
+Hummels ne dit pas que les couts purs n'ont pas baisse, d'ailleurs il montre qu'ils ont baisse, mais qu'au sein d'un mode de transport donne, ils sont en partie compense par les effets de composition.
+Nous, on montre que les couts purs ont baisse, et que les effets de composition, soit ne jouent pas (air), soit amplifient (au lieu de partiellement compenser, vessel) la baisse des couts purs.
+Il me semble qu'il faut reecrire l'interpretation des resultats en ce sens.}
 
@@
 He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped.
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p_{ikt}}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:
+Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:
 
@@
 \beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}}
+&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1}
 \end{eqnarray*}
@@
 The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption.
-This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} One may speculate that making the (empirically relevant) assumption of a varying share of additive costs over time, country and product is not anecdotal.
-Importantly, we confirm the key role of this assumption in analyzing the underlying determinants of transport costs trend patterns.
+This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the importance of this (empirically relevant) in analyzing the underlying determinants of transport costs trend patterns.
 We start presenting our estimation strategy.\smallskip
@@
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
-Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)).
+Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), but once again, they do not change significantly the abovementioned decreasing trend.
 Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
@@
 
-This last result stands in sharp contrast with \citet{hummels2007}, which obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects.
+This decreasing time trend we find for transport costs stands in sharp contrast with \citet{hummels2007}, who obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects.
 If anything, we find the opposite result here.
-This drives us to investigate this difference of result further.
+This drives us to investigate further this difference of result.
 As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.

### 2018-02-23 redaction/dhp_july2017.tex
changement texte

@@
 \begin{abstract}
-This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
-This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components.
-In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports.
-Transport cost ``per se'' did decline.
+This paper investigates the underlying sources behind the downward trend of international transport costs.
+In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in transport costs ``per se'' and changes in trade composition effects.
+This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
+In contrast to \cite{hummels2007}, the main driver of the overall transport costs decrease is the reduction of transport costs ``per se''.
+Further, if trade composition matter, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
@@
 
-\textbf{j'ai enleve les references a Guillaume parce que ca me semble pas bien place ici, je ne sais pas bien ce que ca veut dire a cet endroit. On peut rediscuter de ou les mettre.}
-
 This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others).
@@
 The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}).
-However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have not declined much.
+However, \cite{hummels2007} points out the role of trade composition effects in accounting for the time trends of international transport costs.
+Based on US imports over 1974-2004, he finds that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment.
+However, he also finds that trade composition effects partially compensate this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.
 In this paper, we investigate the robustness of this result.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
-Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip
+Rather, the marked reduction of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
+In maritime shipment, trade composition effects matter much, but they amplify (rather than offset) the reduction in pure transport costs.\medskip
 
@@
 
-Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se'', rather than from trade composition effects.
+Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''.
+Trade composition effects matter much, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs.
 This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
@@
 Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-Our result is more in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip
+Our result thus stands in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip
 
@@
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns).
+Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns).
 Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
@@
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid for additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping.
+While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping.
 This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
@@
 Section \ref{sec:results_trends} is devoted to this analysis.
-Section \ref{sec:robustness} provides some additional robustness checks of our results.
 Section \ref{sec:conclu} concludes.
@@
 This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
-While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage).
+While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
 In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
@@
 With $x = 0.028$ and $0.014$ in vessel and air transport in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport.
-This, on top of our results summarized in Table \ref{tab:summary_results}, point out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.
-
-\textbf{stop here.
-Grosse interrogation sur notre positionnement par rapport a Hummels dans la section suivante.
-Ai l'impresssion qu'on explique mal nos differences de resultats.
-On a l'air de dire que, contrairement a Hummels, les transport costs per se ont bien baisse, et que c'est ca qui compte.
-En regardant la figure "Characterizing the time trends: Applying Hummel's (2007) method", il me semble que c'est mal dit.
-Hummels ne dit pas que les couts purs n'ont pas baisse, d'ailleurs il montre qu'ils ont baisse, mais qu'au sein d'un mode de transport donne, ils sont en partie compense par les effets de composition.
-Nous, on montre que les couts purs ont baisse, et que les effets de composition, soit ne jouent pas (air), soit amplifient (au lieu de partiellement compenser, vessel) la baisse des couts purs.
-Il me semble qu'il faut reecrire l'interpretation des resultats en ce sens.}
+On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.
+In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.
 
@@
 In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.
-However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, yield to cast some doubt on the empirical relevance of this assumption.
+However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, cast some doubt on the empirical relevance of this assumption.
 As reported in Figure \ref{fig:decomp_TC_3d}, if overall transport costs have decreased over time, it is also the case for each additive and multiplicative components.
 In particular, the share of additive costs appears to be varying over time, in both transport modes.
-On top of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
+On light of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
 
@@
 As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution stands around 0.3-0.4 depending of the transport mode, consistently with our previous findings.
-In sharp contradiction with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, that would rather imply a distribution concentrated on a single point, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
-These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the  underlying sources of the overall transport costs trend patterns (between trade composition effects and transport costs ``per se'').
-This is made in next section.
+This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, that would rather imply a distribution concentrated on a single point.
+Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
+These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and transport costs ``per se''.
 
@@
 
-The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption.
-This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the importance of this (empirically relevant) in analyzing the underlying determinants of transport costs trend patterns.
-We start presenting our estimation strategy.\smallskip
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
+We start presenting our estimation strategy before turning to the results.\smallskip
 
 \paragraph{Estimation strategy} Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
-Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below).
+Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect.
+Precisely, starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
 For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time.
@@
 
-The remaining question is then, how to eliminate the composition effects?
-Our reasoning obeys the following rationale.
-Starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-
-\paragraph{Estimation strategy} This empirical strategy can be described as a two-stage process.
+\paragraph{Estimation method} This empirical strategy can be described as a two-stage process.
 First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
@@
 \noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
 Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
@@
 
-\noindent As a result, the two series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
 
@@
 Details are reported in Appendix \ref{app:comp-effects}.
+\textbf{en dire plus la dessus ici?
+Si c'est le point majeur de la comparaison avec Hummels, il me semble que oui.}
 
@@
 \begin{center}
-\includegraphics[height=4in]
+\includegraphics[height=5in]
 {graph_composition_all.pdf}
@@
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
-Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), but once again, they do not change significantly the abovementioned decreasing trend.
+Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), where they amplify the reduction in ``pure'' transport costs (the plain line being lower than the dotted line).
 Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
@@
 
-\textbf{VV: Figure 2 : 1) peu lisible en l’état.
-Soit on conserve ces graphs, et dans ce cas-la, on commente la dynamique, pas seulement le debut et la fin; si on veut rester sur seulement le debut et la fin, mettre un tableau.
-2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas...
+\textbf{VV: 2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas...
 et les couts totaux sont en-dessous !
 Déconcertant.
-Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.
-3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas.
-Ok AI REFORMULE LES CHOSES (SAUF POUR REMARQUE 2).
-OK ainsi?
-}
+Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.}
 
@@
 
-This decreasing time trend we find for transport costs stands in sharp contrast with \citet{hummels2007}, who obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects.
+These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the ``ceteris paribus'' transport costs for both air and maritime shipping.
 If anything, we find the opposite result here.
@@
 This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below.
-\footnote{As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
-Robustness analysis to this point is driven in Section \ref{sec:robustness}.} To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
+composition effects excluded ($\widehat{TC}_t$).
 
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partially offset the decrease in transport costs in both air and vessel shipping (as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel).\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel)} This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl}.
+Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partially offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl} \textbf{preciser le panel du bas?}.
+
 Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
@@
 
-%\section{Robustness checks \label{sec:robustness}}
-
-%\subsection{Measuring transport costs components: Robustness to the separability assumption}
-
-%Ie, estimate
-
-%$$\log(\frac{p_{ik}}{\widetilde{p}_{ik}} -1)= \log(\tau_{is} -1+ \frac{t_{is}}{\widetilde{p}_{ik}})+ \exp(\epsilon_{ik})$$
-
-%To do on 100 products ($s$), 50 countries (the most important?)
-
-\section{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}
-
-In this section, we provide a robustness test to the weighting scheme implemented to extract the trade composition effects implemented in Section \ref{sec:results_trends}.
+\subsection{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}
+
+As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
+In this section, we provide a robustness test to this difference in the weighting scheme implemented to extract the trade composition effects.\smallskip
+
 When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
+Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
 In this section, our objective is to assess that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
@@
 However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}.
-Rather than the weighing scheme, this confirms that the functional form and precisely the modeling of the additive component, turns out to be key in understanding the underlying determinants of the overall transport costs time trends, as we pointed out in Section \ref{sec:results_trends}.
+This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighing scheme, this is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
 
@@
 
-This paper empirically studies the magnitude of additive (or per-kg) costs in international transport costs, by exploiting the differences between the import and the export prices.
+This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
 Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
-Our results may be summarized in three main findings.
-First, we provide a quantitative measure of both the additive and the iceberg transport cost.
+Our results may be summarized in two main findings.
+First, we provide a quantitative measure of both the additive and the ad-valorem transport cost.
 We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%, as mean values over 1974-2013.
 These values are respectively equal to 1.8 and 2.5\% for air transport.
-Second, we show that taking additive costs into account improves the fit of the modeling of transport costs.
-All goodness-of-fit measures point out to this conclusion, which holds for both transport modes and all years considered.
-Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects.
-In all three aspects, our results point the importance of the additive component in accounting for international transport costs.
+Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
+Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
+If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs, in particular in maritime transport.
+In both aspects, our results point the importance of the additive component in accounting for international transport costs.
 

### 2018-02-23 redaction/dhp_july2017.tex
Les changements de Jérôme

@@
 trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\cite{Head_Mayer04} or \cite{Disdier_Head08}, to name a few).
+Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
@@
 Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible.
-If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention.
+If much of trade policy barriers have been removed over the second half of the 20\textsuperscript{th} century, this suggests that the transport costs component remains large and deserves attention.
 International transport costs accordingly stand at the heart of the paper.\smallskip
@@
 Based on US imports over 1974-2004, he finds that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment.
-However, he also finds that trade composition effects partially compensate this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.
+However, he also finds that trade composition effects partly offset this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.
 In this paper, we investigate the robustness of this result.
@@
 Rather, the marked reduction of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
-In maritime shipment, trade composition effects matter much, but they amplify (rather than offset) the reduction in pure transport costs.\medskip
+In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.\medskip
 
@@
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see the references below).
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}).
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
@@
 To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-This database is the same as \cite{hummels2007}, extended over the more recent years (2005-2013).
+This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013).
 \smallskip
@@
 Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''.
-Trade composition effects matter much, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs.
+Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs.
 This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
@@
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns).
+Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns).
 Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
@@
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping.
+While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive transport costs for maritime shipping and 1.4 USD for air transport.
 This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
@@
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
-Higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
+By construction higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
 This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
@@
 Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013.
-Notice that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
+Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
@@
 However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors at the 3-digit level, from approximately 200 countries of origin.
+Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.
 
@@
 
-As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international trade costs, that decomposes into both an additive and an ad-valorem component.
-In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\cite{Irrazabal_2015}, among others).
+As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, that decomposes into both an additive and an ad-valorem component.
+In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others).
 We contribute to this issue by adopting the following strategy.
@@
 Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode.
+We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for model \emph{(a)}, $\widehat{\tau}^{adv}$ and $\widehat{t}$ for model \textit{b)}, for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
-The first step of analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
-This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of transport costs.
+The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter.
 In this respect, these results \textit{per se} also constitute our first original contribution to the literature.
@@
 First, it provides estimation results about the size of the overall transport costs.
-From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping in average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
 Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
@@
 With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons.
-\citet{anderson_wincoop_jel} obtain this 11\% markup as a rough value by aggregating various sources over a bunch of industrialized countries, with data mostly based on the second half of the 1990s.
-By contrast, our estimation of a 6\% markup relies on a single source of data based on the US imports flows over 1974-2013.
-Beyond country coverage and the time sample, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
 
@@
 Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively.
-This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period).
-Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period).
+Put differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
 This shows that the per-kg cost dimension is quantitatively sizeable.
@@
 Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), ie $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and picking up their estimated weighted mean value similarly to us).
+Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), i.e.
+$\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
 One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
-From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect.
+From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect.
 Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport.
@@
 This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
-While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
+While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
 In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
@@
 
-\section{Transport Costs time trends: The role of the additive component}\label{sec:results_trends}
+\section{Transport costs time trends: The role of the additive component}\label{sec:results_trends}
 
@@
 Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner ``ceteris paribus'' transport costs and \textit{ii)} the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighbor countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
+Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
 As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
@@
 In particular, the share of additive costs appears to be varying over time, in both transport modes.
-On light of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).
+On light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country.
 
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
-To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, ie $\beta$, according to:
+To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e.
+$\beta$, according to:
 \begin{eqnarray*}
 \beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1}
+&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
 \end{eqnarray*}
@@
 
-As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution stands around 0.3-0.4 depending of the transport mode, consistently with our previous findings.
-This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, that would rather imply a distribution concentrated on a single point.
+As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending of the transport mode, consistently with our previous findings.
+This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would rather imply a distribution concentrated on a single point.
 Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
@@
 Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
-We start presenting our estimation strategy before turning to the results.\smallskip
+We start with the presentation our estimation strategy before turning to the results.\smallskip
 
@@
 \end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)).
-Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)).
+Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
 Both equations are estimated by transport mode.\smallskip
@@
 This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
-Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components (Panels (a) to (d)).
+Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of the same order for both the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
 Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport.
-For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line are very close to each other, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport.
+For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
@@
 If anything, we find the opposite result here.
-This drives us to investigate further this difference of result.
+This drives us to investigate further this difference.
 As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partially offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as depicted in Figure \ref{fig:totalTC_compeffects_excl} \textbf{preciser le panel du bas?}.
+Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in panels (c) and (f) from Figure \ref{fig:totalTC_compeffects_excl} \textbf{preciser le panel du bas?}.
 
@@
 
-To this aim, we rebuild the fitted values for each of of our transport cost measures (ad-valorem, additive and overall) applying Hummels' s (2007) weighting scheme (ie, taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
+To this aim, we rebuild the fitted values for each of of our transport cost measures (ad-valorem, additive and overall) applying \citet{hummels2007}'s weighting scheme (i.e., taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
 The results are reported in Figure \ref{fig:compeffects_robustness}.
@@
 
-The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives two comments.
+The comparison of Figure \ref{fig:compeffects_robustness} (applying \citet{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) drives two comments.
 First, differences do show up, in particular for the multiplicative component of Air transport.
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}.
-This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighing scheme, this is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
+This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
 
@@
 
+\clearpage
+
+%\appendix
+\setcounter{table}{0}
+\renewcommand{\thefigure}{A.\arabic{figure}}
+\renewcommand{\thetable}{A.\arabic{table}}
+
 \section{Data Appendix \label{app:data}}
@@
 import duties.
+
+\clearpage
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{B.\arabic{figure}}
+\renewcommand{\thetable}{B.\arabic{table}}
 
@@
 
+\clearpage
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{C.\arabic{figure}}
+\renewcommand{\thetable}{C.\arabic{table}}
+
 \section{Estimation at the 4-digit level \label{app:4digit}}
@@
 
+\newpage
+
+\clearpage
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{D.\arabic{figure}}
+\renewcommand{\thetable}{D.\arabic{table}}
+
 \section{Eliminating the composition effects: More details \label{app:comp-effects}}

### 2018-02-23 redaction/dhp_july2017.tex
Travail divers Guillaume

@@
 This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
-In contrast to \cite{hummels2007}, the main driver of the overall transport costs decrease is the reduction of transport costs ``per se''.
-Further, if trade composition matter, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+In contrast to \cite{hummels2007}, trade composition effects (inside each transport mode) do not matter much and tend to amplify (rather than reduce) the decrease in pure transport costs.
+This suggests that transport costs ``per se'' declined even less than argued by \cite{hummels2007}.
 Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
@@
 In particular, they are considered a major obstacle to international economic integration and international trade flows.
-After the Second World War though, the reduction of policy trade costs have been the workhorse of the GATT (WTO after 1995) rounds of multilateral negotiations.
 Based on panel data, \citet{novy13} finds  that U.S.
 trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
+Trade costs declined faster for more developed countries \cite{arvis2016trade}.
+Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
@@
 Equivalently, international transportation costs would represent a 21\% markup over production costs.
-Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible.
-If much of trade policy barriers have been removed over the second half of the 20\textsuperscript{th} century, this suggests that the transport costs component remains large and deserves attention.
+Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+If much of trade policy barriers have been removed over the second half of the twentieth century, this suggests that the transport costs component remains large and deserves attention.
 International transport costs accordingly stand at the heart of the paper.\smallskip
@@
 This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others).
-Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
+Many argue that transport costs have substantially decreased with technological advance in transportation (see \citealp{levinson2016box}), infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
 %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
-The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}).
-However, \cite{hummels2007} points out the role of trade composition effects in accounting for the time trends of international transport costs.
-Based on US imports over 1974-2004, he finds that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment.
-However, he also finds that trade composition effects partly offset this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.
+Yet, distance is still a very important determinant of trade flows and some have noticed that progress in transportation techniques must be compared with technical progress in production.
+The constant-dollar cost of transport has declined, but so has the constant-dollar cost of production.
+Furthermore, the quality of transport services have increased.
+As a result, transport costs as a share of export prices have declined less quickly than expected  (\citealp{hummels2007}).\smallskip
+
+According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by a composition effect.
+Based on US imports over 1974-2004, he finds that transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment, but less so than ``per se'' transport costs because trade composition effects have partly offset this ``per se'' transport costs decrease, thereby attenuating the overall transport costs downward trend.
 In this paper, we investigate the robustness of this result.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
-Rather, the marked reduction of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
-In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.\medskip
+Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
+In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that transport costs ``per se'' declined less than argued by {hummels2007}.\medskip
 
@@
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
-This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts.
+This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
 To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
@@
 Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''.
-Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs.
-This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, suggesting that ``per se'' trade costs declined less than observed trade costs.
+This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
 Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
-Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-Our result thus stands in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip
+Allowing the share of additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+Our result thus deepens the paradox of the less-than-expected decline of transport costs.\smallskip
 
@@
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport: see \citealp{bosker2018ice}).
-As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of trade costs as long as they display an additive component.
+As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs.
 In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
@@
 The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, in contradiction with the ad-valorem hypothesis.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
@@
 
-Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data.
-Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns).
+Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring relative additive trade costs from firm-level trade data.
+Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (expressed in Norwegian crowns).
 Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
 Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
+First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost ; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive transport costs for maritime shipping and 1.4 USD for air transport.
+While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point (for 2004) to a value of 2.8 USD paid in additive transport costs for maritime shipping and 1.5 USD for air transport.
 This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
@@
 Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others).
-The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau, spanning from 1974 to 2013.
+The database we use to construct our measure of transport costs comes from US annual "Imports of Merchandise" provided by the Census bureau, spanning from 1974 to 2013.
 Details on the database are provided in Appendix \ref{app:data}.
@@
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, we refer to the fas price as the price considered by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
+While both terms are closely related, the fas price is the one reported by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
 The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
@@
 Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), i.e.
+Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e.
 $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
@@
 From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect.
-Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport.
+Making use of our estimates for 2004 (see tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport.
 This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
@@
 This obeys the following reasoning.
-\citet{Irrazabal_2015}'s results imply that, for a import price of 100 USD, 6 USD are paid in additive trade costs.
-For the same import price of 100 USD, our results yield the value of 2.7 (1.4) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
-With $x = 0.028$ and $0.014$ in vessel and air transport in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport.
+\citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
+For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
+With $x = 0.029$ and $0.015$ in vessel and air transport in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.028 and 0.015 in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
 On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.
@@
 Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``ceteris paribus'' international transport costs, i.e.
+As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``per se'' or ``ceteris paribus'' international transport costs, i.e.
 per product- and per partner- transport costs.
@@
 Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes (Panels (c) and (f)).
+First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
 International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
 This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
-Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of the same order for both the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
+Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
 Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
@@
 
-These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the ``ceteris paribus'' transport costs for both air and maritime shipping.
+These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the ``per se'' or ``ceteris paribus'' transport costs for both air and maritime shipping.
 If anything, we find the opposite result here.
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in panels (c) and (f) from Figure \ref{fig:totalTC_compeffects_excl} \textbf{preciser le panel du bas?}.
+Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in panels (c) and (f) from Figure \ref{fig:totalTC_compeffects_excl}, panels c and f.
 
@@
 
-As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time.
+As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``per se'' or ``ceteris paribus'' transport costs over time.
 In this section, we provide a robustness test to this difference in the weighting scheme implemented to extract the trade composition effects.\smallskip
@@
 \begin{center}
-    \begin{tabular}{l|cccccc}
+    \begin{tabular}{l|ccccccc}
 \hline\hline
-Year & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
-\hline
-Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.2 & 3.4  \\
-Median & 5.4 & 3.8 & 4.4 & 2.5 & 3.4 & 2.9  \\
+Year & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
+\hline
+\multicolumn{8}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
+\hline
+Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.0 & 4.2 & 3.4  \\
+Median & 5.4 & 3.8 & 4.4 & 2.5 & 2.9 & 3.4 & 2.9  \\
 %Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
 \hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
-Mean & 3.6 & 2.3 & 2.4 &1.7 & 2.6 & 1.7  \\
-Median & 2.7 & 1.6 & 1.6 & 1.2 & 2.2 & 1.7 \\
+\multicolumn{8}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
+Mean & 3.6 & 2.3 & 2.4 &1.7 & 1.9 & 2.6 & 1.7  \\
+Median & 2.7 & 1.6 & 1.6 & 1.2 & 1.4 & 2.2 & 1.7 \\
 %Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
@@
 \hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
-Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.1 & 1.0  \\
-Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.4 & 0.5  \\
+\multicolumn{8}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
+Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.5 & 1.1 & 1.0  \\
+Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.6 & 0.4 & 0.5  \\
 %\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
 \hline
-\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351  \\
+\# observations & 14,955 & 16,118 & 24,958 & 35,027 & ??? & 40,279 & 39,351  \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+\multicolumn{8}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its share in trade (mode-dependent).
 Additive term expressed in fraction of fas price.}}}
@@
 \begin{center}
-    \begin{tabular}{l|cccccc}
+    \begin{tabular}{l|ccccccc}
    \hline\hline
-Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
- \hline
-   \multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
-   \hline
-Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 4.0 & 3.6  \\
-Median & 9.6 & 5.5 & 4.6 & 4.9 & 3.6 & 3.3  \\
+Year         & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
+ \hline
+   \multicolumn{8}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
+   \hline
+Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 5.4 & 4.0 & 3.6  \\
+Median & 9.6 & 5.5 & 4.6 & 4.9 & 5.1  & 3.6 & 3.3  \\
 %Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
 \hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{7}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
-\hline
-Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 1.9 & 2.2 \\
-Median & 4.9 & 2.4 & 2.8 & 2.1 & 1.8 & 1.8  \\
+\multicolumn{8}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{8}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
+\hline
+Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 2.7 & 1.9 & 2.2 \\
+Median & 4.9 & 2.4 & 2.8 & 2.1 & 2.8 & 1.8 & 1.8  \\
 %Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
 \hline
-\multicolumn{7}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
-\hline
-Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.5 & 1.5  \\
-Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 0.8 \\
+\multicolumn{8}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
+\hline
+Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.9 & 2.5 & 1.5  \\
+Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 1.9 & 0.8 \\
 %Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
 \hline
- \# observations & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\
+ \# observations & 19,007 & 17,356 & 28,383 & 36,090 & ??? & 37,748 & 38,473 \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+\multicolumn{8}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its share in trade (mode-dependent).
 Additive term expressed in fraction of fas price.}}}
@@
 \multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
+Statistics are obtained weighting each observation by its share in trade (mode-dependent).
 Additive term expressed in fraction of fas price.}}}

### 2018-03-19 redaction/dhp_july2017.tex
Pour unifier l’apparence des graphiques 3 et 4
Ceux sur l’effet de composition
Graphiques d’Hummels en base 100

@@
 {\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_air.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_ocean.pdf} \\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels.pdf} \\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_base100.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_base100.pdf} \\
+
 \end{tabular}

### 2018-03-19 redaction/dhp_july2017.tex
détails annexe
Chnagement place annexe

@@
 
-\paragraph{Estimation strategy} Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
+\subsubsection{Estimation strategy}
+
+Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
 Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect.
@@
 
-\paragraph{Estimation method} This empirical strategy can be described as a two-stage process.
-First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-
+The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and overall resp.), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and overall respectively).
+We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).
+
+\subsubsection{Estimation method for each additive and multiplicative component}
+
+We start describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series. \smallskip
+
+\paragraph{Obtaining the ``pure'' transport costs component series} For each additive and multiplicative transport costs component, the empirical strategy to get the fitted transport cost measure (i.e., extracting from trade composition effect) can be described as a two-stage process.
+First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
 For the estimated ad-valorem component, we thus estimate the following equation:
@@
 This constitutes the second stage of our procedure.
-As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:\footnote{See details in Appendix \ref{app:comp-effects}.}
+As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
 \begin{equation}
@@
 
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974.
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (See details in Appendix \ref{app:comp-effects}).
 In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
@@
 
-Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded).
-Details are reported in Appendix \ref{app:comp-effects}.
-\textbf{en dire plus la dessus ici?
-Si c'est le point majeur de la comparaison avec Hummels, il me semble que oui.}
+\paragraph{Obtaining the unfitted measures} The objective here is to get the unfitted transport cost component (additive and multiplicative) as as index, with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
+That is, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
+
+$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
+
+\subsubsection{Estimation method for the total transport cost measures}
+
+We also build two measures of the ``total'' transport costs, that agglomerates our estimates of the two additive and ad-valorem components, for both the unfitted series and the ``pure'' transport cost series (i.e., composition effects excluded).
+As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
+Even if obeying to the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
+\smallskip
+
+\paragraph{Obtaining the unfitted total transport cost index} For the unfitted measure, we build for each transport mode, the ``total'' transport cost based on Equation (\ref{eq:base_estimee}), according to:
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ have been estimated (by year) as explained in Section \ref{sec:data_method}.
+Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\widehat{t}_t$ the additive component, both expressed in percentage of the fas price.
+We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
+
+\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to construct the fitted measure of total transport cost (i.e., composition effect excluded), with one notable exception though given that the fitted ad-valorem and additive components have not been estimated in value, but extracted and build as indices.
+In a first step then, we re-build the fitted measures of each transport cost component in value starting from these indices.
+For the ad-valorem component, this can be obtained making use of Equation (\ref{eq:tcadv_compoeffect}), rewritten to get:
+
+$$\widehat{\tau}^{cp, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
+
+with $\widehat{\tau}^{cp, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) ad-valorem cost.
+A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
+
+$$\widehat{t}^{cp}_t = \frac{\Gamma^{add}_t}{100}$$
+
+with $\widehat{t}^{cp}_t \equiv \exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) additive cost.
+
+We then deduce the fitted value of the overall cost according to:
+$$\widehat{tc}^{cp}_t= \widehat{\tau}^{cp, adv}_t  + \widehat{t}^{cp}_t$$
+
+Last, we transform this (fitted) ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{cp}_t -1 }{\widehat{tc}^{cp}_{1974}-1}$$
 
@@
 
-\textbf{VV: 2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas...
-et les couts totaux sont en-dessous !
-Déconcertant.
-Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.}
-
 \subsection{Time trends in transport costs and the modeling of the additive component}
@@
 \hline
-\# observations & 14,955 & 16,118 & 24,958 & 35,027 & ??? & 40,279 & 39,351  \\
+\# observations & 14,955 & 16,118 & 24,958 & 35,027 & 36,990 & 40,279 & 39,351  \\
 \hline\hline
@@
 \hline
- \# observations & 19,007 & 17,356 & 28,383 & 36,090 & ??? & 37,748 & 38,473 \\
+ \# observations & 19,007 & 17,356 & 28,383 & 36,090 & 37,757 & 37,748 & 38,473 \\
 \hline\hline
@@
 Based on this equation, we deduce after estimation that:
-\begin{eqnarray*}
-\text{For the year 1974}&& \widehat{\tau}_{is74} = \exp(\delta +\alpha_i+\beta_s), \\
-\text{For any year}~t> 1974&& \widehat{\tau}_{ist} = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t)
-\end{eqnarray*}
+
+\[\left\{
+  \begin{array}{lcl}
+  \text{For the year 1974:} &&  \widehat{\tau}_{is74} = \exp(\delta +\alpha_i+\beta_s), \\
+  \text{For any year }t> 1974:&& \widehat{\tau}_{ist} = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t) \\
+  \end{array}
+\right.\]
 
@@
 \paragraph{For the additive component} After estimating Equation (\ref{eq:compeffects_add}), we can re-build the additive component according to:
-\begin{eqnarray*}
-\text{For the year 1974}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
-\text{For any year}~t> 1974&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
-\end{eqnarray*}
+\[\left\{
+  \begin{array}{lcl}
+\text{For the year 1974:}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
+\text{For any year}~t> 1974:&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
+\end{array}
+\right.\]
 
@@
 \end{equation}
-
-\paragraph{Obtaining the unfitted measures} It consists in expressing the unfitted transport cost component as as index, with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
-That is, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
-
-$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
-
-\subsubsection{For the total cost measure}
-
-We also build a measure of the ``overall'' transport costs, that agglomerates our estimates of the two additive and iceberg components.
-We construct this measure (by transport mode) on both the unfitted series and the ``pure'' transport cost series (composition effects excluded).
-Even if obeying to the same logic, we proceed slightly differently for the unfitted and the fitted measures though, as we now explain.\smallskip
-
-For the unfitted measure, based on Equation (\ref{eq:base_estimee}), we build for each transport mode, the ``overall'' transport cost as:
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ have been estimated (by year) as explained in Section \ref{sec:data_method}, $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\widehat{t}_t$ the additive component, both expressed in percentage of the fas price.
-For sake of comparison, we transform this ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
-
-We apply the same logic to construct the fitted measure of total transport cost (i.e., composition effect excluded), with a first step that consists in getting the fitted measures of each transport cost component in value (rather than as an index).
-For the ad-valorem component, this can be obtained starting from Equation (\ref{eq:tcadv_compoeffect}), rewritten to get:
-
-$$\widehat{\tau}^{cp, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
-
-with $\widehat{\tau}^{cp, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) ad-valorem cost.
-A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
-
-$$\widehat{t}^{cp}_t = \frac{\Gamma^{add}_t}{100}$$
-
-with $\widehat{t}^{cp}_t \equiv \exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) additive cost.
-
-We then deduce the fitted value of the overall cost according to:
-$$\widehat{tc}^{cp}_t= \widehat{\tau}^{cp, adv}_t  + \widehat{t}^{cp}_t$$
-
-Last, we transform this (fitted) ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{cp}_t -1 }{\widehat{tc}^{cp}_{1974}-1}$$
-
-This finally leaves us with six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and overall resp.), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and overall respectively).
 

### 2018-03-19 redaction/dhp_july2017.tex
Changer les graphiques Hummels

--- 2018-03-19 redaction/dhp_july2017.tex
@@
 & \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels.pdf} \\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_base100.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_base100.pdf} \\
 
 \end{tabular}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Composition effect : comparing Hummels with our best guess}
+\label{fig:comparingHummels_pond}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_comp_trend_hum_nous_.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Composition effect : comparing Hummels with our results with his weighting scheme}
+\label{fig:comparingHummels_np}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_comp_trend_hum_nous__np.pdf}
 \end{center}

### 2018-03-23 redaction/dhp_july2017.tex
changement texte
pas fini, seulement abstract

@@
 \begin{abstract}
-This paper investigates the underlying sources behind the downward trend of international transport costs.
-In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in transport costs ``per se'' and changes in trade composition effects.
-This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
-In contrast to \cite{hummels2007}, trade composition effects (inside each transport mode) do not matter much and tend to amplify (rather than reduce) the decrease in pure transport costs.
+This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
+In the same spirit as \cite{hummels2007}, we disentangle the role of trade composition effects from the reduction in ``pure'' transport costs over time.
+This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components.
+In contrast to \cite{hummels2007}, we find that trade composition effects (inside each transport mode) do not matter much.
+Further, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 This suggests that transport costs ``per se'' declined even less than argued by \cite{hummels2007}.

### 2018-03-29 redaction/dhp_july2017.tex
changement texte

--- 2018-03-19 redaction/dhp_july2017.tex
@@
 This paper investigates the underlying sources behind the downward trend of international transport costs.
-In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in transport costs ``per se'' and changes in trade composition effects.
+In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
 This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
-In contrast to \cite{hummels2007}, trade composition effects (inside each transport mode) do not matter much and tend to amplify (rather than reduce) the decrease in pure transport costs.
-This suggests that transport costs ``per se'' declined even less than argued by \cite{hummels2007}.
+In contrast to \cite{hummels2007}, trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
@@
 This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others).
-Many argue that transport costs have substantially decreased with technological advance in transportation (see \citealp{levinson2016box}), infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-%According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
-Yet, distance is still a very important determinant of trade flows and some have noticed that progress in transportation techniques must be compared with technical progress in production.
+Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
+This view does not make consensus though.
+Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques should be compared with technical progress in production to get a clear picture of the transport costs burden.
 The constant-dollar cost of transport has declined, but so has the constant-dollar cost of production.
-Furthermore, the quality of transport services have increased.
-As a result, transport costs as a share of export prices have declined less quickly than expected  (\citealp{hummels2007}).\smallskip
-
-According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by a composition effect.
-Based on US imports over 1974-2004, he finds that transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment, but less so than ``per se'' transport costs because trade composition effects have partly offset this ``per se'' transport costs decrease, thereby attenuating the overall transport costs downward trend.
-In this paper, we investigate the robustness of this result.
+As well, the quality of transport services has increased.
+In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip
+
+According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries).
+Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
 Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
-In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that transport costs ``per se'' declined less than argued by {hummels2007}.\medskip
-
-To prove this point we provide a careful modelling of transport costs.
+In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip
+
+Proving this point drives us to provide a careful modelling of transport costs.
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}).
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015} or \citealp{martin2012}).
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
@@
 Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''.
-Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, suggesting that ``per se'' trade costs declined less than observed trade costs.
+Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs, suggesting that pure transport costs declined less than observed costs.
 This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
@@
 Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-Our result thus deepens the paradox of the less-than-expected decline of transport costs.\smallskip
+Our result thus deepens the paradox of the less-than-expected decline of transport costs.
+\textbf{j'aime bien parler de ce paradoxe mais la seule source qu'on cite c'est Hummels 2007, d'autres papiers?}\smallskip
 
 In both aspects, our results point the importance of the additive component in accounting for international transport costs.
-In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport: see \citealp{bosker2018ice}).
+In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
 As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs.
@@
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point (for 2004) to a value of 2.8 USD paid in additive transport costs for maritime shipping and 1.5 USD for air transport.
+While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point (for 2004) to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport.
 This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
-\footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
+Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
@@
 Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
-In other words, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
+In other words, this amounts making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
 same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
@@
 We contribute to this issue by adopting the following strategy.
-For each year and transport mode, we estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
-From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.
-
-Under Model  \textit{(a)}, the estimated equation simplifies as:\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
+For each year and transport mode, we estimate two models.
+In Model (a), additive costs are excluded, transport costs being modeled as iceberg costs only (Equation (\ref{eq:model_nlI})).
+In this case, the estimated equation simplifies as:
+
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
+\end{equation}
+
+Under Model (b), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
+From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
 This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
@@
 These results are reported in the Online Appendix, available on the authors' webpages.}
-
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
-\end{equation}
 
@@
 Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for model \emph{(a)}, $\widehat{\tau}^{adv}$ and $\widehat{t}$ for model \textit{b)}, for each year and transportation mode.
+We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (a), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (b), for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter.
-In this respect, these results \textit{per se} also constitute our first original contribution to the literature.
+These results constitute our first original contribution to the literature.
 In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip
@@
 It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
-The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})).
-Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level.
+The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
+Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3- and 4-digit classification level.
 Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
@@
     \multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
+    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
     Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
@@
     \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
-   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
+   \textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} & & & & \\ \hline
     Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
@@
     \textit{Additive term }& & & &   \\ \hline
-    \textit{In \% of the export price} &&&& \\ \hline
+    \textit{In \% of the export price ($\widehat{t}/\widetilde{p}$, in \%)} &&&& \\ \hline
         Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
     Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
-   \textit{ In USD per kg traded }&&&& \\ \hline
+   \textit{In USD per kg traded ($\widehat{t}$)}&&&& \\ \hline
     Mean & 0.09	&1.12	&0.08&	2.16 \\
@@
 The additive term is expressed in fraction of fas price.
-($^\ast$): Four 4-digit estimation: 0n selected years.
-($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}\\
-\parbox[l]{10cm}{\tiny{Notes: For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
-1974, 1981, 1989, 2001, 2009, 2013.}}
+For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
+1974, 1981, 1989, 2001, 2009, 2013.
+($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
 \end{table}%
@@
 The first type is of statistical order.
-As displayed in Table \ref{tab:summary_results}, mean values are systematically higher than medians, by around one percentage point.
-This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median.
+As displayed in Table \ref{tab:summary_results}, estimated mean values are systematically higher than medians, by around one percentage point.
+This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
 One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
@@
 First, it provides estimation results about the size of the overall transport costs.
-From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
-Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
+From Table \ref{tab:summary_results}, transport costs amount to around a 6\% markup in ocean shipping, 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+Under Model (a) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
+Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
 With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible candidate explanation for the difference with \citet{anderson_wincoop_jel}, lays in the difference of empirical methodology.
+\citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
 
@@
 This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period).
-Put differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air.
+Put differently, the additive component amounts to 48\% of total transport costs in Vessel, and 42\% in Air.
 This shows that the per-kg cost dimension is quantitatively sizeable.
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.
-Additionally, we run quality-of-fit diagnostic tests that confirms the importance of the additive cost component.
-The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom, as reported in Appendix \ref{app:diagnostic_test}.
-Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip
+Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests, that confirm the importance of the additive cost component.
+The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom.
+An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip
 
@@
 In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
-This obeys the following reasoning.
 \citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
-For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
-With $x = 0.029$ and $0.015$ in vessel and air transport in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.028 and 0.015 in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
+For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
+Given our estimates for 2004 (see Table \ref{tab:result_air_3d_detail} in Appendix \ref{app:more_results}) such that $x = 0.028$ and $0.015$ in vessel and air transport respectively, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
 On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.
@@
 Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-\cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms.
-However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much.
-In this paper, we investigate the robustness of this result.
+There is no clear evidence about this though, in particular when expressing the transport cost burden in terms of the export price.
+According to \cite{hummels2007}, transport costs in proportion of export prices have thus declined less quickly than expected.
+His results point as an important explanation, that the reduction in pure transport costs in both maritime and air shipping is partially offset by trade composition effects, thereby attenuating the downward trend in overall (observed) transport costs.
+We now investigate the robustness of this result.
+\textbf{stop here}.
 
@@
 & \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels.pdf} \\
+\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_base100.pdf}
+& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_base100.pdf} \\
 
 \end{tabular}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Composition effect : comparing Hummels with our best guess}
-\label{fig:comparingHummels_pond}
-\begin{center}
-\includegraphics[height=8cm]
-{graph_comp_trend_hum_nous_.pdf}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Composition effect : comparing Hummels with our results with his weighting scheme}
-\label{fig:comparingHummels_np}
-\begin{center}
-\includegraphics[height=8cm]
-{graph_comp_trend_hum_nous__np.pdf}
 \end{center}

### 2018-03-30 redaction/dhp_july2017.tex
changement papier

@@
 This paper investigates the underlying sources behind the downward trend of international transport costs.
-In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
-This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
-In contrast to \cite{hummels2007}, trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
-Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
+In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, using information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
+In contrast to \cite{hummels2007} though, we find that trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer.
 Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs.
@@
 This view does not make consensus though.
-Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques should be compared with technical progress in production to get a clear picture of the transport costs burden.
-The constant-dollar cost of transport has declined, but so has the constant-dollar cost of production.
-As well, the quality of transport services has increased.
+Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
 In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip
@@
 Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
-In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip
+In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
+In this respect, our results suggest that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip
 
@@
 In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
-This focus is shared with part of the international trade literature, such as \cite{hummels2007} and \cite{Behar_Venables}.
-Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}).
-There is no clear evidence about this though, in particular when expressing the transport cost burden in terms of the export price.
-According to \cite{hummels2007}, transport costs in proportion of export prices have thus declined less quickly than expected.
-His results point as an important explanation, that the reduction in pure transport costs in both maritime and air shipping is partially offset by trade composition effects, thereby attenuating the downward trend in overall (observed) transport costs.
-We now investigate the robustness of this result.
-\textbf{stop here}.
+This focus is shared with part of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
+Specifically, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, in the same spirit as \cite{hummels2007}.
 
@@
 Two main comments can be made.
-First, in both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time.
+First, in both transport modes, the additive component appears of sizeable importance all over the period.
 Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
 By contrast, they represent an sizeable and structural dimension of international transport costs.
+However, one can also notice that, if the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport over the period, it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered (see Figure \ref{fig:part_cout_additif}, Appendix \ref{app:more_results}).\smallskip
 
@@
 For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
-Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.\smallskip
-
-Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner ``ceteris paribus'' transport costs and \textit{ii)} the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``per se'' or ``ceteris paribus'' international transport costs, i.e.
-per product- and per partner- transport costs.
+Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip
+
+Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
+Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased , independently of any change in transport costs \textit{per se}.
+As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
+Our own results tend to challenge this view, as we further show.
 
 \paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
-He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped.
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:
+He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
+Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes down as:
 
@@
 Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
-Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which Hummels (2007) decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}.
-
-In doing so, Hummels (2007) implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
+Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}.
+In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
 In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.
-However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, cast some doubt on the empirical relevance of this assumption.
-As reported in Figure \ref{fig:decomp_TC_3d}, if overall transport costs have decreased over time, it is also the case for each additive and multiplicative components.
-In particular, the share of additive costs appears to be varying over time, in both transport modes.
-On light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country.
+
+\paragraph{Challenging the $\beta$'s constancy assumption} Our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:part_cout_additif}, cast some doubt on the empirical relevance of this assumption.
+As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes.
+In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
 
@@
 Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
-These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and transport costs ``per se''.
+These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
 
@@
 
-Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country.
-Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect.
-Precisely, starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
+Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
+Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+
 One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time.
-We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}
-
-The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and overall resp.), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and overall respectively).
+For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
 We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).
@@
 
-\paragraph{Obtaining the ``pure'' transport costs component series} For each additive and multiplicative transport costs component, the empirical strategy to get the fitted transport cost measure (i.e., extracting from trade composition effect) can be described as a two-stage process.
+\paragraph{Obtaining the ``pure'' transport costs component series} The empirical strategy to get the fitted transport cost measure (i.e., extracting from trade composition effect) can be described as a two-stage process.
 First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-For the estimated ad-valorem component, we thus estimate the following equation:
+For the estimated ad-valorem component, we estimate the following equation:
 \begin{eqnarray}
@@
 \end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)).
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
 Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
@@
 
-\paragraph{Obtaining the unfitted measures} The objective here is to get the unfitted transport cost component (additive and multiplicative) as as index, with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
-That is, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
+\paragraph{Obtaining the unfitted measures} The objective is to get the unfitted transport cost component (additive and multiplicative) built as as index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
+To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
 
@@
 
-We also build two measures of the ``total'' transport costs, that agglomerates our estimates of the two additive and ad-valorem components, for both the unfitted series and the ``pure'' transport cost series (i.e., composition effects excluded).
+We also build two measures of the ``total'' transport costs, that agglomerates our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e., composition effects excluded).
 As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
@@
 
-\paragraph{Obtaining the unfitted total transport cost index} For the unfitted measure, we build for each transport mode, the ``total'' transport cost based on Equation (\ref{eq:base_estimee}), according to:
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ have been estimated (by year) as explained in Section \ref{sec:data_method}.
-Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\widehat{t}_t$ the additive component, both expressed in percentage of the fas price.
+\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the ``total'' transport cost based on Equation (\ref{eq:base_estimee}) according to:
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
+Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed in percentage of the fas price.
 We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
@@
 
-\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to construct the fitted measure of total transport cost (i.e., composition effect excluded), with one notable exception though given that the fitted ad-valorem and additive components have not been estimated in value, but extracted and build as indices.
+\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to construct the fitted measure of total transport cost (i.e., composition effect excluded), with one notable exception though: Fitted ad-valorem and additive components have not been estimated in value, but extracted and build as indices.
 In a first step then, we re-build the fitted measures of each transport cost component in value starting from these indices.
@@
 
-$$\widehat{\tau}^{cp, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
-
-with $\widehat{\tau}^{cp, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) ad-valorem cost.
+$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
+
+with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
 A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
 
-$$\widehat{t}^{cp}_t = \frac{\Gamma^{add}_t}{100}$$
-
-with $\widehat{t}^{cp}_t \equiv \exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) additive cost.
+$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
+
+with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.
 
 We then deduce the fitted value of the overall cost according to:
-$$\widehat{tc}^{cp}_t= \widehat{\tau}^{cp, adv}_t  + \widehat{t}^{cp}_t$$
-
-Last, we transform this (fitted) ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{cp}_t -1 }{\widehat{tc}^{cp}_{1974}-1}$$
+$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$
+
+Last, we transform this (fitted) measure of total transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
+
+\textbf{stop here}
 
@@
 \hline
-\multicolumn{8}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
+\multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}     \\
 \hline
@@
 \hline
-\multicolumn{8}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
+\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%)}  \\ \hline
 Mean & 3.6 & 2.3 & 2.4 &1.7 & 1.9 & 2.6 & 1.7  \\
@@
  \hline
-   \multicolumn{8}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}  \\
+   \multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}  \\
    \hline
@@
 \hline
-\multicolumn{8}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{8}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} } \\
+\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
+\hline
+\multicolumn{8}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} } \\
 \hline
@@
 \end{table}%
+
+\paragraph{More on additive costs} In Figure , we report the share of additive costs in total transport costs over time in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}, conditional on a year and a transport mode).
+Precisely, each point in time is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent (i.e., in percent of the export price $\widetilde{p}$).
+
+\begin{figure}[htbp]
+\caption{Share of additive costs}
+\label{fig:part_cout_additif}
+\begin{center}
+\includegraphics[width=10cm, height=7cm]{part_cout_additif.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
 

### 2018-04-04 redaction/dhp_july2017.tex

@@
 \begin{abstract}
-This paper investigates the underlying sources behind the downward trend of international transport costs.
-In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, using information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel.
-In contrast to \cite{hummels2007} though, we find that trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+This paper investigates the role of additive costs in international transport costs, using information contained in the US imports flows database over 1974-2013.
+We find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs.
+Further, in the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs.
+In contrast to him, we find that trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer.
-Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs.
 In both aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
 This view does not make consensus though.
-Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
+Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
 In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip
@@
 According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries).
-Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result.
+Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
@@
 Our result thus deepens the paradox of the less-than-expected decline of transport costs.
-\textbf{j'aime bien parler de ce paradoxe mais la seule source qu'on cite c'est Hummels 2007, d'autres papiers?}\smallskip
+\smallskip
 
@@
 From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point (for 2004) to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport.
+While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
 This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
@@
 
+\textbf{a ajuster et remonter figure annexe.}
+
+Two main comments can be made.
+First, in both transport modes, the additive component appears of sizeable importance all over the period.
+Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
+By contrast, they represent an sizeable and structural dimension of international transport costs.
+However, one can also notice that, if the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport over the period, it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered (see Figure \ref{fig:part_cout_additif}, Appendix \ref{app:more_results}).\smallskip
+
 \section{Transport costs time trends: The role of the additive component}\label{sec:results_trends}
@@
 
-As a first step, Figure \ref{fig:decomp_TC_3d} displays the respective shares of additive (black area) and iceberg (grey area) components in total transport costs in total costs (the total of the two, expressed in percentage of the fas price), by transport mode for each year between 1974 and 2013.
+As a first step, Figure \ref{fig:decomp_TC_3d} displays the \textbf{total transport costs (the total of the two, expressed in percentage of the fas price)}, by transport mode for each year between 1974 and 2013.
 
 \begin{figure}[htbp]
-\caption{Decomposing Transport costs (Yearly mean value, 3 digits, in \% of the fas price)}
+\caption{Trend in Transport costs (Yearly mean value, 3 digits, in \% of the fas price)}
 \label{fig:decomp_TC_3d}
@@
 
-Two main comments can be made.
-First, in both transport modes, the additive component appears of sizeable importance all over the period.
-Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-By contrast, they represent an sizeable and structural dimension of international transport costs.
-However, one can also notice that, if the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport over the period, it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered (see Figure \ref{fig:part_cout_additif}, Appendix \ref{app:more_results}).\smallskip
-
-Second, considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+\textbf{figure a changer et texte a adapter}
+
+Figure \ref{fig:decomp_TC_3d} reports that both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
 \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.

### 2018-04-04 redaction/dhp_july2017.tex
Figure 1 more

@@
 \begin{figure}[htbp]
-\caption{Trend in Transport costs (Yearly mean value, 3 digits, in \% of the fas price)}
+\caption{Trend in Transport costs (Yearly mean value, 3 digits, in \% of the fas price, 1989 exclude for "Air")}
 \label{fig:decomp_TC_3d}
 \begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3in, height=2.5in]{Fig2a_decompTC_air_3d.pdf}
-& \includegraphics[width=3in,height=2.5in]{Fig2b_decompTC_vessel_3d.pdf} \\
-\end{tabular}
+\includegraphics[width=6.5in, height=5in]{Figure1_Trend_of_total_TC_by_mode.pdf}
 \end{center}

### 2018-04-05 redaction/dhp_july2017.tex
ajustement
test
changement d'un mot pour voir si j'arrive à puller
changement texte

@@
 From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect.
-Making use of our estimates for 2004 (see tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport.
+Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport.
 This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
@@
 For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
-Given our estimates for 2004 (see Table \ref{tab:result_air_3d_detail} in Appendix \ref{app:more_results}) such that $x = 0.028$ and $0.015$ in vessel and air transport respectively, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
-On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.
+Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
+On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.\smallskip
+
+Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, ie structural phenomenon.
+We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
+The corresponding values reported in detailed tables are available on the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
+Precisely, each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
+
+\begin{figure}[htbp]
+\caption{Share of additive costs (3-digit classification)}
+\label{fig:part_cout_additif}
+\begin{center}
+\includegraphics[width=12cm, height=7cm]{part_cout_additif.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+In both transport modes, the additive component appears of sizeable importance throughout the period.
+Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
+By contrast, they represent an sizeable and structural dimension of international transport costs.
+If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can yet notice that it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
 In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.
-
-\textbf{a ajuster et remonter figure annexe.}
-
-Two main comments can be made.
-First, in both transport modes, the additive component appears of sizeable importance all over the period.
-Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-By contrast, they represent an sizeable and structural dimension of international transport costs.
-However, one can also notice that, if the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport over the period, it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered (see Figure \ref{fig:part_cout_additif}, Appendix \ref{app:more_results}).\smallskip
 
@@
 This focus is shared with part of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
-Specifically, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, in the same spirit as \cite{hummels2007}.
+Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
 
@@
 
-As a first step, Figure \ref{fig:decomp_TC_3d} displays the \textbf{total transport costs (the total of the two, expressed in percentage of the fas price)}, by transport mode for each year between 1974 and 2013.
+As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the fas price,\footnote{Precisely, we use the yearly values of the estimated transport costs components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
+Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2013.
 
 \begin{figure}[htbp]
-\caption{Trend in Transport costs (Yearly mean value, 3 digits, in \% of the fas price, 1989 exclude for "Air")}
-\label{fig:decomp_TC_3d}
-\begin{center}
-\includegraphics[width=6.5in, height=5in]{Figure1_Trend_of_total_TC_by_mode.pdf}
+\caption{Trend in estimated transport costs (Yearly mean value)}
+\label{fig:Trends_in_TC}
+\begin{center}
+\includegraphics[width=14cm, height=8cm]{Figure1_Trend_of_totalTC_bymode.pdf}
 \end{center}
@@
 
-\textbf{figure a changer et texte a adapter}
-
-Figure \ref{fig:decomp_TC_3d} reports that both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+As reported in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
 \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased , independently of any change in transport costs \textit{per se}.
+Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
 As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
@@
 
-\paragraph{Challenging the $\beta$'s constancy assumption} Our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:part_cout_additif}, cast some doubt on the empirical relevance of this assumption.
-As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes.
+\paragraph{Challenging the $\beta$'s constancy assumption} As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes.
 In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
-
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
@@
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
 
@@
 
-\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the ``total'' transport cost based on Equation (\ref{eq:base_estimee}) according to:
+\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
 $$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
@@
 \subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{Figure \ref{fig:totalTC_compeffects_excl} reports the results, all types of goods included.
-In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} Panels (a), (b) and (c), we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
+Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.}\textbf{Et alors?
+Faire un commentaire et mettre l'annexe en ligne} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
 Panels (d), (e) and (f) report the results for Vessel.
@@
 This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
-Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
+Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
 Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport.
-For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)).
-Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for Air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
+For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
+Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the transport costs ``per se''.
 Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), where they amplify the reduction in ``pure'' transport costs (the plain line being lower than the dotted line).
-Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
 This is particularly the case for the multiplicative component (panel (d)).\smallskip
@@
 
-These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the ``per se'' or ``ceteris paribus'' transport costs for both air and maritime shipping.
-If anything, we find the opposite result here.
-This drives us to investigate further this difference.
+These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
+If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward trend pattern.
+This drives us to investigate this difference further.
 As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
@@
 %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below.
+This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
+
 To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in panels (c) and (f) from Figure \ref{fig:totalTC_compeffects_excl}, panels c and f.
+Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).
 
 Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in vessel transport.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in air transport.
 Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
 
-\subsection{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}
-
-As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``per se'' or ``ceteris paribus'' transport costs over time.
-In this section, we provide a robustness test to this difference in the weighting scheme implemented to extract the trade composition effects.\smallskip
+\subsection{Robustness to the weighting scheme \label{sec:robustness}}
+
+As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the pure transport costs over time.
+In this section, we provide a robustness test to this difference of treatment.\smallskip
 
@@
 Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
-In this section, our objective is to assess that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
+In this section, we check that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
 
@@
 Our results could be extended in two main ways.
-On the empirical side, one may want to go deeper in the ``structural'' determinants of trade costs, i.e.
+On the empirical side, one may want to go deeper in the structural determinants of (pure) transport costs, i.e.
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
@@
 
-\paragraph{More on additive costs} In Figure , we report the share of additive costs in total transport costs over time in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}, conditional on a year and a transport mode).
-Precisely, each point in time is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent (i.e., in percent of the export price $\widetilde{p}$).
-
-\begin{figure}[htbp]
-\caption{Share of additive costs}
-\label{fig:part_cout_additif}
-\begin{center}
-\includegraphics[width=10cm, height=7cm]{part_cout_additif.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
 \subsection{Assessing the importance of additive transport costs \label{app:diagnostic_test}}
@@
 The first indicator is through comparing $R^{2}$.
-However, its use is far from being straightforward when evaluating non-linear estimates.\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
+However, its use is far from being straightforward when evaluating non-linear estimates.
+$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
 In a non-linear context, $R^2$ is strictly speaking inappropriate.
-However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.} This drives us to complement the goodness of fit diagnosis with three alternative measures.
+However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.
+This drives us to complement the goodness of fit diagnosis with three alternative measures.
 We provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
@@
 A decrease in the log-likelihood function points to a better quality-of-fit.
-However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.} The preferred model is the one with the minimum AIC value.
+However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.
+Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.
+The preferred model is the one with the minimum AIC value.
 Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including the additive term, i.e.
@@
 \begin{center}
+  \footnotesize{
     \begin{tabular}{l|cccccc}
@@
 Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%) }  \\
+\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
 \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}  \\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}$, in \%) }   \\ \hline
+\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}  \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
 Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)} \\
+\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)} \\
 \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC} }\\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%) }   \\ \hline
+\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC} }\\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
 Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
@@
 \textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &      \\
-Model (A)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Model (B) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
+Model (a)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
+Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
 \textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
-Model (A)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Model (B) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
+Model (a)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
+Model (b) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
 \textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
-Model (B) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\
+Model (a) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
+Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\
 \textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
-Model (A)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
-Model (B) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\
+Model (a)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
+Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\
 \textbf{Test LL} &       &       &       &       &       &       \\
@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& ad-valorem
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs. Model (b) = with additive \& ad-valorem
 transport costs.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
 The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
@@
 \textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (B) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
+Model (a) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
+Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
 \textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
-Model (B) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\
+Model (a)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
+Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\
 \textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-Model (A) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
-Model (B) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\
+Model (a) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
+Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\
 \textbf{Test LL} &       &       & & &  & \\
@@
 
-In this section, we detail our methodology to extract the time trend in the ``ceteris paribus'' (or fitted) transport cost component, for each the ad-valorem and the additive component.
+In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for each the ad-valorem and the additive component.
 
@@
 However, this is a slight difference, as in both cases it ultimately amounts taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-
-A second minor difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}$ in Equation (\ref{eq:Hummels1})) in two separate components (origin country and product dimensions).
-As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
-However, we do not view this as the major cause of difference between our tow methods, as it can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
-
-Third, in contrast to us, \cite{hummels2007} does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effects.
+Second, in contrast to us, \cite{hummels2007} does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effects.
 This is not likely to make a substantial difference though, as these fixed effects are by nature constant over time.
 Accordingly, they only play as a scale effect as the estimate is in log.
+
+A third difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}$ in Equation (\ref{eq:Hummels1})) in two separate components (origin country and product dimensions).
+As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
+However, we do not view this as the major cause of difference between our two methods, \textbf{as it can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
+Ou on fait ca???}
+
 Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeh{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).

### 2018-06-22 redaction/dhp_july2017.tex
changement .tex

@@
 
-\date{February 2018}
+\date{April 2018}
  \maketitle
@@
 trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
-Trade costs declined faster for more developed countries \cite{arvis2016trade}.
+Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
 Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
@@
 
-This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others).
+This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
 Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
@@
 In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
-In this respect, our results suggest that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip
+In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip
 
@@
 
-Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''.
-Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs, suggesting that pure transport costs declined less than observed costs.
+Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, does come from the reduction of transport cost ``per se''.
+Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs.
+Otherwise stated, pure transport costs have declined less than observed costs.
 This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
+Our result thus deepens the paradox of the less-than-expected decline of transport costs.
 Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
 Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-Our result thus deepens the paradox of the less-than-expected decline of transport costs.
 \smallskip
@@
 
-Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others).
+Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
 The database we use to construct our measure of transport costs comes from US annual "Imports of Merchandise" provided by the Census bureau, spanning from 1974 to 2013.
 Details on the database are provided in Appendix \ref{app:data}.
-We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
+We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
 The fas price (for ``Free Alongside Ship'') means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
@@
 \paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+We also provide a robustness analysis to the separability assumption.
+Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries.
+See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}
 
@@
 
-After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.
+After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
+When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is(k)} = \widehat{\tau}_{i} \times \widehat{\tau}_{s(k)}$.
 In this case, notice that the equation could be estimated relying on a linear form.
@@
 These results constitute our first original contribution to the literature.
-In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip
+We also provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation (see results in In Appendix \ref{app:diagnostic_test}).\smallskip
 
@@
 
-\textbf{stop here}
-
 \subsection{Characterizing the time trends in transport costs}
@@
 
-\subsection{Robustness to the weighting scheme \label{sec:robustness}}
-
-As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the pure transport costs over time.
+\section{Robustness analysis \label{sec:robustness}}
+
+\subsection{Robustness to the separability assumption}
+In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, that considers (following so \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
+That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to rather model $t_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
+Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approximatively 200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
+This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold) and without.
+For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+
+Table reports a summary of the results.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Robustness to the separability assumption}
+\begin{center}
+    \begin{tabular}{lcccc}
+    \hline \hline
+    \multicolumn{5}{c}{Mean value over 1974-2013} \\
+ \multicolumn{5}{c}{3 digits} \\    \hline \hline
+   Mode  & \multicolumn{2}{c}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
+    Additive term ($t/\widetilde{p}$. in \%) &       &       &       &  \\
+    Mean  & \multicolumn{1}{c}{1.76} & \multicolumn{1}{c}{1.96} & \multicolumn{1}{c}{2.61} & \multicolumn{1}{c}{2.99} \\
+    Median & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{1.83} & \multicolumn{1}{c}{2.23} \\
+    Ad-valorem term (\$\textbackslash{}tau\$. in \textbackslash{}\%) & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+    Mean  & \multicolumn{1}{c}{1.04} & \multicolumn{1}{c}{0.94} & \multicolumn{1}{c}{2.43} & \multicolumn{1}{c}{2.40} \\
+    Median & \multicolumn{1}{c}{0.56} & \multicolumn{1}{c}{0.83} & \multicolumn{1}{c}{2.06} & \multicolumn{1}{c}{2.04} \\
+    Diagnostic tests & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+    $R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
+    SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
+    AIC criteria & \multicolumn{1}{c}{4657.60} & \multicolumn{1}{c}{4654.23} & \multicolumn{1}{c}{10744.83} & \multicolumn{1}{c}{10690.80} \\
+    Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
+\# observations & \multicolumn{1}{c}{2381} & \multicolumn{1}{c}{2381} & \multicolumn{1}{c}{2798} & \multicolumn{1}{c}{2798} \\
+\# origin country & \multicolumn{1}{c}{13.4} & \multicolumn{1}{c}{13.4} & \multicolumn{1}{c}{19.8} & \multicolumn{1}{c}{19.8} \\
+\# products & \multicolumn{1}{c}{25.5} & \multicolumn{1}{c}{25.5} & \multicolumn{1}{c}{53} & \multicolumn{1}{c}{53} 
+    \end{tabular}%
+\end{center}
+\label{tab:robustness_separability}
+\end{table}
+
+\subsection{Robustness to the weighting scheme }
+
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and Hummels's (2007)dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
 In this section, we provide a robustness test to this difference of treatment.\smallskip
@@
 
-%\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
+%\todo{Changer\begin{table}[htbp]
 \begin{table}[htbp]
-  \centering
   \caption{Air: Transport costs estimates, 3-digits (selected years)}

### 2018-06-22 redaction/dhp_july2017.tex
changement tex

@@
 \begin{center}
-    \begin{tabular}{lcccc}
+    \begin{tabular}{lcc|cc}
     \hline \hline
@@
    Mode  & \multicolumn{2}{c}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
+   Separability & No & Yes & No & Yes \\
     Additive term ($t/\widetilde{p}$. in \%) &       &       &       &  \\

### 2018-06-25 redaction/dhp_july2017.tex
changement texte

@@
 
-\date{April 2018}
+\date{June 2018}
  \maketitle
@@
 Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs.
-\citet{Lashkaripour-mimeo-2017} challenges this view.
-Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result.
-While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the twentieth century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports.
-This result is also confirmed on more aggregate US, sector-level trade data.
-The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+Recent empirical studies based on micro-level data also provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
+Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
+Using firm-level trade data \textbf{from which country?}, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
@@
 \begin{center}
-\includegraphics[width=12cm, height=7cm]{part_cout_additif.pdf}
+\includegraphics[width=8cm, height=5cm]{part_cout_additif.pdf}
 %{\footnotesize {OECD data} }
@@
 Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-By contrast, they represent an sizeable and structural dimension of international transport costs.
-If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can yet notice that it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
-In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.
+By contrast, they represent a sizeable and structural dimension of international transport costs.
+If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can further notice that it substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
+In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs time trends.
 
@@
 \begin{center}
-\includegraphics[width=14cm, height=8cm]{Figure1_Trend_of_totalTC_bymode.pdf}
+\includegraphics[width=14cm, height=6.5cm]{Figure1_Trend_of_totalTC_bymode.pdf}
 \end{center}
@@
 \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
-In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974.
+In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock related transport costs in 1974.
 However, computing the time trends from 1980 does not dramatically change the picture.
@@
 \paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
-He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
+He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per-kg shipped (for a given product $k$ imported from country $i$ in year $t$).
 Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes down as:
@@
 \subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.}\textbf{Et alors?
-Faire un commentaire et mettre l'annexe en ligne} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
+Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
 Panels (d), (e) and (f) report the results for Vessel.
@@
 \begin{center}
-\includegraphics[height=5in]
+\includegraphics[height=4in]
 {graph_composition_all.pdf}
@@
 Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
+First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
 International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
@@
 composition effects excluded ($\widehat{TC}_t$).
+In Panels (a) (for Air) and (b) (for Vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for Air and Vessel respectively.
 
@@
 {\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels.pdf} \\
-\includegraphics[width=3in, height=2.5in]{figure5_comme_hummels_base100.pdf}
-& \includegraphics[width=3in,height=2.5in]{figure6_comme_hummels_base100.pdf} \\
-
+\multicolumn{2}{c}{{\small In percent of the export price}} \\
+\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.pdf}
+& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.pdf} \\
+{\small (c) Ad-valorem Air Freight} & {\small (d) Ad-valorem Vessel Freight}\\
+\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
+\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.pdf}
+& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.pdf} \\
 \end{tabular}
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).
+Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for Vessel (Panel (d)).
+Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).\medskip
 
@@
 Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approximatively 200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
-This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold) and without.
+This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.
 For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
 
-Table reports a summary of the results.
+Table \ref{tab:robustness_separability} reports a summary of the results, considering the average estimated values over the period (by transport mode).
 
@@
   \centering
-  \caption{Robustness to the separability assumption}
+  \caption{Robustness to the separability assumption (Average over the period)}
 \begin{center}
@@
  \multicolumn{5}{c}{3 digits} \\    \hline \hline
-   Mode  & \multicolumn{2}{c}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
-   Separability & No & Yes & No & Yes \\
-    Additive term ($t/\widetilde{p}$. in \%) &       &       &       &  \\
-    Mean  & \multicolumn{1}{c}{1.76} & \multicolumn{1}{c}{1.96} & \multicolumn{1}{c}{2.61} & \multicolumn{1}{c}{2.99} \\
-    Median & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{1.83} & \multicolumn{1}{c}{2.23} \\
-    Ad-valorem term (\$\textbackslash{}tau\$. in \textbackslash{}\%) & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-    Mean  & \multicolumn{1}{c}{1.04} & \multicolumn{1}{c}{0.94} & \multicolumn{1}{c}{2.43} & \multicolumn{1}{c}{2.40} \\
-    Median & \multicolumn{1}{c}{0.56} & \multicolumn{1}{c}{0.83} & \multicolumn{1}{c}{2.06} & \multicolumn{1}{c}{2.04} \\
-    Diagnostic tests & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-    $R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
-    SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
-    AIC criteria & \multicolumn{1}{c}{4657.60} & \multicolumn{1}{c}{4654.23} & \multicolumn{1}{c}{10744.83} & \multicolumn{1}{c}{10690.80} \\
-    Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
-\# observations & \multicolumn{1}{c}{2381} & \multicolumn{1}{c}{2381} & \multicolumn{1}{c}{2798} & \multicolumn{1}{c}{2798} \\
-\# origin country & \multicolumn{1}{c}{13.4} & \multicolumn{1}{c}{13.4} & \multicolumn{1}{c}{19.8} & \multicolumn{1}{c}{19.8} \\
-\# products & \multicolumn{1}{c}{25.5} & \multicolumn{1}{c}{25.5} & \multicolumn{1}{c}{53} & \multicolumn{1}{c}{53}
+   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
+   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
+    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$. in \%)}  \\
+    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
+    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
+    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
+    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
+    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
+\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
+    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
+    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
+ \textit{Diagnostic test} & & &  &  \\
+    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
+    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
+    AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
+    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
+\# observations & 2381 & 2381 & 2798 & 2798 \\
+\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
+\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
     \end{tabular}%
@@
 
+Two main comments can be made.
+First, both models provide a similar quality of fit of the regression (once we take into account the number of explicative variables) as the AIC criteria displays very close values in both models (conditional on transport mode).
+Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs.
+However, the estimated values remain of similar magnitude under both separability and non-separability assumptions, when averaged over the period.
+Before jumping to the conclusion that our results are robust to the separability assumption though, we investigate this comparison deeper on a year-to-year basis, as reported in Figure \ref{fig:robustesse_non_separe}.
+Precisely, Figure \ref{fig:robustesse_non_separe} displays the results for Air transport in panel (a) and for maritime transport in panel (b).
+In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability, for the additive term (upper panel) and the ad-valorem term (lower panel).
+
+\begin{figure}[htbp]
+\caption{Robustness to the separability assumption (Year-to-year basis)}
+\label{fig:robustesse_non_separe}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Air Freight} & {\small (b) Vessel Freight}\\
+\includegraphics[width=2.5in, height=2in]{graph_robustesse_ns_mp_terme_A_air.pdf}
+& \includegraphics[width=2.5in,height=2in]{graph_robustesse_ns_mp_terme_A_ves.pdf} \\
+\includegraphics[width=2.5in, height=2in]{graph_robustesse_ns_mp_terme_I_air.pdf}
+& \includegraphics[width=2.5in,height=2in]{graph_robustesse_ns_mp_terme_I_ves.pdf} \\
+
+\end{tabular}
+\end{center}
+\end{figure}
+
+In line with Table \ref{tab:robustness_separability}, the results reported in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
+However, they also show that the difference is quantitatively small; further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar to the one another, whether they are estimated under the separability assumption or not.
+Along with the quality of fit diagnosis, this confirms the robustness of our estimation results to this assumption.
+
+As last check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
+To this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
+Figure \ref{fig:histogram_beta_robustness} reports the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for Air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a, (c))) and the non-separability scenario (Panels (b), (c)).
+
+\begin{figure}[htbp]
+\caption{Robustness: Histogram of the share of additive costs in total costs}
+\label{fig:histogram_beta_robustness}
+\begin{center}
+\begin{tabular}{cc}
+\multicolumn{2}{c}{Air transport} \\
+{\small (a) Separability (baseline)} & {\small (b) Non-separability}\\
+\includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_air.pdf}
+& \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_air.pdf} \\
+\multicolumn{2}{c}{Maritime transport} \\
+{\small (c) Separability (baseline)} & {\small (d) Non-separability}\\
+\includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_ves.pdf}
+& \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_ves.pdf} \\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+Similarly as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distibution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, the countries of origin and products.
+As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, that establish that the decline of ``pure'' transport costs lies as main reason for the decrease of overall transport costs observed over time.
+
 \subsection{Robustness to the weighting scheme }
 
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and Hummels's (2007)dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and Hummels's (2007) dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
 In this section, we provide a robustness test to this difference of treatment.\smallskip
@@
 This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
+Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
 Our results may be summarized in two main findings.
 First, we provide a quantitative measure of both the additive and the ad-valorem transport cost.
-We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%, as mean values over 1974-2013.
-These values are respectively equal to 1.8 and 2.5\% for air transport.
+We find that additive costs are quantitativeley sizeable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
+This is slighly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
+Further, modeling the additive component significantly improves the quality of fit of the regression.
 Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
+Allowing for a varying share of additive costs in product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
 If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs, in particular in maritime transport.
@@
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles.
+On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows in an international trade theory perspective and in affecting the international transmission of business cycles.
 This is left for further research.
@@
     Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
-    \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
+    \multicolumn{8}{l}{\textbf{$R^2$} }\\ \hline
     Model (A)& 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
@@
 Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
-\multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
+\multicolumn{8}{l}{\textbf{$R^2$} }\\ \hline
 Model (A)& \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & 0.35 & \multicolumn{1}{c}{0.34} & 0.39 \\
@@
 As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
-However, we do not view this as the major cause of difference between our two methods, \textbf{as it can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
-Ou on fait ca???}
+However, we do not view this as the major cause of difference between our two methods, as it can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
 

### 2018-07-02 redaction/dhp_travail.tex
Guillaume sur l’intro

--- 2018-06-25 redaction/dhp_july2017.tex
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
-Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
+Universit\'{e} Paris-Dauphine, Université PSL, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
 Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
 ; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
-
-%\date{July 2017\\PRELIMINARY AND INCOMPLETE \\ DO NOT QUOTE WITHOUT PERMISSION}
+Universit\'{e} Paris-Dauphine, Université PSL, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
 
@@
 \begin{abstract}
-This paper investigates the role of additive costs in international transport costs, using information contained in the US imports flows database over 1974-2013.
-We find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs.
+This paper investigates the role of additive costs in international transport costs, using information contained in a US imports flows database over 1974-2013.
+We find that the additive component of transport costs is sizable, representing between 40 and 50\% of overall transport costs.
 Further, in the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs.
-In contrast to him, we find that trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+In contrast to him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer.
-In both aspects, our results point the importance of the additive component in accounting for international transport costs.
+In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
 \end{abstract}
@@
 Trade costs are central to international economic analysis.
-In particular, they are considered a major obstacle to international economic integration and international trade flows.
+In particular, they are known to be a major obstacle to international economic integration and international trade flows.
 Based on panel data, \citet{novy13} finds  that U.S.
@@
 Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
-Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
-
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
+Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
+
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
 In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.
-Equivalently, international transportation costs would represent a 21\% markup over production costs.
-Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+Equivalently, international transportation costs represent a 21\% markup over production costs.
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
 If much of trade policy barriers have been removed over the second half of the twentieth century, this suggests that the transport costs component remains large and deserves attention.
@@
 Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
-This view does not make consensus though.
+There is no consensus thought.
 Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
-In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip
+In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined more slowly than the common wisdom assumes.\smallskip
 
@@
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
-This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
-To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
+We obtain this by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
+To do so, we exploit information contained in the Census Bureau's database of US imports flows over 1974-2013.
+This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
 This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013).
@@
 Our results may be summarized in two main new findings.
-First, additive transport costs are quantitatively sizeable.
+First, additive transport costs are quantitatively sizable.
 On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
@@
 Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode.
-To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
+\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?} To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
 This represents a valuable insight for calibrating related international trade or business cycle models.
@@
 
-Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, does come from the reduction of transport cost ``per se''.
+Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost ``per se''.
 Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs.
@@
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
-As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs.
+As shown by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs.
 In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
@@
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using firm-level trade data \textbf{from which country?}, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+Using firm-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
@@
 The paper is built as follows.
-First, we estimate the values of international transport costs annually throughout the period 1974-2013 (and by transport mode), explicitly distinguishing between the additive and the ad-valorem components.
-This is made in Section \ref{sec:data_method}, which reports the mean values of the transport costs estimated over the period (by transport mode).
+First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
+Section \ref{sec:data_method} thus reports the mean values of the transport costs estimated over the period (by transport mode).
 Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''.
 Section \ref{sec:results_trends} is devoted to this analysis.
+Secion \ref{sec:robustness} checks its robustness to the separability assumption and the weighting scheme.
 Section \ref{sec:conclu} concludes.
@@
 In this respect, this dataset embraces only a partial view of international transport barriers.
-However, it is also clear that direct transport costs do represent a sizeable share of trade costs.
+However, it is also clear that direct transport costs do represent a sizable share of trade costs.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...), that amount to a  9\% tax equivalent.
@@
 The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
-This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter.
+This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter.
 These results constitute our first original contribution to the literature.
@@
 Put differently, the additive component amounts to 48\% of total transport costs in Vessel, and 42\% in Air.
-This shows that the per-kg cost dimension is quantitatively sizeable.
+This shows that the per-kg cost dimension is quantitatively sizable.
 For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.
@@
 Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
-On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.\smallskip
+On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
 
@@
 
-In both transport modes, the additive component appears of sizeable importance throughout the period.
+In both transport modes, the additive component appears of sizable importance throughout the period.
 Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-By contrast, they represent a sizeable and structural dimension of international transport costs.
+By contrast, they represent a sizable and structural dimension of international transport costs.
 If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can further notice that it substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
@@
 First, we provide a quantitative measure of both the additive and the ad-valorem transport cost.
-We find that additive costs are quantitativeley sizeable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
+We find that additive costs are quantitativeley sizable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
 This is slighly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.

### 2018-07-03 redaction/dhp_travail.tex
Début travail Guillaume 2e partie

@@
 
-\section{Estimating international transports\label{sec:data_method}}
-
-\subsection{Data sources}
+\section{Estimating international transport costs\label{sec:data_method}}
+
+\subsection{Data}
 
@@
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the fas price is the one reported by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
+While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
 The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
@@
 
-Obviously, this dataset is not immune of limitations.
-First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
+Obviously, this dataset has limitations.
+First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
 It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
-Second, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit.
+Second, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
 In this respect, this dataset embraces only a partial view of international transport barriers.
 However, it is also clear that direct transport costs do represent a sizable share of trade costs.
-According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...), that amount to a  9\% tax equivalent.
+According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
 Further, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
@@
 Exploiting this dataset also has (at least) three advantages.
-First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source.
+First, it comes from a single, homogenous and trustworthy customs source.
 This limits the measurement error bias.
-Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
+Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
 This will be crucial to compute transport costs, as we detail below.

### 2018-07-03 redaction/dhp_travail.tex
J’ai mis des sauts de ligne partout…
… pour mieux isoler les changements sous Git

@@
 \title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
+
 Finally, any remaining errors are ours.
@@
 \end{equation*}
-\noindent where $\epsilon_{ik}$ follows a normal law centered on 0. Considered in logs, the above equation becomes:
+\noindent where $\epsilon_{ik}$ follows a normal law centered on 0.
+Considered in logs, the above equation becomes:
 
@@
     %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
-    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
-    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
+    %Min.
+value & 1.003 & 1.001 & 1.003 & 1.000 \\
+    %Max.
+value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
     \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
@@
     %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
-    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
-    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
+    %Min.
+value & 1.001 & 1.000 & 1.000 & 1.000 \\
+    %Max.
+value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
     \textit{Additive term }& & & &   \\ \hline
@@
     %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
-    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
-    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+    %Min.
+value & 0.000 & 0.000 & 0.000 & 0.000 \\
+    %Max.
+value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
     \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
@@
     Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
-    \# obs. & 29279 & 28207 & 29317 & 27680 \\
+    \# obs.
+& 29279 & 28207 & 29317 & 27680 \\
     \# origin country & 188 & 191 & 188 & 189 \\
@@
 
-We start describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series. \smallskip
+We start describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
+\smallskip
 
@@
    Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
-    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$. in \%)}  \\
+    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
+in \%)}  \\
     Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs. Model (b) = with additive \& ad-valorem
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
+Model (b) = with additive \& ad-valorem
 transport costs.
@@
 
-\paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component. Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
+\paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component.
+Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
 
@@
 
-From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$. Given the constraint $t>0$, we then obtain the percentage change from 1974 from:
+From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
+Given the constraint $t>0$, we then obtain the percentage change from 1974 from:
 
@@
 
-What are the main differences with our estimation strategy?
-Five points may be underlined.
+What are the main differences with our estimation strategy? Five points may be underlined.
 First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
@@
 \end{eqnarray*}
-\noindent with $t_{ikt}$ appropriately defined. This exactly corresponds to the case of additive costs only.
+\noindent with $t_{ikt}$ appropriately defined.
+This exactly corresponds to the case of additive costs only.
 
@@
 
-\noindent with $\tau_{ikt}$ appropriately defined. This exactly corresponds to the case of ad-valorem transport costs only.
+\noindent with $\tau_{ikt}$ appropriately defined.
+This exactly corresponds to the case of ad-valorem transport costs only.
 

### 2018-07-03 redaction/dhp_travail.tex
Guillaume continue sa lecture

@@
 \title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
-
 Finally, any remaining errors are ours.
@@
 Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
-This will be crucial to compute transport costs, as we detail below.
-In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air.
+%This will be crucial to compute transport costs, as we detail below.
+In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
 We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode.
-Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country).
+Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
 This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to \cite{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013.
-Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
+Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.\smallskip
-
+In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip
+
+As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
 We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level.
-As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
 Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
@@
 
-\paragraph{The estimated equation} In this section, our purpose is to provide time-varying estimates of the size of ad-valorem and additive costs among total transport costs.
-To do so, we start from the equation that expresses the price $p$ of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}$ (export, or fas price), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:
+\paragraph{The estimated equation} In this section, our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs costs.
+To do so, we start from the equation that expresses the price $p$ (per kg) of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or fas price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
@@
 
-\noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
+\noindent As is usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
  Let us denote $i$ the origin country, and $k$, the product at the 5-digit level.
-Transforming the above equation (\ref{eq:base}) as ratio in order to get the percentage change in prices induced by transportation, we thus get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
-%\footnote{\textbf{Why do we consider the ratio not the difference, with cif price on the LHS and p fas RHS? To be answered?}}
+Transforming the above equation (\ref{eq:base}) into a ratio in order to get the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
 
@@
 
-The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators.
+Equation (\ref{eq:equation0}) is non linear.
+It cannot be estimated using standard linear estimators.
 All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
@@
 Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
-In other words, this amounts making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
+This amounts making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
 same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
@@
 Je crois que ce qu'on veut dire, c'est que le fait que ptilde change en fonction des TC ne change pas les TC, le cif price s'ajuste proportionnellement?
-Si c'est ca, c'est vrai seulement si les couts sont advalorem...
-Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus?
+Si c'est ca, c'est vrai seulement si les couts sont advalorem...Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus?
 Par produit 3d/pays, par secteur 3d tous pays, etc...?}\smallskip
@@
   \centering
-  \caption{Robustness to the separability assumption (Average over the period)}
+  \caption{Robustness to the separability assumption (Average over the period) - restricted sample}
 \begin{center}

### 2018-07-03 redaction/dhp_travail.tex
Premieres_modifs_JH
Modifs Intro, abstract, biblio

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
@@
 This paper investigates the role of additive costs in international transport costs, using information contained in a US imports flows database over 1974-2013.
-We find that the additive component of transport costs is sizable, representing between 40 and 50\% of overall transport costs.
+We find that the additive component of transport costs is sizable, representing between one third and one half of overall transport costs.
 Further, in the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs.
-In contrast to him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+In contrast to \cite{hummels2007}, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer.
@@
 Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
-Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
+Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
 Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
@@
 According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries).
-Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result.
+Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
@@
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015} or \citealp{martin2012}).
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}).
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
@@
 This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period).
-Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode.
-\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?} To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
+Put it differently, additive costs represent between one third and almost one half of the overall transport costs.
+%\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?} 
+To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
 This represents a valuable insight for calibrating related international trade or business cycle models.
@@
 Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost ``per se''.
-Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs.
+Trade composition effects only have a limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
 Otherwise stated, pure transport costs have declined less than observed costs.
@@
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using firm-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+Using transaction-level trade data from Colombia, \cite{Lashkaripour-2017b} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
@@
 While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
-One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}).
+One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
 One contribution of the paper is to fill this gap.}
@@
 Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring relative additive trade costs from firm-level trade data.
-Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (expressed in Norwegian crowns).
+Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
 Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
@@
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
 
@@
 Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''.
-Section \ref{sec:results_trends} is devoted to this analysis.
-Secion \ref{sec:robustness} checks its robustness to the separability assumption and the weighting scheme.
+Section \ref{sec:results_trends} is devoted to this analysis, while section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
 Section \ref{sec:conclu} concludes.

### 2018-07-04 redaction/dhp_travail.tex
Fin_corrections_JH
fin modifs relecture JH. Prêt à envoi à V Vicard après validation Lise et Guillaume

@@
 \usepackage{bbm}
+\usepackage[flushleft]{threeparttable}
 %\usepackage{bbold}
@@
 On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
-This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period).
+This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
 Put it differently, additive costs represent between one third and almost one half of the overall transport costs.
@@
 First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
-It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.
-Second, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
+It is thus silent about the others dimensions of international trade costs.
+Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
 In this respect, this dataset embraces only a partial view of international transport barriers.
-However, it is also clear that direct transport costs do represent a sizable share of trade costs.
+However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
@@
 
-Exploiting this dataset also has (at least) three advantages.
+Exploiting this dataset also has three advantages.
 First, it comes from a single, homogenous and trustworthy customs source.
@@
 Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to \cite{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e.
+This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g.
+\cite{Irrazabal_2015}.
+Third, this dataset is available over a long time span, which we exploit for providing an analysis of the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
@@
 
-\paragraph{The estimated equation} In this section, our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs costs.
+\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs costs.
 To do so, we start from the equation that expresses the price $p$ (per kg) of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or fas price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
@@
  Let us denote $i$ the origin country, and $k$, the product at the 5-digit level.
-Transforming the above equation (\ref{eq:base}) into a ratio in order to get the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification at the root of our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
+Transforming the above equation (\ref{eq:base}) into a ratio, and removing 1 on both sides, in order to get the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
 
@@
 \paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
+Confronted to this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
 This amounts making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
@@
 For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5 percent of the upper and lower tails of the distribution in the regression variables.
+%JH: TO KEEP IN MIND: one day, we could be asked to winsorize tails, instead of cutting them...
 These cut-offs are aimed at eliminating reporting or coding errors.
@@
 The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
-These results are reported in the Online Appendix, available on the authors' webpages.}
-
-One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
+More details of these results in the Online Appendix.}
+
+One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
@@
 
-\subsection{Estimating results: The importance of the additive component \label{sec:results_decomposition}}
-
-Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
-The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+\subsection{The importance of the additive component \label{sec:results_decomposition}}
+
+%Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
+\paragraph{Quantification} The first step of the analysis consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
 This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter.
 These results constitute our first original contribution to the literature.
-We also provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation (see results in In Appendix \ref{app:diagnostic_test}).\smallskip
+We also provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation (see results in Appendix \ref{app:diagnostic_test}).\smallskip
 
@@
 Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
-Descriptive statistics showing that that both observed and predicted values do match in logarithms are available upon request to the authors.\medskip
-
-Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}.
+Descriptive statistics showing that both observed and predicted values do match in logarithms are available upon request to the authors.%\medskip
+
+\paragraph{Interpretation} Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}.
 First, it provides estimation results about the size of the overall transport costs.
@@
 Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests, that confirm the importance of the additive cost component.
-The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom.
-An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip
+The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
+An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.%\smallskip
 
@@
 For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
-Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport.
+Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 47\% (in Vessel) of additive trade costs are attributable to international transport.
 On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
 
-Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, ie structural phenomenon.
+\paragraph{Time dynamics} Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, i.e. structural phenomenon.
 We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
@@
 In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
-In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.
-
-\paragraph{Challenging the $\beta$'s constancy assumption} As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes.
-In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
+Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
+
+\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
+%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
@@
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
 
@@
 
-\paragraph{Obtaining the unfitted measures} The objective is to get the unfitted transport cost component (additive and multiplicative) built as as index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
+\paragraph{Obtaining the unfitted measures} The objective is to get the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
 To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
@@
 Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the transport costs ``per se''.
-Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), where they amplify the reduction in ``pure'' transport costs (the plain line being lower than the dotted line).
+Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
+Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: both trends are very similar.
 Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
@@
 
-These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
-If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward trend pattern.
+The last results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
+If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
 This drives us to investigate this difference further.
@@
 Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
-As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
+Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
 %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
@@
 
-Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs.
+Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
 In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in air transport.
@@
 For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-
 Table \ref{tab:robustness_separability} reports a summary of the results, considering the average estimated values over the period (by transport mode).
@@
 \end{center}
+\parbox[l]{10cm}{\tiny{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+The additive term is expressed in fraction of fas price.
+($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
 \label{tab:robustness_separability}
@@
 
-Two main comments can be made.
-First, both models provide a similar quality of fit of the regression (once we take into account the number of explicative variables) as the AIC criteria displays very close values in both models (conditional on transport mode).
-Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs.
-However, the estimated values remain of similar magnitude under both separability and non-separability assumptions, when averaged over the period.
+Three main comments can be made.
+First, the subset of big trade flows we are considering (around 8\% of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52 and 67\% of total costs on average, versus 40 to 50\% on the whole sample.
+%JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
+Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand side variables) as the AIC criteria displays very close values in both models (conditional on transport mode).
+Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period - the picture is reversed for medians.
 Before jumping to the conclusion that our results are robust to the separability assumption though, we investigate this comparison deeper on a year-to-year basis, as reported in Figure \ref{fig:robustesse_non_separe}.
@@
 
-As last check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
+As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
 To this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
@@
 \begin{figure}[htbp]
-\caption{Robustness: Histogram of the share of additive costs in total costs}
+\caption{Histogram of the share of additive costs in total costs (restricted sample)}
 \label{fig:histogram_beta_robustness}
@@
 & \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_ves.pdf} \\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
 \end{tabular}
@@
 
-Similarly as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distibution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
-Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, the countries of origin and products.
+Similarly as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
 As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, that establish that the decline of ``pure'' transport costs lies as main reason for the decrease of overall transport costs observed over time.
@@
 
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and Hummels's (2007) dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and \citet{hummels2007}'s dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
 In this section, we provide a robustness test to this difference of treatment.\smallskip
@@
 
-The comparison of Figure \ref{fig:compeffects_robustness} (applying \citet{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) drives two comments.
+The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) drives two comments.
 First, differences do show up, in particular for the multiplicative component of Air transport.
@@
 Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
-Our results may be summarized in two main findings.
+Two main findings emerge from our results.
 First, we provide a quantitative measure of both the additive and the ad-valorem transport cost.
-We find that additive costs are quantitativeley sizable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
-This is slighly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
+We find that additive costs are quantitatively sizable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
+This is slightly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
 Further, modeling the additive component significantly improves the quality of fit of the regression.
@@
 Allowing for a varying share of additive costs in product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
-If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs, in particular in maritime transport.
+If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs in maritime transport.
 In both aspects, our results point the importance of the additive component in accounting for international transport costs.
@@
 Table \ref{tab:result_air_3d_detail} reports the results for Air transport.
-The results for Ocean transport are displayed in Table \ref{tab:result_ves_3d_detail}.
+The results for Vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.
 
@@
 However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.
-Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.
+\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.}
 The preferred model is the one with the minimum AIC value.
@@
 
-$$\underbrace{\sum_{k=1}^K \sum_{i=1}^I \left(x_{ij} - \bar{x}_g  \right)^2}_{\text{Total variability}} = \underbrace{\sum_k \sum_i \left(x_{ik} - \bar{x}_k  \right)^2}_{\text{Within-product variability}} + \underbrace{\sum_k \left(\bar{x}_{k} - \bar{x}_g  \right)^2}_{\text{Between-product variability}}$$
+$$\underbrace{\sum_{k=1}^K \sum_{i=1}^I \left(x_{ik} - \bar{x}_g  \right)^2}_{\text{Total variability}} = \underbrace{\sum_k \sum_i \left(x_{ik} - \bar{x}_k  \right)^2}_{\text{Within-product variability}} + \underbrace{\sum_k \left(\bar{x}_{k} - \bar{x}_g  \right)^2}_{\text{Between-product variability}}$$
 
@@
 This suggests that what primarily matters in international transport costs is mostly attributable to the product \textit{per se}, rather than to the country where it comes from.
-By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
+%JH: un peu too much, non ? "By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs."
 

### 2018-07-05 redaction/online_appendix_travail.tex

--- 2017-09-13 redaction/online_appendix_dhp2017.tex
@@
   \label{tab:3models_estimation_results_air}%
-    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \begin{tabular}{l|c c c c c c c c c}
     \hline \hline
@@
   \label{tab:3models_estimation_results_vessel}%
-    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \begin{tabular}{l|c c c c c c c c c}
     \hline \hline
@@
     \label{tab:3models_diagnosis_vessel}%
-   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+   \begin{tabular}{l|c c c c c c c c c}
     \hline \hline
@@
 One can yet mention that this does not make much difference.
-When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component ($\widehat{\tau}^{adv} = 1.9\%$, vs 1.8\% when 1989 is excluded.
-The weighed median value is left unchanged, as the additive component (with weighed mean and median values of $\widehat{t}/\widetilde{p} = 2.5\%$ and 1.8\%, whether 1989 is included or not.
+When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component ($\widehat{\tau}^{adv} = 1.9\%$), vs 1.8\% when 1989 is excluded.
+The weighed median value is left unchanged, as the additive component (with weighed mean and median values of $\widehat{t}/\widetilde{p} = 2.5\%$ and 1.8\%), whether 1989 is included or not.
 
@@
 
-As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, the ``raw'' transport costs are (or the unfitted transport costs, in plain blue line) have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in Air, 60\% in Vessel).
+As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, the ``raw'' transport costs are (or the unfitted transport costs, in the red dashed line plain) have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in Air, 60\% in Vessel).
 However, contrasting these two broad sectors conveys a contrasted message about the role of trade composition effects in accounting for this trend pattern.
 Comparing Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
-In Air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the dotted red line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
+In Air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
 Trade composition effects matter more in maritime transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)), primarily in the ad-valorem component.

### 2018-07-05 redaction/dhp_travail.tex
suppression_lignes_value_Table1
suppression des lignes value apparues dans Table1 (avec un %)
changement texte

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine, Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine, Université PSL, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
+Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
 Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
 ; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-Universit\'{e} Paris-Dauphine, Université PSL, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
+Universit\'{e} Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
 
@@
 \begin{abstract}
-This paper investigates the role of additive costs in international transport costs, using information contained in a US imports flows database over 1974-2013.
-We find that the additive component of transport costs is sizable, representing between one third and one half of overall transport costs.
-Further, in the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs.
-In contrast to \cite{hummels2007}, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
-Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer.
+This paper proposes a detailed investigation of the pattern of international transport costs over time, using information contained the US imports flows over 1974-2013.
+First, we document the importance of the per-unit (additive) component of transport costs.
+We thus find that additive costs are quantitatively sizeable, representing between one third and one half of overall transport costs.
+Second, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
+Contrasting his results, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+Importantly, this difference of results can be attributed to the new way of modeling the per-unit component of transport costs we offer.
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
@@
 First, additive transport costs are quantitatively sizable.
-On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively.
+On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively (at the 3 digit classification level).
 This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
@@
 Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.
+Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip
 
@@
     %Min.
-value & 1.003 & 1.001 & 1.003 & 1.000 \\
+%value & 1.003 & 1.001 & 1.003 & 1.000 \\
     %Max.
-value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
+%value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
     \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
@@
     %Min.
-value & 1.001 & 1.000 & 1.000 & 1.000 \\
+%value & 1.001 & 1.000 & 1.000 & 1.000 \\
     %Max.
-value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
+%value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
     \textit{Additive term }& & & &   \\ \hline
@@
     %Min.
-value & 0.000 & 0.000 & 0.000 & 0.000 \\
+%value & 0.000 & 0.000 & 0.000 & 0.000 \\
     %Max.
-value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+%value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
     \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
@@
 Three main comments can be made.
-First, the subset of big trade flows we are considering (around 8\% of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52 and 67\% of total costs on average, versus 40 to 50\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52 and 67\% of total costs on average, versus 40 to 50\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
@@
 Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period - the picture is reversed for medians.
+
 Before jumping to the conclusion that our results are robust to the separability assumption though, we investigate this comparison deeper on a year-to-year basis, as reported in Figure \ref{fig:robustesse_non_separe}.
 Precisely, Figure \ref{fig:robustesse_non_separe} displays the results for Air transport in panel (a) and for maritime transport in panel (b).
-In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability, for the additive term (upper panel) and the ad-valorem term (lower panel).
+In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
 
@@
 In line with Table \ref{tab:robustness_separability}, the results reported in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
-However, they also show that the difference is quantitatively small; further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar to the one another, whether they are estimated under the separability assumption or not.
+However, they also show that the difference is quantitatively small.
+Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar to the one another, whether they are estimated under the separability assumption or not.
 Along with the quality of fit diagnosis, this confirms the robustness of our estimation results to this assumption.

### 2018-07-05 redaction/dhp_travail.tex
changement abstract

@@
 \begin{abstract}
-This paper proposes a detailed investigation of the pattern of international transport costs over time, using information contained the US imports flows over 1974-2013.
+This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows over 1974-2013.
 First, we document the importance of the per-unit (additive) component of transport costs.
-We thus find that additive costs are quantitatively sizeable, representing between one third and one half of overall transport costs.
+We thus find that additive costs are quantitatively sizable, representing between one third and one half of overall transport costs.
 Second, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
-Contrasting his results, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
 Importantly, this difference of results can be attributed to the new way of modeling the per-unit component of transport costs we offer.

### 2018-08-05 redaction/dhp_travail.tex
Tout petit bug
corrections_SJ
Minor formal corrections after SJ

@@
 We thus find that additive costs are quantitatively sizable, representing between one third and one half of overall transport costs.
-Second, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
+Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
 Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
@@
 
-Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring relative additive trade costs from firm-level trade data.
+Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
 Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
@@
 First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost ; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
-From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect.
+From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
 While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
@@
 %However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.
+%By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.
 

### 2018-09-12 redaction/Presentations/Seminaire_Kiel_Sept2018/slides_Kiel_sept_2018.tex
Présentation KIel
Slides actualisés avec derniers résultats pour la présentation à Kiel

--- 2017-12-11 redaction/Presentations/SeminaireBETA_Nancy_Oct2017/slides_BETA_v1.tex
@@
 \documentclass[10 pt,Helvetica, french]{beamer}
+%\documentclass[10 pt,Helvetica, french, handout]{beamer}
 % option compress pour réduire la taille des bars
@@
 
-\date[October 2017]{BETA Seminar, University of Nancy, October 2017}
-\title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
+\date[September 2018]{Lunch seminar in Interational Economics, Kiel \\ September 7, 2018}
+\title[New Findings from modeling additive costs]{International Transport costs:\\New Findings from modeling additive costs}
 \author[Daudin et al.]{Guillaume Daudin\\
@@
 J\'{e}rome H\'{e}ricourt \\
-{\footnotesize Universit\'{e} de Lille I \& CEPII }\\  \smallskip
+{\footnotesize Universit\'{e} de Lille \& CEPII }\\  \smallskip
 Lise Patureau \\
@@
 \item[-] But still significant: Average international trade costs = a 74\% markup over production costs (Anderson \& Van Wincoop, 2004)
+
+\pause
 \end{itemize}
@@
 \begin{itemize}
-\item[-] Transaction costs, policy costs, time costs, and transport costs \textit{per se}
-\end{itemize}
+\item[-] Transaction costs, policy costs, time costs, and... transport costs \textit{per se}
+\end{itemize}
+\pause
 \item Transport costs are a sizeable share of international trade costs \vspace{0.1cm}
@@
 \end{itemize}
-\textbf{The paper: On international transport costs}
-
-\end{frame}
-
+\textbf{This paper: Focus on international transport costs}
+
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
@@
 \item[$\ast$] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
-\item[$\Rightarrow$] Multiplicative costs:     \footnotesize $pq = \tau \widetilde{p}q$
-\normalsize
-\end{itemize}
-\end{itemize}
-\item Yet... A debated question \vspace{0.1cm}
-\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
+\item[$\Rightarrow$] \textbf{Multiplicative costs: }    \footnotesize $\mathbf{pq = \tau \widetilde{p}q}$
+\normalsize
+\end{itemize}
+\end{itemize}
+\pause
+
+\item Yet... A debated question \vspace{0.1cm}%\bigskip
+\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} 
+\pause
 \begin{itemize}
@@
 \item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
+\pause
 \begin{itemize}
@@
 %\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
-\item[$\Rightarrow$] Additive costs: \footnotesize $pq = (t+\widetilde{p})q  $
-\end{itemize}
-\end{itemize}
-\end{itemize}
-\end{frame}
+\item[$\Rightarrow$] \textbf{Additive costs}: \footnotesize $\mathbf{pq = (t+\widetilde{p})q}  $
+\end{itemize}
+\end{itemize}
+\end{itemize}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{itemize}
+\pause
 \item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.1cm}
@@
 \item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
-
+\pause
 \end{itemize}
 \item[$\Rightarrow$] Transport costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
-\end{itemize}
-\textbf{One objective of the paper: Provide an answer to this question}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Our paper in one question (and 3 answers)}
-
-\begin{itemize}
-\item  \textbf{Do additive transport costs matter?} \vspace{0.1cm}
+\item[$\Rightarrow$] How does this affect our perception of the time dynamics of transport cost?
+\end{itemize}
+\textbf{Objectives of the paper: answer both questions}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Our paper in a nutshell}
+
+\begin{itemize}
+\item  \textbf{What we do} \vspace{0.1cm}
 \begin{itemize}
@@
 \end{itemize}
-\item[$\Rightarrow$] \textbf{Yes, they do} \vspace{0.1cm}
-
-\item[(1)] \textbf{In terms of size}: Additive costs are sizeable \vspace{0.1cm}
+\item[$\Rightarrow$] \textbf{Question 1: do additive costs matter?} \vspace{0.1cm}
+
+\pause
+
+\item[(1)] \textbf{YES, THEY DO!}: Additive costs are sizeable \vspace{0.1cm}
 \begin{itemize}
@@
 \item[-] Roughly 50\% of the overall transport costs \vspace{0.1cm}
-
+\item[-] A substantially better ``goodness-of-fit'' when the additive component included
 %}
 \end{itemize}
-\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
-\begin{itemize}
+%\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
+%\begin{itemize}
 %\footnotesize{
-\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
-\item[-] A substantially better ``goodness-of-fit'' %}
+%\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
+%\item[-] A substantially better ``goodness-of-fit'' %}
 %}
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item[(3)] \textbf{In accounting for the time trend of transport costs} \vspace{0.2cm}
+%\end{itemize}
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+\begin{itemize}
+\item[$\Rightarrow$] \textbf{Question 2: does this change our accounting for the time trend of transport costs?} \vspace{0.1cm}
+\pause
+\item[(2)] \textbf{YES, IT DOES!} \vspace{0.2cm}
 \begin{itemize}
@@
 \item[-] Because it changes the result of excluding the trade composition effects when characterizing the time trend \vspace{0.1cm}
-\begin{itemize}
-\item[$\ast$] The decrease in TC over time is mostly attributable to the decrease in the ``pure'' transport costs,  \vspace{0.1cm}
-\item[$\ast$] Rather than to trade composition effects  \vspace{0.1cm}
-\end{itemize}
-\item[-] A result in sharp contrast with Hummels (2007) \vspace{0.2cm}
-\end{itemize}
+\pause
+\begin{itemize}
+\item[$\ast$] Hummels (2007): composition effects partly offset the decrease in transport costs  \vspace{0.1cm} %XXX CAREFUL to misinterpretations here!!!XXX
+\item[$\ast$] DHP (2018): composition effects do not matter much; if anything, they amplify (slightly) the decrease in pure transport costs.  \vspace{0.1cm}
+\end{itemize}
+%\item[-] A result in sharp contrast with Hummels (2007) \vspace{0.2cm}
+\end{itemize}
+\pause
 \end{itemize}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \item[$\Rightarrow$] In support of the additive cost component assumption \vspace{0.1cm}
-\item[$\neq$] Our results: Quantitatively estimate the \textit{value} of the additive costs \vspace{0.1cm}
+\pause
+\item[$\neq$] \textbf{Our results: Quantitatively estimate the \textit{value} of the additive costs} \vspace{0.1cm}
 \item[-] A broader view of the magnitude of additive costs \vspace{0.1cm}
@@
 %\end{itemize}
-\item[$\Rightarrow$] Additive costs do matter quantitatively, throughout the period } \vspace{0.1cm}
+
+\item[$\Rightarrow$] Additive costs do matter quantitatively, throughout the period} \vspace{0.1cm}
+\pause
 \end{itemize}
@@
 \item[-] Hummels (2007), Behar and Venables (2011) \vspace{0.1cm}
-\item[-] What lies behind the decrease in transportation costs observed in the recent decades? \vspace{0.1cm}
-\item[-] Hummels (2007): Mostly due to trade composition effects \vspace{0.1cm}
-\item[$\neq$] Our results: Trade composition effects only a minor role \vspace{0.1cm}
+\item[-] What is exactly behind the decrease in transportation costs observed in the recent decades? \vspace{0.1cm}
+\item[-] Hummels (2007): trade composition effects partly cancelled decrease in transport costs \vspace{0.1cm}
+\pause
+\item[$\neq$] \textbf{Our results: Trade composition effects only had a (very) minor impact on the trend} \vspace{0.1cm}
 \item[-] Why such a difference?  \vspace{0.1cm}
@@
 
-\begin{frame}
-\frametitle{Plan of the talk}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+\frametitle{Roadmap}
 \begin{itemize}
@@
 \item Results \vspace{0.2cm}
+\item A few Robustness Checks \vspace{0.2cm}
 \item Conclusion
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%ù
 
@@
 \item Database: US Imports of Merchandise database \vspace{0.1cm}
+\pause
 \begin{itemize}
@@
 %(Could go up to HS-10 with recent data)\vspace{0.1cm}
-\end{itemize}
-\item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
+
+\pause
+\end{itemize}
+\item[$\Rightarrow$] \textbf{Our dependent variable: Based on the ratio} $\mathbf{p/\widetilde{p}}$ \vspace{0.1cm}
+
+\pause
+
 \item The RHS is at the 3-digit classification level  \vspace{0.1cm}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
+\pause
 \end{itemize}
@@
 \item[-] Have both the import and the export prices  \vspace{0.1cm}
-\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive transport costs  \vspace{0.1cm}
+\pause
+
+\item[$\Rightarrow$] \textbf{Allows to estimate the levels of both the ad-valorem and the additive transport costs} \vspace{0.1cm}
 \begin{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
 \frametitle{Empirical specification (1)}
-\textbf{The equation at the root of the estimation}
+\textbf{The equation underlying the estimation}
 \begin{itemize}
@@
 
-\begin{frame}[label = slide_variancedecomp]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[label = est_strategy]
 \frametitle{Empirical specification (2)}
@@
 \end{itemize}
+
+\pause
+
 \item About the specification of the error term \vspace{0.1cm}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e., fixed effects) to be estimated \vspace{0.1cm}
+\pause
 \end{itemize}
@@
 \item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
-\begin{itemize}
-\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
-\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
-\end{itemize}
+%\begin{itemize}
+%\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
+%%\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
+%\end{itemize}
+\pause
 \item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \textbf{How to characterize the importance of additive costs relatively to ad-valorem?} \vspace{0.1cm}
+\pause
 \begin{itemize}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{itemize}
+\pause
 \item  Taking the weighted average over the sector-country dimension, we finally get, by year and transport mode: \vspace{0.1cm}
@@
 
-\begin{frame}
-\frametitle{Question, and answers}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+\frametitle{Reminder: two key questions}
 \begin{itemize}
 \item Do additive costs matter? \vspace{0.5cm}
-\item How we answer (yes) to this question
-\end{itemize}
-\end{frame}
+\item How does this impact the analysis of the time dynamics of transport costs?
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}[label = slide_result1]
@@
 \item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
-\begin{itemize}
-\footnotesize{
-\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
-\item[$\ast$] From 5.8\% to 3.2\% in maritime transport (mean value over the period) \vspace{0.1cm}
-\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
-\end{itemize}
-\end{itemize}
+%\begin{itemize}
+%\footnotesize{
+%\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
+%\item[$\ast$] From 5.8\% to 3.2\% in maritime transport (mean value over the period) \vspace{0.1cm}
+%\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
+%\end{itemize}
+\end{itemize}
+\pause
 \item[$\Rightarrow$] Additive costs, a sizeable and structural dimension of international transportation costs \vspace{0.1cm}
@@
 \end{itemize}
+\pause
 \item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[label = slide_result1_more]
+\frametitle{Result 1: Size of additive transport costs over time}
+
+\begin{figure}[htbp]
+\caption{Share of additive costs (3-digit classification)}
+\label{fig:part_cout_additif}
+\begin{center}
+\includegraphics[width=8cm, height=5cm]{part_cout_additif.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+ \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}[label=slide_goodnessfit]
-\frametitle{Result 2: Quality of fit}
+\frametitle{Result 1-bis: Quality of fit}
 \begin{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}[label = slide_fig1]
-\frametitle{Result 3: Characterizing the time trend of TC }
+\frametitle{Result 2: Characterizing the time trend of TC }
 \begin{itemize}
@@
 \begin{figure}[htbp]
-%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
-%\label{fig:decomp_TC_3d}
+\caption{Trend in estimated transport costs (Yearly mean value)}
+\label{fig:Trends_in_TC}
 \begin{center}
-\begin{tabular}{cc}
-{\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
-\includegraphics[width=5cm, height=4cm]{Fig2a_decompTC_air_3d.pdf}
-& \includegraphics[width=5cm,height=4cm]{Fig2b_decompTC_vessel_3d.pdf} \\
-\end{tabular}\end{center}
+\includegraphics[width=10.5cm, height=4.75cm]{Figure1_Trend_of_totalTC_bymode.pdf}
+\end{center}
 \end{figure}
+
 \begin{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
@@
 \item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.2cm}
+\pause
 \item The change in overall transport costs over time: \vspace{0.1cm}
@@
 \end{itemize}
+\pause
 \end{itemize}
 \item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
-\item[$\Rightarrow$] What we do, in accordance with Hummels (2007)
-
-\end{itemize}
-\end{frame}
-
+\item[$\Rightarrow$] What we do, in accordance with Hummels (2007)...
+\pause
+\item ... BUT with one major difference: we assume ($\neq$ Hummels, 2007) a varying share of the additive component over time, product and country partner
+ \vspace{0.1cm}
+
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[label = slide_result2_more]
+\frametitle{Distribution of the additive component}
+
+\begin{figure}[htbp]
+\caption{Share of additive costs in total costs: Histogram}
+\label{fig:histogram_beta}
+\begin{center}
+\begin{tabular}{cc}
+%{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=5.5cm, height=4.5cm]{Etude_beta_pondere_air.pdf}
+& \includegraphics[width=5.5cm, height=4.5cm]{Etude_beta_pondere_ves.pdf} \\
+\multicolumn{2}{l}{{\tiny Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+ \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame} [label = slide_compeffects_strategy]
@@
 \item Start with our estimates of additive/multiplicative transport costs (TC) ($\widehat{\tau}_{ikt}$, $\widehat{t}_{ikt}$)\vspace{0.1cm}
+\pause
 \item Decompose the estimated TC component in three parts \vspace{0.1cm}
@@
 \end{itemize}
+\pause
 \item Extract the ``pure'' TC component by the mean of a time fixed-effect \vspace{0.1cm}
@@
 \end{itemize}
+
+\pause
 \item Build a unified measure of ``total'' transport costs (agglomerate the two components), both fitted and unfitted \vspace{0.1cm}
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%ù
 
@@
 
-\begin{frame}
-\frametitle{Time trends in transport costs: Comments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%ù
+
+\begin{frame}
+\frametitle{Time trends in transport costs: Comments} %XXX ADJUST FROM HERE XXX
 \begin{itemize}
@@
 \end{itemize}
+\pause
+
 \item \textbf{Trade composition effects: A minor role in accounting for the time trend of transport costs}\vspace{0.1cm}
@@
 \end{itemize}
+\pause
 \item In Air transport \vspace{0.1cm}
@@
 \footnotesize
-\item[-] The 50\% decrease is mainly attributable to the reduction of ``pure'' transport costs, for both components  \vspace{0.1cm}
+\item[-] The 50\% decrease is almost exclusively attributable to the reduction of ``pure'' transport costs, for both components  \vspace{0.1cm}
 \end{itemize}
@@
 \item[-] The 60\% decrease is roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
-\item[-] But for the multiplicative component only \vspace{0.1cm}
-\end{itemize}
-\normalsize
-\end{itemize}
-\end{frame}
+%\item[-] But for the multiplicative component only \vspace{0.1cm}
+\end{itemize}
+\normalsize
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \item[-] According to whom pure transport costs have decreased much more than the unfitted ones (over 1974-2004) \vspace{0.1cm}
-\item[-] Suggesting an important role to trade composition effects  \vspace{0.1cm}
-\end{itemize}
+\item[-] Suggesting an important, offsetting role to trade composition effects  \vspace{0.1cm}
+\end{itemize}
+\pause
 \item Not the case here \vspace{0.1cm}
 \item[$\Rightarrow$] \textbf{Where does the difference come from?} \vspace{0.1cm}
+\pause
 \item \textbf{From the treatment of the additive component of transport costs} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{More on this}}\vspace{0.1cm}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%ù
+
 \begin{frame}
@@
 \item[$\Rightarrow$] Transport cost changes decompose as:
+
 \begin{figure}[htbp]
-%\caption{Transport costs (with and without composition effects)}
-%\label{fig:totalTC_compeffects_excl}
+\caption{Characterizing the time trends: Applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
 \begin{center}
 \begin{tabular}{cc}
-{\scriptsize (a) Air } & {\scriptsize (b) Vessel}\\
-\includegraphics[width=4cm, height=3cm]{figure5_comme_hummels_air.pdf}
-& \includegraphics[width=4cm,height=3cm]{figure6_comme_hummels_ocean.pdf} \\
+{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
+\multicolumn{2}{c}{{\small In percent of the export price}} \\
+\includegraphics[width=2in, height=1.5in]{figure5_comme_hummels.pdf}
+& \includegraphics[width=2in, height=1.5in]{figure6_comme_hummels.pdf} \\
+%{\small (c) Ad-valorem Air Freight} & {\small (d) Ad-valorem Vessel Freight}\\
+%\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
+%\includegraphics[width=1.25in, height=0.75in]{figure5_comme_hummels_base100.pdf}
+%& \includegraphics[width=1.25in, height=0.75in]{figure6_comme_hummels_base100.pdf} \\
 \end{tabular}
@@
 \end{figure}
+%\item Composition effects tend to partially offset the decrease in the fitted transport costs, in both Air and Vessel \vspace{0.1cm}
+%\begin{itemize}
+%\item[-] The downward trend being of higher magnitude for the fitted rate, especially for Vessel
+%\end{itemize}
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%ù
+
+\begin{frame}
+%\frametitle{Time trend of transport costs \& the modeling of the additive component}
+%\begin{itemize}
+%\item Assuming (as Hummels, 2007) a constant share of the additive component over time/country partner/product \vspace{0.1cm}
+%\item[$\Rightarrow$] Transport cost changes decompose as:
+
+\begin{figure}[htbp]
+\caption{Characterizing the time trends: Applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
+\begin{center}
+\begin{tabular}{cc}
+%{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
+%\multicolumn{2}{c}{{\small In percent of the export price}} \\
+%\includegraphics[width=1.25in, height=0.75in]{figure5_comme_hummels.pdf}
+%& \includegraphics[width=1.25in, height=0.75in]{figure6_comme_hummels.pdf} \\
+{\small (c) Ad-valorem Air Freight} & {\small (d) Ad-valorem Vessel Freight}\\
+\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
+\includegraphics[width=2in, height=1.5in]{figure5_comme_hummels_base100.pdf}
+& \includegraphics[width=2in, height=1.5in]{figure6_comme_hummels_base100.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+\begin{itemize}
+
 \item Composition effects tend to partially offset the decrease in the fitted transport costs, in both Air and Vessel \vspace{0.1cm}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}[label = slide_compeffects_results]
@@
 \end{itemize}
+\pause
 \item Assuming a varying share of the additive component over time/country partner/product  \vspace{0.1cm}
@@
 \end{itemize}
-\item[$\Rightarrow$] The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs
-\end{itemize}
-\end{frame}
+
+\pause
+
+\item[$\Rightarrow$] \textbf{The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs}
+\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[label = robustness1]
+\frametitle{Robustness 1: Separability assumption}
+\begin{itemize}
+\item Until now: assume separability assumption of country-good fixed effects
+\begin{itemize}
+\item[$=$]separable between the origin country ($i$) and the product ($s(k)$) dimensions.\vspace{0.1cm}
+\pause
+\item[$\Rightarrow$] What if we relax this assumption? Computationally costly (40,000 FE), we restrict to the countries ($i$) and the products constituting 80\% of the total value of trade flows.
+ \end{itemize}
+ \pause
+\begin{table}[htbp]
+  \centering
+   \scriptsize{
+  \caption{Robustness to the separability assumption}
+\begin{center}
+    \begin{tabular}{lcc|cc}
+    \hline \hline
+\multicolumn{5}{c}{Mean value over 1974-2013} \\
+ \multicolumn{5}{c}{3 digits} \\    \hline \hline
+   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
+   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
+    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
+in \%)}  \\
+    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
+    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
+    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
+    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
+    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
+\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
+    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
+    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
+% \textit{Diagnostic test} & & &  &  \\
+    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
+    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
+ %   AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
+    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
+\# observations & 2381 & 2381 & 2798 & 2798 \\
+\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
+\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
+    \end{tabular}%
+\end{center}}
+\end{table}
+   \end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[label = robustness2]
+\frametitle{Robustness 2: Weighting Scheme}
+\begin{itemize}
+\item A difference between DHP and Hummels (2007)'s: the weighting scheme used to get the evolution of the pure TC over time.
+\begin{itemize}
+\item[-]Hummels: unweighed average value over the $i,k$ dimension\vspace{0.1cm}
+\item[-]DHP: each flow weighted by its rel. value in total trade flows in 1974.
+\pause
+\item[$\Rightarrow$]rebuild the fitted values for each TC measure taking the unweighed average value.
+ \end{itemize}
+
+\begin{figure}[htbp]
+\small{\caption{Transport cost time trends: Robustness to the weighting scheme}}
+\label{fig:compeffects_robustness}
+\begin{center}
+\includegraphics[height=5cm]
+{graph_composition_all_np.pdf}
+\end{center}
+\end{figure}
+   \end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \textbf{Our paper: Empirical evidence about the role of the additive component in international transport costs} \vspace{0.1cm}
+\pause
 \begin{itemize}
@@
 \end{itemize}
+\pause
 \item The importance of taking into account additive transport costs \vspace{0.1cm}
@@
 \item[-] Additive costs are sizeable quantitatively \vspace{0.1cm}
-\item[-] A better fit of the model when they are taken into account    \vspace{0.1cm}
+%\item[-] A better fit of the model when they are taken into account    \vspace{0.1cm}
 \item[-] A key role in the understanding of the time trend of transport costs
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \item Hummels (2007) does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effect  - but only a scale effect \vspace{0.1cm}
-\item We separate the country-product fixed effects in two components (country / product), due to the number of fixed effects to estimate in the non-linear setting - robustness analysis to this point (in progress) \vspace{0.1cm}
+\item We separate the country-product fixed effects in two components (country / product), due to the number of fixed effects to estimate in the non-linear setting - robustness analysis to this point \vspace{0.1cm}
 \item We differ in the weighting scheme to obtain the yearly value of the fitted TC measure.  \vspace{0.1cm}

### 2018-09-26 redaction/online_appendix_travail.tex
Travail Guillaume Online Appendix

--- 2017-09-12 redaction/online_appendix_dhp2017.tex
@@
 \urlstyle{rm}
-\author{Guillaume Daudin\footnote{\noindent Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
-SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Corresponding author.
-Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille1.fr}} \qquad \quad Lise Patureau\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France ; email: \url{lise.patureau@dauphine.fr}}}
+\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
+SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille1.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France ; email: \url{lise.patureau@dauphine.fr}}}
 %\date{\vspace{-5ex}}
@@
 
+The paper compares the empirical performances of two models: one with only ad-valorem costs (Model (A)) and one with where both ad-valorem and additive costs (Model (B)).
+In the interest of comprehensiveness, this appendix looks at a model with only additive costs (Model (C)).
+The estimated equation is: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. \\
+We first compare estimates of the transport costs components under the three models. In a second step, we report quality of fit tests for the three models.
+
+\subsubsection{Estimation results}
+
+In this section, we report the estimation results of the three models.
+We also report the estimation results for Model (C), where only additive costs are modeled.
+Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_vessel} for vessel transport.
+
 %%%% RAW TABLE TO BE FOUND AT: "...\Dropbox\Papier_Lise_Guillaume\trade_cost\results\3_models"
@@
 \renewcommand{\thetable}{A.\arabic{table}}
-%\subsubsection{Nominal Exchange Rate Volatility}
 
@@
   \centering
+   \footnotesize{
     \caption{Estimation results of the three models (Air, 3-digit level)}
   \label{tab:3models_estimation_results_air}%
-    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \begin{tabular}{l|c c c c c c c c c}
     \hline \hline
@@
     \end{tabular}
+    }
 \end{table}%
+
+\begin{table}[htbp]
+  \centering
+   \footnotesize{
+    \caption{Estimation results of the three models (Vessel, 3-digit level)}
+  \label{tab:3models_estimation_results_vessel}%
+    \begin{tabular}{l|c c c c c c c c c}
+    \hline \hline
+    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
+    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
+    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
+    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
+    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
+    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
+    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
+     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
+    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
+    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
+    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
+       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
+    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
+    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
+    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
+    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
+    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
+    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
+    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
+   \end{tabular}%
+   }
+\end{table}%
+
+Results for Models (A) (ad-valorem costs only) and (B) (both ad-valorem and additive costs) are identical to those reported in the paper.
+Somewhat unsurprisingly, the estimated size of transport costs under Model (C) is of same order of magnitude as in Models (A) and (B).
+We also observe a downward trend of transport costs over time, in particular since 1980.\\
+
+\subsubsection{Quality of fit diagnostic tests}
+
+To go further into characterizing the empirical relevance of the model with aditive costs only (Model (C)), we compare quality of fit diagnostic tests in this section.
+Table \ref{tab:3models_diagnosis_air} reports the results for air transport, and Table \ref{tab:3models_diagnosis_vessel} those for vessel transport.
+In both tables, we report the values of the $R^2$, the Standard Error of Regression (SER), the AIC criterion  and the log-likelihood (LL) value.
+We also report the value of the log-likelihood ratio that tests the quality of fit of the global model (Model (B)) compared to the other two models.
 
@@
     Model (A) & \multicolumn{1}{c}{0.297} & \multicolumn{1}{c}{0.267} & \multicolumn{1}{c}{0.302} & \multicolumn{1}{c}{0.251} & \multicolumn{1}{c}{0.142} & \multicolumn{1}{c}{0.318} & \multicolumn{1}{c}{0.460} & \multicolumn{1}{c}{0.421} & \multicolumn{1}{c}{0.313} \\
-    Model (B ) & \multicolumn{1}{c}{0.594} & \multicolumn{1}{c}{0.646} & \multicolumn{1}{c}{0.635} & \multicolumn{1}{c}{0.627} & \multicolumn{1}{c}{0.658} & \multicolumn{1}{c}{0.640} & \multicolumn{1}{c}{0.593} & \multicolumn{1}{c}{0.513} & \multicolumn{1}{c}{0.419} \\
+    Model (B) & \multicolumn{1}{c}{0.594} & \multicolumn{1}{c}{0.646} & \multicolumn{1}{c}{0.635} & \multicolumn{1}{c}{0.627} & \multicolumn{1}{c}{0.658} & \multicolumn{1}{c}{0.640} & \multicolumn{1}{c}{0.593} & \multicolumn{1}{c}{0.513} & \multicolumn{1}{c}{0.419} \\
     Model (C) & \multicolumn{1}{c}{0.489} & \multicolumn{1}{c}{0.543} & \multicolumn{1}{c}{0.531} & \multicolumn{1}{c}{0.517} & \multicolumn{1}{c}{0.546} & \multicolumn{1}{c}{0.518} & \multicolumn{1}{c}{0.464} & \multicolumn{1}{c}{0.339} & \multicolumn{1}{c}{0.295} \\ \hline
@@
   \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
-    \label{tab:3models_diagnosis_air}%
-   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
+    \label{tab:3models_diagnosis_vessel}%
+   \begin{tabular}{l|c c c c c c c c c}
     \hline \hline
@@
 
+The model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered.
+That justify our choice to disregard Model (C) in the main text.
+
 %\subsection{Detailed Results \label{secoa:detailed}}
+\newpage
 \setcounter{table}{0}
@@
 \subsection{Transport Cost Estimates: Yearly Detailed Results}
+In this section, we complement Table 1 of the main text by reporting the year-to-year results of the estimation driven at the 3-digit classification level.
+Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for vessel.
+In both cases, we report the estimated values of the transport costs (weighed mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).
 
-\subsubsection{3-Digits Level Product Classification}
 \begin{landscape}
@@
 
+As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000).
+This can be attributed to the presence of outliers in the distribution of the additive costs estimates.
+The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000.
+Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
+Still, this does not make much difference.
+When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
+The weighed median value is left unchanged at 2.9\% whether 1989 is included or not.
+
 \begin{landscape}
@@
 
-\subsection{Eliminating the composition effects: Primary vs. Maufacturing sector} \label{sec_oa:comp-effects}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{C.\arabic{figure}}
 
-%\todo{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}
+\subsection{Eliminating the composition effects: Primary vs. Manufacturing sector} \label{sec_oa:comp-effects}
 
-In this Section, we characterize the time trend in international transport costs at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods.
-The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
+In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows.
+The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
+For easing comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.
+
+The classification retained to categorize trade flows follows the UNCTAD classification (on STIC Revision 3)\footnote{See "UNCTAD product groupings and composition (SITC Rev.
+3)" in http://unctadstat.unctad.org/EN/Classifications.html, accessed Septembre 2018}.
+Are considered as ``primary goods'' all flows recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).
 
@@
 \begin{center}
-\includegraphics[height=4in]
-{graph_composition_manuf.pdf}
+\includegraphics[height=8cm] {graph_composition_manuf.pdf}
 \end{center}
@@
 \begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects)}
+\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_composition_all.pdf}
+\end{center}
+\end{figure}
+
+As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, the ``raw'' transport costs and the unfitted transport costs, in the red dashed line plain have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel).
+However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
+Figure \ref{fig:totalTC_compeffects_excl_manuf} (for manufactures) reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
+In air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
+Trade composition effects matter more in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)), primarily in the ad-valorem component.
+As for the whole range of flows, the 60\% decrease in the unfitted transport costs in vessel can de decomposed in a 50\% decrease in the ``ceteris paribus'' transport costs (fitted), the 10\% remaining to trade composition effects.
+
+The situation is is strikingly for primary goods only.
+In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)).
+Furthermore, in air transport, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).
+
+One explanation of these contrasted results lies in the share of primary goods in total flows.
+It is reported in Figure \ref{fig:Share_prim_goods}.
+In air transport, the share of primary goods in the total value of US imports is very small, around 10\% all over the period from 1974 to 2014.
+Primary goods make a higher proportion of trade flows in vessel transport, especially over 1974-1982 (between 40\% and 60\%).
+On the following sub-period though, their share has fallen to 20-30\%.
+Given the modest proportion of primary goods in total imports flows of the US economy compared to the manufactured sector, it is hence not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the trend patterns that occur within the manufacturing sector.
+\\
+Still, it is interesting to notice that composition effects are not dominated by changes in the primary goods share.
+If they were, we would not observe the same trends for trade as the whole and trade of manufactured goods.
+
+\begin{figure}[htbp]
 \caption{Share of primary goods in the value of total US imports}

### 2018-09-27 redaction/online_appendix_travail.tex
Petits changements divers

@@
 
-One explanation of these contrasted results lies in the share of primary goods in total flows.
-It is reported in Figure \ref{fig:Share_prim_goods}.
-In air transport, the share of primary goods in the total value of US imports is very small, around 10\% all over the period from 1974 to 2014.
+One explanation for the similarity between the results for the manufacturing goods trade and for total trade can be found in the share of primary goods in total flows as reported in Figure \ref{fig:Share_prim_goods}.
+In air transport, the share of primary goods in the total value of US imports is very small, around 10\%.
 Primary goods make a higher proportion of trade flows in vessel transport, especially over 1974-1982 (between 40\% and 60\%).
 On the following sub-period though, their share has fallen to 20-30\%.
-Given the modest proportion of primary goods in total imports flows of the US economy compared to the manufactured sector, it is hence not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the trend patterns that occur within the manufacturing sector.
-\\
-Still, it is interesting to notice that composition effects are not dominated by changes in the primary goods share.
-If they were, we would not observe the same trends for trade as the whole and trade of manufactured goods.
+Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured sector, it is hence not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the trend patterns that occur within the manufacturing sector.
 

### 2018-09-27 redaction/online_appendix_travail.tex
1ers chgts OA par JH
Juste titre et affiliations

--- 2018-09-26 redaction/online_appendix_travail.tex
@@
 \begin{document}
-\title{\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix}
+\title{\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)}
 
 \urlstyle{rm}
-\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
-SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille1.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
-Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France ; email: \url{lise.patureau@dauphine.fr}}}
+\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
+Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
+Universit\'{e} Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE ; email: \url{lise.patureau@dauphine.fr}}}
 %\date{\vspace{-5ex}}

### 2018-09-27 redaction/dhp_travail.tex

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine, Paris-I and Nancy for very useful discussions.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Olger Görg, S\'{e}bastien Jean, Vincent Vicard for their insightful comments.
+The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the Universites of California-Irvine, Paris-Dauphine, Paris-I, Nancy and the the Kiel Institute for the World Economy.
 Finally, any remaining errors are ours.
-This paper features an online appendix containing additional results and available on the authors' websites.}}
+This paper was partly funded by an University grant from Lille-I University.
+It features an online appendix containing additional results and available on the authors' websites.}}
 

### 2018-09-27 redaction/online_appendix_travail.tex
fin modifs OA par JH

@@
     Median (in \%)& \multicolumn{1}{c}{5.43} & \multicolumn{1}{c}{3.79} & \multicolumn{1}{c}{5.47} & \multicolumn{1}{c}{4.37} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{2.47} & \multicolumn{1}{c}{3.12} & \multicolumn{1}{c}{3.41} & \multicolumn{1}{c}{2.92} \\
-    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
+    Standard Deviation & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
     \multicolumn{10}{l}{\textbf{Model (B) - With additive and ad-valorem transport costs}} \\ \hline
@@
     Median (in \%)& \multicolumn{1}{c}{2.71} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{1.79} & \multicolumn{1}{c}{1.60} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.20} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{2.24} & \multicolumn{1}{c}{1.72} \\
-    Std & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline
+    Standard Deviation & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline
       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)} \\ \hline
@@
     Median (in \%)& \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.43} & \multicolumn{1}{c}{0.47} \\
-    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
+    Standard Deviation & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
     \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
@@
     Median (in \%)& \multicolumn{1}{c}{4.45} & \multicolumn{1}{c}{1.82} & \multicolumn{1}{c}{3.36} & \multicolumn{1}{c}{2.27} & \multicolumn{1}{c}{1.72} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.93} & \multicolumn{1}{c}{2.72} & \multicolumn{1}{c}{2.06} \\
-    Std & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline
+    Standard Deviation & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline
     \end{tabular}
@@
     Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
-    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
+    Standard Deviation & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
     \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
@@
     Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
-    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
+    Standard Deviation & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
        \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
@@
     Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
-    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
+    Standard Deviation & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
     \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
@@
     Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
-    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
+    Standard Deviation & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
    \end{tabular}%
@@
 
-To go further into characterizing the empirical relevance of the model with aditive costs only (Model (C)), we compare quality of fit diagnostic tests in this section.
+To go further into the comparison of the empirical relevance of our three empirical models (A), (B) and (C), we compare quality of fit diagnostic tests in this section.
 Table \ref{tab:3models_diagnosis_air} reports the results for air transport, and Table \ref{tab:3models_diagnosis_vessel} those for vessel transport.
@@
     Model (A) & \multicolumn{1}{c}{35675.0} & \multicolumn{1}{c}{41171.0} & \multicolumn{1}{c}{49315.0} & \multicolumn{1}{c}{60715.6} & \multicolumn{1}{c}{74386.4} & \multicolumn{1}{c}{87492.5} & \multicolumn{1}{c}{103983.0} & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{106130.6} \\
-    Model (B ) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\
+    Model (B) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\
     Model (C) & \multicolumn{1}{c}{40808.1} & \multicolumn{1}{c}{45138.5} & \multicolumn{1}{c}{55214.8} & \multicolumn{1}{c}{69458.5} & \multicolumn{1}{c}{83958.6} & \multicolumn{1}{c}{100040.8} & \multicolumn{1}{c}{123592.1} & \multicolumn{1}{c}{129359.0} & \multicolumn{1}{c}{127399.2} \\ \hline
@@
     Model (A) & \multicolumn{1}{c}{-17530.5} & \multicolumn{1}{c}{-20253.5} & \multicolumn{1}{c}{-24315.5} & \multicolumn{1}{c}{-29977.8} & \multicolumn{1}{c}{-36811.2} & \multicolumn{1}{c}{-43341.3} & \multicolumn{1}{c}{-51648.5} & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-52690.3} \\
-    Model (B ) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\
+    Model (B) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\
     Model (C) & \multicolumn{1}{c}{-20074.1} & \multicolumn{1}{c}{-22217.2} & \multicolumn{1}{c}{-27251.4} & \multicolumn{1}{c}{-34355.3} & \multicolumn{1}{c}{-41634.3} & \multicolumn{1}{c}{-49625.4} & \multicolumn{1}{c}{-61533.0} & \multicolumn{1}{c}{-64339.5} & \multicolumn{1}{c}{-63316.6} \\ \hline
@@
     Model (B) & 0.484 & 0.528 & 0.493 & 0.514 & 0.512 & 0.551 & 0.578 & 0.656 & 0.684 \\
-    Model (C ) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
+    Model (C) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
     \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
@@
     Model (B) & 27331.52 & 28067.31 & 34170.52 & 43664.74 & 49275.33 & 60475.91 & 73020.09 & 76161.33 & 80873.72 \\
-    Model (C ) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
+    Model (C) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
     \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
@@
     Model (B) & -12985.76 & -13353.65 & -16398.26 & -21171.37 & -23905.66 & -29490.96 & -35844.04 & -37418.66 & -39751.86 \\
-    Model (C ) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
+    Model (C) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
    \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
@@
 The model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered.
-That justify our choice to disregard Model (C) in the main text.
+That justifies our choice to disregard Model (C) in the main text.
 
@@
 Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for vessel.
-In both cases, we report the estimated values of the transport costs (weighed mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).
+In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).
 
@@
 
-The situation is is strikingly for primary goods only.
+The situation is strikingly for primary goods only.
 In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)).

### 2018-09-28 redaction/dhp_travail.tex
typo
typo sur le ö de Görg

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Olger Görg, S\'{e}bastien Jean, Vincent Vicard for their insightful comments.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Olger G\"{o}rg, S\'{e}bastien Jean, Vincent Vicard for their insightful comments.
 The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the Universites of California-Irvine, Paris-Dauphine, Paris-I, Nancy and the the Kiel Institute for the World Economy.

### 2018-12-17 redaction/TC_new.tex
changements typos graphiques

--- 2018-08-05 redaction/dhp_travail.tex
@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine, Paris-I and Nancy for very useful discussions.
-Finally, any remaining errors are ours.
-This paper features an online appendix containing additional results and available on the authors' websites.}}
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Holger Görg, S\'{e}bastien Jean and Vincent Vicard for their insightful comments.
+The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the universities of California-Irvine, Paris-Dauphine, Paris-I and Nancy and the Kiel Institute for the World Economy.
+Any remaining errors are ours.
+This paper was partly funded by a university grant from Lille-I University.
+It features an online appendix containing additional results, available on the authors' websites.}}
 
 \author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
+University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
 Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
-; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-Universit\'{e} Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
-
-\date{June 2018}
+; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+University Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.psl.eu} } }
+
+\date{December 2018}
  \maketitle
@@
 \begin{abstract}
-This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows over 1974-2013.
+This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2013.
 First, we document the importance of the per-unit (additive) component of transport costs.
-We thus find that additive costs are quantitatively sizable, representing between one third and one half of overall transport costs.
+We find that additive costs are quantitatively sizable, representing between one-third and one-half of overall transport costs.
 Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
 Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
-Importantly, this difference of results can be attributed to the new way of modeling the per-unit component of transport costs we offer.
+Importantly, this difference of results can be attributed to the new method of modeling the per-unit component of transport costs we offer.
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
@@
 In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Based on panel data, \citet{novy13} finds  that U.S.
-trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
+Based on panel data, \citet{novy13} finds  that US
+trade costs with major trading partners declined on average by about 40\% between 1970 and 2000.
 Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
 Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
-Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
-
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs.
+Using data over 1989-2000, \citet{anderson_wincoop_jel} estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
+
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
+In this vein, \citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
 Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
-If much of trade policy barriers have been removed over the second half of the twentieth century, this suggests that the transport costs component remains large and deserves attention.
-International transport costs accordingly stand at the heart of the paper.\smallskip
+With many trade policy barriers removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention.
+International transport costs accordingly stand at the heart of this paper.\smallskip
 
@@
 Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
-There is no consensus thought.
-Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
-In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined more slowly than the common wisdom assumes.\smallskip
-
-According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries).
-Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
+However, there is no consensus.
+Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
+In accordance with this view, \cite{hummels2007} points out that transport costs as a proportion of export prices have declined more slowly than is commonly assumed.\smallskip
+
+According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e.
+changes over time in the baskets of imported goods and/or the origin countries).
+Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e.
+\textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
-Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
+Rather, the evolution of transport costs \textit{per se} constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
 In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
@@
 
-Proving this point drives us to provide a careful modelling of transport costs.
-Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}).
+Proving this point drives us to provide a careful modeling of transport costs.
+Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an ad-valorem tax equivalent (i.e.
+as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} However, a recent strand of the literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015}, or \citealp{martin2012}).
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
@@
 This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013).
+This database is the same as that used by \cite{hummels2007}, extended over more recent years (2005-2013).
 \smallskip
@@
 This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
-Put it differently, additive costs represent between one third and almost one half of the overall transport costs.
+Put differently, additive costs represent between one-third and almost one-half of the overall transport costs.
 %\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?}
@@
 
-Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost ``per se''.
-Trade composition effects only have a limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
+Second, we find that the main source of the downward trend in overall transport costs, by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost \textit{per se}.
+Trade composition effects have only limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
 Otherwise stated, pure transport costs have declined less than observed costs.
@@
 Our result thus deepens the paradox of the less-than-expected decline of transport costs.
-Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer.
-Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
+Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
 \smallskip
 
-In both aspects, our results point the importance of the additive component in accounting for international transport costs.
+In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
-As shown by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs.
-In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out'').
+As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
+In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out'').
 This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
-Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs.
-Recent empirical studies based on micro-level data also provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs.
+Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.
+Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
@@
 This is confirmed by \cite{Irrazabal_2015}.
-While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains.
+While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been made in quantifying such gains.
 One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
-One contribution of the paper is to fill this gap.}
+One contribution of this paper is to fill that gap.}
 
@@
 Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost ; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
+First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
+While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
 This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
-Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013.
+Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2013.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
 
@@
 First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
- Section \ref{sec:data_method} thus reports the mean values of the transport costs estimated over the period (by transport mode).
-Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''.
-Section \ref{sec:results_trends} is devoted to this analysis, while section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
+ Section \ref{sec:data_method} reports the mean values of the transport costs estimated over the period (by transport mode).
+Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
+Section \ref{sec:results_trends} is devoted to this analysis, while Section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
 Section \ref{sec:conclu} concludes.
@@
 
-Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
-The database we use to construct our measure of transport costs comes from US annual "Imports of Merchandise" provided by the Census bureau, spanning from 1974 to 2013.
-Details on the database are provided in Appendix \ref{app:data}.
-We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price.
-The fas price (for ``Free Alongside Ship'') means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
+Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
+The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2013 (details on the database provided in Appendix \ref{app:data}).
+We first use customs values, quantities and freight costs to recover FAS and cost-insurance-freight (CIF) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the FOB (free on board) price rather than the FAS price.
+The FAS price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
 It is also the seller's responsibility to clear the goods for export.
-The fob price (for ``Free on Board'') means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
-Once the goods clear the railing of the vessel the buyer assumes the risk.
+The FOB price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
+Once the goods clear the railing of the vessel, the buyer assumes the risk.
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs.
-The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
-By construction higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent.
-This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
+While both terms are closely related, the US Foreign Trade Statistics reports FAS prices.} More precisely, the (unit) FAS price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the goods net of transportation costs.
+The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+Our dependant variable is finally computed as the ratio of the CIF price divided by the FAS price.
+By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
+This is quite a standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
 
 Obviously, this dataset has limitations.
-First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
-It is thus silent about the others dimensions of international trade costs.
-Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...).
+First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the CIF-FAS price gap only covers freight, insurance and handling costs.
+It is thus silent about the other dimensions of international trade costs.
+Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size, etc.).
 In this respect, this dataset embraces only a partial view of international transport barriers.
@@
 According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
-Further, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers.
-They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs.
+Further, the evidence summarized in \citet{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
+They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
 In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}
@@
 This limits the measurement error bias.
-Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
+Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
 %This will be crucial to compute transport costs, as we detail below.
 In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
-We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode.
+We will make use of this to highlight potential differences in the dynamics of transport costs across transportation mode.
 Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
@@
 \cite{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which we exploit for providing an analysis of the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
+Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
@@
 
-As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level.
-Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
+As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+We estimate international transport costs at the 3-digit classification level, even if data series on the CIF and FAS prices are available at the 5-digit level.
+However, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
 However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.
-
-\subsection{Estimating transport costs: Empirical specification}
-
-\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs costs.
-To do so, we start from the equation that expresses the price $p$ (per kg) of a good paid by the importer (import, or cif price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or fas price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
+Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.
+
+\subsection{Estimating transport costs: empirical specification}
+
+\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
+To do so, we start from the equation that expresses the price $p$ (per kg) of goods paid by the importer (import, or CIF price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or FAS price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
 
@@
 
-\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
+\paragraph{Estimation strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
 In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
@@
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the CIF price $p$ cannot be lower than the FAS price ($p_{ik}>\widetilde{p}_{ik}$).
 Taking into account this constraint in the estimate implies that the error term should be always positive.
@@
 
-Equation (\ref{eq:equation0}) is non linear.
-It cannot be estimated using standard linear estimators.
+Equation (\ref{eq:equation0}) is non-linear. It cannot be estimated using standard linear estimators.
 All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
 The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
-See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Confronted to this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$).
-This amounts making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
-same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
-As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
+See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit level ($k$).
+This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
+same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed CIF-FAS price.
+As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
 on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
-This would not solve the problem though, for three main complementary reasons.
+This would not solve the problem, however, for three main complementary reasons.
 First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
@@
 After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode).
-For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5 percent of the upper and lower tails of the distribution in the regression variables.
+For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5\% of the upper and lower tails of the distribution in the regression variables.
 %JH: TO KEEP IN MIND: one day, we could be asked to winsorize tails, instead of cutting them...
 These cut-offs are aimed at eliminating reporting or coding errors.
-We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$).
+We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
 Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
@@
 
-As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, that decomposes into both an additive and an ad-valorem component.
+As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, which decomposes into both an additive and an ad-valorem component.
 In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others).
@@
 Under Model (b), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
-From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
+From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modeling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
 This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
-The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results.
-More details of these results in the Online Appendix.}
-
-One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden.
-Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
+The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating further results.
+More details of these results are given in the Online Appendix.}
+
+One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
 This is not an issue here, as we are not interested in causal inference.
@@
 
-After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
+After estimating Equation (\ref{eq:model_IetA}), we can rebuild a measure of each component, $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
 When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is(k)} = \widehat{\tau}_{i} \times \widehat{\tau}_{s(k)}$.
 In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.
-Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases.
+Similarly, as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (a), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (b), for each year and transportation mode.
@@
 %Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
-\paragraph{Quantification} The first step of the analysis consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+\paragraph{Quantification} The first step of the analysis consists of providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
 This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter.
 These results constitute our first original contribution to the literature.
-We also provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation (see results in Appendix \ref{app:diagnostic_test}).\smallskip
+We also provide diagnostic tests about the goodness-of-fit measures, which confirm the empirical relevance of adding the additive component in the estimation (see results in Appendix \ref{app:diagnostic_test}).\smallskip
 
@@
 The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
-Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3- and 4-digit classification level.
-Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
-Overall results (over the whole period) are not substantially affected if this year is included though.
-This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}
+Finally, the bottom panel reports the same set of descriptive statistics, but for the actual CIF/FAS ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
+Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for air transport at the 3-digit level, removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
+Overall results (over the whole period) are not substantially affected if this year is included.
+This can be seen in the results detailed year by year in the Online Appendix, available on the authors' webpages.}
 
@@
   \footnotesize{
-  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
+  \caption{Transport costs estimates: summary \label{tab:summary_results}}
   \begin{center}
@@
 \parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
-The additive term is expressed in fraction of fas price.
+The additive term is expressed in fraction of FAS price.
 For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
@@
 As displayed in Table \ref{tab:summary_results}, estimated mean values are systematically higher than medians, by around one percentage point.
-This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
+This result is not surprising, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
 One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
-This is first due to the fact that observed data are \textit{de facto} trade-weighed (i.e., by the weight of each good in total trade) while our estimates are based on simple arithmetic means within sector and country.
+This is first due to the fact that observed data are \textit{de facto} trade-weighted (i.e.
+by the weight of each good in total trade) while our estimates are based on simple arithmetic means within sector and country.
 Furthermore, Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization.
@@
 First, it provides estimation results about the size of the overall transport costs.
-From Table \ref{tab:summary_results}, transport costs amount to around a 6\% markup in ocean shipping, 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+Table \ref{tab:summary_results} shows that transport costs amount to around a 6\% markup in ocean shipping, and 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
 Under Model (a) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
 Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
-With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible candidate explanation for the difference with \citet{anderson_wincoop_jel}, lays in the difference of empirical methodology.
+With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in vessel.
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \citet{anderson_wincoop_jel}, lays in the difference of empirical methodology.
 \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
@@
 On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
-Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively.
-This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period).
-Put differently, the additive component amounts to 48\% of total transport costs in Vessel, and 42\% in Air.
+Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8\% for ocean shipping and air transport respectively.
+This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period).
+Put differently, the additive component amounts to 48\% of total transport costs in vessel, and 42\% in Air.
 This shows that the per-kg cost dimension is quantitatively sizable.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.
-Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests, that confirm the importance of the additive cost component.
+For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 2.
+Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests that confirm the importance of the additive cost component.
 The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
@@
 
-Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question.
+Our estimation results on the size of additive transport costs can be compared to \citet{Irrazabal_2015}, the most recent evidence on this question.
 Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e.
@@
 One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
-From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect.
-Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport.
-This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
+From this, we can rebuild the ratio in similar terms to them, allowing us to gain in generality in this respect.
+Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping and 1.5\% in air transport.
+This may seem surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
 This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
@@
 \citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
-For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
-Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 47\% (in Vessel) of additive trade costs are attributable to international transport.
-On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
-
-\paragraph{Time dynamics} Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, i.e. structural phenomenon.
-We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
-The corresponding values reported in detailed tables are available on the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
-Precisely, each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
+For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
+Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.
+On top of our results summarized in Table \ref{tab:summary_results}, this points to evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
+
+\paragraph{Time dynamics} Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, i.e. a structural phenomenon.
+We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
+The corresponding values reported in detailed tables are available in the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
+Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
 
@@
 \begin{center}
-\includegraphics[width=8cm, height=5cm]{part_cout_additif.pdf}
+\includegraphics[width=8cm, height=5cm]{Figure1_share_of_additive_in_totalTC.pdf}
 %{\footnotesize {OECD data} }
@@
 Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-By contrast, they represent a sizable and structural dimension of international transport costs.
-If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can further notice that it substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
-In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs time trends.
-
-\section{Transport costs time trends: The role of the additive component}\label{sec:results_trends}
+On the contrary, they represent a sizable and structural dimension of international transport costs.
+If the share of additive costs in total costs roughly amounts to 48\% (42\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
+In the next section, we further investigate the importance of the (varying pattern of) additive costs, as regards the determinants of transport costs time trends.
+
+\section{Transport costs time trends: the role of the additive component}\label{sec:results_trends}
 
 In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
-This focus is shared with part of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
+This focus is shared with some of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
 Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
@@
 
-As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the fas price,\footnote{Precisely, we use the yearly values of the estimated transport costs components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
+As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the FAS price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
 Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2013.
@@
 \begin{figure}[htbp]
-\caption{Trend in estimated transport costs (Yearly mean value)}
+\caption{Trend in estimated transport costs (yearly mean value)}
 \label{fig:Trends_in_TC}
 \begin{center}
-\includegraphics[width=14cm, height=6.5cm]{Figure1_Trend_of_totalTC_bymode.pdf}
+\includegraphics[width=14cm, height=6.5cm]{Figure2_Trend_of_totalTC_bymode.pdf}
 \end{center}
@@
 
-As reported in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport.
+\textbf{JH: ATTENTION!!! TYPO à changer dans la Figure 2 ? STILL HERE ? }
+
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
-In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock related transport costs in 1974.
+In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974.
 However, computing the time trends from 1980 does not dramatically change the picture.
 The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996.
-For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \citet{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
+For the same years, we obtain a total decrease from 6.9\% to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
 Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip
 
-Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
+Before making any definite statement, however, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
 Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+As emphasized by \citet{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
 Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
@@
 \paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
-He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per-kg shipped (for a given product $k$ imported from country $i$ in year $t$).
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes down as:
+He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
+Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
 
@@
 Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
-Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}.
+Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
 In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
@@
 \noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-We can henceforth rebuild the value of the elasticity on a per product-year-country basis.
+We can then rebuild the value of the elasticity on a per product-year-country basis.
 The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
@@
 
-As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending of the transport mode, consistently with our previous findings.
-This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would rather imply a distribution concentrated on a single point.
+As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
+This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
 Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
-These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
+These results are not reported for sake of space, but are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
 
@@
 
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
-We start with the presentation our estimation strategy before turning to the results.\smallskip
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
+We start with the presentation of our estimation strategy before turning to the results.\smallskip
 
@@
 
-Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
+The estimation carried out in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
 Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
-
-One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
+We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+
+One advantage of this method is that it yields measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
+For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
 We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).
@@
 
-We start describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
+We begin by describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
 \smallskip
 
-\paragraph{Obtaining the ``pure'' transport costs component series} The empirical strategy to get the fitted transport cost measure (i.e., extracting from trade composition effect) can be described as a two-stage process.
+\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e.
+extracting from trade composition effect) can be described as a two-stage process.
 First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
@@
 
-\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).
-Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
+\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
+the series of $\gamma_t$ differs from one estimation to the other).
+Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
 
@@
 As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
-Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
+Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
 Both equations are estimated by transport mode.\smallskip
 
-In this exercise, we are interested in isolating the change in the time dimension of the each transport cost component.
+In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
 This constitutes the second stage of our procedure.
-As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we built the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
+As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we build the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
 \begin{equation}
@@
 
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (See details in Appendix \ref{app:comp-effects}).
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (see details in Appendix \ref{app:comp-effects}).
 In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
 
-As for the additive cost defined in $[0;+\infty]$, we built the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
+As for the additive cost defined in $[0;+\infty]$, we build the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
 \begin{equation}
@@
 
-\paragraph{Obtaining the unfitted measures} The objective is to get the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
+\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
 To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
@@
 
-We also build two measures of the ``total'' transport costs, that agglomerates our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e., composition effects excluded).
+We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
+composition effects excluded).
 As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
-Even if obeying to the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
+Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
 \smallskip
 
-\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
+\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
 $$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
 \noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
-Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed in percentage of the fas price.
-We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
+Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the FAS price.
+We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
 $$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
 
-\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to construct the fitted measure of total transport cost (i.e., composition effect excluded), with one notable exception though: Fitted ad-valorem and additive components have not been estimated in value, but extracted and build as indices.
-In a first step then, we re-build the fitted measures of each transport cost component in value starting from these indices.
-For the ad-valorem component, this can be obtained making use of Equation (\ref{eq:tcadv_compoeffect}), rewritten to get:
+\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
+composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
+In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
+For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
 
@@
 
-Last, we transform this (fitted) measure of total transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
+Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
 $$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
@@
 \subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974).
-Panels (d), (e) and (f) report the results for Vessel.
+Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
+Panels (d), (e) and (f) report the results for vessel.
 In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
@@
 International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
-This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}).
+This stands in line with the related literature (see \citealp{Hummels_1999}; \citealp{Lafourcade_Thisse}).
 Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
@@
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for Air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
 For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
-Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the transport costs ``per se''.
-Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
-Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: both trends are very similar.
-Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+Air transport costs were reduced by 50\% between 1974 and 2013, which is mainly attributable to a reduction in the transport costs ``per se''.
+Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
+Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
+Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
 This is particularly the case for the multiplicative component (panel (d)).\smallskip
@@
 
-The last results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
+The last results stand in sharp contrast with \citet{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
 If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
 This drives us to investigate this difference further.
-As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
+As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
+Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual CIF-FAS price gap).
 Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
 %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
+This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
 
@@
 composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for Air) and (b) (for Vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for Air and Vessel respectively.
+In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \citeh{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
 
 \begin{figure}[htbp]
-\caption{Characterizing the time trends: Applying Hummel's (2007) method }
+\caption{Characterizing the time trends: applying Hummel's (2007) method }
 \label{fig:comp_effects_as_in_Hummels}
@@
 \begin{tabular}{cc}
-{\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
+{\small (a) Ad-valorem air Freight} & {\small (b) Ad-valorem vessel Freight}\\
 \multicolumn{2}{c}{{\small In percent of the export price}} \\
@@
 & \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.pdf} \\
-{\small (c) Ad-valorem Air Freight} & {\small (d) Ad-valorem Vessel Freight}\\
+{\small (c) Ad-valorem air Freight} & {\small (d) Ad-valorem vessel Freight}\\
 \multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for Vessel (Panel (d)).
-Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).\medskip
+Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
+Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
 
 Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in air transport.
-Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction in international transport costs observed over time, in particular in air transport.
+Complementing the findings of Section \ref{sec:results_decomposition}, these results highlight the importance of integrating the additive dimension of international transport costs, with the aim here of characterizing their time trends.
 
@@
 \subsection{Robustness to the separability assumption}
-In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, that considers (following so \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
-That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to rather model $t_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
-Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approximatively 200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
+In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, which considers (following \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
+That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $t_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
+Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approx.
+200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
 This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.
 For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-Table \ref{tab:robustness_separability} reports a summary of the results, considering the average estimated values over the period (by transport mode).
+Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
 
@@
   \centering
-  \caption{Robustness to the separability assumption (Average over the period) - restricted sample}
+  \caption{Robustness to the separability assumption (average over the period) - restricted sample}
 \begin{center}
@@
 \parbox[l]{10cm}{\tiny{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-The additive term is expressed in fraction of fas price.
+The additive term is expressed in fraction of FAS price.
 ($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52 and 67\% of total costs on average, versus 40 to 50\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52\% and 67\% of total costs on average, versus 40\% to 50\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
-Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand side variables) as the AIC criteria displays very close values in both models (conditional on transport mode).
-Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period - the picture is reversed for medians.
-
-Before jumping to the conclusion that our results are robust to the separability assumption though, we investigate this comparison deeper on a year-to-year basis, as reported in Figure \ref{fig:robustesse_non_separe}.
-Precisely, Figure \ref{fig:robustesse_non_separe} displays the results for Air transport in panel (a) and for maritime transport in panel (b).
+Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
+Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+The picture is reversed for medians.
+
+Before jumping to the conclusion that our results are robust to the separability assumption, however we investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
+Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (a) and for maritime transport in panel (b).
 In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
@@
 \begin{figure}[htbp]
-\caption{Robustness to the separability assumption (Year-to-year basis)}
+\caption{Robustness to the separability assumption (year-to-year basis)}
 \label{fig:robustesse_non_separe}
@@
 
-In line with Table \ref{tab:robustness_separability}, the results reported in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
+In line with Table \ref{tab:robustness_separability}, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
 However, they also show that the difference is quantitatively small.
-Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar to the one another, whether they are estimated under the separability assumption or not.
-Along with the quality of fit diagnosis, this confirms the robustness of our estimation results to this assumption.
+Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
+Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.
 
 As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-To this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
-Figure \ref{fig:histogram_beta_robustness} reports the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for Air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a, (c))) and the non-separability scenario (Panels (b), (c)).
+With this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
+Figure \ref{fig:histogram_beta_robustness} shows the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
 
@@
 
-Similarly as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+Similarly, as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
 Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, that establish that the decline of ``pure'' transport costs lies as main reason for the decrease of overall transport costs observed over time.
+As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.
 
@@
 
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and \citet{hummels2007}'s dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \citet{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
 In this section, we provide a robustness test to this difference of treatment.\smallskip
 
-When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
-In this section, we check that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.
-
-To this aim, we rebuild the fitted values for each of of our transport cost measures (ad-valorem, additive and overall) applying \citet{hummels2007}'s weighting scheme (i.e., taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
+In this section, we check whether this difference of treatment is responsible or not for the difference in results relative to the role of the trade composition effects, which we rather attribute to the modeling of the additive component.
+
+To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \citet{hummels2007} weighting scheme (i.e.
+taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
 The results are reported in Figure \ref{fig:compeffects_robustness}.
@@
 \begin{figure}[htbp]
-\caption{Transport cost time trends: Robustness to the weighting scheme}
+\caption{Transport cost time trends: robustness to the weighting scheme}
 \label{fig:compeffects_robustness}
@@
 
-The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) drives two comments.
-First, differences do show up, in particular for the multiplicative component of Air transport.
+The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
+First, differences do show up, in particular for the multiplicative component of air transport.
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
 This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
@@
 This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
+Using SITC 3 and 4- digit CIF-FAS unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
 Two main findings emerge from our results.
-First, we provide a quantitative measure of both the additive and the ad-valorem transport cost.
-We find that additive costs are quantitatively sizable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
-This is slightly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
+First, we provide a quantitative measure of both the additive and the ad-valorem transport costs.
+We find that additive costs are quantitatively sizable: They amount to 2.8\% of the export price unit values for ocean shipping, and 1.8\% for air transport.
+This is slightly lower that the estimated values for the ad-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
 Further, modeling the additive component significantly improves the quality of fit of the regression.
 Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a varying share of additive costs in product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
-If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs in maritime transport.
-In both aspects, our results point the importance of the additive component in accounting for international transport costs.
+Allowing for a varying share of additive costs in the product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
+If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in maritime transport.
+In both aspects, our results highlight the importance of the additive component in accounting for international transport costs.
 
 Our results could be extended in two main ways.
-On the empirical side, one may want to go deeper in the structural determinants of (pure) transport costs, i.e.
+On the empirical side, one may want to go deeper into the structural determinants of (pure) transport costs - i.e.
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
@@
 The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller.
-In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
-More information on this database is available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}.
+In this respect, the ``custom value'' corresponds to the FAS price (``free-alongside'' price) delivered by the seller.
+More information is available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}.
 
@@
 import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States.
-In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States in Canada or Mexico to the first port of entry.
-
-The cif (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
+In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States in Canada or Mexico) to the first port of entry.
+
+The CIF (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
 It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S.
@@
 In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
-Precisely, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
-Table \ref{tab:result_air_3d_detail} reports the results for Air transport.
-The results for Vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.
+Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
+Table \ref{tab:result_air_3d_detail} reports the results for air transport.
+The results for vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.
 
@@
 \begin{table}[htbp]
-  \caption{Air: Transport costs estimates, 3-digits (selected years)}
+  \caption{Air: Transport costs estimates, 3-digit level (selected years)}
 \begin{center}
@@
 Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
+Additive term expressed in fraction of FAS price.}}}
 \end{tabular}%
@@
   \centering
-  \caption{Vessel: Transport costs estimates, 3 digit (selected years)}
+  \caption{Vessel: Transport costs estimates, 3-digit level (selected years)}
 \begin{center}
@@
 
-In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fas prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs.
+In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed CIF-FAS prices gap, in order to deliver a more systematic diagnosis of the importance of additive costs.
 To do so, we rely on several standard measures of fit.
@@
 However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.
-This drives us to complement the goodness of fit diagnosis with three alternative measures.
+This leads us to complement the goodness-of-fit diagnosis with three alternative measures.
 We provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
@@
 We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test.
-A decrease in the log-likelihood function points to a better quality-of-fit.
+A decrease in the log-likelihood function points to a better quality of fit.
 However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.
-\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.}
+\footnote{Specifically, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.}
 The preferred model is the one with the minimum AIC value.
@@
 The null tested is that the two models are statistically equivalent.
-Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.
+Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for air and vessel respectively, at the 3-digit level.
 
@@
     LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
-    nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
+    Nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
     p-value & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
@@
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.
-The mean statistics calculated as the average value over all years.
+For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.
+The mean statistics are calculated as the average value over all years.
 }}
@@
 Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
-On average over the whole period, the $R^{2}$ doubles when per-kg costs are included for Air, and increases by 50\% for Vessel.
+On average over the whole period, the $R^{2}$ doubles when per-kg costs are included for air, and increases by 50\% for vessel.
 Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
 Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes.
-Both AIC and Log-Likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
+Both AIC and log-likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
 These results holds whatever the considered year.\smallskip
 
-For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in in Appendix \ref{app:4digit}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
+For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Appendix \ref{app:4digit}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
 If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.
@@
 
-In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
-Precisely, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$), \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
+In this section, we provide a variance decomposition exercise on the observed CIF-FAS price gap.
+Specifically, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$) and \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
 The variance decomposition expression at the product level $k$ (5-digit level) is obtained by applying the following formula (by year and transport mode):
@@
 
-with $x$ the observed cif-fas price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
+with $x$ the observed CIF-FAS price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
 
@@
 \begin{figure}[htbp]
-\caption{Variance decomposition (observed cif-fas price gap)}
+\caption{Variance decomposition (observed CIF-FAS price gap)}
 \label{fig:decomp_variance}
@@
 
+\textbf{ATTENTION ERREUR SUR LES FIGURES il y a ecrit "secteur" et non "sector"}
+
 Two interesting results emerge from Figure \ref{fig:decomp_variance}.
-First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of same magnitude of order at the variance between sectors at the 3-digit level.
-Both account for between 30 and 40\% of the total variance in Air transport, depending on the years considered.
+First, the share of the CIF-FAS price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
+Both account for between 30 and 40\% of the total variance in air transport, depending on the years considered.
 This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
-This delivers an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
-Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
+This provides an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
+Second, the variance of the CIF-FAS price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
 This holds throughout the period and for both transport modes.
@@
 
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in Air and Ocean transport respectively.
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively.
 
@@
   \centering
-  \caption{Air: Transport costs estimates, Selected years, 4-digit}
+  \caption{Air: Transport costs estimates, selected years, 4-digit level}
 \begin{center}
@@
   \centering
-\caption{Vessel: Transport costs estimates, Selected years, 4-digit}
+\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
 \begin{center}
@@
 We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
-The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel).
+The results are reported in Tables \ref{tab:good_fit_air_rob} (for air) and \ref{tab:good_fit_ves_rob} (for vessel).
 \begin{table}[htbp]
   \centering
-  \caption{Air: Measures of Goodness-of-fit, 4-digits}
+  \caption{Air: Measures of goodness of fit, 4-digit level}
 \begin{center}
@@
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
+The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
 \end{tabular}%
@@
   \centering
-  \caption{Vessel: Measures of Goodness-of-fit, 4-digits}
+  \caption{Vessel: Measures of goodness of fit, 4-digit level}
 \begin{center}
@@
 
-\section{Eliminating the composition effects: More details \label{app:comp-effects}}
-
-In this section, we explain in more details the method employed to eliminate the country- and product- dimensions of the estimated transport cost.
+\section{Eliminating the composition effects: more details \label{app:comp-effects}}
+
+In this section, we explain in more detail the method employed to eliminate the country and product dimensions of the estimated transport cost.
 
@@
 
-\subsubsection{Additive and ad-valorem transport costs: Excluding the composition effects}
-
-In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for each the ad-valorem and the additive component.
-
-\paragraph{For the ad-valorem component} Consider first the multiplicative transport cost component.
+\subsubsection{Additive and ad-valorem transport costs: excluding the composition effects}
+
+In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for both the ad-valorem and the additive component.
+
+\paragraph{For the ad-valorem component} - Consider first the multiplicative transport cost component.
 Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
@@
 
-\paragraph{For the additive component} After estimating Equation (\ref{eq:compeffects_add}), we can re-build the additive component according to:
+\paragraph{For the additive component} - After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
 \[\left\{
@@
 
-\noindent Note that it is independent of the product-origin country pair, we can thus rewrite the time-trend series for the additive transport cost component as:
+\noindent Note that it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
 \begin{equation}
@@
 Consider now the ``fitted ad-valorem rate''.
-\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels' (2007) terminology) to $i$ (our terminology).} \textit{at time $t$.
+\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
 The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables}''.
@@
 
-After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighed average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
+After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
 
@@
 First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
-However, this is a slight difference, as in both cases it ultimately amounts taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-Second, in contrast to us, \cite{hummels2007} does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effects.
-This is not likely to make a substantial difference though, as these fixed effects are by nature constant over time.
-Accordingly, they only play as a scale effect as the estimate is in log.
-
-A third difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}$ in Equation (\ref{eq:Hummels1})) in two separate components (origin country and product dimensions).
+However, this is a slight difference, as in both cases it ultimately amounts to taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
+Second, in contrast to us, \cite{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
+This is not likely to make a substantial difference, however, as these fixed effects are by nature constant over time. Accordingly, they only amount to a scale effect.
+%Accordingly, \textbf{they only amount to a scale effect. - rather than \textit{they only play as a scale effect as the estimate is in log}. A-t-on besoin de dire que c'est en log?}.
+
+A third difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}$ in Equation (\ref{eq:Hummels1})) into two separate components (origin country and product dimensions).
 As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
-However, we do not view this as the major cause of difference between our two methods, as it can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
+However, we do not view this as the major cause of difference between our two methods, as can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
 
 Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeh{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
-\cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+\cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
@@
 First, this weighting scheme does not overweight the small flows in value.
-Second, considering the 1974 weighting scheme makes sense as starting point of our time trend analysis.
+Second, the 1974 weighting scheme makes sense as the starting point of our time trend analysis.
 However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in Section \ref{sec:robustness}.
 
-One last difference remains to be commented, which relies on the way the additive component of international transport costs is treated.
+One last difference remains to be commented, concerning the way the additive component of international transport costs is treated.
 As we show below, it turns out to be key in accounting for the difference of results with \cite{hummels2007}.
 Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
-To see it clearly, rewrite Equation (\ref{eq:Hummels1}) in a simpler form as:
+Rewriting Equation (\ref{eq:Hummels1}) in a simpler form as:
 
@@
 By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
-This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (ie, somehow allowing for a varying $\beta$ across time, country partner and product).
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (i.e.
+allowing for a varying $\beta$ across time, country partner and product).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007}, outlined above, on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.

### 2018-12-17 redaction/onlineA.tex
changements typos graphiques

--- 2018-09-27 redaction/online_appendix_travail.tex
@@
 \urlstyle{rm}
-\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
+\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
 Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
-Universit\'{e} Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE ; email: \url{lise.patureau@dauphine.fr}}}
+Universit\'{e} Paris-Dauphine, PSL University, LEDa, 75016 PARIS, FRANCE ; email: \url{lise.patureau@dauphine.fr}}}
 %\date{\vspace{-5ex}}
@@
 The paper compares the empirical performances of two models: one with only ad-valorem costs (Model (A)) and one with where both ad-valorem and additive costs (Model (B)).
-In the interest of comprehensiveness, this appendix looks at a model with only additive costs (Model (C)).
-The estimated equation is: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. \\
-We first compare estimates of the transport costs components under the three models. In a second step, we report quality of fit tests for the three models.
+In the interest of comprehensiveness,  we also estimate the model with only additive costs (Model (C)), in which case the estimated equation is:
+$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$$
+
+This section is devoted to presenting the results.
+Precisely, we first compare the estimates of the transport costs components under the three models.
+In a second step, we report quality of fit tests for the three models.
 
@@
 
-In this section, we report the estimation results of the three models.
-We also report the estimation results for Model (C), where only additive costs are modeled.
+In this section, we report the estimation results of the three models: Model (A) (with only ad-valorem costs), Model (B) (with both additive and ad-valorem) and Model (C) (with only additive costs).
 Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_vessel} for vessel transport.
@@
 
-Results for Models (A) (ad-valorem costs only) and (B) (both ad-valorem and additive costs) are identical to those reported in the paper.
-Somewhat unsurprisingly, the estimated size of transport costs under Model (C) is of same order of magnitude as in Models (A) and (B).
+Results for Models (A) and (B) are identical to those reported in the paper.
+Unsurprisingly, the estimated size of overall transport costs under Model (C) is of same order of magnitude as in Models (A) and (B).
 We also observe a downward trend of transport costs over time, in particular since 1980.\\
@@
     Stat LL ratio (B vs A) & \multicolumn{1}{c}{4809.7} & \multicolumn{1}{c}{5980.6} & \multicolumn{1}{c}{7257.3} & \multicolumn{1}{c}{9168.7} & \multicolumn{1}{c}{13548.7} & \multicolumn{1}{c}{13105.7} & \multicolumn{1}{c}{11760.4} & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{6541.2} \\
-    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
+    \textbackslash{}\# restrictions & \multicolumn{1}{c}{355} & \multicolumn{1}{c}{369} & \multicolumn{1}{c}{376} & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{428} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{422} \\
     p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline
     Stat LL ratio (B vs C) & \multicolumn{1}{c}{4948.4} & \multicolumn{1}{c}{4954.0} & \multicolumn{1}{c}{6564.5} & \multicolumn{1}{c}{8961.8} & \multicolumn{1}{c}{11597.4} & \multicolumn{1}{c}{12837.0} & \multicolumn{1}{c}{15764.7} & \multicolumn{1}{c}{17062.0} & \multicolumn{1}{c}{13896.9} \\
-    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
+    \textbackslash{}\# restrictions & \multicolumn{1}{c}{355} & \multicolumn{1}{c}{369} & \multicolumn{1}{c}{376} & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{428} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{422} \\
     p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline \hline
@@
 
-The model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered.
+For all years and whatever the transport mode considered, the model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered.
 That justifies our choice to disregard Model (C) in the main text.
@@
 
-As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, the ``raw'' transport costs and the unfitted transport costs, in the red dashed line plain have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel).
+As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``pure'' transport costs and the unfitted measure, in the red dashed have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)).
 However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
-Figure \ref{fig:totalTC_compeffects_excl_manuf} (for manufactures) reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
+In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf}) reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
 In air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
@@
 
-The situation is strikingly for primary goods only.
+The situation is strikingly different for primary goods only.
 In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)).

### 2018-12-20 redaction/TC_new.tex
Nouvelle version après relecture

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Holger Görg, S\'{e}bastien Jean and Vincent Vicard for their insightful comments.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Holger G\"{o}rg, S\'{e}bastien Jean and Vincent Vicard for their insightful comments.
 The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the universities of California-Irvine, Paris-Dauphine, Paris-I and Nancy and the Kiel Institute for the World Economy.
@@
 This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2013.
-First, we document the importance of the per-unit (additive) component of transport costs.
-We find that additive costs are quantitatively sizable, representing between one-third and one-half of overall transport costs.
-Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
-Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
-Importantly, this difference of results can be attributed to the new method of modeling the per-unit component of transport costs we offer.
-In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
+First, we document the importance of the per-unit component of transport costs.
+It represents between one-third and one-half of overall transport costs.
+Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in “pure” transport costs over time, with trade composition effects playing a minor role.
+In both aspects, our results point to the importance of the additive component in accounting for international transport costs
 \end{abstract}
@@
 This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
-Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.
+Also, their estimates imply that doubling freight costs increases average free alongside ship (FAS) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.
 Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.
@@
 \begin{center}
-\includegraphics[width=14cm, height=6.5cm]{Figure2_Trend_of_totalTC_bymode.pdf}
+\includegraphics[width=14cm, height=7cm]{Figure2_Trend_of_totalTC_bymode.pdf}
 \end{center}
 \end{figure}
-
-\textbf{JH: ATTENTION!!! TYPO à changer dans la Figure 2 ? STILL HERE ? }
 
@@
 \begin{tabular}{cc}
-{\small (a) Ad-valorem air Freight} & {\small (b) Ad-valorem vessel Freight}\\
+{\small (a) Ad-valorem air freight} & {\small (b) Ad-valorem vessel freight}\\
 \multicolumn{2}{c}{{\small In percent of the export price}} \\
@@
 & \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.pdf} \\
-{\small (c) Ad-valorem air Freight} & {\small (d) Ad-valorem vessel Freight}\\
+{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
 \multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
@@
 \begin{tabular}{cc}
-{\small (a) Air Freight} & {\small (b) Vessel Freight}\\
-\includegraphics[width=2.5in, height=2in]{graph_robustesse_ns_mp_terme_A_air.pdf}
-& \includegraphics[width=2.5in,height=2in]{graph_robustesse_ns_mp_terme_A_ves.pdf} \\
-\includegraphics[width=2.5in, height=2in]{graph_robustesse_ns_mp_terme_I_air.pdf}
-& \includegraphics[width=2.5in,height=2in]{graph_robustesse_ns_mp_terme_I_ves.pdf} \\
+{\small (a) Air freight} & {\small (b) Vessel freight}\\
+\includegraphics[width=3in, height=2in]{graph_robustesse_ns_mp_terme_A_air.pdf}
+& \includegraphics[width=3in,height=2in]{graph_robustesse_ns_mp_terme_A_ves.pdf} \\
+\includegraphics[width=3in, height=2in]{graph_robustesse_ns_mp_terme_I_air.pdf}
+& \includegraphics[width=3in,height=2in]{graph_robustesse_ns_mp_terme_I_ves.pdf} \\
 
@@
 \end{center}
+\begin{minipage} [c]  {5in} \scriptsize%
+Notes: FE for (country, sector) fixed effects.
+\end{minipage}
 \end{figure}
@@
 \hline
-\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &      \\
+\multicolumn{7}{l}{\textbf{R$^2$}}      \\ \hline
 Model (a)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
-\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
+Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
+\multicolumn{7}{l}{\textbf{SER} }  \\ \hline
 Model (a)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
 Model (b) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
 Model (a) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
-Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
+Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
+\multicolumn{7}{l}{\textbf{AIC criteria} }  \\ \hline
 Model (a)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
-Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\
-\textbf{Test LL} &       &       &       &       &       &       \\
+Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
+\multicolumn{7}{l}{\textbf{Test LL} }    \\ \hline
 2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
@@
       & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
-\boldmath{}\textbf{R$^{2}$}\unboldmath{} &       &       &       &       &       &         \\
+\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
 Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
+\multicolumn{7}{l}{\textbf{SER} }\\ \hline
 Model (a) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\
-\textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
+\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
 Model (a)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
-Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\
-\textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
+\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
 Model (a) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
-Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\
-\textbf{Test LL} &       &       & & &  & \\
+Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
+\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
 2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\

### 2019-01-08 redaction/TC_new.tex
Changement de graphique

@@
 This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2013.
-First, we document the importance of the per-unit component of transport costs.
-It represents between one-third and one-half of overall transport costs.
-Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in “pure” transport costs over time, with trade composition effects playing a minor role.
-In both aspects, our results point to the importance of the additive component in accounting for international transport costs
+First, we document the importance of the per-unit (additive) component of transport costs.
+We find that additive costs are quantitatively sizable, representing between one-third and one-half of overall transport costs.
+Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
+Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
+Importantly, this difference of results can be attributed to the new method of modeling the per-unit component of transport costs we offer.
+In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
 \end{abstract}

### 2019-01-08 redaction/submissionREStat_DHP.tex

--- 2018-12-20 redaction/TC_new.tex
@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[a4paper,12pt]{article}
 
@@
 
+\usepackage{setspace}
+
 %\usepackage[nolists, figuresfirst]{endfloat}
@@
 
+\renewcommand{\baselinestretch}{1.2}
+\doublespacing
+
+%\linespread{1.6}
+
+\newcommand{\note}[1]{\footnote{\begin{doublespace}#1\end{doublespace}}}
+
 \begin{document}
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
-Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
-; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-University Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.psl.eu} } }
+University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
 
@@
 Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in “pure” transport costs over time, with trade composition effects playing a minor role.
-In both aspects, our results point to the importance of the additive component in accounting for international transport costs
+In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
 \end{abstract}
@@
 
-\noindent \emph{JEL classification}: F14, N70, R40 \\
+\noindent \emph{JEL classification}: F10, N70, R40 \\
 \noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs, trade composition effects
@@
     \end{center}}
-\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
+\parbox[l]{12cm}{\footnotesize{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
 The additive term is expressed in fraction of FAS price.
@@
 \end{center}
-\parbox[l]{10cm}{\tiny{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-The additive term is expressed in fraction of FAS price.
-($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
+\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+The additive term is expressed in fraction of FAS price. 1989 omitted in 3-digit estimation for air.}}
 \label{tab:robustness_separability}
@@
 \hline\hline
-\multicolumn{8}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+\multicolumn{8}{l}{\parbox[l]{15cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
 Statistics are obtained weighting each observation by its share in trade (mode-dependent).
@@
   \label{tab:good_fit_air}%
- \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs.
+ \parbox[l]{15cm}{\footnotesize{Notes: Model (A) = with only ad-valorem transport costs.
 Model (B) = with additive \& ad-valorem transport costs.
@@
   \label{tab:good_fit_vessel}%
-  \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs.
+  \parbox[l]{15cm}{\footnotesize{Notes: Model (A) = with only ad-valorem transport costs.
 Model (B) = with additive \& ad-valorem transport costs.
@@
 \end{figure}
-
-\textbf{ATTENTION ERREUR SUR LES FIGURES il y a ecrit "secteur" et non "sector"}
 

### 2019-01-14 redaction/TC_new.tex
rangement dossier redaction + coquille intro
creation dossier "old" pour vieille version + correction reference manquante en intro (Lashkaripour 2018, on avait un beau "?" en gras au beau milieu dela page 4 ;-))

@@
 Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using transaction-level trade data from Colombia, \cite{Lashkaripour-2017b} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
+Using transaction-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
 \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.

### 2019-01-14 redaction/old/dhp_july2017.tex
rangement dossier redaction + coquille intro
creation dossier "old" pour vieille version + correction reference manquante en intro (Lashkaripour 2018, on avait un beau "?" en gras au beau milieu dela page 4 ;-))

--- 2018-06-25 redaction/dhp_july2017.tex
@@
 
-\date{June 2018}
+\date{April 2018}
  \maketitle
@@
 composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for Air) and (b) (for Vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for Air and Vessel respectively.
 
@@
 {\small (a) Ad-valorem Air Freight} & {\small (b) Ad-valorem Vessel Freight}\\
-\multicolumn{2}{c}{{\small In percent of the export price}} \\
-\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.pdf}
-& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.pdf} \\
-{\small (c) Ad-valorem Air Freight} & {\small (d) Ad-valorem Vessel Freight}\\
-\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
+\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels.pdf}
+& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels.pdf} \\
 \includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.pdf}
 & \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.pdf} \\
+
 \end{tabular}
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for Vessel (Panel (d)).
-Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).\medskip
+Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).
 
@@
 
-In line with Table \ref{tab:robustness_separability}, the results reported in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
+The results reported in Figure \ref{fig:robustesse_non_separe} confirm the findings of Table \ref{tab:robustness_separability}, that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
 However, they also show that the difference is quantitatively small; further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar to the one another, whether they are estimated under the separability assumption or not.
@@
 As last check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-To this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
-Figure \ref{fig:histogram_beta_robustness} reports the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for Air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a, (c))) and the non-separability scenario (Panels (b), (c)).
+To this aim, we replicate Figure \ref{fig:histogram_beta} on the sub-sample, in both specifications of separable (baseline) and non-separable fixed costs.
+Figure reports the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$ for Air transport under the baseline specification (Panel (a)) and the non-separability scenario (Panel (b)).
+Figure reports the hitsograms obtained for maritime transport.
 
 \begin{figure}[htbp]
-\caption{Robustness: Histogram of the share of additive costs in total costs}
-\label{fig:histogram_beta_robustness}
+\caption{Robustness: Histogram of the share of additive costs in total costs, Air}
+\label{fig:histogram_beta_robustness_air}
 \begin{center}
 \begin{tabular}{cc}
-\multicolumn{2}{c}{Air transport} \\
 {\small (a) Separability (baseline)} & {\small (b) Non-separability}\\
@@
 & \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_air.pdf} \\
-\multicolumn{2}{c}{Maritime transport} \\
-{\small (c) Separability (baseline)} & {\small (d) Non-separability}\\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Robustness: Histogram of the share of additive costs in total costs, Vessel}
+\label{fig:histogram_beta_robustness_vessel}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Separability (baseline)} & {\small (b) Non-separability}\\
 \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_ves.pdf}
@@
 
-Similarly as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distibution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
-Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, the countries of origin and products.
-As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, that establish that the decline of ``pure'' transport costs lies as main reason for the decrease of overall transport costs observed over time.
-
 \subsection{Robustness to the weighting scheme }
 
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and Hummels's (2007) dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and Hummels's (2007)dwells on the weighting scheme used to obtain the evolution of the pure transport costs over time.
 In this section, we provide a robustness test to this difference of treatment.\smallskip
@@
 This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
+Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components of transport costs, by transport mode (air or ocean).
 Our results may be summarized in two main findings.
 First, we provide a quantitative measure of both the additive and the ad-valorem transport cost.
-We find that additive costs are quantitativeley sizeable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport.
-This is slighly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
-Further, modeling the additive component significantly improves the quality of fit of the regression.
+We thus find that additive costs amount to 2.8\% of the export price unit values for ocean shipping, and ad-valorem ones 3.2\%, as mean values over 1974-2013.
+These values are respectively equal to 1.8 and 2.5\% for air transport.
 Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a varying share of additive costs in product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
+Allowing for a varying share of additive costs in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
 If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs, in particular in maritime transport.
@@
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows in an international trade theory perspective and in affecting the international transmission of business cycles.
+On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles.
 This is left for further research.

### 2019-01-14 redaction/old/idee.tex
rangement dossier redaction + coquille intro
creation dossier "old" pour vieille version + correction reference manquante en intro (Lashkaripour 2018, on avait un beau "?" en gras au beau milieu dela page 4 ;-))

\documentclass[11pt,english,pdftex]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[centertags]{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage[ansinew]{inputenc}
\usepackage{babel,varioref}
\usepackage{natbib}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2552}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{Created=Friday, December 31, 2004 13:48:08}
%TCIDATA{LastRevised=Friday, April 08, 2011 16:11:47}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{<META NAME="DocumentShell" CONTENT="Standard LaTeX\Blank - Standard LaTeX Article">}
%TCIDATA{Language=American English}
%TCIDATA{CSTFile=40 LaTeX article.cst}

\setlength{\textwidth}{16.5cm} \setlength{\textheight}{21.5cm}
\topmargin-1cm \oddsidemargin1pt \evensidemargin0pt
\baselineskip17pt

%%%%%%%%%%% Here are theorem_related declarations
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{defi}{Definition}
\newtheorem{theo}{Theorem}
\newtheorem{prop}{Proposition}
%\newtheorem{proof}{Proof}
\newtheorem{property}{Property}
\newtheorem{coro}{Corollary}
\newtheorem{propannex}{Proposition}

\newcounter{countAnnexA}
\setcounter{countAnnexA}{0}
\renewcommand{\propannex}{\addtocounter{countAnnexA}{1}
\noindent \large \textbf{Proposition \thesection.\thecountAnnexA}
\normalsize}

%\input{tcilatex}

\begin{document}

\section{Problématique}

Essayer de résumer notre problématique. Point de départ du papier: ``The distance puzzle''

What is it? Start from a standard gravity equation

$$\ln X_{ijt} = - \alpha_t \ln d_{ij} + F_i + F_j + F_t$$

where $X_{ijt}$ represents the aggregate trade flows from country $i$ to country $j$ (at time $t$), $d_{ijt}$ is the geographical distance between $i$ and $j$, and $F_i$, $F_j$ are country-fixed effects, and $F_t$ time fixed effects.

We expect $\alpha$ of positive sign, as you trade less with countries that are far away.

The ``puzzle'': $\alpha$ is relatively constant over time, and even slightly increasing over the recent decades (see the paper of Daudin and Elizaveta). This is striking in light of the increasing wave of globalization, that is expected to reduce trade costs, hence the impact of geography on trade. Hence the ``distance puzzle''.

Our ambition in this paper is to investigate this distance puzzle. To do so, we refer to the theoretical foundations of the gravity equation.

In fact, it is possible using theory, to decompose the distance coefficient $\alpha$, ie the elasticity of trade flows to distance, as the product of two parameters, let's say $\rho$, that measures the elasticity of iceberg transport costs to geographical distance ($\tau_{ij}=(d_{ij})^\rho$) and $\xi$, that measures the elasticity of trade flows to trade costs.

Depending on the underlying model, one can have an economic interpretation of $\xi$, that is referred to as the ``trade elasticity''.

\begin{itemize}
\item In the Armington model, the sensitivity of trade flows to trade costs is equal to $\sigma-1$, with $\sigma>1$ the elasticity of substitution between the domestic and foreign varieties. A $\sigma$ is high, goods are highly substitutable, then an increase in trade costs, strongly reduce trade flows between country $i$ to $j$, as household are ready to easily switch from foreign to home goods.

\item In the Ricardian model (Eaton and Kortum, 2002), the aggregate trade elasticity is equal to $\theta$, that measures the Fréchet distribution parameter that captures the degree of efficiency dispersion across sectors (within countries). Broadly speaking, $\theta$ is inversely related to the strength of the comparative advantage. A higher value of $\theta$ means that there is few heterogeneity across sectors, ie there is a low comparative advantage. A weak comparative advantage exerts a small force against trade barriers, and the elasticity of trade to trade costs is high. The higher $\theta$, the less heterogeneity across sectors, the higher the trade elasticity.

\item In the Chaney-Melitz model, the key explanation of the trade elasticity is given by $\gamma$, that is the Pareto distribution which captures the degree of efficiency dispersion across firms (within the same country, and aggregating over sectors). The intuition is straightforward. The higher $\gamma$, the more homogenous the firms' efficiency distribution. That is, the dispersion of firms is low, with a large number of low-productive firms. As a result, a reduction in trade costs changes the cut-off level above which exporting becomes profitable. The extensive margin of trade increases, all the more as $\gamma$ is high, ie that the fall in trade costs meets an area with a large number of firms. The higher $\gamma$, the lower the efficiency dispersion among firms (within sector),  the larger the response of trade flows to trade costs.
\end{itemize}

Summing up: the distance coefficient is in fact the product of $\rho$ and $\xi$ (whatever the theoretical meaning of $\xi$, depending on the underlying model).

How to account for the distance puzzle, ie the stability of $\alpha$ overt time (over the recent decades)?

\begin{itemize}
\item Document the evolution of $\alpha$ over years: See work of Daudin and Elizaveta
\item Two possible explanations:
\begin{itemize}
\item The evolution of $\rho$, ie the elasticity of trade costs to geographical distance
\item The evolution of $\xi$, ie the elasticity of trade flows to trade costs
\end{itemize}

\item So, to go further in the distance puzzle, to account for the stability of $\alpha$
\begin{itemize}
\item How has evolved $\rho$ over time? Intuitively, it has fallen (less weight of distance-related trade costs in overall (variable) trade barriers)
\item Provided that $\rho$ has decreased over time, stability of $\alpha$ is accounted for by the increase in $\xi$

\end{itemize}
\item Our ambition: to document this, relying on French export data on the empirical side, and on the Mélitz-Chaney model on the theoretical side. If $\xi$ (ie $\gamma$ in terms of Chaney-Mélitz model) has increased over time, it means that the efficiency dispersion of firms has decreased over time (more homogenous firms), inducing more firms to export along the extensive margin of trade for a given reduction in trade costs.

\item In this case then, it means that ultimately, the change of pattern of trade flows is a question of changing productivities (ie, more homogeneous firms).
\end{itemize}

Question, what would be the story in the Ricardian model? The lower the strength of comparative advantage, the larger the trade elasticity. That is, if $\theta$ increases over time, it is because sectors (within countries) become more similar. The result, that comparative advantage has become weaker over time, has been documented by Levchenko and Zhang (2011). In comparison with Levchenko and Zhang (2011), we pursue an alternative (but somehow parallel) road, as we base our investigation on the role of heterogeneity within sectors (rather than between sectors).\bigskip

Point important: Dans le modèle de Chaney, l'élasticité des trade flows aux coûts du commerce est en fait double, il y a l'élasticité aux coûts fixes et l'élasticité aux coûts variables. Ceci repose sur l'hypothèse que le coût fixe à l'exportation $f_{ij}$ dépend de la distance $d_{ij}$. Dans ce cas, l'élasticité des trade flows à la distance ($\alpha$) dépend \textbf{et} de $\gamma$ le degré de dispersion dans la productivité des firmes, \textbf{et} de $\sigma$, qui est l'élasticité de substitution entre variétés de biens différenciées. De plus, on a une élasticité des coûts de commerce à la distance $\rho$, qui est spécifique à chaque type de coût (fixe ou variable).

Nous: éliminer cette dimension en supposant que le coût fixe d'entrée n'est pas fonction de la distance géographique. A ce moment là, le coefficient $\alpha$ estimé sur données agrégées est bien le produit de l'élasticité des flux du commerce aux coûts de commerce variables, et dans le modèle de Chaney correspond au degré de dispersion dans l'efficacité des firmes ($\gamma$), et de l'élasticité des variable trade costs à la distance géographique.

\subsection{Etape 1- Characterizing the evolution of $\rho$}

\begin{itemize}
\item Par COMTRADE, on a les valeurs unitaires du commerce de 1962 à 2009 (2010?), à un niveau de désagrégation sectoriel de 4 digit
\item Par la base CEPII, on a les valeurs unitaires du commerce sur un niveau de désagrégation sectoriel de 6 digit. Mais laissons tomber cette base, au moins pour l'instant
\item On a la distance géographique via la base distance du CEPII
\item On veut estimer l'évolution de $\rho$, ie:

$$\ln\tau_{ijt} = \rho_t \ln d_{ij} + F_i + F_j $$
\item Mais attention, on veut cette estimation au niveau agrégé
\item On peut estimer les trade costs via les valeurs unitaires
\item Mais, donc on aura les trade costs au niveau sectoriel
\item Question: comment agréger les trade costs estimés au niveau sectoriel, au niveau agrégé?
\item[$\Rightarrow$] Réponse: pas besoin d'agréger les trade costs
\end{itemize}

On part donc de la base COMTRADE, sur 1962-2010. On a une estimation des coûts de transport au niveau sectoriel (4-digit), à partir du ratio  des unit values:

$$TC_{ijkt} = \frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}}$$
avec $p^{cif}$ le prix à l'importation, $p^{fob}$ le prix à l'exportation, $i$ le pays exportateur, $j$ le pays importateur, $k$ le secteur et $t$ l'année.

\begin{itemize}
\item On estime l'effet de la distance, selon l'équation
    $$\ln TC_{ijkt} = \rho_t \ln d_{ij} + F_i+ F_j+ X_{ij}+ F_k$$
    avec $X_{ij}$ un ensemble de variables de contrôle spécifiques à la paire ($i,j$) (colonisation, langue, frontière commune).
\item Comme on veut l'évolution temporelle de $\rho$, on estime cette équation année par année
\item[$\Rightarrow$] Soit, 48 points (pour 1962-2009)
\item[] Remarque: Si on part sur 500 secteurs (en 4 digit), supposons que 100 pays exportateurs, 100 pays importateurs, on a alors (par année), 500 000 observations, avec 700 effets fixes
\item[$\Rightarrow$] On obtient alors $\widehat{\rho}_t$

\item Question: cela fait une grosse base, et se pose le pb de la cohérence avec nos données de firme ensuite. Restreindre notre estimation (à ce stade) pour la rendre comparable à l'étape sur données de firmes? Ie
    \begin{itemize}
    \item[-] Ne considérer que la France comme pays exportateur ($i$)
    \item[-] Ne considérer comme pays importateurs que ceux vers laquelle les données de douane françaises nous disent que la France exporte
    \item[-] Et se restreindre à la période 1993-2007 (couverture des données de douane)?
    \end{itemize}
\item Réponse: Ne pas forcément se restreindre a priori au moment de la cosntrauction de la base. Partir large, on pourra toujours vérifier la robustesse de nos résultats sur la base restreintt (France comme exportateur, sur 1993-1997).

\item Une fois qu'on a obtenu $\rho_t$ au niveau agrégé, documenter son évolution temporelle, selon l'équation

$$\rho_t = \beta t + u_t$$

\begin{itemize}
\item Si $\beta$ est significativement positif: élasticité des trade costs à la distance augmente dans le temps, contribue à expliquer le distance puzzle (a priori, pas ce qu'on attend)
\item Si $\beta$ est non-significativement différent de 0, signifie que l'élasticité des trade costs à la distance est constante au cours du temps (ce qu'on aimerait avoir)
\item Si $\beta$ est significativement négatif, élasticité des trade costs à la distance diminue au cours du temps, amplfie le puzzle (encore mieux qu'avant)
\end{itemize}

\end{itemize}

Notre intuition: On s'attend à ce que $\rho$ décroisse au cours du temps (l'élasticité des trade costs à la distance diminue). Dans ce cas, outre apporter une caractérisation empirique de cette élasticité, les résultats de cette étape indiquent qu'il fait aller plus loin pour expliquer le ``distance puzzle''.\bigskip

Question sur cette section: quelle est la littérature qui a estimé cette élasticité? Sur quelles données? Avec quels résultats? A priori peu de papiers mais creuser la question.

\subsection{Etape 2 -- Into the trade costs black box}

S'inscrire dans le débat: Samuelson (1954) vs Alchian and Allen (1964). Ie, les trades costs sont-ils de type iceberg, ou incluent-ils également une composante additive.

Pour résumer, la spécification la plus générale (celle à la Alchian and Allen (1964)) implique la relation suivante entre prix $fob$ et $cif$:
$$p^{cif}_{ijkt} = \tau_{ijkt}~p^{fob}_{ijkt} + T_{ijkt}$$

Si les trade costs sont uniquement de type iceberg, alors $T_{ijkt}=0$, et le ratio des valeurs unitaires $cif-fob$ est indépendant du prix. En revanche, en présence de trade costs additifs, on obtient qu'il décroit avec le prix de vente à l'epxort, selon:
\begin{equation}
\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}} = \tau_{ijkt} + \frac{T_{ijkt}}{p^{fob}_{ijkt}} \label{eq:tradecostsEtape2}
\end{equation}

Dans cette étape, notre ambition est de creuser la spécification des trade costs. Creuser les différents papiers sur cette problématique.

De plus, cette analyse nous permet de tester la robustesse de nos résultats de l'étape 1. Est-ce que notre caractérisation de l'évolution de $\rho$ (l'léasticité des coûts de transport à la distance géographique), est robuste à la spécification des trade costs?

On dispose comme point de départ du ratio des unit values ($\frac{p^{cif}_{ijkt}}p^{fob}_{ijkt}$). Difficultés techniques pour évaluer la pertinence empirique de trade costs additifs (ie, $T_{ijkt} \neq 0$): ça rend les choses additives, donc passage en log (donc, à la forme linéaire) problématique.

\subsubsection{Hypothèse 1}

On fait l'hypothèse que les trade costs iceberg et additifs sont tous les deux fonction de la distance géographique entre $i$ et $j$, ainsi que d'un certain nb de variables spécifiques au pays $i$, au pays $j$ et au secteur $k$, que l'on capture par des effets fixes. De plus, on fait \textbf{ici} l'hypothèse que la forme fonctionnelle est la même, soit:
\begin{eqnarray*}
\tau_{ijkt} &=&f(dist_{ij}, F_i,F_j,F_k) = dist_{ij}^{\rho_t}e^{F_i} e^{F_j}e^{F_k} \\
T_{ijkt} &=&f(dist_{ij}, F_i,F_j,F_k) =dist_{ij}^{\rho_t}e^{F_i} e^{F_j}e^{F_k}
\end{eqnarray*}

Soit finalement, que $\tau_{ijkt} = T_{ijkt}$

Donc, le ratio des unit values (Equation (\ref{eq:tradecostsEtape2}) s'écrit:
\begin{eqnarray*}
&\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}} &= dist_{ij}^{\rho_t}\times e^{F_i}\times e^{F_j}\times e^{F_k}\times \left(1+\frac{1}{p^{fob}_{ijkt}}\right) \\
\Leftrightarrow &\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}} & = dist_{ij}^{\rho_t}\times e^{F_i}\times e^{F_j}\times e^{F_k} \times \left(\frac{p^{fob}_{ijkt}+1}{p^{fob}_{ijkt}}\right)
\end{eqnarray*}

On peut alors considérer le log de cette expression, et l'estimer (année par année), selon
\begin{equation}
\ln\left(\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}}   \right)= \rho_t \ln dist_{ij}+ F_i + F_j + F_k +\ln\left(\frac{p^{fob}_{ijkt}+1}{p^{fob}_{ijkt}} \right) \label{eq:ttt1}
\end{equation}

Plus précisément, cela nous amène à réécrire l'équation estimée en modifiant le membre de gauche (variable expliquée):
\begin{equation}
\ln\left(\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}}   \right)- \ln\left(\frac{p^{fob}_{ijkt}+1}{p^{fob}_{ijkt}} \right)= \rho_t \ln dist_{ij}+ F_i + F_j + F_k   \label{eq:ttt2}
\end{equation}

Dont on déduit la série temporelle de $\rho$.

Pour évaluer la pertinence empirique des coûts de transport additifs, on procède de manière similaire mais \textbf{sans} inclure de coûts de transport additifs, soit en estimant l'équations:

\begin{equation}
\ln\left(\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}}   \right)= \rho_t \ln dist_{ij}+ F_i + F_j + F_k \label{eq:tradecostsICB}
\end{equation}

La comparaison des résultats des deux spécifications nous renseigne sur la valeur des coûts de trasnport additifs (la composante $\ln\left(\frac{p^{fob}_{ijkt}+1}{p^{fob}_{ijkt}} \right)$ étant observation-spécifique)
\begin{itemize}
\item La première régression fait-elle mieux (en termes de variance expliquée) que la seconde? Quel est le meilleur fit? Est-ce que le fait d'ajouter ce terme $\ln\left(\frac{p^{fob}_{ijkt}+1}{p^{fob}_{ijkt}} \right)$ ajoute-t-elle du bruit, ou de l'information?
\item Autre avantage: Selon la spécification qui s'impose, on peut aussi tester la robustesse de nos résultats de l'étape 1, quant à l'évolution temporelle de $\rho$.
\item Limite: on fait l'hypothèse que les trade costs iceberg et additifs sont liés à la distance et aux effets fixes selon la même forme fonctionnelle
\end{itemize}

Mais questions (naïves)
\begin{itemize}
\item Si conceptuellement les équations (\ref{eq:ttt1}) et (\ref{eq:ttt2}) sont identiques, 
\item Peut-on vraiment comparer les deux équations (\ref{eq:ttt2}) et (\ref{eq:tradecostsICB})? Vu qu'on ne considère pas la même variable estimée (puisque $\ln\left(\frac{p^{fob}_{ijkt}+1}{p^{fob}_{ijkt}}\right)$ est observation-spécifique, ce n'est donc pas une constante)?
\end{itemize}

\subsubsection{Hypothèse 2}

On relâche alors l'hypothèse d'une même forme fonctionnelle des trade costs. Plus précisément, on suppose qu'ils sont tjs fonction de la distance et d'effets fixes pays importateurs/exportateur/année, mais avec des coefficients spécifiques, soit:
\begin{eqnarray*}
\tau_{ijkt} &=&f^I(dist_{ij}, F^I_i,F^I_j,F^I_k) = (dist_{ij})^{\rho^I_t}\times F^I_i\times F^I_j\times F^I_k \\
T_{ijkt} &=&f^A(dist_{ij}, F^A_i,F^A_j,F^A_k) =(dist_{ij})^{\rho^A_t}\times F^A_i\times F^A_j\times F^A_k
\end{eqnarray*}

Alors, on ne peut plus factoriser, et l'équation d'intérêt s'écrit sous forme non-linéaire:
\begin{eqnarray}
&\frac{p^{cif}_{ijkt}}{p^{fob}_{ijkt}}= (dist_{ij})^{\rho^I_t}\times e^{F^I_i}\times e^{F^I_j}\times e^{F^I_k}+ \frac{dist_{ij}^{\rho^A_t}\times e^{F^A_i}\times e^{F^A_j}\times e^{F^A_k}}{p^{fob}_{ijkt}} \label{eq:tradecostsNL}
\end{eqnarray}

Avantages de cette méthode
\begin{itemize}
\item On ne se contraint pas à la même forme fonctionnelle sur les trade costs icerberg et additifs
\item En particulier, on autorise une élasticité différente des trade costs à la distance ($\rho^A$ et $\rho^I$), et on produit une estimation différenciée de leur évolution temporelle
\end{itemize}
Questions
\begin{itemize}
\item La toolbox ``nl'' sous stata est-elle convaincante
\item Peut-on estimer en non-linéaire avec des effets fixes? Sous Stata 12?
\end{itemize}

Si on procède ainsi, on peut là aussi évaluer l'importance des coûts de transport addififs, en comparant les résultats de l'estimation de l'équation (\ref{eq:tradecostsICB}) (iceberg) avec ceux de l'équation (\ref{eq:tradecostsNL}) (attention, l'une est en log, l'autre non). Pour voir laquelle a le meilleur fit.

\subsection{Etape 3 -- Caractériser l'évolution de la trade elasticity}

Résumé: On part du ``distance puzzle'', on veut l'expliquer par le fait que, si l'élasticité des trade flows à la distance est constante, voire croissante au cours du temps, c'est parce que l'élasticité des trade flows aux coûts (variables) du commerce elle-même augmente au cours du temps.

On a montré (!) que l'évolution de l'élasticité des couts du commerce à la distance est constante au cours du temps, ou décroissante. Rest donc à explorer l'évolution de $\xi$ = élasticité des trade flows aux trade costs.

On se base pour cela sur le modèle de Chaney-Mélitz, que l'on teste à partir des données de douanes francçaises, que l'on combine aux données EAE.\medskip

Mais, la modélisation adoptée en cette troisième étape, dépend des résultats de l'étape précédente. Si l'étape 2 met en évidence l'existence (au niveau mondial) de côûts de transport additifs ($T_{ijkt} \neq 0$), alors il est nécessaire d'intégrer cette spécification dans le modèle de l'étape 3.

\subsubsection{Sur le plan des données}

\begin{itemize}
\item Données de douane, sur 1993-2010
\begin{itemize}
\item[-] Structure des données de douane (par ordre des colonnes): année / siren de la firme / produit / pays (d'exportation) / valeur en euros / masse en kg / unité supplémentaire
\item[-] Détails,
\begin{itemize}
\item[$\star$] la masse, toujours exprimée en kilos, est obligatoire quelque soit la nomenclature de produit (sauf cas trés particulier, et depuis 2006 en flux intracommunautaire quand les unités supplémentaires sont servies)
\item[$\star$] en plus de la masse, l'opérateur doit aussi fournir 1 chiffre correspondant à une unité supplémentaire (USUP) pour certains produits.
\item[$\star$] les nomenclatures de produits avec USUP obligatoire sont nombreuses :voir le taleau joint NC2012-FR-rev1 qui reprend toutes les nomenclatures en indiquant s'il y a une USUP ou aller sur le site Pro.douane.fr : Deb sous Pro.douane, Documentation, les nomenclatures NC8, liste récapitulative
\item[$\star$] les USUP sont diverses : voir liste des unités supplémentaires jointe
\item[$\star$] Pour résumer, Unités supplémentaires: Cette rubrique n'est fournie que pour autant que la nomenclature du produit l'exige. Ce sont les unités de mesure autres que le kilogramme. Il peut s'agir par exemple: de paires (chaussures de sport NC 64 04 11 00), de litres (eaux minérales NC 22 02 10 00), de grammes (perles de culture NC 71 01 22 00), de $m^2$ (tissus NC 58 03 10 10), etc.
\end{itemize}
\end{itemize}
\item Données EAE: sur 1993-2007
\begin{itemize}
\item[-] Vont permettre de calculer la productivité de la firme (rapport valeur ajoutée/emploi), et de croiser ça avec les données de douane (via siren)
\end{itemize}
\item Donc, on a des données individuelles de firmes sur 1993-2007, avec toutes les exportations par firme/produit, vers quels pays, pour quelle valeur / quantité (donc, à quel prix), et via les EAE, la productivité de chaque firme qui exporte, et plus largement, de toutes les firmes françaises > 20 salariés
\end{itemize}

\subsubsection{Sur le plan du modèle}

Dans le modèle de Chaney-Mélitz, si on raisonne au niveau d'un secteur donné, une baisse des coûts de transport $\tau^k_{ij}$ a un impact d'autant plus grand sur les flux de commerce sectoriels, que $\gamma_h$ est grand. Avec $\gamma_h$, le paramètre de la loi de Pareto qui caractérise la distribution des productivités des firmes (au sein d'un secteur donné).

Deux effets
\begin{itemize}
\item A la marge intensive: Toutes choses égales par ailleurs, la baisse des coûts de transport permet de vendre plus, donc le flux vendu par chaque firme augmente
\item A la marge extensive: plus $\gamma_h$ est élevé, plus la distribution des productivités est homogène, et alors la baisse des coûts de transport modifie la valeur seuil de productivité au-delà de laquelle il devient rentable de payer le coût fixe d'exportation, donc plus de firmes entrent sur le marché d'exportations
\item Notre objectif, caractériser l'évolution du $\gamma$ dans le temps. Notre idée, $\gamma$ augmente au cours du temps, contribue ainsi à expliquer le distance puzzle
\end{itemize}

\subsubsection{Les difficultés: la question de l'agrégation}

On dispose de données de firmes, avec la productivité de chacune (qu'elle exporte ou non). On veut le paramètre $\gamma$ qui est le coefficient qui représente la dispersion dans la distribution de l'efficience des firmes \textbf{au niveau agrégé}. Ie, en ayant agrégé tous les secteurs.

\begin{itemize}
\item On peut estimer un $\gamma_h$ par secteur / par année
\item Question 1: à quel niveau de désagrégation sectorielle? Si on est cohérent avec notre analyse précédente, faut-il estimer un $\gamma_h$ au niveau 4-digit? Combien de secteurs cela représente?

\item Question 2: comment ``remonter'' les $\gamma_h$ sectoriels, au niveau agrégé?

\item Questions connexes: comment gérer si des secteurs apparaissent dans les données d'exportation? ou EAE?
\begin{itemize}
\item[-] Si apparaissent dans les douanes, elles doivent apparaitre aussi dans les EAE, si entreprise a plus de 20 salariés. Si ce n'est pas le cas, où ne peut pas tenir compte de cette information, car on ne peut croiser avec la productivité de la firme.
\item[-] Si n'apparaissent que dans les EAE, et pas les douanes, ça signifie qu'aucune entreprise de ce secteur n'a la productivité seuil suffisante pour exporter. Faut-il tenir compte de ce nouveau secteur?
\end{itemize}
\item Que perd-on si on restreint l'échantillon de firmes
\begin{itemize}
\item[-] On ne garde dans les EAE que les secteurs où les firmes exportent (ie, on élimine les non échangeables). Vérifier la pertinence de cette hypothèse sur le plan théorique.
\item[-] Plus précisément, on ne garde que les secteurs (à quel niveau de désagrégation?) qui sont présents \textbf{tout au long de notre échantillon (ie, toutes les années)} dans les données de douane. Comme ça, on est sûr qu'on a bien, pour toutes les années, uniquement les secteurs ``échangeables'', ie ou au moins une entreprise française exporte.
\end{itemize}

\end{itemize}

Admettons qu'on ait une estimation du $\gamma$ au niveau agrégé. L'objectif final est alors de caractériser son évolution dans le temps, à partir de l'équation

$$\gamma = \beta t + \epsilon_t$$
(ou faut-il prendre le log? a voir)

Et caractériser, comme pour $\rho$, la significativité de $\beta$. Idéalement, on voudrait obtenir $\beta$ significativement positif: On apporterait alors une explication au ``distance puzzle'' par l'évolution (à la hausse) de l'élasticité des flux de commerce aux trade costs, elle-même liée à l'évolution de la dispersion des productivités des firmes, qui sont plus homogènes au cours du temps, ie que c'est donc parce que la marge extensive du commerce joue de plus en plus.

\section{Eléments de modélisation sur l'étape 3}

\subsection{Le cas des coûts de transport iceberg}
On part du modèle de \cite{Chaney-AER-08}.

\subsubsection{Main assumptions}

There a N potentially asymmetric countries that produce goods using only labor. Country $n$ has a population $L_n$ (exogenous and constant). Consumers in each country maximize utility, that derives from the consumption of goods in $H+1$ sectors.

Sector $Z$ provides a single homogeneous good, while each of the other $H$ sectors produce differentiated varieties, with firms in monopolistic competition.

More precisely, we adopt the following specifications. In country $i$, overall utility is given by:

\begin{equation}
\mathcal{U}_i =(C_i^X)^\mu(C_{i}^Z)^{1-\mu} \qquad 0<\mu<1 \label{eq:utility}
\end{equation}

with $C_i^X$ the aggregate bundle of differentiated goods produced in the $H$ sectors. As in \cite{Chaney-AER-08} or \cite{Imbs-Mejean-WP}, we adopt a Cobb-Douglas specification for the composite good of all sectorial bundles ($C_i^X$):

\begin{equation}
C_i^X =\prod_{h=1}^H\left(\frac{C^h_i}{\alpha_h}\right)^{\alpha_h} \qquad \text{with}\sum_{h=1}^H \alpha_h =1 \label{eq:prefgoodX}
\end{equation}

As such, there is an unitary elasticity of substitution across sectorial varieties.

Last, we retain CES preferences for varieties within a given differentiated sector $h$, such as:

\begin{equation}
C_i^h = \left[\int_{\Omega_h}\left(c_i^h(\omega)\right)^{\frac{\sigma_h-1}{\sigma_h}} d\omega \right]^{\frac{\sigma_h}{\sigma_h-1}} \label{eq:prefgoodh}
\end{equation}

with $\sigma_h>1$ the elasticity of substitution between differentiated goods in sector $h$, $\Omega_h$ the set of available varieties (endogenously determined) and $\omega$ a given variety. $c_i^h(\omega)$ thus denotes the consumption by country $i$, of good $\omega$ produced in sector $h$.

As the homogenous good $Z$ is assumed to be freely traded at the worlwide level, with no trade costs, its price is equalized across countries and we normalize its price by 1. Good $Z$ thus serves as numéraire.Good $Z$ is produced under CRS, with one unit of labor in country $n$ producing $w_n$ units of good $Z$. That is, if country $n$ produces this good $Z$ (which is assumed, see Footnote 4 de \cite{Chaney-AER-08}), the wage in country $n$ is $w_n$.

By contrast, goods in the $H$ sectors are subject to trade costs. First, when exporting one variety $\omega$ of sector $h$ from country $i$ to $j$, there are some loss due to iceberg transport cost, ie $\tau^h_{ij}$. Second, firms have to pay a fixed cost of exporting from $i$ to $j$, denoted $f_{ij}^h$. Both fixed and variable costs are assumed to be sector-specific. Due to the presence of fixed costs, firms in all the differentiated sectors operate under increasing-returns-to-scale technology.

Besides, following \cite{Chaney-AER-08}, we assume that firms are heterogenous in productivity. Each firm in a given sector $h$ drams a random unit labor productivity $\varphi$. The efficiency parameter is drawn from a Pareto distribution with shape parameter $\gamma_h$, such that:
\begin{equation}
Pr(\widetilde{\varphi}^h <\varphi)= G_h(\varphi) = 1- \varphi^{-\gamma_h} \qquad \text{with }\gamma_h >\sigma_h-1
\end{equation}

\subsubsection{Programs of the agents}

\paragraph{The household's program: Obtaining the demand functions}

Let's consider the program of the representative household in country $i$. The budget constraint is:
$$C_i^Z + P_i^{1} C_i^{1} + P_i^{2} C_i^{2}+ ... + P_i^{H} C_i^{H} \leq w_i L_i +\pi_i$$
Or

$$C_i^Z + \sum_{h=1}^HP_i^{h} C_i^{h}\leq w_i L_i +\pi_i$$

With $\pi_i$ the dividend received by the i's hh from the global mutual funds (no free-entry condition, so differentiated firms do make profit in equilibrium, given to a worldwide mutual funds, then divided across countries). Precisely, it is assumed that:
$$\pi_i = w_iL_i \pi$$
with $\pi$ the dividend per share of the global mutual fund (endogenous).

In the previous equation, $P_i^h$ represents the price of the composite good of sector $h$ in country $i$.

The budget constraint may equivalently be rewritten as:
\begin{equation}
C_i^Z + P_i^{X} C_i^{X}\leq w_i L_i +\pi_i \label{eq:CB}
\end{equation}

with $P_i^{X}$ the price index of the bundle of differentiated varieties. Let's denote by $Y_i = w_iL_i+\pi_i $ aggregate revenue in country $i$.

\subparagraph{Arbitrage between goods $Z$ and $X$} The optimal choice between the homogenous good $Z$ and the compositie good $X$ is the solution of maximizing utility (Equation \ref{eq:utility}) subject to the budget constraint (\ref{eq:CB}).

We thus obtain the following demand functions:
\begin{eqnarray}
P_i^X C_i^X &=& \mu Y_i \label{eq:dmdgoodX} \\
C_i^Z = (1-\mu) Y_i \notag
\end{eqnarray}

\subparagraph{Arbitrage between sectors} The optimal choice between the various $H$ sectorial varieties is the solution of the following program: Maximize Equation (\ref{eq:prefgoodh}) under the budget constraint:
\begin{eqnarray}
&&P_i^1 C_i^1+... +P_i^H C_i^H \leq P_i^X C_i^X \notag\\
\Leftrightarrow && \sum_{h=1}^H P_i^h C_i^h \leq P_i^X C_i^X \label{eq:CBgoodX}
\end{eqnarray}

The solution of this program maximization yields the following demand function, for the composite good of sector $h$:
$$P_i^h C_i^h = \frac{\alpha_h}{\sum_h \alpha_h}P_i^X C_i^X$$

Recalling our assumption that $\sum_h \alpha_h=1$, and using Equation (\ref{eq:dmdgoodX}, we get:
\begin{equation}
P_i^h C_i^h = \alpha_h \mu Y_B \label{eq:dmdgoodh}
\end{equation}

\subparagraph{Arbitrage between varieties within sectors} Within a given sector $h$, the optimal choice between all varieties solves the program:

$$\max C_i^X = \left[\int_{\Omega_h}\left[c_i^h(\omega)\right]^{\frac{\sigma_h-1}{\sigma_h}} d\omega \right]^{\frac{\sigma_h}{\sigma_h-1}}$$

under the constraint:
\begin{equation}
P_i^h C_i^h = \int_{\Omega_h}p_i^h(\omega)c^h_i(\omega) d\omega \label{eq:CBgoodh}
\end{equation}

Solving this program delivers the optimal demand of country $i$, for the variety $\omega$ produced in sector $h$:
\begin{equation*}
c_i^h(\omega) = \left[\frac{p_i^h(\omega)}{P_i^h}\right]^{-\sigma_h} C_i^h
\end{equation*}

Which becomes, using Equation (\ref{eq:CBgoodh}):
\begin{equation*}
c_i^h(\omega) = \left[\frac{p_i^h(\omega)}{P_i^h}\right]^{-\sigma_h} \alpha_h \mu \frac{Y_i}{P^h_i} \label{eq:dmdgoodi}
\end{equation*}

\subparagraph{Expression of price indices}

The optimal price index of the composite good $X$ ($P_i^X$) is obtained using the previous results. Precisely, when incorporating the optimal demand for any sectorial good $h$ (Equation (\ref{eq:dmdgoodh}) into the constraint (\ref{eq:CBgoodX} gives:

\begin{equation}
P_i^X = \prod_{h=1}^H (P_i^h)^{\alpha-h} \label{eq:indexPX}
\end{equation}

By the same reasoning, one can obtain the expression of the price index for a given sector $h$. Using Equation (\ref{eq:dmdgoodi}) into equation (\ref{eq:CBgoodh}), we obtain that

\begin{equation}
P_i^X = \left[ \int{\Omega_h} (p_i^h(\omega))^{1-\sigma_h}d\omega \right]^{\frac{1}{1-\sigma_h}}\label{eq:indexPh}
\end{equation}

\paragraph{The firms' pricing decisions}

Let's consider now the program of a firm settled in country $i$, from sector $h$, that features a productivity level of $\varphi$. For convenience, let's just assume that there is only two markets to serve, country $i$ and $j$.

The profit expression of firm ($\varphi,h,i)$ is given by:

\begin{equation}
\pi^h_{i}(\varphi) = p^h_{ii}(\varphi) c_{i}^h(\varphi) + p^h_{ij}(\varphi) c_{j}^h(\varphi) - \frac{w_i}{\varphi}\left[c_{i}^h(\varphi)+ \tau^h_{ij}c_{j}^h(\varphi)\right] - f^h_{ij} \label{eq:fprofit}
\end{equation}

with $p^h_{ii}(\varphi)$ the price of good $\varphi$ produced by the firm in sector $h$ located in country $i$ and sold in country $i$, and $p^h_{ij}(\varphi)$ the price of this good when sold in country $y$. Given iceberg costs, the marginal cost of exporting is raised by $\tau^h_{ij}$ in comparison with the local sale (assuming here $\tau^h_{ii}=1$).

Firm are in monopolistic competition in the production of the good, so that each firm $\varphi$ maximizes its profit function (\ref{eq:fprofit}) subject to the demands for its variety coming from each country, recalled here for convenience, coming from countries $i$ and $j$ respectively:
\begin{eqnarray*}
c_i^h(\varphi) &=& \left[\frac{p_{ii}^h(\varphi)}{P_i^h}\right]^{-\sigma_h} C_i^h \\
c_j^h(\varphi) &=& \left[\frac{p_{ij}^h(\varphi)}{P_j^h}\right]^{-\sigma_h} C_j^h
\end{eqnarray*}

Optimal pricing decisions on each domestic and export market respectively yield:
\begin{eqnarray}
p_{ii}^h(\varphi)&=& \frac{\sigma_h}{\sigma_h-1}\frac{w_i}{\varphi} \notag \\
p_{ij}^h(\varphi)&=& \frac{\sigma_h}{\sigma_h-1}\frac{w_i\tau^h_{ij}}{\varphi} \label{eq:prixijh}
\end{eqnarray}

As a result, the exports in value, of firm ($\varphi,h,i$) on market $j$, is equal to
$$x_{ij}^h(\varphi) = p_{ij}^h(\varphi)c_{j}^h(\varphi)$$

Given the optimal pricing decision (\ref{eq:prixijh}) and the demand function (\ref{eq:dmdgoodi}), we obtain:
\begin{equation*}
x_{ij}^h(\varphi)  = \left[\frac{\sigma_h}{\sigma_h-1}\frac{\frac{w_i}{\varphi}\tau^h_{ij}}{P_i^h}  \right]^{1-\sigma_h}\alpha_h \mu Y_j
\end{equation*}

\subsubsection{Skipping the details}

Je n'ai pas totalement tout refait mais bon, publi à l'AER je pense que c'est ok:-)

Donc, on arrive au système donné par l'équation (9) de son papier, plus celle de l'indice de prix $P_i^h$, que je résume ici
\begin{itemize}
\item Export sales of an individual firm ($\phi,h,i$) on export market $j$:
\begin{eqnarray}
x_{ij}^h(\varphi)  &= &\lambda_3^h\left[\frac{Y_j}{Y}\right]^{\frac{\sigma_h-1}{\gamma_h}}\left[\frac{\theta^h_j}{w_i\tau^h_{ij}} \right]^{\sigma_h-1}\varphi^{\sigma_h-1} \qquad \text{if } \varphi\geq\overline{\varphi}^h_{ij}\label{eq:exportsi} \\
&=&0 \qquad \qquad \qquad \qquad \qquad \qquad  \qquad \qquad\text{if }\varphi<\overline{\varphi}^h_{ij} \notag
\end{eqnarray}

\item Cut-off productivity level
\begin{equation}
\overline{\varphi}^h_{ij} = \lambda_4^h \left(\frac{Y}{Y_j}\right)^{\frac{1}{\gamma_h}}\frac{w_i \tau^h_{ij}}{\theta_j}\left[f^h_{ij}  \right]^{\frac{1}{\sigma_h-1}} \label{eq:phibar}
\end{equation}

with $Y$ world production and $\theta_j^h$ an index of remoteness of country $j$, sector $k$ from the rest of the world, whose expression is:
$$\theta^h_j = \left[\sum_{k=1}^N \frac{Y_k}{Y} \left(w_k\tau^h_{kj} \right)^{-\gamma_h}\left(f^h_{kj} \right)^{-(\frac{\gamma_h}{\sigma_h-1}-1)}\right]^{\frac{-1}{\gamma_h}}$$

\item Aggregate revenue in country $i$:
$$Y_i = (1+\lambda_5)w_iL_i$$

\item Dividend per share of the global fund:
$$\pi = \lambda_5$$

\item Price index of sector $h$, in country $j$:
\begin{equation}
P_j^h = \lambda_2\theta^h_j (Y_j)^{\frac{1}{\gamma_h}-\frac{1}{\sigma_h-1}} \label{eq:Phequi}
\end{equation}
\end{itemize}

In the previous system of equations, note that all $\lambda$ are solely functions of deep parameters, hence fully exogenous (assuming $Y$ to be exogenous and constant).

\subsubsection{Obtaining the elasticity of aggregate trade flows to variable trade costs}

From Equation (\ref{eq:exportsi}) and given and the assumption that the mass of potential entrants is proportional to the country size, ie $w_iL_i$), one can obtain the sectoral value of exports of sector $h$ in country $i$, to country $j$:

\begin{equation*}
X_{ij}^h = w_iL_i \int_{\overline{\varphi}_{ij}^h}^{\infty} \lambda_3^h \left[\frac{Y_j}{Y}\right]^{\frac{\sigma_h-1}{\gamma_h}}\left[\frac{\theta^h_j}{w_i\tau^h_{ij}} \right]^{\sigma_h-1}\varphi^{\sigma_h-1} dG(\varphi)
\end{equation*}

Given the properties of the Pareto distribution, we obtain:

\begin{equation}
X_{ij}^h = \alpha_h ~\mu ~\frac{Y_i~Y_j}{Y}\left(\frac{w_i\tau^h_{ij}}{\theta_j^h} \right)^{-\gamma_h}\left(f_{ij}^h\right)^{-\left(\frac{\gamma_h}{\gamma_h-1}-1 \right)} \label{eq:exportsh}
\end{equation}

Total exports of country $i$ to country $j$ are defined as:

$$X_{ij} = \sum_{h=1}^H X^h_{ij}$$

Using our previous results, total exports can be rewritten as:

$$X_{ij} = \sum_{h=1}^H \alpha_h ~\mu ~\frac{Y_i~Y_j}{Y}\left(\frac{w_i\tau^h_{ij}}{\theta_j^h} \right)^{-\gamma_h}\left(f_{ij}^h\right)^{-\left(\frac{\gamma_h}{\gamma_h-1}-1 \right)}$$

Or, equivalently:
\begin{equation}
X_{ij} = \mu \frac{Y_i~Y_j}{Y}~\sum_{h=1}^H \alpha_h \left(\frac{w_i\tau^h_{ij}}{\theta_j^h} \right)^{-\gamma_h}\left(f_{ij}^h\right)^{-\left(\frac{\gamma_h}{\gamma_h-1}-1 \right)} \label{eq:totalexports}
\end{equation}

Au niveau sectoriel, on vérifie bien que l'élasticité des flux de commerce aux variable trade costs est égal à $\gamma_h$:

$$\varepsilon_{X_{ij}^h/\tau^h_{ij}} = -\gamma_h$$

\textbf{Remarque:} Il me semble que cela repose sur l'hypothèse que $\theta^h_j$ est insensible à une variation de $\tau^h_{ij}$. A vérifier.\bigskip

Notre objectif est d'obtenir l'élasticité des flux de commerce \textbf{agrégés}

Deux types de variations de coûts de transport peuvent être étudiés, donc deux élasticités dérivées:
\begin{itemize}
\item Une modification du coût de transport entre le pays $i$ et le pays $j$ qui est identique entre tous les secteurs:
    $$\frac{d\tau^h_{ij}}{\tau^h_{ij}}  = \frac{d\tau^{h^\prime}_{ij}}{\tau^{h^\prime}_{ij}} = \widehat{\tau}_{ij} \qquad \forall h,h^{\prime} \in \Omega_h$$

\item Une modification du coût de transport entre le pays $i$ et le pays $j$ qui est spécifique à un seul secteur, les autres couts de transport sectoriels restant constants:
     $$\frac{d\tau^h_{ij}}{\tau^h_{ij}}  = \widehat{\tau}^h_{ij}$$
\end{itemize}

\paragraph{Simplifying case} Pour simplifier les calculs, supposons deux secteurs (1 et 2). Alors, les exportations totales de $i$ vers $j$ sont:

$$X_{ij} = \mu \frac{Y_i~Y_j}{Y}\left[\alpha_1 \left(\frac{w_i \tau_{ij}^1}{\theta_i^1}\right)^{-\gamma_1}\left(f_{ij}^1  \right)^{-\left(\frac{\gamma_1}{\sigma_1-1}-1 \right)}  + \alpha_2 \left(\frac{w_i \tau_{ij}^2}{\theta_i^2}\right)^{-\gamma_2}\left(f_{ij}^2  \right)^{-\left(\frac{\gamma_2}{\sigma_1-2}-1 \right)} \right] $$

Let's define
$$\kappa_h = \alpha_h \left(\frac{w_i}{\theta_i^h} \right)^{-\gamma_h}\left(f_{ij}^h  \right)^{-\left(\frac{\gamma_h}{\sigma_h-1}-1 \right)}$$

for $h=1,2$ the index of the sector. Thus, aggregate exports can be rewritten as (for a given number $H$ of sectors):

$$X_{ij} = \mu \frac{Y_i~Y_j}{Y} \left[\sum_h=1^H\kappa_h\left(\tau_{ij}^h\right)^{-\gamma_h}\right]$$

In the two-sector case, we get:

$$X_{ij} = \mu \frac{Y_i~Y_j}{Y} \left[\kappa_1\left(\tau_{ij}^1\right)^{-\gamma_1} +\kappa_2\left(\tau_{ij}^2\right)^{-\gamma_2} \right]$$

To obtain the elasticity, I differentiate the previous equation, starting from a given steady state (identified as $\overline{x}$ for any variable $x$).\bigskip

En détaillant les calculs, en supposant le PIB mondial constant (remarque, je me rends compte que j'ai aussi supposé $\kappa_n$ constant, ce qui est faux, mais je fais l'hypothèse que les éléments qui le consitituent ne sont pas affectés par la variation des trade costs. Je pourrai faire pareil pour les PIB, mais là j'ai fait la différentielle totale en supposant qu'ils varient).

\begin{eqnarray*}
X_{ij}-\overline{X}_{ij} &=& \frac{\mu}{Y} \left[\sum_{h=1}^H\kappa_h\left(\tau_{ij}^h\right)^{-\gamma_h}\right]\left[\overline{Y}_j(Y_i-\overline{Y}_i)+ \overline{Y}_i(Y_j-\overline{Y}_j) \right] \\
&&+\mu \frac{\overline{Y}_i\overline{Y}_j}{Y}\left[\kappa_1(-\gamma_1)(\overline{\tau}_{ij}^1)^{-\gamma_1-1}(\tau_{ij}^1- \overline{\tau}_{ij}^1) +\kappa_2(-\gamma_2)(\overline{\tau}_{ij}^2)^{-\gamma_2-1}(\tau_{ij}^2- \overline{\tau}_{ij}^2) \right]
\end{eqnarray*}

Denoting by $\widehat{x}=\frac{x-\overline{x}}{\overline{x}}$ the relative change of variable $x$ from its steady state value, we obtain:

\begin{eqnarray*}
\overline{X}_{ij}\widehat{X}_{ij} &=& \mu \frac{\overline{Y}_i\overline{Y}_j}{Y}\left(\sum_h \kappa_h (\overline{\tau}_{ij}^h \right)^{-\gamma_h}\left[\widehat{Y}_i+\widehat{Y}_j  \right] \\
&&- \mu\frac{\overline{Y}_i\overline{Y}_j}{Y} \left[\kappa_1 \gamma_1 (\overline{\tau}_{ij}^1)^{-\gamma_1}\widehat{\tau}_{ij}^1 +\kappa_2 \gamma_2 (\overline{\tau}_{ij}^2)^{-\gamma_2}\widehat{\tau}_{ij}^2  \right]
\end{eqnarray*}

Given that, in steady state:

$$\overline{X}_{ij} = \mu \frac{\overline{Y}_i\overline{Y}_j}{Y}\left(\sum_h \kappa_h (\overline{\tau}_{ij}^h )^{-\gamma_h}\right)$$

the previous equation can be simplified to give:
$$\widehat{X}_{ij} =\widehat{Y}_i + \widehat{Y}_j - \left[\frac{\kappa_1(\overline{\tau}_{ij}^1)^{-\gamma_1}}{\kappa_1(\overline{\tau}_{ij}^1)^{-\gamma_1}+\kappa_2(\overline{\tau}_{ij}^2)^{-\gamma_2}}\gamma_1\widehat{\tau}_{ij}^1  +\frac{\kappa_2(\overline{\tau}_{ij}^2)^{-\gamma_2}}{\kappa_1(\overline{\tau}_{ij}^1)^{-\gamma_1}+\kappa_2(\overline{\tau}_{ij}^2)^{-\gamma_2}}\gamma_2\widehat{\tau}_{ij}^2  \right]
$$

\paragraph{Soit variation des coûts de transport identiques pour tous les secteurs} Soit $\widehat{\tau}_{ij}^h= \widehat{\tau}_{ij}^{h^\prime} = \widehat{\tau}_{ij}$, et en supposant que $\widehat{Y}_i=\widehat{Y_j}=0$, on peut récrire:

\begin{equation*}
\widehat{X}_{ij} = \frac{-1}{\sum_h \kappa_h (\overline{\tau}_{ij}^h)^{-\gamma_h}}\left[\gamma_1 \kappa_1 \left(\overline{\tau}_{ij}^1\right)^{-\gamma_1}  +\gamma_2 \kappa_2 \left(\overline{\tau}_{ij}^2\right)^{-\gamma_2} \right]\widehat{\tau}_{ij}
\end{equation*}

Soit, l'élasticité des exportations agrégées aux variable trade costs donnée par:

\begin{equation}
\varepsilon_{X_{ij}/\tau_{ij}} = \frac{-1}{\sum_h \kappa_h (\overline{\tau}_{ij}^h)^{-\gamma_h}}\left[\gamma_1 \kappa_1 \left(\overline{\tau}_{ij}^1\right)^{-\gamma_1}  +\gamma_2 \kappa_2 \left(\overline{\tau}_{ij}^2\right)^{-\gamma_2} \right] \label{eq:elast1}
\end{equation}

Problème si on ne peut simplifier
\begin{itemize}
\item Il ne suffit pas d'avoir le paramètre de distribution des productivités par secteur ($\gamma_h$)
\item Il faudrait aussi, le niveau des coûts de transport par secteur ($\overline{\tau}_{ij}^h$)
\item De même que les composantes de $\kappa_h$ (salaire du pays $w_i$, part du secteur $h$ dans le panier de bien différenciés $\alpha_h$, ``remoteness'' du secteur/pays ($\theta_j^h$)
\item Et surtout, le coût fixe à l'export entre le secteur $h$ du pays $i$ vers le pays $j$...
\end{itemize}
\bigskip
Trouver un moyen de simplifier
\bigskip
Appelons $D$ le dénominateur de l'équation (\ref{eq:elast1}) et $N^h$ la partie du numérateur définie telle que:
$$N^h = \kappa_h (\tau_{ij}^h)^{-\gamma_h}$$
L'élasticité des flux de commerce agrégés se réécrit alors:
$$\varepsilon_{X_{ij}/\tau_{ij}}= -\frac{1}{D} \left[\gamma_1 N^1+\gamma_2 N^2 \right]$$

Sachant l'expression des flux agrégés de commerce, on peut récrire $D$ comme

$$D = X_{ij} \frac{1}{\mu}\frac{Y}{Y_i Y_j}$$

Et si on considère l'expression des flux de commerce sectoriels

$$X_{ij}^h = \mu \frac{Y_iY_j}{Y}\underbrace{\kappa_h (\tau_{ij}^h)^{-\gamma_h}}_{N^h}$$

Alors on a
$$N^h = X^h_{ij}\frac{1}{\mu}\frac{Y}{Y_iY_j}$$

Donc, partant de l'élasticité des fux agrégés aux trade costs:
\begin{equation*}
\varepsilon_{X_{ij}/\tau_{ij}} = -\left[\sum_h \frac{N^h}{D}\gamma_h  \right]
\end{equation*}

avec
\begin{eqnarray*}
N^h&=& \alpha_h \left(\frac{w_i \tau_{ij}^h}{\theta_j^h}   \right)^{-\gamma_h}\left(f_{ij}^h\right)^{-\left(\frac{\gamma_h}{\sigma_h-1}-1 \right)} \\
D &= &\sum_{h=1}^H \alpha_h \left(\frac{w_i \tau_{ij}^h}{\theta_j^h}   \right)^{-\gamma_h}\left(f_{ij}^h\right)^{-\left(\frac{\gamma_h}{\sigma_h-1}-1 \right)}
\end{eqnarray*}

On peut simplifier avec
\begin{eqnarray*}
N^h&=&X_{ij}^h \frac{1}{\mu}\frac{Y}{Y_iY_j} \\
D&=&X_{ij} \frac{1}{\mu}\frac{Y}{Y_iY_j}
\end{eqnarray*}

Alors, l'élasticité des flux de commerce agrégés aux trade costs (avraible), est égale à

\begin{equation}
\varepsilon_{X_{ij}/\tau_{ij}}  = - \left(\sum_{h=1}^H \frac{X_{ij}^h}{X_{ij}}\gamma_h \right) \label{eq:elast2}
\end{equation}

L'élasticité des flux de commerce agrégés aux trade costs est (en valeur absolue) égale à la somme pondérée des paramètres de distribution d'efficience sectoriels, la pondération étant donnée par la part de chaque secteur dans les exportations totales.\bigskip

\paragraph{Si on considère une variation des coûts de transport dans un secteur donné} Alors c'est la même chose:

\begin{equation}
\varepsilon_{X_{ij}/\tau^h_{ij}}  = - \frac{X_{ij}^h}{X_{ij}}\gamma_h  \label{eq:elasth}
\end{equation}

\paragraph{Question technique} Attention, on a une nomenclature différente pour les bases de commerce (ie, ce qui nous a servi en étapes 1 et 2 pour calculer le $\rho$) et pour les bases de production (ie, les EAE).

Donc, l'estimation du $\gamma_h$ sectoriel n'a a priori aucune raison de correspondre au même secteur que les secteurs (4-digit) à partir desquels on a estimé l'évolution de $\rho$.

Mais ce n'est pas forcément un problème fondamental car ce qui nous intéresse in fine au niveau agrégé, c'est l'élasticité des trade flows aux trade costs, oe on peut l'obtenir à partir des résultats précédents (\ref{eq:elast2}). Ce qui nous pose problème en revanche, c'est la question de la clé d'agrégation.

On peut obtenir les flux de commerce au niveau sectoriel par les données de douane. En effet, par numéro de siren, on peut apparier les données EAE et les données de douane.
\begin{itemize}
\item On estime un $\gamma_h$ par secteur de production (chaque firme appartient à un secteur productif donné)
\item Pour agréger, on utilise les données de douane pour calculer par firme, le flux de ses exports, on agrège par secteur productif
\item Ie, le secteur est fondamentalement le secteur productif (côté firme) et non pas le secteur du produit exporté (logique, ce qui compte c'est la productivité de la firme qui exporte, qu'elle exporte des bananes ou des tee-shirts ou les deux)
\end{itemize}

{\normalsize %\bibliography{HoursEdhec}
\bibliographystyle{apalike2}
\bibliography{lise}
}

\end{document}

### 2019-01-14 redaction/old/note_secondstage_v1512.tex
rangement dossier redaction + coquille intro
creation dossier "old" pour vieille version + correction reference manquante en intro (Lashkaripour 2018, on avait un beau "?" en gras au beau milieu dela page 4 ;-))

--- 2016-06-13 redaction/dhp_draft_new.tex
@@
 
-\paragraph{Costs} (for handling costs).
+\item Costs (for handling costs).
 The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs).
@@
 
- \paragraph{Freight}.
+\item Freight.
 Quite intuitively, freight costs should depend on the distance between the origin country and the U.S.
@@
 Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.
+
+\end{itemize}
 
@@
 
-%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
+\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively.
 So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ...
@@
 
-%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !
+\textbf{Questions raised} \medskip
+\begin{itemize}
+\item The literature commonly uses gravity variables (such as distance, common language, colonial linkages, common border, etc.) as a proxy for trade costs.
+In the second-step estimation, we only consider distance.
+This choice relies on the fact that we explain the transport gap measured by the gap between the fas price and the caf price.
+Our measure of transport cost is only part of the larger size of trade costs, and there is no a priori reason why the other gravity variables except distance might intervene in the difference between the price at the export departure point and the import arrival point.
 
-\subsection{Results and Interpretation}
-%\section{One step further into interpretation: transportation costs as an indicator of market structure.}
+\item Make a difference between the estimated additive trade cost and the multiplicative trade costs.
+Ie, should we suppose that the explicative variables enter additively when considering the additive trade costs, and multiplicatively regarding the iceberg component?
+No a priori reason.
+In both cases, the dependent variable is expressed as a percentage (of the fob price).
+We decompose this percentage in various dimension which, all added, compose the overall trade cost component.
+This justifies an additive specification, for both the additive and the iceberg trade costs components.
 
-\section{Conclusion}
-
-\newpage
-\bibliographystyle{essaien}
-\bibliography{biblio}
-
-\newpage
+\end{itemize}
 
@@
 A priori non, a verifier.}
-\newpage
-
-\section*{Tables and Figures}
-
-%%%Figures
-% Goodness-of-fit
-
-\begin{figure}[htbp]
-\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-%Economic interpretation
-%usepackage{float} + H is great to keep images on the right page
-\begin{figure}[H]
-\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
-& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-%%%%Tables
-%Goodness-of-fit
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of Goodness-of-fit}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}}\\
-\hline
-\textbf{R$^{2}$} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.31} & \multicolumn{1}{c}{\textbf{0.31}} \\
-Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{\textbf{0.60}} \\
-\textbf{SER} &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & \multicolumn{1}{c}{\textbf{0.85}} \\
-Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & \multicolumn{1}{c}{\textbf{0.73}} \\
-\textbf{Log-likelihood} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & -17530.49 & -20253.49 & -29977.79 & -43341.27 & \multicolumn{1}{c}{-50746.83} & \multicolumn{1}{c}{-52690.29} & \multicolumn{1}{c}{\textbf{-34888.64}} \\
-Terms A \& I & -15125.65 & -17263.20 & -25393.46 & -36788.44 & \multicolumn{1}{c}{-47277.53} & \multicolumn{1}{c}{-49419.70} & \multicolumn{1}{c}{\textbf{-30508.29}} \\
-\textbf{AIC criteria} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{106130.58} & \multicolumn{1}{c}{\textbf{70498.08}} \\
-Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{100155.41} & \multicolumn{1}{c}{\textbf{62284.99}} \\
-\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       &  \\
-2$\times$(ll(UR) -ll(R)) & 4809.68 & 5980.59 & 9168.67 & 13105.67 & \multicolumn{1}{c}{6938.61} & \multicolumn{1}{c}{6541.17} & \multicolumn{1}{c}{\textbf{8760.69}} \\
-\# restrictions  & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426.00} & \multicolumn{1}{c}{422} & \multicolumn{1}{c}{\textbf{401.93}} \\
-p-value & 0.000 & 0.000 & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{\textbf{}} \\
-
-\hline\hline
-   % \bottomrule
-
-    \end{tabular}%
-    \end{center}
-  \label{tab:good_fit_air}%
-  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel, Measures of Goodness-of-fit}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \textbf{Mean stat} \\
-\hline
-\textbf{R$^{2}$}\unboldmath{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &  \\
-Term I only & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{0.34} & \multicolumn{1}{c}{\textbf{0.39}} \\
-Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.56}} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.76} & \multicolumn{1}{c}{\textbf{0.66}} \\
-Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{\textbf{0.57}} \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{-16287.40} & \multicolumn{1}{c}{-16129.13} & \multicolumn{1}{c}{-25169.31} & \multicolumn{1}{c}{-35263.95} & \multicolumn{1}{c}{-41998.95} & \multicolumn{1}{c}{-43692.93} & \multicolumn{1}{c}{\textbf{-28534.30}} \\
-Terms A\& I & \multicolumn{1}{c}{-12985.76} & \multicolumn{1}{c}{-13353.65} & \multicolumn{1}{c}{-21171.37} & \multicolumn{1}{c}{-29490.96} & \multicolumn{1}{c}{-37418.66} & \multicolumn{1}{c}{-39751.86} & \multicolumn{1}{c}{\textbf{-24151.31}} \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & \multicolumn{1}{c}{33328.81} & \multicolumn{1}{c}{33010.27} & \multicolumn{1}{c}{51142.62} & \multicolumn{1}{c}{71365.89} & \multicolumn{1}{c}{84789.89} & \multicolumn{1}{c}{88191.87} & \multicolumn{1}{c}{\textbf{57848.60}} \\
-Terms A \& I & \multicolumn{1}{c}{27331.52} & \multicolumn{1}{c}{28067.31} & \multicolumn{1}{c}{43664.74} & \multicolumn{1}{c}{60475.91} & \multicolumn{1}{c}{76161.33} & \multicolumn{1}{c}{80873.72} & \multicolumn{1}{c}{\textbf{49683.32}} \\
-\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{\textbf{}} \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & \multicolumn{1}{c}{9160.56} & \multicolumn{1}{c}{7882.15} & \multicolumn{1}{c}{\textbf{8765.97}} \\
-\# restrictions  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{427} & \multicolumn{1}{c}{\textbf{417.05}} \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-
-\hline\hline
-   % \bottomrule
-
-    \end{tabular}%
-    \end{center}
-  \label{tab:good_fit_ves}%
-  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of Regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-%Economic Interpretation
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034 & \multicolumn{1}{c}{\textbf{1.051}} \\
-Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029 & \multicolumn{1}{c}{\textbf{1.042}} \\
-Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
-\hline
-\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010 & \multicolumn{1}{c}{\textbf{0.019}} \\
-\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005 & \multicolumn{1}{c}{\textbf{0.007}} \\
-\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
-\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017 & \multicolumn{1}{c}{\textbf{1.025}} \\
-\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 & \multicolumn{1}{c}{\textbf{1.018}} \\
-Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
-\hline
-\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
- \hline
-   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036 & \textbf{1.058} \\
-Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033 & \textbf{1.051} \\
-Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
-\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-\hline
-\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
-Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015 & \textbf{0.029} \\
-Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 & \textbf{0.019} \\
-Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 & \textbf{1.032} \\
-Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018 & \textbf{1.028} \\
-Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
-\hline
-\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\newpage
-
-\appendix
-
-\section{Results with a Higher Level of Products Disaggregation }
-%\vspace{-0.5cm}
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of Goodness-of-fit, 4-digits}
-\begin{center}
-\label{tab:good_fit_air_rob}%
-%\vspace{0.5cm}
-\scalebox{0.97}{
-\begin{tabular}{lccccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                      &  \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\
-\hline
-\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
-Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} & \textbf{0.47} \\
-Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} & \textbf{0.63} \\
-\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93} & \textbf{0.89} \\
-Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86} & \textbf{0.79} \\
-\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87} & \textbf{-34744.40} \\
-Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14} & \textbf{-30243.95} \\
-\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
-Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} & \textbf{70940.07} \\
-Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28} & \textbf{62955.73} \\
-\textbf{Test LL} &       &       &       &       &       &       & \textbf{} \\
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45} & \textbf{9000.89} \\
-\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818} & \textbf{755.73} \\
-p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \textbf{} \\
-\hline\hline
-\end{tabular}}%
-\end{center}
- \vspace{-10pt} \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-\begin{table}[H]
-  \centering
-  \caption{Vessel: Measures of Goodness-of-fit, 4-digits}
-\begin{center}
-\label{tab:good_fit_ves_rob}%
-
-\scalebox{0.97}{\begin{tabular}{lccccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                      &  \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\ \hline
-\boldmath{}\textbf{R$^{2}$}\unboldmath{} &       &       &       &       &       &       &  \\
-Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{\textbf{0.44}} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.59}} \\
-\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} & \multicolumn{1}{c}{\textbf{0.77}} \\
-Terms A \& I &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} & \multicolumn{1}{c}{\textbf{0.68}} \\
-\textbf{Log-likelihood} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & -16460.10 & -16951.61 & \multicolumn{1}{c}{-26771.44} & \multicolumn{1}{c}{-39008.34} & \multicolumn{1}{c}{-43888.90} & \multicolumn{1}{c}{-47161.62} & \multicolumn{1}{c}{\textbf{-29883.62}} \\
-Terms A\& I & -12743.65 & -13546.92 & \multicolumn{1}{c}{-21752.77} & \multicolumn{1}{c}{-33280.96} & \multicolumn{1}{c}{-39078.86} & \multicolumn{1}{c}{-43399.22} & \multicolumn{1}{c}{\textbf{-25303.92}} \\
-\textbf{AIC criteria} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
-Term I only & 34464.19 & 35491.21 & \multicolumn{1}{c}{55272.87} & \multicolumn{1}{c}{79800.67} & \multicolumn{1}{c}{89459.80} & \multicolumn{1}{c}{95987.23} & \multicolumn{1}{c}{\textbf{61425.60}} \\
-Terms A \& I & 28271.29 & 29877.84 & \multicolumn{1}{c}{46595.55} & \multicolumn{1}{c}{69743.91} & \multicolumn{1}{c}{81155.73} & \multicolumn{1}{c}{89692.44} & \multicolumn{1}{c}{\textbf{53573.29}} \\
-\textbf{Test LL} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.75 & \multicolumn{1}{c}{17354.65} & \multicolumn{1}{c}{20113.52} & \multicolumn{1}{c}{16608.16} & \multicolumn{1}{c}{12589.59} & \multicolumn{1}{c}{15704.63} \\
-\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} & \multicolumn{1}{c}{860} \\
-p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-
-\hline\hline
-\end{tabular}}
-\end{center}
-  \vspace{-10pt}\tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion.
-Statistics are obtained weighting each observation by its value in transport (mode-dependent).
-Term A expressed in fraction of fob price.
-R$^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-
-& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
-
-\hline
-\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       &  \\
-Mean  & 1.066 & 1.058 & 1.052 & \multicolumn{1}{c}{1.033} & \multicolumn{1}{c}{1.037} & \multicolumn{1}{c}{1.032} & \multicolumn{1}{c}{\textbf{1.049}} \\
-Median & 1.052 & 1.044 & 1.041 & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.027} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{\textbf{1.037}} \\
-Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{\textbf{0.045}} \\
-\hline
-\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{} \\
-\textit{Additive term} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
-Mean  & 0.026 & 0.021 & 0.017 & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.012} & \multicolumn{1}{c}{0.010} & \multicolumn{1}{c}{\textbf{0.019}} \\
-Median & 0.012 & 0.006 & 0.006 & \multicolumn{1}{c}{0.005} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{0.004} & \multicolumn{1}{c}{\textbf{0.008}} \\
-Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{\textbf{0.034}} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.035 & 1.026 & 1.031 & \multicolumn{1}{c}{1.015} & \multicolumn{1}{c}{1.021} & \multicolumn{1}{c}{1.016} & \multicolumn{1}{c}{\textbf{1.024}} \\
-Median & 1.025 & 1.017 & 1.019 & \multicolumn{1}{c}{1.010} & \multicolumn{1}{c}{1.017} & \multicolumn{1}{c}{1.014} & \multicolumn{1}{c}{\textbf{1.016}} \\
-Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{\textbf{0.026}} \\
-\hline
-\# obs & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460} & \multicolumn{1}{c}{\textbf{}} \\
-
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_air_rob}%
-\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
-
-\begin{table}[htbp]
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-  \centering
-\caption{Vessel: Dynamics of Trade Costs over Time}
-\begin{center}
-    \begin{tabular}{lccccccc}
-  %  \toprule
-   \hline\hline
-          & \multicolumn{6}{c}{Year}                      &  \\
-& 1974  & 1981  & 1989  & 2001  & 2009  & 2013  & \textbf{Mean Stat}\\
-\hline
-\textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.098 & 1.061 & \multicolumn{1}{c}{1.058} & \multicolumn{1}{c}{1.051} & \multicolumn{1}{c}{1.042} & \multicolumn{1}{c}{1.036} & \textbf{1.060} \\
-Median & 1.094 & 1.051 & \multicolumn{1}{c}{1.048} & \multicolumn{1}{c}{1.045} & \multicolumn{1}{c}{1.038} & \multicolumn{1}{c}{1.031} & \textbf{1.052} \\
-Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} & \textbf{0.036} \\
-\hline
-\textbf{With Additive \& Iceberg Trade Costs } &       &       &       &       &       &       & \textbf{}\\
-\textit{Additive term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 0.046 & 0.026 & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.015} & \textbf{0.028} \\
-Median & 0.029 & 0.013 & \multicolumn{1}{c}{0.019} & \multicolumn{1}{c}{0.015} & \multicolumn{1}{c}{0.013} & \multicolumn{1}{c}{0.008} & \textbf{0.017} \\
-Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} & \textbf{0.039} \\
-\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
-Mean  & 1.054 & 1.034 & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.028} & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.021} & \textbf{1.033} \\
-Median & 1.049 & 1.030 & \multicolumn{1}{c}{1.024} & \multicolumn{1}{c}{1.025} & \multicolumn{1}{c}{1.026} & \multicolumn{1}{c}{1.018} & \textbf{1.028} \\
-Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013} & \textbf{0.025} \\
-\# obs & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} & \textbf{}\\
-\hline\hline
-    \end{tabular}%
-  \end{center}
-  \label{tab:result_ves_rob}%
-\vspace{-10pt}\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
-\end{table}%
 

### 2019-01-14 redaction/submission_JoEG.tex
Version JoEG
Version pour soumission au Journal of Econ. Geo, avec Intro modifiée suivant les conseils de F. Robert-Nicoud.

--- 2019-01-14 redaction/TC_new.tex
@@
 
-\date{December 2018}
+\date{January 2019}
  \maketitle
@@
 
-%\begin{spacing}{1.5}
+\begin{spacing}{1.5}
 
@@
 In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Based on panel data, \citet{novy13} finds  that US
-trade costs with major trading partners declined on average by about 40\% between 1970 and 2000.
-Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
-Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
-Using data over 1989-2000, \citet{anderson_wincoop_jel} estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
-
 Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-In this vein, \citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
+However, besides several empirical evidence suggesting that transport costs remain large and deserve attention\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
-With many trade policy barriers removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention.
-International transport costs accordingly stand at the heart of this paper.\smallskip
-
-This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
-Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}).
-However, there is no consensus.
-Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
-In accordance with this view, \cite{hummels2007} points out that transport costs as a proportion of export prices have declined more slowly than is commonly assumed.\smallskip
-
-According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries).
-Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e.
-\textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.}, distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}.
+In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
+The latter is a key issue, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.\smallskip
+
+International transport costs accordingly stand at the heart of this paper.
+We provide a thorough empirical investigation of the pattern of international transport costs over time based on US imports flows from 1974 to 2013.
+Using a similar dataset ending in 2004, \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e.
+\textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects (i.e.
+changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
@@
 
-Proving this point drives us to provide a careful modeling of transport costs.
+These findings are the result of our empirical setting, which carefully distinguishes between multiplicative and additive transport costs.
 Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an ad-valorem tax equivalent (i.e.
 as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} However, a recent strand of the literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015}, or \citealp{martin2012}).
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} However, a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015}, or \citealp{martin2012}).
 Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
@@
 This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-This database is the same as that used by \cite{hummels2007}, extended over more recent years (2005-2013).
+As previously mentioned, this database is the same as that used by \cite{hummels2007}, but extended over more recent years (2005-2013).
 \smallskip
+
+%Based on panel data, \citet{novy13} finds  that US trade costs with major trading partners declined on average by about 40\% between 1970 and 2000.
+%Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
+%Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
+%Using data over 1989-2000, \citet{anderson_wincoop_jel} estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
+
+%This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
 
@@
 Put differently, additive costs represent between one-third and almost one-half of the overall transport costs.
-%\textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?}
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
@@
 In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
-
-%\textbf{enleve ancienne footnote plus haut qui reprenait cette difference, a mettre en footnote ici? en commentaire en dessous}
-%However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$).
-%By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.
 
@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
+\end{spacing}
 

### 2019-01-14 redaction/old/onlineAppendix_DHP.tex
correction version soumise + fin rangement dans dossier old

--- 2018-12-17 redaction/onlineA.tex
@@
 \usepackage{booktabs,caption,fixltx2e}
+\usepackage{setspace}
+\usepackage[flushleft]{threeparttable}
 
-\usepackage[flushleft]{threeparttable}
+\renewcommand{\baselinestretch}{1.2}
+\doublespacing
 
@@
 \urlstyle{rm}
-\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
-Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
-Universit\'{e} Paris-Dauphine, PSL University, LEDa, 75016 PARIS, FRANCE ; email: \url{lise.patureau@dauphine.fr}}}
+\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) \& CEPII; email: \url{jerome.hericourt@univ-lille.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
+Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 Paris, France ; email: \url{lise.patureau@dauphine.psl.eu}}}
 %\date{\vspace{-5ex}}

### 2019-01-14 redaction/submission_JoEG_DHP.tex
correction version soumise + fin rangement dans dossier old

--- 2019-01-14 redaction/submission_JoEG.tex
@@
 This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2013.
-First, we document the importance of the per-unit (additive) component of transport costs.
-We find that additive costs are quantitatively sizable, representing between one-third and one-half of overall transport costs.
-Second, we identify the respective roles of the reduction in ``pure'' transport costs and trade composition effects in the downward trend of international transport costs, in the same spirit as \cite{hummels2007}.
-Unlike him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs.
-Importantly, this difference of results can be attributed to the new method of modeling the per-unit component of transport costs we offer.
+First, we document the importance of the per-unit component of transport costs.
+It represents between one-third and one-half of overall transport costs.
+Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in “pure” transport costs over time, with trade composition effects playing a minor role.
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
@@
 
-\begin{spacing}{1.5}
+%\begin{spacing}{1.5}
 
@@
 \smallskip
-
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-\end{spacing}
 

### 2019-01-14 redaction/submission_JoEG_DHP.tex
unknown
je ne sais pas pourquoi il me demande de commiter, je n'ai rien fait depuis tout à l'heure

@@
 
-Trade costs are central to international economic analysis.
-In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
-However, besides several empirical evidence suggesting that transport costs remain large and deserve attention\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+Trade costs are a major obstacle to international trade flows.
+They take many forms: transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
+Their importance vary enormously between different countries and different sectors.
+They are still the subject of controversy.
+One would think that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
+It is certainly cheaper in constant dollars to transport one ton of goods now than fifty years ago.
+Yet, compared to the value of shipments, several empirical evidence suggesting that transport costs remain large and deserve attention
+\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.}, distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}.
-In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
-The latter is a key issue, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.\smallskip
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+Distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}}.
+Progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
+This an important debate, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.
+This debate is all the more complex as the transport cost burden does not depend only on the size of transport costs, but also on whether they are a proportion of the value of transported goods (and hence do not distort relative prices) or are not related in a linear way to the value of transported goods (and hence do distort relative prices), as would a pure per-ton transport cost schedule .
+\smallskip
 
@@
 \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
+changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
+\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.}
+In this paper, we dig deeper into this issue.
 As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.

### 2019-01-16 redaction/submission_JoEG_DHP.tex
Changement intro

--- 2019-01-14 redaction/submission_JoEG_DHP.tex
@@
 \renewcommand{\baselinestretch}{1.2}
-\doublespacing
+%\doublespacing
 
@@
 
-\newcommand{\note}[1]{\footnote{\begin{doublespace}#1\end{doublespace}}}
+%\newcommand{\note}[1]{\footnote{\begin{doublespace}#1\end{doublespace}}}
 
@@
 
-Trade costs are central to international economic analysis.
-In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
-However, besides several empirical evidence suggesting that transport costs remain large and deserve attention\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+Trade costs are central to international economic analysis. In particular, they are
+known to be a major obstacle to international economic integration and international trade
+flows. Defined as the costs associated with the exchange of goods across
+national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs and non-tariff costs), time costs (time to ship
+goods) and transport costs \textit{per se}. Many argue that the latter have substantially
+decreased with technological advance in transportation, infrastructure development
+and new communication technologies. However, several empirical evidence
+suggest that transport costs remain large and deserve attention,\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.}, distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}.
-In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
-The latter is a key issue, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.\smallskip
-
-International transport costs accordingly stand at the heart of this paper.
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.} with distance as an
+incompressible cost still a very important determinant of trade flows.\footnote{See \citet{Disdier_Head08}.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
+of the transport costs burden.
+The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices).
+If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
+of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring valuable results on this issue.
+
 We provide a thorough empirical investigation of the pattern of international transport costs over time based on US imports flows from 1974 to 2013.
@@
 These findings are the result of our empirical setting, which carefully distinguishes between multiplicative and additive transport costs.
-Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an ad-valorem tax equivalent (i.e.
+Following \citet{samuelson1954}, most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
 as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.

### 2019-09-05 redaction/submission_JEGeo/Initial/submission_JoEG_anonymous.tex
Revision_JoEG
Un peu d'organisation des dossiers JoEG

--- 2019-01-16 redaction/submission_JoEG_DHP.tex
@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Holger G\"{o}rg, S\'{e}bastien Jean and Vincent Vicard for their insightful comments.
-The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the universities of California-Irvine, Paris-Dauphine, Paris-I and Nancy and the Kiel Institute for the World Economy.
-Any remaining errors are ours.
-This paper was partly funded by a university grant from Lille-I University.
-It features an online appendix containing additional results, available on the authors' websites.}}
-
-\author{Guillaume \textsc{Daudin}\thanks{%
-University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}}
 
@@
 flows. Defined as the costs associated with the exchange of goods across
-national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs and non-tariff costs), time costs (time to ship
-goods) and transport costs \textit{per se}. Many argue that the latter have substantially
+national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially
 decreased with technological advance in transportation, infrastructure development
 and new communication technologies. However, several empirical evidence
-suggest that transport costs remain large and deserve attention,\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+suggest that transport costs remain large and deserve attention.\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.} with distance as an
-incompressible cost still a very important determinant of trade flows.\footnote{See \citet{Disdier_Head08}.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+\citet{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
 of the transport costs burden.
@@
 If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
-of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring valuable results on this issue.
+of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.
 

### 2020-06-15 redaction/DP_CEPII/dhp_mise_en_forme_cepii.tex

--- 2018-08-05 redaction/dhp_travail.tex
@@
-\documentclass[a4paper,11pt]{article}
-
+%\documentclass[a4paper,11pt]{article}
+\documentclass{cepiiwp2014}
 %\usepackage[pdftex]{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{geometry}
+\geometry{margin=1in}
+\usepackage[frenchb,english]{babel}
+\usepackage{graphics}
+\usepackage{graphicx}
+%\pagestyle{headings}
+\usepackage{amsmath}
+\usepackage{array}
+\usepackage{booktabs}
+\everymath{\displaystyle}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage[authoryear]{natbib}
+\usepackage{hyperref}
+\usepackage{pict2e}
+\parindent=0cm
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amsthm}
+%\definecolor{perso}{cmyk}{1,0,0,0}
+%\hypersetup{linkcolor=perso}YellowGreen
+\usepackage[usenames,dvipsnames]{color}
+\hypersetup{colorlinks=true,linkcolor=blue, citecolor=Blue, urlcolor=Blue}
+\usepackage{setspace}
+%\onehalfspacing
+\doublespacing
+\usepackage{lscape}
+\setlength{\bibhang}{1.5em}
+\usepackage{relsize,exscale}
+\usepackage{pstricks}
+\setlength{\unitlength}{1cm}
+\usepackage{threeparttable}
+\usepackage{floatrow}
+\usepackage[tableposition=top]{caption}
+\floatsetup[table]{capposition=top}
+\floatsetup[figure]{capposition=top}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \usepackage{amsmath}
@@
 \usepackage{lscape}
-\usepackage[round]{natbib}
+%\usepackage[round]{natbib}
+\usepackage[authoryear]{natbib}
 \usepackage{multirow}
@@
 \usepackage{bbm}
-\usepackage[flushleft]{threeparttable}
+%\usepackage[flushleft]{threeparttable}
 %\usepackage{bbold}
@@
 \usepackage{upgreek}
-\usepackage{comment}
-\includecomment{commentGD}
-\usepackage[draft]{todonotes}
+%\usepackage{comment}
+%\includecomment{commentGD}
+%\usepackage[draft]{todonotes}
 %\usepackage[disable]{todonotes}
@@
 
+\title{International Transport costs:...}
+
 \begin{document}
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine, Paris-I and Nancy for very useful discussions.
+\maketitle
+
+\begin{affiliation}[International Transport costs:\\New Findings from modeling additive costs\footnote{We are grateful to S\'{e}bastien Jean, Vincent Vicard, and participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine, Paris-I and Nancy for very useful discussions.
 Finally, any remaining errors are ours.
-This paper features an online appendix containing additional results and available on the authors' websites.}}
-
-\author{Guillaume \textsc{Daudin}\thanks{%
-Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
-Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
-; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-Universit\'{e} Paris-Dauphine \& PSL University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }
-
-\date{June 2018}
+This paper features an online appendix containing additional results and available on the authors' websites.}]
+
+\urlstyle{rm}
+
+Guillaume Daudin\footnote{Universit\'{e} Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL; Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE); Email: \url{guillaume.daudin@dauphine.fr}} and J\'{e}r\^{o}me Hericourt\footnote{Universit\'{e} de Lille, LEM, UMR CNRS 9221, and CEPII; Email: \url{jerome.hericourt@univ-lille.fr}} and Lise Patureau\footnote{Corresponding author.
+Universit\'{e} Paris-Dauphine \& PSL University, LEDa; Email: \url{lise.patureau@dauphine.fr}}
+
+\end{affiliation}
+
  \maketitle
@@
 
-\begin{abstract}
-This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows over 1974-2013.
+%\begin{abstract}
+%This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows over 1974-2013.
 First, we document the importance of the per-unit (additive) component of transport costs.
@@
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
-\end{abstract}
-
-\thispagestyle{empty} \pagestyle{plain} \setcounter{page}{1}
-
-\bigskip
-
-\noindent \emph{JEL classification}: F14, N70, R40 \\
-\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs, trade composition effects
+%\end{abstract}
+
+%\thispagestyle{empty} \pagestyle{plain} \setcounter{page}{1}
+
+%\noindent \emph{JEL classification}: F14, N70, R40 \\
+%\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs, trade composition effects
 
@@
 \begin{center}
-\includegraphics[width=8cm, height=5cm]{part_cout_additif.pdf}
+\includegraphics[width=10cm, height=7cm]{part_cout_additif.pdf}
 %{\footnotesize {OECD data} }
@@
 To this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
-Figure \ref{fig:histogram_beta_robustness} reports the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for Air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a, (c))) and the non-separability scenario (Panels (b), (c)).
+Figure \ref{fig:histogram_beta_robustness} reports the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$), for Air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c)) and the non-separability scenario (Panels (b), (c)).
 
@@
 \newpage
-\bibliographystyle{essaien}
+\bibliographystyle{elsarticle-harv}
+
+%\section*{References}
 \bibliography{biblio}
@@
 
-\clearpage
-
-%\appendix
+%\clearpage
+
 \setcounter{table}{0}
@@
 
-\section{Data Appendix \label{app:data}}
+\subsection{Data Appendix \label{app:data}}
 
@@
 
-\section{Estimation at the 3-digit classification level \label{app:more_results}}
-
-\subsection{Transport costs estimates: More detailed results}
+\subsection{Estimation at the 3-digit classification level \label{app:more_results}}
+
+\subsubsection{Transport costs estimates: More detailed results}
 
@@
 
-\subsection{Assessing the importance of additive transport costs \label{app:diagnostic_test}}
+\subsubsection{Assessing the importance of additive transport costs \label{app:diagnostic_test}}
 
@@
 
-\subsection{Variance decomposition exercise \label{app:decomp_variance}}
+\subsubsection{Variance decomposition exercise \label{app:decomp_variance}}
 
@@
 
-\section{Estimation at the 4-digit level \label{app:4digit}}
+\subsection{Estimation at the 4-digit level \label{app:4digit}}
 
@@
 
-\subsection{Transport cost estimates}
+\subsubsection{Transport cost estimates}
 
@@
 
-\subsection{Goodness-of-fit tests at the 4-digit level}
+\subsubsection{Goodness-of-fit tests at the 4-digit level}
 
@@
 
-\section{Eliminating the composition effects: More details \label{app:comp-effects}}
+\subsection{Eliminating the composition effects: More details \label{app:comp-effects}}
 
@@
 
-\subsection{More on our methodology}
+%\subsubsection{More on our methodology}
 
@@
 
-\subsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
+\subsubsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
 

### 2020-06-15 redaction/submission_JEGeo/Initial/Title_page_JoEG_DHP.tex

--- 2019-01-14 redaction/submission_JoEG_DHP.tex
@@
 
-%\begin{spacing}{1.5}
-
-\section{Introduction \label{sec:Intro}}
-
-Trade costs are central to international economic analysis.
-In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
-Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
-However, besides several empirical evidence suggesting that transport costs remain large and deserve attention\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
-Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.}, distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}.
-In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
-The latter is a key issue, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.\smallskip
-
-International transport costs accordingly stand at the heart of this paper.
-We provide a thorough empirical investigation of the pattern of international transport costs over time based on US imports flows from 1974 to 2013.
-Using a similar dataset ending in 2004, \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e.
-\textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
-As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
-Rather, the evolution of transport costs \textit{per se} constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
-In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
-In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip
-
-These findings are the result of our empirical setting, which carefully distinguishes between multiplicative and additive transport costs.
-Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an ad-valorem tax equivalent (i.e.
-as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} However, a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015}, or \citealp{martin2012}).
-Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
-We obtain this by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
-To do so, we exploit information contained in the Census Bureau's database of US imports flows over 1974-2013.
-This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-As previously mentioned, this database is the same as that used by \cite{hummels2007}, but extended over more recent years (2005-2013).
-\smallskip
-
-%Based on panel data, \citet{novy13} finds  that US trade costs with major trading partners declined on average by about 40\% between 1970 and 2000.
-%Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
-%Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
-%Using data over 1989-2000, \citet{anderson_wincoop_jel} estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
-
-%This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
-
-Our results may be summarized in two main new findings.
-First, additive transport costs are quantitatively sizable.
-On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively (at the 3 digit classification level).
-This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
-Put differently, additive costs represent between one-third and almost one-half of the overall transport costs.
-To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
-This represents a valuable insight for calibrating related international trade or business cycle models.
-We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.
-Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip
-
-Second, we find that the main source of the downward trend in overall transport costs, by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost \textit{per se}.
-Trade composition effects have only limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
-Otherwise stated, pure transport costs have declined less than observed costs.
-This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
-Our result thus deepens the paradox of the less-than-expected decline of transport costs.
-Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
-Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-\smallskip
-In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
-In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
-As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
-In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out'').
-This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
-Also, their estimates imply that doubling freight costs increases average free alongside ship (FAS) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.
-Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using transaction-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
-\citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
-This is confirmed by \cite{Irrazabal_2015}.
-While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been made in quantifying such gains.
-One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
-One contribution of this paper is to fill that gap.}
-
-Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
-Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
-Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
-Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
-From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
-This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
-Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2013.
-In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
-
-The paper is built as follows.
-First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
- Section \ref{sec:data_method} reports the mean values of the transport costs estimated over the period (by transport mode).
-Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
-Section \ref{sec:results_trends} is devoted to this analysis, while Section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
-Section \ref{sec:conclu} concludes.
-
-\section{Estimating international transport costs\label{sec:data_method}}
-
-\subsection{Data}
-
-Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
-The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2013 (details on the database provided in Appendix \ref{app:data}).
-We first use customs values, quantities and freight costs to recover FAS and cost-insurance-freight (CIF) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the FOB (free on board) price rather than the FAS price.
-The FAS price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
-It is also the seller's responsibility to clear the goods for export.
-The FOB price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
-Once the goods clear the railing of the vessel, the buyer assumes the risk.
-Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports FAS prices.} More precisely, the (unit) FAS price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the goods net of transportation costs.
-The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-Our dependant variable is finally computed as the ratio of the CIF price divided by the FAS price.
-By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
-This is quite a standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
-
-Obviously, this dataset has limitations.
-First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the CIF-FAS price gap only covers freight, insurance and handling costs.
-It is thus silent about the other dimensions of international trade costs.
-Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size, etc.).
-In this respect, this dataset embraces only a partial view of international transport barriers.
-However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
-According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
-Further, the evidence summarized in \citet{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
-They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
-In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}
-
-Exploiting this dataset also has three advantages.
-First, it comes from a single, homogenous and trustworthy customs source.
-This limits the measurement error bias.
-Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
-%This will be crucial to compute transport costs, as we detail below.
-In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
-We will make use of this to highlight potential differences in the dynamics of transport costs across transportation mode.
-Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g.
-\cite{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
-ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip
-
-As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-We estimate international transport costs at the 3-digit classification level, even if data series on the CIF and FAS prices are available at the 5-digit level.
-However, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
-Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.
-
-\subsection{Estimating transport costs: empirical specification}
-
-\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
-To do so, we start from the equation that expresses the price $p$ (per kg) of goods paid by the importer (import, or CIF price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or FAS price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
-
-\begin{equation}
-p = \tau \widetilde{p}+ t \label{eq:base}
-\end{equation}
-
-\noindent As is usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
- Let us denote $i$ the origin country, and $k$, the product at the 5-digit level.
-Transforming the above equation (\ref{eq:base}) into a ratio, and removing 1 on both sides, in order to get the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
-
-\begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
-\end{equation}
-
-\paragraph{Estimation strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
-We also provide a robustness analysis to the separability assumption.
-Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries.
-See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}
-
-\begin{eqnarray}
-\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:ad-valorem}\\
-t_{ik} &=& t_{i} + t_{k} \label{eq:add}
-\end{eqnarray}
-
-\noindent As a result, our underlying structural equation is specified as:
-
-\begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
-\end{equation*}
-
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the CIF price $p$ cannot be lower than the FAS price ($p_{ik}>\widetilde{p}_{ik}$).
-Taking into account this constraint in the estimate implies that the error term should be always positive.
-We ensure that this constraint is fulfilled by specifying the error term as follows:
-
-\begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
-\end{equation*}
-\noindent where $\epsilon_{ik}$ follows a normal law centered on 0.
-Considered in logs, the above equation becomes:
-
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
-\end{equation}
-
-Equation (\ref{eq:equation0}) is non-linear. It cannot be estimated using standard linear estimators.
-All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
-The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
-See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit level ($k$).
-This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
-same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed CIF-FAS price.
-As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
-on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
-This would not solve the problem, however, for three main complementary reasons.
-First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
-Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
-Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
-Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}
-
-This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
-\end{equation}
-where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, strictly speaking, the exact equation that we estimate is the following:
-$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
-After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode).
-For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5\% of the upper and lower tails of the distribution in the regression variables.
-%JH: TO KEEP IN MIND: one day, we could be asked to winsorize tails, instead of cutting them...
-These cut-offs are aimed at eliminating reporting or coding errors.
-We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
-Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
-  \medskip
-
-As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, which decomposes into both an additive and an ad-valorem component.
-In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others).
-We contribute to this issue by adopting the following strategy.
-For each year and transport mode, we estimate two models.
-In Model (a), additive costs are excluded, transport costs being modeled as iceberg costs only (Equation (\ref{eq:model_nlI})).
-In this case, the estimated equation simplifies as:
-
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
-\end{equation}
-
-Under Model (b), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
-From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modeling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
-The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating further results.
-More details of these results are given in the Online Appendix.}
-
-One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
-Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
-This is not an issue here, as we are not interested in causal inference.
-Rather, our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
-
-%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
-Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
-In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
-
-%\textbf{pas encore completement clair.
-Veut-on dire que ce qui est pbtique, c'est que TC reagisse a ptilde?
-Je crois que ce qu'on veut dire, c'est que le fait que ptilde change en fonction des TC ne change pas les TC, le cif price s'ajuste proportionnellement?
-Si c'est ca, c'est vrai seulement si les couts sont advalorem...Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus?
-Par produit 3d/pays, par secteur 3d tous pays, etc...?}\smallskip
-
-After estimating Equation (\ref{eq:model_IetA}), we can rebuild a measure of each component, $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is(k)} = \widehat{\tau}_{i} \times \widehat{\tau}_{s(k)}$.
-In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases.
-Similarly, as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (a), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (b), for each year and transportation mode.
-These results are reported in Section \ref{sec:results_decomposition}.
-
-\subsection{The importance of the additive component \label{sec:results_decomposition}}
-
-%Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
-\paragraph{Quantification} The first step of the analysis consists of providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
-This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter.
-These results constitute our first original contribution to the literature.
-We also provide diagnostic tests about the goodness-of-fit measures, which confirm the empirical relevance of adding the additive component in the estimation (see results in Appendix \ref{app:diagnostic_test}).\smallskip
-
-Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
-The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
-Finally, the bottom panel reports the same set of descriptive statistics, but for the actual CIF/FAS ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
-Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for air transport at the 3-digit level, removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
-Overall results (over the whole period) are not substantially affected if this year is included.
-This can be seen in the results detailed year by year in the Online Appendix, available on the authors' webpages.}
-
-\begin{table}[htbp]
-  \centering
-  \footnotesize{
-  \caption{Transport costs estimates: summary \label{tab:summary_results}}
-  \begin{center}
-    \begin{tabular}{l|cc|cc}
-      \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
-    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
-    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
-    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
-    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
-    %Min.
-%value & 1.003 & 1.001 & 1.003 & 1.000 \\
-    %Max.
-%value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
-   \textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} & & & & \\ \hline
-    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
-    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
-    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
-    %Min.
-%value & 1.001 & 1.000 & 1.000 & 1.000 \\
-    %Max.
-%value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term }& & & &   \\ \hline
-    \textit{In \% of the export price ($\widehat{t}/\widetilde{p}$, in \%)} &&&& \\ \hline
-        Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
-    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
-   \textit{In USD per kg traded ($\widehat{t}$)}&&&& \\ \hline
-    Mean & 0.09	&1.12	&0.08&	2.16 \\
-    Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
-    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
-    %Min.
-%value & 0.000 & 0.000 & 0.000 & 0.000 \\
-    %Max.
-%value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-    \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
-    Transport costs ($p/\widetilde{p} -1$, in \%) & & & & \\ \hline
-    Mean & 5.3 & 5.0& 5.6&3.9 \\
-    Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
-    Export price ($\widetilde{p}$), in USD & & & & \\
-    Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
-    Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
-    \# obs.
-& 29279 & 28207 & 29317 & 27680 \\
-    \# origin country & 188 & 191 & 188 & 189 \\
-    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
-  \end{tabular}
-    \end{center}}
-\parbox[l]{12cm}{\footnotesize{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
-The additive term is expressed in fraction of FAS price.
-For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
-1974, 1981, 1989, 2001, 2009, 2013.
-($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
-\end{table}%
-
-Table \ref{tab:summary_results} calls for two types of comments.
-The first type is of statistical order.
-As displayed in Table \ref{tab:summary_results}, estimated mean values are systematically higher than medians, by around one percentage point.
-This result is not surprising, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
-One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
-This is first due to the fact that observed data are \textit{de facto} trade-weighted (i.e.
-by the weight of each good in total trade) while our estimates are based on simple arithmetic means within sector and country.
-Furthermore, Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization.
-Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
-Descriptive statistics showing that both observed and predicted values do match in logarithms are available upon request to the authors.%\medskip
-
-\paragraph{Interpretation} Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}.
-First, it provides estimation results about the size of the overall transport costs.
-Table \ref{tab:summary_results} shows that transport costs amount to around a 6\% markup in ocean shipping, and 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (a) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
-Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
-With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \citet{anderson_wincoop_jel}, lays in the difference of empirical methodology.
-\citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
-
-Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
-Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8\% for ocean shipping and air transport respectively.
-This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period).
-Put differently, the additive component amounts to 48\% of total transport costs in vessel, and 42\% in Air.
-This shows that the per-kg cost dimension is quantitatively sizable.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 2.
-Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests that confirm the importance of the additive cost component.
-The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
-An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.%\smallskip
-
-Our estimation results on the size of additive transport costs can be compared to \citet{Irrazabal_2015}, the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e.
-$\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
-One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
-From this, we can rebuild the ratio in similar terms to them, allowing us to gain in generality in this respect.
-Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping and 1.5\% in air transport.
-This may seem surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
-This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
-While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
-In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
-\citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
-For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
-Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.
-On top of our results summarized in Table \ref{tab:summary_results}, this points to evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
-
-\paragraph{Time dynamics} Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, i.e. a structural phenomenon.
-We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
-The corresponding values reported in detailed tables are available in the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
-Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
-
-\begin{figure}[htbp]
-\caption{Share of additive costs (3-digit classification)}
-\label{fig:part_cout_additif}
-\begin{center}
-\includegraphics[width=8cm, height=5cm]{Figure1_share_of_additive_in_totalTC.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
-In both transport modes, the additive component appears of sizable importance throughout the period.
-Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-On the contrary, they represent a sizable and structural dimension of international transport costs.
-If the share of additive costs in total costs roughly amounts to 48\% (42\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
-In the next section, we further investigate the importance of the (varying pattern of) additive costs, as regards the determinants of transport costs time trends.
-
-\section{Transport costs time trends: the role of the additive component}\label{sec:results_trends}
-
-In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
-This focus is shared with some of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
-Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
-
-\subsection{A first look at the trend patterns}
-
-As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the FAS price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
-Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2013.
-
-\begin{figure}[htbp]
-\caption{Trend in estimated transport costs (yearly mean value)}
-\label{fig:Trends_in_TC}
-\begin{center}
-\includegraphics[width=14cm, height=7cm]{Figure2_Trend_of_totalTC_bymode.pdf}
-\end{center}
-\end{figure}
-
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
-In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974.
-However, computing the time trends from 1980 does not dramatically change the picture.
-The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \citet{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
-For the same years, we obtain a total decrease from 6.9\% to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
-Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip
-
-Before making any definite statement, however, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
-Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
-Our own results tend to challenge this view, as we further show.
-
-\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
-He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
-
-\begin{equation}
-TC_{ikt}\equiv \frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} = \widetilde{p}_{ikt}^{-\beta}X_{ikt} \label{eq:Hummel_benchmarkeq}
-\end{equation}
-
-Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value).
-As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem.
-At the other extreme, $\beta = 1$ means that all transport costs are additive.
-Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
-Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
-In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
-Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
-
-\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
-%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
-We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
-To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e.
-$\beta$, according to:
-\begin{eqnarray*}
-\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
-\end{eqnarray*}
-
-\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-We can then rebuild the value of the elasticity on a per product-year-country basis.
-The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
-
-\begin{figure}[htbp]
-\caption{Share of additive costs in total costs: Histogram}
-\label{fig:histogram_beta}
-\begin{center}
-\begin{tabular}{cc}
-%{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.0in, height=2.5in]{Etude_beta_pondere_air.pdf}
-& \includegraphics[width=3.0in,height=2.5in]{Etude_beta_pondere_ves.pdf} \\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
-\end{tabular}
-\end{center}
-\end{figure}
-
-As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
-This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
-Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
-These results are not reported for sake of space, but are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
-
-\subsection{Empirical specification}
-
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
-We start with the presentation of our estimation strategy before turning to the results.\smallskip
-
-\subsubsection{Estimation strategy}
-
-The estimation carried out in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
-Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
-
-One advantage of this method is that it yields measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
-We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).
-
-\subsubsection{Estimation method for each additive and multiplicative component}
-
-We begin by describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
-\smallskip
-
-\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e.
-extracting from trade composition effect) can be described as a two-stage process.
-First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-For the estimated ad-valorem component, we estimate the following equation:
-\begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
-\end{eqnarray}
-
-\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
-the series of $\gamma_t$ differs from one estimation to the other).
-Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
-
-%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
-\begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
-\end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
-Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
-Both equations are estimated by transport mode.\smallskip
-
-In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
-This constitutes the second stage of our procedure.
-As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we build the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
-\begin{equation}
-\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
-\end{equation}
-
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (see details in Appendix \ref{app:comp-effects}).
-In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
-
-As for the additive cost defined in $[0;+\infty]$, we build the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
-\begin{equation}
-\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
-\end{equation}
-
-\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
-
-\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
-To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
-
-$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
-
-\subsubsection{Estimation method for the total transport cost measures}
-
-We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
-composition effects excluded).
-As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
-Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
-\smallskip
-
-\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
-Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the FAS price.
-We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
-
-\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
-composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
-In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
-For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
-
-$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
-
-with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
-A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
-
-$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
-
-with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.
-
-We then deduce the fitted value of the overall cost according to:
-$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$
-
-Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
-
-\subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
-Panels (d), (e) and (f) report the results for vessel.
-In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects)}
-\label{fig:totalTC_compeffects_excl}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_all.pdf}
-\end{center}
-\end{figure}
-
-Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
-This stands in line with the related literature (see \citealp{Hummels_1999}; \citealp{Lafourcade_Thisse}).
-Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
-Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
-Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
-For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
-Air transport costs were reduced by 50\% between 1974 and 2013, which is mainly attributable to a reduction in the transport costs ``per se''.
-Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
-Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
-Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
-This is particularly the case for the multiplicative component (panel (d)).\smallskip
-
-\subsection{Time trends in transport costs and the modeling of the additive component}
-
-The last results stand in sharp contrast with \citet{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
-If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
-This drives us to investigate this difference further.
-As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual CIF-FAS price gap).
-Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
-
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
-The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
-composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \citeh{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
-
-\begin{figure}[htbp]
-\caption{Characterizing the time trends: applying Hummel's (2007) method }
-\label{fig:comp_effects_as_in_Hummels}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Ad-valorem air freight} & {\small (b) Ad-valorem vessel freight}\\
-\multicolumn{2}{c}{{\small In percent of the export price}} \\
-\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.pdf}
-& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.pdf} \\
-{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
-\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
-\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.pdf}
-& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.pdf} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
-Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
-
-Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction in international transport costs observed over time, in particular in air transport.
-Complementing the findings of Section \ref{sec:results_decomposition}, these results highlight the importance of integrating the additive dimension of international transport costs, with the aim here of characterizing their time trends.
-
-\section{Robustness analysis \label{sec:robustness}}
-
-\subsection{Robustness to the separability assumption}
-In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, which considers (following \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
-That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $t_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
-Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approx.
-200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
-This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.
-For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
-
-\begin{table}[htbp]
-  \centering
-  \caption{Robustness to the separability assumption (average over the period) - restricted sample}
-\begin{center}
-    \begin{tabular}{lcc|cc}
-    \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2013} \\
- \multicolumn{5}{c}{3 digits} \\    \hline \hline
-   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
-   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
-    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
-in \%)}  \\
-    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
-    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
-    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
-    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
-    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
-\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
-    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
-    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
- \textit{Diagnostic test} & & &  &  \\
-    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
-    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
-    AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
-    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
-\# observations & 2381 & 2381 & 2798 & 2798 \\
-\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
-\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
-    \end{tabular}%
-\end{center}
-\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-The additive term is expressed in fraction of FAS price. 1989 omitted in 3-digit estimation for air.}}
-\label{tab:robustness_separability}
-\end{table}
-
-Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52\% and 67\% of total costs on average, versus 40\% to 50\% on the whole sample.
-%JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
-Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
-The picture is reversed for medians.
-
-Before jumping to the conclusion that our results are robust to the separability assumption, however we investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
-Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (a) and for maritime transport in panel (b).
-In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
-
-\begin{figure}[htbp]
-\caption{Robustness to the separability assumption (year-to-year basis)}
-\label{fig:robustesse_non_separe}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Air freight} & {\small (b) Vessel freight}\\
-\includegraphics[width=3in, height=2in]{graph_robustesse_ns_mp_terme_A_air.pdf}
-& \includegraphics[width=3in,height=2in]{graph_robustesse_ns_mp_terme_A_ves.pdf} \\
-\includegraphics[width=3in, height=2in]{graph_robustesse_ns_mp_terme_I_air.pdf}
-& \includegraphics[width=3in,height=2in]{graph_robustesse_ns_mp_terme_I_ves.pdf} \\
-
-\end{tabular}
-\end{center}
-\begin{minipage} [c]  {5in} \scriptsize%
-Notes: FE for (country, sector) fixed effects.
-\end{minipage}
-\end{figure}
-
-In line with Table \ref{tab:robustness_separability}, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
-However, they also show that the difference is quantitatively small.
-Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
-Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.
-
-As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-With this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
-Figure \ref{fig:histogram_beta_robustness} shows the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
-
-\begin{figure}[htbp]
-\caption{Histogram of the share of additive costs in total costs (restricted sample)}
-\label{fig:histogram_beta_robustness}
-\begin{center}
-\begin{tabular}{cc}
-\multicolumn{2}{c}{Air transport} \\
-{\small (a) Separability (baseline)} & {\small (b) Non-separability}\\
-\includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_air.pdf}
-& \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_air.pdf} \\
-\multicolumn{2}{c}{Maritime transport} \\
-{\small (c) Separability (baseline)} & {\small (d) Non-separability}\\
-\includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_ves.pdf}
-& \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_ves.pdf} \\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
-\end{tabular}
-\end{center}
-\end{figure}
-
-Similarly, as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
-Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.
-
-\subsection{Robustness to the weighting scheme }
-
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \citet{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
-In this section, we provide a robustness test to this difference of treatment.\smallskip
-
-When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
-In this section, we check whether this difference of treatment is responsible or not for the difference in results relative to the role of the trade composition effects, which we rather attribute to the modeling of the additive component.
-
-To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \citet{hummels2007} weighting scheme (i.e.
-taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
-The results are reported in Figure \ref{fig:compeffects_robustness}.
-
-\begin{figure}[htbp]
-\caption{Transport cost time trends: robustness to the weighting scheme}
-\label{fig:compeffects_robustness}
-\begin{center}
-\includegraphics[height=8cm]
-{graph_composition_all_np.pdf}
-\end{center}
-\end{figure}
-
-The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
-First, differences do show up, in particular for the multiplicative component of air transport.
-This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
-This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
-
-\section{Conclusion \label{sec:conclu}}
-
-This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit CIF-FAS unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
-Two main findings emerge from our results.
-First, we provide a quantitative measure of both the additive and the ad-valorem transport costs.
-We find that additive costs are quantitatively sizable: They amount to 2.8\% of the export price unit values for ocean shipping, and 1.8\% for air transport.
-This is slightly lower that the estimated values for the ad-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
-Further, modeling the additive component significantly improves the quality of fit of the regression.
-Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
-Allowing for a varying share of additive costs in the product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
-If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in maritime transport.
-In both aspects, our results highlight the importance of the additive component in accounting for international transport costs.
-
-Our results could be extended in two main ways.
-On the empirical side, one may want to go deeper into the structural determinants of (pure) transport costs - i.e.
-identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows in an international trade theory perspective and in affecting the international transmission of business cycles.
-This is left for further research.
-
-\newpage
-\bibliographystyle{essaien}
-\bibliography{biblio}
-
-\newpage
-
-\appendix
-
-\clearpage
-
-%\appendix
-\setcounter{table}{0}
-\renewcommand{\thefigure}{A.\arabic{figure}}
-\renewcommand{\thetable}{A.\arabic{table}}
-
-\section{Data Appendix \label{app:data}}
-
-The Customs value is the value of imports as appraised by the U.S.
-Customs and Border Protection in accordance with the legal requirements of the Tariff Act of 1930, as amended.
-This value is generally defined as the price actually paid or payable for merchandise when sold for exportation to the United States, excluding U.S.
-import duties, freight, insurance, and other charges incurred in bringing the merchandise to the United States.
-The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller.
-In this respect, the ``custom value'' corresponds to the FAS price (``free-alongside'' price) delivered by the seller.
-More information is available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}.
-
-The import charges represent the aggregate cost of all freight, insurance, and other charges (excluding U.S.
-import duties) incurred in bringing the merchandise from alongside the carrier at the port of exportation in the country of exportation and placing it alongside the carrier at the first port of entry in the United States.
-In the case of overland shipments originating in Canada or Mexico, such costs include freight, insurance, and all other charges, costs and expenses incurred in bringing the merchandise from the point of origin (where the merchandise begins its journey to the United States in Canada or Mexico) to the first port of entry.
-
-The CIF (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
-It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S.
-import duties.
-
-\clearpage
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
-\renewcommand{\thetable}{B.\arabic{table}}
-
-\section{Estimation at the 3-digit classification level \label{app:more_results}}
-
-\subsection{Transport costs estimates: More detailed results}
-
-In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
-Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
-Table \ref{tab:result_air_3d_detail} reports the results for air transport.
-The results for vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.
-
-%\todo{Changer\begin{table}[htbp]
-\begin{table}[htbp]
-  \caption{Air: Transport costs estimates, 3-digit level (selected years)}
-\begin{center}
-    \begin{tabular}{l|ccccccc}
-\hline\hline
-Year & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
-\hline
-\multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}     \\
-\hline
-Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.0 & 4.2 & 3.4  \\
-Median & 5.4 & 3.8 & 4.4 & 2.5 & 2.9 & 3.4 & 2.9  \\
-%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
-\hline
-\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%)}  \\ \hline
-Mean & 3.6 & 2.3 & 2.4 &1.7 & 1.9 & 2.6 & 1.7  \\
-Median & 2.7 & 1.6 & 1.6 & 1.2 & 1.4 & 2.2 & 1.7 \\
-%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
-
-\hline
-\multicolumn{8}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
-Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.5 & 1.1 & 1.0  \\
-Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.6 & 0.4 & 0.5  \\
-%\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
-\hline
-\# observations & 14,955 & 16,118 & 24,958 & 35,027 & 36,990 & 40,279 & 39,351  \\
-\hline\hline
-\multicolumn{8}{l}{\parbox[l]{15cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of FAS price.}}}
-\end{tabular}%
-\end{center}
-\label{tab:result_air_3d_detail}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Transport costs estimates, 3-digit level (selected years)}
-\begin{center}
-    \begin{tabular}{l|ccccccc}
-   \hline\hline
-Year         & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
- \hline
-   \multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}  \\
-   \hline
-Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 5.4 & 4.0 & 3.6  \\
-Median & 9.6 & 5.5 & 4.6 & 4.9 & 5.1  & 3.6 & 3.3  \\
-%Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
-\hline
-\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{8}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} } \\
-\hline
-Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 2.7 & 1.9 & 2.2 \\
-Median & 4.9 & 2.4 & 2.8 & 2.1 & 2.8 & 1.8 & 1.8  \\
-%Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
-\hline
-\multicolumn{8}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
-\hline
-Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.9 & 2.5 & 1.5  \\
-Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 1.9 & 0.8 \\
-%Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
-\hline
- \# observations & 19,007 & 17,356 & 28,383 & 36,090 & 37,757 & 37,748 & 38,473 \\
-\hline\hline
-\multicolumn{8}{l}{\parbox[l]{15cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of FAS price.}}}
-\end{tabular}%
-\end{center} \label{tab:result_ves_3d_detail}
-\end{table}%
-
-\subsection{Assessing the importance of additive transport costs \label{app:diagnostic_test}}
-
-In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed CIF-FAS prices gap, in order to deliver a more systematic diagnosis of the importance of additive costs.
-To do so, we rely on several standard measures of fit.
-The first indicator is through comparing $R^{2}$.
-However, its use is far from being straightforward when evaluating non-linear estimates.
-$R^2$ is based on the underlying assumption that the adjusted model is a linear one.
-In a non-linear context, $R^2$ is strictly speaking inappropriate.
-However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.
-This leads us to complement the goodness-of-fit diagnosis with three alternative measures.
-We provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line.
-The smaller the SER value, the better the quality of fit, as it indicates that the observations are closer to the fitted line.
-We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test.
-A decrease in the log-likelihood function points to a better quality of fit.
-However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.
-\footnote{Specifically, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.}
-The preferred model is the one with the minimum AIC value.
-Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including the additive term, i.e.
-Equation (\ref{eq:model_IetA})) and the Restricted one (\emph{R}, i.e.
-Equation (\ref{eq:model_nlI})).
-The null tested is that the two models are statistically equivalent.
-Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for air and vessel respectively, at the 3-digit level.
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of Goodness-of-fit (3 digits)}
-  \footnotesize{
-\begin{center}
-    \begin{tabular}{l|cccccc|c}
-    \hline \hline
-    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
-    \multicolumn{8}{l}{\textbf{$R^2$} }\\ \hline
-    Model (A)& 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
-    Model (B) & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
-    \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
-     Model (A) & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
-    Model (B) & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
-   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-    Model (A) & 35675.0 & 41171.0 & 60715.6 & 87492.6 & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{88191.9} & 70498.1 \\
-    Model (B)  & 31387.3 & 35738.4 & 52098.9 & 74954.9 & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{80873.7} & 62285.0 \\ \hline
-    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
-    Model (A) & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
-    Model (B) & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
-    LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
-    Nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
-    p-value & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
-    \hline \hline
- \end{tabular}%
-    \end{center}}
-  \label{tab:good_fit_air}%
- \parbox[l]{15cm}{\footnotesize{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& ad-valorem transport costs.
-SER = Standard Error of regression; AIC = Akaike Information Criterion.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.
-The mean statistics are calculated as the average value over all years.
-}}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Measures of Goodness-of-fit (3 digits)}
-    \footnotesize{
-\begin{center}
-\begin{tabular}{l|cccccc|c}
-\hline \hline
-Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
-\multicolumn{8}{l}{\textbf{$R^2$} }\\ \hline
-Model (A)& \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & 0.35 & \multicolumn{1}{c}{0.34} & 0.39 \\
-  Model (B) & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & 0.49 & \multicolumn{1}{c}{0.46} & 0.56 \\ \hline
-\multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
-    Model (A) & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
-    Model (B) & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
-   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-   Model (A) & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
-   Model (B) & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
-    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
-    Model (A) & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
-    Model (B) & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
-    LL ratio & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
-    nb of restrictions & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
-    p-value& \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
-    \hline \hline
-    \end{tabular}%
-    \end{center}}
-  \label{tab:good_fit_vessel}%
-  \parbox[l]{15cm}{\footnotesize{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& ad-valorem transport costs.
-SER = Standard Error of regression; AIC = Akaike Information Criterion.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.
-The mean statistics are calculated as the average value over all years.
-}}
-\end{table}%
-
-Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
-On average over the whole period, the $R^{2}$ doubles when per-kg costs are included for air, and increases by 50\% for vessel.
-Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
-Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes.
-Both AIC and log-likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
-These results holds whatever the considered year.\smallskip
-
-For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Appendix \ref{app:4digit}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
-If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.
-This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
-When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
-In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
-Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.
-
-\subsection{Variance decomposition exercise \label{app:decomp_variance}}
-
-In this section, we provide a variance decomposition exercise on the observed CIF-FAS price gap.
-Specifically, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$) and \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
-The variance decomposition expression at the product level $k$ (5-digit level) is obtained by applying the following formula (by year and transport mode):
-
-$$\underbrace{\sum_{k=1}^K \sum_{i=1}^I \left(x_{ik} - \bar{x}_g  \right)^2}_{\text{Total variability}} = \underbrace{\sum_k \sum_i \left(x_{ik} - \bar{x}_k  \right)^2}_{\text{Within-product variability}} + \underbrace{\sum_k \left(\bar{x}_{k} - \bar{x}_g  \right)^2}_{\text{Between-product variability}}$$
-
-with $x$ the observed CIF-FAS price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
-
-\begin{eqnarray*}
-\bar{x}_g \equiv \frac{1}{n} \sum_{k=1}^K \sum_{i=1}^I x_{ik},&& \bar{x}_k \equiv \frac{1}{K}\sum_{k=1}^K x_{ik}
-\end{eqnarray*}
-
-\noindent with $n$ the total number of observations, $K$ the total number of products and $I$ the total number of country partners.
-We apply the same variance decomposition exercise at the sector level, in which case the sector $s$ index (at the 3-digit level) replaces the $k$ index (at the 5-digit level).
-This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
-We also determine the share of the observed variance that can be attributed to the between-country variance, adapting the variance decomposition formula written above accordingly.
-Results are reported in Figure \ref{fig:decomp_variance}.
-
-\begin{figure}[htbp]
-\caption{Variance decomposition (observed CIF-FAS price gap)}
-\label{fig:decomp_variance}
-\begin{center}
-\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
-Two interesting results emerge from Figure \ref{fig:decomp_variance}.
-First, the share of the CIF-FAS price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
-Both account for between 30 and 40\% of the total variance in air transport, depending on the years considered.
-This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
-This provides an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
-Second, the variance of the CIF-FAS price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
-This holds throughout the period and for both transport modes.
-This suggests that what primarily matters in international transport costs is mostly attributable to the product \textit{per se}, rather than to the country where it comes from.
-%JH: un peu too much, non ? "By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs."
-
-\clearpage
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
-\renewcommand{\thetable}{C.\arabic{table}}
-
-\section{Estimation at the 4-digit level \label{app:4digit}}
-
-In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
-
-\subsection{Transport cost estimates}
-
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively.
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Transport costs estimates, selected years, 4-digit level}
-\begin{center}
-  \footnotesize{
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
-\hline
-Mean  & 6.6 & 5.8 & 5.2 & 3.3 & 3.7 & 3.2 \\
-Median & 5.2 & 4.4 & 4.1 & 2.1 & 2.7 & 2.6 \\
-%Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}  \\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
-Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
-Median & 2.5 & 1.7 & 1.9 & 1.0 & 1.7 & 1.4  \\
-%Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$), in \%}    \\ \hline
-Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
-Median & 1.2 & 0.6 & 0.6 & 0.5 & 0.4 & 0.4  \\
-%Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\
-\hline
-\# observations & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of FAS price.}}}
-\end{tabular}%
-}
-\end{center}
-  \label{tab:result_air_rob}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
-\begin{center}
-  \footnotesize{
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)} \\
-\hline
-Mean  & 9.8 & 6.1 & 5.8 & 5.1 & 4.2 & 3.6  \\
-Median & 9.4 & 5.1 & 4.8 & 4.5 & 3.8 & 3.1 \\
-%Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC} }\\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
-Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
-Median & 4.9 & 3.0 & 2.4 & 2.5 & 2.6 & 1.8 \\
-%Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}^{add}/\widetilde{p}$, in \%) }   \\ \hline
-Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\
-Median & 2.9 & 1.3 & 1.9 & 1.5 & 1.3 & 0.8 \\
-%Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\
-\hline
-\# observations & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of FAS price.}}}
-\end{tabular}%
-}
-\end{center}
-\label{tab:result_ves_rob}%
-\end{table}%
-
-\subsection{Goodness-of-fit tests at the 4-digit level}
-
-We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
-The results are reported in Tables \ref{tab:good_fit_air_rob} (for air) and \ref{tab:good_fit_ves_rob} (for vessel).
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of goodness of fit, 4-digit level}
-\begin{center}
-\label{tab:good_fit_air_rob}%
-%\vspace{0.5cm}
-%\scalebox{0.97}{
-  \footnotesize{
-\begin{tabular}{l|cccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}              \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
-\hline
-\multicolumn{7}{l}{\textbf{R$^2$}}      \\ \hline
-Model (a)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
-\multicolumn{7}{l}{\textbf{SER} }  \\ \hline
-Model (a)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Model (b) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
-\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (a) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
-Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
-\multicolumn{7}{l}{\textbf{AIC criteria} }  \\ \hline
-Model (a)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
-Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
-\multicolumn{7}{l}{\textbf{Test LL} }    \\ \hline
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
-\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
-p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
-Model (b) = with additive \& ad-valorem
-transport costs.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
-\end{tabular}%
-}
-\end{center}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Measures of goodness of fit, 4-digit level}
-\begin{center}
-\label{tab:good_fit_ves_rob}%
-  \footnotesize{
-\begin{tabular}{l|cccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                   \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
-\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
-Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
-\multicolumn{7}{l}{\textbf{SER} }\\ \hline
-Model (a) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
-\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (a)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
-Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
-\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
-Model (a) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
-Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
-\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
-\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
-p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
-Model (b) = with additive \& ad-valorem
-transport costs.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
-\end{tabular}%
-}
-\end{center}
-\end{table}%
-
-\newpage
-
-\clearpage
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
-\renewcommand{\thetable}{D.\arabic{table}}
-
-\section{Eliminating the composition effects: more details \label{app:comp-effects}}
-
-In this section, we explain in more detail the method employed to eliminate the country and product dimensions of the estimated transport cost.
-
-\subsection{More on our methodology}
-
-\subsubsection{Additive and ad-valorem transport costs: excluding the composition effects}
-
-In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for both the ad-valorem and the additive component.
-
-\paragraph{For the ad-valorem component} - Consider first the multiplicative transport cost component.
-Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
-
-\begin{equation*}
-\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbbm{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbbm{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
-\end{equation*}
-
-Based on this equation, we deduce after estimation that:
-
-\[\left\{
-  \begin{array}{lcl}
-  \text{For the year 1974:} &&  \widehat{\tau}_{is74} = \exp(\delta +\alpha_i+\beta_s), \\
-  \text{For any year }t> 1974:&& \widehat{\tau}_{ist} = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t) \\
-  \end{array}
-\right.\]
-
-From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
-Given that $\tau >1$, we can rewrite to get the percentage change between year 1974 and any year $t>1974$:
-\begin{equation*}
-\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}
-\end{equation*}
-
-As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trend.
-At this stage though, it remains specific to a product-origin country pair.
-Next step is to build the index $\Gamma^{adv}_t$ such that:
-\begin{equation}
- \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}  \label{eq:tcadv_compoeffect}
-\end{equation}
-\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
-
-\paragraph{For the additive component} - After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
-\[\left\{
-  \begin{array}{lcl}
-\text{For the year 1974:}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
-\text{For any year}~t> 1974:&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
-\end{array}
-\right.\]
-
-From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
-Given the constraint $t>0$, we then obtain the percentage change from 1974 from:
-
-\begin{equation}
-\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t) \label{eq:indice_add}
-\end{equation}
-
-\noindent Note that it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
-\begin{equation}
-\Gamma^{add}_t  = 100\exp(\gamma_t) \label{eq:tcadd_compoeffect}
-\end{equation}
-
-\subsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
-
-In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
-In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' Stata codes provided on his webpage\footnote{\url{http://www.krannert.purdue.edu/faculty/hummelsd/research/jep-transport-cost-data.php}}), to better identify the precise points of difference between our estimation strategies.
-
-Let us start from Hummels' (2007) quotation (p.
-146) according to which (for air), the ``\textit{unadjusted measure of ad valorem air shipping costs [is] the aggregate expenditures on air shipping divided by the value of airborne imports}''.
-From this, we get the ``raw'' ad-valorem measure of transport costs as the ratio between the imported total value and the exported total value ($(p_{ikt} - \widetilde{p}_{ikt})q_{ikt}/\widetilde{p}_{ikt}q_{ikt})$, the mean yearly value being obtained by weighting each flow by its value in total trade flows (ie, yielding the observed aggregate value of $\tau_t$ for air).
-
-Consider now the ``fitted ad-valorem rate''.
-\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
-The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables}''.
-With the value of the shipment equal to $\widetilde{p}_{ikt}q_{ikt}$, and denoting the air freight cost as $TC_{ikt}$, we can formulate the above sentence according to the following equation (suppressing the transport mode index to alleviate notation):
-
-\begin{eqnarray}
-&&\ln TC_{ikt} = \delta+ \beta \ln \frac{q_{ikt}}{\widetilde{p}_{ikt}q_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t} + \epsilon_{ikt} \notag \\
-\Leftrightarrow && \ln TC_{ikt} = \delta+ \beta \ln \frac{1}{\widetilde{p}_{ikt}} + \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t} + \epsilon_{ikt} \label{eq:Hummels1}
-\end{eqnarray}
-
-\noindent where $TC_{ikt}$ is measured by the ratio between (air) charges, i.e.
-$(p_{ikt} - \widetilde{p}_{ikt})q^{fas}_{ikt}$ and the value of the shipment, i.e.
-$\widetilde{p}_{ikt}q^{fas}_{ikt}$, or equivalently : $TC_{ikt} = \frac{p_{ikt} - \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}$.\medskip
-
-After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
-
-What are the main differences with our estimation strategy? Five points may be underlined.
-First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
-However, this is a slight difference, as in both cases it ultimately amounts to taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-Second, in contrast to us, \cite{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
-This is not likely to make a substantial difference, however, as these fixed effects are by nature constant over time. Accordingly, they only amount to a scale effect.
-%Accordingly, \textbf{they only amount to a scale effect. - rather than \textit{they only play as a scale effect as the estimate is in log}. A-t-on besoin de dire que c'est en log?}.
-
-A third difference is that we separate the country-product fixed effects ($\sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}$ in Equation (\ref{eq:Hummels1})) into two separate components (origin country and product dimensions).
-As discussed in Section \ref{sec:robustness}, this is constrained by the number of fixed effects to estimate.
-However, we do not view this as the major cause of difference between our two methods, as can be inferred from the robustness analysis on this point in Section \ref{sec:robustness}.
-
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeh{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
-\cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
-This choice is made for two reasons.
-First, this weighting scheme does not overweight the small flows in value.
-Second, the 1974 weighting scheme makes sense as the starting point of our time trend analysis.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in Section \ref{sec:robustness}.
-
-One last difference remains to be commented, concerning the way the additive component of international transport costs is treated.
-As we show below, it turns out to be key in accounting for the difference of results with \cite{hummels2007}.
-Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
-Rewriting Equation (\ref{eq:Hummels1}) in a simpler form as:
-
-$$\ln TC_{ikt} \equiv \ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt}- \beta \ln \widetilde{p}_{ikt} $$
-
-\noindent with $\Delta_{ikt} \equiv \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t}$ agglomerates the various fixed effects for reading clarity.
-
-Under the first polar case with $\beta = 1$, it becomes:
-\begin{eqnarray*}
-&&\ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt}- \ln \widetilde{p}_{ikt} \\
-\Leftrightarrow && \ln (p_{ikt}- \widetilde{p}_{ikt}) = \Delta_{ikt} \\
-\end{eqnarray*}
-\noindent Equivalently, we can write:
-
-\begin{eqnarray*}
-&&p_{ikt} = \widetilde{p}_{ikt} + \exp(\Delta_{ikt})\\
-\Leftrightarrow && p_{ikt} = \widetilde{p}_{ikt} + t_{ikt}
-\end{eqnarray*}
-\noindent with $t_{ikt}$ appropriately defined.
-This exactly corresponds to the case of additive costs only.
-
-In the second polar case with $\beta = 0$, Equation (\ref{eq:Hummels1}) becomes:
-\begin{eqnarray*}
-&&\ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt} \\
-\Leftrightarrow && \ln \left(\frac{p_{ikt}}{\widetilde{p}_{ikt}}-1\right) = \Delta_{ikt}\\
-\Leftrightarrow && \frac{p_{ikt}}{\widetilde{p}_{ikt}}  = \exp (\Delta_{ikt})+1
-\end{eqnarray*}
-\noindent that is,
-
-$$p_{ikt} = \tau_{ikt} \widetilde{p}_{ikt}$$
-
-\noindent with $\tau_{ikt}$ appropriately defined.
-This exactly corresponds to the case of ad-valorem transport costs only.
-
-As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
-By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
-This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (i.e.
-allowing for a varying $\beta$ across time, country partner and product).
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007}, outlined above, on our database (which is the same as his until 2004).
-The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-
 \end{document}

### 2020-06-16 redaction/submission_JEGeo/revision/revised_article/DHP_JoEG_rev1.tex
debut de redaction section "endogeneity concerns"

@@
 
-One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
-Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
-This is not an issue here, as we are not interested in causal inference.
-Rather, our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
+%One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+%Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
+%This is not an issue here, as we are not interested in causal inference.
+%Rather, our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
 
 %However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
-Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
-In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
+%Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
+%In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
 
@@
 
+\paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
+Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+
+It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
+However, the latter can be biased by the abovementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+
+XXX propose firss stage equation, explain IVs XXX ICI XXX
+
+XXX provide results 2nd stage
+
+%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
+%Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
+%In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
+
 \section{Transport costs time trends: the role of the additive component}\label{sec:results_trends}

### 2020-06-17 redaction/submission_JEGeo/revision/revised_article/DHP_JoEG_rev1.tex
modif prg FS + section endo
Modification prog first stage pour avoir tableau de stat desc en latex
Premier draft de la sous-section "Endogeneity concerns", pour le moment placée en fin de section 2

@@
 Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
 
 It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
-However, the latter can be biased by the abovementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
-
-XXX propose firss stage equation, explain IVs XXX ICI XXX
-
-XXX provide results 2nd stage
+However, the latter can be biased by the aforementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+This may ultimately harm the estimation of $\widehat{t}/\widetilde{p}$ and consequently, $\widehat{\tau}$.
+
+Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure.
+Put differently, we will replace in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from the following equation:
+
+\begin{equation}
+\ln\left(\widetilde{p}_{ik}\right)= \beta IV_{ik} +\lambda_{is(k)}+ \mu_{ik} \label{eq:first_stage}
+\end{equation}
+
+\noindent where $\lambda_{is(k)}$ represents country-3-digit sector fixed effects, and $IV_{ik} $ the instrumental variable used to purge $\widetilde{p}_{ik}$ from endogeneity issues.
+Following common practice in the literature, we will use as instrument 5-digit product-specific tariff rates, or alternatively, a combination between the latter and lagged fas prices (XXX TO BE DONE XXX).
+Consistently with our overall estimation strategy, Equation (\ref{eq:first_stage}) is estimated on a yearly basis.
+
+Tables \ref{tab: FS_air} and \ref{tab: FS_ves} report summary results of these first stage regressions:
+
+\begin{table}[htbp]\centering
+\caption{Summary Results for First-Stage Regressions - Air\label{tab: FS_air}}
+\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
+\textbf{variable } & \textbf{mean} & \textbf{p25} & \textbf{p50} & \textbf{p75} & \textbf{sd} & \textbf{min} & \textbf{max} \\
+\hline
+   $\beta_{IV}$ &    -0.0858 &    -0.1247 &    -0.0944 &    -0.0362 &     0.0469 &    -0.1741 &    -0.0128 \\
+        $sd_{\beta}$  &     0.0184 &     0.0139 &     0.0197 &     0.0237 &     0.0082 &     0.0048 &     0.0331 \\
+  Student t  &    -5.2255 &    -6.5428 &    -5.2703 &    -3.7943 &     2.4782 &   -11.4557 &    -0.4184 \\
+    F stat  &    33.2938 &    14.4215 &    27.7801 &    42.8707 &    28.1280 &     0.1751 &   131.2323 \\
+  adj R²  &     0.4845 &     0.4712 &     0.4768 &     0.4934 &     0.0227 &     0.4452 &     0.5472 \\
+  R² within &     0.0025 &     0.0015 &     0.0023 &     0.0036 &     0.0016 &     0.0000 &     0.0064 \\
+\hline
+\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} }}
+\end{tabular}
+\end{table}
+
+\begin{table}[htbp]\centering
+\caption{Summary Results for First-Stage Regressions -Ves\label{tab: FS_ves} }
+\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
+\textbf{variable } & \textbf{      mean} & \textbf{       p25} & \textbf{ p50} & \textbf{p75} & \textbf{sd} & \textbf{       min} & \textbf{       max} \\
+\hline
+   $\beta_{IV}$ &     0.0584 &     0.0180 &     0.0733 &     0.0846 &     0.0349 &    -0.0010 &     0.1147 \\
+       $sd_{\beta}$  &     0.0124 &     0.0108 &     0.0138 &     0.0155 &     0.0047 &     0.0041 &     0.0195 \\
+  Student t  &     4.5161 &     3.3592 &     4.8718 &     5.5721 &     1.9626 &    -0.0781 &     8.9563 \\
+    F stat  &    24.1509 &    11.4029 &    23.7486 &    31.0489 &    17.2397 &     0.0061 &    80.2158 \\
+  adj R² &     0.5856 &     0.5767 &     0.5878 &     0.5960 &     0.0172 &     0.5364 &     0.6323 \\
+  R² within &     0.0020 &     0.0006 &     0.0021 &     0.0032 &     0.0015 &     0.0000 &     0.0062 \\
+\hline
+\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} }}
+\end{tabular}
+\end{table}
+
+Both Tables suggest that custom duties are a very good predictor of fas prices $\widetilde{p}_{ik}$, with strong significance.
+The F-statistic for significance of the instrument in the first-stage is on average well above 10, the ``rule-of-thumb'' threshold value recommended by \citet{staiger_stock} as a test for weak instrumentation.
+
+XXX on peut aussi ajouter des graphs des beta estimés ici, peut-être éventuellement avec les IC XXX
+
+XXX Reste ici l'épineux problème de la différence de signe entre les deux modes de transport; avant de creuser, voir si on ne résoudrait pas le problème avec l'IV combinant duties et lagged prices XXX
+
+Provided that the instrument used in Equation \ref{eq:first_stage} displays the right statistical properties, we can confidently re-inject the predictions it produces for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation.
+Results are displayed below, as well as a measure of the gap with our main estimates.
+
+XXX We expect the gap to be very small XXX
 

### 2020-06-23 redaction/topo_first_stage.tex
Topo first stage

\documentclass[11pt,twoside, authoryear]{elsarticle}
\usepackage[frenchb, english]{babel}
\usepackage[utf8]{inputenc}
%\usepackage[utf8]{fontenc}
%\usepackage[ansinew]{inputenc}
\usepackage{setspace}
\usepackage{babel,varioref}
\usepackage{supertabular}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsExtensions{.pdf,.mps,.png,.jpg,.eps}
\usepackage{lscape}
\usepackage[authoryear]{natbib}
%\usepackage[frenchb,english]{babel}
\usepackage{multirow}
\usepackage{ifthen}
%\usepackage{harvard}
\usepackage{vmargin}
\usepackage{verbatim}
\usepackage{array}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=Black, citecolor=blue, urlcolor=blue}
%\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{0cm}
\usepackage{booktabs,caption,fixltx2e}

\usepackage[flushleft]{threeparttable}

%\usepackage{titling}

%\setlength{\droptitle}{-10em}   % This is your set screw

\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
\usepackage{adjustbox}
\usepackage{chngcntr}
\counterwithin{table}{section}
%\usepackage[round]{natbib}

%\renewcommand{\thesection}{\Alph{section}}

% \newcommand\section[1]{%
%   \refstepcounter{section}%
%   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
%   \sectionmark{#1}}
%   }

% % \refstepcounter{section}%
% %   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
% %   \sectionmark{#1}}

%   \newcommand\subsection[1]{%
%   \refstepcounter{subsection}%
%   \addcontentsline{toc}{subsection}{\protect\numberline{\thesubsection}#1}%
%   %\subsectionmark{#1}
%   }

%\EnableSectionsInLOFT

\begin{document}

\section{Some insights about IV specification}

\subsection{Reasoning 1}

The question we deal with, is the functional form of the first stage equation, where we instrument the fas price with duties as suggested by Referee 1.

Denoting $i$ the origin country, $k$ the product at the HS6 level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :

$$\widetilde{p}_{ik} = (1+\tau^d_{is(k)})\bar{p}_{ik}$$

\noindent with $s(k)$ the 3-digit classification (the disaggregation level for tax duties in the database) as a function of the HS6 product classification. Then, taking the total differential around some reference point (referred with the subscript $0$):

\begin{eqnarray}
&&\Delta \widetilde{p}_{ik} = \Delta \bar{p}_{ik}\times (1+\tau^d_{is(k)0}) + \bar{p}_{ik0}\times \Delta \tau^d_{is(k)} \\
\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ik}}{\widetilde{p}_{ik0}} = \Delta \bar{p}_{ik} \frac{1+\tau^d_{is(k)0}}{\bar{p}_{ik0}(1+\tau^d_{is(k)0})} +\frac{\Delta \tau^d_{is(k)0}}{1+\tau_{is(k)0}^d} \\
\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ik0}}{\widetilde{p}_{ik0}} =  \frac{\Delta \bar{p}_{ik}}{\bar{p}_{ik0}} +\frac{\Delta \tau^d_{is(k)}}{1+\tau_{ik0}^d} \label{eq:firststage}
\end{eqnarray}

The intuition behind the Referee 1's endogeneity concerns, is that we need to eliminate from the fas price, any endogenous component that might me related to transport costs. To do so, the referee suggests that we instrument the export price by tariff rates. Put it in plain words, this suggests to run the first-stage regression based on Equation (\ref{eq:firststage}) according to:

$$\frac{\Delta \widetilde{p}_{ik}}{\widetilde{p}_{ik0}} = \alpha \frac{\Delta \tau^d_{is(k)}}{1+\tau_{is(k)0}^d} + \gamma_{i} +\gamma_{k}+\epsilon_{ik}$$

Or equivalently, taking logs:
$$\Delta \log \widetilde{p}_{ik}= \alpha\frac{\Delta \tau^d_{is(k)}}{1+\tau_{is(k)0}^d} +\gamma_{i} +\gamma_{k}+\epsilon_{ik}$$

\noindent with the LHS being the growth rate of the fas price, the first term in the RHS the hange in duty tax rates, the second and third terms fixed effects to capture the ``firm-specific price'' $\bar{p}$, $\epsilon_{ik}$ being the residual.

If this reasoning is correct,
\begin{itemize}
\item we should take as predicted value only the predicted price WITHOUT the fixed effects:
$$\widehat{\frac{\Delta \widetilde{p}_{ik}}{\widetilde{p}_{ik0}}} = \alpha\frac{\Delta \tau^d_{is(k)}}{1+\tau_{is(k)0}^d} $$
\item and we should expect a positive sign between $\Delta \log \widetilde{p}_{ik}$ and $\frac{\Delta \tau^d_{is(k)}}{1+\tau_{is(k)0}^d} $, with the estimated value of $\alpha$ close to 1.
\end{itemize}

Once this is done, rebuilt the instrumented fas price at time $t$ (for clarity, we re-write te temporal index here):

\begin{eqnarray*}
\widehat{\widetilde{p}}_{ikt} = \widetilde{p}_{ikt-1}+ \widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}}
\end{eqnarray*}

with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}}$ the predicted growth rate of the export price based on duty changes.

\subsection{Reasoning 2}

If we come back to my initial idea when discussing together last week, I started from the link between the cif price, the fas price and tax duty.
Denoting $i$ the origin country, $k$ the product at the HS6 level, and reasoning on a yearly basis,

$$p_{ik} = (1+\tau^d_{is(k)})\widetilde{p}_{ik}$$

Taking the total differential around some reference point (referred with the subscript $0$):

\begin{eqnarray*}
&&\Delta p_{ik} = (1+\tau^d_{is(k)0})\Delta \widetilde{p}_{ik} + \widetilde{p}_{ik0}\Delta \tau^d_{is(k)}\\
\Leftrightarrow && \frac{\Delta p_{ik}}{p_{ik0}} = \frac{\Delta \widetilde{p}_{ik}}{\widetilde{p}_{ik0}} + \frac{\Delta \tau^d_{is(k)}}{1+\tau^d_{is(k)0}}
\end{eqnarray*}

The reasoning I held last time was the following: Suppose that the firm does adjust its fas price to duty changes, this mitigates the impact of duty tax changes on the import price. In the extreme case, where the firm fully compensates such that $\frac{\Delta p_{ik}}{p_{ik0}} = 0$, we should get:

$$\frac{\Delta \widetilde{p}_{ik}}{\widetilde{p}_{ik0}} = - \frac{\Delta \tau^d_{is(k)}}{1+\tau^d_{is(k)0}}$$

In this reasoning, we should regress the growth rate of the fas price on the change in tax duty, expecting a negative sign. But if the idea is to take the exogenous component of the fas price, then I am a little bit lost... Shouldn't we take the part of the fas price that is NOT explained by duty changes?

\end{document}

### 2020-06-24 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
rédaction réponse référé 1, Lise

\documentclass[a4paper,12pt]{article}

%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
%J'ai essayé de débuger : je n'y suis pas arrivé
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{amsfonts}
\usepackage{bbold}
\usepackage{bbm}
\usepackage[flushleft]{threeparttable}
%\usepackage{bbold}
%\usepackage[T1]{fontenc}
% JH : impossible de compiler avec le package bbold, remplacé par amsfonts
% LP : impossible de compiler avec le package bbold, remplacé par bbm
\usepackage{upgreek}
\usepackage{comment}
\includecomment{commentGD}
\usepackage[draft]{todonotes}
%\usepackage[disable]{todonotes}
%%GDPour cacher les notes, \usepackage[disable]{todonotes}

\usepackage{setspace}

%\usepackage[nolists, figuresfirst]{endfloat}
%\usepackage{sidefloat}

\setlength{\abovecaptionskip}{-8pt}

%\usepackage[usenames]{color}
%\definecolor{grey}{rgb}{0.35,0.35,0.35}
%\definecolor{webdarkblue}{rgb}{0,0,0.4}
%\definecolor{orange}{rgb}{0.7,0.2,0.05}

%\usepackage[pdfcreator={PDFLaTeX}, pdfproducer={PDFLaTeX}, pdfstartview=FitH, pdfpagemode=UseOutlines, pagebackref=false, colorlinks={true},
%citecolor={webdarkblue}, linkcolor={webdarkblue},
%urlcolor={webdarkblue}]{hyperref}

%\addtolength{\oddsidemargin}{-0.4in}
%\addtolength{\evensidemargin}{-0.4in}
%\addtolength{\textwidth}{0.8in} \addtolength{\topmargin}{-0.85in}
%\addtolength{\textheight}{1.7in}
\renewcommand{\baselinestretch}{1.2}

\newcommand\cites[1]{\citeauthor{#1}'s\ (\citeyear{#1})}

\newcommand\citeh[1]{\citeauthor{#1}'\ (\citeyear{#1})}

\setmarginsrb{3cm}{2cm}{3cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}

% commands

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}
\newcommand{\beqa}{\begin{eqnarray}}
\newcommand{\eeqa}{\end{eqnarray}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bs}{\bigskip}
\newcommand{\A}{$^a$}
\newcommand{\B}{$^b$}
\newcommand{\C}{$^c$}

\renewcommand{\baselinestretch}{1.2}
%\doublespacing

%\linespread{1.6}

%\newcommand{\note}[1]{\footnote{\begin{doublespace}#1\end{doublespace}}}

\begin{document}

\title{\textsc{International Transport costs:\\New Findings from modeling additive costs} \\
Asnwer to Referee 1}

\author{Guillaume \textsc{Daudin}\thanks{%
University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }

\date{July 2020}
 \maketitle
\bigskip

\section{Critique 1: Empirical Strategy}

\subsection{Deriving the share of additive costs in total transport costs \label{ssec:interpret_beta}}
With $i$ the origin country, $k$ the product category at the HS6 level (and reasoning at the year- transport mode level), $p_{ik}$ the import (cif) price and $\tilde{p}_{ik}$ the export (fas) price, transport costs $f_{ik}\equiv \frac{p_{ik}- \tilde{p}_{ik}}{\tilde{p}_{ik}} $ are written in Hummel's terminology as:

\begin{equation}
f_{ik} = X_{is(k)}\tilde{p}_{ik}^{-\beta_{ik}} \label{eq:Hummels}
\end{equation}

It is trivial to show that:
$$\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} \frac{\tilde{p}_{ik}}{f_{ik}}= -\beta_{ik}$$

Then
\begin{itemize}
\item If $\beta_{ik} = 0$, then $p_{ik} = (1+X_{is(k)})\tilde{p}_{ik}$: Only ad-valorem transport costs
\item If $\beta_{ik} = 1$, then $p_{ik}=\tilde{p}_{ik}+X_{is(k)}$: Only additive costs
\end{itemize}

This suggests that, the closer $\beta_{ik}$ to 1, the more prevalent additive costs in total transport costs. Can $\beta_{ik}$ be related to the share of additive costs in total transport costs? The answer is positive. To show this clearly, start from:

\begin{eqnarray*}
f_{ik} &=& \frac{p_{ik}-\tilde{p_{ik}}}{\tilde{p}_{ik}}\\
\text{with}:&& p_{ik} = \tau_{is(k)}\tilde{p}_{ik} + t_{is(k)} \\
\end{eqnarray*}

assuming that transport costs decompose in two components, additive transport costs $t_{is(k)}$ and multiplicative costs (ad-valorem) $\tau_{is(k)}$, supposed to be identical in the product-dimension for any $k$ product within the 3-digit classification $s$ it belongs to.

This gives:

$$f_{ik} = \frac{(\tau_{is(k)}-1)\tilde{p}_{ik}+t_{is(k)}}{\tilde{p}_{ik}}$$

From this, we have

\begin{eqnarray*}
\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} &=& \frac{(\tau_{is(k)}-1)\tilde{p}_{ik} - ((\tau_{is(k)}-1)\tilde{p}_{ik} +t_{is(k)})}{\tilde{p}_{ik}^2} \\
&=& \frac{-t_{is(k)}}{\tilde{p}_{ik}^2}
\end{eqnarray*}

Hence
\begin{eqnarray*}
\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} \frac{\tilde{p}_{ik}}{f_{ik}}&=& \frac{-t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{ik} +t_{is(k)}}\\
&=& \frac{-\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 +\frac{t_{ik}}{\tilde{p}_{ik}}}
\end{eqnarray*}

This shows that $\beta_{ik}$ can be interpreted as the share of additive costs in total transport costs:

$$\beta_{ik} = \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$

\subsection{Coming back to the Referee's 1 concern}

In italics, I report the Referee's comment. Into brackets in bold characters inserted into the referee's statement, my comment.

\textit{Using the notation of the authors, they are intreated
in identifying the share of the specific cost in the total transport cost.} \textbf{\[In fact, he/she forgets that with our identification strategy, we do not only identify the share of additive costs, but also their level. This might be a route to follow for the ``big picture part''\]}.

\textit{ Namely,}
\begin{eqnarray*}
&& \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} \\
\text{or,} &&\frac{t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{is(k)} + t_{is(k)}}
\end{eqnarray*}

\textit{The way they approach the problem is that they assume that (a) $\tau_{ik} = \tau_i\tau_{k}$,
(b) $t_{ik} = t_i +t_k$, and (c) $t_k$ and $\tau_k$ are uniform across products within industry
s. After imposing these assumptions, they estimate the following specification}

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
\end{equation}

\textit{in which $\tau_i$, $\tau_{s(k)}$, $t_i$, and $t_{s(k)}$ are identified as fixed effects coefficients.
In my opinion this choice of strategy is quite sub-optimal, as (i) it relies on
the strong assumptions highlighted above, (ii) it is computationally expensive
as noted by the authors on multiple occasions, and (iii) it is subject to an
endogeneity problem, which the authors disregard with one sentence, but which
is rather detrimental in my opinion.}

The three points raised by the referee are all equally important, and we answer separately to each of them below.

\paragraph{(i) On the assumptions underlining our empirical approach}
Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla (the Review of Economics and Statistics, 2015) to estimate the share of additive costs in a firm-level context. It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''. More importantly, this formulation allows us recovering explicitly the \textbf{levels of}, respectively, additive and multiplicative costs. Despite its formal elegance and simplicity, this is not possible in such a straightforward manner with the Hummels' methodology, which only recovers the \textbf{share} of additive in total transport costs. \textbf{elaborate on this in the Big Picture part.}

Nevertheless, to take full account of the referee’s remark, we decided run an additional set of estimations based on the Hummel's methodogy as proposed by the referee, based on Equation (\ref{eq:estimation_ref1}). This is developed in Section XX of this letter.

\textbf{We need to explain why not in the main text, but in the (online?) appendix}

\paragraph{(ii) On the use of non-linear least squares (NLS)}
As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares. However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators. This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e. $\tau \geq 1$ and $t geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative. Without these restrictions, standard linear, least squares estimates often deliver aberrant values with more than mild quality-of-fit [insert examples here]
We do acknowledge this is very important point was not made clear enough in the initial version, and did our best to make the pint clear in the revised version [see page XX : XX insert quote XX].

\paragraph{(iii) Endogeneity problem}
Indeed, this is a very important point. The referee states that, based on theoretical insights by Melitz (Econometrica, 2003) or Baldwin and Harrigan (AER, 2011), $1/\tilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$. In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
We obviously do not question this conceptual issue. However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices. Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out. That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.

Following the referee's advice, we decided therefore to provide some IV-estimates to provide a clean assessment of the size of the potential bias.

XXX At some point, explain HS10 duty do not exist, contrary to what the referee asserts. At best, it was HS6 or HS8 XXX

\subsection{Exploring Referee's alternative functional form}

\textit{A more natural approach is what the authors, at some point, refer to as the
Hummel's Methodology. That is, one can alternatively estimate the share of the
additive component as:}

$$\frac{t_{ik}}{ (\tau_{ik}-1)\tilde{p}_{ik} + t_{ik}} = \beta_{ik}$$

\textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t. unit price} [\textbf{in absolute value, see Equation \ref{eq:Hummels}}]. \textit{Given the authors' objective and the data they are using, $\beta$ can be separately estimated
for each industry-country pair using the following regression:}

\begin{equation}
\ln f_{ikd} = \beta_{is(k)}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
\end{equation}

\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product. The
identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
obviously require that the authors do not aggregate up the raw Census data
across all districts and all 10-digit products pertaining to the same 5-digit category.}

\textit{The first advantage of this so-called Hummel's Approach is that the above
regression can be estimated separately for various country-industry pairs, without
imposing Assumptions (a) and (b), outlined above.}

\textbf{I report here the referee's footnote: Should we deal with this? If he/she is not convinced because small sub-sample, do it on a much larger sample? This would better convince him that the separability assumption is not an issue.}: \textit{The authors do check the robustness of their results w.r.t. the separability assumption, but this only done for a small sub-sample.}

\textbf{There is also a point here, about the degree of sectoral classification. In our paper, $k$ = 5 digit and $s(k)$ = 3 digit. He/she suggests that, in the estimated equation, $k$ = 10 digit and $s(k)$= 5 digit. Which means, having an estimated $\beta$ specific to the origin country-5-digit level. I need to have a refresher on this: What have we done finally? and why?}

XXX Say something about the fact that in the original US-custom data, the finest degree of sectoral classification is 8-digit, not 10? XXX

\subsubsection{What we do}

Our estimated equation is written as:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
\end{equation}

\noindent with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004). We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel). To sum up our methodoly:
\begin{itemize}
\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
\item We rebuild
\begin{eqnarray*}
\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
\end{eqnarray*}
\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:

$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
\end{itemize}

\subsubsection{What the referee suggests to do}

The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007). As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs. The share of aditive costs can hence be encovered by regressing transport costs on unit (ie, fas) prices. Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).

Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector (\textbf{why 3-digit? and not 5 as suggested by the referee?}, exploiting the variability between sub-sectors at the 8 digit level ($k$) and between ports of entry in the US ($d$).

\subsubsection{Strategy of answer}

\paragraph{a) The advantages are not as important as suggested by the referee}
\begin{itemize}
\item[Advantage 1]: One is not compelled to assume separability between the country-component and the product-component of transport costs.
\item[$\Rightarrow$] Answer can be in 2 parts: 1) This is not a strong assumption - better convince him/her: Run the robustness to the separability assumption on a larger sample; 2) Show that the estimated $\beta$ is not that different between the two methods.
\item[Advantage 2]: There is a handful of instruments (HS-8 product-specific tariff rates or lagged prices) to instrument fas prices
\item[$\Rightarrow$] Answer: When we instrument fas prices at the 5-digit level using tax duties (still using our own methodology), we do not obtain much different results. \textbf{But he/she suggests to do it on HS-8 fas prices, not 5-digit prices, and to imply his estimated equation. Have we done that?}
\item[Advantage 3]: A more transparent comparison with Hummel (2007): How to answer to this point? Re-read the paper to see if we are clear enough?
\end{itemize}

\paragraph{b) The costs are high}
\begin{itemize}
\item[Concern 1] The suggested estimation strategy implies having much less data to exploit, for two reasons.
\begin{enumerate}
\item Information about the port of entry are only available in the recent years (\textbf{since when? why? before, we rely on Hummel's data set, but we could buy the original US custom data for 1974-2004? Or even in those dataset it does not exist?}). In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time, and most importantly how the specific values of the additive and ad-valorem transport costs components have evolved over time. Eliminating this dimension of the paper would be detrimental to its value-added.
\item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 8-d sub-sectors and ports of entry. Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors (even at the HS-6 digit classification) or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run. This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table XXX \textbf{TO DO}. Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry. In this respect, the induced selection bias reduces the general scope of our transport costs estimates. For things to be comparable across methods, we have thus re-run our estimation strategy on the same sample of observations as the one used in the referee's method.

    \end{enumerate}
\item[Concern 2] The suggested estimation method implicitly assumes that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors; in both cases, one might view this as a disputable assumption. By contrast, we suppose that transport costs are sector-country specific. \textbf{FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique?}
\item[Concern 3] The suggested estimation method does not allow to recover the values of each component of transport costs separately, but only the relative shares; while our methodology drives an estimate for each component ($\tau$, $t$) from which we can deduce the relative share of each. In this respect, we gain in generality. \textbf{DEEPEN this for the Big picture part}.
\item[Concern 4] The suggested method features less accuracy in the estimation of the $\beta$. Considering the value of the $\beta$ by itself, there is no criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample). Things are more clear-cut in terms of accuracy of the estimation. Specifically, our method yields a more precise estimation of the $\beta$ than the referee's method. To develop on this, for each year and transport mode:
    \begin{itemize}
    \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$. From this, we can deduce the 5- 95\% threshold values through:
        $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$

    \item  With our method, one estimation allows to calculate each transport cost component $\hat{t}_{is}$, $\hat{\tau}_{is}$ and an associated $\beta_{is}$. We generate a distribution of $\beta$ through bootstrap method (10,000 random draws), from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
    \item We can then evaluate the accuracy of each estimation method by comparing the distributions of the $\beta$. This is reported in the Figures XXX (for Air) and XXX (for Vessel). Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered. For sake of brevity, we report the mean distribution averaged over the period, a similar conclusion applies on a yearly basis. \textbf{Averaged over the two transport modes? For only one?}

\begin{figure}[htbp]
\caption{Accuracy of the estimation of the $\beta$: Comparison}
\label{fig:accuracy_beta}
\begin{center}
\includegraphics[height=4in]{accuracy_beta.pdf}
\end{center}
\end{figure}

Figure \ref{fig:accuracy_beta} reports the log of the interval confidence of the $\beta$ obtained with our methodology (with bootstrap), relative to the one obtained with the referee's method. \textbf{average over the years? the transport modes? unclear}. In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.

    \end{itemize}
\end{itemize}

\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}

\end{document}

### 2020-06-26 redaction/submission_JEGeo/revision/referees/answers/topo_first_stage.tex
changement emplacement topo first stage

\documentclass[11pt,twoside, authoryear]{elsarticle}
\usepackage[frenchb, english]{babel}
\usepackage[utf8]{inputenc}
%\usepackage[utf8]{fontenc}
%\usepackage[ansinew]{inputenc}
\usepackage{setspace}
\usepackage{babel,varioref}
\usepackage{supertabular}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsExtensions{.pdf,.mps,.png,.jpg,.eps}
\usepackage{lscape}
\usepackage[authoryear]{natbib}
%\usepackage[frenchb,english]{babel}
\usepackage{multirow}
\usepackage{ifthen}
%\usepackage{harvard}
\usepackage{vmargin}
\usepackage{verbatim}
\usepackage{array}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=Black, citecolor=blue, urlcolor=blue}
%\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{0cm}
\usepackage{booktabs,caption,fixltx2e}

\usepackage[flushleft]{threeparttable}

%\usepackage{titling}

%\setlength{\droptitle}{-10em}   % This is your set screw

\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
\usepackage{adjustbox}
\usepackage{chngcntr}
\counterwithin{table}{section}
%\usepackage[round]{natbib}

%\renewcommand{\thesection}{\Alph{section}}

% \newcommand\section[1]{%
%   \refstepcounter{section}%
%   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
%   \sectionmark{#1}}
%   }

% % \refstepcounter{section}%
% %   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
% %   \sectionmark{#1}}

%   \newcommand\subsection[1]{%
%   \refstepcounter{subsection}%
%   \addcontentsline{toc}{subsection}{\protect\numberline{\thesubsection}#1}%
%   %\subsectionmark{#1}
%   }

%\EnableSectionsInLOFT

\begin{document}

\section{Some insights about IV specification}

\subsection{A note on prices}

We have three prices: the (US) consumer price, say $p^{c}_{ik}$, the ``import'' or  cif price (at the entrance of the US) $p_{ik}$ and the fas price (at the export gate in the origin country) $\widetilde{p}_{ik}$, denoting $i$ the origin country, $k$ the product at the 8 or 5 digit level, and reasoning on a yearly basis, and $s(k)$ the 3-digit classification the product $k$ belongs to. Further, if we denote $\tau^d_{ik}$ the duty tax rate paid when the good crosses the US border, then we have:

\begin{eqnarray*}
&&p^c_{ik} = (1+\tau^d_{is(k)})p_{ik} \\
with && p_{ik}  = \tau_{is(k)} \widetilde{p}_{ik} +t_{is(k)}
\end{eqnarray*}

\subsection{Specification of the first stage equation}

The question we deal with, is the functional form of the first stage equation, where we instrument the fas price with duties as suggested by Referee 1. The idea is that firms might react to changes in duty tax rates which have nothing to do with transport costs changes. In this respect, considering the predicted part of the fas price related to tax duty is likely to solve potential endogeneity biases.

Denoting $i$ the origin country, $k$ the product at the 8/5 digit level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :

\begin{equation}
\widetilde{p}_{ikt} = (1+\tau^d_{is(k)t})^\alpha \left(\bar{p}_{ikt}\right)^\beta \label{eq:link_fas_duty}
\end{equation}

\noindent with $s(k)$ the 3-digit classification (the disaggregation level for tax duties in the database) as a function of the HS8 (or 5 digit) product classification

\subsubsection{First-stage equation in first difference \label{ssec:first_diff}}

One option to specify the functional form of the first stage is to take the total differential around some reference point at time $t-1$, with $\Delta $ the difference operator ($\Delta\widetilde{p}_{ikt} = \widetilde{p}_{ikt} - \widetilde{p}_{ikt-1}$ and so on):
\begin{eqnarray*}
&&\Delta \widetilde{p}_{ikt} = \beta \bar{p}_{ikt-1}^{\beta-1}(1+\tau^d_{is(k)t-1})\Delta \bar{p}_{ikt} + \alpha \bar{p}^\beta_{ikt-1} (1+\tau^d_{ikt-1})^{\alpha-1}\Delta \tau^d_{is(k)t}  \\
\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \beta \frac{\Delta \bar{p}_{ikt}}{\widetilde{p}_{ikt-1}} \frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} +\alpha \frac{\Delta \tau^d_{is(k)t-1}}{1+\tau_{is(k)t-1}^d}\frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} \end{eqnarray*}

leading to:
\begin{equation}
\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} =  \beta \frac{\Delta \bar{p}_{ikt}}{\bar{p}_{ikt-1}} +\alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{ikt-1}^d} \label{eq:firststage_Deltalog}
\end{equation}

The intuition behind the Referee 1's endogeneity concerns, is that we need to eliminate from the fas price, any endogenous component that might me related to transport costs. To do so, the referee suggests that we instrument the export price by tariff rates. Put it in plain words, this suggests to run the first-stage regression based on Equation (\ref{eq:firststage_Deltalog}) according to:

$$\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \alpha \frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \gamma_{i} +\gamma_{k}+\epsilon_{ik}$$

Or equivalently, taking logs:
$$\Delta \log \widetilde{p}_{ikt}= \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{k}+\epsilon_{ikt}$$

\noindent with the LHS being the growth rate of the fas price (between $t$ and $t-1$), the first term in the RHS the change in duty tax rates, the second and third terms fixed effects to capture changes in the ``firm-specific price'' $\bar{p}_{ik}$, $\epsilon_{ik}$ being the residual. Notice though that the structure of fixed effects should remain consistent between the first and the second stages. This implies to rather consider the following functional form of the first-stage equation:

\begin{equation}
\Delta \log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_Deltalog}
\end{equation}

If this reasoning is correct,
\begin{itemize}
\item we should take as predicted value only the predicted price WITHOUT the fixed effects:
$$\widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $$
\item and we should expect a value of the coefficient $\widehat{\alpha}$ between $\Delta \log \widetilde{p}_{ikt}$ and $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $ between $-1$ and $0$, depending on the degree of ``pricing-to-market'' of firms.
    \begin{itemize}
    \item[-] $\alpha = 0$ corresponds to the case where the firm does not adjust its fas price to the change in tax duty, that would cancel out the influence of the tax duty change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition;
    \item[-] $\alpha <0$ corresponds to the degree of``pricing-to-market'' as the firm offsets the impact of the tax change of the price paid by the final consumer by adjusting her producer price in the opposite direction of the tax change,
    \item[-] with $\alpha = -1$ being the extreme case of ``full PTM'' where the firm fully compensates the tax duty change. As shown by Berman, Martin, Mayer (REStats 2012), this is more likely larger firms.
    \end{itemize}
\end{itemize}

Once this is done, we rebuilt the instrumented fas price at time $t$:

\begin{eqnarray*}
\widehat{\widetilde{p}}_{ikt} = \widetilde{p}_{ikt-1}+ \widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}}
\end{eqnarray*}

with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}}$ the predicted growth rate of the fas price based on duty changes.

\subsection{First-stage equation in level}

Alternatively, one can specify the first-stage equation in log levels. Taking Equation (\ref{eq:link_fas_duty}) into log:

$$\log \widetilde{p}_{ikt} =\alpha \log (1+\tau^d_{is(k)t}) + \beta \left(\bar{p}_{ikt}\right) $$

Leading to the following functional form of the first-stage equation (to be run on a yearly basis, by transport mode)

\begin{equation}
\log \widetilde{p}_{ikt} =\alpha \log (1+\tau^d_{is(k)t}) + \gamma_i +\gamma_s +\epsilon_{ikt}  \label{eq:FS_loglevel_cross_section}
\end{equation}

If we rather adopt a panel approach, the first-stage equation is modified by adding a year fixed effect, according to:

\begin{equation}
\log \widetilde{p}_{ikt} =\alpha \log (1+\tau^d_{is(k)t}) + \gamma_i +\gamma_s +\gamma_t +\epsilon_{ikt}  \label{eq:FS_loglevel_panel}
\end{equation}

\subsubsection{Introducing price inertia}

As suggested by the referee, one cannot exclude price inertia, such that current prices depend on past prices; the ``inertia'' part of the fas price being exogenous to current transport costs changes, it could be also considered in the predicted price at the first stage. One can introduce price inertia in Equation (\ref{eq:link_fas_duty}) by considering that the price component depends on the past price $\widetilde{p}_{ikt-1}$ and the ``firm-specific'' component, according to:
$$\widetilde{p}_{ikt} = (1+\tau^d_{is(k)t})^\alpha (\left[\bar{p}_{ikt-1}(\widetilde{p}_{ikt-1}, \omega_{ikt})\right]^\beta $$
Specifically, using a Cobb-Douglas specification
\begin{equation}
\widetilde{p}_{ikt} = \left[1+\tau^d_{is(k)t}\right]^\alpha \widetilde{p}^{\beta_1}_{ikt-1}\omega_{ikt}^{\beta_2} \label{eq:link_fas_duty_inertia}
\end{equation}

Taking Equation (\ref{eq:link_fas_duty_inertia}) into logs:

$$\log \widetilde{p}_{ikt} = \alpha \log\left(1+\tau^d_{is(k)t}\right) + \beta_1 \log \widetilde{p}_{ikt-1}+ \beta_2\log \omega_{ikt}$$

leading to the following specification of the first-stage equation (in cross section):

\begin{eqnarray}
\log \widetilde{p}_{ikt} &=& \alpha \log\left(1+\tau^d_{is(k)t}\right) + \beta_1 \log \widetilde{p}_{ikt-1}+ \gamma_i +\gamma_s +\epsilon_{ikt} \label{eq:FS_loglevel_cross_section_inertia}   \\
\log \widetilde{p}_{ikt} &=& \alpha \log\left(1+\tau^d_{is(k)t}\right) + \beta_1 \log \widetilde{p}_{ikt-1}+ \gamma_i +\gamma_s + \gamma_t+ \epsilon_{ikt} \label{eq:FS_loglevel_cross_section_panel}
\end{eqnarray}

\subsubsection{Obtaining the instrumented fas price}

As last step, we take as instrument the part of the fas price that is predicted by the sum of the tax duty and the past prices components (if considered):

$$\widehat{\log \widetilde{p}}_{ikt} = \widehat{ \alpha} \log\left(1+\tau^d_{is(k)t}\right) +  \widehat{\beta_1}\ \log \widetilde{p}_{ikt-1} $$

\noindent with $ \widehat{ \alpha}$ and possibly $ \widehat{\beta}_1$ the estimated coefficients at the first stage (depending on the specification considered, with or without price inertia, in panel or in cross-section).

In contrast with the differential approach (Section \ref{ssec:first_diff}), this specification has the advantage to directly yield the instrumented price in log, from which we can deduce the fas price in level (which intervenes in the second-stage equation) by taking the exponential.

\end{document}

### 2020-06-26 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
réponse référé 1, Lise, continued

@@
 \textit{Using the notation of the authors, they are intreated
-in identifying the share of the specific cost in the total transport cost.} \textbf{\[In fact, he/she forgets that with our identification strategy, we do not only identify the share of additive costs, but also their level.
-This might be a route to follow for the ``big picture part''\]}.
-
-\textit{ Namely,}
+in identifying the share of the specific cost in the total transport cost. Namely,}
 \begin{eqnarray*}
@@
 It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
-More importantly, this formulation allows us recovering explicitly the \textbf{levels of}, respectively, additive and multiplicative costs.
+\textbf{Not sure of what it means?
+} More importantly, this formulation allows us recovering explicitly the \textbf{levels of}, respectively, additive and multiplicative costs.
 Despite its formal elegance and simplicity, this is not possible in such a straightforward manner with the Hummels' methodology, which only recovers the \textbf{share} of additive in total transport costs.
@@
 This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
-$\tau \geq 1$ and $t geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative.
-Without these restrictions, standard linear, least squares estimates often deliver aberrant values with more than mild quality-of-fit [insert examples here]
-We do acknowledge this is very important point was not made clear enough in the initial version, and did our best to make the pint clear in the revised version [see page XX : XX insert quote XX].
+$\tau \geq 1$ and $t \geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative.
+Without these restrictions, standard linear, least squares estimates often deliver aberrant values with more than mild quality-of-fit \textbf{[insert examples here]}
+We do acknowledge this is a very important point, which was not made clear enough in the initial version, and did our best to make the pint clear in the revised version \textbf{[see page XX : XX insert quote XX].}
 
@@
 Following the referee's advice, we decided therefore to provide some IV-estimates to provide a clean assessment of the size of the potential bias.
+
+\textbf{careful, there is two points regarding endogeneity: (i), we should estimate the $\beta$ according to the referee's method based on the intrumented fas price (at the HS8 level, if tax duty are available at this disaggregation level) - WE HAVE NOT DONE THAT; (ii), if we maintain our empirical strategy (which we do), we should instrument the fas price at the 5-digit level - WHICH WE ARE TRYING TO DO.}
 
@@
 \item[Advantage 1]: One is not compelled to assume separability between the country-component and the product-component of transport costs.
-\item[$\Rightarrow$] Answer can be in 2 parts: 1) This is not a strong assumption - better convince him/her: Run the robustness to the separability assumption on a larger sample; 2) Show that the estimated $\beta$ is not that different between the two methods.
+\item[$\Rightarrow$] Answer can be in 2 parts: 1) This is not a strong assumption - better convince him/her: Run the robustness to the separability assumption on a larger sample and show that the estimated $\beta$ is not that different between the two methods.
+2) Compare the estimated $\beta$ between our original method and the referee's one.
+\textbf{Where are the results comparing the estimates of the $\beta$, in value?
+I don't find these results anymore...}
+
 \item[Advantage 2]: There is a handful of instruments (HS-8 product-specific tariff rates or lagged prices) to instrument fas prices
@@
 Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
-In this respect, the induced selection bias reduces the general scope of our transport costs estimates.
-For things to be comparable across methods, we have thus re-run our estimation strategy on the same sample of observations as the one used in the referee's method.
+In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
+
+    As a direct consequence of this, for things to be comparable across methods, we re-run our estimation strategy on the same sample of observations as the one used in the referee's method.
 
@@
 
-    \item  With our method, one estimation allows to calculate each transport cost component $\hat{t}_{is}$, $\hat{\tau}_{is}$ and an associated $\beta_{is}$.
+    \item  With our method, one estimation allows to calculate each transport cost component $\hat{t}_{is}$, $\hat{\tau}_{is}$ and an associated $\hat{\beta}_{is}$.
 We generate a distribution of $\beta$ through bootstrap method (10,000 random draws), from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
@@
 
-\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}
+All these elements drive us to maintain our original estimation in the revised version. \textbf{What do we do with the referee's estimation method? online appendix, Appendix?}
+
+\subsection{Other comments}
+
+\textit{If the authors strictly prefer to maintain their current
+approach, [...] they should address two issues either way:}
+
+\begin{itemize}
+\item \textit{\textbf{The endogeneity problem}: quoting Footnote 14 of the paper, the authors
+are estimating ti and ts(k) as coefficients on the industry or country
+dummies times $1/\widetilde{p}_{ik}$. [...] Based on the productivity-sorting model in Melitz (2003) or the quality-sorting
+model in Baldwin and Harrigan (2010),  $1/\widetilde{p}_{ik}$ is either positively
+or negatively correlated with $\epsilon_{ik}$. So, the NNLS estimates are biased; and
+the bias has nothing to with the casual versus accounting interpretation of
+the estimates. Accordingly, the one-line justification the authors provide
+to not address the endogeneity problem is far from convincing.}
+
+\item[\text{Answer}] As stated above, we followed the referee's advice by providing some IV-estimates of the fas price to provide a clean assessment of the size of the potential bias in our original methodoly.
+The IV-strategy is implemented in two steps.
+    \begin{itemize}
+    \item First-stage equation: Following the referee's advice, we regress the fas price $\widetilde{p}_{ik}$ with $k$ = 5-digit product classification and $i$ the origin country, on tax duty and lagged fas price according to Equation (XXX).
+As made clear in the appendix of the revised version (\textbf{Section XXX - take elements of topo-first-stage there}), we run the regression in log level on a yearly basis and by transport mode, consistently with our second-stage empirical strategy.
+
+        Develop according to the results.
+
+    \end{itemize}
+    \item \textit{\textbf{The aggregation problem}: The original annual Census data reports
+trade at the origin country-HS10 product-district level of aggregation,
+whereas the authors are aggregating up the data even further to the origin
+country-HS5 industry-year level. Such an aggregation comes with strong implicit assumptions and sacrifices a lot of useful variation in the data.
+The authors are motivating the aggregation by stating that the problem
+would become computationally expensive without it. But this reasoning
+brings us back to my original point that the authors can use the Hummel's
+Methodology to circumvent the computational burden.}
+
+\item[\text{Answer}] It is true that the original annual Census data reports trade at the origin country-HS8 product-district level of aggregation (not HS10) \textbf{(I am right there?)}.
+We hope we have now convinced the referee of the relevance of our empirical method based on Equation (\ref{eq:equation0}).
+In line with the referee's comment though, we could run the estimation by considering $k$ at the HS8 classification level, not 5.
+Two elements of answer: 1) we can do it on the recent years, 2005-2013, assuming $s(k) = 3 digit and $k= 8 digit.
+Would take time, but technically feasible.
+And compare it with what we obtain in our benchmark results with $k=5$.
+\textbf{Right?} ; 2) on previous years however, we can't run this estimation, because we take Hummel's database directly, in which the finest degree of sectoral classification is at the 5-digit level.
+\textbf{Counter-argument:} Yes, but we could buy the original Census data on those years?
+\textbf{Answer to the counter-argument} No money, but this is not an argument...Better answer in my view: If we follow the referee's advice and run our estimation with $k=8$ digit on the years for which we have the data, compare with benchmark $k=5$, crossing fingers for that hopefully, things are not that different.
+In which case, this would justify NOT wasting time and money on building a new dataset on the previous years.
+Moreover, sticking to the original Hummel's database is interesting as it eliminates any other difference in the results as the estimation strategy.
+
+\end{itemize}
+
+\subsection{To sum up: What remains to be done}
+
+\begin{enumerate}
+\item\label{bp:1} IV-First-stage equation on fas price at the 8-digit level, on recent years; then, run the estimation of the $\beta$ according to the referee's approach (considering $s(k)$ at the 3 digit level)
+\item\label{bp:2} Run our original estimation method but with $k$ = 8-digit level on recent years for which we have the data.
+Also with $\widetilde{p}_{ik}$ instrumented as well?
+To be comparable with the point just above, yes.
+And compare with the $\beta$ obtained above (point \ref{bp:1}).
+\item\label{bp:3} A table comparing the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy.
+Points (\ref{bp:1}, \ref{bp:2} and \ref{bp:3}) aim at providing answers to the general comment of the referee's Critique 1.
+\item\label{bp:4} Comparison of the results of the $t,\tau,\beta$ obtained with our estimation method, for $k=5$ (what we have) and for $k=8$ (Referee's point 2), for recent years - No IV here?
+\item\label{bp:5}IV-First-stage equation on fas price at the 5-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version (Referee's point 1).
+\end{enumerate}
+
+\section{Critique 2: Calculation of Unit Prices}
 

### 2020-06-26 redaction/submission_JEGeo/revision/referees/answers/topo_first_stage.tex
réponse référé 1, Lise, continued

@@
 
-leading to the following specification of the first-stage equation (in cross section):
+leading to the following specification of the first-stage equation (in cross section and in panel respectively):
 
@@
 
+\subsection{To sum up}
+
+What would be my ``preferred'' first-stage equation: Equation (\ref{eq:FS_loglevel_cross_section_inertia}) because i) in log level, so do not need to rebuild the level from the first difference + consistent with our second stage in log level as well; ii) in cross-section, run on a yearly basis, here again consistent with our second-stage equation.
+But need to be pragmatic as well...
+
 \end{document}

### 2020-06-26 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
réponse référé 1, encore

@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:equation0}
 \end{equation}
@@
 
-Our estimated equation is written as:
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
-\end{equation}
-
-\noindent with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
+Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
 We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
-To sum up our methodoly:
+To sum up our methodology:
 \begin{itemize}
@@
 
+\textit{My second critique concerns the way the authors are calculating the unit prices.
+The Census data reports the quantity of goods per observation. So, the authors
+can calculate the unit price as Value/Quantity, which is consistent with how
+price is modeled in standard trade models. Instead, the authors calculate unit
+price as Value/Weight. This used to be a common exercise in the past where
+many data-sets did not report Quantity. But, given their data, there is no
+justification for the authors to calculate the prices this way.}
+
+The referee is right in the sense that Census data do report the quantity of goods.
+However, this information is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
+This explains why we cannot calculate the unit price as suggested by the referee.
+
+At a more fundamental level, one might argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
+For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
+\textbf{Do we have evidence, even anecdotal, on this point?
+Is it an argument we want to make, See what Lashkaripour says about this point also.}
+
+\textit{Calculating the unit price as Value/Weight presents the authors
+with an additional endogenity problem. To elaborate, let $\omega_{ik} = Weight/Quantity$
+denote the unit weight of the goods in observation $ik$. Also let $\widehat{p}_{ik} = Value/Quantity$
+(unlike what the authors assume) denote standard definition of price}.
+As a consequence, making the link with the fas price we consider in the paper: $\widetilde{p}_{ik} = \widehat{p}_{ik} / \omega_{ik}$.
+
+\textit{This paper is essentially estimating the following equation}:
+
+\begin{equation*}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}/\omega_{ik}} \right) + \epsilon_{ik}
+\end{equation*}
+
+\textit{instead of estimating}:
+
+\begin{equation*}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}} \right) + \epsilon_{ik}
+\end{equation*}
+
+\textit{There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
+categories, and (ii) $\omega_{ik}$ is negatively correlated with transport costs. So the
+way the authors are calculating unit prices and estimating the model creates a
+new (but avoidable) source of endogeneity}
+
 \end{document}

### 2020-06-29 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
ajouts mineurs lettre référé 1

@@
 This explains why we cannot calculate the unit price as suggested by the referee.
+\textbf{We make this point clear in Footnote XXX}
+
+\textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
+Et à ce moment là, on peut répondre ``vraiment'' à sa critique?}
 
@@
 way the authors are calculating unit prices and estimating the model creates a
-new (but avoidable) source of endogeneity}
+new (but avoidable) source of endogeneity.}
+
+We thank the referee for pointing out this potential source of bias in our estimates.
+If we cannot follow the referee's advice be replacing weight by quantity for data availability reasons, we can address his/her concern regarding this as source of endogeneity bias by instrumenting the fas price.
+\textbf{il ne faudrait pas dans ce cas mettre de lagged prices, sinon on reintroduit du bruit.
+Non?}
+
+\section{Critique 3: Big Picture Implications}
+
+\textit{My third critique concerns the lack of an exciting punchline. The fact that
+composition effects have not countervailed the reduction in pure transport costs
+(at least not as much as previously believed) is an interesting but minor observation.
+Does this observation revise our understanding of say the gains from
+trade? Does it shed new light on a puzzle many people are thinking about?
+One crude suggestion is to see how the reduction in the industry-specific cost
+terms is related to the industry-level trade elasticities. If the composition effects
+favor low-elasticity industries, the findings in the paper may have first-order
+implications for the gains from trade.
+Another suggestion is to dig deeper into the relative rate at which additive
+and multiplicative transport costs have declined over time. Since additive transport
+costs favor rich (high-quality exporting) countries, the disproportionally
+greater reduction in additive costs can perhaps explain the rise of low-income
+exporter as documented by Hanson (2012, JEP).}
+
+\textbf{Comments about that}
+
+\paragraph{Piste 1} Consider his/her first suggestion.
+What is the idea?
+First, the link between gains from trade and trade-elasticity.
+Am I correct in saying that gains from trade are higher when trade is about low-elasticity goods?
+The idea being, if national goods are low substitutes, then here are larger gains from trading them.
+So, if composition effects (trade shifts between sectors) favor low-elasticity sectors, then one might expect high gains from trade.
+Hummels: trade composition effects matter as they partially offset the reduction in pure transport costs for both air and vessel.
+Over time, tendency to trade goods more costly to export everything else equal.
+So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade.
+But, we disagree with Hummel's findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air.
+Meaning that gains from trade are purely due to reduction in transport costs within sector, but not from switch towards low-elasticity sectors.
+Identify one source of gains from trade, not two.
+
+What to do with this?
+
+\paragraph{Piste 2}
 

### 2020-06-29 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
chnagement encore

@@
 
-\paragraph{Piste 2}
+\paragraph{Piste 2} What is at stake?
+Show that how $\beta$ = the relative share of additive costs has evolved over time.
+Good idea as this is really what differentiates us from Hummels, model a varying $\beta$ over time.
+Specifically, how methodology lets the $\beta$ vary over the three time-sector-partner country dimension.
+Show how $\beta$ varies over time everything else equal?
+Hopefully, goes in the referee's direction (decreases over time) with possibly a role in the understanding in the ``big picture'' of trade patterns?
 

### 2020-06-29 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
erreur push....

@@
 
+\textbf{GD20200629 Ce n’est pas un si petit sample que cela (surtout en terme de valeurs -- comparer avec le sample sur lequel on peut faire sa propre méthode)}
+
 \textbf{There is also a point here, about the degree of sectoral classification.
@@
 and why?}
+
+\textbf{Dans revision_estim_TC_referee1_nonlinear.do, il me semble qu’on fait 8 et 3. on peut effectivement faire sa méthode en 10 et 3 (ou 10 et 5)}
 
@@
 2) Compare the estimated $\beta$ between our original method and the referee's one.
-\textbf{Where are the results comparing the estimates of the $\beta$, in value?
-I don't find these results anymore...}
+\
+\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}
+\textbf{GD20200629 -- Voir le plan des do. Pour comparaison scatter_chronology_baselinesamplereferee1_referee1.pdf (sur le même sample))}
 
@@
 Have we done that?}
+\textbf{GD20200629 Cela ne semble pas vrai. Il y a quelque années en HS8, plus en HS6. Mais pas sur toute la durée}
 \item[Advantage 3]: A more transparent comparison with Hummel (2007): How to answer to this point? Re-read the paper to see if we are clear enough?
+\textbf{GD20200629 changer la manière dont on présente en mettant les beta en avant ?}
 \end{itemize}
@@
 
+\textbf{GD20200629 Les données commencent en 1989. On va demander}
+
     As a direct consequence of this, for things to be comparable across methods, we re-run our estimation strategy on the same sample of observations as the one used in the referee's method.
@@
 \textbf{DEEPEN this for the Big picture part}.
+\textbf{GD20200629 Je ne comprends pas. Si on a les parts et les coûts totaux, on peut reconstruire $\tau$, $t$}
 \item[Concern 4] The suggested method features less accuracy in the estimation of the $\beta$.

### 2020-07-01 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
changement rédaction réponse référé 1

--- 2020-06-29 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
@@
 
-\subsection{Deriving the share of additive costs in total transport costs \label{ssec:interpret_beta}}
+\subsection{Clarifying some technical points \label{ssec:technical_points}}
+
+\subsubsection{Deriving the share of additive costs in total transport costs \label{ssec:interpret_beta}}
+
 With $i$ the origin country, $k$ the product category at the HS6 level (and reasoning at the year- transport mode level), $p_{ik}$ the import (cif) price and $\tilde{p}_{ik}$ the export (fas) price, transport costs $f_{ik}\equiv \frac{p_{ik}- \tilde{p}_{ik}}{\tilde{p}_{ik}} $ are written in Hummel's terminology as:
@@
 
-\subsection{Coming back to the Referee's 1 concern}
+\subsection{Share vs level}
+
+Starting from the above reasoning, the referee proposes to estimate the elasticity of transport costs with respect to the unit price, equivalently the share of additive costs in total costs.
+Specifically, the referee suggests to estimate for each industry-country the following regression:
+
+$$\ln f_{ikd}= \beta_{is(k)} \ln \widetilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd}$$
+
+The question is then, how to recover the levels of additive / multiplicative transports costs $t_{is(k)}$ and $\tau_{is(k)}$.
+Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ from the referee's method for a given sector-country $i,s(k)$, one can solve the following two-equation system:
+
+\begin{eqnarray}
+p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
+\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
+\end{eqnarray}
+
+\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
+With 2 equations and 2 endogenous variables, the system might be solved.
+Specifically, with $\widehat{\beta}$ being estimated at the \textit{sector} $s$- country $i$ level, and $\widetilde{p}_{ik}$ at the \textit{product} $k$-country level, this implies that $t$ and $\tau$ are product-country $i,k$ specific (not sector-country $s,k$ specific).
+\textbf{from this point of view, this is a finer characterization of transport costs, so better from this point of view...}
+
+To show this properly, start from Equation (\ref{eq:system2}) to express additive costs according to:
+$$t_ik = \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
+
+With $\widetilde{p}_{ik} $ being product-country specific, $t_ik$ is necessarily product-country specific ($i,k$, not $i,s(k)$). Plugging this into Equation (\ref{eq:system1}):
+
+$$p_{ik} = \tau_{is(k)} \widetilde{p}_{ik} + \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
+
+Solving for $\tau$, one finally obtains the solutions for the ad-valorem and additive transport costs components:
+
+\begin{eqnarray}
+\tau_{ik} & =& (1-\widehat{\beta}_{is(k)}) \frac{p_{ik}}{\widetilde{p}_{ik} } + \widehat{\beta}_{is(k)} \label{eq:tau_method_ref1}\\
+t_ik &=& \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} \label{eq:t_method_ref1}
+\end{eqnarray}
+
+Two points can be made.
+First, we can check that if $\widehat{\beta}_{is(k)}=0$, then $p_{ik} = \tau_{ik} \widetilde{p}_{ik}$ (Equation \ref{eq:tau_method_ref1}) and $t_{ik} = 0$ (Equation \ref{eq:t_method_ref2}), all transport costs are ad-valorem.
+Conversely, if $\widehat{\beta}_{is(k)}=1$, then $p_{ik} = \widetilde{p}_{ik} + t_{ik} $ (Equation \ref{eq:tau_method_ref2}) and $\tau_{ik} = 0$ (Equation \ref{eq:t_method_ref1}).
+Second, and most importantly, the method allows to recover the levels of both transport costs components at the product-country level $i,k$, even though the share of additive costs in total costs is estimated at the sector-country level.
+
+\subsection{Coming back to the Referee's 1 concerns}
 
@@
 
-\paragraph{(i) On the assumptions underlining our empirical approach}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla (the Review of Economics and Statistics, 2015) to estimate the share of additive costs in a firm-level context.
-It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
-\textbf{Not sure of what it means?
-} More importantly, this formulation allows us recovering explicitly the \textbf{levels of}, respectively, additive and multiplicative costs.
-Despite its formal elegance and simplicity, this is not possible in such a straightforward manner with the Hummels' methodology, which only recovers the \textbf{share} of additive in total transport costs.
-\textbf{elaborate on this in the Big Picture part.}
-
-Nevertheless, to take full account of the referee’s remark, we decided run an additional set of estimations based on the Hummel's methodogy as proposed by the referee, based on Equation (\ref{eq:estimation_ref1}).
-This is developed in Section XX of this letter.
-
-\textbf{We need to explain why not in the main text, but in the (online?) appendix}
-
-\paragraph{(ii) On the use of non-linear least squares (NLS)}
+\paragraph{(i) On the use of non-linear least squares (NLS)}
 As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
@@
 We do acknowledge this is a very important point, which was not made clear enough in the initial version, and did our best to make the pint clear in the revised version \textbf{[see page XX : XX insert quote XX].}
+
+\paragraph{(ii) On the assumptions underlining our empirical approach}
+Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla (the Review of Economics and Statistics, 2015) to estimate the share of additive costs in a firm-level context.
+It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
+\textbf{Not sure of what it means?
+} Nevertheless, to take full account of the referee’s remark, we decided run an additional set of estimations based on the Hummel's methodogy as proposed by the referee, based on Equation (\ref{eq:estimation_ref1}).
+We yet came to the conclusion that this method faces some important limitations, as we develop in Section XXX of this letter.
+As a result, we have decided to maintain our estimation strategy as the benchmark method in the main body of the paper.
+We hope that our letter will convince the referee of the relevance of this choice.
+
+\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier?}
 
@@
 That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
-
 Following the referee's advice, we decided therefore to provide some IV-estimates to provide a clean assessment of the size of the potential bias.
 
-\textbf{careful, there is two points regarding endogeneity: (i), we should estimate the $\beta$ according to the referee's method based on the intrumented fas price (at the HS8 level, if tax duty are available at this disaggregation level) - WE HAVE NOT DONE THAT; (ii), if we maintain our empirical strategy (which we do), we should instrument the fas price at the 5-digit level - WHICH WE ARE TRYING TO DO.}
-
-XXX At some point, explain HS10 duty do not exist, contrary to what the referee asserts. At best, it was HS6 or HS8 XXX
+\paragraph{Summary} We decided to answer to these two last concerns (ii) and (iii) in a separate manner.
+Specifically, we run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as exposed); this way, the results are comparable with our benchmark results of the submitted version, the only change being the estimation strategy.
+In a second step, we implement IV methods on fas prices at the $k=$ 5-digit level, still based on our original estimation strategy.
+This allows us a direct comparison of the results obtained in the submitted version (NLS OLS regression) with those obtained with IV.
 
@@
 
-\textbf{There is also a point here, about the degree of sectoral classification.
-In our paper, $k$ = 5 digit and $s(k)$ = 3 digit.
-He/she suggests that, in the estimated equation, $k$ = 10 digit and $s(k)$= 5 digit.
-Which means, having an estimated $\beta$ specific to the origin country-5-digit level.
-I need to have a refresher on this: What have we done finally?
-and why?}
-
-XXX Say something about the fact that in the original US-custom data, the finest degree of sectoral classification is 8-digit, not 10? XXX
-
-\subsubsection{What we do}
+\paragraph{What we do}
 
@@
 
-\subsubsection{What the referee suggests to do}
+\paragraph{What the referee suggests to do}
 
@@
 As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs.
-The share of aditive costs can hence be encovered by regressing transport costs on unit (ie, fas) prices.
+The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
 Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
 
-Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector (\textbf{why 3-digit?
-and not 5 as suggested by the referee?}, exploiting the variability between sub-sectors at the 8 digit level ($k$) and between ports of entry in the US ($d$).
+Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{Notice that the referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$), while we followed his suggestion considering the 3-digit level.
+We made this choice in view of being able to compare the pros and cons of the referee's method versus ours.
+Both methods require a difference in the sectoral degree at which either the share (Referee's method) or the level (our method) of additive costs are estimated ($s$) and the classification degree at the product level ($k$).
+In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004.
+For comparability reasons, we maintain the same degree of classification $k=5$ for the following years (based on the original Census dataset).
+As a result, transport costs are estimated at the 3-digit level as benchmark estimate, exploiting heterogenity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
+This has driven us to retain $s=3$ when implementing the referee's method, in view of preserving comparability between both strategies.}
 
@@
 2) Compare the estimated $\beta$ between our original method and the referee's one.
-\textbf{Where are the results comparing the estimates of the $\beta$, in value?
-I don't find these results anymore...}
+\
+\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}
+\textbf{GD20200629 -- Voir le plan des do. Pour comparaison scatter-chronology-baselinesamplereferee1-referee1.pdf (sur le même sample))}
 
-\item[Advantage 2]: There is a handful of instruments (HS-8 product-specific tariff rates or lagged prices) to instrument fas prices
+\textbf{stop here}
+
+\item[Advantage 2]: There is a handful of instruments (HS-10 product-specific tariff rates or lagged prices) to instrument fas prices
 \item[$\Rightarrow$] Answer: When we instrument fas prices at the 5-digit level using tax duties (still using our own methodology), we do not obtain much different results.
@@
 Have we done that?}
+\textbf{GD20200629 Cela ne semble pas vrai. Il y a quelque années en HS8, plus en HS6. Mais pas sur toute la durée}
 \item[Advantage 3]: A more transparent comparison with Hummel (2007): How to answer to this point? Re-read the paper to see if we are clear enough?
+\textbf{GD20200629 changer la manière dont on presente en mettant les beta en avant ?}
 \end{itemize}

### 2020-07-02 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
changement réponse référé 1

@@
 
-\subsection{Clarifying some technical points \label{ssec:technical_points}}
+In italics, I report the Referee's comment. Into brackets in bold characters inserted into the referee's statement, my comment.
 
-\subsubsection{Deriving the share of additive costs in total transport costs \label{ssec:interpret_beta}}
+\textit{Using the notation of the authors, they are intreated
+in identifying the share of the specific cost in the total transport cost. Namely,}
+\begin{eqnarray*}
+&& \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} \\
+\text{or,} &&\frac{t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{is(k)} + t_{is(k)}}
+\end{eqnarray*}
+
+\textit{The way they approach the problem is that they assume that (a) $\tau_{ik} = \tau_i\tau_{k}$,
+(b) $t_{ik} = t_i +t_k$, and (c) $t_k$ and $\tau_k$ are uniform across products within industry
+s. After imposing these assumptions, they estimate the following specification}
+
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:equation0}
+\end{equation}
+
+\textit{in which $\tau_i$, $\tau_{s(k)}$, $t_i$, and $t_{s(k)}$ are identified as fixed effects coefficients.
+In my opinion this choice of strategy is quite sub-optimal, as (i) it relies on
+the strong assumptions highlighted above, (ii) it is computationally expensive
+as noted by the authors on multiple occasions, and (iii) it is subject to an
+endogeneity problem, which the authors disregard with one sentence, but which
+is rather detrimental in my opinion.}
+
+The three points raised by the referee are all equally important, and we answer separately to each of them below.
+
+\paragraph{(i) On the use of non-linear least squares (NLS)}
+As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
+However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
+This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
+$\tau \geq 1$ and $t \geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative.
+Without these restrictions, standard linear, least squares estimates often deliver aberrant values with more than mild quality-of-fit \textbf{[insert examples here]}
+We do acknowledge this is a very important point, which was not made clear enough in the initial version, and did our best to make the pint clear in the revised version \textbf{[see page XX : XX insert quote XX].}
+
+\paragraph{(ii) On the assumptions underlining our empirical approach}
+Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla (the Review of Economics and Statistics, 2015) to estimate the share of additive costs in a firm-level context.
+It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
+\textbf{Not sure of what it means?
+} Nevertheless, to take full account of the referee’s remark, we decided run an additional set of estimations based on the Hummel's methodogy as proposed by the referee, based on Equation (\ref{eq:estimation_ref1}).
+We yet came to the conclusion that this method faces some important limitations, that drove us to maintain our initial estimation strategy in the main body of the paper.
+We develop the analysis we held in Section XXX of the letter, and we hope that it will convince the referee of the relevance of this choice.
+
+\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier?}
+
+\paragraph{(iii) Endogeneity problem}
+Indeed, this is a very important point.
+The referee states that, based on theoretical insights by Melitz (Econometrica, 2003) or Baldwin and Harrigan (AER, 2011), $1/\tilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+We obviously do not question this conceptual issue.
+However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices.
+Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out.
+That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
+Following the referee's advice, we decided therefore to provide some IV-estimates to provide a clean assessment of the size of the potential bias.
+
+\paragraph{Summary} We decided to answer to these two last concerns (ii) and (iii) in a separate manner.
+Specifically, we run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as detailed later); this way, the results are comparable with the benchmark results of the submitted version, the only change being the estimation strategy.
+In a second step, we implement IV methods on fas prices at the $k=$ 5-digit level, still based on our original estimation strategy.
+This allows a direct comparison of the results obtained in the submitted version (NLS OLS regression) with those obtained with IV.
+
+\subsection{Exploring Referee's alternative functional form}
+
+\textit{A more natural approach is what the authors, at some point, refer to as the
+Hummel's Methodology. That is, one can alternatively estimate the share of the
+additive component as:}
+
+$$\frac{t_{ik}}{ (\tau_{ik}-1)\tilde{p}_{ik} + t_{ik}} = \beta_{ik}$$
+
+\textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t.
+unit price} [\textbf{in absolute value, see Equation \ref{eq:Hummels}}].
+\textit{Given the authors' objective and the data they are using, $\beta$ can be separately estimated
+for each industry-country pair using the following regression:}
+
+\begin{equation}
+\ln f_{ikd} = \beta_{is(k)}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
+\end{equation}
+
+\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product. The
+identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
+district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
+obviously require that the authors do not aggregate up the raw Census data
+across all districts and all 10-digit products pertaining to the same 5-digit category.}
+
+\textit{The first advantage of this so-called Hummel's Approach is that the above
+regression can be estimated separately for various country-industry pairs, without
+imposing Assumptions (a) and (b), outlined above.}
+
+\textbf{I report here the referee's footnote: Should we deal with this?
+If he/she is not convinced because small sub-sample, do it on a much larger sample?
+This would better convince him that the separability assumption is not an issue.}: \textit{The authors do check the robustness of their results w.r.t.
+the separability assumption, but this only done for a small sub-sample.}
+
+\paragraph{What we do}
+
+Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
+We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
+To sum up our methodology:
+\begin{itemize}
+\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
+\item We rebuild
+\begin{eqnarray*}
+\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
+\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
+\end{eqnarray*}
+\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
+\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
+
+$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
+\end{itemize}
+
+\paragraph{What the referee suggests to do}
+
+The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
+As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs.
+The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
+Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
+
+Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{The referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
+We followed his suggestion but considering the 3-digit level.
+We made this choice in view of being able to compare the pros and cons of the referee's method versus ours.
+In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004.
+For comparability reasons over time, we adopt the same degree of classification $k=5$ for the following years (based on the original Census dataset).
+As a result, transport costs are estimated at the 3-digit level as benchmark estimate, exploiting heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
+This has driven us to retain $s=3$ when implementing the referee's method, in view of preserving comparability between both strategies.}
+
+\subsubsection{Strategy of answer}
+
+\paragraph{a) The advantages of the referee's method are not as important as suggested}
+\begin{itemize}
+\item[Advantage 1]: One is not compelled to assume separability between the country-component and the product-component of transport costs.
+\item[$\Rightarrow$] Answer can be in 2 parts: 1) This is not a strong assumption - better convince him/her: Run the robustness to the separability assumption on a larger sample and show that the estimated $\beta$ is not that different between the two methods.
+\textbf{Show that this is not a so small sample, and compare it with the sample induced by the referee's method} 2) Compare the estimated $\beta$ between our original method and the referee's one.
+\
+\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}
+\textbf{GD20200629 -- Voir le plan des do. Pour comparaison scatter-chronology-baselinesamplereferee1-referee1.pdf (sur le même sample))}
+
+\item[Advantage 2]: There is a handful of instruments (HS-10 product-specific tariff rates or lagged prices) to instrument fas prices
+\item[$\Rightarrow$] Answer: As exposed below, our strategy of answer has been to treat the two points (estimation strategy / potential endogeneity bias) separately.
+Since our investigation on the estimation strategy has led us to discard this method as the benchmark one, we need not to instrument fas prices at the HS 10 level, our method featuring a $k=$5-digit classification level.
+\textbf{However, when run at the HS-5 product level, the IV method does not yield results that are much different from the benchmark OLS results, as shown in Table XXX.}
+
+\item[Advantage 3]: A more transparent comparison with Hummel (2007): \textbf{changer la maniere dont on presente en mettant les beta en avant ?}
+\end{itemize}
+
+\paragraph{b) The costs are high}
+\begin{itemize}
+\item[Concern 1] The suggested estimation strategy implies having much less data to exploit, for two reasons.
+\begin{enumerate}
+\item Information about the port of entry are only available since 1989.
+Implementing the referee's method would hence necessarily reduce the time coverage of our analysis (1974-1988).
+In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
+Eliminating this dimension of the paper would be detrimental to its value-added.
+\item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
+Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors (even at the HS-6 digit classification) or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
+This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table XXX \textbf{TO DO}.
+Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
+
+    As a direct consequence of this, for things to be comparable across methods, we re-run our estimation strategy on the same sample of observations as the one used in the referee's method.
+
+    \end{enumerate}
+\item[Concern 2] The suggested estimation method implicitly assumes that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors; in both cases, one might view this as a disputable assumption.
+By contrast, we suppose that transport costs are sector-country specific.
+\textbf{FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique?}
+\item[Concern 3] The suggested method features less accuracy in the estimation of the $\beta$.
+If we take the value of the $\beta$ by itself, there is no criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
+Things are more clear-cut in terms of accuracy of the estimation.
+Specifically, our method yields a more precise estimation of the $\beta$ than the referee's method.
+To develop on this, for each year and transport mode:
+    \begin{itemize}
+    \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$.
+From this, we can deduce the 5- 95\% threshold values through:
+        $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
+
+    \item  With our method, one estimation allows to calculate each transport cost component $\hat{t}_{is}$, $\hat{\tau}_{is}$ and an associated $\hat{\beta}_{is}$.
+We generate a distribution of $\beta$ through bootstrap method (10,000 random draws), from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
+    \item We can then evaluate the accuracy of each estimation method by comparing the distributions of the $\beta$.
+This is reported in the Figures XXX (for Air) and XXX (for Vessel).
+Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
+For sake of brevity, we report the mean distribution averaged over the period, a similar conclusion applies on a yearly basis.
+\textbf{Averaged over the two transport modes?
+For only one?}
+
+\begin{figure}[htbp]
+\caption{Accuracy of the estimation of the $\beta$: Comparison}
+\label{fig:accuracy_beta}
+\begin{center}
+\includegraphics[height=4in]{accuracy_beta.pdf}
+\end{center}
+\end{figure}
+
+Figure \ref{fig:accuracy_beta} reports the log of the interval confidence of the $\beta$ obtained with our methodology (with bootstrap), relative to the one obtained with the referee's method.
+\textbf{average over the years?
+the transport modes?
+unclear}.
+In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
+
+\end{itemize}
+\end{itemize}
+
+All these elements drive us to maintain our original estimation in the revised version. \textbf{What do we do with the referee's estimation method? online appendix, Appendix?}
+
+\subsection{Potential endogeneity problem}
+
+\textit{\textbf{The endogeneity problem}: quoting Footnote 14 of the paper, the authors
+are estimating ti and ts(k) as coefficients on the industry or country
+dummies times $1/\widetilde{p}_{ik}$. [...] Based on the productivity-sorting model in Melitz (2003) or the quality-sorting
+model in Baldwin and Harrigan (2010),  $1/\widetilde{p}_{ik}$ is either positively
+or negatively correlated with $\epsilon_{ik}$. So, the NNLS estimates are biased; and
+the bias has nothing to with the casual versus accounting interpretation of
+the estimates. Accordingly, the one-line justification the authors provide
+to not address the endogeneity problem is far from convincing.}
+
+We followed the referee's advice by providing some IV-estimates of the fas price to provide a clean assessment of the size of the potential bias in our original methodology.
+The IV-strategy is implemented in two steps.
+
+\begin{itemize}
+\item First-stage equation: We regress the fas price $\widetilde{p}_{ik}$ with $k$ = 5-digit product classification and $i$ the origin country, on tax duty and lagged fas price according to Equation (XXX).
+As made clear in the appendix of the revised version (\textbf{Section XXX - take elements of topo-first-stage there}), we run the regression in log level on a yearly basis and by transport mode, consistently with our second-stage empirical strategy.
+
+\item The second-stage: Explain the difficulty since fas prices are on both sides of the equation; non-trivial way to do it.
+We have decided to not instrument the fas price that intervenes on the LHS.
+Develop.
+
+\end{itemize}
+
+The results are ... DEVELOP.
+
+\subsection{The aggregation problem}
+
+\textit{\textbf{The aggregation problem}: The original annual Census data reports
+trade at the origin country-HS10 product-district level of aggregation,
+whereas the authors are aggregating up the data even further to the origin
+country-HS5 industry-year level. Such an aggregation comes with strong implicit assumptions and sacrifices a lot of useful variation in the data.
+The authors are motivating the aggregation by stating that the problem
+would become computationally expensive without it. But this reasoning
+brings us back to my original point that the authors can use the Hummel's
+Methodology to circumvent the computational burden.}
+
+It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
+We hope we have now convinced the referee of the relevance of our empirical method based on Equation (\ref{eq:equation0}).
+In line with the referee's comment though, we could run the estimation by considering $k$ at the HS10 classification level, not 5.
+Two elements of answer: 1) we can do it since 1989 only \textbf{CAUTIOUS, 1989 is because it is the first year for which we have information about the port of entry, but here we don't need it.
+Potentially we could ask for data since the beginning 1974?
+}, assuming $s(k) = 3 digit and $k= 8 digit.
+Would take time, but technically feasible.
+And compare it with what we obtain in our benchmark results with $k=5$.
+\textbf{Right?.
+If we follow the referee's advice and run our estimation with $k=10$ digit on the years for which we have the data, compare with benchmark $k=5$, crossing fingers for that hopefully, things are not that different.
+In which case, this would justify NOT wasting time and money on building a new dataset on the previous years.
+Moreover, sticking to the original Hummel's database is interesting as it eliminates any other difference in the results as the estimation strategy.}
+
+\subsection{To sum up: What remains to be done}
+
+\begin{enumerate}
+\item\label{bp:1} After running the referee's 1 method with $k=$10, show tables or figures comparing 1) the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy; 2) comparison of the $\beta$ between the two methods; 3) comparison of the precision of the $\beta$ estimation.
+Answer to the critique about the estimation strategy.
+\item\label{bp:2} IV-First-stage equation on fas price at the 5-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version.
+Answer to the critique about the endogeneity bias.
+\item\label{bp:3} Re-run our benchmark method with $s=3$ and $k=10$ over some years and compare the results ($t,\tau,\beta$ with our benchmark ones ($s=3,k=5$).
+Answer to the critique about the aggregation problem.
+\end{enumerate}
+
+\section{Critique 2: Calculation of Unit Prices}
+
+\textit{My second critique concerns the way the authors are calculating the unit prices.
+The Census data reports the quantity of goods per observation. So, the authors
+can calculate the unit price as Value/Quantity, which is consistent with how
+price is modeled in standard trade models. Instead, the authors calculate unit
+price as Value/Weight. This used to be a common exercise in the past where
+many data-sets did not report Quantity. But, given their data, there is no
+justification for the authors to calculate the prices this way.}
+
+The referee is right in the sense that Census data do report the quantity of goods.
+However, this information is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
+This explains why we cannot calculate the unit price as suggested by the referee.
+\textbf{We make this point clear in Footnote XXX}
+
+\textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
+Et à ce moment là, on peut répondre ``vraiment'' à sa critique?}
+
+At a more fundamental level, one might argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
+For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
+\textbf{Do we have evidence, even anecdotal, on this point?
+Is it an argument we want to make, See what Lashkaripour says about this point also.}
+
+\textit{Calculating the unit price as Value/Weight presents the authors
+with an additional endogenity problem. To elaborate, let $\omega_{ik} = Weight/Quantity$
+denote the unit weight of the goods in observation $ik$. Also let $\widehat{p}_{ik} = Value/Quantity$
+(unlike what the authors assume) denote standard definition of price}.
+As a consequence, making the link with the fas price we consider in the paper: $\widetilde{p}_{ik} = \widehat{p}_{ik} / \omega_{ik}$.
+
+\textit{This paper is essentially estimating the following equation}:
+
+\begin{equation*}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}/\omega_{ik}} \right) + \epsilon_{ik}
+\end{equation*}
+
+\textit{instead of estimating}:
+
+\begin{equation*}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}} \right) + \epsilon_{ik}
+\end{equation*}
+
+\textit{There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
+categories, and (ii) $\omega_{ik}$ is negatively correlated with transport costs. So the
+way the authors are calculating unit prices and estimating the model creates a
+new (but avoidable) source of endogeneity.}
+
+\textbf{Je ne comprends pas bien le point (i)}.
+We thank the referee for pointing out this potential source of bias in our estimates.
+If we cannot follow the referee's advice be replacing weight by quantity for data availability reasons, we can address his/her concern regarding this as source of endogeneity bias by instrumenting the fas price.
+\textbf{il ne faudrait pas dans ce cas mettre de lagged prices, sinon on reintroduit du bruit.
+Non?
+Mais alors, on n'explique plus grand chose...
+}
+
+\section{Critique 3: Big Picture Implications}
+
+\textit{My third critique concerns the lack of an exciting punchline. The fact that
+composition effects have not countervailed the reduction in pure transport costs
+(at least not as much as previously believed) is an interesting but minor observation.
+Does this observation revise our understanding of say the gains from
+trade? Does it shed new light on a puzzle many people are thinking about?
+One crude suggestion is to see how the reduction in the industry-specific cost
+terms is related to the industry-level trade elasticities. If the composition effects
+favor low-elasticity industries, the findings in the paper may have first-order
+implications for the gains from trade.
+Another suggestion is to dig deeper into the relative rate at which additive
+and multiplicative transport costs have declined over time. Since additive transport
+costs favor rich (high-quality exporting) countries, the disproportionally
+greater reduction in additive costs can perhaps explain the rise of low-income
+exporter as documented by Hanson (2012, JEP).}
+
+\textbf{Comments about that}
+
+\paragraph{Piste 1} Consider his/her first suggestion.
+What is the idea?
+First, the link between gains from trade and trade-elasticity.
+Am I correct in saying that gains from trade are higher when trade is about low-elasticity goods?
+The idea being, if national goods are low substitutes, then here are larger gains from trading them.
+So, if composition effects (trade shifts between sectors) favor low-elasticity sectors, then one might expect high gains from trade.
+Hummels: trade composition effects matter as they partially offset the reduction in pure transport costs for both air and vessel.
+Over time, tendency to trade goods more costly to export everything else equal.
+So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade.
+But, we disagree with Hummel's findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export).
+Meaning that gains from trade are purely due to reduction in transport costs at the sector, but not from switch towards low-elasticity sectors.
+Identify one source of gains from trade, not two.
+
+What to do with this?
+When composition effects do matter (ocean freight), do they favor low-elasticity industries?
+In which case, on top of the reduction in transport costs per se (which has welfare consequences as well), one might expect additional gains from these composition effects.
+When they don't, this mitigates the gains from trade that could be expected from the picture given by Hummels.
+
+\paragraph{Piste 2} What is at stake?
+Show that how $\beta$ = the relative share of additive costs has evolved over time.
+Good idea as this is really what differentiates us from Hummels, model a varying $\beta$ over time.
+Specifically, how methodology lets the $\beta$ vary over the three time-sector-partner country dimension.
+Show how $\beta$ varies over time everything else equal?
+Hopefully, goes in the referee's direction (decreases over time) with possibly a role in the understanding in the ``big picture'' of trade patterns?
+
+\appendix
+
+\section{Clarifying some technical points \label{ssec:technical_points}}
+
+\subsection{A note on prices}
+
+We have three prices: the (US) consumer price, say $p^{c}_{ik}$, the ``import'' or  cif price (at the entrance of the US) $p_{ik}$ and the fas price (at the export gate in the origin country) $\widetilde{p}_{ik}$, denoting $i$ the origin country, $k$ the product at the 8 or 5 digit level, and reasoning on a yearly basis, and $s(k)$ the 3-digit classification the product $k$ belongs to.
+Further, if we denote $\tau^d_{ik}$ the duty tax rate paid when the good crosses the US border, then we have:
+
+\begin{eqnarray*}
+&&p^c_{ik} = (1+\tau^d_{is(k)})p_{ik} \\
+with && p_{ik}  = \tau_{is(k)} \widetilde{p}_{ik} +t_{is(k)}
+\end{eqnarray*}
+
+\subsection{Deriving the share of additive costs in total transport costs \label{ssec:interpret_beta}}
 
@@
 
-\subsection{Coming back to the Referee's 1 concerns}
+\section{More details on the instrumentation strategy}
 
-In italics, I report the Referee's comment. Into brackets in bold characters inserted into the referee's statement, my comment.
+\subsection{Specification of the first stage equation}
 
-\textit{Using the notation of the authors, they are intreated
-in identifying the share of the specific cost in the total transport cost. Namely,}
-\begin{eqnarray*}
-&& \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} \\
-\text{or,} &&\frac{t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{is(k)} + t_{is(k)}}
-\end{eqnarray*}
+The question we deal with, is the functional form of the first stage equation, where we instrument the fas price with duties as suggested by Referee 1.
+The idea is that firms might react to changes in duty tax rates which have nothing to do with transport costs changes.
+In this respect, considering the predicted part of the fas price related to tax duty is likely to solve potential endogeneity biases.
 
-\textit{The way they approach the problem is that they assume that (a) $\tau_{ik} = \tau_i\tau_{k}$,
-(b) $t_{ik} = t_i +t_k$, and (c) $t_k$ and $\tau_k$ are uniform across products within industry
-s. After imposing these assumptions, they estimate the following specification}
+Denoting $i$ the origin country, $k$ the product at the 5 digit level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
 
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:equation0}
+\widetilde{p}_{ikt} = (1+\tau^d_{is(k)t})^\alpha \left(\bar{p}_{ikt}\right)^\beta \label{eq:link_fas_duty}
 \end{equation}
 
-\textit{in which $\tau_i$, $\tau_{s(k)}$, $t_i$, and $t_{s(k)}$ are identified as fixed effects coefficients.
-In my opinion this choice of strategy is quite sub-optimal, as (i) it relies on
-the strong assumptions highlighted above, (ii) it is computationally expensive
-as noted by the authors on multiple occasions, and (iii) it is subject to an
-endogeneity problem, which the authors disregard with one sentence, but which
-is rather detrimental in my opinion.}
+\noindent with $s(k)$ the 3-digit classification as a function of the 5-digit product classification.
 
-The three points raised by the referee are all equally important, and we answer separately to each of them below.
+\subsubsection{First-stage equation in first difference \label{ssec:first_diff}}
 
-\paragraph{(i) On the use of non-linear least squares (NLS)}
-As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
-However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
-This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
-$\tau \geq 1$ and $t \geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative.
-Without these restrictions, standard linear, least squares estimates often deliver aberrant values with more than mild quality-of-fit \textbf{[insert examples here]}
-We do acknowledge this is a very important point, which was not made clear enough in the initial version, and did our best to make the pint clear in the revised version \textbf{[see page XX : XX insert quote XX].}
+One option to specify the functional form of the first stage is to take the total differential around some reference point at time $t-1$, with $\Delta $ the difference operator ($\Delta\widetilde{p}_{ikt} = \widetilde{p}_{ikt} - \widetilde{p}_{ikt-1}$ and so on):
+\begin{eqnarray*}
+&&\Delta \widetilde{p}_{ikt} = \beta \bar{p}_{ikt-1}^{\beta-1}(1+\tau^d_{is(k)t-1})\Delta \bar{p}_{ikt} + \alpha \bar{p}^\beta_{ikt-1} (1+\tau^d_{ikt-1})^{\alpha-1}\Delta \tau^d_{is(k)t}  \\
+\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \beta \frac{\Delta \bar{p}_{ikt}}{\widetilde{p}_{ikt-1}} \frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} +\alpha \frac{\Delta \tau^d_{is(k)t-1}}{1+\tau_{is(k)t-1}^d}\frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} \end{eqnarray*}
 
-\paragraph{(ii) On the assumptions underlining our empirical approach}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla (the Review of Economics and Statistics, 2015) to estimate the share of additive costs in a firm-level context.
-It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
-\textbf{Not sure of what it means?
-} Nevertheless, to take full account of the referee’s remark, we decided run an additional set of estimations based on the Hummel's methodogy as proposed by the referee, based on Equation (\ref{eq:estimation_ref1}).
-We yet came to the conclusion that this method faces some important limitations, as we develop in Section XXX of this letter.
-As a result, we have decided to maintain our estimation strategy as the benchmark method in the main body of the paper.
-We hope that our letter will convince the referee of the relevance of this choice.
+leading to:
+\begin{equation}
+\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} =  \beta \frac{\Delta \bar{p}_{ikt}}{\bar{p}_{ikt-1}} +\alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{ikt-1}^d} \label{eq:firststage_Deltalog}
+\end{equation}
 
-\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier?}
+The intuition behind the Referee 1's endogeneity concerns, is that we need to eliminate from the fas price, any endogenous component that might me related to transport costs.
+To do so, the referee suggests that we instrument the export price by tariff rates.
+Put it in plain words, this suggests to run the first-stage regression based on Equation (\ref{eq:firststage_Deltalog}) according to:
 
-\paragraph{(iii) Endogeneity problem}
-Indeed, this is a very important point.
-The referee states that, based on theoretical insights by Melitz (Econometrica, 2003) or Baldwin and Harrigan (AER, 2011), $1/\tilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
-In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
-We obviously do not question this conceptual issue.
-However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices.
-Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out.
-That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
-Following the referee's advice, we decided therefore to provide some IV-estimates to provide a clean assessment of the size of the potential bias.
+$$\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \alpha \frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \gamma_{i} +\gamma_{k}+\epsilon_{ik}$$
 
-\paragraph{Summary} We decided to answer to these two last concerns (ii) and (iii) in a separate manner.
-Specifically, we run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as exposed); this way, the results are comparable with our benchmark results of the submitted version, the only change being the estimation strategy.
-In a second step, we implement IV methods on fas prices at the $k=$ 5-digit level, still based on our original estimation strategy.
-This allows us a direct comparison of the results obtained in the submitted version (NLS OLS regression) with those obtained with IV.
+Or equivalently, taking logs:
+$$\Delta \log \widetilde{p}_{ikt}= \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{k}+\epsilon_{ikt}$$
 
-\subsection{Exploring Referee's alternative functional form}
-
-\textit{A more natural approach is what the authors, at some point, refer to as the
-Hummel's Methodology. That is, one can alternatively estimate the share of the
-additive component as:}
-
-$$\frac{t_{ik}}{ (\tau_{ik}-1)\tilde{p}_{ik} + t_{ik}} = \beta_{ik}$$
-
-\textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t.
-unit price} [\textbf{in absolute value, see Equation \ref{eq:Hummels}}].
-\textit{Given the authors' objective and the data they are using, $\beta$ can be separately estimated
-for each industry-country pair using the following regression:}
+\noindent with the LHS being the growth rate of the fas price (between $t$ and $t-1$), the first term in the RHS the change in duty tax rates, the second and third terms fixed effects to capture changes in the ``firm-specific price'' $\bar{p}_{ik}$, $\epsilon_{ik}$ being the residual.
+Notice though that the structure of fixed effects should remain consistent between the first and the second stages.
+This implies to rather consider the following functional form of the first-stage equation:
 
 \begin{equation}
-\ln f_{ikd} = \beta_{is(k)}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
+\Delta \log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_Deltalog}
 \end{equation}
 
-\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product. The
-identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
-district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
-obviously require that the authors do not aggregate up the raw Census data
-across all districts and all 10-digit products pertaining to the same 5-digit category.}
-
-\textit{The first advantage of this so-called Hummel's Approach is that the above
-regression can be estimated separately for various country-industry pairs, without
-imposing Assumptions (a) and (b), outlined above.}
-
-\textbf{I report here the referee's footnote: Should we deal with this?
-If he/she is not convinced because small sub-sample, do it on a much larger sample?
-This would better convince him that the separability assumption is not an issue.}: \textit{The authors do check the robustness of their results w.r.t.
-the separability assumption, but this only done for a small sub-sample.}
-
-\paragraph{What we do}
-
-Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
-We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
-To sum up our methodology:
+If this reasoning is correct,
 \begin{itemize}
-\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
-\item We rebuild
-\begin{eqnarray*}
-\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
-\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
-\end{eqnarray*}
-\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
-\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
-
-$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
-\end{itemize}
-
-\paragraph{What the referee suggests to do}
-
-The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
-As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs.
-The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
-Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
-
-Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{Notice that the referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$), while we followed his suggestion considering the 3-digit level.
-We made this choice in view of being able to compare the pros and cons of the referee's method versus ours.
-Both methods require a difference in the sectoral degree at which either the share (Referee's method) or the level (our method) of additive costs are estimated ($s$) and the classification degree at the product level ($k$).
-In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004.
-For comparability reasons, we maintain the same degree of classification $k=5$ for the following years (based on the original Census dataset).
-As a result, transport costs are estimated at the 3-digit level as benchmark estimate, exploiting heterogenity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
-This has driven us to retain $s=3$ when implementing the referee's method, in view of preserving comparability between both strategies.}
-
-\subsubsection{Strategy of answer}
-
-\paragraph{a) The advantages are not as important as suggested by the referee}
+\item we should take as predicted value only the predicted price without the fixed effects:
+$$\left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $$
+\item and we should expect a value of the coefficient $\widehat{\alpha}$ between $\Delta \log \widetilde{p}_{ikt}$ and $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $ between $-1$ and $0$, depending on the degree of ``pricing-to-market'' of firms.
     \begin{itemize}
-\item[Advantage 1]: One is not compelled to assume separability between the country-component and the product-component of transport costs.
-\item[$\Rightarrow$] Answer can be in 2 parts: 1) This is not a strong assumption - better convince him/her: Run the robustness to the separability assumption on a larger sample and show that the estimated $\beta$ is not that different between the two methods.
-2) Compare the estimated $\beta$ between our original method and the referee's one.
-\
-\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}
-\textbf{GD20200629 -- Voir le plan des do. Pour comparaison scatter-chronology-baselinesamplereferee1-referee1.pdf (sur le même sample))}
-
-\textbf{stop here}
-
-\item[Advantage 2]: There is a handful of instruments (HS-10 product-specific tariff rates or lagged prices) to instrument fas prices
-\item[$\Rightarrow$] Answer: When we instrument fas prices at the 5-digit level using tax duties (still using our own methodology), we do not obtain much different results.
-\textbf{But he/she suggests to do it on HS-8 fas prices, not 5-digit prices, and to imply his estimated equation.
-Have we done that?}
-\textbf{GD20200629 Cela ne semble pas vrai. Il y a quelque années en HS8, plus en HS6. Mais pas sur toute la durée}
-\item[Advantage 3]: A more transparent comparison with Hummel (2007): How to answer to this point? Re-read the paper to see if we are clear enough?
-\textbf{GD20200629 changer la manière dont on presente en mettant les beta en avant ?}
-\end{itemize}
-
-\paragraph{b) The costs are high}
-\begin{itemize}
-\item[Concern 1] The suggested estimation strategy implies having much less data to exploit, for two reasons.
-\begin{enumerate}
-\item Information about the port of entry are only available in the recent years (\textbf{since when?
-why?
-before, we rely on Hummel's data set, but we could buy the original US custom data for 1974-2004?
-Or even in those dataset it does not exist?}).
-In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time, and most importantly how the specific values of the additive and ad-valorem transport costs components have evolved over time.
-Eliminating this dimension of the paper would be detrimental to its value-added.
-\item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 8-d sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors (even at the HS-6 digit classification) or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
-This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table XXX \textbf{TO DO}.
-Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
-In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
-
-\textbf{GD20200629 Les données commencent en 1989. On va demander}
-
-    As a direct consequence of this, for things to be comparable across methods, we re-run our estimation strategy on the same sample of observations as the one used in the referee's method.
-
-    \end{enumerate}
-\item[Concern 2] The suggested estimation method implicitly assumes that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors; in both cases, one might view this as a disputable assumption.
-By contrast, we suppose that transport costs are sector-country specific.
-\textbf{FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique?}
-\item[Concern 3] The suggested estimation method does not allow to recover the values of each component of transport costs separately, but only the relative shares; while our methodology drives an estimate for each component ($\tau$, $t$) from which we can deduce the relative share of each.
-In this respect, we gain in generality.
-\textbf{DEEPEN this for the Big picture part}.
-\textbf{GD20200629 Je ne comprends pas. Si on a les parts et les coûts totaux, on peut reconstruire $\tau$, $t$}
-\item[Concern 4] The suggested method features less accuracy in the estimation of the $\beta$.
-Considering the value of the $\beta$ by itself, there is no criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
-Things are more clear-cut in terms of accuracy of the estimation.
-Specifically, our method yields a more precise estimation of the $\beta$ than the referee's method.
-To develop on this, for each year and transport mode:
-    \begin{itemize}
-    \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$.
-From this, we can deduce the 5- 95\% threshold values through:
-        $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
-
-    \item  With our method, one estimation allows to calculate each transport cost component $\hat{t}_{is}$, $\hat{\tau}_{is}$ and an associated $\hat{\beta}_{is}$.
-We generate a distribution of $\beta$ through bootstrap method (10,000 random draws), from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
-    \item We can then evaluate the accuracy of each estimation method by comparing the distributions of the $\beta$.
-This is reported in the Figures XXX (for Air) and XXX (for Vessel).
-Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
-For sake of brevity, we report the mean distribution averaged over the period, a similar conclusion applies on a yearly basis.
-\textbf{Averaged over the two transport modes?
-For only one?}
-
-\begin{figure}[htbp]
-\caption{Accuracy of the estimation of the $\beta$: Comparison}
-\label{fig:accuracy_beta}
-\begin{center}
-\includegraphics[height=4in]{accuracy_beta.pdf}
-\end{center}
-\end{figure}
-
-Figure \ref{fig:accuracy_beta} reports the log of the interval confidence of the $\beta$ obtained with our methodology (with bootstrap), relative to the one obtained with the referee's method.
-\textbf{average over the years?
-the transport modes?
-unclear}.
-In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
-
+    \item[-] $\alpha = 0$ corresponds to the case where the firm does not adjust its fas price to the change in tax duty, that would cancel out the influence of the tax duty change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition;
+    \item[-] $\alpha <0$ corresponds to the degree of``pricing-to-market'' as the firm offsets the impact of the tax change of the price paid by the final consumer by adjusting her producer price in the opposite direction of the tax change,
+    \item[-] with $\alpha = -1$ being the extreme case of ``full PTM'' where the firm fully compensates the tax duty change.
+As shown by Berman, Martin, Mayer (REStats 2012), this is more likely larger firms.
     \end{itemize}
@@
 
-All these elements drive us to maintain our original estimation in the revised version. \textbf{What do we do with the referee's estimation method? online appendix, Appendix?}
+Once this is done, we can rebuild the instrumented fas price in level (at time $t$):
 
-\subsection{Other comments}
+\begin{eqnarray*}
+\widetilde{p}^{IV}_{ikt} = \widetilde{p}_{ikt-1}\left[1 + \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}\right]
+\end{eqnarray*}
 
-\textit{If the authors strictly prefer to maintain their current
-approach, [...] they should address two issues either way:}
+with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}$ the predicted growth rate of the fas price based on duty changes.
 
-\begin{itemize}
-\item \textit{\textbf{The endogeneity problem}: quoting Footnote 14 of the paper, the authors
-are estimating ti and ts(k) as coefficients on the industry or country
-dummies times $1/\widetilde{p}_{ik}$. [...] Based on the productivity-sorting model in Melitz (2003) or the quality-sorting
-model in Baldwin and Harrigan (2010),  $1/\widetilde{p}_{ik}$ is either positively
-or negatively correlated with $\epsilon_{ik}$. So, the NNLS estimates are biased; and
-the bias has nothing to with the casual versus accounting interpretation of
-the estimates. Accordingly, the one-line justification the authors provide
-to not address the endogeneity problem is far from convincing.}
+\subsection{First-stage equation in level}
 
-\item[\text{Answer}] As stated above, we followed the referee's advice by providing some IV-estimates of the fas price to provide a clean assessment of the size of the potential bias in our original methodoly.
-The IV-strategy is implemented in two steps.
-    \begin{itemize}
-    \item First-stage equation: Following the referee's advice, we regress the fas price $\widetilde{p}_{ik}$ with $k$ = 5-digit product classification and $i$ the origin country, on tax duty and lagged fas price according to Equation (XXX).
-As made clear in the appendix of the revised version (\textbf{Section XXX - take elements of topo-first-stage there}), we run the regression in log level on a yearly basis and by transport mode, consistently with our second-stage empirical strategy.
+Alternatively, one can run the first-stage equation in log level, consistently with the second-stage log-level specification.
+Further, in accordance with the referee's suggestion, this also implies including the lagged price as instrument.
+Starting from Equation (\ref{eq:firststage_Deltalog}), we thus run as first-stage specification:
 
-        Develop according to the results.
+\begin{equation}
+\log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \beta \log\widetilde{p}_{ikt-1}  +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_log}
+\end{equation}
 
-    \end{itemize}
-    \item \textit{\textbf{The aggregation problem}: The original annual Census data reports
-trade at the origin country-HS10 product-district level of aggregation,
-whereas the authors are aggregating up the data even further to the origin
-country-HS5 industry-year level. Such an aggregation comes with strong implicit assumptions and sacrifices a lot of useful variation in the data.
-The authors are motivating the aggregation by stating that the problem
-would become computationally expensive without it. But this reasoning
-brings us back to my original point that the authors can use the Hummel's
-Methodology to circumvent the computational burden.}
+Consistently with the cross-section analysis adopted by now, the first-stage equation is estimated on a yearly basis (and by transport mode). We then build the instrumented fas price
 
-\item[\text{Answer}] It is true that the original annual Census data reports trade at the origin country-HS8 product-district level of aggregation (not HS10) \textbf{(I am right there?)}.
-We hope we have now convinced the referee of the relevance of our empirical method based on Equation (\ref{eq:equation0}).
-In line with the referee's comment though, we could run the estimation by considering $k$ at the HS8 classification level, not 5.
-Two elements of answer: 1) we can do it on the recent years, 2005-2013, assuming $s(k) = 3 digit and $k= 8 digit.
-Would take time, but technically feasible.
-And compare it with what we obtain in our benchmark results with $k=5$.
-\textbf{Right?} ; 2) on previous years however, we can't run this estimation, because we take Hummel's database directly, in which the finest degree of sectoral classification is at the 5-digit level.
-\textbf{Counter-argument:} Yes, but we could buy the original Census data on those years?
-\textbf{Answer to the counter-argument} No money, but this is not an argument...Better answer in my view: If we follow the referee's advice and run our estimation with $k=8$ digit on the years for which we have the data, compare with benchmark $k=5$, crossing fingers for that hopefully, things are not that different.
-In which case, this would justify NOT wasting time and money on building a new dataset on the previous years.
-Moreover, sticking to the original Hummel's database is interesting as it eliminates any other difference in the results as the estimation strategy.
-
-\end{itemize}
-
-\subsection{To sum up: What remains to be done}
-
-\begin{enumerate}
-\item\label{bp:1} IV-First-stage equation on fas price at the 8-digit level, on recent years; then, run the estimation of the $\beta$ according to the referee's approach (considering $s(k)$ at the 3 digit level)
-\item\label{bp:2} Run our original estimation method but with $k$ = 8-digit level on recent years for which we have the data.
-Also with $\widetilde{p}_{ik}$ instrumented as well?
-To be comparable with the point just above, yes.
-And compare with the $\beta$ obtained above (point \ref{bp:1}).
-\item\label{bp:3} A table comparing the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy.
-Points (\ref{bp:1}, \ref{bp:2} and \ref{bp:3}) aim at providing answers to the general comment of the referee's Critique 1.
-\item\label{bp:4} Comparison of the results of the $t,\tau,\beta$ obtained with our estimation method, for $k=5$ (what we have) and for $k=8$ (Referee's point 2), for recent years - No IV here?
-\item\label{bp:5}IV-First-stage equation on fas price at the 5-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version (Referee's point 1).
-\end{enumerate}
-
-\section{Critique 2: Calculation of Unit Prices}
-
-\textit{My second critique concerns the way the authors are calculating the unit prices.
-The Census data reports the quantity of goods per observation. So, the authors
-can calculate the unit price as Value/Quantity, which is consistent with how
-price is modeled in standard trade models. Instead, the authors calculate unit
-price as Value/Weight. This used to be a common exercise in the past where
-many data-sets did not report Quantity. But, given their data, there is no
-justification for the authors to calculate the prices this way.}
-
-The referee is right in the sense that Census data do report the quantity of goods.
-However, this information is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
-This explains why we cannot calculate the unit price as suggested by the referee.
-\textbf{We make this point clear in Footnote XXX}
-
-\textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
-Et à ce moment là, on peut répondre ``vraiment'' à sa critique?}
-
-At a more fundamental level, one might argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
-For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
-\textbf{Do we have evidence, even anecdotal, on this point?
-Is it an argument we want to make, See what Lashkaripour says about this point also.}
-
-\textit{Calculating the unit price as Value/Weight presents the authors
-with an additional endogenity problem. To elaborate, let $\omega_{ik} = Weight/Quantity$
-denote the unit weight of the goods in observation $ik$. Also let $\widehat{p}_{ik} = Value/Quantity$
-(unlike what the authors assume) denote standard definition of price}.
-As a consequence, making the link with the fas price we consider in the paper: $\widetilde{p}_{ik} = \widehat{p}_{ik} / \omega_{ik}$.
-
-\textit{This paper is essentially estimating the following equation}:
-
-\begin{equation*}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}/\omega_{ik}} \right) + \epsilon_{ik}
-\end{equation*}
-
-\textit{instead of estimating}:
-
-\begin{equation*}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}} \right) + \epsilon_{ik}
-\end{equation*}
-
-\textit{There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
-categories, and (ii) $\omega_{ik}$ is negatively correlated with transport costs. So the
-way the authors are calculating unit prices and estimating the model creates a
-new (but avoidable) source of endogeneity.}
-
-We thank the referee for pointing out this potential source of bias in our estimates.
-If we cannot follow the referee's advice be replacing weight by quantity for data availability reasons, we can address his/her concern regarding this as source of endogeneity bias by instrumenting the fas price.
-\textbf{il ne faudrait pas dans ce cas mettre de lagged prices, sinon on reintroduit du bruit.
-Non?}
-
-\section{Critique 3: Big Picture Implications}
-
-\textit{My third critique concerns the lack of an exciting punchline. The fact that
-composition effects have not countervailed the reduction in pure transport costs
-(at least not as much as previously believed) is an interesting but minor observation.
-Does this observation revise our understanding of say the gains from
-trade? Does it shed new light on a puzzle many people are thinking about?
-One crude suggestion is to see how the reduction in the industry-specific cost
-terms is related to the industry-level trade elasticities. If the composition effects
-favor low-elasticity industries, the findings in the paper may have first-order
-implications for the gains from trade.
-Another suggestion is to dig deeper into the relative rate at which additive
-and multiplicative transport costs have declined over time. Since additive transport
-costs favor rich (high-quality exporting) countries, the disproportionally
-greater reduction in additive costs can perhaps explain the rise of low-income
-exporter as documented by Hanson (2012, JEP).}
-
-\textbf{Comments about that}
-
-\paragraph{Piste 1} Consider his/her first suggestion.
-What is the idea?
-First, the link between gains from trade and trade-elasticity.
-Am I correct in saying that gains from trade are higher when trade is about low-elasticity goods?
-The idea being, if national goods are low substitutes, then here are larger gains from trading them.
-So, if composition effects (trade shifts between sectors) favor low-elasticity sectors, then one might expect high gains from trade.
-Hummels: trade composition effects matter as they partially offset the reduction in pure transport costs for both air and vessel.
-Over time, tendency to trade goods more costly to export everything else equal.
-So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade.
-But, we disagree with Hummel's findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air.
-Meaning that gains from trade are purely due to reduction in transport costs within sector, but not from switch towards low-elasticity sectors.
-Identify one source of gains from trade, not two.
-
-What to do with this?
-
-\paragraph{Piste 2} What is at stake?
-Show that how $\beta$ = the relative share of additive costs has evolved over time.
-Good idea as this is really what differentiates us from Hummels, model a varying $\beta$ over time.
-Specifically, how methodology lets the $\beta$ vary over the three time-sector-partner country dimension.
-Show how $\beta$ varies over time everything else equal?
-Hopefully, goes in the referee's direction (decreases over time) with possibly a role in the understanding in the ``big picture'' of trade patterns?
+\begin{eqnarray*}
+&&\log \widetilde{p}_{ikt}^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \widehat{\beta}\log\widetilde{p}_{ikt-1} \\
+\rightarrow &&\widetilde{p}_{ikt}^{IV}  = \exp^{\log \widetilde{p}_{ikt}^{IV}}
+\end{eqnarray*}
 

### 2020-07-02 redaction/submission_JEGeo/revision/referees/answers/topo_first_stage.tex
changement réponse référé 1

@@
 
-Denoting $i$ the origin country, $k$ the product at the 8/5 digit level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
+Denoting $i$ the origin country, $k$ the product at the 5 digit level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
 
@@
 
-\noindent with $s(k)$ the 3-digit classification (the disaggregation level for tax duties in the database) as a function of the HS8 (or 5 digit) product classification
+\noindent with $s(k)$ the 3-digit classification as a function of the 5-digit product classification.
 
@@
 \item we should take as predicted value only the predicted price WITHOUT the fixed effects:
-$$\widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $$
+$$\left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $$
 \item and we should expect a value of the coefficient $\widehat{\alpha}$ between $\Delta \log \widetilde{p}_{ikt}$ and $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $ between $-1$ and $0$, depending on the degree of ``pricing-to-market'' of firms.
@@
 \begin{eqnarray*}
-\widehat{\widetilde{p}}_{ikt} = \widetilde{p}_{ikt-1}+ \widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}}
+\widehat{\widetilde{p}}_{ikt} = \widetilde{p}_{ikt-1}\left[1 + \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}\right]
 \end{eqnarray*}
 
-with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \widehat{\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}}$ the predicted growth rate of the fas price based on duty changes.
+with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}$ the predicted growth rate of the fas price based on duty changes.
 

### 2020-07-08 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
Commentaires Guillaume sur answer referee1

@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+Université Paris-Dauphine, PSL University, CNRS, 8007, IRD, 260, LEDa, DIAL, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
 University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
@@
 
-In italics, I report the Referee's comment. Into brackets in bold characters inserted into the referee's statement, my comment.
+In italics, I report the referee's comment. My comments are bold and between brackets inserted into the referee's statement.
 
@@
 
-The three points raised by the referee are all equally important, and we answer separately to each of them below.
+The three points raised by the referee indeed deserve careful consideration.
 
@@
 However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
-This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
-$\tau \geq 1$ and $t \geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative.
-Without these restrictions, standard linear, least squares estimates often deliver aberrant values with more than mild quality-of-fit \textbf{[insert examples here]}
-We do acknowledge this is a very important point, which was not made clear enough in the initial version, and did our best to make the pint clear in the revised version \textbf{[see page XX : XX insert quote XX].}
+This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
+$\tau \geq 1$ and $t \geq 0$, or $0 \geq  \beta \geq -1$.
+Without these restrictions, standard linear, least squares estimates often deliver negative trade costs \textbf{GD20200706 Je ne suis pas sûr de ce que tu veux dire ici à propos de la quality of fit ?}
+with more than mild quality-of-fit \textbf{[insert examples here]}
+This motivation was not made clear enough in the initial version, and ze did our best to make this very important justification clearer in the revised version \textbf{[see page XX : XX insert quote XX].}
 
 \paragraph{(ii) On the assumptions underlining our empirical approach}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla (the Review of Economics and Statistics, 2015) to estimate the share of additive costs in a firm-level context.
+Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla in the Review of Economics and Statistics (2015) to estimate the share of additive costs in a firm-level context.
 It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
 \textbf{Not sure of what it means?
-} Nevertheless, to take full account of the referee’s remark, we decided run an additional set of estimations based on the Hummel's methodogy as proposed by the referee, based on Equation (\ref{eq:estimation_ref1}).
-We yet came to the conclusion that this method faces some important limitations, that drove us to maintain our initial estimation strategy in the main body of the paper.
-We develop the analysis we held in Section XXX of the letter, and we hope that it will convince the referee of the relevance of this choice.
+} \textbf{GD20200706 Moi non plus...
+(à propos de la phrase juste avant)}
+We already have a robustness check for this separative assumption. The sample of this check is small in terms of observations, but quite large in terms of trade coverage. 
+\textbf{GD20200706 Il faudrait donner les chiffres de coverage}
+Nevertheless, to take full account of the referee’s remark, we  have estimated an additional set of estimations based on the Hummel's methodogy as proposed by the referee in equation (\ref{eq:estimation_ref1}).
+We found that that this method faces some important limitations, that justify maintaining our initial estimation strategy as the baseline.
+We develop the analysis we held in Section XXX of the letter.
 
-\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier?}
+\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier? GD20200706 dans une partie robustesse ?}
+
+\textbf{GD20200706 On pourrait aussi proposer de présenter les résultats en terme de $\beta$ pour garder la comparabilité avec Hummels ?}
 
 \paragraph{(iii) Endogeneity problem}
+
 Indeed, this is a very important point.
-The referee states that, based on theoretical insights by Melitz (Econometrica, 2003) or Baldwin and Harrigan (AER, 2011), $1/\tilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+The referee states that, based on theoretical insights by Melitz (Econometrica, 2003) or Baldwin and Harrigan (AER, 2011), $1/\tilde{p}_{ik}$ is correlated with residuals $\epsilon_{ik}$.
 In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
 We obviously do not question this conceptual issue.
+\textbf{GD20200706 Hum. En quoi est-ce que la productivité des firmes est dans le résidu ?}
 However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices.
@@
 That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
-Following the referee's advice, we decided therefore to provide some IV-estimates to provide a clean assessment of the size of the potential bias.
+Though, we do not know why that should affect differently additive and multiplicative trade costs.
+Following the referee's advice, we decided therefore to provide some IV-estimates to assess the size of the potential bias.
 
-\paragraph{Summary} We decided to answer to these two last concerns (ii) and (iii) in a separate manner.
-Specifically, we run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as detailed later); this way, the results are comparable with the benchmark results of the submitted version, the only change being the estimation strategy.
-In a second step, we implement IV methods on fas prices at the $k=$ 5-digit level, still based on our original estimation strategy.
-This allows a direct comparison of the results obtained in the submitted version (NLS OLS regression) with those obtained with IV.
+\paragraph{Summary} \textbf{GD20200706 Notice plutôt ?} To clarify the discussion, we decided to answer concerns (ii) and (iii) each in turn.
+Specifically, we first run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as detailed later); this way, the results are comparable with the benchmark results of the submitted version, the only change being the estimation strategy.
+Then, we instrument fas prices at the $k=$ 5-digit level, still based on our original estimation strategy.
+This allows a direct comparison of the results obtained in the submitted version (NLS regression) with those obtained with IV.
 
@@
 This would better convince him that the separability assumption is not an issue.}: \textit{The authors do check the robustness of their results w.r.t.
-the separability assumption, but this only done for a small sub-sample.}
+the separability assumption, but this only done for a small sub-sample.} \textbf{GD20200706 À mon avis, tu peux le faire plus haut quand on discute dans le résumé.
+Mais sinon, tu peux en faire un préambule à ce qui suit.
+On peut donner les résultats}
 
@@
 The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
-As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs.
+As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ \textbf{GD20200708 valeurs absolues ?} in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs.
 The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
@@
 \item Information about the port of entry are only available since 1989.
-Implementing the referee's method would hence necessarily reduce the time coverage of our analysis (1974-1988).
+Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by 15 years (1974-1988).
 In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
@@
 Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+It excludes XXX \% of trade flows and XXX \% of trade in value.
 In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
@@
     \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$.
-From this, we can deduce the 5- 95\% threshold values through:
+From this, we can approximate the 5- 95\% threshold values through:
         $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
@@
 We generate a distribution of $\beta$ through bootstrap method (10,000 random draws), from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
-    \item We can then evaluate the accuracy of each estimation method by comparing the distributions of the $\beta$.
+
+    \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$.
 This is reported in the Figures XXX (for Air) and XXX (for Vessel).
@@
 \textbf{Averaged over the two transport modes?
-For only one?}
+For only one?} \textbf{GD20200708 Je ne comprends pas bien ce que tu veux dire.
+Il me semble que les graphiques sont annuels et ne sont pas des moyennes:ce sont des distributions}
 
@@
 All these elements drive us to maintain our original estimation in the revised version. \textbf{What do we do with the referee's estimation method? online appendix, Appendix?}
+\textbf{GD20200708 Appendix à mon avis.
+Mais on peut garder la méthode de mesure de précision des betas dans la baseline dans le corps du texte et le remercier abondamment de nous avoir donné l’occasion de discuter ces questions ?}
 
@@
 Potentially we could ask for data since the beginning 1974?
-}, assuming $s(k) = 3 digit and $k= 8 digit.
+}\textbf{GD20200708 J’ai juste demandé les données en HS10 et on m’a dit qu’elles commençaient en 1989.
+Donc ce n’est pas un problème de port of entry, je crois}, assuming $s(k) = 3 digit and $k= 8 digit.
 Would take time, but technically feasible.
@@
 In which case, this would justify NOT wasting time and money on building a new dataset on the previous years.
-Moreover, sticking to the original Hummel's database is interesting as it eliminates any other difference in the results as the estimation strategy.}
+Moreover, sticking to the original Hummel's database is interesting as it eliminates any other difference in the results as the estimation strategy.} \textbf{GD20200708 Je ne comprends pas ce dernier argument}
 
@@
 \textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
-Et à ce moment là, on peut répondre ``vraiment'' à sa critique?}
+Et à ce moment là, on peut répondre ``vraiment'' à sa critique?} \textbf{GD20200708 Sauf que du coup, on change le sample.
+Il faudrait vérifier si c’est de beaucoup ou pas}
 
@@
 
+\textbf{GD20200708 Pourquoi n’introduit-il pas $\widehat{p}$ à gauche ?}
 \textit{There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
@@
 }
+\textbf{GD20200708 Moi non plus je ne comprends pas très bien}
 
@@
 
+\textbf{GD20200708 Hum...
+Je crois que je comprends le point dur referee et je suis d’accord avec lui.
+Le souci, c’est que de toutes les manières, je ne sais pas si cela favorise les high ou les low elasticity.
+Mais on pourrait voir.}
+
 \paragraph{Piste 2} What is at stake?
@@
 Hopefully, goes in the referee's direction (decreases over time) with possibly a role in the understanding in the ``big picture'' of trade patterns?
+
+\paragraph{Piste 3} \textbf{GD20200708 On sait que les additive costs sont plus distorsifs que les multiplicative costs, puisqu’ils modifient les prix relatifs.
+Donc le fait qu’ils sont une partie croissante des costs signifie qu’on sur-estime l’augmentation des gains de welfare liés à la baisse des coûts de transport}
 

### 2020-07-14 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
changement réponse

@@
 This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
-$\tau \geq 1$ and $t \geq 0$, or $0 \geq  \beta \geq -1$.
-Without these restrictions, standard linear, least squares estimates often deliver negative trade costs \textbf{GD20200706 Je ne suis pas sûr de ce que tu veux dire ici à propos de la quality of fit ?}
+$\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
+Without these restrictions, standard linear, least squares estimates often deliver negative trade costs \textbf{GD20200706 Je ne suis pas sûr de ce que tu veux dire ici à propos de la quality of fit ?
+Moi non plus}
 with more than mild quality-of-fit \textbf{[insert examples here]}
-This motivation was not made clear enough in the initial version, and ze did our best to make this very important justification clearer in the revised version \textbf{[see page XX : XX insert quote XX].}
+This motivation was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version \textbf{[see page XX : XX insert quote XX].}
 
@@
 
-\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier? GD20200706 dans une partie robustesse ?}
+\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier? GD20200706 dans une partie robustesse ? ok}
 
-\textbf{GD20200706 On pourrait aussi proposer de présenter les résultats en terme de $\beta$ pour garder la comparabilité avec Hummels ?}
+\textbf{GD20200706 On pourrait aussi proposer de présenter les résultats en terme de $\beta$ pour garder la comparabilité avec Hummels ? ok je peux m'en charger}
 
@@
 We obviously do not question this conceptual issue.
-\textbf{GD20200706 Hum. En quoi est-ce que la productivité des firmes est dans le résidu ?}
+\textbf{GD20200706 Hum. En quoi est-ce que la productivité des firmes est dans le résidu ? en fait pourquoi parle-t-il de 1/ptilde? }
 However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices.
@@
 
-\paragraph{Summary} \textbf{GD20200706 Notice plutôt ?} To clarify the discussion, we decided to answer concerns (ii) and (iii) each in turn.
+\paragraph{Summary} \textbf{GD20200706 Notice plutôt ?
+a ce stade c'est preliminaire} To clarify the discussion, we decided to answer concerns (ii) and (iii) each in turn.
 Specifically, we first run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as detailed later); this way, the results are comparable with the benchmark results of the submitted version, the only change being the estimation strategy.
@@
 \textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t.
-unit price} [\textbf{in absolute value, see Equation \ref{eq:Hummels}}].
+unit price} in absolute value, as shown with more details in Appendix \ref{app:interpret_beta} through Equation (\ref{eq:Hummels}).
 \textit{Given the authors' objective and the data they are using, $\beta$ can be separately estimated
@@
 This would better convince him that the separability assumption is not an issue.}: \textit{The authors do check the robustness of their results w.r.t.
-the separability assumption, but this only done for a small sub-sample.} \textbf{GD20200706 À mon avis, tu peux le faire plus haut quand on discute dans le résumé.
-Mais sinon, tu peux en faire un préambule à ce qui suit.
-On peut donner les résultats}
+the separability assumption, but this only done for a small sub-sample.} \textbf{GD20200706 a mon avis, tu peux le faire plus haut quand on discute dans le résume.
+Mais sinon, tu peux en faire un preambule à ce qui suit.
+On peut donner les resultats}
 
@@
 The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
-As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ \textbf{GD20200708 valeurs absolues ?} in Equation (\ref{eq:Hummels}), also corresponds to the share of additive costs in total transport costs.
+As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
 The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
@@
 \textbf{GD20200708 Appendix à mon avis.
-Mais on peut garder la méthode de mesure de précision des betas dans la baseline dans le corps du texte et le remercier abondamment de nous avoir donné l’occasion de discuter ces questions ?}
+Mais on peut garder la méthode de mesure de précision des betas dans la baseline dans le corps du texte et le remercier abondamment de nous avoir donné l’occasion de discuter ces questions ?
+ok, appendix du papier plutot}
 
@@
 It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
-We hope we have now convinced the referee of the relevance of our empirical method based on Equation (\ref{eq:equation0}).
-In line with the referee's comment though, we could run the estimation by considering $k$ at the HS10 classification level, not 5.
-Two elements of answer: 1) we can do it since 1989 only \textbf{CAUTIOUS, 1989 is because it is the first year for which we have information about the port of entry, but here we don't need it.
-Potentially we could ask for data since the beginning 1974?
-}\textbf{GD20200708 J’ai juste demandé les données en HS10 et on m’a dit qu’elles commençaient en 1989.
-Donc ce n’est pas un problème de port of entry, je crois}, assuming $s(k) = 3 digit and $k= 8 digit.
-Would take time, but technically feasible.
-And compare it with what we obtain in our benchmark results with $k=5$.
-\textbf{Right?.
-If we follow the referee's advice and run our estimation with $k=10$ digit on the years for which we have the data, compare with benchmark $k=5$, crossing fingers for that hopefully, things are not that different.
-In which case, this would justify NOT wasting time and money on building a new dataset on the previous years.
-Moreover, sticking to the original Hummel's database is interesting as it eliminates any other difference in the results as the estimation strategy.} \textbf{GD20200708 Je ne comprends pas ce dernier argument}
+Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
+We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
+In line with the referee's comment, we now run the estimation by considering $k$ at the HS10 classification level, not 5.
+After checking with the US Bureau of Census, data is only available since 1989, now available to 2019.
+We have hence asked for all available data at the time of revision (1989-2009), and we now present the results considering $s=3$ and $k=10$.
+This however deprives us from being able to exploit the period 1974-1988 at the aggregation level.
+\textbf{Que faire pour la periode 1974-1989}.
+Furthermore, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
+This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
+We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper \textbf{see XXXX}
 
@@
 \begin{enumerate}
-\item\label{bp:1} After running the referee's 1 method with $k=$10, show tables or figures comparing 1) the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy; 2) comparison of the $\beta$ between the two methods; 3) comparison of the precision of the $\beta$ estimation.
+\item\label{bp:3} Re-run our benchmark method with $s=3$ and $k=10$ over available years and compare the results ($t,\tau,\beta$ with our benchmark ones ($s=3,k=5$).
+Hopefully does not change much so that we can keep up the period 1974-1988?
+Consider this as our new benchmark.
+Answer to the critique about the aggregation problem.
+\item\label{bp:1} After running the referee's 1 method with $s=3$, $k=$10, show tables or figures comparing 1) the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy; 2) comparison of the $\beta$ between the two methods; 3) comparison of the precision of the $\beta$ estimation.
 Answer to the critique about the estimation strategy.
-\item\label{bp:2} IV-First-stage equation on fas price at the 5-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version.
+\item\label{bp:2} IV-First-stage equation on fas price at the $s=3$, $k=10$-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version.
 Answer to the critique about the endogeneity bias.
-\item\label{bp:3} Re-run our benchmark method with $s=3$ and $k=10$ over some years and compare the results ($t,\tau,\beta$ with our benchmark ones ($s=3,k=5$).
-Answer to the critique about the aggregation problem.
 \end{enumerate}
@@
 
-\section{Clarifying some technical points \label{ssec:technical_points}}
+\section{Clarifying some technical points \label{app:technical_points}}
 
@@
 
-\subsection{Deriving the share of additive costs in total transport costs \label{ssec:interpret_beta}}
+\subsection{Deriving the share of additive costs in total transport costs \label{app:interpret_beta}}
 

### 2020-09-15 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
Changement réponse référé

@@
 \title{\textsc{International Transport costs:\\New Findings from modeling additive costs} \\
-Asnwer to Referee 1}
+Answer to Referee 1}
 
@@
 
-\date{July 2020}
+\date{September 2020}
  \maketitle
@@
 
-\section{Critique 1: Empirical Strategy}
+We would like to thank you for your insightful comments. They led us to introduce some
+significant changes to the paper that we hope address your concerns. We first give you an overview
+of the revision (Section ) before answering in detail each of your comments (Section ). Number of sections, pages and
+equations refer to the revised version.
 
-In italics, I report the referee's comment. My comments are bold and between brackets inserted into the referee's statement.
+\section{Main changes \label{sec:main_changes}}
+
+\begin{itemize}
+\item Based on your comment 1.2 (\textit{The aggregation problem}), the benchmark set-up now retains $s=3$ and $k=10$ for the industry/product aggregation levels respectively.
+It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
+Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
+We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
+In line with the referee's comment, we now run the estimation by considering $k$ at the HS10 classification level, not 5.
+\textbf{We come back to this point in Section XXX of the letter.}
+
+    This has important consequences in terms of database.
+In the submitted version, we used the original Hummel's database (available on his webpage from 1974 to 2003, which we completed by the US Bureau of census import database over the period 2005-2013.
+In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004.
+For comparability reasons over time, we have adopted the same degree of classification $k=5$ for the following years (based on the original Census dataset).
+As a result, in the submitted version, transport costs are estimated at the 3-digit sector level as benchmark estimate, exploiting product heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
+
+    Switching to $k=10$ as benchmark disaggregation level for the product dimension requires to resort on US Census dataset only.
+As first step, we have re-run the estimation over the years for which we already had the data (2005-2013).
+In parallel, we have bought older years from the Census bureau, as well as the more recent years 2014-2019.
+However, the Census bureau can provide data disaggregated at the $k=10$ product level only starting in 1989.
+This deprives us from the very first years 1974-1988.
+For this period, we can only resort on Hummel's dataset, with $k=5$ the finest product level category.
+This notably induces us to maintain $s=3$ as the benchmark disaggregation level for the sectoral dimension for the whole period.
+This allows us to preserve (as far as possible) the largest time coverage - even though switching from $s=3, k=5$ (1974-1988) to $s=3, k=10$) (1989-2019) undoubtedly constitutes a significant break.
+\textbf{What do we do with that?}
+
+\item Based on your comment 1.1 (\textit{The aggregation problem}), the benchmark regression (non-linear least squares) is now completed by a two-stage procedure, where the fas price is instrumented by tax duties and the lag price \textbf{(See Section XXX).}
+
+\item Based on your main comment 1, the paper has been rewritten so as to be more in line with Hummel's (2004) methodology.
+Specifically, we now start from the share of additive costs $\beta$, from which we can deduce the additive / ad-valorem components ($t,\tau$) rather than the other way round (as in the submitted version).
+We thank the referee for pointing this out, as this contributes to a more straightforward and intuitive understanding of our estimation method.
+
+\item What about the big picture?
+\end{itemize}
+
+We now answer in detail each of your comments given in italics.
+
+\section{Detailed answers \label{sec:detailed_answers}}
+\subsection{Critique 1: Empirical Strategy}
 
@@
 
-\paragraph{(i) On the use of non-linear least squares (NLS)}
+\paragraph{(i) On the assumptions underlining our empirical approach}
+Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla in the Review of Economics and Statistics (2015) to estimate the share of additive costs in a firm-level context.
+It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
+
+Further, we already provide a robustness check for this separative assumption, in Section XXX of the paper (\textbf{Table 2, submitted version}).
+We check the robustness of our results to the separability assumption by re-running the estimation without the separability assumption.
+This comes at the cost of a substantial increase in the number of fixed effects to include in the regression.
+Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted version} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
+This is not tractable computationally.
+For this reason, we have decided to run the robustness check on a reduced sample.
+If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013, vs around 29,000 on the complete sample, see Table 1 ), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+As we conclude at the end of Section XXX , whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
+Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
+
+\paragraph{(ii) On the use of non-linear least squares (NLS)}
 As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
@@
 $\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
-Without these restrictions, standard linear, least squares estimates often deliver negative trade costs \textbf{GD20200706 Je ne suis pas sûr de ce que tu veux dire ici à propos de la quality of fit ?
-Moi non plus}
-with more than mild quality-of-fit \textbf{[insert examples here]}
-This motivation was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version \textbf{[see page XX : XX insert quote XX].}
+When run relaxing these restrictions, standard linear, least squares estimates often deliver negative trade costs, which is meaningless.
+In this respect, implementing the referee's method (see below) does not suppress the requirement of resorting on non-linear estimates (and the computational burden it induces, in particular in terms of time).
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version \textbf{[see page XX : XX insert quote XX].} \smallskip
 
-\paragraph{(ii) On the assumptions underlining our empirical approach}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla in the Review of Economics and Statistics (2015) to estimate the share of additive costs in a firm-level context.
-It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
-\textbf{Not sure of what it means?
-} \textbf{GD20200706 Moi non plus...
-(à propos de la phrase juste avant)}
-We already have a robustness check for this separative assumption. The sample of this check is small in terms of observations, but quite large in terms of trade coverage.
-\textbf{GD20200706 Il faudrait donner les chiffres de coverage}
-Nevertheless, to take full account of the referee’s remark, we  have estimated an additional set of estimations based on the Hummel's methodogy as proposed by the referee in equation (\ref{eq:estimation_ref1}).
-We found that that this method faces some important limitations, that justify maintaining our initial estimation strategy as the baseline.
-We develop the analysis we held in Section XXX of the letter.
-
-\textbf{Methode du refere: en annexe? annexe online? rien du tout dans le papier? GD20200706 dans une partie robustesse ? ok}
-
-\textbf{GD20200706 On pourrait aussi proposer de présenter les résultats en terme de $\beta$ pour garder la comparabilité avec Hummels ? ok je peux m'en charger}
+Of course, this does not prevent us for taking full account of your suggestion to adopt an alternative estimation method (summarized in Equation (\ref{eq:estimation_ref1}).
+It turns out that this method faces some important limitations, that justify maintaining our initial estimation strategy as the baseline in the building block of the paper.
+\textbf{The referee's method is presented in the online appendix, and we refer to it in Footnote XXX of the paper}.
+We develop the analysis we held in \textbf{Section XXX of the letter}.
 
@@
 We obviously do not question this conceptual issue.
-\textbf{GD20200706 Hum. En quoi est-ce que la productivité des firmes est dans le résidu ? en fait pourquoi parle-t-il de 1/ptilde? }
 However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices.
@@
 Following the referee's advice, we decided therefore to provide some IV-estimates to assess the size of the potential bias.
+\textbf{This is made in Section XXX of the paper.}
 
-\paragraph{Summary} \textbf{GD20200706 Notice plutôt ?
-a ce stade c'est preliminaire} To clarify the discussion, we decided to answer concerns (ii) and (iii) each in turn.
-Specifically, we first run the referee's proposed estimation method without relying on instrumental variables (and preserving the same sector-level degree of aggregation, as detailed later); this way, the results are comparable with the benchmark results of the submitted version, the only change being the estimation strategy.
-Then, we instrument fas prices at the $k=$ 5-digit level, still based on our original estimation strategy.
-This allows a direct comparison of the results obtained in the submitted version (NLS regression) with those obtained with IV.
+\paragraph{Strategy of answer} To clarify the discussion, we decided to answer concerns (i) and (iii) each in turn.
+As regards to point (i), we run the referee's proposed estimation method without relying on instrumental variables; this way, the results are comparable with the benchmark results, the only change being the estimation strategy (Section \ref{subsec:functional_form} of the letter).
+As noted above, this drove us to maintain our estimation method as benchmark.
+As regards to point (iii), we instrument fas prices at the $k=$ 10-digit level, still based on our original estimation strategy (but switching to a finest product category disaggregation level $k=10$).
+This allows a direct comparison of the results obtained in the non-linear least squares regression with those obtained with IV (with $s=3$ digit at the sector level and $k=10$ at the product level in both cases).
+This is developed in \textbf{Section ???
+of the letter}.
 
-\subsection{Exploring Referee's alternative functional form}
+\subsubsection{Exploring Referee's alternative functional form \label{subsec:functional_form}}
 
@@
 
-\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product. The
+\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product} ($f_{ikd}$ being the transport costs). \textit{The
 identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
@@
 regression can be estimated separately for various country-industry pairs, without
-imposing Assumptions (a) and (b), outlined above.}
+imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}
 
-\textbf{I report here the referee's footnote: Should we deal with this?
-If he/she is not convinced because small sub-sample, do it on a much larger sample?
-This would better convince him that the separability assumption is not an issue.}: \textit{The authors do check the robustness of their results w.r.t.
-the separability assumption, but this only done for a small sub-sample.} \textbf{GD20200706 a mon avis, tu peux le faire plus haut quand on discute dans le résume.
-Mais sinon, tu peux en faire un preambule à ce qui suit.
-On peut donner les resultats}
+Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
+We have implemented the referee's method over the years for which we already have the data, i.e.
+the years 2005-2013 (already used in the submitted version).
 
-\paragraph{What we do}
+Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset \textbf{over 1974-1988, as the US Bureau of census import database starts in 1989.})
 
-Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
+Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) We develop our results and conclusion in what follows.
+
+%
+%\paragraph{What we do}
+%
+%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
 We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
 To sum up our methodology:
-\begin{itemize}
-\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
-\item We rebuild
-\begin{eqnarray*}
-\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
-\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
-\end{eqnarray*}
-\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
-\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
-
-$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
-\end{itemize}
-
-\paragraph{What the referee suggests to do}
-
-The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
+%\begin{itemize}
+%\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
+%\item We rebuild
+%\begin{eqnarray*}
+%\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
+%\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
+%\end{eqnarray*}
+%\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
+%\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
+%
+%
+%$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
+%\end{itemize}
+%
+%\paragraph{What the referee suggests to do}
+%
+%The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
 As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
@@
 Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
-
-Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{The referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
+%
+%Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{The referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
 We followed his suggestion but considering the 3-digit level.
@@
 
-\subsubsection{Strategy of answer}
+\textbf{stop here}
 
@@
 
-\subsection{Potential endogeneity problem}
+\subsubsection{Potential endogeneity problem}
 
@@
 
-\subsection{The aggregation problem}
+\subsubsection{The aggregation problem}
 
@@
 
-\subsection{To sum up: What remains to be done}
+\subsubsection{To sum up: What remains to be done}
 
@@
 
-\section{Critique 2: Calculation of Unit Prices}
+\subsection{Critique 2: Calculation of Unit Prices}
 
@@
 
-\section{Critique 3: Big Picture Implications}
+\subsection{Critique 3: Big Picture Implications}
 

### 2020-09-18 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex

@@
 
-\textit{The first advantage of this so-called Hummel's Approach is that the above
-regression can be estimated separately for various country-industry pairs, without
-imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}
+The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
+As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
+The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
+Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
 
@@
 
-Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) We develop our results and conclusion in what follows.
+Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) We develop our results and conclusion by answering each argument made by the referee at a time.
 
@@
 
-\textbf{stop here}
+\paragraph{a) The advantages of the referee's method are not as important as suggested}
 
-\paragraph{a) The advantages of the referee's method are not as important as suggested}
 \begin{itemize}
-\item[Advantage 1]: One is not compelled to assume separability between the country-component and the product-component of transport costs.
-\item[$\Rightarrow$] Answer can be in 2 parts: 1) This is not a strong assumption - better convince him/her: Run the robustness to the separability assumption on a larger sample and show that the estimated $\beta$ is not that different between the two methods.
-\textbf{Show that this is not a so small sample, and compare it with the sample induced by the referee's method} 2) Compare the estimated $\beta$ between our original method and the referee's one.
-\
-\textbf{Where are the results comparing the estimates of the $\beta$, in value? I don't find these results anymore...}
-\textbf{GD20200629 -- Voir le plan des do. Pour comparaison scatter-chronology-baselinesamplereferee1-referee1.pdf (sur le même sample))}
+\item \textit{The first advantage of this so-called Hummel's Approach is that the above
+regression can be estimated separately for various country-industry pairs, without
+imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} Our answer is twofold.
 
-\item[Advantage 2]: There is a handful of instruments (HS-10 product-specific tariff rates or lagged prices) to instrument fas prices
-\item[$\Rightarrow$] Answer: As exposed below, our strategy of answer has been to treat the two points (estimation strategy / potential endogeneity bias) separately.
-Since our investigation on the estimation strategy has led us to discard this method as the benchmark one, we need not to instrument fas prices at the HS 10 level, our method featuring a $k=$5-digit classification level.
-\textbf{However, when run at the HS-5 product level, the IV method does not yield results that are much different from the benchmark OLS results, as shown in Table XXX.}
+First, the separability assumption does not seem to be a strong assumption.
+This is based on the conclusion drawn from the robustness check on a reduced sample (in the submitted version, considering $s=3, k=5$, and in the revised version, considering $s=3, k=10$).
+If it is smaller in terms of observations,\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
+Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+As we conclude at the end of Section XXX, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
+Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
 
-\item[Advantage 3]: A more transparent comparison with Hummel (2007): \textbf{changer la maniere dont on presente en mettant les beta en avant ?}
+Second, the referee's method imposes a drastic selection of countries / sectors to be included in the estimation.
+Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects.
+This substantially reduces the number of point estimates, as well as the value of trade flows covered relative to the  benchmark method, as summarized in Table .
+
+\item \textit{The second advantage is that there is a handful of [] instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}  We agree with this point, but we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.
+We thank the referee for pointing this out.
+
+\item \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.} Again, we deeply thank the referee for  \textbf{changer la maniere dont on presente en mettant les beta en avant ?}
 \end{itemize}

### 2020-09-23 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
Changement réponse référé + pgms

@@
 
+Our answer can be articulated in two steps.
+First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
+Second, we emphasize the drawbacks induced by this method.
+
 %
@@
 
-\item \textit{The second advantage is that there is a handful of [] instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}  We agree with this point, but we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.
-We thank the referee for pointing this out.
+\item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}  We thank the referee for this valuable suggestion.
+It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.
 
-\item \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.} Again, we deeply thank the referee for  \textbf{changer la maniere dont on presente en mettant les beta en avant ?}
+\item \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.} We agree with the referee that the comparison with the literature (Hummels (2007) in particular) was not straightforward for the reader in the submitted version.
+We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
+In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
+    \begin{equation}
+    f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta_{ikt}} \label{eq:Hummels}
+    \end{equation}
+
+     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ik}$ the price-elasticity of transport costs, which also corresponds to the share of additive costs in total transport costs (as we now show in the paper's appendix):
+
+     \begin{equation}
+     \beta{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)}-1+\frac{t_{is(k)}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
+     \end{equation}
+
+     As underlined in the paper (both submitted and revised versions), Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method, as he estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, yielding to estimates of $\beta_{is(k)}$ that are invariant over time (but mode-specific).
+
+     Two alternative strategies are yet available.
+Let us start with the referee's method.
+It starts from the functional form linking transport costs and the fas price as specified in the Hummels's equation (\ref{eq:Hummels}).
+Taken in log (on a mode/yearly basis), it is written as:
+     $$\ln f_{ikd} = \beta_{is(k)}\ln \widetilde{p}_{ikd} +\text{Controls}_{ikd}+ \varepsilon_{ikd}$$
+
+     From this, one can then recover the levels of additive / multiplicative transports costs.
+Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ for a given sector-country $i,s(k)$, one can indeed solve the following two-equation system:
+
+\begin{eqnarray}
+p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
+\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
+\end{eqnarray}
+
+\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
+With 2 equations and 2 endogenous variables, the system might be solved.
+
+     Alternatively, our method rather starts from the definition of transport costs as:
+     $$\ln\left(f_{ik}  \right) = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
+from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).
+Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
+
+In this respect, the two methods are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
+Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such.
+We agree that this was not sufficiently transparent in the submitted version.
+We hope that the revised version now fits the referee's expectations on this point.
+
 \end{itemize}
@@
 \item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors (even at the HS-6 digit classification) or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
+Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
 This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table XXX \textbf{TO DO}.
@@
 
-    As a direct consequence of this, for things to be comparable across methods, we re-run our estimation strategy on the same sample of observations as the one used in the referee's method.
+    \end{enumerate}
+\item[Concern 2] Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and across origin countries.
+Put it in plain words, it means that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors.
+In both cases, one might view this as a disputable assumption.
+By contrast, our method exploits the fact that transport costs have a both a country-specific component and a sector-specific component.
+\textbf{This stands in line with ...
+FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique}
 
-    \end{enumerate}
-\item[Concern 2] The suggested estimation method implicitly assumes that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors; in both cases, one might view this as a disputable assumption.
-By contrast, we suppose that transport costs are sector-country specific.
-\textbf{FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique?}
 \item[Concern 3] The suggested method features less accuracy in the estimation of the $\beta$.
@@
     \begin{itemize}
-    \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$.
-From this, we can approximate the 5- 95\% threshold values through:
+    \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
+From this, we can approximate the 5-95\% threshold values through:
         $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
 
-    \item  With our method, one estimation allows to calculate each transport cost component $\hat{t}_{is}$, $\hat{\tau}_{is}$ and an associated $\hat{\beta}_{is}$.
-We generate a distribution of $\beta$ through bootstrap method (10,000 random draws), from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
+        From this, we can deduce the confidence interval $CI^{ref}_{is} = \hat{\beta}_{is}^{max,ref} - \hat{\beta}_{is}^{min,ref}$.
 
-    \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$.
-This is reported in the Figures XXX (for Air) and XXX (for Vessel).
-Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
-For sake of brevity, we report the mean distribution averaged over the period, a similar conclusion applies on a yearly basis.
-\textbf{Averaged over the two transport modes?
-For only one?} \textbf{GD20200708 Je ne comprends pas bien ce que tu veux dire.
-Il me semble que les graphiques sont annuels et ne sont pas des moyennes:ce sont des distributions}
+    \item  Our method does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
+Rather, our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$.
+We then deduce the accuracy of the $\beta$ estimate by relying on bootstrap method.
+Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
+Noticing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
+}
+
+    \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$, for each couple $i,s$ (by year and transport mode).
+Figures XXX (for Air) and XXX (for Vessel) thus report the histogram of the log of the ratio $\frac{CI_{is}}{CI^{ref}_{is}}$, for the year 2012.
 
@@
 
-Figure \ref{fig:accuracy_beta} reports the log of the interval confidence of the $\beta$ obtained with our methodology (with bootstrap), relative to the one obtained with the referee's method.
-\textbf{average over the years?
-the transport modes?
-unclear}.
-In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
+Figure \ref{fig:accuracy_beta} In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
+Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
+We have checked that a similar conclusion applies on other years (they are not reported here for sake of brevity but they are available upon request).
 
@@
 
-All these elements drive us to maintain our original estimation in the revised version. \textbf{What do we do with the referee's estimation method? online appendix, Appendix?}
-\textbf{GD20200708 Appendix à mon avis.
-Mais on peut garder la méthode de mesure de précision des betas dans la baseline dans le corps du texte et le remercier abondamment de nous avoir donné l’occasion de discuter ces questions ?
-ok, appendix du papier plutot}
+All these elements drive us to maintain our original estimation as benchmark method in the revised version, and we deeply hope to convince you of the relevance of this choice.
+However, we also want to keep track of this alternative method as further robustness check.
+Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.
+
+\textbf{stop here}
 
@@
 To show this properly, start from Equation (\ref{eq:system2}) to express additive costs according to:
-$$t_ik = \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
+$$t_{ik} = \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
 
-With $\widetilde{p}_{ik} $ being product-country specific, $t_ik$ is necessarily product-country specific ($i,k$, not $i,s(k)$). Plugging this into Equation (\ref{eq:system1}):
+With $\widetilde{p}_{ik} $ being product-country specific, $t_{ik}$ is necessarily product-country specific ($i,k$, not $i,s(k)$). Plugging this into Equation (\ref{eq:system1}):
 

### 2020-09-25 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement dans organisation des dossiers et du papier

--- 2020-06-17 redaction/submission_JEGeo/revision/revised_article/DHP_JoEG_rev1.tex
@@
 
-\date{January 2019}
+\date{September 2020}
  \maketitle
@@
 \begin{abstract}
-This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2013.
+This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from \textbf{1989 to 2019}.
 First, we document the importance of the per-unit component of transport costs.
@@
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+Further, the sizable elasticity of trade with respect to freight costs obtained by \citet{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
 \citet{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
@@
 If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
-of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.
-
-We provide a thorough empirical investigation of the pattern of international transport costs over time based on US imports flows from 1974 to 2013.
-Using a similar dataset ending in 2004, \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e.
-\textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue.
-As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
-Rather, the evolution of transport costs \textit{per se} constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.
-In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
-In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip
-
-These findings are the result of our empirical setting, which carefully distinguishes between multiplicative and additive transport costs.
-Following \citet{samuelson1954}, most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
-as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} However, a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015}, or \citealp{martin2012}).
-Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs.
-We obtain this by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs.
-To do so, we exploit information contained in the Census Bureau's database of US imports flows over 1974-2013.
-This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis.
-As previously mentioned, this database is the same as that used by \cite{hummels2007}, but extended over more recent years (2005-2013).
-\smallskip
-
-%Based on panel data, \citet{novy13} finds  that US trade costs with major trading partners declined on average by about 40\% between 1970 and 2000.
-%Trade costs declined faster for more developed countries (\citealp{arvis2016trade}).
-%Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few).
-%Using data over 1989-2000, \citet{anderson_wincoop_jel} estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.
-
-%This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}.
+of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip
+
+Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from \textbf{1989 to 2019}.
+Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns.
+While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
+as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\citet{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citet{Irrazabal_2015}, or \citet{martin2012}).
+Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs.
+Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1989-2019 and by transport mode (air or maritime transport).\smallskip
 
@@
 First, additive transport costs are quantitatively sizable.
-On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively (at the 3 digit classification level).
-This is slightly lower than our estimates for the ad-valorem component (2.5\% and 3.2\% for air/vessel respectively on average over the period).
-Put differently, additive costs represent between one-third and almost one-half of the overall transport costs.
+On average over \textbf{1989-2019}, the additive cost is estimated to be \textbf{1.8\% and 2.9\% A AJUSTER} of the export price in air and vessel transport respectively (at the 3-digit classification level).
+This is slightly lower than our estimates for the ad-valorem component \textbf{(2.5\% and 3.2\% for air/vessel} respectively on average over the period).
+Put differently, additive costs represent between \textbf{one-third and almost one-half} of the overall transport costs.
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
@@
 We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.
+Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a \textbf{factor~2}.
 Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip
 
-Second, we find that the main source of the downward trend in overall transport costs, by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost \textit{per se}.
-Trade composition effects have only limited influence, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
-Otherwise stated, pure transport costs have declined less than observed costs.
-This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004).
-Our result thus deepens the paradox of the less-than-expected decline of transport costs.
+Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2013}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
+changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \citet{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
+While Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level.
+Second, as US imports data at the HS-10 level are only available since 1989, we run the analysis over 1989-2019, versus 1974-2004 in \citet{hummels2007}.
+}\textbf{Parler de marge extensive - intensive} \citet{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
+In contrast, we find that trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
 Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
 Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+\textbf{A FAIRE DANS LES REGRESSIONS, METHODE Hummels mais avec s=3, k=10.
+Peut-on le faire?
+For the argument to be convincing, need to do it on the same years with the same sector-product degrees.}
+
 \smallskip
@@
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
-As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
+As shown by \citet{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
 In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out'').
-This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.
-Also, their estimates imply that doubling freight costs increases average free alongside ship (FAS) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.
-Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.
-Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using transaction-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.
-\citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
-This is confirmed by \cite{Irrazabal_2015}.
-While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been made in quantifying such gains.
-One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
-One contribution of this paper is to fill that gap.}
+This is consistent with the empirical findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
+Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e.
+cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
+Using transaction-level trade data from Colombia, \citet{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
+One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
 
@@
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
-This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
-Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2013.
+While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of \textbf{2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.}
+This comparison of results suggests that between \textbf{a third and a half of the additive trade costs} are attributable to the transport sector (depending on the transport mode).
+Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from \textbf{1989 to 2019}.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since \textbf{1989}.
+
+\textbf{Big Picture}.
+Our results have important implications for both international trade and international macroeconomics models.
+Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
+A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
+This is confirmed by \citet{Irrazabal_2015}} These results suggest that important welfare gains can be achieved by reducing additive trade costs.
+On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models.
+In both fields yet, not much progress has been made in quantifying such gains.
+One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
+The quantitative results we provide in the paper allow to make one step in filling that gap.
+\textbf{To develop}.
+\medskip
 
@@
 
-Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
+Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \citet{hummels2007}.
 The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2013 (details on the database provided in Appendix \ref{app:data}).
-We first use customs values, quantities and freight costs to recover FAS and cost-insurance-freight (CIF) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the FOB (free on board) price rather than the FAS price.
-The FAS price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
+We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price.
+The fas price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
 It is also the seller's responsibility to clear the goods for export.
@@
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports FAS prices.} More precisely, the (unit) FAS price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the goods net of transportation costs.
-The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-Our dependant variable is finally computed as the ratio of the CIF price divided by the FAS price.
+While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, \textbf{both items being conditional to the transport mode}; in other words, it is the price for one kg of the goods net of transportation costs, \textbf{by transport mode}.
+The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+\textbf{Cf point 2 refere 1 sur unit prices; avoir quantite plutot que poids, robustesse a faire sur sous-echantillon des produits qui ne passent que par un seul mode de transport).} Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
 By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
@@
 Obviously, this dataset has limitations.
-First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the CIF-FAS price gap only covers freight, insurance and handling costs.
+First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
 It is thus silent about the other dimensions of international trade costs.
@@
 This limits the measurement error bias.
-Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system.
+Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC coding system.
 %This will be crucial to compute transport costs, as we detail below.
@@
 This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g.
-\cite{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period 1974-2013.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
+\citet{Irrazabal_2015}.
+Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period \textbf{1989-2019}.\footnote{Note that, as mentioned by \citet{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
@@
 As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-We estimate international transport costs at the 3-digit classification level, even if data series on the CIF and FAS prices are available at the 5-digit level.
+We estimate international transport costs at the 3-digit classification level, \textbf{exploiting heterogeneity between HS-10 products.} \textbf{La robustesse en 4/10 peut-elle marcher?
+sur les premieres annees pour lesquelles moins de points - je crois?
+Que faire de cette analyse 4d?}
 However, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
@@
 However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.
+Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin.
 
@@
 
-\paragraph{The estimated equation} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
-To do so, we start from the equation that expresses the price $p$ (per kg) of goods paid by the importer (import, or CIF price) as a function of the producer price $\widetilde{p}$ (per kg) (export, or FAS price), given both additive (or per kg) transport costs ($t$) and ad-valorem ($\tau$) transport costs, according to:
+\paragraph{Estimation strategy} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
+To do so, denote by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped.
+In accordance with \citet{hummels2010}, we assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
+In this case, the link between the import, export prices and shipping charges is given by:
+
+\begin{eqnarray}
+&& p = \widetilde{p}+f \notag \\
+\leftrightarrow && p = \widetilde{p}+\widetilde{p}^{1-\beta} X \notag\\
+\leftrightarrow && \frac{p-\widetilde{p}}{\widetilde{p}} = \widetilde{p}^{-\beta} X, \label{eq:tt1}
+\end{eqnarray}
+
+\noindent from which we deduce that $\beta$ is equal to the elasticity of transport costs to the unit price (in absolute value):
 
 \begin{equation}
-p = \tau \widetilde{p}+ t \label{eq:base}
+\frac{\partial TC}{\partial \widetilde{p}} \frac{\widetilde{p}}{TC} = - \beta \label{eq:alternative}
 \end{equation}
 
-\noindent As is usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).
- Let us denote $i$ the origin country, and $k$, the product at the 5-digit level.
-Transforming the above equation (\ref{eq:base}) into a ratio, and removing 1 on both sides, in order to get the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation (skipping the year and transport-mode dimensions in the notations for reading convenience):
+In the first polar case $\beta=0$, then shipping charges are proportional to the fas price, such that $p=\widetilde{p}(1+X)$: Only ad-valorem costs apply.
+At the other extreme, the case $\beta=1$ corresponds to shipping costs being independent of the unit price, with the cif price the sum of the unit price and the shipping cost: Transport costs are fully additive.
+This suggests that, the closer $\beta$ to 1, the more prevalent additive costs in total transport costs.
+One can analytically show that $\beta$ is indeed equal to the share of additive costs in total transport costs.
+To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as:
+
+\begin{eqnarray}
+&& p = (1+\tau) \widetilde{p}+ t \label{eq:tt0} \\
+\leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
+\end{eqnarray}
+
+\noindent with $0\leq \tau<1$ the ad-valorem cost ($1+\tau$ being frequently refereed to as the iceberg trade cost) and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})).
+Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}.
+Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.}
+\begin{eqnarray}
+\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau +\frac{t}{\tilde{p}}} \\
+&=& \frac{-t}{\tau+\widetilde{t}} \label{eq:tt2}
+\end{eqnarray}
+
+Comparing Equations (\ref{eq:tt1}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs (in relative terms) is also equal to the share of additive costs in total transport costs.
+
+\textbf{While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2018}), thereby relying on Equation (\ref{eq:alternative}), we follow another route.
+Specifically, we start from Equation (\ref{eq:tt2}) and exploit our dataset to obtain an explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
+From this, we can deduce the price-elasticity of transport costs, i.e.
+the share of additive costs in total transport costs.}\smallskip
+
+More specifically, let us denote $i$ the origin country, and $k$ the product at the 10-digit level, for each flow that is year- and transport-mode specific.
+For sake of reading convenience though, we skip the year and transport-mode dimensions in the notations.
+Transforming the above equation (\ref{eq:base}) as the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation:
+
+\textbf{stop here. Pb sur 1+ ...}
 
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation}
 
-\paragraph{Estimation strategy} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i-1) + (\tau_k -1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+\paragraph{Obtaining the estimated equation} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $(1+\tau_i)\times (1+\tau_k) -1 \simeq \tau_i + \tau_k$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 \begin{eqnarray}
-\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:ad-valorem}\\
+1+\tau_{ik} &=& (1+ \tau)_{i} \times (1+\tau)_{k} \label{eq:ad-valorem}\\
 t_{ik} &=& t_{i} + t_{k} \label{eq:add}
@@
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the CIF price $p$ cannot be lower than the FAS price ($p_{ik}>\widetilde{p}_{ik}$).
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
 Taking into account this constraint in the estimate implies that the error term should be always positive.
@@
 The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
-See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+See \citet{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
 Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit level ($k$).
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
-same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed CIF-FAS price.
+same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
 As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
@@
 %One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
-%Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
+%Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
 %This is not an issue here, as we are not interested in causal inference.
@@
 The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
-Finally, the bottom panel reports the same set of descriptive statistics, but for the actual CIF/FAS ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
+Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
 Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for air transport at the 3-digit level, removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
@@
 \parbox[l]{12cm}{\footnotesize{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
-The additive term is expressed in fraction of FAS price.
+The additive term is expressed in fraction of fas price.
 For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
@@
 \paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
-Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price).
+Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
 Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
@@
     F stat  &    33.2938 &    14.4215 &    27.7801 &    42.8707 &    28.1280 &     0.1751 &   131.2323 \\
-  adj R²  &     0.4845 &     0.4712 &     0.4768 &     0.4934 &     0.0227 &     0.4452 &     0.5472 \\
-  R² within &     0.0025 &     0.0015 &     0.0023 &     0.0036 &     0.0016 &     0.0000 &     0.0064 \\
+  adj $R^2$  &     0.4845 &     0.4712 &     0.4768 &     0.4934 &     0.0227 &     0.4452 &     0.5472 \\
+  $R^2$ within &     0.0025 &     0.0015 &     0.0023 &     0.0036 &     0.0016 &     0.0000 &     0.0064 \\
 \hline
@@
     F stat  &    24.1509 &    11.4029 &    23.7486 &    31.0489 &    17.2397 &     0.0061 &    80.2158 \\
-  adj R² &     0.5856 &     0.5767 &     0.5878 &     0.5960 &     0.0172 &     0.5364 &     0.6323 \\
-  R² within &     0.0020 &     0.0006 &     0.0021 &     0.0032 &     0.0015 &     0.0000 &     0.0062 \\
-\hline
-\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} }}
+  adj $R^2$ &     0.5856 &     0.5767 &     0.5878 &     0.5960 &     0.0172 &     0.5364 &     0.6323 \\
+  $R^2$ within &     0.0020 &     0.0006 &     0.0021 &     0.0032 &     0.0015 &     0.0000 &     0.0062 \\
+\hline
+\multicolumn{8}{l}{\footnotesize{\emph{Source:} }}
 \end{tabular}
@@
 In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
-This focus is shared with some of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
-Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
+This focus is shared with some of the international trade literature, such as \citet{Lafourcade_Thisse}, \citet{hummels2007} and \citet{Behar_Venables}.
+Specifically, in the same spirit as \citet{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
 
@@
 
-As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the FAS price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
+As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the fas price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
 Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2013.
@@
 Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+As emphasized by \citet{Hummels_1999} and \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
 Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
@@
 He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
+Further, and as in \citet{hummels_skiba}, \citet{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
 
@@
 Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
-Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
-In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
+Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \citet{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
+In doing so, \citet{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
 Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
@@
 \paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
-%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
+%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \citet{hummels2007}.
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
@@
 As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
-This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
+This stands in sharp contrast with the assumption made by  \citet{hummels_skiba} or \citet{hummels2007}, which would imply a distribution concentrated on a single point.
 Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
@@
 \noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
-Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the FAS price.
+Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
 We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
@@
 This drives us to investigate this difference further.
-As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual CIF-FAS price gap).
+As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \citet{hummels2007} in one main dimension.
+Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
 Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+%By contrast, \citet{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
 This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
 
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \citet{hummels2007} exposed above on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
@@
 composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \citeh{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
+In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \citet{hummels2007}.\footnote{Unsurprisingly, this accords with \citet{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
 
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
+Applying \citet{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
 Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
@@
 \parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-The additive term is expressed in fraction of FAS price. 1989 omitted in 3-digit estimation for air.}}
+The additive term is expressed in fraction of fas price. 1989 omitted in 3-digit estimation for air.}}
 \label{tab:robustness_separability}
@@
 
-When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+When aggregating the trade cost measure over the product/country ($i,k$) dimension, \citet{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
@@
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \citet{hummels2007}.
 This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
@@
 This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit CIF-FAS unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
+Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
 Two main findings emerge from our results.
@@
 The term ``price actually paid or payable'' means the total payment (whether direct or indirect, and exclusive of any costs, charges, or expenses incurred for transportation, insurance, and related services incident to the international shipment of the merchandise from the country of exportation to the place of importation in the United States) made, or to be made, for imported merchandise by the buyer to, or for the benefit, of the seller.
-In this respect, the ``custom value'' corresponds to the FAS price (``free-alongside'' price) delivered by the seller.
+In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
 More information is available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}.
@@
 
-The CIF (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
+The cif (cost, insurance, and freight) value represents the landed value of the merchandise at the first port of arrival in the United States.
 It is computed by adding ``Import Charges'' to the ``Customs Value'' (see definitions above) and therefore excludes U.S.
@@
 Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of FAS price.}}}
+Additive term expressed in fraction of fas price.}}}
 \end{tabular}%
@@
 
-In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed CIF-FAS prices gap, in order to deliver a more systematic diagnosis of the importance of additive costs.
+In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fas prices gap, in order to deliver a more systematic diagnosis of the importance of additive costs.
 To do so, we rely on several standard measures of fit.
@@
 
-In this section, we provide a variance decomposition exercise on the observed CIF-FAS price gap.
+In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
 Specifically, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$) and \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
@@
 
-with $x$ the observed CIF-FAS price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
+with $x$ the observed cif-fas price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
 
@@
 \begin{figure}[htbp]
-\caption{Variance decomposition (observed CIF-FAS price gap)}
+\caption{Variance decomposition (observed cif-fas price gap)}
 \label{fig:decomp_variance}
@@
 Two interesting results emerge from Figure \ref{fig:decomp_variance}.
-First, the share of the CIF-FAS price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
+First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
 Both account for between 30 and 40\% of the total variance in air transport, depending on the years considered.
@@
 This provides an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
-Second, the variance of the CIF-FAS price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
+Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
 This holds throughout the period and for both transport modes.
@@
 
-\subsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
-
-In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
-In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' Stata codes provided on his webpage\footnote{\url{http://www.krannert.purdue.edu/faculty/hummelsd/research/jep-transport-cost-data.php}}), to better identify the precise points of difference between our estimation strategies.
+\subsection{Comparing with \citet{hummels2007} \label{app:compare_Hummels}}
+
+In this section, we investigate the difference of results with \citet{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
+In order to do so, we start from the empirical specification implemented by \citet{hummels2007} (also making use of Hummels' Stata codes provided on his webpage\footnote{\url{http://www.krannert.purdue.edu/faculty/hummelsd/research/jep-transport-cost-data.php}}), to better identify the precise points of difference between our estimation strategies.
 
@@
 Consider now the ``fitted ad-valorem rate''.
-\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
+\citet{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
 The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables}''.
@@
 
-After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
+After estimating Equation (\ref{eq:Hummels1}), \citet{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
 
 What are the main differences with our estimation strategy? Five points may be underlined.
-First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
+First, \citet{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
 However, this is a slight difference, as in both cases it ultimately amounts to taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-Second, in contrast to us, \cite{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
+Second, in contrast to us, \citet{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
 This is not likely to make a substantial difference, however, as these fixed effects are by nature constant over time. Accordingly, they only amount to a scale effect.
@@
 
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeh{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
-\cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeth{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
+\citet{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
@@
 Second, the 1974 weighting scheme makes sense as the starting point of our time trend analysis.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in Section \ref{sec:robustness}.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \citet{hummels2007}, as reported in Section \ref{sec:robustness}.
 
 One last difference remains to be commented, concerning the way the additive component of international transport costs is treated.
-As we show below, it turns out to be key in accounting for the difference of results with \cite{hummels2007}.
-Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
+As we show below, it turns out to be key in accounting for the difference of results with \citet{hummels2007}.
+Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \citet{hummels_skiba}.
 Rewriting Equation (\ref{eq:Hummels1}) in a simpler form as:
@@
 
-As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
-By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
+As noted by \citet{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
+By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \citet{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
 This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (i.e.
 allowing for a varying $\beta$ across time, country partner and product).
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007}, outlined above, on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \citet{hummels2007}, outlined above, on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.

### 2020-09-25 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Debug prix fob + chgt texte

@@
 \begin{abstract}
-This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from \textbf{1989 to 2019}.
+This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from \textbf{1974 to 2019}.
 First, we document the importance of the per-unit component of transport costs.
@@
 First, additive transport costs are quantitatively sizable.
-On average over \textbf{1989-2019}, the additive cost is estimated to be \textbf{1.8\% and 2.9\% A AJUSTER} of the export price in air and vessel transport respectively (at the 3-digit classification level).
+On average over \textbf{1974-2019}, the additive cost is estimated to be \textbf{1.8\% and 2.9\% A AJUSTER} of the export price in air and vessel transport respectively (at the 3-digit classification level).
 This is slightly lower than our estimates for the ad-valorem component \textbf{(2.5\% and 3.2\% for air/vessel} respectively on average over the period).
@@
 changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \citet{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
-While Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level.
-Second, as US imports data at the HS-10 level are only available since 1989, we run the analysis over 1989-2019, versus 1974-2004 in \citet{hummels2007}.
-}\textbf{Parler de marge extensive - intensive} \citet{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
+First, we consider a larger period of time, spanning from 1974 (as in \citet{hummels2007}) to 2019 (while he stops in 2004).
+Second, while Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level for the vast majority of the period (1989-2019).
+For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}\textbf{Parler de marge extensive - intensive} \citet{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
 In contrast, we find that trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
@@
 Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \citet{hummels2007}.
-The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2013 (details on the database provided in Appendix \ref{app:data}).
+The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2019 (details on the database provided in Appendix \ref{app:data}).
 We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price.
@@
 This limits the measurement error bias.
-Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC coding system.
+Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level (with a concordance code to the SITC coding system) from \textbf{1989 to 2019}.
+For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.
 %This will be crucial to compute transport costs, as we detail below.
@@
 As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-We estimate international transport costs at the 3-digit classification level, \textbf{exploiting heterogeneity between HS-10 products.} \textbf{La robustesse en 4/10 peut-elle marcher?
+This drives us to estimate international transport costs at the 3-digit classification level, \textbf{exploiting heterogeneity between HS-10 products.} \textbf{La robustesse en 4/10 peut-elle marcher?
 sur les premieres annees pour lesquelles moins de points - je crois?
@@
 To do so, denote by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped.
-In accordance with \citet{hummels2010}, we assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
+In accordance with \citet{hummels2010}, assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
 In this case, the link between the import, export prices and shipping charges is given by:
@@
 
-\textbf{While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2018}), thereby relying on Equation (\ref{eq:alternative}), we follow another route.
-Specifically, we start from Equation (\ref{eq:tt2}) and exploit our dataset to obtain an explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
-From this, we can deduce the price-elasticity of transport costs, i.e.
-the share of additive costs in total transport costs.}\smallskip
+\textbf{While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2018}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
+Based on Equation (\ref{base}), we exploit our dataset to obtain explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
+From this, we can deduce the price-elasticity of transport costs as the share of additive costs in total transport costs, through Equation (\ref{eq:tt2}).}\smallskip
 
@@
 
-\textbf{stop here. Pb sur 1+ ...}
-
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = (1+\tau_{ik}) +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation}
+
+\textbf{stop here}
 

### 2020-09-28 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex

--- 2020-09-23 redaction/submission_JEGeo/revision/referees/answers/Answer_referee1_Lise.tex
@@
      \begin{equation}
-     \beta{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)}-1+\frac{t_{is(k)}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
+     \beta_{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)}-1+\frac{t_{is(k)}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
      \end{equation}
@@
 p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
-\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
+\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
 \end{eqnarray}
@@
      Alternatively, our method rather starts from the definition of transport costs as:
-     $$\ln\left(f_{ik}  \right) = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
+     $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
 from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).

### 2021-01-22 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement mineurs rédaction
la biblio ne compile pas

@@
-\documentclass[a4paper,12pt]{article}
+\documentclass[a4paper,11pt]{article}
 
@@
 and new communication technologies. However, several empirical evidence
-suggest that transport costs remain large and deserve attention.\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \citet{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
-\citet{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+\cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
 of the transport costs burden.
@@
 While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
-as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\citet{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citet{Irrazabal_2015}, or \citet{martin2012}).
+as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\cite{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
 Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs.
@@
 Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2013}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \citet{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
-First, we consider a larger period of time, spanning from 1974 (as in \citet{hummels2007}) to 2019 (while he stops in 2004).
+changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \cite{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
+First, we consider a larger period of time, spanning from 1974 (as in \cite{hummels2007}) to 2019 (while he stops in 2004).
 Second, while Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level for the vast majority of the period (1989-2019).
-For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}\textbf{Parler de marge extensive - intensive} \citet{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
+For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}\textbf{Parler de marge extensive - intensive} \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
 In contrast, we find that trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
@@
 In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
-As shown by \citet{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
+As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
 In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out'').
-This is consistent with the empirical findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
+This is consistent with the empirical findings by \cite{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
 Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e.
-cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using transaction-level trade data from Colombia, \citet{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
-One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
-
-Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
+cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
+Using transaction-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
+One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+
+Closely related to our paper is the work by \cite{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
 Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
-Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
+Our results share in common with \cite{Irrazabal_2015} the important role of the additive component of international trade costs.
 Yet, our paper complements their findings in two main respects.
@@
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of \textbf{2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.}
+While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of \textbf{2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.}
 This comparison of results suggests that between \textbf{a third and a half of the additive trade costs} are attributable to the transport sector (depending on the transport mode).
@@
 Our results have important implications for both international trade and international macroeconomics models.
-Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
+Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \cite{sorensen2014} extends \cite{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
 A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
-This is confirmed by \citet{Irrazabal_2015}} These results suggest that important welfare gains can be achieved by reducing additive trade costs.
+This is confirmed by \cite{Irrazabal_2015}} These results suggest that important welfare gains can be achieved by reducing additive trade costs.
 On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models.
@@
 
-Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \citet{hummels2007}.
+Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
 The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2019 (details on the database provided in Appendix \ref{app:data}).
@@
 By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
-This is quite a standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.
+This is quite a standard and widespread strategy, as emphasized by \cite{anderson_wincoop_jel}.
 
@@
 However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
-According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
-Further, the evidence summarized in \citet{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
-They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
-In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}
+According to \cite{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
+Further, the evidence summarized in \cite{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
+They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \cite{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
+In the same vein, \cite{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}
 
@@
 This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g.
-\citet{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period \textbf{1989-2019}.\footnote{Note that, as mentioned by \citet{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
+\cite{Irrazabal_2015}.
+Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period \textbf{1989-2019}.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
@@
 To do so, denote by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped.
-In accordance with \citet{hummels2010}, assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
+In accordance with \cite{hummels2010}, assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
 In this case, the link between the import, export prices and shipping charges is given by:
@@
 
-\paragraph{Obtaining the estimated equation} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
+\paragraph{Obtaining the estimated equation} We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
 In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $(1+\tau_i)\times (1+\tau_k) -1 \simeq \tau_i + \tau_k$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
@@
 The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
-See \citet{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
 Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit level ($k$).
@@
 To preserve comparability of the results, we keep the same non-linear estimation method in both cases.
-Similarly, as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+Similarly, as \cite{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (a), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (b), for each year and transportation mode.
@@
 With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \citet{anderson_wincoop_jel}, lays in the difference of empirical methodology.
-\citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
+\cite{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
 
@@
 
-Our estimation results on the size of additive transport costs can be compared to \citet{Irrazabal_2015}, the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e.
+Our estimation results on the size of additive transport costs can be compared to \cite{Irrazabal_2015}, the most recent evidence on this question.
+Based on Norwegian trade data for 2004, \cite{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e.
 $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
@@
 Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping and 1.5\% in air transport.
-This may seem surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}.
+This may seem surprisingly low, in contrast to the 6\% obtained by \cite{Irrazabal_2015}.
 This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
-While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
+While the database of \cite{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
 In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
-\citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
+\cite{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
 For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
@@
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+Besides, related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
 In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
@@
 Both Tables suggest that custom duties are a very good predictor of fas prices $\widetilde{p}_{ik}$, with strong significance.
-The F-statistic for significance of the instrument in the first-stage is on average well above 10, the ``rule-of-thumb'' threshold value recommended by \citet{staiger_stock} as a test for weak instrumentation.
+The F-statistic for significance of the instrument in the first-stage is on average well above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation.
 
@@
 In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
-This focus is shared with some of the international trade literature, such as \citet{Lafourcade_Thisse}, \citet{hummels2007} and \citet{Behar_Venables}.
-Specifically, in the same spirit as \citet{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
+This focus is shared with some of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
+Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
 
@@
 As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
-\citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
+\cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
 In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974.
@@
 The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \citet{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
 For the same years, we obtain a total decrease from 6.9\% to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
-Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip
+Overall, our results display trends that are close to those reported by \cite{Hummels_1999}.\smallskip
 
@@
 Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \citet{Hummels_1999} and \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+As emphasized by \cite{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
 Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
@@
 He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
-Further, and as in \citet{hummels_skiba}, \citet{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
+Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
 
@@
 Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
-Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \citet{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
-In doing so, \citet{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
+Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
+In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
 Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
@@
 \paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
-%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \citet{hummels2007}.
+%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
@@
 As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
-This stands in sharp contrast with the assumption made by  \citet{hummels_skiba} or \citet{hummels2007}, which would imply a distribution concentrated on a single point.
+This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
 Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
@@
 
-The last results stand in sharp contrast with \citet{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
+The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
 If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
 This drives us to investigate this difference further.
-As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \citet{hummels2007} in one main dimension.
+As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
 Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
 Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \citet{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
 This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
 
-To establish this point clearly, we replicate the method adopted by \citet{hummels2007} exposed above on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
@@
 composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \citet{hummels2007}.\footnote{Unsurprisingly, this accords with \citet{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
+In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \cite{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
 
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \citet{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
+Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
 Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
@@
 
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \citet{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \cite{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
 In this section, we provide a robustness test to this difference of treatment.\smallskip
 
-When aggregating the trade cost measure over the product/country ($i,k$) dimension, \citet{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
@@
 
-To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \citet{hummels2007} weighting scheme (i.e.
+To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \cite{hummels2007} weighting scheme (i.e.
 taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
@@
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \citet{hummels2007}.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
 This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
@@
 
-\subsection{Comparing with \citet{hummels2007} \label{app:compare_Hummels}}
-
-In this section, we investigate the difference of results with \citet{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
-In order to do so, we start from the empirical specification implemented by \citet{hummels2007} (also making use of Hummels' Stata codes provided on his webpage\footnote{\url{http://www.krannert.purdue.edu/faculty/hummelsd/research/jep-transport-cost-data.php}}), to better identify the precise points of difference between our estimation strategies.
+\subsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
+
+In this section, we investigate the difference of results with \cite{hummels2007} regarding the importance of the composition effects in characterizing the time trends of international transport costs.
+In order to do so, we start from the empirical specification implemented by \cite{hummels2007} (also making use of Hummels' Stata codes provided on his webpage\footnote{\url{http://www.krannert.purdue.edu/faculty/hummelsd/research/jep-transport-cost-data.php}}), to better identify the precise points of difference between our estimation strategies.
 
@@
 Consider now the ``fitted ad-valorem rate''.
-\citet{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
+\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
 The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables}''.
@@
 
-After estimating Equation (\ref{eq:Hummels1}), \citet{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
+After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
 
 What are the main differences with our estimation strategy? Five points may be underlined.
-First, \citet{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
+First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
 However, this is a slight difference, as in both cases it ultimately amounts to taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-Second, in contrast to us, \citet{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
+Second, in contrast to us, \cite{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
 This is not likely to make a substantial difference, however, as these fixed effects are by nature constant over time. Accordingly, they only amount to a scale effect.
@@
 
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \citeth{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
-\citet{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \cite{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
+\cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
@@
 Second, the 1974 weighting scheme makes sense as the starting point of our time trend analysis.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \citet{hummels2007}, as reported in Section \ref{sec:robustness}.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in Section \ref{sec:robustness}.
 
 One last difference remains to be commented, concerning the way the additive component of international transport costs is treated.
-As we show below, it turns out to be key in accounting for the difference of results with \citet{hummels2007}.
-Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \citet{hummels_skiba}.
+As we show below, it turns out to be key in accounting for the difference of results with \cite{hummels2007}.
+Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
 Rewriting Equation (\ref{eq:Hummels1}) in a simpler form as:
@@
 
-As noted by \citet{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
-By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \citet{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
+As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
+By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
 This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (i.e.
 allowing for a varying $\beta$ across time, country partner and product).
-To establish this point clearly, we replicate the method adopted by \citet{hummels2007}, outlined above, on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007}, outlined above, on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.

### 2021-02-19 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte article

@@
 \usepackage{bbold}
+\usepackage{natbib}
+\usepackage{babel,varioref}
 \usepackage{bbm}
@@
 
-\date{September 2020}
+\date{February 2021}
  \maketitle
@@
 \begin{abstract}
-This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from \textbf{1974 to 2019}.
+This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2019.
 First, we document the importance of the per-unit component of transport costs.
 It represents between one-third and one-half of overall transport costs.
-Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in “pure” transport costs over time, with trade composition effects playing a minor role.
+Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs over time, with trade composition effects playing a minor role.
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
@@
 \noindent \emph{JEL classification}: F10, N70, R40 \\
-\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2013, additive costs, trade composition effects
+\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2019, additive costs, trade composition effects
 
@@
 
-Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from \textbf{1989 to 2019}.
+Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
 Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns.
@@
 First, additive transport costs are quantitatively sizable.
-On average over \textbf{1974-2019}, the additive cost is estimated to be \textbf{1.8\% and 2.9\% A AJUSTER} of the export price in air and vessel transport respectively (at the 3-digit classification level).
+On average over 1974-2019, the additive cost is estimated to be \textbf{1.8\% and 2.9\% A AJUSTER} of the export price in air and vessel transport respectively (at the 3-digit classification level).
 This is slightly lower than our estimates for the ad-valorem component \textbf{(2.5\% and 3.2\% for air/vessel} respectively on average over the period).
@@
 Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2013}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \cite{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
+changes over time in the baskets of imported goods and/or the origin countries).} Specifically, trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
+This contrasts with the conclusion obtained by \cite{hummels2007} yet also on US import flows.%\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
 First, we consider a larger period of time, spanning from 1974 (as in \cite{hummels2007}) to 2019 (while he stops in 2004).
 Second, while Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level for the vast majority of the period (1989-2019).
-For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}\textbf{Parler de marge extensive - intensive} \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
-In contrast, we find that trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
+For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}
+\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
 Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
 Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
-\textbf{A FAIRE DANS LES REGRESSIONS, METHODE Hummels mais avec s=3, k=10.
-Peut-on le faire?
-For the argument to be convincing, need to do it on the same years with the same sector-product degrees.}
 
@@
 This comparison of results suggests that between \textbf{a third and a half of the additive trade costs} are attributable to the transport sector (depending on the transport mode).
-Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from \textbf{1989 to 2019}.
+Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2019.
 In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since \textbf{1989}.
+In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed over the last fifty years.
 
@@
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, \textbf{both items being conditional to the transport mode}; in other words, it is the price for one kg of the goods net of transportation costs, \textbf{by transport mode}.
-The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
-\textbf{Cf point 2 refere 1 sur unit prices; avoir quantite plutot que poids, robustesse a faire sur sous-echantillon des produits qui ne passent que par un seul mode de transport).} Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
+While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One might be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity of goods.
+This is due to the fact that in the Census data, information on the quantity of goods is not mode-specific.
+We yet provide a robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.
+\textbf{See Section XXX}.} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
 By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
@@
 This limits the measurement error bias.
-Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level (with a concordance code to the SITC coding system) from \textbf{1989 to 2019}.
-For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.
-%This will be crucial to compute transport costs, as we detail below.
+Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC coding system.
 In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
@@
 \cite{Irrazabal_2015}.
-Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period \textbf{1989-2019}.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
+Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period 1974-2019.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e.
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
@@
 As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-This drives us to estimate international transport costs at the 3-digit classification level, \textbf{exploiting heterogeneity between HS-10 products.} \textbf{La robustesse en 4/10 peut-elle marcher?
-sur les premieres annees pour lesquelles moins de points - je crois?
-Que faire de cette analyse 4d?}
-However, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
+This drives us to estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
+We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i,} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
+Notice that this can be done from 1989 onwards for which HS-10 data are available; \textit{ii,} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
+The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels.
-Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin.
+However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Appendix \ref{app:4digit}.
+\textbf{ONLINE?
+IN TEXT?
+ADD}
+Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin.
 
@@
 
-\paragraph{Estimation strategy} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
-To do so, denote by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped.
-In accordance with \cite{hummels2010}, assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
-In this case, the link between the import, export prices and shipping charges is given by:
-
-\begin{eqnarray}
-&& p = \widetilde{p}+f \notag \\
-\leftrightarrow && p = \widetilde{p}+\widetilde{p}^{1-\beta} X \notag\\
-\leftrightarrow && \frac{p-\widetilde{p}}{\widetilde{p}} = \widetilde{p}^{-\beta} X, \label{eq:tt1}
-\end{eqnarray}
-
-\noindent from which we deduce that $\beta$ is equal to the elasticity of transport costs to the unit price (in absolute value):
+\paragraph{Estimation strategy} Starting from the observed transport costs, we decompose the respective sizes of their two sub-components, ad-valorem and additive transport costs.
+From this, we can deduce the share of the additive component in total transport costs, which also interprets as the elasticity of transport costs with respect to the unit price (in absolute value), as we show below.
+\smallskip
+
+Our starting point relates the import price (at the entry gate) of a given good as equal to the export price (at the export gate) plus freight charges.
+Specifically, denoting by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped, we have: $ p = \widetilde{p} +f$.
+As in \cite{hummels2010}, assume the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
+The link between the import, export prices and shipping charges is hence given by:
+%\begin{eqnarray*}
+$$ p = \widetilde{p}+f ~~ \Leftrightarrow  ~~ p = \widetilde{p}+\widetilde{p}^{1-\beta} X $$
+%\end{eqnarray*}
+Defining transport costs $TC \equiv \frac{p-\widetilde{p}}{\widetilde{p}}$, we then obtain:
+\begin{equation}
+TC \equiv \frac{p-\widetilde{p}}{\widetilde{p}} = \widetilde{p}^{-\beta} X \label{eq:tt1}
+\end{equation}
+From Equation (\ref{eq:tt1}), it is straightforward that $\beta$ is equal to the elasticity of transport costs to the unit price (in absolute value):
 
@@
 
-In the first polar case $\beta=0$, then shipping charges are proportional to the fas price, such that $p=\widetilde{p}(1+X)$: Only ad-valorem costs apply.
+From Equation (\ref{eq:tt1}), we can identify a first polar case $\beta=0$.
+Shipping charges are proportional to the fas price, such that $p=\widetilde{p}(1+X)$: Only ad-valorem costs apply.
 At the other extreme, the case $\beta=1$ corresponds to shipping costs being independent of the unit price, with the cif price the sum of the unit price and the shipping cost: Transport costs are fully additive.
@@
 One can analytically show that $\beta$ is indeed equal to the share of additive costs in total transport costs.
+
 To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as:
@@
 \begin{eqnarray}
-&& p = (1+\tau) \widetilde{p}+ t \label{eq:tt0} \\
-\leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
+&& p = \tau \widetilde{p}+ t \label{eq:tt0} \\
+\Leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
 \end{eqnarray}
 
-\noindent with $0\leq \tau<1$ the ad-valorem cost ($1+\tau$ being frequently refereed to as the iceberg trade cost) and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})).
+\noindent with $\tau \geq 1$ the ad-valorem cost ($\tau =1$ meaning no ad-valorem cost) and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})).
 Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}.
@@
 \begin{eqnarray}
-\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau +\frac{t}{\tilde{p}}} \\
-&=& \frac{-t}{\tau+\widetilde{t}} \label{eq:tt2}
+\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau - 1  +\frac{t}{\tilde{p}}} \\
+&=& \frac{-t}{\tau -1+\widetilde{t}} \label{eq:tt2}
 \end{eqnarray}
@@
 
-\textbf{While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2018}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
-Based on Equation (\ref{base}), we exploit our dataset to obtain explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
-From this, we can deduce the price-elasticity of transport costs as the share of additive costs in total transport costs, through Equation (\ref{eq:tt2}).}\smallskip
-
-More specifically, let us denote $i$ the origin country, and $k$ the product at the 10-digit level, for each flow that is year- and transport-mode specific.
+While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2018}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
+Based on Equation (\ref{eq:base}), we exploit our dataset to obtain explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
+From this, we can deduce the price-elasticity of transport costs as the share of additive costs in total transport costs through Equation (\ref{eq:tt2}).
+
+\paragraph{Obtaining the estimated equation} Let us denote $i$ the origin country, and $k$ the product (at the 5-digit level in our benchmark estimate), for each flow that is year- and transport-mode specific.
 For sake of reading convenience though, we skip the year and transport-mode dimensions in the notations.
@@
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = (1+\tau_{ik}) +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation}
 
-\textbf{stop here}
-
-\paragraph{Obtaining the estimated equation} We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $(1+\tau_i)\times (1+\tau_k) -1 \simeq \tau_i + \tau_k$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i -1)+ (\tau_k-1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 \begin{eqnarray}
-1+\tau_{ik} &=& (1+ \tau)_{i} \times (1+\tau)_{k} \label{eq:ad-valorem}\\
+\tau_{ik} &=&  \tau_{i} \tau_{k} \label{eq:ad-valorem}\\
 t_{ik} &=& t_{i} + t_{k} \label{eq:add}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i}\tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
 \end{equation*}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i}\tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
 \end{equation*}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
 \end{equation}
@@
 See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit level ($k$).
+Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at a more disaggregated product level ($k=5$ in our benchmark estimates).
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
 \end{equation}
@@
   \medskip
+
+\textbf{stop here}
 
@@
 \newpage
-\bibliographystyle{essaien}
+\bibliographystyle{apalike2}
 \bibliography{biblio}

### 2021-02-27 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement rédaction papier

@@
 
-From Equation (\ref{eq:tt1}), we can identify a first polar case $\beta=0$.
+From Equation (\ref{eq:tt1}), we can identify a first polar case where $\beta=0$.
 Shipping charges are proportional to the fas price, such that $p=\widetilde{p}(1+X)$: Only ad-valorem costs apply.
@@
 \begin{eqnarray}
-&& p = \tau \widetilde{p}+ t \label{eq:tt0} \\
+&& p = (1+\tau) \widetilde{p}+ t \label{eq:tt0} \\
 \Leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
@@
 
-\noindent with $\tau \geq 1$ the ad-valorem cost ($\tau =1$ meaning no ad-valorem cost) and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})).
+\noindent with $\tau \geq 0$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})).
 Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}.
@@
 \begin{eqnarray}
-\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau - 1  +\frac{t}{\tilde{p}}} \\
-&=& \frac{-t}{\tau -1+\widetilde{t}} \label{eq:tt2}
+\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau   +\frac{t}{\tilde{p}}} \notag \\
+&=& \frac{-t}{\tau +\widetilde{t}} \label{eq:tt2}
 \end{eqnarray}
@@
 
-\paragraph{Obtaining the estimated equation} Let us denote $i$ the origin country, and $k$ the product (at the 5-digit level in our benchmark estimate), for each flow that is year- and transport-mode specific.
-For sake of reading convenience though, we skip the year and transport-mode dimensions in the notations.
+\paragraph{Obtaining the estimated equation} Let us denote $i$ the origin country, and $k$ the product (at the 5-digit level in our benchmark estimate), for each year-mode specific flow.
+For sake of reading convenience, we skip the year and transport-mode dimensions in the notations.
 Transforming the above equation (\ref{eq:base}) as the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation:
@@
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation}
@@
 We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i -1)+ (\tau_k-1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i+ \tau_k \simeq (1+\tau_i)\times(1+\tau_k) -1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 \begin{eqnarray}
-\tau_{ik} &=&  \tau_{i} \tau_{k} \label{eq:ad-valorem}\\
+1+\tau_{ik} &=& (1+ \tau_{i})(1+ \tau_{k}) \label{eq:ad-valorem}\\
 t_{ik} &=& t_{i} + t_{k} \label{eq:add}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i}\tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =(1+\tau_{i})(1+\tau_{k}) -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
 \end{equation*}
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the import price $p$ cannot be lower than the export price ($p_{ik}\geq\widetilde{p}_{ik}$).
 Taking into account this constraint in the estimate implies that the error term should be always positive.
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i}\tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left((1+\tau_{i})(1+\tau_{k}) -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
 \end{equation*}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left((1+\tau_{i})(1+\tau_{k})+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
 \end{equation}
 
-Equation (\ref{eq:equation0}) is non-linear. It cannot be estimated using standard linear estimators.
+Since Equation (\ref{eq:equation0}) is non-linear, it cannot be estimated using standard linear estimators.
 All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
@@
 Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at a more disaggregated product level ($k=5$ in our benchmark estimates).
-This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
-same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
+This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
 As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
 on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
-This would not solve the problem, however, for three main complementary reasons.
+This would not solve the problem though, for three main complementary reasons.
 First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left((1+\tau_{i})(1+ \tau_{s(k)})-1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
 \end{equation}

### 2021-04-02 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

@@
 
-\paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
-Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-Besides, related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
-In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
-
-It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
-However, the latter can be biased by the aforementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
-This may ultimately harm the estimation of $\widehat{t}/\widetilde{p}$ and consequently, $\widehat{\tau}$.
-
-Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure.
-Put differently, we will replace in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from the following equation:
-
-\begin{equation}
-\ln\left(\widetilde{p}_{ik}\right)= \beta IV_{ik} +\lambda_{is(k)}+ \mu_{ik} \label{eq:first_stage}
-\end{equation}
-
-\noindent where $\lambda_{is(k)}$ represents country-3-digit sector fixed effects, and $IV_{ik} $ the instrumental variable used to purge $\widetilde{p}_{ik}$ from endogeneity issues.
-Following common practice in the literature, we will use as instrument 5-digit product-specific tariff rates, or alternatively, a combination between the latter and lagged fas prices (XXX TO BE DONE XXX).
-Consistently with our overall estimation strategy, Equation (\ref{eq:first_stage}) is estimated on a yearly basis.
-
-Tables \ref{tab: FS_air} and \ref{tab: FS_ves} report summary results of these first stage regressions:
-
-\begin{table}[htbp]\centering
-\caption{Summary Results for First-Stage Regressions - Air\label{tab: FS_air}}
-\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
-\textbf{variable } & \textbf{mean} & \textbf{p25} & \textbf{p50} & \textbf{p75} & \textbf{sd} & \textbf{min} & \textbf{max} \\
-\hline
-   $\beta_{IV}$ &    -0.0858 &    -0.1247 &    -0.0944 &    -0.0362 &     0.0469 &    -0.1741 &    -0.0128 \\
-        $sd_{\beta}$  &     0.0184 &     0.0139 &     0.0197 &     0.0237 &     0.0082 &     0.0048 &     0.0331 \\
-  Student t  &    -5.2255 &    -6.5428 &    -5.2703 &    -3.7943 &     2.4782 &   -11.4557 &    -0.4184 \\
-    F stat  &    33.2938 &    14.4215 &    27.7801 &    42.8707 &    28.1280 &     0.1751 &   131.2323 \\
-  adj $R^2$  &     0.4845 &     0.4712 &     0.4768 &     0.4934 &     0.0227 &     0.4452 &     0.5472 \\
-  $R^2$ within &     0.0025 &     0.0015 &     0.0023 &     0.0036 &     0.0016 &     0.0000 &     0.0064 \\
-\hline
-\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} }}
-\end{tabular}
-\end{table}
-
-\begin{table}[htbp]\centering
-\caption{Summary Results for First-Stage Regressions -Ves\label{tab: FS_ves} }
-\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
-\textbf{variable } & \textbf{      mean} & \textbf{       p25} & \textbf{ p50} & \textbf{p75} & \textbf{sd} & \textbf{       min} & \textbf{       max} \\
-\hline
-   $\beta_{IV}$ &     0.0584 &     0.0180 &     0.0733 &     0.0846 &     0.0349 &    -0.0010 &     0.1147 \\
-       $sd_{\beta}$  &     0.0124 &     0.0108 &     0.0138 &     0.0155 &     0.0047 &     0.0041 &     0.0195 \\
-  Student t  &     4.5161 &     3.3592 &     4.8718 &     5.5721 &     1.9626 &    -0.0781 &     8.9563 \\
-    F stat  &    24.1509 &    11.4029 &    23.7486 &    31.0489 &    17.2397 &     0.0061 &    80.2158 \\
-  adj $R^2$ &     0.5856 &     0.5767 &     0.5878 &     0.5960 &     0.0172 &     0.5364 &     0.6323 \\
-  $R^2$ within &     0.0020 &     0.0006 &     0.0021 &     0.0032 &     0.0015 &     0.0000 &     0.0062 \\
-\hline
-\multicolumn{8}{l}{\footnotesize{\emph{Source:} }}
-\end{tabular}
-\end{table}
-
-Both Tables suggest that custom duties are a very good predictor of fas prices $\widetilde{p}_{ik}$, with strong significance.
-The F-statistic for significance of the instrument in the first-stage is on average well above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation.
-
-XXX on peut aussi ajouter des graphs des beta estimés ici, peut-être éventuellement avec les IC XXX
-
-XXX Reste ici l'épineux problème de la différence de signe entre les deux modes de transport; avant de creuser, voir si on ne résoudrait pas le problème avec l'IV combinant duties et lagged prices XXX
-
-Provided that the instrument used in Equation \ref{eq:first_stage} displays the right statistical properties, we can confidently re-inject the predictions it produces for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation.
-Results are displayed below, as well as a measure of the gap with our main estimates.
-
-XXX We expect the gap to be very small XXX
-
-%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
-%Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
-%In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
-
 \section{Transport costs time trends: the role of the additive component}\label{sec:results_trends}
@@
 
+\subsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
+Besides, related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
+
+It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
+However, the latter can be biased by the aforementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+This may ultimately harm the estimation of $\widehat{t}/\widetilde{p}$ and consequently, $\widehat{\tau}$.
+
+Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure.
+Put differently, we replace in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$  on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
+Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
+In particular, it shows that our main Instrumental Variable displays the right statistical properties, ant that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation.
+Figure \ref{fig:comp_IV_SITC5} below reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+In all cases, these estimates are very similar, when not identical.
+This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
+
+\begin{figure}[htbp]
+\caption{Transport cost time trends: robustness to the weighting scheme}
+\label{fig:comp_IV_SITC5}
+\begin{center}
+\includegraphics[height=8cm]
+{comp_IV_SITC5.pdf}
+\end{center}
+\end{figure}
+
+%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
+%Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
+%In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
+
 \section{Conclusion \label{sec:conclu}}
@@
 
+\subsection{IV estimates: First stage \label{app:first_stage_IV}}
+
+We report here the first-stage procedure and estimates underlying the IV estimates reported in section \ref{sec:results_decomposition}.
+We start by detailing the theoretical intuitions behind the first-stage equation.
+Afterwards, we report the estimates based on 5-digit-product-level data, before moving to the more disaggregate HS10 level for sensitivity checks.
+
+\paragraph{First-stage equation.}
+
+Our main instrumental variable for the fas price $\widetilde{p}_{ikt}$ is the share of tariff duties over the total value imported, at the product line.
+Tariffs may be considered as a (reasonably) exogenous source of variation for fas prices, i.e., uncorrelated with transport costs changes and mostly independent from firm- or sector-level dynamics.
+In this respect, considering the predicted part of the fas price related to tariff duties is likely to solve potential endogeneity biases.
+In other words, tariff shocks should eliminate from the fas price any endogenous component that might me related to transport costs.
+
+Denoting as usual $i$ the origin country, $k$ the product at the 5 digit level, we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
+
+\begin{equation}
+\widetilde{p}_{ikt} = (1+\tau^d_{is(k)t})^\alpha \left(\bar{p}_{ikt}\right)^\beta \label{eq:link_fas_duty}
+\end{equation}
+
+\noindent with $s(k)$ the 3-digit classification as a function of the 5-digit product classification.
+
+We first-differentiate this equation (between $t$ and $t-1$), with $\Delta $ the difference operator ($\Delta\widetilde{p}_{ikt} = \widetilde{p}_{ikt} - \widetilde{p}_{ikt-1}$ and so on):
+\begin{eqnarray*}
+&&\Delta \widetilde{p}_{ikt} = \beta \bar{p}_{ikt-1}^{\beta-1}(1+\tau^d_{is(k)t-1})\Delta \bar{p}_{ikt} + \alpha \bar{p}^\beta_{ikt-1} (1+\tau^d_{ikt-1})^{\alpha-1}\Delta \tau^d_{is(k)t}  \\
+\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \beta \frac{\Delta \bar{p}_{ikt}}{\widetilde{p}_{ikt-1}} \frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} +\alpha \frac{\Delta \tau^d_{is(k)t-1}}{1+\tau_{is(k)t-1}^d}\frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} \end{eqnarray*}
+
+leading to:
+\begin{equation}
+\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} =  \beta \frac{\Delta \bar{p}_{ikt}}{\bar{p}_{ikt-1}} +\alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{ikt-1}^d} \label{eq:firststage_Deltalog}
+\end{equation}
+
+Equation (\ref{eq:firststage_Deltalog}) brings a first-stage regression model as follows:
+
+$$\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \alpha \frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \gamma_{i} +\gamma_{k}+\epsilon_{ik}$$
+
+Or equivalently, taking logs:
+$$\Delta \log \widetilde{p}_{ikt}= \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{k}+\epsilon_{ikt}$$
+
+\noindent with the LHS being the growth rate of the fas price (between $t$ and $t-1$).
+$\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ is the change in the considered tariff rate, the second and third terms are fixed effects designed to capture changes in the ``firm-specific price'' $\bar{p}_{ik}$, $\epsilon_{ikt}$ being the residual.
+Note that, consistently with the second stage, the first stage will be estimated on a year-by-year basis, and will incorporate 3-digit-sector fixed effects.
+This implies to consider the following functional form of the first-stage equation:
+
+\begin{equation}
+\Delta \log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_Deltalog}
+\end{equation}
+
+or equivalently:
+
+\begin{equation}
+\log \widetilde{p}_{ikt} = \log \widetilde{p}_{ikt-1} + \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_Deltalog_bis}
+\end{equation}
+
+Finally, we allow for more flexibility in the empirical set-up, by allowing a variable degree of persistence in the fas price, i.e.
+by relieving the constraint of a unit parameter on $\log \widetilde{p}_{ikt-1}$:
+
+\begin{equation}
+\log \widetilde{p}_{ikt} = \beta\log \widetilde{p}_{ikt-1} + \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\lambda_{s(k)}+\epsilon_{ikt} \label{eq:firststage_final}
+\end{equation}
+
+Note that in such a formulation, $\log \widetilde{p}_{ikt-1}$ appears as a second excluded instrument for $\log \widetilde{p}_{ikt}$.
+The $\alpha$ parameter should fall between $-1$ and $0$, and has a direct theoretical interpretation in terms of ``pricing-to-market'' behavior of firms.
+$\alpha = 0$ corresponds to the case where the considered sector is made of firms not adjusting their fas price to the change in tariff, that would cancel out the influence of the tariff change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition.
+$\alpha = -1$ is the other polar case: the sector here displays full pricing-to-market, because it involves large firms fully compensating the tariff change through their margin (see \citealp{Berman_Martin_Mayer_2012}).
+In-between, $\alpha$ $\in$ $]-1,0[$ corresponds to some variable degree of``pricing-to-market'' in the sector, characterized by firms offsetting partly the impact of the tariff change on the fas price by adjusting the producer price in the opposite direction of the tariff change.
+
+As already mentioned, consistently with the cross-section analysis adopted by now, the first-stage equation is estimated on a yearly basis, and implements 3-digit sector fixed effects ($\lambda_{is(k)}$ ).
+We then replace the observed series of fas prices in the right-hand side of Equation (\ref{eq:model_IetA}) by the predicted fas prices produced by Equation (\ref{eq:firststage_final}).
+We now report results based on our main database at the SITC 5-digit level, before replicating the exercise on HS10 digit level, consistently with one major robustness implemented in the paper.
+
+\paragraph{First-stage estimates based on SITC5 data.}
+
+Here we run Equation (\ref{eq:firststage_final}) on our standard dataset, based on SITC-5 data.
+Consistently, we will use as instrument 5-digit product-specific tariff rates, as well as the one-year-lagged fas price.
+Table \ref{tab: FS_sitc5} reports summary results of these first stage regressions, while Figure \ref{fig: FS_IV_SITC5} displays year estimates of $\beta$ and $\alpha$ parameters, together with 5\% confidence bands:
+
+\begin{table}[htbp]\centering
+\caption{Summary Results for First-Stage Regressions\label{tab: FS_sitc5}}
+\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
+\textbf{variable } & \textbf{Mean} & \textbf{1\textsuperscript{st} Quart.} & \textbf{Median} & \textbf{3\textsuperscript{rd} Quart.} & \textbf{ S.D.} & \textbf{Min} & \textbf{Max} \\
+\hline
+ $\log \widetilde{p}_{ikt-1}$ ($\widehat{\beta}$)&     0.7484 &     0.7400 &     0.7551 &     0.7663 &     0.0248 &     0.6616 &     0.7788 \\
+ % sd_lag~b  &     0.0051 &     0.0040 &     0.0045 &     0.0065 &     0.0015 &     0.0037 &     0.0093 \\
+Student t$_{\widehat{\beta}}$  &   157.07 &   113.24 &   168.82 &   191.15 &    41.23 &    82.88 &   208.23 \\
+$\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)  &    -0.6324 &    -0.8858 &    -0.6079 &    -0.3639 &     0.3962 &    -1.8353 &     0.0534 \\
+ %sd_tar~f  &     0.2216 &     0.2149 &     0.2363 &     0.2599 &     0.0681 &     0.0031 &     0.3399 \\
+ Student t$_{\widehat{\alpha}}$  &    -2.8178 &    -3.7575 &    -2.5741 &    -1.8099 &     1.6808 &    -7.3086 &     0.2280 \\
+ %F_stat  &   1.32e+04 &  6436.5362 &   1.43e+04 &   1.83e+04 &  6174.8488 &  3452.2548 &   2.17e+04 \\
+  F-statistic $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$  &    10.7023 &     3.2757 &     6.6260 &    14.1188 &    11.4529 &     0.0008 &    53.4159 \\
+  %adj_r_~e  &     0.7843 &     0.7809 &     0.7858 &     0.7899 &     0.0105 &     0.7438 &     0.8075 \\
+%  $R^2$ within&     0.5610 &     0.5424 &     0.5738 &     0.5846 &     0.0339 &     0.4373 &     0.5972 \\
+\hline
+%\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} C:\Users\coadministrateur\Dropbox\Papier_Lise_Guillaume\private\revision_JOEG\IV_rev/results/IV_referee1_yearly/FS_parameters_both_yearly_prod5_sect3.dta}}
+\end{tabular}
+\end{table}
+
+\begin{figure}[htbp]
+\caption{Yearly estimates of $\alpha$ and $\beta$ - SITC5}\label{fig: FS_IV_SITC5}
+\begin{center}
+\begin{tabular}{cc}
+%{\small (a) $\beta$} & {\small (b) $\alpha$}\\
+\includegraphics[width=3.2in, height=3.2in,keepaspectratio]{beta_lag_SITC5.pdf}
+& \includegraphics[width=3.2in,height=3.2in,keepaspectratio]{alpha_tariff_SITC5.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+%Staiger and Stock (1997) formalized the definition of “weak instruments” and most researchers seem to have concluded (incorrectly) from that work (orhearsay) that if the F-statistic on the excluded instruments in the first stage is greater than 10, one need worry no further about weak instruments.
+
+%\newpage
+
+Table \ref{tab: FS_sitc5} suggests that custom duties variations are a satisfactory predictor of fas prices $\widetilde{p}_{ik}$, with strong significance.
+The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation.
+In addition, Figure \ref{fig: FS_IV_SITC5} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in 3 or 4 occasions, corresponding to abrupt fluctuations in oil prices.
+This is typically the case in 1975, with $\alpha$=-1.83.
+All in all, these estimates make us confident in the validity of our instrumentation strategy.
+
+\paragraph{First-stage estimates based on HS10 data.}
+
+Here we replicate our first-stage estimations using HS10 data, available between 2002 and 2019.
+In addition to one-year-lagged fas price, we use this time HS10-digit product-specific tariff duties.
+Table \ref{tab: FS_HS10} reports summary results of these first-stage regressions, while Figure \ref{fig: FS_IV_HS10} presents yearly estimates of $\beta$ and $\alpha$ parameters, together with 5\% confidence bands:
+
+\begin{table}[htbp]\centering
+\caption{Summary Results for First-Stage Regressions\label{tab: FS_HS10}}
+\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
+\textbf{variable } & \textbf{Mean} & \textbf{1\textsuperscript{st} Quart.} & \textbf{Median} & \textbf{3\textsuperscript{rd} Quart.} & \textbf{ S.D.} & \textbf{Min} & \textbf{Max} \\
+\hline
+$\log \widetilde{p}_{ikt-1}$ ($\widehat{\beta}$) &     0.7496 &     0.7453 &     0.7504 &     0.7539 &     0.0055 &     0.7385 &     0.7578 \\
+ % sd_lag~t  &     0.0020 &     0.0020 &     0.0020 &     0.0021 &     0.0001 &     0.0019 &     0.0021 \\
+Student t$_{\widehat{\beta}}$  &   369.81 &   361.19 &   366.34 &   376.77 &    12.91 &   352.61 &   395.33 \\
+$\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)  &    -0.3612 &    -0.5636 &    -0.4175 &    -0.1918 &     0.3393 &    -0.9258 &     0.3448 \\
+%  sd_tar~f  &     0.1511 &     0.1451 &     0.1618 &     0.1702 &     0.0325 &     0.0728 &     0.2011 \\
+ Student t$_{\widehat{\alpha}}$   &    -2.2505 &    -3.3518 &    -2.5489 &    -1.3144 &     2.7027 &    -7.9494 &     4.7359 \\
+  %  F_stat  &   6.85e+04 &   6.52e+04 &   6.71e+04 &   7.10e+04 &  4854.7592 &   6.22e+04 &   7.83e+04 \\
+ F-statistic $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ &    11.9399 &     1.7974 &     9.0342 &    14.8620 &    15.1916 &     0.8415 &    63.1932 \\
+%  adj_r_~e  &     0.7697 &     0.7660 &     0.7691 &     0.7725 &     0.0068 &     0.7550 &     0.7826 \\
+%  r_squa~n  &     0.5617 &     0.5591 &     0.5624 &     0.5661 &     0.0057 &     0.5505 &     0.5693 \\
+\hline
+%\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} C:\Users\coadministrateur\Dropbox\Papier_Lise_Guillaume\private\revision_JOEG\IV_rev/results/IV_referee1_yearly/FS_parameters_both_yearly_prod5_sect3.dta}}
+\end{tabular}
+\end{table}
+
+\begin{figure}[htbp]
+\caption{Yearly estimates of $\alpha$ and $\beta$ - HS10}\label{fig: FS_IV_HS10}
+\begin{center}
+\begin{tabular}{cc}
+%{\small (a) $\beta$} & {\small (b) $\alpha$}\\
+\includegraphics[width=3.2in, height=3.2in,keepaspectratio]{beta_lag_HS10.pdf}
+& \includegraphics[width=3.2in,height=3.2in,keepaspectratio]{alpha_tariff_HS10.pdf} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+Results are qualitatively very similar to those reported on the longer-time-span, more-aggregated-SITC5 dataset previously reported.
+The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average close to 12, slightly higher than in our benchmark.
+As previously, Figure \ref{fig: FS_IV_HS10} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a few years which seem again to coincide with significant fluctuations in oil prices - e.g.
+Interestingly, the value of the $\alpha$ parameter is on average -0.36, lower than in our benchmark on SITC5 data.
+At this highly disaggregated level, pricing-to-market behavior seems less strong, indicating that the average firm is smaller than its average SITC5 counterpart.
+Interestingly, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their Table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006.
+In any case, this strengthens our confidence in the validity of our estimation strategy.
+
 \clearpage

### 2021-06-22 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
1eres modifs dans texte (IV)

@@
 Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure.
-Put differently, we replace in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$  on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
+Put differently, we follow earlier literature (see e.g.
+\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$  on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
 Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
-In particular, it shows that our main Instrumental Variable displays the right statistical properties, ant that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation.
+In particular, it shows that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation.
 Figure \ref{fig:comp_IV_SITC5} below reports our benchmark estimates by transport mode, together with their instrumented counterpart.
@@
 In this respect, considering the predicted part of the fas price related to tariff duties is likely to solve potential endogeneity biases.
-In other words, tariff shocks should eliminate from the fas price any endogenous component that might me related to transport costs.
+In other words, tariff shocks should eliminate from the fas price any endogenous component that might be related to transport costs.
+Note that this strategy has been regularly used in the related literature - see for example \citet{Caliendo_Parro_2015} or \citet{Lashkaripour-2017}.
 
@@
 
-Note that in such a formulation, $\log \widetilde{p}_{ikt-1}$ appears as a second excluded instrument for $\log \widetilde{p}_{ikt}$.
+Note that in such a formulation, $\log \widetilde{p}_{ikt-1}$ appears as a second instrument for $\log \widetilde{p}_{ikt}$.
 The $\alpha$ parameter should fall between $-1$ and $0$, and has a direct theoretical interpretation in terms of ``pricing-to-market'' behavior of firms.
@@
 
-Table \ref{tab: FS_sitc5} suggests that custom duties variations are a satisfactory predictor of fas prices $\widetilde{p}_{ik}$, with strong significance.
+Table \ref{tab: FS_sitc5} suggests that custom duties variations are a satisfactory predictor of fas prices $\widetilde{p}_{ik}$, with strong significance and the correct, negative sign.
+The lagged fas price $\log \widetilde{p}_{ikt-1}$ displays also the expected positive sign: with a $\beta$ parameter around 0.75 on average, fas prices exhibit substantial persistence.
 The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation.
-In addition, Figure \ref{fig: FS_IV_SITC5} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in 3 or 4 occasions, corresponding to abrupt fluctuations in oil prices.
+In addition, Figure \ref{fig: FS_IV_SITC5} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a very small number of occasions, corresponding to abrupt fluctuations in oil prices.
 This is typically the case in 1975, with $\alpha$=-1.83.
@@
 The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average close to 12, slightly higher than in our benchmark.
-As previously, Figure \ref{fig: FS_IV_HS10} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a few years which seem again to coincide with significant fluctuations in oil prices - e.g.
+As previously, Figure \ref{fig: FS_IV_HS10} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a few years which seem again to coincide with significant fluctuations in oil prices.
 Interestingly, the value of the $\alpha$ parameter is on average -0.36, lower than in our benchmark on SITC5 data.
 At this highly disaggregated level, pricing-to-market behavior seems less strong, indicating that the average firm is smaller than its average SITC5 counterpart.
-Interestingly, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their Table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006.
+In addition, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their Table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006.
 In any case, this strengthens our confidence in the validity of our estimation strategy.

### 2021-06-24 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
Réponse Critique 2.2

@@
 
-The referee is right in the sense that Census data do report the quantity of goods.
-However, this information is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
-This explains why we cannot calculate the unit price as suggested by the referee.
-\textbf{We make this point clear in Footnote XXX}
-
-\textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
-Et à ce moment là, on peut répondre ``vraiment'' à sa critique?} \textbf{GD20200708 Sauf que du coup, on change le sample.
-Il faudrait vérifier si c’est de beaucoup ou pas}
-
-At a more fundamental level, one might argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
-For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
-\textbf{Do we have evidence, even anecdotal, on this point?
-Is it an argument we want to make, See what Lashkaripour says about this point also.}
-
 \textit{Calculating the unit price as Value/Weight presents the authors
@@
 
-\textbf{GD20200708 Pourquoi n’introduit-il pas $\widehat{p}$ à gauche ?}
+%\textbf{GD20200708 Pourquoi n’introduit-il pas $\widehat{p}$ à gauche ?}
 \textit{There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
@@
 
-\textbf{Je ne comprends pas bien le point (i)}.
+\paragraph{Our answer:}
+\noindent The referee definitely has a point here, in the sense that Census data do report the quantity of goods.
+In addition, referee's remarks echo recent findings by Lashkaripour (\emph{Journal of International Economics}, 2020): based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
+Finally, Lashkaripour (2020) finds accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
+
+We devoted a lot of thought to a potential switch from weight to quantity to get cleaner unit prices. However, the latter appears really uneasy in our context for several, non-trivial reasons:
+
+\begin{enumerate}
+\item The information on quantity is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
+We could of course restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
+
+\item Second, units are not available for many years over our period, especially in the 1970s and the 1980s \textbf{XXX GUILLAUME, TO BE CONFIRMED AND SUPPORTEDXXX}; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
+
+\item Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious \textbf{XXX GUILLAUME: EXAMPLES XXX}.
+At a more fundamental level, one may argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
+For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
+\textbf{XXX GUILLAUME: Do we have evidence, even anecdotal, on this point?}
+
+\end{enumerate}
+
+These motives altogether highlight that it would be too costly and uncertain in our context to switch from weights to quantities to compute unit prices.
+That said, do we have reason to believe this measure of unit prices generate significant deviations from the one relying on quantities?
+
+Lashkaripour (2020) does provide important results related to ours.
+In particular, his result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 20 to 40\% of total costs.
+The two stories are not necessarily incompatible, however.
+First, beyond a longer time span, our database is exhaustive (\textbf{XXX GUILLAUME: TO BE CONFIRMED)}, and encompasses all goods and industries, while Lashkaripour (2020) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
+In this regard, when comparing Tables 1 and 10 in Lashkaripour (2020), one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2021) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+
+%\textbf{We make this point clear in Footnote XXX}
+
+%\textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
+Et à ce moment là, on peut répondre ``vraiment'' à sa critique?} \textbf{GD20200708 Sauf que du coup, on change le sample.
+Il faudrait vérifier si c’est de beaucoup ou pas}
+
+%\textbf{Je ne comprends pas bien le point (i)}.
 We thank the referee for pointing out this potential source of bias in our estimates.
@@
 }
-\textbf{GD20200708 Moi non plus je ne comprends pas très bien}
+%\textbf{GD20200708 Moi non plus je ne comprends pas très bien}
 

### 2021-06-25 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
Modids sur les parties liées aux IV

@@
 However, we also want to keep track of this alternative method as further robustness check.
-Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.
+Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.\textbf{ XXX CHECK THIS SECTION EXISTS IN THE ONLINE APPENDIX XXX}
 
@@
 
-We followed the referee's advice by providing some IV-estimates of the fas price to provide a clean assessment of the size of the potential bias in our original methodology.
+Indeed, this is a very important point.
+The referee states that, based on theoretical insights by Melitz (\emph{Econometrica}, 2003) or Baldwin and Harrigan (\emph{American Economic Review}, 2011), $1/\widetilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+
+We obviously do not question this conceptual issue.
+However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF (p) and the FAS ($\widetilde{p}_{ik}$) prices.
+Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out.
+That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
+
+Following the referee's advice, we decided therefore to provide a full set of IV estimates to provide a clean assessment of the size of the potential bias.
+Section 4.3 in the revised version of the paper provides an overview of the results, while section B.4 in Appendix B provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
+We follow earlier literature (see e.g.
+Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
+First-stage estimates reported in section B.4 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation.
+Figure 9 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+In all cases, these estimates are very similar, when not identical.
+This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
+
+Not this check is performed both on our main dataset (SITC 5 digit), and also at the HS 10 level, to handle simultaneously the referee's concern about aggregation issues.
+In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very, very similar to their non-IV, least squares counterparts.
+%We followed the referee's advice by providing some IV-estimates of the fas price to provide a clean assessment of the size of the potential bias in our original methodology.
 The IV-strategy is implemented in two steps.
 
-\begin{itemize}
-\item First-stage equation: We regress the fas price $\widetilde{p}_{ik}$ with $k$ = 5-digit product classification and $i$ the origin country, on tax duty and lagged fas price according to Equation (XXX).
+%\begin{itemize}
+%\item First-stage equation: We regress the fas price $\widetilde{p}_{ik}$ with $k$ = 5-digit product classification and $i$ the origin country, on tax duty and lagged fas price according to Equation (XXX).
 As made clear in the appendix of the revised version (\textbf{Section XXX - take elements of topo-first-stage there}), we run the regression in log level on a yearly basis and by transport mode, consistently with our second-stage empirical strategy.
 
-\item The second-stage: Explain the difficulty since fas prices are on both sides of the equation; non-trivial way to do it.
+%\item The second-stage: Explain the difficulty since fas prices are on both sides of the equation; non-trivial way to do it.
 We have decided to not instrument the fas price that intervenes on the LHS.
@@
 
-\end{itemize}
-
-The results are ... DEVELOP.
+%\end{itemize}
 

### 2021-06-28 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
Answer Editor + minor changes answer R1

@@
 significant changes to the paper that we hope address your concerns. We first give you an overview
-of the revision (Section ) before answering in detail each of your comments (Section ). Number of sections, pages and
-equations refer to the revised version.
+of the revision (Section\ref{sec:main_changes}) before answering in detail each of your comments (Section\ref{sec:detailed_answers}).
+Number of sections, pages and equations mentioned in the text refer to the revised version.
 
@@
 Though, we do not know why that should affect differently additive and multiplicative trade costs.
-Following the referee's advice, we decided therefore to provide some IV-estimates to assess the size of the potential bias.
-\textbf{This is made in Section XXX of the paper.}
+Following the referee's advice, we decided therefore to provide detailed IV-estimates to assess the size of the potential bias.
+The latter are commented in section 4.3, and detailed in appendix B.4.
 

### 2021-06-28 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
Réécriture réponses à R1

@@
 
-\date{September 2020}
+\date{July 2021}
  \maketitle
@@
 
-We would like to thank you for your insightful comments. They led us to introduce some
+We would like to thank you for your various insightful comments. They led us to introduce some
 significant changes to the paper that we hope address your concerns. We first give you an overview
@@
 \section{Detailed answers \label{sec:detailed_answers}}
+
 \subsection{Critique 1: Empirical Strategy}
@@
 
-The three points raised by the referee indeed deserve careful consideration.
+\textbf{Our answer:}\\
 
-\paragraph{(i) On the assumptions underlining our empirical approach}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla in the Review of Economics and Statistics (2015) to estimate the share of additive costs in a firm-level context.
-It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
-
-Further, we already provide a robustness check for this separative assumption, in Section XXX of the paper (\textbf{Table 2, submitted version}).
-We check the robustness of our results to the separability assumption by re-running the estimation without the separability assumption.
-This comes at the cost of a substantial increase in the number of fixed effects to include in the regression.
-Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted version} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
-This is not tractable computationally.
-For this reason, we have decided to run the robustness check on a reduced sample.
-If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013, vs around 29,000 on the complete sample, see Table 1 ), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
-As we conclude at the end of Section XXX , whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
-Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
-
-\paragraph{(ii) On the use of non-linear least squares (NLS)}
-As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
-However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
-This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
-$\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
-When run relaxing these restrictions, standard linear, least squares estimates often deliver negative trade costs, which is meaningless.
-In this respect, implementing the referee's method (see below) does not suppress the requirement of resorting on non-linear estimates (and the computational burden it induces, in particular in terms of time).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version \textbf{[see page XX : XX insert quote XX].} \smallskip
-
-Of course, this does not prevent us for taking full account of your suggestion to adopt an alternative estimation method (summarized in Equation (\ref{eq:estimation_ref1}).
-It turns out that this method faces some important limitations, that justify maintaining our initial estimation strategy as the baseline in the building block of the paper.
-\textbf{The referee's method is presented in the online appendix, and we refer to it in Footnote XXX of the paper}.
-We develop the analysis we held in \textbf{Section XXX of the letter}.
-
-\paragraph{(iii) Endogeneity problem}
-
-Indeed, this is a very important point.
-The referee states that, based on theoretical insights by Melitz (Econometrica, 2003) or Baldwin and Harrigan (AER, 2011), $1/\tilde{p}_{ik}$ is correlated with residuals $\epsilon_{ik}$.
-In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
-We obviously do not question this conceptual issue.
-However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF ($p$) and the FAS ($\tilde{p}$) prices.
-Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out.
-That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
-Though, we do not know why that should affect differently additive and multiplicative trade costs.
-Following the referee's advice, we decided therefore to provide detailed IV-estimates to assess the size of the potential bias.
-The latter are commented in section 4.3, and detailed in appendix B.4.
-
-\paragraph{Strategy of answer} To clarify the discussion, we decided to answer concerns (i) and (iii) each in turn.
-As regards to point (i), we run the referee's proposed estimation method without relying on instrumental variables; this way, the results are comparable with the benchmark results, the only change being the estimation strategy (Section \ref{subsec:functional_form} of the letter).
+%\paragraph{Strategy of answer} To clarify the discussion, we decided to answer concerns (i) and (iii) each in turn.
+As regards to point (ii), we run the referee's proposed estimation method without relying on instrumental variables; this way, the results are comparable with the benchmark results, the only change being the estimation strategy (Section \ref{subsec:functional_form} of the letter).
 As noted above, this drove us to maintain our estimation method as benchmark.
@@
 
-\subsubsection{Exploring Referee's alternative functional form \label{subsec:functional_form}}
+The three points raised by the referee indeed deserve careful consideration. We provide separate answers to each of them below, starting with points (i) and (iii), and finishing with point (ii).
 
-\textit{A more natural approach is what the authors, at some point, refer to as the
-Hummel's Methodology. That is, one can alternatively estimate the share of the
-additive component as:}
+\subsubsection{On the justifications for our empirical approach}
+
+\paragraph{About the key assumption of separability}
+Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla in the \emph{Review of Economics and Statistics} (2015) to estimate the share of additive costs in a firm-level context.
+It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
+
+Further, we provide a robustness check for this separative assumption, in Section 4.1 of the paper (\textbf{Table 2, previously submitted version and revised version XXX CHECK AT THE END XXX}).
+We check the robustness of our results by re-running the estimation without the separability assumption.
+This comes at the cost of a substantial increase in the number of fixed effects to include in the regression.
+Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted version and revised version  XXX CHECK AT THE ENDXXX} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
+This is computationally not tractable.
+For this reason, we have decided to run the robustness check on a reduced sample.
+If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+As we conclude at the end of Section 4.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not - see in particular Figures 6 and 7 on pp.
+25-26.
+Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
+
+\paragraph{On the use of non-linear least squares (NLS)}
+
+The referee writes \textit{A more natural approach is what the authors, at some point, refer to as the Hummel's Methodology.
+That is, one can alternatively estimate the share of the additive component as:}
 
@@
 across all districts and all 10-digit products pertaining to the same 5-digit category.}
+
+\textbf{Our answer:}
+As highlighted by the referee, our estimated equation requires non-linear estimation methods, such as Non-Linear Least Squares.
+However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
+This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
+$\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
+Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates.
+In this respect, implementing the referee's method (see below) does not suppress the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces).
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version - see footnote 16XXX page 8: ``One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.`on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
+This would not solve the problem though, for three main complementary reasons.
+First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
+Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
+Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
+Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$)''
+
+Of course, this does not prevent us for taking full account of your suggestion to adopt an alternative estimation method (summarized in Equation (\ref{eq:estimation_ref1}).
+It turns out that this method does not go without a few important limitations, that justify maintaining our initial estimation strategy as the baseline in the building block of the paper.\textbf{The referee's method is presented in the online appendix XXX WHERE???
+XXX, and we refer to it in Footnote XXX of the paper}.
+We develop the analysis we held below.
+
+\paragraph{Exploring Referee's alternative functional form \label{subsec:functional_form}}
 
@@
 \begin{itemize}
-\item \textit{The first advantage of this so-called Hummel's Approach is that the above
+\item the referee writes \textit{The first advantage of this so-called Hummel's Approach is that the above
 regression can be estimated separately for various country-industry pairs, without
-imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} Our answer is twofold.
+imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} \\
+
+\textbf{Our answer:}
 
@@
 
-\item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}  We thank the referee for this valuable suggestion.
+\item the referee writes \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.} \\
+
+     \textbf{Our answer:}
+We thank the referee for this valuable suggestion.
 It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.
 
-\item \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.} We agree with the referee that the comparison with the literature (Hummels (2007) in particular) was not straightforward for the reader in the submitted version.
+\item the referee writes \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.}\\
+     \textbf{Our answer:}
+     We agree with the referee that the comparison with the literature (Hummels (2007) in particular) was not straightforward for the reader in the submitted version.
 We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
@@
 
-\textbf{stop here}
+%\textbf{stop here}
 
-\subsubsection{Potential endogeneity problem}
+\subsubsection{Concers about endogeneity}
 
@@
 In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very, very similar to their non-IV, least squares counterparts.
-%We followed the referee's advice by providing some IV-estimates of the fas price to provide a clean assessment of the size of the potential bias in our original methodology.
-The IV-strategy is implemented in two steps.
-
-%\begin{itemize}
-%\item First-stage equation: We regress the fas price $\widetilde{p}_{ik}$ with $k$ = 5-digit product classification and $i$ the origin country, on tax duty and lagged fas price according to Equation (XXX).
-As made clear in the appendix of the revised version (\textbf{Section XXX - take elements of topo-first-stage there}), we run the regression in log level on a yearly basis and by transport mode, consistently with our second-stage empirical strategy.
-
-%\item The second-stage: Explain the difficulty since fas prices are on both sides of the equation; non-trivial way to do it.
-We have decided to not instrument the fas price that intervenes on the LHS.
-Develop.
-
-%\end{itemize}
 
@@
 
-\subsubsection{To sum up: What remains to be done}
+% XXX JH: UNLESS I AM MISTAKING, THE SHOPPING LIST BELOW HAS BEEN IMPLEMENTED 
+%\subsubsection{To sum up: What remains to be done}
 
-\begin{enumerate}
-\item\label{bp:3} Re-run our benchmark method with $s=3$ and $k=10$ over available years and compare the results ($t,\tau,\beta$ with our benchmark ones ($s=3,k=5$).
+%\begin{enumerate}
+%\item\label{bp:3} Re-run our benchmark method with $s=3$ and $k=10$ over available years and compare the results ($t,\tau,\beta$ with our benchmark ones ($s=3,k=5$).
 Hopefully does not change much so that we can keep up the period 1974-1988?
@@
 Answer to the critique about the aggregation problem.
-\item\label{bp:1} After running the referee's 1 method with $s=3$, $k=$10, show tables or figures comparing 1) the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy; 2) comparison of the $\beta$ between the two methods; 3) comparison of the precision of the $\beta$ estimation.
+%\item\label{bp:1} After running the referee's 1 method with $s=3$, $k=$10, show tables or figures comparing 1) the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy; 2) comparison of the $\beta$ between the two methods; 3) comparison of the precision of the $\beta$ estimation.
 Answer to the critique about the estimation strategy.
-\item\label{bp:2} IV-First-stage equation on fas price at the $s=3$, $k=10$-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version.
+%\item\label{bp:2} IV-First-stage equation on fas price at the $s=3$, $k=10$-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version.
 Answer to the critique about the endogeneity bias.
-\end{enumerate}
+%\end{enumerate}
 
@@
 
+% MEMO: DO NOT FORGET TO REPORT THE ARGUMENTATION BELOW IN THE MAIN TEXT XXX 
 \paragraph{Our answer:}

### 2021-06-30 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
Modifications Jérôme

@@
 However, we also want to keep track of this alternative method as further robustness check.
-Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.\textbf{ XXX CHECK THIS SECTION EXISTS IN THE ONLINE APPENDIX XXX}
+Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.\textbf{ XXX TO BE DONE XXX CHECK THIS SECTION EXISTS IN THE ONLINE APPENDIX XXX}
 
@@
 
-\subsubsection{Concers about endogeneity}
+\subsubsection{Concerns about endogeneity}
 
@@
 This however deprives us from being able to exploit the period 1974-1988 at the aggregation level.
-\textbf{Que faire pour la periode 1974-1989}.
+\textbf{Que faire pour la periode 1974-1989 XXX TO BE DISCUSSED XXX}.
 Furthermore, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
-We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper \textbf{see XXXX}
+We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper \textbf{see XXXX TO BE DONE ???
+XXX}
 
-% XXX JH: UNLESS I AM MISTAKING, THE SHOPPING LIST BELOW HAS BEEN IMPLEMENTED 
+% XXX JH: UNLESS I AM MISTAKING, THE SHOPPING LIST BELOW HAS BEEN IMPLEMENTED XXX TO BE CONFIRMED
 %\subsubsection{To sum up: What remains to be done}
@@
 
+% XXX JH: What do we do with the section below?
+
 \section{Clarifying some technical points \label{app:technical_points}}
@@
 
+\end{document}
+
 \section{More details on the instrumentation strategy}
@@
 \end{eqnarray*}
-
-\end{document}

### 2021-07-05 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Reorg

--- 2021-06-22 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to Holger G\"{o}rg, S\'{e}bastien Jean and Vincent Vicard for their insightful comments.
-The paper has greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the universities of California-Irvine, Paris-Dauphine, Paris-I and Nancy and the Kiel Institute for the World Economy.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We thank the referees, as well as the editor, William Kerr, for insightful comments, which helped to significantly improve the paper.
+We are grateful to Holger G\"{o}rg, S\'{e}bastien Jean and Vincent Vicard for their critics and suggestions on earlier drafts.
+The paper has also greatly benefited from remarks from the participants at ETSG 17\textsuperscript{th} (Helsinki), DEGIT XXII (Paris), seminars at the universities of California-Irvine, Paris-Dauphine, Paris-I and Nancy and the Kiel Institute for the World Economy.
 Any remaining errors are ours.
-This paper was partly funded by a university grant from Lille-I University.
+This paper was partly funded by a university grant from the University of Lille.
 It features an online appendix containing additional results, available on the authors' websites.}}
@@
 
-\date{February 2021}
+\date{July 2021}
  \maketitle
@@
 One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
-The quantitative results we provide in the paper allow to make one step in filling that gap.
+The quantitative results we provide in the paper allow making one step in filling that gap.
 \textbf{To develop}.

### 2021-07-09 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Modids sur les parties liées aux IV
rangements et modif mineures (OA)

@@
 
-Trade costs are central to international economic analysis. In particular, they are
-known to be a major obstacle to international economic integration and international trade
-flows. Defined as the costs associated with the exchange of goods across
-national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially
-decreased with technological advance in transportation, infrastructure development
+Trade costs are central to international economic analysis.
+In particular, they are known to be a major obstacle to international economic integration and international trade flows.
+Defined as the costs associated with the exchange of goods across
+national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}.
+Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development
 and new communication technologies. However, several empirical evidence
@@
 If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
-of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip
+of agglomeration, the profitability of outsourcing or the gains from trade liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip
 
@@
 Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs.
-Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1989-2019 and by transport mode (air or maritime transport).\smallskip
+Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip
 

### 2021-07-09 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex

--- 2021-06-30 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
@@
 
-\item Based on your comment 1.1 (\textit{The aggregation problem}), the benchmark regression (non-linear least squares) is now completed by a two-stage procedure, where the fas price is instrumented by tax duties and the lag price \textbf{(See Section XXX).}
+\item Based on your comment 1.1 (\textit{The endogeneity problem}), the benchmark regression (non-linear least squares) is now completed by a two-stage procedure, where the fas price is instrumented by tax duties and its own lag - see section 4.3 and appendix B.4 in the paper, and section 2.1.2 of this letter.
 
@@
 Second, we emphasize the drawbacks induced by this method.
+XXX HERE XXX
 

### 2021-07-09 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_rev1.tex

--- 2018-12-17 redaction/onlineA.tex
@@
 \urlstyle{rm}
-\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
-Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author.
-Universit\'{e} Paris-Dauphine, PSL University, LEDa, 75016 PARIS, FRANCE ; email: \url{lise.patureau@dauphine.fr}}}
+
 %\date{\vspace{-5ex}}

### 2021-07-21 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex

\documentclass[12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.50.0.2960}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{BibliographyScheme=Manual}
%TCIDATA{Created=Monday, February 14, 2011 11:40:51}
%TCIDATA{LastRevised=Thursday, February 17, 2011 16:51:46}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{<META NAME="DocumentShell" CONTENT="Standard LaTeX\Blank - Standard LaTeX Article">}
%TCIDATA{CSTFile=40 LaTeX article.cst}

\usepackage{graphicx}
\usepackage{setspace}
%\usepackage{hyperref}
\usepackage{geometry}
\usepackage{blindtext}
\geometry{hmargin=3cm,vmargin=2.5cm}
\usepackage{lscape}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\ \rule{0.5em}{0.5em}}

\setlength{\parindent}{0pt}

\newcommand*\sepline{%
  \begin{center}
    \rule[1ex]{.5\textwidth}{.5pt}
  \end{center}}
\newcommand*\sepstars{%
  \begin{center}
    $\star\star\star$
  \end{center}}

\begin{document}

\noindent \textbf{Manuscript ID JOEG-2019-027, ``International Transport costs:\\New Findings from modeling additive costs'': revised revision}

\bigskip
\bigskip

\hfill Paris, \today

\bigskip
\noindent Dear Pr. Kerr, dear Bill, \\
\\
\noindent \noindent We would like to thank you for the opportunity to resubmit our paper, and for your patience in waiting the outcome. We are very grateful for the deadline extensions you kindly granted us. Obviously, the health crisis and the various consecutive lockdowns made the access to our offices difficult in several occurrences.

However, the major source of delays lies in our attempts to obtain additional data to accommodate a very important concern by Referee \#1 (see below). We ordered this new data from the Census Bureau at the end of December 2019. However, some of the years we requested (2000 and 2001, as well as years before 1998) are not on their servers: they are back at their office which they have not been able to access since last spring apparently, due to the current health situation. Since then, the situation has never evolved, and we are still currently waiting for these missing years. We can of course make available all our e-mail exchanges with Jeremy Sanchez, our contact at the Census Bureau. Until very recently, Mr. Sanchez never gave us any clear “yes or no” answer to our requests concerning a possible return in his DC office (closed because of the CoViD 19 restrictions), which would have allowed the delivery of the data. We can of course make available all our e-mail exchanges with Jeremy Sanchez. In his last e-mail, he confirmed that access to the additional data was impossible.

Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this e-mail) based on a HS10 dataset over the 2002-2019 period. These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.

and to thank the referees for the comments which have helped us revise it. These suggestions have led to substantial improvements. We enclose detailed answers to the referees. Here we briefly list the major changes brought to the paper. We hope that you will be satisfied by the outcome.

\begin{enumerate}

\item Functional form/use of a non-linear estimator
\item Endogeneity issues/IV estimates
\item Aggregation issues
\item Calculation of unit prices
\item Big picture/usefulness of the exercise
\item XXX
\item XXX

\end{enumerate}

Altogether, we believe that the paper is much clearer and easier to read. The robustness exercises included in the main text focus on the elements that were the most problematic to the referees (XXX). We hope that you will be pleased by the outcome.

\bigskip

Yours sincerely,

\bigskip

\hfill Jérôme Héricourt

\hfill Professor of Economics

\hfill University of Lille (LEM-CNRS) and CEPII

\bigskip
\bigskip
\bigskip
\sepstars

\end{document}

### 2021-07-21 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Première insertion de table réussie dans latex

--- 2018-12-17 redaction/onlineA.tex
@@
     }
+\end{table}%
+
+\begin{table}[htbp]
+	\centering
+	\footnotesize{
+	\caption{Estimation results of the three models (Air, 3-digit level)}
+	\label{tab:3models_estimation_results_air}%
+	\input{TableA1_air.tex}
 \end{table}%

### 2021-07-21 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
La table progresse
Table A1 et A2 qui marchent
 (mais il manque des modèles)
Table B appendix

@@
 	\footnotesize{
-	\caption{Estimation results of the three models (Air, 3-digit level)}
+	\caption{Estimation results of the three models (Air, products at 5-digit level, sectors at 3-digit level)}
 	\label{tab:3models_estimation_results_air}%
 	\input{TableA1_air.tex}
+   \begin{tablenotes}
+	\tiny
+	\item Statistics are weighted by value
+	\item \textbf{Model (B): With additive and ad-valorem transport costs}
+\end{tablenotes}
+
+}	
 \end{table}%
@@
    \end{tabular}%
+   }
+\end{table}%
+
+\begin{table}[htbp]
+	\centering
+	\footnotesize{
+		\caption{Estimation results of the three models (Vessel, products at 5-digit level, sectors at 3-digit level)}
+		\label{tab:3models_estimation_results_ves}%
+		\input{TableA1_ves.tex}
+		\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\end{tablenotes}
    }
@@
 \end{tabular}}%
+	\end{center}
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Vessel: Transport costs estimates, all years, products at 5-digit level, sectors at 3-digit level}
+	\begin{center}
+
+		\input{TableB1974_1987_air.tex}
+
+	\end{center}
+\label{tab_oa:result_air_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB1988_2001_air.tex}
+
+	\end{center}
+	\label{tab_oa:result_air_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB2002_2015_air.tex}
+
+	\end{center}
+	\label{tab_oa:result_air_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB2016_2019_air.tex}
 
@@
 \begin{landscape}
+
 \begin{table}[htbp]
@@
 	\end{center}
-\label{tab_oa:result_vessel_ally3}%
-\end{table}
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Air: Transport costs estimates, all years, products at 5-digit level, sectors at 3-digit level}
+	\begin{center}
+
+		\input{TableB1974_1987_ves.tex}
+
+	\end{center}
+	\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB1988_2001_ves.tex}
+
+	\end{center}
+	\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+\caption{Continued}
+\begin{center}
+
+	\input{TableB2002_2015_ves.tex}
+
+\end{center}
+\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+\caption{Continued}
+\begin{center}
+
+	\input{TableB2016_2019_ves.tex}
+
+\end{center}
+\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
 \end{landscape}

### 2021-07-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
commit de Jérôme
rédaction draft answer big picture

--- 2021-07-09 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex
@@
 \paragraph{About the key assumption of separability}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by Irarrazabal, Moxnes and Opromolla in the \emph{Review of Economics and Statistics} (2015) to estimate the share of additive costs in a firm-level context.
+Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by \citet{Irrazabal_2015} to estimate the share of additive costs in a firm-level context.
 It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
@@
 For this reason, we have decided to run the robustness check on a reduced sample.
-If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013 XXX 2019 ???
+XXX, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
 As we conclude at the end of Section 4.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not - see in particular Figures 6 and 7 on pp.
@@
      \textbf{Our answer:}
-     We agree with the referee that the comparison with the literature (Hummels (2007) in particular) was not straightforward for the reader in the submitted version.
+     We agree with the referee that the comparison with the literature (\citealp{hummels2007}) in particular) was not straightforward for the reader in the submitted version.
 We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
@@
 Indeed, this is a very important point.
-The referee states that, based on theoretical insights by Melitz (\emph{Econometrica}, 2003) or Baldwin and Harrigan (\emph{American Economic Review}, 2011), $1/\widetilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
+The referee states that, based on theoretical insights by \citet{melitz} or \citet{baldwin_harrigan}, $1/\widetilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
 In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
@@
 
-Lashkaripour (2020) does provide important results related to ours.
+\citet{Lashkaripour_JIE2020} does provide important results related to ours.
 In particular, his result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 20 to 40\% of total costs.
 The two stories are not necessarily incompatible, however.
-First, beyond a longer time span, our database is exhaustive (\textbf{XXX GUILLAUME: TO BE CONFIRMED)}, and encompasses all goods and industries, while Lashkaripour (2020) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
-In this regard, when comparing Tables 1 and 10 in Lashkaripour (2020), one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2021) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+First, beyond a longer time span, our database is exhaustive (\textbf{XXX GUILLAUME: TO BE CONFIRMED)}, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
+In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2021) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
 
@@
 
-\textbf{Comments about that}
+\paragraph{Our answer:}
+\noindent We devoted a lot of thought to this point, which directly echoed a similar concern by the other referee.
+Taking stock of your suggestion, we decided to refocus the paper on the welfare implications of our results.
+In this regard, the section 3.4 from the initial version has been moved to Appendix XXX TO BE DONE XXX of the paper, and replaced by a new section 3.4 devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
+\citet{melitz}, featuring additive trade costs.
+The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
+However, \citet{sorensen2014} exclusively performs a theoretical analysis, without any quantitative exercise.
+In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the presence of additive costs, but the latter is based on a calibration for transport costs restricted to 2004, based on the case of Nowrway.
+In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on ``true'' values for the latter - remember that our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the (median) export price.
 
-\paragraph{Piste 1} Consider his/her first suggestion.
+More precisely, we use some of the estimates underlying results reported in sections 2 and 3 concering multiplicative and additive costs, more precisely those for the years 1974 and 2019.
+Based on the latter, we implement several comparative static exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
+To that end, we play with the share of exporting firms in the US, the latter being found in XXX: it reaches 35\% in 2019, versus XX\% in 1974.
+
+Table XX on p.XX in section 3.4XX in the revised version reports the results of these various comparative static exercises for air (Panel A) and Vessel (Panel B) transport modes.
+For each panel, the first two rows of the Table report the figures for total transport costs in 1974 and 2019, and the subsequent rows show the difference between the two years, how it is split (or not) between multiplicative and additive costs and the consequences for the welfare.
+Column (1) reports the results for when costs are assumed to be exclusively multiplicative, while column (2) combines this feature with a increase in the share of exporting firms.
+Column (3) reports results based on the model including additive costs, and column (4) again adds to the picture an increase in the share of exporting firms.
+
+Qualitatively, the conclusions are identical for both transport modes, with two major insights.
+First, for a given decrease in transport costs, the comparison in columns (1) and (3) shows welfare gains around 50\% higher when the underlying framework allows for additive costs - which, in practice, concentrate the major part of transport costs decrease over the period.
+Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing costs, with again a an additional premium coming from the decrease of additive costs.
+Quantitatively, welfare gains are more important for the Vessel mode, which is a natural consequence of a more substantial decrease in transport costs.
+More importantly, we see that the amplification effect coming from the decrease in $f_{x}$ is slightly lower for the specification including additive costs  in Vessel than in Air: a bit more than 40\% additional welfare gains for the former, versus 47\% gains for the latter.
+This reflects a slightly more balanced contribution from each component (additive and multiplicative) in the decrease of transport costs in Vessel compared to Air.
+This also emphasizes that the amplification effect of the decrease in sunk costs affects disproportionately the variations in additive costs.
+
+Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45 years period, and to highlight the respective part of each component in the determination of these gains.
+Note also that not only the inclusion of additive costs in the underlying framework generates big welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
+
+%Gains de bien-être sont plus forts lorsqu’on bouge les coûts additifs. Voir >Table_compWelfare2.xls.
+%Ex 1 : Comparer C (que des add valorem) et E (additifs et ad valorem) : on baisse les coûts de transports ; sheet 1 = air, sheet 2 = Vessel.
+%Ex 2 : si en plus on ajuste fx, cela amplifie le diagnostic. Voir colonnes D et F : la baisse des couts fixes joue plus dans un monde avec ad%ditifs que sans additifs.
+%Voir « C:\Users\coadministrateur\Dropbox\Papier_Lise_Guillaume\trade_cost\model », et les tableaux afférents.
+% Big picture: quantitatively, we are more exhaustive/more complete.
+We find big welfare differences with additive costs, even if we focus only on transport costs.--> Our results are probably a lower bound.
+
+%XXX is chosen to match the share of firms exporting in Norway (38 percent, Moxnes, 2010), while
+
+%IMO : The data cover all Norwegian non-oil exporters in 2004 and originate from customs declarations.
+
+%\textbf{Comments about that}
+
+%\paragraph{Piste 1} Consider his/her first suggestion.
 What is the idea?
@@
 
-What to do with this?
+%What to do with this?
 When composition effects do matter (ocean freight), do they favor low-elasticity industries?
@@
 
-\textbf{GD20200708 Hum...
+%\textbf{GD20200708 Hum...
 Je crois que je comprends le point dur referee et je suis d’accord avec lui.
@@
 
-\paragraph{Piste 2} What is at stake?
+%\paragraph{Piste 2} What is at stake?
 Show that how $\beta$ = the relative share of additive costs has evolved over time.
@@
 
-\paragraph{Piste 3} \textbf{GD20200708 On sait que les additive costs sont plus distorsifs que les multiplicative costs, puisqu’ils modifient les prix relatifs.
+%\paragraph{Piste 3} \textbf{GD20200708 On sait que les additive costs sont plus distorsifs que les multiplicative costs, puisqu’ils modifient les prix relatifs.
 Donc le fait qu’ils sont une partie croissante des costs signifie qu’on sur-estime l’augmentation des gains de welfare liés à la baisse des coûts de transport}
+
+\newpage
+\bibliographystyle{apalike2}
+\bibliography{biblio}
+
+\newpage
 

### 2021-07-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
test debuggage

@@
 If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
-of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip
-
-Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
+of agglomeration, the profitability of outsourcing or the gains from trade liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip
+
+Our paper contributes XXX test XXX to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
 Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns.
@@
 
+\begin{table}
+	\input{Table1.tex}
+\end{table}
+
 Table \ref{tab:summary_results} calls for two types of comments.

### 2021-07-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Test connexion

@@
 
-Our paper contributes XXX test XXX to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
+Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
 Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns.

### 2021-07-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte modele

@@
 
+\section{The role of additive cost: Theoretical insights}
+
+Our estimation results drive two main conclusions: First, additive costs constitute a significant part of international transports costs, in both maritime and air transport.
+Second, the size of international transport costs has decreased since the mid-1970s, in particular in air transport.
+While the share of additive costs is relatively stable in maritime transport, it has substantially decreased in air transport.
+\textbf{METTRE DES CHIFFRES}.
+
+The question we want to explore is about the implications of such results on theoretical grounds.
+While the whole bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive components.
+One potential explanation is the lack of empirical estimates about the size of additive costs.
+Our results may thus usefully palliate this limitation.
+Specifically, we can make use of our empirical results to answer two questions: First, what does it change to introduce additive costs along with standard iceberg trade costs, in particular regarding the pattern of trade flows?
+Second, what are the welfare consequences of a reduction in international transport costs, when such a reduction also occurs through a reduction in the additive component?
+These two key questions (the pattern of trade and the welfare gains of trade liberalization) are overwhelmingly addressed considering ad-valorem trade costs only.
+Given our empirical results pointing out the importance of additive costs, we explore how additive costs modify the picture.
+\textbf{A MIEUX DIRE}.
+
+\subsection{The model}
+
+Two papers have explored the role of additive costs on theoretical grounds: \cite{Irrazabal_2015} and \cite{sorensen2014}.
+As in \cite{sorensen2014}, we investigate the role of additive costs in the trade model of \cite{melitz}, considered in its canonical symmetric two-country version.
+\textbf{discuter de pourquoi on ne retient pas Chaney}.
+\textbf{Faut-il citer Lashkaripour ici aussi?} This allows to highlight the specificities induced by additive versus the standard iceberg (ad-valoram) costs.
+In contrast to \cite{sorensen2014}, we make use of our empirical results to quantify the effects of a reduction in international transports costs on both positive and normative dimensions.
+In accordance with the literature, we focus on steady-state analysis.
+Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
+
+\subsubsection{Households}
+
+In each country, the representative household exogenously supplies $L$ units of labour and maximizes utility $\mathcal{U} = \left[ \int_{\omega \in \Omega}c(\omega)^{\frac{\sigma-1}{\sigma}} d\omega \right]^{\frac{\sigma}{\sigma-1}}$, with $\sigma>1$ the elasticity of substitution between varieties.
+She chooses her consumption $c(\omega)$  of each variety $\omega$ within the set $\Omega$ of available varieties (made of both domestic and foreign varieties, endogenously defined), subject to her budget constraint: $\int_{\omega \in \Omega} p(\omega) c(\omega) dw \leq E$, with $p(\omega)$ the consumer price of variety $\omega$ and $E$ nominal expenditures.
+As standard, the optimal demand function for variety $\omega$ is:
+\begin{eqnarray*}
+c(\omega) &=& \left(\frac{p(\omega)}{P}  \right)^{-\sigma} \frac{E}{P}\\
+\end{eqnarray*}
+\noindent with the associated optimal price index $P$ equal to:
+\begin{equation}
+P = \left[ \int_{\omega \in \Omega}p(\omega)^{1-\sigma}d\omega\right]^{\frac{1}{1-\sigma}} \label{eq:CPI}
+\end{equation}
+
+\subsubsection{Firms behavior}
+
+As in \cite{melitz}, firms are in monopolistic competition, each producing a differentiated variety using labor as sole production factor.
+Both countries being perfectly symmetric, the labor market equilibrium imposes the same wage in both countries, thus considered as the num\'{e}raire.
+Prices and costs are hence expressed in terms of labor.
+At entry, each firm pays a sunk entry cost $f_e$ and subsequently draws its productivity level $\varphi$ from a known distribution $G(\varphi)$.
+Production of $q$ units, a firm should employ $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
+Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depend on the value traded $\tau\geq 1$ and an additive cost depending on the quantity (weight) transported, $t\qeg 0$.
+\textbf{Mettre une footnote, unit vs weight}.
+At each point in time, a firm can be hit by a death shock at the exogenous probability $\delta>0$.
+Following \cite{melitz}, we solve the problem backwards.
+First, we solve for the maximizing profit behavior once entered.
+Second, we determine the entry conditions on the various markets.\medskip
+
+\paragraph{Profit-maximizing behaviors} Once entered, a firm adopts a pricing-to-market behavior (conditional on exporting) by setting a specific price on each destination market.
+
+On the domestic market, the objective is to maximize profit $\pi_d(\omega)$, taken into account the consumer demand function\footnote{In a monopolistic set-up, each firm (of productivity $\varphi$) is the sole producer of variety $\omega$, implying a one-to-one link between $\omega$ and $\varphi$, such that we could write $\omega(\varphi)$.
+For sake of notational simplicity, we omit this link and simply write the consumer demand function as a function of $\varphi$ directly.}:
+\begin{eqnarray*}
+\max_{p(\varphi)} \pi_d(\varphi) &=& \left(p^d(\varphi) -\frac{1}{\varphi}  \right)c(\varphi) \\
+\text{s.t. }&&c(\varphi) = \left(\frac{p(\varphi)}{P}  \right)^{-\sigma} \frac{E}{P}
+\end{eqnarray*}
+The optimal pricing behavior yields:
+$$p^d(\varphi) = \frac{\sigma}{\sigma-1}\frac{1}{\varphi}$$
+With a CES specification on household's preferences, the firm sets its price as a constant mark-up on marginal cost, which is inversely related to its productivity as for the domestic market. \medskip
+
+Conditional on exporting abroad, the profit-maximizing program on the foreign market requires the inclusion of variable international trade costs according to:
+\begin{eqnarray*}
+\max_{p^x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
+\text{s.t. }&&c^x(\varphi) = \left(\frac{p^x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
+\end{eqnarray*}
+
+\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p^x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
+Solving for this yields:
+$$p^x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right]$$
+
+Similarly as for the Home market, the firm sets a constant markup on its marginal cost, which is still a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs.
+
+Under the assumption of symmetric countries (i.e, implying $E=E^\ast$ and $P=P^\ast$), one can deduce the profits on the domestic and export markets ($\pi_d(\varphi)$ and $\pi_x(\varphi)$) respectively as:
+\begin{eqnarray}
+\pi_d(\varphi)  &=& \frac{r_d(\varphi)}{\sigma} - f = B \varphi^{\sigma-1} - f \label{eq:pid_rd}\\
+\pi_x(\varphi) &=& \frac{r_x(\varphi)}{\sigma} - f_x = B \left[ \frac{\tau}{\varphi} + t\right]^{-(\sigma-1)} - f_x \label{eq:pix_rx}\\
+\text{with}&& B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E \notag
+\end{eqnarray}
+\noindent where $r_d(\varphi)$ and $r_x(\varphi)$ denote the sales revenues on each market.
+Profits increase with productivity $\varphi$, and only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
+Further, in presence of trade costs, a given firm of productivity $\varphi$ makes a lower export profit than on its domestic market, conditional on being productive enough to export.
+This drives us to determine the productivity threshold associated to the entry conditions.
+
+Making use of the firm's first-order pricing conditions, revenues on each local and export market respectively read as:
+\begin{eqnarray}
+% \nonumber % Remove numbering (before each equation)
+  r_d(\varphi) &=& \left[\frac{1}{\rho } \frac{1}{\varphi}  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rd}\\
+  r_x(\varphi) &=& \left[\frac{1}{\rho}\left(\frac{\tau}{\varphi} +t\right)  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rx}
+\end{eqnarray}
+\noindent with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup.
+\paragraph{Zero-cutoff profit conditions} Given the fixed costs on both the domestic and the export markets, only firms with a productivity above $\varphi^\ast$ will serve the domestic market, and those with $\varphi>\varphi_x^\ast$ will export abroad.
+The zero cut-off (ZCP) productivity level for successful entry on the domestic and export markets should verify:
+\begin{eqnarray*}
+\frac{\pi_d(\varphi^\ast)}{\delta} = 0 &\Leftrightarrow & \frac{r_d(\varphi^\ast)}{\sigma} - f =0\\
+\frac{\pi_x(\varphi_x^\ast)}{\delta} = 0 &\Leftrightarrow & \frac{r_x(\varphi_x^\ast)}{\sigma} - f_x = 0
+\end{eqnarray*}
+\noindent with $\frac{\pi_d(\varphi^\ast)}{\delta}$ the expected profit on each market. Manipulating equations, we get:
+\begin{eqnarray*}
+% \nonumber % Remove numbering (before each equation)
+  \varphi_x^\ast &=& \frac{\tau}{f_x^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}} -t} \\
+  \varphi^\ast &=& \frac{1}{f_^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}
+\end{eqnarray*}
+
+The two productivity thresholds are accordingly linked according to:
+\begin{equation}
+\frac{\varphi_x^\ast}{\tau+t\varphi_x^\ast} = \varphi^\ast \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} \label{eq:link_thresholds}
+\end{equation}
+
+Following \cite{melitz} and the subsequent literature, we ensure selection on the export market, i.e.
+such that $\varphi_x^\ast > \varphi^\ast$.
+Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as:
+$$\tau+ t\phi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$$
+
+Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\phi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
+We check this condition holds numerically in solving the model.\textbf{A FAIRE ???
+JE CROYAIS LE FAIRE MAIS JE NE LE FAIS QUE DANS LE CAS SANS ADDITIFS.}
+
+\subsubsection{Aggregation, firm entry and exit}
+
+\paragraph{Aggregation} The general equilibrium is characterized by a mass $M$ of producing firms in each country associated with a distribution $\mu(\varphi)$ of productivity levels of a subset of $\left(0,\infty\right)$.
+Given the productivity threshold $\phi^\ast$, $\mu(\varphi)$ is equal to the conditional distribution of $g(\varphi)$ on the subset $\left[\phi^\ast, \infty\right)$:
+\begin{eqnarray}
+\mu(\varphi) &=& \left\{
+\begin{array}{ll}
+\frac{g(\varphi)}{1-G(\varphi^\ast)}, & \text{if } \varphi\geq \varphi^\ast  \\
+0 & \text{otherwise}
+\end{array}
+\right.  \label{eq:muofphi}
+\end{eqnarray}
+\noindent with $1-G(\varphi^\ast)$ the probability of successful entry on the domestic market.
+Conditional on being active on the domestic market, a subset mass $M_x$ of firms export abroad.
+$\mu_x(\varphi)$ can be similarly defined as the conditional distribution of productivity on the export market, e.g.
+on the support $\left[ \varphi^\ast_x,\infty\right)$.
+
+Following \cite{melitz}, and with $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can then rewrite the price index (\ref{eq:CPI}) as:
+\begin{eqnarray*}
+P &=& \left[\int_{\omega \in \Omega_d} p_d^{1-\sigma}(\omega)d\omega + \int_{\omega \in \Omega_x} p_x^{1-\sigma}(\omega)d\omega  \right]^{\frac{1}{1-\sigma}} \\
+&=& \left[\int_{0}^\infty p_d^{1-\sigma}(\varphi)M\mu(\varphi)d\varphi + \int_{0}^\infty p_x^{1-\sigma}(\varphi)M_x \mu_x(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}}
+\end{eqnarray*}
+Making use of Equation (\ref{eq:muofphi}), the price index reads as:
+\begin{eqnarray*}
+P &=& \left[\frac{M}{1-G(\varphi^\ast)}\int_{\phi^\ast}^\infty p_d^{1-\sigma}(\varphi)g(\varphi)d\varphi + \frac{M_x}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast_x}^\infty p_x^{1-\sigma}(\varphi)g(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}} \\
+\text{with}&& \left\{
+\begin{array}{lll}
+p_d(\varphi)&=& \frac{\sigma}{\sigma-1} \frac{1}{\varphi}\\
+p_x(\varphi)&=& \frac{\sigma}{\sigma-1} \left[\frac{\tau}{\varphi} +t\right]
+\end{array}
+\right.
+\end{eqnarray*}
+Accordingly, the price index can be rewritten as:
+$$P = \left(\frac{\sigma}{\sigma-1}\right)\left[\frac{M}{1-G(\varphi^\ast)}\int_{\phi^\ast}^\infty \phi^{\sigma-1}g(\varphi)d\varphi + \frac{M_x}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast_x}^\infty \left[\varphi \left(\frac{\tau}{\tau+t \varphi}  \right)  \right]^{\sigma-1}g(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}}$$
+
+Similarly as in the closed-economy case (\cite{melitz}), let us define $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ as the weighted average of productivities of the domestic firms, on the domestic market and the export market respectively:\footnote{In the symmetric two-country model, $\widetilde{\varphi}_x(\varphi_x^\ast)$ can also be defined as the weighted average of productivities of foreign exporters at Home, e.g.
+associated with the imported varieties.}
+
+\begin{eqnarray}
+\widetilde{\varphi} &=& \left[\frac{1}{1-G(\varphi^\ast)}\int_{\varphi^\ast}^\infty \phi^{\sigma-1}g(\varphi)d\varphi   \right]^{\frac{1}{\sigma-1}} \label{eq:def_phitilde}\\
+\widetilde{\varphi}_x &=& \left[\frac{1}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast}^\infty \left[\varphi\left(\frac{\tau}{\tau+t \varphi}  \right) \right]^{\sigma-1}g(\varphi)d\varphi   \right]^{\frac{1}{\sigma-1}} \label{eq:def_phitildex}
+\end{eqnarray}
+
+Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, it comes that:
+$$P = \frac{\sigma}{\sigma-1}M_T^{\frac{1}{1-\sigma}}\left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$$
+
+With $\widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$ the weighted average of firm productivities on the domestic market (including both local producers and foreign producers exporting at Home), the price index can be expressed as:
+\begin{equation}
+P = M_T^{\frac{1}{1-\sigma}}\frac{1}{\rho \widetilde{\varphi}_T } \label{eq:CPI_vf}
+\end{equation}
+
+\paragraph{Free-entry condition} There is free-entry in the monopolistic industry, imposing that the expected net present value of entry is driven to 0.
+In the open-economy setting, this condition is given by:
+\begin{equation*}
+  \sum_{t=0}^{\infty}(1-\delta)^t \left(\int_{\varphi^\ast}^{\infty} \pi_d(\varphi )dG(\varphi) + \int_{\varphi_x^\ast}^{\infty} \pi_x(\varphi )dG(\varphi)  \right) = f_e
+\end{equation*}
+
+Defining $\bar{\pi}$ as the average profit conditional to entry, this free-entry condition rewrites as:
+\begin{equation}
+\left(1-G(\varphi^\ast)\right)\frac{\bar{\pi}}{\delta} = f_e \label{eq:FEC}
+\end{equation}
+\noindent with $\left(1-G(\varphi^\ast)$ the probability of successful entry. In the open-economy setting, $\bar{\pi}$ can be expressed as:
+$$\bar{pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
+\noindent with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi^}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to:
+$$\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$$
+
+Noticing that, from Equation (\ref{eq:rd}), $\frac{r_d(\varphi)}{r_d(\varphi^\ast)} = \left( \frac{\varphi}{\varphi^\ast} \right)^{\sigma-1}$ and similarly on the export market, Equations (\ref{eq:pid_rd}) and (\ref{eq:pix_rx}) can be rewritten as:
+\begin{eqnarray*}
+\bar{\pi}_d &=& \left\{\left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1}-1 \right\}f \\
+\bar{\pi}_x &=& \left\{\left[\left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)\right]^{\sigma-1}-1 \right\}f_x
+\end{eqnarray*}
+
+Making use of the above elements, we finally get the zero-cut off profit condition in the open economy with additive costs as:
+\begin{equation}
+\bar{\pi} = \left\{\left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1}-1 \right\}f +  \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}\left\{\left[\left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)\right]^{\sigma-1}-1 \right\}f_x \label{eq:ZCP}
+\end{equation}
+\noindent with $\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_x^\ast$ given by Equations (\ref{eq:def_phitilde}) and (\ref{eq:def_phitildex}).
+
+\paragraph{Summary} Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$, from which we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
+Specifically, we will study the consequences of additive costs in terms of welfare, which can be shown to be the inverse of the price level as standard in this literature (the price level $P$ being given by Equation (\ref{eq:CPI_vf})).
+
 \section{Conclusion \label{sec:conclu}}
@@
 \noindent with $\tau_{ikt}$ appropriately defined.
+
 This exactly corresponds to the case of ad-valorem transport costs only.
@@
 
+\section{More on the model \label{app:theoretical_model}}
+
+\subsection{General equilibrium}
+
+Following \cite{melitz} and the subsequent literature, denoting $M_e$ the mass of prospective entrants and $\rho_e$ the probability of successful entry (equal to $1-G(\varphi^\ast)$, considering only steady-state equilibrium needs to verify the entry-exit flow equilibrium condition:
+$$rho_e M_e = \delta M$$.
+
+Given the wage normalization, total revenue in each country should be such that: $E = L_p+\Pi$, with $L_p$ the number of workers of the productive sector and $\Pi$ total profits.
+Labor market equilibrium further requires: $L = L_p+ L_e$ with $L_e = f_eM_e$ as sunk entry costs are paid in terms of labor.
+Combining this with the free-entry condition (\ref{eq:FEC}), we obtain that $\Pi = L_e$: Total profits in the productive sector are used to pay the entry sunk costs.
+Making use of this, we obtain that aggregate income in each country is equal to its labor endowment: $E =L$.
+
+Further, total revenue can de expressed as the product of the average sales and the mass of active producers, according to: $E = M\bar{r}$, with $\bar{r}$ average sales.
+Similarly as for profits, it can be expressed as the sum of average sales on each local and foreign market:
+$$\bar{r} = r_d(\widetilde{\varphi}(\varphi^\ast))+ \rho_x r_d(\widetilde{\varphi}_x(\varphi_x^\ast))$$
+\noindent with the (conditional) probability of export $\rho_x = \frac{1-G(\varphi^\ast_x)}{1-G(\varphi^\ast)}$ and average sales on each market which can be shown to be:
+\begin{eqnarray*}
+\bar{r}_d &=& \left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1} \sigma f \\
+\bar{r}_x &=& \left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)^{\sigma-1}\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)^{\sigma-1} \sigma f_x
+\end{eqnarray*}
+
+Given selection on the export market, the mass of domestic exporters depends on that of domestic active firms according to: $M_x = \rho_x M$.
+In the symmetric equilibrium, the total mass of available varieties is then: $M_T= M+M_x = (1+\rho_x)M$.
+
+\subsection{Summary}
+
+\subsubsection{General form}
+Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, \widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$.
+We can deduce the equilibrium values of the remaining set of variables $\left\{P, M_T, M, M_x, \widetilde{\varphi}_T, \bar{r} \right\}$ by recursively solving the following set of equations:
+
+\begin{eqnarray*}
+% \nonumber % Remove numbering (before each equation)
+ \bar{r}&=& \left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1} \sigma f + \frac{1-G(\varphi^\ast_x)}{1-G(\varphi^\ast)}\left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)^{\sigma-1}\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)^{\sigma-1} \sigma f_x \\
+ M &=& \frac{L}{\bar{r}} \\
+M_x &=& \frac{1-G(\varphi^\ast_x)}{1-G(\varphi^\ast)} M \\
+M_T &=& \left(1+ \frac{1-G(\varphi^\ast_x)}{1-G(\varphi^\ast)} \right)M \\
+\widetilde{\varphi}_T &=& \widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}} \\
+P &=& M_T^{\frac{1}{1-\sigma}} \frac{\rho}{\widetilde{\varphi}_T}
+\end{eqnarray*}
+
 \end{document}

### 2021-07-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Amendement .tex

@@
 Production of $q$ units, a firm should employ $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
-Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depend on the value traded $\tau\geq 1$ and an additive cost depending on the quantity (weight) transported, $t\qeg 0$.
+Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depend on the value traded $\tau\geq 1$ and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
 \textbf{Mettre une footnote, unit vs weight}.
@@
   \varphi_x^\ast &=& \frac{\tau}{f_x^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}} -t} \\
-  \varphi^\ast &=& \frac{1}{f_^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}
+  \varphi^\ast &=& \frac{1}{f^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}
 \end{eqnarray*}
@@
 
-Following \cite{melitz}, and with $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can then rewrite the price index (\ref{eq:CPI}) as:
+Following \cite{melitz}, and decomposing the set of available varieties into $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can rewrite the price index (\ref{eq:CPI}) as:
 \begin{eqnarray*}
@@
 
-Similarly as in the closed-economy case (\cite{melitz}), let us define $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ as the weighted average of productivities of the domestic firms, on the domestic market and the export market respectively:\footnote{In the symmetric two-country model, $\widetilde{\varphi}_x(\varphi_x^\ast)$ can also be defined as the weighted average of productivities of foreign exporters at Home, e.g.
+Similarly as in the closed-economy case (\cite{melitz}), let us define $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ as the weighted average of productivities of the domestic firms, on the domestic market and the export market respectively:\footnote{In the symmetric two-country model, $\widetilde{\varphi}_x(\varphi_x^\ast)$ also corresponds to the weighted average of productivities of foreign exporters at Home, e.g.
 associated with the imported varieties.}
@@
 \end{equation}
-\noindent with $\left(1-G(\varphi^\ast)$ the probability of successful entry. In the open-economy setting, $\bar{\pi}$ can be expressed as:
+
+\noindent with $\left(1-G(\varphi^\ast)\right)$ the probability of successful entry. In the open-economy setting, $\bar{\pi}$ can be expressed as:
 $$\bar{pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
-\noindent with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi^}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to:
+\noindent with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to:
 $$\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$$
@@
 
-\paragraph{Summary} Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$, from which we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
+\paragraph{Summary} Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$.
+From this, we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
 Specifically, we will study the consequences of additive costs in terms of welfare, which can be shown to be the inverse of the price level as standard in this literature (the price level $P$ being given by Equation (\ref{eq:CPI_vf})).
+
+\subsection{Calibration}
+
+As standard in this literature (\cite{Irrazabal_2015}, Melitz fial proofs), the heterogenity in firm productivity is assumed to be given by a Pareto distribution.
+In contrast to the standard case with only ad-valorem costs, the model can no longer be solved analytically due to the presence of additive costs (see Appendix \ref{app:theoretical_model}).
+In this case, we rely on the model's simulations to determine the steady state.
+This drives us to calibrate the structural parameters of the model.
+To do so, we rely on values commonly retained in the literature except for the two additive and ad-valorem transport costs, as reported in Table .
+\begin{table}[htbp]
+  \centering
+  \caption{Calibration (1)} \label{tab:calib_horsTC}
+\begin{center}
+	\input{donnees_calibration.tex}
+\end{center}
+\end{table}
+The calibration of transport costs parameters relies on our empirical estimates.
+We notably illustrate the consequences of trade liberalization by comparing the equilibrium based on the 1974 values of transport costs (at the beginning of our sample), versus those of 2019 (at the very end of our estimation period), in both cases on only ad-valorem transport costs and in the presence of both additive and ad-valorem costs.
+Calibration of transport costs is displayed in Table \ref{tab:calib_TC}.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Calibration (2)}\label{tab:calib_TC}
+\begin{center}
+	\input{table_calibTC_combined.tex}
+\end{center}
+{\parbox[l]{12cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
+In columns (1), (3), (5) and (7) only ad-valorem costs are considered.
+In Columns (2), (4), (6) and (8), we model both ad-valorem and additive costs.
+$t$ is calibrated so that the ratio $t/\widetilde{p}_x$ from the model matches its empirical target, with $widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
+We theoretically define the export fas price through: $p_x = \tau p^{\text{fas}} +t$}}}
+\end{table}
 
@@
 
-\subsection{General equilibrium}
+\subsection{Solving the general equilibrium}
 
@@
 
-\subsection{Summary}
-
-\subsubsection{General form}
+\subsubsection{Summary}
+
 Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, \widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$.
@@
 
+\subsection{With a Pareto distribution}
+
+As in most of the related literature (\cite{Irrazabal_2015}, \textbf{Melitz final proofs}, ...), we adopt a Pareto distribution for firm productivity heterogeneity, with the density and the cumulative functions respectively written as:
+\begin{eqnarray*}
+g(\varphi) &=& k\varphi_{\text{min}}^k \varphi^{-(k+1)}\\
+G(\varphi) &=& 1 -\left( \frac{\varphi_{\text{min}}}{\varphi} \right)^k
+\end{eqnarray*}
+
+With the Pareto distribution, the system of Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) can be rewritten as a system of four equations in four endogenous variables: $\left\{\bar{pi}, g(\varphi_x^\ast),\varphi_x^\ast, \varphi^\ast  \right\}$ as given by:
+
+\begin{eqnarray}
+\bar{\pi}&=& \frac{\sigma-1}{k-(\sigma-1)}f + \left(\varphi^\ast \right)^k f_x\left[k\left(\frac{\tau}{\varphi_x^\ast}+t  \right)^{\sigma-1}\left[  1+(k(\varphi_x^\ast)^k g(\varphi_x^\ast))^{\frac{1}{\sigma-1}}\right]^{1-\sigma}  - \left( \varphi_x^\ast \right)^{-k} \right] \label{eq:systemPareto_1}\\
+g(\varphi_x^\ast)&=&\int_{\varphi_x^\ast}^\infty \varphi^{\sigma-1-(k+1)}(\tau +t\varphi)^{1-\sigma}d\varphi \label{eq:systemPareto_2}\\
+\frac{\varphi_x^\ast}{\tau+ t\varphi_x^\ast}&=&\varphi^\ast \left[ \frac{f}{f_x} \right]^{\frac{1}{1-\sigma}} \label{eq:systemPareto_3}\\
+\bar{\pi}&=& \delta f_e \left( \frac{\varphi^\ast}{\varphi_{\text{min}}} \right)^k \label{eq:systemPareto_4}
+\end{eqnarray}
+
+Notice that, in the absence of additive costs ($t=0$), Equation (\ref{eq:systemPareto_2}) rewrites as:
+$$g(\varphi_x^\ast) = \tau^{1-\sigma}\left[ \frac{(\varphi_x^\ast)^{-(k-(\sigma-1))}}{k-(\sigma-1)}\right]$$
+The system of Equations (\ref{eq:systemPareto_1})-((\ref{eq:systemPareto_4}) thus simplifies to be solved recursively with a full analytical solution.
+This is no longer the case in the presence of additive costs, and we have to rely on the model's simulations to obtain the general equilibrium of the model for a given calibration of the deep parameters.
+
 \end{document}

### 2021-07-27 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
Amendement .tex

\begin{tabular}{lll}
& Value & Reference\\
\hline
sigma & 4 & \cite{Irrazabal_2015}\\
fe & 1 & \cite{ghironi}\\
f & 0.033333 & Share of exporters = 21\% (\cite{ghironi}) \\
L & 1 & Arbitrary\\
$\delta$ & 0.1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
k & 4 \\
$\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
\hline
\end{tabular}

### 2021-07-28 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Continuer sur le modèle

@@
 This drives us to calibrate the structural parameters of the model.
-To do so, we rely on values commonly retained in the literature except for the two additive and ad-valorem transport costs, as reported in Table .
+To do so, we rely on values commonly retained in the literature except for the two additive and ad-valorem transport costs, as reported in Table \ref{tab:calib_horsTC}.
 \begin{table}[htbp]
@@
 \end{table}
+
+$f$ such that all firms produce in autarky (i.e., such that $\varphi^{\ast,a} = \varphi_{\text{min}}$.
+$f_x$ endogenously revealed such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, based on \textbf{???}, following so the calibration method of \cite{ghironi} or \cite{Irrazabal_2015}.
+
 The calibration of transport costs parameters relies on our empirical estimates.
@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs.
 In columns (1), (3), (5) and (7) only ad-valorem costs are considered.
@@
 We theoretically define the export fas price through: $p_x = \tau p^{\text{fas}} +t$}}}
+\end{table}
+
+\textbf{A DEVELOPPER}
+
+\subsection{Trade liberalization: The importance of additive costs}
+
+Our empirical results point to two important conclusions: First, additive costs represent a non-negligible part of overall international transport costs.
+Second, total transport costs are strongly reduced over 1974-2019, and the decrease is particularly marked in the additive component in air transport.
+We exploit the theoretical implications of these empirical findings in two different ways.
+
+First, we run a comparative statics analysis following a reduction in transport costs based on our estimates in 1974 and 2019, by transport mode.
+Specifically, we contrast two cases: In the first scenario, the reduction in total transport costs is solely attributable to the ad-valorem component (e.g., contrasting Table \ref{tab:calib_TC}, Column (3) to (1) for air transport, and Column (7) to (5) for maritime transport); in the second scenario, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
+In this scenario, we contrast Table \ref{tab:calib_TC}, Column (4) to (2) (Air) and Column (8) to (6) (Vessel).
+Importantly, we run this comparative statics analysis \textit{ceteris paribus}, notably with the same value of the export fixed cost $f_x$ as in the initial equilibrium (see Table \ref{tab:calib_horsTC}.
+
+Second, we take into account the fact that the share of exporting firms has substantially increased since the 21\% value estimated by \textbf{???} in \textbf{YEAR YYY}.
+According to \textbf{???}, this share roughly amounts to 35\% in \textbf{???}.
+This can be matched in the model by resetting the (lower) value of the export fixed cost $f_x$ accordingly.
+In the second exercise, we thus run a comparative statics analysis where we contrast the 2014 steady states (e.g., with the 2019 transport costs values) with the two dimensions of trade liberalization: the reduction of transport costs (as in the first exercise) and the reduction of the export fixed cost, consistent with the increase in the share of exporters from 21\% to 35\%.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele}
+\begin{center}
+\begin{tabular}{l|c|c|c|c||c|c|c|c}
+\hline
+Mode & \multicolumn{4}{|c||}{Air} & \multicolumn{4}{c}{Vessel} \\ \hline
+Exercise &  \multicolumn{2}{|c|}{Constant $f_x$} &  \multicolumn{2}{|c||}{Lower $f_x$} & \multicolumn{2}{|c|}{Constant $f_x$} &  \multicolumn{2}{|c}{Lower $f_x$} \\ \hline
+Reduction in & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\ \hline
+$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
+$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
+$\Delta TC$ (in pp) &  -0.034 & -0.034 & -0.034 & -0.034 & -0.064 & -0.064 & -0.064 & -0.064\\
+$\Delta \tau$ (in pp) & -0.0340	&-0.0140&	-0.0340&	-0.0140	&-0.064&	-0.031&	-0.064&	-0.031  \\
+$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-0.02&	0&	-0.02&	0&	-0.033&	0&	-0.033  \\ \hline
+$\Delta \mathcal{U}$ & 0.0202&	0.0295	&0.0362&	0.0527	&0.035&	0.052	&0.047&	0.064 \\
+$\Delta \mathcal{U}/\mathcal{U}$ (in \%) & 1.2319&	1.8187&	2.2073	&3.2477	&2.179	&3.295&	2.872&	4.052   \\
+\hline
+\end{tabular} 
+\end{center}
+{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price.}}}
 \end{table}

### 2021-07-28 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
Continuer sur le modèle

@@
 \hline
-sigma & 4 & \cite{Irrazabal_2015}\\
-fe & 1 & \cite{ghironi}\\
-f & 0.033333 & Share of exporters = 21\% (\cite{ghironi}) \\
-L & 1 & Arbitrary\\
+$\sigma$ & 4 & \cite{Irrazabal_2015}\\
+$f_e$ & 1 & \cite{ghironi}\\
+$f$ & 0.033333 & \cite{ghironi} \\
+$f_x$ & ??? & Share of exporters = 21\% (\cite{ghironi}, \cite{Irrazabal_2015}) \\
+$L$ & 1 & Arbitrary\\
 $\delta$ & 0.1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
-k & 4 \\
+$k$ & 4 \\
 $\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\

### 2021-07-28 redaction/JEGeo/revision_JEGeo/revised_article/table_compWelfare2_air.tex
Continuer sur le modèle

\begin{tabular}{llllll}
& Mode (1 Air, 2 Ves) & Lower tau, same fx & Lower tau AND fx (export share = 35%) & Lower tau and t, same fx & Lower tau, t, AND fx (export share = 35%) \\ 
\hline 
TC(0) & 1 & 0.062 & 0.062 & 0.062 & 0.062 \\ 
TC(1) & 1 & 0.028 & 0.028 & 0.028 & 0.028 \\ 
Delta Total TC (pp) & 1 & -0.034 & -0.034 & -0.034 & -0.034 \\ 
Delta tau (pp) & 1 & -0.034 & -0.034 & -0.014 & -0.014 \\ 
Delta t/pfas (pp) & 1 & 0 & 0 & -0.020007 & -0.020007 \\ 
Delta Welfare (abs) & 1 & 0.020229 & 0.036246 & 0.029487 & 0.052655 \\ 
Rel. Delta welf (in %) & 1 & 1.2319 & 2.2073 & 1.8187 & 3.2477 \\ 
\hline 
\end{tabular}

### 2021-07-29 redaction/JEGeo/revision_JEGeo/revised_article/table_compWelfare2_vessel.tex

\begin{tabular}{llllll}
& Mode (1 Air, 2 Ves) & Lower tau, same fx & Lower tau AND fx (export share = 35%) & Lower tau and t, same fx & Lower tau, t, AND fx (export share = 35%) \\ 
\hline 
TC(0) & 2 & 0.105 & 0.105 & 0.105 & 0.105 \\ 
TC(1) & 2 & 0.041 & 0.041 & 0.041 & 0.041 \\ 
Delta Total TC (pp) & 2 & -0.064 & -0.064 & -0.064 & -0.064 \\ 
Delta tau (pp) & 2 & -0.064 & -0.064 & -0.031 & -0.031 \\ 
Delta t/pfas (pp) & 2 & 0 & 0 & -0.032947 & -0.032947 \\ 
Delta Welfare (abs) & 2 & 0.035413 & 0.046674 & 0.052202 & 0.064189 \\ 
Rel. Delta welf (in %) & 2 & 2.1791 & 2.8719 & 3.2951 & 4.0517 \\ 
\hline 
\end{tabular}

### 2021-07-30 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

@@
 
+The results are reported in Table \ref{tab:resultats_modele}, for Air (Columns (1)-(4)) and Vessel (Columns (5)-(8)) respectively.
+Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
+For each transport mode, the first two columns report the results of reducing transport costs only, maintaining the same value of $f_x$ as in the initial equilibrium (Columns (1)-(2) for Air, (5)-(6) for Vessel), contrasting the case of only ad-valorem costs (Columns (1), (5)) and in presence of transport costs (Columns (2), (6)).
+In Columns (3)-(4) for Air and (7)-(8), we report the results when both the variable transport costs and the fixed export costs are lower in the final steady state (2019) relative to the initial one (1974).
+
 \begin{table}[htbp]
@@
 Mode & \multicolumn{4}{|c||}{Air} & \multicolumn{4}{c}{Vessel} \\ \hline
-Exercise &  \multicolumn{2}{|c|}{Constant $f_x$} &  \multicolumn{2}{|c||}{Lower $f_x$} & \multicolumn{2}{|c|}{Constant $f_x$} &  \multicolumn{2}{|c}{Lower $f_x$} \\ \hline
-Reduction in & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\ \hline
-$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
-$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
-$\Delta TC$ (in pp) &  -0.034 & -0.034 & -0.034 & -0.034 & -0.064 & -0.064 & -0.064 & -0.064\\
-$\Delta \tau$ (in pp) & -0.0340	&-0.0140&	-0.0340&	-0.0140	&-0.064&	-0.031&	-0.064&	-0.031  \\
-$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-0.02&	0&	-0.02&	0&	-0.033&	0&	-0.033  \\ \hline
+Fixed export cost: &  \multicolumn{2}{|c|}{Constant $f_x$} &  \multicolumn{2}{|c||}{Lower $f_x$} & \multicolumn{2}{|c|}{Constant $f_x$} & \multicolumn{2}{|c}{Lower $f_x$} \\ \hline
+Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\ \hline
+%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
+%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
+$\Delta TC$ (in pp) &-3.4 &-3.4 &-3.4 &-3.4 &-6.4 &-6.4 &-6.4 &-6.4  \\ 
+$\Delta \tau$ (in pp) & -3.4	&-1.4&	-3.40&	-1.40	&-6.4&	-3.1&	-6.4&	-3.1  \\ 
+$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3&	0&	-3.3  \\ \hline
 $\Delta \mathcal{U}$ & 0.0202&	0.0295	&0.0362&	0.0527	&0.035&	0.052	&0.047&	0.064 \\
@@
 \end{center}
-{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price.}}}
+{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs. The reduction in transport costs is expressed in percentage points. }}}
 \end{table}
+
+\subsubsection{Reducing international transport costs}
+
+In this first exercise, we contrast the effects of a reduction of international transport costs depending on whether this occurs through a reduction in ad-valorem costs only, or a combined reduction in both the additive and the multiplicative components.
+Accordingly, we focus on the results reported in Table \ref{tab:resultats_modele}, Column (1) versus (2) for Air, Column (5) versus (6) for Vessel.
+
+A first conclusion emerges: Whatever the transport mode (and the magnitude of the transport costs decrease accordingly), welfare gains are higher when this reduction is made (partly) through a reduction of additive costs.
+The effect is substantial: In Air transport, the reduction of 3.4 percentage points of transport costs generates a welfare gain higher by 0.58 percentage point when almost 60\% of such a reduction occurs through the additive component, rather than when it realizes through a decrease of iceberg costs only.
+In maritime transport, where the decrease in total transport costs is larger (by 6.4 percentage points), the extra welfare gain amounts to 1.12 percentage points when 50\% of the transport cost reduction occurs through the additive component.
 

### 2021-07-30 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex

--- 2021-07-27 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
@@
 \hline
-sigma & 4 & \cite{Irrazabal_2015}\\
-fe & 1 & \cite{ghironi}\\
-f & 0.033333 & Share of exporters = 21\% (\cite{ghironi}) \\
-L & 1 & Arbitrary\\
+$\sigma$ & 4 & \cite{Irrazabal_2015}\\
+$f_e$ & 1 & \cite{ghironi}\\
+$f$ & 0.033 & \cite{ghironi} \\
+$f_x$ & 0.09 & Share of exporters = 21\% (\cite{ghironi}, \cite{Irrazabal_2015}) \\
+$L$ & 1 & Arbitrary\\
 $\delta$ & 0.1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
-k & 4 \\
+$k$ & 4 \\
 $\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\

### 2021-07-31 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Les tables sont bonnes

@@
 	\input{Table1.tex}
+		\begin{tablenotes}
+		\tiny
+		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
+		\item The additive term is expressed in fraction of fas price.
+		\item For the  4-digit model, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
+		1974, 1981, 1989, 2001, 2009, 2013, 2017 and 2019
+		\item \textbf{Model (A): With ad-valorem transpost costs only}
+		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+	\end{tablenotes}
 \end{table}

### 2021-08-04 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Figure 1 dans le texte
Amendement texte
Pas encore fini...

--- 2021-07-30 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 
-The two productivity thresholds are accordingly linked according to:
+The two productivity thresholds are linked according to:
 \begin{equation}
@@
 Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as:
-$$\tau+ t\phi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$$
-
-Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\phi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
+$$\tau+ t\varphi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$$
+
+Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
 We check this condition holds numerically in solving the model.\textbf{A FAIRE ???
@@
 \paragraph{Aggregation} The general equilibrium is characterized by a mass $M$ of producing firms in each country associated with a distribution $\mu(\varphi)$ of productivity levels of a subset of $\left(0,\infty\right)$.
-Given the productivity threshold $\phi^\ast$, $\mu(\varphi)$ is equal to the conditional distribution of $g(\varphi)$ on the subset $\left[\phi^\ast, \infty\right)$:
+Given the productivity threshold $\varphi^\ast$, $\mu(\varphi)$ is equal to the conditional distribution of $g(\varphi)$ on the subset $\left[\varphi^\ast, \infty\right)$:
 \begin{eqnarray}
@@
 \noindent with $\left(1-G(\varphi^\ast)\right)$ the probability of successful entry. In the open-economy setting, $\bar{\pi}$ can be expressed as:
-$$\bar{pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
+$$\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
 \noindent with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to:
@@
 From this, we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
-Specifically, we will study the consequences of additive costs in terms of welfare, which can be shown to be the inverse of the price level as standard in this literature (the price level $P$ being given by Equation (\ref{eq:CPI_vf})).
+Specifically, we will study the consequences of additive costs in terms of welfare.
+As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, the price level $P$ being given by Equation (\ref{eq:CPI_vf}).
 
@@
 $\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3&	0&	-3.3  \\ \hline
-$\Delta \mathcal{U}$ & 0.0202&	0.0295	&0.0362&	0.0527	&0.035&	0.052	&0.047&	0.064 \\
-$\Delta \mathcal{U}/\mathcal{U}$ (in \%) & 1.2319&	1.8187&	2.2073	&3.2477	&2.179	&3.295&	2.872&	4.052   \\
+$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\
 \hline
@@
 In this first exercise, we contrast the effects of a reduction of international transport costs depending on whether this occurs through a reduction in ad-valorem costs only, or a combined reduction in both the additive and the multiplicative components.
-Accordingly, we focus on the results reported in Table \ref{tab:resultats_modele}, Column (1) versus (2) for Air, Column (5) versus (6) for Vessel.
-
-A first conclusion emerges: Whatever the transport mode (and the magnitude of the transport costs decrease accordingly), welfare gains are higher when this reduction is made (partly) through a reduction of additive costs.
-The effect is substantial: In Air transport, the reduction of 3.4 percentage points of transport costs generates a welfare gain higher by 0.58 percentage point when almost 60\% of such a reduction occurs through the additive component, rather than when it realizes through a decrease of iceberg costs only.
-In maritime transport, where the decrease in total transport costs is larger (by 6.4 percentage points), the extra welfare gain amounts to 1.12 percentage points when 50\% of the transport cost reduction occurs through the additive component.
+In Table \ref{tab:resultats_modele}, this relies on the results reported in Column (1) versus (2) for Air, Column (5) versus (6) for Vessel.\medskip
+
+Whatever the transport mode (and the associated magnitude of the transport costs decrease), welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
+The effect is substantial: The reduction of 3.4 percentage points of Air transport costs generates a welfare gain higher by 0.58 percentage point when such a reduction occurs through the additive component (accounting for almost 60\% of the total cost decrease), rather than when it realizes through a decrease in iceberg costs only.
+In maritime transport, where the decrease in total transport costs is larger (-6.4 percentage points), the extra welfare gain amounts to 1.12 percentage points when the transport cost reduction occurs through the additive component (by 50\%) rather than through the iceberg component only.\smallskip
+
+Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
+Considering the inverse of Equation (\ref{eq:CPI_vf}), it is straightforward to get the welfare change as:
+\begin{equation}
+\frac{d\mathcal{W}}{\mathcal{W}_0} = \underbrace{\frac{1}{\sigma-1} \frac{d M_T}{M_{T0}}}_{\text{Variety effect}} +  \underbrace{\frac{d\widetilde{\varphi}_T}{\widetilde{\varphi}_{T0}}}_{\text{Price effect}} \label{eq:decompWelfare}
+\end{equation}
+
+Figure \ref{fig:welfare_decomp} displays the welfare change and its decomposition between its two components following the reduction in international transport costs, for the decrease estimated in air transport (left column, Panels (1), (1-a) and (1-b)) and in maritime transport (right column, Panels (2), (2-a) and (2-b)).
+All are expressed in percentage change from the initial steady-state value.
+In each panel, the blue left bar displays the case where the trade liberalization occurs through a reduction in ad-valorem costs; the grey right bar reporting the case where the (same) reduction in international transport costs occurs through a combined reduction in both the ad-valorem and the additive components, based on our estimated results in 1974 and 2019 (for each transport mode), as in Table \ref{tab:calib_TC}).
+
+\begin{figure}[htbp]
+\caption{Welfare decomposition}
+\label{fig:welfare_decomp}
+\includegraphics[width=15cm, height=12cm]{Welfare_change_bothmodes.pdf}
+\begin{minipage} [t]  {7in} \footnotesize%
+Notes: All bars are expressed in percentage change from the initial steady-state value.
+\end{minipage}
+\end{figure}
+
+To go deeper into the analysis, we further decompose each variety and competitiveness effects in its various dimensions, as reported in Figures \ref{fig:variety_air_decomp} (Variety effect) and \ref{fig:competitivity_air_decomp} (Competitiveness effect) for air transport.
+Figures \textbf{XXX} in \textbf{Appendix} report similar results for maritime transport.
+
+\begin{figure}[htbp]
+\caption{Variety effect (Air transport)}
+\label{fig:variety_air_decomp}
+\includegraphics[width=12cm, height=10cm]{VarietyEffect_air.pdf}
+%\begin{minipage} [t]  {7in} \footnotesize%
+%Notes: All bars are expressed in percentage change from the initial steady-state value.
+%\end{minipage}
+\end{figure}
+
+\begin{figure}[htbp]
+\caption{Competitiveness effect (Air transport)}
+\label{fig:competitivity_air_decomp}
+\includegraphics[width=12cm, height=10cm]{CompetEffect_air.pdf}
+%\begin{minipage} [t]  {7in} \footnotesize%
+%Notes: All bars are expressed in percentage change from the initial steady-state value.
+%\end{minipage}
+\end{figure}
+
+To understand the changes induced by the additive costs, let us start from reasoning in the benchmark model with only ad-valorem costs of \cite{melitz}.
+In this case, the reduction in transport costs induces an increase in selection on the domestic market ($\phi^\ast$ increases, Figure \textbf{XXXX STOP HERE }), but more domestic firms can now export ($\phi_x^\ast$ decreases).
+Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases).
+As a whole, the total number of varieties accessible to the consumer reduces ($M_T$ decreases slightly).
+Everything else equal, this reduces welfare.
+This is counteracted by a price competitiveness effect.
+Due to the tougher selection on the domestic market, the less productive firms no longer enter.
+As a result, the average productivity on the domestic market increases ($\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
+This price competitiveness effect pushes welfare upwards;
+
+Reduction in additive costs: More negative variety effect
 

### 2021-08-04 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

@@
 \begin{center}
-\includegraphics[width=8cm, height=5cm]{Figure1_share_of_additive_in_totalTC.pdf}
+ \includegraphics[width=12cm, height=7.5cm]{Figure1_share_of_additive_in_totalTC.jpg}
 %{\footnotesize {OECD data} }

### 2021-08-05 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte modèle encore

@@
 \begin{eqnarray*}
-\max_{p^x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
-\text{s.t. }&&c^x(\varphi) = \left(\frac{p^x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
+\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
+\text{s.t. }&&c^x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
 \end{eqnarray*}
 
-\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p^x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
+\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
 Solving for this yields:
-$$p^x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right]$$
+\begin{equation}
+p_x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right] \label{eq:px}
+\end{equation}
 
@@
 
-To understand the changes induced by the additive costs, let us start from reasoning in the benchmark model with only ad-valorem costs of \cite{melitz}.
-In this case, the reduction in transport costs induces an increase in selection on the domestic market ($\phi^\ast$ increases, Figure \textbf{XXXX STOP HERE }), but more domestic firms can now export ($\phi_x^\ast$ decreases).
-Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases).
-As a whole, the total number of varieties accessible to the consumer reduces ($M_T$ decreases slightly).
-Everything else equal, this reduces welfare.
-This is counteracted by a price competitiveness effect.
+To understand the changes induced by the additive costs, consider first the benchmark model with only ad-valorem costs, as in \cite{melitz}.
+In this case, the reduction in transport costs induces an increase in selection on the domestic market, but more domestic firms can now export (Figure \ref{fig:competitivity_air_decomp} for Air transport).
+Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases, see Figure \ref{fig:variety_air_decomp}).
+As a whole, the total number of varieties accessible to the consumer reduces (Figure \ref{fig:variety_air_decomp}, $M_T$ decreases slightly).
+Everything else equal, this reduces welfare (Figure \ref{fig:welfare_decomp}).
+Yet, this detrimental effect is counteracted by a price competitiveness effect.
 Due to the tougher selection on the domestic market, the less productive firms no longer enter.
-As a result, the average productivity on the domestic market increases ($\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
-This price competitiveness effect pushes welfare upwards;
-
-Reduction in additive costs: More negative variety effect
+As a result, the average productivity on the domestic market increases (Figure \ref{fig:competitivity_air_decomp}, $\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
+As reported in Figure \ref{fig:welfare_decomp}, the strength of the price competitiveness effect dominates, inducing a increase in welfare following the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.
+
+What does it change when the reduction in international transport costs also occurs through a reduction in its additive component?
+As reported in Figure \ref{fig:welfare_decomp} (right grey bars), trade liberalisation still induces welfare gains, and even larger than in the pure ad-valorem case.
+This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though.
+\footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness.
+Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization.
+Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the effects are sizeable.
+
+Through which mechanism does it occur?
+As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (this can be shown from Equation (\ref{eq:px})).
+Within the exporters category, this favors the more productive firms, whose market share, e.g.
+revenues increases relatively more, as can be shown from Equation (\ref{eq:rx}).
+Everything else equal, this pushes the average productivity on the export market upwards.
+In contrast to ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
+At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin, at the heart of the \cite{melitz} model: As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition: Everything else equal,
 

### 2021-08-05 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
texte modèle encore

@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs.
-In columns (1), (3), (5) and (7) only ad-valorem costs are considered.
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model oad-valorem costs only.
 In Columns (2), (4), (6) and (8), we model both ad-valorem and additive costs.
 $t$ is calibrated so that the ratio $t/\widetilde{p}_x$ from the model matches its empirical target, with $widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
-We theoretically define the export fas price through: $p_x = \tau p^{\text{fas}} +t$}}}
+We theoretically define the export fas price through: $p_x = \tau p^{\text{fas}} +t$.
+TC = Transport Costs, expressed in percentage of the average fas export price.}}}
 \end{table}
@@
 \begin{center}
-\begin{tabular}{l|c|c|c|c||c|c|c|c}
+\begin{tabular}{l||c|c|c|c||c|c|c|c}
 \hline
 Mode & \multicolumn{4}{|c||}{Air} & \multicolumn{4}{c}{Vessel} \\ \hline  
-Fixed export cost: &  \multicolumn{2}{|c|}{Constant $f_x$} &  \multicolumn{2}{|c||}{Lower $f_x$} & \multicolumn{2}{|c|}{Constant $f_x$} & \multicolumn{2}{|c}{Lower $f_x$} \\ \hline
+$f_x$ final value &  \multicolumn{2}{|c|}{As in 1974} &  \multicolumn{2}{|c||}{Lower} & \multicolumn{2}{|c|}{As in 1974} & \multicolumn{2}{|c}{Lower } \\ \hline
+& (A) & (B) & (C) & (D) & (E) & (F) &(G) & (H) \\ 
 Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline  
-%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
-%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
+$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
+$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
 $\Delta TC$ (in pp) &-3.4 &-3.4 &-3.4 &-3.4 &-6.4 &-6.4 &-6.4 &-6.4  \\
@@
 \end{center}
-{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs. The reduction in transport costs is expressed in percentage points. }}}
+{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price.
+The reduction in transport costs is expressed in percentage points.
+}}}
 \end{table}
@@
 In this first exercise, we contrast the effects of a reduction of international transport costs depending on whether this occurs through a reduction in ad-valorem costs only, or a combined reduction in both the additive and the multiplicative components.
-In Table \ref{tab:resultats_modele}, this relies on the results reported in Column (1) versus (2) for Air, Column (5) versus (6) for Vessel.\medskip
+In Table \ref{tab:resultats_modele}, this relies on the results reported in Column (A) versus (B) for Air, Column (E) versus (F) for Vessel.\medskip
 
@@
 revenues increases relatively more, as can be shown from Equation (\ref{eq:rx}).
-Everything else equal, this pushes the average productivity on the export market upwards.
-In contrast to ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
-At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin, at the heart of the \cite{melitz} model: As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition: Everything else equal,
+Everything else equal, this pushes the average productivity on the export market upwards.\footnote{This mechanism indirectly appears Figure \ref{fig:competitivity_air_decomp}, panel.
+Given our exercise that assumes a combined decrease in both the additive and the ad-valorem costs, at the general equilibrium, the threshold productivity on the export market $\widetilde{\varphi}^\ast_x$ still decreases, but to a lesser extent when the trade liberalization partly occurs through a reduction in additive costs.
+} In contrast to ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
+At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model: As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition.
+Everything else equal, the threshold productivity on the domestic market increases, as well as the average productivity.
+Accordingly, the competitiveness effect increases by more when the trade liberalization involves a reduction in the additive component.
+
+The more negative variety effect (Figure \ref{fig:welfare_decomp}) can be understood as a consequence of the above mechanism.
+Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a lower mass of exporters (Figure \ref{fig:variety_air_decomp}).
+As the selection on the domestic market is also strengthened when trade liberalization involves a reduction in additive costs, the mass of active domestic firms reduces by a larger amount.
+From the consumer side, this induces a more negative variety effect that is welfare-detrimental everything else equal.
+As for the pure-iceberg case though, the strength of the competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.
+
+\textbf{Faire l'exercice partant d'un monde avec les deux ad-valorem + iceberg, regarder ce qui se passe quand la baisse des coouts totaux est soit soit uniquement iceberg, soit uniquement ad-valorem?
+Plus proche de l'exercice de Sorensen.
+Plus clair sur les effets?
+}
+
+\subsubsection{Reducing fixed export costs}
+
+The above results underline the importance of the additive component of international trade costs.
+Trade liberalization still has positive welfare effects, and even larger when it relies on a reduction in additive costs.
+One might question the robustness of this last result to the type of trade liberalization considered.
+Indeed, the reduction in the fixed export costs seems empirically relevant to account for the larger share of exporting firms nowadays relative to the 1970s (Based on XXX, from 19\% in XXX to around 35\% in XXX).
+As underlined by \cite{melitz} \textbf{l'a-t-il MONTRE VRAIMENT ANALYTIQUEMENT?
+A VERIFIER}, the reduction of the fixed export cost is also welfare-enhancing, in presence of yet only ad-valorem variable trade costs.
+A natural question then emerges: To what extent is this result modified in presence of additive costs?
+Specifically, are the welfare gains of a fixed export costs decrease larger in presence of additive costs?
+
+The answer can be found in Table \ref{tab:resultats_modele}, by comparing Columns (B) to (D) for Air, (F) to (H) for Vessel, the difference between these two columns (for each transport mode) being the lower value of the fixed export cost.
+Here again, the welfare gains of trade liberalization (under the form of a lower fixed export cost) are larger when variable trade costs involve a part of additive costs: If the effect are quantitatively modest in maritime transport (by 0.75 percentage points, from 3.3\% to 4.05\%), the decrease in the fixed export cost is more sizeable when variable costs are lower, as in air transport where the welfare gains almost double (from 1.82\% to 3.25\%).
+
+\begin{table}[htbp]
+  \centering
+  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele_decompwelfare}
+\begin{center}
+\begin{tabular}{l||c|c|c|c||c|c|c|c}
+\hline
+Mode & \multicolumn{4}{|c||}{Air} & \multicolumn{4}{c}{Vessel} \\ \hline
+$f_x$ final value &  \multicolumn{2}{|c|}{As in 1974} &  \multicolumn{2}{|c||}{Lower} & \multicolumn{2}{|c|}{As in 1974} & \multicolumn{2}{|c}{Lower } \\ \hline
+& (A) & (B) & (C) & (D) & (E) & (F) &(G) & (H) \\
+Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline
+%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
+%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
+%$\Delta TC$ (in pp) &-3.4 &-3.4 &-3.4 &-3.4 &-6.4 &-6.4 &-6.4 &-6.4  \\
+%$\Delta \tau$ (in pp) & -3.4	&-1.4&	-3.40&	-1.40	&-6.4&	-3.1&	-6.4&	-3.1  \\
+%$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3&	0&	-3.3  \\ \hline
+%$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\ \hline
+Variety effect (\%)&	-0.827	&0.747	&-2.670&	-1.132	&-1.323	&-0.126&	-3.061	& -1.857 \\
+Price effect (\%)&	2.084&	1.455	&5.202&	4.444&	3.568	&3.001&	6.829&	6.059 \\ 
+\hline
+\end{tabular}
+\end{center}
+%{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: The reduction in transport costs is expressed in percentage points. }}}
+\end{table}
 

### 2021-08-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Sure la table 1

--- 2021-08-04 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 On the contrary, they represent a sizable and structural dimension of international transport costs.
-If the share of additive costs in total costs roughly amounts to 48\% (42\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
+If the share of additive costs in total costs roughly amounts to 36\% (39\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
+The share of additiive costs air transport costs has been steeply declining since the end of the 1980s.
 In the next section, we further investigate the importance of the (varying pattern of) additive costs, as regards the determinants of transport costs time trends.
@@
 
-The two productivity thresholds are linked according to:
+The two productivity thresholds are accordingly linked according to:
 \begin{equation}
@@
 Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as:
-$$\tau+ t\varphi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$$
-
-Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
+$$\tau+ t\phi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$$
+
+Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\phi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
 We check this condition holds numerically in solving the model.\textbf{A FAIRE ???
@@
 \paragraph{Aggregation} The general equilibrium is characterized by a mass $M$ of producing firms in each country associated with a distribution $\mu(\varphi)$ of productivity levels of a subset of $\left(0,\infty\right)$.
-Given the productivity threshold $\varphi^\ast$, $\mu(\varphi)$ is equal to the conditional distribution of $g(\varphi)$ on the subset $\left[\varphi^\ast, \infty\right)$:
+Given the productivity threshold $\phi^\ast$, $\mu(\varphi)$ is equal to the conditional distribution of $g(\varphi)$ on the subset $\left[\phi^\ast, \infty\right)$:
 \begin{eqnarray}
@@
 \noindent with $\left(1-G(\varphi^\ast)\right)$ the probability of successful entry. In the open-economy setting, $\bar{\pi}$ can be expressed as:
-$$\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
+$$\bar{pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
 \noindent with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to:
@@
 From this, we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
-Specifically, we will study the consequences of additive costs in terms of welfare.
-As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, the price level $P$ being given by Equation (\ref{eq:CPI_vf}).
+Specifically, we will study the consequences of additive costs in terms of welfare, which can be shown to be the inverse of the price level as standard in this literature (the price level $P$ being given by Equation (\ref{eq:CPI_vf})).
 
@@
 $\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3&	0&	-3.3  \\ \hline
-$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\
+$\Delta \mathcal{U}$ & 0.0202&	0.0295	&0.0362&	0.0527	&0.035&	0.052	&0.047&	0.064 \\
+$\Delta \mathcal{U}/\mathcal{U}$ (in \%) & 1.2319&	1.8187&	2.2073	&3.2477	&2.179	&3.295&	2.872&	4.052   \\
 \hline
@@
 In this first exercise, we contrast the effects of a reduction of international transport costs depending on whether this occurs through a reduction in ad-valorem costs only, or a combined reduction in both the additive and the multiplicative components.
-In Table \ref{tab:resultats_modele}, this relies on the results reported in Column (1) versus (2) for Air, Column (5) versus (6) for Vessel.\medskip
-
-Whatever the transport mode (and the associated magnitude of the transport costs decrease), welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
-The effect is substantial: The reduction of 3.4 percentage points of Air transport costs generates a welfare gain higher by 0.58 percentage point when such a reduction occurs through the additive component (accounting for almost 60\% of the total cost decrease), rather than when it realizes through a decrease in iceberg costs only.
-In maritime transport, where the decrease in total transport costs is larger (-6.4 percentage points), the extra welfare gain amounts to 1.12 percentage points when the transport cost reduction occurs through the additive component (by 50\%) rather than through the iceberg component only.\smallskip
-
-Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
-Considering the inverse of Equation (\ref{eq:CPI_vf}), it is straightforward to get the welfare change as:
-\begin{equation}
-\frac{d\mathcal{W}}{\mathcal{W}_0} = \underbrace{\frac{1}{\sigma-1} \frac{d M_T}{M_{T0}}}_{\text{Variety effect}} +  \underbrace{\frac{d\widetilde{\varphi}_T}{\widetilde{\varphi}_{T0}}}_{\text{Price effect}} \label{eq:decompWelfare}
-\end{equation}
-
-Figure \ref{fig:welfare_decomp} displays the welfare change and its decomposition between its two components following the reduction in international transport costs, for the decrease estimated in air transport (left column, Panels (1), (1-a) and (1-b)) and in maritime transport (right column, Panels (2), (2-a) and (2-b)).
-All are expressed in percentage change from the initial steady-state value.
-In each panel, the blue left bar displays the case where the trade liberalization occurs through a reduction in ad-valorem costs; the grey right bar reporting the case where the (same) reduction in international transport costs occurs through a combined reduction in both the ad-valorem and the additive components, based on our estimated results in 1974 and 2019 (for each transport mode), as in Table \ref{tab:calib_TC}).
-
-\begin{figure}[htbp]
-\caption{Welfare decomposition}
-\label{fig:welfare_decomp}
-\includegraphics[width=15cm, height=12cm]{Welfare_change_bothmodes.pdf}
-\begin{minipage} [t]  {7in} \footnotesize%
-Notes: All bars are expressed in percentage change from the initial steady-state value.
-\end{minipage}
-\end{figure}
-
-To go deeper into the analysis, we further decompose each variety and competitiveness effects in its various dimensions, as reported in Figures \ref{fig:variety_air_decomp} (Variety effect) and \ref{fig:competitivity_air_decomp} (Competitiveness effect) for air transport.
-Figures \textbf{XXX} in \textbf{Appendix} report similar results for maritime transport.
-
-\begin{figure}[htbp]
-\caption{Variety effect (Air transport)}
-\label{fig:variety_air_decomp}
-\includegraphics[width=12cm, height=10cm]{VarietyEffect_air.pdf}
-%\begin{minipage} [t]  {7in} \footnotesize%
-%Notes: All bars are expressed in percentage change from the initial steady-state value.
-%\end{minipage}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Competitiveness effect (Air transport)}
-\label{fig:competitivity_air_decomp}
-\includegraphics[width=12cm, height=10cm]{CompetEffect_air.pdf}
-%\begin{minipage} [t]  {7in} \footnotesize%
-%Notes: All bars are expressed in percentage change from the initial steady-state value.
-%\end{minipage}
-\end{figure}
-
-To understand the changes induced by the additive costs, let us start from reasoning in the benchmark model with only ad-valorem costs of \cite{melitz}.
-In this case, the reduction in transport costs induces an increase in selection on the domestic market ($\phi^\ast$ increases, Figure \textbf{XXXX STOP HERE }), but more domestic firms can now export ($\phi_x^\ast$ decreases).
-Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases).
-As a whole, the total number of varieties accessible to the consumer reduces ($M_T$ decreases slightly).
-Everything else equal, this reduces welfare.
-This is counteracted by a price competitiveness effect.
-Due to the tougher selection on the domestic market, the less productive firms no longer enter.
-As a result, the average productivity on the domestic market increases ($\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
-This price competitiveness effect pushes welfare upwards;
-
-Reduction in additive costs: More negative variety effect
+Accordingly, we focus on the results reported in Table \ref{tab:resultats_modele}, Column (1) versus (2) for Air, Column (5) versus (6) for Vessel.
+
+A first conclusion emerges: Whatever the transport mode (and the magnitude of the transport costs decrease accordingly), welfare gains are higher when this reduction is made (partly) through a reduction of additive costs.
+The effect is substantial: In Air transport, the reduction of 3.4 percentage points of transport costs generates a welfare gain higher by 0.58 percentage point when almost 60\% of such a reduction occurs through the additive component, rather than when it realizes through a decrease of iceberg costs only.
+In maritime transport, where the decrease in total transport costs is larger (by 6.4 percentage points), the extra welfare gain amounts to 1.12 percentage points when 50\% of the transport cost reduction occurs through the additive component.
 

### 2021-08-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

--- 2021-08-05 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 \begin{eqnarray*}
-\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
-\text{s.t. }&&c^x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
+\max_{p^x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
+\text{s.t. }&&c^x(\varphi) = \left(\frac{p^x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
 \end{eqnarray*}
 
-\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
+\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p^x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
 Solving for this yields:
-\begin{equation}
-p_x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right] \label{eq:px}
-\end{equation}
+$$p^x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right]$$
 
@@
 
-To understand the changes induced by the additive costs, consider first the benchmark model with only ad-valorem costs, as in \cite{melitz}.
-In this case, the reduction in transport costs induces an increase in selection on the domestic market, but more domestic firms can now export (Figure \ref{fig:competitivity_air_decomp} for Air transport).
-Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases, see Figure \ref{fig:variety_air_decomp}).
-As a whole, the total number of varieties accessible to the consumer reduces (Figure \ref{fig:variety_air_decomp}, $M_T$ decreases slightly).
-Everything else equal, this reduces welfare (Figure \ref{fig:welfare_decomp}).
-Yet, this detrimental effect is counteracted by a price competitiveness effect.
+To understand the changes induced by the additive costs, let us start from reasoning in the benchmark model with only ad-valorem costs of \cite{melitz}.
+In this case, the reduction in transport costs induces an increase in selection on the domestic market ($\phi^\ast$ increases, Figure \textbf{XXXX STOP HERE }), but more domestic firms can now export ($\phi_x^\ast$ decreases).
+Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases).
+As a whole, the total number of varieties accessible to the consumer reduces ($M_T$ decreases slightly).
+Everything else equal, this reduces welfare.
+This is counteracted by a price competitiveness effect.
 Due to the tougher selection on the domestic market, the less productive firms no longer enter.
-As a result, the average productivity on the domestic market increases (Figure \ref{fig:competitivity_air_decomp}, $\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
-As reported in Figure \ref{fig:welfare_decomp}, the strength of the price competitiveness effect dominates, inducing a increase in welfare following the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.
-
-What does it change when the reduction in international transport costs also occurs through a reduction in its additive component?
-As reported in Figure \ref{fig:welfare_decomp} (right grey bars), trade liberalisation still induces welfare gains, and even larger than in the pure ad-valorem case.
-This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though.
-\footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness.
-Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization.
-Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the effects are sizeable.
-
-Through which mechanism does it occur?
-As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (this can be shown from Equation (\ref{eq:px})).
-Within the exporters category, this favors the more productive firms, whose market share, e.g.
-revenues increases relatively more, as can be shown from Equation (\ref{eq:rx}).
-Everything else equal, this pushes the average productivity on the export market upwards.
-In contrast to ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
-At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin, at the heart of the \cite{melitz} model: As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition: Everything else equal,
+As a result, the average productivity on the domestic market increases ($\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
+This price competitiveness effect pushes welfare upwards;
+
+Reduction in additive costs: More negative variety effect
 

### 2021-08-07 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte modèle

@@
 $\Delta \tau$ (in pp) & -3.4	&-1.4&	-3.40&	-1.40	&-6.4&	-3.1&	-6.4&	-3.1  \\
-$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3&	0&	-3.3  \\ \hline
-$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\
-\hline
+$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3& 0&	-3.3 \\
+$\Delta \mathcal{W}$ & 0,0202 &  0,0386 & 0,0362 & 0,0540 & 0,0354 & 0,0535 & 0,0467 & 	0.064 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	2,393 &	2.207	&3,348	&2.179	&3,382 &	2.872&	4.00   \\\hline
 \end{tabular}
@@
 
-\subsubsection{Reducing international transport costs}
-
-In this first exercise, we contrast the effects of a reduction of international transport costs depending on whether this occurs through a reduction in ad-valorem costs only, or a combined reduction in both the additive and the multiplicative components.
-In Table \ref{tab:resultats_modele}, this relies on the results reported in Column (A) versus (B) for Air, Column (E) versus (F) for Vessel.\medskip
-
-Whatever the transport mode (and the associated magnitude of the transport costs decrease), welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
+\paragraph{Reduction in international variable transport costs}
+
+We first consider the effects of a reduction of variable international transport costs.
+Specifically, we contrast two different environments: In the first scenario ``\`{a} la M\'{e}litz'', only ad-valorem transport costs apply, which consequently reduce.
+In the second scenario so called ``\`{a} la Hummels'', international transport costs are made of both additive and ad-valorem costs.
+The overall transport costs decreases then comes from a reduction in both components.
+The size of the overall transport costs, as well as their reduction, is identical in both scenarii.
+We calibrate them in our empirical estimates, considering 1974 and 2019 as initial and final values.
+They are transport-mode specific.
+In the scenario ``\`{a} la Hummels'', the initial and finale values of each additive and ad-valorem component is also based on our empirical estimates.
+Importantly, the size of the overall transport costs decrease is kep identical under both scenarii.
+What differs is the composition of this decrease.
+The two scenarii for each transport mode are reported in Table \ref{tab:resultats_modele}.\medskip
+
+A first conclusion emerges.
+Whatever the transport mode (and the associated magnitude of the transport costs decrease), trade liberalization induces welfare gains.
+Second, the welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
 The effect is substantial: The reduction of 3.4 percentage points of Air transport costs generates a welfare gain higher by 0.58 percentage point when such a reduction occurs through the additive component (accounting for almost 60\% of the total cost decrease), rather than when it realizes through a decrease in iceberg costs only.
@@
 
-\subsubsection{Reducing fixed export costs}
+\paragraph{Reduction of the fixed export cost}
 
@@
 
-The answer can be found in Table \ref{tab:resultats_modele}, by comparing Columns (B) to (D) for Air, (F) to (H) for Vessel, the difference between these two columns (for each transport mode) being the lower value of the fixed export cost.
-Here again, the welfare gains of trade liberalization (under the form of a lower fixed export cost) are larger when variable trade costs involve a part of additive costs: If the effect are quantitatively modest in maritime transport (by 0.75 percentage points, from 3.3\% to 4.05\%), the decrease in the fixed export cost is more sizeable when variable costs are lower, as in air transport where the welfare gains almost double (from 1.82\% to 3.25\%).
+The answer can be found in Table \ref{tab:resultats_modele_fx}, by comparing Columns (B) to (D) for Air, (F) to (H) for Vessel, the difference between these two columns (for each transport mode) being the lower value of the fixed export cost.
+Here again, the welfare gains of trade liberalization (under the form of a lower fixed export cost) are larger when variable trade costs involve a part of additive costs.
+\textbf{A FINIR ET A CREUSER}
 
@@
   \centering
-  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele_decompwelfare}
+  \caption{Trade liberalization (2): Reduction of the fixed export cost} \label{tab:resultats_modele_fx}
 \begin{center}
@@
 %$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\ \hline
-Variety effect (\%)&	-0.827	&0.747	&-2.670&	-1.132	&-1.323	&-0.126&	-3.061	& -1.857 \\
-Price effect (\%)&	2.084&	1.455	&5.202&	4.444&	3.568	&3.001&	6.829&	6.059 \\ 
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	2.393 &	2.207	&3.348	&2.179	&3.382&	2.872&	4.00   \\ \hline
+Variety effect (\%)&	-0.827 	&-2.599	&0.747
+&-1.038	&-1.323	& -3.123&	-0.126	&-1.902 \\
+Price effect (\%)&	2.084 &5.201	& 1.455 &	4.444&	3.568	& 6.829&	 3.001&	6.059 \\
 \hline

### 2021-08-08 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
graph 2 and graph 3

@@
 \begin{center}
-\includegraphics[width=14cm, height=7cm]{Figure2_Trend_of_totalTC_bymode.pdf}
+\includegraphics[width=14cm, height=7cm]{Figure2_Trend_of_totalTC_bymode.jpg}
 \end{center}
@@
 %{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.0in, height=2.5in]{Etude_beta_pondere_air.pdf}
-& \includegraphics[width=3.0in,height=2.5in]{Etude_beta_pondere_ves.pdf} \\
+\includegraphics[width=3.0in, height=2.5in]{"Etude_beta_pondere_(a) Air transport".jpg}
+& \includegraphics[width=3.0in,height=2.5in]{"Etude_beta_pondere_(b) Vessel transport".jpg} \\
 \multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\

### 2021-08-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Changement .tex modèle

--- 2021-08-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 Our empirical results point to two important conclusions: First, additive costs represent a non-negligible part of overall international transport costs.
-Second, total transport costs are strongly reduced over 1974-2019, and the decrease is particularly marked in the additive component in air transport.
+Second, total transport costs have strongly reduced over 1974-2019, and the decrease is particularly marked in the additive component in air transport.
 We exploit the theoretical implications of these empirical findings in two different ways.
 
-First, we run a comparative statics analysis following a reduction in transport costs based on our estimates in 1974 and 2019, by transport mode.
-Specifically, we contrast two cases: In the first scenario, the reduction in total transport costs is solely attributable to the ad-valorem component (e.g., contrasting Table \ref{tab:calib_TC}, Column (3) to (1) for air transport, and Column (7) to (5) for maritime transport); in the second scenario, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
-In this scenario, we contrast Table \ref{tab:calib_TC}, Column (4) to (2) (Air) and Column (8) to (6) (Vessel).
+First, we run a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
+Specifically, we contrast two cases: In the first scenario \textit{\grave{a} la M\'{e}litz}, we assume that only ad-valorem costs apply.
+We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
+In the second scenario \textit{\grave{a} la Hummels}, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
+The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
 Importantly, we run this comparative statics analysis \textit{ceteris paribus}, notably with the same value of the export fixed cost $f_x$ as in the initial equilibrium (see Table \ref{tab:calib_horsTC}.
@@
 Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
-For each transport mode, the first two columns report the results of reducing transport costs only, maintaining the same value of $f_x$ as in the initial equilibrium (Columns (1)-(2) for Air, (5)-(6) for Vessel), contrasting the case of only ad-valorem costs (Columns (1), (5)) and in presence of transport costs (Columns (2), (6)).
-In Columns (3)-(4) for Air and (7)-(8), we report the results when both the variable transport costs and the fixed export costs are lower in the final steady state (2019) relative to the initial one (1974).
+For each transport mode, the first two columns report the results of reducing transport costs only, maintaining the same value of $f_x$ as in the initial equilibrium (Columns (A)-(B) for Air, (E)-(F) for Vessel), contrasting the case of only ad-valorem costs (Columns (A), (D)) and in presence of additive costs (Columns (B), (F)).
+Columns (C)-(D) for Air and (G)-(H) for Vessel report the results when both the variable transport costs and the fixed export costs are lower in the final steady state (2019) relative to the initial one (1974).
 
@@
 $\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3& 0&	-3.3 \\ \hline
-$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\
+$\Delta \mathcal{W}$ & 0.0202 &  0.0386 & 0.0362 & 0.0540 & 0.0354 & 0.0535 & 0.0467 & 	0.064 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	2.393 &	2.207	&3.348	&2.179	&3.382 &	2.872&	4.00   \\\hline
+Variety effect (\%)&	-0.827 	&-2.599	&0.747
+&-1.038	&-1.323	& -3.123&	-0.126	&-1.902 \\
+Price effect (\%)&	2.084 &5.201	& 1.455 &	4.444&	3.568	& 6.829&	 3.001&	6.059 \\
 \hline
@@
 The reduction in transport costs is expressed in percentage points.
+In the world including additive costs, we simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.
 }}}
@@
 
-\subsubsection{Reducing international transport costs}
-
-In this first exercise, we contrast the effects of a reduction of international transport costs depending on whether this occurs through a reduction in ad-valorem costs only, or a combined reduction in both the additive and the multiplicative components.
-In Table \ref{tab:resultats_modele}, this relies on the results reported in Column (A) versus (B) for Air, Column (E) versus (F) for Vessel.\medskip
-
-Whatever the transport mode (and the associated magnitude of the transport costs decrease), welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
-The effect is substantial: The reduction of 3.4 percentage points of Air transport costs generates a welfare gain higher by 0.58 percentage point when such a reduction occurs through the additive component (accounting for almost 60\% of the total cost decrease), rather than when it realizes through a decrease in iceberg costs only.
-In maritime transport, where the decrease in total transport costs is larger (-6.4 percentage points), the extra welfare gain amounts to 1.12 percentage points when the transport cost reduction occurs through the additive component (by 50\%) rather than through the iceberg component only.\smallskip
+\paragraph{Reduction in international variable transport costs} Based on our empirical estimates, we simulate a reduction in total variable transport costs by 3.4\% in Air transport and 6.4\% in maritime transport, in two different worlds: In a world with only ad-valorem costs (Columns (A) for Air and (D) for Vessel), versus a world with both iceberg and additive transport costs, and where the overall reduction costs comes from both components (Columns (B) for Air and (E) for Vessel).
+
+A first conclusion emerges.
+Whatever the transport mode (and the associated magnitude of the transport costs decrease), trade liberalization induces welfare gains.
+Second, the welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
+The effect is substantial: The reduction variable transport costs generates a relative welfare gain higher by 1.1 percentage point in Air, 1.2 percentage point in Vessel, when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only.\smallskip
 
@@
 
-Figure \ref{fig:welfare_decomp} displays the welfare change and its decomposition between its two components following the reduction in international transport costs, for the decrease estimated in air transport (left column, Panels (1), (1-a) and (1-b)) and in maritime transport (right column, Panels (2), (2-a) and (2-b)).
+We report the decomposition of the welfare gains in their two components (under the various scenarii) in the last two lines of Table \ref{tab:resultats_modele}, as well as visually through Figure \ref{fig:welfare_decomp}.
+It displays the welfare change and its decomposition between its two components following the reduction in international transport costs, for the decrease estimated in air transport (left column, Panels (1), (1-a) and (1-b)) and in maritime transport (right column, Panels (2), (2-a) and (2-b)).
 All are expressed in percentage change from the initial steady-state value.
@@
 
-\subsubsection{Reducing fixed export costs}
-
-The above results underline the importance of the additive component of international trade costs.
-Trade liberalization still has positive welfare effects, and even larger when it relies on a reduction in additive costs.
-One might question the robustness of this last result to the type of trade liberalization considered.
+\paragraph{Reduction of the fixed export cost} The above results underline the importance of the additive component of international trade costs.
+When trade liberalization occurs through a reduction in variable costs, not only the welfare-enhancing effects still apply, but they are even larger.
+As mentioned by Mélitz final proofs, the reduction in variable costs is not the only form of trade liberalization, which can typically manifest through a reduction in export fixed costs.
+One might then wonder how additive costs do affect the welfare gains of trade liberalization of this form.
 Indeed, the reduction in the fixed export costs seems empirically relevant to account for the larger share of exporting firms nowadays relative to the 1970s (Based on XXX, from 19\% in XXX to around 35\% in XXX).
-As underlined by \cite{melitz} \textbf{l'a-t-il MONTRE VRAIMENT ANALYTIQUEMENT?
+As underlined by \cite{melitz} \textbf{JE CROIS QUE C'est DIT DANS FINAL PROOFS, PAS ECTRA.
+l'a-t-il MONTRE VRAIMENT ANALYTIQUEMENT?
 A VERIFIER}, the reduction of the fixed export cost is also welfare-enhancing, in presence of yet only ad-valorem variable trade costs.
@@
 
-The answer can be found in Table \ref{tab:resultats_modele}, by comparing Columns (B) to (D) for Air, (F) to (H) for Vessel, the difference between these two columns (for each transport mode) being the lower value of the fixed export cost.
-Here again, the welfare gains of trade liberalization (under the form of a lower fixed export cost) are larger when variable trade costs involve a part of additive costs: If the effect are quantitatively modest in maritime transport (by 0.75 percentage points, from 3.3\% to 4.05\%), the decrease in the fixed export cost is more sizeable when variable costs are lower, as in air transport where the welfare gains almost double (from 1.82\% to 3.25\%).
-
-\begin{table}[htbp]
-  \centering
-  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele_decompwelfare}
-\begin{center}
-\begin{tabular}{l||c|c|c|c||c|c|c|c}
-\hline
-Mode & \multicolumn{4}{|c||}{Air} & \multicolumn{4}{c}{Vessel} \\ \hline
-$f_x$ final value &  \multicolumn{2}{|c|}{As in 1974} &  \multicolumn{2}{|c||}{Lower} & \multicolumn{2}{|c|}{As in 1974} & \multicolumn{2}{|c}{Lower } \\ \hline
-& (A) & (B) & (C) & (D) & (E) & (F) &(G) & (H) \\
-Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline
-%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
-%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
-%$\Delta TC$ (in pp) &-3.4 &-3.4 &-3.4 &-3.4 &-6.4 &-6.4 &-6.4 &-6.4  \\
-%$\Delta \tau$ (in pp) & -3.4	&-1.4&	-3.40&	-1.40	&-6.4&	-3.1&	-6.4&	-3.1  \\
-%$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3&	0&	-3.3  \\ \hline
-%$\Delta \mathcal{W}$ & 0.0202&	0.0295	&0.036&	0.053	&0.035&	0.052	&0.047&	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	1.819&	2.207	&3.248	&2.179	&3.295&	2.872&	4.052   \\ \hline
-Variety effect (\%)&	-0.827	&0.747	&-2.670&	-1.132	&-1.323	&-0.126&	-3.061	& -1.857 \\
-Price effect (\%)&	2.084&	1.455	&5.202&	4.444&	3.568	&3.001&	6.829&	6.059 \\ 
-\hline
-\end{tabular}
-\end{center}
-%{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: The reduction in transport costs is expressed in percentage points. }}}
-\end{table}
+The answer can be found in Table \ref{tab:resultats_modele}.
+First, in accordance with \textbf{Mélitz}, the reduction in the fixed export costs adds an extra source of welfare gains.
+In the pure iceberg world, welfare gains amount to 2.2\% when both fixed and variable costs reduce (Column (C) for Air), vs a 1.23\% increase when only ad-valorem costs decrease (Column (A) for Air, a similar result applying for Vessel).
+The larger welfare gains under a combined reduction in the variable and fixed export costs remains in the world with additive costs (Columns (D) vs (B) for Air, (H) vs (E) for Vessel).
+Further, in both worlds (with or without additive costs), the reduction in the fixed export cost induces an extra welfare gains by around 1 percentage point in Air transport, 1.2 percentage point in maritime transport.
+Put it differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
+In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when the reduction in the variable part implies a reduction in the additive component, as observed empirically.
+
+In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results point out the importance of taking into the additive components of international trade costs.
+On the positive side, we show that additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods.
+This has important consequences in normative terms.
+Neglecting the role of additive costs in the theoretical models of trade leads to a substantial underestimation of the welfare gains of trade liberalisation.
+Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined to the reduction in the fixed export costs, has induced an 3.5\% increase in welfare, much higher than if only iceberg costs were modeled (2.4\%).
+A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4\% (as compared to 3.38\% in a pure ad-valorem world).
 

### 2021-08-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Petites corrections d’accents

@@
 First, we run a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
-Specifically, we contrast two cases: In the first scenario \textit{\grave{a} la M\'{e}litz}, we assume that only ad-valorem costs apply.
+Specifically, we contrast two cases: In the first scenario \textit{à la Mélitz}, we assume that only ad-valorem costs apply.
 We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
-In the second scenario \textit{\grave{a} la Hummels}, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
+In the second scenario \textit{à la Hummels}, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
 The consequences of the reduction in international transport costs through both its additive and ad-valorem components.

### 2021-08-23 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Enfin le graphique 4 !
Cela a été compliqué, car il faut prendre en compte terme_Adoll et prix_fob pour faire le terme_A.
Figure 5 ! On les grignote...

@@
 \includegraphics[height=4in]
-{graph_composition_all.pdf}
+{graph_composition_all.jpg}
 \end{center}
@@
 \multicolumn{2}{c}{{\small In percent of the export price}} \\
-\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.pdf}
-& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.pdf} \\
+\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
+& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
 {\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
 \multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
-\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.pdf}
-& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.pdf} \\
+\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.jpg}
+& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.jpg} \\
 \end{tabular}

### 2021-08-25 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Effet sur le papier
Revision texte papier et answer R1

@@
 First, additive transport costs are quantitatively sizable.
-On average over 1974-2019, the additive cost is estimated to be \textbf{1.8\% and 2.9\% A AJUSTER} of the export price in air and vessel transport respectively (at the 3-digit classification level).
-This is slightly lower than our estimates for the ad-valorem component \textbf{(2.5\% and 3.2\% for air/vessel} respectively on average over the period).
-Put differently, additive costs represent between \textbf{one-third and almost one-half} of the overall transport costs.
+On average over 1974-2019, the additive cost is estimated to be 1.7\% and 2.8\% of the export price in air and vessel transport respectively (at the 3-digit classification level).
+This is slightly lower than our estimates for the ad-valorem component (2.4\% and 3\% for air/vessel respectively on average over the period).
+Put differently, additive costs represent on average between 30 and 45\% of the overall transport costs.
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
@@
 We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a \textbf{factor~2}.
+Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by a factor ranging between 1.5 and 1.8.
 Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip
 
-Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2013}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
+Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
 changes over time in the baskets of imported goods and/or the origin countries).} Specifically, trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
@@
 cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using transaction-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
+Using transaction-level trade data from Colombia, \cite{Lashkaripour-2017} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
 One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
@@
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of \textbf{2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.}
-This comparison of results suggests that between \textbf{a third and a half of the additive trade costs} are attributable to the transport sector (depending on the transport mode).
+While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
+This comparison of results suggests that between a quarter and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).%JH : these results should not change, they are for 2004.
 Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2019.
@@
 
-\textbf{Big Picture}.
-Our results have important implications for both international trade and international macroeconomics models.
+%\textbf{Big Picture}.
+Our results are robust to a range of robustness check, including different weighting schemes, the use of instrumental variables and more disaggregated data at the HS10 level.
+They have important implications for both international trade and international macroeconomics models.
 Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \cite{sorensen2014} extends \cite{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
@@
 The quantitative results we provide in the paper allow making one step in filling that gap.
-\textbf{To develop}.
-\medskip
-
-The paper is built as follows.
+
+More precisely, we use our estimates of multiplicative and additive costs for the years 1974 and 2019 to implement several comparative static exercises investigating the different welfare consequences of alterations to multiplicative and additive costs.
+For a given decrease in transport costs, welfare gains are around 50\% higher when the underlying framework allows for additive costs - which, in practice, concentrate the major part of transport costs decrease over the period.
+In addition, we also assess how the latter results are distorted by changes in sunk costs of exports (i.e., increase in the share of exporting firms).
+Welfare gains are proportionally amplified, with again an additional premium coming from the decrease of additive costs.
+
+%\textbf{To develop}. \medskip
+
+The remainder of the paper is built as follows.
 First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
@@
 Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
-Section \ref{sec:results_trends} is devoted to this analysis, while Section \ref{sec:robustness} checks the robustness of our results to the separability assumption and the weighting scheme.
-Section \ref{sec:conclu} concludes.
+Section \ref{sec:results_trends} is devoted to this analysis, while Section \ref{sec:robustness} provides several important robustness checks of our results.
+Section\ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs., and section \ref{sec:conclu} concludes.
 
@@
 
-\section{The role of additive cost: Theoretical insights}
+\section{The role of additive cost: Theoretical insights}\label{sec:theory}
 
@@
 $f$ such that all firms produce in autarky (i.e., such that $\varphi^{\ast,a} = \varphi_{\text{min}}$.
-$f_x$ endogenously revealed such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, based on \textbf{???}, following so the calibration method of \cite{ghironi} or \cite{Irrazabal_2015}.
+$f_x$ endogenously revealed such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%\footnote{Though we could not find exact numbers for the share of exporting firms in 1974, this value of 21\% seems reasonable considering values reported for the 1980s by \citet{Lincoln_McCallum2018}} following so the calibration method of \cite{ghironi} or \cite{Irrazabal_2015}.
 
@@
 
-Second, we take into account the fact that the share of exporting firms has substantially increased since the 21\% value estimated by \textbf{???} in \textbf{YEAR YYY}.
+Second, we take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported by \citet{Lincoln_McCallum2018} for 1987.
 According to \textbf{???}, this share roughly amounts to 35\% in \textbf{???}.
@@
 
-The results are reported in Table \ref{tab:resultats_modele}, for Air (Columns (1)-(4)) and Vessel (Columns (5)-(8)) respectively.
+The results are reported in Table \ref{tab:resultats_modele}, for Air (Columns (A)-(D)) and Vessel (Columns (E)-(H)) respectively.
 Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.

### 2021-08-25 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
presque rien,

@@
 %\textbf{Big Picture}.
-Our results are robust to a range of robustness check, including different weighting schemes, the use of instrumental variables and more disaggregated data at the HS10 level.
+Our results are robust to a range of robustness checks, including different weighting schemes, the use of instrumental variables and more disaggregated data at the HS10 level.
 They have important implications for both international trade and international macroeconomics models.
@@
 
+-------IL FAUT INTRODUIRE LES PRIX FOB. Sinon, les coûts additifs ne sont pas modifiés de l’inflation -----
+
 $$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$

### 2021-08-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
nouvelles modifs main text et appendix 1

@@
 \usepackage{amsmath}
+\usepackage[skip=0.5\baselineskip]{caption}
 %\usepackage[latin1]{inputenc}
@@
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One might be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity of goods.
+While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity of goods.
 This is due to the fact that in the Census data, information on the quantity of goods is not mode-specific.
 We yet provide a robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.
-\textbf{See Section XXX}.} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+\textbf{See Section XXX JH: COULD NOT FIND THESE ANYWHERE.
+HAS IT BEEN DONE?}.} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
@@
 This drives us to estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
+Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin.
 We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i,} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
 Notice that this can be done from 1989 onwards for which HS-10 data are available; \textit{ii,} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
-The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.
+The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2019.
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
 However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Appendix \ref{app:4digit}.
-\textbf{ONLINE?
-IN TEXT?
-ADD}
-Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin.
+\textbf{JH: WHERE ARE THE 3-10/4-5 ESTIMATES?
+NOT IN ANY APPENDICES.}
+
+% JH : reprendre XXX ICI
 
@@
 
-Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
-The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
-Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
-Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for air transport at the 3-digit level, removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
-Overall results (over the whole period) are not substantially affected if this year is included.
-This can be seen in the results detailed year by year in the Online Appendix, available on the authors' webpages.}
-
 \begin{table}[htbp]
  \centering
-  \footnotesize{
-  \caption{Transport costs estimates: summary \label{tab:summary_results}}
-  \begin{center}
-    \begin{tabular}{l|cc|cc}
-      \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
-    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
-    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
-    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
-    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
-    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
-    %Min.
-%value & 1.003 & 1.001 & 1.003 & 1.000 \\
-    %Max.
-%value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
-    \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
-   \textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} & & & & \\ \hline
-    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
-    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
-    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
-    %Min.
-%value & 1.001 & 1.000 & 1.000 & 1.000 \\
-    %Max.
-%value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
-    \textit{Additive term }& & & &   \\ \hline
-    \textit{In \% of the export price ($\widehat{t}/\widetilde{p}$, in \%)} &&&& \\ \hline
-        Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
-    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
-   \textit{In USD per kg traded ($\widehat{t}$)}&&&& \\ \hline
-    Mean & 0.09	&1.12	&0.08&	2.16 \\
-    Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
-    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
-    %Min.
-%value & 0.000 & 0.000 & 0.000 & 0.000 \\
-    %Max.
-%value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
-    \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
-    Transport costs ($p/\widetilde{p} -1$, in \%) & & & & \\ \hline
-    Mean & 5.3 & 5.0& 5.6&3.9 \\
-    Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
-    Export price ($\widetilde{p}$), in USD & & & & \\
-    Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
-    Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
-    \# obs.
-& 29279 & 28207 & 29317 & 27680 \\
-    \# origin country & 188 & 191 & 188 & 189 \\
-    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
-  \end{tabular}
-    \end{center}}
-\parbox[l]{12cm}{\footnotesize{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
-The additive term is expressed in fraction of fas price.
-For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
-1974, 1981, 1989, 2001, 2009, 2013.
-($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
-\end{table}%
-
-\begin{table}
+\caption{Transport costs estimates: summary} \vspace{5mm} \label{tab:summary_results}
 	\input{Table1.tex}
@@
 \end{table}
+
+Table \ref{tab:summary_results} reports a summary of our results.
+It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
+The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
+Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
+Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for air transport at the 3-digit level, removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
+Overall results (over the whole period) are not substantially affected if this year is included.
+This can be seen in the results detailed year by year in the Online Appendix, available on the authors' webpages.}
+
+%\begin{table}[htbp]
+ % \centering
+ % \footnotesize{
+ % \caption{Transport costs estimates: summary \label{tab:summary_results}}
+ % \begin{center}
+  %  \begin{tabular}{l|cc|cc}
+  %    \hline \hline
+  %  \multicolumn{5}{c}{Mean value over 1974-2013}   \\
+  %  \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
+   % Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
+   % \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
+  %  Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
+  %  Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
+    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
+    %Min.
+%value & 1.003 & 1.001 & 1.003 & 1.000 \\
+    %Max.
+%value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
+ %   \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
+ %  \textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} & & & & \\ \hline
+ %   Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
+  %  Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
+    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
+    %Min.
+%value & 1.001 & 1.000 & 1.000 & 1.000 \\
+    %Max.
+%value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
+  %  \textit{Additive term }& & & &   \\ \hline
+  %  \textit{In \% of the export price ($\widehat{t}/\widetilde{p}$, in \%)} &&&& \\ \hline
+  %      Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
+  %  Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
+ %  \textit{In USD per kg traded ($\widehat{t}$)}&&&& \\ \hline
+  %  Mean & 0.09	&1.12	&0.08&	2.16 \\
+  %  Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
+    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
+    %Min.
+%value & 0.000 & 0.000 & 0.000 & 0.000 \\
+    %Max.
+%value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
+  %  \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
+   % Transport costs ($p/\widetilde{p} -1$, in \%) & & & & \\ \hline
+   % Mean & 5.3 & 5.0& 5.6&3.9 \\
+   % Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
+    %Export price ($\widetilde{p}$), in USD & & & & \\
+    %Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
+    %Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
+    %\# obs.
+%& 29279 & 28207 & 29317 & 27680 \\
+%    \# origin country & 188 & 191 & 188 & 189 \\
+%    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
+%  \end{tabular}
+ %   \end{center}}
+%\parbox[l]{12cm}{\footnotesize{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows.
+%The additive term is expressed in fraction of fas price.
+%For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
+1974, 1981, 1989, 2001, 2009, 2013.
+($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
+%\end{table}%
 

### 2021-08-26 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Essai de mettre "Model B" centré dans le tableau
Ne marche pas :(
Tableaux appendix  A et B marche avec modèle I
Petits pas sur les tableaux de l’annexe
Progrès sur les tableaux annexe
Inversion des inputs pour les tables de l’appendix
B2-B5 et B7-B10
nouvelles modifs main text et appendix 1

@@
 In this section, we report the estimation results of the three models: Model (A) (with only ad-valorem costs), Model (B) (with both additive and ad-valorem) and Model (C) (with only additive costs).
-Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_vessel} for vessel transport.
+Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_ves} for vessel transport.
 
@@
 	\footnotesize{
-    \caption{Estimation results of the three models (Air, 3-digit level)}
-  \label{tab:3models_estimation_results_air}%
-    \begin{tabular}{l|c c c c c c c c c}
-    \hline \hline
-    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
-   $\#$ obs.
-& \multicolumn{1}{c}{14955} & \multicolumn{1}{c}{16118} & \multicolumn{1}{c}{19908} & \multicolumn{1}{c}{24958} & \multicolumn{1}{c}{31037} & \multicolumn{1}{c}{35027} & \multicolumn{1}{c}{41806} & \multicolumn{1}{c}{40279} & \multicolumn{1}{c}{39351} \\
-   $\#$  origin countries & \multicolumn{1}{c}{152} & \multicolumn{1}{c}{165} & \multicolumn{1}{c}{169} & \multicolumn{1}{c}{181} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{211} & \multicolumn{1}{c}{210} & \multicolumn{1}{c}{210} \\
-   $\#$  sectors & \multicolumn{1}{c}{203} & \multicolumn{1}{c}{204} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{212} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{218} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{216} & \multicolumn{1}{c}{212} \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
-    Mean (in \%) & \multicolumn{1}{c}{6.93} & \multicolumn{1}{c}{5.41} & \multicolumn{1}{c}{6.08} & \multicolumn{1}{c}{5.03} & \multicolumn{1}{c}{4.61} & \multicolumn{1}{c}{3.60} & \multicolumn{1}{c}{4.10} & \multicolumn{1}{c}{4.19} & \multicolumn{1}{c}{3.36} \\
-    Median (in \%)& \multicolumn{1}{c}{5.43} & \multicolumn{1}{c}{3.79} & \multicolumn{1}{c}{5.47} & \multicolumn{1}{c}{4.37} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{2.47} & \multicolumn{1}{c}{3.12} & \multicolumn{1}{c}{3.41} & \multicolumn{1}{c}{2.92} \\
-    Standard Deviation & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (B) - With additive and ad-valorem transport costs}} \\ \hline
-     \multicolumn{10}{l}{ \textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
-    Mean (in \%) & \multicolumn{1}{c}{3.64} & \multicolumn{1}{c}{2.32} & \multicolumn{1}{c}{2.46} & \multicolumn{1}{c}{2.38} & \multicolumn{1}{c}{2.05} & \multicolumn{1}{c}{1.66} & \multicolumn{1}{c}{2.00} & \multicolumn{1}{c}{2.57} & \multicolumn{1}{c}{1.70} \\
-    Median (in \%)& \multicolumn{1}{c}{2.71} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{1.79} & \multicolumn{1}{c}{1.60} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.20} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{2.24} & \multicolumn{1}{c}{1.72} \\
-    Standard Deviation & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline
-      \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)} \\ \hline
-    Mean (in \%) & \multicolumn{1}{c}{2.56} & \multicolumn{1}{c}{2.04} & \multicolumn{1}{c}{2.83} & \multicolumn{1}{c}{1.83} & \multicolumn{1}{c}{1.64} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{1.43} & \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{1.01} \\
-    Median (in \%)& \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.43} & \multicolumn{1}{c}{0.47} \\
-    Standard Deviation & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
-    Mean (in \%) & \multicolumn{1}{c}{6.88} & \multicolumn{1}{c}{4.82} & \multicolumn{1}{c}{5.99} & \multicolumn{1}{c}{4.44} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{3.09} & \multicolumn{1}{c}{4.00} & \multicolumn{1}{c}{4.49} & \multicolumn{1}{c}{3.29} \\
-    Median (in \%)& \multicolumn{1}{c}{4.45} & \multicolumn{1}{c}{1.82} & \multicolumn{1}{c}{3.36} & \multicolumn{1}{c}{2.27} & \multicolumn{1}{c}{1.72} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.93} & \multicolumn{1}{c}{2.72} & \multicolumn{1}{c}{2.06} \\
-    Standard Deviation & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline
-    \end{tabular}
-    }
-\end{table}%
-
-\begin{table}[htbp]
-	\centering
-	\footnotesize{
 	\caption{Estimation results of the three models (Air, products at 5-digit level, sectors at 3-digit level)}
@@
 	\item Statistics are weighted by value
+	\item \textbf{Model (A): Iceberg transpost costs only}
 	\item \textbf{Model (B): With additive and ad-valorem transport costs}
+				\item \textbf{Model (C): With additive transport costs only}
 \end{tablenotes}
 
-}	
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-   \footnotesize{
-    \caption{Estimation results of the three models (Vessel, 3-digit level)}
-  \label{tab:3models_estimation_results_vessel}%
-    \begin{tabular}{l|c c c c c c c c c}
-    \hline \hline
-    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
-    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
-    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
-    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
-    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
-    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
-    Standard Deviation & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
-     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
-    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
-    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
-    Standard Deviation & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
-       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
-    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
-    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
-    Standard Deviation & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
-    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
-    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
-    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
-    Standard Deviation & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
-   \end{tabular}%
 }	
@@
 		\item Statistics are weighted by value
+		\item \textbf{Model (A): Iceberg transpost costs only}
 		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\item \textbf{Model (C): With additive transport costs only}
 		\end{tablenotes}
@@
 
-\begin{landscape}
 \begin{table}[htbp]
@@
 	\label{tab:3models_diagnosis_air}%
-   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
-    \hline \hline
-    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
-    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
-    Model (A) & \multicolumn{1}{c}{0.297} & \multicolumn{1}{c}{0.267} & \multicolumn{1}{c}{0.302} & \multicolumn{1}{c}{0.251} & \multicolumn{1}{c}{0.142} & \multicolumn{1}{c}{0.318} & \multicolumn{1}{c}{0.460} & \multicolumn{1}{c}{0.421} & \multicolumn{1}{c}{0.313} \\
-    Model (B) & \multicolumn{1}{c}{0.594} & \multicolumn{1}{c}{0.646} & \multicolumn{1}{c}{0.635} & \multicolumn{1}{c}{0.627} & \multicolumn{1}{c}{0.658} & \multicolumn{1}{c}{0.640} & \multicolumn{1}{c}{0.593} & \multicolumn{1}{c}{0.513} & \multicolumn{1}{c}{0.419} \\
-    Model (C) & \multicolumn{1}{c}{0.489} & \multicolumn{1}{c}{0.543} & \multicolumn{1}{c}{0.531} & \multicolumn{1}{c}{0.517} & \multicolumn{1}{c}{0.546} & \multicolumn{1}{c}{0.518} & \multicolumn{1}{c}{0.464} & \multicolumn{1}{c}{0.339} & \multicolumn{1}{c}{0.295} \\ \hline
-    \multicolumn{10}{l}{\textbf{SER}}\\ \hline
-    Model (A) & \multicolumn{1}{c}{0.791} & \multicolumn{1}{c}{0.860} & \multicolumn{1}{c}{0.831} & \multicolumn{1}{c}{0.811} & \multicolumn{1}{c}{0.798} & \multicolumn{1}{c}{0.844} & \multicolumn{1}{c}{0.837} & \multicolumn{1}{c}{0.857} & \multicolumn{1}{c}{0.920} \\
-    Model (B) & \multicolumn{1}{c}{0.674} & \multicolumn{1}{c}{0.715} & \multicolumn{1}{c}{0.692} & \multicolumn{1}{c}{0.675} & \multicolumn{1}{c}{0.641} & \multicolumn{1}{c}{0.697} & \multicolumn{1}{c}{0.727} & \multicolumn{1}{c}{0.787} & \multicolumn{1}{c}{0.847} \\
-    Model (C) & \multicolumn{1}{c}{1.610} & \multicolumn{1}{c}{1.778} & \multicolumn{1}{c}{1.736} & \multicolumn{1}{c}{1.699} & \multicolumn{1}{c}{1.700} & \multicolumn{1}{c}{1.786} & \multicolumn{1}{c}{1.783} & \multicolumn{1}{c}{1.776} & \multicolumn{1}{c}{1.723} \\ \hline
-   \multicolumn{10}{l}{ \textbf{AIC criteria}} \\ \hline
-    Model (A) & \multicolumn{1}{c}{35675.0} & \multicolumn{1}{c}{41171.0} & \multicolumn{1}{c}{49315.0} & \multicolumn{1}{c}{60715.6} & \multicolumn{1}{c}{74386.4} & \multicolumn{1}{c}{87492.5} & \multicolumn{1}{c}{103983.0} & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{106130.6} \\
-    Model (B) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\
-    Model (C) & \multicolumn{1}{c}{40808.1} & \multicolumn{1}{c}{45138.5} & \multicolumn{1}{c}{55214.8} & \multicolumn{1}{c}{69458.5} & \multicolumn{1}{c}{83958.6} & \multicolumn{1}{c}{100040.8} & \multicolumn{1}{c}{123592.1} & \multicolumn{1}{c}{129359.0} & \multicolumn{1}{c}{127399.2} \\ \hline
-    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
-    Model (A) & \multicolumn{1}{c}{-17530.5} & \multicolumn{1}{c}{-20253.5} & \multicolumn{1}{c}{-24315.5} & \multicolumn{1}{c}{-29977.8} & \multicolumn{1}{c}{-36811.2} & \multicolumn{1}{c}{-43341.3} & \multicolumn{1}{c}{-51648.5} & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-52690.3} \\
-    Model (B) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\
-    Model (C) & \multicolumn{1}{c}{-20074.1} & \multicolumn{1}{c}{-22217.2} & \multicolumn{1}{c}{-27251.4} & \multicolumn{1}{c}{-34355.3} & \multicolumn{1}{c}{-41634.3} & \multicolumn{1}{c}{-49625.4} & \multicolumn{1}{c}{-61533.0} & \multicolumn{1}{c}{-64339.5} & \multicolumn{1}{c}{-63316.6} \\ \hline
-    \multicolumn{10}{l}{\textbf{Test LL} }\\ \hline
-    Stat LL ratio (B vs A) & \multicolumn{1}{c}{4809.7} & \multicolumn{1}{c}{5980.6} & \multicolumn{1}{c}{7257.3} & \multicolumn{1}{c}{9168.7} & \multicolumn{1}{c}{13548.7} & \multicolumn{1}{c}{13105.7} & \multicolumn{1}{c}{11760.4} & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{6541.2} \\
-    \textbackslash{}\# restrictions & \multicolumn{1}{c}{355} & \multicolumn{1}{c}{369} & \multicolumn{1}{c}{376} & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{428} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{422} \\
-    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline
-    Stat LL ratio (B vs C) & \multicolumn{1}{c}{4948.4} & \multicolumn{1}{c}{4954.0} & \multicolumn{1}{c}{6564.5} & \multicolumn{1}{c}{8961.8} & \multicolumn{1}{c}{11597.4} & \multicolumn{1}{c}{12837.0} & \multicolumn{1}{c}{15764.7} & \multicolumn{1}{c}{17062.0} & \multicolumn{1}{c}{13896.9} \\
-    \textbackslash{}\# restrictions & \multicolumn{1}{c}{355} & \multicolumn{1}{c}{369} & \multicolumn{1}{c}{376} & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{428} & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{422} \\
-    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline \hline
-\end{tabular}%
+	\input{TableA3_air.tex}
+\begin{tablenotes}
+	\tiny
+	\item SER are weighted by value
+	\item \textbf{Model (A): Iceberg transpost costs only}
+	\item \textbf{Model (B): With additive and ad-valorem transport costs}
+	\item \textbf{Model (C): With additive transport costs only}
+\end{tablenotes}
 }
 \end{table}%
-\end{landscape}
 
-\begin{landscape}
 \begin{table}[htbp]
@@
 	\footnotesize{
-  \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
+		\caption{Quality-of-fit diagnostic tests of the three models (Ves, 3-digit level)}
 		\label{tab:3models_diagnosis_vessel}%
-   \begin{tabular}{l|c c c c c c c c c}
-    \hline \hline
-    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
-    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
-    Model (A) & 0.450 & 0.415 & 0.427 & 0.456 & 0.438 & 0.401 & 0.378 & 0.350 & 0.339 \\
-    Model (B) & 0.612 & 0.575 & 0.571 & 0.590 & 0.611 & 0.571 & 0.541 & 0.491 & 0.462 \\
-    Model (C ) & 0.424 & 0.401 & 0.374 & 0.429 & 0.456 & 0.431 & 0.417 & 0.358 & 0.349 \\ \hline
-    \multicolumn{10}{l}{\textbf{SER} } \\ \hline
-    Model (A) & 0.576 & 0.620 & 0.569 & 0.592 & 0.615 & 0.652 & 0.673 & 0.740 & 0.758 \\
-    Model (B) & 0.484 & 0.528 & 0.493 & 0.514 & 0.512 & 0.551 & 0.578 & 0.656 & 0.684 \\
-    Model (C) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
-    \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
-    Model (A) & 33328.81 & 33010.27 & 40275.70 & 51142.62 & 60414.92 & 71365.89 & 85051.02 & 84789.89 & 88191.87 \\
-    Model (B) & 27331.52 & 28067.31 & 34170.52 & 43664.74 & 49275.33 & 60475.91 & 73020.09 & 76161.33 & 80873.72 \\
-    Model (C) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
-    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
-    Model (A) & -16287.40 & -16129.13 & -19767.85 & -25169.31 & -29790.46 & -35263.95 & -42122.51 & -41998.95 & -43692.93 \\
-    Model (B) & -12985.76 & -13353.65 & -16398.26 & -21171.37 & -23905.66 & -29490.96 & -35844.04 & -37418.66 & -39751.86 \\
-    Model (C) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
-   \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
-    Stat LL ratio (B vs A) & 6603.28 & 5550.96 & 6739.18 & 7995.88 & 11769.59 & 11545.98 & 12556.94 & 9160.56 & 7882.15 \\
-    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
-    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\ \hline
-    Stat LL ratio (B vs C) & 19376,88 & 16920,95 & 25295,20 & 28464,79 & 28439,08 & 28944,59 & 30802,84 & 25859,58 & 24062,55 \\
-    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
-    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\
-\hline \hline
-    \end{tabular}%
+		\input{TableA3_ves.tex}
+		\begin{tablenotes}
+			\tiny
+			\item SER are weighted by value
+			\item \textbf{Model (A): Iceberg transpost costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+\end{tablenotes}
 }
 \end{table}%
-\end{landscape}
 
@@
 \subsection{Transport Cost Estimates: Yearly Detailed Results}
+
 In this section, we complement Table 1 of the main text by reporting the year-to-year results of the estimation driven at the 3-digit classification level.
@@
 
+%JH : reprendre XXX ICI
 \begin{landscape}
-\begin{table}[htbp]
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-  %\centering
-\caption{Air: Transport costs estimates, all years, 3-digit}
-\begin{center}
-\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
-\hline
-\hline
-Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
-\hline
-Mean  & 6.9   & 7.5   & 7.2   & 7.7   & 6.9   & 6.1   & 5.4   & 6.0   & 6.4   & 6.9   & 7.2   & 6.1   & 6.4   & 6.6   & 5.7   & 5.3   & 5.0   & 5.1   & 4.9   & 5.1 \\
-Median & 5.4   & 6.4   & 6.9   & 7.1   & 6.3   & 5.3   & 3.8   & 4.9   & 5.3   & 6.1   & 6.7   & 5.5   & 5.9   & 6.3   & 5.3   & 4.6   & 4.4   & 4.5   & 4.5   & 4.4 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
-\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
-\hline
-Mean  & 3.6   & 3.7   & 3.9   & 3.8   & 3.2   & 3.0   & 2.3   & 2.8   & 2.8   & 2.6   & 3.3   & 2.5   & 3.2   & 2.6   & 3.1   & 3.1   & 2.4   & 2.7   & 2.2   & 2.4 \\
-Median & 2.7   & 2.7   & 2.9   & 2.7   & 2.1   & 2.4   & 1.6   & 1.8   & 1.9   & 1.9   & 2.7   & 1.8   & 2.1   & 2.0   & 2.0   & 1.9   & 1.6   & 1.5   & 1.5   & 1.6 \\
-
-\hline
-\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
-\hline
-Mean  & 2.6   & 3.0   & 2.3   & 3.1   & 2.6   & 2.1   & 2.0   & 2.0   & 2.3   & 2.8   & 2.5   & 2.8   & 2.6   & 2.9   & 1.7   & 4.6   & 1.8   & 1.8   & 1.9   & 1.9 \\
-Median & 1.1   & 1.2   & 0.9   & 1.3   & 1.1   & 0.7   & 0.5   & 0.6   & 0.8   & 1.0   & 1.0   & 1.3   & 1.3   & 1.5   & 1.0   & 0.7   & 0.8   & 0.6   & 0.9   & 0.8 \\
-\hline
-\# observations & 14955 & 15299 & 11397 & 10707 & 15222 & 15684 & 16118 & 16864 & 17322 & 18180 & 20644 & 19908 & 20695 & 20793 & 24663 & 25197 & 24958 & 25156 & 26191 & 28296 \\
-
-%%%%%%%%%%%%%%%%2nd part of the table
-
-\hline\hline
-\multicolumn{20}{c}{ }  \\
-\multicolumn{20}{c}{Continued}  \\
-\hline\hline
-
-Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
-\hline
-Mean  & 4.6   & 4.6   & 4.2   & 4.1   & 3.8   & 3.8   & 3.6   & 3.5   & 3.8   & 3.9   & 4.0   & 4.1   & 3.9   & 4.1   & 4.1   & 4.0   & 4.2   & 3.9   & 3.7   & 3.4 \\
-Median & 3.7   & 3.8   & 3.1   & 3.0   & 2.7   & 2.8   & 2.5   & 2.4   & 2.7   & 2.6   & 2.9   & 3.1   & 2.7   & 3.0   & 3.2   & 3.0   & 3.4   & 3.1   & 3.0   & 2.9 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
-\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
-\hline
-Mean  & 2.3   & 2.1   & 1.9   & 1.8   & 1.8   & 1.8   & 1.7   & 1.6   & 1.6   & 1.9   & 1.9   & 2.0   & 1.8   & 2.3   & 2.3   & 2.3   & 2.6   & 2.2   & 2.2   & 1.7 \\
-Median & 1.3   & 1.4   & 1.4   & 1.3   & 1.3   & 1.5   & 1.2   & 1.1   & 1.2   & 1.4   & 1.4   & 1.6   & 1.4   & 1.9   & 1.9   & 1.8   & 2.2   & 1.7   & 1.9   & 1.7 \\
-\hline
-\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
-\hline
-Mean  & 1.7   & 1.6   & 1.5   & 1.5   & 1.4   & 1.4   & 1.3   & 1.3   & 1.6   & 1.4   & 1.5   & 1.4   & 1.3   & 1.2   & 1.2   & 1.2   & 1.1   & 1.1   & 0.9   & 1.0 \\
-Median & 0.8   & 0.7   & 0.6   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.4   & 0.4   & 0.4   & 0.5 \\
-\hline
-\# observations & 29948 & 31037 & 32187 & 33502 & 33492 & 33523 & 35027 & 34885 & 35159 & 35891 & 36990 & 41806 & 42554 & 40858 & 40159 & 38275 & 40279 & 41190 & 40909 & 39351 \\
-
-\hline
-\hline
-\end{tabular}}%
-	\end{center}
-\end{table}%
 
@@
 
+		\input{TableB1974_1987_ves.tex}
+				\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): With ad-valorem transpost costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
+
+	\end{center}
+\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB1988_2001_ves.tex}
+				\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): With ad-valorem transpost costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
+
+	\end{center}
+	\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB2002_2015_ves.tex}
+				\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+						\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
+
+	\end{center}
+	\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+\end{landscape}
+
+\begin{table}[htbp]
+
+	\caption{Continued}
+	\begin{center}
+
+		\input{TableB2016_2019_ves.tex}
+				\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
+
+	\end{center}
+	\label{tab_oa:result_ves_ally3}%
+\end{table}%
+
+As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000).
+This can be attributed to the presence of outliers in the distribution of the additive costs estimates.
+The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000.
+Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
+Still, this does not make much difference.
+When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
+The weighed median value is left unchanged at 2.9\% whether 1989 is included or not.
+
+\begin{landscape}
+\begin{table}[htbp]
+
+	\caption{Air: Transport costs estimates, all years, products at 5-digit level, sectors at 3-digit level}
+	\begin{center}
+
 		\input{TableB1974_1987_air.tex}
+				\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): With ad-valorem transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
 
@@
 		\input{TableB1988_2001_air.tex}
+				\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): With ad-valorem transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
 
@@
 	\input{TableB2002_2015_air.tex}
+			\begin{tablenotes}
+		\tiny
+		\item Statistics are weighted by value
+		\\item \textbf{Model (A): With ad-valorem transport costs only}
+		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\item \textbf{Model (C): With additive transport costs only}
+	\end{tablenotes}
 
@@
 \end{table}%
+
+\end{landscape}
 
@@
 	\input{TableB2016_2019_air.tex}
+			\begin{tablenotes}
+		\tiny
+		\item Statistics are weighted by value
+			\item \textbf{Model (A): With ad-valorem transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+	\end{tablenotes}
 
@@
 \end{table}%
-\end{landscape}
-
-As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000).
-This can be attributed to the presence of outliers in the distribution of the additive costs estimates.
-The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000.
-Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
-Still, this does not make much difference.
-When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
-The weighed median value is left unchanged at 2.9\% whether 1989 is included or not.
-
-\begin{landscape}
-
-\begin{table}[htbp]
-\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
-\caption{Vessel: Transport costs estimates, all years, 3-digit}
-\begin{center}
-\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
-\hline
-\hline
-Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
-\hline
-Mean  & 9.8   & 9.9   & 8.9   & 8.3   & 8.1   & 7.5   & 6.5   & 6.0   & 6.3   & 7.0   & 7.0   & 6.9   & 6.7   & 6.2   & 6.1   & 5.7   & 5.7   & 5.5   & 5.0   & 5.2 \\
-Median & 9.6   & 8.5   & 8.0   & 7.3   & 7.1   & 6.5   & 5.5   & 5.0   & 5.9   & 5.7   & 6.1   & 6.3   & 7.0   & 6.3   & 5.7   & 4.8   & 4.6   & 4.4   & 4.2   & 4.7 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
-\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
-\hline
-Mean  & 5.4   & 4.8   & 5.4   & 5.2   & 5.9   & 4.6   & 3.1   & 3.3   & 3.4   & 4.2   & 4.1   & 4.0   & 3.9   & 3.6   & 4.0   & 3.0   & 3.3   & 3.0   & 2.6   & 2.9 \\
-Median & 4.9   & 4.1   & 4.8   & 4.4   & 5.4   & 4.0   & 2.4   & 2.9   & 2.9   & 3.9   & 3.5   & 3.6   & 3.6   & 3.0   & 3.5   & 2.6   & 2.8   & 2.7   & 2.3   & 2.6 \\
-\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
-\hline
-Mean  & 5.1   & 5.5   & 3.5   & 3.5   & 2.5   & 3.1   & 3.4   & 2.9   & 3.5   & 2.9   & 3.2   & 3.2   & 2.9   & 2.8   & 2.4   & 2.9   & 2.7   & 2.8   & 2.7   & 2.7 \\
-Median & 2.9   & 3.6   & 1.9   & 1.7   & 1.2   & 1.7   & 2.3   & 1.5   & 2.3   & 2.0   & 2.3   & 2.1   & 1.8   & 1.8   & 1.3   & 2.0   & 1.7   & 1.7   & 1.8   & 1.6 \\
-\hline
-\# observations & 19007 & 18710 & 13615 & 12826 & 16601 & 17274 & 17356 & 17788 & 18075 & 18883 & 21650 & 23348 & 23729 & 23626 & 27661 & 29106 & 28383 & 28095 & 29050 & 30839 \\ %%%%%%%%%%%%%%%%2nd part of the table
-
-\hline\hline
-\multicolumn{20}{c}{ }  \\
-\multicolumn{20}{c}{Continued}  \\
-\hline\hline
-
-Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
-\hline
-Mean  & 5.2   & 5.1   & 4.8   & 4.7   & 4.8   & 5.0   & 5.1   & 5.0   & 4.8   & 5.2   & 5.4   & 5.5   & 4.8   & 4.7   & 4.4   & 4.3   & 4.0   & 3.5   & 3.6   & 3.6 \\
-Median & 4.1   & 4.3   & 3.9   & 3.9   & 3.9   & 4.5   & 4.9   & 4.6   & 4.1   & 4.8   & 5.1   & 4.9   & 4.2   & 4.2   & 3.8   & 4.1   & 3.6   & 3.0   & 3.1   & 3.3 \\
-\hline
-\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
-\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
-\hline
-Mean  & 2.6   & 2.8   & 2.6   & 2.5   & 2.2   & 2.5   & 2.5   & 2.7   & 2.4   & 2.4   & 2.7   & 2.6   & 2.3   & 2.5   & 2.1   & 2.2   & 1.9   & 1.8   & 1.8   & 2.2 \\
-Median & 2.2   & 2.5   & 2.2   & 2.2   & 1.9   & 2.1   & 2.1   & 2.6   & 2.3   & 1.9   & 2.8   & 2.2   & 1.9   & 2.3   & 1.8   & 2.0   & 1.8   & 1.6   & 1.4   & 1.8 \\
-\hline
-\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
-\hline
-Mean  & 2.9   & 2.7   & 2.5   & 2.5   & 3.2   & 2.8   & 2.8   & 2.4   & 2.6   & 3.2   & 2.9   & 3.0   & 2.8   & 2.4   & 2.4   & 2.1   & 2.5   & 1.9   & 1.9   & 1.5 \\
-Median & 2.0   & 1.8   & 1.6   & 1.3   & 2.0   & 2.0   & 2.2   & 1.6   & 2.0   & 2.5   & 1.9   & 2.2   & 1.9   & 1.8   & 2.1   & 1.7   & 1.9   & 1.6   & 1.6   & 0.8 \\
-\hline
-\# observations & 31865 & 32146 & 32344 & 33181 & 33986 & 34585 & 36090 & 36407 & 37255 & 37672 & 37757 & 41431 & 41763 & 39604 & 38950 & 37332 & 37748 & 38562 & 38387 & 38473 \\
-\hline
-\hline
-\end{tabular}}
-
-	\end{center}
-\end{table}%
-
-\begin{table}[htbp]
-
-	\caption{Air: Transport costs estimates, all years, products at 5-digit level, sectors at 3-digit level}
-	\begin{center}
-
-		\input{TableB1974_1987_ves.tex}
-
-	\end{center}
-	\label{tab_oa:result_ves_ally3}%
-\end{table}%
-
-\begin{table}[htbp]
-
-	\caption{Continued}
-	\begin{center}
-
-		\input{TableB1988_2001_ves.tex}
-
-	\end{center}
-	\label{tab_oa:result_ves_ally3}%
-\end{table}%
-
-\begin{table}[htbp]
-
-\caption{Continued}
-\begin{center}
-
-	\input{TableB2002_2015_ves.tex}
-
-\end{center}
-\label{tab_oa:result_ves_ally3}%
-\end{table}%
-
-\begin{table}[htbp]
-
-\caption{Continued}
-\begin{center}
-
-	\input{TableB2016_2019_ves.tex}
-
-\end{center}
-\label{tab_oa:result_ves_ally3}%
-\end{table}%
-
-\end{landscape}
 

### 2021-08-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Première version table 4
Il manque des donnéess.
graphique robustesse séparée

@@
 
+\begin{table}[htbp]
+	\centering
+	\caption{Robustness to the separability assumption (average over the period) - restricted sample}
+	\begin{center}
+			\input{table4.tex}
+	\end{center}
+	\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+			The additive term is expressed in fraction of fas price.}}
+	\label{tab:robustness_separability}
+\end{table}
+
 Three main comments can be made.
@@
 \begin{tabular}{cc}
-{\small (a) Air freight} & {\small (b) Vessel freight}\\
-\includegraphics[width=3in, height=2in]{graph_robustesse_ns_mp_terme_A_air.pdf}
-& \includegraphics[width=3in,height=2in]{graph_robustesse_ns_mp_terme_A_ves.pdf} \\
-\includegraphics[width=3in, height=2in]{graph_robustesse_ns_mp_terme_I_air.pdf}
-& \includegraphics[width=3in,height=2in]{graph_robustesse_ns_mp_terme_I_ves.pdf} \\
+\includegraphics[width=6in, height=4in]{graph_robustesse_ns.jpg}
 

### 2021-08-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Figure 7

@@
 \begin{tabular}{cc}
-\multicolumn{2}{c}{Air transport} \\
-{\small (a) Separability (baseline)} & {\small (b) Non-separability}\\
-\includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_air.pdf}
-& \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_air.pdf} \\
-\multicolumn{2}{c}{Maritime transport} \\
-{\small (c) Separability (baseline)} & {\small (d) Non-separability}\\
-\includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2separe_ves.pdf}
-& \includegraphics[width=7cm, height=5cm]{Etude_beta_pond_sitc2ns_ves.pdf} \\
+\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
 \multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\

### 2021-08-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Graphique 9

@@
 Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
+
+----Pas sûr. Je crois que c’est weighted by value/value total for the year --
+
 In this section, we check whether this difference of treatment is responsible or not for the difference in results relative to the role of the trade composition effects, which we rather attribute to the modeling of the additive component.
@@
 \includegraphics[height=8cm]
-{graph_composition_all_np.pdf}
+{graph_composition_all_np.jpg}
 \end{center}
@@
 \begin{figure}[htbp]
-\caption{Transport cost time trends: robustness to the weighting scheme}
+\caption{Transport cost time trends: robustness to IV procedure}
 \label{fig:comp_IV_SITC5}
@@
 \includegraphics[height=8cm]
-{comp_IV_SITC5.pdf}
+{scatter_chronology_baseline_IV_ref1_y_5_3.png}
 \end{center}

### 2021-08-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 \noindent We devoted a lot of thought to this point, which directly echoed a similar concern by the other referee.
-Taking stock of your suggestion, we decided to refocus the paper on the welfare implications of our results.
-In this regard, the section 3.4 from the initial version has been moved to Appendix XXX TO BE DONE XXX of the paper, and replaced by a new section 3.4 devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
+Taking stock of your suggestion, we decided to offer insights on the welfare implications of our results.
+In this regard,% the section 3.4 from the initial version has been moved to Appendix XXX TO BE DONE XXX of the paper,
+the new section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
 \citet{melitz}, featuring additive trade costs.
@@
 
-More precisely, we use some of the estimates underlying results reported in sections 2 and 3 concering multiplicative and additive costs, more precisely those for the years 1974 and 2019.
+More precisely, we use some of the estimates underlying results reported in sections 2 and 3 concerning multiplicative and additive costs, more precisely those for the years 1974 and 2019.
 Based on the latter, we implement several comparative static exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
 In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
-To that end, we play with the share of exporting firms in the US, the latter being found in XXX: it reaches 35\% in 2019, versus XX\% in 1974.
+To that end, we adjust the share of exporting firms in the US: based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
 
-Table XX on p.XX in section 3.4XX in the revised version reports the results of these various comparative static exercises for air (Panel A) and Vessel (Panel B) transport modes.
-For each panel, the first two rows of the Table report the figures for total transport costs in 1974 and 2019, and the subsequent rows show the difference between the two years, how it is split (or not) between multiplicative and additive costs and the consequences for the welfare.
-Column (1) reports the results for when costs are assumed to be exclusively multiplicative, while column (2) combines this feature with a increase in the share of exporting firms.
-Column (3) reports results based on the model including additive costs, and column (4) again adds to the picture an increase in the share of exporting firms.
+Table 5 in section 5 in the revised version reports the results of these various comparative static exercises for Air and Vessel transport modes.
+Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
+For each transport mode, the first two columns report the results of reducing transport costs only, maintaining the same value of $f_x$ as in the initial equilibrium (Columns (A)-(B) for Air, (E)-(F) for Vessel), contrasting the case of only ad-valorem costs (Columns (A), (D)) and in presence of additive costs (Columns (B), (F)).
+Columns (C)-(D) for Air and (G)-(H) for Vessel report the results when both the variable transport costs and the fixed export costs are lower in the final steady state (2019) relative to the initial one (1974).
 
 Qualitatively, the conclusions are identical for both transport modes, with two major insights.
-First, for a given decrease in transport costs, the comparison in columns (1) and (3) shows welfare gains around 50\% higher when the underlying framework allows for additive costs - which, in practice, concentrate the major part of transport costs decrease over the period.
-Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing costs, with again a an additional premium coming from the decrease of additive costs.
+First, for a given decrease in transport costs,  welfare gains are around 50\% higher when the underlying framework allows for additive costs - which, in practice, concentrate the major part of transport costs decrease over the period.
+Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing costs, with again an additional premium coming from the decrease of additive costs.
 Quantitatively, welfare gains are more important for the Vessel mode, which is a natural consequence of a more substantial decrease in transport costs.
@@
 Note also that not only the inclusion of additive costs in the underlying framework generates big welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
-
-%Gains de bien-être sont plus forts lorsqu’on bouge les coûts additifs. Voir >Table_compWelfare2.xls.
-%Ex 1 : Comparer C (que des add valorem) et E (additifs et ad valorem) : on baisse les coûts de transports ; sheet 1 = air, sheet 2 = Vessel.
-%Ex 2 : si en plus on ajuste fx, cela amplifie le diagnostic. Voir colonnes D et F : la baisse des couts fixes joue plus dans un monde avec ad%ditifs que sans additifs.
-%Voir « C:\Users\coadministrateur\Dropbox\Papier_Lise_Guillaume\trade_cost\model », et les tableaux afférents.
-% Big picture: quantitatively, we are more exhaustive/more complete.
-We find big welfare differences with additive costs, even if we focus only on transport costs.--> Our results are probably a lower bound.
 

### 2021-08-31 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Nouvelles updates et questions texte et appendix 1

@@
 
-% JH : reprendre XXX ICI
-
 \subsection{Estimating transport costs: empirical specification}
@@
 Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}.
-Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.}
+Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.
+XXXJH: WHERE ???
+XXX}
 \begin{eqnarray}
@@
 
-While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2018}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
+While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2017}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
 Based on Equation (\ref{eq:base}), we exploit our dataset to obtain explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
@@
 These cut-offs are aimed at eliminating reporting or coding errors.
-We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
+We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2019, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
 Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
@@
 
-\textbf{stop here}
+\textbf{stop here XXX JH: WHY? WHAT DO YOU MEAN?}
 
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
+It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
 The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
@@
 Overall results (over the whole period) are not substantially affected if this year is included.
-This can be seen in the results detailed year by year in the Online Appendix, available on the authors' webpages.}
+This can be seen in the results detailed year by year in section B of the Online Appendix, available on the authors' webpages.}
 
@@
 This result is not surprising, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
-One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point.
+One can also note that estimated total costs are most of the time higher than observed costs by 0.5 to 1 percentage point.
 This is first due to the fact that observed data are \textit{de facto} trade-weighted (i.e.
@@
 First, it provides estimation results about the size of the overall transport costs.
-Table \ref{tab:summary_results} shows that transport costs amount to around a 6\% markup in ocean shipping, and 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (a) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period).
+Table \ref{tab:summary_results} shows that transport costs amount to around a 6\% markup in ocean shipping, and 4\% markup in air shipping on average over 1974-2019.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+Under Model (a) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 4.6\% in vessel, and 3.9\% in air shipping (considering the mean value over the period).
 Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
-With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in vessel.
-A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
+With $\tau-1= 0.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\% in vessel.
+A similar calculus gives 4.1\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
 \cite{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
@@
 Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
-Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8\% for ocean shipping and air transport respectively.
-This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period).
-Put differently, the additive component amounts to 48\% of total transport costs in vessel, and 42\% in Air.
+On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.1 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
+Expressed in terms of the export price, the additive costs amount to 2.8\% and 1.7\% for ocean shipping and air transport respectively.
+This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3\% and 2.4\% for vessel and air shipping respectively (on average over the period).
+Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as highlighted by the row $\widehat{\beta}$:  \textit{-Share of additive costs}
 This shows that the per-kg cost dimension is quantitatively sizable.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 2.
+For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 1.5/1.7.
 Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests that confirm the importance of the additive cost component.
 The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
-An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.%\smallskip
+An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration (see the exercise performed in section \ref{sec:theory}).%\smallskip
 
@@
 In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
-\cite{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
-For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
+\cite{Irrazabal_2015}'s results on the year 2004 imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
+For the same export price multiplied by the ad-valorem cost of 100 USD, our results for 2004 yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
 Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.
 On top of our results summarized in Table \ref{tab:summary_results}, this points to evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
+
+% JH : reprendre XXX ICI
 
@@
 As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the fas price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
-Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2013.
+Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2019.
 
@@
 
-\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013).
+\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
 To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
@@
 First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping.
+International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
 This stands in line with the related literature (see \citealp{Hummels_1999}; \citealp{Lafourcade_Thisse}).
@@
 For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
-Air transport costs were reduced by 50\% between 1974 and 2013, which is mainly attributable to a reduction in the transport costs ``per se''.
+Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
 Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
@@
     \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2013} \\
+    \multicolumn{5}{c}{Mean value over 1974-2019} \\
  \multicolumn{5}{c}{3 digits} \\    \hline \hline
@@
 This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
+Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2019, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
 Two main findings emerge from our results.
@@
 We find that additive costs are quantitatively sizable: They amount to 2.8\% of the export price unit values for ocean shipping, and 1.8\% for air transport.
-This is slightly lower that the estimated values for the ad-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013.
+This is slightly lower that the estimated values for the ad-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2019.
 Further, modeling the additive component significantly improves the quality of fit of the regression.
@@
 In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
-Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level.
+Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2019, when the degree of classification retained ($s$) is at the 3-digit classification level.
 Table \ref{tab:result_air_3d_detail} reports the results for air transport.

### 2021-08-31 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Nouvelles updates et questions texte et appendix 1

@@
 	\item Statistics are weighted by value
-	\item \textbf{Model (A): Iceberg transpost costs only}
+	\item \textbf{Model (A): Iceberg transport costs only}
 	\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 		\item Statistics are weighted by value
-		\item \textbf{Model (A): Iceberg transpost costs only}
+		\item \textbf{Model (A): Iceberg transport costs only}
 		\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 	\item SER are weighted by value
-	\item \textbf{Model (A): Iceberg transpost costs only}
+	\item \textbf{Model (A): Iceberg transport costs only}
 	\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 			\item SER are weighted by value
-			\item \textbf{Model (A): Iceberg transpost costs only}
+			\item \textbf{Model (A): Iceberg transport costs only}
 			\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 In this section, we complement Table 1 of the main text by reporting the year-to-year results of the estimation driven at the 3-digit classification level.
-Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for vessel.
-In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).
+Table \ref{tab_oa:result_ves_ally3} reports the results for each year over 1974-2019 for vessel; Table \ref{tab_oa:result_air_ally3} reports similar results for air transport.
+In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)), when both additive and ad-valorem costs are modeled (Model (B)) and when only additive costs are modeled (Model (A)).
 
-%JH : reprendre XXX ICI
 \begin{landscape}
@@
 			\item Statistics are weighted by value
-			\item \textbf{Model (A): With ad-valorem transpost costs only}
+			\item \textbf{Model (A): With ad-valorem transport costs only}
 			\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 \end{table}%
+
+\setcounter{table}{0}
 
@@
 \end{landscape}
+
+\setcounter{table}{0}
 
@@
 
+\setcounter{table}{1}
+
 \begin{table}[htbp]
@@
 \end{table}%
+
+\setcounter{table}{1}
 
@@
 \end{table}%
+
+%JH : reprendre XXX ICI
 

### 2021-09-03 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement organisation papier et fond

@@
 
-\date{July 2021}
+\date{September 2021}
  \maketitle
@@
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
+\textbf{A REPRENDRE - BIG PICTURE}
 \end{abstract}
@@
 
-Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
+\textbf{A REPRENDRE - BIG PICTURE EN DEUX POINTS }
+
+Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
 changes over time in the baskets of imported goods and/or the origin countries).} Specifically, trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
@@
 Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
+First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of \textbf{the ratio of the additive cost to the export price multiplied by the ad-valorem cost LE DIRE PLUS SIMPLEMENT}; by contrast, our estimation strategy enables us to estimate each value of the ad-valorem and the additive costs independently from the other.
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
+\textbf{A DEVELOPPER A MINIMA SUR LES IMPLICATIONS EN TERMES DE MODELISATION?}
 While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
@@
 %\textbf{Big Picture}.
-Our results are robust to a range of robustness checks, including different weighting schemes, the use of instrumental variables and more disaggregated data at the HS10 level.
+Our results are robust to a range of robustness checks, including the use of instrumental variables and more disaggregated data at the SITC 4 or HS10 levels.
 They have important implications for both international trade and international macroeconomics models.
-Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \cite{sorensen2014} extends \cite{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
+\textbf{Parler des effets de composition comme une des implications}.
+Beyond the positive aspect, some recent papers also point out the normative implications of additive trade costs.
+\cite{sorensen2014} thus analytically shows that important welfare gains can be achieved by reducing additive trade costs.
+Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
+The quantitative results we provide in the paper allow making one step in filling that gap.
+%\footnote{ \cite{sorensen2014} extends \cite{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
 A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
-This is confirmed by \cite{Irrazabal_2015}} These results suggest that important welfare gains can be achieved by reducing additive trade costs.
-On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models.
-In both fields yet, not much progress has been made in quantifying such gains.
-One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
-The quantitative results we provide in the paper allow making one step in filling that gap.
+This is confirmed by \cite{Irrazabal_2015}.}
+ %On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models.
 
@@
 It is also the seller's responsibility to clear the goods for export.
-The FOB price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
+The fob price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
 Once the goods clear the railing of the vessel, the buyer assumes the risk.
 Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity of goods.
+While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity (the number) of goods.
 This is due to the fact that in the Census data, information on the quantity of goods is not mode-specific.
@@
 \textbf{See Section XXX JH: COULD NOT FIND THESE ANYWHERE.
-HAS IT BEEN DONE?}.} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+HAS IT BEEN DONE?}.
+\textbf{LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
+Le remonter dans le corps PPL?}} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
@@
 Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC coding system.
-In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
-We will make use of this to highlight potential differences in the dynamics of transport costs across transportation mode.
+In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, \textbf{land/ocean (or ``vessel'') POURQUOI LAND ?
+TROUBLANT} and air.
+We will make use of this to highlight potential differences in the magnitude of transport costs across transportation modes.
 Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g.
+This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the levels of both the iceberg costs and of the additive costs, in contrast to e.g.
 \cite{Irrazabal_2015}.
@@
 
-\paragraph{Estimation strategy} Starting from the observed transport costs, we decompose the respective sizes of their two sub-components, ad-valorem and additive transport costs.
-From this, we can deduce the share of the additive component in total transport costs, which also interprets as the elasticity of transport costs with respect to the unit price (in absolute value), as we show below.
-\smallskip
-
-Our starting point relates the import price (at the entry gate) of a given good as equal to the export price (at the export gate) plus freight charges.
+\paragraph{Estimation strategy} Our starting point relates the import price (at the entry gate) of a given good as equal to the export price (at the export gate) plus freight charges.
 Specifically, denoting by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped, we have: $ p = \widetilde{p} +f$.
-As in \cite{hummels2010}, assume the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
+As in \cite{hummels2010}, assume that the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
 The link between the import, export prices and shipping charges is hence given by:
@@
 
-To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as:
-
-\begin{eqnarray}
-&& p = (1+\tau) \widetilde{p}+ t \label{eq:tt0} \\
-\Leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
-\end{eqnarray}
-
-\noindent with $\tau \geq 0$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})).
-Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}.
+To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as $p = (1+\tau) \widetilde{p}+ t$, with $\tau \geq 0$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported.
+We can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, leading to:
+\begin{equation}
+p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
+\end{equation}
 Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.
@@
 For sake of reading convenience, we skip the year and transport-mode dimensions in the notations.
-Transforming the above equation (\ref{eq:base}) as the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation:
+Transforming the above equation (\ref{eq:base}) as the percentage change in prices induced by transportation, we get the following baseline specification underlying our estimation:
 
@@
 We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i+ \tau_k \simeq (1+\tau_i)\times(1+\tau_k) -1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+In other words, $\tau_{ik}$ and $t_{ik}$ are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i+ \tau_k \simeq (1+\tau_i)\times(1+\tau_k) -1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 
-\textbf{stop here XXX JH: WHY? WHAT DO YOU MEAN?}
-
 As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, which decomposes into both an additive and an ad-valorem component.
@@
 For each year and transport mode, we estimate two models.
-In Model (a), additive costs are excluded, transport costs being modeled as iceberg costs only (Equation (\ref{eq:model_nlI})).
-In this case, the estimated equation simplifies as:
+In Model (a), additive costs are excluded, transport costs being modeled as iceberg costs only. In this case, the estimated equation simplifies as:
 
@@
 From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modeling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
-The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating further results.
+This has driven us to estimate this third model (c) as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
+The main result that emerges is that this model (c) is dominated (in terms of quality of fit properties) by the model (a) with multiplicative costs only, which is itself dominated by the complete model (b), anticipating further results.
 More details of these results are given in the Online Appendix.}
@@
 These results constitute our first original contribution to the literature.
-We also provide diagnostic tests about the goodness-of-fit measures, which confirm the empirical relevance of adding the additive component in the estimation (see results in Appendix \ref{app:diagnostic_test}).\smallskip
+\smallskip
 
@@
 \caption{Transport costs estimates: summary} \vspace{5mm} \label{tab:summary_results}
-	\input{Table1.tex}
+	\input{Table1_reshaped.tex}
 		\begin{tablenotes}
@@
 		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
-		\item The additive term is expressed in fraction of fas price.
+		%\item The additive term is expressed in fraction of fas price.
 		\item For the  4-digit model, statistics for observed data have been calculated for the same set of years as used for estimation, i.e.
 		1974, 1981, 1989, 2001, 2009, 2013, 2017 and 2019
-		\item \textbf{Model (A): With ad-valorem transpost costs only}
-		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\item \textbf{Model (a): With ad-valorem transport costs only}
+		\item \textbf{Model (b): With additive and ad-valorem transport costs}
 	\end{tablenotes}
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.
-The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
-Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
-Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for air transport at the 3-digit level, removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year.
-Overall results (over the whole period) are not substantially affected if this year is included.
-This can be seen in the results detailed year by year in section B of the Online Appendix, available on the authors' webpages.}
+It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
+Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}
+The top panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.
+In the medium panel, we report the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})).
+The bottom panel presents our estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
 
@@
 
-Table \ref{tab:summary_results} calls for two types of comments.
-The first type is of statistical order.
-As displayed in Table \ref{tab:summary_results}, estimated mean values are systematically higher than medians, by around one percentage point.
-This result is not surprising, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
-One can also note that estimated total costs are most of the time higher than observed costs by 0.5 to 1 percentage point.
-This is first due to the fact that observed data are \textit{de facto} trade-weighted (i.e.
+\textbf{ATTENTION AI MODIFIE PAR RAPPORT A AVANT, LIRE LE .tex en commentaire. CLARIFICATION: ON MONTRE BIEN LA MOYENNE PONDEREE, non? }.
+Before jumping to the interpretation, Table \ref{tab:summary_results} calls for two comments of statistical order.
+First, the estimated mean values are systematically higher than medians.
+This result is not surprising though, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
+Second, the estimated total costs are not exactly equal to the observed ones, in particular for Model (a).
+This might be accounted for by recalling that all our estimates are based on log-linearization.
+Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
+Descriptive statistics showing that both observed and predicted values do match in logarithms are available upon request to the authors.
+
+%One can also note that estimated total costs are most of the time higher than observed costs by 0.5 to 1 percentage point.
+%This is first due to the fact that observed data are \textit{de facto} trade-weighted (i.e.
 by the weight of each good in total trade) while our estimates are based on simple arithmetic means within sector and country.
-Furthermore, Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization.
-Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values.
-Descriptive statistics showing that both observed and predicted values do match in logarithms are available upon request to the authors.%\medskip
 
@@
 Table \ref{tab:summary_results} shows that transport costs amount to around a 6\% markup in ocean shipping, and 4\% markup in air shipping on average over 1974-2019.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (a) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 4.6\% in vessel, and 3.9\% in air shipping (considering the mean value over the period).
-Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
-With $\tau-1= 0.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\% in vessel.
+Under Model (a) (ad-valorem costs only), this gives a markup of 4.6\% in vessel, and 3.9\% in air shipping (considering the mean value over the period).
+Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau) + \frac{t}{\widetilde{p}}$.
+With $\tau= 0.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\%.
 A similar calculus gives 4.1\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
@@
 Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.1 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
+On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to around 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.13 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
 Expressed in terms of the export price, the additive costs amount to 2.8\% and 1.7\% for ocean shipping and air transport respectively.
@@
 This shows that the per-kg cost dimension is quantitatively sizable.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 1.5/1.7.
-Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests that confirm the importance of the additive cost component.
+\medskip
+
+We also provide diagnostic tests about the goodness-of-fit measures, which confirm the empirical relevance of adding the additive component in the estimation.
+Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests that confirm the importance of the additive cost component.
 The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
-An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration (see the exercise performed in section \ref{sec:theory}).%\smallskip
+For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 1.5/1.7.\smallskip
 
@@
 
-% JH : reprendre XXX ICI
-
-\paragraph{Time dynamics} Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, i.e. a structural phenomenon.
-We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
-The corresponding values reported in detailed tables are available in the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
-Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
-
-\begin{figure}[htbp]
-\caption{Share of additive costs (3-digit classification)}
-\label{fig:part_cout_additif}
-\begin{center}
- \includegraphics[width=12cm, height=7.5cm]{Figure1_share_of_additive_in_totalTC.jpg}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
-In both transport modes, the additive component appears of sizable importance throughout the period.
-Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-On the contrary, they represent a sizable and structural dimension of international transport costs.
-If the share of additive costs in total costs roughly amounts to 36\% (39\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered.
-The share of additiive costs air transport costs has been steeply declining since the end of the 1980s.
-In the next section, we further investigate the importance of the (varying pattern of) additive costs, as regards the determinants of transport costs time trends.
-
-\section{Transport costs time trends: the role of the additive component}\label{sec:results_trends}
+\subsection{A first look at the time dynamics of transport costs}
 
@@
 This focus is shared with some of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
-Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
-
-\subsection{A first look at the trend patterns}
-
-As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the fas price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
-Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2019.
+Along with the pattern of total transport costs over time, our empirical set-up also allows us to investigate whether additive costs we put forward so far looking at the average, is a long-lasting, i.e.
+a structural phenomenon.
+
+\paragraph{Total transport costs: A first look at the trend patterns} Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs expressed in percentage of the fas price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
+Each yearly value is obtained as: $\widehat{\tau}+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2019.
 
@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year} for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 For the same years, we obtain a total decrease from 6.9\% to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
-Overall, our results display trends that are close to those reported by \cite{Hummels_1999}.\smallskip
-
-Before making any definite statement, however, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \cite{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
-Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
-Our own results tend to challenge this view, as we further show.
-
-\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
-He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
-
-\begin{equation}
-TC_{ikt}\equiv \frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} = \widetilde{p}_{ikt}^{-\beta}X_{ikt} \label{eq:Hummel_benchmarkeq}
-\end{equation}
-
-Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value).
-As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem.
-At the other extreme, $\beta = 1$ means that all transport costs are additive.
-Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
-Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
-In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
-Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
-
-\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
-%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
-We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
-To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e.
-$\beta$, according to:
-\begin{eqnarray*}
-\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
-\end{eqnarray*}
-
-\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-We can then rebuild the value of the elasticity on a per product-year-country basis.
-The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
+Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip
+
+\paragraph{The share of additive costs: Time dynamics} As next step, we investigate whether additive costs are a long-lasting, i.e. a structural phenomenon.
+
+We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
+The corresponding values reported in detailed tables are available in the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
+Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
 
 \begin{figure}[htbp]
-\caption{Share of additive costs in total costs: Histogram}
-\label{fig:histogram_beta}
-\begin{center}
-\begin{tabular}{cc}
-%{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.0in, height=2.5in]{"Etude_beta_pondere_(a) Air transport".jpg}
-& \includegraphics[width=3.0in,height=2.5in]{"Etude_beta_pondere_(b) Vessel transport".jpg} \\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
-\end{tabular}
+\caption{Share of additive costs (3-digit classification)}
+\label{fig:part_cout_additif}
+\begin{center}
+ \includegraphics[width=12cm, height=7.5cm]{Figure1_share_of_additive_in_totalTC.jpg}
+%{\footnotesize {OECD data} }
 \end{center}
@@
 
-As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
-This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
-Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
-These results are not reported for sake of space, but are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
-
-\subsection{Empirical specification}
-
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
-We start with the presentation of our estimation strategy before turning to the results.\smallskip
-
-\subsubsection{Estimation strategy}
-
-The estimation carried out in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
-Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
-
-One advantage of this method is that it yields measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
-We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).
-
-\subsubsection{Estimation method for each additive and multiplicative component}
-
-We begin by describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
-\smallskip
-
-\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e.
-extracting from trade composition effect) can be described as a two-stage process.
-First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-For the estimated ad-valorem component, we estimate the following equation:
-\begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
-\end{eqnarray}
-
-\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
-the series of $\gamma_t$ differs from one estimation to the other).
-Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
-
-%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
-\begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
-\end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
-Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
-Both equations are estimated by transport mode.\smallskip
-
-In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
-This constitutes the second stage of our procedure.
-As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we build the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
-\begin{equation}
-\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
-\end{equation}
-
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (see details in Appendix \ref{app:comp-effects}).
-In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
-
-As for the additive cost defined in $[0;+\infty]$, we build the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
-\begin{equation}
-\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
-\end{equation}
-
-\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
-
-\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
-To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
-
--------IL FAUT INTRODUIRE LES PRIX FOB. Sinon, les coûts additifs ne sont pas modifiés de l’inflation -----
-
-$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
-
-\subsubsection{Estimation method for the total transport cost measures}
-
-We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
-composition effects excluded).
-As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
-Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
-\smallskip
-
-\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
-Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
-We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
-
-\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
-composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
-In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
-For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
-
-$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
-
-with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
-A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
-
-$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
-
-with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.
-
-We then deduce the fitted value of the overall cost according to:
-$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$
-
-Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
-
-\subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
-Panels (d), (e) and (f) report the results for vessel.
-In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
-
-\begin{figure}[htbp]
-\caption{Transport costs (with and without composition effects)}
-\label{fig:totalTC_compeffects_excl}
-\begin{center}
-\includegraphics[height=4in]
-{graph_composition_all.jpg}
-\end{center}
-\end{figure}
-
-Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
-This stands in line with the related literature (see \citealp{Hummels_1999}; \citealp{Lafourcade_Thisse}).
-Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
-Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
-Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
-For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
-Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
-Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
-Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
-Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
-This is particularly the case for the multiplicative component (panel (d)).\smallskip
-
-\subsection{Time trends in transport costs and the modeling of the additive component}
-
-The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
-If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
-This drives us to investigate this difference further.
-As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
-Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
-
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
-The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
-composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \cite{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
-
-\begin{figure}[htbp]
-\caption{Characterizing the time trends: applying Hummel's (2007) method }
-\label{fig:comp_effects_as_in_Hummels}
-\begin{center}
-\begin{tabular}{cc}
-{\small (a) Ad-valorem air freight} & {\small (b) Ad-valorem vessel freight}\\
-\multicolumn{2}{c}{{\small In percent of the export price}} \\
-\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
-& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
-{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
-\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
-\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.jpg}
-& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.jpg} \\
-\end{tabular}
-\end{center}
-\end{figure}
-
-These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
-Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
-
-Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction in international transport costs observed over time, in particular in air transport.
-Complementing the findings of Section \ref{sec:results_decomposition}, these results highlight the importance of integrating the additive dimension of international transport costs, with the aim here of characterizing their time trends.
-
-\section{Robustness analysis \label{sec:robustness}}
-
-\subsection{Robustness to the separability assumption}
+In both transport modes, the additive component appears of sizable importance throughout the period.
+Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
+On the contrary, they represent a sizable and structural dimension of international transport costs.
+If the share of additive costs in total costs roughly amounts to 44\% (30\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, with a different pattern across transport modes.
+In maritime transport exhibits, the share of additive costs fluctuates around the average, while it has been steeply declining since the end of the 1980s in air transport.
+
+\paragraph{Summary} Our results so far show that total transport costs have strongly reduced between 1974 and 2019, especially in ocean transport.
+Further, if additive costs represent a substantial share of total transport costs considered on average over the period for both transport modes, a yearly-based inspection reveals marked fluctuations in the share of additive costs in total costs over time.
+In air transport specifically, the share of additive costs has been declining over time since the beginning of the 1990s.
+This suggests that the transport costs decrease is achieved through a more-than-proportional part of additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.
+
+In the next section, we ...
+
+Before turning to the big picture, robustness ...
+
+\textbf{JUSTE CHANGE l'ORGANISATION, PAS LE TEXTE}
+
+\subsection{Robustness analysis \label{sec:robustness}}
+
+\subsubsection{Robustness to the separability assumption}
 In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, which considers (following \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
@@
 
-\subsection{Robustness to the weighting scheme }
-
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \cite{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
-In this section, we provide a robustness test to this difference of treatment.\smallskip
-
-When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
-
-----Pas sûr. Je crois que c’est weighted by value/value total for the year --
-
-In this section, we check whether this difference of treatment is responsible or not for the difference in results relative to the role of the trade composition effects, which we rather attribute to the modeling of the additive component.
-
-To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \cite{hummels2007} weighting scheme (i.e.
-taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
-The results are reported in Figure \ref{fig:compeffects_robustness}.
-
-\begin{figure}[htbp]
-\caption{Transport cost time trends: robustness to the weighting scheme}
-\label{fig:compeffects_robustness}
-\begin{center}
-\includegraphics[height=8cm]
-{graph_composition_all_np.jpg}
-\end{center}
-\end{figure}
-
-The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
-First, differences do show up, in particular for the multiplicative component of air transport.
-This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
-This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
-
 \subsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
@@
 %In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
+
+\section{Additive costs and transport costs time trends}\label{sec:results_trends}
+
+In this section, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
+As underlined by \cite{hummels2007}, the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
+Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
+As emphasized by \cite{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
+Our own results tend to challenge this view, as we further show.
+
+\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
+He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
+Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
+
+\begin{equation}
+TC_{ikt}\equiv \frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} = \widetilde{p}_{ikt}^{-\beta}X_{ikt} \label{eq:Hummel_benchmarkeq}
+\end{equation}
+
+Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value).
+As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem.
+At the other extreme, $\beta = 1$ means that all transport costs are additive.
+Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
+Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
+In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
+Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
+
+\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
+%In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
+We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
+To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e.
+$\beta$, according to:
+\begin{eqnarray*}
+\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
+&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
+\end{eqnarray*}
+
+\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
+We can then rebuild the value of the elasticity on a per product-year-country basis.
+The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
+
+\begin{figure}[htbp]
+\caption{Share of additive costs in total costs: Histogram}
+\label{fig:histogram_beta}
+\begin{center}
+\begin{tabular}{cc}
+%{\small (a) Air } & {\small (b) Vessel}\\
+\includegraphics[width=3.0in, height=2.5in]{"Etude_beta_pondere_(a) Air transport".jpg}
+& \includegraphics[width=3.0in,height=2.5in]{"Etude_beta_pondere_(b) Vessel transport".jpg} \\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
+This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
+Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
+These results are not reported for sake of space, but are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
+
+\subsection{Empirical specification}
+
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
+We start with the presentation of our estimation strategy before turning to the results.\smallskip
+
+\subsubsection{Estimation strategy}
+
+The estimation carried out in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
+Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
+Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+
+One advantage of this method is that it yields measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
+For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
+For sake of reading clarity, we describe the method to extract these series in Appendix \ref{app:comp-effects}, and we now turn to the results.
+
+\subsection{Characterizing the time trends in transport costs}
+Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
+Panels (d), (e) and (f) report the results for vessel.
+In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
+
+\begin{figure}[htbp]
+\caption{Transport costs (with and without composition effects)}
+\label{fig:totalTC_compeffects_excl}
+\begin{center}
+\includegraphics[height=4in]
+{graph_composition_all.jpg}
+\end{center}
+\end{figure}
+
+Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
+First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
+International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
+This stands in line with the related literature (see \citealp{Hummels_1999}; \citealp{Lafourcade_Thisse}).
+Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
+Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
+Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
+For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
+Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
+Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
+Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
+Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+This is particularly the case for the multiplicative component (panel (d)).\smallskip
+
+\subsection{Time trends in transport costs and the modeling of the additive component}
+
+The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
+If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
+This drives us to investigate this difference further.
+As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
+Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
+Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
+%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
+
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
+composition effects excluded ($\widehat{TC}_t$).
+In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \cite{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
+
+\begin{figure}[htbp]
+\caption{Characterizing the time trends: applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
+\begin{center}
+\begin{tabular}{cc}
+{\small (a) Ad-valorem air freight} & {\small (b) Ad-valorem vessel freight}\\
+\multicolumn{2}{c}{{\small In percent of the export price}} \\
+\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
+& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
+{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
+\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
+\includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.jpg}
+& \includegraphics[width=2.5in,height=2in]{figure6_comme_hummels_base100.jpg} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
+Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
+Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
+
+Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction in international transport costs observed over time, in particular in air transport.
+Complementing the findings of Section \ref{sec:results_decomposition}, these results highlight the importance of integrating the additive dimension of international transport costs, with the aim here of characterizing their time trends.
 
@@
 
+\textbf{JE REPORTE ICI LA VERSION DANS LE MAIN TEXT DE LA VERSION SOUMISE. IL FAUT RESUMER.}
+
+\subsubsection{Estimation method for each additive and multiplicative component}
+
+We begin by describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
+\smallskip
+
+\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e.
+extracting from trade composition effect) can be described as a two-stage process.
+First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
+For the estimated ad-valorem component, we estimate the following equation:
+\begin{eqnarray}
+\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\end{eqnarray}
+
+\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
+the series of $\gamma_t$ differs from one estimation to the other).
+Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
+
+%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
+\begin{eqnarray}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\end{eqnarray}
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
+Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
+Both equations are estimated by transport mode.\smallskip
+
+In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
+This constitutes the second stage of our procedure.
+As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we build the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
+\begin{equation}
+\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
+\end{equation}
+
+\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (see details in Appendix \ref{app:comp-effects}).
+In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
+
+As for the additive cost defined in $[0;+\infty]$, we build the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
+\begin{equation}
+\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
+\end{equation}
+
+\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+
+\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
+To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
+
+-------IL FAUT INTRODUIRE LES PRIX FOB. Sinon, les coûts additifs ne sont pas modifiés de l’inflation -----
+
+$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
+
+\subsubsection{Estimation method for the total transport cost measures}
+
+We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
+composition effects excluded).
+As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
+Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
+\smallskip
+
+\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
+Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
+We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
+
+\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
+composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
+In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
+For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
+
+$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
+
+with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
+A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
+
+$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
+
+with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.
+
+We then deduce the fitted value of the overall cost according to:
+$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$
+
+Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
+
+\textbf{STOP HERE POUR LE REPORT DE LA VERSION SOUMISE}
+
 \subsubsection{Additive and ad-valorem transport costs: excluding the composition effects}
@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
+
+\subsection{Robustness to the weighting scheme }
+
+As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \cite{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
+In this section, we provide a robustness test to this difference of treatment.\smallskip
+
+When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
+
+----Pas sûr. Je crois que c’est weighted by value/value total for the year --
+
+In this section, we check whether this difference of treatment is responsible or not for the difference in results relative to the role of the trade composition effects, which we rather attribute to the modeling of the additive component.
+
+To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \cite{hummels2007} weighting scheme (i.e.
+taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
+The results are reported in Figure \ref{fig:compeffects_robustness}.
+
+\begin{figure}[htbp]
+\caption{Transport cost time trends: robustness to the weighting scheme}
+\label{fig:compeffects_robustness}
+\begin{center}
+\includegraphics[height=8cm]
+{graph_composition_all_np.jpg}
+\end{center}
+\end{figure}
+
+The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
+First, differences do show up, in particular for the multiplicative component of air transport.
+This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
+This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
 

### 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
petite modif
Pour exclure 1989-air du tableau 1
chnagement .tex

@@
 
-Our estimation results drive two main conclusions: First, additive costs constitute a significant part of international transports costs, in both maritime and air transport.
-Second, the size of international transport costs has decreased since the mid-1970s, in particular in air transport.
-While the share of additive costs is relatively stable in maritime transport, it has substantially decreased in air transport.
-\textbf{METTRE DES CHIFFRES}.
-
-The question we want to explore is about the implications of such results on theoretical grounds.
-While the whole bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive components.
+\textbf{LIse, 6-09-2021: Je reprends ici}
+
+In this section, we explore the implications of our empirical results on theoretical grounds.
+While the whole bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
 One potential explanation is the lack of empirical estimates about the size of additive costs.
 Our results may thus usefully palliate this limitation.
-Specifically, we can make use of our empirical results to answer two questions: First, what does it change to introduce additive costs along with standard iceberg trade costs, in particular regarding the pattern of trade flows?
-Second, what are the welfare consequences of a reduction in international transport costs, when such a reduction also occurs through a reduction in the additive component?
-These two key questions (the pattern of trade and the welfare gains of trade liberalization) are overwhelmingly addressed considering ad-valorem trade costs only.
-Given our empirical results pointing out the importance of additive costs, we explore how additive costs modify the picture.
-\textbf{A MIEUX DIRE}.
+Specifically, we make use of our empirical results to investigate the welfare consequences of a reduction in international transport costs, when such a reduction also occurs through a reduction in the additive component.
+\cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020} have preceded us on this issue.
+We differentiate from them in the following way.
+\cite{sorensen2014} explores the welfare effects of a trade costs reduction depending on whether they are either fully ad-valorem or fully additive.
+In contrast, we provide a quantification of such welfare gains, as in \cite{Irrazabal_2015} or \cite{Lashkaripour_JIE2020}.
+To do so, we rely on our empirical estimates of the transport costs values, that induce a combined decrease in both the ad-valorem and the additive components over the last decades.
+In a partial-equilibrium model with perfect competition, \cite{Lashkaripour_JIE2020} obtains that welfare gains are not substantially different whether the transport cost decrease implies an additive component or not.
+This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz' type models.
+Anticipating on our results, we also share the result, that the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
+As underlined by \cite{Lashkaripour_JIE2020}, one potential explanation relies on the modeling framework.
+While \cite{Lashkaripour_JIE2020} models a perfectly competitive multi-sector model, we as in \cite{sorensen2014} and \cite{Irrazabal_2015} adopt a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}.
+In this setting, we indeed show that additive costs do modify the selection of firm on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+Put differently, leaving aside the selection issue leads to underestimate the welfare gains that can be achieved by a reduction in the additive component of transport costs.
 
@@
 
-Two papers have explored the role of additive costs on theoretical grounds: \cite{Irrazabal_2015} and \cite{sorensen2014}.
-As in \cite{sorensen2014}, we investigate the role of additive costs in the trade model of \cite{melitz}, considered in its canonical symmetric two-country version.
-\textbf{discuter de pourquoi on ne retient pas Chaney}.
-\textbf{Faut-il citer Lashkaripour ici aussi?} This allows to highlight the specificities induced by additive versus the standard iceberg (ad-valoram) costs.
-In contrast to \cite{sorensen2014}, we make use of our empirical results to quantify the effects of a reduction in international transports costs on both positive and normative dimensions.
-In accordance with the literature, we focus on steady-state analysis.
+As in \cite{sorensen2014}, we adopt the two-country model of \cite{melitz} with monopolistic competition and endogenous firm entry.\footnote{\cite{Irrazabal_2015} adopt the multi-country model of \cite{chaney2008}, consistently with their database that documents the exports of Norwegian firms in various destination markets worldwide.
+We rather choose to retain the \cite{melitz} model, in its simplest form with one sector and two symmetric countries, picking up the average estimated values of transport costs based on our unique importer country, e.g.
+the US.
+Sticking to this canonical set-up allows us to illustrate the role of additive costs in a transparent and intuitive way.} As widely standard in the literature, we focus on steady-state analysis, running comparative statics analysis following a trade cost reduction.
 Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
@@
 
-In each country, the representative household exogenously supplies $L$ units of labour and maximizes utility $\mathcal{U} = \left[ \int_{\omega \in \Omega}c(\omega)^{\frac{\sigma-1}{\sigma}} d\omega \right]^{\frac{\sigma}{\sigma-1}}$, with $\sigma>1$ the elasticity of substitution between varieties.
-She chooses her consumption $c(\omega)$  of each variety $\omega$ within the set $\Omega$ of available varieties (made of both domestic and foreign varieties, endogenously defined), subject to her budget constraint: $\int_{\omega \in \Omega} p(\omega) c(\omega) dw \leq E$, with $p(\omega)$ the consumer price of variety $\omega$ and $E$ nominal expenditures.
+In each country, the representative household exogenously supplies $L$ units of labour and maximizes utility $\mathcal{U} = \left[ \int_{\omega \in \Omega}c(\omega)^{\frac{\sigma-1}{\sigma}} d\omega \right]^{\frac{\sigma}{\sigma-1}}$, with $\sigma>1$ the elasticity of substitution between varieties $\omega$.
+She chooses her consumption $c(\omega)$  within the set $\Omega$ of available varieties (made of both domestic and foreign varieties, endogenously defined), subject to her budget constraint: $\int_{\omega \in \Omega} p(\omega) c(\omega) dw \leq E$, with $p(\omega)$ the consumer price of variety $\omega$ and $E$ nominal expenditures.
 As standard, the optimal demand function for variety $\omega$ is:
@@
 At entry, each firm pays a sunk entry cost $f_e$ and subsequently draws its productivity level $\varphi$ from a known distribution $G(\varphi)$.
-Production of $q$ units, a firm should employ $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
-Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depend on the value traded $\tau\geq 1$ and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
-\textbf{Mettre une footnote, unit vs weight}.
+The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
+Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
 At each point in time, a firm can be hit by a death shock at the exogenous probability $\delta>0$.
@@
 First, we solve for the maximizing profit behavior once entered.
-Second, we determine the entry conditions on the various markets.\medskip
-
-\paragraph{Profit-maximizing behaviors} Once entered, a firm adopts a pricing-to-market behavior (conditional on exporting) by setting a specific price on each destination market.
-
-On the domestic market, the objective is to maximize profit $\pi_d(\omega)$, taken into account the consumer demand function\footnote{In a monopolistic set-up, each firm (of productivity $\varphi$) is the sole producer of variety $\omega$, implying a one-to-one link between $\omega$ and $\varphi$, such that we could write $\omega(\varphi)$.
-For sake of notational simplicity, we omit this link and simply write the consumer demand function as a function of $\varphi$ directly.}:
-\begin{eqnarray*}
-\max_{p(\varphi)} \pi_d(\varphi) &=& \left(p^d(\varphi) -\frac{1}{\varphi}  \right)c(\varphi) \\
-\text{s.t. }&&c(\varphi) = \left(\frac{p(\varphi)}{P}  \right)^{-\sigma} \frac{E}{P}
-\end{eqnarray*}
+Second, we determine the entry conditions on both the domestic and the export markets.\medskip
+
+\paragraph{Profit-maximizing behaviors} Once entered, and conditional on being able to export, a firm adopts a pricing-to-market behavior by setting a specific price on each destination market.
+
+On the domestic market, the objective is to maximize profit $\pi_d(\omega)= \left(p^d(\varphi) -\frac{1}{\varphi}  \right)c(\varphi)$, taken into account the consumer demand function $c(\varphi) = \left(\frac{p(\varphi)}{P}  \right)^{-\sigma} \frac{E}{P}$\footnote{In a monopolistic set-up, each firm (of productivity $\varphi$) is the sole producer of variety $\omega$, implying a one-to-one link between $\omega$ and $\varphi$, such that we could write $\omega(\varphi)$.
+For sake of notational simplicity, we omit this link and simply write the consumer demand function as a function of $\varphi$ directly.}
 The optimal pricing behavior yields:
@@
 
-Conditional on exporting abroad, the profit-maximizing program on the foreign market requires the inclusion of variable international trade costs according to:
+Conditional on exporting abroad, the profit-maximizing program on the export market requires the inclusion of variable international trade costs. The optimizing program accordingly rewrites as: 
 \begin{eqnarray*}
-\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
-\text{s.t. }&&c^x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
+\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p_x(\varphi) -\frac{\tau}{\varphi}  \right)c_x(\varphi) - tc_x(\varphi) \\
+\text{s.t. }&&c_x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
 \end{eqnarray*}
 
-\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
+\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c_x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
 Solving for this yields:
@@
 \pi_d(\varphi)  &=& \frac{r_d(\varphi)}{\sigma} - f = B \varphi^{\sigma-1} - f \label{eq:pid_rd}\\
-\pi_x(\varphi) &=& \frac{r_x(\varphi)}{\sigma} - f_x = B \left[ \frac{\tau}{\varphi} + t\right]^{-(\sigma-1)} - f_x \label{eq:pix_rx}\\
-\text{with}&& B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E \notag
+\pi_x(\varphi) &=& \frac{r_x(\varphi)}{\sigma} - f_x = B \left[ \frac{\tau}{\varphi} + t\right]^{-(\sigma-1)} - f_x \label{eq:pix_rx}
 \end{eqnarray}
-\noindent where $r_d(\varphi)$ and $r_x(\varphi)$ denote the sales revenues on each market.
+\noindent with $B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ and where $r_d(\varphi)$ and $r_x(\varphi)$ denote the sales revenues on each market.
 Profits increase with productivity $\varphi$, and only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
@@
 % \nonumber % Remove numbering (before each equation)
-  r_d(\varphi) &=& \left[\frac{1}{\rho } \frac{1}{\varphi}  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rd}\\
+  r_d(\varphi) &=& \left[\frac{1}{\rho \varphi}  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rd}\\
   r_x(\varphi) &=& \left[\frac{1}{\rho}\left(\frac{\tau}{\varphi} +t\right)  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rx}
 \end{eqnarray}
-\noindent with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup.
+\noindent with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate.
+
 \paragraph{Zero-cutoff profit conditions} Given the fixed costs on both the domestic and the export markets, only firms with a productivity above $\varphi^\ast$ will serve the domestic market, and those with $\varphi>\varphi_x^\ast$ will export abroad.
-The zero cut-off (ZCP) productivity level for successful entry on the domestic and export markets should verify:
+With $\frac{\pi_d(\varphi^\ast)}{\delta}$ and  $\frac{\pi_x(\varphi_x^\ast)}{\delta}$ the expected profit on the domestic and export market respectively, the zero cut-off (ZCP) productivity levels for successful entry on each market should verify:
 \begin{eqnarray*}
-\frac{\pi_d(\varphi^\ast)}{\delta} = 0 &\Leftrightarrow & \frac{r_d(\varphi^\ast)}{\sigma} - f =0\\
-\frac{\pi_x(\varphi_x^\ast)}{\delta} = 0 &\Leftrightarrow & \frac{r_x(\varphi_x^\ast)}{\sigma} - f_x = 0
+\frac{\pi_d(\varphi^\ast)}{\delta} = 0 &\Leftrightarrow & \varphi^\ast = \frac{1}{f^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}\\
+\frac{\pi_x(\varphi_x^\ast)}{\delta} = 0 &\Leftrightarrow & \varphi_x^\ast = \frac{\tau}{f_x^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}} -t}
 \end{eqnarray*}
-\noindent with $\frac{\pi_d(\varphi^\ast)}{\delta}$ the expected profit on each market. Manipulating equations, we get:
-\begin{eqnarray*}
-% \nonumber % Remove numbering (before each equation)
-  \varphi_x^\ast &=& \frac{\tau}{f_x^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}} -t} \\
-  \varphi^\ast &=& \frac{1}{f^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}
-\end{eqnarray*}
-
-The two productivity thresholds are linked according to:
+The two productivity thresholds can be linked according to:
 \begin{equation}
@@
 such that $\varphi_x^\ast > \varphi^\ast$.
-Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as:
-$$\tau+ t\varphi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$$
-
+Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as $\tau+ t\varphi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$.
 Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
-We check this condition holds numerically in solving the model.\textbf{A FAIRE ???
-JE CROYAIS LE FAIRE MAIS JE NE LE FAIS QUE DANS LE CAS SANS ADDITIFS.}
+We check this condition holds numerically in solving the model.
 
@@
 
-Following \cite{melitz}, and decomposing the set of available varieties into $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can rewrite the price index (\ref{eq:CPI}) as:
-\begin{eqnarray*}
-P &=& \left[\int_{\omega \in \Omega_d} p_d^{1-\sigma}(\omega)d\omega + \int_{\omega \in \Omega_x} p_x^{1-\sigma}(\omega)d\omega  \right]^{\frac{1}{1-\sigma}} \\
-&=& \left[\int_{0}^\infty p_d^{1-\sigma}(\varphi)M\mu(\varphi)d\varphi + \int_{0}^\infty p_x^{1-\sigma}(\varphi)M_x \mu_x(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}}
-\end{eqnarray*}
-Making use of Equation (\ref{eq:muofphi}), the price index reads as:
-\begin{eqnarray*}
-P &=& \left[\frac{M}{1-G(\varphi^\ast)}\int_{\phi^\ast}^\infty p_d^{1-\sigma}(\varphi)g(\varphi)d\varphi + \frac{M_x}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast_x}^\infty p_x^{1-\sigma}(\varphi)g(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}} \\
-\text{with}&& \left\{
-\begin{array}{lll}
-p_d(\varphi)&=& \frac{\sigma}{\sigma-1} \frac{1}{\varphi}\\
-p_x(\varphi)&=& \frac{\sigma}{\sigma-1} \left[\frac{\tau}{\varphi} +t\right]
-\end{array}
-\right.
-\end{eqnarray*}
-Accordingly, the price index can be rewritten as:
-$$P = \left(\frac{\sigma}{\sigma-1}\right)\left[\frac{M}{1-G(\varphi^\ast)}\int_{\phi^\ast}^\infty \phi^{\sigma-1}g(\varphi)d\varphi + \frac{M_x}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast_x}^\infty \left[\varphi \left(\frac{\tau}{\tau+t \varphi}  \right)  \right]^{\sigma-1}g(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}}$$
-
-Similarly as in the closed-economy case (\cite{melitz}), let us define $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ as the weighted average of productivities of the domestic firms, on the domestic market and the export market respectively:\footnote{In the symmetric two-country model, $\widetilde{\varphi}_x(\varphi_x^\ast)$ also corresponds to the weighted average of productivities of foreign exporters at Home, e.g.
-associated with the imported varieties.}
-
-\begin{eqnarray}
-\widetilde{\varphi} &=& \left[\frac{1}{1-G(\varphi^\ast)}\int_{\varphi^\ast}^\infty \phi^{\sigma-1}g(\varphi)d\varphi   \right]^{\frac{1}{\sigma-1}} \label{eq:def_phitilde}\\
-\widetilde{\varphi}_x &=& \left[\frac{1}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast}^\infty \left[\varphi\left(\frac{\tau}{\tau+t \varphi}  \right) \right]^{\sigma-1}g(\varphi)d\varphi   \right]^{\frac{1}{\sigma-1}} \label{eq:def_phitildex}
-\end{eqnarray}
-
-Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, it comes that:
+Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, one can show that the consumer price index can be rewritten as: comes that:
 $$P = \frac{\sigma}{\sigma-1}M_T^{\frac{1}{1-\sigma}}\left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$$
 
-With $\widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$ the weighted average of firm productivities on the domestic market (including both local producers and foreign producers exporting at Home), the price index can be expressed as:
-\begin{equation}
-P = M_T^{\frac{1}{1-\sigma}}\frac{1}{\rho \widetilde{\varphi}_T } \label{eq:CPI_vf}
-\end{equation}
-
-\paragraph{Free-entry condition} There is free-entry in the monopolistic industry, imposing that the expected net present value of entry is driven to 0.
+\noindent with $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ the weighted average of productivities of domestic and imported varieties respectively (defined in Appendix \ref{app:theoretical_model}).
+With $\widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$ the weighted average of firm productivities on the domestic market (including both local producers and foreign producers exporting at Home), the price index can be expressed as $P = M_T^{\frac{1}{1-\sigma}}\frac{1}{\rho \widetilde{\varphi}_T }$.
+As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, leading to:
+
+\begin{equation}
+\mathcal{W} = \rho \widetilde{\varphi}_T M_T^{\frac{1}{\sigma-1}}\label{eq:Welfare}
+\end{equation}
+
+\paragraph{Free-entry condition} There is free-entry in the monopolistic industry, imposing that the expected net present value of entry is driven to zero.
 In the open-economy setting, this condition is given by:
@@
 
-\noindent with $\left(1-G(\varphi^\ast)\right)$ the probability of successful entry. In the open-economy setting, $\bar{\pi}$ can be expressed as:
-$$\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$$
-\noindent with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to:
-$$\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$$
+\noindent with $\left(1-G(\varphi^\ast)\right)$ the probability of successful entry.
+$\bar{\pi}$ can be expressed as $\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$, with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{\pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to $\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$.
 
@@
 From this, we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
-Specifically, we will study the consequences of additive costs in terms of welfare.
-As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, the price level $P$ being given by Equation (\ref{eq:CPI_vf}).
 
@@
 
-As standard in this literature (\cite{Irrazabal_2015}, Melitz fial proofs), the heterogenity in firm productivity is assumed to be given by a Pareto distribution.
-In contrast to the standard case with only ad-valorem costs, the model can no longer be solved analytically due to the presence of additive costs (see Appendix \ref{app:theoretical_model}).
-In this case, we rely on the model's simulations to determine the steady state.
+As standard in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the heterogenity in firm productivity is assumed to be given by a Pareto distribution.
+As shown in \cite{melitz-redding-Handbk-IT-2014}, the Pareto distribution indeed allows for a full analytical solving of the M\'{e}litz model.
+This is no longer the case in the presence of additive costs (see Appendix \ref{app:theoretical_model}).
+In this case, we need to rely on the model's simulations to determine the steady state.
 This drives us to calibrate the structural parameters of the model.
-To do so, we rely on values commonly retained in the literature except for the two additive and ad-valorem transport costs, as reported in Table \ref{tab:calib_horsTC}.
+To do so, we dwell on values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
+The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition}.
+
 \begin{table}[htbp]
@@
 \end{center}
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (4\%, based on early 1980s' value) and the final steady state (6\%, based on 2000's value.}}}
 \end{table}
 
-$f$ such that all firms produce in autarky (i.e., such that $\varphi^{\ast,a} = \varphi_{\text{min}}$.
-$f_x$ endogenously revealed such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%\footnote{Though we could not find exact numbers for the share of exporting firms in 1974, this value of 21\% seems reasonable considering values reported for the 1980s by \citet{Lincoln_McCallum2018}} following so the calibration method of \cite{ghironi} or \cite{Irrazabal_2015}.
+Following \cite{ghironi}, $f$ such that all firms produce in autarky (i.e., such that $\varphi^{\ast,a} = \varphi_{\text{min}}$.
+The calibration of the fixed export cost obeys the following logic.
+As in \cite{ghironi}, $f_x$ is endogenously revealed to match a given share of exporting firms at the steady state.
+Further, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
+Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
+Specifically, we endogenously determine the initial value of $f_x$ such that the initial steady state, with 1974 values of transport costs implies a share of exporting firms equal to 4\%, as documented by \cite{BEJK-AER-03}.
+In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 52\% (i.e., from 4\% to 6\%), based on \cite{Lincoln_McCallum2018}.
 
@@
 
-\subsection{Solving the general equilibrium}
-
-Following \cite{melitz} and the subsequent literature, denoting $M_e$ the mass of prospective entrants and $\rho_e$ the probability of successful entry (equal to $1-G(\varphi^\ast)$, considering only steady-state equilibrium needs to verify the entry-exit flow equilibrium condition:
-$$rho_e M_e = \delta M$$.
+\subsection{The model in the general case}
+
+In this section, we provide more details on the model's solving in the general case, i.e. without specifying any functional form for the distribution of firm productivity.
+
+\subsubsection{Aggregation: Price indices}
+
+Following \cite{melitz}, and decomposing the set of available varieties into $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can rewrite the price index (\ref{eq:CPI}) as:
+\begin{eqnarray*}
+P &=& \left[\int_{\omega \in \Omega_d} p_d^{1-\sigma}(\omega)d\omega + \int_{\omega \in \Omega_x} p_x^{1-\sigma}(\omega)d\omega  \right]^{\frac{1}{1-\sigma}} \\
+&=& \left[\int_{0}^\infty p_d^{1-\sigma}(\varphi)M\mu(\varphi)d\varphi + \int_{0}^\infty p_x^{1-\sigma}(\varphi)M_x \mu_x(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}}
+\end{eqnarray*}
+Making use of Equation (\ref{eq:muofphi}), the price index reads as:
+\begin{eqnarray*}
+P &=& \left[\frac{M}{1-G(\varphi^\ast)}\int_{\varphi^\ast}^\infty p_d^{1-\sigma}(\varphi)g(\varphi)d\varphi + \frac{M_x}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast_x}^\infty p_x^{1-\sigma}(\varphi)g(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}} \\
+\text{with}&& \left\{
+\begin{array}{lll}
+p_d(\varphi)&=& \frac{\sigma}{\sigma-1} \frac{1}{\varphi}\\
+p_x(\varphi)&=& \frac{\sigma}{\sigma-1} \left[\frac{\tau}{\varphi} +t\right]
+\end{array}
+\right.
+\end{eqnarray*}
+Accordingly, the price index can be rewritten as:
+$$P = \left(\frac{\sigma}{\sigma-1}\right)\left[\frac{M}{1-G(\varphi^\ast)}\int_{\phi^\ast}^\infty \phi^{\sigma-1}g(\varphi)d\varphi + \frac{M_x}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast_x}^\infty \left[\varphi \left(\frac{\tau}{\tau+t \varphi}  \right)  \right]^{\sigma-1}g(\varphi)d\varphi  \right]^{\frac{1}{1-\sigma}}$$
+
+Similarly as in the closed-economy case (\cite{melitz}), let us define $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ as the weighted average of productivities of the domestic firms, on the domestic market and the export market respectively:\footnote{In the symmetric two-country model, $\widetilde{\varphi}_x(\varphi_x^\ast)$ also corresponds to the weighted average of productivities of foreign exporters at Home, e.g.
+associated with the imported varieties.}
+
+\begin{eqnarray}
+\widetilde{\varphi} &=& \left[\frac{1}{1-G(\varphi^\ast)}\int_{\varphi^\ast}^\infty \phi^{\sigma-1}g(\varphi)d\varphi   \right]^{\frac{1}{\sigma-1}} \label{eq:def_phitilde}\\
+\widetilde{\varphi}_x &=& \left[\frac{1}{1-G(\varphi_x^\ast)}\int_{\varphi^\ast}^\infty \left[\varphi\left(\frac{\tau}{\tau+t \varphi}  \right) \right]^{\sigma-1}g(\varphi)d\varphi   \right]^{\frac{1}{\sigma-1}} \label{eq:def_phitildex}
+\end{eqnarray}
+
+\subsubsection{Solving the general equilibrium}
+
+Following \cite{melitz} and the subsequent literature, denoting $M_e$ the mass of prospective entrants and $\rho_e$ the probability of successful entry (equal to $1-G(\varphi^\ast)$), considering only steady-state equilibrium needs to verify the entry-exit flow equilibrium condition:
+$$\rho_e M_e = \delta M$$.
 
@@
 
-As in most of the related literature (\cite{Irrazabal_2015}, \textbf{Melitz final proofs}, ...), we adopt a Pareto distribution for firm productivity heterogeneity, with the density and the cumulative functions respectively written as:
+As in most of the related literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}, \cite{ghironi}), we adopt a Pareto distribution for firm productivity heterogeneity, with the density and the cumulative functions respectively written as:
 \begin{eqnarray*}
@@
 
-With the Pareto distribution, the system of Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) can be rewritten as a system of four equations in four endogenous variables: $\left\{\bar{pi}, g(\varphi_x^\ast),\varphi_x^\ast, \varphi^\ast  \right\}$ as given by:
+With the Pareto distribution, the system of Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) can be rewritten as a system of four equations in four endogenous variables: $\left\{\bar{\pi}, g(\varphi_x^\ast),\varphi_x^\ast, \varphi^\ast  \right\}$ as given by:
 
@@
 \bar{\pi}&=& \frac{\sigma-1}{k-(\sigma-1)}f + \left(\varphi^\ast \right)^k f_x\left[k\left(\frac{\tau}{\varphi_x^\ast}+t  \right)^{\sigma-1}\left[  1+(k(\varphi_x^\ast)^k g(\varphi_x^\ast))^{\frac{1}{\sigma-1}}\right]^{1-\sigma}  - \left( \varphi_x^\ast \right)^{-k} \right] \label{eq:systemPareto_1}\\
-g(\varphi_x^\ast)&=&\int_{\varphi_x^\ast}^\infty \varphi^{\sigma-1-(k+1)}(\tau +t\varphi)^{1-\sigma}d\varphi \label{eq:systemPareto_2}\\
+g(\varphi_x^\ast)&=&\int_{\varphi_x^\ast}^\infty \varphi^{\sigma-1-(k+1)}(\tau +t\varphi)^{1-\sigma}d\varphi \label{eq:systemPareto_2} \\
 \frac{\varphi_x^\ast}{\tau+ t\varphi_x^\ast}&=&\varphi^\ast \left[ \frac{f}{f_x} \right]^{\frac{1}{1-\sigma}} \label{eq:systemPareto_3}\\

### 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
chnagement .tex

--- 2021-07-27 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
@@
 \hline
-sigma & 4 & \cite{Irrazabal_2015}\\
-fe & 1 & \cite{ghironi}\\
-f & 0.033333 & Share of exporters = 21\% (\cite{ghironi}) \\
-L & 1 & Arbitrary\\
+$\sigma$ & 4 & \cite{Irrazabal_2015}\\
+$f_e$ & 1 & \cite{ghironi}\\
+$f$ & 0.033 & \cite{ghironi} \\
+$L$ & 1 & Arbitrary\\
 $\delta$ & 0.1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
-k & 4 \\
-$\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
+$k$ & 4 \\
+$\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\ \hline
+\multicolumn{3}{l}{Fixed export cost } \\
+$f_x $ & XXX & Share of exporting firms = 4\% (BEJK, 2003) \\
+& XXX& Share of exporting firms = 6\% (\cite{Lincoln_McCallum2018}
 \hline

### 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Note sur 10/3
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
avec le .tex corrigé des bugs

@@
 		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
+		\item 1989 trade by air is excluded
 		%\item The additive term is expressed in fraction of fas price.
@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (4\%, based on early 1980s' value) and the final steady state (6\%, based on 2000's value.}}}
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (4\%, based on early 1980s' value) and the final steady state (6\%, based on 2000's value.)}}}
 \end{table}
@@
 Specifically, we endogenously determine the initial value of $f_x$ such that the initial steady state, with 1974 values of transport costs implies a share of exporting firms equal to 4\%, as documented by \cite{BEJK-AER-03}.
-In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 52\% (i.e., from 4\% to 6\%), based on \cite{Lincoln_McCallum2018}.
+In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 52\% (i.e., from 4\% to 6\%), based on \cite{Lincoln_McCallum2018}.\smallskip
 
@@
 
-\textbf{A DEVELOPPER}
+\textbf{Lise, 06-09-2021: Stop here, changer la calibration}
 

### 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
avec le .tex corrigé des bugs

@@
 $f_x $ & XXX & Share of exporting firms = 4\% (BEJK, 2003) \\
-& XXX& Share of exporting firms = 6\% (\cite{Lincoln_McCallum2018}
+& XXX& Share of exporting firms = 6\% (\cite{Lincoln_McCallum2018}) \\
 \hline

### 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
modification .tex partie modele

@@
 NOT IN ANY APPENDICES.}
+ \textbf{XXXX-GD:4-5 ESTIMATES are in table 1. 3-10 I am not sure... On y fait allusion dans la procédure en IV, c’est sûr.}
 
@@
 such that $\varphi_x^\ast > \varphi^\ast$.
-Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as $\tau+ t\varphi^\ast_x  \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$.
-Notice that, due to additive costs ($t>0$), this condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
-We check this condition holds numerically in solving the model.
+Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as $\left[\tau+ t\varphi^\ast_x \right] \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$.
+Two comments can be made at this point.
+First, additive costs ($t>0$) make the selection condition more likely to happen everything else equal for any given values of ad-valorem and fixed costs.
+Put differently, additive costs somehow relax the constraint of the fixed cost of exporting.
+Second, and most importantly, due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
+We check this condition holds numerically in solving the model.\footnote{This is no longer the case under ad-valorem costs only.
+In this case, with $t=0$, we find the usual condition that $\phi^\ast_x >\phi^\ast$ iif $\tau^{\sigma-1}\frac{f_x}{f}>1$.}
 

### 2021-09-08 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
avancées main text

@@
 We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by a factor ranging between 1.5 and 1.8.
+Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by a factor around 2.
 Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip
@@
 The remainder of the paper is built as follows.
-First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.
-Section \ref{sec:data_method} reports the mean values of the transport costs estimated over the period (by transport mode).
-Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
-Section \ref{sec:results_trends} is devoted to this analysis, while Section \ref{sec:robustness} provides several important robustness checks of our results.
-Section\ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs., and section \ref{sec:conclu} concludes.
+The following section presents the data and the baseline empirical model used to estimate the values of international transport costs.
+Section \ref{sec:basic_results} reports the later on a yearly basis throughout the period 1974-2019 by transport mode, explicitly distinguishing between the additive and the ad-valorem components and providing several important robustness checks of our results.
+In particular, section \ref{sec:basic_results} reports the mean values of the transport costs estimated over the period (by transport mode).
+The two following sections take stock of those results to draw implications for two important debates in economic analysis.
+In section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
+Section\ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs, and section \ref{sec:conclu} concludes.
 
@@
 Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC coding system.
-In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, \textbf{land/ocean (or ``vessel'') POURQUOI LAND ?
-TROUBLANT} and air.
+In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air.
 We will make use of this to highlight potential differences in the magnitude of transport costs across transportation modes.
@@
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Appendix \ref{app:4digit}.
+However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Appendix \ref{app:4digit}).
 \textbf{JH: WHERE ARE THE 3-10/4-5 ESTIMATES?
@@
 
+\section{Overview of transport costs between 1974 and 2019}\label{sec:basic_results}
 \subsection{The importance of the additive component \label{sec:results_decomposition}}
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
+It displays the weighted mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
 Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}
@@
 
-\textbf{ATTENTION AI MODIFIE PAR RAPPORT A AVANT, LIRE LE .tex en commentaire. CLARIFICATION: ON MONTRE BIEN LA MOYENNE PONDEREE, non? }.
 Before jumping to the interpretation, Table \ref{tab:summary_results} calls for two comments of statistical order.
@@
 First, it provides estimation results about the size of the overall transport costs.
-Table \ref{tab:summary_results} shows that transport costs amount to around a 6\% markup in ocean shipping, and 4\% markup in air shipping on average over 1974-2019.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (a) (ad-valorem costs only), this gives a markup of 4.6\% in vessel, and 3.9\% in air shipping (considering the mean value over the period).
+Table \ref{tab:summary_results} shows that transport costs amount to around a 5.5-6\% markup in ocean shipping, and 4\% markup in air shipping on average over 1974-2019.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+Under Model (a) (ad-valorem costs only), this gives a markup of 5.5\% in vessel, and 4.8\% in air shipping (considering the mean value over the period).
 Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau) + \frac{t}{\widetilde{p}}$.
@@
 The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly 1.5/1.7.\smallskip
+For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.\smallskip
 
@@
 We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
-The corresponding values reported in detailed tables are available in the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
+The corresponding values reported in detailed tables are available in section B of the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
 Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
@@
 If the share of additive costs in total costs roughly amounts to 44\% (30\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, with a different pattern across transport modes.
-In maritime transport exhibits, the share of additive costs fluctuates around the average, while it has been steeply declining since the end of the 1980s in air transport.
-
-\paragraph{Summary} Our results so far show that total transport costs have strongly reduced between 1974 and 2019, especially in ocean transport.
+In maritime transport, the share of additive costs fluctuates around the average, while it has been steeply declining since the end of the 1980s in air transport.
+
+\paragraph{Wrap-up} Our results so far show that total transport costs have strongly reduced between 1974 and 2019, especially in ocean transport.
 Further, if additive costs represent a substantial share of total transport costs considered on average over the period for both transport modes, a yearly-based inspection reveals marked fluctuations in the share of additive costs in total costs over time.
 In air transport specifically, the share of additive costs has been declining over time since the beginning of the 1990s.
-This suggests that the transport costs decrease is achieved through a more-than-proportional part of additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.
-
-In the next section, we ...
-
-Before turning to the big picture, robustness ...
-
-\textbf{JUSTE CHANGE l'ORGANISATION, PAS LE TEXTE}
+This suggests that the transport costs decrease is more than proportionally due to additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.
+
+In the following sections, we check the robustness of these results to important checks regarding the separability assumption of country-good fixed effects and potential endogeneity biases.
 
@@
 As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-With this aim, we replicate Figure \ref{fig:histogram_beta} on the reduced sample for both specifications of separable (baseline) and non-separable fixed costs.
-Figure \ref{fig:histogram_beta_robustness} shows the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
+With this aim, Figure \ref{fig:histogram_beta_robustness} shows the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
 
@@
 
-Similarly, as for the whole sample (Figure \ref{fig:histogram_beta}), for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
 Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.
-
-\subsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
+
+\subsubsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).

### 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
note sur la page 14 XXX Jh à GD
actualiser  downward trend in overall transport costs

@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year} for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente} for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.

### 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement .tex sur le modele

@@
 
-\textbf{LIse, 6-09-2021: Je reprends ici}
-
 In this section, we explore the implications of our empirical results on theoretical grounds.
@@
 Anticipating on our results, we also share the result, that the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
-As underlined by \cite{Lashkaripour_JIE2020}, one potential explanation relies on the modeling framework.
+As first explanation, it should be noted that, in \cite{Lashkaripour_JIE2020}'s paper, the case with additive costs relies on a share of additive costs that remains negligible.
+In this context, one might expect that whether additive costs are present or not makes little difference.
+As underlined by \cite{Lashkaripour_JIE2020}, a second explanation may rely on the modeling framework.
 While \cite{Lashkaripour_JIE2020} models a perfectly competitive multi-sector model, we as in \cite{sorensen2014} and \cite{Irrazabal_2015} adopt a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}.
@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (4\%, based on early 1980s' value) and the final steady state (6\%, based on 2000's value.)}}}
+{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.
+BEJK (2003) refers to \cite{BEJK-AER-03}.}}}
 \end{table}
@@
 Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
-Specifically, we endogenously determine the initial value of $f_x$ such that the initial steady state, with 1974 values of transport costs implies a share of exporting firms equal to 4\%, as documented by \cite{BEJK-AER-03}.
-In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 52\% (i.e., from 4\% to 6\%), based on \cite{Lincoln_McCallum2018}.\smallskip
+Specifically, we endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
+In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}.\smallskip
 
@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model oad-valorem costs only.
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model ad-valorem costs only.
 In Columns (2), (4), (6) and (8), we model both ad-valorem and additive costs.
-$t$ is calibrated so that the ratio $t/\widetilde{p}_x$ from the model matches its empirical target, with $widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
-We theoretically define the export fas price through: $p_x = \tau p^{\text{fas}} +t$.
+$t$ is calibrated so that the ratio $t/\widetilde{p}^{\text{fas}}_x$ from the model matches its empirical target, with $\widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
+We theoretically define the export fas price through: $p_x = \tau \widetilde{p}^{\text{fas}} +t$.
 TC = Transport Costs, expressed in percentage of the average fas export price.}}}
@@
 
-\textbf{Lise, 06-09-2021: Stop here, changer la calibration}
-
 \subsection{Trade liberalization: The importance of additive costs}
@@
 Second, total transport costs have strongly reduced over 1974-2019, and the decrease is particularly marked in the additive component in air transport.
-We exploit the theoretical implications of these empirical findings in two different ways.
-
-First, we run a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
-Specifically, we contrast two cases: In the first scenario \textit{à la Mélitz}, we assume that only ad-valorem costs apply.
+We exploit the theoretical implications of these empirical findings by running a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
+Specifically, we contrast two cases: In the first scenario so-called \textit{à la Mélitz}, we assume that only ad-valorem costs apply.
 We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
-In the second scenario \textit{à la Hummels}, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
+In the second scenario so-called \textit{à la Hummels}, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
 The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
-Importantly, we run this comparative statics analysis \textit{ceteris paribus}, notably with the same value of the export fixed cost $f_x$ as in the initial equilibrium (see Table \ref{tab:calib_horsTC}.
-
-Second, we take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported by \citet{Lincoln_McCallum2018} for 1987.
-According to \textbf{???}, this share roughly amounts to 35\% in \textbf{???}.
+In running this exercise, we also take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported in \cite{BEJK-AER-03} (and consistent with the value estimated by \citet{Lincoln_McCallum2018} for 1987 on US data).
+According to \cite{Lincoln_McCallum2018}, this share roughly amounts to 39\% in the mid 2000's.
 This can be matched in the model by resetting the (lower) value of the export fixed cost $f_x$ accordingly.
-In the second exercise, we thus run a comparative statics analysis where we contrast the 2014 steady states (e.g., with the 2019 transport costs values) with the two dimensions of trade liberalization: the reduction of transport costs (as in the first exercise) and the reduction of the export fixed cost, consistent with the increase in the share of exporters from 21\% to 35\%.
-
-The results are reported in Table \ref{tab:resultats_modele}, for Air (Columns (A)-(D)) and Vessel (Columns (E)-(H)) respectively.
+In our benchmark scenario, we thus run a comparative statics analysis under these two dimensions of trade liberalization: the reduction of transport costs along with the reduction of the export fixed cost.
+
+The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.
 Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
-For each transport mode, the first two columns report the results of reducing transport costs only, maintaining the same value of $f_x$ as in the initial equilibrium (Columns (A)-(B) for Air, (E)-(F) for Vessel), contrasting the case of only ad-valorem costs (Columns (A), (D)) and in presence of additive costs (Columns (B), (F)).
-Columns (C)-(D) for Air and (G)-(H) for Vessel report the results when both the variable transport costs and the fixed export costs are lower in the final steady state (2019) relative to the initial one (1974).
+We also run a robustness check of our results by maintaining a fixed export cost value.
+In this case, the welfare changes are solely attributable to the reduction in the variable transport costs.
+These results are reported in Table \ref{tab:resultats_modele}, panel (2).
 
@@
 \begin{center}
-\begin{tabular}{l||c|c|c|c||c|c|c|c}
-\hline
-Mode & \multicolumn{4}{|c||}{Air} & \multicolumn{4}{c}{Vessel} \\ \hline
-$f_x$ final value &  \multicolumn{2}{|c|}{As in 1974} &  \multicolumn{2}{|c||}{Lower} & \multicolumn{2}{|c|}{As in 1974} & \multicolumn{2}{|c}{Lower } \\ \hline
-& (A) & (B) & (C) & (D) & (E) & (F) &(G) & (H) \\
-Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$ & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline
-$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
-$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
-$\Delta TC$ (in pp) &-3.4 &-3.4 &-3.4 &-3.4 &-6.4 &-6.4 &-6.4 &-6.4  \\
-$\Delta \tau$ (in pp) & -3.4	&-1.4&	-3.40&	-1.40	&-6.4&	-3.1&	-6.4&	-3.1  \\
-$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-2.0&	0&	-3.3& 0&	-3.3 \\ \hline
-$\Delta \mathcal{W}$ & 0.0202 &  0.0386 & 0.0362 & 0.0540 & 0.0354 & 0.0535 & 0.0467 & 	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.232&	2.393 &	2.207	&3.348	&2.179	&3.382 &	2.872&	4.00   \\\hline
-Variety effect (\%)&	-0.827 	&-2.599	&0.747
-&-1.038	&-1.323	& -3.123&	-0.126	&-1.902 \\
-Price effect (\%)&	2.084 &5.201	& 1.455 &	4.444&	3.568	& 6.829&	 3.001&	6.059 \\
+\begin{tabular}{l||c|c||c|c}
+\hline
+Mode & \multicolumn{2}{|c||}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
+& (A) & (B) & (C) & (D)  \\
+Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline
+%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
+%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
+$\Delta TC$ (in pp) &-3.4 &-3.4 &-6.4 &-6.4   \\
+$\Delta \tau$ (in pp) & -3.4	&-1.4&-6.4&	-3.1	 \\
+$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-3.3 \\ \hline \hline
+\multicolumn{4}{l}{(1) Lower fixed export cost} \\ \hline
+%$\Delta \mathcal{W}$ & 0.0202 &  0.0386 & 0.0362 & 0.0540 & 0.0354 & 0.0535 & 0.0467 & 	0.064 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.21}	&\textbf{3.35}	&	\textbf{2.87}&	\textbf{4.00 }  \\\hline
+Variety effect (\%)&0.75 &-1.04	&-0.13	&-1.90 \\
+Price effect (\%)&	1.46 &	4.44&		 3.00&	6.06 \\ \hline \hline
+\multicolumn{4}{l}{(2) Constant fixed export cost} \\ \hline
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.39 & 2.18	&3.38 	 \\ \hline
+Variety effect (\%)& -0.83 	&-2.60	&-1.32	& -3.12 \\
+Price effect (\%)&2.08 &5.20	&3.57	& 6.83  \\
 \hline
@@
 \end{center}
-{\parbox[l]{15cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price.
+{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price.
 The reduction in transport costs is expressed in percentage points.
@@
 
-\paragraph{Reduction in international variable transport costs} Based on our empirical estimates, we simulate a reduction in total variable transport costs by 3.4\% in Air transport and 6.4\% in maritime transport, in two different worlds: In a world with only ad-valorem costs (Columns (A) for Air and (D) for Vessel), versus a world with both iceberg and additive transport costs, and where the overall reduction costs comes from both components (Columns (B) for Air and (E) for Vessel).
-
-A first conclusion emerges.
-Whatever the transport mode (and the associated magnitude of the transport costs decrease), trade liberalization induces welfare gains.
-Second, the welfare gains are higher when this reduction is partly achieved through a reduction of additive costs.
-The effect is substantial: The reduction variable transport costs generates a relative welfare gain higher by 1.1 percentage point in Air, 1.2 percentage point in Vessel, when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only.\smallskip
-
-Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
-Considering the inverse of Equation (\ref{eq:CPI_vf}), it is straightforward to get the welfare change as:
-\begin{equation}
-\frac{d\mathcal{W}}{\mathcal{W}_0} = \underbrace{\frac{1}{\sigma-1} \frac{d M_T}{M_{T0}}}_{\text{Variety effect}} +  \underbrace{\frac{d\widetilde{\varphi}_T}{\widetilde{\varphi}_{T0}}}_{\text{Price effect}} \label{eq:decompWelfare}
-\end{equation}
-
-We report the decomposition of the welfare gains in their two components (under the various scenarii) in the last two lines of Table \ref{tab:resultats_modele}, as well as visually through Figure \ref{fig:welfare_decomp}.
-It displays the welfare change and its decomposition between its two components following the reduction in international transport costs, for the decrease estimated in air transport (left column, Panels (1), (1-a) and (1-b)) and in maritime transport (right column, Panels (2), (2-a) and (2-b)).
-All are expressed in percentage change from the initial steady-state value.
-In each panel, the blue left bar displays the case where the trade liberalization occurs through a reduction in ad-valorem costs; the grey right bar reporting the case where the (same) reduction in international transport costs occurs through a combined reduction in both the ad-valorem and the additive components, based on our estimated results in 1974 and 2019 (for each transport mode), as in Table \ref{tab:calib_TC}).
-
-\begin{figure}[htbp]
-\caption{Welfare decomposition}
-\label{fig:welfare_decomp}
-\includegraphics[width=15cm, height=12cm]{Welfare_change_bothmodes.pdf}
-\begin{minipage} [t]  {7in} \footnotesize%
-Notes: All bars are expressed in percentage change from the initial steady-state value.
-\end{minipage}
-\end{figure}
-
-To go deeper into the analysis, we further decompose each variety and competitiveness effects in its various dimensions, as reported in Figures \ref{fig:variety_air_decomp} (Variety effect) and \ref{fig:competitivity_air_decomp} (Competitiveness effect) for air transport.
-Figures \textbf{XXX} in \textbf{Appendix} report similar results for maritime transport.
-
-\begin{figure}[htbp]
-\caption{Variety effect (Air transport)}
-\label{fig:variety_air_decomp}
-\includegraphics[width=12cm, height=10cm]{VarietyEffect_air.pdf}
-%\begin{minipage} [t]  {7in} \footnotesize%
-%Notes: All bars are expressed in percentage change from the initial steady-state value.
-%\end{minipage}
-\end{figure}
-
-\begin{figure}[htbp]
-\caption{Competitiveness effect (Air transport)}
-\label{fig:competitivity_air_decomp}
-\includegraphics[width=12cm, height=10cm]{CompetEffect_air.pdf}
-%\begin{minipage} [t]  {7in} \footnotesize%
-%Notes: All bars are expressed in percentage change from the initial steady-state value.
-%\end{minipage}
-\end{figure}
-
-To understand the changes induced by the additive costs, consider first the benchmark model with only ad-valorem costs, as in \cite{melitz}.
-In this case, the reduction in transport costs induces an increase in selection on the domestic market, but more domestic firms can now export (Figure \ref{fig:competitivity_air_decomp} for Air transport).
-Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases, see Figure \ref{fig:variety_air_decomp}).
-As a whole, the total number of varieties accessible to the consumer reduces (Figure \ref{fig:variety_air_decomp}, $M_T$ decreases slightly).
-Everything else equal, this reduces welfare (Figure \ref{fig:welfare_decomp}).
-Yet, this detrimental effect is counteracted by a price competitiveness effect.
-Due to the tougher selection on the domestic market, the less productive firms no longer enter.
-As a result, the average productivity on the domestic market increases (Figure \ref{fig:competitivity_air_decomp}, $\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase).
-As reported in Figure \ref{fig:welfare_decomp}, the strength of the price competitiveness effect dominates, inducing a increase in welfare following the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.
-
-What does it change when the reduction in international transport costs also occurs through a reduction in its additive component?
-As reported in Figure \ref{fig:welfare_decomp} (right grey bars), trade liberalisation still induces welfare gains, and even larger than in the pure ad-valorem case.
+In line with the literature, we find that trade liberalization induces welfare gains whatever the transport mode (and the associated magnitude of the transport costs decrease).
+The originality of our results rather dwells on comparing the welfare gains depending on the type of cost reduction.
+On that side, the welfare gains are higher when the trade cost reduction occurs through a decrease in the additive component.
+This conclusion stands in accordance with the analytical results of \cite{sorensen2014}.
 This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though.
@@
 Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization.
-Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the effects are sizeable.
+Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable.
+The reduction variable transport costs generates a relative welfare gain higher by around 1.25 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+\cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs.
+In contrast to \cite{Irrazabal_2015} though, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
+\smallskip
 
 Through which mechanism does it occur?
-As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (this can be shown from Equation (\ref{eq:px})).
+Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
+Considering the inverse of Equation (\ref{eq:CPI_vf}), it is straightforward to get the welfare change as:
+\begin{equation}
+\frac{d\mathcal{W}}{\mathcal{W}_0} = \underbrace{\frac{1}{\sigma-1} \frac{d M_T}{M_{T0}}}_{\text{Variety effect}} +  \underbrace{\frac{d\widetilde{\varphi}_T}{\widetilde{\varphi}_{T0}}}_{\text{Price effect}} \label{eq:decompWelfare}
+\end{equation}
+
+We report the decomposition of the welfare gains in their two components in Table \ref{tab:resultats_modele}, under our benchmark scenario (Panel (1)) and under constant fixed export costs (Panel (2)).
+
+To understand the changes induced by the additive costs, we start considering the benchmark model with only ad-valorem costs (Table \ref{tab:resultats_modele}, Columns (A) and (C)).
+As reported in Table \ref{tab:resultats_modele}, the reduction in the iceberg cost induces a negative variety effect, which is yet more than counteracted by a positive competitive effect.
+The reduction in the number of available varieties for the household stands in line with the insights of the canonical \cite{melitz} model.
+With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit.
+This raises the productivity threshold to enter the domestic market.
+Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export.
+As a whole, the total number of varieties accessible to the consumer reduces, which reduces welfare everything else equal.
+
+This detrimental effect is counteracted by a positive price competitiveness effect.
+Due to the tougher selection on the domestic market, the less productive firms no longer enter.
+As a result, the average productivity on the domestic market increases, which pushes the average productivity upwards.
+As reported in Table \ref{tab:resultats_modele}, the strength of the price competitiveness effect dominates, inducing positive welfare gains with the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.
+
+What does it change when the reduction in international transport costs also occurs through a reduction in its additive component?
+As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
 Within the exporters category, this favors the more productive firms, whose market share, e.g.
-revenues increases relatively more, as can be shown from Equation (\ref{eq:rx}).
-Everything else equal, this pushes the average productivity on the export market upwards.\footnote{This mechanism indirectly appears Figure \ref{fig:competitivity_air_decomp}, panel.
-Given our exercise that assumes a combined decrease in both the additive and the ad-valorem costs, at the general equilibrium, the threshold productivity on the export market $\widetilde{\varphi}^\ast_x$ still decreases, but to a lesser extent when the trade liberalization partly occurs through a reduction in additive costs.
-} In contrast to ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
-At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model: As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition.
-Everything else equal, the threshold productivity on the domestic market increases, as well as the average productivity.
+revenues, increases relatively more (see Equation (\ref{eq:rx})).
+Everything else equal, this pushes the average productivity on the export market upwards.
+Unlike the ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
+At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model.
+As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition.
+Everything else equal, the threshold productivity on the domestic market rises, as well as the average productivity.
 Accordingly, the competitiveness effect increases by more when the trade liberalization involves a reduction in the additive component.
 
-The more negative variety effect (Figure \ref{fig:welfare_decomp}) can be understood as a consequence of the above mechanism.
-Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a lower mass of exporters (Figure \ref{fig:variety_air_decomp}).
-As the selection on the domestic market is also strengthened when trade liberalization involves a reduction in additive costs, the mass of active domestic firms reduces by a larger amount.
+The more negative variety effect also results from the above mechanism.
+Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the share of exporters.
+As the selection on the domestic market is also more stringent, the mass of total active firms reduces by a larger amount.
 From the consumer side, this induces a more negative variety effect that is welfare-detrimental everything else equal.
-As for the pure-iceberg case though, the strength of the competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.
-
-\textbf{Faire l'exercice partant d'un monde avec les deux ad-valorem + iceberg, regarder ce qui se passe quand la baisse des coouts totaux est soit soit uniquement iceberg, soit uniquement ad-valorem?
-Plus proche de l'exercice de Sorensen.
-Plus clair sur les effets?
-}
-
-\paragraph{Reduction of the fixed export cost} The above results underline the importance of the additive component of international trade costs.
-When trade liberalization occurs through a reduction in variable costs, not only the welfare-enhancing effects still apply, but they are even larger.
-As mentioned by Mélitz final proofs, the reduction in variable costs is not the only form of trade liberalization, which can typically manifest through a reduction in export fixed costs.
-One might then wonder how additive costs do affect the welfare gains of trade liberalization of this form.
-Indeed, the reduction in the fixed export costs seems empirically relevant to account for the larger share of exporting firms nowadays relative to the 1970s (Based on XXX, from 19\% in XXX to around 35\% in XXX).
-As underlined by \cite{melitz} \textbf{JE CROIS QUE C'est DIT DANS FINAL PROOFS, PAS ECTRA.
-l'a-t-il MONTRE VRAIMENT ANALYTIQUEMENT?
-A VERIFIER}, the reduction of the fixed export cost is also welfare-enhancing, in presence of yet only ad-valorem variable trade costs.
-A natural question then emerges: To what extent is this result modified in presence of additive costs?
-Specifically, are the welfare gains of a fixed export costs decrease larger in presence of additive costs?
-
-The answer can be found in Table \ref{tab:resultats_modele}.
-First, in accordance with \textbf{Mélitz}, the reduction in the fixed export costs adds an extra source of welfare gains.
-In the pure iceberg world, welfare gains amount to 2.2\% when both fixed and variable costs reduce (Column (C) for Air), vs a 1.23\% increase when only ad-valorem costs decrease (Column (A) for Air, a similar result applying for Vessel).
-The larger welfare gains under a combined reduction in the variable and fixed export costs remains in the world with additive costs (Columns (D) vs (B) for Air, (H) vs (E) for Vessel).
-Further, in both worlds (with or without additive costs), the reduction in the fixed export cost induces an extra welfare gains by around 1 percentage point in Air transport, 1.2 percentage point in maritime transport.
+As for the pure-iceberg case though, the strength of the (larger) competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.
+
+As mentioned above, our results dwell on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
+One might yet wonder about the welfare gains induced by the reduction in variable costs only, i.e.
+\textit{ceteris paribus} for a given value of the fixed export cost.
+The answer can be found in Table \ref{tab:resultats_modele}, Panel (2).
+In this case, the fixed export cost is kept constant at its initial value, the only structural change being the decrease in variable transport costs (either ad-valorem only or both additive and ad-valorem).
+Two results emerge.
+First, the welfare gains induced by the trade liberalization are lower in magnitude.
+While this is not a surprising result \textit{per se} (since we only consider the reduction in one source of international frictions, rather than two), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1 percentage point depending on the transport mode.
+Second, the main message of the paper remains valid.
+The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (+ 1.25 percentage point).
 Put it differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
-In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when the reduction in the variable part implies a reduction in the additive component, as observed empirically.
+In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when the reduction in the variable part implies a reduction in the additive component, as observed empirically.\smallskip
 
 In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results point out the importance of taking into the additive components of international trade costs.
-On the positive side, we show that additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods.
-This has important consequences in normative terms.
-Neglecting the role of additive costs in the theoretical models of trade leads to a substantial underestimation of the welfare gains of trade liberalisation.
+Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods.
+This, in turn, has important consequences in normative terms.
 Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined to the reduction in the fixed export costs, has induced an 3.5\% increase in welfare, much higher than if only iceberg costs were modeled (2.4\%).
 A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4\% (as compared to 3.38\% in a pure ad-valorem world).
+In a nutshell, we show that neglecting additive costs leads to a substantial underestimation of the welfare gains of trade liberalisation.
+While this result has been obtained in the canonical \cite{melitz} model, they call for integrating additive costs in New Trade models in view of better assessing the (disputed) question of the welfare gains of trade liberalization.
 

### 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
changement .tex sur le modele

--- 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
@@
 \multicolumn{3}{l}{Fixed export cost } \\
-$f_x $ & XXX & Share of exporting firms = 4\% (BEJK, 2003) \\
-& XXX& Share of exporting firms = 6\% (\cite{Lincoln_McCallum2018}
+Share of exporting firms & 21\% &  \cite{ghironi}, based on BEJK, 2003  \\
+Share of exporting firms & 39\% &  \cite{Lincoln_McCallum2018} \\
 \hline

### 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
New updates in the text

--- 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Univ.
+Lille, CNRS, IESEG School of Management, UMR 9221 - LEM - Lille Économie Management, F-59000 Lille, France, and CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
 University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
@@
 In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
-Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2019, when the degree of classification retained ($s$) is at the 3-digit classification level.
+Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2019, when the degree of classification retained ($s$) is at the 3-digit classification level (products are at 5-digit level).
 Table \ref{tab:result_air_3d_detail} reports the results for air transport.
@@
 
-%\todo{Changer\begin{table}[htbp]
-\begin{table}[htbp]
-  \caption{Air: Transport costs estimates, 3-digit level (selected years)}
-\begin{center}
-    \begin{tabular}{l|ccccccc}
-\hline\hline
-Year & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
-\hline
-\multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}     \\
-\hline
-Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.0 & 4.2 & 3.4  \\
-Median & 5.4 & 3.8 & 4.4 & 2.5 & 2.9 & 3.4 & 2.9  \\
-%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
-\hline
-\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%)}  \\ \hline
-Mean & 3.6 & 2.3 & 2.4 &1.7 & 1.9 & 2.6 & 1.7  \\
-Median & 2.7 & 1.6 & 1.6 & 1.2 & 1.4 & 2.2 & 1.7 \\
-%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
-
-\hline
-\multicolumn{8}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$, in \%)}     \\ \hline
-Mean & 2.6 & 2.0 & 1.8 & 1.3 & 1.5 & 1.1 & 1.0  \\
-Median & 1.1 & 0.5 & 0.8 & 0.5 & 0.6 & 0.4 & 0.5  \\
-%\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020  \\
-\hline
-\# observations & 14,955 & 16,118 & 24,958 & 35,027 & 36,990 & 40,279 & 39,351  \\
-\hline\hline
-\multicolumn{8}{l}{\parbox[l]{15cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-\end{center}
-\label{tab:result_air_3d_detail}
-\end{table}%
-
 \begin{table}[htbp]
 	\centering
-  \caption{Vessel: Transport costs estimates, 3-digit level (selected years)}
-\begin{center}
-    \begin{tabular}{l|ccccccc}
-   \hline\hline
-Year         & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
- \hline
-   \multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}  \\
-   \hline
-Mean  & 9.8 & 6.5 & 5.7 & 5.1 & 5.4 & 4.0 & 3.6  \\
-Median & 9.6 & 5.5 & 4.6 & 4.9 & 5.1  & 3.6 & 3.3  \\
-%Standard Error & 5.3 & 4.0 & 3.2 & 2.8 & 2.0 & 1.8  \\
-\hline
-\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
-\hline
-\multicolumn{8}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} } \\
-\hline
-Mean  & 5.4 & 3.1 & 3.3 & 2.5 & 2.7 & 1.9 & 2.2 \\
-Median & 4.9 & 2.4 & 2.8 & 2.1 & 2.8 & 1.8 & 1.8  \\
-%Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
-\hline
-\multicolumn{8}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)}}  \\
-\hline
-Mean  & 5.1 & 3.4 & 2.7 & 2.8 & 2.9 & 2.5 & 1.5  \\
-Median & 2.9 & 2.3 & 1.7 & 2.2 & 1.9 & 1.9 & 0.8 \\
-%Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\
-\hline
- \# observations & 19,007 & 17,356 & 28,383 & 36,090 & 37,757 & 37,748 & 38,473 \\
-\hline\hline
-\multicolumn{8}{l}{\parbox[l]{15cm}{ \vspace{7pt}\footnotesize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-\end{center} \label{tab:result_ves_3d_detail}
+	\footnotesize{
+	\caption{Air: Transport costs estimates, 3-digit level (selected years)}\vspace{5mm}
+	\label{tab:result_air_3d_detail}%
+	\input{TableA1_air.tex}
+   \begin{tablenotes}
+	\tiny
+	\item Statistics are weighted by value
+	\item \textbf{Model (A): Iceberg transport costs only}
+	\item \textbf{Model (B): With additive and ad-valorem transport costs}
+				\item \textbf{Model (C): With additive transport costs only}
+\end{tablenotes}
+
+}	
+\end{table}%
+
+\begin{table}[htbp]
+	\centering
+	\footnotesize{
+		\caption{Vessel: Transport costs estimates, 3-digit level (selected years)}\vspace{5mm}
+		\label{tab:result_ves_3d_detail}%
+		\input{TableA1_ves.tex}
+		\begin{tablenotes}
+			\tiny
+		\item Statistics are weighted by value
+		\item \textbf{Model (A): Iceberg transport costs only}
+		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\item \textbf{Model (C): With additive transport costs only}
+		\end{tablenotes}
+   }
 \end{table}%
@@
 \centering
-  \caption{Air: Measures of Goodness-of-fit (3 digits)}
 \footnotesize{
-\begin{center}
-    \begin{tabular}{l|cccccc|c}
-    \hline \hline
-    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
-    \multicolumn{8}{l}{\textbf{$R^2$} }\\ \hline
-    Model (A)& 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
-    Model (B) & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
-    \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
-     Model (A) & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
-    Model (B) & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
-   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-    Model (A) & 35675.0 & 41171.0 & 60715.6 & 87492.6 & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{88191.9} & 70498.1 \\
-    Model (B)  & 31387.3 & 35738.4 & 52098.9 & 74954.9 & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{80873.7} & 62285.0 \\ \hline
-    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
-    Model (A) & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
-    Model (B) & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
-    LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
-    Nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
-    p-value & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
-    \hline \hline
- \end{tabular}%
-    \end{center}}
+	\caption{Quality-of-fit diagnostic tests of the three models (Air, 3-digit level)}\vspace{5mm}
 	\label{tab:good_fit_air}%
- \parbox[l]{15cm}{\footnotesize{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& ad-valorem transport costs.
-SER = Standard Error of regression; AIC = Akaike Information Criterion.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.
-The mean statistics are calculated as the average value over all years.
-}}
+	\input{TableA3_air.tex}
+\begin{tablenotes}
+	\tiny
+	\item SER are weighted by value
+	\item \textbf{Model (A): Iceberg transport costs only}
+	\item \textbf{Model (B): With additive and ad-valorem transport costs}
+	\item \textbf{Model (C): With additive transport costs only}
+\end{tablenotes}
+}
 \end{table}%
@@
 	\centering
-  \caption{Vessel: Measures of Goodness-of-fit (3 digits)}
 	\footnotesize{
-\begin{center}
-\begin{tabular}{l|cccccc|c}
-\hline \hline
-Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
-\multicolumn{8}{l}{\textbf{$R^2$} }\\ \hline
-Model (A)& \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & 0.35 & \multicolumn{1}{c}{0.34} & 0.39 \\
-  Model (B) & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & 0.49 & \multicolumn{1}{c}{0.46} & 0.56 \\ \hline
-\multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
-    Model (A) & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
-    Model (B) & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
-   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
-   Model (A) & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
-   Model (B) & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
-    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
-    Model (A) & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
-    Model (B) & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
-    LL ratio & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
-    nb of restrictions & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
-    p-value& \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
-    \hline \hline
-    \end{tabular}%
-    \end{center}}
+		\caption{Quality-of-fit diagnostic tests of the three models (Ves, 3-digit level)}\vspace{5mm}
 		\label{tab:good_fit_vessel}%
-  \parbox[l]{15cm}{\footnotesize{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& ad-valorem transport costs.
-SER = Standard Error of regression; AIC = Akaike Information Criterion.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.
-The mean statistics are calculated as the average value over all years.
-}}
+		\input{TableA3_ves.tex}
+		\begin{tablenotes}
+			\tiny
+			\item SER are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+			\item \textbf{Model (C): With additive transport costs only}
+\end{tablenotes}
+}
 \end{table}%

### 2021-09-10 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
suppression des -

@@
 \beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
+&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}},
 \end{eqnarray*}

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement .tex

--- 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 
-\textbf{JE REPORTE ICI LA VERSION DANS LE MAIN TEXT DE LA VERSION SOUMISE. IL FAUT RESUMER.}
-
 \subsubsection{Estimation method for each additive and multiplicative component}
@@
 First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
+In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).\smallskip
+
 For the estimated ad-valorem component, we estimate the following equation:
@@
 \noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.
-As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
-the series of $\gamma_t$ differs from one estimation to the other).
-Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
-
-%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
-\begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
-\end{eqnarray}
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
-Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
-Both equations are estimated by transport mode.\smallskip
 
@@
 This constitutes the second stage of our procedure.
-As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we build the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
-\begin{equation}
-\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
-\end{equation}
-
-\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (see details in Appendix \ref{app:comp-effects}).
-In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).
-
-As for the additive cost defined in $[0;+\infty]$, we build the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
-\begin{equation}
-\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
-\end{equation}
-
-\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
-
-\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
-To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
-
--------IL FAUT INTRODUIRE LES PRIX FOB. Sinon, les coûts additifs ne sont pas modifiés de l’inflation -----
-
-$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
-
-\subsubsection{Estimation method for the total transport cost measures}
-
-We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
-composition effects excluded).
-As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
-Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
-\smallskip
-
-\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
-Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
-We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
-
-\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
-composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
-In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
-For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
-
-$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$
-
-with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
-A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
-
-$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
-
-with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.
-
-We then deduce the fitted value of the overall cost according to:
-$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$
-
-Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
-
-\textbf{STOP HERE POUR LE REPORT DE LA VERSION SOUMISE}
-
-\subsubsection{Additive and ad-valorem transport costs: excluding the composition effects}
-
-In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for both the ad-valorem and the additive component.
-
-\paragraph{For the ad-valorem component} - Consider first the multiplicative transport cost component.
-Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:
-
-\begin{equation*}
-\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbbm{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbbm{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
-\end{equation*}
-
-Based on this equation, we deduce after estimation that:
+Based on Equation (\ref{eq:compeffects_mult}), we deduce after estimation that:
 
@@
 From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
-Given that $\tau >1$, we can rewrite to get the percentage change between year 1974 and any year $t>1974$:
+Given the constraint $\tau>0$, we rewrite to get the percentage change between year 1974 and any year $t>1974$:
 \begin{equation*}
-\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}
+\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}}{\widehat{\tau}_{is74}} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)}{\widehat{\tau}_{is74}} = 100\exp(\gamma_t)
 \end{equation*}
@@
 As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trend.
-At this stage though, it remains specific to a product-origin country pair.
-Next step is to build the index $\Gamma^{adv}_t$ such that:
+Note that it is independent of the product-origin country pair, such that we can rewrite:
 \begin{equation}
- \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}  \label{eq:tcadv_compoeffect}
+ \Gamma^{adv}_t= 100\exp(\gamma^{adv}_t)  \label{eq:tcadv_compoeffect}
 \end{equation}
-\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
-
-\paragraph{For the additive component} - After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
+
+\noindent where the subscript $^{adv}$ clarifies that this time fixed effect is specific to the ad-valorem component. \medskip
+
+\textbf{REPRENDRE AVEC GUILLAUME. ON ESTIME TAU>1 OU TAU>0? CA CHANGE, CF ANCIENNE VERSION EN ITALIQUE}
+
+%At this stage though, it remains specific to a product-origin country pair.
+%Next step is to build the index $\Gamma^{adv}_t$ such that:
+%\begin{equation}
+% \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)} {\bar{\tau}_{19741}  \label{eq:tcadv_compoeffect}
+%\end{equation}
+%\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
+
+As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
+the series of $\gamma_t$ differs from one estimation to the other).
+Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
+
+%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
+\begin{eqnarray}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\end{eqnarray}
+
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
+Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
+Both equations are estimated by transport mode.\smallskip
+
+After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
 \[\left\{
@@
 From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
-Given the constraint $t>0$, we then obtain the percentage change from 1974 from:
-
+Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)$.
+Again, it is independent of the product-origin country pair.
+We can thus rewrite the time-trend series for the additive transport cost component as:
 \begin{equation}
-\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t) \label{eq:indice_add}
+\Gamma^{add}_t  = 100\exp(\gamma^{add}_t) \label{eq:tcadd_compoeffect}
 \end{equation}
 
-\noindent Note that it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
-\begin{equation}
-\Gamma^{add}_t  = 100\exp(\gamma_t) \label{eq:tcadd_compoeffect}
-\end{equation}
+\noindent where the subscript $^{add}$ clarifies that this time fixed effect is specific to the additive component.
+As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+
+\paragraph{Obtaining the unfitted measures} - The objective here is to express the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
+To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
+
+\textbf{ENLEVER LE TREND LIE A L'INFLATION. POURQUOI SEULEMENT ICI, ET PAS DANS LE FITTED? that exprime en USD aussi. REPRENDRE AVEC GUILLAUME}
+
+$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
+
+\subsubsection{Estimation method for the total transport cost measures}
+
+We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
+composition effects excluded).
+Specifically, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
+Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
+\smallskip
+
+\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
+$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}, \textbf{and $\widetilde{p}_t$ the corresponding average observed fas price}.
+Recall that $\widehat{\tau}^{adv}_t$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
+We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
+%$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$
+
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t  }{\widehat{tc}^{raw}_{1974}}$$
+
+\textbf{POURQUOI UN (-1) AVANT?}
+
+\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
+composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
+In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
+For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
+
+$$\widehat{\tau}^{pure, adv}_t = \bar{\tau}_{1974}\frac{\Gamma^{adv}_t }{100} $$
+
+with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
+A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
+
+$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
+
+with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost. \textbf{POURQUOI PAS tbar 1974? On reconstruit la mesure en USD? PAS CLAIR}
+
+We then deduce the fitted value of the overall cost according to: \textbf{il manque le fas price, je l'ajoute pour la composante additive}
+$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \frac{\widehat{t}^{pure}_t}{\widetilde{p}^{fas}_t}$$
+
+Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
+$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
+
+\textbf{STOP HERE POUR LE REPORT DE LA VERSION SOUMISE}
+
+\subsubsection{Additive and ad-valorem transport costs: excluding the composition effects}
+
+In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for both the ad-valorem and the additive component.
+
+\paragraph{For the additive component} -
 

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Je ne sais pas pourquoi il y a besoin de faire ce commit

@@
 
-To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as $p = (1+\tau) \widetilde{p}+ t$, with $\tau \geq 0$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported.
+To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as $p = \tau \widetilde{p}+ t$, with $\tau \geq 1$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported.
 We can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, leading to:
 \begin{equation}
-p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
+p = \tau \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
 \end{equation}
-Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.
+Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See \textbf{Appendix XXX for a formal demonstration.
 XXXJH: WHERE ???
-XXX}
+XXX A FIARE?}}
 \begin{eqnarray}
-\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau   +\frac{t}{\tilde{p}}} \notag \\
-&=& \frac{-t}{\tau +\widetilde{t}} \label{eq:tt2}
+\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau -1  +\frac{t}{\tilde{p}}} \notag \\
+&=& -\frac{\widetilde{t}}{\tau-1 +\widetilde{t}} \label{eq:tt2}
 \end{eqnarray}
@@
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik}-1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
 \end{equation}
@@
 We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i+ \tau_k \simeq (1+\tau_i)\times(1+\tau_k) -1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+In other words, $\tau_{ik}$ and $t_{ik}$ are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1\simeq (\tau_i-1) + (\tau_k -1) $ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 \begin{eqnarray}
-1+\tau_{ik} &=& (1+ \tau_{i})(1+ \tau_{k}) \label{eq:ad-valorem}\\
+\tau_{ik} &=& \tau_{i} \times\tau_{k} \label{eq:ad-valorem}\\
 t_{ik} &=& t_{i} + t_{k} \label{eq:add}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =(1+\tau_{i})(1+\tau_{k}) -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i}\times\tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
 \end{equation*}
@@
 \begin{equation*}
-\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left((1+\tau_{i})(1+\tau_{k}) -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i}\times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
 \end{equation*}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left((1+\tau_{i})(1+\tau_{k})+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
 \end{equation}
@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left((1+\tau_{i})(1+ \tau_{s(k)})-1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
 \end{equation}
 where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, strictly speaking, the exact equation that we estimate is the following:
-$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
+$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}-1 + \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}\right)+\epsilon_{ik} $$
 After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode).
@@
 
-\section{Overview of transport costs between 1974 and 2019}\label{sec:basic_results}
+\section{Transport costs: Quantifying their size and structure}\label{sec:basic_results}
+
+\textbf{In this section, we... INTRODUIRE CE QU'ON FAIT}
+
 \subsection{The importance of the additive component \label{sec:results_decomposition}}
@@
 Under Model (a) (ad-valorem costs only), this gives a markup of 5.5\% in vessel, and 4.8\% in air shipping (considering the mean value over the period).
-Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau) + \frac{t}{\widetilde{p}}$.
-With $\tau= 0.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\%.
+Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
+With $\tau= 1.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\%.
 A similar calculus gives 4.1\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
@@
 This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3\% and 2.4\% for vessel and air shipping respectively (on average over the period).
-Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as highlighted by the row $\widehat{\beta}$:  \textit{-Share of additive costs}
-This shows that the per-kg cost dimension is quantitatively sizable.
+Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as displayed in the last rows of Table \ref{tab:summary_results}: The per-kg cost dimension is quantitatively sizable.
 \medskip
@@
 \paragraph{Total transport costs: A first look at the trend patterns} Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs expressed in percentage of the fas price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
-Each yearly value is obtained as: $\widehat{\tau}+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2019.
+Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2019.
 
@@
 
-In the following sections, we check the robustness of these results to important checks regarding the separability assumption of country-good fixed effects and potential endogeneity biases.
+In the following section, we check the robustness of these results regarding two important issues, the separability assumption of country-good fixed effects and potential endogeneity biases.
 
@@
 In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, which considers (following \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
-That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $t_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
+That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $\tau_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
 Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approx.
@@
 Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
+%
+%\begin{table}[htbp]
+%  \centering
+%  \caption{Robustness to the separability assumption (average over the period) - restricted sample}
+%\begin{center}
+%    \begin{tabular}{lcc|cc}
+%    \hline \hline
+%    \multicolumn{5}{c}{Mean value over 1974-2019} \\
+% \multicolumn{5}{c}{3 digits} \\    \hline \hline
+%   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
+%   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
+%    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
+%in \%)}  \\
+%    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
+%    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
+%    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
+%    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
+%    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
+%\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
+%    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
+%    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
+% \textit{Diagnostic test} & & &  &  \\
+%    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
+%    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
+%    AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
+%    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
+%\# observations & 2381 & 2381 & 2798 & 2798 \\
+%\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
+%\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
+%    \end{tabular}%
+%\end{center}
+%\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+%The additive term is expressed in fraction of fas price. 1989 omitted in 3-digit estimation for air.}}
+%\label{tab:robustness_separability}
+%\end{table}
 
@@
 	\centering
-  \caption{Robustness to the separability assumption (average over the period) - restricted sample}
+	\caption{Robustness to the separability assumption - restricted sample}
 	\begin{center}
-    \begin{tabular}{lcc|cc}
-    \hline \hline
-    \multicolumn{5}{c}{Mean value over 1974-2019} \\
- \multicolumn{5}{c}{3 digits} \\    \hline \hline
-   Mode  & \multicolumn{2}{c|}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
-   Separability & No  & Yes (baseline) & No & Yes (baseline)\\ \hline
-    \multicolumn{5}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$.
-in \%)}  \\
-    Mean  & 1.76 & 1.96 & 2.61 & 2.99 \\
-    Median &0.65 & 0.84 &1.83 & 2.23 \\ \hline
-    \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}$, in \%)}\\
-    Mean  & 1.04 & 0.94 & 2.43 & 2.40 \\
-    Median & 0.56 & 0.83 & 2.06 & 2.04 \\ \hline
-\multicolumn{5}{l}{\textit{Share of additive costs in total costs (in \%)}} \\
-    Mean  & 62.8  & 67.5  & 51.8  & 55.4 \\
-    Median & 54.1  & 50.5  & 47.0  & 52.2 \\  \hline
- \textit{Diagnostic test} & & &  &  \\
-    %$R^2$    & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.45} \\
-    %SER   & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.72} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.69} \\
-    AIC criteria & 4657.60 & 4654.23 & 10744.83 & 10690.80 \\ \hline
-    %Log-likelihood & \multicolumn{1}{c}{-1757.75} & \multicolumn{1}{c}{-2258.24} & \multicolumn{1}{c}{-4018.51} & \multicolumn{1}{c}{-5220.00} \\ \hline
-\# observations & 2381 & 2381 & 2798 & 2798 \\
-\# origin country & 13.4 & 13.4 & 19.8 & 19.8 \\
-\# products & 25.5 & 25.5 & 53 & 53 \\ \hline \hline
-    \end{tabular}%
-\end{center}
-\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
-The additive term is expressed in fraction of fas price. 1989 omitted in 3-digit estimation for air.}}
-\label{tab:robustness_separability}
-\end{table}
-
-\begin{table}[htbp]
-	\centering
-	\caption{Robustness to the separability assumption (average over the period) - restricted sample}
-	\begin{center}
-			\input{table4.tex}
+			\input{Table4_reshaped.tex}
 	\end{center}
@@
 
+\subsubsection{XXX}
+
+\textbf{Il me semble qu'on avait dit qu'on mettait une 3e robustesse. Non? Faire une transition pour la suite.}
+
 \section{Additive costs and transport costs time trends}\label{sec:results_trends}
@@
 
-Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value).
-As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem.
-At the other extreme, $\beta = 1$ means that all transport costs are additive.
-Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs.
+As already noticed, $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
 Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
@@
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
-To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e.
+To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price on a per product-year-country basis, i.e.
 $\beta$, according to:
 \begin{eqnarray*}
-\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}},
+\beta_{ikt} &\equiv& -\frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
+&=& \frac{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}}{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}+\widehat{\tau}_{is(k)t}-1},
 \end{eqnarray*}
 
-\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-We can then rebuild the value of the elasticity on a per product-year-country basis.
+\textbf{ajoute un s(k) sur nos valeurs estimees}
+
+\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{is(k)t}$ and $\widehat{\tau}_{is(k)t}$).
 The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
@@
 \section{The role of additive cost: Theoretical insights}\label{sec:theory}
+
+\textbf{ATTENTION AJUSTER LES FIGURES ET CHIFFRES A NOUVELLE CALIBRATION}
 
@@
 \begin{eqnarray}
-\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt}-1)&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
   \begin{array}{lcl}
-  \text{For the year 1974:} &&  \widehat{\tau}_{is74} = \exp(\delta +\alpha_i+\beta_s), \\
-  \text{For any year }t> 1974:&& \widehat{\tau}_{ist} = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t) \\
+  \text{For the year 1974:} &&  \widehat{\tau}_{is74} -1 = \exp(\delta +\alpha_i+\beta_s), \\
+  \text{For any year }t> 1974:&& \widehat{\tau}_{ist} -1 = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t) \\
   \end{array}
@@
 
-From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
-Given the constraint $\tau>0$, we rewrite to get the percentage change between year 1974 and any year $t>1974$:
+From this, we obtain the following recursive link: $\widehat{\tau}_{ist} -1 = (\widehat{\tau}_{is74}-1)\exp(\gamma_t)$.
+Given the constraint $\tau>1$, we rewrite to get the percentage change between year 1974 and any year $t>1974$:
 \begin{equation*}
-\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}}{\widehat{\tau}_{is74}} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)}{\widehat{\tau}_{is74}} = 100\exp(\gamma_t)
+\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{(\widehat{\tau}_{is74}-1)\exp(\gamma_t)}{\widehat{\tau}_{is74}} = 100\exp(\gamma_t)
 \end{equation*}
@@
 
-\textbf{REPRENDRE AVEC GUILLAUME. ON ESTIME TAU>1 OU TAU>0? CA CHANGE, CF ANCIENNE VERSION EN ITALIQUE}
+\textbf{REPRENDRE AVEC GUILLAUME. }
 
@@
 %\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
+
+\textbf{STOP HERE}
 

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Transfert en OA des résultats 4-digits de l'annexe C

@@
 Both AIC and log-likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
-These results holds whatever the considered year.\smallskip
-
-For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Appendix \ref{app:4digit}, Tables \ref{tab:good_fit_air_rob} and \ref{tab:good_fit_ves_rob}.
-If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.
-This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
-When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
-In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
-Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.
+These results holds whatever the considered year.
+\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix A.3XXX.
+If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.}
+%This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
+%When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
+%In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
+%Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.
 
@@
 
-\section{Estimation at the 4-digit level \label{app:4digit}}
-
-In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
-
-\subsection{Transport cost estimates}
-
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively.
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Transport costs estimates, selected years, 4-digit level}
-\begin{center}
-  \footnotesize{
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
-\hline
-Mean  & 6.6 & 5.8 & 5.2 & 3.3 & 3.7 & 3.2 \\
-Median & 5.2 & 4.4 & 4.1 & 2.1 & 2.7 & 2.6 \\
-%Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}  \\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
-Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
-Median & 2.5 & 1.7 & 1.9 & 1.0 & 1.7 & 1.4  \\
-%Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$), in \%}    \\ \hline
-Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
-Median & 1.2 & 0.6 & 0.6 & 0.5 & 0.4 & 0.4  \\
-%Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\
-\hline
-\# observations & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-}
-\end{center}
-  \label{tab:result_air_rob}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
-\begin{center}
-  \footnotesize{
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)} \\
-\hline
-Mean  & 9.8 & 6.1 & 5.8 & 5.1 & 4.2 & 3.6  \\
-Median & 9.4 & 5.1 & 4.8 & 4.5 & 3.8 & 3.1 \\
-%Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC} }\\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
-Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
-Median & 4.9 & 3.0 & 2.4 & 2.5 & 2.6 & 1.8 \\
-%Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}^{add}/\widetilde{p}$, in \%) }   \\ \hline
-Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\
-Median & 2.9 & 1.3 & 1.9 & 1.5 & 1.3 & 0.8 \\
-%Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\
-\hline
-\# observations & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-}
-\end{center}
-\label{tab:result_ves_rob}%
-\end{table}%
-
-\subsection{Goodness-of-fit tests at the 4-digit level}
-
-We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
-The results are reported in Tables \ref{tab:good_fit_air_rob} (for air) and \ref{tab:good_fit_ves_rob} (for vessel).
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Measures of goodness of fit, 4-digit level}
-\begin{center}
-\label{tab:good_fit_air_rob}%
-%\vspace{0.5cm}
-%\scalebox{0.97}{
-  \footnotesize{
-\begin{tabular}{l|cccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}              \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
-\hline
-\multicolumn{7}{l}{\textbf{R$^2$}}      \\ \hline
-Model (a)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
-\multicolumn{7}{l}{\textbf{SER} }  \\ \hline
-Model (a)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Model (b) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
-\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (a) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
-Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
-\multicolumn{7}{l}{\textbf{AIC criteria} }  \\ \hline
-Model (a)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
-Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
-\multicolumn{7}{l}{\textbf{Test LL} }    \\ \hline
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
-\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
-p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
-Model (b) = with additive \& ad-valorem
-transport costs.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
-\end{tabular}%
-}
-\end{center}
-\end{table}%
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Measures of goodness of fit, 4-digit level}
-\begin{center}
-\label{tab:good_fit_ves_rob}%
-  \footnotesize{
-\begin{tabular}{l|cccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                   \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
-\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
-Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
-\multicolumn{7}{l}{\textbf{SER} }\\ \hline
-Model (a) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
-\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (a)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
-Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
-\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
-Model (a) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
-Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
-\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
-\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
-p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
-Model (b) = with additive \& ad-valorem
-transport costs.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
-\end{tabular}%
-}
-\end{center}
-\end{table}%
-
-\newpage
-
-\clearpage
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
-\renewcommand{\thetable}{D.\arabic{table}}
-
 \section{Eliminating the composition effects: more details \label{app:comp-effects}}

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Suite comparaison 10/3 - 5/2
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

--- 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 \caption{Transport costs estimates: summary} \vspace{5mm} \label{tab:summary_results}
-	\input{Table1_reshaped.tex}
+	\input{Table1.tex}
 		\begin{tablenotes}
@@
 
-\subsubsection{XXX}
-
-\textbf{Il me semble qu'on avait dit qu'on mettait une 3e robustesse. Non? Faire une transition pour la suite.}
+\subsubsection{Robustness to aggregation}
+
+Current data are available at the 10-digit level for products (HS10).
+Theoretically, they should be available from 1989. 
+But because of the COVID situation, it was only possible to get 1997-1999 and 2001-2019.
+Yet, as \ref{tab: baseline10} shows, the results are similar when using 10-digit product and 5-digit products.
+
+\begin{table}[htbp]
+	\centering
+	\caption{Transport costs estimates: robustess to agregation} \vspace{5mm} \label{tab:baseline10}
+	\input{Table_baseline10.tex}
+	\begin{tablenotes}
+		\tiny
+		\item Statistics for 1997-1999 and 2002-2019
+		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
+		\item \textbf{Model (a): With ad-valorem transport costs only}
+		\item \textbf{Model (b): With additive and ad-valorem transport costs}
+	\end{tablenotes}
+\end{table}
+
+\begin{figure}[htbp]
+	\caption{Transport cost time trends: robustness to agregation}
+	\label{fig:comp_IV_SITC5}
+	\begin{center}
+		\includegraphics[height=8cm]{scatter_chronology_baseline_baseline10.png}
+	\end{center}
+\end{figure}
 

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Sur la comparaison
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Encore des modifs sur le texte

--- 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 52\% and 67\% of total costs on average, versus 40\% to 50\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 300\% to 44\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
-The picture is reversed for medians.
-
-Before jumping to the conclusion that our results are robust to the separability assumption, however we investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
+Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions%, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+The picture is reversed for medians
+.
+
+As a further check, we investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
 Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (a) and for maritime transport in panel (b).
@@
 In line with Table \ref{tab:robustness_separability}, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
-However, they also show that the difference is quantitatively small.
+However, they also show that the difference is quantitatively small in most cases.
 Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
@@
 As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-With this aim, Figure \ref{fig:histogram_beta_robustness} shows the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
+With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
+the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
 
@@
 Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
-In particular, it shows that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation.
+In particular, it shows that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}%\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.}
+ to produce a 2SLS-type of estimation.
 Figure \ref{fig:comp_IV_SITC5} below reports our benchmark estimates by transport mode, together with their instrumented counterpart.
@@
 
-\section{Additive costs and transport costs time trends}\label{sec:results_trends}
+Two main insights emerge up to now.
+First, total transport costs have strongly reduced between 1974 and 2019.
+Second, additive costs represent an important share of those declining transport costs, between 30 and 45\%.
+In the two following sections, we take stock of those results to draw implications for two important debates in economic analysis.
+In the next section, we shed light on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \emph{per se}.
+Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations over time to multiplicative and additive costs.
+
+\section{Transport costs time trends: the key part of additive costs}\label{sec:results_trends}
 
@@
 Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \cite{Hummels_1999} and \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+As emphasized by \cite{Hummels_1999, hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
 Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
@@
 
-The estimation carried out in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
+The estimation carried out in section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
 Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement effets de composition

--- 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 
-\textbf{In this section, we... INTRODUIRE CE QU'ON FAIT}
-
 \subsection{The importance of the additive component \label{sec:results_decomposition}}
@@
 
-\subsubsection{Robustness to aggregation}
-
-Current data are available at the 10-digit level for products (HS10).
-Theoretically, they should be available from 1989. 
-But because of the COVID situation, it was only possible to get 1997-1999 and 2001-2019.
-Yet, as \ref{tab: baseline10} shows, the results are similar when using 10-digit product and 5-digit products.
-
-\begin{table}[htbp]
-	\centering
-	\caption{Transport costs estimates: robustess to agregation} \vspace{5mm} \label{tab:baseline10}
-	\input{Table_baseline10.tex}
-	\begin{tablenotes}
-		\tiny
-		\item Statistics for 1997-1999 and 2002-2019
-		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
-		\item \textbf{Model (a): With ad-valorem transport costs only}
-		\item \textbf{Model (b): With additive and ad-valorem transport costs}
-	\end{tablenotes}
-\end{table}
-
-\begin{figure}[htbp]
-	\caption{Transport cost time trends: robustness to agregation}
-	\label{fig:comp_IV_SITC5}
-	\begin{center}
-		\includegraphics[height=8cm]{scatter_chronology_baseline_baseline10.png}
-	\end{center}
-\end{figure}
+\subsubsection{XXX}
+
+\textbf{Il me semble qu'on avait dit qu'on mettait une 3e robustesse. Non? Faire une transition pour la suite.}
 
@@
 \end{eqnarray*}
-
-\textbf{ajoute un s(k) sur nos valeurs estimees}
 
@@
 
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed exposition of our method and a comparison of Hummel's (2007) methodology.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
 We start with the presentation of our estimation strategy before turning to the results.\smallskip
@@
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
 
@@
 From this, we obtain the following recursive link: $\widehat{\tau}_{ist} -1 = (\widehat{\tau}_{is74}-1)\exp(\gamma_t)$.
-Given the constraint $\tau>1$, we rewrite to get the percentage change between year 1974 and any year $t>1974$:
-\begin{equation*}
-\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{(\widehat{\tau}_{is74}-1)\exp(\gamma_t)}{\widehat{\tau}_{is74}} = 100\exp(\gamma_t)
-\end{equation*}
-
-As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trend.
-Note that it is independent of the product-origin country pair, such that we can rewrite:
+Given the constraint $\tau>1$, we rewrite to get the percentage change between year 1974 and any year $t>1974$: $\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1}  = 100\exp(\gamma_t)$.
+As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the time trend, such that we can rewrite:
 \begin{equation}
@@
 \noindent where the subscript $^{adv}$ clarifies that this time fixed effect is specific to the ad-valorem component. \medskip
-
-\textbf{REPRENDRE AVEC GUILLAUME. }
-
-%At this stage though, it remains specific to a product-origin country pair.
-%Next step is to build the index $\Gamma^{adv}_t$ such that:
-%\begin{equation}
-% \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)} {\bar{\tau}_{19741}  \label{eq:tcadv_compoeffect}
-%\end{equation}
-%\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.
-
-\textbf{STOP HERE}
 
@@
 \noindent where the subscript $^{add}$ clarifies that this time fixed effect is specific to the additive component.
-As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
+However, one should notice that $\Gamma^{add}_t$ has been obtained considering the additive cost in USD, ie subject to the inflation trend.
+To eliminate the influence of inflation in the additive cost index, we extract the time trend (e.g, inflation) from the observed fas price series according to:
+\begin{eqnarray}
+\ln(\widehat{p}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_pfas}
+\end{eqnarray}
+Applying the same logic as above, we get $\Gamma^{\$}_t = 100\exp(\gamma^{\$}_t)$ where the subscript $^{\$}$ clarifies that this time fixed effect is specific to the inflation component of fas prices.
+We then eliminate the inflation trend in the additive cost index through:
+
+\begin{equation}
+\Gamma^{t/\widetilde{p}}_t  = 100 \frac{\Gamma^{add}_t}{\Gamma^{\$}_t} \label{eq:tsptilde_compoeffect}
+\end{equation}
+
+As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{t/\widetilde{p}}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.
 
@@
 
-\textbf{ENLEVER LE TREND LIE A L'INFLATION. POURQUOI SEULEMENT ICI, ET PAS DANS LE FITTED? that exprime en USD aussi. REPRENDRE AVEC GUILLAUME}
-
-$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{add, raw}_t = 100\times\frac{\widehat{t}_t}{\widehat{t}_{1974}}$$
+$$\Gamma^{adv, raw}_t = 100\times\frac{\widehat{\tau}_t}{\widehat{\tau}_{1974}},\qquad \Gamma^{t/\widetilde{p}, raw}_t = 100\times\frac{\widehat{t}_t/\widetilde{p}_t}{\widehat{t}_{1974}/\widetilde{p}_{1974}}$$
+
+\noindent recalling that the estimated series $\widehat{\tau}_t$, $\widehat{t}_t$ and the observed fas price series $\widetilde{p}_t$ are the weighted average series based on the value of each flow in the total value of flows for the relevant year (by transport mode).
 
@@
 
-We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e.
-composition effects excluded).
-Specifically, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
+We also build two measures of the ``total'' transport costs for both the unfitted series and the pure transport cost series (i.e.
+composition effects excluded), with the objective to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
 Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
@@
 \paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
-$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}, \textbf{and $\widetilde{p}_t$ the corresponding average observed fas price}.
-Recall that $\widehat{\tau}^{adv}_t$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
+$$\widehat{TC}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the weighted average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}, and $\widetilde{p}_t$ the corresponding weighted average observed fas price.
 We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
@@
 
-$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t  }{\widehat{tc}^{raw}_{1974}}$$
-
-\textbf{POURQUOI UN (-1) AVANT?}
+$$\Gamma^{tc, raw}_t = 100\frac{\widehat{TC}^{raw}_t  }{\widehat{TC}^{raw}_{1974}}$$
 
@@
 composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
-In a first step then, we rebuild the fitted measures of each transport cost component in value starting from these indices.
-For the ad-valorem component, this can be obtained by rewriting Equation (\ref{eq:tcadv_compoeffect}):
-
-$$\widehat{\tau}^{pure, adv}_t = \bar{\tau}_{1974}\frac{\Gamma^{adv}_t }{100} $$
-
-with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost.
-A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:
-
-$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$
-
-with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost. \textbf{POURQUOI PAS tbar 1974? On reconstruit la mesure en USD? PAS CLAIR}
-
-We then deduce the fitted value of the overall cost according to: \textbf{il manque le fas price, je l'ajoute pour la composante additive}
-$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \frac{\widehat{t}^{pure}_t}{\widetilde{p}^{fas}_t}$$
+In a first step then, starting from these indices, we rebuild the fitted measures of each transport cost component in value (expressed in percentage of the fas price in both cases):
+
+\begin{eqnarray*}
+\widehat{\tau}^{pure, adv}_t -1 &=&\Gamma^{adv}_t \frac{\bar{\tau}_{1974} }{100} \\
+\widehat{t}^{pure}_t &=& \Gamma^{add}_t\frac{\bar{t}_{1974}}{100} 
+\end{eqnarray*}
+
+\noindent with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_s \beta_s)$ the (weighted) mean ad-valorem transport cost in 1974 (based on Equation (\ref{eq:compeffects_mult})) and $\bar{t}_{1974} = \delta + \sum_i \alpha_i + \sum_s \beta_s$ the (weighted) mean additive transport cost in 1974 (based on Equation (\ref{eq:compeffects_add})).
+We then deduce the fitted value of the overall cost according to:
+$$\widehat{TC}^{pure}_t= \widehat{\tau}^{pure, adv}_t -1 + \frac{\widehat{t}^{pure}_t}{\widetilde{p}_t}$$
 
 Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
-$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$
-
-\textbf{STOP HERE POUR LE REPORT DE LA VERSION SOUMISE}
-
-\subsubsection{Additive and ad-valorem transport costs: excluding the composition effects}
-
-In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for both the ad-valorem and the additive component.
-
-\paragraph{For the additive component} -
+$$\Gamma^{tc}_t = 100\frac{\widehat{TC}^{pure}_t}{\widehat{TC}^{pure}_{1974}}$$
 
@@
 
-\noindent where $TC_{ikt}$ is measured by the ratio between (air) charges, i.e.
-$(p_{ikt} - \widetilde{p}_{ikt})q^{fas}_{ikt}$ and the value of the shipment, i.e.
-$\widetilde{p}_{ikt}q^{fas}_{ikt}$, or equivalently : $TC_{ikt} = \frac{p_{ikt} - \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}$.\medskip
-
-After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, that is $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
+\noindent where $TC_{ikt}$ is measured by the ratio between freight charges, i.e.
+$(p_{ikt} - \widetilde{p}_{ikt})q_{ikt}$ and the value of the shipment, i.e.
+$\widetilde{p}_{ikt}q_{ikt}$, or equivalently : $TC_{ikt} = \frac{p_{ikt} - \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}}$.\medskip
+
+After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, denoted $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
 
@@
 \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-We proceed differently, as we weight each flow by its relative value on total trade flows observed in 1974.
-This choice is made for two reasons.
-First, this weighting scheme does not overweight the small flows in value.
-Second, the 1974 weighting scheme makes sense as the starting point of our time trend analysis.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in Section \ref{sec:robustness}.
+We proceed differently, as we weight each flow by its relative value on total trade flows (on the relevant year of observation), based on the argument that this weighting scheme does not overweight the small flows in value.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in the following Section \ref{app:compeffects_weights}.
 
@@
 Coming back to Equation (\ref{eq:Hummels1}), it can be shown that it encompasses the two extreme cases of only ad-valorem costs ($\beta = 0$) and only additive costs ($\beta=1$), as also pointed out in \cite{hummels_skiba}.
-Rewriting Equation (\ref{eq:Hummels1}) in a simpler form as:
+Rewriting the predict of Equation (\ref{eq:Hummels1}) in a simpler form as:
 
@@
 
-\noindent with $\Delta_{ikt} \equiv \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t}$ agglomerates the various fixed effects for reading clarity.
+\noindent where $\Delta_{ikt} \equiv \sum_{i,k}\alpha_{ik}.\mathbbm{1}_{ik}+ \sum_{t}\gamma_{t}.\mathbbm{1}_{t}$ agglomerates the various fixed effects for reading clarity.
 
@@
 &&\ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt}- \ln \widetilde{p}_{ikt} \\
-\Leftrightarrow && \ln (p_{ikt}- \widetilde{p}_{ikt}) = \Delta_{ikt} \\
+\Leftrightarrow && \ln (p_{ikt}- \widetilde{p}_{ikt}) = \Delta_{ikt}
 \end{eqnarray*}
@@
 \end{eqnarray*}
-\noindent with $t_{ikt}$ appropriately defined.
-This exactly corresponds to the case of additive costs only.
+\noindent with $t_{ikt}$ appropriately defined. This exactly corresponds to the case of additive costs only.
 
@@
 
-\noindent with $\tau_{ikt}$ appropriately defined.
-
-This exactly corresponds to the case of ad-valorem transport costs only.
+\noindent with $\tau_{ikt}$ appropriately defined. This exactly corresponds to the case of ad-valorem transport costs only.
 
@@
 allowing for a varying $\beta$ across time, country partner and product).
+We suspect this to be the key driver between Hummels's results and ours regarding the time trend in the trade composition effects.
 To establish this point clearly, we replicate the method adopted by \cite{hummels2007}, outlined above, on our database (which is the same as his until 2004).
@@
 
-\subsection{Robustness to the weighting scheme }
-
-As discussed in Appendix \ref{app:comp-effects}, one source of difference between our methodology and that of \cite{hummels2007} concerns the weighting scheme used to obtain the evolution of the pure transport costs over time.
-In this section, we provide a robustness test to this difference of treatment.\smallskip
+\subsection{Robustness to the weighting scheme \label{app:compeffects_weights}}
 
 When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
-Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974.
-
-----Pas sûr. Je crois que c’est weighted by value/value total for the year --
-
+Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows.
 In this section, we check whether this difference of treatment is responsible or not for the difference in results relative to the role of the trade composition effects, which we rather attribute to the modeling of the additive component.
@@
 
-The comparison of Figure \ref{fig:compeffects_robustness} (applying \citealp{hummels2007}'s weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
+The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummel's weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
 First, differences do show up, in particular for the multiplicative component of air transport.
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
-This confirms that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component).
+This stands in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
+These results confirm that it is the functional form and precisely the modeling of the additive component, rather than the weighting scheme, that is key in understanding the underlying determinants of the overall transport costs time trends, as pointed out in Section \ref{sec:results_trends}.
 

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
petites modfis + nettoyage d'une section en conflit

@@
 
-\subsubsection{XXX}
-
-\textbf{Il me semble qu'on avait dit qu'on mettait une 3e robustesse. Non? Faire une transition pour la suite.}
-
-\section{Additive costs and transport costs time trends}\label{sec:results_trends}
+\subsubsection{Robustness to aggregation}
+
+Current data are available at the 10-digit level for products (HS10).
+Theoretically, they should be available from 1989.
+But because of the COVID situation, it was only possible to get 1997-1999 and 2001-2019.
+Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are similar when using 10-digit product and 5-digit products.
+
+\begin{table}[htbp]
+	\centering
+	\caption{Transport costs estimates: robustess to agregation} \vspace{5mm} \label{tab:baseline10}
+	\input{Table_baseline10.tex}
+	\begin{tablenotes}
+		\tiny
+		\item Statistics for 1997-1999 and 2002-2019
+		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
+		\item \textbf{Model (a): With ad-valorem transport costs only}
+		\item \textbf{Model (b): With additive and ad-valorem transport costs}
+	\end{tablenotes}
+\end{table}
+
+\begin{figure}[htbp]
+	\caption{Transport cost time trends: robustness to agregation}
+	\label{fig:robustness_to_agregation}
+	\begin{center}
+		\includegraphics[height=8cm]{scatter_chronology_baseline_baseline10.png}
+	\end{center}
+\end{figure}
+
+Two main insights emerge up to now.
+First, total transport costs have strongly reduced between 1974 and 2019.
+Second, additive costs represent an important share of those declining transport costs, between 30 and 45\%.
+In the two following sections, we take stock of those results to draw implications for two important debates in economic analysis.
+In the next section, we shed light on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \emph{per se}.
+Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations over time to multiplicative and additive costs.
+
+\section{Transport costs time trends: the key part of additive costs}\label{sec:results_trends}
 

### 2021-09-14 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 
-We would like to thank you for your various insightful comments. They led us to introduce some
+We would like to thank you for your insightful comments. They led us to introduce some
 significant changes to the paper that we hope address your concerns. We first give you an overview
@@
 It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
+\textbf{XXXXX-GD- Bah c’est pas le benchmark, si ?}
 Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
@@
 
-    Switching to $k=10$ as benchmark disaggregation level for the product dimension requires to resort on US Census dataset only.
-As first step, we have re-run the estimation over the years for which we already had the data (2005-2013).
+    Switching to $k=10$ as benchmark disaggregation level for the product dimension would require resorting on US Census dataset only.
+    %As first step, we have re-run the estimation over the years for which we already had the data (2005-2013).
 In parallel, we have bought older years from the Census bureau, as well as the more recent years 2014-2019.
     However, the Census bureau can provide data disaggregated at the $k=10$ product level only starting in 1989.
-This deprives us from the very first years 1974-1988.
+Some of these data are in physical storage and were not accessible because of the Covid-19 pandemic (1989-1996, but also 2000 and 2001).
+In september 2021, they are still not accessible.
+Waiting for them contributed to making this revision quite late.
+This deprives us from the very first years 1974-1996 and 2000-2001.
 For this period, we can only resort on Hummel's dataset, with $k=5$ the finest product level category.
-This notably induces us to maintain $s=3$ as the benchmark disaggregation level for the sectoral dimension for the whole period.
-This allows us to preserve (as far as possible) the largest time coverage - even though switching from $s=3, k=5$ (1974-1988) to $s=3, k=10$) (1989-2019) undoubtedly constitutes a significant break.
-\textbf{What do we do with that?}
+    %This notably induces us to maintain $k=5$ as the benchmark disaggregation level for the sectoral dimension for the whole period.
+    This is a non-negligible cost as we are interested in the long-term evolution of trade costs.
+
+    Comparing the 10/3 ad 5/3 results This allows us to preserve (as far as possible) the largest time coverage.
+
+     % even though switching from $s=3, k=5$ (1974-1988) to $s=3, k=10$) (1989-2019) undoubtedly constitutes a significant break. \textbf{What do we do with that?}
+	\textbf{XXXXXGD - Je crois que c’est caduc puisque nous avons décidé que le benchmark état 5/3. Je ne sais pas qui a écrit "What can we do with that..."}
 

### 2021-09-14 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
New updates in the text

@@
 That justifies our choice to disregard Model (C) in the main text.
+
+\subsection{Estimation at the 4-digit level \label{app:4digit}}
+
+XXX UPDATE XXX In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
+
+\subsubsection{Transport cost estimates}
+
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Air: Transport costs estimates, selected years, 4-digit level}
+\begin{center}
+  \footnotesize{
+    \begin{tabular}{l|cccccc}
+   \hline\hline
+Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
+\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
+\hline
+Mean  & 6.6 & 5.8 & 5.2 & 3.3 & 3.7 & 3.2 \\
+Median & 5.2 & 4.4 & 4.1 & 2.1 & 2.7 & 2.6 \\
+%Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}  \\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
+Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
+Median & 2.5 & 1.7 & 1.9 & 1.0 & 1.7 & 1.4  \\
+%Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$), in \%}    \\ \hline
+Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
+Median & 1.2 & 0.6 & 0.6 & 0.5 & 0.4 & 0.4  \\
+%Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\
+\hline
+\# observations & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its share in trade (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
+\end{tabular}%
+}
+\end{center}
+  \label{tab:result_air_rob}
+\end{table}%
+
+\begin{table}[htbp]
+  \centering
+\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
+\begin{center}
+  \footnotesize{
+    \begin{tabular}{l|cccccc}
+   \hline\hline
+Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)} \\
+\hline
+Mean  & 9.8 & 6.1 & 5.8 & 5.1 & 4.2 & 3.6  \\
+Median & 9.4 & 5.1 & 4.8 & 4.5 & 3.8 & 3.1 \\
+%Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
+\hline
+\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC} }\\ \hline
+\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
+Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
+Median & 4.9 & 3.0 & 2.4 & 2.5 & 2.6 & 1.8 \\
+%Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
+\hline
+\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}^{add}/\widetilde{p}$, in \%) }   \\ \hline
+Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\
+Median & 2.9 & 1.3 & 1.9 & 1.5 & 1.3 & 0.8 \\
+%Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\
+\hline
+\# observations & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+Statistics are obtained weighting each observation by its share in trade (mode-dependent).
+Additive term expressed in fraction of fas price.}}}
+\end{tabular}%
+}
+\end{center}
+\label{tab:result_ves_rob}%
+\end{table}%
+
+\subsubsection{Goodness-of-fit tests at the 4-digit level}
+
+We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years).
+The results are reported in Tables \ref{tab:good_fit_air_rob} (for air) and \ref{tab:good_fit_ves_rob} (for vessel).
+\begin{table}[htbp]
+  \centering
+  \caption{Air: Measures of goodness of fit, 4-digit level}
+\begin{center}
+\label{tab:good_fit_air_rob}%
+%\vspace{0.5cm}
+%\scalebox{0.97}{
+  \footnotesize{
+\begin{tabular}{l|cccccc}
+\hline
+\hline
+      & \multicolumn{6}{c}{Year}              \\
+      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
+\hline
+\multicolumn{7}{l}{\textbf{R$^2$}}      \\ \hline
+Model (a)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
+Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
+\multicolumn{7}{l}{\textbf{SER} }  \\ \hline
+Model (a)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
+Model (b) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
+\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
+Model (a) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
+Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
+\multicolumn{7}{l}{\textbf{AIC criteria} }  \\ \hline
+Model (a)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
+Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
+\multicolumn{7}{l}{\textbf{Test LL} }    \\ \hline
+2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
+\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
+p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
+Model (b) = with additive \& ad-valorem
+transport costs.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
+\end{tabular}%
+}
+\end{center}
+\end{table}%
+
+\begin{table}[htbp]
+  \centering
+  \caption{Vessel: Measures of goodness of fit, 4-digit level}
+\begin{center}
+\label{tab:good_fit_ves_rob}%
+  \footnotesize{
+\begin{tabular}{l|cccccc}
+\hline
+\hline
+      & \multicolumn{6}{c}{Year}                   \\
+      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
+\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
+Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
+Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
+\multicolumn{7}{l}{\textbf{SER} }\\ \hline
+Model (a) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
+Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
+\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
+Model (a)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
+Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
+\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
+Model (a) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
+Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
+\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
+2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
+\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
+p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
+\hline\hline
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
+Model (b) = with additive \& ad-valorem
+transport costs.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
+\end{tabular}%
+}
+\end{center}
+\end{table}%
+
+If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.
+This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
+When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
+In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
+Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.
+
+\newpage
 

### 2021-09-14 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Tavail agregation
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
actulisation chiffres modele

@@
 
-\textbf{ATTENTION AJUSTER LES FIGURES ET CHIFFRES A NOUVELLE CALIBRATION}
-
 In this section, we explore the implications of our empirical results on theoretical grounds.
@@
 \end{center}
-{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.
 BEJK (2003) refers to \cite{BEJK-AER-03}.}}}
@@
 
-The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.
-Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
+The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
+The whole set of results is available to the authors upon request.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
 We also run a robustness check of our results by maintaining a fixed export cost value.
@@
 \multicolumn{4}{l}{(1) Lower fixed export cost} \\ \hline
-%$\Delta \mathcal{W}$ & 0.0202 &  0.0386 & 0.0362 & 0.0540 & 0.0354 & 0.0535 & 0.0467 & 	0.064 \\
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.21}	&\textbf{3.35}	&	\textbf{2.87}&	\textbf{4.00 }  \\\hline
-Variety effect (\%)&0.75 &-1.04	&-0.13	&-1.90 \\
-Price effect (\%)&	1.46 &	4.44&		 3.00&	6.06 \\ \hline \hline
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.50}	&\textbf{3.16}	&	\textbf{3.63}&	\textbf{4.26 }  \\\hline
+Variety effect (\%)&1.36 &0.48	&-0.43	&-1.28 \\
+Price effect (\%)&	1.14 &	2.66&		 4.08&	5.63 \\ \hline \hline
 \multicolumn{4}{l}{(2) Constant fixed export cost} \\ \hline
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.39 & 2.18	&3.38 	 \\ \hline
-Variety effect (\%)& -0.83 	&-2.60	&-1.32	& -3.12 \\
-Price effect (\%)&2.08 &5.20	&3.57	& 6.83  \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.18 & 2.39	&3.38 	 \\ \hline
+Variety effect (\%)& -0.83 	&-1.32	&-2.60	& -3.12 \\
+Price effect (\%)&2.08 &3.56	&5.20	& 6.83  \\
 \hline
@@
 Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable.
-The reduction variable transport costs generates a relative welfare gain higher by around 1.25 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+The reduction variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
 \cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs.
@@
 
-To understand the changes induced by the additive costs, we start considering the benchmark model with only ad-valorem costs (Table \ref{tab:resultats_modele}, Columns (A) and (C)).
-As reported in Table \ref{tab:resultats_modele}, the reduction in the iceberg cost induces a negative variety effect, which is yet more than counteracted by a positive competitive effect.
-The reduction in the number of available varieties for the household stands in line with the insights of the canonical \cite{melitz} model.
+To understand the changes induced by the additive costs, we start considering the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Columns (A) and (C)).
+As can be inferred from Table \ref{tab:resultats_modele}, panel (1), the reduction in international trade costs is driven by positive competitive and variety effects (whatever the transport mode).
+The positive variety effect can be understood as the result of two opposite forces.
+On the one hand, if we consider the reduction in the iceberg cost only (constant $f_x$, as reported in Panel (2)), the variety effect is negative.
 With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit.
@@
 As a whole, the total number of varieties accessible to the consumer reduces, which reduces welfare everything else equal.
-
-This detrimental effect is counteracted by a positive price competitiveness effect.
+The reduction in the number of available varieties for the household stands in line with the insights of the canonical \cite{melitz} model.
+On the other hand, the reduction in the fixed export cost pushes more firms to export everything else equal.
+This increases the number of available imported varieties for the consumer.
+This more than counteracts the reduction in the number of domestic varieties, inducing a positive variety effect when considering both the reduction in $\tau$ and in $f_x$.
+
+The positive price competitiveness effect can be accounted for as follows.
 Due to the tougher selection on the domestic market, the less productive firms no longer enter.
 As a result, the average productivity on the domestic market increases, which pushes the average productivity upwards.
-As reported in Table \ref{tab:resultats_modele}, the strength of the price competitiveness effect dominates, inducing positive welfare gains with the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.
+As reported in Table \ref{tab:resultats_modele}, Panel (1), under the joined reduction in $\tau$ and $f_x$, both the price competitiveness and the variety effects induce positive welfare gains, consistently with the analytical results of \cite{melitz}.
 
 What does it change when the reduction in international transport costs also occurs through a reduction in its additive component?
+As can be inferred from Table \ref{tab:resultats_modele}, the stronger welfare gains of trade liberalization are driven by a more positive competitiveness effect, which more than offsets a variety effect which is now welfare-detrimental, whether $f_x$ remains constant or reduces.
+
+Consider first the price competitiveness effect.
 As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
@@
 Everything else equal, this pushes the average productivity on the export market upwards.
-Unlike the ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
-At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model.
+Unlike pure ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
+At the general equilibrium, this \textit{within-exporters selection effect} also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model.
 As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition.
@@
 
-The more negative variety effect also results from the above mechanism.
-Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the share of exporters.
-As the selection on the domestic market is also more stringent, the mass of total active firms reduces by a larger amount.
-From the consumer side, this induces a more negative variety effect that is welfare-detrimental everything else equal.
-As for the pure-iceberg case though, the strength of the (larger) competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.
+In contrast with the only ad-valorem cost case, the variety effect turns out to be negative, whether $f_x$ is kept constant or reduced.
+This can be accounted for as follows.
+As discussed above, the reduction in additive costs favors the market share of the more productive firms within the set of exporters.
+This makes entry on the export market tougher, thereby moderating the increase in the number of exported varieties in comparison with the pure ad-valorem case.
+This is yet counteracted by the reduction in the export fixed cost, such that the number of exported varieties ultimately raises.
+The negative variety effect is hence mostly attributable to a reduction in the number of domestic incumbents.
+As the selection on the domestic market is more stringent, the mass of total active firms indeed reduces by a larger amount.
+From the consumer side, this induces a negative variety effect that is welfare-detrimental everything else equal.
+The strength of the (larger) competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.\medskip
 
@@
 First, the welfare gains induced by the trade liberalization are lower in magnitude.
-While this is not a surprising result \textit{per se} (since we only consider the reduction in one source of international frictions, rather than two), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1 percentage point depending on the transport mode.
+While this is not a surprising result \textit{per se} (since we only consider the reduction in one source of international frictions, rather than two), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1.25 percentage point depending on the transport mode.
 Second, the main message of the paper remains valid.
-The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (+ 1.25 percentage point).
-Put it differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
-In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when the reduction in the variable part implies a reduction in the additive component, as observed empirically.\smallskip
+The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (+ 1.2 percentage point).
+Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
+In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
 
@@
 This, in turn, has important consequences in normative terms.
-Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined to the reduction in the fixed export costs, has induced an 3.5\% increase in welfare, much higher than if only iceberg costs were modeled (2.4\%).
-A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4\% (as compared to 3.38\% in a pure ad-valorem world).
-In a nutshell, we show that neglecting additive costs leads to a substantial underestimation of the welfare gains of trade liberalisation.
-While this result has been obtained in the canonical \cite{melitz} model, they call for integrating additive costs in New Trade models in view of better assessing the (disputed) question of the welfare gains of trade liberalization.
+Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.16\% increase in welfare, much higher than if only iceberg costs were modeled (2.5\%).
+A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4.3\% (as compared to 3.6\% in a pure ad-valorem world).
+Neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
+While we have obtained these results in the canonical \cite{melitz} model, they call for integrating additive costs in New Trade models in view of better assessing the (disputed) question of the welfare gains of trade liberalization.
 
@@
 From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
+
 Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)$.

### 2021-09-15 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement partie modele + 2-3 points ailleurs

@@
 \end{equation}
-Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See \textbf{Appendix XXX for a formal demonstration.
-XXXJH: WHERE ???
-XXX A FIARE?}}
+Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:
 \begin{eqnarray}
@@
 
-Comparing Equations (\ref{eq:tt1}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs (in relative terms) is also equal to the share of additive costs in total transport costs.
+Comparing Equations (\ref{eq:alternative}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs (in relative terms) is also equal to the share of additive costs in total transport costs.
 
@@
 See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}
-
 \begin{eqnarray}
@@
 As underlined by \cite{Lashkaripour_JIE2020}, a second explanation may rely on the modeling framework.
-While \cite{Lashkaripour_JIE2020} models a perfectly competitive multi-sector model, we as in \cite{sorensen2014} and \cite{Irrazabal_2015} adopt a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}.
-In this setting, we indeed show that additive costs do modify the selection of firm on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+While \cite{Lashkaripour_JIE2020} models a perfectly competitive model, we as in \cite{sorensen2014} and \cite{Irrazabal_2015} adopt a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}.
+In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
 Put differently, leaving aside the selection issue leads to underestimate the welfare gains that can be achieved by a reduction in the additive component of transport costs.
@@
 Put differently, additive costs somehow relax the constraint of the fixed cost of exporting.
-Second, and most importantly, due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
+Second, due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
 We check this condition holds numerically in solving the model.\footnote{This is no longer the case under ad-valorem costs only.
-In this case, with $t=0$, we find the usual condition that $\phi^\ast_x >\phi^\ast$ iif $\tau^{\sigma-1}\frac{f_x}{f}>1$.}
+In this case, with $t=0$, we find the usual condition that $\varphi^\ast_x >\varphi^\ast$ iif $\tau^{\sigma-1}\frac{f_x}{f}>1$.}
 
@@
 
-\paragraph{Summary} Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$.
+\paragraph{Summary} Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, \widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$.
 From this, we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
@@
 The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
+
 In running this exercise, we also take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported in \cite{BEJK-AER-03} (and consistent with the value estimated by \citet{Lincoln_McCallum2018} for 1987 on US data).
@@
 This can be matched in the model by resetting the (lower) value of the export fixed cost $f_x$ accordingly.
-In our benchmark scenario, we thus run a comparative statics analysis under these two dimensions of trade liberalization: the reduction of transport costs along with the reduction of the export fixed cost.
+Our preferred scenario thus embeds the two dimensions of trade liberalization: the reduction of transport costs along with the reduction of the export fixed cost.
 
 The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
-The whole set of results is available to the authors upon request.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
-We also run a robustness check of our results by maintaining a fixed export cost value.
-In this case, the welfare changes are solely attributable to the reduction in the variable transport costs.
+The whole set of results is available upon request to the authors.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
+We also simulate an alternative scenario where only the variable transport costs decrease, the fixed export cost being constant at its initial value.
+This has two purposes.
+It allows us to better highlight the specific mechanisms involved by (a reduction in) additive costs versus the standard iceberg costs.
+In a related manner, it delivers a relevant robustness check on the normative implications of additive costs.
 These results are reported in Table \ref{tab:resultats_modele}, panel (2).
@@
 On that side, the welfare gains are higher when the trade cost reduction occurs through a decrease in the additive component.
-This conclusion stands in accordance with the analytical results of \cite{sorensen2014}.
 This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though.
@@
 Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable.
-The reduction variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+In our preferred scenario, the reduction variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
 \cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs.
@@
 Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
-Considering the inverse of Equation (\ref{eq:CPI_vf}), it is straightforward to get the welfare change as:
+Considering Equation (\ref{eq:Welfare}), it is straightforward to get the welfare change as:
 \begin{equation}
@@
 
-We report the decomposition of the welfare gains in their two components in Table \ref{tab:resultats_modele}, under our benchmark scenario (Panel (1)) and under constant fixed export costs (Panel (2)).
-
-To understand the changes induced by the additive costs, we start considering the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Columns (A) and (C)).
-As can be inferred from Table \ref{tab:resultats_modele}, panel (1), the reduction in international trade costs is driven by positive competitive and variety effects (whatever the transport mode).
-The positive variety effect can be understood as the result of two opposite forces.
-On the one hand, if we consider the reduction in the iceberg cost only (constant $f_x$, as reported in Panel (2)), the variety effect is negative.
-With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit.
-This raises the productivity threshold to enter the domestic market.
-Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export.
-As a whole, the total number of varieties accessible to the consumer reduces, which reduces welfare everything else equal.
-The reduction in the number of available varieties for the household stands in line with the insights of the canonical \cite{melitz} model.
-On the other hand, the reduction in the fixed export cost pushes more firms to export everything else equal.
-This increases the number of available imported varieties for the consumer.
-This more than counteracts the reduction in the number of domestic varieties, inducing a positive variety effect when considering both the reduction in $\tau$ and in $f_x$.
-
-The positive price competitiveness effect can be accounted for as follows.
-Due to the tougher selection on the domestic market, the less productive firms no longer enter.
-As a result, the average productivity on the domestic market increases, which pushes the average productivity upwards.
-As reported in Table \ref{tab:resultats_modele}, Panel (1), under the joined reduction in $\tau$ and $f_x$, both the price competitiveness and the variety effects induce positive welfare gains, consistently with the analytical results of \cite{melitz}.
-
-What does it change when the reduction in international transport costs also occurs through a reduction in its additive component?
-As can be inferred from Table \ref{tab:resultats_modele}, the stronger welfare gains of trade liberalization are driven by a more positive competitiveness effect, which more than offsets a variety effect which is now welfare-detrimental, whether $f_x$ remains constant or reduces.
-
-Consider first the price competitiveness effect.
-As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
-Within the exporters category, this favors the more productive firms, whose market share, e.g.
-revenues, increases relatively more (see Equation (\ref{eq:rx})).
-Everything else equal, this pushes the average productivity on the export market upwards.
-Unlike pure ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
-At the general equilibrium, this \textit{within-exporters selection effect} also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model.
-As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition.
-Everything else equal, the threshold productivity on the domestic market rises, as well as the average productivity.
-Accordingly, the competitiveness effect increases by more when the trade liberalization involves a reduction in the additive component.
-
-In contrast with the only ad-valorem cost case, the variety effect turns out to be negative, whether $f_x$ is kept constant or reduced.
+We report the decomposition of the welfare gains in their two components in Table \ref{tab:resultats_modele}, under our preferred scenario (Panel (1)) and under a constant fixed export cost (Panel (2)).
+
+To understand the changes induced by the additive costs, we start considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (2)).
+In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from the fact that the transport cost decrease involves an additive part or not.
+
+Consider first the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (2), Columns (A) and (C)).
+The reduction in the iceberg cost induces a negative variety effect, which is yet
+more than counteracted by a positive competitive effect. With the reduction in $\tau$, more firms can now export, but this also increases
+the expected average profit. This raises the productivity threshold to enter the domestic
+market. Accordingly, the number of domestic active firms reduces, even if a larger number
+of them is able to export. As a whole, the total number of varieties accessible to the
+consumer reduces, which reduces welfare everything else equal. 
+This detrimental effect is counteracted by a positive price competitiveness effect. Due
+to the tougher selection on the domestic market, the less productive firms no longer enter.
+As a result, the average productivity on the domestic market increases, which pushes
+the average productivity upwards. As reported in Table \ref{tab:resultats_modele}, the strength of the price
+competitiveness effect dominates, inducing positive welfare gains with the reduction in
+iceberg costs, consistently with the analytical results of \cite{melitz}.
+
+What does it change when the reduction in international transport costs also occurs
+through a reduction in its additive component?
+As reported in Table \ref{tab:resultats_modele}, panel (2), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case; and this is driven by a more positive price competitiveness effect and despite the more negative price effect (comparing with Columns (A) and (C), same panel).
 This can be accounted for as follows.
-As discussed above, the reduction in additive costs favors the market share of the more productive firms within the set of exporters.
-This makes entry on the export market tougher, thereby moderating the increase in the number of exported varieties in comparison with the pure ad-valorem case.
-This is yet counteracted by the reduction in the export fixed cost, such that the number of exported varieties ultimately raises.
-The negative variety effect is hence mostly attributable to a reduction in the number of domestic incumbents.
-As the selection on the domestic market is more stringent, the mass of total active firms indeed reduces by a larger amount.
-From the consumer side, this induces a negative variety effect that is welfare-detrimental everything else equal.
-The strength of the (larger) competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.\medskip
-
-As mentioned above, our results dwell on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
-One might yet wonder about the welfare gains induced by the reduction in variable costs only, i.e.
-\textit{ceteris paribus} for a given value of the fixed export cost.
-The answer can be found in Table \ref{tab:resultats_modele}, Panel (2).
-In this case, the fixed export cost is kept constant at its initial value, the only structural change being the decrease in variable transport costs (either ad-valorem only or both additive and ad-valorem).
+
+As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity
+exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
+Within the exporters category, this favors the more productive firms, whose market share,
+e.g. revenues, increases relatively more (see Equation (\ref{eq:rx})). Everything else equal, this
+pushes the average productivity on the export market upwards. Unlike the ad-valorem
+costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of
+the more productive firms. At the general equilibrium, this \textit{within-exporters} selection effect
+also affects the \textit{classic} selection effect at the extensive margin at the heart of the M\'{e}litz
+(2003) model. As the expected export profit increases, it is necessary to be more productive
+to satisfy the free-entry condition. Everything else equal, the threshold productivity
+on the domestic market rises, as well as the average productivity. Accordingly, the
+price competitiveness effect increases by more when the trade liberalization involves a reduction
+in the additive component.
+
+The welfare-dampening impact of the variety effect is also stronger under the scenario \textit{à la Hummels}.
+Since the reduction in additive costs favors the more productive firms in terms of market share,
+entry on the export market is tougher leading to a much more modest increase in the
+share of exporters. As the selection on the domestic market is also more stringent, the
+mass of total active firms reduces by a larger amount. From the consumer side, this induces
+a more negative variety effect that is welfare-detrimental everything else equal. As for the
+pure-iceberg case though, the strength of the (larger) competitiveness effect dominates,
+inducing positive welfare gains, and even larger when the trade liberalization involves a
+reduction in additive costs.\medskip
+
+As mentioned above, our preferred scenario dwells on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
+The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (1).
 Two results emerge.
-First, the welfare gains induced by the trade liberalization are lower in magnitude.
-While this is not a surprising result \textit{per se} (since we only consider the reduction in one source of international frictions, rather than two), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1.25 percentage point depending on the transport mode.
+First, the welfare gains induced by the trade liberalization are stronger in magnitude.
+While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1.25 percentage point depending on the transport mode.
+This can be accounted for as follows.
+The reduction in the fixed export cost pushes more firms to export everything else equal.
+This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market.
+As a result, the joined reduction in variable costs and in $f_x$ lead to a less negative variety effect (which even turns out to be positive in the scenario \textit{à la M\'{e}litz} with only iceberg costs).
+This, combined with a positive price competitiveness effect, accounts for the stronger welfare gains reported in Panel (2).
+
 Second, the main message of the paper remains valid.
-The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (+ 1.2 percentage point).
+The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (around + 1.2 percentage points).
 Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
@@
 Neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
-While we have obtained these results in the canonical \cite{melitz} model, they call for integrating additive costs in New Trade models in view of better assessing the (disputed) question of the welfare gains of trade liberalization.
+While we have obtained these results in the canonical \cite{melitz} model, they call for integrating additive costs in richer New Trade models in view of better assessing the disputed question of the welfare gains of trade liberalization.
 

### 2021-09-18 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
changement texte (papier + annexe en ligne)

@@
 \footnotesize{
-	\caption{Quality-of-fit diagnostic tests of the three models (Air, 3-digit level)}
+	\caption{Quality-of-fit diagnostic tests, Air, 3-digit level}
 	\label{tab:3models_diagnosis_air}%
@@
 	\footnotesize{
-		\caption{Quality-of-fit diagnostic tests of the three models (Ves, 3-digit level)}
+		\caption{Quality-of-fit diagnostic tests, Vessel, 3-digit level}
 		\label{tab:3models_diagnosis_vessel}%
@@
 
-XXX UPDATE XXX In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
+\textbf{XXX UPDATE XXX : Est-ce fait?} 
+In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
 
@@
 Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
+\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
 \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}  \\ \hline
+\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}  \\ \hline
 \multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
-Model (b) = with additive \& ad-valorem
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
+Model (B) = with additive \& ad-valorem
 transport costs.
@@
 In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)), when both additive and ad-valorem costs are modeled (Model (B)) and when only additive costs are modeled (Model (A)).
+In all tables, statistics are weighted by value.
 
@@
 		\input{TableB1974_1987_ves.tex}
-				\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): With ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-			\item \textbf{Model (C): With additive transport costs only}
-		\end{tablenotes}
+				%\begin{tablenotes}
+%			\tiny
+%			\item
+%			\item \textbf{Model (A): With ad-valorem transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%		\end{tablenotes}
 
@@
 
-	\caption{Continued}
+	\caption{Vessel, Yearly estimates, Continued}
 	\begin{center}
@@
 		\input{TableB1988_2001_ves.tex}
-				\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): With ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-			\item \textbf{Model (C): With additive transport costs only}
-		\end{tablenotes}
+				%\begin{tablenotes}
+%			\tiny
+%			\item Statistics are weighted by value
+%			\item \textbf{Model (A): With ad-valorem transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%		\end{tablenotes}
 
@@
 		\input{TableB2002_2015_ves.tex}
-				\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): Iceberg transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-						\item \textbf{Model (C): With additive transport costs only}
-		\end{tablenotes}
+				%\begin{tablenotes}
+%			\tiny
+%			\item Statistics are weighted by value
+%			\item \textbf{Model (A): Iceberg transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%		\end{tablenotes}
 
@@
 		\input{TableB2016_2019_ves.tex}
-				\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): Iceberg transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-			\item \textbf{Model (C): With additive transport costs only}
-		\end{tablenotes}
+			%	\begin{tablenotes}
+%			\tiny
+%			\item Statistics are weighted by value
+%			\item \textbf{Model (A): Iceberg transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%		\end{tablenotes}
 
@@
 		\input{TableB1974_1987_air.tex}
-				\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): With ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-			\item \textbf{Model (C): With additive transport costs only}
-		\end{tablenotes}
-
+				%\begin{tablenotes}
+%			\tiny
+%			\item Statistics are weighted by value
+%			\item \textbf{Model (A): With ad-valorem transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%		\end{tablenotes}
+%		
 	\end{center}
@@
 
-	\caption{Continued}
+	\caption{Air, Vessel, Yearly estimates, Continued}
 	\begin{center}
@@
 		\input{TableB1988_2001_air.tex}
-				\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): With ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-			\item \textbf{Model (C): With additive transport costs only}
-		\end{tablenotes}
-
+				%\begin{tablenotes}
+%			\tiny
+%			\item Statistics are weighted by value
+%			\item \textbf{Model (A): With ad-valorem transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%		\end{tablenotes}
+%		
 	\end{center}
@@
 
-\caption{Continued}
+\caption{Air, yearly estimates, Continued}
 \begin{center}
@@
 	\input{TableB2002_2015_air.tex}
-			\begin{tablenotes}
-		\tiny
-		\item Statistics are weighted by value
-		\\item \textbf{Model (A): With ad-valorem transport costs only}
-		\item \textbf{Model (B): With additive and ad-valorem transport costs}
-		\item \textbf{Model (C): With additive transport costs only}
-	\end{tablenotes}
+			%\begin{tablenotes}
+%		\tiny
+%		\item Statistics are weighted by value
+%		\\item \textbf{Model (A): With ad-valorem transport costs only}
+%		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%		\item \textbf{Model (C): With additive transport costs only}
+%	\end{tablenotes}
 
@@
 	\input{TableB2016_2019_air.tex}
-			\begin{tablenotes}
-		\tiny
-		\item Statistics are weighted by value
-			\item \textbf{Model (A): With ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-			\item \textbf{Model (C): With additive transport costs only}
-	\end{tablenotes}
+			%\begin{tablenotes}
+%		\tiny
+%		\item Statistics are weighted by value
+%			\item \textbf{Model (A): With ad-valorem transport costs only}
+%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+%			\item \textbf{Model (C): With additive transport costs only}
+%	\end{tablenotes}
 

### 2021-09-21 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

@@
 \begin{abstract}
-This paper investigates the pattern of international transport costs over time, using information contained in the US imports flows from 1974 to 2019.
-First, we document the importance of the per-unit component of transport costs.
-It represents between one-third and one-half of overall transport costs.
-Second, allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs over time, with trade composition effects playing a minor role.
-In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
-\textbf{A REPRENDRE - BIG PICTURE}
+This paper demonstrates the magnitude and the role of additive international transport costs.
+Using information contained in the US imports flows from 1974 to 2019, we develop an empirical model that disentangles the ad-valorem and the additive components of international transport costs.
+As main result, we quantify the importance of the per-unit component of transport costs: It represents between 30\% and 45\% of overall transport costs.
+We then emphasize the consequences of modelling additive costs, under two different angles.
+First, modelling time-varying per-unit costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role.
+Second, we evaluate the normative consequences role of trade liberalization in presence of additive costs.
+Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points.
+In both aspects, our results point to the importance of modelling the additive component of international transport costs.
 \end{abstract}
@@
 \noindent \emph{JEL classification}: F10, N70, R40 \\
-\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2019, additive costs, trade composition effects
+\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2019, additive costs, trade composition effects, heterogenous firms
 
@@
 
-Our results may be summarized in two main new findings.
-First, additive transport costs are quantitatively sizable.
+As main result, we show that additive transport costs are quantitatively sizable.
 On average over 1974-2019, the additive cost is estimated to be 1.7\% and 2.8\% of the export price in air and vessel transport respectively (at the 3-digit classification level).
@@
 
-\textbf{A REPRENDRE - BIG PICTURE EN DEUX POINTS }
-
-Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries).} Specifically, trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs.
-This contrasts with the conclusion obtained by \cite{hummels2007} yet also on US import flows.%\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet.
-First, we consider a larger period of time, spanning from 1974 (as in \cite{hummels2007}) to 2019 (while he stops in 2004).
-Second, while Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level for the vast majority of the period (1989-2019).
-For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}
-\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
+A natural question then emerges: To what extent should we care about additive costs? Our paper answers this question from two different sides.
+
+On the empirical side first, we show the importance of incorporating additive costs in the understanding of the international transport costs decrease observed between 1974 and 2019.
+Allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs over time, by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
+changes over time in the baskets of imported goods and/or the origin countries).} Trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
+%\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
 Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
-Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.
+Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
+
+We also show the importance of embedding additive costs when assessing the effects of trade liberalization in normative terms. Specifically, we
+make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: When this reduction solely occurs through a decrease in the ad-valorem cost, as standardly modelled in New Trade models, versus when it is achieved through a combined reduction in the ad-valorem and the additive components, as quantified by our empirical model.
+To this aim, we embed additive costs in the canonical \cite{melitz} model.
+Relying on our empirical estimates of transport costs
+over 1974-2019, our preferred scenario shows an extra welfare gain by around 1.2 percentage point when the transport cost decrease encompasses a reduction in its additive component (as we document empirically) relative to a pure ad-valorem reduction.
+Put differently, neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
 
@@
 Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of \textbf{the ratio of the additive cost to the export price multiplied by the ad-valorem cost LE DIRE PLUS SIMPLEMENT}; by contrast, our estimation strategy enables us to estimate each value of the ad-valorem and the additive costs independently from the other.
+First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to estimate each value of the ad-valorem and the additive costs independently from the other.
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-\textbf{A DEVELOPPER A MINIMA SUR LES IMPLICATIONS EN TERMES DE MODELISATION?}
 While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
-This comparison of results suggests that between a quarter and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).%JH : these results should not change, they are for 2004.
+This comparison of results suggests that between a quarter and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
+%JH : these results should not change, they are for 2004.
 Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2019.
@@
 In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed over the last fifty years.
+\smallskip
 
@@
 They have important implications for both international trade and international macroeconomics models.
-\textbf{Parler des effets de composition comme une des implications}.
-Beyond the positive aspect, some recent papers also point out the normative implications of additive trade costs.
-\cite{sorensen2014} thus analytically shows that important welfare gains can be achieved by reducing additive trade costs.
-Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}).
-The quantitative results we provide in the paper allow making one step in filling that gap.
-%\footnote{ \cite{sorensen2014} extends \cite{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component.
-A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework.
-This is confirmed by \cite{Irrazabal_2015}.}
- %On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models.
-
-More precisely, we use our estimates of multiplicative and additive costs for the years 1974 and 2019 to implement several comparative static exercises investigating the different welfare consequences of alterations to multiplicative and additive costs.
-For a given decrease in transport costs, welfare gains are around 50\% higher when the underlying framework allows for additive costs - which, in practice, concentrate the major part of transport costs decrease over the period.
-In addition, we also assess how the latter results are distorted by changes in sunk costs of exports (i.e., increase in the share of exporting firms).
-Welfare gains are proportionally amplified, with again an additional premium coming from the decrease of additive costs.
+Some recent papers notably point out the normative implications of additive trade costs.
+\cite{sorensen2014} analytically shows that important welfare gains can be achieved by reducing additive trade costs.
+Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs, the one exception being \citealp{Irrazabal_2015}.
+In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade liberalization at stake since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs and by sub-component (additive and ad-valorem).\smallskip
 
@@
 In section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
-Section\ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs, and section \ref{sec:conclu} concludes.
+Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+Section \ref{sec:conclu} concludes.
 
@@
 \caption{Transport costs estimates: summary} \vspace{5mm} \label{tab:summary_results}
-	\input{Table1.tex}
+	%\input{Table1.tex}
+\input{Table1_reshaped.tex}
 		\begin{tablenotes}
-		\tiny
+		\scriptsize
 		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
@@
 		1974, 1981, 1989, 2001, 2009, 2013, 2017 and 2019
-		\item \textbf{Model (a): With ad-valorem transport costs only}
-		\item \textbf{Model (b): With additive and ad-valorem transport costs}
+		\item \textbf{Model (A): With ad-valorem transport costs only}
+		\item \textbf{Model (B): With additive and ad-valorem transport costs}
 	\end{tablenotes}
@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente} for mean vessel transport costs, implying a 50\% decrease in air and a 60\% decrease in vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.
+LP quand je prends les chiffres des tableaux B1 et B2 je ne retrouve pas les graphiques, par ex en air 5.3\% en 1974, sur le graphique quasi 6?} for mean vessel transport costs, implying a \textbf{50\% decrease in air and a 60\% decrease in vessel over the period ACTUALISER, je trouve 55\% environ dans les deux modes, 2.3/5.3 et 4.1/8.9}.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
-For the same years, we obtain a total decrease from 6.9\% to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping.
+For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping.
 Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip
@@
 
-Current data are available at the 10-digit level for products (HS10).
-Theoretically, they should be available from 1989.
-But because of the COVID situation, it was only possible to get 1997-1999 and 2001-2019.
+Current data are available at the 10-digit level for products (HS10) for some specific years.
+The results reported here rely on exploiting data over the years 1997-1999 and 2001-2019.\footnote{Theoretically, they should be available from 1989, which is already smaller than our full sample starting in 1974.
+Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} \textbf{faut-il dire que le sample 10-3 n'est pas exactement le meme que nos baseline results et qu'on refait tourner le 5-3 sur le meme sample exactement?}
+
 Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are similar when using 10-digit product and 5-digit products.
@@
 	\centering
-	\caption{Transport costs estimates: robustess to agregation} \vspace{5mm} \label{tab:baseline10}
-	\input{Table_baseline10.tex}
+	\caption{Transport costs estimates: Robustness to aggregation} \vspace{5mm} \label{tab:baseline10}
+	\input{Table_baseline10_reshaped.tex}
 	\begin{tablenotes}
-		\tiny
+		\scriptsize
 		\item Statistics for 1997-1999 and 2002-2019
 		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
-		\item \textbf{Model (a): With ad-valorem transport costs only}
-		\item \textbf{Model (b): With additive and ad-valorem transport costs}
+        \item Transport costs expressed in \% of the fas price (except for $\widehat{t}$ expressed in USD.
+		%\item \textbf{Model (A): With ad-valorem transport costs only}
+		\item \textbf{Model (B): With additive and ad-valorem transport costs}
 	\end{tablenotes}
@@
 
+\textbf{JE NE COMPRENDS PAS LES RESULTATS EN VESSEL, 5-3 same sample as 10-3, PB?
+c'est un peu court comme commentaire, la comparison en air montre que la part des additifs est tout de meme plus faible, quasiment divisee par 2...
+A REPRENDRE JE PENSE}
+
 \begin{figure}[htbp]
-	\caption{Transport cost time trends: robustness to agregation}
+	\caption{Transport cost time trends: Robustness to aggregation}
 	\label{fig:robustness_to_agregation}
@@
 
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed exposition of our method and a comparison of Hummel's (2007) methodology.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed exposition of our method and a comparison with Hummel's (2007) methodology.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
 We start with the presentation of our estimation strategy before turning to the results.\smallskip
@@
 Our results may thus usefully palliate this limitation.
-Specifically, we make use of our empirical results to investigate the welfare consequences of a reduction in international transport costs, when such a reduction also occurs through a reduction in the additive component.
+Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: When this reduction is achieved through a combined reduction in the ad-valorem and the additive
+components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade models).
+To this aim, we embed additive costs in the canonical M\'{e}litz (2003) model.\smallskip
+
 \cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020} have preceded us on this issue.
 We differentiate from them in the following way.
-\cite{sorensen2014} explores the welfare effects of a trade costs reduction depending on whether they are either fully ad-valorem or fully additive.
+\cite{sorensen2014} explores the welfare effects of a trade costs reduction depending on whether they are either fully ad-valorem or fully additive on analytical grounds.
 In contrast, we provide a quantification of such welfare gains, as in \cite{Irrazabal_2015} or \cite{Lashkaripour_JIE2020}.
-To do so, we rely on our empirical estimates of the transport costs values, that induce a combined decrease in both the ad-valorem and the additive components over the last decades.
 In a partial-equilibrium model with perfect competition, \cite{Lashkaripour_JIE2020} obtains that welfare gains are not substantially different whether the transport cost decrease implies an additive component or not.
@@
 \cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs.
-In contrast to \cite{Irrazabal_2015} though, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
+If we share a similar conclusion regarding the importance of additive costs, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
 \smallskip
@@
 
-This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices.
-Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2019, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
-Two main findings emerge from our results.
-First, we provide a quantitative measure of both the additive and the ad-valorem transport costs.
-We find that additive costs are quantitatively sizable: They amount to 2.8\% of the export price unit values for ocean shipping, and 1.8\% for air transport.
-This is slightly lower that the estimated values for the ad-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2019.
-Further, modeling the additive component significantly improves the quality of fit of the regression.
-Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs.
+This paper assesses the magnitude and the role of additive international transport costs.
+Empirically, we exploit the differences between the import and the export prices taken from the US import database over 1974-2019, to estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
+We find that additive costs are quantitatively sizable: On average over the period, they amount to 2.8\% of the export price unit values for ocean shipping, and 1.7\% for air transport, accounting for 30\% and 45\% of total transport costs in air and maritime transport respectively.
+
+To what extent should we care about additive costs?
+Our paper addresses this question through two different angles.
+First, we show the importance of integrating the additive component in accounting for the time trend of international transport costs observed over 1974-2019.
 Allowing for a varying share of additive costs in the product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
 If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in maritime transport.
+
+Second, we revisit the question of the welfare gains from trade liberalisation to shed light on the role of additive costs on this issue.
+To this aim, we amend the canonical M\'{e}litz (2003) model with additive costs.
+Relying on our empirical estimates of transport
+costs over 1974-2019, our preferred scenario shows an extra welfare gain by around 1.2
+percentage point when the transport cost decrease encompasses a reduction in its additive
+component (as we document empirically) relative to a pure ad-valorem reduction.
 In both aspects, our results highlight the importance of the additive component in accounting for international transport costs.
@@
 identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices.
-On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows in an international trade theory perspective and in affecting the international transmission of business cycles.
+On the theoretical side, our results call for exploring the role of additive costs in shaping international trade flows in richer New Trade models than the canonical M\'{e}litz model used so far, and in affecting the international transmission of business cycles.
 This is left for further research.
@@
 These results holds whatever the considered year.
-\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix A.3XXX.
+\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix A.3 \textbf{CHECK XXX}.
 If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.}
@@
 \begin{eqnarray}
-\bar{\pi}&=& \frac{\sigma-1}{k-(\sigma-1)}f + \left(\varphi^\ast \right)^k f_x\left[k\left(\frac{\tau}{\varphi_x^\ast}+t  \right)^{\sigma-1}\left[  1+(k(\varphi_x^\ast)^k g(\varphi_x^\ast))^{\frac{1}{\sigma-1}}\right]^{1-\sigma}  - \left( \varphi_x^\ast \right)^{-k} \right] \label{eq:systemPareto_1}\\
+\bar{\pi}&=& \kappa f + \left(\varphi^\ast \right)^k f_x\left[k\left(\frac{\tau}{\varphi_x^\ast}+t  \right)^{\sigma-1}\left[  1+(k(\varphi_x^\ast)^k g(\varphi_x^\ast))^{\frac{1}{\sigma-1}}\right]^{1-\sigma}  - \left( \varphi_x^\ast \right)^{-k} \right] \label{eq:systemPareto_1}\\
 g(\varphi_x^\ast)&=&\int_{\varphi_x^\ast}^\infty \varphi^{\sigma-1-(k+1)}(\tau +t\varphi)^{1-\sigma}d\varphi \label{eq:systemPareto_2} \\
@@
 
-Notice that, in the absence of additive costs ($t=0$), Equation (\ref{eq:systemPareto_2}) rewrites as:
+\noindent with $\kappa \equiv \frac{\sigma-1}{k-(\sigma-1)}$. Notice that, in the absence of additive costs ($t=0$), Equation (\ref{eq:systemPareto_2}) rewrites as:
 $$g(\varphi_x^\ast) = \tau^{1-\sigma}\left[ \frac{(\varphi_x^\ast)^{-(k-(\sigma-1))}}{k-(\sigma-1)}\right]$$

### 2021-09-21 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex

--- 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
@@
 \multicolumn{3}{l}{Fixed export cost } \\
-$f_x $ & XXX & Share of exporting firms = 4\% (BEJK, 2003) \\
-& XXX& Share of exporting firms = 6\% (\cite{Lincoln_McCallum2018}
+$f_x $ & 0.08 & Share of exporting plants = 21\% (\cite{ghironi}, based on BEJK, 2003) \\
+& 0.06& Share of exporting plants = 39\% (\cite{Lincoln_McCallum2018}) \\
 \hline

### 2021-09-21 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

--- 2021-08-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
@@
 \begin{itemize}
-\item Based on your comment 1.2 (\textit{The aggregation problem}), the benchmark set-up now retains $s=3$ and $k=10$ for the industry/product aggregation levels respectively.
-It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
+%\item Based on your comment 1.2 (\textit{The aggregation problem}), we have investigated the case with $s=3$ and $k=10$ for the industry/product aggregation levels respectively.
 Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
-We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
-In line with the referee's comment, we now run the estimation by considering $k$ at the HS10 classification level, not 5.
-\textbf{We come back to this point in Section XXX of the letter.}
-
-    This has important consequences in terms of database.
 In the submitted version, we used the original Hummel's database (available on his webpage from 1974 to 2003, which we completed by the US Bureau of census import database over the period 2005-2013.
@@
 As a result, in the submitted version, transport costs are estimated at the 3-digit sector level as benchmark estimate, exploiting product heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
-
-    Switching to $k=10$ as benchmark disaggregation level for the product dimension requires to resort on US Census dataset only.
-As first step, we have re-run the estimation over the years for which we already had the data (2005-2013).
-In parallel, we have bought older years from the Census bureau, as well as the more recent years 2014-2019.
+%
+%    Switching to $k=10$ as benchmark disaggregation level for the product dimension would require resorting on US Census dataset only.
+It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
 However, the Census bureau can provide data disaggregated at the $k=10$ product level only starting in 1989.
-This deprives us from the very first years 1974-1988.
+Some of these data are in physical storage and were not accessible because of the Covid-19 pandemic (1989-1996, but also 2000 and 2001).
+In september 2021, they are still not accessible.
+Waiting for them contributed to making this revision quite late.
+This deprives us from the very first years 1974-1996 and 2000-2001.
 For this period, we can only resort on Hummel's dataset, with $k=5$ the finest product level category.
-This notably induces us to maintain $s=3$ as the benchmark disaggregation level for the sectoral dimension for the whole period.
-This allows us to preserve (as far as possible) the largest time coverage - even though switching from $s=3, k=5$ (1974-1988) to $s=3, k=10$) (1989-2019) undoubtedly constitutes a significant break.
-\textbf{What do we do with that?}
+This is a non-negligible cost as we are interested in the long-term evolution of trade costs.
+%
+%    As a result, we decided to adopt the following strategy in the revised version.
+As benchmark, we still consider the 5/3 product/sector classification (as in the submitted version), as this allows us to preserve (as far as possible) the largest time coverage.
+We check the robustness of our results to the 10/3 classification, for the available years.
+This is made in Section XXX.
+%
+%    \textbf{LP: Pas sure que ce point doive etre mis ici, pas main change du coup}
 
@@
 \item Based on your main comment 1, the paper has been rewritten so as to be more in line with Hummel's (2004) methodology.
-Specifically, we now start from the share of additive costs $\beta$, from which we can deduce the additive / ad-valorem components ($t,\tau$) rather than the other way round (as in the submitted version).
+Specifically, we now start from the elasticity of total transport costs to the unit price $\beta$, which we show also corresponds to the share of additive costs.
+From this, we can deduce the additive / ad-valorem components ($t,\tau$) rather than the other way round (as in the submitted version).
 We thank the referee for pointing this out, as this contributes to a more straightforward and intuitive understanding of our estimation method.

### 2021-09-22 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Chanement partie modèle
 alléger main text

@@
 
-As in \cite{sorensen2014}, we adopt the two-country model of \cite{melitz} with monopolistic competition and endogenous firm entry.\footnote{\cite{Irrazabal_2015} adopt the multi-country model of \cite{chaney2008}, consistently with their database that documents the exports of Norwegian firms in various destination markets worldwide.
+We rely on the model developed by \cite{sorensen2014}, which amends the two-country model of \cite{melitz} with monopolistic competition and endogenous firm entry to include additive costs.\footnote{\cite{Irrazabal_2015} adopt the multi-country model of \cite{chaney2008}, consistently with their database that documents the exports of Norwegian firms in various destination markets worldwide.
 We rather choose to retain the \cite{melitz} model, in its simplest form with one sector and two symmetric countries, picking up the average estimated values of transport costs based on our unique importer country, e.g.
 the US.
-Sticking to this canonical set-up allows us to illustrate the role of additive costs in a transparent and intuitive way.} As widely standard in the literature, we focus on steady-state analysis, running comparative statics analysis following a trade cost reduction.
-Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
+Sticking to this canonical set-up allows us to illustrate the role of additive costs in a transparent and intuitive way.} Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
+For sake of space saving, we only present here the building blocks of the model.
+More details can be found in Appendix \ref{app:theoretical_model}.
 
@@
 She chooses her consumption $c(\omega)$  within the set $\Omega$ of available varieties (made of both domestic and foreign varieties, endogenously defined), subject to her budget constraint: $\int_{\omega \in \Omega} p(\omega) c(\omega) dw \leq E$, with $p(\omega)$ the consumer price of variety $\omega$ and $E$ nominal expenditures.
-As standard, the optimal demand function for variety $\omega$ is:
-\begin{eqnarray*}
-c(\omega) &=& \left(\frac{p(\omega)}{P}  \right)^{-\sigma} \frac{E}{P}\\
-\end{eqnarray*}
-\noindent with the associated optimal price index $P$ equal to:
-\begin{equation}
-P = \left[ \int_{\omega \in \Omega}p(\omega)^{1-\sigma}d\omega\right]^{\frac{1}{1-\sigma}} \label{eq:CPI}
-\end{equation}
+As standard, the optimal demand function for variety $\omega$ is: $c(\omega) = \left(\frac{p(\omega)}{P}  \right)^{-\sigma} \frac{E}{P}$ with the associated optimal price index $P$ equal to $P = \left[ \int_{\omega \in \Omega}p(\omega)^{1-\sigma}d\omega\right]^{\frac{1}{1-\sigma}} $.
 
@@
 \paragraph{Profit-maximizing behaviors} Once entered, and conditional on being able to export, a firm adopts a pricing-to-market behavior by setting a specific price on each destination market.
-
-On the domestic market, the objective is to maximize profit $\pi_d(\omega)= \left(p^d(\varphi) -\frac{1}{\varphi}  \right)c(\varphi)$, taken into account the consumer demand function $c(\varphi) = \left(\frac{p(\varphi)}{P}  \right)^{-\sigma} \frac{E}{P}$\footnote{In a monopolistic set-up, each firm (of productivity $\varphi$) is the sole producer of variety $\omega$, implying a one-to-one link between $\omega$ and $\varphi$, such that we could write $\omega(\varphi)$.
-For sake of notational simplicity, we omit this link and simply write the consumer demand function as a function of $\varphi$ directly.}
-The optimal pricing behavior yields:
-$$p^d(\varphi) = \frac{\sigma}{\sigma-1}\frac{1}{\varphi}$$
-With a CES specification on household's preferences, the firm sets its price as a constant mark-up on marginal cost, which is inversely related to its productivity as for the domestic market. \medskip
-
-Conditional on exporting abroad, the profit-maximizing program on the export market requires the inclusion of variable international trade costs. The optimizing program accordingly rewrites as:
-\begin{eqnarray*}
-\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p_x(\varphi) -\frac{\tau}{\varphi}  \right)c_x(\varphi) - tc_x(\varphi) \\
-\text{s.t. }&&c_x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
-\end{eqnarray*}
-
-\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c_x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
-Solving for this yields:
-\begin{equation}
-p_x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right] \label{eq:px}
-\end{equation}
-
-Similarly as for the Home market, the firm sets a constant markup on its marginal cost, which is still a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs.
+For each destination market, he firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the export market, according to:
+\begin{eqnarray}
+p_d(\varphi) &=& \frac{\sigma}{\sigma-1}\frac{1}{\varphi} \notag \\
+p_x(\varphi) &=& \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right] \label{eq:px}
+\end{eqnarray}
 
@@
 \end{eqnarray}
-\noindent with $B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ and where $r_d(\varphi)$ and $r_x(\varphi)$ denote the sales revenues on each market.
+\noindent with $B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ and where $r_d(\varphi)$ and $r_x(\varphi)$ denote the sales revenues on each market (see Appendix \ref{app:theoretical_model} for a detailed expression).
 Profits increase with productivity $\varphi$, and only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
@@
 
-Making use of the firm's first-order pricing conditions, revenues on each local and export market respectively read as:
-\begin{eqnarray}
-% \nonumber % Remove numbering (before each equation)
-  r_d(\varphi) &=& \left[\frac{1}{\rho \varphi}  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rd}\\
-  r_x(\varphi) &=& \left[\frac{1}{\rho}\left(\frac{\tau}{\varphi} +t\right)  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rx}
-\end{eqnarray}
-\noindent with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate.
-
 \paragraph{Zero-cutoff profit conditions} Given the fixed costs on both the domestic and the export markets, only firms with a productivity above $\varphi^\ast$ will serve the domestic market, and those with $\varphi>\varphi_x^\ast$ will export abroad.
-With $\frac{\pi_d(\varphi^\ast)}{\delta}$ and  $\frac{\pi_x(\varphi_x^\ast)}{\delta}$ the expected profit on the domestic and export market respectively, the zero cut-off (ZCP) productivity levels for successful entry on each market should verify:
-\begin{eqnarray*}
-\frac{\pi_d(\varphi^\ast)}{\delta} = 0 &\Leftrightarrow & \varphi^\ast = \frac{1}{f^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}\\
-\frac{\pi_x(\varphi_x^\ast)}{\delta} = 0 &\Leftrightarrow & \varphi_x^\ast = \frac{\tau}{f_x^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}} -t}
-\end{eqnarray*}
 The two productivity thresholds can be linked according to:
@@
 Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as $\left[\tau+ t\varphi^\ast_x \right] \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$.
-Two comments can be made at this point.
-First, additive costs ($t>0$) make the selection condition more likely to happen everything else equal for any given values of ad-valorem and fixed costs.
-Put differently, additive costs somehow relax the constraint of the fixed cost of exporting.
-Second, due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
+Due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
 We check this condition holds numerically in solving the model.\footnote{This is no longer the case under ad-valorem costs only.
@@
 on the support $\left[ \varphi^\ast_x,\infty\right)$.
-
-Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, one can show that the consumer price index can be rewritten as: comes that:
-$$P = \frac{\sigma}{\sigma-1}M_T^{\frac{1}{1-\sigma}}\left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$$
-
-\noindent with $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ the weighted average of productivities of domestic and imported varieties respectively (defined in Appendix \ref{app:theoretical_model}).
-With $\widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$ the weighted average of firm productivities on the domestic market (including both local producers and foreign producers exporting at Home), the price index can be expressed as $P = M_T^{\frac{1}{1-\sigma}}\frac{1}{\rho \widetilde{\varphi}_T }$.
-As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, leading to:
-
-\begin{equation}
-\mathcal{W} = \rho \widetilde{\varphi}_T M_T^{\frac{1}{\sigma-1}}\label{eq:Welfare}
+Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}):
+\begin{equation}
+\mathcal{W} = \rho \widetilde{\varphi}_T M_T^{\frac{1}{\sigma-1}} \label{eq:Welfare}
 \end{equation}
@@
 \end{equation}
-
-\noindent with $\left(1-G(\varphi^\ast)\right)$ the probability of successful entry.
-$\bar{\pi}$ can be expressed as $\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$, with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{\pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to $\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$.
-
-Noticing that, from Equation (\ref{eq:rd}), $\frac{r_d(\varphi)}{r_d(\varphi^\ast)} = \left( \frac{\varphi}{\varphi^\ast} \right)^{\sigma-1}$ and similarly on the export market, Equations (\ref{eq:pid_rd}) and (\ref{eq:pix_rx}) can be rewritten as:
-\begin{eqnarray*}
-\bar{\pi}_d &=& \left\{\left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1}-1 \right\}f \\
-\bar{\pi}_x &=& \left\{\left[\left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)\right]^{\sigma-1}-1 \right\}f_x
-\end{eqnarray*}
-
-Making use of the above elements, we finally get the zero-cut off profit condition in the open economy with additive costs as:
+As shown in Appendix \ref{app:theoretical_model}, it can be rewritten as:
 \begin{equation}
@@
 \end{equation}
-\noindent with $\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_x^\ast$ given by Equations (\ref{eq:def_phitilde}) and (\ref{eq:def_phitildex}).
-
-\paragraph{Summary} Equations (\ref{eq:def_phitilde}), (\ref{eq:def_phitildex}), (\ref{eq:ZCP}), (\ref{eq:link_thresholds}) and (\ref{eq:FEC}) determine a system of interdependent equations in five endogenous variables $\left\{\bar{\pi}, \widetilde{\varphi}, \widetilde{\varphi}_x, \varphi^\ast, \varphi^\ast_x \right\}$.
-From this, we can deduce the equilibrium values of the remaining set of variables (as detailed in Appendix \ref{app:theoretical_model}).
+\noindent with $\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_x^\ast$ the weighted average productivities of domestic and exporting firms respectively (see Equations (\ref{eq:def_phitilde}) and (\ref{eq:def_phitildex}) in Appendix \ref{app:theoretical_model}).
 
@@
 To do so, we dwell on values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
-The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition}.
+The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC}.
 
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f$ endogenously revealed at the autarky case, such that all firms produce in autarky.
+$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.
 BEJK (2003) refers to \cite{BEJK-AER-03}.}}}
 \end{table}
-
-Following \cite{ghironi}, $f$ such that all firms produce in autarky (i.e., such that $\varphi^{\ast,a} = \varphi_{\text{min}}$.
-The calibration of the fixed export cost obeys the following logic.
-As in \cite{ghironi}, $f_x$ is endogenously revealed to match a given share of exporting firms at the steady state.
-Further, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
-Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
-Specifically, we endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
-In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}.\smallskip
-
-The calibration of transport costs parameters relies on our empirical estimates.
-We notably illustrate the consequences of trade liberalization by comparing the equilibrium based on the 1974 values of transport costs (at the beginning of our sample), versus those of 2019 (at the very end of our estimation period), in both cases on only ad-valorem transport costs and in the presence of both additive and ad-valorem costs.
-Calibration of transport costs is displayed in Table \ref{tab:calib_TC}.
 
@@
 \end{table}
+
+We notably illustrate the consequences of trade liberalization by comparing the equilibrium based on the 1974 values of transport costs (at the beginning of our sample), versus those of 2019 (at the very end of our estimation period), in both cases on only ad-valorem transport costs and in the presence of both additive and ad-valorem costs.
+Further, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
+This can be captured in the model through a reduction in the fixed export cost $f_x$.
+Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.\footnote{We endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
+In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}, along with the 2019 values of variable transport costs}
 
@@
 
+\subsubsection{Firms}
+
+\paragraph{Profit-maximizing behavior}
+
+On the domestic market, the objective is to maximize profit $\pi_d(\omega)= \left(p^d(\varphi) -\frac{1}{\varphi}  \right)c(\varphi)$, taking into account the consumer demand function $c(\varphi) = \left(\frac{p(\varphi)}{P}  \right)^{-\sigma} \frac{E}{P}$.
+In a monopolistic set-up, each firm (of productivity $\varphi$) is the sole producer of variety $\omega$, implying a one-to-one link between $\omega$ and $\varphi$, such that we could write $\omega(\varphi)$.
+For sake of notational simplicity, we omit this link and simply write the consumer demand function as a function of $\varphi$ directly.
+The optimal pricing behavior yields:
+$$p^d(\varphi) = \frac{\sigma}{\sigma-1}\frac{1}{\varphi}$$
+With a CES specification on household's preferences, the firm sets its price as a constant mark-up on marginal cost, which is inversely related to its productivity as for the domestic market. \medskip
+
+Conditional on exporting abroad, the profit-maximizing program on the export market requires the inclusion of variable international trade costs. The optimizing program accordingly rewrites as:
+\begin{eqnarray*}
+\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p_x(\varphi) -\frac{\tau}{\varphi}  \right)c_x(\varphi) - tc_x(\varphi) \\
+\text{s.t. }&&c_x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
+\end{eqnarray*}
+
+\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c_x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market.
+Solving for this yields:
+\begin{equation}
+p_x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right] \label{eq:px}
+\end{equation}
+
+Making use of the firm's first-order pricing conditions, revenues on each local and export market respectively read as:
+\begin{eqnarray}
+% \nonumber % Remove numbering (before each equation)
+  r_d(\varphi) &=& \left[\frac{1}{\rho \varphi}  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rd}\\
+  r_x(\varphi) &=& \left[\frac{1}{\rho}\left(\frac{\tau}{\varphi} +t\right)  \right]^{1-\sigma} P^{\sigma-1} E \label{eq:rx}
+\end{eqnarray}
+\noindent with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate.
+
+\paragraph{Zero-cutoff profit conditions} With $\frac{\pi_d(\varphi^\ast)}{\delta}$ and  $\frac{\pi_x(\varphi_x^\ast)}{\delta}$ the expected profit on the domestic and export market respectively, the zero cut-off (ZCP) productivity levels for successful entry on each market should verify:
+\begin{eqnarray*}
+\frac{\pi_d(\varphi^\ast)}{\delta} = 0 &\Leftrightarrow & \varphi^\ast = \frac{1}{f^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}}}\\
+\frac{\pi_x(\varphi_x^\ast)}{\delta} = 0 &\Leftrightarrow & \varphi_x^\ast = \frac{\tau}{f_x^{\frac{1}{1-\sigma}}B^{\frac{-1}{1-\sigma}} -t}
+\end{eqnarray*}
+
 \subsubsection{Aggregation: Price indices}
 
-Following \cite{melitz}, and decomposing the set of available varieties into $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can rewrite the price index (\ref{eq:CPI}) as:
+Following \cite{melitz}, and decomposing the set of available varieties into $\Omega_d$, $\Omega_x$ the set of varieties respectively produced at Home and imported, one can rewrite the price index as:
 \begin{eqnarray*}
@@
 
+Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, one can show that the consumer price index can be rewritten as: comes that:
+$$P = \frac{\sigma}{\sigma-1}M_T^{\frac{1}{1-\sigma}}\left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$$
+
+\noindent with $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ the weighted average of productivities of domestic and imported varieties respectively defined in Equations (\ref{eq:def_phitilde}) and (\ref{eq:def_phitildex}).
+With $\widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$ the weighted average of firm productivities on the domestic market (including both local producers and foreign producers exporting at Home), the price index can be expressed as $P = M_T^{\frac{1}{1-\sigma}}\frac{1}{\rho \widetilde{\varphi}_T }$.
+As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, leading to:
+
+\begin{equation*}
+\mathcal{W} = \rho \widetilde{\varphi}_T M_T^{\frac{1}{\sigma-1}} 
+\end{equation*}
+\noindent that is, Equation (\ref{eq:Welfare}).
+
+\paragraph{Free-entry condition} $\bar{\pi}$ can be expressed as $\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$, with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{\pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{\pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to $\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$.
+
+Noticing that, from Equation (\ref{eq:rd}), $\frac{r_d(\varphi)}{r_d(\varphi^\ast)} = \left( \frac{\varphi}{\varphi^\ast} \right)^{\sigma-1}$ and similarly on the export market, Equations (\ref{eq:pid_rd}) and (\ref{eq:pix_rx}) can be rewritten as:
+\begin{eqnarray*}
+\bar{\pi}_d &=& \left\{\left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1}-1 \right\}f \\
+\bar{\pi}_x &=& \left\{\left[\left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)\right]^{\sigma-1}-1 \right\}f_x
+\end{eqnarray*}
+
+Making use of the above elements, we finally get the zero-cut off profit condition in the open economy with additive costs as:
+
 \subsubsection{Solving the general equilibrium}

### 2021-09-22 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement papier

--- 2021-09-21 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 \begin{abstract}
-This paper demonstrates the magnitude and the role of additive international transport costs.
+International transport costs do have an additive part.
+By how much?
+To what extent should we care?
+This paper provides new answers to these questions.
 Using information contained in the US imports flows from 1974 to 2019, we develop an empirical model that disentangles the ad-valorem and the additive components of international transport costs.
-As main result, we quantify the importance of the per-unit component of transport costs: It represents between 30\% and 45\% of overall transport costs.
-We then emphasize the consequences of modelling additive costs, under two different angles.
-First, modelling time-varying per-unit costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role.
-Second, we evaluate the normative consequences role of trade liberalization in presence of additive costs.
+The per-unit component of transport costs represents a sizeable share of total transport costs, by between 30\% and 45\% depending on the year and the transport mode considered.
+We then investigate the consequences of modelling additive costs, under two different angles.
+First, modelling varying additive costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role.
+Second, we revisit the welfare gains of trade liberalisation in presence of additive costs.
 Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points.
@@
 On the empirical side first, we show the importance of incorporating additive costs in the understanding of the international transport costs decrease observed between 1974 and 2019.
-Allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs over time, by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e.
-changes over time in the baskets of imported goods and/or the origin countries).} Trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
+Allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.
+Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
 %\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
@@
 This is consistent with the empirical findings by \cite{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
-Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e.
-cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
-Using transaction-level trade data from Colombia, \cite{Lashkaripour-2017} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
-One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip
+Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
+One important contribution of the paper is to provide quantitative estimates of both type of transport costs.
+\smallskip
 
@@
 The remainder of the paper is built as follows.
-The following section presents the data and the baseline empirical model used to estimate the values of international transport costs.
+Section \ref{sec:data_method} presents the data and the baseline empirical model used to estimate the values of international transport costs.
 Section \ref{sec:basic_results} reports the later on a yearly basis throughout the period 1974-2019 by transport mode, explicitly distinguishing between the additive and the ad-valorem components and providing several important robustness checks of our results.
-In particular, section \ref{sec:basic_results} reports the mean values of the transport costs estimated over the period (by transport mode).
 The two following sections take stock of those results to draw implications for two important debates in economic analysis.
@@
 We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price.
-The fas price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
-It is also the seller's responsibility to clear the goods for export.
-The fob price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
-Once the goods clear the railing of the vessel, the buyer assumes the risk.
-Note that this term is used exclusively for maritime and inland waterway transport.
-While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity (the number) of goods.
+While both terms are closely related (see more details in Appendix \ref{app:data}), the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity (the number) of goods.
 This is due to the fact that in the Census data, information on the quantity of goods is not mode-specific.
@@
 
-\subsection{Estimating transport costs: empirical specification}
+\subsection{Estimating transport costs: empirical specification \label{sec:estimation_strat}}
 
@@
 \begin{eqnarray}
-\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau -1  +\frac{t}{\tilde{p}}} \notag \\
-&=& -\frac{\widetilde{t}}{\tau-1 +\widetilde{t}} \label{eq:tt2}
+\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& -\frac{\frac{t}{\tilde{p}}}{\tau -1  +\frac{t}{\tilde{p}}} = -\frac{\widetilde{t}}{\tau-1 +\widetilde{t}} \label{eq:tt2}
 \end{eqnarray}
 
-Comparing Equations (\ref{eq:alternative}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs (in relative terms) is also equal to the share of additive costs in total transport costs.
+Comparing Equations (\ref{eq:alternative}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs is also equal to the share of additive costs in total transport costs.
 
@@
 As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, which decomposes into both an additive and an ad-valorem component.
-In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others).
-We contribute to this issue by adopting the following strategy.
+We do so by adopting the following strategy.
 For each year and transport mode, we estimate two models.
-In Model (a), additive costs are excluded, transport costs being modeled as iceberg costs only. In this case, the estimated equation simplifies as:
-
+In Model (A), additive costs are excluded, transport costs being modeled as iceberg costs only.
+In this case, the estimated equation simplifies as:
 \begin{equation}
@@
 
-Under Model (b), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
+Under Model (B), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
 From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modeling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
 This has driven us to estimate this third model (c) as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
-The main result that emerges is that this model (c) is dominated (in terms of quality of fit properties) by the model (a) with multiplicative costs only, which is itself dominated by the complete model (b), anticipating further results.
+The main result that emerges is that this model (c) is dominated (in terms of quality of fit properties) by the model (A) with multiplicative costs only, which is itself dominated by the complete model (B), anticipating further results.
 More details of these results are given in the Online Appendix.}
@@
 Similarly, as \cite{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
-We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (a), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (b), for each year and transportation mode.
+We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (A), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (B), for each year and transportation mode.
 These results are reported in Section \ref{sec:results_decomposition}.
@@
 The top panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.
-In the medium panel, we report the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})).
-The bottom panel presents our estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})).
+In the medium panel, we report the results for a specification based on ad-valorem costs estimated alone (Model (A), based on Equation (\ref{eq:model_nlI})).
+The bottom panel presents our estimates for a specification involving both ad-valorem and multiplicative components (Model (B), based on Equation (\ref{eq:model_IetA})).
 
@@
    % Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
-   % \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
+   % \multicolumn{5}{l}{\textbf{Model (A) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
   %  Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
@@
 %value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
- %   \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
+ %   \multicolumn{5}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
  %  \textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} & & & & \\ \hline
@@
 This result is not surprising though, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
-Second, the estimated total costs are not exactly equal to the observed ones, in particular for Model (a).
+Second, the estimated total costs are not exactly equal to the observed ones, in particular for Model (A).
 This might be accounted for by recalling that all our estimates are based on log-linearization.
@@
 Table \ref{tab:summary_results} shows that transport costs amount to around a 5.5-6\% markup in ocean shipping, and 4\% markup in air shipping on average over 1974-2019.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
-Under Model (a) (ad-valorem costs only), this gives a markup of 5.5\% in vessel, and 4.8\% in air shipping (considering the mean value over the period).
-Under Model (b), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
+Under Model (A) (ad-valorem costs only), this gives a markup of 5.5\% in vessel, and 4.8\% in air shipping (considering the mean value over the period).
+Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$.
 With $\tau= 1.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\%.
@@
 
-In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database.
-This focus is shared with some of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}.
-Along with the pattern of total transport costs over time, our empirical set-up also allows us to investigate whether additive costs we put forward so far looking at the average, is a long-lasting, i.e.
-a structural phenomenon.
-
-\paragraph{Total transport costs: A first look at the trend patterns} Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs expressed in percentage of the fas price,\footnote{Specifically, we use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix.
-Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2019.
+\paragraph{Total transport costs} Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs expressed in percentage of the fas price, by transport mode for each year between 1974 and 2019.
 
@@
 \includegraphics[width=14cm, height=7cm]{Figure2_Trend_of_totalTC_bymode.jpg}
+
+\begin{minipage} [c]  {5in} \scriptsize%
+Notes: We use the yearly values of the estimated transport cost components averaged over the country/product dimensions (at the 3-digit level).
+Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.
+\end{minipage}
 \end{center}
@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.
-LP quand je prends les chiffres des tableaux B1 et B2 je ne retrouve pas les graphiques, par ex en air 5.3\% en 1974, sur le graphique quasi 6?} for mean vessel transport costs, implying a \textbf{50\% decrease in air and a 60\% decrease in vessel over the period ACTUALISER, je trouve 55\% environ dans les deux modes, 2.3/5.3 et 4.1/8.9}.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.} for mean vessel transport costs, implying a \textbf{50\% decrease in air and a 60\% decrease in vessel over the period ACTUALISER}.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 
-\paragraph{The share of additive costs: Time dynamics} As next step, we investigate whether additive costs are a long-lasting, i.e. a structural phenomenon.
-
-We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}.
-The corresponding values reported in detailed tables are available in section B of the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}.
-Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
+\paragraph{The share of additive costs: Time dynamics} As next step, we investigate whether additive costs are a long-lasting, i.e.
+a structural phenomenon.
+Figure \ref{fig:part_cout_additif} reports the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$.
+.
 
@@
 %{\footnotesize {OECD data} }
+\begin{minipage} [c]  {5in} \scriptsize%
+Notes: Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
+\end{minipage}
 \end{center}
@@
 In maritime transport, the share of additive costs fluctuates around the average, while it has been steeply declining since the end of the 1980s in air transport.
-
-\paragraph{Wrap-up} Our results so far show that total transport costs have strongly reduced between 1974 and 2019, especially in ocean transport.
-Further, if additive costs represent a substantial share of total transport costs considered on average over the period for both transport modes, a yearly-based inspection reveals marked fluctuations in the share of additive costs in total costs over time.
-In air transport specifically, the share of additive costs has been declining over time since the beginning of the 1990s.
-This suggests that the transport costs decrease is more than proportionally due to additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.
-
-In the following section, we check the robustness of these results regarding two important issues, the separability assumption of country-good fixed effects and potential endogeneity biases.
+This suggests that the transport costs decrease is more than proportionally due to additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.\smallskip
+
+In the following section, we check the robustness of these results regarding three important issues, the separability assumption of country-good fixed effects, potential endogeneity biases and the level of aggregation at the product level.
 
@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 300\% to 44\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
@@
 As a further check, we investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
-Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (a) and for maritime transport in panel (b).
+Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (A) and for maritime transport in panel (B).
 In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
@@
 Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
-Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.
-
-As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
-the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (a), (c))) and the non-separability scenario (Panels (b), (c)).
-
-\begin{figure}[htbp]
-\caption{Histogram of the share of additive costs in total costs (restricted sample)}
-\label{fig:histogram_beta_robustness}
-\begin{center}
-\begin{tabular}{cc}
-\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
-\end{tabular}
-\end{center}
-\end{figure}
-
-For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
-Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
-
-\subsubsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.\footnote{As third check, we also study the histogram of the distribution of $\beta$ by transport mode.
+\textbf{See Appendix XXX for details LP A METTRE}.
+}
+
+\subsubsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) is subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-Besides, related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+Related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
 In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
@@
 Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure.
-Put differently, we follow earlier literature (see e.g.
+We follow earlier literature (see e.g.
 \citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
 Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
-In particular, it shows that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}%\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.}
+In particular, it shows that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}) %\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.}
  to produce a 2SLS-type of estimation.
-Figure \ref{fig:comp_IV_SITC5} below reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+Figure \ref{fig:comp_IV_SITC5} reports our benchmark estimates by transport mode, together with their instrumented counterpart.
 In all cases, these estimates are very similar, when not identical.
@@
 
+\textbf{LP: Guillaume, changer les legendes de la figure ecrire beta en toutes lettres pas top...}
 %However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
@@
 The results reported here rely on exploiting data over the years 1997-1999 and 2001-2019.\footnote{Theoretically, they should be available from 1989, which is already smaller than our full sample starting in 1974.
-Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} \textbf{faut-il dire que le sample 10-3 n'est pas exactement le meme que nos baseline results et qu'on refait tourner le 5-3 sur le meme sample exactement?}
+Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} \textbf{REPRENDRE CES RESULTATS PB SUR VESSEL LP.
+}
 
@@
 		%\item \textbf{Model (A): With ad-valorem transport costs only}
-		\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\item Model (B): With additive and ad-valorem transport costs
 	\end{tablenotes}
 \end{table}
-
-\textbf{JE NE COMPRENDS PAS LES RESULTATS EN VESSEL, 5-3 same sample as 10-3, PB?
-c'est un peu court comme commentaire, la comparison en air montre que la part des additifs est tout de meme plus faible, quasiment divisee par 2...
-A REPRENDRE JE PENSE}
 
@@
 
-Two main insights emerge up to now.
+\paragraph{Summary} Two main insights emerge up to now.
 First, total transport costs have strongly reduced between 1974 and 2019.
@@
 In this section, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.
-As underlined by \cite{hummels2007}, the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade.
-Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}.
-As emphasized by \cite{Hummels_1999, hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
+As underlined by \cite{hummels2007}, total transport costs may have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased (composition effects) independently of any change in transport costs \textit{per se}.
+It is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
 Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
@@
 \paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}.
-He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$).
-Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes is written as:
+Starting from a similar specification as in Section \ref{sec:estimation_strat}, \cite{hummels2007} writes his transport cost measure $TC_{ikt}$ as:
 
@@
 
-As already noticed, $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
+\noindent where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) and $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
 Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
@@
 
-\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension) suggests, the share of additive costs is rather varying over time, in both transport modes.
+\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} suggests, the share of additive costs is rather varying over time, in both transport modes.
 %In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
@@
 To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price on a per product-year-country basis, i.e.
-$\beta$, according to:
-\begin{eqnarray*}
-\beta_{ikt} &\equiv& -\frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
-&=& \frac{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}}{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}+\widehat{\tau}_{is(k)t}-1},
-\end{eqnarray*}
+$\beta$, according to $\beta_{ikt} = \frac{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}}{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}+\widehat{\tau}_{is(k)t}-1},
+$
 
 \noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{is(k)t}$ and $\widehat{\tau}_{is(k)t}$).
-The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (a) and (b) for air and maritime transport respectively.
+The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (A) and (B) for air and maritime transport respectively.
 
@@
 \begin{tabular}{cc}
-%{\small (a) Air } & {\small (b) Vessel}\\
-\includegraphics[width=3.0in, height=2.5in]{"Etude_beta_pondere_(a) Air transport".jpg}
-& \includegraphics[width=3.0in,height=2.5in]{"Etude_beta_pondere_(b) Vessel transport".jpg} \\
+%{\small (A) Air } & {\small (B) Vessel}\\
+\includegraphics[width=3.0in, height=2.5in]{"Etude_beta_pondere_(A) Air transport".jpg}
+& \includegraphics[width=3.0in,height=2.5in]{"Etude_beta_pondere_(B) Vessel transport".jpg} \\
 \multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year,product,country), weighted by share of yearly value of flow.}}\\
@@
 
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed exposition of our method and a comparison with Hummel's (2007) methodology.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects.
-We start with the presentation of our estimation strategy before turning to the results.\smallskip
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.
+We start with the presentation of our estimation strategy before turning to the results.
 
@@
 
-The estimation carried out in section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$).
-Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect.
+The estimation carried out in section \ref{sec:results_decomposition} provides us with the 
+Starting from the additive and ad-valorem measures of international transport costs which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$), we extract the pure transport cost measure by the mean of a time fixed effect.
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total cost respectively), and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.
+The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$, and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
 
@@
 \subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
+Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (A), (B) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
 Panels (d), (e) and (f) report the results for vessel.
@@
 First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
-This stands in line with the related literature (see \citealp{Hummels_1999}; \citealp{Lafourcade_Thisse}).
-Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
+International transport costs were reduced by \textbf{50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
+AJUSTER LP}.
+Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (a), (d) and (e)).
 Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (A) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
 For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
@@
 Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
-Afterwards, the dotted line remains below the plain line, but the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
+Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
 Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
@@
 This drives us to investigate this difference further.
+
 As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap).
-Consequently, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions.
-%By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
-This turns out to be of primary importance in the disentangling, in the time trend of transport costs, of what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.
-
+Rather than from the actual cif-fas price gap, we start from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition} at the year/sector/origin country level, by transport mode.
+Put differently, our methodology lets the share of additive costs $\beta$ vary over the three time-product-partner country dimensions.
+By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
+This turns out to be of primary importance in the disentangling of the sources of the transport costs time trend.
 To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
-composition effects excluded ($\widehat{TC}_t$).
-In Panels (a) (for air) and (b) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \cite{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
+.
+In Panels (a) (for air) and (a) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \cite{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
+
+\textbf{LP: A mon avis, enlever l'une des deux, soit ls mettre en niveau, soit en base 100 1974, mais pas les deux. Mettre une footnote pour le dire.}
 
@@
 \begin{tabular}{cc}
-{\small (a) Ad-valorem air freight} & {\small (b) Ad-valorem vessel freight}\\
+{\small (A) Ad-valorem air freight} & {\small (B) Ad-valorem vessel freight}\\
 \multicolumn{2}{c}{{\small In percent of the export price}} \\
@@
 \end{tabular}
+	\begin{minipage}  [c]  {6in}
+		\scriptsize
+The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
+composition effects excluded ($\widehat{TC}_t$).
+\end{minipage}
 \end{center}
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate (plain line) than the unfitted rate (dotted line), especially for vessel (Panel (d)).
+Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate than the unfitted rate, especially for vessel (Panel (d)).
 Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
@@
 Complementing the findings of Section \ref{sec:results_decomposition}, these results highlight the importance of integrating the additive dimension of international transport costs, with the aim here of characterizing their time trends.
+
+\textbf{LP: stop here}
 
@@
 In this respect, the ``custom value'' corresponds to the fas price (``free-alongside'' price) delivered by the seller.
+Let us clarify here the difference with the fob price.
+The fas price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in.
+It is also the seller's responsibility to clear the goods for export.
+The fob price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer.
+Once the goods clear the railing of the vessel, the buyer assumes the risk.
+Note that this term is used exclusively for maritime and inland waterway transport.
 More information is available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}.
+\smallskip
 
@@
 
-\subsection{IV estimates: First stage \label{app:first_stage_IV}}
+\subsection{More on robustness checks}
+
+\subsubsection{Separability assumption}
+
+As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
+With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
+the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (A), (c))) and the non-separability scenario (Panels (B), (c)).
+
+\begin{figure}[htbp]
+\caption{Histogram of the share of additive costs in total costs (restricted sample)}
+\label{fig:histogram_beta_robustness}
+\begin{center}
+\begin{tabular}{cc}
+\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
+\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
+
+\subsubsection{IV estimates: First stage \label{app:first_stage_IV}}
 
@@
 \begin{tabular}{cc}
-%{\small (a) $\beta$} & {\small (b) $\alpha$}\\
+%{\small (A) $\beta$} & {\small (B) $\alpha$}\\
 \includegraphics[width=3.2in, height=3.2in,keepaspectratio]{beta_lag_SITC5.pdf}
@@
 \begin{tabular}{cc}
-%{\small (a) $\beta$} & {\small (b) $\alpha$}\\
+%{\small (A) $\beta$} & {\small (B) $\alpha$}\\
 \includegraphics[width=3.2in, height=3.2in,keepaspectratio]{beta_lag_HS10.pdf}
@@
 \begin{eqnarray}
-\ln(\widehat{\tau}_{ikt}-1)&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt}-1)&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(A)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(A)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
 \end{eqnarray}
 
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (A)), the product dimension (Term (B)) and the pure transport costs time trend (Term (c)).
 Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
@@
 \begin{eqnarray}
-\ln(\widehat{p}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_pfas}
+\ln(\widehat{p}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(A)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_pfas}
 \end{eqnarray}
@@
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more than the fitted component).
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (A), overall transport costs have decreased more than the fitted component).
 This stands in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.

### 2021-09-22 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

@@
 A natural question then emerges: To what extent should we care about additive costs? Our paper answers this question from two different sides.
-
 On the empirical side first, we show the importance of incorporating additive costs in the understanding of the international transport costs decrease observed between 1974 and 2019.
@@
 
-\textbf{LP: stop here}
-
 \section{The role of additive cost: Theoretical insights}\label{sec:theory}
@@
 \paragraph{Free-entry condition} There is free-entry in the monopolistic industry, imposing that the expected net present value of entry is driven to zero. 
-In the open-economy setting, this condition is given by:
-\begin{equation*}
-  \sum_{t=0}^{\infty}(1-\delta)^t \left(\int_{\varphi^\ast}^{\infty} \pi_d(\varphi )dG(\varphi) + \int_{\varphi_x^\ast}^{\infty} \pi_x(\varphi )dG(\varphi)  \right) = f_e
-\end{equation*}
-
 Defining $\bar{\pi}$ as the average profit conditional to entry, this free-entry condition rewrites as:
@@
 To do so, we dwell on values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
-The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition}.
-
-\begin{table}[htbp]
+
+\begin{table}[htb]
   \centering
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.
-BEJK (2003) refers to \cite{BEJK-AER-03}.}}}
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f$ endogenously revealed at the autarky case, such that all firms produce in autarky.
+$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.}}}
 \end{table}
 
-Following \cite{ghironi}, $f$ such that all firms produce in autarky (i.e., such that $\varphi^{\ast,a} = \varphi_{\text{min}}$.
-The calibration of the fixed export cost obeys the following logic.
-As in \cite{ghironi}, $f_x$ is endogenously revealed to match a given share of exporting firms at the steady state.
-Further, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
-Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
-Specifically, we endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
-In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}.\smallskip
-
-The calibration of transport costs parameters relies on our empirical estimates.
-We notably illustrate the consequences of trade liberalization by comparing the equilibrium based on the 1974 values of transport costs (at the beginning of our sample), versus those of 2019 (at the very end of our estimation period), in both cases on only ad-valorem transport costs and in the presence of both additive and ad-valorem costs.
-Calibration of transport costs is displayed in Table \ref{tab:calib_TC}.
-
-\begin{table}[htbp]
+The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC}.
+
+\begin{table}[htb]
   \centering
@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model ad-valorem costs only.
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model ad-valorem costs only.
 In Columns (2), (4), (6) and (8), we model both ad-valorem and additive costs.
@@
 \end{table}
+
+We are interested in studying the normative consequences of trade liberalization in presence or not of additive costs.
+To that aim, we will compare the equilibrium based on the 1974 values of transport costs (at the beginning of our sample), versus those of 2019 (at the very end of our estimation period), in both cases on only ad-valorem transport costs and in the presence of both additive and ad-valorem costs.
+Further, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
+This can be captured in the model through a reduction in the fixed export cost $f_x$.
+Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.\footnote{We endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
+In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}, along with the 2019 values of variable transport costs}
 
@@
 
+Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, one can show that the consumer price index can be rewritten as: comes that:
+$$P = \frac{\sigma}{\sigma-1}M_T^{\frac{1}{1-\sigma}}\left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$$
+
+\noindent with $\widetilde{\varphi}(\varphi^\ast)$ and $\widetilde{\varphi}_x(\varphi_x^\ast)$ the weighted average of productivities of domestic and imported varieties respectively defined in Equations (\ref{eq:def_phitilde}) and (\ref{eq:def_phitildex}).
+With $\widetilde{\varphi}_T \equiv \left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$ the weighted average of firm productivities on the domestic market (including both local producers and foreign producers exporting at Home), the price index can be expressed as $P = M_T^{\frac{1}{1-\sigma}}\frac{1}{\rho \widetilde{\varphi}_T }$.
+As in \cite{melitz}, given our assumptions of wage normalization and a one-sector economy, welfare per worker is simply equal the inverse of the price level: $\mathcal{W} = \frac{w}{P}= P^{-1}$, leading to:
+
+\begin{equation*}
+\mathcal{W} = \rho \widetilde{\varphi}_T M_T^{\frac{1}{\sigma-1}}
+\end{equation*}
+\noindent that is, Equation (\ref{eq:Welfare}).
+
+\paragraph{Free-entry condition} In the open-economy setting, the free-entry condition is given by:
+\begin{equation*}
+  \sum_{t=0}^{\infty}(1-\delta)^t \left(\int_{\varphi^\ast}^{\infty} \pi_d(\varphi )dG(\varphi) + \int_{\varphi_x^\ast}^{\infty} \pi_x(\varphi )dG(\varphi)  \right) = f_e
+\end{equation*}
+$\bar{\pi}$ can be expressed as $\bar{\pi} =\pi_d(\widetilde{\varphi}(\varphi^\ast)) + \rho_x \pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))$, with $\pi_d(\widetilde{\varphi}(\varphi^\ast)) \equiv \bar{\pi}_d$ the average expected profit on the domestic market, $\pi_x(\widetilde{\varphi}_x(\varphi_x^\ast))\equiv \bar{\pi}_x$ the average expected export profit and $\rho_x$ the probability of exporting (conditional on entry on the domestic market), equal to $\rho_x = \frac{1-G(\varphi_x^\ast)}{1-G(\varphi^\ast)}$.
+
+Noticing that, from Equation (\ref{eq:rd}), $\frac{r_d(\varphi)}{r_d(\varphi^\ast)} = \left( \frac{\varphi}{\varphi^\ast} \right)^{\sigma-1}$ and similarly on the export market, Equations (\ref{eq:pid_rd}) and (\ref{eq:pix_rx}) can be rewritten as:
+\begin{eqnarray*}
+\bar{\pi}_d &=& \left\{\left(\frac{\widetilde{\varphi}^\ast}{\varphi^\ast}\right)^{\sigma-1}-1 \right\}f \\
+\bar{\pi}_x &=& \left\{\left[\left(\frac{\widetilde{\varphi}_x^\ast}{\varphi_x^\ast}\right)\left(\frac{\tau+ t \varphi^\ast_x}{\tau+ t \widetilde{\varphi}^\ast_x}\right)\right]^{\sigma-1}-1 \right\}f_x
+\end{eqnarray*}
+
+Making use of the above elements, we finally get the zero-cut off profit condition in the open economy with additive costs as:
+
 \subsubsection{Solving the general equilibrium}

### 2021-09-22 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte papier + online Appendix

@@
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
-As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
+As reported in the Online Appendix, Section \textbf{XXXX}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
 on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
@@
 Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
+\textbf{LP: Pb avec le nb d'observations.
+}
 %
@@
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions%, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the results on a detailed yearly basis and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triple $(i,k,t)$, which confirm the robustness of the results.
+These results are reported in the Online Appendix, \textbf{Section XXX} for sake of space saving.} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
 The picture is reversed for medians
-.
-
-As a further check, we investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
-Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (A) and for maritime transport in panel (B).
-In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
-
-\begin{figure}[htbp]
-\caption{Robustness to the separability assumption (year-to-year basis)}
-\label{fig:robustesse_non_separe}
-\begin{center}
-\begin{tabular}{cc}
-\includegraphics[width=6in, height=4in]{graph_robustesse_ns.jpg}
-
-\end{tabular}
-\end{center}
-\begin{minipage} [c]  {5in} \scriptsize%
-Notes: FE for (country, sector) fixed effects.
-\end{minipage}
-\end{figure}
-
-In line with Table \ref{tab:robustness_separability}, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
-However, they also show that the difference is quantitatively small in most cases.
-Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
-Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.\footnote{As third check, we also study the histogram of the distribution of $\beta$ by transport mode.
-\textbf{See Appendix XXX for details LP A METTRE}.
-}
 
@@
 
-\textbf{LP: Guillaume, changer les legendes de la figure ecrire beta en toutes lettres pas top...}
+\textbf{LP: Guillaume, changer les legendes de la figure enlever beta, mettre en positif la part des additifs et ajouter legende en ordonnee...}
 %However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-.
-In Panels (a) (for air) and (a) (for vessel), we report the transport costs measures (fitted and unfitted) expressed as a percentage of the export price, as in \cite{hummels2007}.\footnote{Unsurprisingly, this accords with \cite{hummels2007} results, see his Figures 5 (for air) and 6 (for vessel).} To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.
-
-\textbf{LP: A mon avis, enlever l'une des deux, soit ls mettre en niveau, soit en base 100 1974, mais pas les deux. Mettre une footnote pour le dire.}
+To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In the Online Appendix, Sextion \textbf{XXX}, we report the two series of transport costs (fitted and unfitted) as percentage of the export price as in \cite{hummels2007}.
+Unsurprisingly, this accords with his results displayed in his Figures 5 (for Air) and 6 (for Vessel) for the years in common (until 2004).
+}
+
+\textbf{LP pour Guillaume: Changer les légendes pour les aligner sur la terminologie du papier, fitted vs unfitted.}
 
@@
 \begin{tabular}{cc}
-{\small (A) Ad-valorem air freight} & {\small (B) Ad-valorem vessel freight}\\
-\multicolumn{2}{c}{{\small In percent of the export price}} \\
-\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
-& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
+%{\small (A) Ad-valorem air freight} & {\small (B) Ad-valorem vessel freight}\\
+%\multicolumn{2}{c}{{\small In percent of the export price}} \\
+%\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
+%& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
 {\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
@@
 		\scriptsize
-The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e.
-composition effects excluded ($\widehat{TC}_t$).
+\textbf{A AJUSTER SELON LEGENDE FINALE}.
 \end{minipage}
@@
 
-\subsection{Variance decomposition exercise \label{app:decomp_variance}}
-
-In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
-Specifically, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$) and \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
-The variance decomposition expression at the product level $k$ (5-digit level) is obtained by applying the following formula (by year and transport mode):
-
-$$\underbrace{\sum_{k=1}^K \sum_{i=1}^I \left(x_{ik} - \bar{x}_g  \right)^2}_{\text{Total variability}} = \underbrace{\sum_k \sum_i \left(x_{ik} - \bar{x}_k  \right)^2}_{\text{Within-product variability}} + \underbrace{\sum_k \left(\bar{x}_{k} - \bar{x}_g  \right)^2}_{\text{Between-product variability}}$$
-
-with $x$ the observed cif-fas price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
-
-\begin{eqnarray*}
-\bar{x}_g \equiv \frac{1}{n} \sum_{k=1}^K \sum_{i=1}^I x_{ik},&& \bar{x}_k \equiv \frac{1}{K}\sum_{k=1}^K x_{ik}
-\end{eqnarray*}
-
-\noindent with $n$ the total number of observations, $K$ the total number of products and $I$ the total number of country partners.
-We apply the same variance decomposition exercise at the sector level, in which case the sector $s$ index (at the 3-digit level) replaces the $k$ index (at the 5-digit level).
-This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
-We also determine the share of the observed variance that can be attributed to the between-country variance, adapting the variance decomposition formula written above accordingly.
-Results are reported in Figure \ref{fig:decomp_variance}.
-
-\begin{figure}[htbp]
-\caption{Variance decomposition (observed cif-fas price gap)}
-\label{fig:decomp_variance}
-\begin{center}
-\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
-%{\footnotesize {OECD data} }
-\end{center}
-\end{figure}
-
-Two interesting results emerge from Figure \ref{fig:decomp_variance}.
-First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
-Both account for between 30 and 40\% of the total variance in air transport, depending on the years considered.
-This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
-This provides an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
-Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
-This holds throughout the period and for both transport modes.
-This suggests that what primarily matters in international transport costs is mostly attributable to the product \textit{per se}, rather than to the country where it comes from.
-%JH: un peu too much, non ? "By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs."
-
-\subsection{More on robustness checks}
-
-\subsubsection{Separability assumption}
-
-As an ultimate check, we investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
-With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
-the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (A), (c))) and the non-separability scenario (Panels (B), (c)).
-
-\begin{figure}[htbp]
-\caption{Histogram of the share of additive costs in total costs (restricted sample)}
-\label{fig:histogram_beta_robustness}
-\begin{center}
-\begin{tabular}{cc}
-\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
-\end{tabular}
-\end{center}
-\end{figure}
-
-For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
-Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
-
-\subsubsection{IV estimates: First stage \label{app:first_stage_IV}}
+\subsection{IV estimates: First stage \label{app:first_stage_IV}}
 
@@
 
-\paragraph{Obtaining the unfitted measures} - The objective here is to express the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
+\paragraph{Obtaining the unfitted measures} The objective here is to express the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
 To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
@@
 
-\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
+\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
 $$\widehat{TC}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
@@
 
-\paragraph{Obtaining the fitted total transport cost index} - The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
+\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
 composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
@@
 \end{eqnarray*}
-\noindent Equivalently, we can write:
-
-\begin{eqnarray*}
-&&p_{ikt} = \widetilde{p}_{ikt} + \exp(\Delta_{ikt})\\
-\Leftrightarrow && p_{ikt} = \widetilde{p}_{ikt} + t_{ikt}
-\end{eqnarray*}
-\noindent with $t_{ikt}$ appropriately defined. This exactly corresponds to the case of additive costs only.
+\noindent Taken the exponential, this becomes: $p_{ikt} = \widetilde{p}_{ikt} + \exp(\Delta_{ikt})$ or, equivalently: $p_{ikt} = \widetilde{p}_{ikt} + t_{ikt}$, with $t_{ikt}$ appropriately defined.
+This exactly corresponds to the case of additive costs only.
 
 In the second polar case with $\beta = 0$, Equation (\ref{eq:Hummels1}) becomes:
-\begin{eqnarray*}
-&&\ln \left(\frac{p_{ikt}- \widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} \right) = \Delta_{ikt} \\
-\Leftrightarrow && \ln \left(\frac{p_{ikt}}{\widetilde{p}_{ikt}}-1\right) = \Delta_{ikt}\\
-\Leftrightarrow && \frac{p_{ikt}}{\widetilde{p}_{ikt}}  = \exp (\Delta_{ikt})+1
-\end{eqnarray*}
-\noindent that is,
-
-$$p_{ikt} = \tau_{ikt} \widetilde{p}_{ikt}$$
-
-\noindent with $\tau_{ikt}$ appropriately defined. This exactly corresponds to the case of ad-valorem transport costs only.
+$$\ln \left(\frac{p_{ikt}}{\widetilde{p}_{ikt}}-1\right) = \Delta_{ikt}$$
+
+Taking the exponential, we get: $\frac{p_{ikt}}{\widetilde{p}_{ikt}}  = \exp (\Delta_{ikt})+1$, that is: $p_{ikt} = \tau_{ikt} \widetilde{p}_{ikt}$, with $\tau_{ikt}$ appropriately defined.
+This exactly corresponds to the case of ad-valorem transport costs only.
 

### 2021-09-23 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex

@@
 
-%JH : reprendre XXX ICI
+\subsection{Robustness checks}
+
+\subsubsection{Variance decomposition exercise \label{app:decomp_variance}}
+
+In this section, we provide a variance decomposition exercise on the observed cif-fas price gap.
+Specifically, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$) and \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
+The variance decomposition expression at the product level $k$ (5-digit level) is obtained by applying the following formula (by year and transport mode):
+
+$$\sum_{k=1}^K \sum_{i=1}^I \left(x_{ik} - \bar{x}_g  \right)^2 = \underbrace{\sum_k \sum_i \left(x_{ik} - \bar{x}_k  \right)^2}_{(1)} + \underbrace{\sum_k \left(\bar{x}_{k} - \bar{x}_g  \right)^2}_{(2)}$$
+
+with $x$ the observed cif-fas price gap and $\bar{x}_g$, $\bar{x}_k$ average values defined as:
+
+\begin{eqnarray*}
+\bar{x}_g \equiv \frac{1}{n} \sum_{k=1}^K \sum_{i=1}^I x_{ik},&& \bar{x}_k \equiv \frac{1}{K}\sum_{k=1}^K x_{ik}
+\end{eqnarray*}
+
+\noindent with $n$ the total number of observations, $K$ the total number of products and $I$ the total number of country partners.
+In the above equation, the left-hand side gives the total variability, which decomposes into the within-product variability (Term (1)) and the between-product variability (Term (2)).
+We apply the same variance decomposition exercise at the sector level, in which case the sector $s$ index (at the 3-digit level) replaces the $k$ index (at the 5-digit level).
+This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
+We also determine the share of the observed variance that can be attributed to the between-country variance, adapting the variance decomposition formula written above accordingly.
+Results are reported in Figure \ref{fig:decomp_variance}.
+
+\begin{figure}[htbp]
+\caption{Variance decomposition (observed cif-fas price gap)}
+\label{fig:decomp_variance}
+\begin{center}
+\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
+%{\footnotesize {OECD data} }
+\end{center}
+\end{figure}
+
+Two interesting results emerge from Figure \ref{fig:decomp_variance}.
+First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
+Both account for between 30 and 40\% of the total variance in air transport, depending on the years considered.
+This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
+This provides an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
+Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
+This holds throughout the period and for both transport modes. By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
 
@@
 
-\subsection{Eliminating the composition effects: Primary vs. Manufacturing sector} \label{sec_oa:comp-effects}
+\subsubsection{Separability Assumption}
 
+This section reports two supplementary checks concerning the separability assumption.
+
+\paragraph{Comparison on a yearly basis} We investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
+Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (A) and for maritime transport in panel (B).
+In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
+
+\begin{figure}[htbp]
+\caption{Robustness to the separability assumption (year-to-year basis)}
+\label{fig:robustesse_non_separe}
+\begin{center}
+\begin{tabular}{cc}
+\includegraphics[width=6in, height=4in]{graph_robustesse_ns.jpg}
+
+\end{tabular}
+\end{center}
+\begin{minipage} [c]  {5in} \scriptsize%
+Notes: FE for (country, sector) fixed effects.
+\end{minipage}
+\end{figure}
+
+In line with Table \ref{tab:robustness_separability}, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
+However, they also show that the difference is quantitatively small in most cases.
+Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
+Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.
+
+\paragraph{Distribution of the $\beta$} As final check, we also investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
+With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
+the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (A), (c))) and the non-separability scenario (Panels (B), (c)).
+
+\begin{figure}[htbp]
+\caption{Histogram of the share of additive costs in total costs (restricted sample)}
+\label{fig:histogram_beta_robustness}
+\begin{center}
+\begin{tabular}{cc}
+\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
+\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
+\end{tabular}
+\end{center}
+\end{figure}
+
+For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
+\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
+
+\subsection{Eliminating the composition effects} \label{sec_oa:comp-effects}
+
+\subsubsection{More on the Hummel's methodology}
+
+This section complements Section \textbf{XXX} in the main paper, where we replicate the method adopted by \cite{hummels2007} on our database.
+Here, we report the unfitted measure of transport costs versus the ``pure'' transport costs series, i.e composition effects excluded
+in percentage of the export price, as in \cite{hummels2007}.
+We also adopt Himmel's (2007) terminology, contrasting the ``expenditure/importe value'' (i.e, the unfitted serie) and the ``fitted ad-valorem rate'' (i.e, corresponding to the ``pure'' transport costs series, composition effects excluded).
+Unsurprisingly, this accords with his Figure 5 (for Air) and 6 (Vessel) for the comme period (until 2014).
+
+\begin{figure}[htbp]
+\caption{Characterizing the time trends: applying Hummel's (2007) method }
+\label{fig:comp_effects_as_in_Hummels}
+\begin{center}
+\begin{tabular}{cc}
+{\small (A) Ad-valorem air freight} & {\small (B) Ad-valorem vessel freight}\\
+\multicolumn{2}{c}{{\small In percent of the export price}} \\
+\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
+& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
+{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight} \\
+\end{tabular}
+\end{center}
+\end{figure}
+
+\subsubsection{Composition effects: Primary vs. Manufacturing sector}
 In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows.

### 2021-09-23 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex

--- 2021-09-06 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
@@
 \multicolumn{3}{l}{Fixed export cost } \\
-$f_x $ & XXX & Share of exporting firms = 4\% (BEJK, 2003) \\
-& XXX& Share of exporting firms = 6\% (\cite{Lincoln_McCallum2018}) \\
+$f_x $ & 0.08 & Share of exporting plants = 21\% (\cite{ghironi}) \\
+& 0.06& Share of exporting plants = 39\% (\cite{Lincoln_McCallum2018}) \\
 \hline

### 2021-09-23 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
relecture intro et partie théorique

@@
 By how much?
-To what extent should we care?
+To what extent it matters?
 This paper provides new answers to these questions.
 Using information contained in the US imports flows from 1974 to 2019, we develop an empirical model that disentangles the ad-valorem and the additive components of international transport costs.
-The per-unit component of transport costs represents a sizeable share of total transport costs, by between 30\% and 45\% depending on the year and the transport mode considered.
-We then investigate the consequences of modelling additive costs, under two different angles.
+The per-unit component of transport costs represents a sizeable share of total transport costs, between 30\% and 45\% depending on the year and the transport mode considered.
+We then investigate the important consequences of modelling additive costs, under two different perspectives.
 First, modelling varying additive costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role.
@@
 Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points.
-In both aspects, our results point to the importance of modelling the additive component of international transport costs.
+%In both aspects, our results point to the importance of modelling the additive component of international transport costs.
 \end{abstract}
@@
 The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices).
-If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
-of agglomeration, the profitability of outsourcing or the gains from trade liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip
+If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade liberalization.
+Our main contribution in this paper is to bring new insights on this matter.\smallskip
 
 Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
-Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns.
+Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which prove to be key in the characterization of international trade patterns.
 While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
@@
 As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
-Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs.
-Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip
+Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip
 
@@
 On average over 1974-2019, the additive cost is estimated to be 1.7\% and 2.8\% of the export price in air and vessel transport respectively (at the 3-digit classification level).
-This is slightly lower than our estimates for the ad-valorem component (2.4\% and 3\% for air/vessel respectively on average over the period).
+This is slightly lower than our estimates for the ad-valorem component (2.4\%/3\% for air/vessel respectively on average over the period).
 Put differently, additive costs represent on average between 30 and 45\% of the overall transport costs.
-To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.
+To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs over such a long period of time.
 This represents a valuable insight for calibrating related international trade or business cycle models.
-We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs.
+We also provide an empirical assessment of what standard international trade models miss by skipping additive transport costs.
 Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by a factor around 2.
-Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.\smallskip
-
-A natural question then emerges: To what extent should we care about additive costs? Our paper answers this question from two different sides.
-On the empirical side first, we show the importance of incorporating additive costs in the understanding of the international transport costs decrease observed between 1974 and 2019.
-Allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2019}.
+Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degrees of freedom.\smallskip
+
+To what extent should we care about additive costs? Our paper answers this question from two different perspectives.
+On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019.
+By allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019.
 Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
@@
 
-We also show the importance of embedding additive costs when assessing the effects of trade liberalization in normative terms. Specifically, we
-make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: When this reduction solely occurs through a decrease in the ad-valorem cost, as standardly modelled in New Trade models, versus when it is achieved through a combined reduction in the ad-valorem and the additive components, as quantified by our empirical model.
+On the theoretical side, we show the importance of embedding additive costs when assessing the effects of trade liberalization in normative terms.
+Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction solely occurs through a decrease in the ad-valorem cost, as standardly modelled in New Trade theories, versus when it is achieved through a combined reduction in the ad-valorem and the additive components, as quantified by our empirical model.
 To this aim, we embed additive costs in the canonical \cite{melitz} model.
 Relying on our empirical estimates of transport costs
-over 1974-2019, our preferred scenario shows an extra welfare gain by around 1.2 percentage point when the transport cost decrease encompasses a reduction in its additive component (as we document empirically) relative to a pure ad-valorem reduction.
+over 1974-2019, our preferred scenario shows an extra welfare gain by around 1.2 percentage point when the transport cost decrease encompasses a reduction in its additive component relative to a pure ad-valorem reduction.
 Put differently, neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
@@
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
-In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
+In this respect, it contributes to the literature challenging the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
 As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
@@
 This is consistent with the empirical findings by \cite{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
-Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs).
-One important contribution of the paper is to provide quantitative estimates of both type of transport costs.
+Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} However, these papers yet do not allow for a quantitative measure of additive costs (neither for the ad-valorem costs).
+Another important contribution of this paper is to provide quantitative estimates of both type of transport costs.
 \smallskip
@@
 Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs, the one exception being \citealp{Irrazabal_2015}.
-In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade liberalization at stake since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs and by sub-component (additive and ad-valorem).\smallskip
+In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade liberalization since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs, as well as by sub-component (additive and ad-valorem).\smallskip
 
@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.} for mean vessel transport costs, implying a \textbf{50\% decrease in air and a 60\% decrease in vessel over the period ACTUALISER}.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.} for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 In this section, we explore the implications of our empirical results on theoretical grounds.
-While the whole bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
+While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
 One potential explanation is the lack of empirical estimates about the size of additive costs.
 Our results may thus usefully palliate this limitation.
-Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: When this reduction is achieved through a combined reduction in the ad-valorem and the additive
-components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade models).
-To this aim, we embed additive costs in the canonical M\'{e}litz (2003) model.\smallskip
+Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive
+components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
+To this aim, we embed additive costs in the canonical \citet{melitz} model.\smallskip
 
@@
 This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz' type models.
-Anticipating on our results, we also share the result, that the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
-As first explanation, it should be noted that, in \cite{Lashkaripour_JIE2020}'s paper, the case with additive costs relies on a share of additive costs that remains negligible.
+We also find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
+
+That said, it should be noted the case with additive costs in \cite{Lashkaripour_JIE2020}'s paper relies on a negligible share of additive costs.
 In this context, one might expect that whether additive costs are present or not makes little difference.
 As underlined by \cite{Lashkaripour_JIE2020}, a second explanation may rely on the modeling framework.
-While \cite{Lashkaripour_JIE2020} models a perfectly competitive model, we as in \cite{sorensen2014} and \cite{Irrazabal_2015} adopt a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}.
+While \cite{Lashkaripour_JIE2020} relies on a perfectly competitive environment, we use a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015}.
 In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
@@
 Sticking to this canonical set-up allows us to illustrate the role of additive costs in a transparent and intuitive way.} Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
-For sake of space saving, we only present here the building blocks of the model.
+For the sake of space, we only present here the building blocks of the model.
 More details can be found in Appendix \ref{app:theoretical_model}.
@@
 The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
-Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
+Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ (as in \citealp{melitz}) and an additive cost depending on the quantity (weight) transported, $t\geq 0$ - not included in the canonical Melitz model.
 At each point in time, a firm can be hit by a death shock at the exogenous probability $\delta>0$.
@@
 \paragraph{Profit-maximizing behaviors} Once entered, and conditional on being able to export, a firm adopts a pricing-to-market behavior by setting a specific price on each destination market.
-For each destination market, he firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the export market, according to:
+For each destination market, the firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the export market, according to:
 \begin{eqnarray}
@@
 Due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
-We check this condition holds numerically in solving the model.\footnote{This is no longer the case under ad-valorem costs only.
+We check this condition holds numerically when solving the model.\footnote{This is no longer the case under ad-valorem costs only.
 In this case, with $t=0$, we find the usual condition that $\varphi^\ast_x >\varphi^\ast$ iif $\tau^{\sigma-1}\frac{f_x}{f}>1$.}
@@
 
-As standard in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the heterogenity in firm productivity is assumed to be given by a Pareto distribution.
+Following a standard assumption in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the heterogeneity in firm productivity is assumed to be given by a Pareto distribution.
 As shown in \cite{melitz-redding-Handbk-IT-2014}, the Pareto distribution indeed allows for a full analytical solving of the M\'{e}litz model.
@@
 {\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f$ endogenously revealed at the autarky case, such that all firms produce in autarky.
-$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above.}}}
+$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2006's value), according to the references provided above.}}}
 \end{table}
@@
 We are interested in studying the normative consequences of trade liberalization in presence or not of additive costs.
-To that aim, we will compare the equilibrium based on the 1974 values of transport costs (at the beginning of our sample), versus those of 2019 (at the very end of our estimation period), in both cases on only ad-valorem transport costs and in the presence of both additive and ad-valorem costs.
-Further, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
+To that aim, we will compare the equilibrium based on the values of transport costs at the beginning (1974) and at the very end of our estimation period (2019), in two seetings: one including only ad-valorem transport costs, and including both additive and ad-valorem costs.
+Besides, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
 This can be captured in the model through a reduction in the fixed export cost $f_x$.
@@
 We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
-In the second scenario so-called \textit{à la Hummels}, we rely of the true estimates of transport costs, with the reduction of total transport costs coming from both additive and ad-valorem components.
+In the second scenario so-called \textit{à la Hummels}, we rely on the estimates of transport costs involving both additive and ad-valorem components.
 The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
@@
 Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
 \cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs.

### 2021-09-24 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Petites modifs

@@
 
+%\citesep{knuth-681,knuth-702}{'s work
+
 \begin{document}
@@
 While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
-as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\cite{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
+as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in \cites{samuelson1954} terminology,\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
 As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
@@
 A similar calculus gives 4.1\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
-\cite{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+\cites{anderson_wincoop_jel} estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
 
@@
 In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
-\cite{Irrazabal_2015}'s results on the year 2004 imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
+\cites{Irrazabal_2015} results on the year 2004 imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
 For the same export price multiplied by the ad-valorem cost of 100 USD, our results for 2004 yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
@@
 It is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs.
-Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
+Specifically, \cites{hummels2007} results support that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
 Our own results tend to challenge this view, as we further show.
@@
 
-That said, it should be noted the case with additive costs in \cite{Lashkaripour_JIE2020}'s paper relies on a negligible share of additive costs.
+That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020} paper relies on a negligible share of additive costs.
 In this context, one might expect that whether additive costs are present or not makes little difference.

### 2021-09-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement lettre référé

--- 2021-08-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
@@
 Number of sections, pages and equations mentioned in the text refer to the revised version.
+When necessary, we refer to number of sections, pages and equations from the submitted version.
+In this case, they are written into brackets.
 
@@
 
+The structure of the paper has been modified after taking into account the referees' comments. 
 \begin{itemize}
-\item Based on your comment 1.2 (\textit{The aggregation problem}), the benchmark set-up now retains $s=3$ and $k=10$ for the industry/product aggregation levels respectively.
-It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
-Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
-We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
-In line with the referee's comment, we now run the estimation by considering $k$ at the HS10 classification level, not 5.
-\textbf{We come back to this point in Section XXX of the letter.}
-
-    This has important consequences in terms of database.
-In the submitted version, we used the original Hummel's database (available on his webpage from 1974 to 2003, which we completed by the US Bureau of census import database over the period 2005-2013.
-In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004.
-For comparability reasons over time, we have adopted the same degree of classification $k=5$ for the following years (based on the original Census dataset).
-As a result, in the submitted version, transport costs are estimated at the 3-digit sector level as benchmark estimate, exploiting product heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
-
-    Switching to $k=10$ as benchmark disaggregation level for the product dimension requires to resort on US Census dataset only.
-As first step, we have re-run the estimation over the years for which we already had the data (2005-2013).
-In parallel, we have bought older years from the Census bureau, as well as the more recent years 2014-2019.
-However, the Census bureau can provide data disaggregated at the $k=10$ product level only starting in 1989.
-This deprives us from the very first years 1974-1988.
-For this period, we can only resort on Hummel's dataset, with $k=5$ the finest product level category.
-This notably induces us to maintain $s=3$ as the benchmark disaggregation level for the sectoral dimension for the whole period.
-This allows us to preserve (as far as possible) the largest time coverage - even though switching from $s=3, k=5$ (1974-1988) to $s=3, k=10$) (1989-2019) undoubtedly constitutes a significant break.
-\textbf{What do we do with that?}
-
-\item Based on your comment 1.1 (\textit{The endogeneity problem}), the benchmark regression (non-linear least squares) is now completed by a two-stage procedure, where the fas price is instrumented by tax duties and its own lag - see section 4.3 and appendix B.4 in the paper, and section 2.1.2 of this letter.
-
-\item Based on your main comment 1, the paper has been rewritten so as to be more in line with Hummel's (2004) methodology.
-Specifically, we now start from the share of additive costs $\beta$, from which we can deduce the additive / ad-valorem components ($t,\tau$) rather than the other way round (as in the submitted version).
-We thank the referee for pointing this out, as this contributes to a more straightforward and intuitive understanding of our estimation method.
-
-\item What about the big picture?
+\item In the submitted version, [Section 2] was devoted to the estimation of international transport costs; specifically, their break-down in two components, additive and ad-valorem.
+In [Section 3], we investigated the role of additive costs in the decomposition of transport costs time trends, between structural changes and composition effects.
+[Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
+\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness to aggregation, robustness to IV estimates).
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and the results, in Section 3, which now includes the robustness checks as final sub-section.
+For sake of space saving, the robustness analysis related to composition effects has been set to the Appendix (Section C.3).
+\item Both referees asked us to put light on the implications of additive costs (the ``Big Picture'').
+We agree that this was not enough emphasized in the submitted version, and we thank the referees for their suggestions on that issue.
+The revised version answers this concern in two points.
+    \begin{enumerate}
+    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we made our best to make this section easier to read, leaving the technical aspects in the Appendix (Section C).
+As such, we hope that the main message of this Section is better catched.
+    \item Most importantly, we now emphasize the ``Big Picture'' implications of additive costs on theoretical grounds.
+Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated (in Section 3 of the revised version), to quantify the extra welfare gains attributable to the reduction of the additive component.
+This analysis is driven in a new Section, Section 5.
+        \end{enumerate}
+\item To keep the paper (both the main text and with appendix included) with a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables in the Online Appendix, to which the reader can refer to.
 \end{itemize}

### 2021-09-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement réponse référé 1 cont'

@@
 
-\textbf{Our answer:}\\
-
 %\paragraph{Strategy of answer} To clarify the discussion, we decided to answer concerns (i) and (iii) each in turn.
@@
 
-The three points raised by the referee indeed deserve careful consideration. We provide separate answers to each of them below, starting with points (i) and (iii), and finishing with point (ii).
+The three points raised by the referee indeed deserve careful consideration. We provide separate answers to each of them below.
 
-\subsubsection{On the justifications for our empirical approach}
+%\subsubsection{On the justifications for our empirical approach}
 
-\paragraph{About the key assumption of separability}
+\subsubsection{Concern (i): About the assumption of separability}
 Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by \citet{Irrazabal_2015} to estimate the share of additive costs in a firm-level context.
@@
 
-\paragraph{On the use of non-linear least squares (NLS)}
+\textbf{WE ALREADY DID THAT BEFORE. SAY THAT WE POUT THAT EARLIER IN THE PAPER AND CLEARER?}
+
+\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is highly demanding computationally when the number of fixed effects increase.
+This constraint notably drove us to impose the separability assumption and retain $s=3$ as the relevant sectoral degree of aggregation.
 
@@
 That is, one can alternatively estimate the share of the additive component as:}
-
 $$\frac{t_{ik}}{ (\tau_{ik}-1)\tilde{p}_{ik} + t_{ik}} = \beta_{ik}$$
@@
 obviously require that the authors do not aggregate up the raw Census data
-across all districts and all 10-digit products pertaining to the same 5-digit category.}
+across all districts and all 10-digit products pertaining to the same 5-digit category.
+[...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-inductry pairs, without imposing Assumptions (a) and (b) outlined above.}
 
-\textbf{Our answer:}
-As highlighted by the referee, our estimated equation requires non-linear estimation methods, such as Non-Linear Least Squares.
+\textbf{Our answer:} First of all, we thank the referee for his relevant suggestion of an alternative estimation method.
+This drove us to question our own empirical specification deeper.
+In the end, we yet decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that you propose does not go without a few important limitations.
+We develop each line of argument in our answer.
+After reading our answer, we hope to convince you that the choice we made is the most relevant one.
+\textbf{The referee's method is presented in the online appendix XXX WHERE???
+XXX, and we refer to it in Footnote XXX of the paper - A FAIRE EN ONLINE APPENDIX?
+JUSTE DANS LA LETTRE?
+}
+
+\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares.
 However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
@@
 In this respect, implementing the referee's method (see below) does not suppress the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version - see footnote 16XXX page 8: ``One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.`on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
-This would not solve the problem though, for three main complementary reasons.
-First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
-Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
-Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
-Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$)''
-
-Of course, this does not prevent us for taking full account of your suggestion to adopt an alternative estimation method (summarized in Equation (\ref{eq:estimation_ref1}).
-It turns out that this method does not go without a few important limitations, that justify maintaining our initial estimation strategy as the baseline in the building block of the paper.\textbf{The referee's method is presented in the online appendix XXX WHERE???
-XXX, and we refer to it in Footnote XXX of the paper}.
-We develop the analysis we held below.
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
+\textbf{LP: A Ai enleve le copie-colle de la footnote dans la lettre; il suffut de dire que c'est dans la version revisee, non?}\smallskip
 
@@
 The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
-As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
+We thank the referee for suggesting to improve the comparison between our method and Hummel's.
+We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
+As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
 The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
@@
 
-Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) We develop our results and conclusion by answering each argument made by the referee at a time.
-
-Our answer can be articulated in two steps.
+Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
 First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
 Second, we emphasize the drawbacks induced by this method.
-XXX HERE XXX
-
 %
@@
 \begin{itemize}
-\item the referee writes \textit{The first advantage of this so-called Hummel's Approach is that the above
+\item \textit{The first advantage of this so-called Hummel's Approach is that the above
 regression can be estimated separately for various country-industry pairs, without
@@
 First, the separability assumption does not seem to be a strong assumption.
-This is based on the conclusion drawn from the robustness check on a reduced sample (in the submitted version, considering $s=3, k=5$, and in the revised version, considering $s=3, k=10$).
+This is based on the conclusion drawn from the robustness check on a reduced sample.
 If it is smaller in terms of observations,\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
-Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
-As we conclude at the end of Section XXX, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
+Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
+\textbf{LP: CHECK THE NUMBERS}} it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
 Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
@@
 Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects.
-This substantially reduces the number of point estimates, as well as the value of trade flows covered relative to the  benchmark method, as summarized in Table .
+This substantially reduces the number of point estimates, as well as the value of trade flows covered relative to the  benchmark method, as summarized in \textbf{Table XXX.
+Il faut reprendre cette reponse.
+Reprendre les resultats sur la methode du refere 1}.
 
-\item the referee writes \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.} \\
-
+\item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}
      \textbf{Our answer:}
@@
 
-\item the referee writes \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.}\\
+\item  \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.}\\
      \textbf{Our answer:}
@@
 
-     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ik}$ the price-elasticity of transport costs, which also corresponds to the share of additive costs in total transport costs (as we now show in the paper's appendix):
+     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ikt}$ the price-elasticity of transport costs, with $i$ the origin country, $k$ the product and $t$ for time.
+As we now show in the paper (Section 2.2), this also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper.
+We directly report them here to ease the comparison with the referee's method afterwards.}
 
      \begin{equation}
-     \beta_{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)}-1+\frac{t_{is(k)}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
+     \beta_{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)t}-1+\frac{t_{is(k)t}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
      \end{equation}
 
-     As underlined in the paper (both submitted and revised versions), Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method, as he estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, yielding to estimates of $\beta_{is(k)}$ that are invariant over time (but mode-specific).
+     Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method.
+In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a constant $\beta$ invariant over time/sector/origin country (even though mode-specific).
+Rather than Equation (\ref{eq:Hummels}), this amounts starting from the functional form: $f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta}$.
 
@@
 Let us start with the referee's method.
-It starts from the functional form linking transport costs and the fas price as specified in the Hummels's equation (\ref{eq:Hummels}).
+It starts from the functional form linking transport costs and the fas price as specified in the Equation (\ref{eq:Hummels}).
 Taken in log (on a mode/yearly basis), it is written as:
@@
 
+     \noindent with $d$ the district of entry.
 From this, one can then recover the levels of additive / multiplicative transports costs.
@@
 
-In this respect, the two methods are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
+In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
 Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such.
+In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
+Section 4 ([Section 3] of the revised version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
 We agree that this was not sufficiently transparent in the submitted version.
@@
 \begin{enumerate}
-\item Information about the port of entry are only available since 1989.
-Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by 15 years (1974-1988).
+\item Information about the port of entry are only available since 1989, which is already smaller than our full sample starting
+in 1974.
+On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
+Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
 In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
@@
 In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
+\textbf{LP: Il faut le faire }
 
@@
 \textbf{This stands in line with ...
-FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique}
+FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique A FAIRE !!!
+}
 
@@
     \begin{itemize}
-    \item With the referee's method, we estimate one value for the share of additive component $beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
+    \item With the referee's method, we estimate one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
 From this, we can approximate the 5-95\% threshold values through:
@@
 
-    \item  Our method does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
-Rather, our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$.
+    \item  Our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$.
+\textbf{NOTE THAT $\beta$ is in fact $i,s,t$ because we use individual fas prices to rebuilt it.
+Should we say tha?}.
+As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
 We then deduce the accuracy of the $\beta$ estimate by relying on bootstrap method.
@@
 Figures XXX (for Air) and XXX (for Vessel) thus report the histogram of the log of the ratio $\frac{CI_{is}}{CI^{ref}_{is}}$, for the year 2012.
+\textbf{IL FAUT ACTUALISER TOUT CA.
+OU DANS LE FICHIER SORTIE?}
 
@@
 
-Figure \ref{fig:accuracy_beta} In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
+In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
 Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
@@
 \textit{\textbf{The endogeneity problem}: quoting Footnote 14 of the paper, the authors
-are estimating ti and ts(k) as coefficients on the industry or country
+are estimating $t_i$ and $t_{s(k)}$ as coefficients on the industry or country
 dummies times $1/\widetilde{p}_{ik}$. [...] Based on the productivity-sorting model in Melitz (2003) or the quality-sorting
@@
 
-Not this check is performed both on our main dataset (SITC 5 digit), and also at the HS 10 level, to handle simultaneously the referee's concern about aggregation issues.
+It should be noted that we performed this check on our main dataset (SITC 5 digit), and also at the HS 10 level, to handle simultaneously the referee's concern about aggregation issues.
 In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very, very similar to their non-IV, least squares counterparts.
+Needless to say that these results are available upon request from the referee.
 
@@
 We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
-In line with the referee's comment, we now run the estimation by considering $k$ at the HS10 classification level, not 5.
 After checking with the US Bureau of Census, data is only available since 1989, now available to 2019.
-We have hence asked for all available data at the time of revision (1989-2009), and we now present the results considering $s=3$ and $k=10$.
-This however deprives us from being able to exploit the period 1974-1988 at the aggregation level.
-\textbf{Que faire pour la periode 1974-1989 XXX TO BE DISCUSSED XXX}.
-Furthermore, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
+As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.
+In front of this arbitrage, we decided to maintain the case $s=3, k=5$ as baseline, but running estimation at the $s=3,k=10$ in the new Sub-Section 3.3.3 of Section 3.3 devoted to robustness analysis.
+
+Additionnally, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
 We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper \textbf{see XXXX TO BE DONE ???
-XXX}
+XXX???
+LE DIRE DANS LA SECTION DE ROBUSTESSE OU ON FAIT DU 3-10??}
 
-% XXX JH: UNLESS I AM MISTAKING, THE SHOPPING LIST BELOW HAS BEEN IMPLEMENTED XXX TO BE CONFIRMED
-%\subsubsection{To sum up: What remains to be done}
-
-%\begin{enumerate}
-%\item\label{bp:3} Re-run our benchmark method with $s=3$ and $k=10$ over available years and compare the results ($t,\tau,\beta$ with our benchmark ones ($s=3,k=5$).
-Hopefully does not change much so that we can keep up the period 1974-1988?
-Consider this as our new benchmark.
-Answer to the critique about the aggregation problem.
-%\item\label{bp:1} After running the referee's 1 method with $s=3$, $k=$10, show tables or figures comparing 1) the number of observations / countries / sectors 3D by year / transport mode when we do our empirical strategy vs the referee's strategy; 2) comparison of the $\beta$ between the two methods; 3) comparison of the precision of the $\beta$ estimation.
-Answer to the critique about the estimation strategy.
-%\item\label{bp:2} IV-First-stage equation on fas price at the $s=3$, $k=10$-digit level, for all years; then 2d-stage estimate of $t,\tau, \beta$ with our methodology; compare with our results obtained in the submitted version.
-Answer to the critique about the endogeneity bias.
-%\end{enumerate}
+\textbf{LP: stop here}
 
@@
 
+\end{document}
+
 \newpage
@@
 
-\begin{equation}
+\begin{equation*}
 f_{ik} = X_{is(k)}\tilde{p}_{ik}^{-\beta_{ik}} \label{eq:Hummels}
-\end{equation}
+\end{equation*}
 

### 2021-09-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement réponse référé 1, cont'

@@
 Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated (in Section 3 of the revised version), to quantify the extra welfare gains attributable to the reduction of the additive component.
-This analysis is driven in a new Section, Section 5.
+This analysis is driven in a new Section 5.
         \end{enumerate}
-\item To keep the paper (both the main text and with appendix included) with a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables in the Online Appendix, to which the reader can refer to.
+\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables in the Online Appendix, to which the reader can refer to.
 \end{itemize}
@@
 
-The referee writes \textit{A more natural approach is what the authors, at some point, refer to as the Hummel's Methodology.
+The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummel's Methodology.
 That is, one can alternatively estimate the share of the additive component as:}
@@
 \textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t.
-unit price} in absolute value, as shown with more details in Appendix \ref{app:interpret_beta} through Equation (\ref{eq:Hummels}).
-\textit{Given the authors' objective and the data they are using, $\beta$ can be separately estimated
-for each industry-country pair using the following regression:}
+unit price.
+Given the authors' objective and the data they are using, $\beta$ can be separately estimated for each industry-country pair using the following regression:}
 
@@
 
-\textbf{Our answer:} First of all, we thank the referee for his relevant suggestion of an alternative estimation method.
+\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummel's.
+We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
+As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
+The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
+
+Now coming to the technical side, we thank the referee for his/her relevant suggestion of an alternative estimation method.
 This drove us to question our own empirical specification deeper.
@@
 Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
-\textbf{LP: A Ai enleve le copie-colle de la footnote dans la lettre; il suffut de dire que c'est dans la version revisee, non?}\smallskip
+\textbf{LP: A Ai enleve le copie-colle de la footnote dans la lettre; il suffit de dire que c'est dans la version revisee, non?}\smallskip
 
@@
 
-The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
-We thank the referee for suggesting to improve the comparison between our method and Hummel's.
-We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
-As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
-The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
-Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
-
-Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
+The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007), yielding Equation (\ref{eq:estimation_ref1}) as estimation equation.
+By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
 We have implemented the referee's method over the years for which we already have the data, i.e.
-the years 2005-2013 (already used in the submitted version).
+the years 2005-2019, at the $s=3, k=10$ sectoral/product levels.
 
@@
 
-Additionnally, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
+Additionally, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
@@
 LE DIRE DANS LA SECTION DE ROBUSTESSE OU ON FAIT DU 3-10??}
-
-\textbf{LP: stop here}
 
@@
 \textit{Calculating the unit price as Value/Weight presents the authors
-with an additional endogenity problem. To elaborate, let $\omega_{ik} = Weight/Quantity$
-denote the unit weight of the goods in observation $ik$. Also let $\widehat{p}_{ik} = Value/Quantity$
-(unlike what the authors assume) denote standard definition of price}.
-As a consequence, making the link with the fas price we consider in the paper: $\widetilde{p}_{ik} = \widehat{p}_{ik} / \omega_{ik}$.
-
-\textit{This paper is essentially estimating the following equation}:
-
-\begin{equation*}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}/\omega_{ik}} \right) + \epsilon_{ik}
-\end{equation*}
-
-\textit{instead of estimating}:
-
-\begin{equation*}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s(k)}}{\widehat{p}_{ik}} \right) + \epsilon_{ik}
-\end{equation*}
-
-%\textbf{GD20200708 Pourquoi n’introduit-il pas $\widehat{p}$ à gauche ?}
-\textit{There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
+with an additional endogeneity problem. To elaborate, let $\omega_{ik} = Weight/Quantity$
+denote the unit weight of the goods in observation $ik$. [...] There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
 categories, and (ii) $\omega_{ik}$ is negatively correlated with transport costs. So the
@@
 new (but avoidable) source of endogeneity.}
+
+\textbf{LP: J'ai enleve ses equations car c'est totalement faux}
 
@@
 In addition, referee's remarks echo recent findings by Lashkaripour (\emph{Journal of International Economics}, 2020): based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
-Finally, Lashkaripour (2020) finds accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
+Finally, Lashkaripour (2020) finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
 
@@
 We could of course restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
+\textbf{LP: DOCUMENT THAT.
+SHOW IN THE LETTER THATWE DID IT?}
 
@@
 
-These motives altogether highlight that it would be too costly and uncertain in our context to switch from weights to quantities to compute unit prices.
+These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression.
 That said, do we have reason to believe this measure of unit prices generate significant deviations from the one relying on quantities?
@@
 \citet{Lashkaripour_JIE2020} does provide important results related to ours.
-In particular, his result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 20 to 40\% of total costs.
+In particular, his result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 30 to 45\% of total costs.
 The two stories are not necessarily incompatible, however.
-First, beyond a longer time span, our database is exhaustive (\textbf{XXX GUILLAUME: TO BE CONFIRMED)}, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
-In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2021) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
-
-%\textbf{We make this point clear in Footnote XXX}
-
-%\textbf{EN MEME TEMPS: Si on écarte de l'analyse les flux qui sont passés à la fois par bateau et par air, en ne gardant que les flux qui ont fait l'un ou l'autre, ne peut-on pas se dire que la quantité (pas mode-spécifique) a totalement voyage par vessel (si le flux indique qqch pour ves-val) ou par avion (s'il indique qqch pour air-val)?
-Et à ce moment là, on peut répondre ``vraiment'' à sa critique?} \textbf{GD20200708 Sauf que du coup, on change le sample.
-Il faudrait vérifier si c’est de beaucoup ou pas}
-
-%\textbf{Je ne comprends pas bien le point (i)}.
-We thank the referee for pointing out this potential source of bias in our estimates.
-If we cannot follow the referee's advice be replacing weight by quantity for data availability reasons, we can address his/her concern regarding this as source of endogeneity bias by instrumenting the fas price.
-\textbf{il ne faudrait pas dans ce cas mettre de lagged prices, sinon on reintroduit du bruit.
-Non?
-Mais alors, on n'explique plus grand chose...
-}
-%\textbf{GD20200708 Moi non plus je ne comprends pas très bien}
+First, beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
+In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2020) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX A QUELLE CRITIQUE FAIT-ON REFERENCE?}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+\textbf{LP: EST-ce qu'on a des choses a lui montrer a lui sur la regression en quantite?
+Est-ce qu'on la met en annexe en ligne?}
 
@@
 Taking stock of your suggestion, we decided to offer insights on the welfare implications of our results.
-In this regard,% the section 3.4 from the initial version has been moved to Appendix XXX TO BE DONE XXX of the paper,
-the new section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
-\citet{melitz}, featuring additive trade costs.
+In this regard, the new section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
+To do so, we start from the canonical \cites{melitz} model where we add additive trade costs.
 The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
@@
 
-More precisely, we use some of the estimates underlying results reported in sections 2 and 3 concerning multiplicative and additive costs, more precisely those for the years 1974 and 2019.
-Based on the latter, we implement several comparative static exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+More precisely, we use some of the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, more precisely those for the years 1974 and 2019.
+Based on the latter, we implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
 In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
 To that end, we adjust the share of exporting firms in the US: based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
+In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/iceberg component, with the reduction in the fixed export cost.
+We compare this with the case where the total transport costs reduction is solely attributed to a decrease in ad-valorem costs.
 
-Table 5 in section 5 in the revised version reports the results of these various comparative static exercises for Air and Vessel transport modes.
-Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
-For each transport mode, the first two columns report the results of reducing transport costs only, maintaining the same value of $f_x$ as in the initial equilibrium (Columns (A)-(B) for Air, (E)-(F) for Vessel), contrasting the case of only ad-valorem costs (Columns (A), (D)) and in presence of additive costs (Columns (B), (F)).
-Columns (C)-(D) for Air and (G)-(H) for Vessel report the results when both the variable transport costs and the fixed export costs are lower in the final steady state (2019) relative to the initial one (1974).
+Table 5 in Section 5 reports the results of these various comparative statics exercises for Air and Vessel transport modes.
+Specifically, we report the change in total transport costs decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
+Qualitatively, the conclusions are identical for both transport modes, with two major insights.
+First, for a given decrease in total transport costs,  welfare gains are around 50\% higher when this reduction is partly achieved through a reduction in the additive costs.
+Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing variable costs, with again an additional premium coming from the decrease of additive costs.
 
-Qualitatively, the conclusions are identical for both transport modes, with two major insights.
-First, for a given decrease in transport costs,  welfare gains are around 50\% higher when the underlying framework allows for additive costs - which, in practice, concentrate the major part of transport costs decrease over the period.
-Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing costs, with again an additional premium coming from the decrease of additive costs.
-Quantitatively, welfare gains are more important for the Vessel mode, which is a natural consequence of a more substantial decrease in transport costs.
-More importantly, we see that the amplification effect coming from the decrease in $f_{x}$ is slightly lower for the specification including additive costs  in Vessel than in Air: a bit more than 40\% additional welfare gains for the former, versus 47\% gains for the latter.
-This reflects a slightly more balanced contribution from each component (additive and multiplicative) in the decrease of transport costs in Vessel compared to Air.
-This also emphasizes that the amplification effect of the decrease in sunk costs affects disproportionately the variations in additive costs.
+Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
 
-Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45 years period, and to highlight the respective part of each component in the determination of these gains.
-Note also that not only the inclusion of additive costs in the underlying framework generates big welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
+\textbf{LISE: A FINIR, A REPRENDRE}
 

### 2021-09-29 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement article partie modele

@@
 In a related manner, it delivers a relevant robustness check on the normative implications of additive costs.
-These results are reported in Table \ref{tab:resultats_modele}, panel (2).
+These results are reported in Table \ref{tab:resultats_modele}, panel (2).\footnote{Under a constant fixed export cost, we also compare two alternative cases.
+Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value, and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974.
+The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces.
+See \textbf{Appendix XXX, Table XXX for the results}.}
 
@@
 In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
-\cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs.
+Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
 If we share a similar conclusion regarding the importance of additive costs, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
@@
 
+\subsection{Some extra theoretical results}
+
+In this section, we run an additional exercise to confirm the importance of the additive dimension in the welfare-detrimental effects of trade costs.
+Under a constant fixed export costs, we simulate the model under two alternative cases.
+Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value, and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974.
+These two scenarii (as well as the initial one based on 1974 values) are reported in Table \ref{tab:calib_TC_appendix}.
+
+\begin{table}[htb]
+  \centering
+  \caption{Exploring alternative scenario: Calibration }\label{tab:calib_TC_appendix}
+\begin{center}
+	\input{table_calibTC_appendix.tex}
+\end{center}
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $t$ is calibrated so that the ratio $t/\widetilde{p}^{\text{fas}}_x$ from the model matches its empirical target, with $\widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
+We theoretically define the export fas price through: $p_x = \tau \widetilde{p}^{\text{fas}} +t$.
+TC = Transport Costs, expressed in percentage of the average fas export price.
+$\beta$ is the share of additive in total transport costs}}}
+\end{table}
+
+The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces.
+
+\begin{table}[htb]
+  \centering
+  \caption{Exploring alternative scenario: Results }\label{tab:calib_TC_appendix}
+\begin{center}
+\begin{tabular}{l|cc|cc}
+& \multicolumn{2}{|c|}{Air transport} & \multicolumn{2}{|c}{Maritime transport} \\ \hline
+$\Delta TC$ (pp)&	-0,02	&-0,014	&-0,033&	-0,031 \\
+$\Delta \tau$ (pp)&	0	&-0,014&	0&	-0,031 \\
+$\Delta t/pfas $(pp)&	-0,020	&0	&-0,033&	0 \\
+$\Delta \beta$ (in pp)	&-27,75&	12,14	&-23,53&	20,36 \\
+%Delta Welfare (abs)&	0,042&	0,007	&0,060&	0,011 \\
+Rel. Delta welf (in \%)&	2,589&	0,411	&3,801	&0,695 \\
+Variety effect (in \%)	&-2,487&	-0,213&	-2,879	&-0,171 \\
+Price effect (in \%)	&5,275	&0,625&	6,973	&0,867 \\
+\hline \hline
+\end{tabular}
+\end{center}
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average fas export price. $\beta$ is the share of additive in total transport costs}}}
+\end{table}
+
 \end{document}

### 2021-10-04 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement réponse et annexe online

@@
 across all districts and all 10-digit products pertaining to the same 5-digit category.
-[...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-inductry pairs, without imposing Assumptions (a) and (b) outlined above.}
+[...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}
 
@@
 By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
-We have implemented the referee's method over the years for which we already have the data, i.e.
-the years 2005-2019, at the $s=3, k=10$ sectoral/product levels.
+As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 only since 2001.
+We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix.
 
 Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset \textbf{over 1974-1988, as the US Bureau of census import database starts in 1989.})
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.)
 

### 2021-10-04 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Premier jet comparaison estimation beta_baseline
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Pour tableaux de comparaison
beta et wgt/qy/baseline

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
changement réponse et annexe online

@@
 
-In line with Table \ref{tab:robustness_separability}, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
+In line with Table 2 of the paper, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
 However, they also show that the difference is quantitatively small in most cases.
@@
 
-\paragraph{Distribution of the $\beta$} As final check, we also investigate whether the varying pattern over time/country of origin/product pointed out in Section \ref{sec:results_decomposition} is also robust to the separability assumption.
+\paragraph{Distribution of the $\beta$} As final check, we also investigate whether the varying pattern over time/country of origin/product is also robust to the separability assumption.
 With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
@@
 Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section \ref{sec:results_trends}, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
+\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section 4 of the paper, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
+
+\subsubsection{Estimating $\widehat{\beta_{i,s}}$ directly}
+
+In this section, we explore an alternative estimation method, which proceeds as follows. $\beta$ can be separately estimated for each industry $s$-origin country $i$ using the following regression:
+
+\begin{equation}
+\ln TC_{ikd} = \beta_{is}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
+\end{equation}
+
+where $d$ denotes the US district of entry and $k$ denotes an HS-10 product, $TC_{ikd}$ being the transport costs. The
+identification of $\beta_{is}$, in this case, would rely on exploiting the variability between sub-sectors at the 10-digit
+level ($k$) and between ports of entry in the US ($d$).
+
+One advantage of this approach is that the above regression can be estimated separately for various country-industry pairs on a linear basis, without having to resort on the separability assumption.
+This does not solve the issue yet, for two main reasons.
+First, we would still be constrained to resort to non-linear estimators.
+This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
+imposing $0 \leq  \beta \leq 1$.
+Should we relax this restriction, standard linear, least squares estimates often deliver negative, meaningless estimates.
+In this respect, implementing this does not suppress the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces).
+
+Second,
+
+\begin{table}[htbp]
+	\caption{Comparison 2005-2013}
+	\begin{center}		
+		\input{Comp_baseline_referee1_reshaped.tex}
+	\end{center}
+	\label{tab_comp_referee1_baseline}%
+\end{table}%
+
+\subsubsection{Weights and quantities}
+
+\begin{table}[htbp]
+	\caption{Comparison 2005-2019}
+	\begin{center}		
+		\input{Comp_baseline_wgt_qy.tex}
+	\end{center}
+	\label{tab_comp_referee1_baseline}%
+\end{table}%
 

### 2021-10-06 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
amendement annexe online + réponse référé 1 + papier

@@
 
-Further, we provide a robustness check for this separative assumption, in Section 4.1 of the paper (\textbf{Table 2, previously submitted version and revised version XXX CHECK AT THE END XXX}).
+Further, we provide a robustness check for this separative assumption, in Section 4.1 of the paper (\textbf{Table 2 in both the submitted and revised versions XXX CHECK AT THE END XXX}).
 We check the robustness of our results by re-running the estimation without the separability assumption.
 This comes at the cost of a substantial increase in the number of fixed effects to include in the regression.
-Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted version and revised version  XXX CHECK AT THE ENDXXX} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
+Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted and revised versions XXX CHECK AT THE END XXX} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
 This is computationally not tractable.
 For this reason, we have decided to run the robustness check on a reduced sample.
-If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013 XXX 2019 ???
-XXX, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
-As we conclude at the end of Section 4.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not - see in particular Figures 6 and 7 on pp.
-25-26.
+If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2019 \textbf{XXX CHECK NUMBERS XXX}, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
+Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
+\textbf{LP: CHECK THE NUMBERS}} As we conclude at the end of Section 4.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not - see in particular Figures 6 and 7 on pp.
+25-26 \textbf{XXX CHECK AT THE END XXX}.
 Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
-
-\textbf{WE ALREADY DID THAT BEFORE. SAY THAT WE POUT THAT EARLIER IN THE PAPER AND CLEARER?}
 
@@
 We develop each line of argument in our answer.
+Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3).
 After reading our answer, we hope to convince you that the choice we made is the most relevant one.
-\textbf{The referee's method is presented in the online appendix XXX WHERE???
-XXX, and we refer to it in Footnote XXX of the paper - A FAIRE EN ONLINE APPENDIX?
-JUSTE DANS LA LETTRE?
-}
 
@@
 Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
-\textbf{LP: A Ai enleve le copie-colle de la footnote dans la lettre; il suffit de dire que c'est dans la version revisee, non?}\smallskip
+\smallskip
 
@@
 Second, we emphasize the drawbacks induced by this method.
+\smallskip
 %
@@
 
-First, the separability assumption does not seem to be a strong assumption.
+First, as already discussed in our letter, the separability assumption does not seem to be a strong assumption.
 This is based on the conclusion drawn from the robustness check on a reduced sample.
-If it is smaller in terms of observations,\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
-Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
-\textbf{LP: CHECK THE NUMBERS}} it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+If it is smaller in terms of observations, it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
 As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
-Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
 
@@
 Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects.
-This substantially reduces the number of point estimates, as well as the value of trade flows covered relative to the  benchmark method, as summarized in \textbf{Table XXX.
-Il faut reprendre cette reponse.
-Reprendre les resultats sur la methode du refere 1}.
+This substantially reduces the number of point estimates, as it can be inferred from Table \ref{tab_comp_referee1_baseline} reported later in our answer and in the Online Appendix, \textbf{Table C.1, Section D.3} ).
 
 \item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}
+
      \textbf{Our answer:}
@@
      $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
-from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).
+\noindent from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).
 Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
@@
 In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
-Section 4 ([Section 3] of the revised version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
+Section 4 ([Section 3] of the submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
 We agree that this was not sufficiently transparent in the submitted version.
@@
 Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
-This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table XXX \textbf{TO DO}.
+This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline}.
 Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
-It excludes XXX \% of trade flows and XXX \% of trade in value.
 In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
-\textbf{LP: Il faut le faire }
 
+    \begin{table}[htbp]
+	\caption{Comparison 2005-2013}
+	\begin{center}		
+		\input{Comp_baseline_referee1_reshaped.tex}
+
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Mean values reported for the 2005 and 2013 values for $\beta$. Time trend coefficient is the annual growth rate.}}}
+\end{center}
+	\label{tab_comp_referee1_baseline}%
+\end{table}%
    \end{enumerate}
-\item[Concern 2] Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and across origin countries.
-Put it in plain words, it means that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors.
-In both cases, one might view this as a disputable assumption.
-By contrast, our method exploits the fact that transport costs have a both a country-specific component and a sector-specific component.
-\textbf{This stands in line with ...
-FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique A FAIRE !!!
-}
 
-\item[Concern 3] The suggested method features less accuracy in the estimation of the $\beta$.
+Notwithstanding these limitations, it remains interesting to run this alternative estimation method as robustness check.
+Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013.
+As noted above, the partners-sectors covered are lower.
+Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
+While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
+In any case yet, the share of additive costs remains substantial, around 40\% on average over the period.
+In this respect, it confirms the main message of our paper shedding light on the importance of the additive component in international transport costs.
+
+\item[Concern 2] The suggested method features less accuracy in the estimation of the $\beta$.
 If we take the value of the $\beta$ by itself, there is no criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
@@
 
-    \item  Our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$.
-\textbf{NOTE THAT $\beta$ is in fact $i,s,t$ because we use individual fas prices to rebuilt it.
-Should we say tha?}.
+    \item  Our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$ (on a year-transport mode basis).
 As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
@@
     \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$, for each couple $i,s$ (by year and transport mode).
-Figures XXX (for Air) and XXX (for Vessel) thus report the histogram of the log of the ratio $\frac{CI_{is}}{CI^{ref}_{is}}$, for the year 2012.
-\textbf{IL FAUT ACTUALISER TOUT CA.
-OU DANS LE FICHIER SORTIE?}
+We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.\textbf{CETTE FIGURE, SUR QUELLE ANNEE, QUEL MODE DE TRANSPORT?
+}
 
@@
 
+One explanation of the limited accuracy with the alternative method might be the following.
+Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and across origin countries.
+Put it in plain words, it means that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
+One might view this as a disputable assumption.
+From a statistical point of view further, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component, that could explain the lower accuracy of the estimation.
+
 \end{itemize}
@@
 However, we also want to keep track of this alternative method as further robustness check.
-Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.\textbf{ XXX TO BE DONE XXX CHECK THIS SECTION EXISTS IN THE ONLINE APPENDIX XXX}
-
-%\textbf{stop here}
+For this purpose, we have amended the online Appendix with a section devoted to the referee's method (\textbf{Section D.3}).
 
@@
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
-We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper \textbf{see XXXX TO BE DONE ???
-XXX???
-LE DIRE DANS LA SECTION DE ROBUSTESSE OU ON FAIT DU 3-10??}
+We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper (see Section 3.3.3 devoted to the robustness to the aggregation level).
+\textbf{XXX A FINIR XXX DANS LE PAPIER ET DANS LA REPONSE}
 
@@
 
-\textbf{LP: J'ai enleve ses equations car c'est totalement faux}
-
 % MEMO: DO NOT FORGET TO REPORT THE ARGUMENTATION BELOW IN THE MAIN TEXT XXX
@@
 \textbf{LP: DOCUMENT THAT.
-SHOW IN THE LETTER THATWE DID IT?}
+SHOW IN THE LETTER THAT WE DID IT?}
 
-\item Second, units are not available for many years over our period, especially in the 1970s and the 1980s \textbf{XXX GUILLAUME, TO BE CONFIRMED AND SUPPORTEDXXX}; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
+\item Second, units are not available for many years over our period, especially in the 1970s and the 1980s \textbf{XXX GUILLAUME, TO BE CONFIRMED AND SUPPORTED XXX}; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
 
@@
 First, beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
-In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2020) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX A QUELLE CRITIQUE FAIT-ON REFERENCE?}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
-\textbf{LP: EST-ce qu'on a des choses a lui montrer a lui sur la regression en quantite?
-Est-ce qu'on la met en annexe en ligne?}
+In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the estimation method implemented by Lashkaripour (2020) (see his Equation (3)) is similar to the alternative method suggested in your Critique 1 above.
+Accordingly, the share of multiplicative costs may produce noisier estimates compared to ours, as we point out in our Answer 2.1.} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+
+Nevertheless, the point made by the referee is an important one and deserves careful investigation.
+This has driven us to run the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis, for the years for which we have the data on quantity.
+\textbf{LP: INCLURE la regression en quantite EN ROBUSTESSE DU PAPIER - A FAIRE}
 
@@
 
-Table 5 in Section 5 reports the results of these various comparative statics exercises for Air and Vessel transport modes.
+Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
 Specifically, we report the change in total transport costs decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
@@
 Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
-
-\textbf{LISE: A FINIR, A REPRENDRE}
 
@@
 
+\newpage
+\bibliographystyle{apalike2}
+\bibliography{biblio}
+
 \end{document}

### 2021-10-06 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
amendement annexe online + réponse référé 1 + papier

@@
 This drives us to estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
-Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin.
-We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i,} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
-Notice that this can be done from 1989 onwards for which HS-10 data are available; \textit{ii,} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
+Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin (hence, around 400 fixed effects).
+We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i),} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
+Notice that this can be done from 1989 onwards for which HS-10 data are available; \textit{ii),} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
 The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2019.
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Appendix \ref{app:4digit}).
-\textbf{JH: WHERE ARE THE 3-10/4-5 ESTIMATES?
-NOT IN ANY APPENDICES.}
- \textbf{XXXX-GD:4-5 ESTIMATES are in table 1. 3-10 I am not sure... On y fait allusion dans la procédure en IV, c’est sûr.}
+However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} and Appendix \ref{app:4digit} for the 4-5 results and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).
+\textbf{ADD 10-3 RESULTS TO BE DONE}
 
@@
 
+Notice that switching to a 3 (sector)-10(product) estimation method does not alter the number of fixed effects to estimate, as they are at the sector $s$ level (along with those at the origin country $i$ level).
+However, within each 3-digit sector category, the 3-10 estimation method has the advantage to exploit a larger source of heterogeneity by considering trade flows at the HS-10 product level.
+
 Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are similar when using 10-digit product and 5-digit products.
@@
 The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces.
-See \textbf{Appendix XXX, Table XXX for the results}.}
+See Appendix \ref{app:theoretical_model}, Table \ref{tab:results_model_appendix} for the results}.
 
@@
 Under a constant fixed export costs, we simulate the model under two alternative cases.
-Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value, and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974.
+Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value (Scenario (A)), and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974 (Scenario (B)).
 These two scenarii (as well as the initial one based on 1974 values) are reported in Table \ref{tab:calib_TC_appendix}.
@@
 
-The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces.
+The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces (Scenario (A)).
+In contrast, under Scenario (B), the reduction in total trade costs is still welfare-enhancing, but the magnitude of the welfare gains is much more modest, because of the larger distortion induced by additive costs - that now account for a larger share of total trade costs.
 
@@
   \centering
-  \caption{Exploring alternative scenario: Results }\label{tab:calib_TC_appendix}
+  \caption{Exploring alternative scenario: Results }\label{tab:results_model_appendix}
 \begin{center}
 \begin{tabular}{l|cc|cc}
+\hline \hline
 & \multicolumn{2}{|c|}{Air transport} & \multicolumn{2}{|c}{Maritime transport} \\ \hline
-$\Delta TC$ (pp)&	-0,02	&-0,014	&-0,033&	-0,031 \\
-$\Delta \tau$ (pp)&	0	&-0,014&	0&	-0,031 \\
-$\Delta t/pfas $(pp)&	-0,020	&0	&-0,033&	0 \\
-$\Delta \beta$ (in pp)	&-27,75&	12,14	&-23,53&	20,36 \\
-%Delta Welfare (abs)&	0,042&	0,007	&0,060&	0,011 \\
-Rel. Delta welf (in \%)&	2,589&	0,411	&3,801	&0,695 \\
-Variety effect (in \%)	&-2,487&	-0,213&	-2,879	&-0,171 \\
-Price effect (in \%)	&5,275	&0,625&	6,973	&0,867 \\
+Scenario & (A) & (B) & (A) & (B) \\ \hline
+$\Delta TC$ (pp)&	-0.02	&-0.014	&-0.033&	-0.031 \\
+$\Delta \tau$ (pp)&	0	&-0.014&	0&	-0.031 \\
+$\Delta t/\widetilde{p}^{fas} $(pp)&	-0.020	&0	&-0.033&	0 \\
+$\Delta \beta$ (in pp)	&-27.75&	12.14	&-23.53&	20.36 \\ \hline
+%Delta Welfare (abs)&	0.042&	0.007	&0.060&	0.011 \\
+Rel. Delta welf (in \%)&	2.589&	0.411	&3.801	&0.695 \\
+Variety effect (in \%)	&-2.487&	-0.213&	-2.879	&-0.171 \\
+Price effect (in \%)	&5.275	&0.625&	6.973	&0.867 \\
 \hline \hline
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average fas export price. $\beta$ is the share of additive in total transport costs}}}
+{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average fas export price. $\beta$ is the share of additive in total transport costs.}}}
 \end{table}

### 2021-10-06 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
amendement annexe online + réponse référé 1 + papier

@@
 
-\textbf{XXX UPDATE XXX : Est-ce fait?}
 In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
@@
 
-\subsubsection{Estimating $\widehat{\beta_{i,s}}$ directly}
+\subsubsection{Estimating $\widehat{\beta}_{i,s}$ directly}
 
@@
 \begin{equation}
-\ln TC_{ikd} = \beta_{is}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
+\ln TC_{ikd} = \beta_{is}\ln \tilde{p}_{ikd} + Z_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
 \end{equation}
 
-where $d$ denotes the US district of entry and $k$ denotes an HS-10 product, $TC_{ikd}$ being the transport costs. The
-identification of $\beta_{is}$, in this case, would rely on exploiting the variability between sub-sectors at the 10-digit
+where $d$ denotes the US district of entry and $k$ denotes an HS-10 product, $TC_{ikd}$ being the transport costs and $Z$ a set of controls variables.
+The identification of $\beta_{is}$, in this case, relies on exploiting the variability between sub-sectors at the 10-digit
 level ($k$) and between ports of entry in the US ($d$).
+From this, one can then recover the levels of additive and multiplicative transports costs.
+Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ for a given sector-country $i,s(k)$, one can indeed solve the following two-equation system:
+
+\begin{eqnarray*}
+p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
+\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
+\end{eqnarray*}
+
+\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
 
@@
 This does not solve the issue yet, for two main reasons.
-First, we would still be constrained to resort to non-linear estimators.
+First, it is still constrained to resort to non-linear estimators.
 This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
@@
 
-Second,
+It should also be noted that this estimation method implies a lower coverage, both in time and in the sectors/countries covered.
+Information about the port of entry are only available since 1989, which is already smaller than our full sample starting in 1974.
+On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
+Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
+In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
+
+With respect to the trade flows coverage, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
+Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
+This can be seen comparing the number of observations by year/ transport mode between this alternative method and our baseline method reported in Table \ref{tab_comp_referee1_baseline}.
+Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
+
+Notwithstanding these limitations, it remains interesting to run this alternative estimation method as robustness check.
+Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013.
+As noted above, the partners-sectors covered are lower.
+Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
+While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
+In any case yet, the share of additive costs remains substantial, around 40\% on average over the period, thereby confirming the main message of our paper shedding light on the importance of the additive component in international transport costs.
 
@@
 		\input{Comp_baseline_referee1_reshaped.tex}
+
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Mean values reported for the 2005 and 2013 values reported for $\beta$. Time trend coefficient is the annual growth rate.}}}
 \end{center}
@@
 
-This section complements Section \textbf{XXX} in the main paper, where we replicate the method adopted by \cite{hummels2007} on our database.
+This section complements Section 4 in the main paper, where we replicate the method adopted by \cite{hummels2007} on our database.
 Here, we report the unfitted measure of transport costs versus the ``pure'' transport costs series, i.e composition effects excluded

### 2021-10-07 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
amendement réponses référés

--- 2021-09-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
@@
 \title{\textsc{International Transport costs:\\New Findings from modeling additive costs} \\
-Answer to Referee 1}
+Answer to Referee 2}
 
@@
 [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
-\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness to aggregation, robustness to IV estimates).
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and the results, in Section 3, which now includes the robustness checks as final sub-section.
+\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to IV estimates have been added).
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, which reports the results and now includes the robustness checks as final sub-section.
 For sake of space saving, the robustness analysis related to composition effects has been set to the Appendix (Section C.3).
@@
 
-\subsection{Critique 1: Empirical Strategy}
+\subsection{Implications/usefulness of the exercise}
 
-\textit{Using the notation of the authors, they are intreated
-in identifying the share of the specific cost in the total transport cost. Namely,}
-\begin{eqnarray*}
-&& \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} \\
-\text{or,} &&\frac{t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{is(k)} + t_{is(k)}}
-\end{eqnarray*}
+\paragraph{``Big Picture''} \textit{[...] The main contribution of this paper thus appears to be to unravel the temporal evolution / relative importance of the ad-valorem and specific transport cost component.
+However, besides showing this evolution, the paper is basically silent on the implication(s) of the patterns found.
+In the conclusion, the authors state that this is left for future research.
+This was quite disappointing, simply documenting the evolution of these two components leaves the reader (at least me) wondering why he/she should care about this.
+The authors should aim to be much more clear on what the trends in ad-valorem / specific transport costs mean/imply from the perspective of these recent theoretical contributions/insights.} \smallskip
 
-\textit{The way they approach the problem is that they assume that (a) $\tau_{ik} = \tau_i\tau_{k}$,
-(b) $t_{ik} = t_i +t_k$, and (c) $t_k$ and $\tau_k$ are uniform across products within industry
-s. After imposing these assumptions, they estimate the following specification}
-
-\begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:equation0}
-\end{equation}
-
-\textit{in which $\tau_i$, $\tau_{s(k)}$, $t_i$, and $t_{s(k)}$ are identified as fixed effects coefficients.
-In my opinion this choice of strategy is quite sub-optimal, as (i) it relies on
-the strong assumptions highlighted above, (ii) it is computationally expensive
-as noted by the authors on multiple occasions, and (iii) it is subject to an
-endogeneity problem, which the authors disregard with one sentence, but which
-is rather detrimental in my opinion.}
-
-%\paragraph{Strategy of answer} To clarify the discussion, we decided to answer concerns (i) and (iii) each in turn.
-As regards to point (ii), we run the referee's proposed estimation method without relying on instrumental variables; this way, the results are comparable with the benchmark results, the only change being the estimation strategy (Section \ref{subsec:functional_form} of the letter).
-As noted above, this drove us to maintain our estimation method as benchmark.
-As regards to point (iii), we instrument fas prices at the $k=$ 10-digit level, still based on our original estimation strategy (but switching to a finest product category disaggregation level $k=10$).
-This allows a direct comparison of the results obtained in the non-linear least squares regression with those obtained with IV (with $s=3$ digit at the sector level and $k=10$ at the product level in both cases).
-This is developed in \textbf{Section ???
-of the letter}.
-
-The three points raised by the referee indeed deserve careful consideration. We provide separate answers to each of them below.
-
-%\subsubsection{On the justifications for our empirical approach}
-
-\subsubsection{Concern (i): About the assumption of separability}
-Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by \citet{Irrazabal_2015} to estimate the share of additive costs in a firm-level context.
-It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.
-
-Further, we provide a robustness check for this separative assumption, in Section 4.1 of the paper (\textbf{Table 2, previously submitted version and revised version XXX CHECK AT THE END XXX}).
-We check the robustness of our results by re-running the estimation without the separability assumption.
-This comes at the cost of a substantial increase in the number of fixed effects to include in the regression.
-Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted version and revised version  XXX CHECK AT THE ENDXXX} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
-This is computationally not tractable.
-For this reason, we have decided to run the robustness check on a reduced sample.
-If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2013 XXX 2019 ???
-XXX, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
-As we conclude at the end of Section 4.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not - see in particular Figures 6 and 7 on pp.
-25-26.
-Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
-
-\textbf{WE ALREADY DID THAT BEFORE. SAY THAT WE POUT THAT EARLIER IN THE PAPER AND CLEARER?}
-
-\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is highly demanding computationally when the number of fixed effects increase.
-This constraint notably drove us to impose the separability assumption and retain $s=3$ as the relevant sectoral degree of aggregation.
-
-The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummel's Methodology.
-That is, one can alternatively estimate the share of the additive component as:}
-$$\frac{t_{ik}}{ (\tau_{ik}-1)\tilde{p}_{ik} + t_{ik}} = \beta_{ik}$$
-
-\textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t.
-unit price.
-Given the authors' objective and the data they are using, $\beta$ can be separately estimated for each industry-country pair using the following regression:}
-
-\begin{equation}
-\ln f_{ikd} = \beta_{is(k)}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
-\end{equation}
-
-\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product} ($f_{ikd}$ being the transport costs). \textit{The
-identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
-district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
-obviously require that the authors do not aggregate up the raw Census data
-across all districts and all 10-digit products pertaining to the same 5-digit category.
-[...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-inductry pairs, without imposing Assumptions (a) and (b) outlined above.}
-
-\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummel's.
-We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
-As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
-The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
-
-Now coming to the technical side, we thank the referee for his/her relevant suggestion of an alternative estimation method.
-This drove us to question our own empirical specification deeper.
-In the end, we yet decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that you propose does not go without a few important limitations.
-We develop each line of argument in our answer.
-After reading our answer, we hope to convince you that the choice we made is the most relevant one.
-\textbf{The referee's method is presented in the online appendix XXX WHERE???
-XXX, and we refer to it in Footnote XXX of the paper - A FAIRE EN ONLINE APPENDIX?
-JUSTE DANS LA LETTRE?
-}
-
-\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares.
-However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
-This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
-$\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
-Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates.
-In this respect, implementing the referee's method (see below) does not suppress the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
-\textbf{LP: A Ai enleve le copie-colle de la footnote dans la lettre; il suffit de dire que c'est dans la version revisee, non?}\smallskip
-
-\paragraph{Exploring Referee's alternative functional form \label{subsec:functional_form}}
-
-The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007), yielding Equation (\ref{eq:estimation_ref1}) as estimation equation.
-By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
-We have implemented the referee's method over the years for which we already have the data, i.e.
-the years 2005-2019, at the $s=3, k=10$ sectoral/product levels.
-
-Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset \textbf{over 1974-1988, as the US Bureau of census import database starts in 1989.})
-
-Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
-First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
-Second, we emphasize the drawbacks induced by this method.
-%
-%\paragraph{What we do}
-%
-%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
-We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
-To sum up our methodology:
-%\begin{itemize}
-%\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
-%\item We rebuild
-%\begin{eqnarray*}
-%\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
-%\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
-%\end{eqnarray*}
-%\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
-%\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
-%
-%
-%$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
-%\end{itemize}
-%
-%\paragraph{What the referee suggests to do}
-%
-%The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007).
-As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
-The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
-Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
-%
-%Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{The referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We followed his suggestion but considering the 3-digit level.
-We made this choice in view of being able to compare the pros and cons of the referee's method versus ours.
-In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004.
-For comparability reasons over time, we adopt the same degree of classification $k=5$ for the following years (based on the original Census dataset).
-As a result, transport costs are estimated at the 3-digit level as benchmark estimate, exploiting heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode).
-This has driven us to retain $s=3$ when implementing the referee's method, in view of preserving comparability between both strategies.}
-
-\paragraph{a) The advantages of the referee's method are not as important as suggested}
-
-\begin{itemize}
-\item \textit{The first advantage of this so-called Hummel's Approach is that the above
-regression can be estimated separately for various country-industry pairs, without
-imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} \\
-
-\textbf{Our answer:}
-
-First, the separability assumption does not seem to be a strong assumption.
-This is based on the conclusion drawn from the robustness check on a reduced sample.
-If it is smaller in terms of observations,\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
-Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
-\textbf{LP: CHECK THE NUMBERS}} it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
-As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
-Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
-
-Second, the referee's method imposes a drastic selection of countries / sectors to be included in the estimation.
-Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects.
-This substantially reduces the number of point estimates, as well as the value of trade flows covered relative to the  benchmark method, as summarized in \textbf{Table XXX.
-Il faut reprendre cette reponse.
-Reprendre les resultats sur la methode du refere 1}.
-
-\item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}
-     \textbf{Our answer:}
-We thank the referee for this valuable suggestion.
-It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.
-
-\item  \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.}\\
-     \textbf{Our answer:}
-     We agree with the referee that the comparison with the literature (\citealp{hummels2007}) in particular) was not straightforward for the reader in the submitted version.
-We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
-In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
-    \begin{equation}
-    f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta_{ikt}} \label{eq:Hummels}
-    \end{equation}
-
-     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ikt}$ the price-elasticity of transport costs, with $i$ the origin country, $k$ the product and $t$ for time.
-As we now show in the paper (Section 2.2), this also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper.
-We directly report them here to ease the comparison with the referee's method afterwards.}
-
-     \begin{equation}
-     \beta_{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)t}-1+\frac{t_{is(k)t}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
-     \end{equation}
-
-     Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method.
-In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a constant $\beta$ invariant over time/sector/origin country (even though mode-specific).
-Rather than Equation (\ref{eq:Hummels}), this amounts starting from the functional form: $f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta}$.
-
-     Two alternative strategies are yet available.
-Let us start with the referee's method.
-It starts from the functional form linking transport costs and the fas price as specified in the Equation (\ref{eq:Hummels}).
-Taken in log (on a mode/yearly basis), it is written as:
-     $$\ln f_{ikd} = \beta_{is(k)}\ln \widetilde{p}_{ikd} +\text{Controls}_{ikd}+ \varepsilon_{ikd}$$
-
-     \noindent with $d$ the district of entry.
-From this, one can then recover the levels of additive / multiplicative transports costs.
-Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ for a given sector-country $i,s(k)$, one can indeed solve the following two-equation system:
-
-\begin{eqnarray}
-p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
-\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
-\end{eqnarray}
-
-\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
-With 2 equations and 2 endogenous variables, the system might be solved.
-
-     Alternatively, our method rather starts from the definition of transport costs as:
-     $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
-from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).
-Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
-
-In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
-Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such.
-In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
-Section 4 ([Section 3] of the revised version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
-We agree that this was not sufficiently transparent in the submitted version.
-We hope that the revised version now fits the referee's expectations on this point.
-
-\end{itemize}
-
-\paragraph{b) The costs are high}
-\begin{itemize}
-\item[Concern 1] The suggested estimation strategy implies having much less data to exploit, for two reasons.
-\begin{enumerate}
-\item Information about the port of entry are only available since 1989, which is already smaller than our full sample starting
-in 1974.
-On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
-Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
-In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
-Eliminating this dimension of the paper would be detrimental to its value-added.
-\item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
-This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table XXX \textbf{TO DO}.
-Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
-It excludes XXX \% of trade flows and XXX \% of trade in value.
-In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
-\textbf{LP: Il faut le faire }
-
-    \end{enumerate}
-\item[Concern 2] Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and across origin countries.
-Put it in plain words, it means that transport costs of say, tissues, have nothing in common whether those goods comes from France or from Bangladesh; or that transport costs that apply to imported goods from France have no common component across sectors.
-In both cases, one might view this as a disputable assumption.
-By contrast, our method exploits the fact that transport costs have a both a country-specific component and a sector-specific component.
-\textbf{This stands in line with ...
-FAIRE REFERENCE A DES PAPIERS EN TRADE QUI ONT DES TRADE COSTS Pays-spécifique A FAIRE !!!
-}
-
-\item[Concern 3] The suggested method features less accuracy in the estimation of the $\beta$.
-If we take the value of the $\beta$ by itself, there is no criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
-Things are more clear-cut in terms of accuracy of the estimation.
-Specifically, our method yields a more precise estimation of the $\beta$ than the referee's method.
-To develop on this, for each year and transport mode:
-    \begin{itemize}
-    \item With the referee's method, we estimate one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
-From this, we can approximate the 5-95\% threshold values through:
-        $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
-
-        From this, we can deduce the confidence interval $CI^{ref}_{is} = \hat{\beta}_{is}^{max,ref} - \hat{\beta}_{is}^{min,ref}$.
-
-    \item  Our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$.
-\textbf{NOTE THAT $\beta$ is in fact $i,s,t$ because we use individual fas prices to rebuilt it.
-Should we say tha?}.
-As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
-We then deduce the accuracy of the $\beta$ estimate by relying on bootstrap method.
-Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
-Noticing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
-}
-
-    \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$, for each couple $i,s$ (by year and transport mode).
-Figures XXX (for Air) and XXX (for Vessel) thus report the histogram of the log of the ratio $\frac{CI_{is}}{CI^{ref}_{is}}$, for the year 2012.
-\textbf{IL FAUT ACTUALISER TOUT CA.
-OU DANS LE FICHIER SORTIE?}
-
-\begin{figure}[htbp]
-\caption{Accuracy of the estimation of the $\beta$: Comparison}
-\label{fig:accuracy_beta}
-\begin{center}
-\includegraphics[height=4in]{accuracy_beta.pdf}
-\end{center}
-\end{figure}
-
-In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
-Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
-We have checked that a similar conclusion applies on other years (they are not reported here for sake of brevity but they are available upon request).
-
-\end{itemize}
-\end{itemize}
-
-All these elements drive us to maintain our original estimation as benchmark method in the revised version, and we deeply hope to convince you of the relevance of this choice.
-However, we also want to keep track of this alternative method as further robustness check.
-Fr this purpose, we have amended the online Appendix with a section devoted to the referee's method.\textbf{ XXX TO BE DONE XXX CHECK THIS SECTION EXISTS IN THE ONLINE APPENDIX XXX}
-
-%\textbf{stop here}
-
-\subsubsection{Concerns about endogeneity}
-
-\textit{\textbf{The endogeneity problem}: quoting Footnote 14 of the paper, the authors
-are estimating $t_i$ and $t_{s(k)}$ as coefficients on the industry or country
-dummies times $1/\widetilde{p}_{ik}$. [...] Based on the productivity-sorting model in Melitz (2003) or the quality-sorting
-model in Baldwin and Harrigan (2010),  $1/\widetilde{p}_{ik}$ is either positively
-or negatively correlated with $\epsilon_{ik}$. So, the NNLS estimates are biased; and
-the bias has nothing to with the casual versus accounting interpretation of
-the estimates. Accordingly, the one-line justification the authors provide
-to not address the endogeneity problem is far from convincing.}
-
-Indeed, this is a very important point.
-The referee states that, based on theoretical insights by \citet{melitz} or \citet{baldwin_harrigan}, $1/\widetilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
-In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
-
-We obviously do not question this conceptual issue.
-However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF (p) and the FAS ($\widetilde{p}_{ik}$) prices.
-Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out.
-That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
-
-Following the referee's advice, we decided therefore to provide a full set of IV estimates to provide a clean assessment of the size of the potential bias.
-Section 4.3 in the revised version of the paper provides an overview of the results, while section B.4 in Appendix B provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
-We follow earlier literature (see e.g.
-Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
-First-stage estimates reported in section B.4 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation.
-Figure 9 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart.
-In all cases, these estimates are very similar, when not identical.
-This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
-
-It should be noted that we performed this check on our main dataset (SITC 5 digit), and also at the HS 10 level, to handle simultaneously the referee's concern about aggregation issues.
-In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very, very similar to their non-IV, least squares counterparts.
-Needless to say that these results are available upon request from the referee.
-
-\subsubsection{The aggregation problem}
-
-\textit{\textbf{The aggregation problem}: The original annual Census data reports
-trade at the origin country-HS10 product-district level of aggregation,
-whereas the authors are aggregating up the data even further to the origin
-country-HS5 industry-year level. Such an aggregation comes with strong implicit assumptions and sacrifices a lot of useful variation in the data.
-The authors are motivating the aggregation by stating that the problem
-would become computationally expensive without it. But this reasoning
-brings us back to my original point that the authors can use the Hummel's
-Methodology to circumvent the computational burden.}
-
-It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
-Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
-We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
-After checking with the US Bureau of Census, data is only available since 1989, now available to 2019.
-As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.
-In front of this arbitrage, we decided to maintain the case $s=3, k=5$ as baseline, but running estimation at the $s=3,k=10$ in the new Sub-Section 3.3.3 of Section 3.3 devoted to robustness analysis.
-
-Additionally, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
-This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
-We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper \textbf{see XXXX TO BE DONE ???
-XXX???
-LE DIRE DANS LA SECTION DE ROBUSTESSE OU ON FAIT DU 3-10??}
-
-\subsection{Critique 2: Calculation of Unit Prices}
-
-\textit{My second critique concerns the way the authors are calculating the unit prices.
-The Census data reports the quantity of goods per observation. So, the authors
-can calculate the unit price as Value/Quantity, which is consistent with how
-price is modeled in standard trade models. Instead, the authors calculate unit
-price as Value/Weight. This used to be a common exercise in the past where
-many data-sets did not report Quantity. But, given their data, there is no
-justification for the authors to calculate the prices this way.}
-
-\textit{Calculating the unit price as Value/Weight presents the authors
-with an additional endogeneity problem. To elaborate, let $\omega_{ik} = Weight/Quantity$
-denote the unit weight of the goods in observation $ik$. [...] There is evidence that (i) $\omega_{ik}$ varies significantly within narrowly-defined product
-categories, and (ii) $\omega_{ik}$ is negatively correlated with transport costs. So the
-way the authors are calculating unit prices and estimating the model creates a
-new (but avoidable) source of endogeneity.}
-
-\textbf{LP: J'ai enleve ses equations car c'est totalement faux}
-
-% MEMO: DO NOT FORGET TO REPORT THE ARGUMENTATION BELOW IN THE MAIN TEXT XXX
-\paragraph{Our answer:}
-\noindent The referee definitely has a point here, in the sense that Census data do report the quantity of goods.
-In addition, referee's remarks echo recent findings by Lashkaripour (\emph{Journal of International Economics}, 2020): based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
-Finally, Lashkaripour (2020) finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
-
-We devoted a lot of thought to a potential switch from weight to quantity to get cleaner unit prices. However, the latter appears really uneasy in our context for several, non-trivial reasons:
-
-\begin{enumerate}
-\item The information on quantity is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
-We could of course restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
-\textbf{LP: DOCUMENT THAT.
-SHOW IN THE LETTER THATWE DID IT?}
-
-\item Second, units are not available for many years over our period, especially in the 1970s and the 1980s \textbf{XXX GUILLAUME, TO BE CONFIRMED AND SUPPORTEDXXX}; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
-
-\item Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious \textbf{XXX GUILLAUME: EXAMPLES XXX}.
-At a more fundamental level, one may argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
-For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
-\textbf{XXX GUILLAUME: Do we have evidence, even anecdotal, on this point?}
-
-\end{enumerate}
-
-These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression.
-That said, do we have reason to believe this measure of unit prices generate significant deviations from the one relying on quantities?
-
-\citet{Lashkaripour_JIE2020} does provide important results related to ours.
-In particular, his result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 30 to 45\% of total costs.
-The two stories are not necessarily incompatible, however.
-First, beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
-In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the share of multiplicative costs is derived from a formulation - Equation (3) in Lashkaripour (2020) - that may produce noisier estimates compared to ours, as we point in answer 2.1.ii to your Critique 1 above XXX \textbf{GUILLAUME : TO BE DISCUSSED AGAIN TOGETHER XXX A QUELLE CRITIQUE FAIT-ON REFERENCE?}} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
-\textbf{LP: EST-ce qu'on a des choses a lui montrer a lui sur la regression en quantite?
-Est-ce qu'on la met en annexe en ligne?}
-
-\subsection{Critique 3: Big Picture Implications}
-
-\textit{My third critique concerns the lack of an exciting punchline. The fact that
-composition effects have not countervailed the reduction in pure transport costs
-(at least not as much as previously believed) is an interesting but minor observation.
-Does this observation revise our understanding of say the gains from
-trade? Does it shed new light on a puzzle many people are thinking about?
-One crude suggestion is to see how the reduction in the industry-specific cost
-terms is related to the industry-level trade elasticities. If the composition effects
-favor low-elasticity industries, the findings in the paper may have first-order
-implications for the gains from trade.
-Another suggestion is to dig deeper into the relative rate at which additive
-and multiplicative transport costs have declined over time. Since additive transport
-costs favor rich (high-quality exporting) countries, the disproportionally
-greater reduction in additive costs can perhaps explain the rise of low-income
-exporter as documented by Hanson (2012, JEP).}
-
-\paragraph{Our answer:}
-\noindent We devoted a lot of thought to this point, which directly echoed a similar concern by the other referee.
+\textbf{Our answer:} We devoted a lot of thought to this point, which directly echoed a similar concern by the other referee.
 Taking stock of your suggestion, we decided to offer insights on the welfare implications of our results.
@@
 
-\textbf{LISE: A FINIR, A REPRENDRE}
+\paragraph{Exploring the heterogeneity of our results} \textit{At the very least the authors could/should make use of the fact that the temporal patterns in specific / ad-valorem transport costs is likely to show substantial heterogeneity across goods as well as countries.
+Exploring this heterogeneity to a fuller extent could already hint at the main drivers (insurance costs / containerization / economies of scale) behind the observed patterns.
+This could be done by e.g.
+not only reporting the mean and median of the ad-valorem / specific transport cost components but exploring more fully the heterogeneity in these transport costs components across sectors/exporters.
+Also, it should be relatively straightforward to provide some analysis of the major correlates with the identified specific and ad-valorem costs (e.g.
+product value/weight ratio, distance to the exporter, etc.).}
 
-%XXX is chosen to match the share of firms exporting in Norway (38 percent, Moxnes, 2010), while
+\textbf{WHAT DO WE DO ABOUT THAT?
+We provide an analysis by sector about the time trends.
+Do more than that: Regress the additive costs/ ad-valorem costs $t_{is}$, $\tau_{is}$ on distance from the US?
+correlation btw each component and the sector/country fixed effect ?}
 
-%IMO : The data cover all Norwegian non-oil exporters in 2004 and originate from customs declarations.
+\subsection{Empirical strategy}
 
-%\textbf{Comments about that}
+\textit{The empirical strategy used to unravel the specific and ad-valorem component in international
+transport costs is quite elaborate. [...] To be more specific, why not simply estimate equation (1) while plugging
+in equation (3) and (4) for $\tau$ and $t$ respectively? Doing this yields the following estimation equation (\ref{eq:ref2_basis}) in
+levels (I have added a time subscript t as well as an additive error term $\epsilon$):}
 
-%\paragraph{Piste 1} Consider his/her first suggestion.
-What is the idea?
-First, the link between gains from trade and trade-elasticity.
-Am I correct in saying that gains from trade are higher when trade is about low-elasticity goods?
-The idea being, if national goods are low substitutes, then here are larger gains from trading them.
-So, if composition effects (trade shifts between sectors) favor low-elasticity sectors, then one might expect high gains from trade.
-Hummels: trade composition effects matter as they partially offset the reduction in pure transport costs for both air and vessel.
-Over time, tendency to trade goods more costly to export everything else equal.
-So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade.
-But, we disagree with Hummel's findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export).
-Meaning that gains from trade are purely due to reduction in transport costs at the sector, but not from switch towards low-elasticity sectors.
-Identify one source of gains from trade, not two.
+\begin{equation}
+p_{ikt}=\tau_{it}\widetilde{p}_{ikt}+ \tau_{kt}\widetilde{p}_{ikt} + t_{it} + t_{kt} + \epsilon_{ikt}   \label{eq:ref2_basis}
+\end{equation}
 
-%What to do with this?
-When composition effects do matter (ocean freight), do they favor low-elasticity industries?
-In which case, on top of the reduction in transport costs per se (which has welfare consequences as well), one might expect additional gains from these composition effects.
-When they don't, this mitigates the gains from trade that could be expected from the picture given by Hummels.
+\textit{I do not see why the authors need the complicated non-linear transformations to do what they aim
+for. Estimating Equation (\ref{eq:ref2}) should be relatively easy, for one it is linear, but it would also not require the
+inclusion of any country-sector fixed effects.}
 
-%\textbf{GD20200708 Hum...
-Je crois que je comprends le point dur referee et je suis d’accord avec lui.
-Le souci, c’est que de toutes les manières, je ne sais pas si cela favorise les high ou les low elasticity.
-Mais on pourrait voir.}
+\begin{equation}
+p_{ikt}= \sum_i \sum_t \alpha_{it}^\tau \mathbbm{1}_{it}\widetilde{p}_{ikt} +  \sum_k \sum_t \alpha_{kt}^\tau \mathbbm{1}_{kt}\widetilde{p}_{ikt} + \sum_i \sum_t \alpha_{it}^t \mathbbm{1}_{it} + \sum_k \sum_t \alpha_{it}^t\mathbbm{1}_{kt} \label{eq:ref2}
+\end{equation}
 
-%\paragraph{Piste 2} What is at stake?
-Show that how $\beta$ = the relative share of additive costs has evolved over time.
-Good idea as this is really what differentiates us from Hummels, model a varying $\beta$ over time.
-Specifically, how methodology lets the $\beta$ vary over the three time-sector-partner country dimension.
-Show how $\beta$ varies over time everything else equal?
-Hopefully, goes in the referee's direction (decreases over time) with possibly a role in the understanding in the ``big picture'' of trade patterns?
+\textbf{Our answer:} As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
+We agree that resorting on a linear form as the one suggested by the referee would substantially alleviate the computational burden.
+It does not seem to solve the issue yet.
 
-%\paragraph{Piste 3} \textbf{GD20200708 On sait que les additive costs sont plus distorsifs que les multiplicative costs, puisqu’ils modifient les prix relatifs.
-Donc le fait qu’ils sont une partie croissante des costs signifie qu’on sur-estime l’augmentation des gains de welfare liés à la baisse des coûts de transport}
+First, it can be noticed that the equation suggested by the referee does not stand in accordance with the assumption we borrow from \cite{Irrazabal_2015}, that ad-valorem costs is decomposed in its two country/sector dimensions multiplicatively (i.e.
+approximating $\tau_{is(k)}$ by the product $\tau_i\times \tau_{s(k)}$.
+
+However, even with another formulation, such as the equations (\ref{eq:ref2_basis}) and (\ref{eq:ref2}) suggested by the referee, we would still be constrained to resort to non-linear estimators.
+This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
+$\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
+Without these restrictions, standard linear, least squares estimates deliver aberrant values with more than mild quality-of-fit.
+Taking into account this constraint implies that the error term should be always positive, which we ensure by specifying the error term such that:
+\begin{equation*}
+\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i}\times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
+\end{equation*}
+\noindent where $\epsilon_{ik}$ follows a normal law centered on 0.
 
@@
 \end{document}
-
-\newpage
-
-\appendix
-
-% XXX JH: What do we do with the section below?
-
-\section{Clarifying some technical points \label{app:technical_points}}
-
-\subsection{A note on prices}
-
-We have three prices: the (US) consumer price, say $p^{c}_{ik}$, the ``import'' or  cif price (at the entrance of the US) $p_{ik}$ and the fas price (at the export gate in the origin country) $\widetilde{p}_{ik}$, denoting $i$ the origin country, $k$ the product at the 8 or 5 digit level, and reasoning on a yearly basis, and $s(k)$ the 3-digit classification the product $k$ belongs to.
-Further, if we denote $\tau^d_{ik}$ the duty tax rate paid when the good crosses the US border, then we have:
-
-\begin{eqnarray*}
-&&p^c_{ik} = (1+\tau^d_{is(k)})p_{ik} \\
-with && p_{ik}  = \tau_{is(k)} \widetilde{p}_{ik} +t_{is(k)}
-\end{eqnarray*}
-
-\subsection{Deriving the share of additive costs in total transport costs \label{app:interpret_beta}}
-
-With $i$ the origin country, $k$ the product category at the HS6 level (and reasoning at the year- transport mode level), $p_{ik}$ the import (cif) price and $\tilde{p}_{ik}$ the export (fas) price, transport costs $f_{ik}\equiv \frac{p_{ik}- \tilde{p}_{ik}}{\tilde{p}_{ik}} $ are written in Hummel's terminology as:
-
-\begin{equation*}
-f_{ik} = X_{is(k)}\tilde{p}_{ik}^{-\beta_{ik}} \label{eq:Hummels}
-\end{equation*}
-
-It is trivial to show that:
-$$\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} \frac{\tilde{p}_{ik}}{f_{ik}}= -\beta_{ik}$$
-
-Then
-\begin{itemize}
-\item If $\beta_{ik} = 0$, then $p_{ik} = (1+X_{is(k)})\tilde{p}_{ik}$: Only ad-valorem transport costs
-\item If $\beta_{ik} = 1$, then $p_{ik}=\tilde{p}_{ik}+X_{is(k)}$: Only additive costs
-\end{itemize}
-
-This suggests that, the closer $\beta_{ik}$ to 1, the more prevalent additive costs in total transport costs.
-Can $\beta_{ik}$ be related to the share of additive costs in total transport costs?
-The answer is positive.
-To show this clearly, start from:
-
-\begin{eqnarray*}
-f_{ik} &=& \frac{p_{ik}-\tilde{p_{ik}}}{\tilde{p}_{ik}}\\
-\text{with}:&& p_{ik} = \tau_{is(k)}\tilde{p}_{ik} + t_{is(k)} \\
-\end{eqnarray*}
-
-assuming that transport costs decompose in two components, additive transport costs $t_{is(k)}$ and multiplicative costs (ad-valorem) $\tau_{is(k)}$, supposed to be identical in the product-dimension for any $k$ product within the 3-digit classification $s$ it belongs to.
-
-This gives:
-
-$$f_{ik} = \frac{(\tau_{is(k)}-1)\tilde{p}_{ik}+t_{is(k)}}{\tilde{p}_{ik}}$$
-
-From this, we have
-
-\begin{eqnarray*}
-\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} &=& \frac{(\tau_{is(k)}-1)\tilde{p}_{ik} - ((\tau_{is(k)}-1)\tilde{p}_{ik} +t_{is(k)})}{\tilde{p}_{ik}^2} \\
-&=& \frac{-t_{is(k)}}{\tilde{p}_{ik}^2}
-\end{eqnarray*}
-
-Hence
-\begin{eqnarray*}
-\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} \frac{\tilde{p}_{ik}}{f_{ik}}&=& \frac{-t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{ik} +t_{is(k)}}\\
-&=& \frac{-\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 +\frac{t_{ik}}{\tilde{p}_{ik}}}
-\end{eqnarray*}
-
-This shows that $\beta_{ik}$ can be interpreted as the share of additive costs in total transport costs:
-
-$$\beta_{ik} = \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
-
-\subsection{Share vs level}
-
-Starting from the above reasoning, the referee proposes to estimate the elasticity of transport costs with respect to the unit price, equivalently the share of additive costs in total costs.
-Specifically, the referee suggests to estimate for each industry-country the following regression:
-
-$$\ln f_{ikd}= \beta_{is(k)} \ln \widetilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd}$$
-
-The question is then, how to recover the levels of additive / multiplicative transports costs $t_{is(k)}$ and $\tau_{is(k)}$.
-Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ from the referee's method for a given sector-country $i,s(k)$, one can solve the following two-equation system:
-
-\begin{eqnarray}
-p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
-\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
-\end{eqnarray}
-
-\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
-With 2 equations and 2 endogenous variables, the system might be solved.
-Specifically, with $\widehat{\beta}$ being estimated at the \textit{sector} $s$- country $i$ level, and $\widetilde{p}_{ik}$ at the \textit{product} $k$-country level, this implies that $t$ and $\tau$ are product-country $i,k$ specific (not sector-country $s,k$ specific).
-\textbf{from this point of view, this is a finer characterization of transport costs, so better from this point of view...}
-
-To show this properly, start from Equation (\ref{eq:system2}) to express additive costs according to:
-$$t_{ik} = \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
-
-With $\widetilde{p}_{ik} $ being product-country specific, $t_{ik}$ is necessarily product-country specific ($i,k$, not $i,s(k)$). Plugging this into Equation (\ref{eq:system1}):
-
-$$p_{ik} = \tau_{is(k)} \widetilde{p}_{ik} + \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
-
-Solving for $\tau$, one finally obtains the solutions for the ad-valorem and additive transport costs components:
-
-\begin{eqnarray}
-\tau_{ik} & =& (1-\widehat{\beta}_{is(k)}) \frac{p_{ik}}{\widetilde{p}_{ik} } + \widehat{\beta}_{is(k)} \label{eq:tau_method_ref1}\\
-t_ik &=& \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} \label{eq:t_method_ref1}
-\end{eqnarray}
-
-Two points can be made.
-First, we can check that if $\widehat{\beta}_{is(k)}=0$, then $p_{ik} = \tau_{ik} \widetilde{p}_{ik}$ (Equation \ref{eq:tau_method_ref1}) and $t_{ik} = 0$ (Equation \ref{eq:t_method_ref2}), all transport costs are ad-valorem.
-Conversely, if $\widehat{\beta}_{is(k)}=1$, then $p_{ik} = \widetilde{p}_{ik} + t_{ik} $ (Equation \ref{eq:tau_method_ref2}) and $\tau_{ik} = 0$ (Equation \ref{eq:t_method_ref1}).
-Second, and most importantly, the method allows to recover the levels of both transport costs components at the product-country level $i,k$, even though the share of additive costs in total costs is estimated at the sector-country level.
-
-\end{document}
-
-\section{More details on the instrumentation strategy}
-
-\subsection{Specification of the first stage equation}
-
-The question we deal with, is the functional form of the first stage equation, where we instrument the fas price with duties as suggested by Referee 1.
-The idea is that firms might react to changes in duty tax rates which have nothing to do with transport costs changes.
-In this respect, considering the predicted part of the fas price related to tax duty is likely to solve potential endogeneity biases.
-
-Denoting $i$ the origin country, $k$ the product at the 5 digit level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
-
-\begin{equation}
-\widetilde{p}_{ikt} = (1+\tau^d_{is(k)t})^\alpha \left(\bar{p}_{ikt}\right)^\beta \label{eq:link_fas_duty}
-\end{equation}
-
-\noindent with $s(k)$ the 3-digit classification as a function of the 5-digit product classification.
-
-\subsubsection{First-stage equation in first difference \label{ssec:first_diff}}
-
-One option to specify the functional form of the first stage is to take the total differential around some reference point at time $t-1$, with $\Delta $ the difference operator ($\Delta\widetilde{p}_{ikt} = \widetilde{p}_{ikt} - \widetilde{p}_{ikt-1}$ and so on):
-\begin{eqnarray*}
-&&\Delta \widetilde{p}_{ikt} = \beta \bar{p}_{ikt-1}^{\beta-1}(1+\tau^d_{is(k)t-1})\Delta \bar{p}_{ikt} + \alpha \bar{p}^\beta_{ikt-1} (1+\tau^d_{ikt-1})^{\alpha-1}\Delta \tau^d_{is(k)t}  \\
-\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \beta \frac{\Delta \bar{p}_{ikt}}{\widetilde{p}_{ikt-1}} \frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} +\alpha \frac{\Delta \tau^d_{is(k)t-1}}{1+\tau_{is(k)t-1}^d}\frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} \end{eqnarray*}
-
-leading to:
-\begin{equation}
-\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} =  \beta \frac{\Delta \bar{p}_{ikt}}{\bar{p}_{ikt-1}} +\alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{ikt-1}^d} \label{eq:firststage_Deltalog}
-\end{equation}
-
-The intuition behind the Referee 1's endogeneity concerns, is that we need to eliminate from the fas price, any endogenous component that might me related to transport costs.
-To do so, the referee suggests that we instrument the export price by tariff rates.
-Put it in plain words, this suggests to run the first-stage regression based on Equation (\ref{eq:firststage_Deltalog}) according to:
-
-$$\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \alpha \frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \gamma_{i} +\gamma_{k}+\epsilon_{ik}$$
-
-Or equivalently, taking logs:
-$$\Delta \log \widetilde{p}_{ikt}= \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{k}+\epsilon_{ikt}$$
-
-\noindent with the LHS being the growth rate of the fas price (between $t$ and $t-1$), the first term in the RHS the change in duty tax rates, the second and third terms fixed effects to capture changes in the ``firm-specific price'' $\bar{p}_{ik}$, $\epsilon_{ik}$ being the residual.
-Notice though that the structure of fixed effects should remain consistent between the first and the second stages.
-This implies to rather consider the following functional form of the first-stage equation:
-
-\begin{equation}
-\Delta \log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_Deltalog}
-\end{equation}
-
-If this reasoning is correct,
-\begin{itemize}
-\item we should take as predicted value only the predicted price without the fixed effects:
-$$\left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $$
-\item and we should expect a value of the coefficient $\widehat{\alpha}$ between $\Delta \log \widetilde{p}_{ikt}$ and $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $ between $-1$ and $0$, depending on the degree of ``pricing-to-market'' of firms.
-    \begin{itemize}
-    \item[-] $\alpha = 0$ corresponds to the case where the firm does not adjust its fas price to the change in tax duty, that would cancel out the influence of the tax duty change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition;
-    \item[-] $\alpha <0$ corresponds to the degree of``pricing-to-market'' as the firm offsets the impact of the tax change of the price paid by the final consumer by adjusting her producer price in the opposite direction of the tax change,
-    \item[-] with $\alpha = -1$ being the extreme case of ``full PTM'' where the firm fully compensates the tax duty change.
-As shown by Berman, Martin, Mayer (REStats 2012), this is more likely larger firms.
-    \end{itemize}
-\end{itemize}
-
-Once this is done, we can rebuild the instrumented fas price in level (at time $t$):
-
-\begin{eqnarray*}
-\widetilde{p}^{IV}_{ikt} = \widetilde{p}_{ikt-1}\left[1 + \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}\right]
-\end{eqnarray*}
-
-with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}$ the predicted growth rate of the fas price based on duty changes.
-
-\subsection{First-stage equation in level}
-
-Alternatively, one can run the first-stage equation in log level, consistently with the second-stage log-level specification.
-Further, in accordance with the referee's suggestion, this also implies including the lagged price as instrument.
-Starting from Equation (\ref{eq:firststage_Deltalog}), we thus run as first-stage specification:
-
-\begin{equation}
-\log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \beta \log\widetilde{p}_{ikt-1}  +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_log}
-\end{equation}
-
-Consistently with the cross-section analysis adopted by now, the first-stage equation is estimated on a yearly basis (and by transport mode). We then build the instrumented fas price
-
-\begin{eqnarray*}
-&&\log \widetilde{p}_{ikt}^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \widehat{\beta}\log\widetilde{p}_{ikt-1} \\
-\rightarrow &&\widetilde{p}_{ikt}^{IV}  = \exp^{\log \widetilde{p}_{ikt}^{IV}}
-\end{eqnarray*}

### 2021-10-07 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex

@@
 Obviously, the health crisis and the various consecutive lockdowns made the access to our offices difficult in several occurrences.
+\medskip
 
@@
 We can of course make available all our e-mail exchanges with Jeremy Sanchez.
-In his last e-mail, he confirmed that access to the additional data was impossible.
+In his last e-mail, he confirmed that access to the additional data was impossible.\medskip
 
@@
 These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.
+\medskip
 
-and to thank the referees for the comments which have helped us revise it.
-These suggestions have led to substantial improvements.
+To get now more precisely to the revisions undertaken, we thank the referees for their various comments which led to substantial improvements of the paper.
 We enclose detailed answers to the referees.
-Here we briefly list the major changes brought to the paper.
-We hope that you will be satisfied by the outcome.
+Below we first detail how we handled two major points raised by both referees, before listing the other major changes brought to the paper.
+
+\section{Two Common Concerns: Estimation Method and Big Picture}
+Both referees questioned (i) our empirical methodology, especially our choice for a non-linear estimator and (ii) the lack of a ``big picture'' in the paper, i.e.
+of a deeper analysis of the implications of our results.
+Below we detail how we answered their concerns:
 

### 2021-10-07 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
amendement réponse référé 2

@@
 
+We do acknowledge this is very important was not made clear enough in the initial version, and did our best to make the point clear in the revised version (see p.8).
+We also devote more time to the robustness of the empirical specification.
+On top of the robustness to the separability assumption (already in the submitted version), we also assess the robustness to the degree of aggregation and endogeneity (Section 3.4).
+We hope that this new version now convinces you that our estimation strategy is both relevant and robust.
+
 \newpage

### 2021-10-08 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
-\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness to aggregation, robustness to IV estimates).
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and the results, in Section 3, which now includes the robustness checks as final sub-section.
+\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to endogeneity have been added).
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, which presents the results and now includes the robustness checks as final sub-section.
 For sake of space saving, the robustness analysis related to composition effects has been set to the Appendix (Section C.3).
@@
 
-\newpage
+%\newpage
 \bibliographystyle{apalike2}

### 2021-10-08 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement divers

@@
 First, beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
-In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.\footnote{Also, note that the estimation method implemented by Lashkaripour (2020) (see his Equation (3)) is similar to the alternative method suggested in your Critique 1 above.
-Accordingly, the share of multiplicative costs may produce noisier estimates compared to ours, as we point out in our Answer 2.1.} It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
+It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
 

### 2021-10-11 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
-In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
+In this respect, the induced selection bias may reduce the general scope of the transport costs estimates.
+This concern might yet be mitigated by noticing that the covered value of trade flows is only slightly lower.
+If it reduces the sample of sectors and origin countries covered, this method preserves the majority of trade flows in value.
 
@@
 
-Notwithstanding these limitations, it remains interesting to run this alternative estimation method as robustness check.
+Following the referee's advice, we thus run this alternative estimation method as robustness check.
 Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013.
@@
 
-All these elements drive us to maintain our original estimation as benchmark method in the revised version, and we deeply hope to convince you of the relevance of this choice.
+All these elements, in particular Concern 2 regarding the lower accuracy of the estimates, drive us to maintain our original estimation as benchmark method in the revised version.
 However, we also want to keep track of this alternative method as further robustness check.
 For this purpose, we have amended the online Appendix with a section devoted to the referee's method (\textbf{Section D.3}).
+We deeply hope to convince you of the relevance of this choice.
 
@@
 \noindent The referee definitely has a point here, in the sense that Census data do report the quantity of goods.
-In addition, referee's remarks echo recent findings by Lashkaripour (\emph{Journal of International Economics}, 2020): based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
-Finally, Lashkaripour (2020) finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
+At a more fundamental level, this raises the following question: Do we have reason to believe our measure of unit prices generate significant deviations from the one relying on quantities?
+This referee's remark echoes recent findings by \cite{Lashkaripour_JIE2020}: based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
+Finally, \cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
+This result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 30 to 45\% of total costs.
+The two stories are not necessarily incompatible, however.
+Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
+In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
+It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
 
-We devoted a lot of thought to a potential switch from weight to quantity to get cleaner unit prices. However, the latter appears really uneasy in our context for several, non-trivial reasons:
+Nevertheless, the point made by the referee is an important one and deserves careful investigation.
+We devoted a lot of thought to a potential switch from weight to quantity.
+Running the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis yet appears really uneasy in our context for several, non-trivial reasons:
 
 \begin{enumerate}
-\item The information on quantity is not mode-specific, which is incompatible with our empirical strategy of estimation transport costs conditional on the transport mode, similarly as in Hummels (2004).
-We could of course restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
-\textbf{LP: DOCUMENT THAT.
-SHOW IN THE LETTER THAT WE DID IT?}
+\item In the original Hummel's database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
+This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
+We could restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
+Investigating this possibility shows that this would reduce the total value of trade flows from around 32,000 billion dollars, to 7,000 billion dollars, i.e.
+a reduction of flows by a factor of 4.5.
 
-\item Second, units are not available for many years over our period, especially in the 1970s and the 1980s \textbf{XXX GUILLAUME, TO BE CONFIRMED AND SUPPORTED XXX}; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
+\item Even on this very reduced sample, units are not available for many years over our period, especially in the 1970s and the 1980s; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
+\textbf{LP: Je ne suis pas sure de cet argument.
+Veut-on dire que dans les "quantity", en fait on trouve de tout mais pas des unites?
+}
 
@@
 For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
-\textbf{XXX GUILLAUME: Do we have evidence, even anecdotal, on this point?}
+\textbf{XXX GUILLAUME: Do we have evidence, even anecdotal, on this point?
+LP: Sinon, laisser tomber cet argument}
+
+    \item Considering the US census data which complete the original \cites{hummels2007} database, it is true that we can retrieve information on quantities.
+Specifically, this can be obtained crossing HS-10 product information with the correspondence table between the HS-6 product classification and the associated units (kg, tons, liters, numbers, etc.) also provided by the US Census.
+We are able to collect information since 2009.
+Investigating this database, it appears that in many 3-digit sectors, there are more than one type of quantities reported (for instance, sector \textbf{XXX} includes goods both recorded in terms of \textbf{YYY} and \textbf{ZZZ}.) To identify a proper unit price, we consequently reduce the flows to the 3-digit sectors reporting a single quantity.
+    \item As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
+As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
+$t_{is(k)} = t_i+t_{s(k)}$ with $i$ the origin country and $s$ the sector.
+In the specific case of quantities, this raises a non-trivial issue yet.
+Specifically, imposing a country $i$ fixed effect would imply that the cost of sending one unit of say, tee-shirt from France is the same than sending one unit of say, car; and that this cost would be added to the sector-fixed effect.
+This does not seem to us as a reasonable assumption; in this case specifically, skipping the separability assumption does appear as the appropriate choice.
+
+        As a consequence, we run the estimation on a reduced sample (\textbf{specify the number of years, value to trade flows in comparison with the full sample}), according to the following equation:
+
+\begin{equation}
+\ln\left(\frac{p_{ik}}{p^u_{ik}}-1 \right)= \ln \left(\tau_{is(k)} -1+\frac{t_{is(k)} }{p^u_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
+\end{equation}
+\noindent where $p^u_{ik}$ refers to the unit price associated with quantity. Depending on the sector, two different units are considered, kilograms and numbers.
+
+The results are the following ... \textbf{To DEVELOP}
 
@@
 
-These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression.
-That said, do we have reason to believe this measure of unit prices generate significant deviations from the one relying on quantities?
-
-\citet{Lashkaripour_JIE2020} does provide important results related to ours.
-In particular, his result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 30 to 45\% of total costs.
-The two stories are not necessarily incompatible, however.
-First, beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
-In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
-It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
-
-Nevertheless, the point made by the referee is an important one and deserves careful investigation.
-This has driven us to run the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis, for the years for which we have the data on quantity.
-\textbf{LP: INCLURE la regression en quantite EN ROBUSTESSE DU PAPIER - A FAIRE}
+These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression. We hope to convince you that
 
@@
 \end{document}
-
-\newpage
-
-\appendix
-
-% XXX JH: What do we do with the section below?
-
-\section{Clarifying some technical points \label{app:technical_points}}
-
-\subsection{A note on prices}
-
-We have three prices: the (US) consumer price, say $p^{c}_{ik}$, the ``import'' or  cif price (at the entrance of the US) $p_{ik}$ and the fas price (at the export gate in the origin country) $\widetilde{p}_{ik}$, denoting $i$ the origin country, $k$ the product at the 8 or 5 digit level, and reasoning on a yearly basis, and $s(k)$ the 3-digit classification the product $k$ belongs to.
-Further, if we denote $\tau^d_{ik}$ the duty tax rate paid when the good crosses the US border, then we have:
-
-\begin{eqnarray*}
-&&p^c_{ik} = (1+\tau^d_{is(k)})p_{ik} \\
-with && p_{ik}  = \tau_{is(k)} \widetilde{p}_{ik} +t_{is(k)}
-\end{eqnarray*}
-
-\subsection{Deriving the share of additive costs in total transport costs \label{app:interpret_beta}}
-
-With $i$ the origin country, $k$ the product category at the HS6 level (and reasoning at the year- transport mode level), $p_{ik}$ the import (cif) price and $\tilde{p}_{ik}$ the export (fas) price, transport costs $f_{ik}\equiv \frac{p_{ik}- \tilde{p}_{ik}}{\tilde{p}_{ik}} $ are written in Hummel's terminology as:
-
-\begin{equation*}
-f_{ik} = X_{is(k)}\tilde{p}_{ik}^{-\beta_{ik}} \label{eq:Hummels}
-\end{equation*}
-
-It is trivial to show that:
-$$\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} \frac{\tilde{p}_{ik}}{f_{ik}}= -\beta_{ik}$$
-
-Then
-\begin{itemize}
-\item If $\beta_{ik} = 0$, then $p_{ik} = (1+X_{is(k)})\tilde{p}_{ik}$: Only ad-valorem transport costs
-\item If $\beta_{ik} = 1$, then $p_{ik}=\tilde{p}_{ik}+X_{is(k)}$: Only additive costs
-\end{itemize}
-
-This suggests that, the closer $\beta_{ik}$ to 1, the more prevalent additive costs in total transport costs.
-Can $\beta_{ik}$ be related to the share of additive costs in total transport costs?
-The answer is positive.
-To show this clearly, start from:
-
-\begin{eqnarray*}
-f_{ik} &=& \frac{p_{ik}-\tilde{p_{ik}}}{\tilde{p}_{ik}}\\
-\text{with}:&& p_{ik} = \tau_{is(k)}\tilde{p}_{ik} + t_{is(k)} \\
-\end{eqnarray*}
-
-assuming that transport costs decompose in two components, additive transport costs $t_{is(k)}$ and multiplicative costs (ad-valorem) $\tau_{is(k)}$, supposed to be identical in the product-dimension for any $k$ product within the 3-digit classification $s$ it belongs to.
-
-This gives:
-
-$$f_{ik} = \frac{(\tau_{is(k)}-1)\tilde{p}_{ik}+t_{is(k)}}{\tilde{p}_{ik}}$$
-
-From this, we have
-
-\begin{eqnarray*}
-\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} &=& \frac{(\tau_{is(k)}-1)\tilde{p}_{ik} - ((\tau_{is(k)}-1)\tilde{p}_{ik} +t_{is(k)})}{\tilde{p}_{ik}^2} \\
-&=& \frac{-t_{is(k)}}{\tilde{p}_{ik}^2}
-\end{eqnarray*}
-
-Hence
-\begin{eqnarray*}
-\frac{\partial f_{ik}}{\partial \tilde{p}_{ik}} \frac{\tilde{p}_{ik}}{f_{ik}}&=& \frac{-t_{is(k)}}{(\tau_{is(k)}-1)\tilde{p}_{ik} +t_{is(k)}}\\
-&=& \frac{-\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 +\frac{t_{ik}}{\tilde{p}_{ik}}}
-\end{eqnarray*}
-
-This shows that $\beta_{ik}$ can be interpreted as the share of additive costs in total transport costs:
-
-$$\beta_{ik} = \frac{\frac{t_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
-
-\subsection{Share vs level}
-
-Starting from the above reasoning, the referee proposes to estimate the elasticity of transport costs with respect to the unit price, equivalently the share of additive costs in total costs.
-Specifically, the referee suggests to estimate for each industry-country the following regression:
-
-$$\ln f_{ikd}= \beta_{is(k)} \ln \widetilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd}$$
-
-The question is then, how to recover the levels of additive / multiplicative transports costs $t_{is(k)}$ and $\tau_{is(k)}$.
-Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ from the referee's method for a given sector-country $i,s(k)$, one can solve the following two-equation system:
-
-\begin{eqnarray}
-p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
-\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
-\end{eqnarray}
-
-\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
-With 2 equations and 2 endogenous variables, the system might be solved.
-Specifically, with $\widehat{\beta}$ being estimated at the \textit{sector} $s$- country $i$ level, and $\widetilde{p}_{ik}$ at the \textit{product} $k$-country level, this implies that $t$ and $\tau$ are product-country $i,k$ specific (not sector-country $s,k$ specific).
-\textbf{from this point of view, this is a finer characterization of transport costs, so better from this point of view...}
-
-To show this properly, start from Equation (\ref{eq:system2}) to express additive costs according to:
-$$t_{ik} = \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
-
-With $\widetilde{p}_{ik} $ being product-country specific, $t_{ik}$ is necessarily product-country specific ($i,k$, not $i,s(k)$). Plugging this into Equation (\ref{eq:system1}):
-
-$$p_{ik} = \tau_{is(k)} \widetilde{p}_{ik} + \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} $$
-
-Solving for $\tau$, one finally obtains the solutions for the ad-valorem and additive transport costs components:
-
-\begin{eqnarray}
-\tau_{ik} & =& (1-\widehat{\beta}_{is(k)}) \frac{p_{ik}}{\widetilde{p}_{ik} } + \widehat{\beta}_{is(k)} \label{eq:tau_method_ref1}\\
-t_ik &=& \frac{ \widehat{\beta}_{is(k)}}{1- \widehat{\beta}_{is(k)}} \left[\tau_{is(k)}-1\right]\widetilde{p}_{ik} \label{eq:t_method_ref1}
-\end{eqnarray}
-
-Two points can be made.
-First, we can check that if $\widehat{\beta}_{is(k)}=0$, then $p_{ik} = \tau_{ik} \widetilde{p}_{ik}$ (Equation \ref{eq:tau_method_ref1}) and $t_{ik} = 0$ (Equation \ref{eq:t_method_ref2}), all transport costs are ad-valorem.
-Conversely, if $\widehat{\beta}_{is(k)}=1$, then $p_{ik} = \widetilde{p}_{ik} + t_{ik} $ (Equation \ref{eq:tau_method_ref2}) and $\tau_{ik} = 0$ (Equation \ref{eq:t_method_ref1}).
-Second, and most importantly, the method allows to recover the levels of both transport costs components at the product-country level $i,k$, even though the share of additive costs in total costs is estimated at the sector-country level.
-
-\newpage
-\bibliographystyle{apalike2}
-\bibliography{biblio}
-
-\end{document}
-
-\section{More details on the instrumentation strategy}
-
-\subsection{Specification of the first stage equation}
-
-The question we deal with, is the functional form of the first stage equation, where we instrument the fas price with duties as suggested by Referee 1.
-The idea is that firms might react to changes in duty tax rates which have nothing to do with transport costs changes.
-In this respect, considering the predicted part of the fas price related to tax duty is likely to solve potential endogeneity biases.
-
-Denoting $i$ the origin country, $k$ the product at the 5 digit level, and reasoning on a yearly basis, if we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
-
-\begin{equation}
-\widetilde{p}_{ikt} = (1+\tau^d_{is(k)t})^\alpha \left(\bar{p}_{ikt}\right)^\beta \label{eq:link_fas_duty}
-\end{equation}
-
-\noindent with $s(k)$ the 3-digit classification as a function of the 5-digit product classification.
-
-\subsubsection{First-stage equation in first difference \label{ssec:first_diff}}
-
-One option to specify the functional form of the first stage is to take the total differential around some reference point at time $t-1$, with $\Delta $ the difference operator ($\Delta\widetilde{p}_{ikt} = \widetilde{p}_{ikt} - \widetilde{p}_{ikt-1}$ and so on):
-\begin{eqnarray*}
-&&\Delta \widetilde{p}_{ikt} = \beta \bar{p}_{ikt-1}^{\beta-1}(1+\tau^d_{is(k)t-1})\Delta \bar{p}_{ikt} + \alpha \bar{p}^\beta_{ikt-1} (1+\tau^d_{ikt-1})^{\alpha-1}\Delta \tau^d_{is(k)t}  \\
-\Leftrightarrow &&\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \beta \frac{\Delta \bar{p}_{ikt}}{\widetilde{p}_{ikt-1}} \frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} +\alpha \frac{\Delta \tau^d_{is(k)t-1}}{1+\tau_{is(k)t-1}^d}\frac{(1+\tau^d_{is(k)t-1})^\alpha \bar{p}_{ikt-1}^\beta}{\bar{p}^\beta_{ikt-1}(1+\tau^d_{is(k)t-1})^\alpha} \end{eqnarray*}
-
-leading to:
-\begin{equation}
-\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} =  \beta \frac{\Delta \bar{p}_{ikt}}{\bar{p}_{ikt-1}} +\alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{ikt-1}^d} \label{eq:firststage_Deltalog}
-\end{equation}
-
-The intuition behind the Referee 1's endogeneity concerns, is that we need to eliminate from the fas price, any endogenous component that might me related to transport costs.
-To do so, the referee suggests that we instrument the export price by tariff rates.
-Put it in plain words, this suggests to run the first-stage regression based on Equation (\ref{eq:firststage_Deltalog}) according to:
-
-$$\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}} = \alpha \frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \gamma_{i} +\gamma_{k}+\epsilon_{ik}$$
-
-Or equivalently, taking logs:
-$$\Delta \log \widetilde{p}_{ikt}= \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{k}+\epsilon_{ikt}$$
-
-\noindent with the LHS being the growth rate of the fas price (between $t$ and $t-1$), the first term in the RHS the change in duty tax rates, the second and third terms fixed effects to capture changes in the ``firm-specific price'' $\bar{p}_{ik}$, $\epsilon_{ik}$ being the residual.
-Notice though that the structure of fixed effects should remain consistent between the first and the second stages.
-This implies to rather consider the following functional form of the first-stage equation:
-
-\begin{equation}
-\Delta \log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_Deltalog}
-\end{equation}
-
-If this reasoning is correct,
-\begin{itemize}
-\item we should take as predicted value only the predicted price without the fixed effects:
-$$\left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $$
-\item and we should expect a value of the coefficient $\widehat{\alpha}$ between $\Delta \log \widetilde{p}_{ikt}$ and $\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} $ between $-1$ and $0$, depending on the degree of ``pricing-to-market'' of firms.
-    \begin{itemize}
-    \item[-] $\alpha = 0$ corresponds to the case where the firm does not adjust its fas price to the change in tax duty, that would cancel out the influence of the tax duty change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition;
-    \item[-] $\alpha <0$ corresponds to the degree of``pricing-to-market'' as the firm offsets the impact of the tax change of the price paid by the final consumer by adjusting her producer price in the opposite direction of the tax change,
-    \item[-] with $\alpha = -1$ being the extreme case of ``full PTM'' where the firm fully compensates the tax duty change.
-As shown by Berman, Martin, Mayer (REStats 2012), this is more likely larger firms.
-    \end{itemize}
-\end{itemize}
-
-Once this is done, we can rebuild the instrumented fas price in level (at time $t$):
-
-\begin{eqnarray*}
-\widetilde{p}^{IV}_{ikt} = \widetilde{p}_{ikt-1}\left[1 + \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}\right]
-\end{eqnarray*}
-
-with $\widehat{\widetilde{p}}_{ikt}$ the instrumented fas price at time $t$, for product $k$ from country $i$; $\widetilde{p}_{ikt-1}$ the observed lagged fas price for the same $i,k$ couple; and $ \left(\frac{\Delta \widetilde{p}_{ikt}}{\widetilde{p}_{ikt-1}}\right)^{IV}$ the predicted growth rate of the fas price based on duty changes.
-
-\subsection{First-stage equation in level}
-
-Alternatively, one can run the first-stage equation in log level, consistently with the second-stage log-level specification.
-Further, in accordance with the referee's suggestion, this also implies including the lagged price as instrument.
-Starting from Equation (\ref{eq:firststage_Deltalog}), we thus run as first-stage specification:
-
-\begin{equation}
-\log \widetilde{p}_{ikt} = \alpha\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \beta \log\widetilde{p}_{ikt-1}  +\gamma_{i} +\gamma_{s}+\epsilon_{ikt} \label{eq:firststage_log}
-\end{equation}
-
-Consistently with the cross-section analysis adopted by now, the first-stage equation is estimated on a yearly basis (and by transport mode). We then build the instrumented fas price
-
-\begin{eqnarray*}
-&&\log \widetilde{p}_{ikt}^{IV} = \widehat{\alpha}\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d} + \widehat{\beta}\log\widetilde{p}_{ikt-1} \\
-\rightarrow &&\widetilde{p}_{ikt}^{IV}  = \exp^{\log \widetilde{p}_{ikt}^{IV}}
-\end{eqnarray*}

### 2021-10-11 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
Avancees lettre editeur + minor stuff answers

@@
 This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
-$\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of trade costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
+$\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of transport costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
 Without these restrictions, standard linear, least squares estimates deliver aberrant values with more than mild quality-of-fit.
 Taking into account this constraint implies that the error term should be always positive, which we ensure by specifying the error term such that:
+
 \begin{equation*}
@@
 
-We do acknowledge this is very important was not made clear enough in the initial version, and did our best to make the point clear in the revised version (see p.8).
+This specification obviously requires a non-linear estimation.
+We did our best to make the point clear in the revised version (see p.8).
 We also devote more time to the robustness of the empirical specification.

### 2021-10-11 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex
petites modfis
Avancees lettre editeur + minor stuff answers

@@
 of a deeper analysis of the implications of our results.
-Below we detail how we answered their concerns:
+Below we detail how we answered their concerns:\medskip
+
+\subsection{Estimation Method}
+Both referees debated our choice for a non-linear estimator, suggesting alternative functional forms estimable with standard linear estimators.
+We do acknowledge that the initial version of the paper did not make a clear enough case for the non-linear estimator.
+We now explain in details, both in the papers and in our answers to reviewers, the reasons why a non-linear estimator is unavoidable in our case.
+This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
+$\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of transport costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
+Without these restrictions, standard linear, least squares estimates deliver aberrant values with more than mild quality-of-fit.
+Taking into account this constraint necessarily brings a functional requiring a non-linear estimation.
+We did our best to make the point clear in the revised version (see p.8).
+\medskip
+
+That said, resorting to non-linear estimations requires some assumptions due to the computational burden.
+The revised version devotes much more space to explicit tests of robustness of the empirical specification.
+On top of the robustness to the separability assumption (already in the initially submitted version), we also assess the robustness to the degree of data aggregation and endogeneity (Section 3.3) - see the list of modifications in section 2 from this letter.
+\medskip
+
+In this regard, we also point in our answers that implementing the methodology suggested by Referee 1 would imply non-negligible costs.
+Firstly, this would leave us much less data to exploit, partly  because of the already mentioned CoViD issues - we would lose half of the years in the sample.
+Secondly, we show in details in our response that our method yields a more precise estimation of transport costs than the referee's method.
+\medskip
+
+\subsection{Big Picture}
+Both referees asked us to shed light on the implications of additive costs (the ``Big Picture''). The revised version answers this concern in two points. \medskip
+
+Firstly, we maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we made our best to make this section easier to read, leaving the technical aspects in the Appendix (Section C).
+As such, we hope that the main message of this Section is better catched.
+\medskip
+
+Secondly, and more importantly, we now emphasize the ``Big Picture'' implications of additive costs on theoretical grounds.
+Both referees suggested various, interesting paths to explore.
+Obviously, all of them could not be handled in a single paper.
+Therefore, we decide to focus on the welfare implications of our results.
+Through the lens of the Melitz's (2003) model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated (in Section 3 of the revised version), to quantify the extra welfare gains attributable to the reduction of the additive component.
+This analysis is driven in a new Section 5.
+We emphasize that the theoretical implications of additive costs have not been fully exhausted in Irarrazabal et al.
+(2015), Hornok and Koren (2015a,b), Alessandria et al.
+(2010), or Sorensen (2014).
+Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantied by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
+To the best of our knowledge, this is the first time that such an assessment is preformed based on long-time, detailed trade data.
+\medskip
+
+Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
+
+\section{List of major revisions in the new version.}
+
+Below we list the main other revisions implemented to answer referees' concerns:
 
@@
 
-\item Functional form/use of a non-linear estimator
-\item Endogeneity issues/IV estimates
-\item Aggregation issues
-\item Calculation of unit prices
-\item Big picture/usefulness of the exercise
-\item XXX
-\item XXX
+\item \emph{Endogeneity issues}: Section 3.3.2 and Appendix B.3 provides detailed developments answering concerns expressed by Referee 1 concerning potential endogeneity in our estimates.
+Relying on instrumental variables suggested by Referee 1, we provide first- and second-stage estimates, based on both at the 5-digit-product level and the more disaggregated HS10 level for sensitivity checks.
+Figure 3 on page XX reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+In all cases, these estimates are very similar, when not identical.
+This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
+
+\item \emph{Aggregation issues}: Referee 1 pointed `` \emph{The original annual Census data reports trade at the origin country-HS10 product-district level of aggregation, whereas the authors are aggregating up the data even further to the origin country-HS5 industry-year level.
+Such an aggregation comes with strong implicit assumptions and sacrifices a lot of useful variation in the data}.'' Firstly, note these data are available only from 1989, whereas our full sample starts in 1974.
+Secondly, as already mentioned at the beginning of this letter, it was only possible to get access to the years 1997-1999 and 2001-2019, due to the COVID situation.
+Considering the importance of the time length for our empirical exercise, especially in the context of the exercise implemented in Section 5, we decided to stick to our initial 5-digit database, and to provide systematic robustness estimates based on the HS10 data available.
+These results are now reported in section 3.3.3 in the revised version.
+In particular, Table 3 and Figure 4 show the results are similar when using 10-digit product sand 5-digit products \textbf{XXX TO BE CONFIRMED XXX}
+
+\item \emph{Calculation of unit prices}: Referee 1 is concerned by the fact that we calculate unit price as Value/Weight, instead of Value/Quantity.
+We explain in details in our answer why a potential switch from weight to quantity for calculating unit prices would be harmful in our context: until 2004, information on quantity \underline{by transport mode} is unavailable XXX + other arguments XXX Therefore, we decided to provide robustness estimates on a reduced sample (\textbf{specify the number of years, value to trade flows in comparison with the full sample}) \textbf{XXX TO BE COMPLETED XXX}
 
@@
 Altogether, we believe that the paper is much clearer and easier to read.
-The robustness exercises included in the main text focus on the elements that were the most problematic to the referees (XXX).
+The robustness exercises included in the main text focus on the elements that were the most problematic to the referees.
 We hope that you will be pleased by the outcome.

### 2021-10-13 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
modifs_answer_R2

@@
 
-\paragraph{``Big Picture''} \textit{[...] The main contribution of this paper thus appears to be to unravel the temporal evolution / relative importance of the ad-valorem and specific transport cost component.
+\paragraph{``Big Picture''} \textit{[...] The fact that transport costs are both specific and ad-valorem is well known.
+Also, the theoretical implications of having specific, instead of ad-valorem (iceberg), transport cost when it comes to analyzing the consequences of further reductions in transport costs are by now quite well understood; see e.g.
+papers by Irarrazabal et al.
+(2015), Hornok and Koren (2015a,b), Alessandria et al.
+(2010), or Sorensen (2014).
+The main contribution of this paper thus appears to be to unravel the temporal evolution / relative importance of the ad-valorem and specific transport cost component.
 However, besides showing this evolution, the paper is basically silent on the implication(s) of the patterns found.
@@
 This was quite disappointing, simply documenting the evolution of these two components leaves the reader (at least me) wondering why he/she should care about this.
-The authors should aim to be much more clear on what the trends in ad-valorem / specific transport costs mean/imply from the perspective of these recent theoretical contributions/insights.} \smallskip
+The authors should aim to be much more clear on what the trends in ad-valorem / specific transport costs mean/imply from the perspective of these recent theoretical contributions/insights.}
+\smallskip
+\paragraph{Exploring the heterogeneity of our results} \textit{At the very least the authors could/should make use of the fact that the temporal patterns in specific / ad-valorem transport costs is likely to show substantial heterogeneity across goods as well as countries.
+Exploring this heterogeneity to a fuller extent could already hint at the main drivers (insurance costs / containerization / economies of scale) behind the observed patterns.
+This could be done by e.g.
+not only reporting the mean and median of the ad-valorem / specific transport cost components but exploring more fully the heterogeneity in these transport costs components across sectors/exporters.
+Also, it should be relatively straightforward to provide some analysis of the major correlates with the identified specific and ad-valorem costs (e.g.
+product value/weight ratio, distance to the exporter, etc.).}
+\smallskip
 
 \textbf{Our answer:} We devoted a lot of thought to this point, which directly echoed a similar concern by the other referee.
-Taking stock of your suggestion, we decided to offer insights on the welfare implications of our results.
-In this regard, the new section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
-To do so, we start from the canonical \cites{melitz} model where we add additive trade costs.
+The latter writes: ``\emph{Does this observation revise our understanding of say the gains from trade?
+Does it shed new light on a puzzle many people are thinking about?
+One crude suggestion is to see how the reduction in the industry-specific cost terms is related to the industry-level trade elasticities.
+If the composition effects favor low-elasticity industries, the findings in the paper may have first-order implications for the gains from trade.
+Another suggestion is to dig deeper into the relative rate at which additive and multiplicative transport costs have declined over time.
+Since additive transport costs favor rich (high-quality exporting) countries, the disproportionally greater reduction in additive costs can perhaps explain the rise of low-income exporter as documented by Hanson (2012, JEP).}''
+
+All in all, your suggestions and those of the other referees pointed in two directions for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the implications regarding grains from trade.
+Both paths were undeniably worth exploring, but could not be followed in a single paper.
+Therefore, we decided to focus on the second path, i.e.
+to  offer insights on the welfare implications of our results.
+In this regard, the new Section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
+
+Our choice to dig into the welfare implications of our results has two main motivations.
+Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} recently finds (though in a different theoretical setting) support for the iceberg trade cost assumption, i.e, iceberg accounting for almost all trade costs.
+Secondly, the theoretical implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, Hornok and Koren (2015a,b), Alessandria et al.
+(2010), or \citet{sorensen2014}.
+More specifically, we start from the canonical \cites{melitz} model where we add additive trade costs.
 The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
@@
 In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on ``true'' values for the latter - remember that our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the (median) export price.
+\textbf{XXX COMPLETE WITH OUR CONTRIBUTIONS WRT TO Hornok and Koren (2015a,b), Alessandria et al.
+(2010)XXX}
 
@@
 
-\paragraph{Exploring the heterogeneity of our results} \textit{At the very least the authors could/should make use of the fact that the temporal patterns in specific / ad-valorem transport costs is likely to show substantial heterogeneity across goods as well as countries.
-Exploring this heterogeneity to a fuller extent could already hint at the main drivers (insurance costs / containerization / economies of scale) behind the observed patterns.
-This could be done by e.g.
-not only reporting the mean and median of the ad-valorem / specific transport cost components but exploring more fully the heterogeneity in these transport costs components across sectors/exporters.
-Also, it should be relatively straightforward to provide some analysis of the major correlates with the identified specific and ad-valorem costs (e.g.
-product value/weight ratio, distance to the exporter, etc.).}
-
-\textbf{WHAT DO WE DO ABOUT THAT?
+%\textbf{WHAT DO WE DO ABOUT THAT?
 We provide an analysis by sector about the time trends.

### 2021-10-14 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Modifs intro pour Hornok et al

@@
 This is consistent with the empirical findings by \cite{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
-Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} However, these papers yet do not allow for a quantitative measure of additive costs (neither for the ad-valorem costs).
-Another important contribution of this paper is to provide quantitative estimates of both type of transport costs.
-\smallskip
-
-Closely related to our paper is the work by \cite{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
+Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
+
+Our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-RES-2015, Hornok-et-al-JIE-2015}.
+While \cite{Alessandria-et-al-AER-2010}  and \cite{Hornok-et-al-RES-2015} point out the role of per-shipment costs in generating some ``lumpiness'' in international trade transactions, \cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment (i.e., additive by nature), and draw welfare conclusions regarding the variations of these costs.
+Also closely related to our paper is the work by \cite{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
 Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
-Our results share in common with \cite{Irrazabal_2015} the important role of the additive component of international trade costs.
-Yet, our paper complements their findings in two main respects.
-First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to estimate each value of the ad-valorem and the additive costs independently from the other.
+Our results share in common with \cite{Irrazabal_2015} the important role of the additive component of international trade costs, and a welfare analysis of the latter.
+Compared to
+
+Our paper complements these various findings in different dimensions.
+First, we provide a careful, long-run characterization of both additive and multiplicative component of transports costs.
+Second, we make use of the latter to calibrate a \cites{melitz} type model of trade, which allows us to infer the welfare gains induced by the decrease of both components between 1974 and 2019.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
+It is true that, compared to \cite{Irrazabal_2015}, our study covers a narrower set of trade costs focused on transport costs.
+However, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost.
+By contrast, our estimation strategy enables us to estimate each value of the ad-valorem and the additive costs independently from the other.
 From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
@@
 This comparison of results suggests that between a quarter and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
-%JH : these results should not change, they are for 2004.
-Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from 1974 to 2019.
-In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
-In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed over the last fifty years.
+
+\textbf{JH : I WOULD DELETE EVERYTHING FROM `` It is true that"}
+
 \smallskip
 
-%\textbf{Big Picture}.
+% nous sommes les premiers à fournir une analyse en termes de welfare basée sur une estimation des "vrais" coûts multiplicatifs et additifs sur une
+
+%Yet, our paper complements their findings in two main respects.
+%In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed over the last fifty years.
+%However, these papers yet do not allow for a quantitative measure of additive costs (neither for the ad-valorem costs).
+Another important contribution of this paper is to provide quantitative estimates of both type of transport costs.
+\smallskip
 Our results are robust to a range of robustness checks, including the use of instrumental variables and more disaggregated data at the SITC 4 or HS10 levels.

### 2021-10-14 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
modifs supplementaires

@@
 Our results share in common with \cite{Irrazabal_2015} the important role of the additive component of international trade costs, and a welfare analysis of the latter.
-Compared to
 
@@
 In this section, we explore the implications of our empirical results on theoretical grounds.
-While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
+Over the last decade, several papers examined the implications of additive/per-shipment costs.
+\cite{Kropf-Saure-JIE-2016} estimate the size and shape of per-shipment costs based on Swiss data.
+\cite{Alessandria-et-al-AER-2010} and \cite{Hornok-et-al-RES-2015} point out the role of per-shipment costs in generating ``lumpiness'' in international trade transactions.
+\cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment (i.e., additive by nature), and draw welfare conclusions regarding the variations of these costs.
+We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transports costs.
+Second, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
+To this aim, we embed additive costs in the canonical \citet{melitz} model.\smallskip
+%While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
 One potential explanation is the lack of empirical estimates about the size of additive costs.
 Our results may thus usefully palliate this limitation.
-Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive
-components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
-To this aim, we embed additive costs in the canonical \citet{melitz} model.\smallskip
 
@@
 
-That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020} paper relies on a negligible share of additive costs.
+That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020}paper relies on a negligible share of additive costs.
 In this context, one might expect that whether additive costs are present or not makes little difference.

### 2021-10-15 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
Modifs réponse referee 2

@@
 Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} recently finds (though in a different theoretical setting) support for the iceberg trade cost assumption, i.e, iceberg accounting for almost all trade costs.
-Secondly, the theoretical implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, Hornok and Koren (2015a,b), Alessandria et al.
-(2010), or \citet{sorensen2014}.
+Secondly, the implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, \cite{Hornok-et-al-RES-2015, Hornok-et-al-JIE-2015}, \cite{Alessandria-et-al-AER-2010} or \citet{sorensen2014}.
+\cite{Kropf-Saure-JIE-2016} estimate the size and shape of per-shipment costs based on Swiss data.
+\cite{Alessandria-et-al-AER-2010} and \cite{Hornok-et-al-RES-2015} point out the role of per-shipment costs in generating ``lumpiness'' in international trade transactions.
+\cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment (i.e., additive by nature), and draw welfare conclusions regarding the variations of these costs.
+We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transports costs.
+Furthermore we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
+
+To this aim, we embed additive costs in the canonical \citet{melitz} model.
 More specifically, we start from the canonical \cites{melitz} model where we add additive trade costs.
@@
 In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on ``true'' values for the latter - remember that our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the (median) export price.
-\textbf{XXX COMPLETE WITH OUR CONTRIBUTIONS WRT TO Hornok and Koren (2015a,b), Alessandria et al.
-(2010)XXX}
 

### 2021-10-17 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Nouveau graphique et tableau comparaison 10/3
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 	\centering
-	\caption{Transport costs estimates: Robustness to aggregation} \vspace{5mm} \label{tab:baseline10}
+	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
+	\input{Table_baseline10.tex}
 	\input{Table_baseline10_reshaped.tex}
@@
 \begin{figure}[htbp]
-	\caption{Transport cost time trends: Robustness to aggregation}
+	\caption{Transport cost time trends: Robustness to aggregation (2005-2019)}
 	\label{fig:robustness_to_agregation}
 	\begin{center}
-		\includegraphics[height=8cm]{scatter_chronology_baseline_baseline10.png}
+		\includegraphics[height=8cm]{scatter_chronology_baseline10_dbsamesample10_5_3.png}
 	\end{center}

### 2021-10-18 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Nouveau graph de comparaison
non_séparé_qy et non_séparé_wgt

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Je change la définition du time trend
... de beta pour les équations de l’appendix

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 \begin{table}[htbp]
-	\caption{Comparison 2005-2019}
+	\caption{Comparison 2005-2019, price per qy or price per kg}
 	\begin{center}		
@@
 \end{table}%
+
+\begin{figure}[htbp]
+	\caption{Comparison 2005-2019, price per qy or price per kg}
+	\begin{center}
+	\includegraphics[height=8cm]
+	{../scatter_chronology_non_séparé_wgt_non_séparé_qy.png}
+	\end{center}
+\end{figure}
+
+J’AI CHANGÉ LA DÉFINITION DU TIME TREND. JE FAIS MAINTENANT LE TREND ENTRE LES BETA MOYENS PAR ANNÉES AU LIEU D’UNE REGRESSIONS SUR TOUS LES BETAS PONDÉRÉ PAR LEUR PART DANS LE COMMERCE. GD.
 

### 2021-10-22 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Avancees lettre editeur + minor stuff answers
Travail sur séparabilité answer referee1 et papie
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 When necessary, we refer to number of sections, pages and equations from the submitted version.
-In this case, they are written into brackets.
+In this case, they are written between brackets.
 
@@
 \item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to endogeneity have been added).
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, which presents the results and now includes the robustness checks as final sub-section.
-For sake of space saving, the robustness analysis related to composition effects has been set to the Appendix (Section C.3).
-\item Both referees asked us to put light on the implications of additive costs (the ``Big Picture'').
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, which presents the results and now includes the robustness checks as a final sub-section.
+For sake of brevity, the robustness analysis related to composition effects has been sent to the Appendix (Section C.3).
+\item Both referees asked us to highlight the implications of additive costs (the ``Big Picture'').
 We agree that this was not enough emphasized in the submitted version, and we thank the referees for their suggestions on that issue.
@@
     \begin{enumerate}
-    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we made our best to make this section easier to read, leaving the technical aspects in the Appendix (Section C).
-As such, we hope that the main message of this Section is better catched.
-    \item Most importantly, we now emphasize the ``Big Picture'' implications of additive costs on theoretical grounds.
-Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated (in Section 3 of the revised version), to quantify the extra welfare gains attributable to the reduction of the additive component.
-This analysis is driven in a new Section 5.
+    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section’s readibility by leaving the technical aspects in the Appendix (Section C).
+As such, we hope that the main message of this Section is easier to understand.
+    \item Most importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds.
+Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated in Section 3, to quantify the extra welfare gains attributable to the reduction of their additive component.
+This analysis is conducted in a new Section 5.
         \end{enumerate}
-\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables in the Online Appendix, to which the reader can refer to.
+\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
 \end{itemize}
@@
 
-\textit{Using the notation of the authors, they are intreated
+\textit{Using the notation of the authors, they are interested
 in identifying the share of the specific cost in the total transport cost. Namely,}
@@
 
-The three points raised by the referee indeed deserve careful consideration. We provide separate answers to each of them below.
+These three points indeed deserve careful consideration. We provide separate answers to each of them below.
 
@@
 
-Further, we provide a robustness check for this separative assumption, in Section 4.1 of the paper (\textbf{Table 2 in both the submitted and revised versions XXX CHECK AT THE END XXX}).
+Further, we provide a robustness check for this separability assumption that $\tau_{ik} = \tau_i\tau_{k}$ and
+ $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (Table 2) \textbf{XXX CHECK AT THE END XXX}).
 We check the robustness of our results by re-running the estimation without the separability assumption.
-This comes at the cost of a substantial increase in the number of fixed effects to include in the regression.
-Based on the numbers of countries and sectors reported in \textbf{Table 1 of the submitted and revised versions XXX CHECK AT THE END XXX} (188 countries, 230 3-digit sectors), it would mean including 86,480 fixed effects ($188\times230\times2$ for the two additive and multiplicative costs) rather than ``only'' 836 ($(188+230)\times 2$).
+This comes at the cost of a substantial increase in the number of fixed in the regression.
+In 2019, for example, there are 14,016 country*sector pairs.
+Adding 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$).
 This is computationally not tractable.
+% use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/data/hummels_tra.dta", clear
+% keep if year==2019
+% keep if mode=="ves"
+% gen sector=substr(sitc2,1,3)
+% gen blouk = iso_o+sector
+% codebook blouk
  For this reason, we have decided to run the robustness check on a reduced sample.
-If it is smaller in terms of observations (2,381 for Air, 2,798 for Vessel on average over the period 1974-2019 \textbf{XXX CHECK NUMBERS XXX}, vs around 29,000 on the complete sample, see Table 1), it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
-Given that the reduced sample makes for 80\% of the total value of trade flows, this suggests that the vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
-\textbf{LP: CHECK THE NUMBERS}} As we conclude at the end of Section 4.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not - see in particular Figures 6 and 7 on pp.
+ We select, for each year, the largest importers that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
+ We keep in the sample all trade observations from these importers and in these sectors.
+ This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1), yet it remains quite large in terms of trade coverage.\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
+The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
+\textbf{LP: CHECK THE NUMBERS}}.
+ As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
+ %- see in particular Figures 6 and 7 on pp.
 25-26 \textbf{XXX CHECK AT THE END XXX}.
@@
 
-These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression. We hope to convince you that
+These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression.
+We hope to convince you that \textbf{XXX TO BE COMPLETED XXX}
 

### 2021-10-22 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Travail sur séparabilité answer referee1 et papie
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
-\textbf{LP: Pb avec le nb d'observations.
-}
+%\textbf{LP: Pb avec le nb d'observations. } GD : CELA A L’AIR BON ET ON LAISSE COMME CELA
 %
@@
 			\input{Table4_reshaped.tex}
+			\input{Table4.tex}
 	\end{center}
-	\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+	\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule:  we select, for each year, the largest importers that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
+We keep in the sample all trade observations from these importers and in these sectors.
 			The additive term is expressed in fraction of fas price.}}
@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but 80\% of the total value of aggregated trade) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade)
+\textbf{Ce n’est pas 80\% du commerce : voir le mode de calcul.
+Je sais que vous allez me demander combien c’est et je n’ai pas de réponse toute faite.
+Je peux passer une heure ou deux à le calculer si vous voulez -- GD}] involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."

### 2021-10-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Travail sur l’hypothèse de séparabilitb
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
  We keep in the sample all trade observations from these importers and in these sectors.
- This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1), yet it remains quite large in terms of trade coverage.\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
+ This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1), yet it remains quite large in terms of trade coverage (mean of 68\%).
+% use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
+% gen cover = cover_non_séparé/cover_baseline
+% sum cover
+ \footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
 The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.

### 2021-10-23 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Travail sur l’hypothèse de séparabilitb
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade)
-\textbf{Ce n’est pas 80\% du commerce : voir le mode de calcul.
-Je sais que vous allez me demander combien c’est et je n’ai pas de réponse toute faite.
-Je peux passer une heure ou deux à le calculer si vous voulez -- GD}] involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade (an annual mean of 68\% )
+%use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
+%gen cover = cover_non_séparé/cover_baseline
+%sum coveruse "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
+%gen cover = cover_non_séparé/cover_baseline
+%sum cover
+involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."

### 2021-10-25 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
modification partie agrégation

@@
 Current data are available at the 10-digit level for products (HS10) for some specific years.
-The results reported here rely on exploiting data over the years 1997-1999 and 2001-2019.\footnote{Theoretically, they should be available from 1989, which is already smaller than our full sample starting in 1974.
-Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} \textbf{REPRENDRE CES RESULTATS PB SUR VESSEL LP.
-}
+The results reported here rely on exploiting data from 2005 to 2019.
+This period is obviously smaller than our main sample, due to two main issues.
+Firstly, we could only access the years 1997-1999 and 2001-2019.\footnote{Theoretically, they should be available from 1989, which is already smaller than our full sample starting in 1974.
+Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} Secondly, for the years 1997-1999 and 2001-2005, compatibility issues between Census and Hummels' data prevented reliable comparisons.
 
@@
 
-Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are similar when using 10-digit product and 5-digit products.
+Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are very similar when using 10-digit product and 5-digit products.
+If anything, the $\beta$ seems on average slightly higher when estimated based on a 3 (sector)-10(product) database, indicating a (small) downward bias in our estimates of additive costs.
+But overall, no major difference arises both regarding levels and trends - results are especially very close for the Air transportation mode.
+This has important implications for the exercise performed in section \ref{sec:theory}: welfare gains computations should not be significantly affected by our use of 5 (sector)-5 (product) digit data.
+This is consistent with e.g.
+\citet{Giri_et_al2021}, who show that sectoral heterogeneity does not always lead to an increase in the gains from trade.
 

### 2021-10-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
modifs textes

@@
 			\input{Table4_reshaped.tex}
-			\input{Table4.tex}
+%			\input{Table4.tex}
 	\end{center}
@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade (an annual mean of 68\% )
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade (an annual mean of 68\% XXX D'OU VIENT CE CHIFFRE DE 68\% ??
+N'est-ce pas 80\% ???).
 %use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
@@
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the results on a detailed yearly basis and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triple $(i,k,t)$, which confirm the robustness of the results.
-These results are reported in the Online Appendix, \textbf{Section XXX} for sake of space saving.} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the results on a detailed yearly basis and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triplet $(i,k,t)$, which confirm the robustness of the results.
+These results are reported in the Online Appendix, respectively in sections C and D.2.} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
 The picture is reversed for medians

### 2021-10-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
autres petites modfis

@@
 We yet provide a robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.
-\textbf{See Section XXX JH: COULD NOT FIND THESE ANYWHERE.
-HAS IT BEEN DONE?}.
+\textbf{See Section D.4 in the Online Appendix}.
 \textbf{LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
@@
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
-As reported in the Online Appendix, Section \textbf{XXXX}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
+As reported in the Online Appendix, Section D.1, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
 on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
@@
 
-Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade (an annual mean of 68\% XXX D'OU VIENT CE CHIFFRE DE 68\% ??
-N'est-ce pas 80\% ???).
 %use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
@@
 %sum cover
-involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
+
+Three main comments can be made.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade (an annual mean of 68\%) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
@@
 	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
-	\input{Table_baseline10.tex}
+	%\input{Table_baseline10.tex}
 	\input{Table_baseline10_reshaped.tex}
@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In the Online Appendix, Sextion \textbf{XXX}, we report the two series of transport costs (fitted and unfitted) as percentage of the export price as in \cite{hummels2007}.
+To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In the Online Appendix, Section \textbf{XXX}, we report the two series of transport costs (fitted and unfitted) as percentage of the export price as in \cite{hummels2007}.
 Unsurprisingly, this accords with his results displayed in his Figures 5 (for Air) and 6 (for Vessel) for the years in common (until 2004).
@@
 These results holds whatever the considered year.
-\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix A.3 \textbf{CHECK XXX}.
+\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix B.2.
 If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.}

### 2021-10-26 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Pour le graphique entre beta direct et baseline
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
autres petites modfis

@@
 \begin{table}[htbp]
-	\caption{Comparison 2005-2013}
+	\caption{Baseline and direct beta estimate 2005-2013}
 	\begin{center}		
@@
 \end{table}%
+
+\begin{figure}[htbp]
+	\caption{Baseline and direct beta estimate }
+	\begin{center}
+		\includegraphics[height=8cm]
+		{../scatter_chronology_baseline_referee1.png}
+	\end{center}
+\end{figure}
 
@@
 in percentage of the export price, as in \cite{hummels2007}.
-We also adopt Himmel's (2007) terminology, contrasting the ``expenditure/importe value'' (i.e, the unfitted serie) and the ``fitted ad-valorem rate'' (i.e, corresponding to the ``pure'' transport costs series, composition effects excluded).
+We also adopt Hummel's (2007) terminology, contrasting the ``expenditure/importe value'' (i.e, the unfitted serie) and the ``fitted ad-valorem rate'' (i.e, corresponding to the ``pure'' transport costs series, composition effects excluded).
 Unsurprisingly, this accords with his Figure 5 (for Air) and 6 (Vessel) for the comme period (until 2014).

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Travail sur la réponse au referee
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the results on a detailed yearly basis and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triplet $(i,k,t)$, which confirm the robustness of the results.
-These results are reported in the Online Appendix, respectively in sections C and D.2.} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+Third, crucially for this robustness check the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the chronolgy of results and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triplet $(i,k,t)$, which confirm the robustness of the results.
+These results are reported in the Online Appendix, respectively in sections C \textbf{Bah non.
+Il n’y a pas de 3/10 dans la section C ?
+Je crois qu’il n’y a que D.2.
+avec quelque chose}and D.2.
+Bon, et je ne sais pas très bien en quoi les histogrammes nous aident ; je ne comprends } %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
 The picture is reversed for medians

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
modifs réunions

--- 2021-10-26 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.} for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -1.9\% for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
-	%\input{Table_baseline10.tex}
+	\input{Table_baseline10.tex}
 	\input{Table_baseline10_reshaped.tex}
@@
 First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by \textbf{50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
-AJUSTER LP}.
+International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
 Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (a), (d) and (e)).

### 2021-10-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
-\textbf{LP: CHECK THE NUMBERS}}.
- As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
+\textbf{LP: CHECK THE NUMBERS}}
+ As we conclude at the end of Section 3.3.1, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
  %- see in particular Figures 6 and 7 on pp.
@@
 
-\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is highly demanding computationally when the number of fixed effects increase.
+\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is highly demanding computationally when the number of fixed effects is high.
 This constraint notably drove us to impose the separability assumption and retain $s=3$ as the relevant sectoral degree of aggregation.
@@
 As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
-The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices.
+The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices.
 
@@
 This drove us to question our own empirical specification deeper.
-In the end, we yet decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that you propose does not go without a few important limitations.
+In the end, we decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that the referee suggests does not go without a few important limitations.
 We develop each line of argument in our answer.
 Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3).
-After reading our answer, we hope to convince you that the choice we made is the most relevant one.
+After reading our answer, we hope to convince you that we made the right choice.
 
@@
 Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates.
-In this respect, implementing the referee's method (see below) does not suppress the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces).
+In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive).
 Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
-\smallskip
+\textbf{Vérifier à la fin  ?
+GD}\smallskip
 
@@
 By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
-As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 only since 2001.
+As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 2001.
 We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix.
@@
 Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.)
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
 
 Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
-First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
+First, we challenge the advantages that the referee's estimation method would bring in comparison with ours.
 Second, we emphasize the drawbacks induced by this method.
@@
 This is based on the conclusion drawn from the robustness check on a reduced sample.
-If it is smaller in terms of observations, it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 80\% of the total value of flows (on a yearly basis).
+If it is smaller in terms of observations, it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 68\% of the total value of flows (on a yearly basis).
 As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex

@@
 
-For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
+\textbf{Je ne comprends pas ce paragraphe.
+La footnote semble être ancienne.
+Mais en quoi est-ce que les histogrammes nous aident ?
+Éventuellement à dire qu’il y a moins de solutions de coin lorsqu’on fait l’estimation en séparé ?
+G.D.}For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
 Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Changements graphiques chronologiques
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
last modifs

@@
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In the Online Appendix, Section \textbf{XXX}, we report the two series of transport costs (fitted and unfitted) as percentage of the export price as in \cite{hummels2007}.
+To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In Section E.1 from Online Appendix, we report the two series of transport costs (fitted and unfitted) as percentage of the export price as in \cite{hummels2007}.
 Unsurprisingly, this accords with his results displayed in his Figures 5 (for Air) and 6 (for Vessel) for the years in common (until 2004).

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex

@@
 \listoftables
+\listoffigures
 
@@
 
+\setcounter{table}{0}
+\renewcommand{\thetable}{D.\arabic{table}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{D.\arabic{figure}}
+
 \subsection{Robustness checks}
@@
 This holds throughout the period and for both transport modes. By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
-
-\setcounter{table}{0}
-\renewcommand{\thetable}{C.\arabic{table}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
 
@@
 \begin{table}[htbp]
-	\caption{Baseline and direct beta estimate 2005-2013}
+	\caption{Baseline and direct $\beta$ estimate, 2005-2013}
 	\begin{center}		
@@
 \begin{figure}[htbp]
-	\caption{Baseline and direct beta estimate }
+	\caption{Baseline and direct $\beta$ estimate }
 	\begin{center}
@@
 \begin{table}[htbp]
-	\caption{Comparison 2005-2019, price per qy or price per kg}
+	\caption{Comparison 2005-2019, Price per quantity or Price per kg}
 	\begin{center}		
@@
 \begin{figure}[htbp]
-	\caption{Comparison 2005-2019, price per qy or price per kg}
+	\caption{$\beta$ Estimate: Price per quantity or price per kg, 2005-2019}
 	\begin{center}
-	\includegraphics[height=8cm]
-	{../scatter_chronology_non_séparé_wgt_non_séparé_qy.png}
+\begin{tabular}{cc}
+	\includegraphics[height=8cm]{../scatter_chronology_non_separe_wgt_non_separe_qy.png}\\
+\multicolumn{2}{l}{{\footnotesize Notes: In both cases, estimation is run without imposing the separability assumption.}}
+\end{tabular}
 	\end{center}
@@
 
+\setcounter{table}{0}
+\renewcommand{\thetable}{E.\arabic{table}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{E.\arabic{figure}}
+
 \subsection{Eliminating the composition effects} \label{sec_oa:comp-effects}
 
-\subsubsection{More on the Hummel's methodology}
+\subsubsection{More on the Hummels's methodology}
 
@@
 in percentage of the export price, as in \cite{hummels2007}.
-We also adopt Hummel's (2007) terminology, contrasting the ``expenditure/importe value'' (i.e, the unfitted serie) and the ``fitted ad-valorem rate'' (i.e, corresponding to the ``pure'' transport costs series, composition effects excluded).
-Unsurprisingly, this accords with his Figure 5 (for Air) and 6 (Vessel) for the comme period (until 2014).
+The parallel with Hummels's (2007) terminology stands as follows: In Figure \ref{fig:comp_effects_as_in_Hummels}, the ``transport costs, excluding composition effects'' series refers to his ``fitted ad-valorem rate''; the ``transport costs'' series refers to his ``expenditure/import value''.
+Unsurprisingly, this accords with his Figure 5 (for Air) and 6 (Vessel) for the same period (until 2014).
 
 \begin{figure}[htbp]
-\caption{Characterizing the time trends: applying Hummel's (2007) method }
+\caption{Characterizing the time trends: applying Hummels's (2007) method }
 \label{fig:comp_effects_as_in_Hummels}
@@
 \subsubsection{Composition effects: Primary vs. Manufacturing sector}
+
 In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows.

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Chnagement mineurs Online appendix

@@
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
+\multicolumn{7}{l}{\parbox[l]{10cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
 Statistics are obtained weighting each observation by its share in trade (mode-dependent).
@@
 
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Mean values reported for the 2005 and 2013 values reported for $\beta$. Time trend coefficient is the annual growth rate.}}}
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Time trend coefficient is the annual growth rate.}}}
 \end{center}
@@
 \begin{table}[htbp]
-	\caption{Comparison 2005-2019, Price per quantity or Price per kg}
+	\caption{Comparison Price per quantity or Price per kg, 2005-2019}
 	\begin{center}		
-		\input{Comp_baseline_wgt_qy.tex}
+		\input{Comp_baseline_wgt_qy_reshaped.tex}
 	\end{center}
@@
 Here, we report the unfitted measure of transport costs versus the ``pure'' transport costs series, i.e composition effects excluded
-in percentage of the export price, as in \cite{hummels2007}.
+in percentage of the export price, as in \cite{hummels2007}, for each transport mode.
 The parallel with Hummels's (2007) terminology stands as follows: In Figure \ref{fig:comp_effects_as_in_Hummels}, the ``transport costs, excluding composition effects'' series refers to his ``fitted ad-valorem rate''; the ``transport costs'' series refers to his ``expenditure/import value''.
@@
 \begin{tabular}{cc}
-{\small (A) Ad-valorem air freight} & {\small (B) Ad-valorem vessel freight}\\
+{\small (A) Air } & {\small (B) Vessel }\\
 \multicolumn{2}{c}{{\small In percent of the export price}} \\
@@
 & \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
-{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight} \\
 \end{tabular}
@@
 \end{figure}
+
+\textbf{CHANGER EXP/IMPORT VALUE POUR TRANSPORT COSTS}
 

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Changement texte + online

@@
 
-\textbf{JH : I WOULD DELETE EVERYTHING FROM `` It is true that"}
+\textbf{XXX JH : I WOULD DELETE EVERYTHING FROM `` It is true that"- LP, OK METTRE EN SECTION 3 Comparaison des resultats XXX }
 
@@
 \textbf{See Section D.4 in the Online Appendix}.
-\textbf{LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
-Le remonter dans le corps PPL?}} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+\textbf{XXX LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
+Le remonter dans le corps PPL?
+XXX }} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
@@
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} and Appendix \ref{app:4digit} for the 4-5 results and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).
-\textbf{ADD 10-3 RESULTS TO BE DONE}
+However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} and Section B of the online appendix for the 4-5 results; and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).
 
@@
 Figure \ref{fig:part_cout_additif} reports the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$.
-.
 
@@
 Three main comments can be made.
-First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade (an annual mean of 68\%) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
+First, the subset of large trade flows we are considering (around 8\% of the observations of our initial sample, but the vast majority of the total value of aggregated trade, with an annual mean of 68\%) involve a higher share of additive costs: between 48\% and 55\% of total costs on average, versus 30\% to 44\% on the whole sample.
 %JH: la phrase suivante est en contradiction avec celle qui suit encore juste après: "Second, assuming that country/product fixed effects are separable in these two dimensions tends to overestimate the value and the share of the additive costs."
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, crucially for this robustness check the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the chronolgy of results and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triplet $(i,k,t)$, which confirm the robustness of the results.
-These results are reported in the Online Appendix, respectively in sections C \textbf{Bah non.
-Il n’y a pas de 3/10 dans la section C ?
-Je crois qu’il n’y a que D.2.
-avec quelque chose}and D.2.
-Bon, et je ne sais pas très bien en quoi les histogrammes nous aident ; je ne comprends } %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+Third, crucially for this robustness check the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the results with/without the separability assumption on a yearly basis and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triplet $(i,k,t)$, which confirm the robustness of the results.
+These results are reported in the Online Appendix, Section C.
+\textbf{XXX ENLEVER LA PARTIE SUR LE LIEN BETA-SEPARABILITE EN ANNEXE ET AJUSTER FOOTNOTE XXX}} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
 The picture is reversed for medians
@@
 \end{figure}
-
-\textbf{LP: Guillaume, changer les legendes de la figure enlever beta, mettre en positif la part des additifs et ajouter legende en ordonnee...}
+%viens de comparaison.do
+
 %However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
@@
 Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are very similar when using 10-digit product and 5-digit products.
-If anything, the $\beta$ seems on average slightly higher when estimated based on a 3 (sector)-10(product) database, indicating a (small) downward bias in our estimates of additive costs.
+If anything, the $\beta$ seems on average slightly higher when estimated based on a 3 (sector)-10 (product) database, indicating a (small) downward bias in our estimates of additive costs.
 But overall, no major difference arises both regarding levels and trends - results are especially very close for the Air transportation mode.
-This has important implications for the exercise performed in section \ref{sec:theory}: welfare gains computations should not be significantly affected by our use of 5 (sector)-5 (product) digit data.
+This has important implications for the exercise performed in Section \ref{sec:theory}: welfare gains computations should not be significantly affected by our use of 3 (sector)-5 (product) digit data.
 This is consistent with e.g.
@@
 	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
-	\input{Table_baseline10.tex}
 	\input{Table_baseline10_reshaped.tex}
@@
 
-\textbf{LP pour Guillaume: Changer les légendes pour les aligner sur la terminologie du papier, fitted vs unfitted.}
-
 \begin{figure}[htbp]
-\caption{Characterizing the time trends: applying Hummel's (2007) method }
+\caption{Characterizing the time trends: Applying Hummels's (2007) method }
 \label{fig:comp_effects_as_in_Hummels}
@@
 %& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
-{\small (c) Ad-valorem air freight} & {\small (d) Ad-valorem vessel freight}\\
-\multicolumn{2}{c}{{\small As indices with reference value 100 in 1974} }\\
+{\small (c) Air } & {\small (d) Vessel}\\
 \includegraphics[width=2.5in, height=2in]{figure5_comme_hummels_base100.jpg}
@@
 \end{tabular}
-	\begin{minipage}  [c]  {6in}
-		\scriptsize
-\textbf{A AJUSTER SELON LEGENDE FINALE}.
+	\begin{minipage}  [c]  {5in}
+		\footnotesize
+Total transport costs expressed in percentage of the value shipped, then built as indices with reference value 100 in 1974.
 \end{minipage}
@@
 
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \cite{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$, Equations (\ref{eq:comp_effects_adv}) and (\ref{eq:comp_effects_add}) in our case).
-\cite{hummels2007} takes the unweighted average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \cite{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$).
+\cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 We proceed differently, as we weight each flow by its relative value on total trade flows (on the relevant year of observation), based on the argument that this weighting scheme does not overweight the small flows in value.
@@
 
-The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummel's weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
+The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighting scheme) prompts two comments.
 First, differences do show up, in particular for the multiplicative component of air transport.

### 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Pour compilation et croix
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
-\textbf{Vérifier à la fin  ?
+\textbf{XXXXVérifier à la fin?
 GD}\smallskip
@@
 Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{XXXX Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 GD?}
 
@@
 We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.\textbf{CETTE FIGURE, SUR QUELLE ANNEE, QUEL MODE DE TRANSPORT?
-}
+xxxxx}
 
@@
 \item Even on this very reduced sample, units are not available for many years over our period, especially in the 1970s and the 1980s; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
-\textbf{LP: Je ne suis pas sure de cet argument.
+\textbf{XXXX LP: Je ne suis pas sure de cet argument.
 Veut-on dire que dans les "quantity", en fait on trouve de tout mais pas des unites?

### 2021-10-28 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Various footnotes main text

@@
 Second, we revisit the welfare gains of trade liberalisation in presence of additive costs.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
 Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points.
@@
 On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019.
-By allowing for time-varying per-unit costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019.
+By allowing for time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019.
 Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
@@
 In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
-It is true that, compared to \cite{Irrazabal_2015}, our study covers a narrower set of trade costs focused on transport costs.
+%It is true that, compared to \cite{Irrazabal_2015}, our study covers a narrower set of trade costs focused on transport costs.
 However, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost.
@@
 
-\textbf{XXX JH : I WOULD DELETE EVERYTHING FROM `` It is true that"- LP, OK METTRE EN SECTION 3 Comparaison des resultats XXX }
+%\textbf{XXX JH : I WOULD DELETE EVERYTHING FROM `` It is true that"- LP, OK METTRE EN SECTION 3 Comparaison des resultats XXX }
 
 \smallskip
-
-% nous sommes les premiers à fournir une analyse en termes de welfare basée sur une estimation des "vrais" coûts multiplicatifs et additifs sur une
 
@@
 This is due to the fact that in the Census data, information on the quantity of goods is not mode-specific.
-We yet provide a robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.
-\textbf{See Section D.4 in the Online Appendix}.
-\textbf{XXX LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
+Nevertheless, Section D.4 in the Online Appendix reports extensive robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.
+%\textbf{XXX LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
 Le remonter dans le corps PPL?
-XXX }} The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
+XXX }
+}
+The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
@@
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, crucially for this robustness check the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also run two other robustness checks: \textit{i}) comparing the results with/without the separability assumption on a yearly basis and \textit{ii}) reporting the histogram of the distribution for $\beta$ over the triplet $(i,k,t)$, which confirm the robustness of the results.
-These results are reported in the Online Appendix, Section C.
-\textbf{XXX ENLEVER LA PARTIE SUR LE LIEN BETA-SEPARABILITE EN ANNEXE ET AJUSTER FOOTNOTE XXX}} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+Third, crucially for this robustness check the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also report in Section C from the Online Appendix a robustness check comparing the results with/without the separability assumption on a yearly basis} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
 The picture is reversed for medians

### 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Travail sur réponse au referee
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now presented in footnote \textbf{12 page 8}.
 \textbf{XXXXVérifier à la fin?
@@
 Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects.
-This substantially reduces the number of point estimates, as it can be inferred from Table \ref{tab_comp_referee1_baseline} reported later in our answer and in the Online Appendix, \textbf{Table C.1, Section D.3} ).
+This substantially reduces the number of point estimates, as it can be inferred from Table \ref{tab_comp_referee1_baseline} reported later in our answer and in the Online Appendix, \textbf{Table D.1, Section D.3} ).
 
@@
      Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method.
-In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a constant $\beta$ invariant over time/sector/origin country (even though mode-specific).
+In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a mode-specific $\beta$ invariant over time/sector/origin country.
 Rather than Equation (\ref{eq:Hummels}), this amounts starting from the functional form: $f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta}$.
 
-     Two alternative strategies are yet available.
-Let us start with the referee's method.
+     Two alternative strategies are possible.
+Let us start with our interpretation of the referee's method.
 It starts from the functional form linking transport costs and the fas price as specified in the Equation (\ref{eq:Hummels}).
@@
 \noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
-With 2 equations and 2 endogenous variables, the system might be solved.
+With 2 equations and 2 endogenous variables, the system can be solved.
 
@@
 
-In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
-Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such.
+In this respect, the referee's method and ours are equivalent, in that they both allow to compute the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
+Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such, as we do now.
 In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
@@
 We agree that this was not sufficiently transparent in the submitted version.
-We hope that the revised version now fits the referee's expectations on this point.
+We hope that the revised version now meets the referee's expectations on this point.
 
@@
 Eliminating this dimension of the paper would be detrimental to its value-added.
-\item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
+\item As underlined before, the method’s regressions are run country by country, and by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
 Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
 This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline}.
-Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+Put it differently, this methodology discards country-sector pairs that export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
 In this respect, the induced selection bias may reduce the general scope of the transport costs estimates.
@@
 	\begin{center}		
-		\input{Comp_baseline_referee1_reshaped.tex}
+		\input{../../revised_article/Online_Appendix/Comp_baseline_referee1_reshaped.tex}
+		\input{../../revised_article/Online_Appendix/Comp_baseline_referee1.tex}
 
@@
 
-Following the referee's advice, we thus run this alternative estimation method as robustness check.
+Following the referee's advice, we thus run this alternative estimation method as a robustness check.
 Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013.

### 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Changement mineur reponse référé 1

--- 2021-10-27 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
@@
 % sum cover
-\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
-The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
-\textbf{LP: CHECK THE NUMBERS}}
+\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the large sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the complete sample).
+The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.}
  As we conclude at the end of Section 3.3.1, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
@@
 
-The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummel's Methodology.
+The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummels' Methodology.
 That is, one can alternatively estimate the share of the additive component as:}
@@
 across all districts and all 10-digit products pertaining to the same 5-digit category.
-[...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}
+[...] The first advantage of this so-called Hummels' approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}
 
-\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummel's.
+\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels'.
 We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
@@
 In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
-\textbf{Vérifier à la fin  ?
-GD}\smallskip
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 12 page 8.
+\textbf{XXX Verifier à la fin?
+GD XXX }\smallskip
 
@@
 Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummel's dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummels' dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
 
@@
 %
-%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
+%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummels' original dataset from 1974-2004).
 We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
@@
 \begin{itemize}
-\item \textit{The first advantage of this so-called Hummel's Approach is that the above
+\item \textit{The first advantage of this so-called Hummels' Approach is that the above
 regression can be estimated separately for various country-industry pairs, without
@@
 In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
-Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such.
+Put it differently, our method is consistent with Hummels' methodology - provided that it is adequately reported as such.
 In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
@@
 would become computationally expensive without it. But this reasoning
-brings us back to my original point that the authors can use the Hummel's
+brings us back to my original point that the authors can use the Hummels'
 Methodology to circumvent the computational burden.}
@@
 It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
-Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
+Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummels' dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
 We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
@@
 \begin{enumerate}
-\item In the original Hummel's database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
+\item In the original Hummels' database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
@@
 So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade.
-But, we disagree with Hummel's findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export).
+But, we disagree with Hummels' findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export).
 Meaning that gains from trade are purely due to reduction in transport costs at the sector, but not from switch towards low-elasticity sectors.

### 2021-10-28 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Changement texte + online

@@
 
-\paragraph{Distribution of the $\beta$} As final check, we also investigate whether the varying pattern over time/country of origin/product is also robust to the separability assumption.
-With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
-the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (A), (c))) and the non-separability scenario (Panels (B), (c)).
-
-\begin{figure}[htbp]
-\caption{Histogram of the share of additive costs in total costs (restricted sample)}
-\label{fig:histogram_beta_robustness}
-\begin{center}
-\begin{tabular}{cc}
-\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
-\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
-\end{tabular}
-\end{center}
-\end{figure}
-
-\textbf{Je ne comprends pas ce paragraphe.
-La footnote semble être ancienne.
-Mais en quoi est-ce que les histogrammes nous aident ?
-Éventuellement à dire qu’il y a moins de solutions de coin lorsqu’on fait l’estimation en séparé ?
-G.D.}For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
-Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
-\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section 4 of the paper, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}
-
 \subsubsection{Estimating $\widehat{\beta}_{i,s}$ directly}
@@
 \end{figure}
+
+\textbf{CHANGER LEGENDE PAS TRES BEAU}
 

### 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement texte réponse référé 1

@@
 significant changes to the paper that we hope address your concerns. We first give you an overview
-of the revision (Section\ref{sec:main_changes}) before answering in detail each of your comments (Section\ref{sec:detailed_answers}).
+of the revision (Section \ref{sec:main_changes}) before answering in detail each of your comments (Section \ref{sec:detailed_answers}).
 Number of sections, pages and equations mentioned in the text refer to the revised version.
@@
 Further, we provide a robustness check for this separability assumption that $\tau_{ik} = \tau_i\tau_{k}$ and
- $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (Table 2) \textbf{XXX CHECK AT THE END XXX}).
+ $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (Table 2 and \textbf{Figure XXX}) \textbf{XXX CHECK AT THE END XXX}).
 We check the robustness of our results by re-running the estimation without the separability assumption.
 This comes at the cost of a substantial increase in the number of fixed in the regression.
-In 2019, for example, there are 14,016 country*sector pairs.
+In 2019, for example, there are 14,016 country$\times$sector pairs.
 Adding 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$).
@@
  For this reason, we have decided to run the robustness check on a reduced sample.
- We select, for each year, the largest importers that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
- We keep in the sample all trade observations from these importers and in these sectors.
- This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1), yet it remains quite large in terms of trade coverage (mean of 68\%).
+For each year, we select the largest exporters that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
+ We keep in the sample all trade observations from these exporters and in these sectors.
+ This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1).
+It yet remains quite large in terms of trade coverage (mean of 68\%).\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the large sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the complete sample).
+The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} On this reduced sample, we re-run the estimation with and without the separability assumption.
+The results are reported in Section 3.3.1 (Table 2 and \textbf{Figure XXX}).
+\textbf{XXX CHECK AT THE END XXX}) As we conclude at the end of this section, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
  % use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
@@
 % sum cover
-\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the large sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the complete sample).
-The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.}
- As we conclude at the end of Section 3.3.1, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
- %- see in particular Figures 6 and 7 on pp.
-25-26 \textbf{XXX CHECK AT THE END XXX}.
-Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.
 
@@
 
-\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels'.
-We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
-As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
-The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices.
+\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels' one.
+We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'' \textbf{XXX Check XX}).
+As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{is(k)}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
+The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices, along the lines you suggest.
 
@@
 We develop each line of argument in our answer.
-Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3).
+Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3 \textbf{XXX CHECK XXX }).
 After reading our answer, we hope to convince you that we made the right choice.
@@
 By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
-As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 2001.
-We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix.
 
-Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummels' dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
-
-Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
-First, we challenge the advantages that the referee's estimation method would bring in comparison with ours.
+Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
+First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
 Second, we emphasize the drawbacks induced by this method.
@@
 
-\paragraph{a) The advantages of the referee's method are not as important as suggested}
+\paragraph{a) The advantages of the referee's method are not as specific to this method as suggested}
 
@@
 \textbf{Our answer:}
-
-First, as already discussed in our letter, the separability assumption does not seem to be a strong assumption.
+As already discussed in our letter, the separability assumption does not seem to be a strong assumption.
 This is based on the conclusion drawn from the robustness check on a reduced sample.
 If it is smaller in terms of observations, it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 68\% of the total value of flows (on a yearly basis).
-As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
-
-Second, the referee's method imposes a drastic selection of countries / sectors to be included in the estimation.
-Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects.
-This substantially reduces the number of point estimates, as it can be inferred from Table \ref{tab_comp_referee1_baseline} reported later in our answer and in the Online Appendix, \textbf{Table C.1, Section D.3} ).
+As we conclude at the end of Section 3.3.1 \textbf{XXX CHECK XXX}, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
 
@@
      \textbf{Our answer:}
-     We agree with the referee that the comparison with the literature (\citealp{hummels2007}) in particular) was not straightforward for the reader in the submitted version.
+     We agree with the referee that the comparison with the literature (\citealp{hummels2007} in particular) was not straightforward for the reader in the submitted version.
 We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
@@
 \noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
-With 2 equations and 2 endogenous variables, the system might be solved.
+With two equations and two endogenous variables, the system can be solved.
 
@@
 
-\paragraph{b) The costs are high}
-\begin{itemize}
-\item[Concern 1] The suggested estimation strategy implies having much less data to exploit, for two reasons.
-\begin{enumerate}
-\item Information about the port of entry are only available since 1989, which is already smaller than our full sample starting
-in 1974.
-On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
-Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
-In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
-Eliminating this dimension of the paper would be detrimental to its value-added.
-\item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
-This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline}.
-Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
-In this respect, the induced selection bias may reduce the general scope of the transport costs estimates.
-This concern might yet be mitigated by noticing that the covered value of trade flows is only slightly lower.
-If it reduces the sample of sectors and origin countries covered, this method preserves the majority of trade flows in value.
+As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 2001.
+We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix, Section D.3, Table D.1 and Figure D.3.\textbf{XXX Check XXX}\footnote{Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the referee's method versus ours.} For sake of reading clarity, we also report here the results of Table D.1 (here Table \ref{tab_comp_referee1_baseline}).
 
@@
 	\begin{center}		
-		\input{Comp_baseline_referee1_reshaped.tex}
+		\input{../../revised_article/Online_Appendix/Comp_baseline_referee1_reshaped.tex}
+		%\input{../../revised_article/Online_Appendix/Comp_baseline_referee1.tex}
 
-{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Mean values reported for the 2005 and 2013 values for $\beta$. Time trend coefficient is the annual growth rate.}}}
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Time trend coefficient is the annual growth rate.}}}
 \end{center}
@@
 \end{table}%
-   \end{enumerate}
 
-Following the referee's advice, we thus run this alternative estimation method as robustness check.
-Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013.
-As noted above, the partners-sectors covered are lower.
 Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
@@
 
+Yet, the suggested method does not go without a few important limitations, that we now develop.
+
+\paragraph{b) The referee's method is not without drawbacks}
+
+\begin{itemize}
+\item[Concern 1] The suggested estimation strategy implies having much less data to cover.
+
+As noted above, the suggested method implies that the estimation is run at the sector-country level (on top of being mode-year specific), exploiting the variability within each origin country-3d sector pair across 10-d sub-sectors and ports of entry.
+Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
+This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline} (also reported in the online appendix, Section D.3.
+\textbf{XXX Check XXX}.
+Put it differently, this methodology discards country-sector pairs that export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+This may induce a selection bias in the sample covered for estimation.
+As reported in Table  \ref{tab_comp_referee1_baseline}, if the number of pairs covered is much lower (especially in the list of countries covered), this concern can yet be mitigated by noticing that the covered value of trade flows is only slightly lower.
+If it reduces the sample of sectors and origin countries covered, this method preserves the majority of trade flows in value.
+
+The most important limitation rather comes from the coverage period.
+Information about the port of entry are only available since 1989, which is already smaller than our full sample starting in 1974.
+On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
+Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
+In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
+Eliminating this dimension of the paper would be detrimental to its value-added.
+
 \item[Concern 2] The suggested method features less accuracy in the estimation of the $\beta$.
-If we take the value of the $\beta$ by itself, there is no criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
+
+If we take the value of the $\beta$ by itself, there is no clear criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
+Further, as shown in Table \ref{tab_comp_referee1_baseline}, no clear conclusion on a potential upwards or downwards bias that would be attached to our method.
 Things are more clear-cut in terms of accuracy of the estimation.
@@
     \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$, for each couple $i,s$ (by year and transport mode).
-We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.\textbf{CETTE FIGURE, SUR QUELLE ANNEE, QUEL MODE DE TRANSPORT?
-}
+We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.\textbf{XXX CETTE FIGURE, SUR QUELLE ANNEE, QUEL MODE DE TRANSPORT?
+xxxxx}
 
@@
 We have checked that a similar conclusion applies on other years (they are not reported here for sake of brevity but they are available upon request).
+    \end{itemize}
 
 One explanation of the limited accuracy with the alternative method might be the following.
-Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and across origin countries.
+Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and \textit{across} origin countries.
 Put it in plain words, it means that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
 One might view this as a disputable assumption.
-From a statistical point of view further, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component, that could explain the lower accuracy of the estimation.
+From a statistical point of view further, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component, that explains the lower accuracy of the estimation.
 
-\end{itemize}
 \end{itemize}
@@
 However, we also want to keep track of this alternative method as further robustness check.
-For this purpose, we have amended the online Appendix with a section devoted to the referee's method (\textbf{Section D.3}).
+For this purpose, we have amended the online Appendix with a section devoted to the referee's method (Section D.3 \textbf{XXX Check XXX}).
 We deeply hope to convince you of the relevance of this choice.
@@
 We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
-After checking with the US Bureau of Census, data is only available since 1989, now available to 2019.
+After checking with the US Bureau of Census, HS-10 data is only available since 1989, now available to 2019.
 As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.
-In front of this arbitrage, we decided to maintain the case $s=3, k=5$ as baseline, but running estimation at the $s=3,k=10$ in the new Sub-Section 3.3.3 of Section 3.3 devoted to robustness analysis.
+In front of this arbitrage, we decided to maintain the case $s=3, k=5$ as baseline, but running estimation at the $s=3,k=10$ in the new Sub-Section 3.3.3 \textbf{XXX CHECK XXX }devoted to robustness analysis.
 
@@
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
-We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper (see Section 3.3.3 devoted to the robustness to the aggregation level).
-\textbf{XXX A FINIR XXX DANS LE PAPIER ET DANS LA REPONSE}
+We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper (see Section 3.3.3 \textbf{XXX CHECK XXX }).
 
@@
 a reduction of flows by a factor of 4.5.
+\textbf{XXX LP: Donne l'impression que ce n'est pas ce qu'on fait, alors que si, en fait?
+XXX}
 
 \item Even on this very reduced sample, units are not available for many years over our period, especially in the 1970s and the 1980s; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
-\textbf{LP: Je ne suis pas sure de cet argument.
+\textbf{XXXX LP: Je ne suis pas sure de cet argument.
 Veut-on dire que dans les "quantity", en fait on trouve de tout mais pas des unites?
@@
 
-The results are the following ... \textbf{To DEVELOP}
-
 \end{enumerate}
 
+Because the referee's concern yet seems important to address, we run a robustness check to the use of quantity versus weight.
+These results, obtained on a reduced sample given the above-mentioned constraints, have been added in the online appendix, Section D.4, Table D.2 and Figure D.4\textbf{XXX CHECK XXX }.
+For sake of reading clarity, we report here the results of Table D.2 (Table \ref{tab_comp_wgt_qty} here).
+For each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019).
+
+\begin{table}[htbp]
+	\caption{Comparison: Price per quantity versus per kg, 2005-2019}
+	\begin{center}		
+		\input{../../revised_article/Online_Appendix/Comp_baseline_wgt_qy_reshaped.tex}
+		%\input{../../revised_article/Online_Appendix/Comp_baseline_referee1.tex}
+
+{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Time trend coefficient is the annual growth rate.}}}
+\end{center}
+	\label{tab_comp_wgt_qty}%
+\end{table}%
+
+Two main results emerge.
+First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
+In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
+In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
+Second, in terms of $\beta$ estimates, it is true that our method seems to yield a higher share of additive costs than when considering the price per quantity.
+This is somehow an expected result, as we do agree that there should be a link between the transport costs and the weight, or the volume of the goods transported.
+In all cases though, the share of additive costs remains substantial on average.
+In this respect, one might view these results as the lower bound for the size of additive costs in total transport costs.\smallskip
+
 These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression.
-We hope to convince you that \textbf{XXX TO BE COMPLETED XXX}
+We then decided to maintain our original empirical strategy covering the whole period 1974-2019 in the revised version, and to extend the online Appendix with this supplementary robustness check.
+We hope to convince you that this is a relevant choice.
+\textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
 

### 2021-10-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
chnagement réponse référé 1

@@
 We devoted a lot of thought to a potential switch from weight to quantity.
-Running the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis yet appears really uneasy in our context for several, non-trivial reasons:
+Running the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis yet appears really uneasy in our context for several, non-trivial reasons, that we develop below.
 
-\begin{enumerate}
-\item In the original Hummels' database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
+In the original Hummels' database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
@@
 
-\item Even on this very reduced sample, units are not available for many years over our period, especially in the 1970s and the 1980s; this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
+Even on this very reduced sample, information on the \textit{units} considered for \textit{quantities} are not available for many years over our period, especially in the 1970s and the 1980s; discarding information on this would lead to blurred results as many types of units can be found within a same sector (dozens, liters, etc...) which are hardly comparable with each other.
+The option of not considering these observations would mean a drastic reduction in the estimation period that could be covered.
+Again, this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
 \textbf{XXXX LP: Je ne suis pas sure de cet argument.
 Veut-on dire que dans les "quantity", en fait on trouve de tout mais pas des unites?
-}
+Je crois que je comprends, check with Guillaume }
 
-\item Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious \textbf{XXX GUILLAUME: EXAMPLES XXX}.
-At a more fundamental level, one may argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
-For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
-\textbf{XXX GUILLAUME: Do we have evidence, even anecdotal, on this point?
-LP: Sinon, laisser tomber cet argument}
+\textbf{SUPPRIMER CE QUI EST EN GRAS Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious} \textbf{XXX GUILLAUME: EXAMPLES XXX}.
 
-    \item Considering the US census data which complete the original \cites{hummels2007} database, it is true that we can retrieve information on quantities.
+Considering the US census data which complete the original \cites{hummels2007} database, it is true that we can retrieve information on quantities.
 Specifically, this can be obtained crossing HS-10 product information with the correspondence table between the HS-6 product classification and the associated units (kg, tons, liters, numbers, etc.) also provided by the US Census.
 We are able to collect information since 2009.
-Investigating this database, it appears that in many 3-digit sectors, there are more than one type of quantities reported (for instance, sector \textbf{XXX} includes goods both recorded in terms of \textbf{YYY} and \textbf{ZZZ}.) To identify a proper unit price, we consequently reduce the flows to the 3-digit sectors reporting a single quantity.
-    \item As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
+Investigating this database, it appears that in many 3-digit sectors, there are more than one type of quantities reported (for instance, sector \textbf{XXX} includes goods both recorded in terms of \textbf{YYY} and \textbf{ZZZ}.) To identify a proper unit price, we consequently reduce the flows to the 3-digit sectors reporting a single quantity \textbf{at the 3-digit level}.
+
+Getting deeper into the referee's suggestion, we were yet confronted to another issue.
+As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
 As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
@@
 In the specific case of quantities, this raises a non-trivial issue yet.
-Specifically, imposing a country $i$ fixed effect would imply that the cost of sending one unit of say, tee-shirt from France is the same than sending one unit of say, car; and that this cost would be added to the sector-fixed effect.
-This does not seem to us as a reasonable assumption; in this case specifically, skipping the separability assumption does appear as the appropriate choice.
+Specifically, imposing a country $i$ fixed effect would imply that the cost of sending one unit of say, tee-shirt from France is the same than sending one unit of say, car; and that this cost would be \textit{added} to the sector-fixed effect.
+\textbf{XXX PAS CLAIR IL FAUT MIEUX EXPLIQUER } This does not seem to us as a reasonable assumption; in this case specifically, skipping the separability assumption does appear as the appropriate choice.
 
@@
 \noindent where $p^u_{ik}$ refers to the unit price associated with quantity. Depending on the sector, two different units are considered, kilograms and numbers.
-
-\end{enumerate}
 
@@
 \textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
+
+At a more fundamental level, one may argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
+For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
+\textbf{XXX Guillaume : Do we have evidence on this point ?
+XXX}
 

### 2021-10-29 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Sur un graphique pour non separe
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
chnagement réponse référé 1

@@
 \begin{figure}[htbp]
-\caption{Share of additive costs (3-digit classification)}
+\caption{Share of additive costs $\beta$ (3-digit classification)}
 \label{fig:part_cout_additif}
@@
 \begin{minipage} [c]  {5in} \scriptsize%
-Notes: Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
+Notes:The share of additive costs $\beta$ is reported on the $y-$axis.
+Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
 \end{minipage}
@@
 
+\begin{figure}[htbp]
+	\caption{Transport cost time trends: robustness to the separability assumption}
+	\label{fig:comp_separability_SITC5}
+	\begin{center}
+		\includegraphics[height=8cm]
+		{scatter_chronology_non_separe_pour_robustesse_ns.png}
+	\end{center}
+\end{figure}
+%viens de comparaison.do
+
 \subsubsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) is subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.

### 2021-10-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Pour le graphique de comparaison des amplitudes...
... de ci de l’estimation des betas

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
modifs answer R1

--- 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
@@
 % sum cover
-\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the large sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the complete sample).
-The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.}
+\footnote{In the restricted sample, we have around 25 sectors from 14 countries for Air transport, vs 230 sectors from 190 countries in the large sample; and around 50 sectors from 20 countries for Vessel, vs 600 sectors from 190 countries in the complete sample).
+The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.
+\textbf{LP: CHECK THE NUMBERS}}
  As we conclude at the end of Section 3.3.1, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
@@
 
-\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is highly demanding computationally when the number of fixed effects is high.
+\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is computationally highly demanding when the number of fixed effects is high.
 This constraint notably drove us to impose the separability assumption and retain $s=3$ as the relevant sectoral degree of aggregation.
 
-The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummels' Methodology.
+The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummel's Methodology.
 That is, one can alternatively estimate the share of the additive component as:}
@@
 across all districts and all 10-digit products pertaining to the same 5-digit category.
-[...] The first advantage of this so-called Hummels' approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}
+[...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}
 
-\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels'.
+\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummel's.
 We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'').
@@
 
-\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares.
+\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on Non-Linear Least Squares.
 However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
@@
 $\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
-Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates.
-In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 12 page 8.
-\textbf{XXX Verifier à la fin?
-GD XXX }\smallskip
+Should we relax these restrictions, standard linear least squares estimates often deliver negative, meaningless estimates.
+In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee's method is indeed less computational intensive).
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote \textbf{12 page 8}.
+\textbf{Vérifier à la fin ?
+GD}\smallskip
 
@@
 
-The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007), yielding Equation (\ref{eq:estimation_ref1}) as estimation equation.
-By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
+The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in \citet{hummels2010}, yielding for estimation Equation (\ref{eq:estimation_ref1}).
+This implies running the estimation by year and transport mode, for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
 As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 2001.
@@
 Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummels' dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original \cites{hummels2007} dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) %\textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}
 
 Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
-First, we challenge the advantages that the referee's estimation method would bring in comparison with ours.
+First, we detail the reasons why the advantages the referee's estimation method would bring in comparison with ours may not be as important as expected.
 Second, we emphasize the drawbacks induced by this method.
@@
 %
-%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummels' original dataset from 1974-2004).
+%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummel's original dataset from 1974-2004).
 We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel).
@@
 \begin{itemize}
-\item \textit{The first advantage of this so-called Hummels' Approach is that the above
+\item \textit{The first advantage of this so-called Hummel's Approach is that the above
 regression can be estimated separately for various country-industry pairs, without
@@
 We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
-In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
+In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in \citet{hummels2007}:
     \begin{equation}
@@
 
-     Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method.
+     Equation (\ref{eq:Hummels}) lies at the root of \cites{hummels2007} method.
 In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a constant $\beta$ invariant over time/sector/origin country (even though mode-specific).
@@
 In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
-Put it differently, our method is consistent with Hummels' methodology - provided that it is adequately reported as such.
-In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
+Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such.
+In comparison with \cite{hummels2007}, one additional advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
 Section 4 ([Section 3] of the submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
 We agree that this was not sufficiently transparent in the submitted version.
-We hope that the revised version now fits the referee's expectations on this point.
+We hope that the revised version now meets the referee's expectations on this point.
 
@@
 Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
-In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
-Eliminating this dimension of the paper would be detrimental to its value-added.
+We hope the revised version of our paper makes clear that historical coverage is very important \emph{per se} for our analysis, especially in light of the new Section 5.
+In the latter, we revisit the welfare gains of trade liberalization in presence of additive costs.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
+Therefore, eliminating the long time length from the paper would be detrimental to its value-added.
+
 \item As underlined before, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
@@
 We follow earlier literature (see e.g.
-Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
+\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
 First-stage estimates reported in section B.4 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation.
@@
 would become computationally expensive without it. But this reasoning
-brings us back to my original point that the authors can use the Hummels'
+brings us back to my original point that the authors can use the Hummel's
 Methodology to circumvent the computational burden.}
@@
 It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation.
-Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummels' dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
+Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummel's dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
 We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
@@
 \begin{enumerate}
-\item In the original Hummels' database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
+\item In the original Hummel's database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
@@
 So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade.
-But, we disagree with Hummels' findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export).
+But, we disagree with Hummel's findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export).
 Meaning that gains from trade are purely due to reduction in transport costs at the sector, but not from switch towards low-elasticity sectors.

### 2021-10-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Travail sur quantity units
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

--- 2021-10-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
@@
     \begin{center}
-    \includegraphics[height=4in]{accuracy_beta.pdf}
+    \includegraphics[height=4in]{../../revised_article/comparaison_amplitude_baseline_referee1_2013_ves.png"}
+    \begin{minipage} [c]  {5in} \scriptsize%
+    	    Note: Ratio of the confidence intervals = $ _{}ln(CI_{is}/CI^{ref}_{is})$
+    \end{minipage}
+
     \end{center}
@@
 
-    In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
+   \textbf{XXXX J’ai laissé en plan le calcul sur d’autres modes et d’autres années à reprendre ?
+commit 16b215ca44c3f4cfb0fd43d3aaec6e6532aa7f9d  GD} In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
 Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
@@
 \paragraph{Our answer:}
-\noindent The referee definitely has a point here, in the sense that Census data do report the quantity of goods.
-At a more fundamental level, this raises the following question: Do we have reason to believe our measure of unit prices generate significant deviations from the one relying on quantities?
-This referee's remark echoes recent findings by \cite{Lashkaripour_JIE2020}: based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
-Finally, \cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
-This result of transport costs close to be totally iceberg is quite different from our own findings of additive costs representing 30 to 45\% of total costs.
-The two stories are not necessarily incompatible, however.
+\noindent This is a very important point we had not thought about, and we thank the referee for raising it.
+The referee definitely has a point here, in the sense that Census data do report the quantity of goods (and sometimes even two quantities).
+%At a more fundamental level, this raises the following question: Do we have reason to believe our measure of unit prices generate significant deviations from the one relying on quantities?
+To that extent, the findings by \cite{Lashkaripour_JIE2020} were an eye-opener.
+Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
+\cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
+This result of transport costs close to be totally iceberg contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
+We believe the two sets of findings are complementary.
+
+We now report results based on per-unit prices.
+They confirm the results of \cite{Lashkaripour_JIE2020}
+Yet, first, looking at per-unit prices leads to leaving aside a large part of the data, both sectorally and chronologically.
+Second  it is interesting to look at additive costs on a per-weight basis even there is a correlation between unit price and weight.
+
+%The two stories are not necessarily incompatible, however.
 Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
@@
 
-Nevertheless, the point made by the referee is an important one and deserves careful investigation.
-We devoted a lot of thought to a potential switch from weight to quantity.
-Running the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis yet appears really uneasy in our context for several, non-trivial reasons, that we develop below.
+%Nevertheless, the point made by the referee is an important one and deserves careful investigation. We devoted a lot of thought to a potential switch from weight to quantity.
+%Running the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis yet appears really uneasy in our context for several, non-trivial reasons, that we develop below.
 
-In the original Hummels' database which we use from 1974 to 2004, we don't have information on quantity by transport mode.
+\subsubsection{Data issues}
+
+We do not know why that is, but the Census Bureau data do not report quantities by transport mode.
+Nor do the 1974-2004 Hummels’s data, which is based on the Census Bureau data. 
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
+As a result, to compute the unit price of all trade would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
+This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
+Hummels data are reported at the  sitc rev2 5-digit x  country level. In that context, 41.1 \% of trade flows and 79\% of trade value are multimode.
+Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail of large loss of useful information. 
+%data from "vérification quantities.do"
+Original census data, when accessible (normally from 1989, but actually only for 1997-1999 and from 2002 because of the Covid situation), are more promising.
+They are reported at the HS-10 digit x district of entry x district of unleading x rate provision code.
+In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+This is still a large share, but not too large to perclude any work. 
+So we decided to work only with data obtained directly from the Bureau of census.
+
+%This would be a non-trivial reduction.
+for Hummels).
 We could restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
@@
 
-Even on this very reduced sample, information on the \textit{units} considered for \textit{quantities} are not available for many years over our period, especially in the 1970s and the 1980s; discarding information on this would lead to blurred results as many types of units can be found within a same sector (dozens, liters, etc...) which are hardly comparable with each other.
+%Even on this very reduced sample
+Another difficulty is that our data, comming from Hummels or the Bureau of Census, include quantity, but not units.
+To the best of our knowledge, these can only be found in the "Harmonized Tariff Schedule of the United State".
+These schedules change many time per year.
+Schedules from 1987 can be found on the webside of the bureau (https://www.usitc.gov/tata/hts/archive/index.htm), but the ones from 1987 to 2008 are under pdf or html format.
+We could only find them in .csv format in a continuous manner from 2009.
+Even these files need cleaning up as the units are not always coherent (eg weight can be given in "kg" "kg." "Kg" and, also, "g" and "t").
+We did not try to transform the older pdf and html ones into csv, as that would have been too much work compared to our objectives.
+This is important, because, at our sector level (3-digit  sitc rev2 ), goods (either 5-digit sitc rev 2 or 10-digit hts) are measured in different units.
+In 2019, only 125 3-digit sitc sectors out of 230 are measured in a single unit (after cleaning up the unit names).
+For example, 3-digit sitc sector "061.
+Sugar, molasses and honey" includes among other products, 5-digit sitc "06150 Molasses, whether or not decolourized", measured in liters and 5-digit sitc "06160 Natural Honey" measured in kg.
+Actually worse than this, the 5-digit sitc product "06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel" includes both 10-digit hts product "1702903500 Other: invert molasses", measured in liters and 10-digit hts product " 1702602200 Blended syrups described in additional u.s.
+note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions"  measured in kilograms.
+3-digit sitc sector "291 Crude animal materials, n.e.s." includes 5-digit sitc products measured in kg ("29250 Seeds, fruit and spores, n.e.s., of a.
+kind used for sowing), in pieces ("29269 Other live plants (including trees, shrubs, bushes, roots, cuttings and slips)" )or measured in kg or not measured at all ("29230 Vegetable materials of a.
+kind used primarily for plaiting (e.g.
+cereal straw, cleaned, bleached or dyed, osier, reeds, rushes, rattans, bamboos, raffia and lime bark").
+
+So we are limited to using quantity units from 2009 to 2019.
+hese have to be found information on the \textit{units} considered for \textit{quantities} are not available for many years over our period, especially in the 1970s and the 1980s; discarding information on this would lead to blurred results as many types of units can be found within a same sector (dozens, liters, etc...) which are hardly comparable with each other.
 The option of not considering these observations would mean a drastic reduction in the estimation period that could be covered.

### 2021-10-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Discussion quantity units
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 %At a more fundamental level, this raises the following question: Do we have reason to believe our measure of unit prices generate significant deviations from the one relying on quantities?
-To that extent, the findings by \cite{Lashkaripour_JIE2020} were an eye-opener.
+To that extent, the findings in \cite{Lashkaripour_JIE2020} were an eye-opener.
 Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
@@
 We could only find them in .csv format in a continuous manner from 2009.
-Even these files need cleaning up as the units are not always coherent (eg weight can be given in "kg" "kg." "Kg" and, also, "g" and "t").
+Even these files need cleaning up as the units are not always coherent (eg weight can be given in ``kg’’, ``kg." ``Kg" and, also, ``g" and ``t").
 We did not try to transform the older pdf and html ones into csv, as that would have been too much work compared to our objectives.
@@
 In 2019, only 125 3-digit sitc sectors out of 230 are measured in a single unit (after cleaning up the unit names).
-For example, 3-digit sitc sector "061.
-Sugar, molasses and honey" includes among other products, 5-digit sitc "06150 Molasses, whether or not decolourized", measured in liters and 5-digit sitc "06160 Natural Honey" measured in kg.
-Actually worse than this, the 5-digit sitc product "06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel" includes both 10-digit hts product "1702903500 Other: invert molasses", measured in liters and 10-digit hts product " 1702602200 Blended syrups described in additional u.s.
+For example, 3-digit sitc sector ``061 Sugar, molasses and honey" includes among other products, 5-digit sitc product ``6150 Molasses, whether or not decolourized", measured in liters and 5-digit sitc product ``06160 Natural Honey" measured in kg.
+Actually worse than this, the 5-digit sitc product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel" includes both 10-digit hts product ````1702903500 Other: invert molasses", measured in liters and 10-digit hts product ``1702602200 Blended syrups described in additional u.s.
 note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions"  measured in kilograms.
-3-digit sitc sector "291 Crude animal materials, n.e.s." includes 5-digit sitc products measured in kg ("29250 Seeds, fruit and spores, n.e.s., of a.
-kind used for sowing), in pieces ("29269 Other live plants (including trees, shrubs, bushes, roots, cuttings and slips)" )or measured in kg or not measured at all ("29230 Vegetable materials of a.
-kind used primarily for plaiting (e.g.
+3-digit sitc sector ``291 Crude animal materials, n.e.s." includes 5-digit sitc products measured in kg (``29250 Seeds, fruit and spores, n.e.s., of a.
+kind used for sowing), in pieces (``29269 Other live plants (including trees, shrubs, bushes, roots, cuttings and slips)" )or measured in kg or not measured at all (``29230 Vegetable materials of a kind used primarily for plaiting (e.g.
 cereal straw, cleaned, bleached or dyed, osier, reeds, rushes, rattans, bamboos, raffia and lime bark").
+Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece" to be the same as the transport cost of a ``kg’’, even in the same sector.
+So we are limited to using quantity units from 2009 to 2019.
 
-So we are limited to using quantity units from 2009 to 2019.
-hese have to be found information on the \textit{units} considered for \textit{quantities} are not available for many years over our period, especially in the 1970s and the 1980s; discarding information on this would lead to blurred results as many types of units can be found within a same sector (dozens, liters, etc...) which are hardly comparable with each other.
+%
+%hese have to be found information on the \textit{units} considered for \textit{quantities} are not available for many years over our period, especially in the 1970s and the 1980s; discarding information on this would lead to blurred results as many types of units can be found within a same sector (dozens, liters, etc...) which are hardly comparable with each other.
 The option of not considering these observations would mean a drastic reduction in the estimation period that could be covered.
@@
 Je crois que je comprends, check with Guillaume }
-
-\textbf{SUPPRIMER CE QUI EST EN GRAS Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious} \textbf{XXX GUILLAUME: EXAMPLES XXX}.
-
-Considering the US census data which complete the original \cites{hummels2007} database, it is true that we can retrieve information on quantities.
+%
+%\textbf{SUPPRIMER CE QUI EST EN GRAS Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious} \textbf{XXX GUILLAUME: EXAMPLES XXX}.
+%
+%
+%Considering the US census data which complete the original \cites{hummels2007} database, it is true that we can retrieve information on quantities.
 Specifically, this can be obtained crossing HS-10 product information with the correspondence table between the HS-6 product classification and the associated units (kg, tons, liters, numbers, etc.) also provided by the US Census.
@@
 
-Getting deeper into the referee's suggestion, we were yet confronted to another issue.
+The same issue imposes the use of (sector,quantity unit) fixed effect rather than sector fixed effects
+But this is not enough to adapt our equation to unit prices rather than weight price.
 As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
-As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
-$t_{is(k)} = t_i+t_{s(k)}$ with $i$ the origin country and $s$ the sector.
-In the specific case of quantities, this raises a non-trivial issue yet.
-Specifically, imposing a country $i$ fixed effect would imply that the cost of sending one unit of say, tee-shirt from France is the same than sending one unit of say, car; and that this cost would be \textit{added} to the sector-fixed effect.
-\textbf{XXX PAS CLAIR IL FAUT MIEUX EXPLIQUER } This does not seem to us as a reasonable assumption; in this case specifically, skipping the separability assumption does appear as the appropriate choice.
+As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e. $t_{is(k)} = t_i+t_{s(k)}$ with $i$ the origin country and $s$ the sector.
+In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
+Sector-level fixed effects can be easily replaced by (sector,quantity unit) fixed effects.
+But the same is not possible for country fixed effects.
+In the per-weight price, the existence of a single country fixed effect applicable to all sector could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport from, say, Germany.
+But a unit of tee-shirt is quite different from an unit of car.
+Assume imports of tee-shirts and cars from Germany and China.
+Importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries).
+Idem for importing a car.
+Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same (in dollars) as difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
+This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
+As discussed before, that dramatically increases the computational difficulty of our method and yet does not yield different results.
 
@@
 We then decided to maintain our original empirical strategy covering the whole period 1974-2019 in the revised version, and to extend the online Appendix with this supplementary robustness check.
-We hope to convince you that this is a relevant choice.
-\textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
+We hope to convince you that this is a relevant choice. \textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
 

### 2021-10-29 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Nouveau graphique wgt / qy1
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
discussion qty-wgt dans online appendix

@@
 Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013.
-As noted above, the partners-sectors covered are lower.
-Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
-While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
-In any case yet, the share of additive costs remains substantial, around 40\% on average over the period, thereby confirming the main message of our paper shedding light on the importance of the additive component in international transport costs.
+Figure \ref{fig:beta_comp_direct_estimate} displays the estimated share of additive costs $\beta$ on a yearly basis over the estimation period, under both methods.
 
@@
 
+As noted above, the partners-sectors covered are lower.
+Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
+While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
+In any case yet, the share of additive costs remains substantial, around 40\% on average over the period, thereby confirming the main message of our paper shedding light on the importance of the additive component in international transport costs.
+
 \begin{figure}[htbp]
 	\caption{Baseline and direct $\beta$ estimate }
+\label{fig:beta_comp_direct_estimate}
 	\begin{center}
@@
 
+One may be concerned by the fact that unit prices are calculated as the value divided by the shipping
+weight, rather than to the quantity (the number) of goods.
+\cite{Lashkaripour_JIE2020} notably points out the importance of this measurement issue.
+Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
+\cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
+This result of transport costs close to be totally iceberg contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
+In this section, we then investigate the robustness of our results where unit values are built as the ratio between value and quantity of goods per observation.
+
+As we develop below, this is not without difficulties.
+The first one refers to the availability of information on quantities.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels' data (based on the Census Bureau data) do report the quantity by transport mode.
+This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
+As a result, to compute the unit price of all trade would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
+This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
+The alternative would then be to only consider single-mode flows.
+In Hummels' data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
+Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail of large loss of useful information.
+Original census data, when accessible (from 2002 because of the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+This is still a large share, but not too large to preclude any work.
+Accordingly, we dig deeper into this issue with the US Census dataset starting in 2002, restricted to single transport mode flows.
+\smallskip
+
+The second difficulty dwells on the units used to measure quantities.
+The Census data, if it provides some information on quantity, remains silent on the units considered.
+To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
+Unfortunately, these schedules change many time per year and only the schedules from 2009 onwards were available in a .csv format.
+On the years for which the match could be done (from 2009), it turns out that at our sector level (3-digit), goods (either at the 5-digit or the 10-digit level) are measured in different units.
+In 2019, only 125 3-digit sectors out of 230 are measured in a single unit (after cleaning up the unit names\footnote{As units are not always coherent (eg weight can be given in ``kg’’, ``kg.'' ``Kg'' and, also, ``g'' and ``t'').}).\footnote{For instance, 3-digit SITC sector ``061 Sugar, molasses and honey'' includes among other products, 5-digit SITC product ``6150 Molasses, whether or not decolourized'', measured in liters and 5-digit SITC product ``06160 Natural Honey'' measured in kg.
+Actually worse than this, the 5-digit SITC product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel'' includes both 10-digit HTS product ``1702903500 Other: invert molasses'', measured in liters and 10-digit HTS product ``1702602200 Blended syrups described in additional u.s.
+note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions''  measured in kilograms.} Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece'' to be the same as the transport cost of a ``kg'', even in the same sector.
+So we are limited to using quantity units from 2009 to 2019, amending our structure of fixed effects to consider the triplet (origin country, sector, quantity unit).
+
+Getting deeper on the estimated functional form raised a third difficulty in adapting our equation to unit prices rather than weight price.
+As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
+As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
+$t_{is(k)} = t_i+t_{s(k)}$ with $i$ the origin country and $s$ the sector.
+In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
+Sector-level fixed effects could indeed be easily replaced by (sector,quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Let us develop on that.
+
+In the per-weight price, the existence of a single country fixed effect applicable to all sector could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport from, say, Germany.
+In contrast, an unit of tee-shirt is quite different from an unit of car.
+To clarify this point, assume imports of tee-shirts and cars from Germany and China.
+Under the separability assumption, importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries); and similarly for importing a car.
+Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same (in dollars) as the difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
+This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
+As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper).\smallskip
+
+Integrating these various constraints, we run this robustness check (weight versus quantity) on a reduced sample over 2009-2019, considering single transport mode flows and retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ where $u$ refers to the unit considered to measure the quantity.
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{fig:beta_comp_wgt_qty}.
+In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
+
 \begin{table}[htbp]
@@
 	\end{center}
-	\label{tab_comp_referee1_baseline}%
+	\label{tab_comp_wgt_qty}%
 \end{table}%
@@
 	\caption{$\beta$ Estimate: Price per quantity or price per kg, 2005-2019}
+\label{fig:beta_comp_wgt_qty}
 	\begin{center}
@@
 
-\textbf{CHANGER LEGENDE PAS TRES BEAU}
+\textbf{CHANGER LEGENDE PAS TRES BEAU - ENLEVER unweigthed}
 
-J’AI CHANGÉ LA DÉFINITION DU TIME TREND. JE FAIS MAINTENANT LE TREND ENTRE LES BETA MOYENS PAR ANNÉES AU LIEU D’UNE REGRESSIONS SUR TOUS LES BETAS PONDÉRÉ PAR LEUR PART DANS LE COMMERCE. GD.
+Two main results emerge.
+First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
+In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
+In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
+Second, in terms of $\beta$ estimates, it is true that our method seems to yield a higher share of additive costs than when considering the price per quantity.
+This is somehow an expected result, as we do agree that there should be a link between the transport costs and the weight, or the volume of the goods transported.
+In all cases though, the share of additive costs remains substantial on average.
+In this respect, one might view these results as the lower bound for the size of additive costs in total transport costs.\smallskip
 
@@
 
-\textbf{CHANGER EXP/IMPORT VALUE POUR TRANSPORT COSTS}
+\textbf{XXX CHANGER legende pour transport costs, excluding composition effects; EXP/IMPORT VALUE = transport costs. On parle en \% de la valeur (y-axis) ou du prix fas (title)? XXX }
 

### 2021-10-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Encore un peu d’answers au referee1
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 Yet, first, looking at per-unit prices leads to leaving aside a large part of the data, both sectorally and chronologically.
-Second  it is interesting to look at additive costs on a per-weight basis even there is a correlation between unit price and weight.
+Second,  it is interesting to look at additive costs on a per-weight basis even there is a correlation between unit price and weight.
 
@@
 
-As a consequence, we run the estimation on a reduced sample (\textbf{specify the number of years, value to trade flows in comparison with the full sample}), according to the following equation:
+\subsubsection{Estimation}
+
+As a consequence, we run the following estimation on a reduced sample (2009-2019, 66\% of the value of the baseline sample)
 
@@
 
-Because the referee's concern yet seems important to address, we run a robustness check to the use of quantity versus weight.
-These results, obtained on a reduced sample given the above-mentioned constraints, have been added in the online appendix, Section D.4, Table D.2 and Figure D.4\textbf{XXX CHECK XXX }.
-For sake of reading clarity, we report here the results of Table D.2 (Table \ref{tab_comp_wgt_qty} here).
-For each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019).
+These results, in the online appendix, Section D.4, Table D.2 and Figure D.4\textbf{XXX CHECK XXX }.
+For sake of reading clarity, we report here the results of Table D.2 (Table \ref{tab_comp_wgt_qty} here and figure \ref{graph_comp_wgt_qty} ).
+For each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2009-2019).
 
 \begin{table}[htbp]
-	\caption{Comparison: Price per quantity versus per kg, 2005-2019}
+	\caption{Comparison: Price per quantity versus per kg, 2009-2019}
 	\begin{center}		
 		\input{../../revised_article/Online_Appendix/Comp_baseline_wgt_qy_reshaped.tex}
-		%\input{../../revised_article/Online_Appendix/Comp_baseline_referee1.tex}
-
+%		\input{../../revised_article/Online_Appendix/Comp_baseline_wgt_qy.tex}
 {\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Time trend coefficient is the annual growth rate.}}}
@@
 \end{table}%
+
+\begin{figure}[htbp]
+	\caption{$\beta$ Estimate: Price per quantity or price per kg, 2009-2019}
+	\begin{center}
+		\begin{tabular}{cc}
+			\includegraphics[height=8cm]{../../revised_article/scatter_chronology_non_separe_wgt_non_separe_qy.png}\\
+			\multicolumn{2}{l}{{\footnotesize Notes: In both cases, estimation is run without imposing the separability assumption.}}
+		\end{tabular}
+	\end{center}
+		\label{graph_comp_wgt_qty}%
+\end{figure}
+
+These results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
+They confirm the results of \citet{Lashkaripour_JIE2020} that there is a correlation between weight per unit and unit price.
+Still, the share of additive cost is not reduced to zero, except if one looks at the median for vessel).
+Thus these results do not disqualify our approach, especially as we cover a much larger period and sample.
+
+\subsubsection{Why is price per weight interesting}
+It is true that to study welfare, the relevant variable that would create distortions is the cost per unit. 
+However, our primary aim is descriptive. 
+From that point of view, transport cost schedule are expressed historically and currently in a mix per-weight and per-volume basis. 
+Fo example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight’’ or ``dimensional weight).
+See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).
+Maritime freight costs are expressed in volume.
+Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
+The existence of per-volume or per-weight shipping cost is a fact of (second) nature, directly linked to the technical constraints of the transportation sector.
+
+The notion of per-unit shipping cost is one step removed for technical constraints and the functionning of the transportation sector.
+The weight per unit is at least partly the result of a choice by firms.
 

### 2021-10-30 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Toujours du travail sur les biens dénombrables
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
     \item We can then evaluate the accuracy of each estimation method by comparing the size of the confidence intervals of the $\beta$, for each couple $i,s$ (by year and transport mode).
-We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.\textbf{XXX CETTE FIGURE, SUR QUELLE ANNEE, QUEL MODE DE TRANSPORT?
-xxxxx}
+We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.
 
@@
 
-The notion of per-unit shipping cost is one step removed for technical constraints and the functionning of the transportation sector.
+The notion of per-unit shipping cost is one step removed from technical constraints and the functionning of the transportation sector.
 The weight per unit is at least partly the result of a choice by firms. 
-
-Two main results emerge.
-First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
-In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
-In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
-Second, in terms of $\beta$ estimates, it is true that our method seems to yield a higher share of additive costs than when considering the price per quantity.
-This is somehow an expected result, as we do agree that there should be a link between the transport costs and the weight, or the volume of the goods transported.
-In all cases though, the share of additive costs remains substantial on average.
-In this respect, one might view these results as the lower bound for the size of additive costs in total transport costs.\smallskip
+To understand the context in which they operate, we find that studying the technical per-volume or per-weight shipping cost is interesting.
 

### 2021-10-31 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Fin de la discussion weight / quantity
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 The existence of per-volume or per-weight shipping cost is a fact of (second) nature, directly linked to the technical constraints of the transportation sector.
-
 The notion of per-unit shipping cost is one step removed from technical constraints and the functionning of the transportation sector.
@@
 
-These motives altogether highlight that it would be too costly in our context to switch from weights to quantities to compute unit prices for our baseline regression.
-We then decided to maintain our original empirical strategy covering the whole period 1974-2019 in the revised version, and to extend the online Appendix with this supplementary robustness check.
+Another argument is illustrated by graph \ref{graph_Share_of_discrete_goods.png}.
+After computing the share of discrete goods per 5-digit sitc import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports.
+According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, only over 50\% for two years.
+This justifies not switching the paper from the study of price per weight to the study of price per quantity.
+
+\begin{figure}[htbp]
+	\caption{Estimated share of discrete goods in US imports, 1974-2019}
+	\begin{center}
+		\begin{tabular}{cc}
+			\includegraphics[height=8cm]{../../revised_article/Share_of_discrete_goods.png}\\
+		\end{tabular}
+	\end{center}
+	\label{graph_Share_of_discrete_goods.png}%
+\end{figure}
+
+We have in consequence then decided to maintain our original empirical strategy covering the whole period 1974-2019 in the revised version, and to extend the online Appendix with this supplementary robustness check.
 We hope to convince you that this is a relevant choice. \textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
-
-At a more fundamental level, one may argue that considering the unit price as value/weight is relevant in our setting where we seek to identify transport costs.
-For instance, it makes sense that the shipment costs of cars do depend not only on the quantity of cars exported, but on the weight it makes, which is related to the volume it takes in the plane or the vessel.
-\textbf{XXX Guillaume : Do we have evidence on this point ?
-XXX}
 

### 2021-11-01 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Graphique E1
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

Et sans le bilf

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 \begin{table}[htbp]
-	\caption{Comparison Price per quantity or Price per kg, 2005-2019}
+	\caption{Comparison Price per quantity or Price per kg, 2009-2019}
 	\begin{center}		
@@
 This is somehow an expected result, as we do agree that there should be a link between the transport costs and the weight, or the volume of the goods transported.
-In all cases though, the share of additive costs remains substantial on average.
-In this respect, one might view these results as the lower bound for the size of additive costs in total transport costs.\smallskip
+In all cases though, the share of additive costs remains substantial (except if one looks at the median fro vessel).
+Thus these results do not disqualify our approach, especially as our baseline estimate covers a much larger period and trade flows.\smallskip
 
@@
 
-\subsubsection{More on the Hummels's methodology}
+\subsubsection{More on the Hummels' methodology}
 
@@
 {\small (A) Air } & {\small (B) Vessel }\\
-\multicolumn{2}{c}{{\small In percent of the export price}} \\
-\includegraphics[width=2.5in, height=1.8in]{figure5_comme_hummels.jpg}
-& \includegraphics[width=2.5in,height=1.8in]{figure6_comme_hummels.jpg} \\
+\multicolumn{2}{c}{{\small In percent of value shipped (i.e. the export price)}} \\
+\includegraphics[width=2.5in, height=1.8in]{../figure5_comme_hummels.jpg}
+& \includegraphics[width=2.5in,height=1.8in]{../figure6_comme_hummels.jpg} \\
 \end{tabular}
@@
 
-\textbf{XXX CHANGER legende pour transport costs, excluding composition effects; EXP/IMPORT VALUE = transport costs. On parle en \% de la valeur (y-axis) ou du prix fas (title)? XXX }
+%\textbf{XXX CHANGER legende pour transport costs, excluding composition effects; EXP/IMPORT VALUE = transport costs. On parle en \% de la valeur (y-axis) ou du prix fas (title)? XXX }
+%Fait GD
 

### 2021-11-01 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
graphs E2-E4
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 \begin{center}
-\includegraphics[height=8cm] {graph_composition_manuf.pdf}
+\includegraphics[height=8cm] {../graph_composition_manuf.jpg}
 \end{center}
@@
 \includegraphics[height=8cm]
-{graph_composition_primary.pdf}
+{../graph_composition_primary.jpg}
 \end{center}
@@
 \includegraphics[height=8cm]
-{graph_composition_all.pdf}
+{../graph_composition_all.jpg}
 \end{center}

### 2021-11-01 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
reprise argument Guillaume unit vs weight

@@
     \begin{center}
-    \includegraphics[height=4in]{../../revised_article/comparaison_amplitude_baseline_referee1_2013_ves.png"}
+    \includegraphics[height=4in]{../../revised_article/comparaison_amplitude_baseline_referee1_2013_ves.png}
     \begin{minipage} [c]  {5in} \scriptsize%
@@
 This result of transport costs close to be totally iceberg contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
-We believe the two sets of findings are complementary.
-
-We now report results based on per-unit prices.
-They confirm the results of \cite{Lashkaripour_JIE2020}
-Yet, first, looking at per-unit prices leads to leaving aside a large part of the data, both sectorally and chronologically.
-Second,  it is interesting to look at additive costs on a per-weight basis even there is a correlation between unit price and weight.
-
-%The two stories are not necessarily incompatible, however.
+Some elements can yet be put forward to account for the difference of results.
 Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
@@
 It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+
+Nevertheless, we did'nt want to stop at this very first argument.
+We indeed took the referee's remark very seriously and we did our best to address it to give you a convincing answer.
+It was not without difficulty yet, in particular due to data constraints.
+At a more fundamental level, we also believe that looking at transport costs on a per-weight basis do add relevant information.
+We develop each element of answer in what follows.
+
+%We now report results based on per-unit prices.
+%They confirm the results of \cite{Lashkaripour_JIE2020}
+%Yet, first, looking at per-unit prices leads to leaving aside a large part of the data, both sectorally and chronologically.
+%Second,  it is interesting to look at additive costs on a per-weight basis even there is a correlation between unit price and weight.
+
+%The two stories are not necessarily incompatible, however.
 
@@
 
-We do not know why that is, but the Census Bureau data do not report quantities by transport mode.
-Nor do the 1974-2004 Hummels’s data, which is based on the Census Bureau data. 
+\paragraph{Availability issue} In our attempt to address the referee's remark, we faced a first difficulty regarding the availability of information on quantities.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels' data (based on the Census Bureau data) do report the quantity by transport mode.
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
@@
 This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
-Hummels data are reported at the  sitc rev2 5-digit x  country level. In that context, 41.1 \% of trade flows and 79\% of trade value are multimode.
+The alternative would then be to only consider single-mode flows.
+In Hummels' data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
 Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail of large loss of useful information.
-%data from "vérification quantities.do"
-Original census data, when accessible (normally from 1989, but actually only for 1997-1999 and from 2002 because of the Covid situation), are more promising.
-They are reported at the HS-10 digit x district of entry x district of unleading x rate provision code.
-In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
-This is still a large share, but not too large to perclude any work. 
-So we decided to work only with data obtained directly from the Bureau of census.
+Original census data, when accessible (from 2002 because of the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+This is still a large share, but not too large to preclude any work.
+Accordingly, we dig deeper into this issue with the US Census dataset starting in 2002, restricted to single transport mode flows.
+\smallskip
 
-%This would be a non-trivial reduction.
-for Hummels).
-We could restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases.
-Investigating this possibility shows that this would reduce the total value of trade flows from around 32,000 billion dollars, to 7,000 billion dollars, i.e.
-a reduction of flows by a factor of 4.5.
-\textbf{XXX LP: Donne l'impression que ce n'est pas ce qu'on fait, alors que si, en fait?
-XXX}
+The second difficulty dwells on the units used to measure quantities.
+The Census data, if it provides some information on quantity, remains silent on the units considered.
+To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
+Unfortunately, these schedules change many time per year and only the schedules from 2009 onwards were available in a .csv format.
+On the years for which the match could be done (from 2009), it turns out that at our sector level (3-digit), goods (either at the 5-digit or the 10-digit level) are measured in different units.
+In 2019, only 125 3-digit sectors out of 230 are measured in a single unit (after cleaning up the unit names\footnote{As units are not always coherent (eg weight can be given in ``kg’’, ``kg.'' ``Kg'' and, also, ``g'' and ``t'').}).\footnote{For instance, 3-digit SITC sector ``061 Sugar, molasses and honey'' includes among other products, 5-digit SITC product ``6150 Molasses, whether or not decolourized'', measured in liters and 5-digit SITC product ``06160 Natural Honey'' measured in kg.
+Actually worse than this, the 5-digit SITC product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel'' includes both 10-digit HTS product ``1702903500 Other: invert molasses'', measured in liters and 10-digit HTS product ``1702602200 Blended syrups described in additional u.s.
+note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions''  measured in kilograms.} Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece'' to be the same as the transport cost of a ``kg'', even in the same sector.
+So we are limited to using quantity units from 2009 to 2019, amending our structure of fixed effects to consider the triplet (origin country, sector, quantity unit).
 
-%Even on this very reduced sample
-Another difficulty is that our data, comming from Hummels or the Bureau of Census, include quantity, but not units.
-To the best of our knowledge, these can only be found in the "Harmonized Tariff Schedule of the United State".
-These schedules change many time per year.
-Schedules from 1987 can be found on the webside of the bureau (https://www.usitc.gov/tata/hts/archive/index.htm), but the ones from 1987 to 2008 are under pdf or html format.
-We could only find them in .csv format in a continuous manner from 2009.
-Even these files need cleaning up as the units are not always coherent (eg weight can be given in ``kg’’, ``kg." ``Kg" and, also, ``g" and ``t").
-We did not try to transform the older pdf and html ones into csv, as that would have been too much work compared to our objectives.
-This is important, because, at our sector level (3-digit  sitc rev2 ), goods (either 5-digit sitc rev 2 or 10-digit hts) are measured in different units.
-In 2019, only 125 3-digit sitc sectors out of 230 are measured in a single unit (after cleaning up the unit names).
-For example, 3-digit sitc sector ``061 Sugar, molasses and honey" includes among other products, 5-digit sitc product ``6150 Molasses, whether or not decolourized", measured in liters and 5-digit sitc product ``06160 Natural Honey" measured in kg.
-Actually worse than this, the 5-digit sitc product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel" includes both 10-digit hts product ````1702903500 Other: invert molasses", measured in liters and 10-digit hts product ``1702602200 Blended syrups described in additional u.s.
-note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions"  measured in kilograms.
-3-digit sitc sector ``291 Crude animal materials, n.e.s." includes 5-digit sitc products measured in kg (``29250 Seeds, fruit and spores, n.e.s., of a.
-kind used for sowing), in pieces (``29269 Other live plants (including trees, shrubs, bushes, roots, cuttings and slips)" )or measured in kg or not measured at all (``29230 Vegetable materials of a kind used primarily for plaiting (e.g.
-cereal straw, cleaned, bleached or dyed, osier, reeds, rushes, rattans, bamboos, raffia and lime bark").
-Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece" to be the same as the transport cost of a ``kg’’, even in the same sector.
-So we are limited to using quantity units from 2009 to 2019.
+\paragraph{Estimation issue} Getting deeper on the estimated functional form raised a third difficulty in adapting our equation to per-unit rather than per-weight prices.
+As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
+As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
+$t_{is(k)} = t_i+t_{s(k)}$ with $i$ the origin country and $s$ the sector.
+In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
+Sector-level fixed effects could indeed be easily replaced by (sector,quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Let us develop on that.
 
-%
-%hese have to be found information on the \textit{units} considered for \textit{quantities} are not available for many years over our period, especially in the 1970s and the 1980s; discarding information on this would lead to blurred results as many types of units can be found within a same sector (dozens, liters, etc...) which are hardly comparable with each other.
-The option of not considering these observations would mean a drastic reduction in the estimation period that could be covered.
-Again, this is particularly problematic considering that the time length is crucial for the assessment of transport costs long-run dynamics, which is one of the main objectives of the paper.
-\textbf{XXXX LP: Je ne suis pas sure de cet argument.
-Veut-on dire que dans les "quantity", en fait on trouve de tout mais pas des unites?
-Je crois que je comprends, check with Guillaume }
-%
-%\textbf{SUPPRIMER CE QUI EST EN GRAS Time length is especially important considering that, in the 1970s, trade composition was significantly different from today, with more goods for which correlation between weight and quality is likely to be less obvious} \textbf{XXX GUILLAUME: EXAMPLES XXX}.
-%
-%
-%Considering the US census data which complete the original \cites{hummels2007} database, it is true that we can retrieve information on quantities.
-Specifically, this can be obtained crossing HS-10 product information with the correspondence table between the HS-6 product classification and the associated units (kg, tons, liters, numbers, etc.) also provided by the US Census.
-We are able to collect information since 2009.
-Investigating this database, it appears that in many 3-digit sectors, there are more than one type of quantities reported (for instance, sector \textbf{XXX} includes goods both recorded in terms of \textbf{YYY} and \textbf{ZZZ}.) To identify a proper unit price, we consequently reduce the flows to the 3-digit sectors reporting a single quantity \textbf{at the 3-digit level}.
+In the per-weight price, the existence of a single country fixed effect applicable to all sectors could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport from, say, Germany.
+In contrast, an unit of tee-shirt is quite different from an unit of car.
+To clarify this point, assume imports of tee-shirts and cars from Germany and China.
+Under the separability assumption, importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries); and similarly for importing a car.
+Hence, the difference in the cost of importing a car from Germany versus China would be the same (in dollars) as the difference between the cost of importing a tee-shirt from Germany relative to China.
+This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
+As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper).\smallskip
 
-The same issue imposes the use of (sector,quantity unit) fixed effect rather than sector fixed effects
-But this is not enough to adapt our equation to unit prices rather than weight price.
-As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
-As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e. $t_{is(k)} = t_i+t_{s(k)}$ with $i$ the origin country and $s$ the sector.
-In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects can be easily replaced by (sector,quantity unit) fixed effects.
-But the same is not possible for country fixed effects.
-In the per-weight price, the existence of a single country fixed effect applicable to all sector could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport from, say, Germany.
-But a unit of tee-shirt is quite different from an unit of car.
-Assume imports of tee-shirts and cars from Germany and China.
-Importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries).
-Idem for importing a car.
-Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same (in dollars) as difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
-This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
-As discussed before, that dramatically increases the computational difficulty of our method and yet does not yield different results.
+Integrating these various constraints, we run this robustness check (weight versus quantity) \textit{i)} on a sample reduced over 2009-2019, considering single transport mode flows; all in all, this sample amounts to 66\% of the value of the baseline sample, \textit{ii)} retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ where , according to the following equation:
+\begin{equation}
+\ln\left(\frac{p_{ik}}{p^u_{ik}}-1 \right)= \ln \left(\tau_{is(k)u} -1+\frac{t_{is(k)u} }{p^u_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
+\end{equation}
+\noindent where $p^u_{ik}$ refers to the unit price associated with quantity and $u$ to the unit considered to measure the quantity.
+Depending on the sector, two different units are considered, kilograms and numbers.
 
-\subsubsection{Estimation}
-
-As a consequence, we run the following estimation on a reduced sample (2009-2019, 66\% of the value of the baseline sample)
-
-\begin{equation}
-\ln\left(\frac{p_{ik}}{p^u_{ik}}-1 \right)= \ln \left(\tau_{is(k)} -1+\frac{t_{is(k)} }{p^u_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
-\end{equation}
-\noindent where $p^u_{ik}$ refers to the unit price associated with quantity. Depending on the sector, two different units are considered, kilograms and numbers.
-
-These results, in the online appendix, Section D.4, Table D.2 and Figure D.4\textbf{XXX CHECK XXX }.
-For sake of reading clarity, we report here the results of Table D.2 (Table \ref{tab_comp_wgt_qty} here and figure \ref{graph_comp_wgt_qty} ).
-For each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2009-2019).
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{fig:beta_comp_wgt_qty}.
+In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
 
@@
 These results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
-They confirm the results of \citet{Lashkaripour_JIE2020} that there is a correlation between weight per unit and unit price.
+In this respect, they stand in line with the results of \citet{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
 Still, the share of additive cost is not reduced to zero, except if one looks at the median for vessel).
-Thus these results do not disqualify our approach, especially as we cover a much larger period and sample.
+Thus these results do not disqualify our approach, especially as we cover a much larger period and sample with our baseline estimate.
 
@@
 It is true that to study welfare, the relevant variable that would create distortions is the cost per unit.
+\textbf{Je trouve l'argument delicat car dans notre section modele, c'est justement le welfare qu'on regarde.}
 However, our primary aim is descriptive.
-From that point of view, transport cost schedule are expressed historically and currently in a mix per-weight and per-volume basis. 
-Fo example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight’’ or ``dimensional weight).
+
+Our main objective in this paper is to provide a quantitative assessment of international transport costs over time; on that matter, one may start with the observation that transport cost schedules are expressed historically and currently in a mix per-weight and per-volume basis.
+\textbf{stop here}
+For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight’’ or ``dimensional weight).
 See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).

### 2021-11-01 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Graphique E5
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Graphique D1
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

--- 2021-11-01 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
@@
 \begin{center}
-\includegraphics[width=10cm, height=7cm]{variance_decomposition.pdf}
+\includegraphics[width=10cm, height=7cm]{../Décomposition de la variance à la mimine_ss_val_ext.png}
 %{\footnotesize {OECD data} }
@@
 \begin{figure}[htbp]
-	\caption{$\beta$ Estimate: Price per quantity or price per kg, 2005-2019}
+	\caption{$\beta$ Estimate: Price per quantity or price per kg, 2009-2019}
 \label{fig:beta_comp_wgt_qty}
@@
 
-\textbf{CHANGER LEGENDE PAS TRES BEAU - ENLEVER unweigthed}
+%\textbf{CHANGER LEGENDE PAS TRES BEAU - ENLEVER unweigthed}. Je crois que c’est bon.
 
@@
 \includegraphics[height=6cm]
-{Share_of_primary.pdf}
+{../Share_of_primary.jpg}
 \end{center}

### 2021-11-01 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement réponse référé 1 + mineurs online appendix

@@
 Nevertheless, we did'nt want to stop at this very first argument.
-We indeed took the referee's remark very seriously and we did our best to address it to give you a convincing answer.
+We indeed took the referee's remark very seriously and we investigated this question in depth.
 It was not without difficulty yet, in particular due to data constraints.
 At a more fundamental level, we also believe that looking at transport costs on a per-weight basis do add relevant information.
-We develop each element of answer in what follows.
+We develop each line of argument in what follows.
 
@@
 
-\paragraph{Availability issue} In our attempt to address the referee's remark, we faced a first difficulty regarding the availability of information on quantities.
+\paragraph{Data availability issue} In our attempt to address the referee's remark, we faced a first difficulty regarding the availability of information on quantities.
 Neither the US Bureau of Census data nor the 1974-2004 Hummels' data (based on the Census Bureau data) do report the quantity by transport mode.
@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could indeed be easily replaced by (sector,quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Sector-level fixed effects could indeed be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
 Let us develop on that.
@@
 Under the separability assumption, importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries); and similarly for importing a car.
-Hence, the difference in the cost of importing a car from Germany versus China would be the same (in dollars) as the difference between the cost of importing a tee-shirt from Germany relative to China.
+Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany relative to China.
 This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
@@
 
-Integrating these various constraints, we run this robustness check (weight versus quantity) \textit{i)} on a sample reduced over 2009-2019, considering single transport mode flows; all in all, this sample amounts to 66\% of the value of the baseline sample, \textit{ii)} retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ where , according to the following equation:
+Integrating these various constraints, we run this robustness check (weight versus quantity) \textit{i)} on a sample reduced over 2009-2019, considering single transport mode flows - amounting to 66\% of the value of the baseline sample; \textit{ii)} retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ where , according to the following equation:
 \begin{equation}
@@
 
-The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{fig:beta_comp_wgt_qty}.
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{fig:beta_comp_wgt_qty} (as well as in the online appendix D.4 \textbf{XXX Check in the end XXX}.
 In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
@@
 In this respect, they stand in line with the results of \citet{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
-Still, the share of additive cost is not reduced to zero, except if one looks at the median for vessel).
+Still, the share of additive cost is not reduced to zero (except if one looks at the median for vessel).
 Thus these results do not disqualify our approach, especially as we cover a much larger period and sample with our baseline estimate.
@@
 
-Our main objective in this paper is to provide a quantitative assessment of international transport costs over time; on that matter, one may start with the observation that transport cost schedules are expressed historically and currently in a mix per-weight and per-volume basis.
-\textbf{stop here}
-For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight’’ or ``dimensional weight).
-See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).
-Maritime freight costs are expressed in volume.
+\textbf{XXX LP: ma version plutot XXXX}
+At a more fundamental level, we do believe that the study of transport costs measured on a per-weight add some relevant information on the international transport cost burden faced by exporters.
+Let us explain the underlying rationale.
+
+Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mix per-weight and per-volume basis, both historically and nowadays.
+For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight'' or ``dimensional weight'')\footnote{See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).} while maritime freight costs are expressed in volume.
 Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
-The existence of per-volume or per-weight shipping cost is a fact of (second) nature, directly linked to the technical constraints of the transportation sector.
-The notion of per-unit shipping cost is one step removed from technical constraints and the functionning of the transportation sector.
-The weight per unit is at least partly the result of a choice by firms.
-To understand the context in which they operate, we find that studying the technical per-volume or per-weight shipping cost is interesting.
+The existence of per-volume or per-weight shipping cost hence appears as a fact of (second) nature, directly linked to the technical constraints of the transportation sector.
+In contrast, the notion of per-unit shipping cost is one step removed from technical constraints and the functioning of the transportation sector.
+While one may argue that the weight per unit is the result of the firm's choice, this can only be partly as firms in a given sector are very likely to have a limited set of options to alleviate the weight per unit of good; in this respect, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
 
 Another argument is illustrated by graph \ref{graph_Share_of_discrete_goods.png}.
-After computing the share of discrete goods per 5-digit sitc import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports.
+After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
 According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, only over 50\% for two years.
-This justifies not switching the paper from the study of price per weight to the study of price per quantity.
+Notice that this magnitude of order stands in line with \cite{Lashkaripour_JIE2020}.
+Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
 
@@
 
-We have in consequence then decided to maintain our original empirical strategy covering the whole period 1974-2019 in the revised version, and to extend the online Appendix with this supplementary robustness check.
-We hope to convince you that this is a relevant choice. \textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
+After considering all the above argument in depth, we decided not switching the paper from the study of price per weight to the study of price per quantity.
+We then maintained our original empirical strategy covering the whole period 1974-2019 in the revised version, while we extended the online Appendix with this supplementary robustness check.
+We hope to convince you that this is a relevant choice.
+\textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
 

### 2021-11-02 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
changement annexe online + réponse référé 1 unit vs weight

@@
 
-\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels' one.
+\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels's one.
 We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'' \textbf{XXX Check XX}).
@@
 
-Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
+Despite its interest, implementing the referee's method uncovers some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
 First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
@@
 \noindent from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).
-Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
+Then, relying on the Hummels's equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
 
@@
     \begin{center}
-    \includegraphics[height=4in]{../../revised_article/comparaison_amplitude_baseline_referee1_2013_ves.png}
+    \includegraphics[height=3in]{../../revised_article/comparaison_amplitude_baseline_referee1_2013_ves.png}
     \begin{minipage} [c]  {5in} \scriptsize%
@@
 
-   \textbf{XXXX J’ai laissé en plan le calcul sur d’autres modes et d’autres années à reprendre ?
-commit 16b215ca44c3f4cfb0fd43d3aaec6e6532aa7f9d  GD} In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
+   \textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
 Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
@@
 Some elements can yet be put forward to account for the difference of results.
-Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods representing a bit more 56\% of US total imports of goods.
+Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods that represent a bit more 56\% of US total imports of goods in his dataset.
 In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
@@
 
-Nevertheless, we did'nt want to stop at this very first argument.
+Nevertheless, we did not want to stop at this very first argument.
 We indeed took the referee's remark very seriously and we investigated this question in depth.
-It was not without difficulty yet, in particular due to data constraints.
+This was not without difficulty yet, in particular due to data and estimation constraints.
 At a more fundamental level, we also believe that looking at transport costs on a per-weight basis do add relevant information.
@@
 
-\subsubsection{Data issues}
+\subsubsection{Data availability issue}
 
-\paragraph{Data availability issue} In our attempt to address the referee's remark, we faced a first difficulty regarding the availability of information on quantities.
-Neither the US Bureau of Census data nor the 1974-2004 Hummels' data (based on the Census Bureau data) do report the quantity by transport mode.
+In our attempt to address the referee's remark, we faced a first difficulty regarding the availability of information on quantities.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) do report the quantity by transport mode.
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
@@
 The alternative would then be to only consider single-mode flows.
-In Hummels' data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
+In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
 Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail of large loss of useful information.
@@
 
-\paragraph{Estimation issue} Getting deeper on the estimated functional form raised a third difficulty in adapting our equation to per-unit rather than per-weight prices.
+\subsubsection{Estimation issue}
+
+Getting deeper on the estimated functional form raised another difficulty in adapting our equation to per-unit rather than per-weight prices.
 As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
@@
 This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
-As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper).\smallskip
+As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1.
+in the main paper \textbf{XXX check XXX}).\smallskip
 
-Integrating these various constraints, we run this robustness check (weight versus quantity) \textit{i)} on a sample reduced over 2009-2019, considering single transport mode flows - amounting to 66\% of the value of the baseline sample; \textit{ii)} retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ where , according to the following equation:
-\begin{equation}
-\ln\left(\frac{p_{ik}}{p^u_{ik}}-1 \right)= \ln \left(\tau_{is(k)u} -1+\frac{t_{is(k)u} }{p^u_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
-\end{equation}
-\noindent where $p^u_{ik}$ refers to the unit price associated with quantity and $u$ to the unit considered to measure the quantity.
+Integrating these various constraints, we run the robustness check (weight versus quantity) \textit{i)} on a sample reduced over 2009-2019, considering single transport mode flows - amounting to 66\% of the value of the baseline sample; \textit{ii)} retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ according to the following equation:
+\begin{equation*}
+\ln\left(\frac{p_{iku}}{\widetilde{p}_{iku}}-1 \right)= \ln \left(\tau_{is(k)u} -1+\frac{t_{is(k)u} }{\widetilde{p}_{iku}} \right) + \epsilon_{iku} 
+\end{equation*}
+\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$, $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
 Depending on the sector, two different units are considered, kilograms and numbers.
 
-The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{fig:beta_comp_wgt_qty} (as well as in the online appendix D.4 \textbf{XXX Check in the end XXX}.
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} (as well as in the online appendix D.4 \textbf{XXX Check in the end XXX}.
 In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
@@
 
-These results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
-In this respect, they stand in line with the results of \citet{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
+Two main results emerge.
+First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
+In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
+In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
+Second, in terms of $\beta$ estimates, the results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
+In this respect, they stand in line with the results of \cite{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
 Still, the share of additive cost is not reduced to zero (except if one looks at the median for vessel).
@@
 
-\subsubsection{Why is price per weight interesting}
-It is true that to study welfare, the relevant variable that would create distortions is the cost per unit.
-\textbf{Je trouve l'argument delicat car dans notre section modele, c'est justement le welfare qu'on regarde.}
-However, our primary aim is descriptive.
-
-\textbf{XXX LP: ma version plutot XXXX}
-At a more fundamental level, we do believe that the study of transport costs measured on a per-weight add some relevant information on the international transport cost burden faced by exporters.
+\subsubsection{On the relevance of the price per-weight approach}
+At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.
 Let us explain the underlying rationale.
@@
 Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
-The existence of per-volume or per-weight shipping cost hence appears as a fact of (second) nature, directly linked to the technical constraints of the transportation sector.
+The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector.
 In contrast, the notion of per-unit shipping cost is one step removed from technical constraints and the functioning of the transportation sector.
-While one may argue that the weight per unit is the result of the firm's choice, this can only be partly as firms in a given sector are very likely to have a limited set of options to alleviate the weight per unit of good; in this respect, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
+The shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
+Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
 
-Another argument is illustrated by graph \ref{graph_Share_of_discrete_goods.png}.
+Another argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png}.
 After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
-According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, only over 50\% for two years.
-Notice that this magnitude of order stands in line with \cite{Lashkaripour_JIE2020}.
+According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
 Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
@@
 
-After considering all the above argument in depth, we decided not switching the paper from the study of price per weight to the study of price per quantity.
-We then maintained our original empirical strategy covering the whole period 1974-2019 in the revised version, while we extended the online Appendix with this supplementary robustness check.
+After considering all the above arguments in depth, we decided to not switch the paper from the study of price per weight to the study of price per quantity.
+We thus maintained our original empirical strategy covering the whole period 1974-2019 in the revised version, while we extended the online Appendix with this supplementary robustness check.
 We hope to convince you that this is a relevant choice.
-\textbf{XXX CHECK RELECTURE TO BE COMPLETED XXX}
 
@@
 In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
-To that end, we adjust the share of exporting firms in the US: based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
+To that end, we adjust the share of exporting firms in the US: Based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
 In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/iceberg component, with the reduction in the fixed export cost.
@@
 
-Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+Overall, these results appear to add a substantial contribution to the paper : We are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
 Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.

### 2021-11-02 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
changement réponse référé 1 + mineurs online appendix
changement annexe online + réponse référé 1 unit vs weight

@@
 \begin{center}
-\includegraphics[width=10cm, height=7cm]{../Décomposition de la variance à la mimine_ss_val_ext.png}
+\includegraphics[width=10cm, height=7cm]{../Decomposition_de_la_variance_a_la_mimine_ss_val_ext.png}
 %{\footnotesize {OECD data} }
@@
 This result of transport costs close to be totally iceberg contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
-In this section, we then investigate the robustness of our results where unit values are built as the ratio between value and quantity of goods per observation.
+Some elements can yet be put forward to account for the difference of results.
+Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods that represent a bit more 56\% of US total imports of goods in his dataset.
+In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
+It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
 
-As we develop below, this is not without difficulties.
-The first one refers to the availability of information on quantities.
-Neither the US Bureau of Census data nor the 1974-2004 Hummels' data (based on the Census Bureau data) do report the quantity by transport mode.
+Nevertheless, it remains that the question deserves further investigation.
+In this section, we thus study the robustness of our results when unit values are built as the ratio between value and quantity of goods per observation.
+This was not without difficulty yet, in particular due to data and estimation constraints.
+At a more fundamental level, we also believe that looking at transport costs on a per-weight basis do add relevant information.
+
+\paragraph{Data availability issue} The first difficulty is tied to the availability of information on quantities.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) do report the quantity by transport mode.
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
@@
 note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions''  measured in kilograms.} Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece'' to be the same as the transport cost of a ``kg'', even in the same sector.
-So we are limited to using quantity units from 2009 to 2019, amending our structure of fixed effects to consider the triplet (origin country, sector, quantity unit).
+So we are limited to using quantity units from 2009 to 2019, amending our structure of fixed effects to consider the triplet (origin country, sector, quantity unit).\medskip
 
-Getting deeper on the estimated functional form raised a third difficulty in adapting our equation to unit prices rather than weight price.
+\paragraph{Estimation issue} Getting deeper on the estimated functional form raised another difficulty in adapting our equation to unit prices rather than weight price.
 As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
@@
 Under the separability assumption, importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries); and similarly for importing a car.
-Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same (in dollars) as the difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
+Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
 This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
-As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper).\smallskip
+As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1.
+in the main paper \textbf{XXX check XXX}).\smallskip
 
-Integrating these various constraints, we run this robustness check (weight versus quantity) on a reduced sample over 2009-2019, considering single transport mode flows and retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ where $u$ refers to the unit considered to measure the quantity.
+Integrating these various constraints, we run the robustness check (weight versus quantity) \textit{i)} on a sample reduced over 2009-2019, considering single transport mode flows - amounting to 66\% of the value of the baseline sample; \textit{ii)} retaining a non-separable structure of three-dimensional fixed effects $t_{is(k)u}$, $\tau_{is(k)u}$ according to the following equation:
+$$
+\ln\left(\frac{p_{iku}}{\widetilde{p}_{iku}}-1 \right)= \ln \left(\tau_{is(k)u} -1+\frac{t_{is(k)u} }{\widetilde{p}_{iku}} \right) + \epsilon_{iku}
+$$
+\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$, $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
+Depending on the sector, two different units are considered, kilograms and numbers.
+
 The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{fig:beta_comp_wgt_qty}.
@@
 In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
-Second, in terms of $\beta$ estimates, it is true that our method seems to yield a higher share of additive costs than when considering the price per quantity.
-This is somehow an expected result, as we do agree that there should be a link between the transport costs and the weight, or the volume of the goods transported.
-In all cases though, the share of additive costs remains substantial (except if one looks at the median fro vessel).
-Thus these results do not disqualify our approach, especially as our baseline estimate covers a much larger period and trade flows.\smallskip
+Second, in terms of $\beta$ estimates, the results clearly show that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
+In this respect, they stand in line with the results of \cite{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
+Still, the share of additive cost is not reduced to zero (except if one looks at the median for vessel).
+Thus these results do not disqualify our approach, especially as we cover a much larger period and sample with our baseline estimate.
+
+\paragraph{On the relevance of the price per-weight approach}
+At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.
+Let us explain the underlying rationale.
+
+Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mix per-weight and per-volume basis, both historically and nowadays.
+For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight'' or ``dimensional weight'')\footnote{See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).} while maritime freight costs are expressed in volume.
+Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
+The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector.
+In contrast, the notion of per-unit shipping cost is one step removed from technical constraints and the functioning of the transportation sector.
+The shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
+Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
+
+Another argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png}.
+After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
+According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
+Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
+
+\begin{figure}[htbp]
+	\caption{Estimated share of discrete goods in US imports, 1974-2019}
+	\begin{center}
+		\begin{tabular}{cc}
+			\includegraphics[height=8cm]{../../revised_article/Share_of_discrete_goods.png}\\
+		\end{tabular}
+	\end{center}
+	\label{graph_Share_of_discrete_goods.png}%
+\end{figure}
 
@@
 
+\bibliographystyle{apalike2}
+\bibliography{biblio}
+
 \end{document}

### 2021-11-10 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
relecture answer R1

@@
 \paragraph{Our answer:}
-\noindent This is a very important point we had not thought about, and we thank the referee for raising it.
-The referee definitely has a point here, in the sense that Census data do report the quantity of goods (and sometimes even two quantities).
+\noindent %This is a very important point, and we thank the referee for raising it.
+The referee definitely raised a very important point here, and indeed, Census data do report the quantity of goods (and sometimes even two quantities).
 %At a more fundamental level, this raises the following question: Do we have reason to believe our measure of unit prices generate significant deviations from the one relying on quantities?
-To that extent, the findings in \cite{Lashkaripour_JIE2020} were an eye-opener.
+In this context, the findings in \cite{Lashkaripour_JIE2020} are specifically enlightening.
 Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
@@
 In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
-It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs far from 100\%.
+It is therefore reasonable to think that the inclusion of all goods, like in our setting, would move the average share of multiplicative costs away from 100\%.
 
-Nevertheless, we did not want to stop at this very first argument.
-We indeed took the referee's remark very seriously and we investigated this question in depth.
+Nevertheless, considering the importance of the referee's remark, we went beyond this very first argument to investigate this question in depth.
 This was not without difficulty yet, in particular due to data and estimation constraints.
-At a more fundamental level, we also believe that looking at transport costs on a per-weight basis do add relevant information.
+At a more fundamental level, we also believe that looking at transport costs on a per-weight basis does add relevant information.
 We develop each line of argument in what follows.
@@
 This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
-As a result, to compute the unit price of all trade would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
+As a result, computing the unit price of all trade would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
 This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
@@
 In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
-Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail of large loss of useful information.
+This would reduce the total value of the sample from 12,500 billion dollars to 2,650 billion dollars, entailing a large loss of useful information.
 Original census data, when accessible (from 2002 because of the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
@@
 Sector-level fixed effects could indeed be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
-Let us develop on that.
+Let us elaborate on this issue.
 
@@
 Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany relative to China.
-This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
-As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1.
-in the main paper \textbf{XXX check XXX}).\smallskip
+This does not seem a sensible assumption; in this case specifically, we cannot keep the separability assumption.
+As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper).\smallskip
 
@@
 
-The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} (as well as in the online appendix D.4 \textbf{XXX Check in the end XXX}.
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} (as well as in the online appendix D.4.
 In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
@@
 Second, in terms of $\beta$ estimates, the results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
-In this respect, they stand in line with the results of \cite{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
+In this regard, they stand in line with the results of \cite{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
 Still, the share of additive cost is not reduced to zero (except if one looks at the median for vessel).
@@
 
-Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mix per-weight and per-volume basis, both historically and nowadays.
+Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mixed, per-weight and per-volume basis, both historically and nowadays.
 For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight'' or ``dimensional weight'')\footnote{See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).} while maritime freight costs are expressed in volume.
 Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
-The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector.
-In contrast, the notion of per-unit shipping cost is one step removed from technical constraints and the functioning of the transportation sector.
+The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions.
+In contrast, per-unit shipping cost depends on the behavior of exporting firms.
 The shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
@@
 
-After considering all the above arguments in depth, we decided to not switch the paper from the study of price per weight to the study of price per quantity.
+After considering all the above arguments in depth, we decided not to switch the paper from the study of price per weight to the study of price per quantity.
 We thus maintained our original empirical strategy covering the whole period 1974-2019 in the revised version, while we extended the online Appendix with this supplementary robustness check.
-We hope to convince you that this is a relevant choice.
+We hope to have convinced you that this is a relevant choice.
 

### 2021-11-10 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Requête
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
nouvelle tentative
Travail Guillaume - Lise sur le modèle
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 \begin{minipage} [c]  {5in} \scriptsize%
-Notes:The share of additive costs $\beta$ is reported on the $y-$axis.
+Note:The share of additive costs $\beta$ is reported on the $y-$axis.
 Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
@@
 In this section, we explore the implications of our empirical results on theoretical grounds.
-Over the last decade, several papers examined the implications of additive/per-shipment costs.
+Over the last decade, several papers examined the implications of non-iceberg costs.
 \cite{Kropf-Saure-JIE-2016} estimate the size and shape of per-shipment costs based on Swiss data.
 \cite{Alessandria-et-al-AER-2010} and \cite{Hornok-et-al-RES-2015} point out the role of per-shipment costs in generating ``lumpiness'' in international trade transactions.
-\cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment (i.e., additive by nature), and draw welfare conclusions regarding the variations of these costs.
-We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transports costs.
-Second, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
-To this aim, we embed additive costs in the canonical \citet{melitz} model.\smallskip
+\cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment, and draw welfare conclusions regarding the variations of these costs.
+We have complemented these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
+In this section, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two measures.
+One measure assumes that the reduction was achieved solely through the decline of ad-valorem costs, as modeled in standard New Trade theoretical approaches.
+The other measures builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
 %While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
@@
 
-\cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020} have preceded us on this issue.
-We differentiate from them in the following way.
+Our aim is related to \cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020}.
+We proceed in a different way.
 \cite{sorensen2014} explores the welfare effects of a trade costs reduction depending on whether they are either fully ad-valorem or fully additive on analytical grounds.
-In contrast, we provide a quantification of such welfare gains, as in \cite{Irrazabal_2015} or \cite{Lashkaripour_JIE2020}.
-In a partial-equilibrium model with perfect competition, \cite{Lashkaripour_JIE2020} obtains that welfare gains are not substantially different whether the transport cost decrease implies an additive component or not.
-This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz' type models.
-We also find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
-
-That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020}paper relies on a negligible share of additive costs.
+In contrast, as in \cite{Irrazabal_2015}, we provide a quantification of such welfare gains.
+\cite{Irrazabal_2015}, though, does not look into the effect of the reduction of trade costs it provides estimates of transport costs only for one year.
+We ground our study in the observed reduction of transport costs from 1974 to 2019.
+\cite{Lashkaripour_JIE2020} uses a partial-equilibrium model with perfect competition and builds on its result that for  denumerable goods, per-unit additive transport costs are very small to find a very small welfare effect of additive costs.
+We look at the whole range of goods over a longer period and use a \citet{melitz} model like \cite{sorensen2014} and \cite{Irrazabal_2015}.
+In that context,
+%\cite{Lashkaripour_JIE2020} obtains that welfare gains are not substantially different whether the transport cost decrease implies an additive component or not.
+This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz's type models.
+we find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
+
+%That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020}paper relies on a negligible share of additive costs.
 In this context, one might expect that whether additive costs are present or not makes little difference.
 As underlined by \cite{Lashkaripour_JIE2020}, a second explanation may rely on the modeling framework.
-While \cite{Lashkaripour_JIE2020} relies on a perfectly competitive environment, we use a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015}.
+While \cite{Lashkaripour_JIE2020} relies on a perfectly competitive environment,
+Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
 In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
-Put differently, leaving aside the selection issue leads to underestimate the welfare gains that can be achieved by a reduction in the additive component of transport costs.
+Leaving aside the selection issue leads to an underestimation of the welfare gains that can be achieved by a reduction in additive transport costs.
 
@@
 
-We rely on the model developed by \cite{sorensen2014}, which amends the two-country model of \cite{melitz} with monopolistic competition and endogenous firm entry to include additive costs.\footnote{\cite{Irrazabal_2015} adopt the multi-country model of \cite{chaney2008}, consistently with their database that documents the exports of Norwegian firms in various destination markets worldwide.
-We rather choose to retain the \cite{melitz} model, in its simplest form with one sector and two symmetric countries, picking up the average estimated values of transport costs based on our unique importer country, e.g.
-the US.
-Sticking to this canonical set-up allows us to illustrate the role of additive costs in a transparent and intuitive way.} Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
-For the sake of space, we only present here the building blocks of the model.
-More details can be found in Appendix \ref{app:theoretical_model}.
+We rely on the model developed by \cite{sorensen2014}, which amends the two-country model of \cite{melitz} with monopolistic competition and endogenous firm entry to include additive costs.\footnote{\cite{Irrazabal_2015} adopt the multi-country model of \cite{chaney2008}.
+%, consistently with their database that documents the exports of Norwegian firms in various destination markets worldwide.
+We choose to retain the \cite{melitz} model, in its simplest form with one sector and two symmetric countries.
+%, picking up the average estimated values of transport costs based on our unique importer country, e.g. the US.
+Sticking to this canonical set-up allows us to illustrate the role of additive costs in a transparent and intuitive way.}
+%Since our focus in on the role of additive versus ad-valorem transport costs, we directly present the two-country version of the model, with the two countries (Home and Foreign) being perfectly symmetric.
+For the sake of space, we  present here only the building blocks of the model. More details can be found in Appendix \ref{app:theoretical_model}.
 
@@
 In each country, the representative household exogenously supplies $L$ units of labour and maximizes utility $\mathcal{U} = \left[ \int_{\omega \in \Omega}c(\omega)^{\frac{\sigma-1}{\sigma}} d\omega \right]^{\frac{\sigma}{\sigma-1}}$, with $\sigma>1$ the elasticity of substitution between varieties $\omega$.
-She chooses her consumption $c(\omega)$  within the set $\Omega$ of available varieties (made of both domestic and foreign varieties, endogenously defined), subject to her budget constraint: $\int_{\omega \in \Omega} p(\omega) c(\omega) dw \leq E$, with $p(\omega)$ the consumer price of variety $\omega$ and $E$ nominal expenditures.
-As standard, the optimal demand function for variety $\omega$ is: $c(\omega) = \left(\frac{p(\omega)}{P}  \right)^{-\sigma} \frac{E}{P}$ with the associated optimal price index $P$ equal to $P = \left[ \int_{\omega \in \Omega}p(\omega)^{1-\sigma}d\omega\right]^{\frac{1}{1-\sigma}} $.
-
-\subsubsection{Firms behavior}
+She chooses her consumption $c(\omega)$  within the set $\Omega$ of available varieties (both domestic and foreign), subject to her budget constraint: $\int_{\omega \in \Omega} p(\omega) c(\omega) dw \leq E$, with $p(\omega)$ the consumer price of variety $\omega$ and $E$ nominal expenditures.
+The optimal demand function for variety $\omega$ is: $c(\omega) = \left(\frac{p(\omega)}{P}  \right)^{-\sigma} \frac{E}{P}$ with the associated optimal price index $P$ equal to $P = \left[ \int_{\omega \in \Omega}p(\omega)^{1-\sigma}d\omega\right]^{\frac{1}{1-\sigma}} $.
+
+\subsubsection{Firms}
 
 As in \cite{melitz}, firms are in monopolistic competition, each producing a differentiated variety using labor as sole production factor.
-Both countries being perfectly symmetric, the labor market equilibrium imposes the same wage in both countries, thus considered as the num\'{e}raire.
-Prices and costs are hence expressed in terms of labor.
+Both countries are perfectly symmetric, the labor market equilibria impose an unique international wage that we use as the num\'{e}raire.
 At entry, each firm pays a sunk entry cost $f_e$ and subsequently draws its productivity level $\varphi$ from a known distribution $G(\varphi)$.
-The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
-Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ (as in \citealp{melitz}) and an additive cost depending on the quantity (weight) transported, $t\geq 0$ - not included in the canonical Melitz model.
-At each point in time, a firm can be hit by a death shock at the exogenous probability $\delta>0$.
-Following \cite{melitz}, we solve the problem backwards.
-First, we solve for the maximizing profit behavior once entered.
-Second, we determine the entry conditions on both the domestic and the export markets.\medskip
-
-\paragraph{Profit-maximizing behaviors} Once entered, and conditional on being able to export, a firm adopts a pricing-to-market behavior by setting a specific price on each destination market.
-For each destination market, the firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the export market, according to:
+At each point in time, a firm can be terminated
+%be hit by a death shock
+at the exogenous probability $\delta>0$.
+The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$, units of labor, with $f$ the fixed cost of production.
+Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ (as in \citealp{melitz}) and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
+This $t$ is our innovation compared to the canonical Mélitz model.
+
+Following \cite{melitz}, we solve the program of firms backwards.
+First, we solve for the maximizing profit behavior after entry.
+Second, we determine the entry conditions on both domestic and exporting markets.\medskip
+\paragraph{Profit-maximizing behaviors} After entry, and conditional on being able to export, firms set a specific price on each destination market.
+The firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the destination market (see  Appendix \ref{app:theoretical_model}):
 \begin{eqnarray}
@@
 
-Under the assumption of symmetric countries (i.e, implying $E=E^\ast$ and $P=P^\ast$), one can deduce the profits on the domestic and export markets ($\pi_d(\varphi)$ and $\pi_x(\varphi)$) respectively as:
+Under the assumption of symmetric countries (implying $E=E^\ast$ and $P=P^\ast$), one can compute the profits on the domestic and export markets ($\pi_d(\varphi)$ and $\pi_x(\varphi)$) respectively as:
 \begin{eqnarray}
@@
 \end{eqnarray}
-\noindent with $B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ and where $r_d(\varphi)$ and $r_x(\varphi)$ denote the sales revenues on each market (see Appendix \ref{app:theoretical_model} for a detailed expression).
-Profits increase with productivity $\varphi$, and only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
-Further, in presence of trade costs, a given firm of productivity $\varphi$ makes a lower export profit than on its domestic market, conditional on being productive enough to export.
-This drives us to determine the productivity threshold associated to the entry conditions.
-
-\paragraph{Zero-cutoff profit conditions} Given the fixed costs on both the domestic and the export markets, only firms with a productivity above $\varphi^\ast$ will serve the domestic market, and those with $\varphi>\varphi_x^\ast$ will export abroad.
+\noindent $r_d(\varphi)$ and $r_x(\varphi)$ are the sales revenues on each market (see Appendix \ref{app:theoretical_model} for a detailed expression).
+$B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ .
+Profits increase with productivity $\varphi$.Only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
+Further, in presence of trade costs, profits for a given productivity are lower on the export market than on the domestic market.
+%a given firm of productivity $\varphi$ makes a lower export profit than on its domestic market, conditional on being productive enough to export.
+We now determine the productivity threshold associated to the entry conditions.
+
+\paragraph{Zero-cutoff profit conditions} Given the fixed costs on both the domestic and the export markets, only firms with a productivity above $\varphi^\ast$ will serve the domestic market, and only firms with a productivity above $\varphi_x^\ast$ will export.
 The two productivity thresholds can be linked according to:
 \begin{equation}
-\frac{\varphi_x^\ast}{\tau+t\varphi_x^\ast} = \varphi^\ast \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} \label{eq:link_thresholds}
-\end{equation}
-
-Following \cite{melitz} and the subsequent literature, we ensure selection on the export market, i.e.
-such that $\varphi_x^\ast > \varphi^\ast$.
-Making use of Equation (\ref{eq:link_thresholds}), this condition writes down as $\left[\tau+ t\varphi^\ast_x \right] \left( \frac{f}{f_x} \right)^{\frac{1}{1-\sigma}} >1$.
-Due to additive costs ($t>0$), the export selection condition depends on the general equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on the space parameter.
-We check this condition holds numerically when solving the model.\footnote{This is no longer the case under ad-valorem costs only.
-In this case, with $t=0$, we find the usual condition that $\varphi^\ast_x >\varphi^\ast$ iif $\tau^{\sigma-1}\frac{f_x}{f}>1$.}
+\frac{\varphi_x^\ast}{\tau+t\varphi_x^\ast} = \varphi^\ast \left( \frac{f_x}{f} \right)^{\frac{1}{\sigma-1}} \label{eq:link_thresholds}
+\end{equation}
+
+Following \cite{melitz} and the subsequent literature, we impose selection on the export market, (i.e.
+$\varphi_x^\ast > \varphi^\ast$).
+From Equation (\ref{eq:link_thresholds}), this condition writes down as $\left[\tau+ t\varphi^\ast_x \right] \left( \frac{f_x}{f} \right)^{\frac{1}{\sigma-1}} >1$.
+In contrast to the ad-valorem only case, due to additive costs ($t>0$), the export selection condition depends on the equilibrium of the model (through $\varphi^\ast_x $) and cannot be imposed ex-ante by an appropriate restriction on parameters.
+We check this condition holds numerically when solving the model.
+%\footnote{This is no longer the case under ad-valorem costs only. In this case, with $t=0$, we find the usual condition that $\varphi^\ast_x >\varphi^\ast$ iif $\tau^{\sigma-1}\frac{f_x}{f}>1$.}
 
@@
 
-\paragraph{Aggregation} The general equilibrium is characterized by a mass $M$ of producing firms in each country associated with a distribution $\mu(\varphi)$ of productivity levels of a subset of $\left(0,\infty\right)$.
-Given the productivity threshold $\varphi^\ast$, $\mu(\varphi)$ is equal to the conditional distribution of $g(\varphi)$ on the subset $\left[\varphi^\ast, \infty\right)$:
+\paragraph{Aggregation} The general equilibrium is characterized by a mass $M$ of producing firms in each country associated with a distribution $\mu(\varphi)$.
+Given the productivity threshold $\varphi^\ast$, $\mu(\varphi)$ is equal to:
+%the conditional distribution of $g(\varphi)$ on the subset $\left[\varphi^\ast, \infty\right)$:
 \begin{eqnarray}
@@
 on the support $\left[ \varphi^\ast_x,\infty\right)$.
-Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}):
+Further defining $M_T = M+M_x$ as the total mass of varieties accessible to the consumers, $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}), with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate:
 \begin{equation}
@@
 
-\paragraph{Free-entry condition} There is free-entry in the monopolistic industry, imposing that the expected net present value of entry is driven to zero.
+\paragraph{Free-entry condition} Assuming free entry, the expected net present value of entry is driven to zero.
 Defining $\bar{\pi}$ as the average profit conditional to entry, this free-entry condition rewrites as:
@@
 \end{equation}
-As shown in Appendix \ref{app:theoretical_model}, it can be rewritten as:
+As shown in Appendix \ref{app:theoretical_model}, the free-entry market condition can be rewritten as:
 \begin{equation}
@@
 
-Following a standard assumption in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the heterogeneity in firm productivity is assumed to be given by a Pareto distribution.
-As shown in \cite{melitz-redding-Handbk-IT-2014}, the Pareto distribution indeed allows for a full analytical solving of the M\'{e}litz model.
+Following a standard assumption in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the distribution in firm productivity is assumed to follow a Pareto distribution with parameters $(\varphi_{min},k)$.
+This allows a full analytical solving of the M\'{e}litz model.
 This is no longer the case in the presence of additive costs (see Appendix \ref{app:theoretical_model}).
@@
 This drives us to calibrate the structural parameters of the model.
-To do so, we dwell on values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
+To do so, we usevalues commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
 
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f$ endogenously revealed at the autarky case, such that all firms produce in autarky.
-$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2006's value), according to the references provided above.}}}
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
+$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s value) and the final steady state (39\%, based on 2006 value), according to the references provided above.}}}
 \end{table}
 
-The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC}.
+The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC} xxxGD pas idéal, je pense.
+À mon avis, je prendrais les points du début et de la fin selon le trendxxx
 
@@
 $t$ is calibrated so that the ratio $t/\widetilde{p}^{\text{fas}}_x$ from the model matches its empirical target, with $\widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
-We theoretically define the export fas price through: $p_x = \tau \widetilde{p}^{\text{fas}} +t$.
-TC = Transport Costs, expressed in percentage of the average fas export price.}}}
+We theoretically define the export fas price through: $p_x = \tau \widetilde{p}_x^{\text{fas}} +t$.
+TC = Transport Costs.}}}
 \end{table}
@@
 This can be captured in the model through a reduction in the fixed export cost $f_x$.
-Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.\footnote{We endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
-In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}, along with the 2019 values of variable transport costs}
+Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
+%\footnote{We endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
+In the final steady state, $f_x$ will be revealed such that this share of exporting plants has increased by 85\% (i.e., from 21\% to 39\%), based on the evidence reported in \cite{Lincoln_McCallum2018}, along with the 2019 values of variable transport costs xxxx GD redondant avec la note du tableau, non ?xxxx }
 
@@
 In running this exercise, we also take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported in \cite{BEJK-AER-03} (and consistent with the value estimated by \citet{Lincoln_McCallum2018} for 1987 on US data).
-According to \cite{Lincoln_McCallum2018}, this share roughly amounts to 39\% in the mid 2000's.
+According to \cite{Lincoln_McCallum2018}, this share roughly amounts to 39\% in the mid 2000s.
 This can be matched in the model by resetting the (lower) value of the export fixed cost $f_x$ accordingly.
@@
 This has two purposes.
-It allows us to better highlight the specific mechanisms involved by (a reduction in) additive costs versus the standard iceberg costs.
-In a related manner, it delivers a relevant robustness check on the normative implications of additive costs.
+It allows us to better highlight the specific mechanisms involved by a reduction in additive costs versus a reduction in ad-valorem costs.
+And it delivers a relevant robustness check on the welfare implications of additive costs.
 These results are reported in Table \ref{tab:resultats_modele}, panel (2).\footnote{Under a constant fixed export cost, we also compare two alternative cases.
-Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value, and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974.
-The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces.
+Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^{fas}$, maintaining the ad-valorem component at its initial (1974) value, and  \textit{(ii)} the opposite case, the observed reduction of ad-valorem costs ($\tau$), maintaining the additive component at its initial value.
+The quantitative results strongly confirm that, keeping cost level constant, additive costs are more detrimental for welfare than ad-valorem costs.
+%, and the larger welfare gains when this specific component of international trade costs is reduced.
 See Appendix \ref{app:theoretical_model}, Table \ref{tab:results_model_appendix} for the results}.
@@
 The reduction in transport costs is expressed in percentage points.
-In the world including additive costs, we simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.
+In Columns B and D, we first simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.
 }}}
@@
 
-In line with the literature, we find that trade liberalization induces welfare gains whatever the transport mode (and the associated magnitude of the transport costs decrease).
+In line with the literature, we find that a reduction of trade costs  induces welfare gains
+% whatever the transport mode (and the associated magnitude of the transport costs decrease).
 The originality of our results rather dwells on comparing the welfare gains depending on the type of cost reduction.
-On that side, the welfare gains are higher when the trade cost reduction occurs through a decrease in the additive component.
-This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though.
+Welfare gains are higher when the trade cost reduction includes a decrease in additive trade costs.
+This is consistent with the analytical results of \cite{sorensen2014} and the quantitative ones of \cite{Irrazabal_2015}.
+%in a different exercise of trade liberalization though.
 \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness.
 Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization.
-Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
-Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
+Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.}
+Quantitatively, the welfare gains are sizeable.
+In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 25\% percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+xxxGD ajuster avec les nouveaux chiffresxxx.
+%Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
 If we share a similar conclusion regarding the importance of additive costs, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
@@
 
-Through which mechanism does it occur?
+What are the mechanisms?
 Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
@@
 The optimal pricing behavior yields:
-$$p^d(\varphi) = \frac{\sigma}{\sigma-1}\frac{1}{\varphi}$$
+$$p_d(\varphi) = \frac{\sigma}{\sigma-1}\frac{1}{\varphi}$$
 With a CES specification on household's preferences, the firm sets its price as a constant mark-up on marginal cost, which is inversely related to its productivity as for the domestic market. \medskip

### 2021-11-10 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
réponse référé 2

@@
 
-\date{September 2021}
+\date{November 2021}
  \maketitle
@@
 Does it shed new light on a puzzle many people are thinking about?
-One crude suggestion is to see how the reduction in the industry-specific cost terms is related to the industry-level trade elasticities.
-If the composition effects favor low-elasticity industries, the findings in the paper may have first-order implications for the gains from trade.
-Another suggestion is to dig deeper into the relative rate at which additive and multiplicative transport costs have declined over time.
-Since additive transport costs favor rich (high-quality exporting) countries, the disproportionally greater reduction in additive costs can perhaps explain the rise of low-income exporter as documented by Hanson (2012, JEP).}''
+[One] suggestion is to dig deeper into the relative rate at which additive and multiplicative transport costs have declined over time.}
+''
 
-All in all, your suggestions and those of the other referees pointed in two directions for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the implications regarding grains from trade.
+All in all, your suggestion and the other referee's one pointed in two directions for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the implications regarding gains from trade.
 Both paths were undeniably worth exploring, but could not be followed in a single paper.
@@
 We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transports costs.
-Furthermore we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
+Based on these empirical results, we investigate the welfare consequences of the reduction in international transport costs as observed over 1974-2019.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the disproportionate part of additive transport costs in determining those welfare variations.\smallskip
 
@@
 However, \citet{sorensen2014} exclusively performs a theoretical analysis, without any quantitative exercise.
-In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the presence of additive costs, but the latter is based on a calibration for transport costs restricted to 2004, based on the case of Nowrway.
-In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on ``true'' values for the latter - remember that our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the (median) export price.
+In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the reduction of additive costs (vs iceberg costs).
+For this, they rely on a calibration for transport costs based on a single-year estimation (for the year 2004), from which they depart by assuming an ad-hoc hypothetical reduction of trade costs.
+In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on the values actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).\smallskip
 
-More precisely, we use some of the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, more precisely those for the years 1974 and 2019.
-Based on the latter, we implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
-In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
+Specifically, we use the estimation results reported in Section 3 concerning multiplicative and additive costs, corresponding to the values at the beginning and at the end of the sample.
+We investigate the different welfare consequences of alterations to multiplicative and additive costs, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
+In addition, we also assess how the results are modified by changes in sunk costs of exports, $f_{x}$.
 To that end, we adjust the share of exporting firms in the US: based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
-In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/iceberg component, with the reduction in the fixed export cost.
+In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/iceberg component, along with the reduction in the fixed export cost.
 We compare this with the case where the total transport costs reduction is solely attributed to a decrease in ad-valorem costs.

### 2021-11-10 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex

@@
 
-Nevertheless, it remains that the question deserves further investigation.
-In this section, we thus study the robustness of our results when unit values are built as the ratio between value and quantity of goods per observation.
+Nevertheless, this section investigates the robustness of our results when unit values are built as the ratio between value and quantity of goods per observation.
 This was not without difficulty yet, in particular due to data and estimation constraints.
-At a more fundamental level, we also believe that looking at transport costs on a per-weight basis do add relevant information.
+At a more fundamental level, we also believe that looking at transport costs on a per-weight basis does add relevant information.
 
@@
 Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) do report the quantity by transport mode.
-This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
-As a result, to compute the unit price of all trade would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
+This is incompatible with our empirical strategy for estimating transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
+As a result, to compute the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
 This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
@@
 In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
-Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail of large loss of useful information.
+Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail a large loss of useful information.
 Original census data, when accessible (from 2002 because of the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
@@
 On the years for which the match could be done (from 2009), it turns out that at our sector level (3-digit), goods (either at the 5-digit or the 10-digit level) are measured in different units.
-In 2019, only 125 3-digit sectors out of 230 are measured in a single unit (after cleaning up the unit names\footnote{As units are not always coherent (eg weight can be given in ``kg’’, ``kg.'' ``Kg'' and, also, ``g'' and ``t'').}).\footnote{For instance, 3-digit SITC sector ``061 Sugar, molasses and honey'' includes among other products, 5-digit SITC product ``6150 Molasses, whether or not decolourized'', measured in liters and 5-digit SITC product ``06160 Natural Honey'' measured in kg.
+In 2019, only 125 3-digit sectors out of 230 are measured in a single unit (after cleaning up the unit names\footnote{As units are not always coherent (e.g.
+weight can be given in ``kg’’, ``kg.'' ``Kg'' and, also, ``g'' and ``t'').}).\footnote{For instance, 3-digit SITC sector ``061 Sugar, molasses and honey'' includes among other products, 5-digit SITC product ``6150 Molasses, whether or not decolourized'', measured in liters and 5-digit SITC product ``06160 Natural Honey'' measured in kg.
 Actually worse than this, the 5-digit SITC product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel'' includes both 10-digit HTS product ``1702903500 Other: invert molasses'', measured in liters and 10-digit HTS product ``1702602200 Blended syrups described in additional u.s.
@@
 Sector-level fixed effects could indeed be easily replaced by (sector,quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
-Let us develop on that.
+Let us elaborate on this issue.
 
@@
 Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
-This does not seem to us as a reasonable assumption; in this case specifically, we cannot keep the separability assumption.
-As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1.
-in the main paper \textbf{XXX check XXX}).\smallskip
+This does not seem a sensible assumption; in this case specifically, we cannot keep the separability assumption.
+As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper.\smallskip
 
@@
 In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
-In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
+This raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
 Second, in terms of $\beta$ estimates, the results clearly show that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
@@
 At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.
-Let us explain the underlying rationale.
+We detail below the underlying rationale.
 
@@
 Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
-The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector.
-In contrast, the notion of per-unit shipping cost is one step removed from technical constraints and the functioning of the transportation sector.
-The shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
+The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions - in contrast, per-unit shipping cost depends on the behavior of exporting firms.
+Put differently, the shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
 Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.

### 2021-11-12 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Révision légère figure 1 et 2. Sortie des valeurs ...
Révision légère figure 1 et 2. Sortie des valeurs prédites total TC et beta

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Guillaume sur le texte (surtout partie théorique)
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices).
-If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade liberalization.
+If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction.
 Our main contribution in this paper is to bring new insights on this matter.\smallskip
@@
 
-On the theoretical side, we show the importance of embedding additive costs when assessing the effects of trade liberalization in normative terms.
+On the theoretical side, we show the importance of embedding additive costs when assessing the effects of trade cost reduction in welfare terms.
 Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction solely occurs through a decrease in the ad-valorem cost, as standardly modelled in New Trade theories, versus when it is achieved through a combined reduction in the ad-valorem and the additive components, as quantified by our empirical model.
@@
 They have important implications for both international trade and international macroeconomics models.
-Some recent papers notably point out the normative implications of additive trade costs.
+Some recent papers notably point out the welfare implications of additive trade costs.
 \cite{sorensen2014} analytically shows that important welfare gains can be achieved by reducing additive trade costs.
 Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs, the one exception being \citealp{Irrazabal_2015}.
-In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade liberalization since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs, as well as by sub-component (additive and ad-valorem).\smallskip
+In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade cost reduction since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs, as well as by sub-component (additive and ad-valorem).\smallskip
 
@@
 Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the levels of both the iceberg costs and of the additive costs, in contrast to e.g.
+This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the levels of both the ad-valorem costs and of the additive costs, in contrast to e.g.
 \cite{Irrazabal_2015}.
@@
 For each year and transport mode, we estimate two models.
-In Model (A), additive costs are excluded, transport costs being modeled as iceberg costs only.
+In Model (A), additive costs are excluded, transport costs being modeled as ad-valorem costs only.
 In this case, the estimated equation simplifies as:
@@
 After estimating Equation (\ref{eq:model_IetA}), we can rebuild a measure of each component, $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode.
-When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is(k)} = \widehat{\tau}_{i} \times \widehat{\tau}_{s(k)}$.
+When assuming ad-valorem costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is(k)} = \widehat{\tau}_{i} \times \widehat{\tau}_{s(k)}$.
 In this case, notice that the equation could be estimated relying on a linear form.
@@
 
+%%Sur le trend :      +------------------------------+
+%| trend_~t         mode   year |
+%|------------------------------|
+%1. | 3.652468   (b) Vessel   2019 |
+%46. | 8.546226   (b) Vessel   1974 |
+%47. | 2.488636      (a) Air   2019 |
+%92. | 6.289694      (a) Air   1974 |
+%+------------------------------+
+
 As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -1.9\% for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
@@
 \end{figure}
+
+%Le beta sur le trend quadratic (  reg ln_meanshareA year year2 )
+%
+%   +------------------------------+
+%| trend_~a         mode   year |
+%|------------------------------|
+%1. |  .438216   (b) Vessel   2019 |
+%46. | .3468959   (b) Vessel   1974 |
+%47. | .2301836      (a) Air   2019 |
+%92. | .3839365      (a) Air   1974 |
+%+------------------------------+
+%))
 
@@
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
-We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and iceberg) in a unified measure of transport costs.
+We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and ad-valorem) in a unified measure of transport costs.
 The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$, and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
@@
 In this section, we explore the implications of our empirical results on theoretical grounds.
-Over the last decade, several papers examined the implications of non-iceberg costs.
+Over the last decade, several papers examined the implications of non-ad-valorem costs.
 \cite{Kropf-Saure-JIE-2016} estimate the size and shape of per-shipment costs based on Swiss data.
@@
 The other measures builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
-%While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
+%While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., ad-valorem trade costs), not much is said about the role of the additive component.
 One potential explanation is the lack of empirical estimates about the size of additive costs.
@@
 Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
-In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade cost reduction lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
 Leaving aside the selection issue leads to an underestimation of the welfare gains that can be achieved by a reduction in additive transport costs.
@@
 
-We are interested in studying the normative consequences of trade liberalization in presence or not of additive costs.
+We are interested in studying the welfare consequences of trade cost reduction in presence or not of additive costs.
 To that aim, we will compare the equilibrium based on the values of transport costs at the beginning (1974) and at the very end of our estimation period (2019), in two seetings: one including only ad-valorem transport costs, and including both additive and ad-valorem costs.
@@
 This can be captured in the model through a reduction in the fixed export cost $f_x$.
-Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
+Accordingly, we will assess the effects of the trade cost reduction assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
 %\footnote{We endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
@@
 
-\subsection{Trade liberalization: The importance of additive costs}
+\subsection{Trade cost reduction: The importance of additive costs}
 
@@
 We exploit the theoretical implications of these empirical findings by running a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
-Specifically, we contrast two cases: In the first scenario so-called \textit{à la Mélitz}, we assume that only ad-valorem costs apply.
+Specifically, we contrast two cases: In the first scenario, we assume that only ad-valorem costs apply.
 We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
-In the second scenario so-called \textit{à la Hummels}, we rely on the estimates of transport costs involving both additive and ad-valorem components.
+In the second scenario, we rely on the estimates of transport costs involving both additive and ad-valorem components.
 The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
@@
 This can be matched in the model by resetting the (lower) value of the export fixed cost $f_x$ accordingly.
-Our preferred scenario thus embeds the two dimensions of trade liberalization: the reduction of transport costs along with the reduction of the export fixed cost.
-
-The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
+Our preferred scenario thus embeds the two dimensions of trade cost reduction: the reduction of transport costs along with the reduction of the export fixed cost.
+
+The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade cost reduction, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
 The whole set of results is available upon request to the authors.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
@@
   \centering
-  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele}
+  \caption{Trade cost reduction: The role of additive costs} \label{tab:resultats_modele}
 \begin{center}
@@
 
+\textbf{Inverser panel 1 et 2 pour avoir le même ordre dans le tableau et le texte xxxxGD}
 In line with the literature, we find that a reduction of trade costs  induces welfare gains
@@
 This is consistent with the analytical results of \cite{sorensen2014} and the quantitative ones of \cite{Irrazabal_2015}.
-%in a different exercise of trade liberalization though.
+%in a different exercise of trade cost reduction though.
 \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness.
-Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization.
+Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade cost reduction.
 Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.}
 Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 25\% percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 25\% percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
 xxxGD ajuster avec les nouveaux chiffresxxx.
@@
 What are the mechanisms?
-Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
+Recalling that welfare is the inverse of the price index, the explanation can be understood by decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
 Considering Equation (\ref{eq:Welfare}), it is straightforward to get the welfare change as:
@@
 
-To understand the changes induced by the additive costs, we start considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (2)).
-In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from the fact that the transport cost decrease involves an additive part or not.
-
-Consider first the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (2), Columns (A) and (C)).
-The reduction in the iceberg cost induces a negative variety effect, which is yet
-more than counteracted by a positive competitive effect. With the reduction in $\tau$, more firms can now export, but this also increases
-the expected average profit. This raises the productivity threshold to enter the domestic
-market. Accordingly, the number of domestic active firms reduces, even if a larger number
-of them is able to export. As a whole, the total number of varieties accessible to the
-consumer reduces, which reduces welfare everything else equal.
-This detrimental effect is counteracted by a positive price competitiveness effect. Due
-to the tougher selection on the domestic market, the less productive firms no longer enter.
-As a result, the average productivity on the domestic market increases, which pushes
-the average productivity upwards. As reported in Table \ref{tab:resultats_modele}, the strength of the price
-competitiveness effect dominates, inducing positive welfare gains with the reduction in
-iceberg costs, consistently with the analytical results of \cite{melitz}.
-
-What does it change when the reduction in international transport costs also occurs
-through a reduction in its additive component?
-As reported in Table \ref{tab:resultats_modele}, panel (2), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case; and this is driven by a more positive price competitiveness effect and despite the more negative price effect (comparing with Columns (A) and (C), same panel).
+\paragraph{Holding fixed costs constant} To understand the changes induced by the additive costs, we start by considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (2)).
+In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from whether transport cost involves an additive part or not.
+
+Consider first the benchmark model \textit{\`{a} la} M\'{e}litz with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (2), Columns (A) and (C)).
+The reduction in the ad-valorem cost induces a negative variety effect, which is yet more than counteracted by a positive price effect.
+With the reduction in $\tau$, the expected average export profit increases and more firms can now export.
+More exporting firms means more competition in each domestic market.
+A well known result is that this raises the productivity threshold to enter the domestic markets.
+Accordingly, the number of domestic active firms is reduced, even if a larger number of them is able to export.
+As a whole, the total number of varieties accessible to the consumer is reduced in some of the cases we study here.
+This reduces welfare everything else equal.
+This detrimental effect is counteracted by a positive price competitiveness effect.
+Due to the tougher selection on the domestic market, the less productive firms no longer enter.
+As a result, the average productivity on the domestic market increases, which pushes the average productivity upwards.
+As reported in Table \ref{tab:resultats_modele}, the strength of the price competitiveness effect dominates, inducing positive welfare from the reduction in ad-valorem costs, consistently with the analytical results of \cite{melitz}.
+
+The situation changes when the reduction in international transport costs also occurs through the reduction in an additive component.
+As reported in Table \ref{tab:resultats_modele}, panel (2), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case; and this is driven by a more positive price competitiveness effect and despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
 This can be accounted for as follows.
 
-As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity
-exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
-Within the exporters category, this favors the more productive firms, whose market share,
-e.g. revenues, increases relatively more (see Equation (\ref{eq:rx})). Everything else equal, this
-pushes the average productivity on the export market upwards. Unlike the ad-valorem
-costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of
-the more productive firms. At the general equilibrium, this \textit{within-exporters} selection effect
-also affects the \textit{classic} selection effect at the extensive margin at the heart of the M\'{e}litz
-(2003) model. As the expected export profit increases, it is necessary to be more productive
-to satisfy the free-entry condition. Everything else equal, the threshold productivity
-on the domestic market rises, as well as the average productivity. Accordingly, the
-price competitiveness effect increases by more when the trade liberalization involves a reduction
-in the additive component.
-
-The welfare-dampening impact of the variety effect is also stronger under the scenario \textit{à la Hummels}.
-Since the reduction in additive costs favors the more productive firms in terms of market share,
-entry on the export market is tougher leading to a much more modest increase in the
-share of exporters. As the selection on the domestic market is also more stringent, the
-mass of total active firms reduces by a larger amount. From the consumer side, this induces
-a more negative variety effect that is welfare-detrimental everything else equal. As for the
-pure-iceberg case though, the strength of the (larger) competitiveness effect dominates,
-inducing positive welfare gains, and even larger when the trade liberalization involves a
-reduction in additive costs.\medskip
-
-As mentioned above, our preferred scenario dwells on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
+As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
+Within exporters, this favors the more productive firms, whose market share, e.g. revenues, increases relatively more (see Equation (\ref{eq:rx})).
+Everything else equal, this pushes the average productivity of exporters upwards.
+Unlike the ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms.
+At the general equilibrium, this \textit{within-exporters} selection effect also affects the \textit{classic} selection effect at the extensive margin at the heart of the M\'{e}litz (2003) model.
+Exporting firms are now more competitive. Accordingly, this raises the threshold productivity on the domestic market rises, as well as the average productivity.
+Accordingly, the price competitiveness effect is larger when trade cost reduction involves a reduction in the additive component.
+
+The welfare-dampening impact of the variety effect is also stronger in this scenario.
+Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the share of exporters.
+As the selection on the domestic market is also more stringent, the mass of total active firms reduces by a larger amount.
+From the consumer side, this induces a more negative variety effect.
+As for the pure-ad-valorem case though, the strength of the (larger) competitiveness effect dominates, inducing positive welfare gains, and even larger ones than in the pure-ad-valorem case.\medskip
+
+\paragraph{Taking into account the reduction in fixed costs}As mentioned above, our preferred scenario includes, in addition to the reduction of variable costs, a reduction in fixed export costs, based on \cite{Lincoln_McCallum2018}’s empirical evidence.
 The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (1).
 Two results emerge.
-First, the welfare gains induced by the trade liberalization are stronger in magnitude.
-While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1.25 percentage point depending on the transport mode.
+First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
+While this is not a surprising result \textit{per se} (we now consider two sources of internation frictions: fixed and variable costs)), this exercise delivers also a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial and amount to between 0.6 to 1.25 percentage point depending on the transport mode.
+\textbf{passer en \% GD xxxxx}
 This can be accounted for as follows.
-The reduction in the fixed export cost pushes more firms to export everything else equal.
+The reduction in the fixed export cost pushes more firms to export.
 This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market.
-As a result, the joined reduction in variable costs and in $f_x$ lead to a less negative variety effect (which even turns out to be positive in the scenario \textit{à la M\'{e}litz} with only iceberg costs).
+As a result, the joined reduction in variable costs and in $f_x$ lead to a less negative variety effect (which even turns out to be positive in the scenario with only ad-valorem costs).
 This, combined with a positive price competitiveness effect, accounts for the stronger welfare gains reported in Panel (2).
+ \textbf{panel 1 ou 2 ? Voir plus haut GD xxxxx}.
 
 Second, the main message of the paper remains valid.
-The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (around + 1.2 percentage points).
-Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
-In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
-
-In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results point out the importance of taking into the additive components of international trade costs.
-Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods.
-This, in turn, has important consequences in normative terms.
-Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.16\% increase in welfare, much higher than if only iceberg costs were modeled (2.5\%).
-A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4.3\% (as compared to 3.6\% in a pure ad-valorem world).
-Neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
-While we have obtained these results in the canonical \cite{melitz} model, they call for integrating additive costs in richer New Trade models in view of better assessing the disputed question of the welfare gains of trade liberalization.
+The welfare gains of the trade cost reduction are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (around + 1.2 percentage points) \textbf{passer en \% GD xxxxx}.
+Put differently, the balance between the effect of the reduction of additive costs and the reduction of ad-valorem costs does does not seem to depend on reduction of fixed costs.
+In the end, the welfare gains are the highest when trade cost reduction induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
+
+In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results show the importance of taking into account the additive components of international trade costs.
+Additive costs have a substantial impact effect on the composition of the baskets of exported goods.
+This, in turn, has important consequences for welfare.
+Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.16\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.5\%).
+\textbf{passer en \% GD xxxxx}.
+We find a similar result in vessel transport with the observed 6.4\% in total variable costs along with the reduction in the fixed export costs.
+The relative increase in welfare is 4.3\% (as compared to 3.6\% in a pure ad-valorem world).
+Neglecting the existence and long-term evolution additive costs thus leads to a substantial underestimation of the welfare gains of trade cost reduction.
 
@@
 	\item Statistics are weighted by value
-	\item \textbf{Model (A): Iceberg transport costs only}
+	\item \textbf{Model (A): ad-valorem transport costs only}
 	\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 		\item Statistics are weighted by value
-		\item \textbf{Model (A): Iceberg transport costs only}
+		\item \textbf{Model (A): Ad-valorem transport costs only}
 		\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 	\item SER are weighted by value
-	\item \textbf{Model (A): Iceberg transport costs only}
+	\item \textbf{Model (A): Ad-valorem transport costs only}
 	\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 			\item SER are weighted by value
-			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (A): Ad-valorem transport costs only}
 			\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (A)), the product dimension (Term (B)) and the pure transport costs time trend (Term (c)).
-Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
+Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the ad-valorem cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
 Both equations are estimated by transport mode.\smallskip

### 2021-11-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement lise partie theorique

@@
 \begin{tabular}{l||c|c||c|c}
-\hline
+\hline \hline
 Mode & \multicolumn{2}{|c||}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
@@
 %$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
-$\Delta TC$ (in pp) &-3.4 &-3.4 &-6.4 &-6.4   \\
-$\Delta \tau$ (in pp) & -3.4	&-1.4&-6.4&	-3.1	 \\
-$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-3.3 \\ \hline \hline
-\multicolumn{4}{l}{(1) Lower fixed export cost} \\ \hline
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.50}	&\textbf{3.16}	&	\textbf{3.63}&	\textbf{4.26 }  \\\hline
-Variety effect (\%)&1.36 &0.48	&-0.43	&-1.28 \\
-Price effect (\%)&	1.14 &	2.66&		 4.08&	5.63 \\ \hline \hline
-\multicolumn{4}{l}{(2) Constant fixed export cost} \\ \hline
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.18 & 2.39	&3.38 	 \\ \hline
-Variety effect (\%)& -0.83 	&-1.32	&-2.60	& -3.12 \\
-Price effect (\%)&2.08 &3.56	&5.20	& 6.83  \\
-\hline
+$\Delta TC$ (in pp) &-3.73 &-3.73 &-4.9 &-4.9   \\
+$\Delta \tau$ (in pp) & -3.73	&-1.95&-4.9&	-3.53	 \\
+$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-1.84&	0&	-1.36 \\ \hline \hline
+\multicolumn{4}{l}{(1) Constant fixed export cost} \\ \hline
+$\Delta \beta$ (in pp) & 0&	-15.4 & 0	&+9.3 	 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.38&	2.46 & 1.71	&2.20 	 \\ \hline
+Variety effect (\%)& -0.92 	&-2.57	&-1.09	& -1.87 \\
+Price effect (\%)&2.34 &5.24	&2.85	& 4.19  \\ \hline
+\multicolumn{4}{l}{(2) Lower fixed export cost} \\ \hline
+$\Delta \beta$ (in pp) & 0&	-15.2 & 0	&+9.2 	 \\
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.62}	&\textbf{3.66}	&	\textbf{2.83}&	\textbf{3.23 }  \\\hline
+Variety effect (\%)&1.20 &-0.45	&0.92	&0.23 \\
+Price effect (\%)&	1.42 &	4.14&		1.90&	3.00 \\ \hline \hline
 \end{tabular}
@@
 
-\textbf{Inverser panel 1 et 2 pour avoir le même ordre dans le tableau et le texte xxxxGD}
 In line with the literature, we find that a reduction of trade costs  induces welfare gains
@@
 Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 25\% percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
-xxxGD ajuster avec les nouveaux chiffresxxx.
+In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 40\% (14\%) in air (vessel) transport when such a reduction partly occurs through the additive component, rather than when it realizes through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
+One can also notice that the welfare gains are much larger when the trade cost reduction is achieved through a larger reduction in the additive component (such that $\beta$ reduces, as for air transport, Columns (B) vs (A)) than when it mostly occurs through the ad-valorem component (in which case $\beta$ increases, as for vessel transport, Columns (D) vs (C)).
 %Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
@@
 
-\paragraph{Holding fixed costs constant} To understand the changes induced by the additive costs, we start by considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (2)).
+\paragraph{Holding fixed costs constant} To understand the changes induced by the additive costs, we start by considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (1)).
 In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from whether transport cost involves an additive part or not.
 
-Consider first the benchmark model \textit{\`{a} la} M\'{e}litz with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (2), Columns (A) and (C)).
+Consider first the benchmark model \textit{\`{a} la} M\'{e}litz with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (1), Columns (A) and (C)).
 The reduction in the ad-valorem cost induces a negative variety effect, which is yet more than counteracted by a positive price effect.
@@
 The situation changes when the reduction in international transport costs also occurs through the reduction in an additive component.
-As reported in Table \ref{tab:resultats_modele}, panel (2), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case; and this is driven by a more positive price competitiveness effect and despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
+As reported in Table \ref{tab:resultats_modele}, Panel (1), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case; and this is driven by a more positive price competitiveness effect and despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
 This can be accounted for as follows.
@@
 
-\paragraph{Taking into account the reduction in fixed costs}As mentioned above, our preferred scenario includes, in addition to the reduction of variable costs, a reduction in fixed export costs, based on \cite{Lincoln_McCallum2018}’s empirical evidence.
-The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (1).
+\paragraph{Taking into account the reduction in fixed costs} As mentioned above, our preferred scenario includes, in addition to the reduction of variable costs, a reduction in fixed export costs, based on \cite{Lincoln_McCallum2018}’s empirical evidence.
+The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (2).
 Two results emerge.
 First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
-While this is not a surprising result \textit{per se} (we now consider two sources of internation frictions: fixed and variable costs)), this exercise delivers also a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial and amount to between 0.6 to 1.25 percentage point depending on the transport mode.
-\textbf{passer en \% GD xxxxx}
-This can be accounted for as follows.
+While this is not a surprising result \textit{per se} (we now consider two sources of international frictions: fixed and variable costs)), this exercise delivers also a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by inducing welfare gains higher by around 60-80\% depending on the transport mode.\footnote{For instance, consider the case of Air transport (Column (A)).
+The supplementary welfare gains associated with the reduction in $f_x$ is obtained as the ratio 2.62/1.38, i.e.
+89\%.
+\textbf{xxx check calculus}} This can be accounted for as follows.
 The reduction in the fixed export cost pushes more firms to export.
 This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market.
-As a result, the joined reduction in variable costs and in $f_x$ lead to a less negative variety effect (which even turns out to be positive in the scenario with only ad-valorem costs).
+As a result, the joined reduction in variable costs and in $f_x$ lead to a less negative variety effect (which may even turn out to be positive, e.g.
+in vessel transport).
 This, combined with a positive price competitiveness effect, accounts for the stronger welfare gains reported in Panel (2).
- \textbf{panel 1 ou 2 ? Voir plus haut GD xxxxx}.
 
 Second, the main message of the paper remains valid.
-The welfare gains of the trade cost reduction are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (around + 1.2 percentage points) \textbf{passer en \% GD xxxxx}.
-Put differently, the balance between the effect of the reduction of additive costs and the reduction of ad-valorem costs does does not seem to depend on reduction of fixed costs.
+The welfare gains of the trade cost reduction are still higher when part of this structural change is achieved through a reduction in additive transport costs, even if lower in magnitude (around + 50\%).
+Put differently, the balance between the effect of the reduction of additive costs and the reduction of ad-valorem costs does not seem to depend on reduction of fixed costs.
 In the end, the welfare gains are the highest when trade cost reduction induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
@@
 This, in turn, has important consequences for welfare.
-Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.16\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.5\%).
-\textbf{passer en \% GD xxxxx}.
-We find a similar result in vessel transport with the observed 6.4\% in total variable costs along with the reduction in the fixed export costs.
-The relative increase in welfare is 4.3\% (as compared to 3.6\% in a pure ad-valorem world).
+Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.7\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.6\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.6\%).
+We find a similar result in vessel transport with the observed 4.9\% in total variable costs along with the reduction in the fixed export costs.
+The relative increase in welfare is 3.2\% (as compared to 2.8\% in a pure ad-valorem world).
 Neglecting the existence and long-term evolution additive costs thus leads to a substantial underestimation of the welfare gains of trade cost reduction.

### 2021-11-13 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
-The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions.
-In contrast, per-unit shipping cost depends on the behavior of exporting firms.
-The shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
+The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions - in contrast, per-unit shipping cost depends on the behavior of exporting firms.
+Put differently, the shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
 Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.

### 2021-11-14 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte

--- 2021-11-10 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 \cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment, and draw welfare conclusions regarding the variations of these costs.
-We have complemented these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
-In this section, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two measures.
-One measure assumes that the reduction was achieved solely through the decline of ad-valorem costs, as modeled in standard New Trade theoretical approaches.
-The other measures builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
+Our results so far complement these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
+In this section, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases.
+One case assumes that the reduction was achieved solely through the decline of ad-valorem costs, as modeled in standard New Trade theoretical approaches.
+The other measure builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
 %While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
@@
 Our aim is related to \cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020}.
-We proceed in a different way.
+We proceed in a different way though.
 \cite{sorensen2014} explores the welfare effects of a trade costs reduction depending on whether they are either fully ad-valorem or fully additive on analytical grounds.
 In contrast, as in \cite{Irrazabal_2015}, we provide a quantification of such welfare gains.
-\cite{Irrazabal_2015}, though, does not look into the effect of the reduction of trade costs it provides estimates of transport costs only for one year.
+As \cite{Irrazabal_2015} provide estimates of transport costs only for one year (2004), their assessment of the welfare effects of a trade cost increase is based on a hypothetical assumption.
 We ground our study in the observed reduction of transport costs from 1974 to 2019.
@@
 This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz's type models.
-we find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
+we find that the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
 
@@
 While \cite{Lashkaripour_JIE2020} relies on a perfectly competitive environment,
-Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
-In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+Using a model with monopolistic competition and endogenous firm selection, as \cite{sorensen2014} and \cite{Irrazabal_2015}, does matter in this result.
+Indeed, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
 Leaving aside the selection issue leads to an underestimation of the welfare gains that can be achieved by a reduction in additive transport costs.
@@
 This drives us to calibrate the structural parameters of the model.
-To do so, we usevalues commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
+To do so, we use values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
 
@@
 
-The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC} xxxGD pas idéal, je pense.
-À mon avis, je prendrais les points du début et de la fin selon le trendxxx
+The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC}.
 
@@
 We theoretically define the export fas price through: $p_x = \tau \widetilde{p}_x^{\text{fas}} +t$.
-TC = Transport Costs.}}}
+TC = Transport Costs.
+The empirical targets for 1974 and 2019 are based on the estimated linear trend over the period.}}}
 \end{table}
@@
 We are interested in studying the normative consequences of trade liberalization in presence or not of additive costs.
-To that aim, we will compare the equilibrium based on the values of transport costs at the beginning (1974) and at the very end of our estimation period (2019), in two seetings: one including only ad-valorem transport costs, and including both additive and ad-valorem costs.
+To that aim, we will compare the equilibrium based on the values of transport costs at the beginning (1974) and at the very end of our estimation period (2019), in two settings: the first one including only ad-valorem transport costs (uneven columns in Table \ref{tab:calib_TC}).
+We here apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
+The second scenario includes both additive and ad-valorem costs (even columns in Table \ref{tab:calib_TC}).
+In this case, we rely on the estimates of transport costs involving both additive and ad-valorem components.
+The reduction in international transport costs channels through both its additive and ad-valorem components.
 Besides, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
-This can be captured in the model through a reduction in the fixed export cost $f_x$.
+This can be captured in the model through a reduction in the fixed export cost $f_x$ (as reported in Table \ref{tab:calib_horsTC}).
 Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
@@
 We exploit the theoretical implications of these empirical findings by running a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
-Specifically, we contrast two cases: In the first scenario so-called \textit{à la Mélitz}, we assume that only ad-valorem costs apply.
-We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
-In the second scenario so-called \textit{à la Hummels}, we rely on the estimates of transport costs involving both additive and ad-valorem components.
-The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
-
 In running this exercise, we also take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported in \cite{BEJK-AER-03} (and consistent with the value estimated by \citet{Lincoln_McCallum2018} for 1987 on US data).
@@
 
-The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
+The results are reported in Table \ref{tab:resultats_modele}, panel (2), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
 The whole set of results is available upon request to the authors.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
@@
 It allows us to better highlight the specific mechanisms involved by a reduction in additive costs versus a reduction in ad-valorem costs.
-And it delivers a relevant robustness check on the welfare implications of additive costs.
-These results are reported in Table \ref{tab:resultats_modele}, panel (2).\footnote{Under a constant fixed export cost, we also compare two alternative cases.
+It also delivers a relevant robustness check on the welfare implications of additive costs.
+These results are reported in Table \ref{tab:resultats_modele}, panel (1).\footnote{Under a constant fixed export cost, we also compare two alternative cases.
 Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^{fas}$, maintaining the ad-valorem component at its initial (1974) value, and  \textit{(ii)} the opposite case, the observed reduction of ad-valorem costs ($\tau$), maintaining the additive component at its initial value.
@@
 %$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
-$\Delta TC$ (in pp) &-3.4 &-3.4 &-6.4 &-6.4   \\
-$\Delta \tau$ (in pp) & -3.4	&-1.4&-6.4&	-3.1	 \\
-$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-3.3 \\ \hline \hline
-\multicolumn{4}{l}{(1) Lower fixed export cost} \\ \hline
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.50}	&\textbf{3.16}	&	\textbf{3.63}&	\textbf{4.26 }  \\\hline
-Variety effect (\%)&1.36 &0.48	&-0.43	&-1.28 \\
-Price effect (\%)&	1.14 &	2.66&		 4.08&	5.63 \\ \hline \hline
-\multicolumn{4}{l}{(2) Constant fixed export cost} \\ \hline
-$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.18 & 2.39	&3.38 	 \\ \hline
-Variety effect (\%)& -0.83 	&-1.32	&-2.60	& -3.12 \\
-Price effect (\%)&2.08 &3.56	&5.20	& 6.83  \\
+$\Delta TC$ (in pp) &-3.8 &-3.8 &-4.9 &-4.9   \\
+$\Delta \tau$ (in pp) & -3.8	&-1.95&-4.9&	-3.53	 \\
+$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-1.8&	0&	-1.4\\ \hline 
+\multicolumn{4}{l}{(1) Constant fixed export cost} \\ \hline
+$\Delta \beta$ (in pp) &0&-15.4&0&9.3 \\ 
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.38&	2.46 & 1.71	&2.20 	 \\ 
+\hspace{1em} Variety effect (\%)& -0.93 	&-2.57	&-1.09	& -1.88 \\
+\hspace{1em} Price effect (\%)&2.34 &5.24	&2.85	& 4.19  \\
 \hline 
+\multicolumn{4}{l}{(2) Lower fixed export cost} \\ \hline
+$\Delta \beta$ (in pp) &0&-15.15&0&9.16 \\ 
+$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.62}	&\textbf{3.66}	&	\textbf{2.83}&	\textbf{3.23 }  \\
+\hspace{1em} Variety effect (\%)&1.19 &-0.45	&0.92	&0.23 \\
+\hspace{1em} Price effect (\%)&	1.42 &	4.14&		 1.90&	3.0 \\ \hline \hline
 \end{tabular}
@@
 The reduction in transport costs is expressed in percentage points.
-In Columns B and D, we first simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.
+In Columns (B) and (D), we first simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.
 }}}
@@
 Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction in variable transport costs generates a relative welfare gain higher by around 25\% percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel).
-xxxGD ajuster avec les nouveaux chiffresxxx.
+In our preferred scenario, the reduction in variable transport costs in Air transport generates a relative welfare gain 40\% higher when such a reduction partly occurs through the additive component, rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)).
+In maritime transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
+This can be set in connection with the fact that in maritime transport, ad-valorem costs have reduced more than the additive ones, implying a larger $\beta$ at the end of the period.\footnote{See Appendix \ref{app:theoretical_model} for a similar result.}
 %Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
@@
 
-We report the decomposition of the welfare gains in their two components in Table \ref{tab:resultats_modele}, under our preferred scenario (Panel (1)) and under a constant fixed export cost (Panel (2)).
-
-To understand the changes induced by the additive costs, we start considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (2)).
-In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from the fact that the transport cost decrease involves an additive part or not.
-
-Consider first the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (2), Columns (A) and (C)).
+We report the decomposition of the welfare gains in their two components in Table \ref{tab:resultats_modele}, under a constant fixed export cost (Panel (1)) and under our preferred scenario (Panel (2)).
+\medskip
+
+\paragraph{Under a constant fixed export cost} In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from the fact that the transport cost decrease involves an additive part or not.
+
+Consider first the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (1), Columns (A) and (C)).
 The reduction in the iceberg cost induces a negative variety effect, which is yet
-more than counteracted by a positive competitive effect. With the reduction in $\tau$, more firms can now export, but this also increases
-the expected average profit. This raises the productivity threshold to enter the domestic
+more than counteracted by a positive competitive effect.
+This is accounted for as follows.
+With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit.
+This raises the productivity threshold to enter the domestic
 market. Accordingly, the number of domestic active firms reduces, even if a larger number
@@
 through a reduction in its additive component?
-As reported in Table \ref{tab:resultats_modele}, panel (2), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case; and this is driven by a more positive price competitiveness effect and despite the more negative price effect (comparing with Columns (A) and (C), same panel).
+As reported in Table \ref{tab:resultats_modele}, panel (1), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case.
+Again, the effect is quantitatively sizeable: Welfare gains are thus between 30\% and 78\% higher, depending on the transport mode.
+This is driven by a more positive price competitiveness effect and despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
 This can be accounted for as follows.
 
-As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity
+As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productive
 exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
@@
 
-The welfare-dampening impact of the variety effect is also stronger under the scenario \textit{à la Hummels}.
-Since the reduction in additive costs favors the more productive firms in terms of market share,
-entry on the export market is tougher leading to a much more modest increase in the
+The welfare-dampening impact of the variety effect is also stronger under this scenario.
+Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the
 share of exporters. As the selection on the domestic market is also more stringent, the
@@
 
-As mentioned above, our preferred scenario dwells on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
-The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (1).
+\paragraph{Reduction in both variable and fixed trade costs} As mentioned above, our preferred scenario dwells on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
+The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (2).
 Two results emerge.
-First, the welfare gains induced by the trade liberalization are stronger in magnitude.
-While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1.25 percentage point depending on the transport mode.
-This can be accounted for as follows.
+First, the welfare gains induced by the trade costs reduction are stronger in magnitude.
+While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, higher by around 50\% relative to the case when only the two components of the variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively.
+Put differently, the reduction of the sunk export cost accounts for the majority of welfare gains.
+This is not the case under additive variable costs; in this case, the reduction of the distortion due to the additive costs continues to substantially matter in the welfare-enhancing effect of the globalization process.}  This can be accounted for as follows.
 The reduction in the fixed export cost pushes more firms to export everything else equal.
 This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market.
-As a result, the joined reduction in variable costs and in $f_x$ lead to a less negative variety effect (which even turns out to be positive in the scenario \textit{à la M\'{e}litz} with only iceberg costs).
+As a result, the joined reduction in variable costs and in $f_x$ dampens the negativity of the variety effect (which even turns out to be positive in some cases).
 This, combined with a positive price competitiveness effect, accounts for the stronger welfare gains reported in Panel (2).
@@
 Second, the main message of the paper remains valid.
-The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (around + 1.2 percentage points).
+The welfare gains of the trade cost reduction are still higher when part of this structural change is achieved through a reduction in additive transport costs.
+Quantitatively, the effect remains sizeable: Welfare gains are between 14\% and 40\% higher when the globalization process involves a reduction in additive costs.
 Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
-In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
+In the end, the welfare gains are the highest when trade costs reduction induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
 
@@
 This, in turn, has important consequences in normative terms.
-Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.16\% increase in welfare, much higher than if only iceberg costs were modeled (2.5\%).
-A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4.3\% (as compared to 3.6\% in a pure ad-valorem world).
+Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 60\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.66\% increase in welfare, much higher than if only iceberg costs were modeled (2.6\%).
+A similar result applies with the observed 57\% reduction in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world).
 Neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
@@
 Relying on our empirical estimates of transport
-costs over 1974-2019, our preferred scenario shows an extra welfare gain by around 1.2
-percentage point when the transport cost decrease encompasses a reduction in its additive
+costs over 1974-2019, our preferred scenario shows an extra welfare gain higher by 14\% and 40\% respectively in vessel and air transport when the transport cost decrease encompasses a reduction in its additive
 component (as we document empirically) relative to a pure ad-valorem reduction.
@@
   \centering
-  \caption{Exploring alternative scenario: Calibration }\label{tab:calib_TC_appendix}
+  \caption{Exploring alternative scenarios: Calibration }\label{tab:calib_TC_appendix}
 \begin{center}
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $t$ is calibrated so that the ratio $t/\widetilde{p}^{\text{fas}}_x$ from the model matches its empirical target, with $\widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
+{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $\beta$ is the share of additive in total transport costs.
+$t$ is calibrated so that the ratio $t/\widetilde{p}^{\text{fas}}_x$ from the model matches its empirical target, with $\widetilde{p}^{\text{fas}}_x$ the average fas price set in place by those domestic firms that export.
 We theoretically define the export fas price through: $p_x = \tau \widetilde{p}^{\text{fas}} +t$.
 TC = Transport Costs, expressed in percentage of the average fas export price.
-$\beta$ is the share of additive in total transport costs}}}
+}}}
 \end{table}
@@
   \centering
-  \caption{Exploring alternative scenario: Results }\label{tab:results_model_appendix}
+  \caption{Exploring alternative scenarios: Results }\label{tab:results_model_appendix}
 \begin{center}
@@
 \hline \hline
-& \multicolumn{2}{|c|}{Air transport} & \multicolumn{2}{|c}{Maritime transport} \\ \hline
+& \multicolumn{2}{|c|}{Air transport} & \multicolumn{2}{|c}{Vessel transport} \\ \hline
 Scenario & (A) & (B) & (A) & (B) \\ \hline
-$\Delta TC$ (pp)&	-0.02	&-0.014	&-0.033&	-0.031 \\
-$\Delta \tau$ (pp)&	0	&-0.014&	0&	-0.031 \\
-$\Delta t/\widetilde{p}^{fas} $(pp)&	-0.020	&0	&-0.033&	0 \\
-$\Delta \beta$ (in pp)	&-27.75&	12.14	&-23.53&	20.36 \\ \hline
+$\Delta TC$ (pp)&	-0.02	&-0.02	&-0.014&	-0.014 \\
+$\Delta \tau$ (pp)&	0	&-0.02&	0&	-0.014 \\
+$\Delta t/\widetilde{p}^{fas} $(pp)&	-0.02	&0	&-0.014&	0 \\
+$\Delta \beta$ (in pp)	&-25.37&	17.45	&-12.45&	24.45 \\ \hline
 %Delta Welfare (abs)&	0.042&	0.007	&0.060&	0.011 \\
-Rel. Delta welf (in \%)&	2.589&	0.411	&3.801	&0.695 \\
-Variety effect (in \%)	&-2.487&	-0.213&	-2.879	&-0.171 \\
-Price effect (in \%)	&5.275	&0.625&	6.973	&0.867 \\
+Rel. Delta welf (in \%)&	2.72&	0.56	&2.68	&0.97 \\
+Variety effect (in \%)	&-2.40&	-0.28&	-1.60	&-0.42 \\
+Price effect (in \%)	&5.31 &0.84&	4.38	&1.393 \\
 \hline \hline

### 2021-11-14 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changements texte avant relecture finale

@@
 Relying on our empirical estimates of transport costs
-over 1974-2019, our preferred scenario shows an extra welfare gain by around 1.2 percentage point when the transport cost decrease encompasses a reduction in its additive component relative to a pure ad-valorem reduction.
+over 1974-2019, our preferred scenario shows an extra welfare gain higher by 14\% to 40\% (in vessel and air transport respectively) when the transport cost decrease encompasses a reduction in its additive component relative to a pure ad-valorem reduction.
 Put differently, neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
@@
 Second, we make use of the latter to calibrate a \cites{melitz} type model of trade, which allows us to infer the welfare gains induced by the decrease of both components between 1974 and 2019.
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the key role of additive transport costs in determining those welfare variations.
 %It is true that, compared to \cite{Irrazabal_2015}, our study covers a narrower set of trade costs focused on transport costs.
@@
 \cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment, and draw welfare conclusions regarding the variations of these costs.
-Our results so far complement these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
+We have complemented these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
 In this section, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases.
 One case assumes that the reduction was achieved solely through the decline of ad-valorem costs, as modeled in standard New Trade theoretical approaches.
-The other measure builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
-%While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., iceberg trade costs), not much is said about the role of the additive component.
+The other case builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
+%While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., ad-valorem trade costs), not much is said about the role of the additive component.
 One potential explanation is the lack of empirical estimates about the size of additive costs.
@@
 In contrast, as in \cite{Irrazabal_2015}, we provide a quantification of such welfare gains.
-As \cite{Irrazabal_2015} provide estimates of transport costs only for one year (2004), their assessment of the welfare effects of a trade cost increase is based on a hypothetical assumption.
-We ground our study in the observed reduction of transport costs from 1974 to 2019.
+As \cite{Irrazabal_2015} provide estimates of transport costs only for one year (2004), their assessment of the welfare effects of a trade cost increase is based on a ad-hoc assumption (specifically, of a 5\% increase in the ratio of additive cost to the median fas price).
+Unlike them, we can exploit our database time span to ground our study in the observed reduction of transport costs from 1974 to 2019.
 \cite{Lashkaripour_JIE2020} uses a partial-equilibrium model with perfect competition and builds on its result that for denumerable goods, per-unit additive transport costs are very small to find a very small welfare effect of additive costs.
@@
 This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz's type models.
-we find that the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
+we find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
 
@@
 While \cite{Lashkaripour_JIE2020} relies on a perfectly competitive environment,
-Using a model with monopolistic competition and endogenous firm selection, as \cite{sorensen2014} and \cite{Irrazabal_2015}, does matter in this result.
-Indeed, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade liberalization lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
+In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade cost reduction lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
 Leaving aside the selection issue leads to an underestimation of the welfare gains that can be achieved by a reduction in additive transport costs.
@@
 This drives us to calibrate the structural parameters of the model.
-To do so, we use values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
+To do so, we usevalues commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
 
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
+{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
 $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s value) and the final steady state (39\%, based on 2006 value), according to the references provided above.}}}
@@
 \end{center}
-{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model ad-valorem costs only.
+{\parbox[l]{14cm}{ \vspace{4pt}\footnotesize{Notes: In Columns (1), (3), (5) and (7), we model ad-valorem costs only.
 In Columns (2), (4), (6) and (8), we model both ad-valorem and additive costs.
@@
 We exploit the theoretical implications of these empirical findings by running a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
+%Specifically, we contrast two cases: In the first scenario, we assume that only ad-valorem costs apply.
+We hence apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
+In the second scenario, we rely on the estimates of transport costs involving both additive and ad-valorem components.
+The consequences of the reduction in international transport costs through both its additive and ad-valorem components.
+
 In running this exercise, we also take into account the fact that the share of exporting firms has substantially increased since the 21\% value reported in \cite{BEJK-AER-03} (and consistent with the value estimated by \citet{Lincoln_McCallum2018} for 1987 on US data).
@@
 This can be matched in the model by resetting the (lower) value of the export fixed cost $f_x$ accordingly.
-Our preferred scenario thus embeds the two dimensions of trade liberalization: the reduction of transport costs along with the reduction of the export fixed cost.
+Our preferred scenario thus embeds the two dimensions of trade cost reduction: the reduction of transport costs along with the reduction of the export fixed cost.
 
@@
 \medskip
-
-\paragraph{Under a constant fixed export cost} In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from the fact that the transport cost decrease involves an additive part or not.
+%Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
+If we share a similar conclusion regarding the importance of additive costs, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
+
+\paragraph{Holding fixed costs constant} To understand the changes induced by the additive costs, we start by considering the scenario with a constant fixed export cost (Table  \ref{tab:resultats_modele}, Panel (1)).
+In this case, the welfare changes are solely attributable to the reduction in the variable transport costs, and the difference of results stems from the fact that the transport cost decrease involves an additive part or not.
 
@@
 
-\paragraph{Reduction in both variable and fixed trade costs} As mentioned above, our preferred scenario dwells on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}.
+\paragraph{Taking into account the reduction in fixed costs} As mentioned above, our preferred scenario includes, in addition to the reduction of variable costs, a reduction in fixed export costs, based on \cite{Lincoln_McCallum2018}'s empirical evidence.
 The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (2).
 Two results emerge.
-First, the welfare gains induced by the trade costs reduction are stronger in magnitude.
+First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
 While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, higher by around 50\% relative to the case when only the two components of the variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively.
@@
 A similar result applies with the observed 57\% reduction in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world).
-Neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
-While we have obtained these results in the canonical \cite{melitz} model, they call for integrating additive costs in richer New Trade models in view of better assessing the disputed question of the welfare gains of trade liberalization.
+Neglecting the existence and long-term evolution additive costs thus leads to a substantial underestimation of the welfare gains of trade cost reduction.
 
@@
 Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value (Scenario (A)), and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974 (Scenario (B)).
-These two scenarii (as well as the initial one based on 1974 values) are reported in Table \ref{tab:calib_TC_appendix}.
+These two scenarios (as well as the initial one based on 1974 values) are reported in Table \ref{tab:calib_TC_appendix}.
 
@@
 Rel. Delta welf (in \%)&	2.72&	0.56	&2.68	&0.97 \\
-Variety effect (in \%)	&-2.40&	-0.28&	-1.60	&-0.42 \\
-Price effect (in \%)	&5.31 &0.84&	4.38	&1.393 \\
+\hspace{1em}Variety effect (in \%)	&-2.40&	-0.28&	-1.60	&-0.42 \\
+\hspace{1em}Price effect (in \%)	&5.31 &0.84&	4.38	&1.393 \\
 \hline \hline

### 2021-11-15 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
début relecture

@@
 Second, we revisit the welfare gains of trade liberalisation in presence of additive costs.
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the disproportionate part of additive transport costs in determining those welfare variations.
 Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points.
-%In both aspects, our results point to the importance of modelling the additive component of international transport costs.
 \end{abstract}
@@
 In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Defined as the costs associated with the exchange of goods across
-national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}.
-Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development
-and new communication technologies. However, several empirical evidence
-suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}.
+Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
+However, several empirical evidence suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
-Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.
+Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) attests to the impact of transport costs on trade flows.
 \cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
@@
 If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction.
-Our main contribution in this paper is to bring new insights on this matter.\smallskip
+Among others, an important contribution in this paper is to bring new insights on the latter issue.\smallskip
 
@@
 By allowing for time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019.
-Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
+Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows.
 %\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.

### 2021-11-17 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte vfinale (1)

@@
 This paper was partly funded by a university grant from the University of Lille.
-It features an online appendix containing additional results, available on the authors' websites.}}
+It features %an online appendix containing additional results, available on the authors' websites and on https://github.com/gdaudin/trade_costs.
+}}
 
@@
 International transport costs do have an additive part.
-By how much?
-To what extent it matters?
+How large is it?
+Does it matter?
 This paper provides new answers to these questions.
@@
 The per-unit component of transport costs represents a sizeable share of total transport costs, between 30\% and 45\% depending on the year and the transport mode considered.
-We then investigate the important consequences of modelling additive costs, under two different perspectives.
+We then investigate the important consequences of additive costs, under two different perspectives.
 First, modelling varying additive costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role.
-Second, we revisit the welfare gains of trade liberalisation in presence of additive costs.
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the disproportionate part of additive transport costs in determining those welfare variations.
-Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points.
+Second, we revisit the welfare gains of the transport cost reduction in presence of additive costs.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the key role of additive transport costs in determining those welfare variations.
+Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease.
 \end{abstract}
@@
 \noindent \emph{JEL classification}: F10, N70, R40 \\
-\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2019, additive costs, trade composition effects, heterogenous firms
+\noindent \emph{Keywords}: Transport costs estimates, non-linear econometrics, period 1974-2019, additive costs, trade composition effects, gains from trade
 
@@
 Trade costs are central to international economic analysis.
-In particular, they are known to be a major obstacle to international economic integration and international trade flows.
-Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}.
-Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
-However, several empirical evidence suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
+In particular, they are a major obstacle to international trade flows.
+Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs.
+Many argue that the latter have substantially decreased thanks to technological advance in transportation, infrastructure development and new communication technologies.
+However, several empirical studies suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs.
 Equivalently, international transportation costs represent a 21\% markup over production costs.
@@
 of the transport costs burden.
-The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices).
-If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction.
+The latter also depends on whether the transport costs are in proportion to the value of goods traded (hence do not distort relative prices) or are not related to the value of goods traded (hence do distort relative prices).
+Having a clear view of the international transport cost burden is a complex task, but of key importance.
+Indeed, the answers to several important (both academic and policy) economic issues depend on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction.
 Among others, an important contribution in this paper is to bring new insights on the latter issue.\smallskip
@@
 Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019.
-Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which prove to be key in the characterization of international trade patterns.
+Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs.
 While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e.
-as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in \cites{samuelson1954} terminology,\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
-As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
+as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis (\cites{samuelson1954}),\footnote{``Iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.} a recent strand of the empirical literature points to the existence of an additive component, dependent on the physical quantities traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
 Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip
 
-As main result, we show that additive transport costs are quantitatively sizable.
+We show that additive transport costs are quantitatively sizable.
 On average over 1974-2019, the additive cost is estimated to be 1.7\% and 2.8\% of the export price in air and vessel transport respectively (at the 3-digit classification level).
@@
 Put differently, additive costs represent on average between 30 and 45\% of the overall transport costs.
+Further, the goodness of fit is systematically better when the additive component is included in the regression.
 To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs over such a long period of time.
-This represents a valuable insight for calibrating related international trade or business cycle models.
-We also provide an empirical assessment of what standard international trade models miss by skipping additive transport costs.
-Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by a factor around 2.
-Further, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degrees of freedom.\smallskip
+\smallskip
 
@@
 On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019.
-By allowing for time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019.
+By allowing for country-product-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel.
 Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows.
-%\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
-Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer.
+Importantly, the difference can be attributed to our more flexible modeling of the additive component.
 Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
 
-On the theoretical side, we show the importance of embedding additive costs when assessing the effects of trade cost reduction in welfare terms.
-Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction solely occurs through a decrease in the ad-valorem cost, as standardly modelled in New Trade theories, versus when it is achieved through a combined reduction in the ad-valorem and the additive components, as quantified by our empirical model.
-To this aim, we embed additive costs in the canonical \cite{melitz} model.
+On the theoretical side, we show the importance of additive costs when assessing the welfare effects of trade cost reduction.
+Specifically, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019.
+We contrast two cases: when this reduction solely occurs through a decrease in ad-valorem costs, as standardly modelled in New Trade theories, versus when it is achieved through a combined reduction in the ad-valorem and the additive components, as quantified by our empirical model.
+To this aim, we introduce additive costs in the canonical \cite{melitz} model.
 Relying on our empirical estimates of transport costs
-over 1974-2019, our preferred scenario shows an extra welfare gain higher by 14\% to 40\% (in vessel and air transport respectively) when the transport cost decrease encompasses a reduction in its additive component relative to a pure ad-valorem reduction.
-Put differently, neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation.
+over 1974-2019, our preferred scenario shows an extra welfare gain between 14\% and 40\% (in vessel and air transport respectively) when the transport cost decrease encompasses a reduction in its additive component compared to a pure ad-valorem reduction.
+Put differently, neglecting additive costs leads to a substantial underestimation of the welfare gains of globalization.
 
@@
 In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
-In this respect, it contributes to the literature challenging the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
-As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs.
+In this respect, it contributes to the literature challenging the dominant role of ad-valorem costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}).
+As shown by \cite{alchian}, the relative price of two varieties will depend on the level of additive trade costs.
 In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out'').
@@
 Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
+In contrast, \cite{Lashkaripour_JIE2020} finds evidence supporting the ad-valorem assumption for denumerable goods.
+Meanwhile, he provides evidence that the share of ad-valorem costs tends to decrease with the share of discrete goods.
+It is therefore reasonable to think that the inclusion of all goods would move the average share of multiplicative costs far from 100\%.
 
@@
 Also closely related to our paper is the work by \cite{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
-Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
-Our results share in common with \cite{Irrazabal_2015} the important role of the additive component of international trade costs, and a welfare analysis of the latter.
+Using Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs amount to 6\% of the export price multiplied by the ad-valorem cost.
+Like \cite{Irrazabal_2015}, we show the important role of the additive component of international trade costs.
+The literature also explores the welfare implications of additive trade costs.
+\cite{sorensen2014} analytically shows that important welfare gains can be achieved by reducing additive trade costs.
+Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs, the one exceptions being \cite{Irrazabal_2015} and \cite{Hornok-et-al-JIE-2015}.
+In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade cost reduction since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs, as well as by sub-component (additive and ad-valorem).\smallskip
 
@@
 First, we provide a careful, long-run characterization of both additive and multiplicative component of transports costs.
-Second, we make use of the latter to calibrate a \cites{melitz} type model of trade, which allows us to infer the welfare gains induced by the decrease of both components between 1974 and 2019.
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the key role of additive transport costs in determining those welfare variations.
-%It is true that, compared to \cite{Irrazabal_2015}, our study covers a narrower set of trade costs focused on transport costs.
-However, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost.
-By contrast, our estimation strategy enables us to estimate each value of the ad-valorem and the additive costs independently from the other.
-From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
-While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.
-This comparison of results suggests that between a quarter and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode).
-
-%\textbf{XXX JH : I WOULD DELETE EVERYTHING FROM `` It is true that"- LP, OK METTRE EN SECTION 3 Comparaison des resultats XXX }
-
+We show that additive costs represent a substantial part of total transport costs.
+Our results are robust to a range of robustness checks, including the use of instrumental variables and more disaggregated data at the SITC 4 or HS10 levels.
+Second, additive costs matter in measuring the role of composition effects in the long-term decline of transport costs.
+Third, we make use of our empirical estimates to calibrate a \cites{melitz} type model of trade, which allows us to infer the welfare gains induced by the decrease of transport costs between 1974 and 2019.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the key role of additive transport costs in determining those welfare variations.
 \smallskip
 
-%Yet, our paper complements their findings in two main respects.
-%In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed over the last fifty years.
-%However, these papers yet do not allow for a quantitative measure of additive costs (neither for the ad-valorem costs).
-Another important contribution of this paper is to provide quantitative estimates of both type of transport costs.
-\smallskip
-Our results are robust to a range of robustness checks, including the use of instrumental variables and more disaggregated data at the SITC 4 or HS10 levels.
-They have important implications for both international trade and international macroeconomics models.
-Some recent papers notably point out the welfare implications of additive trade costs.
-\cite{sorensen2014} analytically shows that important welfare gains can be achieved by reducing additive trade costs.
-Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs, the one exception being \citealp{Irrazabal_2015}.
-In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade cost reduction since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs, as well as by sub-component (additive and ad-valorem).\smallskip
-
-%\textbf{To develop}. \medskip
-
 The remainder of the paper is built as follows.
-Section \ref{sec:data_method} presents the data and the baseline empirical model used to estimate the values of international transport costs.
+Section \ref{sec:data_method} presents the data and the baseline empirical model we use to estimate the values of international transport costs.
 Section \ref{sec:basic_results} reports the later on a yearly basis throughout the period 1974-2019 by transport mode, explicitly distinguishing between the additive and the ad-valorem components and providing several important robustness checks of our results.
 The two following sections take stock of those results to draw implications for two important debates in economic analysis.
-In section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \textit{per se}.
-Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+In Section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and partner country), and what arises from changes in transport costs \textit{per se}.
+Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of the observed decrease in ad-valorem and additive costs.
 Section \ref{sec:conclu} concludes.
@@
 
-Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}.
-The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2019 (details on the database provided in Appendix \ref{app:data}).
+Our analysis of transportation costs exploits the difference between commodity-level export and import prices, as in \cite{hummels2007}.
+We use his database from 1974 to 2004.
+We extend this database to 2019, using his original source, the US annual ``Imports of Merchandise'' provided by the US Census Bureau (details on the database provided in Appendix \ref{app:data}).
 We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price.
-While both terms are closely related (see more details in Appendix \ref{app:data}), the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity (the number) of goods.
+While both terms are closely related (see more details in Appendix \ref{app:data}), the US Foreign Trade Statistics reports fas prices.} More precisely, the fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity (the number) of goods.
 This is due to the fact that in the Census data, information on the quantity of goods is not mode-specific.
-Nevertheless, Section D.4 in the Online Appendix reports extensive robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.
-%\textbf{XXX LE METTRE EN FOOTNOTE ME PARAIT LEGER, POINT IMPORTANT DU REFERE 1 + bcp de footnotes.
-Le remonter dans le corps PPL?
-XXX }
-}
+Furthermore, many goods are measured only by weight or volume, especially in the 1970s.
+Nevertheless, Section D.4 in the Online Appendix reports extensive robustness check where unit values are built as the ratio between value and quantity of goods per observation, on a specific sub-sample restricted to US imports flows that only use one single transport mode.}
 The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight.
 Our dependant variable is finally computed as the ratio of the cif price divided by the fas price.
-By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
+By construction higher than 1, the variable provides a measure of transport costs as a proportion of the fas price, an ad-valorem equivalent.
 This is quite a standard and widespread strategy, as emphasized by \cite{anderson_wincoop_jel}.
@@
 It is thus silent about the other dimensions of international trade costs.
-Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size, etc.).
+Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory costs, preparation costs associated with shipment size, etc.).
 In this respect, this dataset embraces only a partial view of international transport barriers.
@@
 This limits the measurement error bias.
-Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level, with a concordance code to the SITC coding system.
+Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level.
+We use a concordance table with the SITC coding system to be consistent with Hummels's (2007) database.
 In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air.
@@
 As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-This drives us to estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
+hence, our benchmark results estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
 Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin (hence, around 400 fixed effects).
 We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i),} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
-Notice that this can be done from 1989 onwards for which HS-10 data are available; \textit{ii),} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
-The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2019.
+Notice that this can theoretically be done from 1989 onwards for which HS-10 data are available; \textit{ii),} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
@@
 
-\paragraph{Estimation strategy} Our starting point relates the import price (at the entry gate) of a given good as equal to the export price (at the export gate) plus freight charges.
+\paragraph{Estimation strategy} Our starting point relates the import price of a given good as equal to the export price plus freight charges.
 Specifically, denoting by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped, we have: $ p = \widetilde{p} +f$.
-As in \cite{hummels2010}, assume that the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.).
+As in \cite{hummels2010}, assume that the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents the transport costs level (depending on elements such as distance, port quality, etc.).
 The link between the import, export prices and shipping charges is hence given by:
@@
 At the other extreme, the case $\beta=1$ corresponds to shipping costs being independent of the unit price, with the cif price the sum of the unit price and the shipping cost: Transport costs are fully additive.
-This suggests that, the closer $\beta$ to 1, the more prevalent additive costs in total transport costs.
+The closer $\beta$ to 1, the more prevalent additive costs in total transport costs.
+
 One can analytically show that $\beta$ is indeed equal to the share of additive costs in total transport costs.
-
 To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as $p = \tau \widetilde{p}+ t$, with $\tau \geq 1$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported.

### 2021-11-17 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Correction url
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
changement texte (encore)

@@
 This paper was partly funded by a university grant from the University of Lille.
-It features %an online appendix containing additional results, available on the authors' websites and on https://github.com/gdaudin/trade_costs.
-}}
+It features an online appendix containing additional results, available on the authors' websites and on \url{https://github.com/gdaudin/trade_costs}.}}
 
@@
 \end{equation}
-Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:
+Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ is equal to:
 \begin{eqnarray}
@@
 
-Comparing Equations (\ref{eq:alternative}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs is also equal to the share of additive costs in total transport costs.
-
-While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour-2017}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
+Comparing Equations (\ref{eq:alternative}) and (\ref{eq:tt2}), it is straightforward that the unit price-elasticity of transport costs (in absolute value, $\beta$) is also equal to the share of additive costs in total transport costs.
+
+While many in the related literature provide an indirect measure of additive costs by estimating the elasticity of freight charges to the unit price (\citealp{hummels_skiba}, \citealp{Lashkaripour_JIE2020}), thereby relying on Equation (\ref{eq:alternative}), we propose an alternative route.
 Based on Equation (\ref{eq:base}), we exploit our dataset to obtain explicit measures of both transport costs components $\tau$ and $t$ (or $\widetilde{t}$).
@@
 \begin{equation}
-\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik}-1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
+\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik}-1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}, ~~\text{with } \tau_{ik}\geq 1 ~\text{and }t_{ik}\geq 0 \label{eq:base_estimee}
 \end{equation}
@@
 We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
-In other words, $\tau_{ik}$ and $t_{ik}$ are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1\simeq (\tau_i-1) + (\tau_k -1) $ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
+In other words, $\tau_{ik}$ and $t_{ik}$ are written as:\footnote{Notice that, given the small magnitude of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1\simeq (\tau_i-1) + (\tau_k -1) $ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
 We also provide a robustness analysis to the separability assumption.
@@
 
-The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the import price $p$ cannot be lower than the export price ($p_{ik}\geq\widetilde{p}_{ik}$).
-Taking into account this constraint in the estimate implies that the error term should be always positive.
+The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the import price cannot be lower than the export price ($p_{ik}\geq\widetilde{p}_{ik}$).
+Taking into account this constraint in the estimation implies that the error term should be always positive.
 We ensure that this constraint is fulfilled by specifying the error term as follows:
@@
 \noindent where $\epsilon_{ik}$ follows a normal law centered on 0.
-Considered in logs, the above equation becomes:
+Considering that the LHS is positive by nature, we need to take into account a positivity constraint on the RHS.
+We fulfill this requirement by taking logs.
+The above equation then becomes:
 
@@
 \end{equation}
+
+\textbf{STOP HERE XXXX }
 

### 2021-11-18 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte

@@
 \begin{equation}
-\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
-\end{equation}
-
-\textbf{STOP HERE XXXX }
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}, ~\text{with } \tau_{i}\geq 1, \tau_{k}\geq 1, t_{i}\geq 0  ~\text{and }t_{k}\geq 0  \label{eq:equation0}
+\end{equation}
+
+\noindent where we ensure the positivity constraint on $\tau_x-1$, $t_x$ (with $x =i,k$) by taking logs.
 
@@
 See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at a more disaggregated product level ($k=5$ in our benchmark estimates).
+Confronted with this trade-off, we estimate 3-digit level sectoral fixed effects as our benchmark.
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
@@
 on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
-This would not solve the problem though, for three main complementary reasons.
+This would not solve the problem though, for three complementary reasons.
 First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
@@
 Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
+In a related manner, we could not impose positivity constraints on the coefficients.
 Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}
@@
 We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2019, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
-Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
+Depending on the year considered, this leaves us with around 400 fixed effects to estimate by transport mode at the 3-digit level.
 \medskip
 
-As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, which decomposes into both an additive and an ad-valorem component.
-We do so by adopting the following strategy.
+To drive home the importance of additive costs, we adopt the following strategy.
 For each year and transport mode, we estimate two models.
@@
 
+In this case, notice that the equation could be estimated relying on a linear form.
+To preserve comparability of the results, we keep the same non-linear estimation method in both cases.
 Under Model (B), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})).
 From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modeling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs.
-This has driven us to estimate this third model (c) as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
-The main result that emerges is that this model (c) is dominated (in terms of quality of fit properties) by the model (A) with multiplicative costs only, which is itself dominated by the complete model (B), anticipating further results.
-More details of these results are given in the Online Appendix.}
-
-%One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
-%Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-%This is not an issue here, as we are not interested in causal inference.
-%Rather, our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
-
-%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis).
-%Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control.
-%In this respect, we are confident that our estimation strategy is immune from endogeneity problems.
-
-%\textbf{pas encore completement clair.
-Veut-on dire que ce qui est pbtique, c'est que TC reagisse a ptilde?
-Je crois que ce qu'on veut dire, c'est que le fait que ptilde change en fonction des TC ne change pas les TC, le cif price s'ajuste proportionnellement?
-Si c'est ca, c'est vrai seulement si les couts sont advalorem...Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus?
-Par produit 3d/pays, par secteur 3d tous pays, etc...?}\smallskip
+This has driven us to estimate this third model (C) as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
+The main result that emerges is that this model (C) is dominated (in terms of quality of fit properties) by the model (A) with multiplicative costs only, which is itself dominated by the complete model (B).
+More details of these results are given in the Online Appendix, Section A.}
 
@@
 When assuming ad-valorem costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is(k)} = \widehat{\tau}_{i} \times \widehat{\tau}_{s(k)}$.
-In this case, notice that the equation could be estimated relying on a linear form.
-To preserve comparability of the results, we keep the same non-linear estimation method in both cases.
-Similarly, as \cite{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
+As \cite{Irrazabal_2015}, we take the average over the country-sector dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme.
 We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (A), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (B), for each year and transportation mode.
@@
 %Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components.
-\paragraph{Quantification} The first step of the analysis consists of providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
+\paragraph{Quantification} The first step of the analysis consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})).
 This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter.
@@
 It displays the weighted mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
-Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}
+Results for all years (available at the 3-digit level) are reported in the Online Appendix.}
 The top panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.
 In the medium panel, we report the results for a specification based on ad-valorem costs estimated alone (Model (A), based on Equation (\ref{eq:model_nlI})).
-The bottom panel presents our estimates for a specification involving both ad-valorem and multiplicative components (Model (B), based on Equation (\ref{eq:model_IetA})).
+The bottom panel presents our estimates for our baseline specification (Model (B), based on Equation (\ref{eq:model_IetA})).
 
@@
 
-Before jumping to the interpretation, Table \ref{tab:summary_results} calls for two comments of statistical order.
+Before jumping to the interpretation, Table \ref{tab:summary_results} calls for two comments.
 First, the estimated mean values are systematically higher than medians.
-This result is not surprising though, given that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median.
+This result is not surprising though, given that our key variable is positive by nature; therefore, outliers can only raise the mean upward compared to the median.
 Second, the estimated total costs are not exactly equal to the observed ones, in particular for Model (A).
@@
 
-\paragraph{Interpretation} Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}.
-First, it provides estimation results about the size of the overall transport costs.
-Table \ref{tab:summary_results} shows that transport costs amount to around a 5.5-6\% markup in ocean shipping, and 4\% markup in air shipping on average over 1974-2019.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
+\paragraph{Interpretation} We now come to Table \ref{tab:summary_results}.
+First, the estimated size of overall transport costs amount to around a 5.5-5.8\% markup in ocean shipping, and 4.1-4.8\% markup in air shipping on average over the period.
+%\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$.
 Under Model (A) (ad-valorem costs only), this gives a markup of 5.5\% in vessel, and 4.8\% in air shipping (considering the mean value over the period).
@@
 With $\tau= 1.030$ and $\frac{t}{\widetilde{p}}= 0.028$ in vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 5.8\%.
-A similar calculus gives 4.1\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
+%A similar calculus gives 4.1\% in air transport.} 
+If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
 \cites{anderson_wincoop_jel} estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
@@
 Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as displayed in the last rows of Table \ref{tab:summary_results}: The per-kg cost dimension is quantitatively sizable.
-\medskip
-
-We also provide diagnostic tests about the goodness-of-fit measures, which confirm the empirical relevance of adding the additive component in the estimation.
-Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests that confirm the importance of the additive cost component.
-The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the removed degree of freedom.
-For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2.\smallskip
+
+We also provide diagnostic tests about the goodness-of-fit measures, which confirm the relevance of adding the additive component in the estimation.
+As reported in Appendix \ref{app:diagnostic_test}, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the fewer degrees of freedom.\smallskip
 
 Our estimation results on the size of additive transport costs can be compared to \cite{Irrazabal_2015}, the most recent evidence on this question.
-Based on Norwegian trade data for 2004, \cite{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e.
+Based on Norwegian trade data for 2004, \cite{Irrazabal_2015} estimate that additive costs amount to 6\% of export price multiplied by the ad-valorem cost, i.e.
 $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
 One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
-From this, we can rebuild the ratio in similar terms to them, allowing us to gain in generality in this respect.
+This allows building the ratio in similar terms to them, allowing us to gain in generality in this respect.
 Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping and 1.5\% in air transport.
@@
 Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.
-On top of our results summarized in Table \ref{tab:summary_results}, this points to evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip
 

### 2021-11-18 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte

@@
 Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.
+1989 trade by air is excluded.
 \end{minipage}
@@
 The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
-We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
-For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping.
+We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.}
+
+%Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
+%For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping.
 Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip
 
-\paragraph{The share of additive costs: Time dynamics} As next step, we investigate whether additive costs are a long-lasting, i.e.
-a structural phenomenon.
-Figure \ref{fig:part_cout_additif} reports the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$.
+\paragraph{The share of additive costs: Time dynamics} Figure \ref{fig:part_cout_additif} reports the share of additive costs in total transport costs on a yearly basis, in both air and vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$.
 
 \begin{figure}[htbp]
-\caption{Share of additive costs $\beta$ (3-digit classification)}
+\caption{Share of additive costs $\beta$ (baseline at 3-digit level)}
 \label{fig:part_cout_additif}
@@
 \begin{minipage} [c]  {5in} \scriptsize%
-Note:The share of additive costs $\beta$ is reported on the $y-$axis.
-Each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).
+Note: The dotted line is the quadratic trend.
 \end{minipage}
@@
 In both transport modes, the additive component appears of sizable importance throughout the period.
-Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon.
-On the contrary, they represent a sizable and structural dimension of international transport costs.
+Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs play a substantial role in total transport costs throughout the period.
 If the share of additive costs in total costs roughly amounts to 44\% (30\%) in vessel (air) transport on average, one can further notice that it varies substantially over time, with a different pattern across transport modes.
-In maritime transport, the share of additive costs fluctuates around the average, while it has been steeply declining since the end of the 1980s in air transport.
-This suggests that the transport costs decrease is more than proportionally due to additive costs in air transport, but more equally shared between ad-valorem and additive costs in ocean transport.\smallskip
+After a period of common increase in air and vessel from the mid 1970s to the end of the 1980s, it has steadily decreased for air transport.
+In contrast, the share of additive costs has continued to increase until the 2000s.
+Put differently, the total transport costs decrease is more than proportionally due to additive costs in air transport, while it is mostly driven by ad-valorem cost reduction in vessel transport.\smallskip
 
@@
 \subsubsection{Robustness to the separability assumption}
-In this section, we provide a robustness analysis to the separability assumption of country-good fixed effects, which considers (following \citealp{Irrazabal_2015}) that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the product ($s(k)$) dimensions.
+Following \cite{Irrazabal_2015}, we have assumed that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the sector ($s(k)$) dimensions.
+In this section, we provide a robustness analysis to this separability assumption.
 That is, we relax Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to model $\tau_{is(k)}$ and $t_{is(k)}$ (on a yearly/transport mode basis).
-Yet, this comes at the cost of having to estimate a much larger number of fixed effects (for approx.
-200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation (all the more given our non-linear setting) intractable in practice.
+This comes at the cost of having to estimate a much larger number of fixed effects (for approx.
+200 countries of origin and 200 sectors at the 3-digit level, this means estimating 40,000 versus 400 fixed effects, per year and per transport mode), making the estimation intractable in practice (all the more given our non-linear setting).
 This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.

### 2021-11-18 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
corrections diverses

@@
 ignoring those flows that did not happen because of presumably prohibitive transport costs.
-In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip
+In light of this, the estimated values that we obtain can be viewed as lower bounds of the ``potential'' transport costs.} \smallskip
 
@@
 Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} and Section B of the online appendix for the 4-5 results; and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).
+However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} below and Section B of the online appendix for the 4-5 results; and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).
 
@@
 Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries.
-See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}
+See Section \ref{sec:rob_sep} for a detailed presentation of this robustness check.}
 \begin{eqnarray}
@@
 This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter.
-These results constitute our first original contribution to the literature.
+These results constitute the first original contribution to the literature of this paper.
 \smallskip
@@
 It displays the weighted mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
-Results for all years (available at the 3-digit level) are reported in the Online Appendix.}
+Results for all years (available at the 3-digit level) are reported in Section C from the Online Appendix.}
 The top panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.
@@
 If this seems modest in comparison with the 11\% of markup obtained by \cite{anderson_wincoop_jel} over a set of industrialized countries, it stands in line with the results obtained by \cite{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible explanation for the difference with \cite{anderson_wincoop_jel}, lays in the difference of empirical methodology.
-\cites{anderson_wincoop_jel} estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}
+\cites{anderson_wincoop_jel} estimates are based on standard gravity equations, while ours come from a non-linear estimation of a structural decomposition of the difference between export and import prices.}
 
@@
 This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3\% and 2.4\% for vessel and air shipping respectively (on average over the period).
-Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as displayed in the last rows of Table \ref{tab:summary_results}: The per-kg cost dimension is quantitatively sizable.
+Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as displayed in the last rows of Table \ref{tab:summary_results}.
+The per-kg cost dimension is quantitatively sizable.
 
@@
 In contrast, the share of additive costs has continued to increase until the 2000s.
-Put differently, the total transport costs decrease is more than proportionally due to additive costs in air transport, while it is mostly driven by ad-valorem cost reduction in vessel transport.\smallskip
-
-In the following section, we check the robustness of these results regarding three important issues, the separability assumption of country-good fixed effects, potential endogeneity biases and the level of aggregation at the product level.
+Put differently, the total transport costs decrease is disproportionately related to additive costs in air transport, while it is mostly driven by ad-valorem cost reduction in vessel transport.\smallskip
+
+In the following section, we check the robustness of these results regarding three important issues: the separability assumption of country-good fixed effects, potential endogeneity biases and the level of aggregation at the product level.
 
@@
 
-\subsubsection{Robustness to the separability assumption}
+\subsubsection{Robustness to the separability assumption}\label{sec:rob_sep}
 Following \cite{Irrazabal_2015}, we have assumed that both the ad-valorem and the additive costs are separable between the origin country ($i$) and the sector ($s(k)$) dimensions.
@@
 
-\subsubsection{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) is subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
+\subsubsection{Endogeneity concerns}\label{sec:IV_main_text}
+
+ One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) is subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
@@
 We follow earlier literature (see e.g.
-\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
+\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties based on tariffs at the product line, together with one-year lagged fas prices.
 Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
@@
 
-\subsubsection{Robustness to aggregation}
+\subsubsection{Robustness to aggregation}\label{sec:agg_main_text}
 
@@
 
-The estimation carried out in section \ref{sec:results_decomposition} provides us with the
-Starting from the additive and ad-valorem measures of international transport costs which vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$), we extract the pure transport cost measure by the mean of a time fixed effect.
+Using the origin country-product-time-varying additive and ad-valorem measures of international transport costs (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$) estimated in section \ref{sec:results_decomposition}, we extract the pure transport cost measure by the mean of a time fixed effect.
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
@@
 \subsection{Characterizing the time trends in transport costs}
-Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (A), (B) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
+Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for air transport (starting from the reference value 100 in 1974).
 Panels (d), (e) and (f) report the results for vessel.
@@
 International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
-Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (a), (d) and (e)).
+Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
 Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (A) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
 For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
@@
 The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
-If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
+If anything, we find the opposite result here: trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
 This drives us to investigate this difference further.
@@
 As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Rather than from the actual cif-fas price gap, we start from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition} at the year/sector/origin country level, by transport mode.
+Rather than from the actual cif-fas price gap, we start from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition} at the year/sector/origin country level, by transport mode).
 Put differently, our methodology lets the share of additive costs $\beta$ vary over the three time-product-partner country dimensions.
@@
 Unlike them, we can exploit our database time span to ground our study in the observed reduction of transport costs from 1974 to 2019.
-\cite{Lashkaripour_JIE2020} uses a partial-equilibrium model with perfect competition and builds on its result that for denumerable goods, per-unit additive transport costs are very small to find a very small welfare effect of additive costs.
-We look at the whole range of goods over a longer period and use a \citet{melitz} model like \cite{sorensen2014} and \cite{Irrazabal_2015}.
+Based on his main result that per-unit additive transport costs are very small for denumerable goods, \cite{Lashkaripour_JIE2020} consistently finds a very small welfare effect of additive costs within a partial-equilibrium model with perfect competition.
+In this paper, we look at the whole range of goods over a longer period and use a \citet{melitz} model like \cite{sorensen2014} and \cite{Irrazabal_2015}.
 In that context,
@@
 we find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
-
+In contrast, \cite{Lashkaripour_JIE2020} finds evidence supporting the ad-valorem assumption for denumerable goods
 %That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020}paper relies on a negligible share of additive costs.
@@
 
-In line with the literature, we find that a reduction of trade costs  induces welfare gains
+In line with the literature, we find that a reduction of trade costs  induces welfare gains.
 % whatever the transport mode (and the associated magnitude of the transport costs decrease).
@@
 Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction in variable transport costs in Air transport generates a relative welfare gain 40\% higher when such a reduction partly occurs through the additive component, rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)).
+In our preferred scenario, the reduction in variable transport costs in Air transport generates a relative welfare gain 40\% higher when such a reduction partly occurs through the additive component, rather than when it happens through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)).
 In maritime transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
@@
 Again, the effect is quantitatively sizeable: Welfare gains are thus between 30\% and 78\% higher, depending on the transport mode.
-This is driven by a more positive price competitiveness effect and despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
+This is driven by a more positive price competitiveness effect, despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
 This can be accounted for as follows.
@@
 As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productive
-exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})).
+exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px}) in Appendix \ref{app: model_general}).
 Within the exporters category, this favors the more productive firms, whose market share,
-e.g. revenues, increases relatively more (see Equation (\ref{eq:rx})). Everything else equal, this
+e.g. revenues, increases relatively more (see Equation (\ref{eq:rx}) in Appendix \ref{app: model_general}). Everything else equal, this
 pushes the average productivity on the export market upwards. Unlike the ad-valorem
@@
 Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
-In the end, the welfare gains are the highest when trade costs reduction induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
-
-In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results point out the importance of taking into the additive components of international trade costs.
+In the end, the welfare gains are the highest when trade costs reduction relies on a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
+
+In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results emphasize the importance of taking into account the additive components of international trade costs.
 Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods.
@@
 
-This paper assesses the magnitude and the role of additive international transport costs.
+This paper assesses the magnitude and the impact of additive international transport costs.
 Empirically, we exploit the differences between the import and the export prices taken from the US import database over 1974-2019, to estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
@@
 On the theoretical side, our results call for exploring the role of additive costs in shaping international trade flows in richer New Trade models than the canonical M\'{e}litz model used so far, and in affecting the international transmission of business cycles.
-This is left for further research.
 
@@
 A decrease in the log-likelihood function points to a better quality of fit.
-However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.
-\footnote{Specifically, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.}
+However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.\footnote{Specifically, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $, the number of parameters being given by the number of restrictions.}
 The preferred model is the one with the minimum AIC value.
-Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including the additive term, i.e.
-Equation (\ref{eq:model_IetA})) and the Restricted one (\emph{R}, i.e.
-Equation (\ref{eq:model_nlI})).
+Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the unrestricted model (including the additive term, i.e.
+Equation (\ref{eq:model_IetA})) and the restricted one (i.e. Equation (\ref{eq:model_nlI})).
 The null tested is that the two models are statistically equivalent.
@@
 
-We report here the first-stage procedure and estimates underlying the IV estimates reported in section \ref{sec:results_decomposition}.
+We report here the first-stage procedure and estimates underlying the IV estimates reported in section \ref{sec:IV_main_text}.
 We start by detailing the theoretical intuitions behind the first-stage equation.
-Afterwards, we report the estimates based on 5-digit-product-level data, before moving to the more disaggregate HS10 level for sensitivity checks.
+Afterwards, we report the estimates based on 5-digit-product-level data, before moving to the more disaggregated HS10 level for sensitivity checks.
 
@@
 We then replace the observed series of fas prices in the right-hand side of Equation (\ref{eq:model_IetA}) by the predicted fas prices produced by Equation (\ref{eq:firststage_final}).
-We now report results based on our main database at the SITC 5-digit level, before replicating the exercise on HS10 digit level, consistently with one major robustness implemented in the paper.
+We now report results based on our main database at the SITC 5-digit level, before replicating the exercise on HS10 digit level, consistently with another major robustness implemented in section \ref{sec:agg_main_text} from the main text.
 
@@
 Consistently, we will use as instrument 5-digit product-specific tariff rates, as well as the one-year-lagged fas price.
-Table \ref{tab: FS_sitc5} reports summary results of these first stage regressions, while Figure \ref{fig: FS_IV_SITC5} displays year estimates of $\beta$ and $\alpha$ parameters, together with 5\% confidence bands:
+Table \ref{tab: FS_sitc5} reports summary results of these first stage regressions, while Figure \ref{fig: FS_IV_SITC5} displays yearly estimates of $\beta$ and $\alpha$ parameters, together with 5\% confidence bands:
 
@@
 At this highly disaggregated level, pricing-to-market behavior seems less strong, indicating that the average firm is smaller than its average SITC5 counterpart.
-In addition, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their Table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006.
+In addition, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006.
 In any case, this strengthens our confidence in the validity of our estimation strategy.
@@
 
+\clearpage
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{D.\arabic{figure}}
+\renewcommand{\thetable}{D.\arabic{table}}
+
 \section{More on the model \label{app:theoretical_model}}
 
-\subsection{The model in the general case}
+\subsection{The model in the general case} \label{app: model_general}
 

### 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Quelques modifications Guillaume
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 \begin{figure}[htbp]
-\caption{Share of additive costs $\beta$ (baseline at 3-digit level)}
+\caption{Share of additive costs $\beta$: baseline at 3-digit level}
 \label{fig:part_cout_additif}
@@
 This drives us to conduct the analysis on a sub-sample of trade flows, upon which two estimations are conducted (for every year and each transport mode): with the separability assumption (Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) hold, our baseline case) and without.
-For each year and transport mode, we select the countries ($i$) and the products (at the 3-digit level, $s(k)$) that constitute 80\% of the total value of trade flows.
+For each year and transport mode, we select the most important countries ($i$) and sectors (at the 3-digit level, $s(k)$).
 Table \ref{tab:robustness_separability} shows a summary of the results, considering the average estimated values over the period (by transport mode).
@@
 	\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule:  we select, for each year, the largest importers that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
-We keep in the sample all trade observations from these importers and in these sectors.
-			The additive term is expressed in fraction of fas price.}}
+We keep in the sample all trade observations from these importers and in these sectors.}}
 	\label{tab:robustness_separability}
@@
 Second, both models provide a similar quality of fit of the regression (once we take into account the number of right-hand-side variables) as the AIC criteria display very close values in both models (conditional on transport mode).
-Third, crucially for this robustness check the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also report in Section C from the Online Appendix a robustness check comparing the results with/without the separability assumption on a yearly basis} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
+Third, crucially for this robustness check, the estimated values remain of similar magnitude under both separability and non-separability assumptions.\footnote{We also report in Section C from the Online Appendix a robustness check comparing the results with/without the separability assumption on a yearly basis} %, even if one can note that under the separability assumption, the share of additive costs seems slightly biased upward, at least for averages over the period.
 The picture is reversed for medians
@@
 \begin{figure}[htbp]
-	\caption{Transport cost time trends: robustness to the separability assumption}
+	\caption{Share of additive costs $\beta$: robustness to the separability assumption}
 	\label{fig:comp_separability_SITC5}
@@
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-Related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+Related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-sector pair.
 In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
@@
 We follow earlier literature (see e.g.
-\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties based on tariffs at the product line, together with one-year lagged fas prices.
+\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties at the product-country level, together with one-year lagged fas prices.
 Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
@@
 Figure \ref{fig:comp_IV_SITC5} reports our benchmark estimates by transport mode, together with their instrumented counterpart.
-In all cases, these estimates are very similar, when not identical.
+In all cases, these estimates are very similar.
 This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
@@
 \begin{figure}[htbp]
-\caption{Transport cost time trends: robustness to IV procedure}
+\caption{Share of additive costs $\beta$: robustness to IV procedure}
 \label{fig:comp_IV_SITC5}
@@
 Notice that switching to a 3 (sector)-10(product) estimation method does not alter the number of fixed effects to estimate, as they are at the sector $s$ level (along with those at the origin country $i$ level).
-However, within each 3-digit sector category, the 3-10 estimation method has the advantage to exploit a larger source of heterogeneity by considering trade flows at the HS-10 product level.
+However, within each 3-digit sector category, the 3-10 estimation method has the advantage of exploiting a larger source of heterogeneity by considering trade flows at the HS-10 product level.
 
@@
 If anything, the $\beta$ seems on average slightly higher when estimated based on a 3 (sector)-10 (product) database, indicating a (small) downward bias in our estimates of additive costs.
-But overall, no major difference arises both regarding levels and trends - results are especially very close for the Air transportation mode.
+But overall, no major difference arises in levels and trends - results are especially very close for the Air transportation mode.
 This has important implications for the exercise performed in Section \ref{sec:theory}: welfare gains computations should not be significantly affected by our use of 3 (sector)-5 (product) digit data.
@@
 	\centering
-	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
+	\caption{Share of additive costs $\beta$: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
 	\input{Table_baseline10_reshaped.tex}
@@
 
-\paragraph{Summary} Two main insights emerge up to now.
-First, total transport costs have strongly reduced between 1974 and 2019.
-Second, additive costs represent an important share of those declining transport costs, between 30 and 45\%.
+\paragraph{Summary}
+%Two main insights emerge up to now. First, total transport costs have strongly reduced between 1974 and 2019. Second, a
+\textbf{À valider xxxxx GD}
+From 1974 to 2019, additive costs represent an important share of declining transport costs, between 30 and 45\%.
 In the two following sections, we take stock of those results to draw implications for two important debates in economic analysis.
@@
 Specifically, \cites{hummels2007} results support that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974.
-Our own results tend to challenge this view, as we further show.
+Our own results challenge this view.
 
@@
 
-\noindent where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) and $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
+\noindent where $X_{ikt}$ represents transport costs level due to distance, port quality, etc.
+and $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
 Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
 In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year.
-Put differently, this means that the share of additive costs in total costs does not vary across these dimensions.
-
-\paragraph{Challenging the $\beta$'s constancy assumption} As Figure \ref{fig:part_cout_additif} suggests, the share of additive costs is rather varying over time, in both transport modes.
+
+\paragraph{Challenging the  constancy of $\beta$ assumption} As Figure \ref{fig:part_cout_additif} suggests, the share of additive costs is rather varying over time, in both transport modes.
 %In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.
@@
 
-\noindent making use of our first-stage estimates for both the additive and the ad-valorem components ($\widehat{t}_{is(k)t}$ and $\widehat{\tau}_{is(k)t}$).
-The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}, in panels (A) and (B) for air and maritime transport respectively.
+\noindent making use of our estimates for both the additive and the ad-valorem components ($\widehat{t}_{is(k)t}$ and $\widehat{\tau}_{is(k)t}$).
+The histogram of the distribution of the estimated $\beta_{ikt}$ is reported in Figure \ref{fig:histogram_beta}.
 
@@
 
-As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending on the transport mode, consistently with our previous findings.
+As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$.
 This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
-Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
-These results are not reported for sake of space, but are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
+Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{This it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
+These are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.
 
@@
 
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, which explicitly takes into account the varying share of additive costs over time, country partner and product.
+Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, explicitly taking into account the varying share of additive costs over time, country partner and product.
 We start with the presentation of our estimation strategy before turning to the results.
@@
 
-Using the origin country-product-time-varying additive and ad-valorem measures of international transport costs (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$) estimated in section \ref{sec:results_decomposition}, we extract the pure transport cost measure by the mean of a time fixed effect.
+Using the origin country-product-time-varying additive and ad-valorem measures of international transport costs %(i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$)
+estimated in section \ref{sec:results_decomposition}, we extract the pure transport cost by estimating time fixed effects.
 Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
@@
 One advantage of this method is that it yields measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time.
+For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (keeping country and sector constant) that have reduced over time.
 For sake of reading clarity, we describe the method to extract these series in Appendix \ref{app:comp-effects}, and we now turn to the results.
@@
 Panels (d), (e) and (f) report the results for vessel.
-In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.
+In each panel, we report the evolution of transport costs for both the unfitted (plain line) and the fitted (dotted line) measures.
 
@@
 First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping.
+International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in vessel shipping.
 Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
-Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
+Further, the transport cost reduction is much smoother in vessel transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
 Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
@@
 In this section, we explore the implications of our empirical results on theoretical grounds.
-Over the last decade, several papers examined the implications of non-ad-valorem costs.
+Over the last decade, several papers have examined the implications of non-ad-valorem costs.
 \cite{Kropf-Saure-JIE-2016} estimate the size and shape of per-shipment costs based on Swiss data.
@@
 This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz's type models.
-we find (see below) the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
+we find (see below) that transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
 In contrast, \cite{Lashkaripour_JIE2020} finds evidence supporting the ad-valorem assumption for denumerable goods
@@
 Both countries are perfectly symmetric, the labor market equilibria impose an unique international wage that we use as the num\'{e}raire.
-At entry, each firm pays a sunk entry cost $f_e$ and subsequently draws its productivity level $\varphi$ from a known distribution $G(\varphi)$.
+Upon entry, each firm pays a sunk entry cost $f_e$ and subsequently draws its productivity level $\varphi$ from a known distribution $G(\varphi)$.
 At each point in time, a firm can be terminated
 %be hit by a death shock
-at the exogenous probability $\delta>0$.
+with an exogenous probability $\delta>0$.
 The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$, units of labor, with $f$ the fixed cost of production.
 Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ (as in \citealp{melitz}) and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
-This $t$ is our innovation compared to the canonical Mélitz model.
+Introducting $t$ is our innovation compared to the canonical Mélitz model.
 
@@
 Second, we determine the entry conditions on both domestic and exporting markets.\medskip
-\paragraph{Profit-maximizing behaviors} After entry, and conditional on being able to export, firms set a specific price on each destination market.
+\paragraph{Profit-maximizing behaviors} After successful entry on a market (home or foreign), firms set a market-specific price.
 The firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the destination market (see  Appendix \ref{app:theoretical_model}):
@@
 $B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ .
-Profits increase with productivity $\varphi$.Only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
+Profits increase with productivity $\varphi$.
+Only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
 Further, in presence of trade costs, profits for a given productivity are lower on the export market than on the domestic market.
@@
 on the support $\left[ \varphi^\ast_x,\infty\right)$.
-Further defining $M_T = M+M_x$ as the total mass of varieties accessible to the consumers, $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}), with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate:
+Further defining $M_T = M+M_x$ as the total mass of varieties accessible to the consumers and $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}), with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate:
 \begin{equation}
@@
 This is no longer the case in the presence of additive costs (see Appendix \ref{app:theoretical_model}).
-In this case, we need to rely on the model's simulations to determine the steady state.
+In this case, we need to rely on simulations to determine the steady state.
 This drives us to calibrate the structural parameters of the model.
-To do so, we usevalues commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
+To do so, we use values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
 
@@
 
-We are interested in studying the normative consequences of trade liberalization in presence or not of additive costs.
+We are interested in studying the welfare consequences of trade liberalization in presence or not of additive costs.
 To that aim, we will compare the equilibrium based on the values of transport costs at the beginning (1974) and at the very end of our estimation period (2019), in two settings: the first one including only ad-valorem transport costs (uneven columns in Table \ref{tab:calib_TC}).
@@
 Second, total transport costs have strongly reduced over 1974-2019, and the decrease is particularly marked in the additive component in air transport.
-We exploit the theoretical implications of these empirical findings by running a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
+We study the theoretical implications of these empirical findings by running a comparative statics analysis following a reduction in variable transport costs based on our estimates in 1974 and 2019, by transport mode.
 %Specifically, we contrast two cases: In the first scenario, we assume that only ad-valorem costs apply.
@@
 
-The results are reported in Table \ref{tab:resultats_modele}, panel (2), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
+The results are reported in Table \ref{tab:resultats_modele}, panel (2), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade cost reduction, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
 The whole set of results is available upon request to the authors.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
@@
 % whatever the transport mode (and the associated magnitude of the transport costs decrease).
-The originality of our results rather dwells on comparing the welfare gains depending on the type of cost reduction.
+The originality of our results rather rests on comparing the welfare gains depending on the type of cost reduction.
 Welfare gains are higher when the trade cost reduction includes a decrease in additive trade costs.
@@
 Quantitatively, the welfare gains are sizeable.
-In our preferred scenario, the reduction in variable transport costs in Air transport generates a relative welfare gain 40\% higher when such a reduction partly occurs through the additive component, rather than when it happens through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)).
+In our preferred scenario, the reduction in variable transport costs in Air transport generates a 40\% higher welfare gain  when such a reduction partly occurs through the additive component, rather than when it happens through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)).
 In maritime transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
@@
 What are the mechanisms?
-Recalling that welfare is the inverse of the price index, the explanation can be understood decomposing the welfare change in its two components, the variety effect and the price competitiveness effect.
+Recalling that welfare is the inverse of the price index, it is useful to decompose the welfare change in its two components, the variety effect and the price competitiveness effect.
 Considering Equation (\ref{eq:Welfare}), it is straightforward to get the welfare change as:
@@
 
-Consider first the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (1), Columns (A) and (C)).
-The reduction in the iceberg cost induces a negative variety effect, which is yet
+Consider first the benchmark model \textit{\`{a} la} M\'{e}litz with only ad-valorem costs (Table \ref{tab:resultats_modele}, Panel (1), Columns (A) and (C)).
+The reduction in the ad-valorem cost induces a negative variety effect, which is yet
 more than counteracted by a positive competitive effect.
 This is accounted for as follows.
-With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit.
+With the reduction in $\tau$, more foreign firms can enter the domestic market, and this decrease the expected average profit.
 This raises the productivity threshold to enter the domestic
-market. Accordingly, the number of domestic active firms reduces, even if a larger number
+market. Accordingly, the number of domestic active firms decreases, even if a larger number
 of them is able to export. As a whole, the total number of varieties accessible to the
@@
 competitiveness effect dominates, inducing positive welfare gains with the reduction in
-iceberg costs, consistently with the analytical results of \cite{melitz}.
+ad-valorem costs, consistently with the analytical results of \cite{melitz}.
 
@@
 also affects the \textit{classic} selection effect at the extensive margin at the heart of the M\'{e}litz
-(2003) model. As the expected export profit increases, it is necessary to be more productive
-to satisfy the free-entry condition. Everything else equal, the threshold productivity
+(2003) model. As the expected export profit increases, more foreign firms enter the domestic market and it is necessary to be more productive
+to enter it. Everything else equal, the threshold productivity
 on the domestic market rises, as well as the average productivity. Accordingly, the
@@
 a more negative variety effect that is welfare-detrimental everything else equal. As for the
-pure-iceberg case though, the strength of the (larger) competitiveness effect dominates,
+pure-ad-valorem case though, the strength of the (larger) competitiveness effect dominates,
 inducing positive welfare gains, and even larger when the trade liberalization involves a
@@
 This, in turn, has important consequences in normative terms.
-Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 60\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.66\% increase in welfare, much higher than if only iceberg costs were modeled (2.6\%).
+Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 60\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.66\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.6\%).
 A similar result applies with the observed 57\% reduction in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world).
@@
 Interestingly, the value of the $\alpha$ parameter is on average -0.36, lower than in our benchmark on SITC5 data.
-At this highly disaggregated level, pricing-to-market behavior seems less strong, indicating that the average firm is smaller than its average SITC5 counterpart.
+At this highly disaggregated level, pricing-to-market behavior seems weaker, indicating that the per-product average firm is smaller than the average per-sector one.
 In addition, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006.

### 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Encore des changements
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis (\cites{samuelson1954}),\footnote{``Iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.} a recent strand of the empirical literature points to the existence of an additive component, dependent on the physical quantities traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
-Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip
+Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or vessel transport: only waterborne and airborne shipments are included).\smallskip
 
@@
 By allowing for country-product-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel.
-Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows.
+Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in vessel transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows.
 Importantly, the difference can be attributed to our more flexible modeling of the additive component.
@@
 Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
-On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to around 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.13 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
+On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to around 10 cents (0.1 USD) per kilogram exported in vessel transport, and 1.13 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
 Expressed in terms of the export price, the additive costs amount to 2.8\% and 1.7\% for ocean shipping and air transport respectively.
@@
 \cites{Irrazabal_2015} results on the year 2004 imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
-For the same export price multiplied by the ad-valorem cost of 100 USD, our results for 2004 yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
+For the same export price multiplied by the ad-valorem cost of 100 USD, our results for 2004 yield the value of 2.8 (1.5) USD paid in additive transport costs for vessel (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
 Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.
@@
 Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
-Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+Considering the raw series (plain line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
 This is particularly the case for the multiplicative component (panel (d)).\smallskip
@@
 
-The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping.
+The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and vessel shipping.
 If anything, we find the opposite result here: trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
@@
 In our preferred scenario, the reduction in variable transport costs in Air transport generates a 40\% higher welfare gain  when such a reduction partly occurs through the additive component, rather than when it happens through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)).
-In maritime transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
-This can be set in connection with the fact that in maritime transport, ad-valorem costs have reduced more than the additive ones, implying a larger $\beta$ at the end of the period.\footnote{See Appendix \ref{app:theoretical_model} for a similar result.}
+In vessel transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
+This can be set in connection with the fact that in vessel transport, ad-valorem costs have reduced more than the additive ones, implying a larger $\beta$ at the end of the period.\footnote{See Appendix \ref{app:theoretical_model} for a similar result.}
 %Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem.
@@
 pure-ad-valorem case though, the strength of the (larger) competitiveness effect dominates,
-inducing positive welfare gains, and even larger when the trade liberalization involves a
+inducing positive welfare gains. These gains are larger when the trade cost reduction involves a
 reduction in additive costs.\medskip
@@
 First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
-While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, higher by around 50\% relative to the case when only the two components of the variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively.
+While this is not a surprising result \textit{per se} (we now reduce two sources of international frictions, rather than one), this exercise delivers a quantification of the welfare gains associated with a reduction in the fixed export cost.
+They are substantial, higher by around 50\% relative to the case when only variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively.
 Put differently, the reduction of the sunk export cost accounts for the majority of welfare gains.
 This is not the case under additive variable costs; in this case, the reduction of the distortion due to the additive costs continues to substantially matter in the welfare-enhancing effect of the globalization process.}  This can be accounted for as follows.
-The reduction in the fixed export cost pushes more firms to export everything else equal.
+The reduction in the fixed export cost pushes more firms to export, everything else being equal.
 This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market.
@@
 Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods.
-This, in turn, has important consequences in normative terms.
+This, in turn, has important consequences in welfare terms.
 Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 60\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.66\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.6\%).
-A similar result applies with the observed 57\% reduction in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world).
+A similar result applies with the observed 57\% reduction in total variable costs in vessel transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world).
 Neglecting the existence and long-term evolution additive costs thus leads to a substantial underestimation of the welfare gains of trade cost reduction.
@@
 Empirically, we exploit the differences between the import and the export prices taken from the US import database over 1974-2019, to estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean).
-We find that additive costs are quantitatively sizable: On average over the period, they amount to 2.8\% of the export price unit values for ocean shipping, and 1.7\% for air transport, accounting for 30\% and 45\% of total transport costs in air and maritime transport respectively.
+We find that additive costs are quantitatively sizable: On average over the period, they amount to 2.8\% of the export price unit values for ocean shipping, and 1.7\% for air transport, accounting for 30\% and 45\% of total transport costs in air and vessel transport respectively.
 
@@
 Allowing for a varying share of additive costs in the product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role.
-If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in maritime transport.
+If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in vessel transport.
 
@@
 
-\section{Data Appendix \label{app:data}}
+\section{Data Appendix: customs value and import charges\label{app:data}}
 
@@
 The results for vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.
+Full annual data are available in the Online Appendix.
 
@@
 Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
-On average over the whole period, the $R^{2}$ doubles when per-kg costs are included for air, and increases by 50\% for vessel.
+On average over the whole period, the $R^{2}$ doubles when additive costs are included for air, and increases by 50\% for vessel.
 Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
@@
 \footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix B.2.
-If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.}
+It is slightly more favorable to the model including additive costs.}
 %This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
@@
 The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation.
+It is true that the median is at 6.63, indicating that estimates may be noisier for more than half of the considered years.
+That said, the use of the lagged fas price as additional instrument ensures an overall strong predictive power of the first-stage equation, and consequently, accurate (i.e.
+less noisy) second-stage estimates.
+
 In addition, Figure \ref{fig: FS_IV_SITC5} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a very small number of occasions, corresponding to abrupt fluctuations in oil prices.
@@
 Results are qualitatively very similar to those reported on the longer-time-span, more-aggregated-SITC5 dataset previously reported.
-The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average close to 12, slightly higher than in our benchmark.
+The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average close to 12, with a median closer to 10.
+In both cases, these numbers are slightly higher than in our benchmark.
+
 As previously, Figure \ref{fig: FS_IV_HS10} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a few years which seem again to coincide with significant fluctuations in oil prices.

### 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Quelques révisions de plus
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

--- 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 
-Denoting as usual $i$ the origin country, $k$ the product at the 5 digit level, we assume that the fas price $\widetilde{p}_{ik}$ decomposes in two components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
+Denoting as usual $i$ the origin country, $k$ the product at the 5 digit level, we assume that the fas price $\widetilde{p}_{ik}$ is a function of components, say $\bar{p}_{ik}$ the ``firm-specific'' price (related to its cost and pricing strategy) and $\tau^d_{ik}$ the tax duty (out of the firm's hands) according to :
 
@@
 
-\noindent with $s(k)$ the 3-digit classification as a function of the 5-digit product classification.
+\noindent with $s(k)$ the 3-digit sector as a function of the 5-digit product.
 
@@
 Note that, consistently with the second stage, the first stage will be estimated on a year-by-year basis, and will incorporate 3-digit-sector fixed effects.
-This implies to consider the following functional form of the first-stage equation:
+This considering the following functional form of the first-stage equation:
 
@@
 
-Finally, we allow for more flexibility in the empirical set-up, by allowing a variable degree of persistence in the fas price, i.e.
-by relieving the constraint of a unit parameter on $\log \widetilde{p}_{ikt-1}$:
+Finally, we allow a variable degree of persistence in the fas price, i.e. by relieving the constraint of a unit parameter on $\log \widetilde{p}_{ikt-1}$:
 

### 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Encore des coquilles
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 
-\noindent with $s(k)$ the 3-digit sector as a function of the 5-digit product.
+\noindent with $s(k)$ the 3-digit sector of the 5-digit product.
+
+The $\alpha$ parameter should fall between $-1$ and $0$, and has a direct theoretical interpretation in terms of ``pricing-to-market'' behavior of firms.
+$\alpha = 0$ corresponds to the case where the considered sector is made of firms not adjusting their fas price to the change in tariff, to cancel out the influence of the tariff change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition.
+$\alpha = -1$ is the other polar case: the sector here displays full pricing-to-market, because it involves large firms fully compensating the tariff change through their margin (see \citealp{Berman_Martin_Mayer_2012}).
+In-between, $\alpha$ $\in$ $]-1,0[$ corresponds to some variable degree of``pricing-to-market'' in the sector, characterized by firms offsetting partly the impact of the tariff change on the fas price by adjusting the producer price in the opposite direction of the tariff change.
 
@@
 Note that in such a formulation, $\log \widetilde{p}_{ikt-1}$ appears as a second instrument for $\log \widetilde{p}_{ikt}$.
-The $\alpha$ parameter should fall between $-1$ and $0$, and has a direct theoretical interpretation in terms of ``pricing-to-market'' behavior of firms.
-$\alpha = 0$ corresponds to the case where the considered sector is made of firms not adjusting their fas price to the change in tariff, that would cancel out the influence of the tariff change on the price paid by the US consumers; this rather corresponds to small firms which do not have enough market power to manipulate their prices following changes in international competition.
-$\alpha = -1$ is the other polar case: the sector here displays full pricing-to-market, because it involves large firms fully compensating the tariff change through their margin (see \citealp{Berman_Martin_Mayer_2012}).
-In-between, $\alpha$ $\in$ $]-1,0[$ corresponds to some variable degree of``pricing-to-market'' in the sector, characterized by firms offsetting partly the impact of the tariff change on the fas price by adjusting the producer price in the opposite direction of the tariff change.
-
-As already mentioned, consistently with the cross-section analysis adopted by now, the first-stage equation is estimated on a yearly basis, and implements 3-digit sector fixed effects ($\lambda_{is(k)}$ ).
+
+As already mentioned, consistently with our cross-section analysis adopted, the first-stage equation is estimated on a yearly basis, and implements 3-digit sector fixed effects ($\lambda_{is(k)}$).
 We then replace the observed series of fas prices in the right-hand side of Equation (\ref{eq:model_IetA}) by the predicted fas prices produced by Equation (\ref{eq:firststage_final}).
-We now report results based on our main database at the SITC 5-digit level, before replicating the exercise on HS10 digit level, consistently with another major robustness implemented in section \ref{sec:agg_main_text} from the main text.
+We first report results based on our main database at the SITC 5-digit level, before replicating the exercise on HS10 digit level, consistently with another major robustness implemented in section \ref{sec:agg_main_text} from the main text.
 
@@
 First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).\smallskip
-
-For the estimated ad-valorem component, we estimate the following equation:
+Our aim is to obtain time fixed effects measuring how costs have changed over time.\smallskip
+
+For the estimated ad-valorem component, we estimate the following equation by transport mode:
 \begin{eqnarray}
-\ln(\widehat{\tau}_{ikt}-1)&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(A)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
+\ln(\widehat{\tau}_{ikt}-1)&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(A)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(C)}+\epsilon_{ikt} \label{eq:compeffects_mult}
 \end{eqnarray}
@@
 the series of $\gamma_t$ differs from one estimation to the other).
-Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
+Note that we apply the same weighting scheme throughout (based on the relative value of the flow).}
 
@@
 \begin{eqnarray}
-\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(A)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
+\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(A)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(C)}+\epsilon_{ikt} \label{eq:compeffects_add}
 \end{eqnarray}
 
-As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (A)), the product dimension (Term (B)) and the pure transport costs time trend (Term (c)).
+As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (A)), the product dimension (Term (B)) and the pure transport costs time trend (Term (C)).
 Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the ad-valorem cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
@@
 \noindent where the subscript $^{add}$ clarifies that this time fixed effect is specific to the additive component.
-However, one should notice that $\Gamma^{add}_t$ has been obtained considering the additive cost in USD, ie subject to the inflation trend.
+However, one should notice that $\Gamma^{add}_t$ has been obtained considering the additive cost in USD, and thus is subject to the inflation trend.
 To eliminate the influence of inflation in the additive cost index, we extract the time trend (e.g, inflation) from the observed fas price series according to:
@@
 $$\widehat{TC}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
-\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the weighted average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}, and $\widetilde{p}_t$ the corresponding weighted average observed fas price.
+\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the weighted average values (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}, and $\widetilde{p}_t$ the corresponding weighted average observed fas price.
 We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
@@
 \paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
-composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices.
+composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices: we cannot add them together.
 In a first step then, starting from these indices, we rebuild the fitted measures of each transport cost component in value (expressed in percentage of the fas price in both cases):
@@
 
-One last difference remains to be commented, concerning the way the additive component of international transport costs is treated.
+One last difference remains, concerning the way the additive component of international transport costs is treated.
 As we show below, it turns out to be key in accounting for the difference of results with \cite{hummels2007}.
@@
 \end{eqnarray*}
-\noindent Taken the exponential, this becomes: $p_{ikt} = \widetilde{p}_{ikt} + \exp(\Delta_{ikt})$ or, equivalently: $p_{ikt} = \widetilde{p}_{ikt} + t_{ikt}$, with $t_{ikt}$ appropriately defined.
+\noindent Taking the exponential, this becomes: $p_{ikt} = \widetilde{p}_{ikt} + \exp(\Delta_{ikt})$ or, equivalently: $p_{ikt} = \widetilde{p}_{ikt} + t_{ikt}$, with $t_{ikt}$ appropriately defined.
 This exactly corresponds to the case of additive costs only.
@@
 
-In this section, we provide more details on the model's solving in the general case, i.e. without specifying any functional form for the distribution of firm productivity.
+In this section, we provide more details on solving in the model in general case, i.e. without specifying any functional form for the distribution of firm productivity.
 
@@
 \end{eqnarray*}
+\noindent with $B$ defined in Equation (\ref{eq:muofphi}).
 
@@
 \end{eqnarray*}
-Making use of Equation (\ref{eq:muofphi}), the price index reads as:
+Using Equation (\ref{eq:muofphi}), the price index reads as:
 \begin{eqnarray*}
@@
 
-Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, one can show that the consumer price index can be rewritten as: comes that:
+Further defining $M_T = M+M_x$ as the total number of varieties accessible to the consumers, one can show that the consumer price index can be rewritten as:
 $$P = \frac{\sigma}{\sigma-1}M_T^{\frac{1}{1-\sigma}}\left[\frac{1}{M_T}\left(M\widetilde{\varphi}^{\sigma-1} + M_x \left( \frac{\widetilde{\varphi}_x}{\tau} \right)^{\sigma-1} \right)  \right]^{\frac{1}{1-\sigma}}$$
@@
 
-Further, total revenue can de expressed as the product of the average sales and the mass of active producers, according to: $E = M\bar{r}$, with $\bar{r}$ average sales.
+Further, total revenue can be expressed as the product of the average sales and the mass of active producers, according to: $E = M\bar{r}$, with $\bar{r}$ average sales.
 Similarly as for profits, it can be expressed as the sum of average sales on each local and foreign market:
@@
 
-Given selection on the export market, the mass of domestic exporters depends on that of domestic active firms according to: $M_x = \rho_x M$.
-In the symmetric equilibrium, the total mass of available varieties is then: $M_T= M+M_x = (1+\rho_x)M$.
+Given selection on the export market, the mass of domestic exporters depends on that of domestic active firms according to: $M_x = \rho_x M$ ($\rho_x$ is defined in Equation (\ref{eq:px})In the symmetric equilibrium, the total mass of available varieties is then: $M_T= M+M_x = (1+\rho_x)M$.
 

### 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte

@@
 Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
-Related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-sector pair.
+Related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge different prices, all other things equal – in our case, for a given country-sector pair.
 In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
@@
 Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure.
-We follow earlier literature (see e.g.
-\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}) by replacing in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties at the product-country level, together with one-year lagged fas prices.
+Following earlier literature (see e.g.
+\citealp{Caliendo_Parro_2015} or \citealp{Lashkaripour-2017}), we replace the export price $\widetilde{p}$ in the right-hand side of Equation (\ref{eq:model_IetA}) by its predicted values stemming from a first-stage equation regressing $\widetilde{p}$ on custom duties at the product-country level, together with one-year lagged fas prices.
 Section \ref{app:first_stage_IV} in Appendix \ref{app:more_results} provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
@@
 This is consistent with e.g.
-\citet{Giri_et_al2021}, who show that sectoral heterogeneity does not always lead to an increase in the gains from trade.
+\citet{Giri_et_al2021}, who show that taking into account sectoral heterogeneity does not always lead to an increase in the gains from trade.
 
@@
 %Two main insights emerge up to now. First, total transport costs have strongly reduced between 1974 and 2019. Second, a
-\textbf{À valider xxxxx GD}
 From 1974 to 2019, additive costs represent an important share of declining transport costs, between 30 and 45\%.

### 2021-11-20 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Relecture Lise article finie

@@
 In all cases, these estimates are very similar.
-This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
+This supports that our benchmark estimates do not suffer from any substantial biases arising from endogeneity concerns.
 
@@
 
+While our benchmark estimates exploit product variability at the 5-digit level, one may wonder about the importance of the disaggregation level in driving the results.
+In this section, we check the robustness to the use of more disaggregate data within sector.
 Current data are available at the 10-digit level for products (HS10) for some specific years.
@@
 Firstly, we could only access the years 1997-1999 and 2001-2019.\footnote{Theoretically, they should be available from 1989, which is already smaller than our full sample starting in 1974.
-Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} Secondly, for the years 1997-1999 and 2001-2005, compatibility issues between Census and Hummels' data prevented reliable comparisons.
+Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} Secondly, for the years 1997-1999 and 2001-2005, compatibility issues between Census and Hummels' data (which we use until 2004) prevented reliable comparisons.
 
@@
 Yet, as Table \ref{tab:baseline10} and 	Figure \ref{fig:robustness_to_agregation} show, the results are very similar when using 10-digit product and 5-digit products.
-If anything, the $\beta$ seems on average slightly higher when estimated based on a 3 (sector)-10 (product) database, indicating a (small) downward bias in our estimates of additive costs.
-But overall, no major difference arises in levels and trends - results are especially very close for the Air transportation mode.
+If anything, the share of additive costs $\beta$ seems on average slightly higher when estimated based on a 3 (sector)-10 (product) database, indicating a (small) downward bias in our estimates of additive costs.
+Overall, no major difference arises in levels and trends - results are especially very close for the Air transportation mode.
 This has important implications for the exercise performed in Section \ref{sec:theory}: welfare gains computations should not be significantly affected by our use of 3 (sector)-5 (product) digit data.
@@
 \noindent where $X_{ikt}$ represents transport costs level due to distance, port quality, etc.
-and $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
+and $\beta$ is the elasticity of transport costs to the export price (in absolute value), also equal to the share of additive costs in total transport costs.
 Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described in more detail in Appendix \ref{app:compare_Hummels}.
@@
 estimated in section \ref{sec:results_decomposition}, we extract the pure transport cost by estimating time fixed effects.
-Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
+Precisely, we extract the changes over time in the pure transport cost series by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974.
 We conduct the same analysis on an overall transport costs measure, built by combining the two estimated components (additive and ad-valorem) in a unified measure of transport costs.
-The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the unfitted transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$, and three series for the fitted values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
-
-One advantage of this method is that it yields measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
-For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) makes it possible to identify if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (keeping country and sector constant) that have reduced over time.
+The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: three time series for the raw transport costs measures $\{\Gamma^{t/\widetilde{p}, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$, and three series for the ``pure'' values $\{\Gamma^{t/\widetilde{p}}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).
+
+One advantage of this method is that it yields measures of transport costs (raw and pure, i.e.
+composition effects excluded) that are easily comparable between transport modes and transport cost components.
+For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the raw measure and the pure transport costs series makes it possible to identify if the decrease observed over the period in the raw series is due to trade composition effects (for instance, changes in the country partners, in the type/quality of products traded), or if it is the pure transport costs (keeping country and sector constant) that have reduced over time.
 For sake of reading clarity, we describe the method to extract these series in Appendix \ref{app:comp-effects}, and we now turn to the results.
@@
 Panels (d), (e) and (f) report the results for vessel.
-In each panel, we report the evolution of transport costs for both the unfitted (plain line) and the fitted (dotted line) measures.
+In each panel, we report the evolution of transport costs for both the raw series (dotted line) and the pure transport costs series (plain line).
 
@@
 
-Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
-First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
-International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in vessel shipping.
-Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
-Further, the transport cost reduction is much smoother in vessel transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
-Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
+The main result from Figure \ref{fig:totalTC_compeffects_excl} is that composition effects do not play a major role in accounting for the time trend of overall transport costs.
+Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the raw transport costs measure and the pure transport costs.
 For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
-Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
+Air transport costs were reduced by around 60\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
 Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
 Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
-Considering the raw series (plain line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
 This is particularly the case for the multiplicative component (panel (d)).\smallskip
@@
 As we explain in more detail in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension.
-Rather than from the actual cif-fas price gap, we start from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition} at the year/sector/origin country level, by transport mode).
+Rather than from the actual cif-fas price gap, we start from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}) at the year/sector/origin country level, by transport mode).
 Put differently, our methodology lets the share of additive costs $\beta$ vary over the three time-product-partner country dimensions.
@@
 This turns out to be of primary importance in the disentangling of the sources of the transport costs time trend.
-To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004).
+To establish this point clearly, we replicate the method adopted by \cite{hummels2007} on our database (which is the same as his until 2004).
 The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}.
-To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In Section E.1 from Online Appendix, we report the two series of transport costs (fitted and unfitted) as percentage of the export price as in \cite{hummels2007}.
+To ease the interpretation of the trends, we express them as indices with the reference value 100 in 1974, in Panels (c) and (d) for air and vessel respectively.\footnote{In Section E.1 from Online Appendix, we report the two series of transport costs (raw and pure) as percentage of the export price as in \cite{hummels2007}.
 Unsurprisingly, this accords with his results displayed in his Figures 5 (for Air) and 6 (for Vessel) for the years in common (until 2004).
@@
 These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}.
-Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted rate than the unfitted rate, especially for vessel (Panel (d)).
-Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panels (c) and (f).\medskip
+Applying \cite{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in vessel shipping, as the downward trend is of higher magnitude for the pure transport cost measure than the raw one (Panel (d)).
+Yet, this result is overturned when we allow for more flexibility in the role of the additive component, as indicated in Figure \ref{fig:totalTC_compeffects_excl}, Panel (f).\medskip
 
@@
 
-In this section, we explore the implications of our empirical results on theoretical grounds.
+In this section, we explore the welfare implications of our empirical results.
 Over the last decade, several papers have examined the implications of non-ad-valorem costs.
@@
 \cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment, and draw welfare conclusions regarding the variations of these costs.
-We have complemented these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
+Our results so far complement these studies by providing a long-run characterization of both the additive and the multiplicative component of transports costs.
 In this section, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases.
-One case assumes that the reduction was achieved solely through the decline of ad-valorem costs, as modeled in standard New Trade theoretical approaches.
-The other case builds on the results of our empirical investigation to assume that the reduction was achieved by a combined reduction in the ad-valorem and the additive components, as modeled by the addition of additive costs to the canonical \citet{melitz} model.\smallskip
+One case assumes that the reduction was achieved solely through the decline of ad-valorem costs, as standardly assumed in New Trade models.
+The other case builds on the results of our empirical investigation to assume that the transport cost reduction was achieved by a combined decrease in the ad-valorem and the additive components.
+To this aim, we amend the canonical \citet{melitz} model with additive costs.\smallskip
 %While the bulk of the trade literature investigates the role of trade costs modelled as ad-valorem costs (e.g., ad-valorem trade costs), not much is said about the role of the additive component.
@@
 In contrast, as in \cite{Irrazabal_2015}, we provide a quantification of such welfare gains.
-As \cite{Irrazabal_2015} provide estimates of transport costs only for one year (2004), their assessment of the welfare effects of a trade cost increase is based on a ad-hoc assumption (specifically, of a 5\% increase in the ratio of additive cost to the median fas price).
+As \cite{Irrazabal_2015} provide estimates of transport costs only for one year (2004), their assessment of the welfare effects of a trade cost increase is based on a ad-hoc assumption (specifically, a 5\% increase in the ratio of additive cost to the median fas price).
 Unlike them, we can exploit our database time span to ground our study in the observed reduction of transport costs from 1974 to 2019.
@@
 This stands in contrast with the conclusion reached by \cite{sorensen2014} (analytically) and \cite{Irrazabal_2015} (quantitatively) in M\'{e}litz's type models.
-we find (see below) that transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
-In contrast, \cite{Lashkaripour_JIE2020} finds evidence supporting the ad-valorem assumption for denumerable goods
+we find that the transport cost reduction implies substantially larger welfare gains when they are achieved through a reduction in the additive component.
 %That said, it should be noted the case with additive costs in \cites{Lashkaripour_JIE2020}paper relies on a negligible share of additive costs.
@@
 While \cite{Lashkaripour_JIE2020} relies on a perfectly competitive environment,
-Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
-In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade cost reduction lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
-Leaving aside the selection issue leads to an underestimation of the welfare gains that can be achieved by a reduction in additive transport costs.
+Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as also did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
+In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+Leaving aside the selection issue might underestimate the welfare gains that can be achieved by a reduction in additive transport costs.
 
@@
 \paragraph{Profit-maximizing behaviors} After successful entry on a market (home or foreign), firms set a market-specific price.
-The firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the destination market (see  Appendix \ref{app:theoretical_model}):
+The firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the foreign destination market (see  Appendix \ref{app:theoretical_model}):
 \begin{eqnarray}
@@
 %a given firm of productivity $\varphi$ makes a lower export profit than on its domestic market, conditional on being productive enough to export.
-We now determine the productivity threshold associated to the entry conditions.
+We now determine the productivity thresholds associated to the entry conditions.
 
@@
 
-Following \cite{melitz} and the subsequent literature, we impose selection on the export market, (i.e.
+Following \cite{melitz} and the subsequent literature, we impose selection on the export market (i.e.
 $\varphi_x^\ast > \varphi^\ast$).
@@
 \begin{equation}
-\left(1-G(\varphi^\ast)\right)\frac{\bar{\pi}}{\delta} = f_e \label{eq:FEC}
-\end{equation}
-As shown in Appendix \ref{app:theoretical_model}, the free-entry market condition can be rewritten as:
+\left(1-G(\varphi^\ast)\right)\frac{\bar{\pi}}{\delta} = f_e, \label{eq:FEC}
+\end{equation}
+\noindent with $f_e$ the sunk entry cost. As shown in Appendix \ref{app:theoretical_model}, the free-entry market condition can be rewritten as:
 \begin{equation}
@@
 
-Following a standard assumption in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the distribution in firm productivity is assumed to follow a Pareto distribution with parameters $(\varphi_{min},k)$.
+Following a standard assumption in this literature (\cite{Irrazabal_2015}, \cite{melitz-redding-Handbk-IT-2014}), the distribution in firm productivity is assumed to follow a Pareto distribution with parameters $(\varphi_{\text{min}},k)$.
 This allows a full analytical solving of the M\'{e}litz model.
@@
 \end{center}
-{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
-$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s value) and the final steady state (39\%, based on 2006 value), according to the references provided above.}}}
+{\parbox[l]{12cm}{ \vspace{2pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
+$f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, from \cite{BEJK-AER-03} (BEKJ 2003 for short)) based on early 1980s value) and the final steady state (39\%, based on 2006 value), according to the references provided above.
+}}}
 \end{table}
@@
 
-We are interested in studying the welfare consequences of trade liberalization in presence or not of additive costs.
-To that aim, we will compare the equilibrium based on the values of transport costs at the beginning (1974) and at the very end of our estimation period (2019), in two settings: the first one including only ad-valorem transport costs (uneven columns in Table \ref{tab:calib_TC}).
+We are interested in studying the welfare consequences of a trade cost reduction in presence or not of additive costs.
+To that aim, we will compare the equilibrium based on the values of transport costs at the beginning of our estimation period (1974) and at the very end (2019), in two settings: the first one including only ad-valorem transport costs (uneven columns in Table \ref{tab:calib_TC}).
 We here apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
@@
 The reduction in international transport costs channels through both its additive and ad-valorem components.
-Besides, \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
+Besides, \cite{BEJK-AER-03} and \cite{Lincoln_McCallum2018} documents the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
 This can be captured in the model through a reduction in the fixed export cost $f_x$ (as reported in Table \ref{tab:calib_horsTC}).
@@
 
-The results are reported in Table \ref{tab:resultats_modele}, panel (2), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade cost reduction, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
+The results are reported in Table \ref{tab:resultats_modele}, panel (2), by transport mode (for Air, Columns (A)-(B) and Vessel, Columns (C)-(D) respectively).\footnote{For sake of space saving, we only report the welfare consequences of trade cost reduction, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
 The whole set of results is available upon request to the authors.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
@@
 %, and the larger welfare gains when this specific component of international trade costs is reduced.
-See Appendix \ref{app:theoretical_model}, Table \ref{tab:results_model_appendix} for the results}.
+See Appendix \ref{app:theoretical_model}, Table \ref{tab:results_model_appendix} for the results.}.
 
@@
 This is accounted for as follows.
-With the reduction in $\tau$, more foreign firms can enter the domestic market, and this decrease the expected average profit.
-This raises the productivity threshold to enter the domestic
+With the reduction in $\tau$, more foreign firms can enter the domestic market, and this decreases the expected average profit.
+This, in turn, requires a higher productivity threshold to enter the domestic
 market. Accordingly, the number of domestic active firms decreases, even if a larger number
@@
 also affects the \textit{classic} selection effect at the extensive margin at the heart of the M\'{e}litz
-(2003) model. As the expected export profit increases, more foreign firms enter the domestic market and it is necessary to be more productive
-to enter it. Everything else equal, the threshold productivity
+(2003) model. As the expected export profit increases, more foreign firms enter the domestic market which increases the entry difficulty. Everything else equal, the threshold productivity
 on the domestic market rises, as well as the average productivity. Accordingly, the
@@
 Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the
-share of exporters. As the selection on the domestic market is also more stringent, the
+share of exporters. As selection on the domestic market is also more stringent, the
 mass of total active firms reduces by a larger amount. From the consumer side, this induces
@@
 First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
-While this is not a surprising result \textit{per se} (we now reduce two sources of international frictions, rather than one), this exercise delivers a quantification of the welfare gains associated with a reduction in the fixed export cost.
-They are substantial, higher by around 50\% relative to the case when only variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively.
+While this is not a surprising result \textit{per se} (we now reduce the reductions of two sources of international frictions, rather than one), this exercise delivers a quantification of the welfare gains associated with a reduction in the fixed export cost.
+They are substantial, higher by around 50\% relative to the case when only variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are even higher, by 66\% (Vessel) and 89\% (Air) respectively.
 Put differently, the reduction of the sunk export cost accounts for the majority of welfare gains.
@@
 Quantitatively, the effect remains sizeable: Welfare gains are between 14\% and 40\% higher when the globalization process involves a reduction in additive costs.
-Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
-In the end, the welfare gains are the highest when trade costs reduction relies on a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
+Put differently, the effect of the reduction in the fixed export cost does not seem to depend on the composition of the variable costs.
+In the end, the welfare gains are the highest when the globalization process materializes through a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip
 
@@
 Both AIC and log-likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
-These results holds whatever the considered year.
-\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix B.2.
+These results holds whatever the considered year.\footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix B.2.
 It is slightly more favorable to the model including additive costs.}
@@
 
-We report here the first-stage procedure and estimates underlying the IV estimates reported in section \ref{sec:IV_main_text}.
+We report here the first-stage procedure and estimates underlying the IV estimates reported in Section \ref{sec:IV_main_text}.
 We start by detailing the theoretical intuitions behind the first-stage equation.
@@
 $\alpha = -1$ is the other polar case: the sector here displays full pricing-to-market, because it involves large firms fully compensating the tariff change through their margin (see \citealp{Berman_Martin_Mayer_2012}).
-In-between, $\alpha$ $\in$ $]-1,0[$ corresponds to some variable degree of``pricing-to-market'' in the sector, characterized by firms offsetting partly the impact of the tariff change on the fas price by adjusting the producer price in the opposite direction of the tariff change.
+In-between, $\alpha$ $\in$ $]-1,0[$ corresponds to some variable degree of ``pricing-to-market'' in the sector, characterized by firms offsetting partly the impact of the tariff change on the fas price by adjusting the producer price in the opposite direction of the tariff change.
 
@@
 
-We begin by describing the estimation method to extract the fitted and unfitted series for both the additive and the multiplicative components, before turning to the `total costs'' series.
+We begin by describing the estimation method to obtain the pure and raw series for both the additive and the multiplicative components, before turning to the `total costs'' series.
 \smallskip
 
-\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e.
-extracting from trade composition effect) can be described as a two-stage process.
+\paragraph{Obtaining the ``pure'' transport costs component series.} The empirical strategy to find the pure transport cost measure (i.e.
+exclude trade composition effect) can be described as a two-stage process.
 First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
-Our aim is to obtain time fixed effects measuring how costs have changed over time.\smallskip
+Our aim is to obtain time fixed effects measuring how costs have changed over time, controlling for changes in the country-product composition structure.\smallskip
 
@@
 
-In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
-This constitutes the second stage of our procedure.
-Based on Equation (\ref{eq:compeffects_mult}), we deduce after estimation that:
-
-\[\left\{
-  \begin{array}{lcl}
-  \text{For the year 1974:} &&  \widehat{\tau}_{is74} -1 = \exp(\delta +\alpha_i+\beta_s), \\
-  \text{For any year }t> 1974:&& \widehat{\tau}_{ist} -1 = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t) \\
-  \end{array}
-\right.\]
-
-From this, we obtain the following recursive link: $\widehat{\tau}_{ist} -1 = (\widehat{\tau}_{is74}-1)\exp(\gamma_t)$.
-Given the constraint $\tau>1$, we rewrite to get the percentage change between year 1974 and any year $t>1974$: $\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1}  = 100\exp(\gamma_t)$.
-As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the time trend, such that we can rewrite:
-\begin{equation}
- \Gamma^{adv}_t= 100\exp(\gamma^{adv}_t)  \label{eq:tcadv_compoeffect}
-\end{equation}
-
-\noindent where the subscript $^{adv}$ clarifies that this time fixed effect is specific to the ad-valorem component. \medskip
-
 As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g.
@@
 Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the ad-valorem cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
-Both equations are estimated by transport mode.\smallskip
-
-After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
+Both equations are estimated by transport mode.
+
+In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
+This constitutes the second stage of our procedure.
+Based on Equation (\ref{eq:compeffects_mult}), we deduce after estimation that:
+
 \[\left\{
   \begin{array}{lcl}
-\text{For the year 1974:}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
-\text{For any year}~t> 1974:&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
+  \text{For the year 1974:} &&  \widehat{\tau}^p_{is74} -1 = \exp(\delta +\alpha_i+\beta_s), \\
+  \text{For any year }t> 1974:&& \widehat{\tau}^p_{ist} -1 = \exp(\delta +\alpha_i+\beta_s)\times \exp(\gamma_t), \\
   \end{array}
@@
 
-From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
-
-Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)$.
+\noindent where the subscript $^{p}$ indicates the predicted value.
+From this, we obtain the following recursive link: $\widehat{\tau}^p_{ist} -1 = (\widehat{\tau}^p_{is74}-1)\exp(\gamma_t)$.
+Given the constraint $\tau>1$, we rewrite to get the percentage change between year 1974 and any year $t>1974$: $\Gamma_{ist} = 100.\frac{\widehat{\tau}^p_{ist}-1}{\widehat{\tau}^p_{is74}-1}  = 100\exp(\gamma_t)$.
+As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the time trend, such that we can rewrite:
+\begin{equation}
+ \Gamma^{adv}_t= 100\exp(\gamma^{adv}_t)  \label{eq:tcadv_compoeffect}
+\end{equation}
+
+\noindent where the subscript $^{adv}$ clarifies that this time fixed effect is specific to the ad-valorem component.
+
+We adopt a similar reasoning for the additive cost component. After estimating Equation (\ref{eq:compeffects_add}), it can be rebuilt according to:
+\[\left\{
+  \begin{array}{lcl}
+\text{For the year 1974:}&&\widehat{t}^p_{is74}=  \delta + \alpha_i+ \beta_s, \\
+\text{For any year}~t> 1974:&&\widehat{t}^p_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t),
+\end{array}
+\right.\]
+
+\noindent where the subscript $^{p}$ indicates the predicted value. From this, we deduce the recursive link: $\widehat{t}^p_{ist} = \widehat{t}^p_{is74} \times \exp(\gamma_t)$. 
+Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}^p_{ist}}{\widehat{t}^p_{ik74}} = 100\exp(\gamma_t)$.
 Again, it is independent of the product-origin country pair.
@@
 \begin{eqnarray}
-\ln(\widehat{p}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(A)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_pfas}
+\ln(\widetilde{p}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(A)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(B)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_pfas}
 \end{eqnarray}
@@
 
-\paragraph{Obtaining the unfitted measures} The objective here is to express the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
+\paragraph{Obtaining the raw measures.} The objective here is to express the raw transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
 To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:
@@
 
-We also build two measures of the ``total'' transport costs for both the unfitted series and the pure transport cost series (i.e.
-composition effects excluded), with the objective to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
-Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
+We also build two measures of the ``total'' transport costs for both the raw and the pure transport cost series, with the objective to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
+Even if obeying the same logic, we proceed slightly differently for the raw versus the pure transport costs measures, as we now explain.
 \smallskip
 
-\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
+\paragraph{Obtaining the raw total transport cost index} For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
 $$\widehat{TC}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
@@
 
-\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to constructing the fitted measure of total transport cost (i.e.
-composition effect excluded), with one notable exception: Fitted ad-valorem and additive components have not been estimated in value, but extracted and built as indices: we cannot add them together.
-In a first step then, starting from these indices, we rebuild the fitted measures of each transport cost component in value (expressed in percentage of the fas price in both cases):
+\paragraph{Obtaining the pure total transport cost index} The same logic as above applies to constructing the pure measure of total transport cost (i.e.
+composition effect excluded), with one notable exception: Pure ad-valorem and additive components have not been estimated in value, but extracted and built as indices; accordingly, we cannot add them together.
+In a first step then, starting from these indices, we rebuild the pure measures of each transport cost component in value (expressed in percentage of the fas price in both cases):
 
@@
 \noindent with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_s \beta_s)$ the (weighted) mean ad-valorem transport cost in 1974 (based on Equation (\ref{eq:compeffects_mult})) and $\bar{t}_{1974} = \delta + \sum_i \alpha_i + \sum_s \beta_s$ the (weighted) mean additive transport cost in 1974 (based on Equation (\ref{eq:compeffects_add})).
-We then deduce the fitted value of the overall cost according to:
+We then deduce the value of the pure overall cost according to:
 $$\widehat{TC}^{pure}_t= \widehat{\tau}^{pure, adv}_t -1 + \frac{\widehat{t}^{pure}_t}{\widetilde{p}_t}$$
 
-Last, we transform this (fitted) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
+Last, we transform this (pure) measure of total transport cost in an index with base year 1974, applying a similar formula as above (by transport mode):
 $$\Gamma^{tc}_t = 100\frac{\widehat{TC}^{pure}_t}{\widehat{TC}^{pure}_{1974}}$$
@@
 
-Consider now the ``fitted ad-valorem rate''.
-\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels, 2007 terminology) to $i$ (our terminology).} \textit{at time $t$.
+Consider now the ``pure ad-valorem rate''.
+\cite{hummels2007} uses ``\textit{a regression in which the dependent variable is the ad-valorem air freight cost in logs for commodity $k$ shipped from exporter} $i$\footnote{To be consistent with our notations, we change the country subscript from $j$ (Hummels' (2007) terminology) to $i$ (our terminology).
+Also notice that the ``unfitted'' (``fitted'') measure in Hummels (2007) refers to the raw (pure) measure in our terminology.} \textit{at time $t$.
 The independent variables include a separate intercept for each exporter-commodity shipped, the weight/value ratio in logs for each shipment, and year dummy variables}''.
@@
 
-After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, denoted $\widehat{TC}_{t}$, to be compared to the unfitted ad-valorem rate $TC_{t}$ (by transport mode).\medskip
+After estimating Equation (\ref{eq:Hummels1}), \cite{hummels2007} uses the predicted value of the regression, denoted $\widehat{TC}_{ikt}$, to obtain the unweighted average in the product/origin country dimension $i,k$, denoted $\widehat{TC}_{t}$, to be compared to the raw ad-valorem rate $TC_{t}$ (by transport mode).\medskip
 
 What are the main differences with our estimation strategy? Five points may be underlined.
-First, \cite{hummels2007} obtains the fitted ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the fitted transport cost (ie, composition effect excluded) from the (already estimated) unfitted rate.
+First, \cite{hummels2007} obtains the pure ad-valorem rate in one step (considering the observed export-import price gap on the left-hand side of Equation (\ref{eq:Hummels1})), while we use our two-step approach to extract the pure transport cost series (ie, composition effect excluded) from the (already estimated) raw rate.
 However, this is a slight difference, as in both cases it ultimately amounts to taking the predicted value of Equation (\ref{eq:Hummels1}) (which eliminates the changes specific to the origin country/product/year triplet).
-Second, in contrast to us, \cite{hummels2007} does not purge his measure of the fitted ad-valorem rate by the country-sector fixed effects.
+Second, in contrast to us, \cite{hummels2007} does not purge his measure of the pure ad-valorem rate by the country-sector fixed effects.
 This is not likely to make a substantial difference, however, as these fixed effects are by nature constant over time. Accordingly, they only amount to a scale effect.
@@
 
-Fourth, we differ in the weighting scheme retained to obtain the yearly value of the fitted transport cost (in the terms of \cite{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$).
+Fourth, we differ in the weighting scheme retained to obtain the yearly value of the pure transport cost (in the terms of \cites{hummels2007} method, the switch from $\widehat{TC}_{ikt}$ to $\widehat{TC}_{t}$).
 \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow.
 We proceed differently, as we weight each flow by its relative value on total trade flows (on the relevant year of observation), based on the argument that this weighting scheme does not overweight the small flows in value.
-However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average fitted transport costs rate as in \cite{hummels2007}, as reported in the following Section \ref{app:compeffects_weights}.
+However, we ensure that our results are not sensitive to an alternative weighting scheme, by building the average pure transport costs rate as in \cite{hummels2007}, as reported in the following Section \ref{app:compeffects_weights}.
 
@@
 As noted by \cite{hummels_skiba}, $0<\beta<1$ represents the elasticity of freight costs to the export prices, increasing with the relative weight of the additive component in international transport costs.
-By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, country partner or product.
+By estimating Equation (\ref{eq:Hummels1}) with a constant $\beta$, \cite{hummels2007} assumes that the share of additive costs does not vary over time, exporter country or product.
 This is an important difference with our method, as we measure transport costs by explicitly allowing for an additive component that varies in the three dimensions (i.e.
@@
 
-To this aim, we rebuild the fitted values for each of our transport cost measures (ad-valorem, additive and overall) applying the \cite{hummels2007} weighting scheme (i.e.
+To this aim, we rebuild the pure values for each of our transport cost measures (ad-valorem, additive and overall) applying the \cite{hummels2007} weighting scheme (i.e.
 taking the unweighed average value over the $i,k$ dimension), which we then express as indices, with the reference value 100 in 1974.
@@
 This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise.
-However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (A), overall transport costs have decreased more than the fitted component).
+However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in air transport (in Figure \ref{fig:compeffects_robustness}, panel (A), overall transport costs have decreased more than the pure component).
 This stands in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than partly offsetting this decline, as argued by \cite{hummels2007}.
@@
 
-Making use of the above elements, we finally get the zero-cut off profit condition in the open economy with additive costs as:
+Making use of the above elements, we finally get the zero-cut off profit condition in the open economy with additive costs as in Equation (\ref{eq:ZCP}).
 
@@
 
-Following \cite{melitz} and the subsequent literature, denoting $M_e$ the mass of prospective entrants and $\rho_e$ the probability of successful entry (equal to $1-G(\varphi^\ast)$), considering only steady-state equilibrium needs to verify the entry-exit flow equilibrium condition:
+Following \cite{melitz}, we denote $M_e$ the mass of prospective entrants and $\rho_e$ the probability of successful entry (equal to $1-G(\varphi^\ast)$).
+At the steady-state equilibrium, $\rho_e$ needs to verify the entry-exit flow equilibrium condition:
 $$\rho_e M_e = \delta M$$.
@@
 Under a constant fixed export costs, we simulate the model under two alternative cases.
-Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^fas$, maintaining the ad-valorem component at its initial (1974) value (Scenario (A)), and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974 (Scenario (B)).
+Starting from the values observed in 1974, we simulate \textit{(i)} the observed reduction in the additive cost $t/\widetilde{p}^{fas}$, maintaining the ad-valorem component at its initial (1974) value (Scenario (A)), and  \textit{(ii)} the opposite case where $\tau$ takes its 2019 value with the same additive cost (in percentage of the fas price) as in 1974 (Scenario (B)).
 These two scenarios (as well as the initial one based on 1974 values) are reported in Table \ref{tab:calib_TC_appendix}.
@@
 
-The quantitative results strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces (Scenario (A)).
+The welfare results are reported in Table \ref{tab:results_model_appendix}.
+They strongly confirm the welfare-detrimental effects of additive costs, and the larger welfare gains when this specific component of international trade costs reduces (Scenario (A)).
 In contrast, under Scenario (B), the reduction in total trade costs is still welfare-enhancing, but the magnitude of the welfare gains is much more modest, because of the larger distortion induced by additive costs - that now account for a larger share of total trade costs.
@@
 \hspace{1em}Variety effect (in \%)	&-2.40&	-0.28&	-1.60	&-0.42 \\
-\hspace{1em}Price effect (in \%)	&5.31 &0.84&	4.38	&1.393 \\
+\hspace{1em}Price effect (in \%)	&5.31 &0.84&	4.38	&1.39 \\
 \hline \hline

### 2021-11-22 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
corrections diverses

@@
 The paper compares the empirical performances of two models: one with only ad-valorem costs (Model (A)) and one with where both ad-valorem and additive costs (Model (B)).
-In the interest of comprehensiveness,  we also estimate the model with only additive costs (Model (C)), in which case the estimated equation is:
+For comparison purposes,  we also estimate the model with only additive costs (Model (C)), in which case the estimated equation is:
 $$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$$
@@
 This section is devoted to presenting the results.
-Precisely, we first compare the estimates of the transport costs components under the three models.
+We first compare the estimates of the transport costs components under the three models.
 In a second step, we report quality of fit tests for the three models.
@@
 
+\newpage
+\setcounter{table}{0}
+\renewcommand{\thetable}{B.\arabic{table}}
+
 \subsection{Estimation at the 4-digit level \label{app:4digit}}
@@
 Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively.
+Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs, already established at the 3-digit level.
 
@@
 Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
+\multicolumn{7}{l}{\textbf{Model (A)} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
 \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}  \\ \hline
+\multicolumn{7}{l}{\textbf{Model (B)}}  \\ \hline
 \multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)} \\
+\multicolumn{7}{l}{\textbf{Model (A)} ($\widehat{\tau}^{ice}-1$, in \%)} \\
 \hline
@@
 \hline
-\multicolumn{7}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC} }\\ \hline
+\multicolumn{7}{l}{\textbf{Model (B)} }\\ \hline
 \multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
@@
 The results are reported in Tables \ref{tab:good_fit_air_rob} (for air) and \ref{tab:good_fit_ves_rob} (for vessel).
+
 \begin{table}[htbp]
@@
 \multicolumn{7}{l}{\textbf{R$^2$}}      \\ \hline
-Model (a)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Model (b) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
+Model (A)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
+Model (B) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
 \multicolumn{7}{l}{\textbf{SER} }  \\ \hline
-Model (a)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Model (b) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
+Model (A)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
+Model (B) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
 \multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (a) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
-Model (b) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
+Model (A) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
+Model (B) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
 \multicolumn{7}{l}{\textbf{AIC criteria} }  \\ \hline
-Model (a)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
-Model (b) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
+Model (A)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
+Model (B) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
 \multicolumn{7}{l}{\textbf{Test LL} }    \\ \hline
-2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
-\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
-p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
+Stat LL ratio (B vs A) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
+\# of restrictions (B vs A) & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
+p-value (B vs A) & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
 \hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (a) = with only ad-valorem transport costs.
-Model (b) = with additive \& ad-valorem
-transport costs.
+\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
+Model (B) = with additive \& ad-valorem transport costs.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
@@
 \multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
-Term I only & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
-Terms A \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
+Model (A) & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
+Model (B) \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
 \multicolumn{7}{l}{\textbf{SER} }\\ \hline
-Model (a) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (b) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
+Model (A) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
+Model (B) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
 \multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (a)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
-Model (b) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
+Model (A)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
+Model (B) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
 \multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
-Model (a) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
-Model (b) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
+Model (A) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
+Model (B) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
 \multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
-2$\times$(ll(UR) -ll(R)) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
-\# restrictions  & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
-p-value & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
+Stat LL ratio (B vs A) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
+\# of restrictions (B vs A) & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
+p-value (B vs A) & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
 \hline\hline
 \multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& ad-valorem
-transport costs.
+Model (B) = with additive \& ad-valorem transport costs.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
@@
 When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
-In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
-Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.
+In other words, if using a more disaggregated product classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
 
@@
 %\subsection{Detailed Results \label{secoa:detailed}}
+
 \newpage
 \setcounter{table}{0}
-\renewcommand{\thetable}{B.\arabic{table}}
+\renewcommand{\thetable}{C.\arabic{table}}
 
@@
 Table \ref{tab_oa:result_ves_ally3} reports the results for each year over 1974-2019 for vessel; Table \ref{tab_oa:result_air_ally3} reports similar results for air transport.
-In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)), when both additive and ad-valorem costs are modeled (Model (B)) and when only additive costs are modeled (Model (A)).
+In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)), when both additive and ad-valorem costs are modeled (Model (B)) and when only additive costs are modeled (Model (C)).
 In all tables, statistics are weighted by value.
@@
 Still, this does not make much difference.
-When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
-The weighed median value is left unchanged at 2.9\% whether 1989 is included or not.
+When 1989 is included, the weighted mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
+The weighted median value is left unchanged at 2.9\% whether 1989 is included or not.
 
@@
 This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
-This provides an indirect robustness check to the degree of classification we have retained to estimate international transport costs.
+This provides an indirect robustness check to the degree of classification we have selected to estimate international transport costs.
 Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
@@
 
-\paragraph{Comparison on a yearly basis} We investigate this comparison deeper on a year-to-year basis, as shown in Figure \ref{fig:robustesse_non_separe}.
-Specifically, Figure \ref{fig:robustesse_non_separe} displays the results for air transport in panel (A) and for maritime transport in panel (B).
+\paragraph{Comparison on a yearly basis} Figure \ref{fig:robustesse_non_separe} displays the results on a yearly basis for air transport in panel (A) and for maritime transport in panel (B).
 In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
@@
 Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
-Along with the quality-of-fit diagnosis, this confirms the robustness of our estimation results to this assumption.
+Along with the quality-of-fit analysis, this confirms the robustness of our estimation results to this assumption.
 
@@
 This does not solve the issue yet, for two main reasons.
-First, it is still constrained to resort to non-linear estimators.
+First, it still requires a non-linear estimator.
 This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
@@
 Should we relax this restriction, standard linear, least squares estimates often deliver negative, meaningless estimates.
-In this respect, implementing this does not suppress the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces).
+In this respect, the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces) remains.
 
 It should also be noted that this estimation method implies a lower coverage, both in time and in the sectors/countries covered.
-Information about the port of entry are only available since 1989, which is already smaller than our full sample starting in 1974.
-On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
-Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
-In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
+Information about the port of entry is only available since 1989, which is already smaller than our full sample starting in 1974.
+On top of that, because of the Covid situation, the US Bureau of Census was only able to provide us the data for the years 1997-1999 and 2001-2019.
+Implementing the referee's method would hence reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
+In our view, the historical coverage is interesting \emph{per se}: it provides useful insights about how transport costs have evolved over several decades and more importantly, it allows us to highlight the welfare alterations induced over the ``hyper-globalization" period by the relative dynamics of additive and multiplicative costs.
 
 With respect to the trade flows coverage, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
+Yet, it appears that for many pairs (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
 This can be seen comparing the number of observations by year/ transport mode between this alternative method and our baseline method reported in Table \ref{tab_comp_referee1_baseline}.
-Put it differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+Put differently, this methodology discards countries which export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
 In this respect, the induced selection bias reduces the general scope of the transport costs estimates.
@@
 
-As noted above, the partners-sectors covered are lower.
-Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
+As noted above, the number of partners-sectors covered are lower.
+Regarding the share of additive costs $\beta$, the comparison between the two methods delivers opposite conclusions depending on the transport mode.
 While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
-In any case yet, the share of additive costs remains substantial, around 40\% on average over the period, thereby confirming the main message of our paper shedding light on the importance of the additive component in international transport costs.
+In any case yet, the share of additive costs remains substantial, around 40\% on average across transport modes over the period, thereby confirming the main message of our paper regarding the importance of the additive component in international transport costs.
 
@@
 Some elements can yet be put forward to account for the difference of results.
-Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods that represent a bit more 56\% of US total imports of goods in his dataset.
+Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020} restricts to indivisible/discrete goods that represent a bit more than 56\% of US total imports of goods in his dataset.
 In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
@@
 \paragraph{Data availability issue} The first difficulty is tied to the availability of information on quantities.
-Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) do report the quantity by transport mode.
-This is incompatible with our empirical strategy for estimating transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
-As a result, to compute the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) report the quantity by transport mode.
+This is incompatible with our empirical strategy for estimating transport costs conditionally on the transport mode, similarly as in \cite{hummels2007}.
+As a result, computing the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
 This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
@@
 Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail a large loss of useful information.
-Original census data, when accessible (from 2002 because of the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+Original census data, when accessible (from 2002 due to the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
 This is still a large share, but not too large to preclude any work.
-Accordingly, we dig deeper into this issue with the US Census dataset starting in 2002, restricted to single transport mode flows.
+Accordingly, we dig deeper into this issue with the US Census dataset starting in 2002, restricting to single transport mode flows.
 \smallskip
@@
 The second difficulty dwells on the units used to measure quantities.
-The Census data, if it provides some information on quantity, remains silent on the units considered.
+The Census data does provide some information on quantity but remains silent on the units considered.
 To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
@@
 weight can be given in ``kg’’, ``kg.'' ``Kg'' and, also, ``g'' and ``t'').}).\footnote{For instance, 3-digit SITC sector ``061 Sugar, molasses and honey'' includes among other products, 5-digit SITC product ``6150 Molasses, whether or not decolourized'', measured in liters and 5-digit SITC product ``06160 Natural Honey'' measured in kg.
-Actually worse than this, the 5-digit SITC product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel'' includes both 10-digit HTS product ``1702903500 Other: invert molasses'', measured in liters and 10-digit HTS product ``1702602200 Blended syrups described in additional u.s.
+Even worse, the 5-digit SITC product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel'' includes both 10-digit HTS product ``1702903500 Other: invert molasses'', measured in liters and 10-digit HTS product ``1702602200 Blended syrups described in additional u.s.
 note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions''  measured in kilograms.} Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece'' to be the same as the transport cost of a ``kg'', even in the same sector.
@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could indeed be easily replaced by (sector,quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Sector-level fixed effects could indeed be easily replaced by (sector times quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
 Let us elaborate on this issue.
@@
 First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
-In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
+In contrast to the aggregation and the separability issues, these restrictions strongly limit both the number of sectors and of countries covered, and do also impact the value of trade flows covered.
 This raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
@@
 \paragraph{On the relevance of the price per-weight approach}
-At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.
+At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price adds some relevant information on the international transport cost burden faced by exporters.
 We detail below the underlying rationale.
 
-Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mix per-weight and per-volume basis, both historically and nowadays.
+Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mixed per-weight and per-volume basis, both historically and nowadays.
 For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight'' or ``dimensional weight'')\footnote{See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).} while maritime freight costs are expressed in volume.
-Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
+Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measures of mean density for different sectors.
 The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions - in contrast, per-unit shipping cost depends on the behavior of exporting firms.
@@
 However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
-In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf}) reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
+In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
 In air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
@@
 The situation is strikingly different for primary goods only.
-In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)).
+In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)).
 Furthermore, in air transport, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).

### 2021-11-22 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Travail sur réponse au referee 1

@@
 \item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to endogeneity have been added).
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, which presents the results and now includes the robustness checks as a final sub-section.
-For sake of brevity, the robustness analysis related to composition effects has been sent to the Appendix (Section C.3).
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
+Section 3 now includes the robustness checks as a final sub-section.
 \item Both referees asked us to highlight the implications of additive costs (the ``Big Picture'').
@@
     \item Most importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds.
-Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated in Section 3, to quantify the extra welfare gains attributable to the reduction of their additive component.
-This analysis is conducted in a new Section 5.
+Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs estimated in Section 3.
+This allows us to quantify the extra welfare gains attributable to reduction of additive trade costs.
+This analysis forms Section 5.
         \end{enumerate}
-\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
+\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
 \end{itemize}
@@
 In 2019, for example, there are 14,016 country$\times$sector pairs.
-Adding 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$).
-This is computationally not tractable.
+Without the separability assumption, one must estimate 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$).
+This is computationally impossible.
 % use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/data/hummels_tra.dta", clear
@@
  This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1).
-It yet remains quite large in terms of trade coverage (mean of 68\%).\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the large sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the complete sample).
+It yet remains quite large in terms of trade coverage (mean of 68\%).\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the baseline sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the baseline sample).
 The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} On this reduced sample, we re-run the estimation with and without the separability assumption.
@@
 This drove us to question our own empirical specification deeper.
-In the end, we decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that the referee suggests does not go without a few important limitations.
+In the end, we decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight its costs, and the alternative that the referee suggests does not go without a few important limitations.
 We develop each line of argument in our answer.
-Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3 \textbf{XXX CHECK XXX }).
-After reading our answer, we hope to convince you that we made the right choice.
+Yet, because the referee's estimation method is also an interesting alternative, we have added a new section in the Online appendix devoted to it (Section D.3 \textbf{XXX CHECK XXX }).
 
@@
 By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
-
-Despite its interest, implementing the referee's method uncovers some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
-First, we challenge the advantages that the referee's estimation method would bring in comparison with our's.
+Despite its interest, the referee's method has some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
+First, we challenge the advantages that the referee's estimation method would bring in comparison with ours.
 Second, we emphasize the drawbacks induced by this method.
@@
 \textbf{Our answer:}
-As already discussed in our letter, the separability assumption does not seem to be a strong assumption.
+As already discussed in this answer, the separability assumption does not seem to be a strong assumption.
 This is based on the conclusion drawn from the robustness check on a reduced sample.
-If it is smaller in terms of observations, it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 68\% of the total value of flows (on a yearly basis).
 As we conclude at the end of Section 3.3.1 \textbf{XXX CHECK XXX}, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
@@
      We agree with the referee that the comparison with the literature (\citealp{hummels2007} in particular) was not straightforward for the reader in the submitted version.
-We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method.
+We did our best to make the comparison clearer in the revised version, in a way consistent with our estimation method.
 In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
@@
      with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ikt}$ the price-elasticity of transport costs, with $i$ the origin country, $k$ the product and $t$ for time.
-As we now show in the paper (Section 2.2), this also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper.
+As we now show in the paper (Section 2.2), $\beta_{ikt}$ also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper.
 We directly report them here to ease the comparison with the referee's method afterwards.}
@@
 
-In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode).
+In this respect, the referee's method and ours are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time, country, sector and transport mode.
 Put it differently, our method is consistent with Hummels' methodology - provided that it is adequately reported as such.
@@
 We agree that this was not sufficiently transparent in the submitted version.
-We hope that the revised version now fits the referee's expectations on this point.
+We hope that the revised version now fulfills the referee's expectations on this point.
 
@@
 
-As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 2001.
+As previously noted, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 1989 (though we could not obtain the whole data because of  COVID).
 We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix, Section D.3, Table D.1 and Figure D.3.\textbf{XXX Check XXX}\footnote{Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
@@
 Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
-While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
-In any case yet, the share of additive costs remains substantial, around 40\% on average over the period.
+While the estimated $\beta$ is lower for vessel transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
+In any case, the share of additive costs remains substantial, around 40\% on average over the period.
 In this respect, it confirms the main message of our paper shedding light on the importance of the additive component in international transport costs.
@@
 This may induce a selection bias in the sample covered for estimation.
-As reported in Table  \ref{tab_comp_referee1_baseline}, if the number of pairs covered is much lower (especially in the list of countries covered), this concern can yet be mitigated by noticing that the covered value of trade flows is only slightly lower.
+As reported in Table  \ref{tab_comp_referee1_baseline}, the number of pairs covered is much lower (especially in the list of countries covered).
+This concern is mitigated by noticing that the covered value of trade flows is only slightly lower.
 If it reduces the sample of sectors and origin countries covered, this method preserves the majority of trade flows in value.
@@
 Information about the port of entry are only available since 1989, which is already smaller than our full sample starting in 1974.
-On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019.
+On top of that, because of the Covid situation, the US Bureau of Census was only able to send us the data relative to the years 1997-1999 and 2001-2019.
 Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
 In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
-Eliminating this dimension of the paper would be detrimental to its value-added.
+Eliminating this dimension of the paper would be detrimental to its contribution.
 
@@
     \item With the referee's method, we estimate one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
-From this, we can approximate the 5-95\% threshold values through:
+From this, we can compute the size of the 5-95\% confidence interval $CI^{ref}_{is} = \hat{\beta}_{is}^{max,ref} - \hat{\beta}_{is}^{min,ref}$ where:
         $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
-
-        From this, we can deduce the confidence interval $CI^{ref}_{is} = \hat{\beta}_{is}^{max,ref} - \hat{\beta}_{is}^{min,ref}$.
 
@@
 As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
-We then deduce the accuracy of the $\beta$ estimate by relying on bootstrap method.
-Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 threshold values for each couple $i,s$.
-Noticing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
+We compute the accuracy of the $\beta$ estimate by relying on bootstrap method.
+Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.
+Noticing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
 }
@@
 
-   \textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a lower interval confidence of the $\beta$ estimate with our methodology.
-Our method undoubtedly yields more accurate estimations of the share of additive components, whatever the transport mode considered.
+   \textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a smaller interval confidence of the $\beta$ estimate with our methodology.
+Our method thus yields more accurate estimations of the share of additive components, whatever the transport mode considered.
 We have checked that a similar conclusion applies on other years (they are not reported here for sake of brevity but they are available upon request).
@@
 
-One explanation of the limited accuracy with the alternative method might be the following.
+One explanation of the limited accuracy of the alternative method might be the following.
 Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and \textit{across} origin countries.
 Put it in plain words, it means that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
-One might view this as a disputable assumption.
-From a statistical point of view further, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component, that explains the lower accuracy of the estimation.
+This is a disputable assumption.
+From a statistical point of view, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component.
+We believe that explains the better accuracy of our baseline estimation.
 
@@
 
-All these elements, in particular Concern 2 regarding the lower accuracy of the estimates, drive us to maintain our original estimation as benchmark method in the revised version.
+All these elements, in particular Concern 2 regarding the lower accuracy of the estimates, drive us to maintain our original estimation as baseline method in the revised version.
 However, we also want to keep track of this alternative method as further robustness check.
 For this purpose, we have amended the online Appendix with a section devoted to the referee's method (Section D.3 \textbf{XXX Check XXX}).
-We deeply hope to convince you of the relevance of this choice.
 
@@
 The referee states that, based on theoretical insights by \citet{melitz} or \citet{baldwin_harrigan}, $1/\widetilde{p}_{ik}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
-In other words, more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.
+In other words, more productive firms and/or firms selling high-quality products will charge systematically different prices, all other things equal – in our case, for a given country-product pair.
 
 We obviously do not question this conceptual issue.
-However, it is worth noting that a good deal of the bias (actually, the part relating to the quality effect) is going to appear identically in the CIF (p) and the FAS ($\widetilde{p}_{ik}$) prices.
+However, it is worth noting that a good deal of the bias (such as the part relating to the quality effect) is going to appear identically in the CIF (p) and the FAS ($\widetilde{p}_{ik}$) prices.
 Consequently, since our dependent variable is based on a ratio between the former and the latter, the (reverse causality) bias cancels out.
-That said, remains the possibility that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
+That said, the possibility remains that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
 
-Following the referee's advice, we decided therefore to provide a full set of IV estimates to provide a clean assessment of the size of the potential bias.
+Following the referee's advice, we therefore decided to provide a full set of IV estimates to provide a clean assessment of the size of the potential bias.
 Section 4.3 in the revised version of the paper provides an overview of the results, while section B.4 in Appendix B provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
@@
 Figure 9 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart.
-In all cases, these estimates are very similar, when not identical.
-This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
+In all cases, these estimates are very similar.
+This suggests that our baseline estimates do not suffer any substantial biases arising from endogeneity concerns.
 
 It should be noted that we performed this check on our main dataset (SITC 5 digit), and also at the HS 10 level, to handle simultaneously the referee's concern about aggregation issues.
-In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very, very similar to their non-IV, least squares counterparts.
-Needless to say that these results are available upon request from the referee.
+In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very similar to their non-IV, least squares counterparts.
+Needless to say that these results are available upon request.
 
@@
 Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummels' dataset available over 1974-2004 with $k=5$ the finest degree of aggregation.
-We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data.
+We agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variations in the data.
 After checking with the US Bureau of Census, HS-10 data is only available since 1989, now available to 2019.
-As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.
-In front of this arbitrage, we decided to maintain the case $s=3, k=5$ as baseline, but running estimation at the $s=3,k=10$ in the new Sub-Section 3.3.3 \textbf{XXX CHECK XXX }devoted to robustness analysis.
+As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.We decided to maintain the case $s=3, k=5$ as baseline, and to run the estimation using 10-digit products in the new Sub-Section 3.3.3 \textbf{XXX CHECK XXX }devoted to robustness analysis.
 
@@
 In this context, the findings in \cite{Lashkaripour_JIE2020} are specifically enlightening.
-Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
-\cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
-This result of transport costs close to be totally iceberg contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
-Some elements can yet be put forward to account for the difference of results.
+Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and that the cost of transportation increases more rapidly with unit weight than the cost of production.
+\cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the ad-valorem cost assumption regarding transport costs.
+This result of transport costs close to be totally ad-valorem contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
+There are explanations for this difference.
 Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods that represent a bit more 56\% of US total imports of goods in his dataset.
@@
 Nevertheless, considering the importance of the referee's remark, we went beyond this very first argument to investigate this question in depth.
-This was not without difficulty yet, in particular due to data and estimation constraints.
-At a more fundamental level, we also believe that looking at transport costs on a per-weight basis does add relevant information.
-We develop each line of argument in what follows.
+We first discuss data and estimation difficulties before giving the results.
+At a more fundamental level, we argue that looking at transport costs on a per-weight basis does add relevant information.
 
@@
 
-In our attempt to address the referee's remark, we faced a first difficulty regarding the availability of information on quantities.
-Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) do report the quantity by transport mode.
-This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}.
-As a result, computing the unit price of all trade would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
+The first difficulty is availability of information on quantities.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) report the quantity by transport mode.
+This is incompatible with our empirical strategy of estimation which is conditional on the transport mode, similarly as in \cite{hummels2007}.
+As a result, computing the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
 This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
-The alternative would then be to only consider single-mode flows.
+The alternative is only to consider single-mode flows.
 In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
 This would reduce the total value of the sample from 12,500 billion dollars to 2,650 billion dollars, entailing a large loss of useful information.
-Original census data, when accessible (from 2002 because of the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+Original census data, when accessible, are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
 This is still a large share, but not too large to preclude any work.
-Accordingly, we dig deeper into this issue with the US Census dataset starting in 2002, restricted to single transport mode flows.
+Accordingly, we dig deeper into this issue with the US Census dataset, restricted to single transport mode flows.
 \smallskip
 
-The second difficulty dwells on the units used to measure quantities.
-The Census data, if it provides some information on quantity, remains silent on the units considered.
+Quantity units are a second difficult.
+The Census data remain silent on that matter.
 To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
-Unfortunately, these schedules change many time per year and only the schedules from 2009 onwards were available in a .csv format.
+Unfortunately, these schedules change many time per year and only the schedules from 2009 onwards are available in a .csv format on the Census website.
 On the years for which the match could be done (from 2009), it turns out that at our sector level (3-digit), goods (either at the 5-digit or the 10-digit level) are measured in different units.
@@
 Actually worse than this, the 5-digit SITC product ``06190 Other sugars; sugar syrups; artificial honey (whether or not mixed with natural honey); caramel'' includes both 10-digit HTS product ``1702903500 Other: invert molasses'', measured in liters and 10-digit HTS product ``1702602200 Blended syrups described in additional u.s.
-note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions''  measured in kilograms.} Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece'' to be the same as the transport cost of a ``kg'', even in the same sector.
+note 4 to chapter 17: described in general note 15 of the tariff schedule and entered pursuant to its provisions''  measured in kg.} Neglecting these differences would invalidate our functional forms, as no one expects the transport cost of a ``piece'' to be the same as the transport cost of a ``kg'', even in the same sector.
 So we are limited to using quantity units from 2009 to 2019, amending our structure of fixed effects to consider the triplet (origin country, sector, quantity unit).
@@
 Getting deeper on the estimated functional form raised another difficulty in adapting our equation to per-unit rather than per-weight prices.
-As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
+As discussed above, in our baseline regression we separate each transport cost component in two distinct sector-origin country dimensions.
 As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could indeed be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Sector-level fixed effects could be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
 Let us elaborate on this issue.
 
-In the per-weight price, the existence of a single country fixed effect applicable to all sectors could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport from, say, Germany.
+In the per-weight price, the existence of a single country fixed effect applicable to all sectors is justified, as a kilogram is always a kilogram and one can assume they both add the same to transport costs from, say, Germany.
 In contrast, an unit of tee-shirt is quite different from an unit of car.
-To clarify this point, assume imports of tee-shirts and cars from Germany and China.
-Under the separability assumption, importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries); and similarly for importing a car.
+To clarify this point, consider imports of tee-shirts and cars from Germany and China.
+Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by both countries fixed effect (the same for all sectors) and the cost given by the sector fixed effect ; and similarly for importing a car.
 Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany relative to China.
-This does not seem a sensible assumption; in this case specifically, we cannot keep the separability assumption.
+This does not seem a sensible assumption.
+In this setting, we cannot keep the separability assumption.
 As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper).\smallskip
@@
 \noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$, $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
-Depending on the sector, two different units are considered, kilograms and numbers.
 
-The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} (as well as in the online appendix D.4.
-In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity; while the second column reports the results obtained considering the price per kg on the same sample; the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} (as well as in the online appendix D.4).
+In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity, the second column reports the results obtained considering the price per kg on the same sample and the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
 
@@
 In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
-In our view, this raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
+In our view, this raises a non-trivial sample selection bias that reduces the general scope of the transport cost estimates.
 Second, in terms of $\beta$ estimates, the results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
-In this regard, they stand in line with the results of \cite{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
+In this regard, they stand in line with the results of \cite{Lashkaripour_JIE2020}, and confirm the correlation between weight per unit and unit price.
 Still, the share of additive cost is not reduced to zero (except if one looks at the median for vessel).
@@
 At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.
-Let us explain the underlying rationale.
+Let us explain why.
 
-Our main objective in this paper is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mixed, per-weight and per-volume basis, both historically and nowadays.
+Our main objective is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mixed, per-weight and per-volume basis, both historically and nowadays.
 For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight'' or ``dimensional weight'')\footnote{See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).} while maritime freight costs are expressed in volume.
@@
 Put differently, the shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
-Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
+Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which firms operate.
 
@@
 After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
-According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
+According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value above 50\% for only two years.
 Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
@@
 Taking stock of your suggestion, we decided to offer insights on the welfare implications of our results.
-In this regard, the new section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
+The new section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
 To do so, we start from the canonical \cites{melitz} model where we add additive trade costs.
@@
 To that end, we adjust the share of exporting firms in the US: Based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
-In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/iceberg component, with the reduction in the fixed export cost.
+In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component, with the reduction in the fixed export cost.
 We compare this with the case where the total transport costs reduction is solely attributed to a decrease in ad-valorem costs.

### 2021-11-22 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Travail guillaume réponse referee 1
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
-However, \citet{sorensen2014} exclusively performs a theoretical analysis, without any quantitative exercise.
-In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the presence of additive costs, but the latter is based on a calibration for transport costs restricted to 2004, based on the case of Nowrway.
-In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on ``true'' values for the latter - remember that our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the (median) export price.
+However, \citet{sorensen2014} dose not perform an empirically based quantitative exercise.
+In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the presence of additive costs, but the latter is based on a calibration for transport costs restricted to 2004, based on the case of Norway.
+In contrast, our own exercise relies on a several decades of US data.
+It allows us to compute the welfare effect of  taking intor account the relative dynamics of additive and multiplicative costs.
+Furthermore, our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the sum of prices and multiplicative costs.
 
-More precisely, we use some of the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, more precisely those for the years 1974 and 2019.
-Based on the latter, we implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+More precisely, we use the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, to implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
 In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
-To that end, we adjust the share of exporting firms in the US: Based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
-In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component, with the reduction in the fixed export cost.
-We compare this with the case where the total transport costs reduction is solely attributed to a decrease in ad-valorem costs.
+To that end, we adjust the share of exporting firms in the US.
+Based on \citet{BEJK-AER-03, Lincoln_McCallum2018}.
+In our preferred exercise, we quantify the welfare gains that derive from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component in a context of fixed export cost reduction.
+We compare this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
 
 Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
-Specifically, we report the change in total transport costs decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
-Qualitatively, the conclusions are identical for both transport modes, with two major insights.
-First, for a given decrease in total transport costs,  welfare gains are around 50\% higher when this reduction is partly achieved through a reduction in the additive costs.
+Specifically, we report the welfare effect of change in total transport costs decomposed in its two dimensions (additive and ad-valorem),both in absolute and in relative terms.
+We have two major insights.
+First, using the reduction of transport costs in vessel (air),  welfare gains are around 50\% (14\%) higher when this reduction is achieved through the estimated reduction in additive and ad-valorem costs.
 Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing variable costs, with again an additional premium coming from the decrease of additive costs.
 
-Overall, these results appear to add a substantial contribution to the paper : We are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
-Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
+Overall, these results add a substantial contribution to the paper : We are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+%Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
 

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Suite au mail de Jérôme

@@
 In contrast, an unit of tee-shirt is quite different from an unit of car.
-To clarify this point, consider imports of tee-shirts and cars from Germany and China.
-Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by both countries fixed effect (the same for all sectors) and the cost given by the sector fixed effect ; and similarly for importing a car.
-Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany relative to China.
+To clarify this point, consider imports of tee-shirts and cars from
+Germany and China.
+Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by the  country fixed effect (the same for all sectors) plus the cost given by
+the sector fixed effect ; and similarly for importing a car.
+Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee- shirt from Germany relative to China.
 This does not seem a sensible assumption.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
petite coquille OA
Modification des tableaux de l’appendice
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Trois des tableaux supplémentaires apparaissent
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Les tableaux sont là
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Suite au mail de Jérôme

@@
 
-This section is devoted to presenting the results.
-We first compare the estimates of the transport costs components under the three models.
-In a second step, we report quality of fit tests for the three models.
+In this section, we first compare the estimates of the transport costs components under the three models. In a second step, we report quality of fit tests for the three models.
 
@@
 
-In this section, we report the estimation results of the three models: Model (A) (with only ad-valorem costs), Model (B) (with both additive and ad-valorem) and Model (C) (with only additive costs).
+We report estimations of the three models for selected years.
+Section \ref{secoa:yearly_results} provides estimations for every year.
 Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_ves} for vessel transport.
@@
 
-Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively.
-Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs, already established at the 3-digit level.
+Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and vessel transport respectively.
+Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs we found at the 3-digit level.
 
@@
 	\centering
+%	\footnotesize{
+		\caption{Air: Transport costs estimates, selected years, 4-digit level}
+		\label{tab:result_air_rob}%
+		\input{TableB1_air.tex}
+		\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\end{tablenotes}
+\end{table}%
+
+\begin{table}[htbp]
+  \centering
 \caption{Vessel: Transport costs estimates, selected years, 4-digit level}
@@
 \end{table}%
+
+\begin{table}[htbp]
+	\centering
+%	\footnotesize{
+		\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
+		\label{tab:result_air_rob}%
+		\input{TableB1_ves.tex}
+		\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\end{tablenotes}
+\end{table}
 
@@
 	\centering
+%	\footnotesize{
+		\caption{Air: Measures of goodness of fit, 4-digit level}
+		\label{tab:good_fit_air_rob}%
+		\input{TableB3_air.tex}
+		\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\end{tablenotes}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
   \caption{Vessel: Measures of goodness of fit, 4-digit level}
@@
 Model (A) & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
-Model (B) \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
+Model (B) & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
 \multicolumn{7}{l}{\textbf{SER} }\\ \hline
@@
 
+\begin{table}[htbp]
+	\centering
+%	\footnotesize{
+		\caption{Vessel: Measures of goodness of fit, 4-digit level}
+		\label{tab:good_fit_air_rob}%
+		\input{TableB3_ves.tex}
+		\begin{tablenotes}
+			\tiny
+			\item Statistics are weighted by value
+			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (B): With additive and ad-valorem transport costs}
+		\end{tablenotes}
+	\end{table}
+
 If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.
@@
 When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
-In other words, if using a more disaggregated product classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
+Using a more disaggregated product classification unsurprisingly adds some statistical precision, but the gain is not large enough to disqualify the use of slightly more aggregated data.
 
@@
 
-\subsection{Transport Cost Estimates: Yearly Detailed Results}
+\subsection{Transport Cost Estimates: Yearly Detailed Results \label{secoa:yearly_results}}
 
@@
 
-		\input{TableB1974_1987_ves.tex}
+		\input{TableC1974_1987_ves.tex}
 				%\begin{tablenotes}
@@
 
-		\input{TableB1988_2001_ves.tex}
+		\input{TableC1988_2001_ves.tex}
 				%\begin{tablenotes}
@@
 
-		\input{TableB2002_2015_ves.tex}
+		\input{TableC2002_2015_ves.tex}
 				%\begin{tablenotes}
@@
 
-		\input{TableB2016_2019_ves.tex}
+		\input{TableC2016_2019_ves.tex}
 			%	\begin{tablenotes}
@@
 As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000).
-This can be attributed to the presence of outliers in the distribution of the additive costs estimates.
+This can be attributed to the presence of outliers in the distribution of the additive costs estimates, despite the exclusion of outliers in the data.
 The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000.
 Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
-Still, this does not make much difference.
-When 1989 is included, the weighted mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
+Excluding or excluding a single year out of 45 does not make much difference in the weighted mean transport cost.
+%When 1989 is included, the weighted mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
 The weighted median value is left unchanged at 2.9\% whether 1989 is included or not.
@@
 
-		\input{TableB1974_1987_air.tex}
+		\input{TableC1974_1987_air.tex}
 				%\begin{tablenotes}
@@
 
-		\input{TableB1988_2001_air.tex}
+		\input{TableC1988_2001_air.tex}
 				%\begin{tablenotes}
@@
 
-	\input{TableB2002_2015_air.tex}
+	\input{TableC2002_2015_air.tex}
 			%\begin{tablenotes}
@@
 
-	\input{TableB2016_2019_air.tex}
+	\input{TableC2016_2019_air.tex}
 			%\begin{tablenotes}
@@
 Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
-This holds throughout the period and for both transport modes. By extension, one can expect a limited role of distance and other country-related variables as determinants of transport costs.
+This holds throughout the period and for both transport modes.
 
@@
 In contrast, an unit of tee-shirt is quite different from an unit of car.
-To clarify this point, assume imports of tee-shirts and cars from Germany and China.
-Under the separability assumption, importing a tee-shirt from either country implies paying the fixed effect for both countries (the same for all sectors) and the fixed effect for the sector (the same for all countries); and similarly for importing a car.
-Hence, the difference in the cost of importing a car from Germany and the cost of importing a car from China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany and the cost of importing a tee-shirt from China.
-This does not seem a sensible assumption; in this case specifically, we cannot keep the separability assumption.
+In the per-weight price, the existence of a single country fixed effect applicable to all sectors is justified, as a kilogram is always a kilogram and one can assume they both add the same to transport costs from, say, Germany.
+In contrast, an unit of tee-shirt is quite different from an unit of car.
+To clarify this point, consider imports of tee-shirts and cars from Germany and China.
+Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by the  country fixed effect (the same for all sectors) plus the cost given by the sector fixed effect ; and similarly for importing a car.
+Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee- shirt from Germany relative to China.
+This does not seem a sensible assumption.
+In this setting, we cannot keep the separability assumption.
 As discussed before, this dramatically increases the computational difficulty of our method (with no substantial different results, see Subsection 3.3.1. in the main paper.\smallskip

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Petite chose dans l’estimation directe

@@
 If we take the value of the $\beta$ by itself, there is no clear criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
-Further, as shown in Table \ref{tab_comp_referee1_baseline}, no clear conclusion on a potential upwards or downwards bias that would be attached to our method.
+Further, as shown in Table \ref{tab_comp_referee1_baseline}, no clear sign of a potential upwards or downwards bias that would be attached to our method.
 Things are more clear-cut in terms of accuracy of the estimation.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Encore sur la méthode directe

@@
 As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
-We compute the accuracy of the $\beta$ estimate by relying on bootstrap method.
+We compute the accuracy of the $\beta$ estimation through a bootstrap method.
 Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.
-Noticing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
+Writing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
 }
@@
 
-   \textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a smaller interval confidence of the $\beta$ estimate with our methodology.
+   \textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a smaller interval confidence of the baseline $\beta$ estimation.
 Our method thus yields more accurate estimations of the share of additive components, whatever the transport mode considered.
@@
 
-One explanation of the limited accuracy of the alternative method might be the following.
+One explanation of the limited accuracy of the direct method might be the following.
 Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and \textit{across} origin countries.
-Put it in plain words, it means that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
+In plain words, it assumes that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
 This is a disputable assumption.
 From a statistical point of view, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component.
-We believe that explains the better accuracy of our baseline estimation.
+We believe that explains the better accuracy of the baseline estimation.
 

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Réponse referee 1

@@
 As a result, computing the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
-This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
-The alternative is only to consider single-mode flows.
+This is unlikely, as one would expect the higher value-to-weight varieties to be transported by air rather than by vessel.
+The alternative is to consider only single-mode flows.
 In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
@@
 Getting deeper on the estimated functional form raised another difficulty in adapting our equation to per-unit rather than per-weight prices.
-As discussed above, in our baseline regression we separate each transport cost component in two distinct sector-origin country dimensions.
+As discussed above, in our baseline regression we separate each transport cost component in two distinct sector-country dimensions.
 As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
encore des petites choses sur la réponse (per-kg)

@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Sector-level fixed effects could be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear possible for country fixed effects.
 Let us elaborate on this issue.
@@
 \end{equation*}
-\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$, $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
+\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$ (eg kg or number), $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
 
@@
 
-\subsubsection{On the relevance of the price per-weight approach}
+\subsubsection{On the relevance of the per-weight price approach}
 At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
OA : jusqu’à D

@@
 %%%% RAW TABLE TO BE FOUND AT: "...\Dropbox\Papier_Lise_Guillaume\trade_cost\results\3_models"
-% modele iceberg tout seul : terme nlI
+% modele ad-valorem tout seul : terme nlI
 % colonnes L à U: modèle avec les deux
 % termes nla, de V à Z, modèles avec que de l'additif
-%nlI quer iceberg
+%nlI quer ad-valorem
 %nla que additif
@@
 	\item Statistics are weighted by value
-	\item \textbf{Model (A): Iceberg transport costs only}
+	\item \textbf{Model (A): Ad-valorem transport costs only}
 	\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 			\item Statistics are weighted by value
-			\item \textbf{Model (A): Iceberg transport costs only}
+			\item \textbf{Model (A): Ad-valorem transport costs only}
 			\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 %			\item Statistics are weighted by value
-%			\item \textbf{Model (A): Iceberg transport costs only}
+%			\item \textbf{Model (A): Ad-valorem transport costs only}
 %			\item \textbf{Model (B): With additive and ad-valorem transport costs}
@@
 We apply the same variance decomposition exercise at the sector level, in which case the sector $s$ index (at the 3-digit level) replaces the $k$ index (at the 5-digit level).
-This gives us an alternative way to ensure the robustness of the estimation results to the degree of classification retained to estimate international transport costs.
+This gives us an alternative way to ensure the robustness of the estimation results to the precision of  the sector classification retained to estimate international transport costs.
 We also determine the share of the observed variance that can be attributed to the between-country variance, adapting the variance decomposition formula written above accordingly.
@@
 Two interesting results emerge from Figure \ref{fig:decomp_variance}.
-First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is of the same magnitude of order as the variance between sectors at the 3-digit level.
+First, the share of the cif-fas price gap variance that comes from the variance between products (5-digit level) is close to the variance between sectors at the 3-digit level.
 Both account for between 30 and 40\% of the total variance in air transport, depending on the years considered.
-This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is more pronounced (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
-This provides an indirect robustness check to the degree of classification we have selected to estimate international transport costs.
+This is also the case for vessel transport, even if the difference between the between-product variance share and the between-sector share is larger (30\% for the between-sector vs 40\% for the between-product variance at the beginning of the period).
+This suggests our procedure should be robust to the choice of the precision of the sector classification.
 Second, the variance of the cif-fas price gap that can be attributed to the product (or sector) dimension is much larger than the between-country variance.
@@
 
-\paragraph{Comparison on a yearly basis} Figure \ref{fig:robustesse_non_separe} displays the results on a yearly basis for air transport in panel (A) and for maritime transport in panel (B).
+\paragraph{Comparison on a yearly basis} Figure \ref{fig:robustesse_non_separe} displays the results on a yearly basis for air transport in panel (A) and for vessel transport in panel (B).
 In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
@@
 
-In line with Table 2 of the paper, the results shown in Figure \ref{fig:robustesse_non_separe} confirm that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
-However, they also show that the difference is quantitatively small in most cases.
-Further, and most importantly, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs, whether estimated under the separability assumption or not, are very similar.
-Along with the quality-of-fit analysis, this confirms the robustness of our estimation results to this assumption.
+In line with Table 2 of the paper, Figure \ref{fig:robustesse_non_separe} confirms that the separability assumption (retained in our baseline estimation) tends to overestimate the value of the additive cost component (and underestimate the value of the ad-valorem cost).
+However, it also shows that the difference is quantitatively small in most cases.
+Further, and most importantly, whatever the transport mode and for both types of transport costs, the trends, whether estimated under the separability assumption or not, are very similar.
+This confirms the robustness of our estimation results to this assumption.
 
@@
 
-One advantage of this approach is that the above regression can be estimated separately for various country-industry pairs on a linear basis, without having to resort on the separability assumption.
-This does not solve the issue yet, for two main reasons.
-First, it still requires a non-linear estimator.
-This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
+One advantage of this approach is that the above equation is linear and can be estimated separately for various country-industry pairs, without having to resort on the separability assumption.
+Actually, it still requires a non-linear estimator due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
 imposing $0 \leq  \beta \leq 1$.
@@
 
-It should also be noted that this estimation method implies a lower coverage, both in time and in the sectors/countries covered.
-Information about the port of entry is only available since 1989, which is already smaller than our full sample starting in 1974.
+This estimation method also implies a lower coverage, both in time and in the sectors/countries covered.
+With respect to time coverage, information about the port of entry is only available since 1989, which is already smaller than our full sample starting in 1974.
 On top of that, because of the Covid situation, the US Bureau of Census was only able to provide us the data for the years 1997-1999 and 2001-2019.
-Implementing the referee's method would hence reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
+Using it would hence reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
 In our view, the historical coverage is interesting \emph{per se}: it provides useful insights about how transport costs have evolved over several decades and more importantly, it allows us to highlight the welfare alterations induced over the ``hyper-globalization" period by the relative dynamics of additive and multiplicative costs.
@@
 With respect to the trade flows coverage, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many pairs (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
+Yet, it appears that for many pairs (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, to estimate the equation..
 This can be seen comparing the number of observations by year/ transport mode between this alternative method and our baseline method reported in Table \ref{tab_comp_referee1_baseline}.
@@
 Regarding the share of additive costs $\beta$, the comparison between the two methods delivers opposite conclusions depending on the transport mode.
-While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
-In any case yet, the share of additive costs remains substantial, around 40\% on average across transport modes over the period, thereby confirming the main message of our paper regarding the importance of the additive component in international transport costs.
+While the estimated $\beta$ is lower for vessel transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
+In any case, the share of additive costs remains substantial, around 40\% on average across transport modes over the period, thereby confirming the main message of our paper regarding the importance of the additive component in international transport costs.
 
@@
 
+If we take the value of the $\beta$ by itself, there is no clear criteria to discriminate between the value estimated with the baseline method and estimating directly $\beta$ (when, of course, run on the same sample).
+Further, as shown in Table \ref{tab_comp_referee1_baseline}, there is no clear sign of a potential upwards or downwards bias that would be attached to either method.
+Things are more clear-cut in terms of accuracy of the estimation.
+Specifically, the baseline method yields a more precise estimation of the $\beta$ than the direct method.
+To develop on this, for each year and transport mode:
+\begin{itemize}
+	\item The direct method estimates one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{dir}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
+From this, we can compute the size of the 5-95\% confidence interval $CI^{dir}_{is} = \hat{\beta}_{is}^{max,dir} - \hat{\beta}_{is}^{min,dir}$ where:
+	$$\hat{\beta}_{is}^{min,dir} = \hat{\beta}^{dir}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,dir} = \hat{\beta}^{dir}_{is(k)} + 1.96 SD_{is}$$
+
+	\item  The baseline method estimates the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$ (on a year-transport mode basis).
+As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
+We compute the accuracy of the $\beta$ estimation through a bootstrap method.
+Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.
+Writing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the direct method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
+}
+
+	\item We can then evaluate the accuracy of each  method by comparing the size of the confidence intervals of the $\beta$, for each couple $i,s$ (by year and transport mode).
+We summarize this comparison in Figure \ref{fig:accuracy_beta}, which reports the distribution of the ratio of confidence intervals, taken in log.
+
+	\begin{figure}[htbp]
+		\caption{Accuracy of the estimation of the $\beta$: Comparison}
+		\label{fig:accuracy_beta}
+		\begin{center}
+			\includegraphics[height=3in]{../../revised_article/comparaison_amplitude_baseline_referee1_2013_ves.png}
+			\begin{minipage} [c]  {5in} \scriptsize%
+				Note: Ratio of the confidence intervals = $ _{}ln(CI_{is}/CI^{ref}_{is})$
+			\end{minipage}
+
+		\end{center}
+	\end{figure}
+
+	\textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a smaller interval confidence of the baseline $\beta$ estimation.
+Our method thus yields more accurate estimations of the share of additive components, whatever the transport mode considered.
+We have checked that a similar conclusion applies on other years (they are not reported here for sake of brevity but they are available upon request).
+\end{itemize}
+
+One explanation of the limited accuracy of the alternative method might be the following.
+Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and \textit{across} origin countries.
+Put in plain words, it assumes that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
+This is a disputable assumption.
+From a statistical point of view, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component.
+We believe that explains the better accuracy of our baseline estimation.
+
 \subsubsection{Weights and quantities}
 
-One may be concerned by the fact that unit prices are calculated as the value divided by the shipping
-weight, rather than to the quantity (the number) of goods.
-\cite{Lashkaripour_JIE2020} notably points out the importance of this measurement issue.
-Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than the cost of production.
-\cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the iceberg cost assumption regarding transport costs.
-This result of transport costs close to be totally iceberg contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
-Some elements can yet be put forward to account for the difference of results.
+In the baseline method (as in most of the literature), the explained variable is a per-kg price rather than an unit price.
+\cite{Lashkaripour_JIE2020} notably points out the importance of this issue.
+Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is substantially heterogeneous even within narrowly-defined product categories and the cost of transportation increases more rapidly with unit weight than with the cost of production.
+\cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the ad-valorem cost assumption regarding transport costs.
+This result of transport costs close to be totally ad-valorem contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
+There are explanations for this difference.
 Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020} restricts to indivisible/discrete goods that represent a bit more than 56\% of US total imports of goods in his dataset.
@@
 Nevertheless, this section investigates the robustness of our results when unit values are built as the ratio between value and quantity of goods per observation.
-This was not without difficulty yet, in particular due to data and estimation constraints.
-At a more fundamental level, we also believe that looking at transport costs on a per-weight basis does add relevant information.
-
-\paragraph{Data availability issue} The first difficulty is tied to the availability of information on quantities.
+We first discuss data and estimation difficulties before giving the results.
+At a more fundamental level, we argue that looking at transport costs on a per-weight basis does add relevant information.
+
+\paragraph{Data availability issue} The first difficulty is availability of information on quantities.
 Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) report the quantity by transport mode.
-This is incompatible with our empirical strategy for estimating transport costs conditionally on the transport mode, similarly as in \cite{hummels2007}.
+This is incompatible with our empirical strategy for estimation which is conditional on the transport mode, similarly as in \cite{hummels2007}.
 As a result, computing the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
-This is unlikely, as one would expect the higher-to-weight value to be transported by air rather than by vessel.
-The alternative would then be to only consider single-mode flows.
+This is unlikely, as one would expect the higher value-to-weight varieties to be transported by air rather than by vessel.
+The alternative is to consider only single-mode flows.
 In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
 Reducing the total value of the sample from 12,500 billion dollars to 2,650 billion dollars would entail a large loss of useful information.
-Original census data, when accessible (from 2002 due to the Covid situation), are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+Original census data, when accessible, are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
 This is still a large share, but not too large to preclude any work.
-Accordingly, we dig deeper into this issue with the US Census dataset starting in 2002, restricting to single transport mode flows.
+Accordingly, we dig deeper into this issue with the US Census dataset, restricting to single transport mode flows.
 \smallskip
 
-The second difficulty dwells on the units used to measure quantities.
-The Census data does provide some information on quantity but remains silent on the units considered.
+Quantity units are a second difficulty.
+The Census data remain silent on that matter.
 To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
@@
 
-\paragraph{Estimation issue} Getting deeper on the estimated functional form raised another difficulty in adapting our equation to unit prices rather than weight price.
-As discussed above, in our baseline regression we separate each transport cost component in its two distinct sector-origin country dimensions.
+\paragraph{Estimation issue} Getting deeper on the estimated functional form raised another difficulty in adapting our equation to per-unit rather than per-weight price.
+As discussed above, in our baseline regression we separate each transport cost component in two distinct sector-country dimensions.
 As in \cite{Irrazabal_2015}, we assume an additive form for the additive cost, i.e.
@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could indeed be easily replaced by (sector times quantity unit) fixed effects; but the same does not appear relevant for country fixed effects.
+Sector-level fixed effects could indeed be easily replaced by (sector times quantity unit) fixed effects; but the same does not appear possible for country fixed effects.
 Let us elaborate on this issue.
 
-In the per-weight price, the existence of a single country fixed effect applicable to all sector could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport from, say, Germany.
-In contrast, an unit of tee-shirt is quite different from an unit of car.
-In the per-weight price, the existence of a single country fixed effect applicable to all sectors is justified, as a kilogram is always a kilogram and one can assume they both add the same to transport costs from, say, Germany.
+In the per-weight price, the existence of a single country fixed effect applicable to all sector could be justified, as a kilogram is always a kilogram and one can assume they both add the same to transport costs from, say, Germany.
 In contrast, an unit of tee-shirt is quite different from an unit of car.
 To clarify this point, consider imports of tee-shirts and cars from Germany and China.
-Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by the  country fixed effect (the same for all sectors) plus the cost given by the sector fixed effect ; and similarly for importing a car.
-Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee- shirt from Germany relative to China.
+Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by the  country fixed effect (the same for all sectors) plus the cost given by the sector fixed effect; and similarly for importing a car.
+Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany relative to China.
 This does not seem a sensible assumption.
@@
 $$
-\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$, $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
-Depending on the sector, two different units are considered, kilograms and numbers.
+\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$ (e.g.
+kg or number), $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
 
@@
 Second, in terms of $\beta$ estimates, the results clearly show that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
-In this respect, they stand in line with the results of \cite{Lashkaripour_JIE2020}, stressing the correlation between weight per unit and unit price.
+In this respect, they stand in line with the results of \cite{Lashkaripour_JIE2020}, and confirm the correlation between weight per unit and unit price.
 Still, the share of additive cost is not reduced to zero (except if one looks at the median for vessel).
@@
 
-\paragraph{On the relevance of the price per-weight approach}
+\paragraph{On the relevance of the per-weight price approach}
 At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price adds some relevant information on the international transport cost burden faced by exporters.
-We detail below the underlying rationale.
+We detail below why.
 

### 2021-11-23 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Fin Guillaume OA

@@
 \paragraph{Comparison on a yearly basis} Figure \ref{fig:robustesse_non_separe} displays the results on a yearly basis for air transport in panel (A) and for vessel transport in panel (B).
-In each case, the estimated values (as well as the fitted regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
+In each case, the estimated values (as well as the regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).
 
@@
 This section complements Section 4 in the main paper, where we replicate the method adopted by \cite{hummels2007} on our database.
-Here, we report the unfitted measure of transport costs versus the ``pure'' transport costs series, i.e composition effects excluded
-in percentage of the export price, as in \cite{hummels2007}, for each transport mode.
-The parallel with Hummels's (2007) terminology stands as follows: In Figure \ref{fig:comp_effects_as_in_Hummels}, the ``transport costs, excluding composition effects'' series refers to his ``fitted ad-valorem rate''; the ``transport costs'' series refers to his ``expenditure/import value''.
+Here, we report the raw measure of transport costs versus the ``ceteris paribus'' transport costs series, i.e.
+excluding composition effects, in percentage of the export price, as in \cite{hummels2007}, for each transport mode.
+The parallel with Hummels's (2007) terminology stands as follows: In Figure \ref{fig:comp_effects_as_in_Hummels}, the ``transport costs, excluding composition effects'' series refers to his ``fitted ad-valorem rate''; the ``raw transport costs'' series refers to his ``expenditure/import value''.
 Unsurprisingly, this accords with his Figure 5 (for Air) and 6 (Vessel) for the same period (until 2014).
@@
 In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows.
-The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.
-For easing comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.
+The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary sector.
+To ease the comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.
 
@@
 
-As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``pure'' transport costs and the unfitted measure, in the red dashed have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)).
+As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``ceteris paribus'' transport costs and the raw measure have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)).
 However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
 In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
-In air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)).
-Trade composition effects matter more in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)), primarily in the ad-valorem component.
-As for the whole range of flows, the 60\% decrease in the unfitted transport costs in vessel can de decomposed in a 50\% decrease in the ``ceteris paribus'' transport costs (fitted), the 10\% remaining to trade composition effects.
-
-The situation is strikingly different for primary goods only.
-In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)).
-Furthermore, in air transport, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).
+In vessel transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line): trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)).
+Trade composition effects matter more in air transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)), primarily in the ad-valorem component.
+As for the whole range of flows, the 60\% decrease in the raw transport costs in vessel can de decomposed in a 50\% decrease in the ``ceteris paribus'' transport costs, the 10\% remaining to trade composition effects.
+
+The situation is similar in the primary sector.
+In this case, in air transport te composition effect matters more (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)).
 
@@
 On the following sub-period though, their share has fallen to 20-30\%.
-Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured sector, it is hence not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the trend patterns that occur within the manufacturing sector.
+Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured sector, it is not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the manufacturing sector.
 

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Détails referee 1

@@
 [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
-\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to endogeneity have been added).
+\item In the revised version, we have strengthened the robustness checks regarding the estimation of international transport costs (robustness checks to aggregation and to endogeneity have been added).
 Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
 Section 3 now includes the robustness checks as a final sub-section.
+For sake of saving space, the robustness analysis related to composition effects has been sentt to the Appendix (Section C.3).
 \item Both referees asked us to highlight the implications of additive costs (the ``Big Picture'').
-We agree that this was not enough emphasized in the submitted version, and we thank the referees for their suggestions on that issue.
+We agree that this was not emphasized enough in the submitted version, and we thank the referees for their suggestions on that issue.
 The revised version answers this concern in two points.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
Fin referee 2, début éditeur

@@
 [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
-\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to IV estimates have been added).
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, which reports the results and now includes the robustness checks as final sub-section.
-For sake of space saving, the robustness analysis related to composition effects has been set to the Appendix (Section C.3).
-\item Both referees asked us to put light on the implications of additive costs (the ``Big Picture'').
-We agree that this was not enough emphasized in the submitted version, and we thank the referees for their suggestions on that issue.
+	\item In the revised version, we have strengthened the robustness checks regarding the estimation of international transport costs (robustness checks to aggregation and to endogeneity have been added).
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
+Section 3 now includes the robustness checks as a final sub-section.
+For sake of saving space, the robustness analysis related to composition effects has been sentt to the Appendix (Section C.3).
+	\item Both referees asked us to highlight the implications of additive costs (the ``Big Picture'').
+We agree that this was not emphasized enough in the submitted version, and we thank the referees for their suggestions on that issue.
 The revised version answers this concern in two points.
 	\begin{enumerate}
-    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we made our best to make this section easier to read, leaving the technical aspects in the Appendix (Section C).
-As such, we hope that the main message of this Section is better catched.
-    \item Most importantly, we now emphasize the ``Big Picture'' implications of additive costs on theoretical grounds.
-Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated (in Section 3 of the revised version), to quantify the extra welfare gains attributable to the reduction of the additive component.
-This analysis is driven in a new Section 5.
+		\item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section’s readibility by leaving the technical aspects in the Appendix (Section C).
+As such, we hope that the main message of this Section is easier to understand.
+		\item Most importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds.
+Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs estimated in Section 3.
+This allows us to quantify the extra welfare gains attributable to reduction of additive trade costs.
+This analysis forms Section 5.
 	\end{enumerate}
-\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, yet to be fully transparent about our results, we have added numerous detailed tables in the Online Appendix, to which the reader can refer to.
+	\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
 \end{itemize}
@@
 
-All in all, your suggestion and the other referee's one pointed in two directions for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the implications regarding gains from trade.
+All in all, your suggestion and the other referee's one pointed toward two possible directions for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the welfare implications.
 Both paths were undeniably worth exploring, but could not be followed in a single paper.
@@
 to  offer insights on the welfare implications of our results.
-In this regard, the new Section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the alterations to welfare gains involved by the relative variations of additive and multiplicative transport costs over our period of analysis.
+In this regard, the new Section 5 ``The role of additive cost: Theoretical insights'' is devoted to a theory-based analysis of the changes to welfare gains implied by the relative variations of additive and multiplicative transport costs over our period of analysis.
 
 Our choice to dig into the welfare implications of our results has two main motivations.
-Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} recently finds (though in a different theoretical setting) support for the iceberg trade cost assumption, i.e, iceberg accounting for almost all trade costs.
+Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} recently finds (though in a different theoretical setting) support for the ad-valorem trade cost assumption, i.e, ad-valorem accounting for almost all trade costs.
 Secondly, the implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, \cite{Hornok-et-al-RES-2015, Hornok-et-al-JIE-2015}, \cite{Alessandria-et-al-AER-2010} or \citet{sorensen2014}.
@@
 We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transports costs.
-Based on these empirical results, we investigate the welfare consequences of the reduction in international transport costs as observed over 1974-2019.
+Based on these empirical results, we investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019.
 In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the disproportionate part of additive transport costs in determining those welfare variations.\smallskip
 
-To this aim, we embed additive costs in the canonical \citet{melitz} model.
-More specifically, we start from the canonical \cites{melitz} model where we add additive trade costs.
+To this aim, we introduce additive costs in the canonical \citet{melitz} model.
 The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
 However, \citet{sorensen2014} exclusively performs a theoretical analysis, without any quantitative exercise.
-In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the reduction of additive costs (vs iceberg costs).
+In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the reduction of additive costs (vs ad-valorem costs).
 For this, they rely on a calibration for transport costs based on a single-year estimation (for the year 2004), from which they depart by assuming an ad-hoc hypothetical reduction of trade costs.
-In contrast, our own exercise relies on a several decades time span for the US, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on the values actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).\smallskip
+In contrast, our own exercise relies on a several decades of US data, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on the values actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).\smallskip
 
-Specifically, we use the estimation results reported in Section 3 concerning multiplicative and additive costs, corresponding to the values at the beginning and at the end of the sample.
-We investigate the different welfare consequences of alterations to multiplicative and additive costs, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
-In addition, we also assess how the results are modified by changes in sunk costs of exports, $f_{x}$.
-To that end, we adjust the share of exporting firms in the US: based on \citet{Lincoln_McCallum2018}, we set the latter to 35\% in 2019, versus 21\% in 1974.
-In our preferred exercise, we quantify the welfare gains that derive from the reduction in transport costs as documented in Section 3, relying on a combined reduction in each additive/iceberg component, along with the reduction in the fixed export cost.
-We compare this with the case where the total transport costs reduction is solely attributed to a decrease in ad-valorem costs.
+More precisely, we use the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, to implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
+To that end, we adjust the share of exporting firms in the US.
+Based on \citet{BEJK-AER-03, Lincoln_McCallum2018}.
+In our preferred exercise, we quantify the welfare gains that derive from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component in a context of fixed export cost reduction.
+We compare this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
 
 Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
-Specifically, we report the change in total transport costs decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
-Qualitatively, the conclusions are identical for both transport modes, with two major insights.
-First, for a given decrease in total transport costs,  welfare gains are around 50\% higher when this reduction is partly achieved through a reduction in the additive costs.
+Specifically, we report the welfare effect of change in total transport costs decomposed in its two dimensions (additive and ad-valorem),both in absolute and in relative terms.
+We have two major insights.
+First, using the reduction of transport costs in vessel (air),  welfare gains are around 50\% (14\%) higher when this reduction is achieved through the estimated reduction in additive and ad-valorem costs.
 Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing variable costs, with again an additional premium coming from the decrease of additive costs.
 
-Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
-Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
+Overall, these results add a substantial contribution to the paper : We are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+%Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
 
@@
 We agree that resorting on a linear form as the one suggested by the referee would substantially alleviate the computational burden.
-It does not seem to solve the issue yet.
+It would have some drawbacks.
 
@@
 $\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of transport costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
-Without these restrictions, standard linear, least squares estimates deliver aberrant values with more than mild quality-of-fit.
-Taking into account this constraint implies that the error term should be always positive, which we ensure by specifying the error term such that:
+Without these restrictions, standard linear, least squares estimates deliver aberrant values with mediocre quality-of-fit.
+Taking into account $p_{ikt} < \widetilde{p}_{ikt}$ also implies that the error term should be always positive, which we ensure by specifying the error term such that:
 
@@
 This specification obviously requires a non-linear estimation.
-We did our best to make the point clear in the revised version (see p.8).
+We did our best to make the point clearer in the revised version (see p.8).
 We also devote more time to the robustness of the empirical specification.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex
Fin referee 2, début éditeur

@@
 Until very recently, Mr.
-Sanchez never gave us any clear “yes or no” answer to our requests concerning a possible return in his DC office (closed because of the CoViD 19 restrictions), which would have allowed the delivery of the data.
-We can of course make available all our e-mail exchanges with Jeremy Sanchez.
+Sanchez did not gave us any clear “yes or no” answer to our requests concerning a possible return in his DC office (closed because of the CoViD 19 restrictions), which would have allowed the delivery of the data.
 In his last e-mail, he confirmed that access to the additional data was impossible.\medskip
 
-Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this e-mail) based on a HS10 dataset over the 2002-2019 period.
+Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this e-mail) based on a HS10 dataset over the 2005-2019 period.
 These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex
Fin réponse à l’éditeur.
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 We enclose detailed answers to the referees.
-Below we first detail how we handled two major points raised by both referees, before listing the other major changes brought to the paper.
+Below we detail how we handled two major points raised by both referees, before listing the other major changes brought to the paper.
 
 \section{Two Common Concerns: Estimation Method and Big Picture}
-Both referees questioned (i) our empirical methodology, especially our choice for a non-linear estimator and (ii) the lack of a ``big picture'' in the paper, i.e.
+Both referees questioned (i) our empirical methodology, especially our choice of a non-linear estimator and (ii) the lack of a ``big picture'' in the paper, i.e.
 of a deeper analysis of the implications of our results.
@@
 \subsection{Estimation Method}
-Both referees debated our choice for a non-linear estimator, suggesting alternative functional forms estimable with standard linear estimators.
+Both referees challenged our choice of a non-linear estimator, suggesting alternative functional forms that could be estimated with standard linear estimators.
 We do acknowledge that the initial version of the paper did not make a clear enough case for the non-linear estimator.
-We now explain in details, both in the papers and in our answers to reviewers, the reasons why a non-linear estimator is unavoidable in our case.
+We now explain in details, both in the papers and in our answers to referees, the reasons why a non-linear estimator is unavoidable in our case.
 This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
-$\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of transport costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
-Without these restrictions, standard linear, least squares estimates deliver aberrant values with more than mild quality-of-fit.
-Taking into account this constraint necessarily brings a functional requiring a non-linear estimation.
-We did our best to make the point clear in the revised version (see p.8).
+$\tau \geq 1$ and $t\geq 0$.
+This restriction is there because we know that both types of transport costs are positive and we do not want to neglect that information.
+By construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
+Without these restrictions, standard linear, least squares estimates deliver aberrant values with mediocre quality-of-fit.
+Taking into account this constraint necessarily brings up a functional requiring a non-linear estimation.
+We did our best to make the point clearer in the revised version (see p.8).
 \medskip
@@
 
-Firstly, we maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we made our best to make this section easier to read, leaving the technical aspects in the Appendix (Section C).
-As such, we hope that the main message of this Section is better catched.
+First, we maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we made our best to make this section easier to read, leaving the technical aspects in the Appendix (Section C).
+As such, we hope that the main message of this Section is easier to understand.
 \medskip
 
-Secondly, and more importantly, we now emphasize the ``Big Picture'' implications of additive costs on theoretical grounds.
+Second, and probably more importantly, we now emphasize the ``Big Picture'' implications of additive costs on theoretical grounds.
 Both referees suggested various, interesting paths to explore.
 Obviously, all of them could not be handled in a single paper.
-Therefore, we decide to focus on the welfare implications of our results.
+We have decided to focus on the welfare implications of our results.
 Through the lens of the Melitz's (2003) model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs that we have estimated (in Section 3 of the revised version), to quantify the extra welfare gains attributable to the reduction of the additive component.
-This analysis is driven in a new Section 5.
+This analysis is conducted in a new Section 5.
 We emphasize that the theoretical implications of additive costs have not been fully exhausted in Irarrazabal et al.
@@
 (2010), or Sorensen (2014).
-Specifically, we make use of our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantied by our empirical model) or solely through the ad-valorem cost (as standardly modelled in New Trade theoretical approaches).
+Specifically, we use our empirical results to investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019, contrasting two cases: when this reduction is achieved through a combined reduction in the ad-valorem and the additive components (as quantified by our empirical model) or solely through the ad-valorem cost (as modelled in standard New Trade theoretical approaches).
 To the best of our knowledge, this is the first time that such an assessment is preformed based on long-time, detailed trade data.
@@
 
-Overall, these results appear to add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
-Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
+Overall, we believe these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+%Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
 
@@
 
-\item \emph{Endogeneity issues}: Section 3.3.2 and Appendix B.3 provides detailed developments answering concerns expressed by Referee 1 concerning potential endogeneity in our estimates.
+\item \emph{Endogeneity issues}: Section 3.3.2 and Appendix B.3 provides detailed developments addressing concerns expressed by Referee 1 concerning potential endogeneity in our estimates.
 Relying on instrumental variables suggested by Referee 1, we provide first- and second-stage estimates, based on both at the 5-digit-product level and the more disaggregated HS10 level for sensitivity checks.
-Figure 3 on page XX reports our benchmark estimates by transport mode, together with their instrumented counterpart.
-In all cases, these estimates are very similar, when not identical.
-This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
+Figure 4 (page 17 \textbf{À vérifier xxxx} in the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+In all cases, these estimates are very similar.
+This suggests that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.
 
@@
 Secondly, as already mentioned at the beginning of this letter, it was only possible to get access to the years 1997-1999 and 2001-2019, due to the COVID situation.
-Considering the importance of the time length for our empirical exercise, especially in the context of the exercise implemented in Section 5, we decided to stick to our initial 5-digit database, and to provide systematic robustness estimates based on the HS10 data available.
+Considering the importance of having a long time span for our empirical exercise, especially in the context of the welfare analysis, we decided to stick to our initial 5-digit database, and to provide systematic robustness estimates based on the HS10 data available.
 These results are now reported in section 3.3.3 in the revised version.
-In particular, Table 3 and Figure 4 show the results are similar when using 10-digit product sand 5-digit products \textbf{XXX TO BE CONFIRMED XXX}
+In particular, Table 3 and Figure ) show the results are similar when using 10-digit product sand 5-digit products \textbf{XXX TO BE CONFIRMED XXX}
 
 \item \emph{Calculation of unit prices}: Referee 1 is concerned by the fact that we calculate unit price as Value/Weight, instead of Value/Quantity.
-We explain in details in our answer why a potential switch from weight to quantity for calculating unit prices would be harmful in our context: until 2004, information on quantity \underline{by transport mode} is unavailable XXX + other arguments XXX Therefore, we decided to provide robustness estimates on a reduced sample (\textbf{specify the number of years, value to trade flows in comparison with the full sample}) \textbf{XXX TO BE COMPLETED XXX}
+We explain in details in our answer and the online appendix why we decided against a switch from per-kg price to per-unit prices.
+First, there are data availability issues.
+Second, the estimation is even more computationaly intensive and third, we believe that the study of per-kg prices adds relevant information to the study of international transport costs, both because actual shipping costs are expressed in a per-weight/volume basis and because it allows us to integrate a larger share of trade in our analysis.
 
@@
 The robustness exercises included in the main text focus on the elements that were the most problematic to the referees.
-We hope that you will be pleased by the outcome.
+Its relevance to big picture debates is more obvious.
+We hope that you will be pleased by the outcome, and we apology again for the delay.
 

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
Relecture Answer Referee 2

@@
 
-The structure of the paper has been modified after taking into account the referees' comments.
+The structure of the paper has been significantly modified along the revision process, in particular to take into account the referees' comments.
 \begin{itemize}
@@
 [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
-	\item In the revised version, we have strengthened the robustness checks regarding the estimation of international transport costs (robustness checks to aggregation and to endogeneity have been added).
+	\item In the revised version, we have substantially strengthened the robustness checks regarding the estimation of international transport costs, with the addition of detailed checks to aggregation and endogeneity.
 Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
 Section 3 now includes the robustness checks as a final sub-section.
-For sake of saving space, the robustness analysis related to composition effects has been sentt to the Appendix (Section C.3).
-	\item Both referees asked us to highlight the implications of additive costs (the ``Big Picture'').
-We agree that this was not emphasized enough in the submitted version, and we thank the referees for their suggestions on that issue.
+For the sake of space, the robustness analysis related to composition effects has been sent to the Appendix (Section C.3).
+	\item Both referees asked us to devote more developments to the implications of additive costs (the ``Big Picture'').
+We do agree that this was not emphasized enough in the submitted version, and we thank the referees for their suggestions on that issue.
 The revised version answers this concern in two points.
 	\begin{enumerate}
-		\item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section’s readibility by leaving the technical aspects in the Appendix (Section C).
+		\item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section's readability by leaving the technical aspects in the Appendix (Section C).
 As such, we hope that the main message of this Section is easier to understand.
 		\item Most importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds.
-Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs estimated in Section 3.
-This allows us to quantify the extra welfare gains attributable to reduction of additive trade costs.
-This analysis forms Section 5.
+Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains deriving from the reduction in international transport costs estimated in Section 3 over 1974-2019.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period.
+This also allows us to quantify the extra welfare gains attributable to reduction of additive trade costs.
+This analysis is performed in Section 5.
+
 	\end{enumerate}
-	\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
+	\item To keep the paper (both the main text and its various appendices) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
 \end{itemize}
@@
 
-All in all, your suggestion and the other referee's one pointed toward two possible directions for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the welfare implications.
+All in all, your suggestion and the other referee's one pointed toward two possible options for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the welfare implications.
 Both paths were undeniably worth exploring, but could not be followed in a single paper.
@@
 Our choice to dig into the welfare implications of our results has two main motivations.
-Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} recently finds (though in a different theoretical setting) support for the ad-valorem trade cost assumption, i.e, ad-valorem accounting for almost all trade costs.
+Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} finds evidence supporting the ad-valorem assumption for denumerable goods, and consistently finds a very small welfare effect of additive costs within a partial-equilibrium model with perfect competition.
 Secondly, the implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, \cite{Hornok-et-al-RES-2015, Hornok-et-al-JIE-2015}, \cite{Alessandria-et-al-AER-2010} or \citet{sorensen2014}.
@@
 In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
-To that end, we adjust the share of exporting firms in the US.
-Based on \citet{BEJK-AER-03, Lincoln_McCallum2018}.
-In our preferred exercise, we quantify the welfare gains that derive from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component in a context of fixed export cost reduction.
-We compare this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
+To that end, we adjust the share of exporting firms in the US based on \citet{BEJK-AER-03} and \citet{Lincoln_McCallum2018}.
+In our preferred exercise, we quantify the welfare gains deriving from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component in a context of fixed export cost reduction.
+We compare setting this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
 
 Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
-Specifically, we report the welfare effect of change in total transport costs decomposed in its two dimensions (additive and ad-valorem),both in absolute and in relative terms.
+Specifically, we report the welfare effect of change in total transport costs decomposed in its two dimensions (additive and ad-valorem), both in absolute and in relative terms.
 We have two major insights.
@@
 
-Overall, these results add a substantial contribution to the paper : We are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+Overall, these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period covering the hyper-globalization moment, and to highlight the respective part of each component in the determination of these gains.
 %Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
@@
 We agree that resorting on a linear form as the one suggested by the referee would substantially alleviate the computational burden.
-It would have some drawbacks.
+However, it would have some important drawbacks.
 
 First, it can be noticed that the equation suggested by the referee does not stand in accordance with the assumption we borrow from \cite{Irrazabal_2015}, that ad-valorem costs is decomposed in its two country/sector dimensions multiplicatively (i.e.
-approximating $\tau_{is(k)}$ by the product $\tau_i\times \tau_{s(k)}$.
+approximating $\tau_{is(k)}$ by the product $\tau_i\times \tau_{s(k)}$).
 
@@
 This specification obviously requires a non-linear estimation.
-We did our best to make the point clearer in the revised version (see p.8).
+We did our best to make the point clearer in the revised version (see section 2.2, more specifically p.
+8).
 We also devote more time to the robustness of the empirical specification.
-On top of the robustness to the separability assumption (already in the submitted version), we also assess the robustness to the degree of aggregation and endogeneity (Section 3.4).
-We hope that this new version now convinces you that our estimation strategy is both relevant and robust.
+On top of the robustness to the separability assumption (already in the previously submitted version), we also assess the robustness to endogeneity and the level of aggregation (Sections 3.3.2 and 3.3.3).
+We hope that this new version will convince you that our estimation strategy is both relevant and robust.
 

### 2021-11-23 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement papier fin + OA en cours

@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-University Paris-Dauphine \& PSL University, IRD, LEDa, UMR 225, DIAL, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Univ.
-Lille, CNRS, IESEG School of Management, UMR 9221 - LEM - Lille Économie Management, F-59000 Lille, France, and CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805- UMR IRD 260, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
+Univ.
+Lille, CNRS, IESEG School of Management, UMR 9221 - LEM - Lille Économie Management, F-59000 Lille, France, and CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{
+University Paris Dauphine, PSL University, LEDa, UMR CNRS 8805- UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
 
@@
 The latter also depends on whether the transport costs are in proportion to the value of goods traded (hence do not distort relative prices) or are not related to the value of goods traded (hence do distort relative prices).
-Having a clear view of the international transport cost burden is a complex task, but of key importance.
+Having a clear view of the international transport cost burden hence appears as a complex task, but of key importance.
 Indeed, the answers to several important (both academic and policy) economic issues depend on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction.
@@
 as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis (\cites{samuelson1954}),\footnote{``Iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.} a recent strand of the empirical literature points to the existence of an additive component, dependent on the physical quantities traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}).
-Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or vessel transport: only waterborne and airborne shipments are included).\smallskip
+Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode.\smallskip
 
@@
 On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019.
-By allowing for country-product-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel.
-Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in vessel transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows.
+By allowing for country-sector-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given sector/country partner) over time, by roughly 50\% to 60\%.
+Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in vessel transport), composition effects amplify the reduction of pure transport costs.
+This contrasts with the conclusion obtained by \cite{hummels2007} also on US import flows.
 Importantly, the difference can be attributed to our more flexible modeling of the additive component.
@@
 Our paper complements these various findings in different dimensions.
-First, we provide a careful, long-run characterization of both additive and multiplicative component of transports costs.
+First, we provide a careful, long-run characterization of both additive and multiplicative components of transports costs.
 We show that additive costs represent a substantial part of total transport costs.
@@
 Section \ref{sec:data_method} presents the data and the baseline empirical model we use to estimate the values of international transport costs.
-Section \ref{sec:basic_results} reports the later on a yearly basis throughout the period 1974-2019 by transport mode, explicitly distinguishing between the additive and the ad-valorem components and providing several important robustness checks of our results.
-The two following sections take stock of those results to draw implications for two important debates in economic analysis.
-In Section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and partner country), and what arises from changes in transport costs \textit{per se}.
+Section \ref{sec:basic_results} reports the estimation results on a yearly basis throughout the period 1974-2019 by transport mode, explicitly distinguishing between the additive and the ad-valorem components and providing several important robustness checks of our results.
+The following sections take stock of those results to draw implications for two important debates in economic analysis.
+In Section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by sector and partner country), and what arises from changes in transport costs \textit{per se}.
 Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of the observed decrease in ad-valorem and additive costs.
@@
 We extend this database to 2019, using his original source, the US annual ``Imports of Merchandise'' provided by the US Census Bureau (details on the database provided in Appendix \ref{app:data}).
-We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price.
+We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode, air or vessel (specifically, only waterborne and airborne shipments are included).\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price.
 While both terms are closely related (see more details in Appendix \ref{app:data}), the US Foreign Trade Statistics reports fas prices.} More precisely, the fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, both items being conditional to the transport mode; in other words, it is the price for one kg of the goods net of transportation costs, by transport mode.\footnote{One may be concerned by the fact that unit prices are calculated as the value divided by the shipping weight, rather than to the quantity (the number) of goods.
@@
 Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
-This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the levels of both the ad-valorem costs and of the additive costs, in contrast to e.g.
+This is highly valuable, as it gives us a direct measure of international transport costs, from which we can estimate separately the levels of both the ad-valorem costs and of the additive costs, in contrast to e.g.
 \cite{Irrazabal_2015}.
@@
 
-As detailed below, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
-hence, our benchmark results estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
-Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin (hence, around 400 fixed effects).
+As detailed below, the use of a non-linear estimator triggers computational limitations that constrain the level of possible detail, especially when covering a long period of time.
+Hence, our benchmark results estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
+Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin (hence, around 400 fixed effects to estimate).
 We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i),} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
@@
 Specifically, denoting by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped, we have: $ p = \widetilde{p} +f$.
-As in \cite{hummels2010}, assume that the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents the transport costs level (depending on elements such as distance, port quality, etc.).
+As in \cite{hummels2010}, we assume that the shipping charges may depend on the price $\widetilde{p}$ according to the functional form $f = \widetilde{p}^{1-\beta} X$, where $X$ represents the transport costs level (depending on elements such as distance, port quality, etc.).
 The link between the import, export prices and shipping charges is hence given by:
@@
 At the other extreme, the case $\beta=1$ corresponds to shipping costs being independent of the unit price, with the cif price the sum of the unit price and the shipping cost: Transport costs are fully additive.
-The closer $\beta$ to 1, the more prevalent additive costs in total transport costs.
-
+
+This suggests that the closer $\beta$ to 1, the more prevalent additive costs in total transport costs.
 One can analytically show that $\beta$ is indeed equal to the share of additive costs in total transport costs.
-To do so, start from the link between the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as $p = \tau \widetilde{p}+ t$, with $\tau \geq 1$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported.
+To do so, start from the link between cif, fas prices and transport costs by explicitly decomposing between the ad-valorem and the additive costs as $p = \tau \widetilde{p}+ t$, with $\tau \geq 1$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported.
 We can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, leading to:
@@
 Considering that the LHS is positive by nature, we need to take into account a positivity constraint on the RHS.
-We fulfill this requirement by taking logs.
-The above equation then becomes:
+We fulfill this requirement by taking the log of the equation, which thus becomes:
 
@@
 
-\noindent where we ensure the positivity constraint on $\tau_x-1$, $t_x$ (with $x =i,k$) by taking logs.
+\noindent where we ensure the positivity constraint on $\tau_x-1$, $t_x$ (with $x =i,k$) by taking the logs of the parameters.
 
@@
 The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
-See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
+See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that constrain the level of possible detail, especially when covering a long period of time.
 Confronted with this trade-off, we estimate 3-digit level sectoral fixed effects as our benchmark.
@@
 This has driven us to estimate this third model (C) as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$.
-The main result that emerges is that this model (C) is dominated (in terms of quality of fit properties) by the model (A) with multiplicative costs only, which is itself dominated by the complete model (B).
+The main result that emerges is that the model (C) is dominated (in terms of quality of fit properties) by the model (A) with multiplicative costs only, which is itself dominated by the complete model (B).
 More details of these results are given in the Online Appendix, Section A.}
@@
 		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
-		\item 1989 trade by air is excluded
+		\item 1989 trade by air is excluded (see Section C of the Online Appendix).
 		%\item The additive term is expressed in fraction of fas price.
@@
 Table \ref{tab:summary_results} reports a summary of our results.
-It displays the weighted mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
+It displays the weighted mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2019, for both the baseline and the 4-digit sectorial level estimation.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport modes, at the 3- and 4-digit classification level.
 Results for all years (available at the 3-digit level) are reported in Section C from the Online Appendix.}
-The top panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.
+The top panel reports total transport costs statistics in the data.
 In the medium panel, we report the results for a specification based on ad-valorem costs estimated alone (Model (A), based on Equation (\ref{eq:model_nlI})).
@@
 Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.
-1989 trade by air is excluded.
+1989 trade by air is excluded (see the Online Appendix, Section C).
 \end{minipage}
@@
 
-As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -1.9\% for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
+As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by roughly the same magnitude of -2\% per year.
+This implies a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
 \cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
@@
 	\end{center}
-	\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule:  we select, for each year, the largest importers that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
-We keep in the sample all trade observations from these importers and in these sectors.}}
+	\parbox[l]{12cm}{\footnotesize{Notes: Estimations are performed on a restricted sample, based on the following rule: for each year, we identify \textit{(i)}, the largest importers that form together at least 80\% of annual trade and \textit{(ii)}, the largest traded sectors that form together at least 80\% of annual trade.
+We keep all trade observations that satisfy both conditions in the restricted sample.}}
 	\label{tab:robustness_separability}
@@
 
-It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
+It is true that we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.
 However, the latter can be biased by the aforementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.
@@
 In the two following sections, we take stock of those results to draw implications for two important debates in economic analysis.
-In the next section, we shed light on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \emph{per se}.
+In Section \ref{sec:results_trends}, we go deeper in the decrease of transport costs over time by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs \emph{per se}.
 Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of alterations over time to multiplicative and additive costs.
@@
 We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs.
-To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price on a per product-year-country basis, i.e.
+To do so, we compute from Equation (\ref{eq:tt2}) the elasticity of transport costs to the export price on a per product-year-country basis, i.e.
 $\beta$, according to $\beta_{ikt} = \frac{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}}{\widehat{t}_{is(k)t}/\widetilde{p}_{ikt}+\widehat{\tau}_{is(k)t}-1},
@@
 As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$.
-This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would imply a distribution concentrated on a single point.
+This stands in sharp contrast with the assumption made by \cite{hummels2007}, which would imply a distribution concentrated on a single point.
 Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{This it is also the case when we look at the distribution over the range of origin countries or products, for a given year.
@@
 
-Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, explicitly taking into account the varying share of additive costs over time, country partner and product.
+In this section, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, explicitly taking into account the varying share of additive costs over time, country partner and product.
 We start with the presentation of our estimation strategy before turning to the results.
@@
 
-The main result from Figure \ref{fig:totalTC_compeffects_excl} is that composition effects do not play a major role in accounting for the time trend of overall transport costs.
+The main result from Figure \ref{fig:totalTC_compeffects_excl} is that composition effects do not play a major role in accounting for the time trend of transport costs.
 Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the raw transport costs measure and the pure transport costs.
@@
 
-The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and vessel shipping.
+These results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and vessel shipping.
 If anything, we find the opposite result here: trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern.
@@
 
-Our aim is related to \cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020}.
+Our objective here relates to \cite{Irrazabal_2015}, \cite{sorensen2014} and \cite{Lashkaripour_JIE2020}.
 We proceed in a different way though.
 \cite{sorensen2014} explores the welfare effects of a trade costs reduction depending on whether they are either fully ad-valorem or fully additive on analytical grounds.
-In contrast, as in \cite{Irrazabal_2015}, we provide a quantification of such welfare gains.
+In contrast, as in \cite{Irrazabal_2015} and \cite{Lashkaripour_JIE2020}, we provide a quantification of such welfare gains.
 As \cite{Irrazabal_2015} provide estimates of transport costs only for one year (2004), their assessment of the welfare effects of a trade cost increase is based on a ad-hoc assumption (specifically, a 5\% increase in the ratio of additive cost to the median fas price).
@@
 Using a model with monopolistic competition and firm selection \`{a} la \cite{melitz} or \cite{chaney2008}, as also did \cite{sorensen2014} and \cite{Irrazabal_2015} makes a difference.
-In this setting, we show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
+We indeed show that additive costs do modify the selection of firms on both the domestic and the export markets, e.g the extensive margin of trade, whose role in the welfare gains of trade lies at the heart of the New Trade models  \`{a} la \cite{melitz}.
 Leaving aside the selection issue might underestimate the welfare gains that can be achieved by a reduction in additive transport costs.
@@
 with an exogenous probability $\delta>0$.
-The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$, units of labor, with $f$ the fixed cost of production.
+The production of $q$ units requires employing $l(q|\varphi) = f+\frac{q}{\varphi}$ units of labor, with $f$ the fixed cost of production.
 Exporting is subject to an extra fixed cost $f_x$ and variable exporting costs of two types: an ad-valorem cost that depends on the value traded $\tau\geq 1$ (as in \citealp{melitz}) and an additive cost depending on the quantity (weight) transported, $t\geq 0$.
-Introducting $t$ is our innovation compared to the canonical Mélitz model.
 
@@
 \paragraph{Profit-maximizing behaviors} After successful entry on a market (home or foreign), firms set a market-specific price.
-The firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of trade costs on the foreign destination market (see  Appendix \ref{app:theoretical_model}):
+Given CES preferences, the firm sets a constant markup on its marginal cost, which is a decreasing function of the firm's productivity $\varphi$, but also takes into account both components of variable trade costs on the foreign destination market (see  Appendix \ref{app:theoretical_model}):
 \begin{eqnarray}
@@
 \end{eqnarray}
-\noindent $r_d(\varphi)$ and $r_x(\varphi)$ are the sales revenues on each market (see Appendix \ref{app:theoretical_model} for a detailed expression).
-$B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$ .
-Profits increase with productivity $\varphi$.
-Only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
+\noindent where $r_d(\varphi)$ and $r_x(\varphi)$ are the sales revenues on each market (see Appendix \ref{app:theoretical_model} for a detailed expression) and $B \equiv \frac{1}{\sigma-1}\left(\frac{\sigma}{\sigma-1}\right)^{-\sigma}P^{\sigma-1}E$.
+Since profits increase with the productivity level $\varphi$, only sufficiently productive firms are able to cover the fixed costs of operating in a given market.
 Further, in presence of trade costs, profits for a given productivity are lower on the export market than on the domestic market.
@@
 on the support $\left[ \varphi^\ast_x,\infty\right)$.
-Further defining $M_T = M+M_x$ as the total mass of varieties accessible to the consumers and $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}), with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate:
+Further defining $M_T = M+M_x$ as the total mass of varieties accessible to the consumers and $\widetilde{\varphi}_T$ the weighted average of firms productivities on the domestic market, one can show that the welfare per worker can be written as (see Appendix \ref{app:theoretical_model}):
 \begin{equation}
@@
 \end{equation}
+\noindent with $\rho =  \frac{\sigma-1}{\sigma}$ inversely related to the markup rate.
 
@@
 This allows a full analytical solving of the M\'{e}litz model.
-This is no longer the case in the presence of additive costs (see Appendix \ref{app:theoretical_model}).
-In this case, we need to rely on simulations to determine the steady state.
-This drives us to calibrate the structural parameters of the model.
+This is no longer the case in the presence of additive costs.
+In this case, we need to rely on simulations to determine the steady state, for which we calibrate the structural parameters of the model.
 To do so, we use values commonly retained in the literature, as reported in Table \ref{tab:calib_horsTC}.
+The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC}.
 
@@
 \end{center}
-{\parbox[l]{12cm}{ \vspace{2pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
+{\parbox[l]{14cm}{ \vspace{2pt}\footnotesize{Notes: $f$ is endogenously revealed such that all firms produce in autarky.
 $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, from \cite{BEJK-AER-03} (BEKJ 2003 for short)) based on early 1980s value) and the final steady state (39\%, based on 2006 value), according to the references provided above.
@@
 \end{table}
-
-The one exception is the value for the two additive and ad-valorem transport costs, for which we use our empirical estimates of Section \ref{sec:results_decomposition} for the years 1974 and 2019, as reported in Table \ref{tab:calib_TC}.
 
@@
 We are interested in studying the welfare consequences of a trade cost reduction in presence or not of additive costs.
-To that aim, we will compare the equilibrium based on the values of transport costs at the beginning of our estimation period (1974) and at the very end (2019), in two settings: the first one including only ad-valorem transport costs (uneven columns in Table \ref{tab:calib_TC}).
+To that aim, we compare the equilibrium based on the values of transport costs at the beginning of our estimation period (1974) and at the very end (2019), in two settings: the first one including only ad-valorem transport costs (uneven columns in Table \ref{tab:calib_TC}).
 We here apply the estimated reduction in total variable transport costs ``as if'' it was solely attributable to the ad-valorem component.
@@
 The reduction in international transport costs channels through both its additive and ad-valorem components.
-Besides, \cite{BEJK-AER-03} and \cite{Lincoln_McCallum2018} documents the large increase in the number of exporting firms in the US between the mid 1980s et the mid 2000s, with a share of exporting firms raising by 52\% over this period.
+Besides, \cite{BEJK-AER-03} and \cite{Lincoln_McCallum2018} document the large increase in the number of exporting firms in the US between the mid 1980s and the mid 2000s, with a share of exporting firms raising by 52\% over this period.
 This can be captured in the model through a reduction in the fixed export cost $f_x$ (as reported in Table \ref{tab:calib_horsTC}).
-Accordingly, we will assess the effects of the trade liberalization assuming a joined reduction in both the variable transport costs ($\tau, t$, based on our empirical estimates) and the fixed export cost $f_x$.
 %\footnote{We endogenously determine the initial value of $f_x$ such that the initial steady state (with 1974 values of transport costs) implies a share of exporting firms equal to 21\%, as in  \cite{ghironi} based on the value estimated by \cite{BEJK-AER-03} in the early 1990s.
@@
 
-The results are reported in Table \ref{tab:resultats_modele}, panel (2), by transport mode (for Air, Columns (A)-(B) and Vessel, Columns (C)-(D) respectively).\footnote{For sake of space saving, we only report the welfare consequences of trade cost reduction, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
-The whole set of results is available upon request to the authors.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.
+The results are reported in Table \ref{tab:resultats_modele}, by transport mode (for Air, Columns (A)-(B) and Vessel, Columns (C)-(D) respectively).
+Specifically, we report the change in total variable transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms.\footnote{For sake of space saving, we only report the welfare consequences of trade cost reduction, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms.
+The whole set of results is available upon request to the authors.} Panel (2) displays the results of our preferred scenario where both variable and fixed export costs decrease.
 We also simulate an alternative scenario where only the variable transport costs decrease, the fixed export cost being constant at its initial value.
@@
 %, and the larger welfare gains when this specific component of international trade costs is reduced.
-See Appendix \ref{app:theoretical_model}, Table \ref{tab:results_model_appendix} for the results.}.
+See Appendix \ref{app:theoretical_model}, Table \ref{tab:results_model_appendix} for the results.}
 
@@
 As reported in Table \ref{tab:resultats_modele}, panel (1), columns (B) and (D), the welfare gains are stronger than in the pure ad-valorem case.
-Again, the effect is quantitatively sizeable: Welfare gains are thus between 30\% and 78\% higher, depending on the transport mode.
+The effect is quantitatively sizeable: Welfare gains are thus between 30\% and 78\% higher, depending on the transport mode.
 This is driven by a more positive price competitiveness effect, despite the more negative variety effect (comparing with Columns (A) and (C), same panel).
@@
 on the domestic market rises, as well as the average productivity. Accordingly, the
-price competitiveness effect increases by more when the trade liberalization involves a reduction
+price competitiveness effect increases by more when the trade cost decrease is partly achieved through a reduction
 in the additive component.
@@
 First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
-While this is not a surprising result \textit{per se} (we now reduce the reductions of two sources of international frictions, rather than one), this exercise delivers a quantification of the welfare gains associated with a reduction in the fixed export cost.
+While this is not a surprising result \textit{per se} (we now reduce two sources of international frictions, rather than one), this exercise delivers a quantification of the welfare gains associated with a reduction in the fixed export cost.
 They are substantial, higher by around 50\% relative to the case when only variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are even higher, by 66\% (Vessel) and 89\% (Air) respectively.
@@
 
-In this section, we explain in more detail the method employed to eliminate the country and product dimensions of the estimated transport cost.
+In this section, we explain the method employed to eliminate the country and product dimensions of the estimated transport cost in more detail.
 
@@
 As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (A)), the product dimension (Term (B)) and the pure transport costs time trend (Term (C)).
-Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the ad-valorem cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
+Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the ad-valorem cost is the product of the country of origin and the goods dimensions, while the additive cost is the sum of the two dimensions.
 Both equations are estimated by transport mode.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
changement papier fin + OA en cours

@@
 The paper compares the empirical performances of two models: one with only ad-valorem costs (Model (A)) and one with where both ad-valorem and additive costs (Model (B)).
-For comparison purposes, we also estimate the model with only additive costs (Model (C)), in which case the estimated equation is:
+For comparison purposes, we also estimate a third model with only additive costs (Model (C)), in which case the estimated equation is:
 $$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$$
@@
 Results for Models (A) and (B) are identical to those reported in the paper.
-Unsurprisingly, the estimated size of overall transport costs under Model (C) is of same order of magnitude as in Models (A) and (B).
+Unsurprisingly, the estimated size of overall transport costs under Model (C) is of same magnitude as in Models (A) and (B).
 We also observe a downward trend of transport costs over time, in particular since 1980.\\
@@
 
-To go further into the comparison of the empirical relevance of our three empirical models (A), (B) and (C), we compare quality of fit diagnostic tests in this section.
+To go further into the comparison of the empirical relevance of our three empirical models, we compare quality of fit diagnostic tests.
 Table \ref{tab:3models_diagnosis_air} reports the results for air transport, and Table \ref{tab:3models_diagnosis_vessel} those for vessel transport.
@@
 
-For all years and whatever the transport mode considered, the model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered.
+For all years and whatever the transport mode considered, the model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)).
+This conclusion holds whatever the type of diagnostic test considered.
 That justifies our choice to disregard Model (C) in the main text.
@@
 
-In this section, we report the estimation results when we retain the 4-digit classification level ($s$=4-digit).
+In this section, we report the estimation results when we retain a sectorial classification at the 4-digit level ($s$=4-digit).
 
@@
 Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and vessel transport respectively.
-Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs we found at the 3-digit level.
-
-\begin{table}[htbp]
-  \centering
-  \caption{Air: Transport costs estimates, selected years, 4-digit level}
-\begin{center}
-  \footnotesize{
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\ \hline
-\multicolumn{7}{l}{\textbf{Model (A)} ($\widehat{\tau}^{ice}-1$, in \%) }  \\
-\hline
-Mean  & 6.6 & 5.8 & 5.2 & 3.3 & 3.7 & 3.2 \\
-Median & 5.2 & 4.4 & 4.1 & 2.1 & 2.7 & 2.6 \\
-%Standard Error & 0.056 & 0.054 & 0.046 & \multicolumn{1}{c}{0.040} & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.025}  \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (B)}}  \\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term }($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
-Mean  & 3.5 & 2.6 & 3.1 & 1.5 & 2.1 & 1.6  \\
-Median & 2.5 & 1.7 & 1.9 & 1.0 & 1.7 & 1.4  \\
-%Standard Error & 0.036 & 0.028 & 0.030 & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.024} & \multicolumn{1}{c}{0.015}  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$), in \%}    \\ \hline
-Mean  & 2.6 & 2.1 & 1.7 & 1.2 &1.2 & 1.0 \\
-Median & 1.2 & 0.6 & 0.6 & 0.5 & 0.4 & 0.4  \\
-%Standard Error & 0.039 & 0.042 & 0.033 & \multicolumn{1}{c}{0.027} & \multicolumn{1}{c}{0.029} & \multicolumn{1}{c}{0.019} \\
-\hline
-\# observations & 14944 & 16844 & 25307 & \multicolumn{1}{c}{35005} & \multicolumn{1}{c}{38475} & \multicolumn{1}{c}{39460}  \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{10cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-}
-\end{center}
-  \label{tab:result_air_rob}
-\end{table}%
+Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs that we found at the 3-digit level.
 
@@
 		\label{tab:result_air_rob}%
-		\input{TableB1_air.tex}
+		\input{TableB1_air_reshaped.tex}
 		\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
+			\scriptsize
+			\item Statistics are obtained weighting each observation by its share in trade (mode-dependent).
 			\item \textbf{Model (A): Ad-valorem transport costs only}
@@
 	\centering
-\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
-\begin{center}
-  \footnotesize{
-    \begin{tabular}{l|cccccc}
-   \hline\hline
-Year & 1974  & 1981  & 1989  & 2001  & 2009  & 2013 \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (A)} ($\widehat{\tau}^{ice}-1$, in \%)} \\
-\hline
-Mean  & 9.8 & 6.1 & 5.8 & 5.1 & 4.2 & 3.6  \\
-Median & 9.4 & 5.1 & 4.8 & 4.5 & 3.8 & 3.1 \\
-%Standard Error & 0.060 & 0.038 & \multicolumn{1}{c}{0.036} & \multicolumn{1}{c}{0.030} & \multicolumn{1}{c}{0.023} & \multicolumn{1}{c}{0.020} \\
-\hline
-\multicolumn{7}{l}{\textbf{Model (B)} }\\ \hline
-\multicolumn{7}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%) }   \\ \hline
-Mean  & 5.4 & 3.4 & 2.8 & 2.8 & 2.4 & 2.1  \\
-Median & 4.9 & 3.0 & 2.4 & 2.5 & 2.6 & 1.8 \\
-%Standard Error & 0.043 & 0.026 & \multicolumn{1}{c}{0.025} & \multicolumn{1}{c}{0.021} & \multicolumn{1}{c}{0.016} & \multicolumn{1}{c}{0.013}  \\
-\hline
-\multicolumn{7}{l}{\textit{Additive term} ($\widehat{t}^{add}/\widetilde{p}$, in \%) }   \\ \hline
-Mean  & 4.6 & 2.6 & 3.1 & 2.4 & 2.1 & 1.5  \\
-Median & 2.9 & 1.3 & 1.9 & 1.5 & 1.3 & 0.8 \\
-%Standard Error & 0.068 & 0.044 & \multicolumn{1}{c}{0.037} & \multicolumn{1}{c}{0.035} & \multicolumn{1}{c}{0.031} & \multicolumn{1}{c}{0.023} \\
-\hline
-\# observations & 19196 & 17916 & \multicolumn{1}{c}{29387} & \multicolumn{1}{c}{36677} & \multicolumn{1}{c}{37643} & \multicolumn{1}{c}{38820} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{10cm}{ \vspace{7pt}\scriptsize{Notes: TC = Transport Costs.
-Statistics are obtained weighting each observation by its share in trade (mode-dependent).
-Additive term expressed in fraction of fas price.}}}
-\end{tabular}%
-}
-\end{center}
-\label{tab:result_ves_rob}%
-\end{table}%
-
-\begin{table}[htbp]
-	\centering
 %	\footnotesize{
@@
 		\label{tab:result_air_rob}%
-		\input{TableB1_ves.tex}
+		\input{TableB1_ves_reshaped.tex}
 		\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
+			\scriptsize
+			\item Statistics are obtained weighting each observation by its share in trade (mode-dependent).
 			\item \textbf{Model (A): Ad-valorem transport costs only}
@@
 	\centering
-  \caption{Air: Measures of goodness of fit, 4-digit level}
-\begin{center}
-\label{tab:good_fit_air_rob}%
-%\vspace{0.5cm}
-%\scalebox{0.97}{
-  \footnotesize{
-\begin{tabular}{l|cccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}              \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
-\hline
-\multicolumn{7}{l}{\textbf{R$^2$}}      \\ \hline
-Model (A)& \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
-Model (B) & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\ \hline
-\multicolumn{7}{l}{\textbf{SER} }  \\ \hline
-Model (A)& \multicolumn{1}{c}{0.8} & \multicolumn{1}{c}{0.9} & \multicolumn{1}{c}{0.83} &   0.87    & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
-Model (B) & \multicolumn{1}{c}{0.67} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.69} &  0.80     & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
-\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (A) & \multicolumn{1}{c}{-17505.6} & \multicolumn{1}{c}{-21813.5} & \multicolumn{1}{c}{-30960.6} & \multicolumn{1}{c}{-44067.6} & \multicolumn{1}{c}{-49375.6} & \multicolumn{1}{c}{-53197.9}  \\
-Model (B) & \multicolumn{1}{c}{-14895.8} & \multicolumn{1}{c}{-18589.9} & \multicolumn{1}{c}{-26553.5} & \multicolumn{1}{c}{-37297.9} & \multicolumn{1}{c}{-45747.6} & \multicolumn{1}{c}{-49899.1}  \\ \hline
-\multicolumn{7}{l}{\textbf{AIC criteria} }  \\ \hline
-Model (A)& \multicolumn{1}{c}{36243.1} & \multicolumn{1}{c}{44966.9} & \multicolumn{1}{c}{63417.1} & \multicolumn{1}{c}{89747.2} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.7} \\
-Model (B) & \multicolumn{1}{c}{31873.6} & \multicolumn{1}{c}{39495.8} & \multicolumn{1}{c}{55777.1} & \multicolumn{1}{c}{77439.9} & \multicolumn{1}{c}{94059.1} & \multicolumn{1}{c}{102224.3}  \\ \hline
-\multicolumn{7}{l}{\textbf{Test LL} }    \\ \hline
-Stat LL ratio (B vs A) & \multicolumn{1}{c}{5219.5} & \multicolumn{1}{c}{6447.1} & \multicolumn{1}{c}{8814.1} & \multicolumn{1}{c}{13539.4} & \multicolumn{1}{c}{7256.0} & \multicolumn{1}{c}{6597.5}  \\
-\# of restrictions (B vs A) & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
-p-value (B vs A) & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.000} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
-Model (B) = with additive \& ad-valorem transport costs.
-$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
-\end{tabular}%
-}
-\end{center}
-\end{table}%
-
-\begin{table}[htbp]
-	\centering
 %	\footnotesize{
@@
 		\label{tab:good_fit_air_rob}%
-		\input{TableB3_air.tex}
-		\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): Ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-		\end{tablenotes}
-\end{table}
-
-\begin{table}[htbp]
-  \centering
-  \caption{Vessel: Measures of goodness of fit, 4-digit level}
-\begin{center}
-\label{tab:good_fit_ves_rob}%
-  \footnotesize{
-\begin{tabular}{l|cccccc}
-\hline
-\hline
-      & \multicolumn{6}{c}{Year}                   \\
-      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
-\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
-Model (A) & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
-Model (B) & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
-\multicolumn{7}{l}{\textbf{SER} }\\ \hline
-Model (A) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
-Model (B) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
-\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
-Model (A)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
-Model (B) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
-\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
-Model (A) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
-Model (B) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
-\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
-Stat LL ratio (B vs A) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
-\# of restrictions (B vs A) & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
-p-value (B vs A) & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
-\hline\hline
-\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
+		\input{TableB3_air_reshaped.tex}
+{\parbox[l]{12cm}{
+			\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
 Model (B) = with additive \& ad-valorem transport costs.
 $R^{2}$ between the log of predicted ratio and the log of the observed ratio.
-The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
-\end{tabular}%
-}
-\end{center}
-\end{table}%
+The number \# of restrictions is equal to the number of parameters estimated, i.e.
+the number of partner countries plus the number of sectors.
+}}}
+\end{table}
 
@@
 		\label{tab:good_fit_air_rob}%
-		\input{TableB3_ves.tex}
-		\begin{tablenotes}
-			\tiny
-			\item Statistics are weighted by value
-			\item \textbf{Model (A): Ad-valorem transport costs only}
-			\item \textbf{Model (B): With additive and ad-valorem transport costs}
-		\end{tablenotes}
+		\input{TableB3_ves_reshaped.tex}
+		%\begin{tablenotes}
+{\parbox[l]{12cm}{
+			\scriptsize{ Notes: Model (A) = with only ad-valorem transport costs.
+Model (B) = with additive \& ad-valorem transport costs.
+$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
+The number \# of restrictions is equal to the number of parameters estimated, i.e.
+the number of partner countries plus the number of sectors.
+}}}
+		%\end{tablenotes}
 	\end{table}
@@
 
-As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000).
+As it can be inferred from Table \ref{tab_oa:result_air_ally3}, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000).
 This can be attributed to the presence of outliers in the distribution of the additive costs estimates, despite the exclusion of outliers in the data.
 The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000.
-Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
-Excluding or excluding a single year out of 45 does not make much difference in the weighted mean transport cost.
-%When 1989 is included, the weighted mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded.
-The weighted median value is left unchanged at 2.9\% whether 1989 is included or not.
+Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport, all the more as excluding or not excluding a single year out of 45 ones does not make much difference in the weighted mean transport cost.
 
@@
 Specifically, we determine the share of the observed variance in the ratio $\ln(\frac{p_{ik}}{\widetilde{p}_{ik}}-1)$ that comes from \textit{i)} the between-product variance (at the 5-digit level, $k$) and \textit{ii)} the between-sector variance (at the 3-digit level, $s$).
-The variance decomposition expression at the product level $k$ (5-digit level) is obtained by applying the following formula (by year and transport mode):
+The variance decomposition expression at the product level $k$ is obtained by applying the following formula (by year and transport mode):
 
@@
 
-In this section, we explore an alternative estimation method, which proceeds as follows. $\beta$ can be separately estimated for each industry $s$-origin country $i$ using the following regression:
+In this section, we explore an alternative estimation method (so called ``direct method'' for sake of clarity), which proceeds as follows.
+$\beta$ can be separately estimated for each sector $s$-origin country $i$ using the following regression:
 
@@
 
-where $d$ denotes the US district of entry and $k$ denotes an HS-10 product, $TC_{ikd}$ being the transport costs and $Z$ a set of controls variables.
-The identification of $\beta_{is}$, in this case, relies on exploiting the variability between sub-sectors at the 10-digit
+\noindent where $d$ denotes the US district of entry and $k$ denotes an HS-10 product, $TC_{ikd}$ being the transport costs and $Z$ a set of controls variables.
+The identification of $\beta_{is}$, in this case, relies on exploiting the variability between products at the 10-digit
 level ($k$) and between ports of entry in the US ($d$).
@@
 With respect to the trade flows coverage, the method is run country by country, and 3-digit sector by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many pairs (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, to estimate the equation..
+Yet, it appears that for many pairs (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, to estimate the equation.
 This can be seen comparing the number of observations by year/ transport mode between this alternative method and our baseline method reported in Table \ref{tab_comp_referee1_baseline}.
@@
 
-If we take the value of the $\beta$ by itself, there is no clear criteria to discriminate between the value estimated with the baseline method and estimating directly $\beta$ (when, of course, run on the same sample).
+If we take the value of $\beta$ by itself, there is no clear criteria to discriminate between the value estimated with the baseline method and estimating directly $\beta$ (when, of course, run on the same sample).
 Further, as shown in Table \ref{tab_comp_referee1_baseline}, there is no clear sign of a potential upwards or downwards bias that would be attached to either method.
+
 Things are more clear-cut in terms of accuracy of the estimation.
-Specifically, the baseline method yields a more precise estimation of the $\beta$ than the direct method.
-To develop on this, for each year and transport mode:
+Specifically, we show that the baseline method yields a more precise estimation of the $\beta$ than the direct method.
+We reach this conclusion by implementing the following steps, for each year and transport mode.
 \begin{itemize}
-	\item The direct method estimates one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{dir}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
+	\item The direct method estimates one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{dir}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode (where the subscript $^{dir}$ identifies the direct method).
 From this, we can compute the size of the 5-95\% confidence interval $CI^{dir}_{is} = \hat{\beta}_{is}^{max,dir} - \hat{\beta}_{is}^{min,dir}$ where:
@@
 As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
-We compute the accuracy of the $\beta$ estimation through a bootstrap method.
-Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.
-Writing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the direct method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
+We hence compute the accuracy of the $\beta$ estimation through a bootstrap method.
+Specifically, on a yearly/mode basis we draw a distribution of transport costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.
+Denoting $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the direct method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
 }
@@
 From a statistical point of view, it implies leaving aside information contained in the fact that transport costs have a both a country-specific component and a sector-specific component.
-We believe that explains the better accuracy of our baseline estimation.
+We believe that this explains the better accuracy of our baseline estimation.
 
+\textbf{XXX Lise stop here relecture XXX }
 \subsubsection{Weights and quantities}

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex

@@
 We also devote more time to the robustness of the empirical specification.
-On top of the robustness to the separability assumption (already in the previously submitted version), we also assess the robustness to endogeneity and the level of aggregation (Sections 3.3.2 and 3.3.3).
+On top of the robustness to the separability assumption (already in the previously submitted version), we also assess the robustness to potential endogeneity and the level of aggregation (Sections 3.3.2 and 3.3.3).
 We hope that this new version will convince you that our estimation strategy is both relevant and robust.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex

@@
 \\
-\noindent \noindent We would like to thank you for the opportunity to resubmit our paper, and for your patience in waiting the outcome.
+\noindent \noindent We would like to sincerely thank you for the opportunity to resubmit our paper, and for your patience in waiting the outcome.
 We are very grateful for the deadline extensions you kindly granted us.
-Obviously, the health crisis and the various consecutive lockdowns made the access to our offices difficult in several occurrences.
+Obviously, the CoViD-19 crisis and the various consecutive lockdowns made the access to our offices difficult in several occurrences, sometimes for several months.
 \medskip
@@
 We ordered this new data from the Census Bureau at the end of December 2019.
-However, some of the years we requested (2000 and 2001, as well as years before 1998) are not on their servers: they are back at their office which they have not been able to access since last spring apparently, due to the current health situation.
+However, some of the years we requested (2000 and 2001, as well as years before 1998) are not on their servers: these data are back at their office which they have not been able to access since last spring apparently, due to the current health situation.
 Since then, the situation has never evolved, and we are still currently waiting for these missing years.
@@
 
-Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this e-mail) based on a HS10 dataset over the 2005-2019 period.
+Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this letter) based on a HS10 dataset over the 2005-2019 period.
 These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.
@@
 We do acknowledge that the initial version of the paper did not make a clear enough case for the non-linear estimator.
-We now explain in details, both in the papers and in our answers to referees, the reasons why a non-linear estimator is unavoidable in our case.
+We now explain in details, both in the paper and in our answers to referees, the reasons why a non-linear estimator is unavoidable in our case.
 This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
 $\tau \geq 1$ and $t\geq 0$.
-This restriction is there because we know that both types of transport costs are positive and we do not want to neglect that information.
+These restrictions are required because we know that both types of transport costs are positive and we cannot neglect this characteristic.
 By construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
 Without these restrictions, standard linear, least squares estimates deliver aberrant values with mediocre quality-of-fit.
-Taking into account this constraint necessarily brings up a functional requiring a non-linear estimation.
-We did our best to make the point clearer in the revised version (see p.8).
+Taking into account this constraint necessarily brings up a functional form requiring a non-linear estimation.
+We did our best to make the point clearer in the revised version (see section 2.2, more specifically p.
+8).
 \medskip
@@
 The revised version devotes much more space to explicit tests of robustness of the empirical specification.
-On top of the robustness to the separability assumption (already in the initially submitted version), we also assess the robustness to the degree of data aggregation and endogeneity (Section 3.3) - see the list of modifications in section 2 from this letter.
+On top of the robustness to the separability assumption (already in the initially submitted version), we also assess the robustness to potential endogeneity and the level of aggregation (Sections 3.3.2 and 3.3.3) - see the list of modifications in section 2 from this letter.
 \medskip
 
-In this regard, we also point in our answers that implementing the methodology suggested by Referee 1 would imply non-negligible costs.
+Finally, we also point in our answers that implementing the estimation methodology suggested by Referee 1 would imply non-negligible costs.
 Firstly, this would leave us much less data to exploit, partly  because of the already mentioned CoViD issues - we would lose half of the years in the sample.
@@
 
-Overall, we believe these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+Overall, we believe these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period covering the hyper-globalization moment, and to highlight the respective part of each component in the determination of these gains.
 %Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
@@
 \item \emph{Endogeneity issues}: Section 3.3.2 and Appendix B.3 provides detailed developments addressing concerns expressed by Referee 1 concerning potential endogeneity in our estimates.
-Relying on instrumental variables suggested by Referee 1, we provide first- and second-stage estimates, based on both at the 5-digit-product level and the more disaggregated HS10 level for sensitivity checks.
-Figure 4 (page 17 \textbf{À vérifier xxxx} in the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+Relying on instrumental variables suggested by Referee 1, we provide first- and second-stage estimates, based on both the 5-digit-product level and the more disaggregated HS10 level for sensitivity checks.
+Figure 4 on page 17 in the paper reports our benchmark estimates by transport mode, together with their instrumented counterparts.
 In all cases, these estimates are very similar.
@@
 These results are now reported in section 3.3.3 in the revised version.
-In particular, Table 3 and Figure ) show the results are similar when using 10-digit product sand 5-digit products \textbf{XXX TO BE CONFIRMED XXX}
+In particular, Table 3 and Figure 5 show the results are similar when using 10-digit product sand 5-digit products.
 
 \item \emph{Calculation of unit prices}: Referee 1 is concerned by the fact that we calculate unit price as Value/Weight, instead of Value/Quantity.
-We explain in details in our answer and the online appendix why we decided against a switch from per-kg price to per-unit prices.
+We explain in details in our answer and the Online Appendix why we decided against a switch from per-kg price to per-unit prices.
 First, there are data availability issues.
-Second, the estimation is even more computationaly intensive and third, we believe that the study of per-kg prices adds relevant information to the study of international transport costs, both because actual shipping costs are expressed in a per-weight/volume basis and because it allows us to integrate a larger share of trade in our analysis.
+Second, the estimation is even more computationally intensive.
+Third, we believe that the study of per-kg prices adds relevant information to the study of international transport costs, both because actual shipping costs are expressed in a per-weight/volume basis and because it allows us to integrate a larger share of trade in our analysis.
 
@@
 The robustness exercises included in the main text focus on the elements that were the most problematic to the referees.
-Its relevance to big picture debates is more obvious.
-We hope that you will be pleased by the outcome, and we apology again for the delay.
+The paper's relevance to big picture debates is more obvious.
+We hope that you will be pleased by the outcome, and we sincerely apologize again for the delay.
 

### 2021-11-24 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Lise fin relecture sur OA

@@
 
-This section reports two supplementary checks concerning the separability assumption.
-
-\paragraph{Comparison on a yearly basis} Figure \ref{fig:robustesse_non_separe} displays the results on a yearly basis for air transport in panel (A) and for vessel transport in panel (B).
+In this section, we report more results on the separability assumption robustness check.
+Figure \ref{fig:robustesse_non_separe} displays the results on a yearly basis for air transport in panel (A) and for vessel transport in panel (B).
 In each case, the estimated values (as well as the regression line) under both separability (baseline) and non-separability are reported, for the additive term (upper panel) and the ad-valorem term (lower panel).

### 2021-11-24 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
relecture OA (???)

@@
 Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and \textit{across} origin countries.
-Put in plain words, it assumes that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
+Put in plain words, it assumes that transport costs of say, cars, have nothing in common whether those goods come from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
 This is a disputable assumption.

### 2021-11-24 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
relecture OA Lise fin

@@
 Using it would hence reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
-In our view, the historical coverage is interesting \emph{per se}: it provides useful insights about how transport costs have evolved over several decades and more importantly, it allows us to highlight the welfare alterations induced over the ``hyper-globalization" period by the relative dynamics of additive and multiplicative costs.
+In our view, the historical coverage is interesting \emph{per se}: it provides useful insights about how transport costs have evolved over several decades and more importantly, it allows us to highlight the welfare alterations induced over the ``hyper-globalization'' period by the relative dynamics of additive and multiplicative costs.
 
@@
 
-\textbf{XXX Lise stop here relecture XXX }
 \subsubsection{Weights and quantities}
@@
 In contrast to the aggregation and the separability issues, these restrictions strongly limit both the number of sectors and of countries covered, and do also impact the value of trade flows covered.
-This raises a non-trivial sample selection bias that may reduce the general scope of the transport cost estimates.
+A related argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png}.
+After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
+According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
+Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
+
+\begin{figure}[htbp]
+	\caption{Estimated share of discrete goods in US imports, 1974-2019}
+	\begin{center}
+		\begin{tabular}{cc}
+			\includegraphics[height=8cm]{../../revised_article/Share_of_discrete_goods.png}\\
+		\end{tabular}
+	\end{center}
+	\label{graph_Share_of_discrete_goods.png}%
+\end{figure}
+
 Second, in terms of $\beta$ estimates, the results clearly show that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
@@
 Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
-
-Another argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png}.
-After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
-According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
-Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
-
-\begin{figure}[htbp]
-	\caption{Estimated share of discrete goods in US imports, 1974-2019}
-	\begin{center}
-		\begin{tabular}{cc}
-			\includegraphics[height=8cm]{../../revised_article/Share_of_discrete_goods.png}\\
-		\end{tabular}
-	\end{center}
-	\label{graph_Share_of_discrete_goods.png}%
-\end{figure}
 
@@
 
-The situation is similar in the primary sector.
-In this case, in air transport te composition effect matters more (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)).
+The situation is different in the primary sector.
+In this case, in air transport the composition effects matter more (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)).
 

### 2021-11-24 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
changement papier + OA - ENCORE A RELIRE

@@
 In contrast to the aggregation and the separability issues, these restrictions strongly limit both the number of sectors and of countries covered, and do also impact the value of trade flows covered.
-A related argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png}.
-After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
-According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
-Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
-
-\begin{figure}[htbp]
-	\caption{Estimated share of discrete goods in US imports, 1974-2019}
-	\begin{center}
-		\begin{tabular}{cc}
-			\includegraphics[height=8cm]{../../revised_article/Share_of_discrete_goods.png}\\
-		\end{tabular}
-	\end{center}
-	\label{graph_Share_of_discrete_goods.png}%
-\end{figure}
+This raises a non-trivial sample selection bias that would limit the general scope of our transport costs estimates.
 
@@
 Accordingly, studying the technical per-volume or per-weight shipping cost is interesting to understand the context in which they operate.
+
+In a related manner, restricting the sample to denumerable goods would defeat our purpose of estimating international transport costs for the whole range of US imports.
+This is illustrated in Figure \ref{graph_Share_of_discrete_goods.png}.
+After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
+According to this rough approximation, discrete goods are between 23\% and 52\% of US imports, reaching a value only above 50\% for two years.
+Switching to a per-unit approach would hence mean a marked reduction in the trade flows coverage, thereby limiting the general scope of our estimates.
+
+\begin{figure}[htbp]
+	\caption{Estimated share of discrete goods in US imports, 1974-2019}
+	\begin{center}
+		\begin{tabular}{cc}
+			\includegraphics[height=8cm]{../../revised_article/Share_of_discrete_goods.png}\\
+		\end{tabular}
+	\end{center}
+	\label{graph_Share_of_discrete_goods.png}%
+\end{figure}
 
@@
 In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
-In vessel transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line): trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)).
-Trade composition effects matter more in air transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)), primarily in the ad-valorem component.
-As for the whole range of flows, the 60\% decrease in the raw transport costs in vessel can de decomposed in a 50\% decrease in the ``ceteris paribus'' transport costs, the 10\% remaining to trade composition effects.
+One can further notice that in vessel transport, composition effects play virtually no role at the end of the period, as most of the decrease can be imputed to the reduction of ``ceteris paribus'' transport costs (the blue continuous line)
 
 The situation is different in the primary sector.
-In this case, in air transport the composition effects matter more (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)).
+In air transport, the role of composition effects exhibits an irregular pattern over time.
+They dampen the reduction in pure transport costs up to the 2000s.
+Afterwards, they conversely more than compensate the increase in pure transport costs (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)).
+In vessel transport, the composition effects play a limited role in the primary sector as for the whole range of goods matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)), by slightly amplifying the reduction in the pure transport costs in both cases.
 

### 2021-11-24 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Relecture answer R1

@@
 We would like to thank you for your insightful comments. They led us to introduce some
-significant changes to the paper that we hope address your concerns. We first give you an overview
+significant changes to the paper that we hope address your concerns. We first give you a overview
 of the revision (Section \ref{sec:main_changes}) before answering in detail each of your comments (Section \ref{sec:detailed_answers}).
@@
 [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
-\item In the revised version, we have strengthened the robustness checks regarding the estimation of international transport costs (robustness checks to aggregation and to endogeneity have been added).
+\item In the revised version, we have strengthened the robustness checks regarding the estimation of international transport costs (robustness checks to the level of aggregation and to endogeneity have been added).
 Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
 Section 3 now includes the robustness checks as a final sub-section.
-For sake of saving space, the robustness analysis related to composition effects has been sentt to the Appendix (Section C.3).
-\item Both referees asked us to highlight the implications of additive costs (the ``Big Picture'').
+For the sake of saving space, the robustness analysis related to composition effects has been sent to the Appendix (Section C.3).
+\item Both referees asked us to devote more developments to the implications of additive costs (the ``Big Picture'').
 We agree that this was not emphasized enough in the submitted version, and we thank the referees for their suggestions on that issue.
@@
     \begin{enumerate}
-    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section’s readibility by leaving the technical aspects in the Appendix (Section C).
+    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section's readability by leaving the technical aspects in the Appendix (Section C).
 As such, we hope that the main message of this Section is easier to understand.
-    \item Most importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds.
-Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs estimated in Section 3.
-This allows us to quantify the extra welfare gains attributable to reduction of additive trade costs.
-This analysis forms Section 5.
+    \item More importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds.
+Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains deriving from the reduction in international transport costs estimated in Section 3 over 1974-2019.
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period.
+This also allows us to quantify the extra welfare gains attributable to reduction of additive trade costs.
+This analysis is performed in Section 5.
+
         \end{enumerate}
-\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
+\item To keep the paper (both the main text and its various appendices) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
 \end{itemize}
@@
 Further, we provide a robustness check for this separability assumption that $\tau_{ik} = \tau_i\tau_{k}$ and
- $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (Table 2 and \textbf{Figure XXX}) \textbf{XXX CHECK AT THE END XXX}).
+ $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (see Table 2 and Figure 3 on pp.
+15-16).
 We check the robustness of our results by re-running the estimation without the separability assumption.
@@
 In 2019, for example, there are 14,016 country$\times$sector pairs.
-Without the separability assumption, one must estimate 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$).
+Without the separability assumption, one must estimate 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixed effects) rather than ``only'' 836 ($(188+230)\times 2$).
 This is computationally impossible.
@@
  We keep in the sample all trade observations from these exporters and in these sectors.
- This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1).
+ This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2 in the paper), vs more than 30,000 on the complete sample (see Table 1 in the paper).
 It yet remains quite large in terms of trade coverage (mean of 68\%).\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the baseline sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the baseline sample).
 The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} On this reduced sample, we re-run the estimation with and without the separability assumption.
-The results are reported in Section 3.3.1 (Table 2 and \textbf{Figure XXX}).
-\textbf{XXX CHECK AT THE END XXX}) As we conclude at the end of this section, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
+The results are reported in Section 3.3.1 (Table 2 and Figure 3 on pp.
+15-16).
+As we conclude at the end of this section, the trend pattern of the share of additive costs are very similar whether estimated under the separability assumption or not.
  % use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
@@
 \textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels's one.
-We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'' \textbf{XXX Check XX}).
-As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{is(k)}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
+We took this remark into account by adding a new subsection in the revised version (Section 2.2, paragraph ``Estimation strategy'').
+As we now show in this subsection, the elasticity of transport costs to unit prices $\beta_{is(k)}$ (in absolute value) also corresponds to the share of additive costs in total transport costs.
 The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices, along the lines you suggest.
@@
 We develop each line of argument in our answer.
-Yet, because the referee's estimation method is also an interesting alternative, we have added a new section in the Online appendix devoted to it (Section D.3 \textbf{XXX CHECK XXX }).
+Yet, because the referee's estimation method is also an interesting alternative, we have added a new section in the Online appendix devoted to the latter (Section D.3).
 
@@
 In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 12 page 8.
-\textbf{XXX Verifier à la fin?
-GD XXX }\smallskip
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 11 page 8.\smallskip
 
@@
 The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007), yielding Equation (\ref{eq:estimation_ref1}) as estimation equation.
-By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
+By year and transport mode, this implies running the estimation for each country of origin $i$ and each $s(k)$ = 3-digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
 Despite its interest, the referee's method has some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
-First, we challenge the advantages that the referee's estimation method would bring in comparison with ours.
+First, we debate the advantages that the referee's estimation method would bring in comparison with ours.
 Second, we emphasize the drawbacks induced by this method.
@@
 This is based on the conclusion drawn from the robustness check on a reduced sample.
-As we conclude at the end of Section 3.3.1 \textbf{XXX CHECK XXX}, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
+As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.
 
@@
 We thank the referee for this valuable suggestion.
-It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.
+It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version (see section B.3 in the paper's appendix).
 
@@
      \textbf{Our answer:}
-     We agree with the referee that the comparison with the literature (\citealp{hummels2007} in particular) was not straightforward for the reader in the submitted version.
+     We agree with the referee that the comparison with the literature (\citealp{hummels2007}, in particular) was not straightforward for the reader in the initially submitted version.
 We did our best to make the comparison clearer in the revised version, in a way consistent with our estimation method.
-In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
+In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in \citet{hummels2007}:
     \begin{equation}
@@
      with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ikt}$ the price-elasticity of transport costs, with $i$ the origin country, $k$ the product and $t$ for time.
-As we now show in the paper (Section 2.2), $\beta_{ikt}$ also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper.
+As we now show in the paper (Section 2.2, pp.
+6-7), $\beta_{ikt}$ also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper.
 We directly report them here to ease the comparison with the referee's method afterwards.}
@@
 
-\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
+\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ are respectively the cif and fas prices observed in our dataset (conditional on a given year-transport mode).
 With two equations and two endogenous variables, the system can be solved.
@@
 In this respect, the referee's method and ours are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time, country, sector and transport mode.
-Put it differently, our method is consistent with Hummels' methodology - provided that it is adequately reported as such.
+In other words, our method is consistent with Hummels' methodology - provided that it is adequately reported as such.
 In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
-Section 4 ([Section 3] of the submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
+Section 4 ([Section 3] in the initially submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend.
 We agree that this was not sufficiently transparent in the submitted version.
@@
 As previously noted, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 1989 (though we could not obtain the whole data because of  COVID).
-We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix, Section D.3, Table D.1 and Figure D.3.\textbf{XXX Check XXX}\footnote{Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the referee's method versus ours.} For sake of reading clarity, we also report here the results of Table D.1 (here Table \ref{tab_comp_referee1_baseline}).
+We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix, Section D.3, Table D.1 and Figure D.3.\footnote{Note that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the referee's method versus ours.} For the referee's convenience, we also report here the results of Table D.1 (here Table \ref{tab_comp_referee1_baseline}).
 
@@
 
-Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode.
+Regarding the share of additive costs $\beta$, the comparison between the two methods delivers opposite conclusions depending on the transport mode.
 While the estimated $\beta$ is lower for vessel transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
-In any case, the share of additive costs remains substantial, around 40\% on average over the period.
+In any case, the share of additive costs remains substantial, around 40\% on average across transport modes over the period.
 In this respect, it confirms the main message of our paper shedding light on the importance of the additive component in international transport costs.
@@
 
-As noted above, the suggested method implies that the estimation is run at the sector-country level (on top of being mode-year specific), exploiting the variability within each origin country-3d sector pair across 10-d sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run.
-This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline} (also reported in the online appendix, Section D.3.
-\textbf{XXX Check XXX}.
+As noted above, the suggested method implies that the estimation is run at the sector-country level (on top of being mode-year specific), exploiting the variability within each origin country-3digit sector pair across 10-digit sub-sectors and ports of entry.
+Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports of entry given the number of fixed effects included in the regression, such that estimation cannot be run.
+This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline} (also reported in the online appendix, Table D.1 of Section D.3).
 Put it differently, this methodology discards country-sector pairs that export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
@@
 The most important limitation rather comes from the coverage period.
-Information about the port of entry are only available since 1989, which is already smaller than our full sample starting in 1974.
+Information about the port of entry is only available since 1989, which is already smaller than our full sample starting in 1974.
 On top of that, because of the Covid situation, the US Bureau of Census was only able to send us the data relative to the years 1997-1999 and 2001-2019.
 Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular).
-In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time.
+In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over a period including years when globalization was specifically intense, and the induced long-run welfare variations.
 Eliminating this dimension of the paper would be detrimental to its contribution.
@@
 
-If we take the value of the $\beta$ by itself, there is no clear criteria to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
-Further, as shown in Table \ref{tab_comp_referee1_baseline}, no clear sign of a potential upwards or downwards bias that would be attached to our method.
+If we take the value of the $\beta$ by itself, there is no clear criterion to discriminate between the value estimated with our method and the one obtained with the suggested method (when, of course, run on the same sample).
+Further, Table \ref{tab_comp_referee1_baseline} does not show any clear sign of a potential upwards or downwards bias that would be attached to our method.
 Things are more clear-cut in terms of accuracy of the estimation.
-Specifically, our method yields a more precise estimation of the $\beta$ than the referee's method.
+Specifically, our method yields a more accurate estimation of the $\beta$ than the referee's method.
 To develop on this, for each year and transport mode:
@@
 However, we also want to keep track of this alternative method as further robustness check.
-For this purpose, we have amended the online Appendix with a section devoted to the referee's method (Section D.3 \textbf{XXX Check XXX}).
+For this purpose, we have amended the online Appendix with a section devoted to the referee's method (Section D.3).
 
@@
 Following the referee's advice, we therefore decided to provide a full set of IV estimates to provide a clean assessment of the size of the potential bias.
-Section 4.3 in the revised version of the paper provides an overview of the results, while section B.4 in Appendix B provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
+Section 3.3.2 in the revised version of the paper provides an overview of the results, while section B.3 in Appendix B provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates.
 We follow earlier literature (see e.g.
-Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices.
-First-stage estimates reported in section B.4 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation.
-Figure 9 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart.
+Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties based on tariffs at the product line, together with one-year lagged fas prices.
+First-stage estimates reported in section B.3 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation.
+Figure 4 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterparts.
 In all cases, these estimates are very similar.
@@
 We agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variations in the data.
-After checking with the US Bureau of Census, HS-10 data is only available since 1989, now available to 2019.
-As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.We decided to maintain the case $s=3, k=5$ as baseline, and to run the estimation using 10-digit products in the new Sub-Section 3.3.3 \textbf{XXX CHECK XXX }devoted to robustness analysis.
+After checking with the US Bureau of Census, HS-10 data is only available from 1989 until 2019.
+As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.We decided to maintain the case $s=3, k=5$ as baseline, and to run the estimation using 10-digit products in the new Sub-Section 3.3.3.
 
@@
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
-We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper (see Section 3.3.3 \textbf{XXX CHECK XXX }).
+We thank the referee for pointing this ambiguity in our paper, which drove us to rewrite the associated paragraph in the revised version of the paper (see Section 3.3.3).
 
@@
 In this context, the findings in \cite{Lashkaripour_JIE2020} are specifically enlightening.
-Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories and that the cost of transportation increases more rapidly with unit weight than the cost of production.
+Based on US data very similar to ours over the period 1995-2015, the paper highlights, among other results, that the unit weight of imported goods is indeed substantially heterogeneous even within narrowly-defined product categories, and that the cost of transportation increases more rapidly with unit weight than the cost of production.
 \cite{Lashkaripour_JIE2020} finds that accounting for the heterogeneity in export unit weights provides evidence in favor of the ad-valorem cost assumption regarding transport costs.
 This result of transport costs close to be totally ad-valorem contrasts from our own findings of additive costs representing 30 to 45\% of total costs.
-There are explanations for this difference.
-Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020}) restricts to indivisible/discrete goods that represent a bit more 56\% of US total imports of goods in his dataset.
+There are several explanations for this difference.
+Beyond a longer time span, our database is exhaustive, and encompasses all goods and industries, while \citet{Lashkaripour_JIE2020} restricts to indivisible/discrete goods that represent a bit more 56\% of US total imports of goods in his dataset.
 In this regard, when comparing Tables 1 and 10 in \citet{Lashkaripour_JIE2020}, one can see that the share of multiplicative costs tends to decrease with the share of discrete goods.
@@
 Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) report the quantity by transport mode.
-This is incompatible with our empirical strategy of estimation which is conditional on the transport mode, similarly as in \cite{hummels2007}.
+This is incompatible with our empirical strategy of estimation which is conditional to the transport mode, as in \cite{hummels2007}.
 As a result, computing the unit price of all trade flows would require the assumption that the weight per unit is the same for vessel shipments and air shipments at the level of the observation.
@@
 This would reduce the total value of the sample from 12,500 billion dollars to 2,650 billion dollars, entailing a large loss of useful information.
-Original census data, when accessible, are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unleading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+Original census data, when accessible, are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unloading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
 This is still a large share, but not too large to preclude any work.
@@
 
-Quantity units are a second difficult.
+Quantity units are a second difficulty.
 The Census data remain silent on that matter.
 To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
-Unfortunately, these schedules change many time per year and only the schedules from 2009 onwards are available in a .csv format on the Census website.
+Unfortunately, these schedules change many times per year and only the schedules from 2009 onwards are available in a .csv format on the Census website.
 On the years for which the match could be done (from 2009), it turns out that at our sector level (3-digit), goods (either at the 5-digit or the 10-digit level) are measured in different units.
@@
 
-Getting deeper on the estimated functional form raised another difficulty in adapting our equation to per-unit rather than per-weight prices.
+We also faced another difficulty in adapting our equation to per-unit rather than per-weight prices.
 As discussed above, in our baseline regression we separate each transport cost component in two distinct sector-country dimensions.
@@
 \end{equation*}
-\noindent where $p_{iku}$ refers to the cif price associated with quantity of goof $k$ imported from country $i$, measured in unit $u$ (eg kg or number), $\widetilde{p}_{iku}$ the corresponding unit value (fas price).
+\noindent where $p_{iku}$ refers to the cif price associated with the quantity of good $k$ imported from country $i$, measured in unit $u$ (eg kg or number), $\widetilde{p}_{iku}$ being the corresponding unit value (fas price).
 
-The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} (as well as in the online appendix D.4).
-In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity, the second column reports the results obtained considering the price per kg on the same sample and the last column reported our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
+The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} below (as well as in the Online Appendix D.4).
+In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity, the second column reports the results obtained considering the price per kg on the same sample and the last column reports our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
 
@@
 First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
-In contrast to the aggregation and the separability issues, these restrictions that strongly limit both the number of sectors and of countries covered, do also impact the value of trade flows covered.
+In contrast to the aggregation and the separability issues, these restrictions strongly limit both the number of sectors and of countries covered, and do also impact the value of trade flows covered.
 In our view, this raises a non-trivial sample selection bias that reduces the general scope of the transport cost estimates.
@@
 \subsubsection{On the relevance of the per-weight price approach}
-At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price add some relevant information on the international transport cost burden faced by exporters.
+At a more fundamental level, we do believe that the study of transport costs measured on a per-weight price adds some relevant information on the international transport cost burden faced by exporters.
 Let us explain why.
@@
 
-Another argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png}.
+Another argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png} below.
 After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
@@
 After considering all the above arguments in depth, we decided not to switch the paper from the study of price per weight to the study of price per quantity.
-We thus maintained our original empirical strategy covering the whole period 1974-2019 in the revised version, while we extended the online Appendix with this supplementary robustness check.
+We thus maintained our original empirical strategy covering the whole period 1974-2019 in the revised version, while we extended the Online Appendix with this supplementary robustness check.
 We hope to have convinced you that this is a relevant choice.
@@
 In contrast, our own exercise relies on a several decades of US data.
-It allows us to compute the welfare effect of  taking intor account the relative dynamics of additive and multiplicative costs.
+It allows us to compute the welfare effect of  taking into account the relative dynamics of additive and multiplicative costs.
 Furthermore, our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the sum of prices and multiplicative costs.
@@
 In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
-To that end, we adjust the share of exporting firms in the US.
-Based on \citet{BEJK-AER-03, Lincoln_McCallum2018}.
-In our preferred exercise, we quantify the welfare gains that derive from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component in a context of fixed export cost reduction.
+To that end, we adjust the share of exporting firms in the US, based on \citet{BEJK-AER-03} and \citet{Lincoln_McCallum2018}.
+In our preferred exercise, we quantify the welfare gains deriving from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component together with a fixed export cost reduction.
 We compare this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
@@
 Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
-Specifically, we report the welfare effect of change in total transport costs decomposed in its two dimensions (additive and ad-valorem),both in absolute and in relative terms.
+Specifically, we report the welfare effect of changes in total transport costs decomposed in its two dimensions (additive and ad-valorem), both in absolute and in relative terms.
 We have two major insights.
@@
 
-Overall, these results add a substantial contribution to the paper : We are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period, and to highlight the respective part of each component in the determination of these gains.
+Overall, these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period covering a moment of unprecedented acceleration in international trade, and to highlight the respective part of each component in the determination of these gains.
 %Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.

### 2021-11-24 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
je ne comprends pas pourquoi tower me sort ces commit

@@
 Answer to Referee 2}
-
-\author{Guillaume \textsc{Daudin}\thanks{%
-Université Paris-Dauphine, PSL University, CNRS, 8007, IRD, 260, LEDa, DIAL, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+%
+%\author{Guillaume \textsc{Daudin}\thanks{%
+%Université Paris-Dauphine, PSL University, CNRS, 8007, IRD, 260, LEDa, DIAL, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
+%University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
 

### 2021-11-24 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
je ne comprends pas pourquoi tower me sort ces commit

@@
 By allowing for country-sector-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given sector/country partner) over time, by roughly 50\% to 60\%.
-Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in vessel transport), composition effects amplify the reduction of pure transport costs.
+Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs.
+When they matter, composition effects amplify the reduction of pure transport costs.
 This contrasts with the conclusion obtained by \cite{hummels2007} also on US import flows.
@@
 
-The main result from Figure \ref{fig:totalTC_compeffects_excl} is that composition effects do not play a major role in accounting for the time trend of transport costs.
-Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the raw transport costs measure and the pure transport costs.
-For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
-Air transport costs were reduced by around 60\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
-Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing).
+The main result from Figure \ref{fig:totalTC_compeffects_excl} is that composition effects do not play a major role in accounting for the time trend of transport costs; when they matter, they rather tend to amplify the decrease in pure transport costs.
+
+Going into a more detailed analysis by transport mode, we do not find much evidence of a substantial trend difference for air transport between the raw transport costs measure and the pure transport costs until the mid-2000s.
+Inspecting the panels (a) to (c) of Figure \ref{fig:totalTC_compeffects_excl}, for all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely until around 2005.
+Afterwards however, composition effects are the main driver of the reduction in the raw series, while pure transport costs are roughly constant, especially for the additive component.
+
+Composition effects do also play a role in vessel transport, yet limited in magnitude.
+For all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), they amplify the reduction in ``pure'' transport costs at the beginning of the period in the 1970s (the dotted line is below the plain line, and the gap is increasing).
 Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
-Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in pure transport costs (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
 This is particularly the case for the multiplicative component (panel (d)).\smallskip
@@
 Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``pure'' transport costs.
-In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction in international transport costs observed over time, in particular in air transport.
+In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction in international transport costs observed over time.
 Complementing the findings of Section \ref{sec:results_decomposition}, these results highlight the importance of integrating the additive dimension of international transport costs, with the aim here of characterizing their time trends.

### 2021-11-24 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
papier + réponse référé 1 + 2 + OA FINIS

@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-Université Paris-Dauphine, PSL University, CNRS, 8007, IRD, 260, LEDa, DIAL, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author.
-University Paris-Dauphine \& PSL University, LEDa, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+Université Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
+Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{
+University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
 
@@
 \item In the revised version, we have strengthened the robustness checks regarding the estimation of international transport costs (robustness checks to the level of aggregation and to endogeneity have been added).
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the results.
 Section 3 now includes the robustness checks as a final sub-section.
@@
  $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (see Table 2 and Figure 3 on pp.
-15-16).
+14-15).
 We check the robustness of our results by re-running the estimation without the separability assumption.
@@
  For this reason, we have decided to run the robustness check on a reduced sample.
-For each year, we select the largest exporters that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
- We keep in the sample all trade observations from these exporters and in these sectors.
+For each year, we identify \textit{(i)} the largest exporters that form together at least 80\% of annual trade and \textit{(ii)} the largest traded sectors that form together at least 80\% of annual trade.
+ We keep in the sample all trade observations that belong to both sets of exporters and sectors.
+
  This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2 in the paper), vs more than 30,000 on the complete sample (see Table 1 in the paper).
 It yet remains quite large in terms of trade coverage (mean of 68\%).\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the baseline sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the baseline sample).
-The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} On this reduced sample, we re-run the estimation with and without the separability assumption.
-The results are reported in Section 3.3.1 (Table 2 and Figure 3 on pp.
-15-16).
-As we conclude at the end of this section, the trend pattern of the share of additive costs are very similar whether estimated under the separability assumption or not.
+The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} As we conclude at the end of Section 3.3.1, the trend pattern of the share of additive costs are very similar whether estimated under the separability assumption or not.
  % use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
@@
 
-\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels's one.
+\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels' one.
 We took this remark into account by adding a new subsection in the revised version (Section 2.2, paragraph ``Estimation strategy'').
@@
 We develop each line of argument in our answer.
-Yet, because the referee's estimation method is also an interesting alternative, we have added a new section in the Online appendix devoted to the latter (Section D.3).
+Yet, because the referee's estimation method is also an interesting alternative, we have added a new section in the Online Appendix devoted to the latter (Section D.3).
 
-\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares.
+\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods.
 However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators.
-This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e.
+This is due to the necessity of imposing an \textit{ex-ante} restrictions on parameters, i.e.
 $\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$.
@@
 In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive).
-Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 11 page 8.\smallskip
+Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version.
+We hope that the new explanation provided in Section 2.2 (specifically pp.
+7-8) now fulfills this requirement.
+\smallskip
 
@@
 By year and transport mode, this implies running the estimation for each country of origin $i$ and each $s(k)$ = 3-digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$).
-Despite its interest, the referee's method has some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps.
+Despite its interest, the referee's method has some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$}).
+Our answer can be articulated in two steps.
 First, we debate the advantages that the referee's estimation method would bring in comparison with ours.
@@
 We thank the referee for this valuable suggestion.
-It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version (see section B.3 in the paper's appendix).
+It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version (see sections 3.3.2 and  B.3 in the paper).
 
@@
 
-     Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method.
+     Equation (\ref{eq:Hummels}) lies at the root of Hummels' (2007) method.
 In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a mode-specific $\beta$ invariant over time/sector/origin country.
@@
 \noindent from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis).
-Then, relying on the Hummels's equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
+One can then deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.
 
@@
 In other words, our method is consistent with Hummels' methodology - provided that it is adequately reported as such.
+We agree that this was not the case in the submitted version and we hope that this revised version is now convincing enough on that side.
+
 In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector.
@@
 
-As previously noted, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 1989 (though we could not obtain the whole data because of  COVID).
+As previously noted, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 1989 (though we could not obtain the whole data because of Covid).
 We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix, Section D.3, Table D.1 and Figure D.3.\footnote{Note that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$).
-We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the referee's method versus ours.} For the referee's convenience, we also report here the results of Table D.1 (here Table \ref{tab_comp_referee1_baseline}).
+We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the referee's method versus ours.} For the referee's convenience, we also report the results of Table D.1 here (Table \ref{tab_comp_referee1_baseline}).
 
@@
 Regarding the share of additive costs $\beta$, the comparison between the two methods delivers opposite conclusions depending on the transport mode.
-While the estimated $\beta$ is lower for vessel transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport.
+While the estimated $\beta$ is lower for vessel transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for air transport.
 In any case, the share of additive costs remains substantial, around 40\% on average across transport modes over the period.
@@
 As noted above, the suggested method implies that the estimation is run at the sector-country level (on top of being mode-year specific), exploiting the variability within each origin country-3digit sector pair across 10-digit sub-sectors and ports of entry.
-Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports of entry given the number of fixed effects included in the regression, such that estimation cannot be run.
-This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline} (also reported in the online appendix, Table D.1 of Section D.3).
-Put it differently, this methodology discards country-sector pairs that export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
+Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports of entry given the number of fixed effects included in the regression to estimate the equation.
+This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in the online appendix, Table D.1 of Section D.3 (and here in Table \ref{tab_comp_referee1_baseline}).
+Put differently, this methodology discards country-sector pairs that export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry.
 This may induce a selection bias in the sample covered for estimation.
@@
 Specifically, our method yields a more accurate estimation of the $\beta$ than the referee's method.
-To develop on this, for each year and transport mode:
+We reach this conclusion by implementing the following reasoning (for each year and transport mode):
     \begin{itemize}
-    \item With the referee's method, we estimate one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{ref}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
-From this, we can compute the size of the 5-95\% confidence interval $CI^{ref}_{is} = \hat{\beta}_{is}^{max,ref} - \hat{\beta}_{is}^{min,ref}$ where:
-        $$\hat{\beta}_{is}^{min,ref} = \hat{\beta}^{ref}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,ref} = \hat{\beta}^{ref}_{is(k)} + 1.96 SD_{is}$$
-
+    \item With the referee's method (referred to as the ``direct method'' in the Online Appendix), we estimate one value for the share of additive component $\beta$ at the $i,s$ level denoted $\hat{\beta}^{dir}_{is(k)}$ associated with a standard deviation $SD_{is}$, by year and transport mode.
+From this, we can compute the size of the 5-95\% confidence interval $CI^{dir}_{is} = \hat{\beta}_{is}^{max,dir} - \hat{\beta}_{is}^{min,dir}$ where:
+        $$\hat{\beta}_{is}^{min,dir} = \hat{\beta}^{dir}_{is(k)} - 1.96 SD_{is},\quad \hat{\beta}_{is}^{max,dir} = \hat{\beta}^{dir}_{is(k)} + 1.96 SD_{is}$$
+\noindent where the $^{dir}$ subscript identifies the direct method.
         \item  Our method provides estimates of the underlying trade costs components ($\widehat{\tau}_i$, $\widehat{\tau}_s$, $\widehat{t}_i$, $\widehat{t}_s$) with an associated matrix of variance-covariance, from which we can rebuild $\beta_{is}$ (on a year-transport mode basis).
 As such, it does not yield an estimate of $\beta_{is}$ and an associated standard deviation, so that we cannot directly compare the precision of the estimation.
-We compute the accuracy of the $\beta$ estimation through a bootstrap method.
-Specifically, on a yearly/mode basis we draw a distribution of trade costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.
-Writing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries / sectors at first.
-}
+We hence compute the accuracy of the $\beta$ estimation through a bootstrap method.
+Specifically, on a yearly/mode basis we draw a distribution of transport costs components and associated $\beta_{is}$ (10,000 random draws) from which we can compute the mean, the median and the 5-95 confidence interval for each couple $i,s$.\footnote{Notice that this should be made on the same sample as the one obtained with the referee's method, since the goal is to compare the accuracy of the $\beta_{is}$ estimate - implying to have the same sample of countries-sectors at first.
+} Writing $\beta_{is}^{95}$ and $\beta_{is}^{05}$ the associated thresholds, we then obtain the size of the confidence interval $CI_{is}= \beta_{is}^{95}- \beta_{is}^{05}$.
 
@@
 Given that the estimation method is run at the country-sector level (exploiting heterogeneity \textit{within} a given country-sector pair), it implicitly assumes that transport costs are independent \textit{across} sectors and \textit{across} origin countries.
-In plain words, it assumes that transport costs of say, cars, have nothing in common whether those goods comes from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
+In plain words, it assumes that transport costs of say, cars, have nothing in common whether those goods come from France or from Germany; or that transport costs that apply to imported goods from France have no common component across sectors.
 This is a disputable assumption.
@@
 That said, the possibility remains that bigger firms may impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
+In presence of pricing-to-market behavior of firms (see Krugman, 1987), the export price set by the firm ($\widetilde{p}_{ik}$) may also be partly endogenous to the size of transport costs if for instance, the exporting firm absorbs (part of) the transport costs by reducing the fas price.
 
@@
 We follow earlier literature (see e.g.
-Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties based on tariffs at the product line, together with one-year lagged fas prices.
-First-stage estimates reported in section B.3 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation.
+Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing the fas price $\widetilde{p}$ on custom duties based on tariffs at the product line, together with one-year lagged fas prices.
+First-stage estimates reported in section B.3 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of our estimated equation (Equation (9) in the paper) to produce a 2SLS-type of estimation.
 Figure 4 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterparts.
@@
 After checking with the US Bureau of Census, HS-10 data is only available from 1989 until 2019.
-As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.We decided to maintain the case $s=3, k=5$ as baseline, and to run the estimation using 10-digit products in the new Sub-Section 3.3.3.
+As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards.
+As previously mentioned, we view as important to preserve the historical coverage of our analysis.
+Accordingly, we decided to maintain the case $s=3, k=5$ as baseline, and to run the estimation using 10-digit products as robustness check in the new Section 3.3.3.
 
-Additionally, the computational burden mentioned by the referee is not attributable to the product classification level $k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
+Additionally, the computational burden mentioned by the referee is not attributable to the product classification level ($k=5$ or 10) but rather to the degree of sectoral classification ($s=3$ or 4) as it conditions the number of fixed effects.
 This explains why we consider the $s=4$ digit- sectoral classification level only for some years.
@@
 The first difficulty is availability of information on quantities.
-Neither the US Bureau of Census data nor the 1974-2004 Hummels's data (based on the Census Bureau data) report the quantity by transport mode.
+Neither the US Bureau of Census data nor the 1974-2004 Hummels' data (based on the Census Bureau data) report the quantity by transport mode.
 This is incompatible with our empirical strategy of estimation which is conditional to the transport mode, as in \cite{hummels2007}.
@@
 The alternative is to consider only single-mode flows.
-In Hummels's data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
+In Hummels' data (1974-2014), on average 41.1 \% of trade flows and 79\% of trade value are multi-mode.
 This would reduce the total value of the sample from 12,500 billion dollars to 2,650 billion dollars, entailing a large loss of useful information.
-Original census data, when accessible, are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unloading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
+Original Census data, when accessible, are more promising.\footnote{Original Census data are  reported at the HS-10 digit $\times$ district of entry $\times$ district of unloading $\times$ rate provision code.} In 2019, multi-modal flows are 570 billion dollars out of 1,790 billion dollars, or 31\% of the data.
 This is still a large share, but not too large to preclude any work.
@@
 The Census data remain silent on that matter.
-To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United State''.
+To the best of our knowledge, these can only be found in the ``Harmonized Tariff Schedule of the United States''.
 Unfortunately, these schedules change many times per year and only the schedules from 2009 onwards are available in a .csv format on the Census website.
@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could be easily replaced by (sector, quantity unit) fixed effects; but the same does not appear possible for country fixed effects.
+Sector-level fixed effects could be easily replaced by (\textbf{sector, quantity unit} ou \textbf{sector $\times$ quantity unit}) fixed effects; but the same does not appear possible for country fixed effects.
 Let us elaborate on this issue.
+Let us develop the reasoning in more details.
 
@@
 Under the separability assumption, importing a tee-shirt from either country implies paying the cost given by the country fixed effect (the same for all sectors) plus the cost given by
-the sector fixed effect ; and similarly for importing a car.
-Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee- shirt from Germany relative to China.
+the sector fixed effect; and similarly for importing a car.
+Hence, the difference in the cost of importing a car from Germany versus China would be the same, in dollars, as the difference between the cost of importing a tee-shirt from Germany relative to China.
 This does not seem a sensible assumption.
@@
 \end{equation*}
-\noindent where $p_{iku}$ refers to the cif price associated with the quantity of good $k$ imported from country $i$, measured in unit $u$ (eg kg or number), $\widetilde{p}_{iku}$ being the corresponding unit value (fas price).
+\noindent where $p_{iku}$ refers to the cif price associated with the quantity of good $k$ imported from country $i$, measured in unit $u$ (e.g.
+kg or number), $\widetilde{p}_{iku}$ being the corresponding unit value (fas price).
 
-The results are reported in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} below (as well as in the Online Appendix D.4).
-In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity, the second column reports the results obtained considering the price per kg on the same sample and the last column reports our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
+The results are reported in the Online Appendix D.4, as well as here in Table \ref{tab_comp_wgt_qty} and in Figure \ref{graph_comp_wgt_qty} for reading convenience.
+In Table \ref{tab_comp_wgt_qty}, for each transport mode, the first column reports the estimation based on the price per quantity, the second column reports the results obtained considering the price per kg on the same (limited) sample and the last column reports our baseline results on the large sample (considering the average value over 2005-2019, under the separability assumption).
 
@@
 Two main results emerge.
-First, given the constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
+First, given the previously underlined constraints on the sample selection for quantities, the estimation based on quantity can only be run on a reduced sample.
 In contrast to the aggregation and the separability issues, these restrictions strongly limit both the number of sectors and of countries covered, and do also impact the value of trade flows covered.
 In our view, this raises a non-trivial sample selection bias that reduces the general scope of the transport cost estimates.
-Second, in terms of $\beta$ estimates, the results show clearly that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
+Second, in terms of $\beta$ estimates, the results clearly show that using price per unit (whatever that unit is) instead of price per weight decreases the share of additive costs (except if one look at the median for air).
 In this regard, they stand in line with the results of \cite{Lashkaripour_JIE2020}, and confirm the correlation between weight per unit and unit price.
@@
 
-Our main objective is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mixed, per-weight and per-volume basis, both historically and nowadays.
+Our main objective is to provide a quantitative assessment of the international transport costs burden over time; on that matter, one may start with the observation that transport cost schedules are expressed through a mixed per-weight and per-volume basis, both historically and nowadays.
 For example, in 2021, air freight trade costs are expressed in weights, with a minimal weight per volume (``volumetric weight'' or ``dimensional weight'')\footnote{See for example \href{https://transporteca.co.uk/shipping-costs}{https://transporteca.co.uk/shipping-costs}, \href{http://wap.dhl.com/serv/volweight.html}{http://wap.dhl.com/serv/volweight.html} and \href{https://en.wikipedia.org/wiki/Dimensional_weight}{ https://en.wikipedia.org/wiki/Dimensional\_weight}).} while maritime freight costs are expressed in volume.
 Hence, looking at per-kg transport cost can be interpreted straightforwardly as trying to estimate the actual existing shipping costs, with sector fixed effects interpretable as measure of mean density for different sectors.
-The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions - in contrast, per-unit shipping cost depends on the behavior of exporting firms.
+The existence of per-volume or per-weight shipping cost is directly linked to the technical constraints of the transportation sector, hence exogenous to exporting firms' decisions.
+In contrast, per-unit shipping cost depends on the behavior of exporting firms.
 Put differently, the shipping cost per weight is one of the exogenous parameters firms take into account in their production choices, including weight per unit.
@@
 
-Another argument is illustrated by Figure \ref{graph_Share_of_discrete_goods.png} below.
+In a related manner, restricting the sample to denumerable goods would defeat our purpose of estimating international transport costs for the whole range of US imports.
+This argument is illustrated by Figure D.6 in the online appendix (replicated here with Figure \ref{graph_Share_of_discrete_goods.png}).
 After computing the share of discrete goods per 5-digit SITC import in 2009, we have used the changing composition of US imports to approximate the share of discrete goods in the total value of US imports on a yearly basis.
@@
 To do so, we start from the canonical \cites{melitz} model where we add additive trade costs.
-The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
-However, \citet{sorensen2014} dose not perform an empirically based quantitative exercise.
-In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the presence of additive costs, but the latter is based on a calibration for transport costs restricted to 2004, based on the case of Norway.
+The inclusion of additive costs in a New Trade model has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
+However, \citet{sorensen2014} exclusively provides a theoretical analysis, without any quantitative exercise.
+In a version of the \citet{chaney2008} model amended for additive costs, \citet{Irrazabal_2015} perform a quantitative simulation to assess the welfare variations induced by the increase in such additive costs (vs ad-valorem costs).
+For this, they rely on a calibration for trade costs based on their single-year estimation (for the year 2004), from which they depart by assuming an ad-hoc assumption (specifically, a 5\% increase in the ratio of additive cost to the median fas price).
 In contrast, our own exercise relies on a several decades of US data.
-It allows us to compute the welfare effect of  taking into account the relative dynamics of additive and multiplicative costs.
-Furthermore, our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the sum of prices and multiplicative costs.
+It allows us to assess the welfare changes induced over time by the relative dynamics of additive and multiplicative costs, based on the values actually observed in the data over the period 1974-2019 in all their dimensions (additive and ad-valorem).\footnote{On a more positive side, our methodology allows for the identification of both multiplicative and additive costs as a share of total costs, whereas additive costs in \citet{Irrazabal_2015} are expressed relatively to the product of prices and multiplicative costs.}
 
-More precisely, we use the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, to implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
-In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
+In the new Section 5, we use our estimation results over 1974-2019 (Section 3 of the paper) to implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+In addition, we also assess how the latter results are affected by changes in sunk costs of exports $f_{x}$.
 To that end, we adjust the share of exporting firms in the US, based on \citet{BEJK-AER-03} and \citet{Lincoln_McCallum2018}.
-In our preferred exercise, we quantify the welfare gains deriving from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component together with a fixed export cost reduction.
-We compare this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
+In our preferred exercise, we quantify the welfare gains deriving from the joined reduction in the fixed export cost $f_x$ and in the variable transport costs as documented in Section 3 (where the two additive/ad-valorem components decrease).
+We contrast this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
 
-Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
-Specifically, we report the welfare effect of changes in total transport costs decomposed in its two dimensions (additive and ad-valorem), both in absolute and in relative terms.
-We have two major insights.
-First, using the reduction of transport costs in vessel (air),  welfare gains are around 50\% (14\%) higher when this reduction is achieved through the estimated reduction in additive and ad-valorem costs.
-Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing variable costs, with again an additional premium coming from the decrease of additive costs.
+Table 6 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
+Specifically, we report the welfare effect of a change in variable transport costs, under both scenarios of a constant and fixed export cost.
+In our preferred scenario, the reduction in variable transport costs in Air transport generates a 40\% higher welfare gain when such a reduction partly
+occurs through the additive component, rather than when it happens through a decrease in ad-valorem costs only. In vessel transport, the
+extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
 
-Overall, these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period covering a moment of unprecedented acceleration in international trade, and to highlight the respective part of each component in the determination of these gains.
+Overall, these results add a substantial contribution to the paper: we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period covering a moment of unprecedented acceleration in international trade, and to highlight the respective part of each component in the determination of these gains.
 %Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.

### 2021-11-24 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
papier + réponse référé 1 + 2 + OA FINIS

@@
 \author{Guillaume \textsc{Daudin}\thanks{%
-University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805- UMR IRD 260, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
+University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
 Univ.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Bricoles sur les effets de composition
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 
-Going into a more detailed analysis by transport mode, we do not find much evidence of a substantial trend difference for air transport between the raw transport costs measure and the pure transport costs until the mid-2000s.
-Inspecting the panels (a) to (c) of Figure \ref{fig:totalTC_compeffects_excl}, for all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely until around 2005.
+Going into a more detailed analysis by transport mode, we do not find much evidence of a substantial trend difference for air transport between the raw transport costs measure and the pure transport costs until the mid-2000s: Inspecting the panels (a) to (c) of Figure \ref{fig:totalTC_compeffects_excl}, for all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line stay close together until around 2005.
 Afterwards however, composition effects are the main driver of the reduction in the raw series, while pure transport costs are roughly constant, especially for the additive component.
 
-Composition effects do also play a role in vessel transport, yet limited in magnitude.
-For all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), they amplify the reduction in ``pure'' transport costs at the beginning of the period in the 1970s (the dotted line is below the plain line, and the gap is increasing).
+Composition effects do also play a role in vessel transport.
+It is limited in magnitude.
+For all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), they amplify the reduction in ``pure'' transport costs in the 1970s (the dotted line is below the plain line, and the gap is increasing).
 Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
-Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in pure transport costs (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
+Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in pure transport costs (dotted line), and a 10\% reduction that comes from composition effects (the difference between the two line).
 This is particularly the case for the multiplicative component (panel (d)).\smallskip

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Bricoles sur les effets de composition
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 
-In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows.
-The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary sector.
+In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary and manufactured goods trade flows.
+The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) is reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary sector.
 To ease the comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.
@@
 3)" in http://unctadstat.unctad.org/EN/Classifications.html, accessed Septembre 2018}.
-Are considered as ``primary goods'' all flows recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).
+``Primary goods'' flows are recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).
+All the other flows are part of ``Manufacturing''.
 
@@
 
-As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``ceteris paribus'' transport costs and the raw measure have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)).
-However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
+As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``pure'' transport costs and the raw measure have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)).
+However the role of trade composition effects in accounting for this trend depends on the sector.\\
 In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).
-One can further notice that in vessel transport, composition effects play virtually no role at the end of the period, as most of the decrease can be imputed to the reduction of ``ceteris paribus'' transport costs (the blue continuous line)
+One can further notice that in vessel transport, composition effects play virtually no role at the end of the period, as most of the decrease can be imputed to the reduction of ``pure'' transport costs (the blue continuous line)
 
@@
 Afterwards, they conversely more than compensate the increase in pure transport costs (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)).
-In vessel transport, the composition effects play a limited role in the primary sector as for the whole range of goods matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)), by slightly amplifying the reduction in the pure transport costs in both cases.
+In vessel transport, the composition effects play a limited role in the primary sector, similarly to their role for the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)), by slightly amplifying the reduction in the pure transport costs.
 
@@
 On the following sub-period though, their share has fallen to 20-30\%.
-Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured sector, it is not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the manufacturing sector.
+Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured goods, it is not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the manufacturing.
 

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
J’enlève l’espoir d’un 2e graph
J’enlève l’espoir d’un 2e graph pour la comparaison des betas

Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 
-	\textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a smaller interval confidence of the baseline $\beta$ estimation.
+In most cases, the log of the ratio is negative, implying a smaller interval confidence of the baseline $\beta$ estimation.
 Our method thus yields more accurate estimations of the share of additive components, whatever the transport mode considered.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Encore un commit que m'impose Tower

@@
 
-   \textbf{XXXX AJouter le complement sur Air et stop XXX} In most cases, the log of the ratio is negative, implying a smaller interval confidence of the baseline $\beta$ estimation.
+In most cases, the log of the ratio is negative, implying a smaller interval confidence of the baseline $\beta$ estimation.
 Our method thus yields more accurate estimations of the share of additive components, whatever the transport mode considered.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
Encore un commit que m'impose Tower

--- 2021-11-23 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
@@
 	\item In the revised version, we have substantially strengthened the robustness checks regarding the estimation of international transport costs, with the addition of detailed checks to aggregation and endogeneity.
-Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results.
+Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the results.
 Section 3 now includes the robustness checks as a final sub-section.
@@
 product value/weight ratio, distance to the exporter, etc.).}
-\smallskip
+\medskip
 
@@
 
-All in all, your suggestion and the other referee's one pointed toward two possible options for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying of our results, or focusing on the welfare implications.
+All in all, your suggestion and the other referee's one pointed toward two possible options for the ``Big Picture'' implications of our exercise: either exploring the compositional country-product dynamics underlying our results, or focusing on the welfare implications.
 Both paths were undeniably worth exploring, but could not be followed in a single paper.
@@
 Firstly, even if ``\emph{the fact that transport costs are both specific and ad-valorem is well known}'', some recent literature still debates the practical importance of additive costs - for example, \cite{Lashkaripour_JIE2020} finds evidence supporting the ad-valorem assumption for denumerable goods, and consistently finds a very small welfare effect of additive costs within a partial-equilibrium model with perfect competition.
-Secondly, the implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, \cite{Hornok-et-al-RES-2015, Hornok-et-al-JIE-2015}, \cite{Alessandria-et-al-AER-2010} or \citet{sorensen2014}.
+Secondly, the implications of additive costs have not been fully exhausted in \citet{Irrazabal_2015}, \cite{Hornok-et-al-JIE-2015,Hornok-et-al-RES-2015}, \cite{Alessandria-et-al-AER-2010} or \citet{sorensen2014}.
 \cite{Kropf-Saure-JIE-2016} estimate the size and shape of per-shipment costs based on Swiss data.
@@
 \cite{Hornok-et-al-JIE-2015} focus on administrative costs incurring with every shipment (i.e., additive by nature), and draw welfare conclusions regarding the variations of these costs.
-We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transports costs.
+We complement these various outcome by providing a long-run characterization of both additive and multiplicative component of transport costs.
 Based on these empirical results, we investigate the welfare consequences of the reduction in international transport costs observed over 1974-2019.
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the disproportionate part of additive transport costs in determining those welfare variations.\smallskip
+In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the key role of additive transport costs in determining those welfare variations.\smallskip
 
 To this aim, we introduce additive costs in the canonical \citet{melitz} model.
-The inclusion of additive costs in a \citet{melitz} setting has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
-However, \citet{sorensen2014} exclusively performs a theoretical analysis, without any quantitative exercise.
-In addition to a partial equilibrium extension of \citet{melitz}, \citet{Irrazabal_2015} do perform a quantitative simulation to assess the welfare variations induced by the reduction of additive costs (vs ad-valorem costs).
-For this, they rely on a calibration for transport costs based on a single-year estimation (for the year 2004), from which they depart by assuming an ad-hoc hypothetical reduction of trade costs.
-In contrast, our own exercise relies on a several decades of US data, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on the values actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).\smallskip
+The inclusion of additive costs in a New Trade model has already been performed in \citet{sorensen2014} and \citet{Irrazabal_2015}.
+However, \citet{sorensen2014} exclusively provides a theoretical analysis, without any quantitative exercise.
+In a version of the \citet{chaney2008} model amended for additive costs, \citet{Irrazabal_2015} perform a quantitative simulation to assess the welfare variations induced by the increase in such additive costs (vs ad-valorem costs).
+For this, they rely on a calibration for trade costs based on their single-year estimation (for the year 2004), from which they depart by assuming an ad-hoc assumption (specifically, a 5\% increase in the ratio of additive cost to the median fas price).
+In contrast, our own exercise relies on a several decades of US data, allowing us to highlight the welfare alterations induced over time by the relative dynamics of additive and multiplicative costs, based on the values actually observed in the data over the period 1974-2019 in all their dimensions (additive and ad-valorem).\smallskip
 
-More precisely, we use the estimates underlying results reported in Section 3 concerning multiplicative and additive costs, to implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
-In addition, we also assess how the latter results are distorted by changes in sunk costs of exports, $f_{x}$.
+In the new Section 5, we use our estimation results over 1974-2019 (Section 3 of the paper) to implement several comparative statics exercises to investigate the different welfare consequences of alterations to multiplicative and additive costs.
+In addition, we also assess how the latter results are modified by changes in sunk costs of exports $f_{x}$.
 To that end, we adjust the share of exporting firms in the US based on \citet{BEJK-AER-03} and \citet{Lincoln_McCallum2018}.
-In our preferred exercise, we quantify the welfare gains deriving from the reduction in variable transport costs as documented in Section 3, relying on a combined reduction in each additive/ad-valorem component in a context of fixed export cost reduction.
-We compare setting this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
+In our preferred exercise, we quantify the welfare gains deriving from the joined reduction in the fixed export cost $f_x$ and in the variable transport costs as documented in Section 3 (where the two additive/ad-valorem components decrease).
+We contrast this with the case where the reduction of total transport costs  is solely attributed to a decrease in ad-valorem costs.
 
-Table 5 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
-Specifically, we report the welfare effect of change in total transport costs decomposed in its two dimensions (additive and ad-valorem), both in absolute and in relative terms.
-We have two major insights.
-First, using the reduction of transport costs in vessel (air),  welfare gains are around 50\% (14\%) higher when this reduction is achieved through the estimated reduction in additive and ad-valorem costs.
-Second, the decrease in export sunk costs (i.e., increase in the share of exporting firms) proportionally amplifies the gains from decreasing variable costs, with again an additional premium coming from the decrease of additive costs.
+Table 6 in Section 5 of the paper reports the results of these various comparative statics exercises for Air and Vessel transport modes.
+Specifically, we report the welfare effect of a change in variable transport costs, under both scenarios of a constant and fixed export cost.
+In our preferred scenario, the reduction in variable transport costs in Air transport generates a 40\% higher welfare gain when such a reduction partly
+occurs through the additive component, rather than when it happens through a decrease in ad-valorem costs only. In vessel transport, the
+extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%.
 
-Overall, these results add a substantial contribution to the paper : we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types (additive and multiplicative) of costs over a 45-year period covering the hyper-globalization moment, and to highlight the respective part of each component in the determination of these gains.
+Overall, these results add a substantial contribution to the paper: we are able to provide a quantitative assessment of the welfare gains induced by the decrease in both types of costs (additive and multiplicative) over a 45-year period covering the hyper-globalization moment, and to highlight the respective part of each component in the determination of these gains.
 %Note also that not only the inclusion of additive costs in the underlying framework generates large welfare differences, but also that the latter are probably a lower bound of the welfare variations induced by changes in additive \emph{trade} costs, larger than the sole transport costs.
@@
 for. Estimating Equation (\ref{eq:ref2}) should be relatively easy, for one it is linear, but it would also not require the
-inclusion of any country-sector fixed effects.}
+inclusion of any country-sector fixed effects [...] which can be easily estimated [...] using the authors' notations [as]:}
 
@@
 
-\textbf{Our answer:} As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods, such as Non-Linear Least Squares.
+\textbf{Our answer:} As highlighted by the referee, our estimated equation imposes to use non-linear estimation methods.
 We agree that resorting on a linear form as the one suggested by the referee would substantially alleviate the computational burden.
@@
 
-First, it can be noticed that the equation suggested by the referee does not stand in accordance with the assumption we borrow from \cite{Irrazabal_2015}, that ad-valorem costs is decomposed in its two country/sector dimensions multiplicatively (i.e.
+First, it can be noticed that the equation suggested by the referee does not stand in accordance with the assumption we borrow from \cite{Irrazabal_2015}, that the ad-valorem cost is decomposed in its two country/sector dimensions multiplicatively (i.e.
 approximating $\tau_{is(k)}$ by the product $\tau_i\times \tau_{s(k)}$).
@@
 This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
-$\tau \geq 1$ and $t\geq 0$, the latter meaning simply we constrain both types of transport costs to be non-negative, as by construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
+$\tau \geq 1$ and $t\geq 0$, or $0 \leq \beta \leq 1$.
 Without these restrictions, standard linear, least squares estimates deliver aberrant values with mediocre quality-of-fit.
@@
 
-This specification obviously requires a non-linear estimation.
+Taking into account the various positivity constraints finally leads us to estimate:
+\begin{equation}
+\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
+\end{equation}
+\noindent which corresponds to Equation (9) in the paper, and where we ensure the positivity constraint on $\tau_x\geq 1$, $t_x \geq 0$ (with  $x = i, s$) by taking the logs of the parameters.
+This specification requires a non-linear estimation method.
+We agree that this point was not clarified enough in the submitted version and we thank the referee for pointing this out.
 We did our best to make the point clearer in the revised version (see section 2.2, more specifically p.
-8).
+7-8).
+
 We also devote more time to the robustness of the empirical specification.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex
Encore un commit que m'impose Tower

@@
 We ordered this new data from the Census Bureau at the end of December 2019.
-However, some of the years we requested (2000 and 2001, as well as years before 1998) are not on their servers: these data are back at their office which they have not been able to access since last spring apparently, due to the current health situation.
+However, some of the years we requested (2000 and 2001, as well as years before 1998) are not on their servers: these data are back at their office which they have not been able to access since Spring 2020 apparently, due to the ongoing health situation.
 Since then, the situation has never evolved, and we are still currently waiting for these missing years.
@@
 Until very recently, Mr.
-Sanchez did not gave us any clear “yes or no” answer to our requests concerning a possible return in his DC office (closed because of the CoViD 19 restrictions), which would have allowed the delivery of the data.
+Sanchez did not gave us any clear ``yes or no'' answer to our requests concerning a possible return in his DC office (closed because of the CoViD 19 restrictions), which would have allowed the delivery of the data.
 In his last e-mail, he confirmed that access to the additional data was impossible.\medskip
 
-Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this letter) based on a HS10 dataset over the 2005-2019 period.
-These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.
+Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this letter) based on a HS10 dataset over the 2005-2019 period.\footnote{We uncovered some mismatches between the original Hummels' dataset (whose finest degree of classification is at the 5-digit level, and which we use until 2004) and the Census data with detailed information at the HS10 level for the years 1997-1999 and 2001-2004 which we could access to.
+Since 2005 onwards, we use original Census data with perfect match between the 5- and 10-classification degrees.
+Accordingly, we run our robustness check to aggregation over this period 2005-2019 (Section 3.3.3).} These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.
 \medskip
@@
 The revised version devotes much more space to explicit tests of robustness of the empirical specification.
-On top of the robustness to the separability assumption (already in the initially submitted version), we also assess the robustness to potential endogeneity and the level of aggregation (Sections 3.3.2 and 3.3.3) - see the list of modifications in section 2 from this letter.
+On top of the robustness to the separability assumption (already in the initially submitted version), we also assess the robustness to potential endogeneity and to the level of aggregation (Sections 3.3.2 and 3.3.3) - see the list of modifications in section 2 from this letter.
 \medskip
@@
 
-\item \emph{Endogeneity issues}: Section 3.3.2 and Appendix B.3 provides detailed developments addressing concerns expressed by Referee 1 concerning potential endogeneity in our estimates.
+\item \emph{Endogeneity issues}: Section 3.3.2 and Appendix B.3 provide detailed developments addressing concerns expressed by Referee 1 concerning potential endogeneity in our estimates.
 Relying on instrumental variables suggested by Referee 1, we provide first- and second-stage estimates, based on both the 5-digit-product level and the more disaggregated HS10 level for sensitivity checks.
@@
 These results are now reported in section 3.3.3 in the revised version.
-In particular, Table 3 and Figure 5 show the results are similar when using 10-digit product sand 5-digit products.
+In particular, Table 3 and Figure 5 show the results are similar when using 10-digit products sand 5-digit products.
 
-\item \emph{Calculation of unit prices}: Referee 1 is concerned by the fact that we calculate unit price as Value/Weight, instead of Value/Quantity.
-We explain in details in our answer and the Online Appendix why we decided against a switch from per-kg price to per-unit prices.
-First, there are data availability issues.
+\item \emph{Calculation of unit prices}: Referee 1 is concerned by the fact that we calculate unit price as Value/Weight, rather than Value/Quantity.
+We extensively explain why we decided against a switch from per-kg price to per-unit prices in our answer to the referee and the Online Appendix (which has been extended with a section on this issue).
+To sum up in a few words, first, there are data availability issues.
 Second, the estimation is even more computationally intensive.
-Third, we believe that the study of per-kg prices adds relevant information to the study of international transport costs, both because actual shipping costs are expressed in a per-weight/volume basis and because it allows us to integrate a larger share of trade in our analysis.
+Third, we believe that the study of per-kg prices adds relevant information to the study of international transport costs, both because actual shipping costs are expressed in a per-weight/volume basis and because it allows us to integrate a larger share of trade flows in our analysis.
 
@@
 The robustness exercises included in the main text focus on the elements that were the most problematic to the referees.
-The paper's relevance to big picture debates is more obvious.
+The paper's relevance to big picture debates is now more straightforward.
 We hope that you will be pleased by the outcome, and we sincerely apologize again for the delay.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex
Travail collectif
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could be easily replaced by (\textbf{sector, quantity unit} ou \textbf{sector $\times$ quantity unit}) fixed effects; but the same does not appear possible for country fixed effects.
+Sector-level fixed effects could be easily replaced by (sector $\times$ quantity unit) fixed effects; but the same does not appear possible for country fixed effects.
 Let us elaborate on this issue.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex
Travail collectif
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 
-Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this letter) based on a HS10 dataset over the 2005-2019 period.\footnote{We uncovered some mismatches between the original Hummels' dataset (whose finest degree of classification is at the 5-digit level, and which we use until 2004) and the Census data with detailed information at the HS10 level for the years 1997-1999 and 2001-2004 which we could access to.
+Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this letter) based on a HS10 dataset over the 2005-2019 period.
+%\footnote{We uncovered some mismatches between the original Hummels' dataset (whose finest degree of classification is at the 5-digit level, and which we use until 2004) and the Census data with detailed information at the HS10 level for the years 1997-1999 and 2001-2004 which we could access to.
 Since 2005 onwards, we use original Census data with perfect match between the 5- and 10-classification degrees.
-Accordingly, we run our robustness check to aggregation over this period 2005-2019 (Section 3.3.3).} These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.
+Accordingly, we run our robustness check to aggregation over this period 2005-2019 (Section 3.3.3).}
+These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.
 \medskip
@@
 We now explain in details, both in the paper and in our answers to referees, the reasons why a non-linear estimator is unavoidable in our case.
-This is due to the restrictions imposed \textit{ex-ante} on parameters, i.e.
+This is due to structural restrictions on parameters, i.e.
 $\tau \geq 1$ and $t\geq 0$.
 These restrictions are required because we know that both types of transport costs are positive and we cannot neglect this characteristic.
-By construction we cannot have $p_{ikt} < \widetilde{p}_{ikt}$.
+It also ensures that $p_{ikt} \geq  \widetilde{p}_{ikt}$.
 Without these restrictions, standard linear, least squares estimates deliver aberrant values with mediocre quality-of-fit.

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
could / would note 11
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Travail collectif
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 %\usepackage[latin1]{inputenc}
-%\usepackage{hyperref}
+%\usepackage{ref}
 %\usepackage[T1]{fontenc}
@@
 Second, we revisit the welfare gains of the transport cost reduction in presence of additive costs.
-In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the key role of additive transport costs in determining those welfare variations.
+In this regard, we shed light on the welfare variations induced by the international trade acceleration and the ``hyper-globalization'', as well as the key role of additive transport costs in determining those welfare variations.
 Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease.
@@
 Using Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs amount to 6\% of the export price multiplied by the ad-valorem cost.
-Like \cite{Irrazabal_2015}, we show the important role of the additive component of international trade costs.
+Like \cite{Irrazabal_2015}, we show the important role of the additive component in international trade costs.
 The literature also explores the welfare implications of additive trade costs.
 \cite{sorensen2014} analytically shows that important welfare gains can be achieved by reducing additive trade costs.
-Not much progress has yet been made in quantifying such gains, one potential reason being the lack of an empirical characterization of the additive component of trade costs, the one exceptions being \cite{Irrazabal_2015} and \cite{Hornok-et-al-JIE-2015}.
+Providing an actual quantification is difficult, one potential reason being the lack of an empirical characterization of such costs.
+Some recent progress has been made, e.g. by \cite{Irrazabal_2015}, \cite{Hornok-et-al-JIE-2015} and \cite{Lashkaripour_JIE2020}.
 In contrast to them, we can exploit the time dimension of our database to quantify the welfare gains of the trade cost reduction since 1974, based on our empirical yearly estimates over 1974-2019 of overall transport costs, as well as by sub-component (additive and ad-valorem).\smallskip
@@
 The following sections take stock of those results to draw implications for two important debates in economic analysis.
-In Section \ref{sec:results_trends}, we offer new insights on the origins of the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by sector and partner country), and what arises from changes in transport costs \textit{per se}.
+In Section \ref{sec:results_trends}, we offer new insights on the decrease of transport costs over time, by providing a decomposition of the transport costs trend patterns, between what arises from changes in the trade composition (by sector and partner country), and what arises from changes in transport costs \textit{per se}.
 Section \ref{sec:theory} implements several theory-based calibration exercises to investigate the different welfare consequences of the observed decrease in ad-valorem and additive costs.
@@
 As detailed below, the use of a non-linear estimator triggers computational limitations that constrain the level of possible detail, especially when covering a long period of time.
-Hence, our benchmark results estimate international transport costs at the 3-digit classification level, exploiting heterogeneity between products at the SITC-5 classification level.
+Hence, our benchmark estimation retains the 3-digit sectoral classification level, exploiting heterogeneity between products at the SITC-5 classification level.
 Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin (hence, around 400 fixed effects to estimate).
@@
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
-As reported in the Online Appendix, Section D.1, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
-on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
+As reported in the Online Appendix, Section D.1, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could run an OLS estimation on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
 This would not solve the problem though, for three complementary reasons.
@@
 Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
-In a related manner, we could not impose positivity constraints on the coefficients.
-Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}
+Third, in a related manner, we would not be able to impose the necessary constraints on the coefficients ($\tau\geq 1, t\geq 0 )$.}
+%Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).
 
@@
 After a period of common increase in air and vessel from the mid 1970s to the end of the 1980s, it has steadily decreased for air transport.
-In contrast, the share of additive costs has continued to increase until the 2000s.
-Put differently, the total transport costs decrease is disproportionately related to additive costs in air transport, while it is mostly driven by ad-valorem cost reduction in vessel transport.\smallskip
+In contrast, the share of additive costs has continued to increase until the 2000s for vessel.
+Put differently, the total transport costs decrease mostly comes from the reduction of additive costs in air transport, while it is mainly driven by ad-valorem cost reduction in vessel transport.\smallskip
 
@@
 Related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge different prices, all other things equal – in our case, for a given country-sector pair.
-In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.
+In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability for bargaining discounts for larger shipped volumes.
 
@@
 Firstly, we could only access the years 1997-1999 and 2001-2019.\footnote{Theoretically, they should be available from 1989, which is already smaller than our full sample starting in 1974.
-Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} Secondly, for the years 1997-1999 and 2001-2005, compatibility issues between Census and Hummels' data (which we use until 2004) prevented reliable comparisons.
+Because of the COVID situation, it was only possible to get access to the years 1997-1999 and 2001-2019.} Secondly, for the years 1997-1999 and 2001-2004, compatibility issues between Census and Hummels' data (which we use until 2004) prevented reliable comparisons.
 

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Travail collectif
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

@@
 		\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
-		\label{tab:result_air_rob}%
+		\label{tab:result_ves_rob}%
 		\input{TableB1_ves_reshaped.tex}
@@
 		\caption{Vessel: Measures of goodness of fit, 4-digit level}
-		\label{tab:good_fit_air_rob}%
+		\label{tab:good_fit_ves_rob}%
 		\input{TableB3_ves_reshaped.tex}
@@
 In the specific case of quantities, this raises a non-trivial issue for country-level fixed effects.
-Sector-level fixed effects could indeed be easily replaced by (sector times quantity unit) fixed effects; but the same does not appear possible for country fixed effects.
+Sector-level fixed effects could indeed be easily replaced by (sector $\times$ quantity unit) fixed effects; but the same does not appear possible for country fixed effects.
 Let us elaborate on this issue.

### 2021-11-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee2.tex
Micro modifs liées à la soumission + passage en WP

@@
 
-\author{Guillaume \textsc{Daudin}\thanks{%
-Université Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
+%\author{Guillaume \textsc{Daudin}\thanks{%
+%Université Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
 Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{
-University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+%University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
 

### 2021-11-29 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

@@
 
-\author{Guillaume \textsc{Daudin}\thanks{%
-Université Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
-Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{
-University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+%\author{Guillaume \textsc{Daudin}\thanks{% Université Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016, Paris, France ; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
+Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{ University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
 

### 2021-11-29 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We thank the referees, as well as the editor, William Kerr, for insightful comments, which helped to significantly improve the paper.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{%We thank the referees, as well as the editor, William Kerr, for insightful comments, which helped to significantly improve the paper.
 We are grateful to Holger G\"{o}rg, S\'{e}bastien Jean and Vincent Vicard for their critics and suggestions on earlier drafts.
@@
 This paper was partly funded by a university grant from the University of Lille.
-It features an online appendix containing additional results, available on the authors' websites and on \url{https://github.com/gdaudin/trade_costs}.}}
+It features an online appendix containing additional results, available on the authors' websites%and on \url{https://github.com/gdaudin/trade_costs}
+.}}
 

### 2021-11-29 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex

@@
 
-%Modèle additif seul DONE
-%+Tableaux avec toutes les années en 3 digit DONE
-% Tableaux avec toutes les années en 4 digit ???
-%Harmoniser la présentation des résultats (cf table 1 du papier) DONE
-%Les résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix? DONE, mais à compléter
+\author{Guillaume \textsc{Daudin}\thanks{%
+University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Corresponding author.
+Univ.
+Lille, CNRS, IESEG School of Management, UMR 9221 - LEM - Lille Économie Management, F-59000 Lille, France, and CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{
+University Paris Dauphine, PSL University, LEDa, UMR CNRS 8805- UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+
+\date{November 2021}
 

### 2021-12-11 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Resolution pb auteurs OA

--- 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage[symbol]{footmisc}
 %\usepackage[utf8]{fontenc}
@@
 \renewcommand{\thesection}{\Alph{section}}
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+%\footnote[num]{text}
 
@@
 \begin{document}
-\title{\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)}
+\title{\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix}
 
@@
 
-%Modèle additif seul DONE
-%+Tableaux avec toutes les années en 3 digit DONE
-% Tableaux avec toutes les années en 4 digit ???
-%Harmoniser la présentation des résultats (cf table 1 du papier) DONE
-%Les résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix? DONE, mais à compléter
+\author{Guillaume \textsc{Daudin}\footnote[2]{%
+University Paris Dauphine, PSL University, LEDa-DIAL, UMR CNRS 8805-UMR IRD 260, 75016 Paris, France; email: \url{guillaume.daudin@dauphine.psl.eu}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \footnote[3]{Corresponding author.
+Univ.
+Lille, CNRS, IESEG School of Management, UMR 9221 - LEM - Lille Économie Management, F-59000 Lille, France, and CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\footnote[5]{
+University Paris Dauphine, PSL University, LEDa, UMR CNRS 8805- UMR IRD 260, 75016 Paris, France;  email: \url{lise.patureau@dauphine.psl.eu} } }
+
+\date{November 2021}
 
@@
 	\renewcommand\thesubsection{\Alph{subsection}}
+
+\setcounter{footnote}{0}
 

### 2022-03-11 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_accepted_version.tex
Amendement DOI + changement paragraphe version finale

--- 2021-11-29 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
@@
 \usepackage{bbold}
-\usepackage{natbib}
+\usepackage{natbib,har2nat}
 \usepackage{babel,varioref}
 \usepackage{bbm}
+\RequirePackage{doi}
 \usepackage[flushleft]{threeparttable}
@@
 
-As detailed below, the use of a non-linear estimator triggers computational limitations that constrain the level of possible detail, especially when covering a long period of time.
-Hence, our benchmark estimation retains the 3-digit sectoral classification level, exploiting heterogeneity between products at the SITC-5 classification level.
-Depending on the year, this leaves us with around 200 3-digit sectors, from around 200 countries of origin (hence, around 400 fixed effects to estimate).
-We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i),} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
-Notice that this can theoretically be done from 1989 onwards for which HS-10 data are available; \textit{ii),} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
-Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
-However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} below and Section B of the online appendix for the 4-5 results; and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).
-
 \subsection{Estimating transport costs: empirical specification \label{sec:estimation_strat}}
@@
 See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that constrain the level of possible detail, especially when covering a long period of time.
-Confronted with this trade-off, we estimate 3-digit level sectoral fixed effects as our benchmark.
+Confronted with this trade-off, we estimate 3-digit level sectoral fixed effects as our benchmark, exploiting heterogeneity between products at the SITC-5 classification level.
 This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
@@
 Depending on the year considered, this leaves us with around 400 fixed effects to estimate by transport mode at the 3-digit level.
-\medskip
+
+We ensure the robustness of these results by conducting two types of alternative estimates: \textit{i),} within each 3-digit sector, exploiting heterogeneity between HS-10 products.
+Notice that this can theoretically be done from 1989 onwards for which HS-10 data are available; \textit{ii),} running the estimation at the 4-digit sector level (still with 5-digit products) for some selected years.
+Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation.
+However, we obtain no substantial difference between the results obtained for the benchmark 3-5 sector-product classification levels and those considering either the 3-10 and 4-5 ones (See Table \ref{tab:summary_results} below and Section B of the online appendix for the 4-5 results; and Table \ref{tab:baseline10}, Section \ref{sec:robustness} for the 3-10 estimation results).\medskip
 
@@
 
+\section{Data accessibility statement}
+
+The raw data supporting our results are openly available at: \texttt{https://doi.org/10.5281/zenodo.6344001}.
+The associated Stata programs are openly available at: \texttt{https://doi.org/10.5281/zenodo.6344353}.
+See also \cite{DHP2022-DOI-data} and \cite{DHP2022-DOI-do}.
+
 \newpage
-\bibliographystyle{apalike2}
+\bibliographystyle{erae}
 \bibliography{biblio}

### 2022-03-11 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_accepted_version.tex
Ajour remerciements référés-éditeur

@@
 
-\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{%We thank the referees, as well as the editor, William Kerr, for insightful comments, which helped to significantly improve the paper.
+\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We warmly thank the referees, as well as the editor, William Kerr, for insightful comments, which helped to significantly improve the paper.
 We are grateful to Holger G\"{o}rg, S\'{e}bastien Jean and Vincent Vicard for their critics and suggestions on earlier drafts.
